### 2017-07-18 etc/uWUE_gw_notes/wue_gs.tex
added medlyn model, not need to test if scipy optmiziers can solve efficiently

\input{def}
%\usepackage{amsmath}
\usepackage{graphics, graphicx}
\graphicspath{ {./} } %directory for figures 
%\pagestyle{empty}

\begin{document}
\begin{center}
\large{Using uWUE to remove dependence of ET on GPP}\end{center}

\bigskip

Penman montieth is given by:

\begin{equation}
  \lambda E = \frac{\Delta A + \frac{\rho_a c_p D_{ref}}{r_a}}{\Delta + \gamma(1 + \frac{r_s}{r_a})}
\end{equation}

where:
\begin{equation}
  r_s = \frac{1}{LAI \cdot  g_{st}}
\end{equation}

Using the Medlyn model (from Frank et. al. 2017) we have:

\begin{equation}
g_{st} = g_{w} = 1.6 \left( 1 + \frac{g_{1M}}{\sqrt{D_s}} \right) \frac{A}{c_s}
\end{equation}

Here, $g_{1M}$ has units of kPa$^{0.5}$, $A$ is the net CO$_2$ assimilation rate in $\mu$ mol m$^{-2}$ s$^{-1}$, $c_s$ is the CO$_2$ concentration in ppm (supposed to be at leaf surface). The problem is the $A$ term that we need to get rid of. Also $g_w$ as above has units of mol (air?) m$^{-2}$ s$^{-1}$.

we can use:

\begin{equation}
  \text{uWUE} = \frac{\text{GPP} \cdot \sqrt{\text{VPD}}}{ET}
\end{equation}

This has units of $\frac{\text{g C  hPA}^{0.5}}{\text{kg H$_2$O}}$.

So let's do these unit conversions, starting with $g_w$ unit converstion. Note that the mean molecular mass of dry air is $28.9645$ g/mol. So then we have:

\begin{equation}
  g_{st\; -\; m/s} = g_w \frac{mol}{m^2 s^1} \frac{28.9645 g air}{ mol air} \frac{1 kg}{1000 g} \frac{m^3}{\rho \; kg} = \frac{m}{s}
\end{equation}

so:
\begin{equation}
  g_{st\; -\; m/s} = g_w \frac{R_{air}}{\rho_{air}}
\end{equation}

where:
\begin{equation}
  R_{air} = .0289645 \frac{\text{kg}}{\text{mol}}
\end{equation}

Now we just need to convert uWUE from gC to $\mu$ mol C:

\begin{equation}
uWUE_{\mu \; \text{mol C}} = uWUE \frac{ \text{mol C}}{12.011 \text{g C}} \frac{1.e6 \; \mu \; \text{mol}}{mol}
\end{equation}

so:
\begin{equation}
  uWUE = \frac{uWUE \;1.e6}{ 12.011}
\end{equation}

\end{document}

### 2017-07-18 src/codebase/penman_monteith.py

@@ -13,6 +13,8 @@ VP_FACTOR = 100. #convert hPa -> Pa
 K = 0.41 # vonkarmans constant
 CP = 1004. # specific heat air
 GAMMA = 66. # psychrometric constant
+LV = 2.5e6
+R_AIR = .0289645 # mean kg/mol dry air
 
 # From Changjie's oren model for stomatal resistance
 # see "Survey and synthesis of intra- and interspecific variation
@@ -21,12 +23,33 @@ GAMMA = 66. # psychrometric constant
 OREN = pd.read_csv('../dat/orens_model.csv')
 # convert to m/s
 OREN.iloc[:, 2:6] = OREN.iloc[:, 2:6]/1000.
-OREN.index = OREN.PFTs
+OREN.index = OREN.PFT
 
-#below need to get actual values for Chiangjie
-MEDLYN = pd.read_csv('../dat/orens_model.csv')
-MEDLYN['g0'] = MEDLYN.G1mean
-MEDLYN['g1'] = MEDLYN.G2mean
+#below taken from Franks et al (with Berry)
+MEDLYN = pd.read_csv('../dat/franks_et_al_table2.csv',\
+                     comment='#', delimiter=',')
+MEDLYN.index = MEDLYN.PFT
+
+WUE = pd.read_csv('../dat/zhou_et_al_table_4.csv',\
+                     comment='#', delimiter=',')
+WUE.index = WUE.PFT
+
+
+def medlyn_g_w(vpd, co2, pft, et):
+    """
+    returns leaf stomatal conductance in m/s
+    et : W/m2
+    pft : planf functional type
+    vpd : Pa
+    co2 : ppm
+    """
+    #convert g C -> mu mol C
+    wue = WUE.loc[atmos['pft'],'u_wue_yearly']*1.e6/12.011
+    # note bellow assumes that atmos co2 is same as leaf, might be bad
+    _g1 = MEDLYN.loc[atmos['pft'],'g1M'] # note this sqrt(kPa)
+    g_w = 1.6*(1. + _g1/np.sqrt(vpd/1000.))*wue*et/LV/sqrt(vpd/100.)/co2
+    return g_w
+
 
 def oren_r_l(vpd, pft):
     """
@@ -49,7 +72,7 @@ def r_a(atmos, canopy):
             /K**2/atmos['u_z']
 
 def r_s(atmos, canopy):
-    """returns stomatal resistance in s/m"""
[…]

### 2017-07-18 src/test_medlyn_solver.py

@@ -1,0 +1,5 @@
+#! ~/edward/bin/python
+"""
+This is a module to test if vpd solver is working.
+"""
+

### 2017-07-18 src/vpd_driver_reducer.py

@@ -72,19 +72,21 @@ if str(__name__) == "__main__":
     ATMOS['rho_a'] = 1.205 #density kg/m3
     ATMOS['t_a'] = np.linspace(15., 35.) # C
     ATMOS['rh'] = np.linspace(20., 99.) # rel humdidity
-    ATMOS['t_a'], ATMOS['rh'] = np.meshgrid(ATMOS['t_a'], ATMOS['rh'])
+
     ATMOS['u_z'] = 2. #wind speed at meas. hiehgt (m/s)
     ATMOS['vpd'] = met.vapor_pres(ATMOS['t_a'])*(1.-ATMOS['rh']/100.)*100.
 
-    print(ATMOS['vpd'].max())
-    PERT = copy.deepcopy(ATMOS)
-    PERT['vpd'] += 1. # add 1 hpa for perturbation
-    print(ATMOS['vpd'].max())
-    print(PERT['vpd'].max())
 
     CANOPY = {}
     CANOPY['pft'] = 'EBF'
     CANOPY['height'] = 10. # plant heigh m
     CANOPY['lai'] = 1. # leaf area index pierre says 1 max feasible
+
+    ATMOS['t_a'], ATMOS['rh'],  = np.meshgrid(ATMOS['t_a'], ATMOS['rh'])
+
+    PERT = copy.deepcopy(ATMOS)
+    PERT['vpd'] += 1. # add 1 hpa for perturbation
+
+
 
     RESULT = d_et_d_vpd(ATMOS, CANOPY, PERT)

### 2017-07-19 src/vpd_driver_reducer.py
got global warming experiment working, still seems way too sensitive to co2 concentration. also made it so looping for medlyn penman monteith is in penman_moneith moduel. this is much better organziationally, but appeared to slow down the script (dictionary allocation is slow?).

@@ -7,12 +7,12 @@ import os
 import copy
 import importlib
 from collections import OrderedDict
+import time
 import numpy as np
 import matplotlib as mpl
 import matplotlib.pyplot as plt
 import codebase.penman_monteith as pm
 import metcalcs as met
-
 
 importlib.reload(pm)
 
@@ -58,7 +58,7 @@ def plot_results(result, atmos):
     return
 
 
-def d_et_d_vpd(atmos, canopy, pert, var=('vpd', 'vpd_leaf'), dvar=1.):
+def d_et_d_vpd(atmos, canopy, pert, dvar=1.):
     """
     numerically calculate dET/dpert
 
@@ -73,37 +73,26 @@ def d_et_d_vpd(atmos, canopy, pert, var=('vpd', 'vpd_leaf'), dvar=1.):
         pft :: plant functional type
         height :: crop height (m)
         lai :: leaf area index (pierre says 1 is max feasible)
-    pert :: dictionary of atmos variables with 1 changed
+    pert :: dictionary of atmos variables with changed
     var :: variable key derivative is taken wrt
     thresh :: this is a p/m min/max threshold for the output
              sometimes the solver struggles, so this hopes to fix
     dvar :: this is what the difference in the variable is (defaul 1 hPa)
     """
-    # there is probably a much better way to do below
-    _atmos = atmos.copy()
-    _canopy = canopy.copy()
-    _it = np.nditer([atmos['t_a'], atmos['rh'], atmos['vpd'],\
-                    atmos['co2'], pert[var[0]], pert[var[1]], None])
-    for _atmos['t_a'], _atmos['rh'], _atmos['vpd'], _atmos['co2'],\
-        pert_vpd, pert_vpd2, result in _it:
-        _pert = _atmos.copy()
-        _pert[var[0]] = pert_vpd
-        _pert[var[1]] = pert_vpd2
-        # for key in _atmos:
-        #     print(key,_atmos[key])
-        # for key in _pert:
-        #     print(key,_pert[key])
-        result[...] = (pm.medlyn_penman_monteith(_pert, _canopy)\
-                       -pm.medlyn_penman_monteith(_atmos, _canopy))\
+
+    result = (pm.medlyn_penman_monteith(pert, canopy)\
[…]

### 2017-08-28 etc/2017-08-28_VPD_uWUE_notes/explicit.tex
added updated notes on analytical ET and dET/dVPD, also think I like the LAI model so now need to decide if want to calcualte unique LAI for each timestep and then calcualte dET/DVPD derivatives for that. Could be noisy and unphysical, but who know maybe LAI relatively constant

\input{def}
%\usepackage{amsmath}
\usepackage{graphics, graphicx}
\graphicspath{ {./} } %directory for figures 
%\pagestyle{empty}

\begin{document}
\begin{center}
\large{Using uWUE to remove dependence of ET on GPP}\end{center}

\bigskip

This document is to analytically caluclate ET and $\frac{\partial ET} {\partial VPD}$.

Penman montieth is given by:

\begin{equation}
  \lambda E = \frac{\Delta R + g_a \rho_a c_p D_{s}}{\Delta + \gamma(1 + \frac{g_a}{g_s})}
\end{equation}

where from Medlyn we have:
\begin{equation}
  g_s = \text{LAI } 1.6 \left(1 + \frac{g_1}{\sqrt{D_{s}}}\right) \frac{A}{c_s}
\end{equation}

Here, $g_{1M}$ has units of kPa$^{0.5}$, $A$ is the net CO$_2$ assimilation rate in $\mu$ mol m$^{-2}$ s$^{-1}$, $c_s$ is the CO$_2$ concentration in ppm (supposed to be at leaf surface). The problem is the $A$ term that we need to get rid of. Also $g_w$ as above has units of mol (air?) m$^{-2}$ s$^{-1}$.

we can use:

\begin{equation}
  \text{uWUE} = \frac{\text{GPP} \cdot \sqrt{\text{VPD}}}{ET}
\end{equation}

As long as we put uWUE into $\mu mol$, we can use the uWUE provided by Zhou et. al. To do this just multiply be a factor $ \frac{1 mol C}{12.011 g C} \frac{1.e6 \mu mol}{1 mol}$. Then make sure that units of VPD match. If using Franks et al. 2017 and Zhou et al. then VPD within the medlyn model should be in kPa, and VPD cancelling uWUE should be in hPa.

Then we can show that if we multiply $g_s$ by $\frac{ uWUE \; ET}{A \sqrt{D_s}}$ (which is equivalent to 1 with unit conversion above), we can rearrange Penman Moneith to get:

\begin{equation}
  \lambda E = \frac{\Delta R + g_a \left( \rho_a c_p D_{s} - \frac{\gamma c_s \sqrt{D_s}}{\text{LAI }1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{D_s}})} \right) }{ \Delta + \gamma}
\end{equation}

If we want to solve for LAI as a function of all other knowns, we have:
\begin{equation}
LAI  = \frac{g_a \gamma c_s \sqrt{D_s}}{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p D_{s}\right) 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{D_s}})}
\end{equation}

Now taking the derivative w.r.t. $D_s$ we get:

\begin{equation}
\frac{\partial \;  \lambda E}{\partial \; D_s} = \frac{g_a}{\Delta + \gamma}   \left( \rho_a c_p - \frac{\gamma c_s }{\text{LAI }1.6 \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D_s}}{2 (g_1 + \sqrt{D_s})^2}\right) \right)
\end{equation}

\end{document}

### 2017-08-28 src/calc_et_vpd_derivative.py
redoing calc derivative to be correct

@@ -23,31 +23,39 @@ SITELIST = pd.read_csv('%s/changjie/fluxnet_algorithm/'\
                        delimiter=',')
 SITELIST.index = SITELIST.Site
 
+def calc_derivative(atmos, canopy, data):
+  """adds various derivative fields to data, given atmos and canopy"""
+  data['et'] = pm.penman_monteith_uwue(atmos, canopy)
+  data['scaling'] = 1./(atmos['r_a']*(atmos['delta'] + atmos['gamma']))
+  data['vpd_atm'] = atmos['rho_a']*pm.CP
+  data['vpd_leaf'] = atmos['gamma']*atmos['c_a']/\
+                     (canopy['lai']*1.6*canopy['uwue'])\
+                     *(2.*canopy['g1'] + np.sqrt(atmos['vpd']))\
+                     /(2.*(canopy['g1'] + np.sqrt(atmos['vpd']))**2)
+  data['lai_fit'] = atmos['gamma']*atmos['c_a']*np.sqrt(atmos['vpd'])\
+                    /(atmos['r_a']\
+                      *(data['et_obs']*(atmos['gamma']+atmos['delta'])\
+                        -atmos['delta']*(atmos['r_n']-canopy['g_flux'])\
+                        -1./atmos['r_a']*atmos['rho_a']*pm.CP*atmos['vpd'])
+                      *1.6*canopy['uwue']*(1.+canopy['g1']/atmos['vpd']))
+  return
 #def main():
 """wrapper for main script"""
 
 
 start = time.time()
-coef = pd.read_csv('../dat/site_coef_mm_s_W_m2_medlyn_lai.csv')
+coef = pd.read_csv('../dat/site_coef_mm_s_W_m2_medlyn_lai_nolim.csv')
 coef.index = coef['Unnamed: 0']
-outdir = '%s/changjie/pandas_data_lai_fit/' % os.environ['DATA']
-for i, index in enumerate(coef.index[:]):
+outdir = '%s/changjie/pandas_data_lai_fit_nolim/' % os.environ['DATA']
+for i, index in enumerate(coef.index[:1]):
   print('Working on %s, file number %d, time elapsed: %f m' \
         % (index, i, (time.time()-start)/60.))
   atmos, canopy, data = d_io.load_mat_data(index)
   canopy['et_stomatal_model'] = 'adam_medlyn'
[…]

### 2017-08-28 src/calc_et_vpd_derivative.py
put all units into Pa, better this way but could be some lingering errors

@@ -23,22 +23,31 @@ SITELIST = pd.read_csv('%s/changjie/fluxnet_algorithm/'\
                        delimiter=',')
 SITELIST.index = SITELIST.Site
 
+def leaf_vpd(atmos, canopy, lai):
+  """calculates the leaf term in dET/dVPD (see doc folder)"""
+  return atmos['gamma']*atmos['c_a']*atmos['p_a']*/\
+    (lai*1.6*pm.R_STAR*(273.15+atmos['t_a'])*canopy['uwue'])\
+    *(2.*canopy['g1'] + np.sqrt(atmos['vpd']))\
+    /(2.*(canopy['g1'] + np.sqrt(atmos['vpd']))**2)
+
+
 def calc_derivative(atmos, canopy, data):
   """adds various derivative fields to data, given atmos and canopy"""
   data['et'] = pm.penman_monteith_uwue(atmos, canopy)
   data['scaling'] = 1./(atmos['r_a']*(atmos['delta'] + atmos['gamma']))
   data['vpd_atm'] = atmos['rho_a']*pm.CP
-  data['vpd_leaf'] = atmos['gamma']*atmos['c_a']/\
-                     (canopy['lai']*1.6*canopy['uwue'])\
-                     *(2.*canopy['g1'] + np.sqrt(atmos['vpd']))\
-                     /(2.*(canopy['g1'] + np.sqrt(atmos['vpd']))**2)
-  data['lai_fit'] = atmos['gamma']*atmos['c_a']*np.sqrt(atmos['vpd'])\
+  data['vpd_leaf'] = leaf_vpd(atmos, canopy, canopy['lai'])
+  data['lai_fit'] = atmos['gamma']*atmos['c_a']\
+                    *np.sqrt(atmos['vpd'])*atmos['p_a']\
                     /(atmos['r_a']\
                       *(data['et_obs']*(atmos['gamma']+atmos['delta'])\
                         -atmos['delta']*(atmos['r_n']-canopy['g_flux'])\
                         -1./atmos['r_a']*atmos['rho_a']*pm.CP*atmos['vpd'])
-                      *1.6*canopy['uwue']*(1.+canopy['g1']/atmos['vpd']))
-  return
+                      *1.6*pm.R_STAR*(273.15 + canopy['t_a'])\
+                      *canopy['uwue']*(1.+canopy['g1']/atmos['vpd']))
[…]

### 2017-08-28 src/fit_et_medlyn_model.py

@@ -29,8 +29,8 @@ def medlyn_fit_et(g_coef, *args):
   canopy['lai'] = g_coef[0]
   # canopy['g1'] = g_coef[1]
   data['et'] = pm.penman_monteith_uwue(atmos, canopy)
-  data.loc[data.et > 1000., 'et'] = 1000.
-  data.loc[data.et <= 0., 'et'] = 0.
+  # data.loc[data.et > 1000., 'et'] = 1000.
+  # data.loc[data.et <= 0., 'et'] = 0.
   return data['et'] - data['et_obs']
 
 def calc_coef():
@@ -75,7 +75,7 @@ def calc_coef():
     else:
       print('ERRORR!!!! filename %s failed!' % filename)
       #_coef.loc[filename, 'g1'] = result['x'][1]
-      _coef.loc[filename, 'g1'] = canopy['g1'].iloc[0]
+    _coef.loc[filename, 'g1'] = canopy['g1'].iloc[0]
     _coef.loc[filename, 'PFT'] = canopy['pft'].iloc[0]
     _coef.loc[filename, 'r2'] = 1. - \
                                 np.sum(medlyn_fit_et(result['x'], atmos,\
@@ -105,10 +105,10 @@ def generate_coef_stats(_coef):
 # def main():
 """wrapper for main script"""
 coef = calc_coef()
-coef.to_csv('../dat/site_coef_mm_s_W_m2_medlyn_lai.csv')
+coef.to_csv('../dat/site_coef_mm_s_W_m2_medlyn_lai_nolim.csv')
 statistics = coef.groupby('PFT').apply(generate_coef_stats)
 statistics.index = statistics.index.droplevel(1)
-outdir = '../dat/adam_mm_s_W_m2_medlyn_lai.csv'
+outdir = '../dat/adam_mm_s_W_m2_medlyn_lai_nolim.csv'
 statistics.to_csv(outdir)
 #   return coef
 

### 2017-08-29 etc/2017-08-28_VPD_uWUE_notes/explicit.tex
redoing calc derivative to be correct
put all units into Pa, better this way but could be some lingering errors
NOTE DELETING OLD PLOTTING CODE THAT COULD BE USEFUL, also found order of mangitude error in new model but still not 1:1

@@
 \begin{equation}
-  g_s = \text{LAI } 1.6 \left(1 + \frac{g_1}{\sqrt{D_{s}}}\right) \frac{A}{c_s}
+  g_s = \text{LAI } \frac{R T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{D_{s}}}\right) \frac{A}{c_s}
 \end{equation}
@@
 
-Then we can show that if we multiply $g_s$ by $\frac{ uWUE \; ET}{A \sqrt{D_s}}$ (which is equivalent to 1 with unit conversion above), we can rearrange Penman Moneith to get:
+Then we can show that if we multiply $g_s$ by $\frac{ uWUE \; ET}{A \sqrt{D_s} L_v}$ (which is equivalent to 1 with unit conversion above), we can rearrange Penman Moneith to get:
 
 \begin{equation}
-  \lambda E = \frac{\Delta R + g_a \left( \rho_a c_p D_{s} - \frac{\gamma c_s \sqrt{D_s}}{\text{LAI }1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{D_s}})} \right) }{ \Delta + \gamma}
+  \lambda E = \frac{\Delta R + g_a \left( \rho_a c_p D_{s} - \frac{\gamma c_s \sqrt{D_s} L_v P }{\text{LAI } R \; T \; 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{D_s}})} \right) }{ \Delta + \gamma}
 \end{equation}
@@
 \begin{equation}
-LAI  = \frac{g_a \gamma c_s \sqrt{D_s}}{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p D_{s}\right) 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{D_s}})}
+LAI  = \frac{g_a \gamma c_s \sqrt{D_s} L_v P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p D_{s}\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{D_s}})}
 \end{equation}
@@
 \begin{equation}
-\frac{\partial \;  \lambda E}{\partial \; D_s} = \frac{g_a}{\Delta + \gamma}   \left( \rho_a c_p - \frac{\gamma c_s }{\text{LAI }1.6 \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D_s}}{2 (g_1 + \sqrt{D_s})^2}\right) \right)
+\frac{\partial \;  \lambda E}{\partial \; D_s} = \frac{g_a}{\Delta + \gamma}   \left( \rho_a c_p - \frac{\gamma c_s L_v  P }{\text{LAI }1.6 \; R\; T \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D_s}}{2 (g_1 + \sqrt{D_s})^2}\right) \right)
 \end{equation}

### 2017-08-29 src/calc_et_vpd_derivative.py

@@ -27,7 +27,7 @@ SITELIST.index = SITELIST.Site
 
 def leaf_vpd(atmos, canopy, lai):
   """calculates the leaf term in dET/dVPD (see doc folder)"""
-  return atmos['gamma']*atmos['c_a']*atmos['p_a']/\
+  return atmos['gamma']*atmos['c_a']*pm.LV*atmos['p_a']/\
     (lai*1.6*pm.R_STAR*(273.15+atmos['t_a'])*canopy['uwue'])\
     *(2.*canopy['g1'] + np.sqrt(atmos['vpd']))\
     /(2.*(canopy['g1'] + np.sqrt(atmos['vpd']))**2)
@@ -35,14 +35,14 @@ def leaf_vpd(atmos, canopy, lai):
 
 def calc_derivative(atmos, canopy, data):
   """adds various derivative fields to data, given atmos and canopy"""
-  canopy['lai'] = atmos['gamma']*atmos['c_a']\
+  canopy['lai'] = atmos['gamma']*atmos['c_a']*pm.LV\
                 *np.sqrt(atmos['vpd'])*atmos['p_a']\
                 /(atmos['r_a']\
                   *(data['et_obs']*(atmos['gamma']+atmos['delta'])\
                     -atmos['delta']*(atmos['r_n']-canopy['g_flux'])\
                     -1./atmos['r_a']*atmos['rho_a']*pm.CP*atmos['vpd'])
                   *1.6*pm.R_STAR*(273.15 + atmos['t_a'])\
-                  *canopy['uwue']*(1.+canopy['g1']/atmos['vpd']))
+                  *canopy['uwue']*(1.+canopy['g1']/np.sqrt(atmos['vpd'])))
   data['et'] = pm.penman_monteith_uwue(atmos, canopy)
   data['scaling'] = 1./(atmos['r_a']*(atmos['delta'] + atmos['gamma']))
   data['vpd_atm'] = atmos['rho_a']*pm.CP

### 2017-08-29 src/calc_et_vpd_derivative.py
sign error in LAI,fixed now all debugging should be done

@@ -35,7 +35,7 @@ def leaf_vpd(atmos, canopy, lai):
 
 def calc_derivative(atmos, canopy, data):
   """adds various derivative fields to data, given atmos and canopy"""
-  canopy['lai'] = atmos['gamma']*atmos['c_a']*pm.LV\
+  canopy['lai'] = - atmos['gamma']*atmos['c_a']*pm.LV\
                 *np.sqrt(atmos['vpd'])*atmos['p_a']\
                 /(atmos['r_a']\
                   *(data['et_obs']*(atmos['gamma']+atmos['delta'])\

### 2017-08-30 etc/2017-08-28_VPD_uWUE_notes/explicit.tex
sign error in LAI,fixed now all debugging should be done
debugged scatter plot, and added code to make panels. sent new plots to pierre

@@
 \begin{equation}
-  \lambda E = \frac{\Delta R + g_a \rho_a c_p D_{s}}{\Delta + \gamma(1 + \frac{g_a}{g_s})}
+  ET = \frac{\Delta R + g_a \rho_a c_p D_{s}}{\Delta + \gamma(1 + \frac{g_a}{g_s})}
 \end{equation}
@@
 Then make sure that units of VPD match.
-If using Franks et al.
+Note the table in Franks et al.
 2017 and Zhou et al.
-then VPD within the medlyn model should be in kPa, and VPD cancelling uWUE should be in hPa.
+have VPD in kPa and hPa, respectively.
 
@@
 \begin{equation}
-  \lambda E = \frac{\Delta R + g_a \left( \rho_a c_p D_{s} - \frac{\gamma c_s \sqrt{D_s} L_v P }{\text{LAI } R \; T \; 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{D_s}})} \right) }{ \Delta + \gamma}
+  ET = \frac{\Delta R + g_a \left( \rho_a c_p D_{s} - \frac{\gamma c_s \sqrt{D_s} L_v P }{\text{LAI } R \; T \; 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{D_s}})} \right) }{ \Delta + \gamma}
 \end{equation}
 
-If we want to solve for LAI as a function of all other knowns, we have:
+If we want to solve for a ``pseudo-LAI'' (in practice this will also be a function of noise and model error in addition to leaf area) as a function of all other knowns, we have:
 \begin{equation}
-LAI  = \frac{g_a \gamma c_s \sqrt{D_s} L_v P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p D_{s}\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{D_s}})}
+LAI  = - \frac{g_a \gamma c_s \sqrt{D_s} L_v P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p D_{s}\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{D_s}})}
 \end{equation}
 
-Now taking the derivative w.r.t. $D_s$ we get:
+Now also taking the derivative of ET w.r.t. $D_s$ we get:
 
 \begin{equation}
-\frac{\partial \;  \lambda E}{\partial \; D_s} = \frac{g_a}{\Delta + \gamma}   \left( \rho_a c_p - \frac{\gamma c_s L_v  P }{\text{LAI }1.6 \; R\; T \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D_s}}{2 (g_1 + \sqrt{D_s})^2}\right) \right)
+\frac{\partial \;  ET}{\partial \; D_s} = \frac{g_a}{\Delta + \gamma}   \left( \rho_a c_p - \frac{\gamma c_s L_v  P }{\text{LAI }1.6 \; R\; T \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D_s}}{2 (g_1 + \sqrt{D_s})^2}\right) \right)
 \end{equation}

### 2017-08-30 src/plot_et_vpd_derivative.py

@@ -62,19 +62,16 @@ def make_ax_plot(_ax, var, _df, plot_meta):
     vmax = 0.8 # *_df.et_obs.mean()
   nstd = 1.0
   print(_df['pft'][0])
-  print(var.std())
+  # print(var.std())
   print('mean', var.mean())
-  # var[var > (var.mean() + nstd*var.std())] = np.nan
-  # var[var < (var.mean() - nstd*var.std())] = np.nan
-  # vmax = 5.*var.mean() #var.mean() + nstd*var.std()
-
   vmin = -vmax
   if plot_meta['cmap'] == 'viridis':
     vmax = var.mean() + 0.8 # 5.*var.mean() #nstd*var.std()
     vmin = var.mean() - 0.8 # 5.*var.mean() #nstd*var.std()
 
   color = _ax.scatter(_df[plot_meta['x_axis']], _df['t_a'], c=var, alpha=0.5,\
-                      s=1, cmap=plot_meta['cmap'], vmin=vmin, vmax=vmax)
+                      s=plot_meta['size'], cmap=plot_meta['cmap'],\
+                      vmin=vmin, vmax=vmax)
   if (plot_meta['x_axis'] == 'vpd'):
     t_a = np.linspace(_df['t_a'].min(),_df['t_a'].max(), 200.)
     test = met.vapor_pres(t_a)*100.*(1. - 0.90)
@@ -93,28 +90,27 @@ def make_ax_plot(_ax, var, _df, plot_meta):
 
 def soil_moisture_scatter(_df, plot_meta):
   """
-  plots sam plots as scatter_plot, but going down is different
+  plots scatter_plots, but going down is different
   percentiels of soil moisture insetad of leaf VPD, etc.
   """
   fig = plt.figure()
+  plot_meta['size'] = 16
   percentiles = [ .25, .50, .75, 1.00]
   nplots = len(percentiles)
   _df['var'] = _df['scaling']*(_df['vpd_atm'] + _df['vpd_leaf'])
-  plot_meta['folder_label'] = '%s_%s' % (plot_meta['folder_label'], 'swc')
   fig.set_figheight(fig.get_figheight()*nplots)
   ax = []
 
   for i, percentile in enumerate(percentiles):
-    ax.append(fig.add_subplot(nplots, i+1, 1))
-    var = _df['var'].loc[(_df['swc'] > _df.swc.quantile(q=percentile-0.25)) &\
[…]

### 2017-08-31 src/codebase/data_io.py
forgot to add typo fixes in explicit and formatting consistency change in data_io

@@ -53,7 +53,7 @@ def load_mat_data(filename):
   # below mm/s
   _data['et_obs'] = np.squeeze(data['LE'])
   _data['swc'] = np.squeeze(data['SWC'])
-  _data['gpp'] = np.squeeze(data['GEP'])
+  _data['gpp_obs'] = np.squeeze(data['GEP'])
 
   atmos = atmos_dict(data)
   canopy = canopy_dict(data)

### 2017-08-31 src/plot_et_vpd_derivative.py
everything looks good except for d_gpp

@@ -10,10 +10,21 @@ import matplotlib.pyplot as plt
 import numpy as np
 import metcalcs as met
 import seaborn as sns
+import resource
 from scipy.stats import spearmanr
 # import matplotlib as mpl
 # mpl.rcParams.update(mpl.rcParamsDefault)
 
+resource.setrlimit(resource.RLIMIT_AS, (4000e6, 4000e6))
+
+def test_savefig(fname):
+  """tries to save a figure and makes a folder if it doesn't exist"""
+  try:
+    plt.savefig(fname)
+  except FileNotFoundError:
+    os.system('mkdir %s' % '/'.join(fname.split('/')[:-1]))
+    plt.savefig(fname)
+  return
 
 def split_df(_df):
   """
@@ -50,12 +61,12 @@ def concat_dfs(folder='pandas_data_v2', fname='full_pandas_v2'):
   return full_df
 
 
-def make_ax_plot(_ax, var, _df, plot_meta):
+def make_ax_plot(_ax, var, _df, meta):
   """makes an axis plot"""
-  if plot_meta['log'] == 'log':
+  if meta['log'] == 'log':
     var = var/_df['et_obs']
     vmax = 0.8/175.
-  elif plot_meta['log'] == 'scaling':
+  elif meta['log'] == 'scaling':
     var = var/_df['scaling']
     vmax = 2000.
   else:
@@ -65,36 +76,36 @@ def make_ax_plot(_ax, var, _df, plot_meta):
   # print(var.std())
   print('mean', var.mean())
   vmin = -vmax
-  if plot_meta['cmap'] == 'viridis':
+  if meta['cmap'] == 'viridis':
     vmax = var.mean() + 0.8 # 5.*var.mean() #nstd*var.std()
     vmin = var.mean() - 0.8 # 5.*var.mean() #nstd*var.std()
 
-  color = _ax.scatter(_df[plot_meta['x_axis']], _df['t_a'], c=var, alpha=0.5,\
-                      s=plot_meta['size'], cmap=plot_meta['cmap'],\
+  color = _ax.scatter(_df[meta['x_axis']], _df['t_a'], c=var, alpha=0.5,\
+                      s=meta['size'], cmap=meta['cmap'],\
                       vmin=vmin, vmax=vmax)
-  if (plot_meta['x_axis'] == 'vpd'):
+  if (meta['x_axis'] == 'vpd'):
[…]

### 2017-09-01 src/codebase/penman_monteith.py
tweaked formulation to break up roe, really makes atm term constant and better to analyze

@@ -17,7 +17,7 @@ GAMMA = 66. # psychrometric constant
 LV = 2.5e6
 R_AIR = .0289645 # mean kg/mol dry air
 R_STAR = 8.3144598 #J /mol/K
-
+R_DRY = 287.058
 # From Changjie's oren model for stomatal resistance
 # see "Survey and synthesis of intra- and interspecific variation
 # in stomatal sensitivity to vapour pressure deficit" - Oren
@@ -282,10 +282,14 @@ def gamma(_atmos):
   """calculates gamma in Pa/C from pressure and T"""
   return 1.e-6*CP*_atmos['p_a']/(0.622*(2.501-0.00236*_atmos['t_a']))
 
+def r_moist(_atmos):
+  """calculates the gas constant of moist air"""
+  return R_DRY/(1.-(0.378*_atmos['e_s']\
+                    /_atmos['p_a']*_atmos['rh']/100.))
+
 def rho_air(_atmos):
   """returns rho in kg/m3"""
-  return (_atmos['p_a']/(287.05*(_atmos['t_a']+273.15)))\
-    *(1.-(0.378*_atmos['e_s']/_atmos['p_a']*_atmos['rh']/100.))
+  return _atmos['p_a']/(_atmos['r_moist']*(_atmos['t_a']+273.15))
 
 def penman_monteith_prep(_atmos, _canopy):
   """
@@ -297,6 +301,7 @@ def penman_monteith_prep(_atmos, _canopy):
   _atmos['e_s'] = met.vapor_pres(_atmos['t_a'])*VP_FACTOR
   _atmos['delta'] = delta(_atmos)
   _atmos['gamma'] = gamma(_atmos)
+  _atmos['r_moist'] = r_moist(_atmos)
   _atmos['rho_a'] = rho_air(_atmos)
 
   # below allows us to provide vpd

### 2017-09-15 etc/2017-08-28_VPD_uWUE_notes/explicit.tex
added pierre's comments and GPP notes, now need to add WUE and then code it all up
added WUE to document, now need to code up
forgot to add typo fixes in explicit and formatting consistency change in data_io
everything looks good except for d_gpp
tweaked formulation to break up roe, really makes atm term constant and better to analyze
minor typo in notes
added other derivatives to d_et

@@
 \begin{center}
-\large{Using uWUE to remove dependence of ET on GPP}\end{center}
+\large{Using uWUE to analyticaly evaluate VPD effect on WUE, GPP, and ET}\end{center}
 
@@
 Penman montieth is given by:
-
+n
 \begin{equation}
@@
 \begin{equation}
-  ET = \frac{\Delta R + g_a \left( \rho_a c_p D_{s} - \frac{\gamma c_s \sqrt{D_s} L_v P }{\text{LAI } R \; T \; 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{D_s}})} \right) }{ \Delta + \gamma}
+  ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p D_{s}}{R_{air}} - \frac{\gamma c_s \sqrt{D_s} L_v }{\text{LAI } R* \; 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{D_s}})} \right) }{ \Delta + \gamma}
 \end{equation}
@@
 \begin{equation}
-\frac{\partial \;  ET}{\partial \; D_s} = \frac{g_a}{\Delta + \gamma}   \left( \rho_a c_p - \frac{\gamma c_s L_v  P }{\text{LAI }1.6 \; R\; T \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D_s}}{2 (g_1 + \sqrt{D_s})^2}\right) \right)
+\frac{\partial \;  ET}{\partial \; D_s} = \frac{g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s L_v  }{\text{LAI }1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D_s}}{2 (g_1 + \sqrt{D_s})^2}\right) \right)
+\end{equation}
+
+Now that we have ET, we can also use uWUE to get GPP:
+
+\begin{equation}
+  \frac{ET}{L_v} = \frac{\text{GPP} \cdot \sqrt{D_s}}{  \text{uWUE}}
+\end{equation}
+
+Plugging this into Equation (4) gives:
+
+\begin{equation}
+  \text{GPP} = \frac{\text{uWUE}\; \Delta R + g_a \left( \text{uWUE}\; \rho_a c_p D_{s} - \frac{\gamma c_s \sqrt{D_s} L_v P }{\text{LAI } R \; T \; 1.6  (1 + \frac{g_1}{\sqrt{D_s}})} \right) }{L_v \; \sqrt{D_s} \left(\Delta + \gamma \right)}
+\end{equation}
+
+Taking the derivative w.r.t $D_s$ gives:
+
+\begin{equation}
+  \frac{\partial GPP}{\partial D_s} = \frac{1}{2 L_v D_s^{3/2} \left(\Delta + \gamma \right)} \left( -\text{uWUE}\; \Delta R + g_a \text{uWUE}\; \rho_a c_p  D_s - \frac{g_a g_1  \gamma c_s L_v P }{\text{LAI } R \; T \; 1.6 ( 1 + \frac{g_1}{\sqrt{D_s}})^2 } \right)
+\end{equation}
+
+We can also solve for WUE, again using uWUE (just make sure units are correct: mol C/mol H$_2$O):
+
+\begin{equation}
+\frac{GPP}{ET} = \frac{\text{uWUE}}{\sqrt{D_s}}
+\end{equation}
+
+And the derivative:
+
+\begin{equation}
+  \frac{\partial \text{WUE}}{\partial D_s} = -\frac{\text{uWUE}}{2 D_s^{3/2}}
+\end{equation}
+
+We can also take the derivative with respect to $g_a$, $\Delta$, and $LAI$. First LAI:
+
+\begin{equation}
+  \frac{\partial \; ET}{\partial \; \text{LAI}} = \frac{ g_a\; P }{T \left( \Delta + \gamma \right)}  \frac{\gamma c_s \sqrt{D_s} L_v }{\text{LAI$^2$ } R* \; 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{D_s}})}
+\end{equation}
+
+Now w.r.t. $g_a$:
+
+\begin{equation}
+  \frac{\partial \; ET}{\partial \; g_a} = \frac{ P }{T \left( \Delta + \gamma \right)} \left(\frac{c_p D_s}{R_{air}} -  \frac{\gamma c_s \sqrt{D_s} L_v }{\text{LAI } R* \; 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{D_s}})}\right)
+\end{equation}
+
+Now w.r.t. $\Delta$:
+
+\begin{equation}
+\frac{\partial \;  ET}{\partial \Delta} = \frac{\gamma R - \frac{g_a\; P}{T} \left( \frac{ c_p D_{s}}{R_{air}} - \frac{\gamma c_s \sqrt{D_s} L_v }{\text{LAI } R* \; 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{D_s}})} \right) }{ (\Delta + \gamma)^2}
 \end{equation}

### 2017-09-18 doc/eee_symposium/eee_abstract.tex
foudn typo in agu abstract (auto correct on evapotranspiration?) and added eee abstract

\documentclass[a4paper,11pt]{article}
\usepackage{amsmath,amsfonts,amsthm,dsfont,amssymb}
\usepackage[blocks]{authblk}
\newenvironment{keywords}{\noindent\textbf{Keywords:}}{}
\newenvironment{classification}{\noindent\textbf{AMS subject classifications.}}{}
\date{}
\newcommand{\email}[1]{\texttt{\small #1}}

\begin{document}
% % % % %--------------------------------------------------------------------
% % % % %          Title of the Paper and Acknowledgement
% % % % %--------------------------------------------------------------------
  \title{When does vapor pressure deficit drive or reduce evapotranspiration?\footnote{Oral presentation preferred}}
% % % % %--------------------------------------------------------------------
% % % % %         Authors,, Affiliations and email ids
% % % % %--------------------------------------------------------------------

\author[1]{\underline{Adam Massmann}}
\author[1]{Pierre Gentine}
\author[2]{Changjie Lin}

\affil[1]{Department of Earth and Environmental Engineering, Columbia University, New York, NY 10027

  \textit{Contact:} \email{akm2203@columbia.edu}}
\affil[2]{Department of Hydraulic Engineering, Tsinghua University, Beijing, CN}
% % % % %--------------------------------------------------------------------

	\maketitle

\begin{abstract}
  Depending on plant response (e.g. stomatal closure), ecosystem-scale evapotranspiration can either increase or decrease with changes in vapor pressure deficit. This ecosystem response drives evapotranspiration and atmospheric moisture feedbacks. We use data from 75 FluxNet sites within a Penman-Monteith framework to examine when ecosystem evapotranspiration is suppressed or enhanced by increases in vapor pressure deficit. Evapotranspiration response is quantified as a function of soil moisture, atmospheric conditions, and plant functional type. The framework is designed so model and observational uncertainties are represented in the results. This in-situ observation-based analysis aids understanding for how ecosystems will respond and/or contribute to future shifts in atmospheric water demand.

\end{abstract}

% \begin{keywords}
% ginverse, drazin inverse, diffrential equation
% \end{keywords}

% \begin{classification}
% 13C10; 15A09; 15A24; 15B57
% \end{classification}
%  \begin{thebibliography}{100}

%  \bibitem{einstein} 
%  Albert Einstein. 
%  {\em Zur Elektrodynamik bewegter K{\"o}rper}. 
%  Journal of Mathematics, 322(10):89--921, 1905.

%   \bibitem{impj}  The Japan Reader {\em Imperial Japan 1800-1945} 1973:
%   Random House, N.Y.

%   \bibitem{norman} E. H. Norman {\em Japan's emergence as a modern
%   state} 1940: International Secretariat, Institute of Pacific
%   Relations.

%   \bibitem{fo} Bob Tadashi Wakabayashi {\em Anti-Foreignism and Western
%   Learning in Early-Modern Japan} 1986: Harvard University Press.

%   \end{thebibliography}

\end{document}

### 2017-09-18 src/idealized_analysis_et.py

@@ -190,5 +190,6 @@ def site_plot(_df, folder_label=''):
   plt.close('all')
   return
 
-non_mean_pft.apply(site_plot, folder_label='mean', axis=1)
+non_mean_pft.apply(site_plot, folder_label='mean_non-linear', axis=1)
+non_mean.apply(site_plot, folder_label='non-linear', axis=1)
 var_vector.apply(site_plot, axis=1)

### 2017-09-20 doc/paper/vpd_et_paper.tex
added paper start

%% 11/23/2015
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
%
% This file includes commands and instructions
% given in the order necessary to produce a final output that will
% satisfy AGU requirements. 
%
% You may copy this file and give it your
% article name, and enter your text.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PLEASE DO NOT USE YOUR OWN MACROS
% DO NOT USE \newcommand, \renewcommand, or \def, etc.
%
% FOR FIGURES, DO NOT USE \psfrag or \subfigure.
% DO NOT USE \psfrag or \subfigure commands.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Step 1: Set the \documentclass
%
% There are two options for article format:
%
% 1) PLEASE USE THE DRAFT OPTION TO SUBMIT YOUR PAPERS.
% The draft option produces double spaced output.
% 
% 2) numberline will give you line numbers.

%% To submit your paper:
\documentclass[draft,linenumbers]{agujournal}
\draftfalse

%% For final version.
% \documentclass{agujournal}

% Now, type in the journal name: \journalname{<Journal Name>}

% ie, \journalname{Journal of Geophysical Research}
%% Choose from this list of Journals:
%
% JGR-Atmospheres
% JGR-Biogeosciences
% JGR-Earth Surface
% JGR-Oceans
% JGR-Planets
% JGR-Solid Earth
% JGR-Space Physics
% Global Biochemical Cycles
% Geophysical Research Letters
% Paleoceanography
% Radio Science
% Reviews of Geophysics
% Tectonics
% Space Weather
% Water Resource Research
% Geochemistry, Geophysics, Geosystems
% Journal of Advances in Modeling Earth Systems (JAMES)
% Earth's Future
% Earth and Space Science
%
%

\journalname{Water Resource Research}

\begin{document}

%% ------------------------------------------------------------------------ %%
%  Title
% 
% (A title should be specific, informative, and brief. Use
% abbreviations only if they are defined in the abstract. Titles that
% start with general keywords then specific terms are optimized in
% searches)
%
%% ------------------------------------------------------------------------ %%

% Example: \title{This is a test title}

\title{When does vapor pressure deficit drive or reduce evapotranspiration?}

%% ------------------------------------------------------------------------ %%
%
%  AUTHORS AND AFFILIATIONS
%
%% ------------------------------------------------------------------------ %%

% Authors are individuals who have significantly contributed to the
% research and preparation of the article. Group authors are allowed, if
% each author in the group is separately identified in an appendix.)

% List authors by first name or initial followed by last name and
% separated by commas. Use \affil{} to number affiliations, and
% \thanks{} for author notes.  
% Additional author notes should be indicated with \thanks{} (for
% example, for current addresses).

% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
% Author\affil{3,4}\thanks{Also funded by Monsanto.}}

\authors{A. Massmann\affil{1}, P. Gentine\affil{1}, C. Lin\affil{2}}

% \affiliation{1}{First Affiliation}
% \affiliation{2}{Second Affiliation}
% \affiliation{3}{Third Affiliation}
% \affiliation{4}{Fourth Affiliation}

\affiliation{1}{Department of Earth and Environmental Engineering, Columbia University, New York, NY 10027}
\affiliation{2}{Department of Hydraulic Engineering, Tsinghua University, Beijing, CN}

  % (repeat as many times as is necessary)

%% Corresponding Author:
% Corresponding author mailing address and e-mail address:

% (include name and email addresses of the corresponding author.  More
% than one corresponding author is allowed in this LaTeX file and for
% publication; but only one corresponding author is allowed in our
% editorial system.)

% Example: \correspondingauthor{First and Last Name}{email@address.edu}

\correspondingauthor{Adam Massmann}{akm2203@columbia.edu}

%% Keypoints, final entry on title page.

% Example: 
% \begin{keypoints}
% \item	List up to three key points (at least one is required)
% \item	Key Points summarize the main points and conclusions of the article
% \item	Each must be 100 characters or less with no special characters or punctuation 
% \end{keypoints}

%  List up to three key points (at least one is required)
%  Key Points summarize the main points and conclusions of the article
%  Each must be 100 characters or less with no special characters or punctuation

\begin{keypoints}
\item = enter point 1 here = 
\item = enter point 2 here = 
\item = enter point 3 here = 
\end{keypoints}

%% ------------------------------------------------------------------------ %%
%
%  ABSTRACT
%
% A good abstract will begin with a short description of the problem
% being addressed, briefly describe the new data or analyses, then
% briefly states the main conclusion(s) and how they are supported and
% uncertainties. 
%% ------------------------------------------------------------------------ %%

%% \begin{abstract} starts the second page

\begin{abstract}
= enter abstract here =
\end{abstract}

%% ------------------------------------------------------------------------ %%
%
%  TEXT
%
%% ------------------------------------------------------------------------ %%

%%% Suggested section heads:
% \section{Introduction}
% 
% The main text should start with an introduction. Except for short
% manuscripts (such as comments and replies), the text should be divided
% into sections, each with its own heading.

% Headings should be sentence fragments and do not begin with a
% lowercase letter or number. Examples of good headings are:

% \section{Materials and Methods}
% Here is text on Materials and Methods.
%
% \subsection{A descriptive heading about methods}
% More about Methods.
% 
% \section{Data} (Or section title might be a descriptive heading about data)
% 
% \section{Results} (Or section title might be a descriptive heading about the
% results)
% 
% \section{Conclusions}

\section{Introduction}

Changes to vapor pressure deficit (VPD) alter the atmospheric demand for water from the land surface. However, plant stomata have evolved to optimally regulate the exchange of water and carbon between vegetation and the atmosphere \citep{Franks_2017}. Therfore, an increase (decrease) in VPD may not correspond to an increase (decrease) in evapotranspiration (ET) because stomatal closure (opening) can cancel the effects of shifts to atmospheric demand.

Quantifying the plant response to a perturbation to atmospheric VPD increases our understanidng of feedbacks between the land surface and the atmosphere. If plant reponse reduces ET in response to an increase in VPD, the land surface will contribute a positive feedback in reponse to atmospheric drying. Conversely, if plant response increases ET in response to increase in VPD, then the land surface will contribute a negative feedback to atmospheric drying. The sign of these feedbacks drives the evolution of the atmosphere and landsurface at many timescales, from diurnal to interdecadal.

Here we use a Penman-Monteith framework to quantify plant reponse to perturbations to atmospheric demand for water. Section 2 derives the framework, Section 3 describes the data used, Section 4 presents results, and Section 5 discusses conclusions.

\section{Methods}

The Penman-Monteith equation (hereafter PM) estimates ET as a function of atmospheric and land-surface variaibles:

\begin{linenomath*}
  \begin{equation}
      ET = \frac{\Delta R + g_a \rho_a c_p D_{s}}{\Delta + \gamma(1 + \frac{g_a}{g_s})},
  \end{equation}
\end{linenomath*}

 where variable definitions are given in Table 1. \cite{MEDLYN_2011} developed a model for $g_s$ by combining optimal photosynthesis theory with empiracle approaches. The result for leaf-scale stomatal resistance was:

\begin{linenomath*}
  \begin{equation}
  g_{l-s} = g_0 + 1.6 \left(1 + \frac{g_1}{\sqrt{D_{s}}}\right) \frac{A}{c_s}
  \end{equation}
\end{linenomath*}

This can be adapted to an ecosystem-scale stomatal resistance by multiplying by leaf area index (LAI) and converting units to m s$^{-1}$:
\begin{linenomath*}
  \begin{equation}
  g_s = \text{LAI } \frac{R* T}{P} \left( g_0 + 1.6 \left(1 + \frac{g_1}{\sqrt{D_{s}}}\right) \frac{A}{c_s}\right)
  \end{equation}
\end{linenomath*}

While Equation 3 can be used in PM, it will make analytical work with the function intractable because $A$ is a relatively strong function of ET. To remove dependence of ET on $A$ we can use the semi-impiracle results of \cite{Zhou_2015}. \cite{Zhou_2015} showed that:

\begin{linenomath*}
  \begin{equation}
uWUE = \frac{GPP \cdot \sqrt{D}}{ET}
  \end{equation}
\end{linenomath*}
is relatively constant across time and space (within plant functional type). If, following \cite{Lin_2015}, we approximate $g_0$ as $0$, we can use uWUE to remove $A$ from $g_s$ in a way that makes PM analytically tractable:

\begin{linenomath*}
  \begin{equation}
  g_s = \text{LAI } \frac{R* T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{D_{s}}}\right) \frac{uWUE \; ET}{c_s \; \sqrt{D}}
  \end{equation}
\end{linenomath*}

Plugging Equation 5 into Equation 1 and rearranging gives:

\begin{linenomath*}
  \begin{equation}
  ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p D_{s}}{R_{air}} - \frac{\gamma c_s \sqrt{D} }{\text{LAI } R* \; 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{D}})} \right) }{ \Delta + \gamma}
  \end{equation}
\end{linenomath*}

We can then take the derivative with respect to $D$ to determine ecosystem reponse to atmospheric demand perturbations:

\begin{linenomath*}
  \begin{equation}
\frac{\partial \;  ET}{\partial \; D} = \frac{g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{\text{LAI }1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right) \right)
  \end{equation}
\end{linenomath*}
Note that given yearly uWUE from \cite{Zhou_2015}, $g_1$ from \cite{Lin_2015} \citep[as presented in ][]{Franks_2017}, and observations of R, T, P, D$_s$, and wind speed (WS), the only unknown is LAI. With flux tower observations of ET, LAI will then be uniquely determined for each observation through Equation 6:

\begin{linenomath*}
  \begin{equation}
LAI  = - \frac{g_a \gamma c_s \sqrt{D_s} P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p D_{s}\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{D_s}})}
  \end{equation}
\end{linenomath*}

\begin{table}
\caption{Definition of symbols and variables}
\centering
\begin{tabular}{l c c}
\hline
 Variable & Description & Units  \\
\hline
$e_s$  & saturation vapor pressure & Pa  \\ 
$T$  & temperature  & K \\
$\Delta$  & $\frac{\partial e_s}{\partial T}$ & Pa K$^{-1}$ \\
$R$  & net radiation at land surface minus ground heat flux & W m$^{-2}$   \\
  $g_a$  & atmospheric conductance & m s$^{-1}$  \\
  $\rho_a$  & air density & kg m$^{-3}$  \\
  $c_p$  & specific heat capacity of air at constant pressure & J K$^{-1}$ kg$^{-1}$ \\
  $D$  & VPD & Pa  \\
  $\gamma$  & psychrometric constant & Pa K$^{-1}$   \\
  $g_s$  & stomatal conductance & m s$^{-1}$  \\
  $g_{l-s}$  & leaf-scale stomatal conductance & mol m$^{-2}$ s$^{-1}$  \\
  $R*$ & universal gas constant & J mol$^{-1}$ K$^{-1}$ \\
  $LAI$ & leaf area idex & -\\
\hline
\multicolumn{2}{l}{$^{a}$Footnote text here.}
\end{tabular}
\end{table}

%%n
 \begin{linenomath*}
 \begin{equation}
 \end{equation}
 \end{linenomath*}

%% Enter Figures and Tables near as possible to where they are first mentioned:
%
% DO NOT USE \psfrag or \subfigure commands.
%
% Figure captions go below the figure.
% Table titles go above tables;  other caption information
%  should be placed in last line of the table, using
% \multicolumn2l{$^a$ This is a table note.}
%
%----------------
% EXAMPLE FIGURE
%
% \begin{figure}[h]
% \centering
% when using pdflatex, use pdf file:
% \includegraphics[width=20pc]{figsamp.pdf}
%
% when using dvips, use .eps file:
% \includegraphics[width=20pc]{figsamp.eps}
%
% \caption{Short caption}
% \label{figone}
%  \end{figure}
%
% ---------------
% EXAMPLE TABLE
%
% \begin{table}
% \caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
% \centering
% \begin{tabular}{l c}
% \hline
%  Run  & Time (min)  \\
% \hline
%   $l1$  & 260   \\
%   $l2$  & 300   \\
%   $l3$  & 340   \\
%   $h1$  & 270   \\
%   $h2$  & 250   \\
%   $h3$  & 380   \\
%   $r1$  & 370   \\
%   $r2$  & 390   \\
% \hline
% \multicolumn{2}{l}{$^{a}$Footnote text here.}
% \end{tabular}
% \end{table}

%% SIDEWAYS FIGURE and TABLE 
% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
%
% \begin{sidewaysfigure}
% \includegraphics[width=20pc]{figsamp}
% \caption{caption here}
% \label{newfig}
% \end{sidewaysfigure}
% 
%  \begin{sidewaystable}
%  \caption{Caption here}
% \label{tab:signif_gap_clos}
%  \begin{tabular}{ccc}
% one&two&three\\
% four&five&six
%  \end{tabular}
%  \end{sidewaystable}

%% If using numbered lines, please surround equations with \begin{linenomath*}...\end{linenomath*}
%\begin{linenomath*}
%\begin{equation}
%y|{f} \sim g(m, \sigma),
%\end{equation}
%\end{linenomath*}

%%% End of body of article

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional Appendix goes here
%
% The \appendix command resets counters and redefines section heads
%
% After typing \appendix
%
%\section{Here Is Appendix Title}
% will show
% A: Here Is Appendix Title
%
%\appendix
%\section{Here is a sample appendix}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Optional Glossary, Notation or Acronym section goes here:
%
%%%%%%%%%%%%%%  
% Glossary is only allowed in Reviews of Geophysics
%  \begin{glossary}
%  \term{Term}
%   Term Definition here
%  \term{Term}
%   Term Definition here
%  \term{Term}
%   Term Definition here
%  \end{glossary}

%
%%%%%%%%%%%%%%
% Acronyms
%   \begin{acronyms}
%   \acro{Acronym}
%   Definition here
%   \acro{EMOS}
%   Ensemble model output statistics 
%   \acro{ECMWF}
%   Centre for Medium-Range Weather Forecasts
%   \end{acronyms}

%
%%%%%%%%%%%%%%
% Notation 
%   \begin{notation}
%   \notation{$a+b$} Notation Definition here
%   \notation{$e=mc^2$} 
%   Equation in German-born physicist Albert Einstein's theory of special
%  relativity that showed that the increased relativistic mass ($m$) of a
%  body comes from the energy of motion of the body—that is, its kinetic
%  energy ($E$)—divided by the speed of light squared ($c^2$).
%   \end{notation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  ACKNOWLEDGMENTS
%
% The acknowledgments must list:
%
% •	All funding sources related to this work from all authors
%
% •	Any real or perceived financial conflicts of interests for any
%	author
%
% •	Other affiliations for any author that may be perceived as
% 	having a conflict of interest with respect to the results of this
% 	paper.
%
% •	A statement that indicates to the reader where the data
% 	supporting the conclusions can be obtained (for example, in the
% 	references, tables, supporting information, and other databases).
%
% It is also the appropriate place to thank colleagues and other contributors. 
% AGU does not normally allow dedications.

\acknowledgments
 = enter acknowledgments here =

%% ------------------------------------------------------------------------ %%
%% Citations

% Please use ONLY \citet and \citep for reference citations.
% DO NOT use other cite commands (e.g., \cite, \citeyear, \nocite, \citealp, etc.).

%% Example \citet and \citep:
%  ...as shown by \citet{Boug10}, \citet{Buiz07}, \citet{Fra10},
%  \citet{Ghel00}, and \citet{Leit74}.

%  ...as shown by \citep{Boug10}, \citep{Buiz07}, \citep{Fra10},
%  \citep{Ghel00, Leit74}.

%  ...has been shown \citep [e.g.,][]{Boug10,Buiz07,Fra10}.

%%  REFERENCE LIST AND TEXT CITATIONS
%
% Either type in your references using
%
% \begin{thebibliography}{}
% \bibitem[{\textit{Kobayashi et~al.}}(2003)]{R2013} Kobayashi, T.,
% Tran, A.~H., Nishijo, H., Ono, T., and Matsumoto, G.  (2003).
% Contribution of hippocampal place cell activity to learning and
% formation of goal-directed navigation in rats. \textit{Neuroscience}
% 117, 1025--1035.
%
% \bibitem{}
% Text
% \end{thebibliography}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Or, to use BibTeX:
%
% Follow these steps
%
% 1. Type in \bibliography{<name of your .bib file>} 
%    Run LaTeX on your LaTeX file.
%
% 2. Run BiBTeX on your LaTeX file.
%
% 3. Open the new .bbl file containing the reference list and
%   copy all the contents into your LaTeX file here.
%
% 4. Run LaTeX on your new file which will produce the citations.
%
% AGU does not want a .bib or a .bbl file. Please copy in the contents of your .bbl file here.

%% After you run BibTeX, Copy in the contents of the .bbl file here:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Track Changes:
% To add words, \added{<word added>}
% To delete words, \deleted{<word deleted>}
% To replace words, \replace{<word to be replaced>}{<replacement word>}
% To explain why change was made: \explain{<explanation>} This will put
% a comment into the right margin.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% At the end of the document, use \listofchanges, which will list the
% changes and the page and line number where the change was made.

% When final version, \listofchanges will not produce anything,
% \added{<word or words>} word will be printed, \deleted{<word or words} will take away the word,
% \replaced{<delete this word>}{<replace with this word>} will print only the replacement word.
%  In the final version, \explain will not print anything.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%
\listofchanges
%%%

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Supporting Information
%% (Optional) See AGUSuppInfoSamp.tex/pdf for requirements 
%% for Supporting Information.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

More Information and Advice:

%% ------------------------------------------------------------------------ %%
%
%  SECTION HEADS
%
%% ------------------------------------------------------------------------ %%

% Capitalize the first letter of each word (except for
% prepositions, conjunctions, and articles that are
% three or fewer letters).

% AGU follows standard outline style; therefore, there cannot be a section 1 without
% a section 2, or a section 2.3.1 without a section 2.3.2.
% Please make sure your section numbers are balanced.
% ---------------
% Level 1 head
%
% Use the \section{} command to identify level 1 heads;
% type the appropriate head wording between the curly
% brackets, as shown below.
%
%An example:
%\section{Level 1 Head: Introduction}
%
% ---------------
% Level 2 head
%
% Use the \subsection{} command to identify level 2 heads.
%An example:
%\subsection{Level 2 Head}
%
% ---------------
% Level 3 head
%
% Use the \subsubsection{} command to identify level 3 heads
%An example:
%\subsubsection{Level 3 Head}
%
%---------------
% Level 4 head
%
% Use the \subsubsubsection{} command to identify level 3 heads
% An example:
%\subsubsubsection{Level 4 Head} An example.
%
%% ------------------------------------------------------------------------ %%
%
%  IN-TEXT LISTS
%
%% ------------------------------------------------------------------------ %%
%
% Do not use bulleted lists; enumerated lists are okay.
% \begin{enumerate}
% \item
% \item
% \item
% \end{enumerate}
%
%% ------------------------------------------------------------------------ %%
%
%  EQUATIONS
%
%% ------------------------------------------------------------------------ %%

% Single-line equations are centered.
% Equation arrays will appear left-aligned.

Math coded inside display math mode \[ ...\]
 will not be numbered, e.g.,:
 \[ x^2=y^2 + z^2\]

 Math coded inside \begin{equation} and \end{equation} will
 be automatically numbered, e.g.,:
 \begin{equation}
 x^2=y^2 + z^2
 \end{equation}

% To create multiline equations, use the
% \begin{eqnarray} and \end{eqnarray} environment
% as demonstrated below.
\begin{eqnarray}
  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
        && + (y - y_{0}) \sin \Theta  \nonumber \\
  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
        && + (y - y_{0}) \cos \Theta.
\end{eqnarray}

%If you don't want an equation number, use the star form:
%\begin{eqnarray*}...\end{eqnarray*}

% Break each line at a sign of operation
% (+, -, etc.) if possible, with the sign of operation
% on the new line.

% Indent second and subsequent lines to align with
% the first character following the equal sign on the
% first line.

% Use an \hspace{} command to insert horizontal space
% into your equation if necessary. Place an appropriate
% unit of measure between the curly braces, e.g.
% \hspace{1in}; you may have to experiment to achieve
% the correct amount of space.

%% ------------------------------------------------------------------------ %%
%
%  EQUATION NUMBERING: COUNTER
%
%% ------------------------------------------------------------------------ %%

% You may change equation numbering by resetting
% the equation counter or by explicitly numbering
% an equation.

% To explicitly number an equation, type \eqnum{}
% (with the desired number between the brackets)
% after the \begin{equation} or \begin{eqnarray}
% command.  The \eqnum{} command will affect only
% the equation it appears with; LaTeX will number
% any equations appearing later in the manuscript
% according to the equation counter.
%

% If you have a multiline equation that needs only
% one equation number, use a \nonumber command in
% front of the double backslashes (\\) as shown in
% the multiline equation above.

% If you are using line numbers, remember to surround
% equations with \begin{linenomath*}...\end{linenomath*}

%  To add line numbers to lines in equations:
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}

### 2017-09-20 src/lai_inference.py

@@ -11,7 +11,7 @@ import metcalcs as met
 import seaborn as sns
 import resource
 from scipy.stats import spearmanr
-nfrom scipy.optimize import least_squares
+from scipy.optimize import least_squares
 import matplotlib.pyplot as plt
 import matplotlib as mpl
 import codebase.penman_monteith as pm
@@ -64,7 +64,7 @@ def fit_curves(_df, ax):
   result = least_squares(seasonal_lai, x0, bounds=bounds)
   if result['success']:
     seasonal = seasonal_lai(result['x']) + lai
-    df_out = pd.DataFrame(data={'modelled_cycle': seasonal}, index=_df.index)
+    df_out = pd.DataFrame(data={'seasonal_lai': seasonal}, index=_df.index)
     for i, _x in enumerate(result['x']):
       df_out['x%d' % i] = _x
     ax.plot(_df.time.values, seasonal, 'k-')
@@ -89,6 +89,7 @@ def plot_lai(_df, fit_plot=False, suff=''):
 # # below is to generate datta
 # x = df.groupby('site').apply(plot_lai, fit_plot=True)
 # full_df = pd.concat([df, x], axis=1)
+# full_df['residual_lai'] = df['lai'] - df['seasonal_lai']
 # full_df.to_pickle('%s/changjie/full_pandas_seasonal_fit.pkl'\
 #                   % os.environ['DATA'])
 
@@ -97,7 +98,7 @@ df = pd.read_pickle('%s/changjie/full_pandas_seasonal_fit.pkl'\
 
 # bestfit = df.loc[df.x0.argmax()]
 
-# df['residual'] = df.lai-df.modelled_cycle
+# df['residual'] = df.lai-df.seasonal_lai
 
 # def residual_plot(_df):
 #   """plots residuals"""
@@ -118,10 +119,9 @@ df = pd.read_pickle('%s/changjie/full_pandas_seasonal_fit.pkl'\
 
 
 columns = ['lai', 'jd', 'year',\
-           'modelled_cycle', 'x0', 'x1', 'x2', 'site', 'pft']
+           'seasonal_lai', 'x0', 'x1', 'x2', 'site', 'pft']
 subset = df.loc[(df.year == 2003) & (df.site == 'US-MMS'), columns]
 
-subset['residual'] = subset.lai-subset.modelled_cycle
 # well-fit subset to test with edward approahc
 # plot_lai(subset, suff='_subset')
 
[…]

### 2017-09-20 doc/paper/vpd_et_paper.tex
got map working, now I just nee to do a better projection

@@
 
-\section{Introduction}
+\section{Introduction}\explain{This section needs to be fleshed out, and I definately need to add more citations}
 
@@
 
+This ``pseudo-LAI'' is some part ``true'' LAI (a measure of leaf area), and some part model and observational error, including error involving our assumption of constant uWUE.
+By calculating a unique LAI for each observation we will progate any model and observational uncertainty forward into our expression for $\frac{\partial \; ET}{\partial D}$.
+
 \begin{table}
@@
   $LAI$ & leaf area idex & -\\
+\hline
+\multicolumn{2}{l}{$^{a}$Footnote text here.}
+\end{tabular}
+\end{table}
+
+\section{Data}
+
+We use data from FLUXNET2015.
+Because $g_1$ coefficients \citep{Lin_2015} and uWUE were only both available for five plant functional types (PFTs - see Table 2),  only 56 of the 77 sites were used.
+Figure 1 presents each site and its plant functional type.
+
+\begin{table}
+\caption{Plant functional types, their abbreviation, Medlyn coefficient \citep[from ][]{Lin_2015}, and uWUE \citep[from ][]{Zhou_2015}.
+Note that units are converted such that the quantities fit into Equations 1-8}
+\centering
+\begin{tabular}{l c c c}
+  \hline
+  Abbreviation & PFT & $g_1$ (Pa$^{0.5}$) & uWUE ($\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET])  \\
+  \hline
+  CRO & cropland & 183.1 & 3.80 \\
+  CSH & closed shrub & 148.6 & 2.18 \\
+  DBF & deciduous broadleaf forest & 140.7 & 3.12 \\
+  ENF & evergreen needleleaf forest & 74.3 & 3.30 \\
+  GRA & grassland (C3) & 166.0 & 2.68 \\
 \hline
@@
 \acknowledgments
- = enter acknowledgments here =
+This work used eddy covariance data acquired and shared by the FLUXNET community, including these networks: AmeriFlux, AfriFlux, AsiaFlux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont, ChinaFlux, Fluxnet-Canada, GreenGrass, ICOS, KoFlux, LBA, NECC, OzFlux-TERN, TCOS-Siberia, and USCCC.
+The ERA-Interim reanalysis data are provided by ECMWF and processed by LSCE.
+The FLUXNET eddy covariance data processing and harmonization was carried out by the European Fluxes Database Cluster, AmeriFlux Management Project, and Fluxdata project of FLUXNET, with the support of CDIAC and ICOS Ecosystem Thematic Center, and the OzFlux, ChinaFlux and AsiaFlux offices.
 

### 2017-09-22 doc/paper/vpd_et_paper.tex
I broke the plots

@@
 \begin{table}
-\caption{Plant functional types, their abbreviation, Medlyn coefficient \citep[from ][]{Lin_2015}, and uWUE \citep[from ][]{Zhou_2015}.
-Note that units are converted such that the quantities fit into Equations 1-8}
+\caption{Plant functional types, their abbreviation, Medlyn coefficient \citep[from ][]{Lin_2015}, and uWfUE \citep[from ][]{Zhou_2015}.
+Note that units are converted such that the quantities fit into Equations 1-8 with the variables in Table 1.}
 \centering
@@
 \end{table}
+
+We restrict our analysis to the daytime (sensible heat > 5 W m$^{-1}$ and shortwave radiation > 50 W m$^{-2}$) when there is no precipitation and the plants are growing (GPP > 10\% of the 95th pecentile).
+Also, because some sites use half hourly data but some use hourly, we aggregate all data to hourly averages.
+Only times with good quality control flags are used.
+
+\section{Results}
 

### 2017-09-22 src/idealized_analysis_et.py

@@ -20,7 +20,6 @@ mpl.rcParams.update(mpl.rcParamsDefault)
 
 importlib.reload(util)
 
-
 df = pd.read_pickle('%s/changjie/full_pandas_seasonal_fit.pkl'\
                     % os.environ['DATA'])
 df['g_a'] = 1./df['r_a']

### 2017-09-22 src/make_map_paper_fig1.py

@@ -36,14 +36,24 @@ def make_map(_df):
   """makes a map given a _df with Lat, Lon, and Cover"""
   pfts = _df.Cover_type.drop_duplicates()
   fig = plt.figure()
-  ax = fig.add_subplot(111)
-  m = Basemap(projection='npaeqd',boundinglat=10,lon_0=270,resolution='l')
-  m.drawcoastlines()
-  for pft in pfts:
-    print(pft)
-    _ds = _df.loc[(_df.Cover_type==pft), ['Latitude', 'Longitude']]
-    x, y = m(_ds.Longitude.values, _ds.Latitude.values)
-    sc = ax.scatter(x, y, s=50, label=pft)
+  fig.set_figheight(fig.get_figheight()*3)
+  axs = [fig.add_subplot(3, 1, i+1) for i in range(3)]
+  xlims = [(-180., 180.), (-140., -50.), (-20., 40.)]
+  ylims = [(-90., 90.), (15., 60.), (30., 60.)]
+  for ax, xlim, ylim in zip(axs, xlims, ylims):
+    print(ylim)
+    print(xlim)
+    m = Basemap(projection='merc', llcrnrlat=ylim[0], urcrnrlat=ylim[1],\
+            llcrnrlon=xlim[0], urcrnrlon=xlim[1], lat_ts=np.mean(ylims),\
+                resolution='c', ax=ax)#res was l
+    m.drawcoastlines()
+    for pft in pfts:
+      print(pft)
+      _ds = _df.loc[(_df.Cover_type==pft), ['Latitude', 'Longitude']]
+      x, y = m(_ds.Longitude.values, _ds.Latitude.values)
+      sc = ax.scatter(x, y, s=8, label=pft)
+    # m.drawparallels(np.arange(-90.,120.,30.))
+    # m.drawmeridians(np.arange(0.,420.,60.))
   plt.legend(loc='best')
   util.test_savefig('../doc/paper/figs/fig01.png')
   return

### 2017-09-25 src/make_map_paper_fig1.py
added vpd paper

@@ -7,11 +7,12 @@ import os
 import pandas as pd
 import numpy as np
 import metcalcs as met
+import matplotlib as mpl
+mpl.use('Pdf')
 import seaborn as sns
 import resource
 from scipy.stats import spearmanr
 import matplotlib.pyplot as plt
-import matplotlib as mpl
 import codebase.penman_monteith as pm
 import util
 import importlib
@@ -55,7 +56,7 @@ def make_map(_df):
     # m.drawparallels(np.arange(-90.,120.,30.))
     # m.drawmeridians(np.arange(0.,420.,60.))
   plt.legend(loc='best')
-  util.test_savefig('../doc/paper/figs/fig01.png')
+  util.test_savefig('../doc/paper/fig01.pdf')
   return
 
 make_map(meta)

### 2017-09-25 doc/paper/vpd_et_paper.tex
added vpd paper
back to writing

@@
     \frac{\partial \;  ET}{\partial \; D} = \frac{g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{\text{LAI }1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right) \right)
+    \label{d_et}
   \end{equation}
@@
     LAI  = - \frac{g_a \gamma c_s \sqrt{D_s} P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p D_{s}\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{D_s}})}
+    \label{lai}
   \end{equation}
@@
 Because $g_1$ coefficients \citep{Lin_2015} and uWUE were only both available for five plant functional types (PFTs - see Table 2),  only 56 of the 77 sites were used.
-Figure 1 presents each site and its plant functional type.
+Figure \ref{map_fig} presents each site and its plant functional type.
 
@@
 
+\begin{figure}[h]
+\centering
+\includegraphics[width=20pc]{./fig01.pdf}
+\caption{Plant functional type and location of sites used in analysis.}
+\label{map_fig}
+ \end{figure}
+
 We restrict our analysis to the daytime (sensible heat > 5 W m$^{-1}$ and shortwave radiation > 50 W m$^{-2}$) when there is no precipitation and the plants are growing (GPP > 10\% of the 95th pecentile).
@@
 \section{Results}
+
+By construction, the variability in the LAI term (Equation \ref{lai}) contains all model and observational uncertities.
+LAI also has physical meaning corresponding to ``true'' leaf area, and we expect that it would be approximately $O(1)$.
+We can have some confidence in our framework, including the assumption of constant uWUE, if calculated LAIs are generally $O(1)$.
+Figure \ref{lai_fig} presents the histogram of calculated LAIs with outliers (lowest and highest 5\% percent) and unphysical values (LAI $<$ 0.) removed.
 

### 2017-09-25 src/make_paper_plots.py
lint collected paper plots and seperated plot functions from script (now in codebase)

@@ -2,26 +2,20 @@
 """
 This script makes a map of PFT for the paper, fig1.
 """
-import glob
 import os
+import importlib
 import pandas as pd
 import numpy as np
-import metcalcs as met
 import matplotlib as mpl
-mpl.use('Pdf')
-import seaborn as sns
-import resource
-from scipy.stats import spearmanr
 import matplotlib.pyplot as plt
-import codebase.penman_monteith as pm
-import util
-import importlib
 from mpl_toolkits.basemap import Basemap
-from matplotlib.patches import Polygon
+import codebase.plot_tools as plot_tools
+import util
+
+mpl.use('Pdf')
 mpl.rcParams.update(mpl.rcParamsDefault)
-
 importlib.reload(util)
-
+importlib.reload(plot_tools)
 # grab sites actually used in analysis
 df = pd.read_pickle('%s/changjie/full_pandas_seasonal_fit.pkl'\
                     % os.environ['DATA']).loc[:, 'site'].drop_duplicates()
@@ -50,15 +44,29 @@ def make_map(_df):
     m.drawcoastlines()
     for pft in pfts:
       print(pft)
-      _ds = _df.loc[(_df.Cover_type==pft), ['Latitude', 'Longitude']]
+      _ds = _df.loc[(_df.Cover_type == pft), ['Latitude', 'Longitude']]
       x, y = m(_ds.Longitude.values, _ds.Latitude.values)
-      sc = ax.scatter(x, y, s=16, label=pft)
+      ax.scatter(x, y, s=16, label=pft)
     # m.drawparallels(np.arange(-90.,120.,30.))
     # m.drawmeridians(np.arange(0.,420.,60.))
   plt.legend(loc='best')
   util.test_savefig('../doc/paper/fig01.pdf')
   return
 
+#make a map
 make_map(meta)
 
+#for final npaper should make this .pdf, and probably
+df = pd.read_pickle('%s/changjie/full_pandas_lai_clean.pkl'\
+                    % os.environ['DATA'])
+df['g_a'] = 1./df['r_a']
+df['d_et_leaf'] = df['scaling']*df['vpd_leaf']
+df['d_et_atm'] = df['scaling']*df['vpd_atm']
 
+meta = {}
+meta['var'] = 'lai'
+meta['folder'] = ''
+meta['folder_label'] = ''
+plot_tools.histogram(df, meta)
+os.system('cp %s/climate_et/histogram/full_lai.png '\
+          '../doc/paper/fig01.png' % (os.environ['PLOTS']))

### 2017-09-25 doc/paper/vpd_et_paper.tex
lint collected paper plots and seperated plot functions from script (now in codebase)
seeking easy solution to overlaps

@@
 \includegraphics[width=20pc]{./fig01.pdf}
-\caption{Plant functional type and location of sites used in analysis.}
+\caption{Plant functional type and location of sites used in analysis. Note should probably just split this into 4 continents (US, Europe, Africa, Australia).}
 \label{map_fig}
@@
 Figure \ref{lai_fig} presents the histogram of calculated LAIs with outliers (lowest and highest 5\% percent) and unphysical values (LAI $<$ 0.) removed.
+All remaining LAI values are $O(1)$ which provides confidence in model framework.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=20pc]{./fig02.png}
+\caption{Histogram of LAI values calculated for each site and time according to Equation \ref{lai}.
+The lowest and highest 5\% are removed as outliers, as well as any values below 0.
+The curve is normalized such that its area is 1.}
+\label{lai_fig}
+ \end{figure}
 

### 2017-09-25 src/make_paper_plots.py

@@ -1,6 +1,6 @@
 #! ~/edward/bin/python
 """
-This script makes a map of PFT for the paper, fig1.
+This script makes all figs for the paper
 """
 import os
 import importlib
@@ -31,29 +31,31 @@ def make_map(_df):
   """makes a map given a _df with Lat, Lon, and Cover"""
   pfts = _df.Cover_type.drop_duplicates()
   fig = plt.figure()
-  fig.set_figheight(fig.get_figheight()*3)
-  axs = [fig.add_subplot(3, 1, i+1) for i in range(3)]
-  xlims = [(-180., 180.), (-140., -50.), (-20., 40.)]
-  ylims = [(-50., 70.), (15., 60.), (30., 60.)]
-  for ax, xlim, ylim in zip(axs, xlims, ylims):
+  fig.set_figheight(fig.get_figheight()*4)
+  axs = [fig.add_subplot(4, 1, i+1) for i in range(4)]
+  xlims = [(-140., -50.), (-20., 40.), (10., 40), (110., 155.)]
+  ylims = [(15., 60.), (30., 70.), (-20., -10.), (-40., -10.)]
+  sizes = [24, 12, 40, 40]
+  for ax, xlim, ylim, size in zip(axs, xlims, ylims, sizes):
     print(ylim)
     print(xlim)
     m = Basemap(projection='merc', llcrnrlat=ylim[0], urcrnrlat=ylim[1],\
             llcrnrlon=xlim[0], urcrnrlon=xlim[1], lat_ts=np.mean(ylims),\
-                resolution='c', ax=ax)#res was l
+                resolution='l', ax=ax)#res was l
     m.drawcoastlines()
     for pft in pfts:
       print(pft)
       _ds = _df.loc[(_df.Cover_type == pft), ['Latitude', 'Longitude']]
       x, y = m(_ds.Longitude.values, _ds.Latitude.values)
-      ax.scatter(x, y, s=16, label=pft)
+      ax.scatter(x, y, s=size, label=pft)
     # m.drawparallels(np.arange(-90.,120.,30.))
     # m.drawmeridians(np.arange(0.,420.,60.))
   plt.legend(loc='best')
+  plt.tight_layout()
   util.test_savefig('../doc/paper/fig01.pdf')
   return
 
-#make a map
+#make a map fig 1
 make_map(meta)
 
 #for final npaper should make this .pdf, and probably
@@ -69,4 +71,4 @@ meta['folder'] = ''
[…]

### 2017-09-26 doc/paper/vpd_et_paper.tex
added vpd dependence plot to paper

@@
 \end{figure}
+
+An additional concern is that the LAI term may in fact be some function of $D$, in which case the dependence would need to be accounted for when taking the derivative.
+Figure \ref{lai_vpd_fig} plots the jount distrivution of LAI and VPD, and shows that LAI is very weakly a function of VPD.
+Given this weak dependence, we argue that Equation \ref{d_et} is a valid approxiamtion for ET response to $D$.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=20pc]{./fig03.png}
+\caption{The joint distribution of $D$ and LAI. LAI has only a weak dependence on $D$}
+\label{lai_vpd_fig}
+\end{figure}
+
+Before diving into calcualted values of $\frac{\partial \; ET}{\partial \; D}$, it is useful to consider the functional form of Equation \ref{d_et}.
+There are three terms: a scaling term for the full expression we will call Term 1 ($\frac{g_a \; P}{T(\Delta + \gamma)}$), a relatively constant offset we will call Term 2 ($\frac{c_p}{R_{air}}$), and a variable term we will call Term 3 ($- \frac{\gamma c_s }{\text{LAI }1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right)$).
+All variables are constant, so the relative magnitude between Term 2 and Term 3 will determine the sign of the derivative, while Term 1 will scale the expression larger or smaller.
 

### 2017-09-26 src/codebase/plot_tools.py

@@ -36,8 +36,6 @@ def unpack_df(filename):
   site = ''.join(filename.split('/')[-1].split('.')[:-1])
   return atmos, canopy, data, site
 
-
-
 def make_ax_plot(_ax, var, _df, meta):
   """makes an axis plot"""
   if meta['vmax'] is None:

### 2017-09-26 src/plot_et_vpd_derivative.py

@@ -29,29 +29,29 @@ df['g_a'] = 1./df['r_a']
 df['d_et_leaf'] = df['scaling']*df['vpd_leaf']
 df['d_et_atm'] = df['scaling']*df['vpd_atm']
 
-meta = {}
-meta['vmax'] = None
-meta['var'] = ''
-meta['label'] = ''
-meta['x_axis'] = 'rh'
-meta['log'] = ''
-meta['size'] = 8
+# meta = {}
+# meta['vmax'] = None
+# meta['var'] = ''
+# meta['label'] = ''
+# meta['x_axis'] = 'rh'
+# meta['log'] = ''
+# meta['size'] = 8
 
-meta = {}
-meta['var'] = 'lai'
-meta['folder'] = 'site_et'
-meta['folder_label'] = 'site'
-df.groupby('site').apply(plot_tools.histogram, meta)
-meta['folder'] = 'pft_et'
-meta['folder_label'] = 'pft'
-df.groupby('site').apply(plot_tools.histogram, meta)
-meta['folder'] = ''
-meta['folder_label'] = ''
-plot_tools.histogram(df, meta)
-meta['var'] = 'lai_gpp'
-meta['folder'] = 'gpp'
-df.groupby('site').apply(plot_tools.histogram, meta)
-plt.close('all')
+# meta = {}
+# meta['var'] = 'lai'
+# meta['folder'] = 'site_et'
+# meta['folder_label'] = 'site'
+# df.groupby('site').apply(plot_tools.histogram, meta)
+# meta['folder'] = 'pft_et'
+# meta['folder_label'] = 'pft'
+# df.groupby('site').apply(plot_tools.histogram, meta)
+# meta['folder'] = ''
+# meta['folder_label'] = ''
+# plot_tools.histogram(df, meta)
+# meta['var'] = 'lai_gpp'
+# meta['folder'] = 'gpp'
+# df.groupby('site').apply(plot_tools.histogram, meta)
+# plt.close('all')
 
 # # meta['folder_label'] = 'site'
 # for meta['var'], meta['vmax'] in zip(['d_et', 'd_gpp'],\
@@ -73,7 +73,7 @@ df['jd'] = time.dayofyear
 #   meta['plot_type'] = '' #'simple'
 #   meta['x_var'] = 'swc'
 #   meta['y_var'] = name
-#   scatter_wrapper(df, meta)
+#   plot_tools.scatter_wrapper(df, meta)
 #   meta['group'] = 'site'
[…]

### 2017-09-27 doc/paper/vpd_et_paper.tex
okay going to change plotting around

@@
 An additional concern is that the LAI term may in fact be some function of $D$, in which case the dependence would need to be accounted for when taking the derivative.
-Figure \ref{lai_vpd_fig} plots the jount distrivution of LAI and VPD, and shows that LAI is very weakly a function of VPD.
+Figure \ref{lai_vpd_fig} plots the joint distrivution of LAI and VPD, and shows that LAI is very weakly a function of VPD.
 Given this weak dependence, we argue that Equation \ref{d_et} is a valid approxiamtion for ET response to $D$.
@@
 
-%%n
+In Term 1, $\frac{P}{T} \propto \rho$, so this should vary little relative to $g_a$ and $\Delta$.
+$\gamma$ should also be relatively constant, So the scaling term, Term 1, should be primarily a function of $g_a$ and temperature (though the function $\Delta$).
+Figure \ref{scale_vary} shows how the magnitude of the scaling varies with $T$ and $g_a$.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=20pc]{./fig04.pdf}
+\caption{Variability of Term 1 for each PFT. Each colored line correspondes to a different aerodynamic resistance.}
+\label{lai_vpd_fig}
+\end{figure}
+
+%%
  \begin{linenomath*}

### 2017-09-27 doc/paper/vpd_et_paper.tex
going to make a coffe and think about fig 5

@@
 Before diving into calcualted values of $\frac{\partial \; ET}{\partial \; D}$, it is useful to consider the functional form of Equation \ref{d_et}.
-There are three terms: a scaling term for the full expression we will call Term 1 ($\frac{g_a \; P}{T(\Delta + \gamma)}$), a relatively constant offset we will call Term 2 ($\frac{c_p}{R_{air}}$), and a variable term we will call Term 3 ($- \frac{\gamma c_s }{\text{LAI }1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right)$).
+There are three terms: a scaling term for the full expression we will call Term 1 ($\frac{g_a \; P}{T(\Delta + \gamma)}$), a relatively constant offset we will call Term 2 ($\frac{c_p}{R_{air}}$), and a variable term we will call Term 3 ($\frac{\gamma c_s }{\text{LAI }1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right)$).
 All variables are constant, so the relative magnitude between Term 2 and Term 3 will determine the sign of the derivative, while Term 1 will scale the expression larger or smaller.
@@
 In Term 1, $\frac{P}{T} \propto \rho$, so this should vary little relative to $g_a$ and $\Delta$.
-$\gamma$ should also be relatively constant, So the scaling term, Term 1, should be primarily a function of $g_a$ and temperature (though the function $\Delta$).
-Figure \ref{scale_vary} shows how the magnitude of the scaling varies with $T$ and $g_a$.
+$\gamma$ should also be relatively constant, So the scaling term, Term 1, should be primarily a function of $g_a$ and temperature (through the function $\Delta$).
+While temperature range may vary for PFT, the fucntional form of $\Delta$ will be the same.
+$g_a$ will vary strongly with PFT due to the importance of surface roughness.
+However, the coeffiecent of variability for $g_a$ is relatively constant across PFT.
+So, the influence of $g_a$ on the relative (to the mean) variability of Term 1 is approximately similiar across PFT.
+
+Figure \ref{scale_vary}A shows Term 1 normalized by mean $g_a$ (calcualted for each plant functional type), and confirms that much of the relative variability of Term 1 is contained in the $g_a$ term's relative variability.
+Additionally, the impact of $T$ on the relative variability increases with increasing $g_a$.
+
+While the relative variability of Term 1 is similiar across PFT, the absolute value of Term 1 varies strongly across PFT.
+Figure \ref{scale_vary}B shows Term 1 evaluated with the mean $g_a$ for each PFT, and at the range of observed temperatures for each PFT.
+As expected, for the tree PFTs, the Term 1 is much larger and the temperature dependence is much stronger.
+Additionally, because std$(g_a) \propto \overline{g_a}$, the spread of Term 1 due to $g_a$ variability will be larger, although this is not shown for simplicity.
 
@@
 \includegraphics[width=20pc]{./fig04.pdf}
-\caption{Variability of Term 1 for each PFT. Each colored line correspondes to a different aerodynamic resistance.}
-\label{lai_vpd_fig}
+\caption{Primary sources of variability for Term 1.
+A) Term 1 normalized by mean $g_a$ for each PFT.
+B) Term 1 evaluated at mean $g_a$ for each PFT.
+Temperature range is 5-95th percentile for each PFT.
+Additionally, stars denote the location of the 25th, 50th, and 75th percentiles.}
+\label{scale_vary}
 \end{figure}

### 2017-09-27 doc/paper/vpd_et_paper.tex
forgot to add calc tools, and some final changes for the evening on vpd_paper

@@
 
-Before diving into calcualted values of $\frac{\partial \; ET}{\partial \; D}$, it is useful to consider the functional form of Equation \ref{d_et}.
+Before diving into calculated values of $\frac{\partial \; ET}{\partial \; D}$, it is useful to consider the functional form of Equation \ref{d_et}.
 There are three terms: a scaling term for the full expression we will call Term 1 ($\frac{g_a \; P}{T(\Delta + \gamma)}$), a relatively constant offset we will call Term 2 ($\frac{c_p}{R_{air}}$), and a variable term we will call Term 3 ($\frac{\gamma c_s }{\text{LAI }1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right)$).
@@
 Figure \ref{scale_vary}B shows Term 1 evaluated with the mean $g_a$ for each PFT, and at the range of observed temperatures for each PFT.
-As expected, for the tree PFTs, the Term 1 is much larger and the temperature dependence is much stronger.
+As expected, for the tree PFTs the Term 1 is much larger and the temperature dependence is much stronger.
 Additionally, because std$(g_a) \propto \overline{g_a}$, the spread of Term 1 due to $g_a$ variability will be larger, although this is not shown for simplicity.
+To summarize, the variability of Term 1 will look like Figure \ref{scale_vary}A for each PFT, but the scale of the x and y-axis will increase or decrease according to Figure \ref{scale_vary}B.
 

### 2017-09-29 doc/paper/vpd_et_paper.tex
made fig5 two panels so show converse with lai on x axis, debugging now
about to swithc up legend terms

@@
 There are three terms: a scaling term for the full expression we will call Term 1 ($\frac{g_a \; P}{T(\Delta + \gamma)}$), a relatively constant offset we will call Term 2 ($\frac{c_p}{R_{air}}$), and a variable term we will call Term 3 ($\frac{\gamma c_s }{\text{LAI }1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right)$).
-All variables are constant, so the relative magnitude between Term 2 and Term 3 will determine the sign of the derivative, while Term 1 will scale the expression larger or smaller.
+All variables are positive, so the relative magnitude between Term 2 and Term 3 will determine the sign of the derivative, while Term 1 will scale the expression larger or smaller.
 
@@
 Additionally, the impact of $T$ on the relative variability increases with increasing $g_a$.
+% plot every PFT and show they collapse onto the same vurve?
 
@@
 Figure \ref{scale_vary}B shows Term 1 evaluated with the mean $g_a$ for each PFT, and at the range of observed temperatures for each PFT.
-As expected, for the tree PFTs the Term 1 is much larger and the temperature dependence is much stronger.
+As expected, for the tree PFTs (DBF, ENF) Term 1 is much larger and the temperature dependence is much stronger.
+Systematic differences in observed temperatures also cause difference in the average magnitude of Term 1.
+For example, ENF experiences on average colder temperatures and is thus more likely to have a larger scaling term.
 Additionally, because std$(g_a) \propto \overline{g_a}$, the spread of Term 1 due to $g_a$ variability will be larger, although this is not shown for simplicity.
@@
 \caption{Primary sources of variability for Term 1.
-A) Term 1 normalized by mean $g_a$ for each PFT.
-B) Term 1 evaluated at mean $g_a$ for each PFT.
+A) Variability within each PFT: Term 1 normalized by mean $g_a$ for each PFT.
+B) Variability between each PFT: Term 1 evaluated at mean $g_a$ for each PFT.
 Temperature range is 5-95th percentile for each PFT.
@@
 \label{scale_vary}
+\end{figure}
+
+Term 2 minus Term 3 determines the sign and magnitude that the scaling Term 1 is multiplied.
+If we assume that $c_s$ variability is relatively less than LAI and $D$ variability, then variability within PFT will be soley detmined by LAI and $D$.
+Figure \ref{term3_within} shows how (Term 2 - Term 3) varies within PFT.
+Each row is a different PFT, and the first column depects $D$ on the x-axis, evaluated at 5th, 25th, 50th, 75th, and 95th percentiles.
+The second column has LAI on the x-axis, evaluated at 5th, 25th, 50th, and 95th percentiles of VPD.
+Within each PFT LAI variability is generally of greater importance to (Term 2 - Term 3)'s variaibility than VPD variaibility.
+Low VPD and LAI increase the probability of a negative ET response to VPD, and the dependence of the response increases as VPD and LAI get smaller.
+
+Generally, functional forms within each PFT are similiar between each PFT.
+However, as uWUE increases Term 3 decreases, causing (Term 2 - Term 3) to shift towards positive values.
+Additionally, the smaller g1 the more important VPD is for the magnitude of Term 3.
+Fo example, ENF has the smallest g1 (74.31), and the VPD-induced variaibility is of similiar magnitude to the LAI-induced variability.
+% introduce quantitiy uWUE*\overline{LAI}?
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=20pc]{./fig05.pdf}
+\caption{}
+\label{term3_within}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=20pc]{./fig06.pdf}
+\caption{}
+\label{term3_between}
 \end{figure}

### 2017-09-29 doc/paper/vpd_et_paper.tex
thinking about completely distroying penman_monteith to clean out all fo the old calc methods that were useless

@@
 
+Figure \ref{term3_between} more clearly illustrates differences between PFT.
+Term 3 can be broken into a scaling (Figure \ref{term3_between}b), and a $D$ term (Figure \ref{term3_between}d).
+PFTs with smaller uWUE will have larger magnitude scaling, and PFTs with a smaller g1 will have a greater VPD dependence and a larger magnitude $D$ term.
+When combined into Term 2 - Term 3 (Figure \ref{term3_between}a,c) the consequences are that CRO is the most likely to have a postive $D$ response, and CSH is most likely to have a negative $D$ response.
+DBF, ENF, and GRA have similiar liklihoods of positive and negative responses, depnding on LAI and VPD.
+ENF has by far the strongest VPD dependence of response, as observen in Figure \ref{term3_within}.
+
 \begin{figure}[h]
@@
 \end{figure}
+
+Now time for full plot as a function of t and rh, then break up into single scaling (g, T), and pft-specific plots of (T, RH), or try 3d maps?
 

### 2017-09-29 src/make_paper_plots.py

@@ -30,7 +30,7 @@ site_list = pd.read_csv('%s/changjie/fluxnet_algorithm/'\
 site_list.index = site_list.Site
 site_list = site_list.loc[df.values]
 
-# below is the label 
+# below is the label
 paren_string = r'(Term 2 - Term 3) $\left(\frac{ c_p}{R_{air}} '\
                r'- \frac{\gamma c_s }{LAI \; 1.6 \; R\; uWUE  }'\
                r'\left( \frac{2 g_1 + \sqrt{D}}'\
@@ -346,26 +346,28 @@ def pft_leaf(_df, axs):
   lai = np.linspace(_df.lai.quantile(q=0.05), _df.lai.quantile(q=0.95))
   vpd = _df.vpd.mean()
   axs[0].plot(lai, term_2(_df, lai, vpd),\
-              label=r"%s, $\overline{VPD}$*uWUE=%4.0f Pa, uWUE=%4.2f, g1=%4.1f"\
-              % (_df.pft.iloc[0], vpd*_df.uwue_norm.iloc[0],\
+              label=r"$\overline{VPD}$=%4.0f Pa, uWUE=%4.2f, g1=%4.1f"\
+              % (vpd,\
                  _df.uwue_norm.iloc[0], _df.g1.iloc[0]))
   axs[1].plot(lai, first_half(_df, lai),\
               label='%s, uWUE = %4.2f'\
               % (_df.pft.iloc[0], _df.uwue_norm.iloc[0]))
   ptiles = np.array([_df.lai.quantile(q=_p/100.)\
                      for _p in [25., 50., 75.]])
+  # axs[0].plot(ptiles, term_2(_df, ptiles, vpd), 'k*')
   axs[1].plot(ptiles, first_half(_df, ptiles), 'k*')
   # now second half
   vpd = np.linspace(_df.vpd.quantile(q=0.05), _df.vpd.quantile(q=0.95))
   lai = _df.lai.mean()
   axs[2].plot(vpd, term_2(_df, lai, vpd),\
-              label=r"%s, $\overline{LAI}$*uWUE*g1=%4.2f, g1=%4.1f"\
-              % (_df.pft.iloc[0],\
-                 lai*_df.uwue_norm.iloc[0]*_df.g1.iloc[0], _df.g1.iloc[0]))
[…]

### 2017-10-03 doc/paper/vpd_et_paper.tex
just added last idea to work in next session, (thanks to beer for the inspiration
think some changse in lsat commit didn't make it because I hadn't saved

--- 2017-09-29 doc/paper/vpd_et_paper.tex
@@
 
-Term 2 minus Term 3 determines the sign and magnitude that the scaling Term 1 is multiplied.
+%%%%% replace below with idea:
+
+``just say that vpd effect is approximately linear (except at low VPD), and then say so we are most concerned about when d et/dvpd is zero, as this is the minima.
+then you can just look at this treshold as a function of LAI, uWUE, and g1.''
+
+Term 2 minus Term 3 determines the sign and magnitude that the scaling Term 1 is multiplied by.
 If we assume that $c_s$ variability is relatively less than LAI and $D$ variability, then variability within PFT will be soley detmined by LAI and $D$.
-Figure \ref{term3_within} shows how (Term 2 - Term 3) varies within PFT.
-Each row is a different PFT, and the first column depects $D$ on the x-axis, evaluated at 5th, 25th, 50th, 75th, and 95th percentiles.
-The second column has LAI on the x-axis, evaluated at 5th, 25th, 50th, and 95th percentiles of VPD.
-Within each PFT LAI variability is generally of greater importance to (Term 2 - Term 3)'s variaibility than VPD variaibility.
-Low VPD and LAI increase the probability of a negative ET response to VPD, and the dependence of the response increases as VPD and LAI get smaller.
-
-Generally, functional forms within each PFT are similiar between each PFT.
-However, as uWUE increases Term 3 decreases, causing (Term 2 - Term 3) to shift towards positive values.
-Additionally, the smaller g1 the more important VPD is for the magnitude of Term 3.
-Fo example, ENF has the smallest g1 (74.31), and the VPD-induced variaibility is of similiar magnitude to the LAI-induced variability.
-% introduce quantitiy uWUE*\overline{LAI}?
+Figure \ref{term3} shows how (Term 2 - Term 3) varies with $D$ and LAI, as a function of PFT.
+In Figure \ref{term3}a lower uWUE and LAI shift the disrtribution of (Term 2 - Term 3) towards negative values.
+Additionally, the smaller $g1$, the greater the relative $D$ dependence of of (Term 2 - Term 3).
+This is observed most strongly for the ENF PFT, which has the smallest $g1$ (74.31).
 
@@
 \caption{}
-\label{term3_within}
+\label{term3}
 \end{figure}
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=20pc]{./fig06.pdf}
-\caption{}
-\label{term3_between}
-\end{figure}
+Figure \ref{term3}b shows the location of the minima of ET, as a function of LAI and $D$.
+For any LAI or VPD less (more) than these curves, Term 2 - Term 3 will be negative (positive).
+It is clear that the portion of VPD observations below these curves will be a strong function of $LAI$.
+However, we can see some general trends.
+For CSH, $\frac{\partial \; ET}{\partial \; D}$ should be negative for the vast majority of observed LAI and VPD.
+The split appears to be more even among ENF, GRA, and DBF, and we might expect a greater frequency of positive $\frac{\partial \; ET}{\partial \; D}$ for CRO.
+
+I would say add a table of det dd bar, also evaluated at mean values of each pft, also normalized by std of vpd, and finally
+
+Now time for full plot as a function of t and rh, then break up into single scaling (g, T), and pft-specific plots of (T, RH), or try 3d maps?
 

### 2017-10-03 doc/paper/vpd_et_paper.tex
added table to latex document and formatted, now just need to get back to writing, shoudl really try to get this done tomorrow

@@
   $R*$ & universal gas constant & J mol$^{-1}$ K$^{-1}$ \\
-  $LAI$ & leaf area idex & -\\
+  $LAI$ & leaf area index & -\\
+  $c_s$ & CO$_2$ concentration & $\mu$ mol CO$_2$ mol$^{-1}$ air\\
 \hline
@@
 \label{map_fig}
- \end{figure}
+ \end{figure}p
 
@@
 Now time for full plot as a function of t and rh, then break up into single scaling (g, T), and pft-specific plots of (T, RH), or try 3d maps?
+
+\begin{table}
+\caption{Statistics of $\frac{\partial \; ET}{\partial \; D}$ as a function of PFT.}
+\centering
+\begin{tabular}{l c c c c c}
+  \hline
+PFT & $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ & $\frac{\partial \; ET}{\partial \; D}\left(\overline{T, \ldots , D}\right)$ & $\frac{\partial \; ET}{\partial \; D}\left(\overline{T, \ldots , D}\right)*\text{std}(D)$ & $\frac{\frac{\partial \; ET}{\partial \; D}\left(\overline{T, \ldots , D}\right)*\text{std}(D)}{ \frac{\partial \; ET}{\partial \; R}\left(\overline{T, \ldots , D}\right)*\text{std}(R)}$ & fraction $\frac{\partial \; ET}{\partial \; VPD} < 0.$ \\
+  \hline
+CRO & 0.000853 & 0.026241 & 18.523659 & 0.203022 & 0.473311\\
+CSH & -0.108234 & -0.091526 & 50.861613 & 0.439379 & 0.931660\\
+DBF & -0.012727 & 0.013794 & 19.734435 & 0.164241 & 0.461674\\
+ENF & -0.034087 & 0.000706 & 16.611852 & 0.148548 & 0.534425\\
+GRA & -0.019637 & -0.000921 & 16.798083 & 0.173552 & 0.631735\\
+\hline
+\multicolumn{2}{l}{$^{a}$Footnote text here.}
+
+\end{tabular}
+\end{table}
 

### 2017-10-03 src/codebase/calc_tools.py

@@ -81,9 +81,6 @@ def d_et_d_delta(atmos, canopy):
               *(1. + canopy['g1']/np.sqrt(atmos['vpd'])))))\
               /(atmos['delta'] + atmos['gamma'])**2
 
-def d_et_d_r_net(_df):
-  """derivantve w.r.t net radiation"""
-  return _df['delta']/(_df['delta']+_df['gamma'])
 
 def scaling(atmos):
   """calcualtes the scaling term"""
@@ -133,3 +130,46 @@ def calc_derivative(atmos, canopy, data):
   #retun lai to that used for ET
   canopy['lai'] = lai_true
   return atmos, canopy, data
+
+#### below fucntions added from jacobian/varaibility plot
+def get_uwue(_df):
+  """lazy way to get uwue"""
+  return _df.loc[:, ['uwue']].iloc[0]
+
+def d_et(_df):
+  """calcs the full d ET/d Ds, confirmed correct vs df"""
+  return _df['g_a']*_df['p_a']/\
+    ((_df['t_a']+ 273.15)*(_df['gamma']+_df['delta']))*\
+    (pm.CP/_df['r_moist']-_df['gamma']*_df['c_a']*pm.LV/\
+     (_df['lai']*1.6*pm.R_STAR*_df['uwue'])*\
+     (2.*_df['g1']+np.sqrt(_df['vpd']))\
+     /(2.*(_df['g1']+np.sqrt(_df['vpd']))**2))
+
+def d_et_d_r_net(_df):
+  """derivantve w.r.t net radiation"""
+  return _df['delta']/(_df['delta']+_df['gamma'])
+
+def d_et_d_lai(_df):
+  """calc derivative d et/ dlai"""
+  return _df['g_a']*_df['p_a']*_df['gamma']*_df['c_a']\
+    *np.sqrt(_df['vpd'])*pm.LV\
+    /(_df['t_a_k']*(_df['delta']+_df['gamma'])*_df['lai']**2\
+      *pm.R_STAR*1.6*_df['uwue']*(1. + _df['g1']/np.sqrt(_df['vpd'])))
+
+def d_et_d_g_a(_df):
+  """calc derivative w.r.t. g_a"""
+  return _df['p_a']/(_df['t_a_k']*(_df['delta'] + _df['gamma']))\
+    *(pm.CP*_df['vpd']/_df['r_moist']\
[…]

### 2017-10-03 src/make_paper_plots.py

@@ -79,6 +79,14 @@ df['g_a'] = 1./df['r_a']
 df['d_et_leaf'] = df['scaling']*df['vpd_leaf']
 df['d_et_atm'] = df['scaling']*df['vpd_atm']
 df['uwue_norm'] = df.uwue/pm.LV
+df['r_net'] = df['r_n'] - df['g_flux']
+jacobians = {'vpd' : calc.d_et,\
+             'lai' : calc.d_et_d_lai,\
+             'seasonal_lai' : calc.d_et_d_lai,\
+             'residual_lai' : calc.d_et_d_lai,\
+             'g_a' : calc.d_et_d_g_a,\
+             'delta' : calc.d_et_d_delta,\
+             'r_net' : calc.d_et_d_r_net}
 
 meta = {}
 meta['var'] = 'lai'
@@ -471,26 +479,30 @@ plt.savefig('../doc/paper/fig05.pdf')
 
 
 ###### table 5 ####
+importlib.reload(calc)
 def frequency(_df):
   """return fraction fos amples d_et < 0"""
   return _df.d_et[_df.d_et < 0.].count() / _df.d_et.count()
 
-importlib.reload(calc)
-df['net_rad'] = df.r_n - df.g_flux
-pft_grouped = df.groupby('pft')
-mean_df = pft_grouped.mean()
-std_df = pft_grouped.std()
-counts = pft_grouped.apply(frequency)
 
-mean_df['d_et_bar'] = calc.scaling(mean_df)\
-                      *(pm.CP/mean_df.r_moist +\
-                        calc.leaf_vpd(mean_df, mean_df, mean_df.lai))
-mean_df['d_rn_bar'] = calc.d_et_d_r_net(mean_df)
+pft = df.groupby('site').apply(get_pft)
+mean = df.groupby('site').mean()
+mean['pft'] = pft
+std = df.groupby('site').std()
+#jacobian = site_analysis(mean)
+mean['d_et_bar'] = calc.d_et(mean)
+mean['d_et_bar_std'] = np.absolute(calc.d_et(mean))*std['vpd']
+mean['d_rn_bar'] = np.absolute(calc.d_et_d_r_net(mean))*std['r_net']
+mean['d_et_bar_norm_rn'] = mean['d_et_bar_std']\
+                           /mean['d_rn_bar']
+
+mean_df = mean.groupby('pft').mean()
+counts = df.groupby('pft').apply(frequency)
[…]

### 2017-10-04 doc/paper/vpd_et_paper.tex
I took another look at the scatters of LAI vs Delta, VPD, g_a. I think an independent approximation might be okay on for LAI,e ven though at site level there are some correlations (for a few of the sites).
think it's looking realistic to be done with this today

@@
 As expected, for the tree PFTs (DBF, ENF) Term 1 is much larger and the temperature dependence is much stronger.
-Systematic differences in observed temperatures also cause difference in the average magnitude of Term 1.
+Systematic differences in observed temperatures also cause differences in the average magnitude of Term 1.
 For example, ENF experiences on average colder temperatures and is thus more likely to have a larger scaling term.
@@
 
-I would say add a table of det dd bar, also evaluated at mean values of each pft, also normalized by std of vpd, and finally
+Table 3 confirms these expectations for PFT behaviour of $\frac{\partial \; ET}{\partial \; D}$.
+For all PFTs except for CRO, average $\frac{\partial \; ET}{\partial \; D}$ is less than zero.
+However, $\frac{\partial \; ET}{\partial \; D}$ evaluated at the average of all variables (e.g.
+LAI, $T$, $c_s$, $D$) is only negative for CSH and GRA.
+And, DBF in addition to CRO experiences $\frac{\partial \; ET}{\partial \; D}$ < 0 less than half the time.
+These observations highlight the effect of the nonlinear function in Figure \ref{term3}: $\frac{\partial \; ET}{\partial \; D}$ has a much steeper slope when the function is negative, and is thus more likely to be large magnitude.
 
-Now time for full plot as a function of t and rh, then break up into single scaling (g, T), and pft-specific plots of (T, RH), or try 3d maps?
+The units of $\frac{\partial \; ET}{\partial \; D}$ make it difficult to interpret if $D$ is even a meaningful contributor to ET's variability.
+To understand $D$'s contribution better, we use a linear approximation and present $\frac{\partial \; ET}{\partial \; D}$ multiplied by $D$'s standard deviation.
+The range of $D$'s contribution to ET's variability ranges between 16 - 20 W m$^{-2}$ for all PFTs except for CSH, which is about 51 W m$^{-2}$.
+Another meaningful comparison is to $\frac{\partial \; ET}{\partial \; R} * std(R)$, as net radiation generally the driver of ET (cite joe berry here).
+For all PFTs except for CSH $D$ contributes between 14.5 - 20.5 \% of $R$'s contribution to variability.
+For CSH the portion is much larger, about 44 \%.
+However it is important to note that a linear approximation about a mean base state is probably not a very good approximation across the range of variabilility, so these values are just estimates of $D$'s contribution to ET's variability.
 
-\begin{table}
+So far, idealized plots and statistics have illuminated the form of $\frac{\partial \; ET}{\partial \; D}$ and how it varies with PFT.
+Large mean LAI and uWUE shifts CRO and DBF towards positive $\frac{\partial \; ET}{\partial \; D}$.
+However, the strongly nonlinear funtion of $\frac{\partial \; ET}{\partial \; D}$ at $\frac{\partial \; ET}{\partial \; D} < 0$ pushes $\overline{\frac{\partial \; ET}{\partial \; D}}$ negative for DBF (it does not do this for CRO because of CRO's high $g1$).
+ENF's low $g1$ value increases the dependence of $\frac{\partial \; ET}{\partial \; D}$ on $D$, and makes the function more strongly nonlinear.
+This has the side effect of pushing $\overline{\frac{\partial \; ET}{\partial \; D}}$ negative futher than other PFTs for a given fraction $\frac{\partial \; ET}{\partial \; D} < 0$ and magntiude $\frac{\partial \; ET}{\partial \; D}(\overline{T,\ldots,D})$.
+GRA shows the opposite behavior; a relatively high $g1$  makes the function more linear, decreasing the magnitude of $\overline{\frac{\partial \; ET}{\partial \; D}}$ for a given fraction $\frac{\partial \; ET}{\partial \; D} < 0$ and magntiude $\frac{\partial \; ET}{\partial \; D}(\overline{T,\ldots,D})$ (although $g_a$ and Term 1 also probably have a role in this).
+Finally, low $uWUE$ of CSH pushes to toward by far the lowest values $\frac{\partial \; ET}{\partial \; D}$ (Figure \ref{term3}).
+Variability in $D$ accounts for the largest about of $ET$ variability for CSH.
+For the other PFTs, $D$ contributes less to $ET$ variaibiliyt, but still represents about 15-20 \% of $R$'s contributions to ET variability.
+
+\Begin{table}
 \caption{Statistics of $\frac{\partial \; ET}{\partial \; D}$ as a function of PFT.}
@@
 \end{table}
+
+\subsection{Direct observations of $\frac{\partial \; ET}{\partial \; D}$}
+
+Now that we have an intuitive understanding of $\frac{\partial \; ET}{\partial \; D}$'s behavior, we are equipped to intepret fully realistic plots of $\frac{\partial \; ET}{\partial \; D}$ for each PFT.
+Figure \ref{fig} presents calculated $\frac{\partial \; ET}{\partial \; D}$ where, unless otherwise noted, all variables in Equation \ref{d_et} are allowed to vary.
 

### 2017-10-04 doc/paper/vpd_et_paper.tex
think I have final plot 7 ready
okay heading home to cook. first draft done now just need to shell check, proofread, format nicely. then I can send to Pierre

@@
 \documentclass[draft,linenumbers]{agujournal}
-\draftfalse
+% \draftfalse
+\drafttrue
 
@@
 Section 2 derives the framework, Section 3 describes the data used, Section 4 presents results, and Section 5 discusses conclusions.
+The goal of this paper is to use reasonable approximations as a tool to increase intuition for plant response to atmospheric drying.
+This intuition will aid interpretation of observations and full complexity climate models.
 
@@
 Because $g_1$ coefficients \citep{Lin_2015} and uWUE were only both available for five plant functional types (PFTs - see Table 2),  only 56 of the 77 sites were used.
-Figure \ref{map_fig} presents each site and its plant functional type.
+Figure \ref{map_fig} \explain{map needs to be fleshed out} presents each site and its plant functional type.
 
@@
 \section{Results}
+\label{results}
 
@@
 
+% below show histogram of site specific spearmannr? think would be useful 
 \begin{figure}[h]
@@
 \includegraphics[width=20pc]{./fig05.pdf}
-\caption{}
+\caption{Sources of variability for Term 2 - Term 3.
+Top: Term 2 - Term 3 as a function of VPD, with LAI held fixed at PFT averages.
+The observed range of VPD for each PFT is also shown below the x-axis.
+Line extent correspondes to 5th and 95th precentiles, while stars denote the location of the 25th, 50th, and 75th percentiles.
+
+Bottom: The location of the minima of ET, as a function of VPD and LAI. Lines and stars denote the distribution of VPD and LAI next to axis, following the same percentiles as above.}
 \label{term3}
@@
 These observations highlight the effect of the nonlinear function in Figure \ref{term3}: $\frac{\partial \; ET}{\partial \; D}$ has a much steeper slope when the function is negative, and is thus more likely to be large magnitude.
-
+n
 The units of $\frac{\partial \; ET}{\partial \; D}$ make it difficult to interpret if $D$ is even a meaningful contributor to ET's variability.
@@
 
-\Begin{table}
+\begin{table}
 \caption{Statistics of $\frac{\partial \; ET}{\partial \; D}$ as a function of PFT.}
@@
 
-\subsection{Direct observations of $\frac{\partial \; ET}{\partial \; D}$}
+\subsection{Full observations of $\frac{\partial \; ET}{\partial \; D}$}
 
 Now that we have an intuitive understanding of $\frac{\partial \; ET}{\partial \; D}$'s behavior, we are equipped to intepret fully realistic plots of $\frac{\partial \; ET}{\partial \; D}$ for each PFT.
-Figure \ref{fig} presents calculated $\frac{\partial \; ET}{\partial \; D}$ where, unless otherwise noted, all variables in Equation \ref{d_et} are allowed to vary.
+Figure \ref{real} presents calculated $\frac{\partial \; ET}{\partial \; D}$ where, unless otherwise noted, all variables in Equation \ref{d_et} are allowed to vary.
+Each column is a different quantity related to $\frac{\partial \; ET}{\partial \; D}$, and each row is a different PFT.
+
+The full observations generally confirm expectations from Section \ref{results}.
+CRO has the most postitive values of $\frac{\partial \; ET}{\partial \; D}$, $\frac{\partial \; ET}{\partial \; D}$ is almost always negative for CSH, and reponse depends more with the environmental conditions for the other PFTs (especially ENF).
+Through the columns of Figure \ref{real} we can see the impact of LAI and $g_a$ on the variability of $\frac{\partial \; ET}{\partial \; D}$.
+$g_a$'s (columns 1 and 3) scaling alters the magnitude considerably.
+$LAI$ (columns 2 and 3) variability adds a lot of additional noise to the signal of $\frac{\partial \; ET}{\partial \; D}$, which is slightly undesirable given LAI's role in representing model and observational uncertainty.
+However, the general story with the noise appears to match the cleaner signal when $LAI$ is help constant and $D_{ETmin}$ is clearly visible \explain{I really need to make these pltos better - way too much overlapping of points that hurts the story}.
+One exception is possibly with GRA, for which uncertainty is high and causes the full complexity plots (Columns 1 and 2) to not match well with LAI held fixed (Columns 3 and 4).
+
+For ENF and GRA $D_{ETmin}$ does not appear to be only a function of LAI.
+It turns out that the site to site variaibility in $\gamma$ causes $D_{ETmin}$ to vary, which is not discussed in the pregious section.
+The impact is observable in both ENF and GRA, but especially for ENF which has a larger $\frac{\partial^2 \; ET}{\partial^2 \; D}$ than the other PFTs.
+
+In general the full complexity plots of $\frac{\partial \; ET}{\partial \; D}$ match our expectations, even with the large sensitivty to LAI measures of uncertainty observed in Figure \ref{term3}.
+Our LAI-based method of uncertainty propogation blurrs the idealized expectations the most for GRA, and also has a considerable effect for CRO.
+We therefor have the most confidence in our conclusion based on Equation \ref{d_et} for PFTS CSH, DBF, and ENF, as the full complexity plots with uncertainty included closely match the story when LAI is held fixed.
+**see somewhat preferred alternate figure \ref{real2} \explain{I think I like this more as thinking in terms of T and RH is easier.
+However, I used the other plot because Fig 5 does not discuss things in terms of temperature, as this would make things more complicated (adding another diminsion).
+I could just include both versions of figure 6 though (using figure 6a as a bridge to figure 6b)}.
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{./fig06.png}
+\caption{Scatter plots of $\frac{\partial \; ET}{\partial \; D}$.
+Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; D}$, as labelled: Column 1 - $\frac{\partial \; ET}{\partial \; D}$; Column 2 - $\frac{\partial \; ET}{\partial \; D}$ normalized by $g_a$; Column 3 - $\frac{\partial \; ET}{\partial \; D}$ with LAI held fixed at PFT average; and Column 4 - $\frac{\partial \; ET}{\partial \; D}$ normalized by $g_a$ and with LAI held fixed.
+For reference, lines corresponding to RH = 20\% and RH = 90 \% are drawn.
+Please note differences in the colorbar scale.}
+\label{real}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{./fig06b.png}
+\caption{****alternate Fig 06****  Scatter plots of $\frac{\partial \; ET}{\partial \; D}$.
+Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; D}$, as labelled.
+If I end up using this, I could also draw on the curve of $D_{ETmin}$ with $\overline{LAI}$.
+}
+\label{real2}
+\end{figure}
+
+\section{Conclusions} \explain{also need to flesh this section out}
+
+The idealized representation of ET used here is successful in developing intuition for how ET responds to changes in $D$.
+This intuition will aid the community in interpreting observations and output from sophisticated full complexity climate models.
+
+The idealized framework leads to the following general conclusions:
+\begin{itemize}
+  \item Aerodynamic resistance plays an important role of scaling $\frac{\partial \; ET}{\partial \; D}$. This is a leading order effect for observing higher magnitude reponses in DBF and ENF.
+\item In general, CSH has the most negative (i.e.
+ET reduced) response to increases in $D$ (atmospheric drying).
+So CSH plants will almost always try and conserve water, effectivey reducing ET with dry atmospheric perturbation.
+\item CRO has the most postive response (i.e. ET increased) in response to increases in $D$. This is consistent with CROs that may be evolved or bred to thrive in non-water-limited environments.
+\item The reponse is more a function of the environment for DBF, ENF, and GRA.
+Because as VPD increases the response is more likely to be postive, if RH is fixed then the reponse will be more likely to be positive at warmer T, or if T is fixed the response is more likely to be positive with decreasing RH.
+\item ENF has the strongest dependence on environmental conditions due to its small $g1$.
+\item Model and observational uncertainty is highest for GRA and CRO, so conlusions about those PFTs should be tempered.
+\item However, inclusion of uncertainty doesn't alter conclusions about DBF, ENF, and CSH.
+\end{itemize}
+
+The intuition developed using this framework can be used to understand how the land surface will respond and contribute to changes in the environment.
 
 %%
- \begin{linenomath*}
- \begin{equation}
- \end{equation}
- \end{linenomath*}
-
 %% Enter Figures and Tables near as possible to where they are first mentioned:

### 2017-10-04 doc/paper/vpd_et_paper.tex
finished with spell check
minor formatting changes

@@
 
-\section{Introduction}\explain{This section needs to be fleshed out, and I definately need to add more citations}
+\section{Introduction}\explain{This section needs to be fleshed out, and I definitely need to add more citations}
 
@@
 However, plant stomata have evolved to optimally regulate the exchange of water and carbon between vegetation and the atmosphere \citep{Franks_2017}.
-Therfore, an increase (decrease) in VPD may not correspond to an increase (decrease) in evapotranspiration (ET) because stomatal closure (opening) can cancel the effects of shifts to atmospheric demand.
+Therefore, an increase (decrease) in VPD may not correspond to an increase (decrease) in evapotranspiration (ET) because stomatal closure (opening) can cancel the effects of shifts to atmospheric demand.
 
-Quantifying the plant response to a perturbation to atmospheric VPD increases our understanidng of feedbacks between the land surface and the atmosphere.
-If plant reponse reduces ET in response to an increase in VPD, the land surface will contribute a positive feedback in reponse to atmospheric drying.
+Quantifying the plant response to a perturbation to atmospheric VPD increases our understanding of feedbacks between the land surface and the atmosphere.
+If plant response reduces ET in response to an increase in VPD, the land surface will contribute a positive feedback in response to atmospheric drying.
 Conversely, if plant response increases ET in response to increase in VPD, then the land surface will contribute a negative feedback to atmospheric drying.
-The sign of these feedbacks drives the evolution of the atmosphere and landsurface at many timescales, from diurnal to interdecadal.
+The sign of these feedbacks drives the evolution of the atmosphere and land-surface at many timescales, from diurnal to interdecadal.
 
-Here we use a Penman-Monteith framework to quantify plant reponse to perturbations to atmospheric demand for water.
+Here we use a Penman-Monteith framework to quantify plant response to perturbations to atmospheric demand for water.
 Section 2 derives the framework, Section 3 describes the data used, Section 4 presents results, and Section 5 discusses conclusions.
@@
 
-The Penman-Monteith equation (hereafter PM) estimates ET as a function of atmospheric and land-surface variaibles:
+The Penman-Monteith equation (hereafter PM) estimates ET as a function of atmospheric and land-surface variables:
 
@@
  where variable definitions are given in Table 1.
-\cite{MEDLYN_2011} developed a model for $g_s$ by combining optimal photosynthesis theory with empiracle approaches.
+\cite{MEDLYN_2011} developed a model for $g_s$ by combining optimal photosynthesis theory with empirical approaches.
 The result for leaf-scale stomatal resistance was:
@@
 While Equation 3 can be used in PM, it will make analytical work with the function intractable because $A$ is a relatively strong function of ET.
-To remove dependence of ET on $A$ we can use the semi-impiracle results of \cite{Zhou_2015}.
+To remove dependence of ET on $A$ we can use the semi-empirical results of \cite{Zhou_2015}.
 \cite{Zhou_2015} showed that:
@@
 
-We can then take the derivative with respect to $D$ to determine ecosystem reponse to atmospheric demand perturbations:
+We can then take the derivative with respect to $D$ to determine ecosystem response to atmospheric demand perturbations:
 
@@
 This ``pseudo-LAI'' is some part ``true'' LAI (a measure of leaf area), and some part model and observational error, including error involving our assumption of constant uWUE.
-By calculating a unique LAI for each observation we will progate any model and observational uncertainty forward into our expression for $\frac{\partial \; ET}{\partial D}$.
+By calculating a unique LAI for each observation we will propagate any model and observational uncertainty forward into our expression for $\frac{\partial \; ET}{\partial D}$.
 
@@
 
-We restrict our analysis to the daytime (sensible heat > 5 W m$^{-1}$ and shortwave radiation > 50 W m$^{-2}$) when there is no precipitation and the plants are growing (GPP > 10\% of the 95th pecentile).
+We restrict our analysis to the daytime (sensible heat > 5 W m$^{-1}$ and shortwave radiation > 50 W m$^{-2}$) when there is no precipitation and the plants are growing (GPP > 10\% of the 95th percentile).
 Also, because some sites use half hourly data but some use hourly, we aggregate all data to hourly averages.
@@
 
-By construction, the variability in the LAI term (Equation \ref{lai}) contains all model and observational uncertities.
+By construction, the variability in the LAI term (Equation \ref{lai}) contains all model and observational uncertainties.
 LAI also has physical meaning corresponding to ``true'' leaf area, and we expect that it would be approximately $O(1)$.
 We can have some confidence in our framework, including the assumption of constant uWUE, if calculated LAIs are generally $O(1)$.
-Figure \ref{lai_fig} presents the histogram of calculated LAIs with outliers (lowest and highest 5\% percent) and unphysical values (LAI $<$ 0.) removed.
+Figure \ref{lai_fig} presents the histogram of calculated LAIs with outliers (lowest and highest 5\% percent) and nonphysical values (LAI $<$ 0.) removed.
 All remaining LAI values are $O(1)$ which provides confidence in model framework.
@@
 An additional concern is that the LAI term may in fact be some function of $D$, in which case the dependence would need to be accounted for when taking the derivative.
-Figure \ref{lai_vpd_fig} plots the joint distrivution of LAI and VPD, and shows that LAI is very weakly a function of VPD.
-Given this weak dependence, we argue that Equation \ref{d_et} is a valid approxiamtion for ET response to $D$.
+Figure \ref{lai_vpd_fig} plots the joint distribution of LAI and VPD, and shows that LAI is very weakly a function of VPD.
+Given this weak dependence, we argue that Equation \ref{d_et} is a valid approximation for ET response to $D$.
 
@@
 $\gamma$ should also be relatively constant, So the scaling term, Term 1, should be primarily a function of $g_a$ and temperature (through the function $\Delta$).
-While temperature range may vary for PFT, the fucntional form of $\Delta$ will be the same.
+While temperature range may vary for PFT, the functional form of $\Delta$ will be the same.
 $g_a$ will vary strongly with PFT due to the importance of surface roughness.
-However, the coeffiecent of variability for $g_a$ is relatively constant across PFT.
-So, the influence of $g_a$ on the relative (to the mean) variability of Term 1 is approximately similiar across PFT.
+However, the coefficient of variability for $g_a$ is relatively constant across PFT.
+So, the influence of $g_a$ on the relative (to the mean) variability of Term 1 is approximately similar across PFT.
 
-Figure \ref{scale_vary}A shows Term 1 normalized by mean $g_a$ (calcualted for each plant functional type), and confirms that much of the relative variability of Term 1 is contained in the $g_a$ term's relative variability.
+Figure \ref{scale_vary}A shows Term 1 normalized by mean $g_a$ (calculated for each plant functional type), and confirms that much of the relative variability of Term 1 is contained in the $g_a$ term's relative variability.
 Additionally, the impact of $T$ on the relative variability increases with increasing $g_a$.
@@
 
-While the relative variability of Term 1 is similiar across PFT, the absolute value of Term 1 varies strongly across PFT.
+While the relative variability of Term 1 is similar across PFT, the absolute value of Term 1 varies strongly across PFT.
 Figure \ref{scale_vary}B shows Term 1 evaluated with the mean $g_a$ for each PFT, and at the range of observed temperatures for each PFT.
@@
 Term 2 minus Term 3 determines the sign and magnitude that the scaling Term 1 is multiplied by.
-If we assume that $c_s$ variability is relatively less than LAI and $D$ variability, then variability within PFT will be soley detmined by LAI and $D$.
+If we assume that $c_s$ variability is relatively less than LAI and $D$ variability, then variability within PFT will be solely determined by LAI and $D$.
 Figure \ref{term3} shows how (Term 2 - Term 3) varies with $D$ and LAI, as a function of PFT.
-In Figure \ref{term3}a lower uWUE and LAI shift the disrtribution of (Term 2 - Term 3) towards negative values.
+In Figure \ref{term3}a lower uWUE and LAI shift the distribution of (Term 2 - Term 3) towards negative values.
 Additionally, the smaller $g1$, the greater the relative $D$ dependence of of (Term 2 - Term 3).
@@
 The observed range of VPD for each PFT is also shown below the x-axis.
-Line extent correspondes to 5th and 95th precentiles, while stars denote the location of the 25th, 50th, and 75th percentiles.
+Line extent corresponds to 5th and 95th percentiles, while stars denote the location of the 25th, 50th, and 75th percentiles.
 
@@
 
-Table 3 confirms these expectations for PFT behaviour of $\frac{\partial \; ET}{\partial \; D}$.
+Table 3 confirms these expectations for PFT behavior of $\frac{\partial \; ET}{\partial \; D}$.
 For all PFTs except for CRO, average $\frac{\partial \; ET}{\partial \; D}$ is less than zero.
@@
 For CSH the portion is much larger, about 44 \%.
-However it is important to note that a linear approximation about a mean base state is probably not a very good approximation across the range of variabilility, so these values are just estimates of $D$'s contribution to ET's variability.
+However it is important to note that a linear approximation about a mean base state is probably not a very good approximation across the range of variability, so these values are just estimates of $D$'s contribution to ET's variability.
 
@@
 Large mean LAI and uWUE shifts CRO and DBF towards positive $\frac{\partial \; ET}{\partial \; D}$.
-However, the strongly nonlinear funtion of $\frac{\partial \; ET}{\partial \; D}$ at $\frac{\partial \; ET}{\partial \; D} < 0$ pushes $\overline{\frac{\partial \; ET}{\partial \; D}}$ negative for DBF (it does not do this for CRO because of CRO's high $g1$).
+However, the strongly nonlinear function of $\frac{\partial \; ET}{\partial \; D}$ at $\frac{\partial \; ET}{\partial \; D} < 0$ pushes $\overline{\frac{\partial \; ET}{\partial \; D}}$ negative for DBF (it does not do this for CRO because of CRO's high $g1$).
 ENF's low $g1$ value increases the dependence of $\frac{\partial \; ET}{\partial \; D}$ on $D$, and makes the function more strongly nonlinear.
-This has the side effect of pushing $\overline{\frac{\partial \; ET}{\partial \; D}}$ negative futher than other PFTs for a given fraction $\frac{\partial \; ET}{\partial \; D} < 0$ and magntiude $\frac{\partial \; ET}{\partial \; D}(\overline{T,\ldots,D})$.
-GRA shows the opposite behavior; a relatively high $g1$  makes the function more linear, decreasing the magnitude of $\overline{\frac{\partial \; ET}{\partial \; D}}$ for a given fraction $\frac{\partial \; ET}{\partial \; D} < 0$ and magntiude $\frac{\partial \; ET}{\partial \; D}(\overline{T,\ldots,D})$ (although $g_a$ and Term 1 also probably have a role in this).
+This has the side effect of pushing $\overline{\frac{\partial \; ET}{\partial \; D}}$ negative further than other PFTs for a given fraction $\frac{\partial \; ET}{\partial \; D} < 0$ and magnitude $\frac{\partial \; ET}{\partial \; D}(\overline{T,\ldots,D})$.
+GRA shows the opposite behavior; a relatively high $g1$  makes the function more linear, decreasing the magnitude of $\overline{\frac{\partial \; ET}{\partial \; D}}$ for a given fraction $\frac{\partial \; ET}{\partial \; D} < 0$ and magnitude $\frac{\partial \; ET}{\partial \; D}(\overline{T,\ldots,D})$ (although $g_a$ and Term 1 also probably have a role in this).
 Finally, low $uWUE$ of CSH pushes to toward by far the lowest values $\frac{\partial \; ET}{\partial \; D}$ (Figure \ref{term3}).
 Variability in $D$ accounts for the largest about of $ET$ variability for CSH.
-For the other PFTs, $D$ contributes less to $ET$ variaibiliyt, but still represents about 15-20 \% of $R$'s contributions to ET variability.
+For the other PFTs, $D$ contributes less to $ET$ variability, but still represents about 15-20 \% of $R$'s contributions to ET variability.
 
@@
 
-Now that we have an intuitive understanding of $\frac{\partial \; ET}{\partial \; D}$'s behavior, we are equipped to intepret fully realistic plots of $\frac{\partial \; ET}{\partial \; D}$ for each PFT.
+Now that we have an intuitive understanding of $\frac{\partial \; ET}{\partial \; D}$'s behavior, we are equipped to interpret fully realistic plots of $\frac{\partial \; ET}{\partial \; D}$ for each PFT.
 Figure \ref{real} presents calculated $\frac{\partial \; ET}{\partial \; D}$ where, unless otherwise noted, all variables in Equation \ref{d_et} are allowed to vary.
@@
 The full observations generally confirm expectations from Section \ref{results}.
-CRO has the most postitive values of $\frac{\partial \; ET}{\partial \; D}$, $\frac{\partial \; ET}{\partial \; D}$ is almost always negative for CSH, and reponse depends more with the environmental conditions for the other PFTs (especially ENF).
+CRO has the most positive values of $\frac{\partial \; ET}{\partial \; D}$, $\frac{\partial \; ET}{\partial \; D}$ is almost always negative for CSH, and response depends more with the environmental conditions for the other PFTs (especially ENF).
 Through the columns of Figure \ref{real} we can see the impact of LAI and $g_a$ on the variability of $\frac{\partial \; ET}{\partial \; D}$.
@@
 $LAI$ (columns 2 and 3) variability adds a lot of additional noise to the signal of $\frac{\partial \; ET}{\partial \; D}$, which is slightly undesirable given LAI's role in representing model and observational uncertainty.
-However, the general story with the noise appears to match the cleaner signal when $LAI$ is help constant and $D_{ETmin}$ is clearly visible \explain{I really need to make these pltos better - way too much overlapping of points that hurts the story}.
+However, the general story with the noise appears to match the cleaner signal when $LAI$ is help constant and $D_{ETmin}$ is clearly visible \explain{I really need to make these plots better - way too much overlapping of points that hurts the story}.
 One exception is possibly with GRA, for which uncertainty is high and causes the full complexity plots (Columns 1 and 2) to not match well with LAI held fixed (Columns 3 and 4).
@@
 For ENF and GRA $D_{ETmin}$ does not appear to be only a function of LAI.
-It turns out that the site to site variaibility in $\gamma$ causes $D_{ETmin}$ to vary, which is not discussed in the pregious section.
+It turns out that the site to site variability in $\gamma$ causes $D_{ETmin}$ to vary, which is not discussed in the previous section.
 The impact is observable in both ENF and GRA, but especially for ENF which has a larger $\frac{\partial^2 \; ET}{\partial^2 \; D}$ than the other PFTs.
 
-In general the full complexity plots of $\frac{\partial \; ET}{\partial \; D}$ match our expectations, even with the large sensitivty to LAI measures of uncertainty observed in Figure \ref{term3}.
-Our LAI-based method of uncertainty propogation blurrs the idealized expectations the most for GRA, and also has a considerable effect for CRO.
+In general the full complexity plots of $\frac{\partial \; ET}{\partial \; D}$ match our expectations, even with the large sensitivity to LAI measures of uncertainty observed in Figure \ref{term3}.
+Our LAI-based method of uncertainty propagation blurs the idealized expectations the most for GRA, and also has a considerable effect for CRO.
 We therefor have the most confidence in our conclusion based on Equation \ref{d_et} for PFTS CSH, DBF, and ENF, as the full complexity plots with uncertainty included closely match the story when LAI is held fixed.
 **see somewhat preferred alternate figure \ref{real2} \explain{I think I like this more as thinking in terms of T and RH is easier.
-However, I used the other plot because Fig 5 does not discuss things in terms of temperature, as this would make things more complicated (adding another diminsion).
+However, I used the other plot because Fig 5 does not discuss things in terms of temperature, as this would make things more complicated (adding another dimension).
 I could just include both versions of figure 6 though (using figure 6a as a bridge to figure 6b)}.
@@
 \caption{Scatter plots of $\frac{\partial \; ET}{\partial \; D}$.
-Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; D}$, as labelled: Column 1 - $\frac{\partial \; ET}{\partial \; D}$; Column 2 - $\frac{\partial \; ET}{\partial \; D}$ normalized by $g_a$; Column 3 - $\frac{\partial \; ET}{\partial \; D}$ with LAI held fixed at PFT average; and Column 4 - $\frac{\partial \; ET}{\partial \; D}$ normalized by $g_a$ and with LAI held fixed.
+Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; D}$, as labeled: Column 1 - $\frac{\partial \; ET}{\partial \; D}$; Column 2 - $\frac{\partial \; ET}{\partial \; D}$ normalized by $g_a$; Column 3 - $\frac{\partial \; ET}{\partial \; D}$ with LAI held fixed at PFT average; and Column 4 - $\frac{\partial \; ET}{\partial \; D}$ normalized by $g_a$ and with LAI held fixed.
 For reference, lines corresponding to RH = 20\% and RH = 90 \% are drawn.
@@
 \caption{****alternate Fig 06****  Scatter plots of $\frac{\partial \; ET}{\partial \; D}$.
-Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; D}$, as labelled.
+Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; D}$, as labeled.
 If I end up using this, I could also draw on the curve of $D_{ETmin}$ with $\overline{LAI}$.
@@
 \begin{itemize}
-  \item Aerodynamic resistance plays an important role of scaling $\frac{\partial \; ET}{\partial \; D}$. This is a leading order effect for observing higher magnitude reponses in DBF and ENF.
+  \item Aerodynamic resistance plays an important role of scaling $\frac{\partial \; ET}{\partial \; D}$. This is a leading order effect for observing higher magnitude responses in DBF and ENF.
 \item In general, CSH has the most negative (i.e.
 ET reduced) response to increases in $D$ (atmospheric drying).
-So CSH plants will almost always try and conserve water, effectivey reducing ET with dry atmospheric perturbation.
-\item CRO has the most postive response (i.e. ET increased) in response to increases in $D$. This is consistent with CROs that may be evolved or bred to thrive in non-water-limited environments.
-\item The reponse is more a function of the environment for DBF, ENF, and GRA.
-Because as VPD increases the response is more likely to be postive, if RH is fixed then the reponse will be more likely to be positive at warmer T, or if T is fixed the response is more likely to be positive with decreasing RH.
+So CSH plants will almost always try and conserve water, effectively reducing ET with dry atmospheric perturbation.
+\item CRO has the most positive response (i.e. ET increased) in response to increases in $D$. This is consistent with CROs that may be evolved or bred to thrive in non-water-limited environments.
+\item The response is more a function of the environment for DBF, ENF, and GRA.
+Because as VPD increases the response is more likely to be positive, if RH is fixed then the response will be more likely to be positive at warmer T, or if T is fixed the response is more likely to be positive with decreasing RH.
 \item ENF has the strongest dependence on environmental conditions due to its small $g1$.
-\item Model and observational uncertainty is highest for GRA and CRO, so conlusions about those PFTs should be tempered.
+\item Model and observational uncertainty is highest for GRA and CRO, so conclusions about those PFTs should be tempered.
 \item However, inclusion of uncertainty doesn't alter conclusions about DBF, ENF, and CSH.

### 2017-10-05 doc/paper/vpd_et_paper.tex
VERSION 1 ROUGH DRAFT, send off to pierre

@@
 
-\journalname{Water Resource Research}
+\journalname{Agricultural and Forest Meteorology}
 
@@
  where variable definitions are given in Table 1.
-\cite{MEDLYN_2011} developed a model for $g_s$ by combining optimal photosynthesis theory with empirical approaches.
+\citet{MEDLYN_2011} developed a model for $g_s$ by combining optimal photosynthesis theory with empirical approaches.
 The result for leaf-scale stomatal resistance was:
@@
 While Equation 3 can be used in PM, it will make analytical work with the function intractable because $A$ is a relatively strong function of ET.
-To remove dependence of ET on $A$ we can use the semi-empirical results of \cite{Zhou_2015}.
-\cite{Zhou_2015} showed that:
+To remove dependence of ET on $A$ we can use the semi-empirical results of \citet{Zhou_2015}.
+\citet{Zhou_2015} showed that:
 
@@
 is relatively constant across time and space (within plant functional type).
-If, following \cite{Lin_2015}, we approximate $g_0$ as $0$, we can use uWUE to remove $A$ from $g_s$ in a way that makes PM analytically tractable:
+If, following \citet{Lin_2015}, we approximate $g_0$ as $0$, we can use uWUE to remove $A$ from $g_s$ in a way that makes PM analytically tractable:
 
@@
 \end{linenomath*}
-Note that given yearly uWUE from \cite{Zhou_2015}, $g_1$ from \cite{Lin_2015} \citep[as presented in ][]{Franks_2017}, and observations of R, T, P, D$_s$, and wind speed (WS), the only unknown is LAI.
+Note that given yearly uWUE from \citet{Zhou_2015}, $g_1$ from \citet{Lin_2015} \citep[as presented in ][]{Franks_2017}, and observations of R, T, P, D$_s$, and wind speed (WS), the only unknown is LAI.
 With flux tower observations of ET, LAI will then be uniquely determined for each observation through Equation 6:
@@
 Because $g_1$ coefficients \citep{Lin_2015} and uWUE were only both available for five plant functional types (PFTs - see Table 2),  only 56 of the 77 sites were used.
-Figure \ref{map_fig} \explain{map needs to be fleshed out} presents each site and its plant functional type.
+Figure \ref{map_fig} \explain{map needs to be improved - it's a placeholder for now} presents each site and its plant functional type.
 
@@
 \includegraphics[width=20pc]{./fig01.pdf}
-\caption{Plant functional type and location of sites used in analysis. Note should probably just split this into 4 continents (US, Europe, Africa, Australia).}
+\caption{Plant functional type and location of sites used in analysis. ***This is just a placeholder for now and needs to be improved i.e. with lat lon, better placement of continents, etc.)***}
 \label{map_fig}
- \end{figure}p
+ \end{figure}
 
-We restrict our analysis to the daytime (sensible heat > 5 W m$^{-1}$ and shortwave radiation > 50 W m$^{-2}$) when there is no precipitation and the plants are growing (GPP > 10\% of the 95th percentile).
+We restrict our analysis to the daytime (sensible heat > 5 W m$^{-1}$ and shortwave radiation > 50 W m$^{-2}$) when there is no precipitation and the plants are growing (GPP > 10\% of the 95th percentile) \explain{This GPP thresholding was used by Changjie, do you know if there is a citation for it?
+Otherwise it seems like something a reviewer would have issue with as it is arbitray}.
 Also, because some sites use half hourly data but some use hourly, we aggregate all data to hourly averages.
@@
 By construction, the variability in the LAI term (Equation \ref{lai}) contains all model and observational uncertainties.
-LAI also has physical meaning corresponding to ``true'' leaf area, and we expect that it would be approximately $O(1)$.
+LAI also has physical meaning corresponding to ``true'' leaf area, and we expect that it would be approximately $O(1)$, especially because uWUE is fit in \citet{Zhou_2015} at the ecosystem-level.
 We can have some confidence in our framework, including the assumption of constant uWUE, if calculated LAIs are generally $O(1)$.
@@
 The lowest and highest 5\% are removed as outliers, as well as any values below 0.
-The curve is normalized such that its area is 1.}
+The curve is normalized such that its area is 1.
+}
 \label{lai_fig}
@@
 An additional concern is that the LAI term may in fact be some function of $D$, in which case the dependence would need to be accounted for when taking the derivative.
-Figure \ref{lai_vpd_fig} plots the joint distribution of LAI and VPD, and shows that LAI is very weakly a function of VPD.
+Figure \ref{lai_vpd_fig}\explain{Figs 2 and 3 can probably be combined - the histogram of LAI is shown in Fig 3}.
+plots the joint distribution of LAI and VPD, and shows that LAI is very weakly a function of VPD.
 Given this weak dependence, we argue that Equation \ref{d_et} is a valid approximation for ET response to $D$.
@@
 \includegraphics[width=20pc]{./fig03.png}
-\caption{The joint distribution of $D$ and LAI. LAI has only a weak dependence on $D$}
+\caption{The joint distribution of $D$ and LAI.
+LAI has only a weak dependence on $D$.
+***This plot could proabably benefit from a box plot of site specific correlations, because some sites do have stronger depednence than others.
+Note also Figs 3 and 2 can probably be combined because this figure shoes LAI's histogram.***}
 \label{lai_vpd_fig}
@@
 $g_a$ will vary strongly with PFT due to the importance of surface roughness.
-However, the coefficient of variability for $g_a$ is relatively constant across PFT.
-So, the influence of $g_a$ on the relative (to the mean) variability of Term 1 is approximately similar across PFT.
+However, the coefficient of variability for both $g_a$ and Term 1 is relatively constant across PFT, suggesting that the influence of $g_a$ on the relative (to the mean) variability of Term 1 is approximately similar across PFT.
 
 Figure \ref{scale_vary}A shows Term 1 normalized by mean $g_a$ (calculated for each plant functional type), and confirms that much of the relative variability of Term 1 is contained in the $g_a$ term's relative variability.
+Generally, $T$ has less of a role.
 Additionally, the impact of $T$ on the relative variability increases with increasing $g_a$.
@@
 For example, ENF experiences on average colder temperatures and is thus more likely to have a larger scaling term.
-Additionally, because std$(g_a) \propto \overline{g_a}$, the spread of Term 1 due to $g_a$ variability will be larger, although this is not shown for simplicity.
-To summarize, the variability of Term 1 will look like Figure \ref{scale_vary}A for each PFT, but the scale of the x and y-axis will increase or decrease according to Figure \ref{scale_vary}B.
+Additionally, because std$(g_a) \propto \overline{g_a}$, the spread of Term 1 due to $g_a$ variability will be larger for the tree PFTs, although this is not shown for simplicity.
+To summarize, the variability of Term 1 within each PFT will look like Figure \ref{scale_vary}A for each PFT, but the scale of the x and y-axis will increase or decrease according to mean $g_a$ oberved in Figure \ref{scale_vary}B.
 
@@
 %%%%% replace below with idea:
-
-``just say that vpd effect is approximately linear (except at low VPD), and then say so we are most concerned about when d et/dvpd is zero, as this is the minima.
-then you can just look at this treshold as a function of LAI, uWUE, and g1.''
 
@@
 
-Bottom: The location of the minima of ET, as a function of VPD and LAI. Lines and stars denote the distribution of VPD and LAI next to axis, following the same percentiles as above.}
+Bottom: The location of the minima of ET, as a function of VPD and LAI. Lines and stars denote the distribution of VPD and LAI next to each axis, following the same percentiles as above.}
 \label{term3}
@@
 For any LAI or VPD less (more) than these curves, Term 2 - Term 3 will be negative (positive).
-It is clear that the portion of VPD observations below these curves will be a strong function of $LAI$.
+It is clear that the portion of VPD observations below/above these curves will be a strong function of $LAI$.
 However, we can see some general trends.
@@
 These observations highlight the effect of the nonlinear function in Figure \ref{term3}: $\frac{\partial \; ET}{\partial \; D}$ has a much steeper slope when the function is negative, and is thus more likely to be large magnitude.
-n
+
 The units of $\frac{\partial \; ET}{\partial \; D}$ make it difficult to interpret if $D$ is even a meaningful contributor to ET's variability.
 To understand $D$'s contribution better, we use a linear approximation and present $\frac{\partial \; ET}{\partial \; D}$ multiplied by $D$'s standard deviation.
-The range of $D$'s contribution to ET's variability ranges between 16 - 20 W m$^{-2}$ for all PFTs except for CSH, which is about 51 W m$^{-2}$.
+The $D$'s contribution to ET's variability ranges between 16 - 20 W m$^{-2}$ for all PFTs except for CSH, which is about 51 W m$^{-2}$.
 Another meaningful comparison is to $\frac{\partial \; ET}{\partial \; R} * std(R)$, as net radiation generally the driver of ET (cite joe berry here).
@@
 However it is important to note that a linear approximation about a mean base state is probably not a very good approximation across the range of variability, so these values are just estimates of $D$'s contribution to ET's variability.
+Regardless, $D$'s variability is certiantly a non-negligable contributor to $ET$'s variability.
 
@@
 This has the side effect of pushing $\overline{\frac{\partial \; ET}{\partial \; D}}$ negative further than other PFTs for a given fraction $\frac{\partial \; ET}{\partial \; D} < 0$ and magnitude $\frac{\partial \; ET}{\partial \; D}(\overline{T,\ldots,D})$.
-GRA shows the opposite behavior; a relatively high $g1$  makes the function more linear, decreasing the magnitude of $\overline{\frac{\partial \; ET}{\partial \; D}}$ for a given fraction $\frac{\partial \; ET}{\partial \; D} < 0$ and magnitude $\frac{\partial \; ET}{\partial \; D}(\overline{T,\ldots,D})$ (although $g_a$ and Term 1 also probably have a role in this).
+GRA shows the opposite behavior; a relatively high $g1$  makes the function more linear, decreasing the magnitude of $-\overline{\frac{\partial \; ET}{\partial \; D}}$ for a given  [large] fraction $\frac{\partial \; ET}{\partial \; D} < 0$ and negative $\frac{\partial \; ET}{\partial \; D}(\overline{T,\ldots,D})$ (although $g_a$ and Term 1 also probably have a role in this).
 Finally, low $uWUE$ of CSH pushes to toward by far the lowest values $\frac{\partial \; ET}{\partial \; D}$ (Figure \ref{term3}).
@@
 Through the columns of Figure \ref{real} we can see the impact of LAI and $g_a$ on the variability of $\frac{\partial \; ET}{\partial \; D}$.
-$g_a$'s (columns 1 and 3) scaling alters the magnitude considerably.
-$LAI$ (columns 2 and 3) variability adds a lot of additional noise to the signal of $\frac{\partial \; ET}{\partial \; D}$, which is slightly undesirable given LAI's role in representing model and observational uncertainty.
+$g_a$'s scaling (included in columns 1 and 3) alters the magnitude considerably.
+$LAI$ variability (included in columns 1 and 2) adds a lot of additional noise to the signal of $\frac{\partial \; ET}{\partial \; D}$, which is slightly undesirable given LAI's role in representing model and observational uncertainty.
 However, the general story with the noise appears to match the cleaner signal when $LAI$ is help constant and $D_{ETmin}$ is clearly visible \explain{I really need to make these plots better - way too much overlapping of points that hurts the story}.
-One exception is possibly with GRA, for which uncertainty is high and causes the full complexity plots (Columns 1 and 2) to not match well with LAI held fixed (Columns 3 and 4).
+One exception is possibly with GRA, for which uncertainty represented in LAI is high and causes the full complexity plots (Columns 1 and 2) to not match well with LAI held fixed (Columns 3 and 4).
 
-For ENF and GRA $D_{ETmin}$ does not appear to be only a function of LAI.
+For ENF and GRA $D_{ETmin}$ does not appear to be only a function of LAI (most observable in Column 4).
 It turns out that the site to site variability in $\gamma$ causes $D_{ETmin}$ to vary, which is not discussed in the previous section.
@@
 We therefor have the most confidence in our conclusion based on Equation \ref{d_et} for PFTS CSH, DBF, and ENF, as the full complexity plots with uncertainty included closely match the story when LAI is held fixed.
-**see somewhat preferred alternate figure \ref{real2} \explain{I think I like this more as thinking in terms of T and RH is easier.
+**see somewhat preferred alternate figure \ref{real2} \explain{I think I like the alternate plot much  more as thinking in terms of T and RH is easier, and it makes the story easier to see at relatively low temperatures.
 However, I used the other plot because Fig 5 does not discuss things in terms of temperature, as this would make things more complicated (adding another dimension).
@@
 For reference, lines corresponding to RH = 20\% and RH = 90 \% are drawn.
-Please note differences in the colorbar scale.}
+Please note differences in the colorbar scale.
+***see alternate (or additional) plot below.***}
 \label{real}
@@
 So CSH plants will almost always try and conserve water, effectively reducing ET with dry atmospheric perturbation.
+    \item Additionally for CSH, $D$ variability contributes the most to $ET$ variability.
 \item CRO has the most positive response (i.e. ET increased) in response to increases in $D$. This is consistent with CROs that may be evolved or bred to thrive in non-water-limited environments.
@@
 
+\input{vpd_et_paper.bbl}
+%\bibliography{references.bib}
+
 %% After you run BibTeX, Copy in the contents of the .bbl file here:

### 2017-10-10 doc/paper/vpd_et_paper.tex
starting to work on Pierre's comments

@@
 Changes to vapor pressure deficit (VPD) alter the atmospheric demand for water from the land surface.
+Atmospheric scientists and hydrometeorologists generally think that an increase in atmospheric demand induces an increases in evapotranspiration (ET).
 However, plant stomata have evolved to optimally regulate the exchange of water and carbon between vegetation and the atmosphere \citep{Franks_2017}.
-Therefore, an increase (decrease) in VPD may not correspond to an increase (decrease) in evapotranspiration (ET) because stomatal closure (opening) can cancel the effects of shifts to atmospheric demand.
+Therefore, an increase (decrease) in VPD may not correspond to an increase (decrease) in ET stomatal closure (opening) can cancel the effects of shifts to atmospheric demand.
 

### 2017-10-12 doc/paper/vpd_et_paper.tex
finished with superficial comments up to page 7, need to start working on meatier conecptual comments, read monteith's oroginal 1965 paper.

--- 2017-10-05 doc/paper/vpd_et_paper.tex
@@
 Changes to vapor pressure deficit (VPD) alter the atmospheric demand for water from the land surface.
-However, plant stomata have evolved to optimally regulate the exchange of water and carbon between vegetation and the atmosphere \citep{Franks_2017}.
-Therefore, an increase (decrease) in VPD may not correspond to an increase (decrease) in evapotranspiration (ET) because stomatal closure (opening) can cancel the effects of shifts to atmospheric demand.
+Traditionally, atmospheric scientists and hydrometeorologists generally think that an increase in atmospheric demand induces an increases in evapotranspiration (ET) (citations?).
+This possible misconception developed in part due to the proliferation of studies examing potential ET rather than estimates of ET itself (citations?).
+In contrast, plant physiologists know that stomata have evolved to optimally regulate the exchange of water and carbon, and tend to close in response to increased atmospheric dryness \citep{Ball_1987, Leuning_1990, MEDLYN_2011}.
+Therefore, an increase (decrease) in VPD may not correspond to an increase (decrease) in ET because stomatal closure (opening) can cancel the effects of shifts to atmospheric demand.
 
-Quantifying the plant response to a perturbation to atmospheric VPD increases our understanding of feedbacks between the land surface and the atmosphere.
-If plant response reduces ET in response to an increase in VPD, the land surface will contribute a positive feedback in response to atmospheric drying.
-Conversely, if plant response increases ET in response to increase in VPD, then the land surface will contribute a negative feedback to atmospheric drying.
-The sign of these feedbacks drives the evolution of the atmosphere and land-surface at many timescales, from diurnal to interdecadal.
+Quantifying the plant response to a perturbation to atmospheric VPD increases our understanding of land surface response to shifts in atmospheric conditions.
+If plant response reduces ET in response to atmospheric drying then soil moisture will be conserved.
+An increase in ET in reponse to atmospheric drying will reduce soil moisture, but contribute increased moisenting to the atmosperhere.
+Clearly, the sign and magnitude of land-surface responsedrives the co-evolution of the atmosphere and land-surface at many timescales, from diurna to interdecadal.
 
-Here we use a Penman-Monteith framework to quantify plant response to perturbations to atmospheric demand for water.
-Section 2 derives the framework, Section 3 describes the data used, Section 4 presents results, and Section 5 discusses conclusions.
+In order to quantify plant response to perturbations to atmospheric demand for water, we apply a Penman-Monteith framework to eddy-covariance observation spanning various biomes and climates.
+Section 2 describes the data used, Section 3 derives the framework, Section 4 presents results, and Section 5 discusses conclusions.
 The goal of this paper is to use reasonable approximations as a tool to increase intuition for plant response to atmospheric drying.
@@
 \begin{linenomath*}
+  \label{medlyn}
   \begin{equation}
-  g_s = \text{LAI } \frac{R* T}{P} \left( g_0 + 1.6 \left(1 + \frac{g_1}{\sqrt{D_{s}}}\right) \frac{A}{c_s}\right)
+  g_s = \text{LAI } \frac{R \,T}{P} \left( g_0 + 1.6 \left(1 + \frac{g_1}{\sqrt{D_{s}}}\right) \frac{A}{c_s}\right)
   \end{equation}
@@
 
-While Equation 3 can be used in PM, it will make analytical work with the function intractable because $A$ is a relatively strong function of ET.
+While Equation 3 can be used in PM, it will make analytical work with the function intractable because $A$ is a function of ET itself.
 To remove dependence of ET on $A$ we can use the semi-empirical results of \citet{Zhou_2015}.
@@
   \begin{equation}
-  g_s = \text{LAI } \frac{R* T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{D_{s}}}\right) \frac{uWUE \; ET}{c_s \; \sqrt{D}}
+  g_s = \frac{\text{LAI }}{\text{LAI$_{ref}$}} \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{D_{s}}}\right) \frac{uWUE \; ET}{c_s \; \sqrt{D}}
   \end{equation}
@@
 
-Plugging Equation 5 into Equation 1 and rearranging gives:
+Note that $uWUE$ is fit on the ecosystem scale in \citet{Zhou_2015} so it already has an implicit LAI$_{ref}$ reflective of the average $LAI$ in \citet{Zhou_2015}'s fits.
+So the LAI term in Equation \ref{medlyn} becomes a term we call $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$, which is a measure of how close the observed LAI is to the LAI$_{ref}$ in \citet{Zhou_2015}.
+
+Next, plugging Equation 5 into Equation 1 and rearranging gives:
 
@@
   \begin{equation}
-  ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p D_{s}}{R_{air}} - \frac{\gamma c_s \sqrt{D} }{\text{LAI } R* \; 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{D}})} \right) }{ \Delta + \gamma}
+  ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p D_{s}}{R_{air}} - \frac{\text{LAI$_{ref}$}}{\text{LAI}}  \frac{\gamma c_s \sqrt{D} }{ R* \; 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{D}})} \right) }{ \Delta + \gamma}
   \end{equation}
@@
 
-We can then take the derivative with respect to $D$ to determine ecosystem response to atmospheric demand perturbations:
+We can then take the derivative with respect to $D$ to analytically determine ecosystem response to atmospheric demand perturbations:
 
@@
   \begin{equation}
-    \frac{\partial \;  ET}{\partial \; D} = \frac{g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{\text{LAI }1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right) \right)
+    \frac{\partial \;  ET}{\partial \; D} = \frac{g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} - \frac{\text{LAI$_{ref}$}}{\text{LAI}} \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right) \right)
     \label{d_et}
@@
 \end{linenomath*}
-Note that given yearly uWUE from \citet{Zhou_2015}, $g_1$ from \citet{Lin_2015} \citep[as presented in ][]{Franks_2017}, and observations of R, T, P, D$_s$, and wind speed (WS), the only unknown is LAI.
-With flux tower observations of ET, LAI will then be uniquely determined for each observation through Equation 6:
+Note that given yearly uWUE from \citet{Zhou_2015}, $g_1$ from \citet{Lin_2015} \citep[as presented in ][]{Franks_2017}, and observations of R, T, P, D$_s$, and wind speed (WS), the only unknown is $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$.
+With flux tower observations of ET, $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ will then be uniquely determined for each observation through Equation 6:
 
@@
   \begin{equation}
-    LAI  = - \frac{g_a \gamma c_s \sqrt{D_s} P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p D_{s}\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{D_s}})}
+    \frac{\text{LAI}}{\text{LAI$_{ref}$}}  = - \frac{g_a \gamma c_s \sqrt{D_s} P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p D_{s}\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{D_s}})}
     \label{lai}
@@
 
-This ``pseudo-LAI'' is some part ``true'' LAI (a measure of leaf area), and some part model and observational error, including error involving our assumption of constant uWUE.
-By calculating a unique LAI for each observation we will propagate any model and observational uncertainty forward into our expression for $\frac{\partial \; ET}{\partial D}$.
+%% see registers!!!!!! - help for emacs productivity
+This $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ is some part a ``true'' deviation of  LAI from LAI$_{ref}$ in \citet{Zhou_2015}, and some part model and observational error, including assumptions about constant $g1$ and $uWUE$.
+By calculating a unique $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ for each observation we will propagate any model and observational uncertainty forward into our expression for $\frac{\partial \; ET}{\partial D}$.
 
@@
   $LAI$ & leaf area index & -\\
+  $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ & ratio of LAI to LAI$_{ref}$ in \citet{Zhou_2015} & -\\
   $c_s$ & CO$_2$ concentration & $\mu$ mol CO$_2$ mol$^{-1}$ air\\
@@
 
-By construction, the variability in the LAI term (Equation \ref{lai}) contains all model and observational uncertainties.
-LAI also has physical meaning corresponding to ``true'' leaf area, and we expect that it would be approximately $O(1)$, especially because uWUE is fit in \citet{Zhou_2015} at the ecosystem-level.
-We can have some confidence in our framework, including the assumption of constant uWUE, if calculated LAIs are generally $O(1)$.
-Figure \ref{lai_fig} presents the histogram of calculated LAIs with outliers (lowest and highest 5\% percent) and nonphysical values (LAI $<$ 0.) removed.
-All remaining LAI values are $O(1)$ which provides confidence in model framework.
+By construction, the variability in the $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ term (Equation \ref{lai}) contains all model and observational uncertainties.
+We expect LAI to be cloase to LAI$_{ref}$, so $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ should be $O(1)$.
+We can have some confidence in our framework, including the assumption of constant uWUE, if calculated $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$s are generally $O(1)$.
+Figure \ref{lai_fig} presents the histogram of calculated $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$s with outliers (lowest and highest 5\% percent) and nonphysical values ($\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ $<$ 0.) removed.
+All remaining $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ values are $O(1)$ which provides confidence in model framework.
 
@@
 \includegraphics[width=20pc]{./fig02.png}
-\caption{Histogram of LAI values calculated for each site and time according to Equation \ref{lai}.
+\caption{Histogram of $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ values calculated for each site and time according to Equation \ref{lai}.
 The lowest and highest 5\% are removed as outliers, as well as any values below 0.
@@
 
-An additional concern is that the LAI term may in fact be some function of $D$, in which case the dependence would need to be accounted for when taking the derivative.
-Figure \ref{lai_vpd_fig}\explain{Figs 2 and 3 can probably be combined - the histogram of LAI is shown in Fig 3}.
-plots the joint distribution of LAI and VPD, and shows that LAI is very weakly a function of VPD.
+An additional concern is that the $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ term may in fact be some function of $D$, in which case the dependence would need to be accounted for when taking the derivative.
+Figure \ref{lai_vpd_fig}\explain{Figs 2 and 3 can probably be combined - the histogram of $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ is shown in Fig 3}.
+plots the joint distribution of $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ and VPD, and shows that $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ is very weakly a function of VPD.
 Given this weak dependence, we argue that Equation \ref{d_et} is a valid approximation for ET response to $D$.
@@
 \includegraphics[width=20pc]{./fig03.png}
-\caption{The joint distribution of $D$ and LAI.
-LAI has only a weak dependence on $D$.
+\caption{The joint distribution of $D$ and $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$.
+$\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ has only a weak dependence on $D$.
 ***This plot could proabably benefit from a box plot of site specific correlations, because some sites do have stronger depednence than others.
-Note also Figs 3 and 2 can probably be combined because this figure shoes LAI's histogram.***}
+Note also Figs 3 and 2 can probably be combined because this figure shoes $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$'s histogram.***}
 \label{lai_vpd_fig}
@@
 
-Before diving into calculated values of $\frac{\partial \; ET}{\partial \; D}$, it is useful to consider the functional form of Equation \ref{d_et}.
-There are three terms: a scaling term for the full expression we will call Term 1 ($\frac{g_a \; P}{T(\Delta + \gamma)}$), a relatively constant offset we will call Term 2 ($\frac{c_p}{R_{air}}$), and a variable term we will call Term 3 ($\frac{\gamma c_s }{\text{LAI }1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right)$).
+Before calculating the sensitivity of ET to VPDn, it is useful to consider the functional form of Equation \ref{d_et}.
+There are three terms: a scaling term for the full expression we will call Term 1 ($\frac{g_a \; P}{T(\Delta + \gamma)}$), a relatively constant offset we will call Term 2 ($\frac{c_p}{R_{air}}$), and a variable term we will call Term 3 ($\frac{\text{LAI$_{ref}$}}{\text{LAI}}\frac{\gamma c_s }{1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right)$).
 All variables are positive, so the relative magnitude between Term 2 and Term 3 will determine the sign of the derivative, while Term 1 will scale the expression larger or smaller.
@@
 
-Term 2 minus Term 3 determines the sign and magnitude that the scaling Term 1 is multiplied by.
-If we assume that $c_s$ variability is relatively less than LAI and $D$ variability, then variability within PFT will be solely determined by LAI and $D$.
-Figure \ref{term3} shows how (Term 2 - Term 3) varies with $D$ and LAI, as a function of PFT.
-In Figure \ref{term3}a lower uWUE and LAI shift the distribution of (Term 2 - Term 3) towards negative values.
+Term 2 minus Term 3 determines the sign of the sensitivty of ET to VPD and is thus crucial.
+$c_s$ variability is relatively less than $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ and $D$ variability, then variability within PFT will be solely determined by $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ and $D$ at the different fluxnet sites.
+Figure \ref{term3} shows how (Term 2 - Term 3) varies with $D$ and $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$, as a function of PFT.
+In Figure \ref{term3}a lower uWUE and $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ shift the distribution of (Term 2 - Term 3) towards negative values.
 Additionally, the smaller $g1$, the greater the relative $D$ dependence of of (Term 2 - Term 3).
@@
 \caption{Sources of variability for Term 2 - Term 3.
-Top: Term 2 - Term 3 as a function of VPD, with LAI held fixed at PFT averages.
+Top: Term 2 - Term 3 as a function of VPD, with $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ held fixed at PFT averages.
 The observed range of VPD for each PFT is also shown below the x-axis.
@@
 
-Bottom: The location of the minima of ET, as a function of VPD and LAI. Lines and stars denote the distribution of VPD and LAI next to each axis, following the same percentiles as above.}
+Bottom: The location of the minima of ET, as a function of VPD and $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$.
+Lines and stars denote the distribution of VPD and $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ next to each axis, following the same percentiles as above.}
 \label{term3}
@@
 
-Figure \ref{term3}b shows the location of the minima of ET, as a function of LAI and $D$.
-For any LAI or VPD less (more) than these curves, Term 2 - Term 3 will be negative (positive).
-It is clear that the portion of VPD observations below/above these curves will be a strong function of $LAI$.
+Figure \ref{term3}b shows the location of the minima of ET, as a function of $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ and $D$.
+For any $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ or VPD less (more) than these curves, Term 2 - Term 3 will be negative (positive).
+It is clear that the portion of VPD observations below/above these curves will be a strong function of $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$.
 However, we can see some general trends.
-For CSH, $\frac{\partial \; ET}{\partial \; D}$ should be negative for the vast majority of observed LAI and VPD.
-The split appears to be more even among ENF, GRA, and DBF, and we might expect a greater frequency of positive $\frac{\partial \; ET}{\partial \; D}$ for CRO.
+For CSH, $\frac{\partial \; ET}{\partial \; D}$ should be negative for the vast majority of observed $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ and VPD.
+The fraciton of positive $\frac{\partial \; ET}{\partial \; D}$ appears to be more even for ENF, GRA, and DBF, and we might expect a greater frequency of positive $\frac{\partial \; ET}{\partial \; D}$ for CRO.
 
@@
 However, $\frac{\partial \; ET}{\partial \; D}$ evaluated at the average of all variables (e.g.
-LAI, $T$, $c_s$, $D$) is only negative for CSH and GRA.
+$\frac{\text{LAI}}{\text{LAI$_{ref}$}}$, $T$, $c_s$, $D$) is only negative for CSH and GRA.
 And, DBF in addition to CRO experiences $\frac{\partial \; ET}{\partial \; D}$ < 0 less than half the time.
-These observations highlight the effect of the nonlinear function in Figure \ref{term3}: $\frac{\partial \; ET}{\partial \; D}$ has a much steeper slope when the function is negative, and is thus more likely to be large magnitude.
+These observations highlight the effect of the nonlinear function in Figure \ref{term3}: $\frac{\partial \; ET}{\partial \; D}$ has a much steeper slope when the function is negative, and is thus more likely to be large.
 
 The units of $\frac{\partial \; ET}{\partial \; D}$ make it difficult to interpret if $D$ is even a meaningful contributor to ET's variability.
-To understand $D$'s contribution better, we use a linear approximation and present $\frac{\partial \; ET}{\partial \; D}$ multiplied by $D$'s standard deviation.
-The $D$'s contribution to ET's variability ranges between 16 - 20 W m$^{-2}$ for all PFTs except for CSH, which is about 51 W m$^{-2}$.
-Another meaningful comparison is to $\frac{\partial \; ET}{\partial \; R} * std(R)$, as net radiation generally the driver of ET (cite joe berry here).
+To better understand $D$'s contribution, we normalize $\frac{\partial \; ET}{\partial \; D}$ with $D$'s standard deviation to define a (linearized) relative change in ET for variations in $D$ .
+$D$'s contribution to ET's variability ranges between 16 - 20 W m$^{-2}$ for all PFTs except for CSH, which is about 51 W m$^{-2}$.
+Another meaningful comparison is to $\frac{\partial \; ET}{\partial \; R} * std(R)$, as net radiationis generally the driver of ET (cite joe berry here).
 For all PFTs except for CSH $D$ contributes between 14.5 - 20.5 \% of $R$'s contribution to variability.
 For CSH the portion is much larger, about 44 \%.
-However it is important to note that a linear approximation about a mean base state is probably not a very good approximation across the range of variability, so these values are just estimates of $D$'s contribution to ET's variability.
-Regardless, $D$'s variability is certiantly a non-negligable contributor to $ET$'s variability.
+$D$'s variability is certiantly a non-negligable contributor to $ET$'s variability.
 
 So far, idealized plots and statistics have illuminated the form of $\frac{\partial \; ET}{\partial \; D}$ and how it varies with PFT.
-Large mean LAI and uWUE shifts CRO and DBF towards positive $\frac{\partial \; ET}{\partial \; D}$.
+Large mean $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ and uWUE shifts CRO and DBF towards positive $\frac{\partial \; ET}{\partial \; D}$.
 However, the strongly nonlinear function of $\frac{\partial \; ET}{\partial \; D}$ at $\frac{\partial \; ET}{\partial \; D} < 0$ pushes $\overline{\frac{\partial \; ET}{\partial \; D}}$ negative for DBF (it does not do this for CRO because of CRO's high $g1$).
@@
 CRO has the most positive values of $\frac{\partial \; ET}{\partial \; D}$, $\frac{\partial \; ET}{\partial \; D}$ is almost always negative for CSH, and response depends more with the environmental conditions for the other PFTs (especially ENF).
-Through the columns of Figure \ref{real} we can see the impact of LAI and $g_a$ on the variability of $\frac{\partial \; ET}{\partial \; D}$.
+Through the columns of Figure \ref{real} we can see the impact of $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ and $g_a$ on the variability of $\frac{\partial \; ET}{\partial \; D}$.
 $g_a$'s scaling (included in columns 1 and 3) alters the magnitude considerably.
-$LAI$ variability (included in columns 1 and 2) adds a lot of additional noise to the signal of $\frac{\partial \; ET}{\partial \; D}$, which is slightly undesirable given LAI's role in representing model and observational uncertainty.
-However, the general story with the noise appears to match the cleaner signal when $LAI$ is help constant and $D_{ETmin}$ is clearly visible \explain{I really need to make these plots better - way too much overlapping of points that hurts the story}.
-One exception is possibly with GRA, for which uncertainty represented in LAI is high and causes the full complexity plots (Columns 1 and 2) to not match well with LAI held fixed (Columns 3 and 4).
+$\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ variability (included in columns 1 and 2) adds a lot of additional noise to the signal of $\frac{\partial \; ET}{\partial \; D}$, which is slightly undesirable given $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$'s role in representing model and observational uncertainty.
+However, the general story with the noise appears to match the cleaner signal when $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ is help constant and $D_{ETmin}$ is clearly visible \explain{I really need to make these plots better - way too much overlapping of points that hurts the story}.
+One exception is possibly with GRA, for which uncertainty represented in $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ is high and causes the full complexity plots (Columns 1 and 2) to not match well with $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ held fixed (Columns 3 and 4).
 
-For ENF and GRA $D_{ETmin}$ does not appear to be only a function of LAI (most observable in Column 4).
+For ENF and GRA $D_{ETmin}$ does not appear to be only a function of $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ (most observable in Column 4).
 It turns out that the site to site variability in $\gamma$ causes $D_{ETmin}$ to vary, which is not discussed in the previous section.
@@
 
-In general the full complexity plots of $\frac{\partial \; ET}{\partial \; D}$ match our expectations, even with the large sensitivity to LAI measures of uncertainty observed in Figure \ref{term3}.
-Our LAI-based method of uncertainty propagation blurs the idealized expectations the most for GRA, and also has a considerable effect for CRO.
-We therefor have the most confidence in our conclusion based on Equation \ref{d_et} for PFTS CSH, DBF, and ENF, as the full complexity plots with uncertainty included closely match the story when LAI is held fixed.
+In general the full complexity plots of $\frac{\partial \; ET}{\partial \; D}$ match our expectations, even with the large sensitivity to $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ measures of uncertainty observed in Figure \ref{term3}.
+Our $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$-based method of uncertainty propagation blurs the idealized expectations the most for GRA, and also has a considerable effect for CRO.
+We therefor have the most confidence in our conclusion based on Equation \ref{d_et} for PFTS CSH, DBF, and ENF, as the full complexity plots with uncertainty included closely match the story when $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ is held fixed.
 **see somewhat preferred alternate figure \ref{real2} \explain{I think I like the alternate plot much  more as thinking in terms of T and RH is easier, and it makes the story easier to see at relatively low temperatures.
@@
 \caption{Scatter plots of $\frac{\partial \; ET}{\partial \; D}$.
-Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; D}$, as labeled: Column 1 - $\frac{\partial \; ET}{\partial \; D}$; Column 2 - $\frac{\partial \; ET}{\partial \; D}$ normalized by $g_a$; Column 3 - $\frac{\partial \; ET}{\partial \; D}$ with LAI held fixed at PFT average; and Column 4 - $\frac{\partial \; ET}{\partial \; D}$ normalized by $g_a$ and with LAI held fixed.
+Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; D}$, as labeled: Column 1 - $\frac{\partial \; ET}{\partial \; D}$; Column 2 - $\frac{\partial \; ET}{\partial \; D}$ normalized by $g_a$; Column 3 - $\frac{\partial \; ET}{\partial \; D}$ with $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ held fixed at PFT average; and Column 4 - $\frac{\partial \; ET}{\partial \; D}$ normalized by $g_a$ and with $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ held fixed.
 For reference, lines corresponding to RH = 20\% and RH = 90 \% are drawn.
@@
 Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; D}$, as labeled.
-If I end up using this, I could also draw on the curve of $D_{ETmin}$ with $\overline{LAI}$.
+If I end up using this, I could also draw on the curve of $D_{ETmin}$ with $\overline{$\frac{\text{LAI}}{\text{LAI$_{ref}$}}$}$.
 }

### 2017-10-19 doc/glab_meeting_2017-10-20/vpd.tex
added glab presentation

\documentclass{beamer}
%
% Choose how your presentation looks.
%
% For more themes, color themes and font themes, see:
% http://deic.uab.es/~iblanes/beamer_gallery/index_by_theme.html
%
\mode<presentation>
{
  \usetheme{default}      % or try Darmstadt, Madrid, Warsaw, ...
  \usecolortheme{default} % or try albatross, beaver, crane, ...
  \usefonttheme{default}  % or try serif, structurebold, ...
  \setbeamertemplate{navigation symbols}{}
  \setbeamertemplate{caption}[numbered]
}

\usepackage[english]{babel}
 % \usepackage[utf8x]{inputenc}
\usepackage[style=authoryear]{biblatex}
\addbibresource{references.bib}

\title[Your Short Title]{Bayesian Neural  Networks and Probalistic Programming}
\author{Adam Massmann and Laureline Josset}
\institute{Water Center NN Meetings}
\date{Week 5: Oct. 17th, 2017}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\section{References}
\begin{frame}{References}
  \AtNextBibliography{\small}
  \printbibliography
\end{frame}

% \begin{figure}
%   \includegraphics[width=3in]{ranganath.png}
%   \caption{Reproduced from \cite{ranganath2014}.}
% \end{figure}

% \begin{table}
% \centering
% \begin{tabular}{l|r}
% Item & Quantity \\\hline
% Widgets & 42 \\
% Gadgets & 13
% \end{tabular}
% \caption{\label{tab:widgets}An example table.}
% \end{table}

\end{document}

### 2017-10-20 doc/glab_meeting_2017-10-20/vpd.tex
added vpd, calling it quits fo rthe night

@@
 
-\title[Your Short Title]{Bayesian Neural  Networks and Probalistic Programming}
-\author{Adam Massmann and Laureline Josset}
-\institute{Water Center NN Meetings}
-\date{Week 5: Oct. 17th, 2017}
+\title[Your Short Title]{When does VPD drive or reduce ET?}
+\author{Adam with Pierre and Changjie}
+\institute{G-Lab Meeting}
+\date{October 20th, 2017}
 
@@
   \titlepage
+\end{frame}
+
+\begin{frame}{When does VPD drive or reduce ET?}
+  \begin{itemize}
+  \item \textbf{Hydrometeorologists} would say that an increase in VPD (increase in \textbf{atmospheric demand}) would drive an \textbf{increase in ET}.
+  \item However, \textbf{plant physioligists} know that plants have evolved to use stomata to conserve and regulate water use.
+So \textbf{stomata closure} in response to increases in VPD may \textbf{decrease ET}.
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{The question is, which effect dominates with an increase in VPD: plant response (decrease in ET) or atmospheric demand (increase in ET)?}
+  \begin{itemize}
+  \item We hypothesize it will be a function of plant type and the environment:
+    \begin{itemize}
+    \item Plants that are evolved to conserve water will tend to reduce ET with increases in VPD.
+    \item However the environment can overwhelm plant response: at very high VPD the atmospheric demand will dominate and plants will not be able to consernve water, no matter how much they have evolved to do so.
+    \end{itemize}
+  \end{itemize}
 \end{frame}

### 2017-10-20 doc/glab_meeting_2017-10-20/vpd.tex
progress on presentation

@@
 \title[Your Short Title]{When does VPD drive or reduce ET?}
-\author{Adam with Pierre and Changjie}
+\author{Adam (with Pierre and Changjie)}
 \institute{G-Lab Meeting}
@@
 
+\section{Introduction}
 \begin{frame}{When does VPD drive or reduce ET?}
@@
   \begin{itemize}
-  \item We hypothesize it will be a function of plant type and the environment:
+  \item It should be a function of plant type and the environment:
     \begin{itemize}
     \item Plants that are evolved to conserve water will tend to reduce ET with increases in VPD.
-    \item However the environment can overwhelm plant response: at very high VPD the atmospheric demand will dominate and plants will not be able to consernve water, no matter how much they have evolved to do so.
+    \item However the environment can overwhelm plant response: at some threshold (i.e.
+very high VPD) the atmospheric demand should dominate and plants will not be able to consernve water, no matter how much they have evolved to do so.
     \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\section{Method}
+\begin{frame}{We will try an analytical approach to this question:}
+  We can use Penman-Monteith (PM) to estimate ET:
+  \[ET = \frac{\Delta R + g_a \rho_a c_p D}{\Delta + \gamma(1 + \frac{g_a}{g_s})},\]
+  But in:
+  \[  g_{s} = g_0 + 1.6 \left(1 + \frac{g_1}{\sqrt{D}}\right) \frac{A}{c_s}\]
+  $A$ is a function of $ET$. So $ET$ in PM is really an implicit function of itself.
+\end{frame}
+
+\begin{frame}{Remove dependence of $ET$ on $A$}
+  \cite{Zhou_2016} shows that uWUE:
+  \[uWUE = \frac{GPP \cdot \sqrt{D}}{ET}\]
+  is conserved across space and time (within PFT).\\
+  \medskip
+  So we can use uWUE to remove dependence of $g_s$ on $A$:
+  \[g_s = \frac{\text{LAI }}{\text{LAI$_{ref}$}} \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{D_{s}}}\right) \frac{uWUE \; ET}{c_s \; \sqrt{D}}\]
+\end{frame}
+
+\begin{frame}{Derive new form of PM}
+  If we plug in our new expression for $g_s$ we can solve for ET to get a new form of PM:
+  \[  ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p D_{s}}{R_{air}} - \frac{\text{LAI$_{ref}$}}{\text{LAI}}  \frac{\gamma c_s \sqrt{D} }{ R* \; 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{D}})} \right)\footnote{note that all terms are known, except for $\frac{\text{LAI$_{ref}$}}{\text{LAI}}$} }{ \Delta + \gamma}\]
+  Now, we can take derivatives, and:
+  \[\frac{\partial \;  ET}{\partial \; D} = \frac{g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} - \frac{\text{LAI$_{ref}$}}{\text{LAI}} \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right) \right)\]
+\end{frame}
+
+\section{Results}
+\begin{frame}{Scaling Term}
+  \[\frac{\partial \;  ET}{\partial \; D} = \mathbf{ \frac{g_a \; P}{T(\Delta + \gamma)} }   \left(\frac{ c_p}{R_{air}} - \frac{\text{LAI$_{ref}$}}{\text{LAI}} \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right) \right)\]
+\begin{figure}
+  \includegraphics[width=3.5in]{scaling.png}
+  % \caption{}
+\end{figure}
+\end{frame}
+
+\begin{frame}{``Sign'' Term}
+  \begin{figure}
+    \includegraphics[width=3.5in]{sign.png}
+    % \caption{}
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Summary of theory}
+  \begin{itemize}
+  \item We use \cite{Zhou_2016}'s uWUE to derive a new analytically tractable form of PM.
+  \item This new analysis suggests that the ``tipping'' point for which atmospheric demand overwhelms plant response will be almost exclusively a function of VPD.
+    \begin{itemize}
+    \item For each PFT there will be a VPD$_{crit}$ above which atmospheric demand will dominate and ET will increase with VPD.
+    \end{itemize}
+  \item Plant types evolved to conserve water (CSH) have a higher VPD$_{crit}$ than plants evolved (or bred) to use water and prioritize GPP (CRO).
+Trees and grasslands are somewhere between these two extremes.
+  \item Aerodynamic conductance scales the response, so plants with large surface roughness will be more likely to have a larger response as there is less resistance between the surface and the atmosphere.
   \end{itemize}
@@
 \end{frame}
-
-% \begin{figure}
-%   \includegraphics[width=3in]{ranganath.png}
-%   \caption{Reproduced from \cite{ranganath2014}.}
-% \end{figure}
 

### 2017-10-20 doc/glab_meeting_2017-10-20/gen_images.sh

@@ -1,0 +1,2 @@
+convert ../paper/fig04.pdf -gravity South -crop 100x50%+0+0 ./scaling.png
+convert ../paper/fig05.pdf -crop 100x50%+0+0 ./sign.png

### 2017-10-20 doc/glab_meeting_2017-10-20/vpd.tex
rough draft done, as is spellcheck

@@
 \addbibresource{references.bib}
+\usepackage{tikz}
 
@@
 
-\section{Results}
+\section{Results - theory}
 \begin{frame}{Scaling Term}
@@
 
+\section{Results - reality}
+{ % all template changes are local to this group.
+    \setbeamertemplate{navigation symbols}{}
+    \begin{frame}[plain]{Idealized results - CRO}
+        \begin{tikzpicture}[remember picture,overlay]
+            \node[at=(current page.center)] {
+                \includegraphics[width=\paperwidth]{cro_ideal.png}
+            };
+        \end{tikzpicture}
+     \end{frame}
+}
+
+{ % all template changes are local to this group.
+    \setbeamertemplate{navigation symbols}{}
+    \begin{frame}[plain]{Results with uncertainty - CRO}
+        \begin{tikzpicture}[remember picture,overlay]
+            \node[at=(current page.center)] {
+                \includegraphics[width=\paperwidth]{cro.png}
+            };
+        \end{tikzpicture}
+     \end{frame}
+}
+
+{ % all template changes are local to this group.
+    \setbeamertemplate{navigation symbols}{}
+    \begin{frame}[plain]{Idealized results - CSH}
+        \begin{tikzpicture}[remember picture,overlay]
+            \node[at=(current page.center)] {
+                \includegraphics[width=\paperwidth]{csh_ideal.png}
+            };
+        \end{tikzpicture}
+     \end{frame}
+}
+
+{ % all template changes are local to this group.
+    \setbeamertemplate{navigation symbols}{}
+    \begin{frame}[plain]{Results with uncertainty - CSH}
+        \begin{tikzpicture}[remember picture,overlay]
+            \node[at=(current page.center)] {
+                \includegraphics[width=\paperwidth]{csh.png}
+            };
+        \end{tikzpicture}
+     \end{frame}
+}
+
+{ % all template changes are local to this group.
+    \setbeamertemplate{navigation symbols}{}
+    \begin{frame}[plain]{Idealized results - DBF}
+        \begin{tikzpicture}[remember picture,overlay]
+            \node[at=(current page.center)] {
+                \includegraphics[width=\paperwidth]{dbf_ideal.png}
+            };
+        \end{tikzpicture}
+     \end{frame}
+}
+
+{ % all template changes are local to this group.
+    \setbeamertemplate{navigation symbols}{}
+    \begin{frame}[plain]{Results with uncertainty - DBF}
+        \begin{tikzpicture}[remember picture,overlay]
+            \node[at=(current page.center)] {
+                \includegraphics[width=\paperwidth]{dbf.png}
+            };
+        \end{tikzpicture}
+     \end{frame}
+   }
+
+{ % all template changes are local to this group.
+    \setbeamertemplate{navigation symbols}{}
+    \begin{frame}[plain]{Idealized results - ENF}
+        \begin{tikzpicture}[remember picture,overlay]
+            \node[at=(current page.center)] {
+                \includegraphics[width=\paperwidth]{enf_ideal.png}
+            };
+        \end{tikzpicture}
+     \end{frame}
+}
+
+{ % all template changes are local to this group.
+    \setbeamertemplate{navigation symbols}{}
+    \begin{frame}[plain]{Results with uncertainty - ENF}
+        \begin{tikzpicture}[remember picture,overlay]
+            \node[at=(current page.center)] {
+                \includegraphics[width=\paperwidth]{enf.png}
+            };
+        \end{tikzpicture}
+     \end{frame}
+   }
+
+{ % all template changes are local to this group.
+    \setbeamertemplate{navigation symbols}{}
+    \begin{frame}[plain]{Idealized results - GRA}
+        \begin{tikzpicture}[remember picture,overlay]
+            \node[at=(current page.center)] {
+                \includegraphics[width=\paperwidth]{gra_ideal.png}
+            };
+        \end{tikzpicture}
+     \end{frame}
+}
+
+{ % all template changes are local to this group.
+    \setbeamertemplate{navigation symbols}{}
+    \begin{frame}[plain]{Results with uncertainty - GRA}
+        \begin{tikzpicture}[remember picture,overlay]
+            \node[at=(current page.center)] {
+                \includegraphics[width=\paperwidth]{gra.png}
+            };
+        \end{tikzpicture}
+     \end{frame}
+   }
+
+   \begin{frame}{Summary statistics}
+     \begin{table}
+       \caption{Statistics of $\frac{\partial \; ET}{\partial \; D}$ as a function of PFT.}
+       \centering
+       \begin{tabular}{l c c}
+         \hline
+         PFT & $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ & fraction $\frac{\partial \; ET}{\partial \; VPD} < 0.$ \\
+         \hline
+         CRO & 0.000853  & 0.473311\\
+         CSH & -0.108234 & 0.931660\\
+         DBF & -0.012727 & 0.461674\\
+         ENF & -0.034087 & 0.534425\\
+         GRA & -0.019637 & 0.631735\\
+         \hline
+         \multicolumn{2}{l}{}  
+       \end{tabular}
+     \end{table}
+   \end{frame}
+
+\section{Conclusions}
+\begin{frame}{Summary}
+  \begin{itemize}
+  \item Theory suggests that the ``tipping'' point for which atmospheric demand overwhelms plant response will be almost exclusively a function of VPD.
+Plant types evolved to conserve water (CSH) have a higher VPD$_{crit}$ (and more negative ET response) than plants evolved (or bred) to use water and prioritize GPP (CRO).
+  \item On average, ecosystem response to VPD follows roughly what we might expect: CRO (prioritize GPP) has postive ET response to VPD, while all others have a negative response.
+Ordering by increasing magnitude of negative reponse gives: DBF, GRA, ENF, CSH; which might seem to correspond to expectations for increasing water conservation as a function of PFT..
+  \item Uncertainty is high, expecially for CRO and GRA. However, inclusion of uncertainty does not change the story for ENF or CSH.
+  \end{itemize}
+\end{frame}
+
 \section{References}

### 2017-10-20 doc/glab_meeting_2017-10-20/gen_images.sh

@@ -1,2 +1,51 @@
+rm -rf ./*.png
 convert ../paper/fig04.pdf -gravity South -crop 100x50%+0+0 ./scaling.png
 convert ../paper/fig05.pdf -crop 100x50%+0+0 ./sign.png
+convert ../paper/fig06.png -gravity NorthEast -crop 25x20%+0+0 ./temp_1.png
+convert ../paper/fig06b.png -gravity NorthEast -crop 25x20%+0+0 ./temp_2.png
+convert +append temp*.png cro_ideal.png
+rm -rf ./temp*.png
+convert ../paper/fig06.png -crop 25x20%+0+0 ./temp_1.png
+convert ../paper/fig06b.png -crop 25x20%+0+0 ./temp_2.png
+convert +append temp*.png cro.png
+rm -rf ./temp*.png
+
+# idealized plots
+convert ../paper/fig06.png -gravity East -crop 25x20%+0+0 ./temp_1.png
+convert ../paper/fig06b.png -gravity East -crop 25x20%+0+0 ./temp_2.png
+convert +append temp*.png dbf_ideal.png
+rm -rf ./temp*.png
+convert ../paper/fig06.png -gravity West -crop 25x20%+0+0 ./temp_1.png
+convert ../paper/fig06b.png -gravity West -crop 25x20%+0+0 ./temp_2.png
+convert +append temp*.png dbf.png
+rm -rf ./temp*.png
+
+
+# idealized plots
+convert ../paper/fig06.png -crop 640x480+0+480 ./temp_1.png
+convert ../paper/fig06b.png -crop 640x480+0+480 ./temp_2.png
+convert +append temp*.png csh.png
+rm -rf ./temp*.png
+convert ../paper/fig06.png -gravity NorthEast -crop 640x480+0+480 ./temp_1.png
+convert ../paper/fig06b.png -gravity NorthEast -crop 640x480+0+480 ./temp_2.png
+convert +append temp*.png csh_ideal.png
+rm -rf ./temp*.png
+
+
+convert ../paper/fig06.png -crop 640x480+0+1440 ./temp_1.png
+convert ../paper/fig06b.png -crop 640x480+0+1440 ./temp_2.png
+convert +append temp*.png enf.png
+rm -rf ./temp*.png
+convert ../paper/fig06.png -gravity NorthEast -crop 640x480+0+1440 ./temp_1.png
[…]

### 2017-10-20 doc/glab_meeting_2017-10-20/vpd.tex
what I presented at glab, see pierre's notes and ask dan about his thoughts on sap flow to diagnose hydraulic failure

@@
   \item \textbf{Hydrometeorologists} would say that an increase in VPD (increase in \textbf{atmospheric demand}) would drive an \textbf{increase in ET}.
-  \item However, \textbf{plant physioligists} know that plants have evolved to use stomata to conserve and regulate water use.
+  \item However, \textbf{plant physiologists} know that plants have evolved to use stomata to conserve and regulate water use.
 So \textbf{stomata closure} in response to increases in VPD may \textbf{decrease ET}.
@@
     \item However the environment can overwhelm plant response: at some threshold (i.e.
-very high VPD) the atmospheric demand should dominate and plants will not be able to consernve water, no matter how much they have evolved to do so.
+very high VPD) the atmospheric demand should dominate and plants will not be able to conserve water, no matter how much they have evolved to do so.
     \end{itemize}
@@
   \[ET = \frac{\Delta R + g_a \rho_a c_p D}{\Delta + \gamma(1 + \frac{g_a}{g_s})},\]
-  But in:
+  But in (\cite{MEDLYN_2011}):
   \[  g_{s} = g_0 + 1.6 \left(1 + \frac{g_1}{\sqrt{D}}\right) \frac{A}{c_s}\]
@@
   So we can use uWUE to remove dependence of $g_s$ on $A$:
-  \[g_s = \frac{\text{LAI }}{\text{LAI$_{ref}$}} \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{D_{s}}}\right) \frac{uWUE \; ET}{c_s \; \sqrt{D}}\]
+  \[g_s =  \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{D_{s}}}\right) \frac{uWUE \; ET}{c_s \; \sqrt{D}}\]
 \end{frame}
@@
   \begin{itemize}
-  \item Theory suggests that the ``tipping'' point for which atmospheric demand overwhelms plant response will be almost exclusively a function of VPD.
+  \item Theory finds that the ``tipping'' point for which atmospheric demand overwhelms plant response will be almost exclusively a function of VPD.
 Plant types evolved to conserve water (CSH) have a higher VPD$_{crit}$ (and more negative ET response) than plants evolved (or bred) to use water and prioritize GPP (CRO).
-  \item On average, ecosystem response to VPD follows roughly what we might expect: CRO (prioritize GPP) has postive ET response to VPD, while all others have a negative response.
-Ordering by increasing magnitude of negative reponse gives: DBF, GRA, ENF, CSH; which might seem to correspond to expectations for increasing water conservation as a function of PFT..
-  \item Uncertainty is high, expecially for CRO and GRA. However, inclusion of uncertainty does not change the story for ENF or CSH.
+  \item On average, ecosystem response to VPD follows roughly what we might expect: CRO (prioritize GPP) has positive ET response to VPD, while all others have a negative response.
+Ordering by increasing magnitude of negative response gives: DBF, GRA, ENF, CSH; which roughly correspond to expectations for increasing water conservation as a function of PFT.
+  \item Uncertainty is high, especially for CRO and GRA. However, inclusion of uncertainty does not change the story for ENF or CSH.
   \end{itemize}

### 2017-10-23 doc/paper/vpd_et_paper.tex
added glab presentation
did asymptptoc approximation and changed language with uncertainty parameter

@@
 
-This can be adapted to an ecosystem-scale stomatal resistance by multiplying by leaf area index (LAI) and converting units to m s$^{-1}$:
+This can be adapted to an ecosystem-scale stomatal resistance by multiplying by leaf area index (LAI) and converting units to m s$^{-1}$
+
 \begin{linenomath*}
@@
   \begin{equation}
-  g_s = \text{LAI } \frac{R \,T}{P} \left( g_0 + 1.6 \left(1 + \frac{g_1}{\sqrt{D_{s}}}\right) \frac{A}{c_s}\right)
+  g_s = \text{LAI} \frac{R \,T}{P} \left( g_0 + 1.6 \left(1 + \frac{g_1}{\sqrt{D_{s}}}\right) \frac{A}{c_s}\right)
   \end{equation}
@@
   \begin{equation}
+    \label{uwue}
 uWUE = \frac{GPP \cdot \sqrt{D}}{ET}
@@
   \begin{equation}
-  g_s = \frac{\text{LAI }}{\text{LAI$_{ref}$}} \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{D_{s}}}\right) \frac{uWUE \; ET}{c_s \; \sqrt{D}}
+  g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{D_{s}}}\right) \frac{\sigma \; uWUE \; ET}{c_s \; \sqrt{D}}
   \end{equation}
@@
 
-Note that $uWUE$ is fit on the ecosystem scale in \citet{Zhou_2015} so it already has an implicit LAI$_{ref}$ reflective of the average $LAI$ in \citet{Zhou_2015}'s fits.
-So the LAI term in Equation \ref{medlyn} becomes a term we call $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$, which is a measure of how close the observed LAI is to the LAI$_{ref}$ in \citet{Zhou_2015}.
+Note that $uWUE$ is fit on the ecosystem scale in \citet{Zhou_2015} so GPP in \ref{uwue} is really $A\cdot \text{LAI}$.
+We also recognize that our use of $uWUE$ introudces considerable uncertainty, so we added an uncertainty parameter $\sigma = f(t, \text{site})$ modifying $uWUE$.
+This will be elaborated on below.
 
@@
   \begin{equation}
-  ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p D_{s}}{R_{air}} - \frac{\text{LAI$_{ref}$}}{\text{LAI}}  \frac{\gamma c_s \sqrt{D} }{ R* \; 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{D}})} \right) }{ \Delta + \gamma}
+  ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p D_{s}}{R_{air}} -  \frac{\gamma c_s \sqrt{D} }{ R* \; 1.6\; \sigma\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{D}})} \right) }{ \Delta + \gamma}
   \end{equation}
@@
   \begin{equation}
-    \frac{\partial \;  ET}{\partial \; D} = \frac{g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} - \frac{\text{LAI$_{ref}$}}{\text{LAI}} \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right) \right)
+    \frac{\partial \;  ET}{\partial \; D} = \frac{g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_s }{1.6 \; R*\; \sigma \; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right) \right)
     \label{d_et}
@@
 \end{linenomath*}
-Note that given yearly uWUE from \citet{Zhou_2015}, $g_1$ from \citet{Lin_2015} \citep[as presented in ][]{Franks_2017}, and observations of R, T, P, D$_s$, and wind speed (WS), the only unknown is $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$.
-With flux tower observations of ET, $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ will then be uniquely determined for each observation through Equation 6:
+
+The $D$ dependence in Equation \ref{d_et} is a little opaque.
+However, mean $D$ is 1062 Pa, so $\sqrt{D}$ is 32.6 Pa$^{1/2}$, which is much less than $g_1$ (with the exception of ENF; Table \ref{pft}).
+So a series expansion in the limit $\frac{\sqrt{D}}{g_1} \to 0$ gives an approximation which makes the functional form more clear:
 
@@
   \begin{equation}
-    \frac{\text{LAI}}{\text{LAI$_{ref}$}}  = - \frac{g_a \gamma c_s \sqrt{D_s} P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p D_{s}\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{D_s}})}
+    \frac{\partial \;  ET}{\partial \; D} \approx \frac{g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_s }{1.6 \; R*\; \sigma \; \text{ uWUE }} \left( \frac{1}{g_{1}} - \frac{3 \sqrt{D}}{2 g_{1}^{2}} + \frac{2 \sqrt{D}^{2}}{g_{1}^{3}} - \frac{5 \sqrt{D}^{3}}{2 g_{1}^{4}} + \mathcal{O}\left(\left(\frac{\sqrt{D}}{g_1}\right)^{4}\right) \right) \right)
+    \label{d_et_approx}
+  \end{equation}
+\end{linenomath*}
+
+Note that given yearly uWUE from \citet{Zhou_2015}, $g_1$ from \citet{Lin_2015} \citep[as presented in ][]{Franks_2017}, and observations of R, T, P, D$_s$, and wind speed (WS), all quantities in Equations \ref{d_et} and \ref{d_et_approx} are known except for the uncertainty parameter $\sigma$.
+With flux tower observations of ET, $\sigma$ will then be uniquely determined for each observation through Equation 6:
+
+\begin{linenomath*}
+  \begin{equation}
+    \sigma = -\frac{g_a \gamma c_s \sqrt{D_s} P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p D_{s}\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{D_s}})}
     \label{lai}
@@
 %% see registers!!!!!! - help for emacs productivity
-This $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ is some part a ``true'' deviation of  LAI from LAI$_{ref}$ in \citet{Zhou_2015}, and some part model and observational error, including assumptions about constant $g1$ and $uWUE$.
-By calculating a unique $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ for each observation we will propagate any model and observational uncertainty forward into our expression for $\frac{\partial \; ET}{\partial D}$.
+This $\sigma$ will be some part uncertainty associated with assumptions about constant $g1$ and $uWUE$, including differences between the data used here and in the fits in \cite{Zhou_2016} and \cite{Lin_2015}, but also some general PM model and FLUXNET observationsal uncertainty.
+By calculating a unique $\sigma$ for each observation we will propagate any model and observational uncertainty forward into our expression for $\frac{\partial \; ET}{\partial D}$.
 
@@
   $LAI$ & leaf area index & -\\
-  $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ & ratio of LAI to LAI$_{ref}$ in \citet{Zhou_2015} & -\\
+  $\sigma$ & ratio of LAI to LAI$_{ref}$ in \citet{Zhou_2015} & -\\
   $c_s$ & CO$_2$ concentration & $\mu$ mol CO$_2$ mol$^{-1}$ air\\
@@
 We use data from FLUXNET2015.
-Because $g_1$ coefficients \citep{Lin_2015} and uWUE were only both available for five plant functional types (PFTs - see Table 2),  only 56 of the 77 sites were used.
+Because $g_1$ coefficients \citep{Lin_2015} and uWUE were only both available for five plant functional types (PFTs - see Table \ref{pft}),  only 56 of the 77 sites were used.
 Figure \ref{map_fig} \explain{map needs to be improved - it's a placeholder for now} presents each site and its plant functional type.
@@
 \begin{table}
+  \label{pft}
 \caption{Plant functional types, their abbreviation, Medlyn coefficient \citep[from ][]{Lin_2015}, and uWfUE \citep[from ][]{Zhou_2015}.
@@
 
-By construction, the variability in the $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ term (Equation \ref{lai}) contains all model and observational uncertainties.
-We expect LAI to be cloase to LAI$_{ref}$, so $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ should be $O(1)$.
-We can have some confidence in our framework, including the assumption of constant uWUE, if calculated $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$s are generally $O(1)$.
-Figure \ref{lai_fig} presents the histogram of calculated $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$s with outliers (lowest and highest 5\% percent) and nonphysical values ($\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ $<$ 0.) removed.
-All remaining $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ values are $O(1)$ which provides confidence in model framework.
+By construction, the variability in the $\sigma$ term (Equation \ref{lai}) contains all model and observational uncertainties.
+We expect LAI to be cloase to LAI$_{ref}$, so $\sigma$ should be $O(1)$.
+We can have some confidence in our framework, including the assumption of constant uWUE, if calculated $\sigma$s are generally $O(1)$.
+Figure \ref{lai_fig} presents the histogram of calculated $\sigma$s with outliers (lowest and highest 5\% percent) and nonphysical values ($\sigma$ $<$ 0.) removed.
+All remaining $\sigma$ values are $O(1)$ which provides confidence in model framework.
 
@@
 \includegraphics[width=20pc]{./fig02.png}
-\caption{Histogram of $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ values calculated for each site and time according to Equation \ref{lai}.
+\caption{Histogram of $\sigma$ values calculated for each site and time according to Equation \ref{lai}.
 The lowest and highest 5\% are removed as outliers, as well as any values below 0.
@@
 
-An additional concern is that the $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ term may in fact be some function of $D$, in which case the dependence would need to be accounted for when taking the derivative.
-Figure \ref{lai_vpd_fig}\explain{Figs 2 and 3 can probably be combined - the histogram of $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ is shown in Fig 3}.
-plots the joint distribution of $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ and VPD, and shows that $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ is very weakly a function of VPD.
+An additional concern is that the $\sigma$ term may in fact be some function of $D$, in which case the dependence would need to be accounted for when taking the derivative.
+Figure \ref{lai_vpd_fig}\explain{Figs 2 and 3 can probably be combined - the histogram of $\sigma$ is shown in Fig 3}.
+plots the joint distribution of $\sigma$ and VPD, and shows that $\sigma$ is very weakly a function of VPD.
 Given this weak dependence, we argue that Equation \ref{d_et} is a valid approximation for ET response to $D$.
@@
 \includegraphics[width=20pc]{./fig03.png}
-\caption{The joint distribution of $D$ and $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$.
-$\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ has only a weak dependence on $D$.
+\caption{The joint distribution of $D$ and $\sigma$.
+$\sigma$ has only a weak dependence on $D$.
 ***This plot could proabably benefit from a box plot of site specific correlations, because some sites do have stronger depednence than others.
-Note also Figs 3 and 2 can probably be combined because this figure shoes $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$'s histogram.***}
+Note also Figs 3 and 2 can probably be combined because this figure shoes $\sigma$'s histogram.***}
 \label{lai_vpd_fig}
@@
 Before calculating the sensitivity of ET to VPDn, it is useful to consider the functional form of Equation \ref{d_et}.
-There are three terms: a scaling term for the full expression we will call Term 1 ($\frac{g_a \; P}{T(\Delta + \gamma)}$), a relatively constant offset we will call Term 2 ($\frac{c_p}{R_{air}}$), and a variable term we will call Term 3 ($\frac{\text{LAI$_{ref}$}}{\text{LAI}}\frac{\gamma c_s }{1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right)$).
+There are three terms: a scaling term for the full expression we will call Term 1 ($\frac{g_a \; P}{T(\Delta + \gamma)}$), a relatively constant offset we will call Term 2 ($\frac{c_p}{R_{air}}$), and a variable term we will call Term 3 ($\frac{\gamma c_s }{1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right)$).
 All variables are positive, so the relative magnitude between Term 2 and Term 3 will determine the sign of the derivative, while Term 1 will scale the expression larger or smaller.
@@
 Term 2 minus Term 3 determines the sign of the sensitivty of ET to VPD and is thus crucial.
-$c_s$ variability is relatively less than $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ and $D$ variability, then variability within PFT will be solely determined by $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ and $D$ at the different fluxnet sites.
-Figure \ref{term3} shows how (Term 2 - Term 3) varies with $D$ and $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$, as a function of PFT.
-In Figure \ref{term3}a lower uWUE and $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ shift the distribution of (Term 2 - Term 3) towards negative values.
+$c_s$ variability is relatively less than $\sigma$ and $D$ variability, then variability within PFT will be solely determined by $\sigma$ and $D$ at the different fluxnet sites.
+Figure \ref{term3} shows how (Term 2 - Term 3) varies with $D$ and $\sigma$, as a function of PFT.
+In Figure \ref{term3}a lower uWUE and $\sigma$ shift the distribution of (Term 2 - Term 3) towards negative values.
 Additionally, the smaller $g1$, the greater the relative $D$ dependence of of (Term 2 - Term 3).
@@
 \caption{Sources of variability for Term 2 - Term 3.
-Top: Term 2 - Term 3 as a function of VPD, with $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ held fixed at PFT averages.
+Top: Term 2 - Term 3 as a function of VPD, with $\sigma$ held fixed at PFT averages.
 The observed range of VPD for each PFT is also shown below the x-axis.
@@
 
-Bottom: The location of the minima of ET, as a function of VPD and $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$.
-Lines and stars denote the distribution of VPD and $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ next to each axis, following the same percentiles as above.}
+Bottom: The location of the minima of ET, as a function of VPD and $\sigma$. Lines and stars denote the distribution of VPD and $\sigma$ next to each axis, following the same percentiles as above.}
 \label{term3}
@@
 
-Figure \ref{term3}b shows the location of the minima of ET, as a function of $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ and $D$.
-For any $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ or VPD less (more) than these curves, Term 2 - Term 3 will be negative (positive).
-It is clear that the portion of VPD observations below/above these curves will be a strong function of $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$.
+Figure \ref{term3}b shows the location of the minima of ET, as a function of $\sigma$ and $D$.
+For any $\sigma$ or VPD less (more) than these curves, Term 2 - Term 3 will be negative (positive).
+It is clear that the portion of VPD observations below/above these curves will be a strong function of $\sigma$.
 However, we can see some general trends.
-For CSH, $\frac{\partial \; ET}{\partial \; D}$ should be negative for the vast majority of observed $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ and VPD.
+For CSH, $\frac{\partial \; ET}{\partial \; D}$ should be negative for the vast majority of observed $\sigma$ and VPD.
 The fraciton of positive $\frac{\partial \; ET}{\partial \; D}$ appears to be more even for ENF, GRA, and DBF, and we might expect a greater frequency of positive $\frac{\partial \; ET}{\partial \; D}$ for CRO.
@@
 However, $\frac{\partial \; ET}{\partial \; D}$ evaluated at the average of all variables (e.g.
-$\frac{\text{LAI}}{\text{LAI$_{ref}$}}$, $T$, $c_s$, $D$) is only negative for CSH and GRA.
+$\sigma$, $T$, $c_s$, $D$) is only negative for CSH and GRA.
 And, DBF in addition to CRO experiences $\frac{\partial \; ET}{\partial \; D}$ < 0 less than half the time.
@@
 So far, idealized plots and statistics have illuminated the form of $\frac{\partial \; ET}{\partial \; D}$ and how it varies with PFT.
-Large mean $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ and uWUE shifts CRO and DBF towards positive $\frac{\partial \; ET}{\partial \; D}$.
+Large mean $\sigma$ and uWUE shifts CRO and DBF towards positive $\frac{\partial \; ET}{\partial \; D}$.
 However, the strongly nonlinear function of $\frac{\partial \; ET}{\partial \; D}$ at $\frac{\partial \; ET}{\partial \; D} < 0$ pushes $\overline{\frac{\partial \; ET}{\partial \; D}}$ negative for DBF (it does not do this for CRO because of CRO's high $g1$).
@@
 CRO has the most positive values of $\frac{\partial \; ET}{\partial \; D}$, $\frac{\partial \; ET}{\partial \; D}$ is almost always negative for CSH, and response depends more with the environmental conditions for the other PFTs (especially ENF).
-Through the columns of Figure \ref{real} we can see the impact of $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ and $g_a$ on the variability of $\frac{\partial \; ET}{\partial \; D}$.
+Through the columns of Figure \ref{real} we can see the impact of $\sigma$ and $g_a$ on the variability of $\frac{\partial \; ET}{\partial \; D}$.
 $g_a$'s scaling (included in columns 1 and 3) alters the magnitude considerably.
-$\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ variability (included in columns 1 and 2) adds a lot of additional noise to the signal of $\frac{\partial \; ET}{\partial \; D}$, which is slightly undesirable given $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$'s role in representing model and observational uncertainty.
-However, the general story with the noise appears to match the cleaner signal when $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ is help constant and $D_{ETmin}$ is clearly visible \explain{I really need to make these plots better - way too much overlapping of points that hurts the story}.
-One exception is possibly with GRA, for which uncertainty represented in $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ is high and causes the full complexity plots (Columns 1 and 2) to not match well with $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ held fixed (Columns 3 and 4).
+$\sigma$ variability (included in columns 1 and 2) adds a lot of additional noise to the signal of $\frac{\partial \; ET}{\partial \; D}$, which is slightly undesirable given $\sigma$'s role in representing model and observational uncertainty.
+However, the general story with the noise appears to match the cleaner signal when $\sigma$ is help constant and $D_{ETmin}$ is clearly visible \explain{I really need to make these plots better - way too much overlapping of points that hurts the story}.
+One exception is possibly with GRA, for which uncertainty represented in $\sigma$ is high and causes the full complexity plots (Columns 1 and 2) to not match well with $\sigma$ held fixed (Columns 3 and 4).
 
-For ENF and GRA $D_{ETmin}$ does not appear to be only a function of $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ (most observable in Column 4).
+For ENF and GRA $D_{ETmin}$ does not appear to be only a function of $\sigma$ (most observable in Column 4).
 It turns out that the site to site variability in $\gamma$ causes $D_{ETmin}$ to vary, which is not discussed in the previous section.
@@
 
-In general the full complexity plots of $\frac{\partial \; ET}{\partial \; D}$ match our expectations, even with the large sensitivity to $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ measures of uncertainty observed in Figure \ref{term3}.
-Our $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$-based method of uncertainty propagation blurs the idealized expectations the most for GRA, and also has a considerable effect for CRO.
-We therefor have the most confidence in our conclusion based on Equation \ref{d_et} for PFTS CSH, DBF, and ENF, as the full complexity plots with uncertainty included closely match the story when $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ is held fixed.
+In general the full complexity plots of $\frac{\partial \; ET}{\partial \; D}$ match our expectations, even with the large sensitivity to $\sigma$ measures of uncertainty observed in Figure \ref{term3}.
+Our $\sigma$-based method of uncertainty propagation blurs the idealized expectations the most for GRA, and also has a considerable effect for CRO.
+We therefor have the most confidence in our conclusion based on Equation \ref{d_et} for PFTS CSH, DBF, and ENF, as the full complexity plots with uncertainty included closely match the story when $\sigma$ is held fixed.
 **see somewhat preferred alternate figure \ref{real2} \explain{I think I like the alternate plot much  more as thinking in terms of T and RH is easier, and it makes the story easier to see at relatively low temperatures.
@@
 \caption{Scatter plots of $\frac{\partial \; ET}{\partial \; D}$.
-Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; D}$, as labeled: Column 1 - $\frac{\partial \; ET}{\partial \; D}$; Column 2 - $\frac{\partial \; ET}{\partial \; D}$ normalized by $g_a$; Column 3 - $\frac{\partial \; ET}{\partial \; D}$ with $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ held fixed at PFT average; and Column 4 - $\frac{\partial \; ET}{\partial \; D}$ normalized by $g_a$ and with $\frac{\text{LAI}}{\text{LAI$_{ref}$}}$ held fixed.
+Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; D}$, as labeled: Column 1 - $\frac{\partial \; ET}{\partial \; D}$; Column 2 - $\frac{\partial \; ET}{\partial \; D}$ normalized by $g_a$; Column 3 - $\frac{\partial \; ET}{\partial \; D}$ with $\sigma$ held fixed at PFT average; and Column 4 - $\frac{\partial \; ET}{\partial \; D}$ normalized by $g_a$ and with $\sigma$ held fixed.
 For reference, lines corresponding to RH = 20\% and RH = 90 \% are drawn.
@@
 Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; D}$, as labeled.
-If I end up using this, I could also draw on the curve of $D_{ETmin}$ with $\overline{$\frac{\text{LAI}}{\text{LAI$_{ref}$}}$}$.
+If I end up using this, I could also draw on the curve of $D_{ETmin}$ with $\overline{\frac{\text{LAI}}{\text{LAI$_{ref}$}}}$.
 }

### 2017-10-23 src/make_paper_plots.py

@@ -15,7 +15,17 @@ import util
 import metcalcs as met
 import codebase.penman_monteith as pm
 import codebase.calc_tools as calc
+from sympy import Symbol, sqrt, series, latex, limit
 
+
+#try some analytis with sympy
+g1 = Symbol('g_1')
+# x = Symbol('\frac{\sqrt{D}}{g_1}')
+# func = g1*(2 + x)/(2*g1**2*(1 + x)**2)
+# print(latex(series(func, x, x0=0., dir='+')))
+x = Symbol('\sqrt{D}')
+func  = (2*g1 + x)/(2*(g1 + x)**2)
+print(latex(series(func, x, x0=0., dir='+')))
 
 mpl.use('Pdf')
 mpl.rcParams.update(mpl.rcParamsDefault)
@@ -261,6 +271,25 @@ def term_2(_df, lai, vpd):
   canopy = {'uwue' : _df.uwue.mean(), 'g1' : _df.g1.mean()}
   return pm.CP/_df.r_moist.mean() +calc.leaf_vpd(atmos, canopy, lai)
 
+
+def term_2_approx(_df, lai, vpd):
+  """calculates term 2"""
+  atmos = {'gamma' : _df.gamma.mean(), 'c_a' : _df.c_a.mean(),\
+           'vpd' : vpd}
+  if _df.uwue.std() > 1.e-8:
+    print('error, uWUE is variable: %f!!!!' % _df.uwue.std())
+  elif _df.g1.std() > 1.e-8:
+    print('error, g1 is variabile: %f!!!!!' % _df.g1.std())
+  canopy = {'uwue' : _df.uwue.mean(), 'g1' : _df.g1.mean()}
+  return pm.CP/_df.r_moist.mean() \
+    -atmos['gamma']*atmos['c_a']*pm.LV/\
+    (lai*1.6*pm.R_STAR*canopy['uwue'])\
+    *(1./canopy['g1'] - 3.*np.sqrt(atmos['vpd'])/(2.*canopy['g1']**2)
+      + 2.*atmos['vpd']/canopy['g1']**3\
+      - 5.*np.sqrt(atmos['vpd'])**3/(2.*canopy['g1']**4))
+
+
+
 # def plot_leaf_vpd(_df, ax, savefig=False):
 #   """makes idealized plots of plant term as a function of lai and vpd"""
[…]

### 2017-10-23 doc/paper/vpd_et_paper.tex
some writing changse

@@
 Traditionally, atmospheric scientists and hydrometeorologists generally think that an increase in atmospheric demand induces an increases in evapotranspiration (ET) (citations?).
-This possible misconception developed in part due to the proliferation of studies examing potential ET rather than estimates of ET itself (citations?).
+This possible misconception developed in part due to the proliferation of studies examing potential ET (PET) rather than estimates of ET itself (citations?).
 In contrast, plant physiologists know that stomata have evolved to optimally regulate the exchange of water and carbon, and tend to close in response to increased atmospheric dryness \citep{Ball_1987, Leuning_1990, MEDLYN_2011}.
@@
 Clearly, the sign and magnitude of land-surface responsedrives the co-evolution of the atmosphere and land-surface at many timescales, from diurna to interdecadal.
+
+We hypothesize that for most plant types a common resonse to increase in VPD will actually be a decrease in ET.
+The excpetion would be plants such as crops that are evolved (or bred) to prioritize gross primary production (GPP) over water conservation.
+However, for all otehr plants types, our hypothesis calls into question the validity of PET-based drought metrics deceloped by hydrometeorologists and used extensively in operations \citep[e.g.
+PDSI, P-PET,][]{2002, Alley_1984} \explain{more citations needed, including recent PET climate studies like Jack Scheff}.
+These metrics ignore the role of plants as gatekeepers for surface water loss to the atmosphere and have limited physical meaning for drought of vegetated land types.
+Additionally, plants evolved in arid climates should prioritize water conservation and we would expect a very negative ET response to increase in VPD.
+Therefore, vegetated locations most likely to experience droughts should show the strongest deviation between reality and a PET-based approximation.
 
@@
 Note that $uWUE$ is fit on the ecosystem scale in \citet{Zhou_2015} so GPP in \ref{uwue} is really $A\cdot \text{LAI}$.
+This leads to the cancelation of LAI in additoin to uWUE in Equation \ref{medlyn}.
 We also recognize that our use of $uWUE$ introudces considerable uncertainty, so we added an uncertainty parameter $\sigma = f(t, \text{site})$ modifying $uWUE$.
-This will be elaborated on below.
+The meaning of this parameter will be elaborated on after the derivation of $\frac{\partial \; ET}{\partial \; D}$.
 
@@
     ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p D_{s}}{R_{air}} -  \frac{\gamma c_s \sqrt{D} }{ R* \; 1.6\; \sigma\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{D}})} \right) }{ \Delta + \gamma}
+    \label{et}
   \end{equation}
@@
 
-We can then take the derivative with respect to $D$ to analytically determine ecosystem response to atmospheric demand perturbations:
+Note that if we set the uncertaitny parameter $\sigma$ to 1, and approximage $c_s$ at a global mean CO$_2$ concentration, then Equation \ref{et} is fully defined using commonly available weather station data and the constants published in \citet{Zhou_2015, Lin_2015}.
+This then begs the question, why use PET at all for drought metrics in vegetated areas?
+A much more physically realistic estimate of ET can be had with the same information reqiured to calculate PET.
+
+We can now take the derivative with respect to $D$ to analytically determine ecosystem response to atmospheric demand perturbations:
 
@@
 
-%% see registers!!!!!! - help for emacs productivity
-This $\sigma$ will be some part uncertainty associated with assumptions about constant $g1$ and $uWUE$, including differences between the data used here and in the fits in \cite{Zhou_2016} and \cite{Lin_2015}, but also some general PM model and FLUXNET observationsal uncertainty.
+This $\sigma$ will be some part uncertainty associated with assumptions about constant $g1$ and $uWUE$, including differences between the data used here and in the fits in \citet{Zhou_2015} and \citet{Lin_2015}, but also some part general PM model and FLUXNET observationsal uncertainty.
 By calculating a unique $\sigma$ for each observation we will propagate any model and observational uncertainty forward into our expression for $\frac{\partial \; ET}{\partial D}$.
@@
 By construction, the variability in the $\sigma$ term (Equation \ref{lai}) contains all model and observational uncertainties.
-We expect LAI to be cloase to LAI$_{ref}$, so $\sigma$ should be $O(1)$.
-We can have some confidence in our framework, including the assumption of constant uWUE, if calculated $\sigma$s are generally $O(1)$.
+For an observation that perfectly matches our model and assumptions $\sigma$ will be one.
+Therefore, if for our assumptions and framework to be reasonable $\sigma$ should be $O(1)$.
 Figure \ref{lai_fig} presents the histogram of calculated $\sigma$s with outliers (lowest and highest 5\% percent) and nonphysical values ($\sigma$ $<$ 0.) removed.
@@
 An additional concern is that the $\sigma$ term may in fact be some function of $D$, in which case the dependence would need to be accounted for when taking the derivative.
-Figure \ref{lai_vpd_fig}\explain{Figs 2 and 3 can probably be combined - the histogram of $\sigma$ is shown in Fig 3}.
-plots the joint distribution of $\sigma$ and VPD, and shows that $\sigma$ is very weakly a function of VPD.
+Figure \ref{lai_vpd_fig}\expplain{Figs 2 and 3 can probably be combined - the histogram of $\sigma$ is shown in Fig 3.} plots the joint distribution of $\sigma$ and VPD, and shows that $\sigma$ is very weakly a function of VPD.
 Given this weak dependence, we argue that Equation \ref{d_et} is a valid approximation for ET response to $D$.
@@
 
-Before calculating the sensitivity of ET to VPDn, it is useful to consider the functional form of Equation \ref{d_et}.
+Before calculating the sensitivity of ET to VPD, it is useful to consider the functional form of Equation \ref{d_et}.
 There are three terms: a scaling term for the full expression we will call Term 1 ($\frac{g_a \; P}{T(\Delta + \gamma)}$), a relatively constant offset we will call Term 2 ($\frac{c_p}{R_{air}}$), and a variable term we will call Term 3 ($\frac{\gamma c_s }{1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right)$).
@@
 Additionally, the impact of $T$ on the relative variability increases with increasing $g_a$.
-% plot every PFT and show they collapse onto the same vurve?
+\explain{plot every PFT and show they collapse onto the same curve?
+- might just be too messy though}.
 
@@
 Term 2 minus Term 3 determines the sign of the sensitivty of ET to VPD and is thus crucial.
-$c_s$ variability is relatively less than $\sigma$ and $D$ variability, then variability within PFT will be solely determined by $\sigma$ and $D$ at the different fluxnet sites.
+$c_s$ variability is relatively less than $\sigma$ and $D$ variability, so variability within PFT will be solely determined by $\sigma$ and $D$ at the different fluxnet sites.
 Figure \ref{term3} shows how (Term 2 - Term 3) varies with $D$ and $\sigma$, as a function of PFT.
-In Figure \ref{term3}a lower uWUE and $\sigma$ shift the distribution of (Term 2 - Term 3) towards negative values.
-Additionally, the smaller $g1$, the greater the relative $D$ dependence of of (Term 2 - Term 3).
-This is observed most strongly for the ENF PFT, which has the smallest $g1$ (74.31).
+Equation \ref{d_et_approx} aids interpretation of Figure \ref{term3}.
+Larger $uWUE$, $\overline{\sigma}$, and $g1$ shift the leading-order constant term (\frac{1}{g_1}) towards smaller values, and (Term 2 - Term 3) towards positive values.
+$uWUE$ and $g1$ are both water-use effiency type constants.
+Higher values corresponde to plants that are more willing to spend water on primary production and less evolved to conserve water.
+Figure \ref{term3} confirms our physical intuition: CROs are the least water conservative so have the smallest constant portion of Term 3, while CSH are the most water conservative and have the largest constant portion of Term 3.
+For the VPD-dependent terms in Equation \ref{d_et_approx}, differences in $g1$ between PFTs exerts a greater influence than difference $uWUE$, as the power of $g1$ increases.
+Increasing $uWUE$ and $g1$ decreases the VPD-dependence, but $g1$ has the bigger effect due to its increasing powers.
+ENF ($g1 = 74.31$) has by far the VPD dependence of response, while CRO ($g1 = 183.1$) has the smallest VPD dependence.
 
@@
 The intuition developed using this framework can be used to understand how the land surface will respond and contribute to changes in the environment.
+Additionally, Equation \ref{et} gives provides an estimate of ET that requires no additional information beyond that required to calculate PET.
+Given that for all PFTs, with the exception of CRO, we found a high frequency and magnitude of negative $\frac{\partial \; ET}{\partial \; D}$, PET is a physically unrealisitic representation of ET for vegetated surfaces and PET-based drought metrics are not usefull.
+We advocate for drought metrics using Equation \ref{et} instead of PET.
 
@@
 % AGU does not want a .bib or a .bbl file. Please copy in the contents of your .bbl file here.
-
+\bibliography{references.bib}
 \input{vpd_et_paper.bbl}
-%\bibliography{references.bib}
 

### 2017-10-23 doc/paper/vpd_et_paper.tex
took a closer look at PM and added some more intuitive languge on the scaling term

@@
 $R$  & net radiation at land surface minus ground heat flux & W m$^{-2}$   \\
-  $g_a$  & atmospheric conductance & m s$^{-1}$  \\
+  $g_a$  & aerodynamic conductance & m s$^{-1}$  \\
   $\rho_a$  & air density & kg m$^{-3}$  \\
@@
   $LAI$ & leaf area index & -\\
-  $\sigma$ & ratio of LAI to LAI$_{ref}$ in \citet{Zhou_2015} & -\\
+  $\sigma$ & uncertainty parameter & -\\
   $c_s$ & CO$_2$ concentration & $\mu$ mol CO$_2$ mol$^{-1}$ air\\
@@
 An additional concern is that the $\sigma$ term may in fact be some function of $D$, in which case the dependence would need to be accounted for when taking the derivative.
-Figure \ref{lai_vpd_fig}\expplain{Figs 2 and 3 can probably be combined - the histogram of $\sigma$ is shown in Fig 3.} plots the joint distribution of $\sigma$ and VPD, and shows that $\sigma$ is very weakly a function of VPD.
+Figure \ref{lai_vpd_fig}\explain{Figs 2 and 3 can probably be combined - the histogram of $\sigma$ is shown in Fig 3.} plots the joint distribution of $\sigma$ and VPD, and shows that $\sigma$ is very weakly a function of VPD.
 Given this weak dependence, we argue that Equation \ref{d_et} is a valid approximation for ET response to $D$.
@@
 
-In Term 1, $\frac{P}{T} \propto \rho$, so this should vary little relative to $g_a$ and $\Delta$.
-$\gamma$ should also be relatively constant, So the scaling term, Term 1, should be primarily a function of $g_a$ and temperature (through the function $\Delta$).
-While temperature range may vary for PFT, the functional form of $\Delta$ will be the same.
-$g_a$ will vary strongly with PFT due to the importance of surface roughness.
-However, the coefficient of variability for both $g_a$ and Term 1 is relatively constant across PFT, suggesting that the influence of $g_a$ on the relative (to the mean) variability of Term 1 is approximately similar across PFT.
+In Term 1, $\frac{P}{T} \propto \rho$, so this should vary little relative to aerodynamic conductance and $\Delta$.
+$\gamma$ should also be relatively constant, so the scaling term, Term 1, should be primarily a function of aerodynamic conductance and temperature (through the function $\Delta$).
+This makes sense, as aerodynamic conductance represents how efficently response at the surface is communicated to the atmosphere.
+As it increases, any plant response will be communicated more strongly to the atmosphere (and vice-versa).
 
-Figure \ref{scale_vary}A shows Term 1 normalized by mean $g_a$ (calculated for each plant functional type), and confirms that much of the relative variability of Term 1 is contained in the $g_a$ term's relative variability.
+$\Delta$'s presence in the scaling term also matches physical intition.
+Evaporative cooling will will dampen the ability of the atmosphere to take more moisture, as $e_{sat}$ decreases with temperature..
+The decrease in $e_{sat}$ is proportional to $\Delta$ ($\delta e_{sat} = \Delta \delta T$).
+So as $\Delta$ increases, you will get a larger damping of ET due to evaporative cooling.
+The functional from of $\Delta$ will be the same across PFT, but the temperature range may vary slightly.
+In contrast, aerodynamic conductance will vary strongly with PFT due to the importance of surface roughness.
+So most of the differences in scaling between PFT should be in the aerodynamic conductance term.
+One interesting side note is that the coefficient of variability for both aerodynamic conductance and Term 1 is relatively constant across PFT, suggesting that the influence of aerodynamic conductance on the relative (to the PFT mean) variability of Term 1 is approximately similar across PFT.
+
+Figure \ref{scale_vary}A shows Term 1 normalized by mean aerodynamic conductance (calculated for each plant functional type), and confirms that much of the relative variability of Term 1 is contained in the aerodynamic conductance variability.
 Generally, $T$ has less of a role.
-Additionally, the impact of $T$ on the relative variability increases with increasing $g_a$.
+Additionally, the impact of $T$ on the relative variability increases with increasing aerodynamic conductance.
 \explain{plot every PFT and show they collapse onto the same curve?
@@
 While the relative variability of Term 1 is similar across PFT, the absolute value of Term 1 varies strongly across PFT.
-Figure \ref{scale_vary}B shows Term 1 evaluated with the mean $g_a$ for each PFT, and at the range of observed temperatures for each PFT.
+Figure \ref{scale_vary}B shows Term 1 evaluated with the mean aerodynamic conductance for each PFT, and at the range of observed temperatures for each PFT.
 As expected, for the tree PFTs (DBF, ENF) Term 1 is much larger and the temperature dependence is much stronger.
@@
 For example, ENF experiences on average colder temperatures and is thus more likely to have a larger scaling term.
-Additionally, because std$(g_a) \propto \overline{g_a}$, the spread of Term 1 due to $g_a$ variability will be larger for the tree PFTs, although this is not shown for simplicity.
-To summarize, the variability of Term 1 within each PFT will look like Figure \ref{scale_vary}A for each PFT, but the scale of the x and y-axis will increase or decrease according to mean $g_a$ oberved in Figure \ref{scale_vary}B.
+Additionally, because the variability of aerodynamic conductance increases proportionally to the mean, the spread of Term 1 due to aerodynamic conductance variability will be larger for the tree PFTs, although this is not shown for simplicity.
+To summarize, the variability of Term 1 within each PFT will look like Figure \ref{scale_vary}A for each PFT, but the scale of the y-axis will increase or decrease according to mean aerodynamic conductance oberved in Figure \ref{scale_vary}B.
 
@@
 Equation \ref{d_et_approx} aids interpretation of Figure \ref{term3}.
-Larger $uWUE$, $\overline{\sigma}$, and $g1$ shift the leading-order constant term (\frac{1}{g_1}) towards smaller values, and (Term 2 - Term 3) towards positive values.
+Larger $uWUE$, $\overline{\sigma}$, and $g1$ shift the leading-order constant term ($\frac{1}{g_1}$) towards smaller values, and (Term 2 - Term 3) towards positive values.
 $uWUE$ and $g1$ are both water-use effiency type constants.
@@
 Figure \ref{term3} confirms our physical intuition: CROs are the least water conservative so have the smallest constant portion of Term 3, while CSH are the most water conservative and have the largest constant portion of Term 3.
-For the VPD-dependent terms in Equation \ref{d_et_approx}, differences in $g1$ between PFTs exerts a greater influence than difference $uWUE$, as the power of $g1$ increases.
+For the VPD-dependent terms in Equation \ref{d_et_approx}, differences in $g1$ between PFTs exert a greater influence than difference $uWUE$, as the power of $g1$ increases.
 Increasing $uWUE$ and $g1$ decreases the VPD-dependence, but $g1$ has the bigger effect due to its increasing powers.
-ENF ($g1 = 74.31$) has by far the VPD dependence of response, while CRO ($g1 = 183.1$) has the smallest VPD dependence.
+ENF ($g1 = 74.31$) has by far the largest VPD dependence of response, while CRO ($g1 = 183.1$) has the smallest VPD dependence.
 
@@
 This has the side effect of pushing $\overline{\frac{\partial \; ET}{\partial \; D}}$ negative further than other PFTs for a given fraction $\frac{\partial \; ET}{\partial \; D} < 0$ and magnitude $\frac{\partial \; ET}{\partial \; D}(\overline{T,\ldots,D})$.
-GRA shows the opposite behavior; a relatively high $g1$  makes the function more linear, decreasing the magnitude of $-\overline{\frac{\partial \; ET}{\partial \; D}}$ for a given  [large] fraction $\frac{\partial \; ET}{\partial \; D} < 0$ and negative $\frac{\partial \; ET}{\partial \; D}(\overline{T,\ldots,D})$ (although $g_a$ and Term 1 also probably have a role in this).
+GRA shows the opposite behavior; a relatively high $g1$  makes the function more linear, decreasing the magnitude of $-\overline{\frac{\partial \; ET}{\partial \; D}}$ for a given  [large] fraction $\frac{\partial \; ET}{\partial \; D} < 0$ and negative $\frac{\partial \; ET}{\partial \; D}(\overline{T,\ldots,D})$ (although aerodynamic conductance and Term 1 also probably have a role in this).
 Finally, low $uWUE$ of CSH pushes to toward by far the lowest values $\frac{\partial \; ET}{\partial \; D}$ (Figure \ref{term3}).

### 2017-10-23 src/make_paper_plots.py

@@ -700,3 +700,10 @@ os.system('convert -append %s/climate_et/paper_plots/scatter/*_%s.png '\
           % (os.environ['PLOTS'], meta['x_axis']))
 
 
+##### Table 5  #####
+columns = []
+for i in range(1, 5):
+  columns.append('term_%d' % i)
+  mean_df[columns[-1]] = 1./(mean_df.lai*mean_df.uwue_norm*mean_df.g1**i)
+
+print(mean_df.loc[:, columns])

### 2017-10-25 doc/eee_symposium/vpd.tex
added presentation and did analytical work

--- 2017-10-20 doc/glab_meeting_2017-10-20/vpd.tex
@@
 \title[Your Short Title]{When does VPD drive or reduce ET?}
-\author{Adam (with Pierre and Changjie)}
-\institute{G-Lab Meeting}
-\date{October 20th, 2017}
+\author{Adam Massmann,  Pierre Gentine and Changjie Lin}
+\institute{EEE Graduate Student Symposium}
+\date{October 27th, 2017}
 
@@
 
+\begin{frame}
+  \[\frac{g_{a} \rho}{R* R_{air} \sigma uWUE \left(g_{1} + 1.0 \sqrt{- e_{s} \left(RH - 1.0\right)}\right)^{2} \left(\gamma + \frac{d}{d T} e_{s}\right)} \left(RH - 1.0\right) \left(- R* c_{p} \sigma uWUE \left(g_{1} + 1.0 \sqrt{- e_{s} \left(RH - 1.0\right)}\right)^{2} + 0.3125 R_{air} c_{s} g_{1} \gamma + 0.3125 R_{air} c_{s} \gamma \left(g_{1} + 1.0 \sqrt{- e_{s} \left(RH - 1.0\right)}\right)\right)\]
+\end{frame}
+
 \section{Introduction}
-\begin{frame}{When does VPD drive or reduce ET?}
+\begin{frame}{When does vapor pressure deficit (VPD) drive or reduce evapotranspiration (ET)?}
+  \[VPD = (1-RH)\cdot e_s (T)\]
+\begin{figure}
+  \includegraphics[width=0.25in]{stomata.png}
+ \caption{from evolution.berkeley.edu}
+\end{figure}
+
   \begin{itemize}
@@
   So we can use uWUE to remove dependence of $g_s$ on $A$:
-  \[g_s =  \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{D_{s}}}\right) \frac{uWUE \; ET}{c_s \; \sqrt{D}}\]
+p  \[g_s =  \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{D_{s}}}\right) \frac{uWUE \; ET}{c_s \; \sqrt{D}}\]
 \end{frame}

### 2017-10-25 doc/paper/vpd_et_paper.tex
added presentation and did analytical work

@@
 The Penman-Monteith equation (hereafter PM) estimates ET as a function of atmospheric and land-surface variables:
+\begin{multline}
+   - \frac{g_{a} \rho}{D_{s}^{2} R* R_{air} \sigma uWUE \left(\sqrt{D_{s}} + g_{1}\right)^{2} \left(\Delta + \gamma\right) \left(RH - 1.0\right)} \left(D_{s}^{2} \left(RH - 1\right) - e_{s}^{2} \left(RH - 1.0\right)\right) \left(- R* c_{p} \sigma uWUE \left(\sqrt{D_{s}} + g_{1}\right)^{2} + 0.3125 R_{air} c_{s} g_{1} \gamma + 0.3125 R_{air} c_{s} \gamma \left(\sqrt{D_{s}} + g_{1}\right)\right)
+\end{multline}
 

### 2017-10-25 etc/2017-08-28_VPD_uWUE_notes/explicit.tex
added presentation and did analytical work

@@
 
-Here, $g_{1M}$ has units of kPa$^{0.5}$, $A$ is the net CO$_2$ assimilation rate in $\mu$ mol m$^{-2}$ s$^{-1}$, $c_s$ is the CO$_2$ concentration in ppm (supposed to be at leaf surface).
+Here, $g_{1M}$ has units of kPa$^{0.5}$, $A$ is the net CO$_2$ assimilation rate in $\mu$ mnol m$^{-2}$ s$^{-1}$, $c_s$ is the CO$_2$ concentration in ppm (supposed to be at leaf surface).
 The problem is the $A$ term that we need to get rid of.

### 2017-10-25 src/paper_figure_code/fig06.py

@@ -3,6 +3,7 @@
 This script makes fig 6
 """
 from shared_functions import *
+import time
 
 ### Figure 6 ###
 # note below really takes a long time t
@@ -41,8 +42,8 @@ def vpd_d_et_min(_df):
   """returns rh where ET is minimized, as a function of T and LAI"""
   mean_df = _df.loc[:, ['r_moist', 'gamma', 'c_a',\
                         'lai', 'uwue_norm', 'g1']].mean()
-  vpd = et_min_vpd(mean_df, mean_df.lai)
   t = np.linspace(_df.t_a.min(), _df.t_a.max())
+  vpd = np.ones(t.shape)*et_min_vpd(mean_df, mean_df.lai)
   return t, vpd
 
 def meshgrid_apply(_df, column='var', sample=False):
@@ -67,8 +68,8 @@ def make_ax_plot(_ax, var, _df, meta):
   #                     vmin=vmin, vmax=vmax)
   #below is a hack, should have done by name to begin with
   _df['var'] = var
-  _df = _df.assign(x_cut=pd.cut(_df[meta['x_axis']], 1000),\
-                   t_a_cut=pd.cut(_df['t_a'], 1000))
+  _df = _df.assign(x_cut=pd.cut(_df[meta['x_axis']], 100),\
+                   t_a_cut=pd.cut(_df['t_a'], 100))
   if meta['sample'] == 'sampled':
     grouped = _df.groupby(['x_cut', 't_a_cut']).apply(meshgrid_apply,\
                                                       sample=True)
@@ -181,7 +182,7 @@ def scatter_plot_paper(_df, meta):
 
   plt.tight_layout()
 
-  fname = '%s/climate_et/paper_plots/scatter/%s_%s.png'\
+  fname = '%s/climate_et/paper_plots/scatter/%s_%s.pdf'\
           % (os.environ['PLOTS'], _df.pft.iloc[0], meta['x_axis'])
 
   try:
@@ -192,21 +193,25 @@ def scatter_plot_paper(_df, meta):
   plt.show(block=False)
   return
 
+start = time.time()
+
 plt.close('all')
 meta = {}
 meta['nplots'] = 4 # 5 4
-meta['x_axis'] = 'rh'
 # meta['sample'] = 'sampled'
 meta['sample'] = ''
-df.groupby('pft').apply(scatter_plot_paper, meta)
-os.system('convert -append %s/climate_et/paper_plots/scatter/*_%s.png '\
[…]

### 2017-10-25 src/plot_et_vpd_derivative.py

@@ -138,6 +138,13 @@ meta['ylim'] = None
 plot_tools.scatter_wrapper(df, meta)
 
 meta = {}
+meta['x_var'] = 'e_s'
+meta['y_var'] = 'rh'
+meta['xlim'] = None
+meta['ylim'] = None
+plot_tools.scatter_wrapper(df, meta)
+
+meta = {}
 meta['x_var'] = 'rh'
 meta['y_var'] = 'delta'
 meta['xlim'] = None

### 2017-10-26 doc/eee_symposium/vpd.tex
reorganized to share figures for presentations

@@
 \usepackage{tikz}
-
+\graphicspath{{figures/}}
 \title[Your Short Title]{When does VPD drive or reduce ET?}
@@
   \titlepage
-\end{frame}
-
-\begin{frame}
-  \[\frac{g_{a} \rho}{R* R_{air} \sigma uWUE \left(g_{1} + 1.0 \sqrt{- e_{s} \left(RH - 1.0\right)}\right)^{2} \left(\gamma + \frac{d}{d T} e_{s}\right)} \left(RH - 1.0\right) \left(- R* c_{p} \sigma uWUE \left(g_{1} + 1.0 \sqrt{- e_{s} \left(RH - 1.0\right)}\right)^{2} + 0.3125 R_{air} c_{s} g_{1} \gamma + 0.3125 R_{air} c_{s} \gamma \left(g_{1} + 1.0 \sqrt{- e_{s} \left(RH - 1.0\right)}\right)\right)\]
 \end{frame}

### 2017-10-26 src/analytical_work.py
think may have fixed all code with factor 2 scaling, need to check that modified df in make papaer plots matches new def generated in calc_et_vpd_derivative

@@ -62,4 +62,5 @@ delta_es = Derivative(e_s, t_a)
 term1 = g_a*rho/(gamma + delta_es)
 test2 = Derivative(term1, e_s)
 
+
 test2

### 2017-10-26 src/calc_et_vpd_derivative.py

@@ -29,7 +29,7 @@ H2O = 18.01528e-3 #molecular mass kg/mol
 
 
 start = time.time()
-outdir = '%s/changjie/pandas_data_lai/' % os.environ['DATA']
+outdir = '%s/changjie/pandas_data_lai_fix_scaling/' % os.environ['DATA']
 
 filenames = glob.glob('%s/changjie/MAT_DATA/*.mat' % os.environ['DATA'])
 
@@ -81,8 +81,10 @@ def site_clean(_df, var='lai'):
 
 reload_data = True
 if reload_data:
-  concat_dfs(folder='pandas_data_lai', fname='full_pandas_lai')
-  df = pd.read_pickle('%s/changjie/full_pandas_lai.pkl' % os.environ['DATA'])
+  concat_dfs(folder='pandas_data_lai_fix_scaling',\
+             fname='full_pandas_fix_scaling')
+  df = pd.read_pickle('%s/changjie/full_pandas_fix_scaling.pkl'\
+                      % os.environ['DATA'])
   meta = {}
   meta['folder_label'] = 'site'
   meta['folder'] = 'hist_plots'
@@ -94,7 +96,8 @@ if reload_data:
   df = clean_df(df, var='lai_gpp')
   # test = df.groupby('site').apply(site_clean, 'lai_gpp')
   # test = clean_df(test, var='lai_gpp')
-  df.to_pickle('%s/changjie/full_pandas_lai_clean.pkl' % os.environ['DATA'])
+  df.to_pickle('%s/changjie/full_pandas_fix_scaling_clean.pkl'\
+               % os.environ['DATA'])
   print(df.shape)
   #df.groupby('site').apply(histogram, meta)
   # histogram(df, meta)

### 2017-10-26 src/paper_figure_code/fig05.py

@@ -199,4 +199,4 @@ for ax in axs[:1]:
   ax.legend(h, l, loc='best', fontsize=9)
 # plt.legend(loc='best')
 plt.tight_layout()
-plt.savefig('../doc/paper/fig05.pdf')
+plt.savefig('../../doc/paper/fig05.pdf')

### 2017-10-26 src/paper_figure_code/runall.py

@@ -1,0 +1,8 @@
+#! ~/edward/bin/python
+
+for i in range(5,7):
+  print('working on fig %02d' % i)
+  exec(open('fig%02d.py' % i).read())
+exec(open('tables.py' % i).read())
+
+

### 2017-10-26 doc/eee_symposium/vpd.tex
starting to make progress on presnentation

--- 2017-10-20 doc/glab_meeting_2017-10-20/vpd.tex
@@
 \usepackage{tikz}
+\graphicspath{{figures/}}
 
-\title[Your Short Title]{When does VPD drive or reduce ET?}
-\author{Adam (with Pierre and Changjie)}
-\institute{G-Lab Meeting}
-\date{October 20th, 2017}
+\usepackage{caption}
+\captionsetup{font=scriptsize, labelfont=scriptsize}
+
+\title[Your Short Title]{When does atmospheric dryness drive or reduce evapotranspiration?}
+\author{Adam Massmann,  Pierre Gentine and Changjie Lin}
+\institute{EEE Graduate Student Symposium}
+\date{October 27th, 2017}
 
@@
 \section{Introduction}
-\begin{frame}{When does VPD drive or reduce ET?}
+\begin{frame}{When does atmospheric dryness (VPD) drive or reduce evapotranspiration (ET)?}
+  \begin{Huge}
+  \[VPD = (1-RH)\cdot e_s (T)\]
+\end{Huge}
   \begin{itemize}
   \item \textbf{Hydrometeorologists} would say that an increase in VPD (increase in \textbf{atmospheric demand}) would drive an \textbf{increase in ET}.
+  \end{itemize}
+\end{frame}
+
+\begin{frame}%{When does atmospheric dryness (VPD) drive or reduce evapotranspiration (ET)?}
+  \begin{figure}
+  \includegraphics[width=2.70in]{stomata.png}% was O(1.5in)
+  \caption{from evolution.berkeley.edu}
+\end{figure}
+  \begin{itemize}
   \item However, \textbf{plant physiologists} know that plants have evolved to use stomata to conserve and regulate water use.
@@
 \end{frame}
-
+% develop theory, present vpd figure (add in vpd = 0 locations on line plot below), then test theory with data. add vertical line on vpd plots
 \section{Method}
-\begin{frame}{We will try an analytical approach to this question:}
+\begin{frame}{Use physically reasonable assumptions to develop a theory.}
   We can use Penman-Monteith (PM) to estimate ET:
   \[ET = \frac{\Delta R + g_a \rho_a c_p D}{\Delta + \gamma(1 + \frac{g_a}{g_s})},\]
-  But in (\cite{MEDLYN_2011}):
-  \[  g_{s} = g_0 + 1.6 \left(1 + \frac{g_1}{\sqrt{D}}\right) \frac{A}{c_s}\]
-  $A$ is a function of $ET$. So $ET$ in PM is really an implicit function of itself.
+\textbf{Problem}: $g_s$ (stomatal conductance) is a function of photosythesis, which is a function of ET itself.  So ET in PM is really an implicit function of itself and we cannot take derivatives!
 \end{frame}
 
-\begin{frame}{Remove dependence of $ET$ on $A$}
+\begin{frame}{Use physically reasonable assumptions to develop a theory.}
+Apply a constant uWUE assumption (conserved within plant type; see \cite{Zhou_2016}):
+\[uWUE = \frac{GPP \cdot \sqrt{D}}{ET},\]
+To derive a new form of Penman-Monteith without implicit ET depedence:
+  \[  ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p D_{s}}{R_{air}} - \frac{\text{LAI$_{ref}$}}{\text{LAI}}  \frac{\gamma c_s \sqrt{D} }{ R* \; 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{D}})} \right)\footnote{note that all terms are known }}{ \Delta + \gamma}\]
+\end{frame}
+
+\begin{frame}{Recent result (\cite{Zhou_2016}) allows }
   \cite{Zhou_2016} shows that uWUE:
   \[uWUE = \frac{GPP \cdot \sqrt{D}}{ET}\]
-  is conserved across space and time (within PFT).\\
+  is conserved across space and time (within a plant type).\\
   \medskip

### 2017-10-26 doc/eee_symposium/vpd.tex
serious progress, woudl say almost done with slides:

@@
 \graphicspath{{figures/}}
+\DeclareMathOperator{\sign}{sign}
 
@@
 To derive a new form of Penman-Monteith without implicit ET depedence:
-  \[  ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p D_{s}}{R_{air}} - \frac{\text{LAI$_{ref}$}}{\text{LAI}}  \frac{\gamma c_s \sqrt{D} }{ R* \; 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{D}})} \right)\footnote{note that all terms are known }}{ \Delta + \gamma}\]
+  \[  ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p D_{s}}{R_{air}} -  \frac{\gamma c_s \sqrt{D} }{ R* \; 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{D}})} \right)\footnote{note that all terms are known }}{ \Delta + \gamma}\]
 \end{frame}
 
-\begin{frame}{Recent result (\cite{Zhou_2016}) allows }
-  \cite{Zhou_2016} shows that uWUE:
-  \[uWUE = \frac{GPP \cdot \sqrt{D}}{ET}\]
-  is conserved across space and time (within a plant type).\\
-  \medskip
-  So we can use uWUE to remove dependence of $g_s$ on $A$:
-  \[g_s =  \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{D_{s}}}\right) \frac{uWUE \; ET}{c_s \; \sqrt{D}}\]
-\end{frame}
+\begin{frame}{Use physically reasonable assumptions to develop a theory.}
+  With our new form of Penman-Monteith we can now take derivatives, giving:
+  \[\frac{\partial \;  ET}{\partial \; D} = \frac{2 g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right) \right)\]
+  In the interest of time, we will just focus in the ``sign'' term:
+  \[\sign \left[\frac{\partial \;  ET}{\partial \; D}\right] = \sign \left[  \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right) \right) \right] \]
 
-\begin{frame}{Derive new form of PM}
-  If we plug in our new expression for $g_s$ we can solve for ET to get a new form of PM:
-  \[  ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p D_{s}}{R_{air}} - \frac{\text{LAI$_{ref}$}}{\text{LAI}}  \frac{\gamma c_s \sqrt{D} }{ R* \; 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{D}})} \right)\footnote{note that all terms are known, except for $\frac{\text{LAI$_{ref}$}}{\text{LAI}}$} }{ \Delta + \gamma}\]
-  Now, we can take derivatives, and:
-  \[\frac{\partial \;  ET}{\partial \; D} = \frac{g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} - \frac{\text{LAI$_{ref}$}}{\text{LAI}} \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right) \right)\]
 \end{frame}
@@
 \section{Results - theory}
-\begin{frame}{Scaling Term}
-  \[\frac{\partial \;  ET}{\partial \; D} = \mathbf{ \frac{g_a \; P}{T(\Delta + \gamma)} }   \left(\frac{ c_p}{R_{air}} - \frac{\text{LAI$_{ref}$}}{\text{LAI}} \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right) \right)\]
-\begin{figure}
-  \includegraphics[width=3.5in]{scaling.png}
-  % \caption{}
-\end{figure}
+\begin{frame}{Consequences of theory - ``sign'' term}
+  \[ \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right) \right)\]
+  \begin{itemize}
+  \item $c_p$ and $R*$ are constants
+  \item $R_{air}$, $c_s$, and $\gamma$ are approximately constant (relative to $\sqrt{D}$)
+  \item uWUE and g_1 are constants within plant type (i.e. grass, grops, deciduous forest, evergreen needleleaf forest, shrub)
+  \item Which leaves D = VPD
+  \end{itemize}
+  So within each plant type, whether the atmospheric demand (ET increasing with VPD) or plant response (ET decreasing with VPD) dominates is essentially just a function of VPD!
+\end{frame}
+
+\begin{frame}{Consequences of theory - ``sign'' term}
+  shoe expression for when d et dvpd < 0
+  \begin{itemize}
+  \item $c_p$ and $R*$ are constants
+  \item $R_{air}$, $c_s$, and $\gamma$ are approximately constant (relative to $\sqrt{D}$)
+  \item uWUE and g_1 are constants within plant type (i.e. grass, grops, deciduous forest, evergreen needleleaf forest, shrub)
+  \item Which leaves D = VPD
+  \end{itemize}
+  So within each plant type, whether the atmospheric demand (ET increasing with VPD) or plant response (ET decreasing with VPD) dominates is essentially just a function of VPD!
 \end{frame}

### 2017-10-27 doc/eee_symposium/vpd.tex
figured otu cropping pdf, going to be will finish in morn

@@
   \item $R_{air}$, $c_s$, and $\gamma$ are approximately constant (relative to $\sqrt{D}$)
-  \item uWUE and g_1 are constants within plant type (i.e. grass, grops, deciduous forest, evergreen needleleaf forest, shrub)
+  \item uWUE and $g_1$ are constants within plant type (i.e. grass, grops, deciduous forest, evergreen needleleaf forest, shrub)
   \item Which leaves D = VPD
@@
 \begin{frame}{Consequences of theory - ``sign'' term}
-  shoe expression for when d et dvpd < 0
-  \begin{itemize}
-  \item $c_p$ and $R*$ are constants
-  \item $R_{air}$, $c_s$, and $\gamma$ are approximately constant (relative to $\sqrt{D}$)
-  \item uWUE and g_1 are constants within plant type (i.e. grass, grops, deciduous forest, evergreen needleleaf forest, shrub)
-  \item Which leaves D = VPD
-  \end{itemize}
-  So within each plant type, whether the atmospheric demand (ET increasing with VPD) or plant response (ET decreasing with VPD) dominates is essentially just a function of VPD!
+  \scriptsize
+  \[ VPD_{et_min} = \frac{R_{air}}{4 \, c_p} \left[ \frac{\gamma  c_s}{R* \, 1.6\, \text{uWUE}} + \sqrt{\left(g_1 + 8\, \frac{c_p}{R_{air}}\right) \frac{\gamma c_s}{R* \, 1.6\, \text{uWUE}}} - 4 \, g_1 \frac{ c_p}{R_{air}}\right] \]
+\scriptsize
 \end{frame}
@@
   \begin{figure}
-    \includegraphics[width=3.5in]{sign.png}
+    \includegraphics[width=3.5in]{sign.pdf}
     % \caption{}
@@
 
-\section{Results - reality}
+\section{Results -n reality}
 { % all template changes are local to this group.

### 2017-10-27 doc/eee_symposium/vpd.tex
think prenentation done, just need to spell check proof read and do runthoughs

--- 2017-10-26 doc/eee_symposium/vpd.tex
@@
 \usepackage[english]{babel}
+\usepackage[export]{adjustbox}
  % \usepackage[utf8x]{inputenc}
@@
   We can use Penman-Monteith (PM) to estimate ET:
-  \[ET = \frac{\Delta R + g_a \rho_a c_p D}{\Delta + \gamma(1 + \frac{g_a}{g_s})},\]
+  \[ET = \frac{\Delta R + g_a \rho_a c_p \, VPD}{\Delta + \gamma(1 + \frac{g_a}{g_s})},\]
 \textbf{Problem}: $g_s$ (stomatal conductance) is a function of photosythesis, which is a function of ET itself.  So ET in PM is really an implicit function of itself and we cannot take derivatives!
@@
 Apply a constant uWUE assumption (conserved within plant type; see \cite{Zhou_2016}):
-\[uWUE = \frac{GPP \cdot \sqrt{D}}{ET},\]
+\[uWUE = \frac{GPP \cdot \sqrt{VPD}}{ET},\]
 To derive a new form of Penman-Monteith without implicit ET depedence:
-  \[  ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p D_{s}}{R_{air}} -  \frac{\gamma c_s \sqrt{D} }{ R* \; 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{D}})} \right)\footnote{note that all terms are known }}{ \Delta + \gamma}\]
+  \[  ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p \, VPD}{R_{air}} -  \frac{\gamma c_s \sqrt{VPD} }{ R* \; 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right)\footnote{note that all terms are known }}{ \Delta + \gamma}\]
 \end{frame}
@@
   With our new form of Penman-Monteith we can now take derivatives, giving:
-  \[\frac{\partial \;  ET}{\partial \; D} = \frac{2 g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right) \right)\]
+  \[\frac{\partial \;  ET}{\partial \, VPD} = \frac{2 g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]
   In the interest of time, we will just focus in the ``sign'' term:
-  \[\sign \left[\frac{\partial \;  ET}{\partial \; D}\right] = \sign \left[  \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right) \right) \right] \]
+  \[\sign \left[\frac{\partial \;  ET}{\partial \, VPD}\right] = \sign \left[  \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right) \right] \]
 
@@
 \begin{frame}{Consequences of theory - ``sign'' term}
-  \[ \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right) \right)\]
+  \[\frac{\partial \, ET}{\partial \, VPD} = \text{scaling} \cdot \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]
   \begin{itemize}
   \item $c_p$ and $R*$ are constants
-  \item $R_{air}$, $c_s$, and $\gamma$ are approximately constant (relative to $\sqrt{D}$)
-  \item uWUE and g_1 are constants within plant type (i.e. grass, grops, deciduous forest, evergreen needleleaf forest, shrub)
+  \item $R_{air}$, $c_s$, and $\gamma$ are approximately constant (relative to $\sqrt{VPD}$)
+  \item uWUE and $g_1$ are constants within plant type (i.e. grass, grops, deciduous forest, evergreen needleleaf forest, shrub)
   \item Which leaves D = VPD
@@
 \begin{frame}{Consequences of theory - ``sign'' term}
-  shoe expression for when d et dvpd < 0
-  \begin{itemize}
-  \item $c_p$ and $R*$ are constants
-  \item $R_{air}$, $c_s$, and $\gamma$ are approximately constant (relative to $\sqrt{D}$)
-  \item uWUE and g_1 are constants within plant type (i.e. grass, grops, deciduous forest, evergreen needleleaf forest, shrub)
-  \item Which leaves D = VPD
-  \end{itemize}
-  So within each plant type, whether the atmospheric demand (ET increasing with VPD) or plant response (ET decreasing with VPD) dominates is essentially just a function of VPD!
+Below is a little opaque (at least to me):
+\[\frac{\partial \, ET}{\partial \, VPD} = \text{scaling} \cdot  \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]
+so do expansion about $g_1 >> \sqrt{VPD}$:
+\footnotesize
+\[ \frac{\partial \, ET}{\partial \, VPD} = \text{scaling} \cdot  \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{1}{g_1} - \frac{3\sqrt{VPD}}{2 g_1^2} +  \mathcal{O}\left(\frac{\sqrt{VPD}}{g_1}\right)^{2} \right) \right)\]
+\large
+and:
+  \[ \sqrt{VPD} \Big|_{\frac{\partial ET}{\partial VPD} = 0} \approx  \frac{2 g_1}{3}  \left( -\frac{ c_p \, g_1\,  1.6  \; R*\; \text{ uWUE }}{R_{air} \gamma c_s } + 1  \right) \]
+
+  % \[ VPD \Big|_{\frac{\partial ET}{\partial VPD} = 0} = \frac{R_{air}}{4 \, c_p} \left[ \frac{\gamma  c_s}{R* \, 1.6\, \text{uWUE}} + \sqrt{\left(g_1 + 8\, \frac{c_p}{R_{air}}\right) \frac{\gamma c_s}{R* \, 1.6\, \text{uWUE}}} - 4 \, g_1 \frac{ c_p}{R_{air}}\right] \]
 \end{frame}
@@
   \begin{figure}
-    \includegraphics[width=3.5in]{sign.png}
-    % \caption{}
+    \adjincludegraphics[width=3.5in, trim={0 {0.51\height} 0 0}, clip]{fig05.pdf}
   \end{figure}
@@
 
-\begin{frame}{Summary of theory}
-  \begin{itemize}
-  \item We use \cite{Zhou_2016}'s uWUE to derive a new analytically tractable form of PM.
-  \item This new analysis suggests that the ``tipping'' point for which atmospheric demand overwhelms plant response will be almost exclusively a function of VPD.
-    \begin{itemize}
-    \item For each PFT there will be a VPD$_{crit}$ above which atmospheric demand will dominate and ET will increase with VPD.
-    \end{itemize}
-  \item Plant types evolved to conserve water (CSH) have a higher VPD$_{crit}$ than plants evolved (or bred) to use water and prioritize GPP (CRO).
-Trees and grasslands are somewhere between these two extremes.
-  \item Aerodynamic conductance scales the response, so plants with large surface roughness will be more likely to have a larger response as there is less resistance between the surface and the atmosphere.
-  \end{itemize}
-\end{frame}
-
-\section{Results - reality}
+\section{Test Theory}
 { % all template changes are local to this group.
     \setbeamertemplate{navigation symbols}{}
-    \begin{frame}[plain]{Idealized results - CRO}
+    \begin{frame}[plain]{Test theory with FLUXNET data - CSH}
         \begin{tikzpicture}[remember picture,overlay]
             \node[at=(current page.center)] {
-                \includegraphics[width=\paperwidth]{cro_ideal.png}
+                \includegraphics[width=\paperwidth]{csh_eee.png}
             };
@@
     \setbeamertemplate{navigation symbols}{}
-    \begin{frame}[plain]{Results with uncertainty - CRO}
+    \begin{frame}[plain]{Test theory with FLUXNET data - ENF}
         \begin{tikzpicture}[remember picture,overlay]
             \node[at=(current page.center)] {
-                \includegraphics[width=\paperwidth]{cro.png}
+                \includegraphics[width=\paperwidth]{enf_eee.png}
             };
@@
     \setbeamertemplate{navigation symbols}{}
-    \begin{frame}[plain]{Idealized results - CSH}
+    \begin{frame}[plain]{Test theory with FLUXNET data - DBF}
         \begin{tikzpicture}[remember picture,overlay]
             \node[at=(current page.center)] {
-                \includegraphics[width=\paperwidth]{csh_ideal.png}
+                \includegraphics[width=\paperwidth]{dbf_eee.png}
             };
@@
     \setbeamertemplate{navigation symbols}{}
-    \begin{frame}[plain]{Results with uncertainty - CSH}
+    \begin{frame}[plain]{Test theory with FLUXNET data - CRO}
         \begin{tikzpicture}[remember picture,overlay]
             \node[at=(current page.center)] {
-                \includegraphics[width=\paperwidth]{csh.png}
+                \includegraphics[width=\paperwidth]{cro_eee.png}
             };
@@
     \setbeamertemplate{navigation symbols}{}
-    \begin{frame}[plain]{Idealized results - DBF}
+    \begin{frame}[plain]{Test theory with FLUXNET data - GRA}
         \begin{tikzpicture}[remember picture,overlay]
             \node[at=(current page.center)] {
-                \includegraphics[width=\paperwidth]{dbf_ideal.png}
+                \includegraphics[width=\paperwidth]{gra_eee.png}
             };
@@
 
-{ % all template changes are local to this group.
-    \setbeamertemplate{navigation symbols}{}
-    \begin{frame}[plain]{Results with uncertainty - DBF}
-        \begin{tikzpicture}[remember picture,overlay]
-            \node[at=(current page.center)] {
-                \includegraphics[width=\paperwidth]{dbf.png}
-            };
-        \end{tikzpicture}
-     \end{frame}
-   }
+% \begin{frame}{Summary of theory}
+%   \begin{itemize}
+%   \item We use \cite{Zhou_2016}'s uWUE to derive a new analytically tractable form of PM.
+%   \item This new analysis suggests that the ``tipping'' point for which atmospheric demand overwhelms plant response will be almost exclusively a function of VPD.
+%     \begin{itemize}
+%     \item For each PFT there will be a VPD$_{crit}$ above which atmospheric demand will dominate and ET will increase with VPD.
+%     \end{itemize}
+%   \item Plant types evolved to conserve water (CSH) have a higher VPD$_{crit}$ than plants evolved (or bred) to use water and prioritize GPP (CRO).
+Trees and grasslands are somewhere between these two extremes.
+%   \item Aerodynamic conductance scales the response, so plants with large surface roughness will be more likely to have a larger response as there is less resistance between the surface and the atmosphere.
+%   \end{itemize}
+% \end{frame}
 
-{ % all template changes are local to this group.
-    \setbeamertemplate{navigation symbols}{}
-    \begin{frame}[plain]{Idealized results - ENF}
-        \begin{tikzpicture}[remember picture,overlay]
-            \node[at=(current page.center)] {
-                \includegraphics[width=\paperwidth]{enf_ideal.png}
-            };
-        \end{tikzpicture}
-     \end{frame}
-}
+\section{Conclusions}
+% \begin{frame}{Summary}
+%   \begin{itemize}
+%   \item Theory finds that the ``tipping'' point for which atmospheric demand overwhelms plant response will be almost exclusively a function of VPD.
+Plant types evolved to conserve water (CSH) have a higher VPD$_{crit}$ (and more negative ET response) than plants evolved (or bred) to use water and prioritize GPP (CRO).
+%   \item On average, ecosystem response to VPD follows roughly what we might expect: CRO (prioritize GPP) has positive ET response to VPD, while all others have a negative response.
+Ordering by increasing magnitude of negative response gives: DBF, GRA, ENF, CSH; which roughly correspond to expectations for increasing water conservation as a function of PFT.
+%   \item Uncertainty is high, especially for CRO and GRA. However, inclusion of uncertainty does not change the story for ENF or CSH.
+%   \end{itemize}
+% \end{frame}
 
-{ % all template changes are local to this group.
-    \setbeamertemplate{navigation symbols}{}
-    \begin{frame}[plain]{Results with uncertainty - ENF}
-        \begin{tikzpicture}[remember picture,overlay]
-            \node[at=(current page.center)] {
-                \includegraphics[width=\paperwidth]{enf.png}
-            };
-        \end{tikzpicture}
-     \end{frame}
-   }
-
-{ % all template changes are local to this group.
-    \setbeamertemplate{navigation symbols}{}
-    \begin{frame}[plain]{Idealized results - GRA}
-        \begin{tikzpicture}[remember picture,overlay]
-            \node[at=(current page.center)] {
-                \includegraphics[width=\paperwidth]{gra_ideal.png}
-            };
-        \end{tikzpicture}
-     \end{frame}
-}
-
-{ % all template changes are local to this group.
-    \setbeamertemplate{navigation symbols}{}
-    \begin{frame}[plain]{Results with uncertainty - GRA}
-        \begin{tikzpicture}[remember picture,overlay]
-            \node[at=(current page.center)] {
-                \includegraphics[width=\paperwidth]{gra.png}
-            };
-        \end{tikzpicture}
-     \end{frame}
-   }
-
-   \begin{frame}{Summary statistics}
-     \begin{table}
-       \caption{Statistics of $\frac{\partial \; ET}{\partial \; D}$ as a function of PFT.}
-       \centering
-       \begin{tabular}{l c c}
-         \hline
-         PFT & $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ & fraction $\frac{\partial \; ET}{\partial \; VPD} < 0.$ \\
-         \hline
-         CRO & 0.000853  & 0.473311\\
-         CSH & -0.108234 & 0.931660\\
-         DBF & -0.012727 & 0.461674\\
-         ENF & -0.034087 & 0.534425\\
-         GRA & -0.019637 & 0.631735\\
-         \hline
-         \multicolumn{2}{l}{}  
-       \end{tabular}
-     \end{table}
+\begin{frame}{Summary}
+  \begin{itemize}
+  \item Derived a new form of Penman-Monteith that explicitly calculates ET given published constants.
+  \item Developed a theory for when atmospheric dryness will increase ET (atmospheric demand dominates) vs.
+decrease ET (plant response dominates).
+Theory tests well using real data from FLUXNET2015 on plant types CSH, ENF, and DBF.
+Applicability to CRO and GRA is more uncertain.
+  \item For plants evolved to prioritize production over water conservation (high uWUE and $g_1$), atmospheric demand is more likely to dominate.
+  \item But, for all plant types except for crops decreases in ET (plant response dominates) with increases in VPD are common (plant response dominates).
+Therefor, the hydrometeorlogists perspective is often wrong, and drought indices built upon assumptions about atmospepheric demand are unrealistic for vegetated surfaces.
+  \item Our form of Penman-Monteith is capable of estimating ET more realistically and could be used as a replacement for PET in drought indices.
+  \end{itemize}
 \end{frame}
 
-\section{Conclusions}
-\begin{frame}{Summary}
-  \begin{itemize}
-  \item Theory finds that the ``tipping'' point for which atmospheric demand overwhelms plant response will be almost exclusively a function of VPD.
-Plant types evolved to conserve water (CSH) have a higher VPD$_{crit}$ (and more negative ET response) than plants evolved (or bred) to use water and prioritize GPP (CRO).
-  \item On average, ecosystem response to VPD follows roughly what we might expect: CRO (prioritize GPP) has positive ET response to VPD, while all others have a negative response.
-Ordering by increasing magnitude of negative response gives: DBF, GRA, ENF, CSH; which roughly correspond to expectations for increasing water conservation as a function of PFT.
-  \item Uncertainty is high, especially for CRO and GRA. However, inclusion of uncertainty does not change the story for ENF or CSH.
-  \end{itemize}
-\end{frame}
+% \begin{frame}{Summary statistics - extra sl}
+%      \begin{table}
+%        \caption{Statistics of $\frac{\partial \; ET}{\partial \, VPD}$ as a function of PFT.}
+%        \centering
+%        \begin{tabular}{l c c}
+%          \hline
+%          PFT & $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ & fraction $\frac{\partial \; ET}{\partial \; VPD} < 0.$ \\
+%          \hline
+%          CRO & 0.000853  & 0.473311\\
+%          CSH & -0.108234 & 0.931660\\
+%          DBF & -0.012727 & 0.461674\\
+%          ENF & -0.034087 & 0.534425\\
+%          GRA & -0.019637 & 0.631735\\
+%          \hline
+%          \multicolumn{2}{l}{}  
+%        \end{tabular}
+%      \end{table}
+%    \end{frame}
 

### 2017-10-27 src/paper_figure_code/fig06.py

@@ -91,7 +91,7 @@ def make_ax_plot(_ax, var, _df, meta):
   grouped = grouped.sort_index(axis=1)
   # print('grouped sortex axis 1', grouped)
   grouped = grouped.sort_index(axis=0)
-  # print('grouped sort axis 2', grouped)n
+  # print('grouped sort axis 2', grouped)
   _x = grouped.columns.values
   _y = grouped.index.values
   _x = np.append(_x, _xmax)
@@ -113,7 +113,7 @@ def make_ax_plot(_ax, var, _df, meta):
     # _ax.plot(test, t_a, 'k-')
     # test = met.vapor_pres(t_a)*100.*(1. - 0.2)
     # _ax.plot(test, t_a, 'k-')
-  _ax.set_xlabel(meta['x_axis'])
+  _ax.set_xlabel('%s (Pa)' % meta['x_axis'].upper())
   _ax.set_ylabel('T (C)')
   _ax.set_title('PFT: %s; %s'\
                 % (str(_df['pft'][0]),\

### 2017-10-27 doc/eee_symposium/vpd.tex
about to just include uncertainty plots

@@
   \[ET = \frac{\Delta R + g_a \rho_a c_p \, VPD}{\Delta + \gamma(1 + \frac{g_a}{g_s})},\]
-\textbf{Problem}: $g_s$ (stomatal conductance) is a function of photosythesis, which is a function of ET itself.  So ET in PM is really an implicit function of itself and we cannot take derivatives!
+\textbf{Problem}: $g_s$ (stomatal conductance) is a function of photosynthesis, which is a function of ET itself.  So ET in PM is really an implicit function of itself and we cannot take derivatives!
 \end{frame}
@@
 \[uWUE = \frac{GPP \cdot \sqrt{VPD}}{ET},\]
-To derive a new form of Penman-Monteith without implicit ET depedence:
+To derive a new form of Penman-Monteith without implicit ET dependence:
   \[  ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p \, VPD}{R_{air}} -  \frac{\gamma c_s \sqrt{VPD} }{ R* \; 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right)\footnote{note that all terms are known }}{ \Delta + \gamma}\]
@@
   \item $R_{air}$, $c_s$, and $\gamma$ are approximately constant (relative to $\sqrt{VPD}$)
-  \item uWUE and $g_1$ are constants within plant type (i.e. grass, grops, deciduous forest, evergreen needleleaf forest, shrub)
-  \item Which leaves D = VPD
+  \item uWUE and $g_1$ are constants within plant type (i.e. grass, crops, deciduous forest, evergreen needleleaf forest, shrub)
+  \item Which just leaves VPD
   \end{itemize}
@@
 \footnotesize
-\[ \frac{\partial \, ET}{\partial \, VPD} = \text{scaling} \cdot  \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{1}{g_1} - \frac{3\sqrt{VPD}}{2 g_1^2} +  \mathcal{O}\left(\frac{\sqrt{VPD}}{g_1}\right)^{2} \right) \right)\]
-\large
-and:
-  \[ \sqrt{VPD} \Big|_{\frac{\partial ET}{\partial VPD} = 0} \approx  \frac{2 g_1}{3}  \left( -\frac{ c_p \, g_1\,  1.6  \; R*\; \text{ uWUE }}{R_{air} \gamma c_s } + 1  \right) \]
+\[ \frac{\partial \, ET}{\partial \, VPD} \approx \text{scaling} \cdot  \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{1}{g_1} - \frac{3\sqrt{VPD}}{2 g_1^2} +  \mathcal{O}\left(\frac{\sqrt{VPD}}{g_1}\right)^{2} \right) \right)\]
+% \large
+% and:
+%   \[ \sqrt{VPD} \Big|_{\frac{\partial ET}{\partial VPD} = 0} \approx  \frac{2 g_1}{3}  \left( -\frac{ c_p \, g_1\,  1.6  \; R*\; \text{ uWUE }}{R_{air} \gamma c_s } + 1  \right) \]
 
@@
 \begin{frame}{Summary}
+  \small
   \begin{itemize}
@@
 Applicability to CRO and GRA is more uncertain.
-  \item For plants evolved to prioritize production over water conservation (high uWUE and $g_1$), atmospheric demand is more likely to dominate.
-  \item But, for all plant types except for crops decreases in ET (plant response dominates) with increases in VPD are common (plant response dominates).
-Therefor, the hydrometeorlogists perspective is often wrong, and drought indices built upon assumptions about atmospepheric demand are unrealistic for vegetated surfaces.
-  \item Our form of Penman-Monteith is capable of estimating ET more realistically and could be used as a replacement for PET in drought indices.
+  \item For plants evolved to prioritize water conservation over production (low uWUE and $g_1$), atmospherc demand is less likely to dominate.
+  \item All plant types except for crops had common occurence of decreases in ET (plant response dominates) with increases in VPD (plant response dominates).
+Therefore, the hydrometeorologists perspective is often wrong, and drought indices built upon assumptions about atmospheric demand are unrealistic for vegetated surfaces.
+  \item Our form of Penman-Monteith is capable of estimating ET more realistically and could be used as a replacement for PET in drought indices (e.g. Palmer Drought Severity Index - PDSI).
   \end{itemize}
@@
   \printbibliography
+  \scriptsize
+  \begin{itemize}
+  \item This work used eddy covariance data acquired and shared by the FLUXNET community, including these networks: AmeriFlux, AfriFlux, AsiaFlux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont, ChinaFlux, Fluxnet-Canada, GreenGrass, ICOS, KoFlux, LBA, NECC, OzFlux-TERN, TCOS-Siberia, and USCCC.
+The ERA-Interim reanalysis data are provided by ECMWF and processed by LSCE.
+The FLUXNET eddy covariance data processing and harmonization was carried out by the European Fluxes Database Cluster, AmeriFlux Management Project, and Fluxdata project of FLUXNET, with the support of CDIAC and ICOS Ecosystem Thematic Center, and the OzFlux, ChinaFlux and AsiaFlux offices.
+    \end{itemize}
 \end{frame}

### 2017-10-27 src/paper_figure_code/fig05.py

@@ -4,8 +4,6 @@ This script makes fig 5
 """
 from shared_functions import *
 
-
-
 # below is whether to plot taylor series approximation
 PLOT_SERIES = False
 #below is to make slight changes for the talk
@@ -13,7 +11,7 @@ PLOT_TALK = True
 
 if PLOT_TALK:
   paren_string = r'$\left(\frac{ c_p}{R_{air}} '\
-                 r'- \frac{\gamma c_s }{LAI \; 1.6 \; R\; uWUE  }'\
+                 r'- \frac{\gamma c_s }{1.6 \; R\; uWUE  }'\
                  r'\left( \frac{2 g_1 + \sqrt{D}}'\
                  r'{2 (g_1 + \sqrt{D})^2}\right)\right)$'
 else:
@@ -196,11 +194,11 @@ for pft in ['CRO', 'DBF', 'GRA', 'ENF', 'CSH']:
 axs[1].set_xlabel('VPD (Pa)')
 axs[3].set_xlabel('LAI')
 axs[0].set_ylabel(paren_string)
+axs[0].plot(axs[0].get_xlim(), [0., 0.], 'k--', linewidth=1.0)
 plt.setp(axs[2].get_yticklabels(), visible=False)
 axs[-1].set_ylabel(r'VPD')#$_{ETmin}$')
 axs[2].set_title('VPD where ET = Min(ET) '\
                  r'($\frac{\partial \; ET}{\partial \; D} = 0$)')
-axs[0].plot(axs[2].get_xlim(), [0., 0.], 'k--', linewidth=0.2)
 # axs[2].set_ylim([0., np.around(df.vpd.quantile(q=0.95), decimals=-2)])
 axs[2].set_ylim([0., 4000.])
 axs[1].set_ylim([0.5,5.5])

### 2017-10-27 doc/eee_symposium/vpd.tex
edition I'm senind goff to james

@@
   \item $c_p$ and $R*$ are constants
-  \item $R_{air}$, $c_s$, and $\gamma$ are approximately constant (relative to $\sqrt{VPD}$)
+  \item $R_{air}$, $\gamma$, and $c_s$ are approximately constant (relative to $\sqrt{VPD}$)
   \item uWUE and $g_1$ are constants within plant type (i.e. grass, crops, deciduous forest, evergreen needleleaf forest, shrub)
@@
   \end{itemize}
-  So within each plant type, whether the atmospheric demand (ET increasing with VPD) or plant response (ET decreasing with VPD) dominates is essentially just a function of VPD!
+  So within each plant type, whether the atmospheric demand or plant response (ET decreasing with VPD) dominates is essentially just a function of VPD!
 \end{frame}
@@
 \section{Test Theory}
-{ % all template changes are local to this group.
-    \setbeamertemplate{navigation symbols}{}
-    \begin{frame}[plain]{Test theory with FLUXNET data - CSH}
-        \begin{tikzpicture}[remember picture,overlay]
-            \node[at=(current page.center)] {
-                \includegraphics[width=\paperwidth]{csh_eee.png}
-            };
-        \end{tikzpicture}
+    \begin{frame}{Test theory with FLUXNET data - CSH}
+                \includegraphics[width=3.5in]{csh_eee.png}
      \end{frame}
-}
 
-{ % all template changes are local to this group.
-    \setbeamertemplate{navigation symbols}{}
-    \begin{frame}[plain]{Test theory with FLUXNET data - ENF}
-        \begin{tikzpicture}[remember picture,overlay]
-            \node[at=(current page.center)] {
-                \includegraphics[width=\paperwidth]{enf_eee.png}
-            };
-        \end{tikzpicture}
+    \begin{frame}{Test theory with FLUXNET data - ENF}
+                \includegraphics[width=3.5in]{enf_eee.png}
      \end{frame}
-}
 
-{ % all template changes are local to this group.
-    \setbeamertemplate{navigation symbols}{}
-    \begin{frame}[plain]{Test theory with FLUXNET data - DBF}
-        \begin{tikzpicture}[remember picture,overlay]
-            \node[at=(current page.center)] {
-                \includegraphics[width=\paperwidth]{dbf_eee.png}
-            };
-        \end{tikzpicture}
+    \begin{frame}{Test theory with FLUXNET data - DBF}
+                \includegraphics[width=3.5in]{dbf_eee.png}
      \end{frame}
-}
 
-{ % all template changes are local to this group.
-    \setbeamertemplate{navigation symbols}{}
-    \begin{frame}[plain]{Test theory with FLUXNET data - CRO}
-        \begin{tikzpicture}[remember picture,overlay]
-            \node[at=(current page.center)] {
-                \includegraphics[width=\paperwidth]{cro_eee.png}
-            };
-        \end{tikzpicture}
+    \begin{frame}{Test theory with FLUXNET data - CRO}
+                \includegraphics[width=3.5in]{cro_eee.png}
      \end{frame}
-}
 
-{ % all template changes are local to this group.
-    \setbeamertemplate{navigation symbols}{}
-    \begin{frame}[plain]{Test theory with FLUXNET data - GRA}
-        \begin{tikzpicture}[remember picture,overlay]
-            \node[at=(current page.center)] {
-                \includegraphics[width=\paperwidth]{gra_eee.png}
-            };
-        \end{tikzpicture}
+    \begin{frame}{Test theory with FLUXNET data - GRA}
+                \includegraphics[width=3.5in]{gra_eee.png}
      \end{frame}
-}
 
@@
 Applicability to CRO and GRA is more uncertain.
-  \item For plants evolved to prioritize water conservation over production (low uWUE and $g_1$), atmospherc demand is less likely to dominate.
+  \item For plants evolved to prioritize water conservation over production, atmospherc demand is less likely to dominate.
   \item All plant types except for crops had common occurence of decreases in ET (plant response dominates) with increases in VPD (plant response dominates).
 Therefore, the hydrometeorologists perspective is often wrong, and drought indices built upon assumptions about atmospheric demand are unrealistic for vegetated surfaces.
-  \item Our form of Penman-Monteith is capable of estimating ET more realistically and could be used as a replacement for PET in drought indices (e.g. Palmer Drought Severity Index - PDSI).
+  \item Our form of Penman-Monteith is capable of estimating ET realistically and could be used as a replacement for PET in drought indices (e.g. Palmer Drought Severity Index - PDSI).
   \end{itemize}

### 2017-10-27 doc/gen_shared_images.sh

@@ -13,10 +13,10 @@ convert ./paper/fig06.png -crop 25x20%+0+0 ./shared_figs/temp_1.png
 convert ./paper/fig06b.png -crop 25x20%+0+0 ./shared_figs/temp_2.png
 convert +append ./shared_figs/temp*.png ./shared_figs/cro.png
 rm -rf ./shared_figs/temp*.png
-convert ./paper/fig06.png -gravity NorthEast -crop 25x20%+0+0 ./shared_figs/temp_1.png
-convert ./paper/fig06.png -crop 25x20%+0+0 ./shared_figs/temp_2.png
-convert +append ./shared_figs/temp*.png ./shared_figs/cro_eee.png
-rm -rf ./shared_figs/temp*.png
+convert ./paper/fig06.png -crop 25x20%+0+0 ./shared_figs/cro_eee.png
+
+
+
 
 # idealized plots
 convert ./paper/fig06.png -gravity East -crop 25x20%+0+0 ./shared_figs/temp_1.png
@@ -27,10 +27,10 @@ convert ./paper/fig06.png -gravity West -crop 25x20%+0+0 ./shared_figs/temp_1.pn
 convert ./paper/fig06b.png -gravity West -crop 25x20%+0+0 ./shared_figs/temp_2.png
 convert +append ./shared_figs/temp*.png ./shared_figs/dbf.png
 rm -rf ./shared_figs/temp*.png
-convert ./paper/fig06.png -gravity East -crop 25x20%+0+0 ./shared_figs/temp_1.png
-convert ./paper/fig06.png -gravity West -crop 25x20%+0+0 ./shared_figs/temp_2.png
-convert +append ./shared_figs/temp*.png ./shared_figs/dbf_eee.png
-rm -rf ./shared_figs/temp*.png
+convert ./paper/fig06.png -gravity West -crop 25x20%+0+0 ./shared_figs/dbf_eee.png
+
+
+
 
 
 # idealized plots
@@ -42,10 +42,9 @@ convert ./paper/fig06.png -gravity NorthEast -crop 640x480+0+480 ./shared_figs/t
 convert ./paper/fig06b.png -gravity NorthEast -crop 640x480+0+480 ./shared_figs/temp_2.png
 convert +append ./shared_figs/temp*.png ./shared_figs/csh_ideal.png
 rm -rf ./shared_figs/temp*.png
-convert ./paper/fig06.png -crop 640x480+0+480 ./shared_figs/temp_2.png
[…]

### 2017-10-27 src/paper_figure_code/fig06.py

@@ -136,33 +136,33 @@ def scatter_plot_paper(_df, meta):
   fig.set_figwidth(fig.get_figwidth()*nplots)
 
   if nplots == 4:
-    titles = [r'$\frac{\partial \; ET}{\partial \; D}$',\
-              r'$\frac{\partial \; ET}{g_a \partial \; D}$',\
-              r'$\frac{\partial \; ET}{\partial \; D}(\overline{\sigma})$',\
-              r'$\frac{\partial \; ET}{g_a \partial \; D}(\overline{\sigma})$']
+    titles = [r'$\frac{\partial \; ET}{\partial \, VPD}$',\
+              r'$\frac{\partial \; ET}{g_a \partial \, VPD}$',\
+              r'$\frac{\partial \; ET}{\partial \, VPD}(\overline{\sigma})$',\
+              r'$\frac{\partial \; ET}{g_a \partial \, VPD}(\overline{\sigma})$']
     _vars = [_df['d_et'],\
              _df['d_et']/_df['g_a'],\
              _df['d_et_lai_fixed'],\
              _df['d_et_lai_fixed']/_df['g_a']]# ,\
   elif nplots == 5:
-    titles = [r'$\frac{\partial \; ET}{\partial \; D}$',\
-              r'$\frac{\partial \; ET}{g_a \partial \; D}$',\
-              r'$\frac{\partial \; ET}{\partial \; D}(\overline{\sigma})$',\
-              r'$\frac{\partial \; ET}{g_a \partial \; D}(\overline{\sigma})$',\
+    titles = [r'$\frac{\partial \; ET}{\partial \, VPD}$',\
+              r'$\frac{\partial \; ET}{g_a \partial \, VPD}$',\
+              r'$\frac{\partial \; ET}{\partial \, VPD}(\overline{\sigma})$',\
+              r'$\frac{\partial \; ET}{g_a \partial \, VPD}(\overline{\sigma})$',\
               r'$\frac{\partial \; ET}'\
-              r'{g_a \partial \; D}(\overline{LAI, \gamma})$']
+              r'{g_a \partial \, VPD}(\overline{LAI, \gamma})$']
     _vars = [_df['d_et'],\
              _df['d_et']/_df['g_a'],\
              _df['d_et_lai_fixed'],\
              _df['d_et_lai_fixed']/_df['g_a'],\
              _df['d_et_lai_gamma_fixed']/_df['g_a']]
   else:
[…]

### 2017-10-27 doc/eee_symposium/vpd.tex
actually what I send

@@
   \end{itemize}
-  So within each plant type, whether the atmospheric demand or plant response (ET decreasing with VPD) dominates is essentially just a function of VPD!
+  So within each plant type, whether the atmospheric demand (ET increasing with VPD) or plant response (ET decreasing with VPD) dominates is essentially just a function of VPD!
 \end{frame}
@@
   \item For plants evolved to prioritize water conservation over production, atmospherc demand is less likely to dominate.
-  \item All plant types except for crops had common occurence of decreases in ET (plant response dominates) with increases in VPD (plant response dominates).
+  \item All plant types except for crops had common occurence of decreases in ET increases in VPD (plant response dominates).
 Therefore, the hydrometeorologists perspective is often wrong, and drought indices built upon assumptions about atmospheric demand are unrealistic for vegetated surfaces.

### 2017-11-13 src/paper_figure_code/fig06.py
tried to do calcs with c4 grass, but it seems like uncertainty is always higher so maybe all these grasses are in fact c3?

@@ -139,7 +139,8 @@ def scatter_plot_paper(_df, meta):
     titles = [r'$\frac{\partial \; ET}{\partial \, VPD}$',\
               r'$\frac{\partial \; ET}{g_a \partial \, VPD}$',\
               r'$\frac{\partial \; ET}{\partial \, VPD}(\overline{\sigma})$',\
-              r'$\frac{\partial \; ET}{g_a \partial \, VPD}(\overline{\sigma})$']
+              r'$\frac{\partial \; ET}{g_a \partial \, VPD}'\
+              r'(\overline{\sigma})$']
     _vars = [_df['d_et'],\
              _df['d_et']/_df['g_a'],\
              _df['d_et_lai_fixed'],\
@@ -148,7 +149,8 @@ def scatter_plot_paper(_df, meta):
     titles = [r'$\frac{\partial \; ET}{\partial \, VPD}$',\
               r'$\frac{\partial \; ET}{g_a \partial \, VPD}$',\
               r'$\frac{\partial \; ET}{\partial \, VPD}(\overline{\sigma})$',\
-              r'$\frac{\partial \; ET}{g_a \partial \, VPD}(\overline{\sigma})$',\
+              r'$\frac{\partial \; ET}{g_a \partial \, VPD}'\
+              r'(\overline{\sigma})$',\
               r'$\frac{\partial \; ET}'\
               r'{g_a \partial \, VPD}(\overline{LAI, \gamma})$']
     _vars = [_df['d_et'],\
@@ -160,7 +162,8 @@ def scatter_plot_paper(_df, meta):
     titles = [r'$\frac{\partial \; ET}{\partial \, VPD}$',\
               r'$\frac{\partial \; ET}{g_a \partial \, VPD}$',\
               r'$\frac{\partial \; ET}{\partial \, VPD}(\overline{\sigma})$',\
-              r'$\frac{\partial \; ET}{g_a \partial \, VPD}(\overline{\sigma})$',\
+              r'$\frac{\partial \; ET}{g_a \partial \, VPD}'\
+              r'(\overline{\sigma})$',\
               r'$\frac{\partial \;n ET}'\
               r'{g_a \partial \, VPD}(\overline{LAI, \gamma})$',\
               'c_a_fixed',\

### 2017-11-27 doc/paper/vpd_et_paper.tex
think may have fixed all code with factor 2 scaling, need to check that modified df in make papaer plots matches new def generated in calc_et_vpd_derivative
starting edits to paper again

--- 2017-10-23 doc/paper/vpd_et_paper.tex
@@
-%% 11/23/2015
+% 11/23/2015
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-In order to quantify plant response to perturbations to atmospheric demand for water, we apply a Penman-Monteith framework to eddy-covariance observation spanning various biomes and climates.
+In order to quantify plant response to perturbations to atmospheric demand for water, we apply a Penman-Monteith framework to eddy-covariance observations spanning various biomes and climates.
 Section 2 describes the data used, Section 3 derives the framework, Section 4 presents results, and Section 5 discusses conclusions.
@@
 The Penman-Monteith equation (hereafter PM) estimates ET as a function of atmospheric and land-surface variables:
-
 \begin{linenomath*}
@@
 Note that $uWUE$ is fit on the ecosystem scale in \citet{Zhou_2015} so GPP in \ref{uwue} is really $A\cdot \text{LAI}$.
-This leads to the cancelation of LAI in additoin to uWUE in Equation \ref{medlyn}.
+This leads to the cancelation of LAI in addition to uWUE in Equation \ref{medlyn}.
 We also recognize that our use of $uWUE$ introudces considerable uncertainty, so we added an uncertainty parameter $\sigma = f(t, \text{site})$ modifying $uWUE$.
@@
 % AGU does not want a .bib or a .bbl file. Please copy in the contents of your .bbl file here.
-\bibliography{references.bib}
+% \bibliography{references.bib}
 \input{vpd_et_paper.bbl}

### 2017-11-27 src/calc_et_vpd_derivative.py

@@ -103,6 +103,13 @@ def site_clean(_df, var='lai'):
                 (_df[var] > _df[var].quantile(q=0.05))), :]
   return out
 
+def print_cv_lai_lai_gpp(_df):
+  """prints the cv of lai and lai_gpp"""
+  print(_df.site.iloc[0], _df.pft.iloc[0])
+  print('cv lai: %f' % (_df.lai.std()/_df.lai.mean()))
+  print('cv lai_gpp: %f'  %(_df.lai_gpp.std()/_df.lai_gpp.mean()))
+  return (_df.lai.std()/_df.lai.mean())# ,\
+          # (_df.lai_gpp.std()/_df.lai_gpp.mean())
 reload_data = True
 if reload_data:
   concat_dfs(folder='pandas_data_c3_c4',\
@@ -122,6 +129,8 @@ if reload_data:
   # test = clean_df(test, var='lai_gpp')
   df.to_pickle('%s/changjie/full_pandas_c3_c4_clean.pkl'\
                % os.environ['DATA'])
+
+  cvs = df.groupby('site').apply(print_cv_lai_lai_gpp)
   print(df.shape)
   #df.groupby('site').apply(histogram, meta)
   # histogram(df, meta)

### 2017-11-27 doc/paper/vpd_et_paper.tex
some progress on paper edits, hopefully can finish by end of today and send out to Pierre and Changjie

@@
 
+\section{Data}
+\label{data}
+We use data from FLUXNET2015.
+Because $g_1$ coefficients \citep{Lin_2015} and uWUE were only both available for five plant functional types (PFTs - see Table \ref{pft}),  only 56 of the 77 sites were used.
+Figure \ref{map_fig} \explain{map needs to be improved - it's a placeholder for now} presents each site and its plant functional type.
+
 \section{Methods}
-
+n
 The Penman-Monteith equation (hereafter PM) estimates ET as a function of atmospheric and land-surface variables:
@@
   \begin{equation}
-  g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{D_{s}}}\right) \frac{\sigma \; uWUE \; ET}{c_s \; \sqrt{D}}
+  g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{D_{s}}}\right) \frac{uWUE \; ET}{c_s \; \sqrt{D}}
   \end{equation}
@@
 This leads to the cancelation of LAI in addition to uWUE in Equation \ref{medlyn}.
-We also recognize that our use of $uWUE$ introudces considerable uncertainty, so we added an uncertainty parameter $\sigma = f(t, \text{site})$ modifying $uWUE$.
-The meaning of this parameter will be elaborated on after the derivation of $\frac{\partial \; ET}{\partial \; D}$.
-
-Next, plugging Equation 5 into Equation 1 and rearranging gives:
+Plugging Equation 5 into Equation 1 and rearranging gives:
 
@@
   \begin{equation}
-    ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p D_{s}}{R_{air}} -  \frac{\gamma c_s \sqrt{D} }{ R* \; 1.6\; \sigma\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{D}})} \right) }{ \Delta + \gamma}
+    ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p D_{s}}{R_{air}} -  \frac{\gamma c_s \sqrt{D} }{ R* \; 1.6\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{D}})} \right) }{ \Delta + \gamma}
     \label{et}
@@
 
-Note that if we set the uncertaitny parameter $\sigma$ to 1, and approximage $c_s$ at a global mean CO$_2$ concentration, then Equation \ref{et} is fully defined using commonly available weather station data and the constants published in \citet{Zhou_2015, Lin_2015}.
-This then begs the question, why use PET at all for drought metrics in vegetated areas?
-A much more physically realistic estimate of ET can be had with the same information reqiured to calculate PET.
-
-We can now take the derivative with respect to $D$ to analytically determine ecosystem response to atmospheric demand perturbations:
+Given FLUXNET data described in Section \ref{data}, every term in Equation \ref{et} is known.
+However, our sampling of sites and our focus on the growing season may intoduce some deviations of $uWUE$ from those observed in \citet{Zhou_2015}.
+Also, we wish to include some measure of uncertainty in our analysis to guide if our many assumptions and simplifications are reasonable.
+To account for both mean deviations of $uWUE$ and uncertainty, we will introduce an uncertainty parameter $\sigma$ modifying $uWUE$:
 
@@
   \begin{equation}
-    \frac{\partial \;  ET}{\partial \; D} = \frac{g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_s }{1.6 \; R*\; \sigma \; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right) \right)
+    ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p D_{s}}{R_{air}} -  \frac{\gamma c_s \sqrt{D} }{ R* \; 1.6\; \sigma \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{D}})} \right) }{ \Delta + \gamma}
+    \label{et_sigma}
+  \end{equation}
+\end{linenomath*}
+
+Now, from each FLUXNET observation we can calculate single value of $\sigma$:
+
+\begin{linenomath*}
+  \begin{equation}
+\sigma = - \frac{g_a \gamma c_s \sqrt{D_s} L_v P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p D_{s}\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{D_s}})}
+    \label{et_sigma}
+  \end{equation}
+\end{linenomath*}
+
+The variablility of $\sigma$ across sites and time will provides some measure of uncertainty in our model, assumptions, as well as the fluxnet observations themselves.
+Mean shifts in $\sigma$ can be attributed to differences in out sampling from those used to calcualte $uWUE$ in \citet{Zhou_2015}.
+The influence of $\sigma$ will propogate through any uncertainty to our derivative of Equation \ref{et_sigma}:
+
+\begin{linenomath*}
+  \begin{equation}
+    \frac{\partial \;  ET}{\partial \; D} = \frac{2\; g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_s }{1.6 \; R*\; \sigma \; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right) \right)
     \label{d_et}
@@
 
+With Equation \ref{d_et} we have provided an analytical franework for ecosystem reponse to atmospheric demand perturbations.
+There are a few subtelties to taking the derivative in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$) and $D$ are functionally related, so while taking the derivative we evaluate $\frac{\partial \; ET}{\partial \; D} = \frac{\partial \; ET} {\partial \; e_s} \frac{\partial \; e_s}{\partial \; D} \Big|_{\text{RH fixed}} + \frac{\partial \; ET}{partial \; RH} \frac{\partial \; RH}{\partial \; D} \Big|_{\text{$e_s$ fixed}}$.
+$RH$ and $e_s$ are assumed to be approximately orthogonal.
+
 The $D$ dependence in Equation \ref{d_et} is a little opaque.
 However, mean $D$ is 1062 Pa, so $\sqrt{D}$ is 32.6 Pa$^{1/2}$, which is much less than $g_1$ (with the exception of ENF; Table \ref{pft}).
-So a series expansion in the limit $\frac{\sqrt{D}}{g_1} \to 0$ gives an approximation which makes the functional form more clear:
+So a series expansion in the limit $\frac{\sqrt{D}}{g_1} \to 0$ gives an approximation which makes the functional form more clear: \explain{Should I even include the series expansion?}
 
@@
 
-Note that given yearly uWUE from \citet{Zhou_2015}, $g_1$ from \citet{Lin_2015} \citep[as presented in ][]{Franks_2017}, and observations of R, T, P, D$_s$, and wind speed (WS), all quantities in Equations \ref{d_et} and \ref{d_et_approx} are known except for the uncertainty parameter $\sigma$.
-With flux tower observations of ET, $\sigma$ will then be uniquely determined for each observation through Equation 6:
-
-\begin{linenomath*}
-  \begin{equation}
-    \sigma = -\frac{g_a \gamma c_s \sqrt{D_s} P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p D_{s}\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{D_s}})}
-    \label{lai}
-  \end{equation}
-\end{linenomath*}
-
-This $\sigma$ will be some part uncertainty associated with assumptions about constant $g1$ and $uWUE$, including differences between the data used here and in the fits in \citet{Zhou_2015} and \citet{Lin_2015}, but also some part general PM model and FLUXNET observationsal uncertainty.
-By calculating a unique $\sigma$ for each observation we will propagate any model and observational uncertainty forward into our expression for $\frac{\partial \; ET}{\partial D}$.
+One final comment on our derivation which will not be discussed further but is relevant for future analysis: if we  approximage $c_s$ at a global mean CO$_2$ concentration, then the RHS of Equation \ref{et} is fully defined using commonly available weather station data and the constants published in \citet{Zhou_2015, Lin_2015}.
+This then begs the question, why use PET for drought metrics in vegetated areas?
+It appears a much more physically realistic estimate of ET can be had with the same information reqiured to calculate PET.
 
@@
 \end{table}
-
-\section{Data}
-
-We use data from FLUXNET2015.
-Because $g_1$ coefficients \citep{Lin_2015} and uWUE were only both available for five plant functional types (PFTs - see Table \ref{pft}),  only 56 of the 77 sites were used.
-Figure \ref{map_fig} \explain{map needs to be improved - it's a placeholder for now} presents each site and its plant functional type.
 
@@
 An additional concern is that the $\sigma$ term may in fact be some function of $D$, in which case the dependence would need to be accounted for when taking the derivative.
-Figure \ref{lai_vpd_fig}\explain{Figs 2 and 3 can probably be combined - the histogram of $\sigma$ is shown in Fig 3.} plots the joint distribution of $\sigma$ and VPD, and shows that $\sigma$ is very weakly a function of VPD.
+Figure \ref{lai_vpd_fig} plots the joint distribution of $\sigma$ and VPD, and shows that $\sigma$ is very weakly a function of VPD.
 Given this weak dependence, we argue that Equation \ref{d_et} is a valid approximation for ET response to $D$.
@@
 
-In Term 1, $\frac{P}{T} \propto \rho$, so this should vary little relative to aerodynamic conductance and $\Delta$.
-$\gamma$ should also be relatively constant, so the scaling term, Term 1, should be primarily a function of aerodynamic conductance and temperature (through the function $\Delta$).
+Term 2 minus Term 3's role in determining the sign of the sensitivty of ET to VPD makes it crucuial for answering our question ``When does VPD drive or reduce ET?'' Exploring these terms more, $c_s$ and $\gamma$ variability is relatively less than $\sigma$ and $D$ variability, so variability within PFT will be solely determined by $\sigma$ and $D$.
+If we fix uncertainty $\sigma$ at PFT averages, then Term 2 minus Term 3 is just a function of $D$.
+We can further dertmine a $D_{crit}$ where $\frac{\partial \; ET}{\partial \; D} = 0$:
+
+\begin{linenomath*}
+  \begin{equation}
+D_{crit} = \frac{R_{air}}{4 c_p} \left( \frac{\gamma c_s}{1.6\; R \; \overline{\sigma} uWUE} + \sqrt{\frac{\gamma c_s}{1.6\; R \; \overline{\sigma} uWUE}\left( \frac{\gamma c_s}{1.6\; R \; \overline{\sigma} uWUE} + 8 g_1 \frac{c_p}{R_{air}}\right)} - 4 g_1 \frac{c_p}{R_{air}} \right)
+\label{vpd_min_et}
+  \end{equation}
+\end{linenomath*}
+
+Values of $D_{crit}$ as a function of PFT are shown in Table \ref{vpd_crit}.
+For any values of $D$ less than $D_{crit}$, $\frac{\partial \; ET}{\partial \; D}$ will be negative, and for values of $D$ greater than $D_{crit}$, $\frac{\partial \; ET}{\partial \; D}$ will be positive.
+
+\begin{table}
+  \label{vpd_crit}
+\caption{Values of $D_{crit}$, where $\frac{\partial \; ET}{\partial \; D} = 0$, evaluated at PFT average values for $R_{air}$, $\sigma$, $\gamma$, and $c_s$.
+For reference, these values are also provided.
+For values of $D$ less than $D_{crit}$, $\frac{\partial \; ET}{\partial \; D}$ will be negative, and for values of $D$ greater than $D_{crit}$, $\frac{\partial \; ET}{\partial \; D}$ will be positive.}
+\centering
+\begin{tabular}{l c c c c c c}
+  \hline
+  PFT & $R_{air}$ & $c_s$ (ppm) & $\gamma$ & $\overline{\sigma}$ & $\overline{\sigma} \cdot uWUE$ & \textbf{$D_{crit}$ (Pa)} \\
+  \hline
+  CRO &  288.680920 & 372.567691& 65.351523& 0.684394&       2.602873&  \textbf{133.165438} \\
+  CSH &   289.067152& 381.593622& 67.613172& 0.997224&       2.175278& \textbf{4439.564212} \\
+  DBF &   288.624437& 377.449849& 63.421812& 0.881061&       2.746393&  \textbf{888.773243} \\
+  ENF &  288.183849& 377.676463& 61.559242& 1.217892&       4.015362&  \textbf{978.084845} \\
+  GRA &  288.425651& 377.264645& 61.598768& 0.850869&       2.281074& \textbf{1141.630778} \\
+\hline
+\multicolumn{2}{l}{$^{a}$Footnote text here.}
+\end{tabular}
+\end{table}
+
+Figure \ref{term3} provides greater description for how (Term 2 - Term 3) varies with $D$, as a function of PFT.
+Equation \ref{d_et_approx} aids interpretation of Figure \ref{term3}.
+Larger $uWUE\cdot\overline{\sigma}$, and $g1$ shift the leading-order constant term ($\frac{1}{g_1}$) towards smaller values, and (Term 2 - Term 3) towards positive values.
+$uWUE$ and $g1$ are both water-use effiency type constants.
+Higher values corresponde to plants that are more willing to spend water on primary production and less evolved to conserve water.
+Figure \ref{term3} confirms our physical intuition: CROs are the least water conservative so have the smallest constant portion of Term 3, while CSH are the most water conservative and have the largest constant portion of Term 3.
+For the VPD-dependent terms in Equation \ref{d_et_approx}, differences in $g1$ between PFTs exert a greater influence than difference $uWUE$, as the power of $g1$ increases.
+Increasing $uWUE$ and $g1$ decreases the VPD-dependence, but $g1$ has the bigger effect due to its increasing powers.
+ENF ($g1 = 74.31$) has by far the largest VPD dependence of response, while CRO ($g1 = 183.1$) has the smallest VPD dependence.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=20pc]{./fig05.pdf}
+\caption{Sources of variability for Term 2 - Term 3.
+Top: Term 2 - Term 3 as a function of VPD, with $\sigma$ held fixed at PFT averages.
+The observed range of VPD for each PFT is also shown below the x-axis.
+Line extent corresponds to 5th and 95th percentiles, while stars denote the location of the 25th, 50th, and 75th percentiles.
+
+Bottom: The location of the minima of ET, as a function of VPD and $\sigma$. Lines and stars denote the distribution of VPD and $\sigma$ next to each axis, following the same percentiles as above.}
+\label{term3}
+\end{figure}
+
+Figure \ref{term3}b shows the location of the minima of ET, as a function of $\sigma$ and $D$.
+For any $\sigma$ or VPD less (more) than these curves, Term 2 - Term 3 will be negative (positive).
+It is clear that the portion of VPD observations below/above these curves will be a strong function of $\sigma$.
+However, we can see some general trends.
+For CSH, $\frac{\partial \; ET}{\partial \; D}$ should be negative for the vast majority of observed $\sigma$ and VPD.
+The fraciton of positive $\frac{\partial \; ET}{\partial \; D}$ appears to be more even for ENF, GRA, and DBF, and we might expect a greater frequency of positive $\frac{\partial \; ET}{\partial \; D}$ for CRO.
+
+While the above discussion of the sign of $\frac{\partial \; ET}{\partial \; D}$ is important to answer our research question, the magnitude of $\frac{\partial \; ET}{\partial \; D}$ will also impact statistics of $\frac{\partial \; ET}{\partial \; D}$ and the importance of $D$ variablility for $ET$ variability.
+So we will more closely examine the scaling Term 1: $\frac{P}{T} \propto \rho$, so this should vary little relative to aerodynamic conductance and $\Delta$.
+$\gamma$ should also be relatively constant, so Term 1 should be primarily a function of aerodynamic conductance and temperature (through the function $\Delta$).
 This makes sense, as aerodynamic conductance represents how efficently response at the surface is communicated to the atmosphere.
@@
 $\Delta$'s presence in the scaling term also matches physical intition.
-Evaporative cooling will will dampen the ability of the atmosphere to take more moisture, as $e_{sat}$ decreases with temperature..
-The decrease in $e_{sat}$ is proportional to $\Delta$ ($\delta e_{sat} = \Delta \delta T$).
+Evaporative cooling will dampen the ability of the atmosphere to take more moisture, because $e_{s}$ decreases with temperature.
+The decrease in $e_{s}$ is proportional to $\Delta$ ($\delta e_{s} = \Delta \delta T$).
 So as $\Delta$ increases, you will get a larger damping of ET due to evaporative cooling.
@@
 %%%%% replace below with idea:
+\subsection{Bulk statistics of $\frac{\partial \; ET}{\partial \; D}$}
 
-Term 2 minus Term 3 determines the sign of the sensitivty of ET to VPD and is thus crucial.
-$c_s$ variability is relatively less than $\sigma$ and $D$ variability, so variability within PFT will be solely determined by $\sigma$ and $D$ at the different fluxnet sites.
-Figure \ref{term3} shows how (Term 2 - Term 3) varies with $D$ and $\sigma$, as a function of PFT.
-Equation \ref{d_et_approx} aids interpretation of Figure \ref{term3}.
-Larger $uWUE$, $\overline{\sigma}$, and $g1$ shift the leading-order constant term ($\frac{1}{g_1}$) towards smaller values, and (Term 2 - Term 3) towards positive values.
-$uWUE$ and $g1$ are both water-use effiency type constants.
-Higher values corresponde to plants that are more willing to spend water on primary production and less evolved to conserve water.
-Figure \ref{term3} confirms our physical intuition: CROs are the least water conservative so have the smallest constant portion of Term 3, while CSH are the most water conservative and have the largest constant portion of Term 3.
-For the VPD-dependent terms in Equation \ref{d_et_approx}, differences in $g1$ between PFTs exert a greater influence than difference $uWUE$, as the power of $g1$ increases.
-Increasing $uWUE$ and $g1$ decreases the VPD-dependence, but $g1$ has the bigger effect due to its increasing powers.
-ENF ($g1 = 74.31$) has by far the largest VPD dependence of response, while CRO ($g1 = 183.1$) has the smallest VPD dependence.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=20pc]{./fig05.pdf}
-\caption{Sources of variability for Term 2 - Term 3.
-Top: Term 2 - Term 3 as a function of VPD, with $\sigma$ held fixed at PFT averages.
-The observed range of VPD for each PFT is also shown below the x-axis.
-Line extent corresponds to 5th and 95th percentiles, while stars denote the location of the 25th, 50th, and 75th percentiles.
-
-Bottom: The location of the minima of ET, as a function of VPD and $\sigma$. Lines and stars denote the distribution of VPD and $\sigma$ next to each axis, following the same percentiles as above.}
-\label{term3}
-\end{figure}
-
-Figure \ref{term3}b shows the location of the minima of ET, as a function of $\sigma$ and $D$.
-For any $\sigma$ or VPD less (more) than these curves, Term 2 - Term 3 will be negative (positive).
-It is clear that the portion of VPD observations below/above these curves will be a strong function of $\sigma$.
-However, we can see some general trends.
-For CSH, $\frac{\partial \; ET}{\partial \; D}$ should be negative for the vast majority of observed $\sigma$ and VPD.
-The fraciton of positive $\frac{\partial \; ET}{\partial \; D}$ appears to be more even for ENF, GRA, and DBF, and we might expect a greater frequency of positive $\frac{\partial \; ET}{\partial \; D}$ for CRO.
-
-Table 3 confirms these expectations for PFT behavior of $\frac{\partial \; ET}{\partial \; D}$.
+Table 3 confirms our expectations for PFT behavior of $\frac{\partial \; ET}{\partial \; D}$.
 For all PFTs except for CRO, average $\frac{\partial \; ET}{\partial \; D}$ is less than zero.
@@
 To better understand $D$'s contribution, we normalize $\frac{\partial \; ET}{\partial \; D}$ with $D$'s standard deviation to define a (linearized) relative change in ET for variations in $D$ .
-$D$'s contribution to ET's variability ranges between 16 - 20 W m$^{-2}$ for all PFTs except for CSH, which is about 51 W m$^{-2}$.
-Another meaningful comparison is to $\frac{\partial \; ET}{\partial \; R} * std(R)$, as net radiationis generally the driver of ET (cite joe berry here).
-For all PFTs except for CSH $D$ contributes between 14.5 - 20.5 \% of $R$'s contribution to variability.
-For CSH the portion is much larger, about 44 \%.
-$D$'s variability is certiantly a non-negligable contributor to $ET$'s variability.
+$D$'s contribution to ET's variability ranges between 30 - 40 W m$^{-2}$ for all PFTs except for CSH, which is about 100 W m$^{-2}$.
+Another meaningful comparison is to $\frac{\partial \; ET}{\partial \; R} \cdot std(R)$, as net radiation is generally the driver of ET (cite joe berry here).
+For all PFTs except for CSH $D$ contributes between 30 - 40 \% of $R$'s contribution to variability.
+For CSH the portion is much larger, about 88 \%.
+$D$'s variability is certiantly a primary contributor to $ET$'s variability.
 

### 2017-11-27 src/paper_figure_code/shared_functions.py

@@ -95,3 +95,6 @@ def et_min_vpd(_df, lai):
     if sqrt_vpd < 0.:
       sqrt_vpd = np.nan
   return sqrt_vpd**2
+
+def get_pft(_df):
+  return _df['pft'].iloc[0]

### 2017-11-27 src/paper_figure_code/tables.py

@@ -37,10 +37,18 @@ test = mean_df.loc[:, ['d_et', 'd_et_bar', 'd_et_bar_std',\
 
 print(test)
 
-# antoher table, nto sure which one
+# think below is jsut for reference for the series, but not used in paper
 columns = []
 for i in range(1, 5):
   columns.append('term_%d' % i)
   mean_df[columns[-1]] = 1./(mean_df.lai*mean_df.uwue_norm*mean_df.g1**i)
 
 print(mean_df.loc[:, columns])
+
+mean_df['vpd_crit'] = et_min_vpd(mean_df, mean_df.lai)
+mean_df['sigma_dot_uwue'] = mean_df.lai*mean_df.uwue_norm
+columns = ['r_moist', 'c_a', 'gamma', 'lai',\
+           'sigma_dot_uwue', 'vpd_crit']
+print('\n vpd critical table')
+print(mean_df.loc[:, columns])
+

### 2017-11-27 doc/paper/vpd_et_paper.tex
finished first paragraph of testing theory; should now flow with current text and probably just needs editing

@@
 
-\subsection{Full observations of $\frac{\partial \; ET}{\partial \; D}$}
-
-Now that we have an intuitive understanding of $\frac{\partial \; ET}{\partial \; D}$'s behavior, we are equipped to interpret fully realistic plots of $\frac{\partial \; ET}{\partial \; D}$ for each PFT.
-Figure \ref{real} presents calculated $\frac{\partial \; ET}{\partial \; D}$ where, unless otherwise noted, all variables in Equation \ref{d_et} are allowed to vary.
+\subsection{Testing the theory - does the data match the model?}
+
+So far we have developed a theory for $\frac{\partial \; ET}{\partial \; D}$'s behavior.
+In particular, we determined a critical threshold $D_{crit}$ for $\frac{\partial \; ET}{\partial \; D}$.
+For $D < D_{crit}$, $\frac{\partial \; ET}{\partial \; D}  < 0$, and for $D > D_{crit}$, $\frac{\partial \; ET}{\partial \; D} > 0$.
+This $D_{crit}$ is only a function of PFT.
+However, many assumptions, including constant uWUE, were made in deriving $D_{crit}$.
+So, we need to test if our theory and $D_{crit}$ hold up in the face of uncertainty.
+Figure \ref{real} presents calculated $\frac{\partial \; ET}{\partial \; D}$ where, unless otherwise noted, all variables in Equation \ref{d_et} are allowed to vary, including uncertainty.
 Each column is a different quantity related to $\frac{\partial \; ET}{\partial \; D}$, and each row is a different PFT.
@@
 The full observations generally confirm expectations from Section \ref{results}.
-CRO has the most positive values of $\frac{\partial \; ET}{\partial \; D}$, $\frac{\partial \; ET}{\partial \; D}$ is almost always negative for CSH, and response depends more with the environmental conditions for the other PFTs (especially ENF).
+CRO has the most pospitive values of $\frac{\partial \; ET}{\partial \; D}$, $\frac{\partial \; ET}{\partial \; D}$ is almost always negative for CSH, and response depends more with the environmental conditions for the other PFTs (especially ENF).
 Through the columns of Figure \ref{real} we can see the impact of $\sigma$ and $g_a$ on the variability of $\frac{\partial \; ET}{\partial \; D}$.

### 2017-11-27 doc/paper/vpd_et_paper.tex
hopefully changed all d to vpd, much easier to udnerstand now

--- 2017-11-27 doc/paper/vpd_et_paper.tex
@@
   \begin{equation}
-      ET = \frac{\Delta R + g_a \rho_a c_p D_{s}}{\Delta + \gamma(1 + \frac{g_a}{g_s})},
+      ET = \frac{\Delta R + g_a \rho_a c_p VPD}{\Delta + \gamma(1 + \frac{g_a}{g_s})},
   \end{equation}
@@
   \begin{equation}
-  g_{l-s} = g_0 + 1.6 \left(1 + \frac{g_1}{\sqrt{D_{s}}}\right) \frac{A}{c_s}
+  g_{l-s} = g_0 + 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{A}{c_s}
   \end{equation}
@@
   \begin{equation}
-  g_s = \text{LAI} \frac{R \,T}{P} \left( g_0 + 1.6 \left(1 + \frac{g_1}{\sqrt{D_{s}}}\right) \frac{A}{c_s}\right)
+  g_s = \text{LAI} \frac{R \,T}{P} \left( g_0 + 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{A}{c_s}\right)
   \end{equation}
@@
     \label{uwue}
-uWUE = \frac{GPP \cdot \sqrt{D}}{ET}
+uWUE = \frac{GPP \cdot \sqrt{VPD}{ET}
   \end{equation}
@@
   \begin{equation}
-  g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{D_{s}}}\right) \frac{uWUE \; ET}{c_s \; \sqrt{D}}
+  g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{uWUE \; ET}{c_s \; \sqrt{VPD}}
   \end{equation}
@@
   \begin{equation}
-    ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p D_{s}}{R_{air}} -  \frac{\gamma c_s \sqrt{D} }{ R* \; 1.6\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{D}})} \right) }{ \Delta + \gamma}
+    ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_s \sqrt{VPD} }{ R* \; 1.6\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
     \label{et}
@@
   \begin{equation}
-    ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p D_{s}}{R_{air}} -  \frac{\gamma c_s \sqrt{D} }{ R* \; 1.6\; \sigma \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{D}})} \right) }{ \Delta + \gamma}
+    ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_s \sqrt{VPD} }{ R* \; 1.6\; \sigma \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
     \label{et_sigma}
@@
   \begin{equation}
-\sigma = - \frac{g_a \gamma c_s \sqrt{D_s} L_v P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p D_{s}\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{D_s}})}
+\sigma = - \frac{g_a \gamma c_s \sqrt{VPD} L_v P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p VPD\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})}
     \label{et_sigma}
@@
   \begin{equation}
-    \frac{\partial \;  ET}{\partial \; D} = \frac{2\; g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_s }{1.6 \; R*\; \sigma \; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right) \right)
+    \frac{\partial \;  ET}{\partial \; VPD} = \frac{2\; g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_s }{1.6 \; R*\; \sigma \; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)
     \label{d_et}
@@
 With Equation \ref{d_et} we have provided an analytical franework for ecosystem reponse to atmospheric demand perturbations.
-There are a few subtelties to taking the derivative in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$) and $D$ are functionally related, so while taking the derivative we evaluate $\frac{\partial \; ET}{\partial \; D} = \frac{\partial \; ET} {\partial \; e_s} \frac{\partial \; e_s}{\partial \; D} \Big|_{\text{RH fixed}} + \frac{\partial \; ET}{partial \; RH} \frac{\partial \; RH}{\partial \; D} \Big|_{\text{$e_s$ fixed}}$.
+There are a few subtelties to taking the derivative in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$) and $VPD$ are functionally related, so while taking the derivative we evaluate $\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET} {\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD} \Big|_{\text{RH fixed}} + \frac{\partial \; ET}{partial \; RH} \frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$ fixed}}$.
 $RH$ and $e_s$ are assumed to be approximately orthogonal.
 
-The $D$ dependence in Equation \ref{d_et} is a little opaque.
-However, mean $D$ is 1062 Pa, so $\sqrt{D}$ is 32.6 Pa$^{1/2}$, which is much less than $g_1$ (with the exception of ENF; Table \ref{pft}).
-So a series expansion in the limit $\frac{\sqrt{D}}{g_1} \to 0$ gives an approximation which makes the functional form more clear: \explain{Should I even include the series expansion?}
+The $VPD$ dependence in Equation \ref{d_et} is a little opaque.
+However, mean $VPD$ is 1062 Pa, so $\sqrt{VPD}$ is 32.6 Pa$^{1/2}$, which is much less than $g_1$ (with the exception of ENF; Table \ref{pft}).
+So a series expansion in the limit $\frac{\sqrt{VPD}}{g_1} \to 0$ gives an approximation which makes the functional form more clear: \explain{Should I even include the series expansion?}
 
@@
   \begin{equation}
-    \frac{\partial \;  ET}{\partial \; D} \approx \frac{g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_s }{1.6 \; R*\; \sigma \; \text{ uWUE }} \left( \frac{1}{g_{1}} - \frac{3 \sqrt{D}}{2 g_{1}^{2}} + \frac{2 \sqrt{D}^{2}}{g_{1}^{3}} - \frac{5 \sqrt{D}^{3}}{2 g_{1}^{4}} + \mathcal{O}\left(\left(\frac{\sqrt{D}}{g_1}\right)^{4}\right) \right) \right)
+    \frac{\partial \;  ET}{\partial \; VPD} \approx \frac{g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_s }{1.6 \; R*\; \sigma \; \text{ uWUE }} \left( \frac{1}{g_{1}} - \frac{3 \sqrt{VPD}}{2 g_{1}^{2}} + \frac{2 \sqrt{VPD}^{2}}{g_{1}^{3}} - \frac{5 \sqrt{VPD}^{3}}{2 g_{1}^{4}} + \mathcal{O}\left(\left(\frac{\sqrt{VPD}}{g_1}\right)^{4}\right) \right) \right)
     \label{d_et_approx}
@@
   $c_p$  & specific heat capacity of air at constant pressure & J K$^{-1}$ kg$^{-1}$ \\
-  $D$  & VPD & Pa  \\
+  $VPD$  & VPD & Pa  \\
   $\gamma$  & psychrometric constant & Pa K$^{-1}$   \\
@@
 
-An additional concern is that the $\sigma$ term may in fact be some function of $D$, in which case the dependence would need to be accounted for when taking the derivative.
+An additional concern is that the $\sigma$ term may in fact be some function of $VPD$, in which case the dependence would need to be accounted for when taking the derivative.
 Figure \ref{lai_vpd_fig} plots the joint distribution of $\sigma$ and VPD, and shows that $\sigma$ is very weakly a function of VPD.
-Given this weak dependence, we argue that Equation \ref{d_et} is a valid approximation for ET response to $D$.
+Given this weak dependence, we argue that Equation \ref{d_et} is a valid approximation for ET response to $VPD$.
 
@@
 \includegraphics[width=20pc]{./fig03.png}
-\caption{The joint distribution of $D$ and $\sigma$.
-$\sigma$ has only a weak dependence on $D$.
+\caption{The joint distribution of $VPD$ and $\sigma$.
+$\sigma$ has only a weak dependence on $VPD$.
 ***This plot could proabably benefit from a box plot of site specific correlations, because some sites do have stronger depednence than others.
@@
 Before calculating the sensitivity of ET to VPD, it is useful to consider the functional form of Equation \ref{d_et}.
-There are three terms: a scaling term for the full expression we will call Term 1 ($\frac{g_a \; P}{T(\Delta + \gamma)}$), a relatively constant offset we will call Term 2 ($\frac{c_p}{R_{air}}$), and a variable term we will call Term 3 ($\frac{\gamma c_s }{1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right)$).
+There are three terms: a scaling term for the full expression we will call Term 1 ($\frac{g_a \; P}{T(\Delta + \gamma)}$), a relatively constant offset we will call Term 2 ($\frac{c_p}{R_{air}}$), and a variable term we will call Term 3 ($\frac{\gamma c_s }{1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right)$).
 All variables are positive, so the relative magnitude between Term 2 and Term 3 will determine the sign of the derivative, while Term 1 will scale the expression larger or smaller.
 
-Term 2 minus Term 3's role in determining the sign of the sensitivty of ET to VPD makes it crucuial for answering our question ``When does VPD drive or reduce ET?'' Exploring these terms more, $c_s$ and $\gamma$ variability is relatively less than $\sigma$ and $D$ variability, so variability within PFT will be solely determined by $\sigma$ and $D$.
-If we fix uncertainty $\sigma$ at PFT averages, then Term 2 minus Term 3 is just a function of $D$.
-We can further dertmine a $D_{crit}$ where $\frac{\partial \; ET}{\partial \; D} = 0$:
+Term 2 minus Term 3's role in determining the sign of the sensitivty of ET to VPD makes it crucuial for answering our question ``When does VPD drive or reduce ET?'' Exploring these terms more, $c_s$ and $\gamma$ variability is relatively less than $\sigma$ and $VPD$ variability, so variability within PFT will be solely determined by $\sigma$ and $VPD$.
+If we fix uncertainty $\sigma$ at PFT averages, then Term 2 minus Term 3 is just a function of $VPD$.
+We can further dertmine a $VPD_{crit}$ where $\frac{\partial \; ET}{\partial \; VPD} = 0$:
 
@@
   \begin{equation}
-D_{crit} = \frac{R_{air}}{4 c_p} \left( \frac{\gamma c_s}{1.6\; R \; \overline{\sigma} uWUE} + \sqrt{\frac{\gamma c_s}{1.6\; R \; \overline{\sigma} uWUE}\left( \frac{\gamma c_s}{1.6\; R \; \overline{\sigma} uWUE} + 8 g_1 \frac{c_p}{R_{air}}\right)} - 4 g_1 \frac{c_p}{R_{air}} \right)
+VPD_{crit} = \frac{R_{air}}{4 c_p} \left( \frac{\gamma c_s}{1.6\; R \; \overline{\sigma} uWUE} + \sqrt{\frac{\gamma c_s}{1.6\; R \; \overline{\sigma} uWUE}\left( \frac{\gamma c_s}{1.6\; R \; \overline{\sigma} uWUE} + 8 g_1 \frac{c_p}{R_{air}}\right)} - 4 g_1 \frac{c_p}{R_{air}} \right)
 \label{vpd_min_et}
@@
 
-Values of $D_{crit}$ as a function of PFT are shown in Table \ref{vpd_crit}.
-For any values of $D$ less than $D_{crit}$, $\frac{\partial \; ET}{\partial \; D}$ will be negative, and for values of $D$ greater than $D_{crit}$, $\frac{\partial \; ET}{\partial \; D}$ will be positive.
+Values of $VPD_{crit}$ as a function of PFT are shown in Table \ref{vpd_crit}.
+For any values of $VPD$ less than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be negative, and for values of $VPD$ greater than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be positive.
 
@@
   \label{vpd_crit}
-\caption{Values of $D_{crit}$, where $\frac{\partial \; ET}{\partial \; D} = 0$, evaluated at PFT average values for $R_{air}$, $\sigma$, $\gamma$, and $c_s$.
+\caption{Values of $VPD_{crit}$, where $\frac{\partial \; ET}{\partial \; VPD} = 0$, evaluated at PFT average values for $R_{air}$, $\sigma$, $\gamma$, and $c_s$.
 For reference, these values are also provided.
-For values of $D$ less than $D_{crit}$, $\frac{\partial \; ET}{\partial \; D}$ will be negative, and for values of $D$ greater than $D_{crit}$, $\frac{\partial \; ET}{\partial \; D}$ will be positive.}
+For values of $VPD$ less than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be negative, and for values of $VPD$ greater than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be positive.}
 \centering
@@
   \hline
-  PFT & $R_{air}$ & $c_s$ (ppm) & $\gamma$ & $\overline{\sigma}$ & $\overline{\sigma} \cdot uWUE$ & \textbf{$D_{crit}$ (Pa)} \\
+  PFT & $R_{air}$ & $c_s$ (ppm) & $\gamma$ & $\overline{\sigma}$ & $\overline{\sigma} \cdot uWUE$ & \textbf{$VPD_{crit}$ (Pa)} \\
   \hline
@@
 
-Figure \ref{term3} provides greater description for how (Term 2 - Term 3) varies with $D$, as a function of PFT.
+Figure \ref{term3} provides greater description for how (Term 2 - Term 3) varies with $VPD$, as a function of PFT.
 Equation \ref{d_et_approx} aids interpretation of Figure \ref{term3}.
@@
 
-Figure \ref{term3}b shows the location of the minima of ET, as a function of $\sigma$ and $D$.
+Figure \ref{term3}b shows the location of the minima of ET, as a function of $\sigma$ and $VPD$.
 For any $\sigma$ or VPD less (more) than these curves, Term 2 - Term 3 will be negative (positive).
@@
 However, we can see some general trends.
-For CSH, $\frac{\partial \; ET}{\partial \; D}$ should be negative for the vast majority of observed $\sigma$ and VPD.
-The fraciton of positive $\frac{\partial \; ET}{\partial \; D}$ appears to be more even for ENF, GRA, and DBF, and we might expect a greater frequency of positive $\frac{\partial \; ET}{\partial \; D}$ for CRO.
-
-While the above discussion of the sign of $\frac{\partial \; ET}{\partial \; D}$ is important to answer our research question, the magnitude of $\frac{\partial \; ET}{\partial \; D}$ will also impact statistics of $\frac{\partial \; ET}{\partial \; D}$ and the importance of $D$ variablility for $ET$ variability.
+For CSH, $\frac{\partial \; ET}{\partial \; VPD}$ should be negative for the vast majority of observed $\sigma$ and VPD.
+The fraciton of positive $\frac{\partial \; ET}{\partial \; VPD}$ appears to be more even for ENF, GRA, and DBF, and we might expect a greater frequency of positive $\frac{\partial \; ET}{\partial \; VPD}$ for CRO.
+
+While the above discussion of the sign of $\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our research question, the magnitude of $\frac{\partial \; ET}{\partial \; VPD}$ will also impact statistics of $\frac{\partial \; ET}{\partial \; VPD}$ and the importance of $VPD$ variablility for $ET$ variability.
 So we will more closely examine the scaling Term 1: $\frac{P}{T} \propto \rho$, so this should vary little relative to aerodynamic conductance and $\Delta$.
@@
 %%%%% replace below with idea:
-\subsection{Bulk statistics of $\frac{\partial \; ET}{\partial \; D}$}
-
-Table 3 confirms our expectations for PFT behavior of $\frac{\partial \; ET}{\partial \; D}$.
-For all PFTs except for CRO, average $\frac{\partial \; ET}{\partial \; D}$ is less than zero.
-However, $\frac{\partial \; ET}{\partial \; D}$ evaluated at the average of all variables (e.g.
-$\sigma$, $T$, $c_s$, $D$) is only negative for CSH and GRA.
-And, DBF in addition to CRO experiences $\frac{\partial \; ET}{\partial \; D}$ < 0 less than half the time.
-These observations highlight the effect of the nonlinear function in Figure \ref{term3}: $\frac{\partial \; ET}{\partial \; D}$ has a much steeper slope when the function is negative, and is thus more likely to be large.
-
-The units of $\frac{\partial \; ET}{\partial \; D}$ make it difficult to interpret if $D$ is even a meaningful contributor to ET's variability.
-To better understand $D$'s contribution, we normalize $\frac{\partial \; ET}{\partial \; D}$ with $D$'s standard deviation to define a (linearized) relative change in ET for variations in $D$ .
-$D$'s contribution to ET's variability ranges between 30 - 40 W m$^{-2}$ for all PFTs except for CSH, which is about 100 W m$^{-2}$.
+\subsection{Bulk statistics of $\frac{\partial \; ET}{\partial \; VPD}$}
+
+Table 3 confirms our expectations for PFT behavior of $\frac{\partial \; ET}{\partial \; VPD}$.
+For all PFTs except for CRO, average $\frac{\partial \; ET}{\partial \; VPD}$ is less than zero.
+However, $\frac{\partial \; ET}{\partial \; VPD}$ evaluated at the average of all variables (e.g.
+$\sigma$, $T$, $c_s$, $VPD$) is only negative for CSH and GRA.
+And, DBF in addition to CRO experiences $\frac{\partial \; ET}{\partial \; VPD}$ < 0 less than half the time.
+These observations highlight the effect of the nonlinear function in Figure \ref{term3}: $\frac{\partial \; ET}{\partial \; VPD}$ has a much steeper slope when the function is negative, and is thus more likely to be large.
+
+The units of $\frac{\partial \; ET}{\partial \; VPD}$ make it difficult to interpret if $VPD$ is even a meaningful contributor to ET's variability.
+To better understand $VPD$'s contribution, we normalize $\frac{\partial \; ET}{\partial \; VPD}$ with $VPD$'s standard deviation to define a (linearized) relative change in ET for variations in $VPD$ .
+$VPD$'s contribution to ET's variability ranges between 30 - 40 W m$^{-2}$ for all PFTs except for CSH, which is about 100 W m$^{-2}$.
 Another meaningful comparison is to $\frac{\partial \; ET}{\partial \; R} \cdot std(R)$, as net radiation is generally the driver of ET (cite joe berry here).
-For all PFTs except for CSH $D$ contributes between 30 - 40 \% of $R$'s contribution to variability.
+For all PFTs except for CSH $VPD$ contributes between 30 - 40 \% of $R$'s contribution to variability.
 For CSH the portion is much larger, about 88 \%.
-$D$'s variability is certiantly a primary contributor to $ET$'s variability.
-
-So far, idealized plots and statistics have illuminated the form of $\frac{\partial \; ET}{\partial \; D}$ and how it varies with PFT.
-Large mean $\sigma$ and uWUE shifts CRO and DBF towards positive $\frac{\partial \; ET}{\partial \; D}$.
-However, the strongly nonlinear function of $\frac{\partial \; ET}{\partial \; D}$ at $\frac{\partial \; ET}{\partial \; D} < 0$ pushes $\overline{\frac{\partial \; ET}{\partial \; D}}$ negative for DBF (it does not do this for CRO because of CRO's high $g1$).
-ENF's low $g1$ value increases the dependence of $\frac{\partial \; ET}{\partial \; D}$ on $D$, and makes the function more strongly nonlinear.
-This has the side effect of pushing $\overline{\frac{\partial \; ET}{\partial \; D}}$ negative further than other PFTs for a given fraction $\frac{\partial \; ET}{\partial \; D} < 0$ and magnitude $\frac{\partial \; ET}{\partial \; D}(\overline{T,\ldots,D})$.
-GRA shows the opposite behavior; a relatively high $g1$  makes the function more linear, decreasing the magnitude of $-\overline{\frac{\partial \; ET}{\partial \; D}}$ for a given  [large] fraction $\frac{\partial \; ET}{\partial \; D} < 0$ and negative $\frac{\partial \; ET}{\partial \; D}(\overline{T,\ldots,D})$ (although aerodynamic conductance and Term 1 also probably have a role in this).
-Finally, low $uWUE$ of CSH pushes to toward by far the lowest values $\frac{\partial \; ET}{\partial \; D}$ (Figure \ref{term3}).
-Variability in $D$ accounts for the largest about of $ET$ variability for CSH.
-For the other PFTs, $D$ contributes less to $ET$ variability, but still represents about 15-20 \% of $R$'s contributions to ET variability.
+$VPD$'s variability is certiantly a primary contributor to $ET$'s variability.
+
+So far, idealized plots and statistics have illuminated the form of $\frac{\partial \; ET}{\partial \; VPD}$ and how it varies with PFT.
+Large mean $\sigma$ and uWUE shifts CRO and DBF towards positive $\frac{\partial \; ET}{\partial \; VPD}$.
+However, the strongly nonlinear function of $\frac{\partial \; ET}{\partial \; VPD}$ at $\frac{\partial \; ET}{\partial \; VPD} < 0$ pushes $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ negative for DBF (it does not do this for CRO because of CRO's high $g1$).
+ENF's low $g1$ value increases the dependence of $\frac{\partial \; ET}{\partial \; VPD}$ on $VPD$, and makes the function more strongly nonlinear.
+This has the side effect of pushing $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ negative further than other PFTs for a given fraction $\frac{\partial \; ET}{\partial \; VPD} < 0$ and magnitude $\frac{\partial \; ET}{\partial \; VPD}(\overline{T,\ldots,VPD})$.
+GRA shows the opposite behavior; a relatively high $g1$  makes the function more linear, decreasing the magnitude of $-\overline{\frac{\partial \; ET}{\partial \; VPD}}$ for a given  [large] fraction $\frac{\partial \; ET}{\partial \; VPD} < 0$ and negative $\frac{\partial \; ET}{\partial \; VPD}(\overline{T,\ldots,VPD})$ (although aerodynamic conductance and Term 1 also probably have a role in this).
+Finally, low $uWUE$ of CSH pushes to toward by far the lowest values $\frac{\partial \; ET}{\partial \; VPD}$ (Figure \ref{term3}).
+Variability in $VPD$ accounts for the largest about of $ET$ variability for CSH.
+For the other PFTs, $VPD$ contributes less to $ET$ variability, but still represents about 15-20 \% of $R$'s contributions to ET variability.
 
 \begin{table}
-\caption{Statistics of $\frac{\partial \; ET}{\partial \; D}$ as a function of PFT.}
+\caption{Statistics of $\frac{\partial \; ET}{\partial \; VPD}$ as a function of PFT.}
 \centering
@@
   \hline
-PFT & $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ & $\frac{\partial \; ET}{\partial \; D}\left(\overline{T, \ldots , D}\right)$ & $\frac{\partial \; ET}{\partial \; D}\left(\overline{T, \ldots , D}\right)*\text{std}(D)$ & $\frac{\frac{\partial \; ET}{\partial \; D}\left(\overline{T, \ldots , D}\right)*\text{std}(D)}{ \frac{\partial \; ET}{\partial \; R}\left(\overline{T, \ldots , D}\right)*\text{std}(R)}$ & fraction $\frac{\partial \; ET}{\partial \; VPD} < 0.$ \\
+PFT & $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ & $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{T, \ldots , VPD}\right)$ & $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{T, \ldots , VPD}\right)*\text{std}(VPD)$ & $\frac{\frac{\partial \; ET}{\partial \; VPD}\left(\overline{T, \ldots , VPD}\right)*\text{std}(VPD)}{ \frac{\partial \; ET}{\partial \; R}\left(\overline{T, \ldots , VPD}\right)*\text{std}(R)}$ & fraction $\frac{\partial \; ET}{\partial \; VPD} < 0.$ \\
   \hline
@@
 
-\subsection{Full observations of $\frac{\partial \; ET}{\partial \; D}$}
-
-Now that we have an intuitive understanding of $\frac{\partial \; ET}{\partial \; D}$'s behavior, we are equipped to interpret fully realistic plots of $\frac{\partial \; ET}{\partial \; D}$ for each PFT.
-Figure \ref{real} presents calculated $\frac{\partial \; ET}{\partial \; D}$ where, unless otherwise noted, all variables in Equation \ref{d_et} are allowed to vary.
-Each column is a different quantity related to $\frac{\partial \; ET}{\partial \; D}$, and each row is a different PFT.
+\subsection{Testing the theory - does the data match the model?}
+
+So far we have developed a theory for $\frac{\partial \; ET}{\partial \; VPD}$'s behavior.
+In particular, we determined a critical threshold $VPD_{crit}$ for $\frac{\partial \; ET}{\partial \; VPD}$.
+For $VPD < VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}  < 0$, and for $VPD > VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD} > 0$.
+This $VPD_{crit}$ is only a function of PFT.
+However, many assumptions, including constant uWUE, were made in deriving $VPD_{crit}$.
+So, we need to test if our theory and $VPD_{crit}$ hold up in the face of uncertainty.
+
+Figure \ref{real} presents calculated $\frac{\partial \; ET}{\partial \; VPD}$ where, unless otherwise noted, all variables in Equation \ref{d_et} are allowed to vary, including uncertainty.
+Each column is a different quantity related to $\frac{\partial \; ET}{\partial \; VPD}$, and each row is a different PFT.
 
 The full observations generally confirm expectations from Section \ref{results}.
-CRO has the most positive values of $\frac{\partial \; ET}{\partial \; D}$, $\frac{\partial \; ET}{\partial \; D}$ is almost always negative for CSH, and response depends more with the environmental conditions for the other PFTs (especially ENF).
-Through the columns of Figure \ref{real} we can see the impact of $\sigma$ and $g_a$ on the variability of $\frac{\partial \; ET}{\partial \; D}$.
+CRO has the most positive values of $\frac{\partial \; ET}{\partial \; VPD}$, $\frac{\partial \; ET}{\partial \; VPD}$ is almost always negative for CSH, and response depends more with the environmental conditions for the other PFTs (especially ENF).
+Through the columns of Figure \ref{real} we can see the impact of $\sigma$ and $g_a$ on the variability of $\frac{\partial \; ET}{\partial \; VPD}$.
 $g_a$'s scaling (included in columns 1 and 3) alters the magnitude considerably.
-$\sigma$ variability (included in columns 1 and 2) adds a lot of additional noise to the signal of $\frac{\partial \; ET}{\partial \; D}$, which is slightly undesirable given $\sigma$'s role in representing model and observational uncertainty.
-However, the general story with the noise appears to match the cleaner signal when $\sigma$ is help constant and $D_{ETmin}$ is clearly visible \explain{I really need to make these plots better - way too much overlapping of points that hurts the story}.
+$\sigma$ variability (included in columns 1 and 2) adds a lot of additional noise to the signal of $\frac{\partial \; ET}{\partial \; VPD}$, which is slightly undesirable given $\sigma$'s role in representing model and observational uncertainty.
+However, the general story with the noise appears to match the cleaner signal when $\sigma$ is help constant and $VPD_{crit}$ is clearly visible.
 One exception is possibly with GRA, for which uncertainty represented in $\sigma$ is high and causes the full complexity plots (Columns 1 and 2) to not match well with $\sigma$ held fixed (Columns 3 and 4).
 
-For ENF and GRA $D_{ETmin}$ does not appear to be only a function of $\sigma$ (most observable in Column 4).
-It turns out that the site to site variability in $\gamma$ causes $D_{ETmin}$ to vary, which is not discussed in the previous section.
-The impact is observable in both ENF and GRA, but especially for ENF which has a larger $\frac{\partial^2 \; ET}{\partial^2 \; D}$ than the other PFTs.
-
-In general the full complexity plots of $\frac{\partial \; ET}{\partial \; D}$ match our expectations, even with the large sensitivity to $\sigma$ measures of uncertainty observed in Figure \ref{term3}.
+For ENF and GRA $VPD_{crit}$ does not appear to be only a function of $\sigma$ (most observable in Column 4).
+It turns out that the site to site variability in $\gamma$ causes $VPD_{crit}$ to vary, which is not discussed in the previous section.
+The impact is observable in both ENF and GRA, but especially for ENF which has a larger $\frac{\partial^2 \; ET}{\partial^2 \; VPD}$ than the other PFTs.
+
+In general the full complexity plots of $\frac{\partial \; ET}{\partial \; VPD}$ match our expectations, even with the large sensitivity to $\sigma$ measures of uncertainty observed in Figure \ref{term3}.
 Our $\sigma$-based method of uncertainty propagation blurs the idealized expectations the most for GRA, and also has a considerable effect for CRO.
@@
 \includegraphics[width=\textwidth]{./fig06.png}
-\caption{Scatter plots of $\frac{\partial \; ET}{\partial \; D}$.
-Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; D}$, as labeled: Column 1 - $\frac{\partial \; ET}{\partial \; D}$; Column 2 - $\frac{\partial \; ET}{\partial \; D}$ normalized by $g_a$; Column 3 - $\frac{\partial \; ET}{\partial \; D}$ with $\sigma$ held fixed at PFT average; and Column 4 - $\frac{\partial \; ET}{\partial \; D}$ normalized by $g_a$ and with $\sigma$ held fixed.
+\caption{Scatter plots of $\frac{\partial \; ET}{\partial \; VPD}$.
+Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; VPD}$, as labeled: Column 1 - $\frac{\partial \; ET}{\partial \; VPD}$; Column 2 - $\frac{\partial \; ET}{\partial \; VPD}$ normalized by $g_a$; Column 3 - $\frac{\partial \; ET}{\partial \; VPD}$ with $\sigma$ held fixed at PFT average; and Column 4 - $\frac{\partial \; ET}{\partial \; VPD}$ normalized by $g_a$ and with $\sigma$ held fixed.
 For reference, lines corresponding to RH = 20\% and RH = 90 \% are drawn.
@@
 \includegraphics[width=\textwidth]{./fig06b.png}
-\caption{****alternate Fig 06****  Scatter plots of $\frac{\partial \; ET}{\partial \; D}$.
-Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; D}$, as labeled.
-If I end up using this, I could also draw on the curve of $D_{ETmin}$ with $\overline{\frac{\text{LAI}}{\text{LAI$_{ref}$}}}$.
+\caption{****alternate Fig 06****  Scatter plots of $\frac{\partial \; ET}{\partial \; VPD}$.
+Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; VPD}$, as labeled.
+If I end up using this, I could also draw on the curve of $VPD_{crit}$ with $\overline{\frac{\text{LAI}}{\text{LAI$_{ref}$}}}$.
 }
@@
 
-The idealized representation of ET used here is successful in developing intuition for how ET responds to changes in $D$.
+The idealized representation of ET used here is successful in developing intuition for how ET responds to changes in $VPD$.
 This intuition will aid the community in interpreting observations and output from sophisticated full complexity climate models.
@@
 \begin{itemize}
-  \item Aerodynamic resistance plays an important role of scaling $\frac{\partial \; ET}{\partial \; D}$. This is a leading order effect for observing higher magnitude responses in DBF and ENF.
+  \item Aerodynamic resistance plays an important role of scaling $\frac{\partial \; ET}{\partial \; VPD}$. This is a leading order effect for observing higher magnitude responses in DBF and ENF.
   \item In general, CSH has the most negative (i.e.
-ET reduced) response to increases in $D$ (atmospheric drying).
+ET reduced) response to increases in $VPD$ (atmospheric drying).
 So CSH plants will almost always try and conserve water, effectively reducing ET with dry atmospheric perturbation.
-    \item Additionally for CSH, $D$ variability contributes the most to $ET$ variability.
-\item CRO has the most positive response (i.e. ET increased) in response to increases in $D$. This is consistent with CROs that may be evolved or bred to thrive in non-water-limited environments.
+    \item Additionally for CSH, $VPD$ variability contributes the most to $ET$ variability.
+\item CRO has the most positive response (i.e. ET increased) in response to increases in $VPD$. This is consistent with CROs that may be evolved or bred to thrive in non-water-limited environments.
 \item The response is more a function of the environment for DBF, ENF, and GRA.
@@
 Additionally, Equation \ref{et} gives provides an estimate of ET that requires no additional information beyond that required to calculate PET.
-Given that for all PFTs, with the exception of CRO, we found a high frequency and magnitude of negative $\frac{\partial \; ET}{\partial \; D}$, PET is a physically unrealisitic representation of ET for vegetated surfaces and PET-based drought metrics are not usefull.
+Given that for all PFTs, with the exception of CRO, we found a high frequency and magnitude of negative $\frac{\partial \; ET}{\partial \; VPD}$, PET is a physically unrealisitic representation of ET for vegetated surfaces and PET-based drought metrics are not usefull.
 We advocate for drought metrics using Equation \ref{et} instead of PET.

### 2017-11-27 doc/paper/vpd_et_paper.tex
trouble with the optmizier just for getting optmized vpd to compare to theory

@@
 % PLEASE DO NOT USE YOUR OWN MACROS
-% DO NOT USE \newcommand, \renewcommand, or \def, etc.
+% DSO NOT USE \newcommand, \renewcommand, or \def, etc.
 %
@@
   $c_p$  & specific heat capacity of air at constant pressure & J K$^{-1}$ kg$^{-1}$ \\
-  $VPD$  & VPD & Pa  \\
+  $VPD$  & vapor pressure deficit & Pa  \\
   $\gamma$  & psychrometric constant & Pa K$^{-1}$   \\

### 2017-11-27 src/paper_figure_code/fig05.py

@@ -197,7 +197,8 @@ axs[0].set_ylabel(paren_string)
 axs[0].plot(axs[0].get_xlim(), [0., 0.], 'k--', linewidth=1.0)
 plt.setp(axs[2].get_yticklabels(), visible=False)
 axs[-1].set_ylabel(r'VPD')#$_{ETmin}$')
-axs[2].set_title('D_{crit} (where $\frac{\partial \; ET}{\partial \; D} = 0$)')
+axs[2].set_title(r'D$_{crit}$'\
+                 r' (where $\frac{\partial \; ET}{\partial \; D} = 0$)')
 # axs[2].set_ylim([0., np.around(df.vpd.quantile(q=0.95), decimals=-2)])
 axs[2].set_ylim([0., 4000.])
 axs[1].set_ylim([0.5,5.5])

### 2017-11-27 src/paper_figure_code/tables.py

@@ -4,6 +4,7 @@ This script makes all figs for the paper
 """
 
 from shared_functions import *
+import scipy.optimize
 
 ###### table 5 ####
 def frequency(_df):
@@ -51,4 +52,50 @@ columns = ['r_moist', 'c_a', 'gamma', 'lai',\
            'sigma_dot_uwue', 'vpd_crit']
 print('\n vpd critical table')
 print(mean_df.loc[:, columns])
+
+def optimizer(vpd, *args):
+  """this finds the otopmism vpd to get the most true hits"""
+  print('did it!')
+  _dff = args[0]
+  print('was args the problem?', _dff)
+  try:
+    count_true_less = float(_dff.loc[((_dff.vpd < vpd)\
+                                      & (_dff.d_et < 0.0)), 'd_et'].count())
+  except ValueError:
+    count_true_less = 0.0
+  try:
+    count_true_more = float(_dff.loc[((_dff.vpd > vpd)\
+                                      & (_dff.d_et > 0.0)), 'd_et'].count())
+  except ValueError:
+    count_true_more = 0.0
+  output = (count_true_less+count_true_more)
+  return output
+
+def vpd_statistics(_df):
+  """computes fraction of time theory is correct, and mean det/dvpd as a
+  function of theory"""
+  columns = ['frac_correct_less', 'frac_incorrect_less',\
+             'frac_correct_more', 'frac_incorrect_more',\
+             'mean_less', 'mean_more', 'mean'] # ,\
+             # 'optimized_vpd', 'true_vpd']
+  vpd_crit = mean_df.loc[_df.pft.iloc[0], 'vpd_crit']
+  less_df = _df.d_et.loc[(_df.vpd < vpd_crit)]
+  more_df = _df.d_et.loc[(_df.vpd > vpd_crit)]
+  # print('test', vpd_crit)
+  # vpd_opt = scipy.optimize.golden(optimizer, args=(_df))# ,\
+  #                                 # brack=(0.0, 5.0*vpd_crit),\
+  # print(vpd_opt)
+  vpd = np.linspace(0.0, 2.0*vpd_crit)
+  hits = 
+
+  data = [float(less_df.loc[less_df < 0.0].count())/float(less_df.count()),
+          float(less_df.loc[less_df > 0.0].count())/float(less_df.count()),
+          float(more_df.loc[more_df > 0.0].count())/float(more_df.count()),
[…]

### 2017-11-29 doc/agu_2017/vpd.tex
think done with rough draft

--- 2017-10-27 doc/eee_symposium/vpd.tex
@@
 
-\title[Your Short Title]{When does atmospheric dryness drive or reduce evapotranspiration?}
+\newcommand{\specialcell}[2][c]{%
+  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
+
+\title[Your Short Title]{When does vapor pressure deficit drive or reduce evapotranspiration?}
 \author{Adam Massmann,  Pierre Gentine and Changjie Lin}
-\institute{EEE Graduate Student Symposium}
+\institute{AGU Fall Meeting}
 \date{October 27th, 2017}
@@
 \section{Introduction}
-\begin{frame}{When does atmospheric dryness (VPD) drive or reduce evapotranspiration (ET)?}
+\begin{frame}{Does VPD drive or reduce ET? - hydrometeorologists' perspective}
   \begin{Huge}
@@
 
-\begin{frame}%{When does atmospheric dryness (VPD) drive or reduce evapotranspiration (ET)?}
+\begin{frame}{Does VPD drive or reduce ET? - physiologists' perspective}
   \begin{figure}
-  \includegraphics[width=2.70in]{stomata.png}% was O(1.5in)
+  \includegraphics[width=2.25in]{stomata.png}% was O(1.5in)
   \caption{from evolution.berkeley.edu}
@@
 \end{frame}
-% develop theory, present vpd figure (add in vpd = 0 locations on line plot below), then test theory with data. add vertical line on vpd plots
+
 \section{Method}
@@
 
-\begin{frame}{Consequences of theory - ``sign'' term}
-Below is a little opaque (at least to me):
-\[\frac{\partial \, ET}{\partial \, VPD} = \text{scaling} \cdot  \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]
-so do expansion about $g_1 >> \sqrt{VPD}$:
-\footnotesize
-\[ \frac{\partial \, ET}{\partial \, VPD} \approx \text{scaling} \cdot  \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{1}{g_1} - \frac{3\sqrt{VPD}}{2 g_1^2} +  \mathcal{O}\left(\frac{\sqrt{VPD}}{g_1}\right)^{2} \right) \right)\]
-% \large
-% and:
-%   \[ \sqrt{VPD} \Big|_{\frac{\partial ET}{\partial VPD} = 0} \approx  \frac{2 g_1}{3}  \left( -\frac{ c_p \, g_1\,  1.6  \; R*\; \text{ uWUE }}{R_{air} \gamma c_s } + 1  \right) \]
-
-%   \[ VPD \Big|_{\frac{\partial ET}{\partial VPD} = 0} = \frac{R_{air}}{4 \, c_p} \left[ \frac{\gamma  c_s}{R* \, 1.6\, \text{uWUE}} + \sqrt{\left(g_1 + 8\, \frac{c_p}{R_{air}}\right) \frac{\gamma c_s}{R* \, 1.6\, \text{uWUE}}} - 4 \, g_1 \frac{ c_p}{R_{air}}\right] \]
-\end{frame}
-
 \begin{frame}{``Sign'' Term}
@@
 \section{Test Theory}
+\begin{frame}{The theory seems nice, but we should test with data!}
+    \begin{itemize}
+    \item Introduce a free uncertainty parameter $\sigma$ to Penman Monteith:
+      \[ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_s \sqrt{VPD} }{ R* \; 1.6\; \boldsymbol{\sigma} \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}\]
+    \item At each observation calculate a unique $\sigma$:
+      \small
+      \[\sigma = - \frac{g_a \gamma c_s \sqrt{VPD} P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p VPD\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})}\]
+      \normalsize
+    \item Then propagate uncertainty forward by including $\sigma$ in the derivative:
+    \end{itemize}
+      \[\frac{\partial \;  ET}{\partial \; VPD} = \text{scaling} \cdot \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_s }{1.6 \; R*\; \sigma \; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]
+
+  \end{frame}
+
     \begin{frame}{Test theory with FLUXNET data - CSH}
@@
 Applicability to CRO and GRA is more uncertain.
-  \item For plants evolved to prioritize water conservation over production, atmospherc demand is less likely to dominate.
-  \item All plant types except for crops had common occurence of decreases in ET increases in VPD (plant response dominates).
+  \item For plants evolved to prioritize water conservation over production, atmospheric demand is less likely to dominate.
+  \item All plant types except for crops had common occurrence of decreases in ET with increases in VPD (plant response dominates).
 Therefore, the hydrometeorologists perspective is often wrong, and drought indices built upon assumptions about atmospheric demand are unrealistic for vegetated surfaces.
@@
   \end{itemize}
+\end{frame}
+
+\section{References}
+\begin{frame}{References}
+  \AtNextBibliography{\small}
+  \printbibliography
+  \scriptsize
+  \begin{itemize}
+  \item This work used eddy covariance data acquired and shared by the FLUXNET community, including these networks: AmeriFlux, AfriFlux, AsiaFlux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont, ChinaFlux, Fluxnet-Canada, GreenGrass, ICOS, KoFlux, LBA, NECC, OzFlux-TERN, TCOS-Siberia, and USCCC.
+The ERA-Interim reanalysis data are provided by ECMWF and processed by LSCE.
+The FLUXNET eddy covariance data processing and harmonization was carried out by the European Fluxes Database Cluster, AmeriFlux Management Project, and Fluxdata project of FLUXNET, with the support of CDIAC and ICOS Ecosystem Thematic Center, and the OzFlux, ChinaFlux and AsiaFlux offices.
+    \end{itemize}
+\end{frame}
+
+\section{Extra Slides}
+
+\begin{frame}{Extra slide - statistics}
+  \begin{table}
+  \label{vpd_crit}
+\caption{More quantitative test of theory.}
+\centering
+\begin{tabular}{l c c c c c c}
+  \hline
+  PFT & \specialcell{Fraction of Obs.\\Theory is Correct} & Mean ($\frac{\partial \; ET}{\partial \; VPD} < 0$) & Mean ($\frac{\partial \; ET}{\partial \; VPD} > 0$)\\
+  \hline
+  CRO & 0.566517 & -0.209152 &  0.005856\\
+  CSH & 0.931660 & -0.264746 &       NaN\\
+  DBF & 0.633363 & -0.135679 &  0.042910\\
+  ENF & 0.633138 & -0.150665 &  0.029281\\
+  GRA & 0.442306 & -0.042158 & -0.042480\\
+  \hline
+\end{tabular}
+\end{table}
+\end{frame}
+
+\begin{frame}{Extra slide - is theory VPD$_{crit}$ optimum?}
+  \includegraphics[width=\textwidth]{for_agu.pdf}
 \end{frame}
@@
 
-\section{References}
-\begin{frame}{References}
-  \AtNextBibliography{\small}
-  \printbibliography
-  \scriptsize
-  \begin{itemize}
-  \item This work used eddy covariance data acquired and shared by the FLUXNET community, including these networks: AmeriFlux, AfriFlux, AsiaFlux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont, ChinaFlux, Fluxnet-Canada, GreenGrass, ICOS, KoFlux, LBA, NECC, OzFlux-TERN, TCOS-Siberia, and USCCC.
-The ERA-Interim reanalysis data are provided by ECMWF and processed by LSCE.
-The FLUXNET eddy covariance data processing and harmonization was carried out by the European Fluxes Database Cluster, AmeriFlux Management Project, and Fluxdata project of FLUXNET, with the support of CDIAC and ICOS Ecosystem Thematic Center, and the OzFlux, ChinaFlux and AsiaFlux offices.
-    \end{itemize}
-\end{frame}
-
 % \begin{table}

### 2017-11-29 doc/agu_2017/vpd.tex
what I presented

@@
 \author{Adam Massmann,  Pierre Gentine and Changjie Lin}
-\institute{AGU Fall Meeting}
-\date{October 27th, 2017}
+\institute{AGU Fall Meeting (Rough Draft)}
+\date{December 14th, 2017}
 
@@
       \[ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_s \sqrt{VPD} }{ R* \; 1.6\; \boldsymbol{\sigma} \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}\]
-    \item At each observation calculate a unique $\sigma$:
+    \item At each observation from FLUXNET (56 sites across globe) calculate a unique $\sigma$:
       \small
@@
 The FLUXNET eddy covariance data processing and harmonization was carried out by the European Fluxes Database Cluster, AmeriFlux Management Project, and Fluxdata project of FLUXNET, with the support of CDIAC and ICOS Ecosystem Thematic Center, and the OzFlux, ChinaFlux and AsiaFlux offices.
+    \item This material is based upon work supported by the National Science Foundation Graduate Research Fellowship under Grant No.
+DGE 16-44869.
+Any opinion, findings, and conclusions or recommendations expressed in this material are those of the authors(s) and do not necessarily reflect the views of the National Science Foundation.
     \end{itemize}

### 2017-12-07 doc/qualifying_exam/vpd.tex
think done with qualifyin exam presentation

--- 2017-11-29 doc/agu_2017/vpd.tex
@@
 
-\title[Your Short Title]{When does vapor pressure deficit drive or reduce evapotranspiration?}
-\author{Adam Massmann,  Pierre Gentine and Changjie Lin}
-\institute{AGU Fall Meeting (Rough Draft)}
-\date{December 14th, 2017}
+\title[Your Short Title]{Qualifying Exam: Examples of my approach and research}
+\author{Adam Massmann}
+\date{December 7th, 2017}
 
@@
 \section{Introduction}
-\begin{frame}{Does VPD drive or reduce ET? - hydrometeorologists' perspective}
-  \begin{Huge}
-  \[VPD = (1-RH)\cdot e_s (T)\]
-\end{Huge}
+\begin{frame}{Current project - When does vapor pressure deficit (VPD) drive or reduce evapotranspiration ET?}
+  \begin{columns}[T,onlytextwidth]
+    \begin{column}{0.5\textwidth}
+      \begin{minipage}{\textwidth}
+
         \begin{itemize}
-  \item \textbf{Hydrometeorologists} would say that an increase in VPD (increase in \textbf{atmospheric demand}) would drive an \textbf{increase in ET}.
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Does VPD drive or reduce ET? - physiologists' perspective}
-  \begin{figure}
-  \includegraphics[width=2.25in]{stomata.png}% was O(1.5in)
-  \caption{from evolution.berkeley.edu}
-\end{figure}
-  \begin{itemize}
+        \item A \textbf{hydrometeorologist} might say that an increase in VPD (increase in \textbf{atmospheric demand}) would drive an \textbf{increase in ET}.
         \item However, \textbf{plant physiologists} know that plants have evolved to use stomata to conserve and regulate water use.
@@
         \end{itemize}
+      \end{minipage}
+    \end{column}
+    \begin{column}{0.5\textwidth}
+      \begin{minipage}{\textwidth}
+        \begin{figure}
+          \includegraphics[width=2.00in]{stomata.png}% was O(1.5in)
+          \caption{from evolution.berkeley.edu}
+        \end{figure}
+      \end{minipage}
+    \end{column}
+  \end{columns}
 \end{frame}
 
-\begin{frame}{The question is, which effect dominates with an increase in VPD: plant response (decrease in ET) or atmospheric demand (increase in ET)?}
+\section{Method}
+
+\begin{frame}{My approach}
   \begin{itemize}
-  \item It should be a function of plant type and the environment:
+  \item Use physically reasonable assumptions guided by intuition and literature to simplify the problem and develop a transparent and intuitive theory (or hypothesis).
+The goal is to capture the leading order behavior of the system.
     \begin{itemize}
-    \item Plants that are evolved to conserve water will tend to reduce ET with increases in VPD.
-    \item However the environment can overwhelm plant response: at some threshold (i.e.
-very high VPD) the atmospheric demand should dominate and plants will not be able to conserve water, no matter how much they have evolved to do so.
+    \item Here we start with Penman Monteith, and apply an assumption of constant uWUE (recently identified by \cite{Zhou_2016}) which gives a differentiable function of ET.
+    \end{itemize}
+  \item Test theory with observational or sophisticated model data to see if theory is valid, and to what extent it may generalize.
+    \begin{itemize}
+    \item Here we use FLUXNET2015 data.
     \end{itemize}
@@
 
-\section{Method}
-\begin{frame}{Use physically reasonable assumptions to develop a theory.}
-  We can use Penman-Monteith (PM) to estimate ET:
-  \[ET = \frac{\Delta R + g_a \rho_a c_p \, VPD}{\Delta + \gamma(1 + \frac{g_a}{g_s})},\]
-\textbf{Problem}: $g_s$ (stomatal conductance) is a function of photosynthesis, which is a function of ET itself.  So ET in PM is really an implicit function of itself and we cannot take derivatives!
-\end{frame}
+\section{Results - theory}
+\begin{frame}{Results - theory's form for ET response to VPD perturbations}
+  \begin{tikzpicture}
+    \node[anchor=south west,inner sep=0] (image) at (0,0) {\adjincludegraphics[width=3.5in, trim={0 {0.51\height} 0 0}, clip]{fig05.pdf}};
+        \begin{scope}[x={(image.south east)},y={(image.north west)}]
+        %% next four lines will help you to locate the point needed by forming a grid. comment these four lines in the final picture.↓
+       % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
+       % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
+       % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
+       % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
+       %% upto here
+        \draw[-latex, line width=2pt] (0.83,0.725) --++(0.0in,0.3in)node[anchor=south] {ET Increasing};
+        \draw[-latex, line width=2pt] (0.83,0.725) --++(0.0in,-0.3in)node[anchor=north] {ET Decreasing};
 
-\begin{frame}{Use physically reasonable assumptions to develop a theory.}
-Apply a constant uWUE assumption (conserved within plant type; see \cite{Zhou_2016}):
-\[uWUE = \frac{GPP \cdot \sqrt{VPD}}{ET},\]
-To derive a new form of Penman-Monteith without implicit ET dependence:
-  \[  ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p \, VPD}{R_{air}} -  \frac{\gamma c_s \sqrt{VPD} }{ R* \; 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right)\footnote{note that all terms are known }}{ \Delta + \gamma}\]
-\end{frame}
-
-\begin{frame}{Use physically reasonable assumptions to develop a theory.}
-  With our new form of Penman-Monteith we can now take derivatives, giving:
-  \[\frac{\partial \;  ET}{\partial \, VPD} = \frac{2 g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]
-  In the interest of time, we will just focus in the ``sign'' term:
-  \[\sign \left[\frac{\partial \;  ET}{\partial \, VPD}\right] = \sign \left[  \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right) \right] \]
-
-\end{frame}
-
-\section{Results - theory}
-\begin{frame}{Consequences of theory - ``sign'' term}
-  \[\frac{\partial \, ET}{\partial \, VPD} = \text{scaling} \cdot \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]
-  \begin{itemize}
-  \item $c_p$ and $R*$ are constants
-  \item $R_{air}$, $\gamma$, and $c_s$ are approximately constant (relative to $\sqrt{VPD}$)
-  \item uWUE and $g_1$ are constants within plant type (i.e. grass, crops, deciduous forest, evergreen needleleaf forest, shrub)
-  \item Which just leaves VPD
-  \end{itemize}
-  So within each plant type, whether the atmospheric demand (ET increasing with VPD) or plant response (ET decreasing with VPD) dominates is essentially just a function of VPD!
-\end{frame}
-
-\begin{frame}{``Sign'' Term}
-  \begin{figure}
-    \adjincludegraphics[width=3.5in, trim={0 {0.51\height} 0 0}, clip]{fig05.pdf}
-  \end{figure}
+        % \draw[dashed,-latex, line width=5pt] (0.8,0.75) --++(0.0in,+0.5in)node[anchor=north] {ET Decreasing};
+    \end{scope}
+    % \node[align=center,red,font={\large}] at (image.southwest) {ET decreasing};
+  \end{tikzpicture}
 \end{frame}
@@
 \section{Test Theory}
-\begin{frame}{The theory seems nice, but we should test with data!}
-    \begin{itemize}
-    \item Introduce a free uncertainty parameter $\sigma$ to Penman Monteith:
-      \[ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_s \sqrt{VPD} }{ R* \; 1.6\; \boldsymbol{\sigma} \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}\]
-    \item At each observation from FLUXNET (56 sites across globe) calculate a unique $\sigma$:
-      \small
-      \[\sigma = - \frac{g_a \gamma c_s \sqrt{VPD} P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p VPD\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})}\]
-      \normalsize
-    \item Then propagate uncertainty forward by including $\sigma$ in the derivative:
-    \end{itemize}
-      \[\frac{\partial \;  ET}{\partial \; VPD} = \text{scaling} \cdot \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_s }{1.6 \; R*\; \sigma \; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]
-
-  \end{frame}
-
-    \begin{frame}{Test theory with FLUXNET data - CSH}
-                \includegraphics[width=3.5in]{csh_eee.png}
-     \end{frame}
-
-    \begin{frame}{Test theory with FLUXNET data - ENF}
+\begin{frame}{Testing Theory - the good: evergreen forest}
   \includegraphics[width=3.5in]{enf_eee.png}
@@
 
-    \begin{frame}{Test theory with FLUXNET data - DBF}
-                \includegraphics[width=3.5in]{dbf_eee.png}
-     \end{frame}
-
-    \begin{frame}{Test theory with FLUXNET data - CRO}
-                \includegraphics[width=3.5in]{cro_eee.png}
-     \end{frame}
-
-    \begin{frame}{Test theory with FLUXNET data - GRA}
+\begin{frame}{Testing Theory - the bad: grassland}
   \includegraphics[width=3.5in]{gra_eee.png}
@@
 
-% \begin{frame}{Summary of theory}
-%   \begin{itemize}
-%   \item We use \cite{Zhou_2016}'s uWUE to derive a new analytically tractable form of PM.
-%   \item This new analysis suggests that the ``tipping'' point for which atmospheric demand overwhelms plant response will be almost exclusively a function of VPD.
-%     \begin{itemize}
-%     \item For each PFT there will be a VPD$_{crit}$ above which atmospheric demand will dominate and ET will increase with VPD.
-%     \end{itemize}
-%   \item Plant types evolved to conserve water (CSH) have a higher VPD$_{crit}$ than plants evolved (or bred) to use water and prioritize GPP (CRO).
-Trees and grasslands are somewhere between these two extremes.
-%   \item Aerodynamic conductance scales the response, so plants with large surface roughness will be more likely to have a larger response as there is less resistance between the surface and the atmosphere.
-%   \end{itemize}
-% \end{frame}
+\section{Conclusions}
 
-\section{Conclusions}
-% \begin{frame}{Summary}
-%   \begin{itemize}
-%   \item Theory finds that the ``tipping'' point for which atmospheric demand overwhelms plant response will be almost exclusively a function of VPD.
-Plant types evolved to conserve water (CSH) have a higher VPD$_{crit}$ (and more negative ET response) than plants evolved (or bred) to use water and prioritize GPP (CRO).
-%   \item On average, ecosystem response to VPD follows roughly what we might expect: CRO (prioritize GPP) has positive ET response to VPD, while all others have a negative response.
-Ordering by increasing magnitude of negative response gives: DBF, GRA, ENF, CSH; which roughly correspond to expectations for increasing water conservation as a function of PFT.
-%   \item Uncertainty is high, especially for CRO and GRA. However, inclusion of uncertainty does not change the story for ENF or CSH.
-%   \end{itemize}
-% \end{frame}
-
-\begin{frame}{Summary}
+\begin{frame}{Summary - When does VPD reduce ET?}
   \small
   \begin{itemize}
-  \item Derived a new form of Penman-Monteith that explicitly calculates ET given published constants.
-  \item Developed a theory for when atmospheric dryness will increase ET (atmospheric demand dominates) vs.
-decrease ET (plant response dominates).
-Theory tests well using real data from FLUXNET2015 on plant types CSH, ENF, and DBF.
-Applicability to CRO and GRA is more uncertain.
-  \item For plants evolved to prioritize water conservation over production, atmospheric demand is less likely to dominate.
-  \item All plant types except for crops had common occurrence of decreases in ET with increases in VPD (plant response dominates).
-Therefore, the hydrometeorologists perspective is often wrong, and drought indices built upon assumptions about atmospheric demand are unrealistic for vegetated surfaces.
-  \item Our form of Penman-Monteith is capable of estimating ET realistically and could be used as a replacement for PET in drought indices (e.g. Palmer Drought Severity Index - PDSI).
+  \item Successfully simplified the problem to develop a theory for when atmospheric dryness will increase ET (atmospheric demand dominates) vs. decrease ET (plant response dominates).
+  \item Tested the theory with real data from FLUXNET2015: shrubs and forests tested well, while the theory's applicability to crops and grass is more uncertain.
+  \item All plant types except for crops has a common occurrence of decreases in ET with increases in VPD (plant response dominates).
+Therefore, any drought indices which assume that atmospheric demand dominates (e.g PET-based metrics) are likely unrealistic for vegetated surfaces.
   \end{itemize}
@@
 \section{Extra Slides}
+
+    \begin{frame}{Test theory with FLUXNET data - CSH}
+                \includegraphics[width=3.5in]{csh_eee.png}
+     \end{frame}
+
+    \begin{frame}{Test theory with FLUXNET data - DBF}
+                \includegraphics[width=3.5in]{dbf_eee.png}
+     \end{frame}
+
+    \begin{frame}{Test theory with FLUXNET data - CRO}
+                \includegraphics[width=3.5in]{cro_eee.png}
+     \end{frame}
 

### 2017-12-12 doc/agu_2017/vpd.tex
starting final edits for AGU
some small progress

@@
-\documentclass{beamer}
+\documentclass[aspectratio=169]{beamer}
 %
@@
 \author{Adam Massmann,  Pierre Gentine and Changjie Lin}
-\institute{AGU Fall Meeting (Rough Draft)}
+\institute{AGU Fall Meeting}
 \date{December 14th, 2017}
@@
 \section{Introduction}
-\begin{frame}{Does VPD drive or reduce ET? - hydrometeorologists' perspective}
+\begin{frame}{Does VPD drive or reduce ET? - atmospheric demand perspective}
   \begin{Huge}
@@
   \begin{itemize}
-  \item \textbf{Hydrometeorologists} would say that an increase in VPD (increase in \textbf{atmospheric demand}) would drive an \textbf{increase in ET}.
+  \item Increase in VPD (\textbf{increase in atmospheric demand}) drives  an \textbf{increase in ET}.
   \end{itemize}
@@
 
-\begin{frame}{Does VPD drive or reduce ET? - physiologists' perspective}
+\begin{frame}{Does VPD drive or reduce ET? - plant response perspective}
+  \begin{columns}
+    \begin{column}{0.5\textwidth}
+      \begin{itemize}
+      \item However, plants evolved to use stomata to conserve and regulate water use. So \textbf{stomata closure} in response to increases in VPD may \textbf{decrease ET}.
+      \end{itemize}
+    \end{column}
+    \begin{column}{0.5\textwidth}
       \begin{figure}
@@
       \end{figure}
-  \begin{itemize}
-  \item However, \textbf{plant physiologists} know that plants have evolved to use stomata to conserve and regulate water use.
-So \textbf{stomata closure} in response to increases in VPD may \textbf{decrease ET}.
-  \end{itemize}
+    \end{column}
+  \end{columns}
 \end{frame}
@@
   \begin{itemize}
-  \item It should be a function of plant type and the environment:
+  \item Apriori intuition:
     \begin{itemize}
-    \item Plants that are evolved to conserve water will tend to reduce ET with increases in VPD.
-    \item However the environment can overwhelm plant response: at some threshold (i.e.
-very high VPD) the atmospheric demand should dominate and plants will not be able to conserve water, no matter how much they have evolved to do so.
+    \item Should be a function of plant type. e.g. plants that are evolved to conserve water will tend to reduce ET with increases in VPD.
+    \item But the environment still matters: if the atmosphere dries enough  e.g. in very dry conditions the plants will not be able to conserve water, no matter how much they have evolved to do so.
     \end{itemize}

### 2017-12-12 doc/agu_2017/vpd.tex
back to the conference

@@
   \begin{itemize}
-  \item Apriori intuition:
+  \item Prior intuition:
     \begin{itemize}
-    \item Should be a function of plant type. e.g. plants that are evolved to conserve water will tend to reduce ET with increases in VPD.
-    \item But the environment still matters: if the atmosphere dries enough  e.g. in very dry conditions the plants will not be able to conserve water, no matter how much they have evolved to do so.
+    \item Should be a function of plant type: plants that are evolved to conserve water will tend to reduce ET with increases in VPD.
+    \item But the environment still matters: if the atmosphere dries enough plant water conservation strategies will max out and atmospheric demand drives increases in ET with increases in VPD.
     \end{itemize}

### 2017-12-13 doc/agu_2017/vpd.tex
think I got the poster plots pretty much done

@@
 \usepackage{tikz}
+\usepackage{xcolor}
 \graphicspath{{figures/}}
@@
   \begin{itemize}
-  \item Prior intuition:
+  \item Prior expectation:
     \begin{itemize}
     \item Should be a function of plant type: plants that are evolved to conserve water will tend to reduce ET with increases in VPD.
-    \item But the environment still matters: if the atmosphere dries enough plant water conservation strategies will max out and atmospheric demand drives increases in ET with increases in VPD.
+    \item But the environment still matters: if the atmosphere dries enough, plant water conservation strategies will reach their limit and atmospheric demand will drive increases in ET with increases in VPD.
     \end{itemize}
@@
 \section{Method}
-\begin{frame}{Use physically reasonable assumptions to develop a theory.}
-  We can use Penman-Monteith (PM) to estimate ET:
-  \[ET = \frac{\Delta R + g_a \rho_a c_p \, VPD}{\Delta + \gamma(1 + \frac{g_a}{g_s})},\]
-\textbf{Problem}: $g_s$ (stomatal conductance) is a function of photosynthesis, which is a function of ET itself.  So ET in PM is really an implicit function of itself and we cannot take derivatives!
+\begin{frame}{Develop a theory to quantify VPD effects}
+  \begin{itemize}
+  \item The goal is to simply the problem (increase transparency) while still capturing the leading order behavior the system. (intrinsic knowledge)
+    \begin{itemize}
+    \item While simplifying the system aids intrinsic understanding, simplifying assumptions sacrifice physical realism.
+    \end{itemize}
+  \item Because we run the risk of over-simplification, we will use FLUXNET2015 data to test how well our theory matches the data. 
+  \end{itemize}
 \end{frame}
 
-\begin{frame}{Use physically reasonable assumptions to develop a theory.}
+\begin{frame}{Simple theory - start with Penman-Monteith}
+  We can use Penman-Monteith (PM) to estimate ET:
+  \[ET = \frac{\Delta R + g_a \rho_a c_p \, VPD}{\Delta + \gamma(1 + \frac{g_a}{\textcolor{magenta}{\boldsymbol{g_s}}})},\]
+\textbf{Problem}: \Large $\textcolor{magenta}{\boldsymbol{g_s}}$ \normalsize (stomatal conductance) is a function of photosynthesis, which is a function of ET itself.
+So ET in Penman-Monteith is really an implicit function of itself and we cannot take derivatives!
+\end{frame}
+
+\begin{frame}{Use physically reasonable assumptions remove implicit dependence}
 Apply a constant uWUE assumption (conserved within plant type; see \cite{Zhou_2016}):
@@
 To derive a new form of Penman-Monteith without implicit ET dependence:
-  \[  ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p \, VPD}{R_{air}} -  \frac{\gamma c_s \sqrt{VPD} }{ R* \; 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right)\footnote{note that all terms are known }}{ \Delta + \gamma}\]
+  \[  ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p \, VPD}{R_{air}} -  \frac{\gamma c_s \sqrt{VPD} }{ R* \; 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right)}{ \Delta + \gamma}\]
 \end{frame}
 
-\begin{frame}{Use physically reasonable assumptions to develop a theory.}
+\begin{frame}{Now just take $\frac{\partial \; ET}{\partial \; VPD}$}
   With our new form of Penman-Monteith we can now take derivatives, giving:
@@
 \begin{frame}{Consequences of theory - ``sign'' term}
-  \[\frac{\partial \, ET}{\partial \, VPD} = \text{scaling} \cdot \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]
+  \[\frac{\partial \, ET}{\partial \, VPD} = \text{scaling} \cdot \left(\frac{\textcolor{magenta}{\boldsymbol{ c_p}}}{R_{air}} - \frac{\gamma c_s }{1.6 \; \textcolor{magenta}{\boldsymbol{R*}}\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]
+  \begin{itemize}
+  \item $\textcolor{magenta}{\boldsymbol{c_p}}$ and $\textcolor{magenta}{\boldsymbol{R*}}$ are constants
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Consequences of theory - ``sign'' term}
+  \[\frac{\partial \, ET}{\partial \, VPD} = \text{scaling} \cdot \left(\frac{ c_p}{\textcolor{magenta}{\boldsymbol{R_{air}}}} - \frac{\textcolor{magenta}{\boldsymbol{\gamma}} \textcolor{magenta}{\boldsymbol{c_s}} }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]
+  \begin{itemize}
+  \item $c_p$ and $R*$ are constants
+  \item $\textcolor{magenta}{\boldsymbol{R_{air}}}$, $\textcolor{magenta}{\boldsymbol{\gamma}}$, and $\textcolor{magenta}{\boldsymbol{c_s}}$ are approximately constant (relative to $\sqrt{VPD}$)
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Consequences of theory - ``sign'' term}
+  \[\frac{\partial \, ET}{\partial \, VPD} = \text{scaling} \cdot \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ \textcolor{magenta}{\textbf{uWUE }}}} \left( \frac{2 \textcolor{magenta}{\boldsymbol{g1}} + \sqrt{VPD}}{2 (\textcolor{magenta}{\boldsymbol{g1}} + \sqrt{VPD})^2}\right) \right)\]
   \begin{itemize}
@@
   \item $R_{air}$, $\gamma$, and $c_s$ are approximately constant (relative to $\sqrt{VPD}$)
-  \item uWUE and $g_1$ are constants within plant type (i.e. grass, crops, deciduous forest, evergreen needleleaf forest, shrub)
-  \item Which just leaves VPD
+  \item \textcolor{magenta}{\textbf{uWUE}} and $\textcolor{magenta}{\boldsymbol{g1}}$ are constants within plant type (i.e. grass, crops, deciduous forest, evergreen needleleaf forest, shrub)
   \end{itemize}
@@
 
-\begin{frame}{``Sign'' Term}
+\begin{frame}{``Sign'' term as a function of VPD and PFT}
   \begin{figure}
-    \adjincludegraphics[width=3.5in, trim={0 {0.51\height} 0 0}, clip]{fig05.pdf}
+      \begin{tikzpicture}
+    \node[anchor=south west,inner sep=0] (image) at (0,0) {\adjincludegraphics[width=3.3in, trim={0 {0.51\height} 0 0}, clip]{fig05.pdf}};
+        \begin{scope}[x={(image.south east)},y={(image.north west)}]
+        %% next four lines will help you to locate the point needed by forming a grid. comment these four lines in the final picture.↓
+       % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
+       % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
+       % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
+       % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
+       %% upto here
+        \draw[-latex, line width=2pt] (0.83,0.725) --++(0.0in,0.3in)node[anchor=south] {ET Increasing};
+        \draw[-latex, line width=2pt] (0.83,0.725) --++(0.0in,-0.3in)node[anchor=north] {ET Decreasing};
+
+        % \draw[dashed,-latex, line width=5pt] (0.8,0.75) --++(0.0in,+0.5in)node[anchor=north] {ET Decreasing};
+    \end{scope}
+    % \node[align=center,red,font={\large}] at (image.southwest) {ET decreasing};
+  \end{tikzpicture}
   \end{figure}

### 2017-12-13 src/paper_figure_code/fig05.py

@@ -3,12 +3,13 @@
 This script makes fig 5
 """
 from shared_functions import *
-
+fontsize=11
 # below is whether to plot taylor series approximation
 PLOT_SERIES = False
 #below is to make slight changes for the talk
 PLOT_TALK = True
 
+PLOT_PET = False
 
 name_dict = {'CRO': 'crops',\
              'DBF': 'deciduous forest',
@@ -150,7 +151,7 @@ def pft_leaf(_df, axs):
   axs[1].plot(vpd, np.ones(vpd.shape)*I)
   axs[1].plot(ptiles, np.ones(ptiles.shape)*I, 'k*')
   vpd_crit = et_min_vpd(_df.mean(), lai)
-  axs[1].plot([vpd_crit], [I], 'k|', markersize=15)
+  axs[1].plot([vpd_crit], [I], 'k|', markersize=15, mew=3)
   axs[-1].plot(np.ones(vpd.shape)*I, vpd)
   axs[-1].plot(np.ones(ptiles.shape)*I, ptiles, 'k*')
 
@@ -198,10 +199,13 @@ for pft in ['CRO', 'DBF', 'GRA', 'ENF', 'CSH']:
   pft_leaf(_df, axs)
 
 # df.groupby('pft').apply(pft_leaf, axs)
-axs[1].set_xlabel('VPD (Pa)')
+axs[1].set_xlabel('VPD (Pa)', fontsize=fontsize)
 axs[3].set_xlabel('$\sigma$')
-axs[0].set_ylabel(paren_string)
+axs[0].set_ylabel(paren_string, fontsize=fontsize)
 axs[0].plot(axs[0].get_xlim(), [0., 0.], 'k--', linewidth=1.0)
+if PLOT_PET:
+  axs[0].plot(axs[0].get_xlim(), [pm.CP/287.0, pm.CP/287.0],\
+              'm--', linewidth=2.0, label='PET')
 plt.setp(axs[2].get_yticklabels(), visible=False)
 axs[-1].set_ylabel(r'VPD')#$_{ETmin}$')
 axs[2].set_title(r'D$_{crit}$'\
@@ -227,14 +231,20 @@ axs[2].text(1.4, 3500., r'$\frac{\partial \; ET}{\partial \; D} > 0$',\
 # axs[3].set_xlabel('VPD (Pa)')
 # axs[2].set_ylabel(paren_string)
[…]

### 2017-12-13 src/paper_figure_code/shared_functions.py

@@ -66,16 +66,17 @@ site_list = site_list.loc[df.values]
 # for column1, column2,  _diff in zip(_df.columns, _df2.columns, difference):
 #   print(column1, column2, _diff)
 
-# df = pd.read_pickle('%s/changjie/full_pandas_fix_scaling_clean.pkl'\
-#                     % os.environ['DATA'])
-df = pd.read_pickle('%s/changjie/full_pandas_calc_uwue_clean.pkl'\
-             % os.environ['DATA'])
+df = pd.read_pickle('%s/changjie/full_pandas_fix_scaling_clean.pkl'\
+                    % os.environ['DATA'])
+# df = pd.read_pickle('%s/changjie/full_pandas_calc_uwue_clean.pkl'\
+#              % os.environ['DATA'])
 
 df['g_a'] = 1./df['r_a']
 df['d_et_leaf'] = df['scaling']*df['vpd_leaf']
 df['d_et_atm'] = df['scaling']*df['vpd_atm']
 df['uwue_norm'] = df.uwue/pm.LV
 df['r_net'] = df['r_n'] - df['g_flux']
+df['sign'] = df['vpd_atm'] + df['vpd_leaf']
 jacobians = {'vpd' : calc.d_et,\
              'lai' : calc.d_et_d_lai,\
              'seasonal_lai' : calc.d_et_d_lai,\
@@ -89,7 +90,7 @@ def et_min_vpd(_df, lai):
   """calculates theoretical max vpd as functoin of -df and lai"""
   c3 = pm.CP/_df.r_moist
   # c1 = _df.gamma*_df.c_a/(lai*pm.R_STAR*1.6*_df.uwue_norm)
-  c1 = _df.gamma*_df.c_a/(pm.R_STAR*1.6*_df.uwue_norm)
+  c1 = _df.gamma*_df.c_a/(lai*pm.R_STAR*1.6*_df.uwue_norm)
   c2 = _df.g1
   sqrt_vpd = (c1 + np.sqrt(c1 + 8.*c2*c3)*np.sqrt(c1)-4.*c2*c3)/(4.*c3)
   try:
@@ -102,3 +103,5 @@ def et_min_vpd(_df, lai):
 
 def get_pft(_df):
   return _df['pft'].iloc[0]
+
+# df.index = df.reset_index(drop=True)

### 2017-12-13 doc/agu_2017/vpd.tex
pretty much done with slides, just proofread, see if columbia template, maybe change to kPa, and add some more extras

@@
 \usepackage{xcolor}
+\usepackage{relsize}
+\usepackage[absolute,overlay]{textpos}
 \graphicspath{{figures/}}
@@
   \item $R_{air}$, $\gamma$, and $c_s$ are approximately constant (relative to $\sqrt{VPD}$)
-  \item \textcolor{magenta}{\textbf{uWUE}} and $\textcolor{magenta}{\boldsymbol{g1}}$ are constants within plant type (i.e. grass, crops, deciduous forest, evergreen needleleaf forest, shrub)
+  \item \textcolor{magenta}{\textbf{uWUE}} and $\textcolor{magenta}{\boldsymbol{g1}}$ are constants within plant type (e.g.
+grass, crops, deciduous broadleaf forest, evergreen needleleaf forest, shrub)
   \end{itemize}
@@
 
-\section{Test Theory}
-\begin{frame}{The theory seems nice, but we should test with data!}
-    \begin{itemize}
-    \item Introduce a free uncertainty parameter $\sigma$ to Penman Monteith:
-      \[ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_s \sqrt{VPD} }{ R* \; 1.6\; \boldsymbol{\sigma} \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}\]
-    \item At each observation from FLUXNET (56 sites across globe) calculate a unique $\sigma$:
-      \small
-      \[\sigma = - \frac{g_a \gamma c_s \sqrt{VPD} P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p VPD\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})}\]
-      \normalsize
-    \item Then propagate uncertainty forward by including $\sigma$ in the derivative:
-    \end{itemize}
-      \[\frac{\partial \;  ET}{\partial \; VPD} = \text{scaling} \cdot \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_s }{1.6 \; R*\; \sigma \; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]
-
+\begin{frame}{``Sign'' term as a function of VPD and PFT}
+  \begin{figure}
+      \begin{tikzpicture}
+    \node[anchor=south west,inner sep=0] (image) at (0,0) {\adjincludegraphics[width=3.3in, trim={0 {0.51\height} 0 0}, clip]{fig05_pet.pdf}};
+        \begin{scope}[x={(image.south east)},y={(image.north west)}]
+        %% next four lines will help you to locate the point needed by forming a grid. comment these four lines in the final picture.↓
+       % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
+       % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
+       % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
+       % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
+       %% upto here
+        % \draw[-latex, line width=2pt] (0.83,0.725) --++(0.0in,0.3in)node[anchor=south] {ET Increasing};
+        % \draw[-latex, line width=2pt] (0.83,0.725) --++(0.0in,-0.3in)node[anchor=north] {ET Decreasing};
+        % \draw[dashed,-latex, line width=5pt] (0.8,0.75) --++(0.0in,+0.5in)node[anchor=north] {ET Decreasing};
+    \end{scope}
+    % \node[align=center,red,font={\large}] at (image.southwest) {ET decreasing};
+  \end{tikzpicture}
+\end{figure}
+\begin{textblock*}{0.25\textwidth}(0.02\textwidth,0.3\textheight)
+  \textcolor{magenta}{Dashed line} gives response for potential evapotranspiration (PET).\\
+\medskip
+  Plants are crucial for land response!
+\end{textblock*}
 \end{frame}
 
-    \begin{frame}{Test theory with FLUXNET data - CSH}
-                \includegraphics[width=3.5in]{csh_eee.png}
+\section{Test Theory}
+\begin{frame}{The theory seems nice, but we need to test with data!}
+  \begin{itemize}
+  \item Introduce a free uncertainty parameter \Large $\textcolor{magenta}{\boldsymbol{\sigma}}$ \normalsize to Penman Monteith:
+    \[ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_s \sqrt{VPD} }{ R* \; 1.6\; \textcolor{magenta}{\boldsymbol{\mathlarger{\sigma}}} \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}\]
+  \item At each observation from FLUXNET (56 sites) calculate a unique \Large $\textcolor{magenta}{\boldsymbol{\sigma}}$ \normalsize:
+    \[\textcolor{magenta}{\boldsymbol{\mathlarger{\sigma}}}(t, \text{site}) = f(ET_{obs})\]
+
+    \item Then propagate uncertainty forward by including \Large $\textcolor{magenta}{\boldsymbol{\sigma}}$ \normalsize in the derivative:
+    \end{itemize}
+      \[\frac{\partial \;  ET}{\partial \; VPD} = \text{scaling} \cdot \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_s }{1.6 \; R*\; \textcolor{magenta}{\boldsymbol{\mathlarger{\sigma}}} \; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]
 \end{frame}
 
-    \begin{frame}{Test theory with FLUXNET data - ENF}
-                \includegraphics[width=3.5in]{enf_eee.png}
+\begin{frame}{Test theory with FLUXNET data - the good}
+  \begin{columns}
+    \begin{column}{0.33\textwidth}
+      \includegraphics[width=1.2\textwidth]{DBF_box.pdf}
+    \end{column}
+    \begin{column}{0.33\textwidth}
+      \includegraphics[width=1.2\textwidth]{ENF_box.pdf}
+    \end{column}
+    \begin{column}{0.33\textwidth}
+      \includegraphics[width=1.2\textwidth]{CSH_box.pdf}
+    \end{column}
+  \end{columns}
 \end{frame}
 
-    \begin{frame}{Test theory with FLUXNET data - DBF}
-                \includegraphics[width=3.5in]{dbf_eee.png}
-     \end{frame}
-
-    \begin{frame}{Test theory with FLUXNET data - CRO}
-                \includegraphics[width=3.5in]{cro_eee.png}
-     \end{frame}
-
-    \begin{frame}{Test theory with FLUXNET data - GRA}
-                \includegraphics[width=3.5in]{gra_eee.png}
+\begin{frame}{Test theory with FLUXNET data - the bad}
+    \begin{columns}
+    \begin{column}{0.5\textwidth}
+      \includegraphics[width=\textwidth]{GRA_box.pdf}
+    \end{column}
+    \begin{column}{0.5\textwidth}
+      \includegraphics[width=\textwidth]{CRO_box.pdf}
+    \end{column}
+  \end{columns}
 \end{frame}
@@
 \begin{frame}{Extra slide - is theory VPD$_{crit}$ optimum?}
-  \includegraphics[width=\textwidth]{for_agu.pdf}
+  \includegraphics[width=0.7\textwidth]{for_agu.pdf}
+\end{frame}
+
+\begin{frame}{Test theory with FLUXNET data - CSH}
+  \includegraphics[width=3.5in]{csh_eee.png}
+\end{frame}
+
+\begin{frame}{Test theory with FLUXNET data - ENF}
+  \includegraphics[width=3.5in]{enf_eee.png}
+\end{frame}
+
+\begin{frame}{Test theory with FLUXNET data - DBF}
+  \includegraphics[width=3.5in]{dbf_eee.png}
+\end{frame}
+
+\begin{frame}{Test theory with FLUXNET data - CRO}
+  \includegraphics[width=3.5in]{cro_eee.png}
+\end{frame}
+
+\begin{frame}{Test theory with FLUXNET data - GRA}
+  \includegraphics[width=3.5in]{gra_eee.png}
 \end{frame}

### 2017-12-13 doc/agu_2017/vpd.tex
refactoring slides

@@
 \section{Introduction}
+
 \begin{frame}{Does VPD drive or reduce ET? - atmospheric demand perspective}
-  \begin{Huge}
+  \begin{columns}
+    \begin{column}{0.7\textwidth}
+      Increase in VPD (\textbf{increase in atmospheric demand}) drives  an \textbf{increase in ET}.
       \[VPD = (1-RH)\cdot e_s (T)\]
-\end{Huge}
-  \begin{itemize}
-  \item Increase in VPD (\textbf{increase in atmospheric demand}) drives  an \textbf{increase in ET}.
-  \end{itemize}
+    \end{column}
+    \begin{column}{0.3\textwidth}<2>      
+      \includegraphics[width=\textwidth]{towel.jpg}
+    \end{column}
+  \end{columns}
 \end{frame}
@@
     \begin{column}{0.5\textwidth}
-      \begin{itemize}
-      \item However, plants evolved to use stomata to conserve and regulate water use. So \textbf{stomata closure} in response to increases in VPD may \textbf{decrease ET}.
-      \end{itemize}
+      However, plants evolved to use stomata to conserve and regulate water use. So \textbf{stomata closure} in response to increases in VPD may \textbf{decrease ET}.
     \end{column}
@@
   \begin{itemize}
-  \item The goal is to simply the problem (increase transparency) while still capturing the leading order behavior the system. (intrinsic knowledge)
+  \item The goal is to simply the problem (increase transparency) while still capturing the leading order behavior the system.
     \begin{itemize}
@@
     \[ET = \frac{\Delta R + g_a \rho_a c_p \, VPD}{\Delta + \gamma(1 + \frac{g_a}{\textcolor{magenta}{\boldsymbol{g_s}}})},\]
+  \begin{overprint}
+    \onslide<2>
     \textbf{Problem}: \Large $\textcolor{magenta}{\boldsymbol{g_s}}$ \normalsize (stomatal conductance) is a function of photosynthesis, which is a function of ET itself.
-So ET in Penman-Monteith is really an implicit function of itself and we cannot take derivatives!
+So ET in Penman-Monteith is really an implicit function of itself and we cannot take derivatives!<2>
+  \end{overprint}
 \end{frame}
@@
   So within each plant type, whether the atmospheric demand (ET increasing with VPD) or plant response (ET decreasing with VPD) dominates is essentially just a function of VPD!
+\end{frame}
+
+\begin{frame}{Consequences of theory - ``sign'' term}
+  \[\frac{\partial \, ET}{\partial \, VPD} = \text{scaling} \cdot \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g1 + \sqrt{VPD}}{2 (g1 + \sqrt{VPD})^2}\right) \right)\]
+  \begin{itemize}
+  \item $c_p$ and $R*$ are constants
+  \item $R_{air}$, $\gamma$, and $c_s$ are approximately constant (relative to $\sqrt{VPD}$)
+  \item uWUE and $g1$ are constants within plant type (e.g. grass, crops, deciduous broadleaf forest, evergreen needleleaf forest, shrub)
+  \end{itemize}
+  So \textbf{within each plant type}, whether the atmospheric demand (ET increasing with VPD) or plant response (ET decreasing with VPD) dominates is approximately \textbf{just a function of VPD}!
 \end{frame}
@@
     \begin{column}{0.33\textwidth}
+    \end{column}
+    \begin{column}{0.33\textwidth}
+    \end{column}
+  \end{columns}
+\end{frame}
+
+\begin{frame}{Test theory with FLUXNET data - the good}
+  \begin{columns}
+    \begin{column}{0.33\textwidth}
+      \includegraphics[width=1.2\textwidth]{DBF_box.pdf}
+    \end{column}
+    \begin{column}{0.33\textwidth}
       \includegraphics[width=1.2\textwidth]{ENF_box.pdf}
@@
 
-\begin{frame}{Summary}
-  \small
+\begin{frame}{Summary - When does VPD drive or reduce ET?}
   \begin{itemize}
-  \item Derived a new form of Penman-Monteith that explicitly calculates ET given published constants.
-  \item Developed a theory for when atmospheric dryness will increase ET (atmospheric demand dominates) vs.
-decrease ET (plant response dominates).
-Theory tests well using real data from FLUXNET2015 on plant types CSH, ENF, and DBF.
-Applicability to CRO and GRA is more uncertain.
-  \item For plants evolved to prioritize water conservation over production, atmospheric demand is less likely to dominate.
-  \item All plant types except for crops had common occurrence of decreases in ET with increases in VPD (plant response dominates).
-Therefore, the hydrometeorologists perspective is often wrong, and drought indices built upon assumptions about atmospheric demand are unrealistic for vegetated surfaces.
-  \item Our form of Penman-Monteith is capable of estimating ET realistically and could be used as a replacement for PET in drought indices (e.g. Palmer Drought Severity Index - PDSI).
+  \item Theory predicts that each plant type has a \textbf{critical VPD} \textbf{below which ET will decrease} (plant response dominates), and \textbf{above which ET will increase} (atmosheric demand dominsates).
+  \item For forest sites, environmental VPD approximately straddles the critical VPD.
+  \item For shrubs environmental VPD never exceeds the critical VPD.
+  \item Theory tested poorly with FLUXNET data for for crops and grass. 
+  \item All plant types exhibited a response far below that of PET.
+  \item The new uWUE-version of Penman-Monteith we derived could be used as a replacement for PET in drought indices over vegetated surfaces.
   \end{itemize}

### 2017-12-13 doc/agu_2017/vpd.tex
slides looking very good, ready for walkthough

@@
 \usepackage{relsize}
+\usepackage{textcomp}
 \usepackage[absolute,overlay]{textpos}
@@
   \[\frac{\partial \, ET}{\partial \, VPD} = \text{scaling} \cdot \left(\frac{\textcolor{magenta}{\boldsymbol{ c_p}}}{R_{air}} - \frac{\gamma c_s }{1.6 \; \textcolor{magenta}{\boldsymbol{R*}}\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]
+  \begin{overprint}
+    \onslide<1>
     \begin{itemize}
@@
     \end{itemize}
-\end{frame}
-
-\begin{frame}{Consequences of theory - ``sign'' term}
-  \[\frac{\partial \, ET}{\partial \, VPD} = \text{scaling} \cdot \left(\frac{ c_p}{\textcolor{magenta}{\boldsymbol{R_{air}}}} - \frac{\textcolor{magenta}{\boldsymbol{\gamma}} \textcolor{magenta}{\boldsymbol{c_s}} }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]
+    \onslide<2>
     \begin{itemize}
@@
     \end{itemize}
-\end{frame}
-
-\begin{frame}{Consequences of theory - ``sign'' term}
-  \[\frac{\partial \, ET}{\partial \, VPD} = \text{scaling} \cdot \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ \textcolor{magenta}{\textbf{uWUE }}}} \left( \frac{2 \textcolor{magenta}{\boldsymbol{g1}} + \sqrt{VPD}}{2 (\textcolor{magenta}{\boldsymbol{g1}} + \sqrt{VPD})^2}\right) \right)\]
+    \onslide<3>
     \begin{itemize}
@@
     \end{itemize}
-  So within each plant type, whether the atmospheric demand (ET increasing with VPD) or plant response (ET decreasing with VPD) dominates is essentially just a function of VPD!
-\end{frame}
-
-\begin{frame}{Consequences of theory - ``sign'' term}
-  \[\frac{\partial \, ET}{\partial \, VPD} = \text{scaling} \cdot \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g1 + \sqrt{VPD}}{2 (g1 + \sqrt{VPD})^2}\right) \right)\]
+    \onslide<4>
     \begin{itemize}
@@
     So \textbf{within each plant type}, whether the atmospheric demand (ET increasing with VPD) or plant response (ET decreasing with VPD) dominates is approximately \textbf{just a function of VPD}!
+  \end{overprint}    
 \end{frame}
@@
 \begin{frame}{Test theory with FLUXNET data - the good}
+    \begin{textblock*}{1.0\textwidth}(0.3\textwidth,0.8\textheight)
+      \textcolor{blue}{Blue line} is theory's critical VPD.\\
+  \end{textblock*}
+
   \begin{columns}
-    \begin{column}{0.33\textwidth}
+    \begin{column}{0.33\textwidth}<1-2>
       \includegraphics[width=1.2\textwidth]{DBF_box.pdf}
     \end{column}
-    \begin{column}{0.33\textwidth}
-    \end{column}
-    \begin{column}{0.33\textwidth}
-    \end{column}
-  \end{columns}
-\end{frame}
-
-\begin{frame}{Test theory with FLUXNET data - the good}
-  \begin{columns}
-    \begin{column}{0.33\textwidth}
-      \includegraphics[width=1.2\textwidth]{DBF_box.pdf}
-    \end{column}
-    \begin{column}{0.33\textwidth}
+    \begin{column}{0.33\textwidth}<2>
       \includegraphics[width=1.2\textwidth]{ENF_box.pdf}
     \end{column}
-    \begin{column}{0.33\textwidth}
+    \begin{column}{0.33\textwidth}<2>
       \includegraphics[width=1.2\textwidth]{CSH_box.pdf}
@@
   \end{columns}
+      \begin{textblock*}{1.0\textwidth}(0.3\textwidth,0.85\textheight)
+      \textcolor{blue}{Blue line} is theory's critical VPD.\\
+  \end{textblock*}
+
 \end{frame}
@@
 \section{References}
-\begin{frame}{References}
-  \AtNextBibliography{\small}
-  \printbibliography
-  \scriptsize
-  \begin{itemize}
-  \item This work used eddy covariance data acquired and shared by the FLUXNET community, including these networks: AmeriFlux, AfriFlux, AsiaFlux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont, ChinaFlux, Fluxnet-Canada, GreenGrass, ICOS, KoFlux, LBA, NECC, OzFlux-TERN, TCOS-Siberia, and USCCC.
+\begin{frame}{Acknowledgements - Thank you NSF and FLUXNET!!!}
+ This work used eddy covariance data acquired and shared by the FLUXNET community, including these networks: AmeriFlux, AfriFlux, AsiaFlux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont, ChinaFlux, Fluxnet-Canada, GreenGrass, ICOS, KoFlux, LBA, NECC, OzFlux-TERN, TCOS-Siberia, and USCCC.
 The ERA-Interim reanalysis data are provided by ECMWF and processed by LSCE.
-The FLUXNET eddy covariance data processing and harmonization was carried out by the European Fluxes Database Cluster, AmeriFlux Management Project, and Fluxdata project of FLUXNET, with the support of CDIAC and ICOS Ecosystem Thematic Center, and the OzFlux, ChinaFlux and AsiaFlux offices.
-    \item This material is based upon work supported by the National Science Foundation Graduate Research Fellowship under Grant No.
+The FLUXNET eddy covariance data processing and harmonization was carried out by the European Fluxes Database Cluster, AmeriFlux Management Project, and Fluxdata project of FLUXNET, with the support of CDIAC and ICOS Ecosystem Thematic Center, and the OzFlux, ChinaFlux and AsiaFlux offices.\\
+  \medskip
+  This material is based upon work supported by the National Science Foundation Graduate Research Fellowship under Grant No.
 DGE 16-44869.
 Any opinion, findings, and conclusions or recommendations expressed in this material are those of the authors(s) and do not necessarily reflect the views of the National Science Foundation.
-    \end{itemize}
+\end{frame}
+
+\begin{frame}{References}
+  \begin{enumerate}
+  \item Franks, P.
+J., Berry, J.
+A., Lombardozzi, D.
+L., \& Bonan, G.
+B.
+(2017).
+Stomatal function across temporal and spatial scales: deep-time trends, land-atmosphere coupling and global models.
+\textit{Plant Physiology}, 174(2), 583-602.
+    \item Lin, Y.
+S., Medlyn, B.
+E., Duursma, R.
+A., Prentice, I.
+C., Wang, H., Baig, S., ...
+\& De Beeck, M.
+O.
+(2015).
+Optimal stomatal behaviour around the world.
+Nature Climate Change, 5(5), 459-464.
+    \item Medlyn, B.
+E., Duursma, R.
+A., Eamus, D., Ellsworth, D.
+S., Prentice, I.
+C., Barton, C.
+V., ...
+\& Wingate, L.
+(2011).
+Reconciling the optimal and empirical approaches to modelling stomatal conductance.
+\textit{Global Change Biology}, 17(6), 2134-2144.
+    \item Zhou, S., Yu, B., Huang, Y., \& Wang, G. (2015). Daily underlying water use efficiency for AmeriFlux sites. \textit{Journal of Geophysical Research: Biogeosciences}, 120(5), 887-902.
+    \end{enumerate}
 \end{frame}
@@
   \end{table}
+\end{frame}
+
+\begin{frame}{ideas}
+  full summary statistics
+  variance statistics
+  joint distributions of LAI and VPD, and RH and esat
+  aerodynamic scaling theory.
 \end{frame}

### 2017-12-13 doc/agu_2017/vpd.tex
spell check done

@@
 \begin{frame}{Consequences of theory - ``sign'' term}
-  \[\frac{\partial \, ET}{\partial \, VPD} = \text{scaling} \cdot \left(\frac{\textcolor{magenta}{\boldsymbol{ c_p}}}{R_{air}} - \frac{\gamma c_s }{1.6 \; \textcolor{magenta}{\boldsymbol{R*}}\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]
   \begin{overprint}
     \onslide<1>
+    \[\frac{\partial \, ET}{\partial \, VPD} = \text{scaling} \cdot \left(\frac{\textcolor{magenta}{\boldsymbol{ c_p}}}{R_{air}} - \frac{\gamma c_s }{1.6 \; \textcolor{magenta}{\boldsymbol{R*}}\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]
     \begin{itemize}
@@
     \onslide<2>
+    \[\frac{\partial \, ET}{\partial \, VPD} = \text{scaling} \cdot \left(\frac{ c_p}{\textcolor{magenta}{\boldsymbol{R_{air}}}} - \frac{\textcolor{magenta}{\boldsymbol{\gamma c_s}} }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]
     \begin{itemize}
@@
     \onslide<3>
+    \[\frac{\partial \, ET}{\partial \, VPD} = \text{scaling} \cdot \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ \textcolor{magenta}{\textbf{ uWUE}} }} \left( \frac{2  \textcolor{magenta}{\boldsymbol{g_1}} + \sqrt{VPD}}{2 ( \textcolor{magenta}{\boldsymbol{g_1}} + \sqrt{VPD})^2}\right) \right)\]
+
     \begin{itemize}
@@
     \onslide<4>
+    \[\frac{\partial \, ET}{\partial \, VPD} = \text{scaling} \cdot \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]
+
     \begin{itemize}
@@
   \begin{itemize}
-  \item Theory predicts that each plant type has a \textbf{critical VPD} \textbf{below which ET will decrease} (plant response dominates), and \textbf{above which ET will increase} (atmosheric demand dominsates).
+  \item Theory predicts that each plant type has a \textbf{critical VPD} \textbf{below which ET will decrease} (plant response dominates), and \textbf{above which ET will increase} (atmospheric demand dominates).
   \item For forest sites, environmental VPD approximately straddles the critical VPD.
@@
 \section{References}
-\begin{frame}{Acknowledgements - Thank you NSF and FLUXNET!!!}
+\begin{frame}{Acknowledgments - Thank you NSF and FLUXNET!!!}
  This work used eddy covariance data acquired and shared by the FLUXNET community, including these networks: AmeriFlux, AfriFlux, AsiaFlux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont, ChinaFlux, Fluxnet-Canada, GreenGrass, ICOS, KoFlux, LBA, NECC, OzFlux-TERN, TCOS-Siberia, and USCCC.
@@
   aerodynamic scaling theory.
+  put vpd in kpa
+  add pictures to first slide
 \end{frame}

### 2017-12-13 doc/agu_2017/vpd.tex
done with slides, now jsut run through and maybe adjust figure linewidth, font and vpd to kpa

@@
 
-\begin{frame}{``Sign'' term as a function of VPD and PFT}
+\begin{frame}{``Sign'' term as a function of VPD and plant type}
   \begin{figure}
@@
 
+\begin{frame}{Quantitative $VPD_{crit}$}
+
+  \scriptsize
+  \begin{table}
+    \label{vpd_crit}
+    \caption{Values of $VPD_{crit}$, where $\frac{\partial \; ET}{\partial \; VPD} = 0$, evaluated at PFT average values for $R_{air}$, $\sigma$, $\gamma$, and $c_s$.
+For reference, these values are also provided.
+For values of $VPD$ less than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be negative, and for values of $VPD$ greater than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be positive.}
+    \centering
+    \begin{tabular}{l c c c c c}
+      \hline
+      PFT & $R_{air}$ & $c_s$ (ppm) & $\gamma$ &  uWUE & \textbf{$VPD_{crit}$ (Pa)} \\
+      \hline
+      CRO &  288.680920 & 372.567691& 65.351523& 2.602873&  \textbf{133.165438} \\
+      CSH &   289.067152& 381.593622& 67.613172& 2.175278& \textbf{4439.564212} \\
+      DBF &   288.624437& 377.449849& 63.421812& 2.746393&  \textbf{888.773243} \\
+      ENF &  288.183849& 377.676463& 61.559242& 4.015362&  \textbf{978.084845} \\
+      GRA &  288.425651& 377.264645& 61.598768& 2.281074& \textbf{1141.630778} \\
+      \hline
+      \multicolumn{2}{l}{}
+    \end{tabular}
+  \end{table}
+\end{frame}
+
+\begin{frame}{Statistics with uncertainty}
+  \scriptsize
+  \begin{table}
+    \caption{Statistics of $\frac{\partial \; ET}{\partial \; VPD}$ as a function of PFT.}
+    \centering
+    \begin{tabular}{l c c c c c}
+      \hline
+      PFT & $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ & $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)$ & $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)*\text{std}(VPD)$ & $\frac{\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)*\text{std}(VPD)}{ \frac{\partial \; ET}{\partial \; R}\left(\overline{env}\right)*\text{std}(R)}$ & fraction $\frac{\partial \; ET}{\partial \; VPD} < 0.$ \\
+      \hline
+      CRO & 0.000853 & 0.026241 & 37.05 & 0.41 & 0.473311\\
+      CSH & -0.108234 & -0.091526 & 101.72 & 0.88 & 0.931660\\
+      DBF & -0.012727 & 0.013794 & 39.47 & 0.33 & 0.461674\\
+      ENF & -0.034087 & 0.000706 & 33.22 & 0.30 & 0.534425\\
+      GRA & -0.019637 & -0.000921 & 33.60 & 0.35 & 0.631735\\
+      \hline
+      \multicolumn{2}{l}{}
+
+    \end{tabular}
+  \end{table}
+\end{frame}
 \begin{frame}{ideas}
-  full summary statistics
-  variance statistics
   joint distributions of LAI and VPD, and RH and esat
-  aerodynamic scaling theory.
   put vpd in kpa
   add pictures to first slide
+\end{frame}
+
+\begin{frame}{Scaling Term}
+  \[\frac{\partial \;  ET}{\partial \; D} = \mathbf{ \frac{g_a \; P}{T(\Delta + \gamma)} }   \left(\frac{ c_p}{R_{air}} - \frac{\text{LAI$_{ref}$}}{\text{LAI}} \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right) \right)\]
+  \begin{figure}
+    \includegraphics[width=3.0in]{scaling.png}
+    % \caption{}
+  \end{figure}
 \end{frame}
@@
   \includegraphics[width=3.5in]{gra_eee.png}
+\end{frame}
+
+\begin{frame}{How to take VPD derivative?}
+  % below gnereated in fig03.py
+  \includegraphics[width=3.0in]{e_s_rh.png}
+\end{frame}
+
+\begin{frame}{Is uncertainty a function of VPD?}
+  \includegraphics[width=3.0in]{vpd_sigma.png}
 \end{frame}

### 2017-12-13 src/codebase/plot_tools.py

@@ -230,7 +230,7 @@ def test_trend(_df, meta, fig=None):
     g = sns.jointplot(x=_df[meta['x_var']], y=_df[meta['y_var']], kind='hex',\
                       xlim=meta['xlim'], ylim=meta['ylim'], stat_func=spearmanr)
     g.set_axis_labels(meta['x_var'],meta['y_var'])
-  plt.title('pft: %s' % _df['pft'].iloc[0])
+
   if meta['full_ds']:
     util.test_savefig('%s/climate_et/scatters/%s_%s.png'\
                 % (os.environ['PLOTS'], meta['x_var'], meta['y_var']))
@@ -238,10 +238,12 @@ def test_trend(_df, meta, fig=None):
     util.test_savefig('%s/climate_et/scatters/%s_%s_site/%s.png'\
                  % (os.environ['PLOTS'], meta['x_var'],\
                     meta['y_var'],  _df['site'].iloc[0]))
+    plt.title('pft: %s' % _df['pft'].iloc[0])
   else:
     util.test_savefig('%s/climate_et/scatters/%s_%s/%s.png'\
                 % (os.environ['PLOTS'], meta['x_var'],\
                    meta['y_var'],  _df['pft'].iloc[0]))
+    plt.title('pft: %s' % _df['pft'].iloc[0])
   return spearmanr(_df[meta['x_var']], _df[meta['y_var']]).correlation
 
 def scatter_wrapper(df, meta):

### 2017-12-14 doc/agu_2017/vpd.tex
took care of all ideas, unless I want to do a CU theme or add CU and NSF pictures to my slides

@@
   \[ET = \frac{\Delta R + g_a \rho_a c_p \, VPD}{\Delta + \gamma(1 + \frac{g_a}{\textcolor{magenta}{\boldsymbol{g_s}}})},\]
-  \begin{overprint}
-    \onslide<2>
+  % \begin{overprint}
+  %   \onslide<2>
     \textbf{Problem}: \Large $\textcolor{magenta}{\boldsymbol{g_s}}$ \normalsize (stomatal conductance) is a function of photosynthesis, which is a function of ET itself.
-So ET in Penman-Monteith is really an implicit function of itself and we cannot take derivatives!<2>
-  \end{overprint}
+So ET in Penman-Monteith is really an implicit function of itself and we cannot take derivatives!
+  % \end{overprint}
 \end{frame}
@@
   \item Introduce a free uncertainty parameter \Large $\textcolor{magenta}{\boldsymbol{\sigma}}$ \normalsize to Penman Monteith:
-    \[ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_s \sqrt{VPD} }{ R* \; 1.6\; \textcolor{magenta}{\boldsymbol{\mathlarger{\sigma}}} \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}\]
+    \[ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_s \sqrt{VPD} }{ R* \; 1.6\; \textcolor{magenta}{\boldsymbol{\mathlarger{\mathlarger{\sigma}}}} \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}\]
   \item At each observation from FLUXNET (56 sites) calculate a unique \Large $\textcolor{magenta}{\boldsymbol{\sigma}}$ \normalsize:
@@
   \end{itemize}
-  \[\frac{\partial \;  ET}{\partial \; VPD} = \text{scaling} \cdot \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_s }{1.6 \; R*\; \textcolor{magenta}{\boldsymbol{\mathlarger{\sigma}}} \; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]
+  \[\frac{\partial \;  ET}{\partial \; VPD} = \text{scaling} \cdot \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_s }{1.6 \; R*\; \textcolor{magenta}{\boldsymbol{\mathlarger{\mathlarger{\sigma}}}} \; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]
 \end{frame}
@@
     \begin{column}{0.33\textwidth}<1-2>
-      \includegraphics[width=1.2\textwidth]{DBF_box.pdf}
+      \includegraphics[width=1.1\textwidth]{DBF_box.pdf}
     \end{column}
     \begin{column}{0.33\textwidth}<2>
-      \includegraphics[width=1.2\textwidth]{ENF_box.pdf}
+      \includegraphics[width=1.1\textwidth]{ENF_box.pdf}
     \end{column}
     \begin{column}{0.33\textwidth}<2>
-      \includegraphics[width=1.2\textwidth]{CSH_box.pdf}
+      \includegraphics[width=1.1\textwidth]{CSH_box.pdf}
     \end{column}
@@
 \end{frame}
-\begin{frame}{ideas}
-  joint distributions of LAI and VPD, and RH and esat
-  put vpd in kpa
-  add pictures to first slide
-\end{frame}
 

### 2017-12-14 doc/agu_2017/vpd.tex
going to bed

@@
 \begin{frame}{The theory seems nice, but we need to test with data!}
-  \begin{itemize}
-  \item Introduce a free uncertainty parameter \Large $\textcolor{magenta}{\boldsymbol{\sigma}}$ \normalsize to Penman Monteith:
+  Introduce a free uncertainty parameter \Large $\textcolor{magenta}{\boldsymbol{\sigma}}$ \normalsize to Penman Monteith:
   \[ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_s \sqrt{VPD} }{ R* \; 1.6\; \textcolor{magenta}{\boldsymbol{\mathlarger{\mathlarger{\sigma}}}} \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}\]
-  \item At each observation from FLUXNET (56 sites) calculate a unique \Large $\textcolor{magenta}{\boldsymbol{\sigma}}$ \normalsize:
+  At each observation from FLUXNET (56 sites) calculate a unique \Large $\textcolor{magenta}{\boldsymbol{\sigma}}$ \normalsize:
   \[\textcolor{magenta}{\boldsymbol{\mathlarger{\sigma}}}(t, \text{site}) = f(ET_{obs})\]
 
-  \item Then propagate uncertainty forward by including \Large $\textcolor{magenta}{\boldsymbol{\sigma}}$ \normalsize in the derivative:
-  \end{itemize}
+  Then propagate uncertainty forward by including \Large $\textcolor{magenta}{\boldsymbol{\sigma}}$ \normalsize in the derivative:
+
   \[\frac{\partial \;  ET}{\partial \; VPD} = \text{scaling} \cdot \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_s }{1.6 \; R*\; \textcolor{magenta}{\boldsymbol{\mathlarger{\mathlarger{\sigma}}}} \; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]
@@
 
+\begin{frame}{Is uncertainty a function of VPD?}
+  \includegraphics[width=3.0in]{vpd_sigma.png}
+\end{frame}
+
 \begin{frame}{How to take VPD derivative?}
@@
   \includegraphics[width=3.0in]{e_s_rh.png}
-\end{frame}
-
-\begin{frame}{Is uncertainty a function of VPD?}
-  \includegraphics[width=3.0in]{vpd_sigma.png}
 \end{frame}

### 2017-12-14 doc/agu_2017/vpd.tex
before first run through
small edits

@@
 \DeclareMathOperator{\sign}{sign}
-
 \usepackage{caption}
@@
 \date{December 14th, 2017}
+\titlegraphic{\vspace*{1.0cm}\hspace*{-0.5cm}\includegraphics[width=6.0cm]{cu_logo.pdf}\hspace*{6.00cm}~%
+   \includegraphics[width=1.5cm]{nsf.png}
+}
 
@@
     \begin{itemize}
-    \item While simplifying the system aids intrinsic understanding, simplifying assumptions sacrifice physical realism.
+    \item Simplifying the system aids intrinsic understanding, but at a physical realism.
     \end{itemize}
   \item Because we run the risk of over-simplification, we will use FLUXNET2015 data to test how well our theory matches the data.
+    \begin{itemize}
+    \item We will test in the growing season of 5 plant types: Deciduous broadleaf forest, evergreen needleleaf forest, shrub, grass, and crops.
+    \end{itemize}
   \end{itemize}
@@
   \[uWUE = \frac{GPP \cdot \sqrt{VPD}}{ET},\]
+  \begin{overprint}
+    \onslide<2>
     To derive a new form of Penman-Monteith without implicit ET dependence:
     \[  ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p \, VPD}{R_{air}} -  \frac{\gamma c_s \sqrt{VPD} }{ R* \; 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right)}{ \Delta + \gamma}\]
+    \end{overprint}
 \end{frame}
@@
     \item $R_{air}$, $\gamma$, and $c_s$ are approximately constant (relative to $\sqrt{VPD}$)
-    \item \textcolor{magenta}{\textbf{uWUE}} and $\textcolor{magenta}{\boldsymbol{g1}}$ are constants within plant type (e.g.
-grass, crops, deciduous broadleaf forest, evergreen needleleaf forest, shrub)
+    \item \textcolor{magenta}{\textbf{uWUE}} and $\textcolor{magenta}{\boldsymbol{g1}}$ are constants within plant type.
     \end{itemize}
@@
     \item $R_{air}$, $\gamma$, and $c_s$ are approximately constant (relative to $\sqrt{VPD}$)
-    \item uWUE and $g1$ are constants within plant type (e.g. grass, crops, deciduous broadleaf forest, evergreen needleleaf forest, shrub)
+    \item uWUE and $g1$ are constants within plant type
     \end{itemize}
@@
   \item All plant types exhibited a response far below that of PET.
-  \item The new uWUE-version of Penman-Monteith we derived could be used as a replacement for PET in drought indices over vegetated surfaces.
+  \item The new uWUE-version of Penman-Monteith we derived could be used as a complement for PET in drought indices over vegetated surfaces.
   \end{itemize}

### 2017-12-15 doc/agu_2017/vpd.tex
first commit of messy

@@
 \section{Method}
-\begin{frame}{Develop a theory to quantify VPD effects}
+\begin{frame}{Develop a theory to quantify ET response to VPD}
   \begin{itemize}
@@
     \begin{itemize}
-    \item We will test in the growing season of 5 plant types: Deciduous broadleaf forest, evergreen needleleaf forest, shrub, grass, and crops.
+    \item We will test in the growing season of 5 plant types: deciduous broadleaf forest, evergreen needleleaf forest, shrub, grass, and crops.
     \end{itemize}
@@
   \begin{textblock*}{1.0\textwidth}(0.3\textwidth,0.8\textheight)
-    \textcolor{blue}{Blue line} is theory's critical VPD.\\
+    \textcolor{blue}{Blue line} is theory.\\
   \end{textblock*}
@@
   \begin{columns}
-    \begin{column}{0.33\textwidth}<1-2>
+    \begin{column}{0.33\textwidth}<1-3>
       \includegraphics[width=1.1\textwidth]{DBF_box.pdf}
     \end{column}
-    \begin{column}{0.33\textwidth}<2>
+    \begin{column}{0.33\textwidth}<2-3>
+      \includegraphics[width=1.1\textwidth]{CSH_box.pdf}
+    \end{column}
+    \begin{column}{0.33\textwidth}<3>
       \includegraphics[width=1.1\textwidth]{ENF_box.pdf}
-    \end{column}
-    \begin{column}{0.33\textwidth}<2>
-      \includegraphics[width=1.1\textwidth]{CSH_box.pdf}
     \end{column}
@@
   \begin{textblock*}{1.0\textwidth}(0.3\textwidth,0.85\textheight)
-    \textcolor{blue}{Blue line} is theory's critical VPD.\\
+    \textcolor{blue}{Blue line} is theory.\\
   \end{textblock*}
@@
   \end{enumerate}
+\end{frame}
+
+\begin{frame}{Summary - When does VPD drive or reduce ET?}
+  \begin{itemize}
+  \item Theory predicts that each plant type has a \textbf{critical VPD} \textbf{below which ET will decrease} (plant response dominates), and \textbf{above which ET will increase} (atmospheric demand dominates).
+  \item For forest sites, environmental VPD approximately straddles the critical VPD.
+  \item For shrubs environmental VPD never exceeds the critical VPD.
+  \item Theory tested poorly with FLUXNET data for for crops and grass. 
+  \item All plant types exhibited a response far below that of PET.
+  \item The new uWUE-version of Penman-Monteith we derived could be used as a complement for PET in drought indices over vegetated surfaces.
+  \end{itemize}
+  \bigskip
+  \medskip
+  Adam Massmann:\textbf{ akm2203@columbia.edu }
 \end{frame}
@@
 
+\begin{frame}{Consequences of theory - ``sign'' term}
+Below is a little opaque (at least to me):
+\[\frac{\partial \, ET}{\partial \, VPD} = \text{scaling} \cdot  \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]
+so do expansion about $g_1 >> \sqrt{VPD}$:
+\footnotesize
+\[ \frac{\partial \, ET}{\partial \, VPD} \approx \text{scaling} \cdot  \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{1}{g_1} - \frac{3\sqrt{VPD}}{2 g_1^2} +  \mathcal{O}\left(\frac{\sqrt{VPD}}{g_1}\right)^{2} \right) \right)\]
+% \large
+% and:
+%   \[ \sqrt{VPD} \Big|_{\frac{\partial ET}{\partial VPD} = 0} \approx  \frac{2 g_1}{3}  \left( -\frac{ c_p \, g_1\,  1.6  \; R*\; \text{ uWUE }}{R_{air} \gamma c_s } + 1  \right) \]
+
+%   \[ VPD \Big|_{\frac{\partial ET}{\partial VPD} = 0} = \frac{R_{air}}{4 \, c_p} \left[ \frac{\gamma  c_s}{R* \, 1.6\, \text{uWUE}} + \sqrt{\left(g_1 + 8\, \frac{c_p}{R_{air}}\right) \frac{\gamma c_s}{R* \, 1.6\, \text{uWUE}}} - 4 \, g_1 \frac{ c_p}{R_{air}}\right] \]
+\end{frame}
+
 \begin{frame}{Is uncertainty a function of VPD?}

### 2017-12-15 src/paper_figure_code/fig05.py

@@ -33,17 +33,7 @@ else:
 
 
 # ##### Figure 5 #####
-def term_2(_df, lai, vpd):
-  """calculates term 2"""
-  atmos = {'gamma' : _df.gamma.mean(), 'c_a' : _df.c_a.mean(),\
-           'vpd' : vpd}
-  if _df.uwue.std() > 1.e-8:
-    print('error, uWUE is variable: %f!!!!' % _df.uwue.std())
-  elif _df.g1.std() > 1.e-8:
-    print('error, g1 is variabile: %f!!!!!' % _df.g1.std())
-  canopy = {'uwue' : _df.uwue.mean(), 'g1' : _df.g1.mean()}
-  return pm.CP/_df.r_moist.mean() +calc.leaf_vpd(atmos, canopy, lai)
-
+## term 2 movwd to shared_figs
 
 def term_2_approx(_df, lai, vpd, order=4):
   """calculates term 2"""

### 2017-12-15 src/paper_figure_code/shared_functions.py

@@ -104,4 +104,15 @@ def et_min_vpd(_df, lai):
 def get_pft(_df):
   return _df['pft'].iloc[0]
 
+def term_2(_df, lai, vpd):
+  """calculates term 2, takes in a _df of one pft"""
+  atmos = {'gamma' : _df.gamma.mean(), 'c_a' : _df.c_a.mean(),\
+           'vpd' : vpd}
+  if _df.uwue.std() > 1.e-8:
+    print('error, uWUE is variable: %f!!!!' % _df.uwue.std())
+  elif _df.g1.std() > 1.e-8:
+    print('error, g1 is variabile: %f!!!!!' % _df.g1.std())
+  canopy = {'uwue' : _df.uwue.mean(), 'g1' : _df.g1.mean()}
+  return pm.CP/_df.r_moist.mean() +calc.leaf_vpd(atmos, canopy, lai)
+
 # df.index = df.reset_index(drop=True)

### 2017-12-18 src/codebase/test_functions.py
found typo in agu presentation, fixed now

@@ -8,6 +8,8 @@ import seaborn as sns
 import numpy as np
 from codebase.data_calc import *
 
+plt.close('all')
+
 PLOTDIR = '%s/climate_et/test_plots' % os.environ['PLOTS']
 
 def max_diff(quant1, quant2):
@@ -23,8 +25,8 @@ def test_et_model(_df):
   sigma = _df['uwue']/_df['uwue_zhou']
   plt.figure()
   sns.distplot(sigma)
-  plt.xlabel("(calc'ed uwue)/(zhou's uwue)")
-  plt.title("Middle 90\% of data")
+  plt.xlabel("sigma") #"(calc'ed uwue)/(zhou's uwue)")
+  plt.title("PDF for middle 90\% of data")
   plt.savefig('%s/sigma_hist.png' % PLOTDIR)
   return
 
@@ -104,9 +106,12 @@ def compare_et_wrapper(dfs):
   pfts = list(dfs['mean'].index)
   bias = {'new' : [], 'original' : [], 'gppfixed' : []}
   rmse = {'new' : [], 'original' : [], 'gppfixed' :[]}
+  bias = {'new' : [], 'original' : []}
+  rmse = {'new' : [], 'original' : []}
+
   for pft in pfts:
     _bias, _rmse = compare_et(dfs['full'].loc[(dfs['full'].pft == pft), :],\
-                            dfs['mean'])
+                              dfs['mean'])
     for key in bias:
       bias[key].append(_bias[key])
       rmse[key].append(_rmse[key])

### 2017-12-18 doc/agu_2017/vpd.tex
found typo in agu presentation, fixed now
confirmed fig05 good and refactored, but ntoe that I broke the paper version a little for the agu talk so I need to fix this (conversion from Pa to kPa)

--- 2017-12-14 doc/agu_2017/vpd.tex
@@
     \begin{itemize}
-    \item Simplifying the system aids intrinsic understanding, but at a physical realism.
+    \item Simplifying the system aids intrinsic understanding, but at a physical realism cost.
     \end{itemize}
@@
 
-\section{Extra Slides}
-
-\begin{frame}{Extra slide - statistics}
-  \begin{table}
-    \label{vpd_crit}
-    \caption{More quantitative test of theory.}
-    \centering
-    \begin{tabular}{l c c c c c c}
-      \hline
-      PFT & \specialcell{Fraction of Obs.\\Theory is Correct} & Mean ($\frac{\partial \; ET}{\partial \; VPD} < 0$) & Mean ($\frac{\partial \; ET}{\partial \; VPD} > 0$)\\
-      \hline
-      CRO & 0.566517 & -0.209152 &  0.005856\\
-      CSH & 0.931660 & -0.264746 &       NaN\\
-      DBF & 0.633363 & -0.135679 &  0.042910\\
-      ENF & 0.633138 & -0.150665 &  0.029281\\
-      GRA & 0.442306 & -0.042158 & -0.042480\\
-      \hline
-    \end{tabular}
-  \end{table}
+\begin{frame}{Summary - When does VPD drive or reduce ET?}
+  \begin{itemize}
+  \item Theory predicts that each plant type has a \textbf{critical VPD} \textbf{below which ET will decrease} (plant response dominates), and \textbf{above which ET will increase} (atmospheric demand dominates).
+  \item For forest sites, environmental VPD approximately straddles the critical VPD.
+  \item For shrubs environmental VPD never exceeds the critical VPD.
+  \item Theory tested poorly with FLUXNET data for for crops and grass. 
+  \item All plant types exhibited a response far below that of PET.
+  \item The new uWUE-version of Penman-Monteith we derived could be used as a complement for PET in drought indices over vegetated surfaces.
+  \end{itemize}
+  \bigskip
+  \medskip
+  Adam Massmann:\textbf{ akm2203@columbia.edu }
 \end{frame}
 
-\begin{frame}{Quantitative $VPD_{crit}$}
+% \section{Extra Slides}
 
-  \scriptsize
-  \begin{table}
-    \label{vpd_crit}
-    \caption{Values of $VPD_{crit}$, where $\frac{\partial \; ET}{\partial \; VPD} = 0$, evaluated at PFT average values for $R_{air}$, $\sigma$, $\gamma$, and $c_s$.
+% \begin{frame}{Extra slide - statistics}
+%   \begin{table}
+%     \label{vpd_crit}
+%     \caption{More quantitative test of theory.}
+%     \centering
+%     \begin{tabular}{l c c c c c c}
+%       \hline
+%       PFT & \specialcell{Fraction of Obs.\\Theory is Correct} & Mean ($\frac{\partial \; ET}{\partial \; VPD} < 0$) & Mean ($\frac{\partial \; ET}{\partial \; VPD} > 0$)\\
+%       \hline
+%       CRO & 0.566517 & -0.209152 &  0.005856\\
+%       CSH & 0.931660 & -0.264746 &       NaN\\
+%       DBF & 0.633363 & -0.135679 &  0.042910\\
+%       ENF & 0.633138 & -0.150665 &  0.029281\\
+%       GRA & 0.442306 & -0.042158 & -0.042480\\
+%       \hline
+%     \end{tabular}
+%   \end{table}
+% \end{frame}
+
+% \begin{frame}{Quantitative $VPD_{crit}$}
+
+%   \scriptsize
+%   \begin{table}
+%     \label{vpd_crit}
+%     \caption{Values of $VPD_{crit}$, where $\frac{\partial \; ET}{\partial \; VPD} = 0$, evaluated at PFT average values for $R_{air}$, $\sigma$, $\gamma$, and $c_s$.
 For reference, these values are also provided.
 For values of $VPD$ less than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be negative, and for values of $VPD$ greater than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be positive.}
-    \centering
-    \begin{tabular}{l c c c c c}
-      \hline
-      PFT & $R_{air}$ & $c_s$ (ppm) & $\gamma$ &  uWUE & \textbf{$VPD_{crit}$ (Pa)} \\
-      \hline
-      CRO &  288.680920 & 372.567691& 65.351523& 2.602873&  \textbf{133.165438} \\
-      CSH &   289.067152& 381.593622& 67.613172& 2.175278& \textbf{4439.564212} \\
-      DBF &   288.624437& 377.449849& 63.421812& 2.746393&  \textbf{888.773243} \\
-      ENF &  288.183849& 377.676463& 61.559242& 4.015362&  \textbf{978.084845} \\
-      GRA &  288.425651& 377.264645& 61.598768& 2.281074& \textbf{1141.630778} \\
-      \hline
-      \multicolumn{2}{l}{}
-    \end{tabular}
-  \end{table}
-\end{frame}
-
-\begin{frame}{Statistics with uncertainty}
-  \scriptsize
-  \begin{table}
-    \caption{Statistics of $\frac{\partial \; ET}{\partial \; VPD}$ as a function of PFT.}
-    \centering
-    \begin{tabular}{l c c c c c}
-      \hline
-      PFT & $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ & $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)$ & $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)*\text{std}(VPD)$ & $\frac{\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)*\text{std}(VPD)}{ \frac{\partial \; ET}{\partial \; R}\left(\overline{env}\right)*\text{std}(R)}$ & fraction $\frac{\partial \; ET}{\partial \; VPD} < 0.$ \\
-      \hline
-      CRO & 0.000853 & 0.026241 & 37.05 & 0.41 & 0.473311\\
-      CSH & -0.108234 & -0.091526 & 101.72 & 0.88 & 0.931660\\
-      DBF & -0.012727 & 0.013794 & 39.47 & 0.33 & 0.461674\\
-      ENF & -0.034087 & 0.000706 & 33.22 & 0.30 & 0.534425\\
-      GRA & -0.019637 & -0.000921 & 33.60 & 0.35 & 0.631735\\
-      \hline
-      \multicolumn{2}{l}{}
-
-    \end{tabular}
-  \end{table}
-\end{frame}
-
-\begin{frame}{Scaling Term}
-  \[\frac{\partial \;  ET}{\partial \; D} = \mathbf{ \frac{g_a \; P}{T(\Delta + \gamma)} }   \left(\frac{ c_p}{R_{air}} - \frac{\text{LAI$_{ref}$}}{\text{LAI}} \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right) \right)\]
-  \begin{figure}
-    \includegraphics[width=3.0in]{scaling.png}
-    % \caption{}
-  \end{figure}
-\end{frame}
-
-\begin{frame}{Extra slide - is theory VPD$_{crit}$ optimum?}
-  \includegraphics[width=0.7\textwidth]{for_agu.pdf}
-\end{frame}
-
-\begin{frame}{Test theory with FLUXNET data - CSH}
-  \includegraphics[width=3.5in]{csh_eee.png}
-\end{frame}
-
-\begin{frame}{Test theory with FLUXNET data - ENF}
-  \includegraphics[width=3.5in]{enf_eee.png}
-\end{frame}
-
-\begin{frame}{Test theory with FLUXNET data - DBF}
-  \includegraphics[width=3.5in]{dbf_eee.png}
-\end{frame}
-
-\begin{frame}{Test theory with FLUXNET data - CRO}
-  \includegraphics[width=3.5in]{cro_eee.png}
-\end{frame}
-
-\begin{frame}{Test theory with FLUXNET data - GRA}
-  \includegraphics[width=3.5in]{gra_eee.png}
-\end{frame}
-
-\begin{frame}{Is uncertainty a function of VPD?}
-  \includegraphics[width=3.0in]{vpd_sigma.png}
-\end{frame}
-
-\begin{frame}{How to take VPD derivative?}
-  % below gnereated in fig03.py
-  \includegraphics[width=3.0in]{e_s_rh.png}
-\end{frame}
-
-% \begin{frame}{Summary statistics - extra sl}
-%   \begin{table}
-%     \caption{Statistics of $\frac{\partial \; ET}{\partial \, VPD}$ as a function of PFT.}
 %     \centering
-%     \begin{tabular}{l c c}
+%     \begin{tabular}{l c c c c c}
 %       \hline
-%       PFT & $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ & fraction $\frac{\partial \; ET}{\partial \; VPD} < 0.$ \\
+%       PFT & $R_{air}$ & $c_s$ (ppm) & $\gamma$ &  uWUE & \textbf{$VPD_{crit}$ (Pa)} \\
 %       \hline
-%       CRO & 0.000853  & 0.473311\\
-%       CSH & -0.108234 & 0.931660\\
-%       DBF & -0.012727 & 0.461674\\
-%       ENF & -0.034087 & 0.534425\\
-%       GRA & -0.019637 & 0.631735\\
+%       CRO &  288.680920 & 372.567691& 65.351523& 2.602873&  \textbf{133.165438} \\
+%       CSH &   289.067152& 381.593622& 67.613172& 2.175278& \textbf{4439.564212} \\
+%       DBF &   288.624437& 377.449849& 63.421812& 2.746393&  \textbf{888.773243} \\
+%       ENF &  288.183849& 377.676463& 61.559242& 4.015362&  \textbf{978.084845} \\
+%       GRA &  288.425651& 377.264645& 61.598768& 2.281074& \textbf{1141.630778} \\
 %       \hline
@@
 
+% \begin{frame}{Statistics with uncertainty}
+%   \scriptsize
 %   \begin{table}
+%     \caption{Statistics of $\frac{\partial \; ET}{\partial \; VPD}$ as a function of PFT.}
 %     \centering
-%   \begin{tabular}{l|r}
-%     Item & Quantity \\\hline
-%     Widgets & 42 \\
-%     Gadgets & 13
+%     \begin{tabular}{l c c c c c}
+%       \hline
+%       PFT & $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ & $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)$ & $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)*\text{std}(VPD)$ & $\frac{\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)*\text{std}(VPD)}{ \frac{\partial \; ET}{\partial \; R}\left(\overline{env}\right)*\text{std}(R)}$ & fraction $\frac{\partial \; ET}{\partial \; VPD} < 0.$ \\
+%       \hline
+%       CRO & 0.000853 & 0.026241 & 37.05 & 0.41 & 0.473311\\
+%       CSH & -0.108234 & -0.091526 & 101.72 & 0.88 & 0.931660\\
+%       DBF & -0.012727 & 0.013794 & 39.47 & 0.33 & 0.461674\\
+%       ENF & -0.034087 & 0.000706 & 33.22 & 0.30 & 0.534425\\
+%       GRA & -0.019637 & -0.000921 & 33.60 & 0.35 & 0.631735\\
+%       \hline
+%       \multicolumn{2}{l}{}
+
 %     \end{tabular}
-%   \caption{\label{tab:widgets}An example table.}
 %   \end{table}
+% \end{frame}
+
+% \begin{frame}{Scaling Term}
+%   \[\frac{\partial \;  ET}{\partial \; D} = \mathbf{ \frac{g_a \; P}{T(\Delta + \gamma)} }   \left(\frac{ c_p}{R_{air}} - \frac{\text{LAI$_{ref}$}}{\text{LAI}} \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D}}{2 (g_1 + \sqrt{D})^2}\right) \right)\]
+%   \begin{figure}
+%     \includegraphics[width=3.0in]{scaling.png}
+%     % \caption{}
+%   \end{figure}
+% \end{frame}
+
+% \begin{frame}{Extra slide - is theory VPD$_{crit}$ optimum?}
+%   \includegraphics[width=0.7\textwidth]{for_agu.pdf}
+% \end{frame}
+
+% \begin{frame}{Test theory with FLUXNET data - CSH}
+%   \includegraphics[width=3.5in]{csh_eee.png}
+% \end{frame}
+
+% \begin{frame}{Test theory with FLUXNET data - ENF}
+%   \includegraphics[width=3.5in]{enf_eee.png}
+% \end{frame}
+
+% \begin{frame}{Test theory with FLUXNET data - DBF}
+%   \includegraphics[width=3.5in]{dbf_eee.png}
+% \end{frame}
+
+% \begin{frame}{Test theory with FLUXNET data - CRO}
+%   \includegraphics[width=3.5in]{cro_eee.png}
+% \end{frame}
+
+% \begin{frame}{Test theory with FLUXNET data - GRA}
+%   \includegraphics[width=3.5in]{gra_eee.png}
+% \end{frame}
+
+% \begin{frame}{Consequences of theory - ``sign'' term}
+% Below is a little opaque (at least to me):
+% \[\frac{\partial \, ET}{\partial \, VPD} = \text{scaling} \cdot  \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]
+% so do expansion about $g_1 >> \sqrt{VPD}$:
+% \footnotesize
+% \[ \frac{\partial \, ET}{\partial \, VPD} \approx \text{scaling} \cdot  \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{1}{g_1} - \frac{3\sqrt{VPD}}{2 g_1^2} +  \mathcal{O}\left(\frac{\sqrt{VPD}}{g_1}\right)^{2} \right) \right)\]
+% % \large
+% % and:
+% %   \[ \sqrt{VPD} \Big|_{\frac{\partial ET}{\partial VPD} = 0} \approx  \frac{2 g_1}{3}  \left( -\frac{ c_p \, g_1\,  1.6  \; R*\; \text{ uWUE }}{R_{air} \gamma c_s } + 1  \right) \]
+
+% %   \[ VPD \Big|_{\frac{\partial ET}{\partial VPD} = 0} = \frac{R_{air}}{4 \, c_p} \left[ \frac{\gamma  c_s}{R* \, 1.6\, \text{uWUE}} + \sqrt{\left(g_1 + 8\, \frac{c_p}{R_{air}}\right) \frac{\gamma c_s}{R* \, 1.6\, \text{uWUE}}} - 4 \, g_1 \frac{ c_p}{R_{air}}\right] \]
+% \end{frame}
+
+% \begin{frame}{Is uncertainty a function of VPD?}
+%   \includegraphics[width=3.0in]{vpd_sigma.png}
+% \end{frame}
+
+% \begin{frame}{How to take VPD derivative?}
+%   % below gnereated in fig03.py
+%   \includegraphics[width=3.0in]{e_s_rh.png}
+% \end{frame}
+
+% % \begin{frame}{Summary statistics - extra sl}
+% %   \begin{table}
+% %     \caption{Statistics of $\frac{\partial \; ET}{\partial \, VPD}$ as a function of PFT.}
+% %     \centering
+% %     \begin{tabular}{l c c}
+% %       \hline
+% %       PFT & $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ & fraction $\frac{\partial \; ET}{\partial \; VPD} < 0.$ \\
+% %       \hline
+% %       CRO & 0.000853  & 0.473311\\
+% %       CSH & -0.108234 & 0.931660\\
+% %       DBF & -0.012727 & 0.461674\\
+% %       ENF & -0.034087 & 0.534425\\
+% %       GRA & -0.019637 & 0.631735\\
+% %       \hline
+% %       \multicolumn{2}{l}{}  
+% %     \end{tabular}
+% %   \end{table}
+% % \end{frame}
+
+% % \begin{table}
+% %   \centering
+% %   \begin{tabular}{l|r}
+% %     Item & Quantity \\\hline
+% %     Widgets & 42 \\
+% %     Gadgets & 13
+% %   \end{tabular}
+% %   \caption{\label{tab:widgets}An example table.}
+% % \end{table}
 

### 2017-12-18 doc/paper/vpd_et_paper.tex
think done with rough draft
confirmed fig05 good and refactored, but ntoe that I broke the paper version a little for the agu talk so I need to fix this (conversion from Pa to kPa)

@@
     \label{uwue}
-uWUE = \frac{GPP \cdot \sqrt{VPD}{ET}
+uWUE = \frac{GPP \cdot \sqrt{VPD}}{ET}
   \end{equation}
@@
   \hline
-PFT & $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ & $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{T, \ldots , VPD}\right)$ & $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{T, \ldots , VPD}\right)*\text{std}(VPD)$ & $\frac{\frac{\partial \; ET}{\partial \; VPD}\left(\overline{T, \ldots , VPD}\right)*\text{std}(VPD)}{ \frac{\partial \; ET}{\partial \; R}\left(\overline{T, \ldots , VPD}\right)*\text{std}(R)}$ & fraction $\frac{\partial \; ET}{\partial \; VPD} < 0.$ \\
-  \hline
-CRO & 0.000853 & 0.026241 & 18.523659 & 0.203022 & 0.473311\\
-CSH & -0.108234 & -0.091526 & 50.861613 & 0.439379 & 0.931660\\
-DBF & -0.012727 & 0.013794 & 19.734435 & 0.164241 & 0.461674\\
-ENF & -0.034087 & 0.000706 & 16.611852 & 0.148548 & 0.534425\\
-GRA & -0.019637 & -0.000921 & 16.798083 & 0.173552 & 0.631735\\
+PFT & $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ & $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)$ & $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)*\text{std}(VPD)$ & $\frac{\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)*\text{std}(VPD)}{ \frac{\partial \; ET}{\partial \; R}\left(\overline{env}\right)*\text{std}(R)}$ & fraction $\frac{\partial \; ET}{\partial \; VPD} < 0.$ \\
+  \hline
+CRO & 0.000853 & 0.026241 & 37.05 & 0.41 & 0.473311\\
+CSH & -0.108234 & -0.091526 & 101.72 & 0.88 & 0.931660\\
+DBF & -0.012727 & 0.013794 & 39.47 & 0.33 & 0.461674\\
+ENF & -0.034087 & 0.000706 & 33.22 & 0.30 & 0.534425\\
+GRA & -0.019637 & -0.000921 & 33.60 & 0.35 & 0.631735\\
 \hline
@@
 We therefor have the most confidence in our conclusion based on Equation \ref{d_et} for PFTS CSH, DBF, and ENF, as the full complexity plots with uncertainty included closely match the story when $\sigma$ is held fixed.
-**see somewhat preferred alternate figure \ref{real2} \explain{I think I like the alternate plot much  more as thinking in terms of T and RH is easier, and it makes the story easier to see at relatively low temperatures.
+% **see somewhat preferred alterate figure \ref{real2} \explain{I think I like the alternate plot much  more as thinking in terms of T and RH is easier, and it makes the story easier to see at relatively low temperatures.
 However, I used the other plot because Fig 5 does not discuss things in terms of temperature, as this would make things more complicated (adding another dimension).
@@
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{./fig06b.png}
-\caption{****alternate Fig 06****  Scatter plots of $\frac{\partial \; ET}{\partial \; VPD}$.
+% \begin{figure}[h]
+% \centering
+% \includegraphics[width=\textwidth]{./fig06b.png}
+% \caption{****alternate Fig 06****  Scatter plots of $\frac{\partial \; ET}{\partial \; VPD}$.
 Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; VPD}$, as labeled.
@@
 }
-\label{real2}
-\end{figure}
+% \label{real2}
+% \end{figure}
 

### 2017-12-18 src/codebase/data_calc.py

@@ -55,13 +55,15 @@ def pm_et(_df, vpd=None, uwue=None):
         *(1.0+_df['g1']/np.sqrt(vpd)))))\
         /(_df['delta']+_df['gamma'])
 
-def sign(_df, vpd=None):
+def sign(_df, vpd=None, uwue=None):
   """calculates the 'sign' term of et derivative w.r.t. vpd"""
   if vpd is None:
     vpd = _df['vpd']
+  if uwue is None:
+    uwue = _df['uwue']
   return CP/_df['r_moist']\
                 -_df['gamma']*_df['c_a']\
-                /(1.6*R_STAR*_df['uwue'])\
+                /(1.6*R_STAR*uwue)\
                 *((2.*_df['g1']+np.sqrt(vpd))\
                   /(2.0*(_df['g1']+np.sqrt(vpd))**2))
 

### 2017-12-18 src/paper_figure_code/fig05.py

@@ -3,14 +3,16 @@
 This script makes fig 5
 """
 from shared_functions import *
+
+# mean_df defined her at module-level
+
 fontsize=14
 linewidth=2.5
 dashedlinewidth=3.0
 markersize=8
 # below is whether to plot taylor series approximation
-PLOT_SERIES = False
 #below is to make slight changes for the talk
-PLOT_TALK = True
+PLOT_TALK = False
 
 PLOT_PET = False
 
@@ -32,153 +34,52 @@ else:
                r'{2 (g_1 + \sqrt{D})^2}\right)\right)$'
 
 
-# ##### Figure 5 #####
-## term 2 movwd to shared_figs
-
-def term_2_approx(_df, lai, vpd, order=4):
-  """calculates term 2"""
-  atmos = {'gamma' : _df.gamma.mean(), 'c_a' : _df.c_a.mean(),\
-           'vpd' : vpd}
-  if _df.uwue.std() > 1.e-8:
-    print('error, uWUE is variable: %f!!!!' % _df.uwue.std())
-  elif _df.g1.std() > 1.e-8:
-    print('error, g1 is variabile: %f!!!!!' % _df.g1.std())
-  canopy = {'uwue' : _df.uwue.mean(), 'g1' : _df.g1.mean()}
-  if order == 4:
-    return d_calc.CP/_df.r_moist.mean() \
-      -atmos['gamma']*atmos['c_a']*d_calc.LV/\
-      (lai*1.6*d_calc.R_STAR*canopy['uwue'])\
-      *(1./canopy['g1'] - 3.*np.sqrt(atmos['vpd'])/(2.*canopy['g1']**2)
-        + 2.*atmos['vpd']/canopy['g1']**3\
-        - 5.*np.sqrt(atmos['vpd'])**3/(2.*canopy['g1']**4))
-  elif order == 2:
-    return d_calc.CP/_df.r_moist.mean() \
-      -atmos['gamma']*atmos['c_a']*d_calc.LV/\
-      (lai*1.6*d_calc.R_STAR*canopy['uwue'])\
-      *(1./canopy['g1'] - 3.*np.sqrt(atmos['vpd'])/(2.*canopy['g1']**2))
-  else:
-    print('error uncoded order number %d' % order)
-    return
-
-
-grouped = df.groupby('pft')
[…]

### 2017-12-18 src/paper_figure_code/shared_functions.py

@@ -29,11 +29,12 @@ with open('%s/changjie/diagnosed_data.pkl' % os.environ['DATA'],\
 df = dfs['full']
 mean_df = dfs['mean']
 
-def et_min_vpd(_df, lai):
+def et_min_vpd(_df, uwue=None):
   """calculates theoretical max vpd as functoin of -df and lai"""
-  c3 = pm.CP/_df.r_moist
-  # c1 = n_df.gamma*_df.c_a/(lai*pm.R_STAR*1.6*_df.uwue_norm)
-  c1 = _df.gamma*_df.c_a/(lai*pm.R_STAR*1.6*_df.uwue_norm)
+  if uwue is None:
+    uwue = _df.uwue
+  c3 = d_calc.CP/_df.r_moist
+  c1 = _df.gamma*_df.c_a/(d_calc.R_STAR*1.6*uwue)
   c2 = _df.g1
   sqrt_vpd = (c1 + np.sqrt(c1 + 8.*c2*c3)*np.sqrt(c1)-4.*c2*c3)/(4.*c3)
   try:
@@ -41,21 +42,10 @@ def et_min_vpd(_df, lai):
   except TypeError:
     if sqrt_vpd < 0.:
       sqrt_vpd = np.nan
-      print("setting to nan")
   return sqrt_vpd**2
+
 
 def get_pft(_df):
   return _df['pft'].iloc[0]
-
-def term_2(_df, lai, vpd):
-  """calculates term 2, takes in a _df of one pft"""
-  atmos = {'gamma' : _df.gamma.mean(), 'c_a' : _df.c_a.mean(),\
-           'vpd' : vpd}
-  if _df.uwue.std() > 1.e-8:
-    print('error, uWUE is variable: %f!!!!' % _df.uwue.std())
-  elif _df.g1.std() > 1.e-8:
-    print('error, g1 is variabile: %f!!!!!' % _df.g1.std())
-  canopy = {'uwue' : _df.uwue.mean(), 'g1' : _df.g1.mean()}
-  return pm.CP/_df.r_moist.mean() +calc.leaf_vpd(atmos, canopy, lai)
 
 # df.index = df.reset_index(drop=True)

### 2017-12-19 doc/paper/vpd_et_paper.tex
done with writing portion, one important thing to iron out is whether zhou et al 2015 abosolute values has any information relatevant to our data sampling, or is it just the form that's importnatn?

@@
 \label{data}
-We use data from FLUXNET2015.
+We use data from FLUXNET2015, including all sites with more than four years of data.
+Each site's plant functional type (PFT) was classified using the International Geosphere-Biosphere Programme vegetation classification scheme (\cite{Loveland_1999}).
 Because $g_1$ coefficients \citep{Lin_2015} and uWUE were only both available for five plant functional types (PFTs - see Table \ref{pft}),  only 56 of the 77 sites were used.
@@
 
+The purpose of this study is examine ecosystem response to atmospheric drying during the growing season.
+To accomplish this, we filter and quality control the data using a similiar procedure as \cite{Zhou_2015}:
+\begin{itemize}
+\item Only measured or highest (``good'') quality gapfilled data, according to qualtiy control flags, are used.
+\item To isolate the growing season, we only use days in which the average GPP exceeds 10\% of the observed 95th percentile of GPP for a given site.
+\item We remove days with rain and the day following to avoid issues with rain interveption and sensor satruation and high relative humidity (Medlyn et. al. 2017).
+\end{itemize}
+Additionally, we restrict data to the daytime which is identified when downwelling shortwave radation is greater than 50 W m$^{-2}$ and sensible heat flux is greater than 5 W m$^{-2}$.
+To reduce further reduce the chance of sensor saturation at high relative humidity, we remove all time steps for which VPD is less than .01 kPa.
+Timesteps with negative observed GPP or ET are also removed.
+After these quality control proceedures, 332556 upscaled hourly observations remain.
+
 \section{Methods}
-n
 The Penman-Monteith equation (hereafter PM) estimates ET as a function of atmospheric and land-surface variables:
@@
 
-While Equation 3 can be used in PM, it will make analytical work with the function intractable because $A$ is a function of ET itself.
+While Equation 3 can be used in PM, it will make analytical work with the function intractable because $A$, net CO$_2$ assimilation, is closely functionally related to ET itself.
 To remove dependence of ET on $A$ we can use the semi-empirical results of \citet{Zhou_2015}.
@@
 The variablility of $\sigma$ across sites and time will provides some measure of uncertainty in our model, assumptions, as well as the fluxnet observations themselves.
-Mean shifts in $\sigma$ can be attributed to differences in out sampling from those used to calcualte $uWUE$ in \citet{Zhou_2015}.
-The influence of $\sigma$ will propogate through any uncertainty to our derivative of Equation \ref{et_sigma}:
+To correct for differences in sampling between \cite{Zhou_2015} and our data, we set uWUE such that $\overline{\sigma} = 1$.
+The variability of $\sigma$ then propogates through any uncertainty to our derivative of Equation \ref{et_sigma}:
 
@@
 With Equation \ref{d_et} we have provided an analytical franework for ecosystem reponse to atmospheric demand perturbations.
-There are a few subtelties to taking the derivative in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$) and $VPD$ are functionally related, so while taking the derivative we evaluate $\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET} {\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD} \Big|_{\text{RH fixed}} + \frac{\partial \; ET}{partial \; RH} \frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$ fixed}}$.
+There are a few subtelties to taking the derivative in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$) and $VPD$ are functionally related, so while taking the derivative we evaluate $\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET} {\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD} \Big|_{\text{RH fixed}} + \frac{\partial \; ET}{\partial \; RH} \frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$ fixed}}$\explain{Should I take the derivative this way?
+I think it's more mathematically correct, but it makes the explanation more complicated and only changes the solution by a factor of two.}.
 $RH$ and $e_s$ are assumed to be approximately orthogonal.
@@
 \begin{table}
-\caption{Definition of symbols and variables}
+\caption{Definition of symbols and variables, with citation for calculation if applicable.}
 \centering
-\begin{tabular}{l c c}
-\hline
- Variable & Description & Units  \\
-\hline
-$e_s$  & saturation vapor pressure & Pa  \\ 
-$T$  & temperature  & K \\
-$\Delta$  & $\frac{\partial e_s}{\partial T}$ & Pa K$^{-1}$ \\
-$R$  & net radiation at land surface minus ground heat flux & W m$^{-2}$   \\
-  $g_a$  & aerodynamic conductance & m s$^{-1}$  \\
-  $\rho_a$  & air density & kg m$^{-3}$  \\
-  $c_p$  & specific heat capacity of air at constant pressure & J K$^{-1}$ kg$^{-1}$ \\
-  $VPD$  & vapor pressure deficit & Pa  \\
-  $\gamma$  & psychrometric constant & Pa K$^{-1}$   \\
-  $g_s$  & stomatal conductance & m s$^{-1}$  \\
-  $g_{l-s}$  & leaf-scale stomatal conductance & mol m$^{-2}$ s$^{-1}$  \\
-  $R*$ & universal gas constant & J mol$^{-1}$ K$^{-1}$ \\
-  $LAI$ & leaf area index & -\\
-  $\sigma$ & uncertainty parameter & -\\
-  $c_s$ & CO$_2$ concentration & $\mu$ mol CO$_2$ mol$^{-1}$ air\\
+\begin{tabular}{l c c c}
+\hline
+ Variable & Description & Units & Citation \\
+\hline
+$e_s$  & saturation vapor pressure & Pa  & - \\ 
+$T$  & temperature  & K & - \\
+$\Delta$  & $\frac{\partial e_s}{\partial T}$ & Pa K$^{-1}$ & - \\
+$R$  & net radiation at land surface minus ground heat flux & W m$^{-2}$   & - \\
+  $g_a$  & aerodynamic conductance & m s$^{-1}$  & \cite{Thom_1977}, \cite{Paulson_1970}, \cite{Beljaars_1991} \\
+  $\rho_a$  & air density & kg m$^{-3}$  & - \\
+  $c_p$  & specific heat capacity of air at constant pressure & J K$^{-1}$ kg$^{-1}$ & - \\
+  $VPD$  & vapor pressure deficit & Pa  & - \\
+  $\gamma$  & psychrometric constant & Pa K$^{-1}$   & - \\
+  $g_s$  & stomatal conductance & m s$^{-1}$  & \cite{MEDLYN_2011} \\
+  $g_{l-s}$  & leaf-scale stomatal conductance & mol m$^{-2}$ s$^{-1}$  & \cite{MEDLYN_2011} \\
+  $R*$ & universal gas constant & J mol$^{-1}$ K$^{-1}$ & - \\
+  $LAI$ & leaf area index & -& - \\
+  $\sigma$ & uncertainty parameter & -& - \\
+  $c_s$ & CO$_2$ concentration & $\mu$ mol CO$_2$ mol$^{-1}$ air& - \\
 \hline
@@
 
-We restrict our analysis to the daytime (sensible heat > 5 W m$^{-1}$ and shortwave radiation > 50 W m$^{-2}$) when there is no precipitation and the plants are growing (GPP > 10\% of the 95th percentile) \explain{This GPP thresholding was used by Changjie, do you know if there is a citation for it?
-Otherwise it seems like something a reviewer would have issue with as it is arbitray}.
-Also, because some sites use half hourly data but some use hourly, we aggregate all data to hourly averages.
-Only times with good quality control flags are used.
-
 \section{Results}
@@
 By construction, the variability in the $\sigma$ term (Equation \ref{lai}) contains all model and observational uncertainties.
-For an observation that perfectly matches our model and assumptions $\sigma$ will be one.
-Therefore, if for our assumptions and framework to be reasonable $\sigma$ should be $O(1)$.
-Figure \ref{lai_fig} presents the histogram of calculated $\sigma$s with outliers (lowest and highest 5\% percent) and nonphysical values ($\sigma$ $<$ 0.) removed.
+For an observation that perfectly matches our model and constant uWUE assumption $\sigma$ will be one.
+Therefore, for our assumptions and framework to be reasonable $\sigma$ should be $O(1)$.
+Figure \ref{lai_fig} presents the histogram of calculated $\sigma$s with outliers (lowest and highest 5\% percent).
 All remaining $\sigma$ values are $O(1)$ which provides confidence in model framework.

### 2017-12-19 src/analysis.py

@@ -12,14 +12,19 @@ importlib.reload(d_io)
 importlib.reload(d_prep)
 importlib.reload(d_calc)
 
-# df = d_io.load_mat_data()
-# df.to_pickle('%s/changjie/loaded_data.pkl' % os.environ['DATA'])
+df = d_io.load_mat_data()
+df.to_pickle('%s/changjie/loaded_data.pkl' % os.environ['DATA'])
 df = pd.read_pickle('%s/changjie/loaded_data.pkl' % os.environ['DATA'])
+print('pre-prep data shape:', df.shape)
 
 prep_df = d_prep.generate_vars(df)
+print('pre-calc data shape:', prep_df.shape)
 df.to_pickle('%s/changjie/prepped_data.pkl' % os.environ['DATA'])
 importlib.reload(d_calc)
 dfs = d_calc.all_diagnostics(prep_df)
+prep_df = d_prep.generate_vars(df)
+print('post-calc data shape:', dfs['full'].shape)
+
 with open('%s/changjie/diagnosed_data.pkl' % os.environ['DATA'],\
           mode='wb') as file :
   pickle.dump(dfs, file)

### 2017-12-19 src/codebase/data_io.py

@@ -25,6 +25,8 @@ WUE.index = WUE.PFT
 # and from kg H20 to joules
 WUE.loc[:, 'u_wue_yearly'] = WUE.loc[:, 'u_wue_yearly']\
                              *1.e6/12.011*np.sqrt(100.)/LV
+WUE.loc[:, 'u_wue_yearly_std'] = WUE.loc[:, 'u_wue_yearly_std']\
+                             *1.e6/12.011*np.sqrt(100.)/LV
 
 
 SITELIST = pd.read_csv('%s/changjie/fluxnet_algorithm/'\
@@ -96,8 +98,10 @@ def load_file(filename):
     return None
   try:
     data_out['uwue_zhou'] = WUE.loc[pft, 'u_wue_yearly']
+    data_out['uwue_zhou_std'] = WUE.loc[pft, 'u_wue_yearly_std']
   except KeyError:
     print('Setting Zhou uWUE to NaN for PFT: %s' % pft)
+    data_out['uwue_zhou_std'] = np.nan
     data_out['uwue_zhou'] = np.nan
   data_out['site'] = ''.join(filename.split('/')[-1].split('.')[:-1])
   data_out = data_out.dropna()

### 2017-12-20 doc/paper/vpd_et_paper.tex
done with wed's writing

@@
 Additionally, we restrict data to the daytime which is identified when downwelling shortwave radation is greater than 50 W m$^{-2}$ and sensible heat flux is greater than 5 W m$^{-2}$.
-To reduce further reduce the chance of sensor saturation at high relative humidity, we remove all time steps for which VPD is less than .01 kPa.
+To further reduce the chance of sensor saturation at high relative humidity, we remove all time steps for which VPD is less than .01 kPa.
 Timesteps with negative observed GPP or ET are also removed.
@@
 
-This can be adapted to an ecosystem-scale stomatal resistance by multiplying by leaf area index (LAI) and converting units to m s$^{-1}$
+This can be adapted to an ecosystem-scale stomatal resistance by multiplying by leaf area index (LAI) and converting units to m s$^{-1}$:
 
@@
 Before calculating the sensitivity of ET to VPD, it is useful to consider the functional form of Equation \ref{d_et}.
-There are three terms: a scaling term for the full expression we will call Term 1 ($\frac{g_a \; P}{T(\Delta + \gamma)}$), a relatively constant offset we will call Term 2 ($\frac{c_p}{R_{air}}$), and a variable term we will call Term 3 ($\frac{\gamma c_s }{1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right)$).
-All variables are positive, so the relative magnitude between Term 2 and Term 3 will determine the sign of the derivative, while Term 1 will scale the expression larger or smaller.
+There are two main terms: a ``scaling'' term which modifies the mangnitude but not the sign:
+
+\[\frac{g_a \; P}{T(\Delta + \gamma)},\]
+
+and a ``sign'' term which determines whether the derivative is postive or negative:
+
+\[\frac{c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right).\]
+ All variables are positive, so the relative magnitude between the first term and the second term will deterimine whether ET increases or decreases with increasing VPD.
+If the second term is larger magnitude then plant response dominates, while if the first term is larger then atmospheric demand dominats.
+Given that the sign term is crucial to answering our question ``When does VPD drive or reduce ET?'', we will explore its functional form first.
+
+\subsection{Functional Form of the Sign Term}
 
 Term 2 minus Term 3's role in determining the sign of the sensitivty of ET to VPD makes it crucuial for answering our question ``When does VPD drive or reduce ET?'' Exploring these terms more, $c_s$ and $\gamma$ variability is relatively less than $\sigma$ and $VPD$ variability, so variability within PFT will be solely determined by $\sigma$ and $VPD$.
-If we fix uncertainty $\sigma$ at PFT averages, then Term 2 minus Term 3 is just a function of $VPD$.
+If we fix uncertainty $\sigma$ at 1, then Term 2 minus Term 3 is just a function of $VPD$.
 We can further dertmine a $VPD_{crit}$ where $\frac{\partial \; ET}{\partial \; VPD} = 0$:
@@
 For any values of $VPD$ less than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be negative, and for values of $VPD$ greater than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be positive.
+Intuitively this makes sense; as VPD increases ecosystems will start to run out of strategies to hold back water, and atmospheric demand will increase.
+Any further increases in atmospheric demand will then result in an increase in VPD.
+The extent to which a given ecosystem has evolved to conserve water will determine at what VPD atmospheric demand takes over.
 
@@
 
-Figure \ref{term3} provides greater description for how (Term 2 - Term 3) varies with $VPD$, as a function of PFT.
+Figure \ref{term3} provides greater description for how the sign term varies with $VPD$, as a function of PFT.
 Equation \ref{d_et_approx} aids interpretation of Figure \ref{term3}.
-Larger $uWUE\cdot\overline{\sigma}$, and $g1$ shift the leading-order constant term ($\frac{1}{g_1}$) towards smaller values, and (Term 2 - Term 3) towards positive values.
+Larger $uWUE$, and $g1$ shift the leading-order constant term ($\frac{1}{g_1}$) towards smaller values, and the sign term towards overall positive values.
 $uWUE$ and $g1$ are both water-use effiency type constants.
-Higher values corresponde to plants that are more willing to spend water on primary production and less evolved to conserve water.
-Figure \ref{term3} confirms our physical intuition: CROs are the least water conservative so have the smallest constant portion of Term 3, while CSH are the most water conservative and have the largest constant portion of Term 3.
+Higher values corresponde to plants that are more willing to spend water on primary production and less evolved to conserve water.\explain{really need to think about this sentence more, tehre is more nuance and I actually thing g1 and uwue really partially cancel.} Figure \ref{term3} confirms our physical intuition: CROs are the least water conservative so have the smallest constant portion of Term 3, while CSH are the most water conservative and have the largest constant portion of Term 3.
 For the VPD-dependent terms in Equation \ref{d_et_approx}, differences in $g1$ between PFTs exert a greater influence than difference $uWUE$, as the power of $g1$ increases.
 Increasing $uWUE$ and $g1$ decreases the VPD-dependence, but $g1$ has the bigger effect due to its increasing powers.
-ENF ($g1 = 74.31$) has by far the largest VPD dependence of response, while CRO ($g1 = 183.1$) has the smallest VPD dependence.
+ENF ($g1 = 74.31$) has by far the largest VPD dependence of response, while CRO ($g1 = 183.1$) has the smallest VPD dependence.\explain{could probably simplify and just say that g1 controls VPD dependence.}
 
@@
 \includegraphics[width=20pc]{./fig05.pdf}
-\caption{Sources of variability for Term 2 - Term 3.
-Top: Term 2 - Term 3 as a function of VPD, with $\sigma$ held fixed at PFT averages.
+\caption{Sources of variability for sign term.
+Top: sign term as a function of VPD, with $\sigma$ held fixed at PFT averages.
 The observed range of VPD for each PFT is also shown below the x-axis.
@@
 Figure \ref{term3}b shows the location of the minima of ET, as a function of $\sigma$ and $VPD$.
-For any $\sigma$ or VPD less (more) than these curves, Term 2 - Term 3 will be negative (positive).
+For any $\sigma$ or VPD less (more) than these curves, sign term will be negative (positive).
 It is clear that the portion of VPD observations below/above these curves will be a strong function of $\sigma$.
@@
 For CSH, $\frac{\partial \; ET}{\partial \; VPD}$ should be negative for the vast majority of observed $\sigma$ and VPD.
-The fraciton of positive $\frac{\partial \; ET}{\partial \; VPD}$ appears to be more even for ENF, GRA, and DBF, and we might expect a greater frequency of positive $\frac{\partial \; ET}{\partial \; VPD}$ for CRO.
-
-While the above discussion of the sign of $\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our research question, the magnitude of $\frac{\partial \; ET}{\partial \; VPD}$ will also impact statistics of $\frac{\partial \; ET}{\partial \; VPD}$ and the importance of $VPD$ variablility for $ET$ variability.
+The fraction of positive $\frac{\partial \; ET}{\partial \; VPD}$ appears to be more even for ENF, GRA, and DBF, and we might expect a greater frequency of positive $\frac{\partial \; ET}{\partial \; VPD}$ for CRO.
+
+\subsection{Functional Form Scaling Term}
+
+While the above discussion of the sign of $\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our research question, the magnitude of $\frac{\partial \; ET}{\partial \; VPD}$ will also impact statistics of $\frac{\partial \; ET}{\partial \; VPD}$ and the importance of $VPD$ variablility for overall $ET$ variability.
 So we will more closely examine the scaling Term 1: $\frac{P}{T} \propto \rho$, so this should vary little relative to aerodynamic conductance and $\Delta$.
@@
 $\Delta$'s presence in the scaling term also matches physical intition.
-Evaporative cooling will dampen the ability of the atmosphere to take more moisture, because $e_{s}$ decreases with temperature.
+Evaporative cooling will dampen the ability of the atmosphere to take more moisture, because $e_{s}$ decreases with decreasing temperature.
 The decrease in $e_{s}$ is proportional to $\Delta$ ($\delta e_{s} = \Delta \delta T$).
@@
 So most of the differences in scaling between PFT should be in the aerodynamic conductance term.
-One interesting side note is that the coefficient of variability for both aerodynamic conductance and Term 1 is relatively constant across PFT, suggesting that the influence of aerodynamic conductance on the relative (to the PFT mean) variability of Term 1 is approximately similar across PFT.
-
-Figure \ref{scale_vary}A shows Term 1 normalized by mean aerodynamic conductance (calculated for each plant functional type), and confirms that much of the relative variability of Term 1 is contained in the aerodynamic conductance variability.
+One interesting side note is that the coefficient of variability for both aerodynamic conductance and the scaling term is relatively constant across PFT, suggesting that the influence of aerodynamic conductance on the relative (to the PFT mean) variability of the scaling term is approximately similar across PFT.
+
+Figure \ref{scale_vary}A shows scaling term normalized by mean aerodynamic conductance (calculated for each plant functional type), and confirms that much of the relative variability of the scaling term is contained in the aerodynamic conductance variability.
 Generally, $T$ has less of a role.
 Additionally, the impact of $T$ on the relative variability increases with increasing aerodynamic conductance.
-\explain{plot every PFT and show they collapse onto the same curve?
-- might just be too messy though}.
-
-While the relative variability of Term 1 is similar across PFT, the absolute value of Term 1 varies strongly across PFT.
-Figure \ref{scale_vary}B shows Term 1 evaluated with the mean aerodynamic conductance for each PFT, and at the range of observed temperatures for each PFT.
-As expected, for the tree PFTs (DBF, ENF) Term 1 is much larger and the temperature dependence is much stronger.
-Systematic differences in observed temperatures also cause differences in the average magnitude of Term 1.
+
+While the relative variability of scaling term is similar across PFT, the absolute value of scaling term varies strongly across PFT.
+Figure \ref{scale_vary}B shows scaling term evaluated with the mean aerodynamic conductance for each PFT, and at the range of observed temperatures for each PFT.
+As expected, for the tree PFTs (DBF, ENF) scaling term is much larger and the temperature dependence is much stronger.
+Systematic differences in observed temperatures also cause differences in the average magnitude of scaling term.
 For example, ENF experiences on average colder temperatures and is thus more likely to have a larger scaling term.
-Additionally, because the variability of aerodynamic conductance increases proportionally to the mean, the spread of Term 1 due to aerodynamic conductance variability will be larger for the tree PFTs, although this is not shown for simplicity.
-To summarize, the variability of Term 1 within each PFT will look like Figure \ref{scale_vary}A for each PFT, but the scale of the y-axis will increase or decrease according to mean aerodynamic conductance oberved in Figure \ref{scale_vary}B.
+Additionally, because the variability of aerodynamic conductance increases proportionally to the mean, the spread of scaling term due to aerodynamic conductance variability will be larger for the tree PFTs, although this is not shown for simplicity.
+To summarize, the variability of scaling term within each PFT will look like Figure \ref{scale_vary}A for each PFT, but the scale of the y-axis will increase or decrease according to mean aerodynamic conductance oberved in Figure \ref{scale_vary}B.
 
@@
 \includegraphics[width=20pc]{./fig04.pdf}
-\caption{Primary sources of variability for Term 1.
-A) Variability within each PFT: Term 1 normalized by mean $g_a$ for each PFT.
-B) Variability between each PFT: Term 1 evaluated at mean $g_a$ for each PFT.
+\caption{Primary sources of variability for scaling term.
+A) Variability within each PFT: scaling term normalized by mean $g_a$ for each PFT.
+B) Variability between each PFT: scaling term evaluated at mean $g_a$ for each PFT.
 Temperature range is 5-95th percentile for each PFT.
@@
 And, DBF in addition to CRO experiences $\frac{\partial \; ET}{\partial \; VPD}$ < 0 less than half the time.
-These observations highlight the effect of the nonlinear function in Figure \ref{term3}: $\frac{\partial \; ET}{\partial \; VPD}$ has a much steeper slope when the function is negative, and is thus more likely to be large.
+These observations highlight the effect of the nonlinear function in Figure \ref{term3}: $\frac{\partial \; ET}{\partial \; VPD}$ has a much steeper slope when the function is negative, and is thus more likely to be large magnitude.
 
 The units of $\frac{\partial \; ET}{\partial \; VPD}$ make it difficult to interpret if $VPD$ is even a meaningful contributor to ET's variability.
-To better understand $VPD$'s contribution, we normalize $\frac{\partial \; ET}{\partial \; VPD}$ with $VPD$'s standard deviation to define a (linearized) relative change in ET for variations in $VPD$ .
+To better understand $VPD$'s contribution, we mulitply $\frac{\partial \; ET}{\partial \; VPD}$ with $VPD$'s standard deviation to define a (linearized) relative change in ET for variations in $VPD$ .
 $VPD$'s contribution to ET's variability ranges between 30 - 40 W m$^{-2}$ for all PFTs except for CSH, which is about 100 W m$^{-2}$.
@@
 So far, idealized plots and statistics have illuminated the form of $\frac{\partial \; ET}{\partial \; VPD}$ and how it varies with PFT.
-Large mean $\sigma$ and uWUE shifts CRO and DBF towards positive $\frac{\partial \; ET}{\partial \; VPD}$.
+Large mean uWUE shifts CRO and DBF towards positive $\frac{\partial \; ET}{\partial \; VPD}$.
 However, the strongly nonlinear function of $\frac{\partial \; ET}{\partial \; VPD}$ at $\frac{\partial \; ET}{\partial \; VPD} < 0$ pushes $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ negative for DBF (it does not do this for CRO because of CRO's high $g1$).
@@
 This has the side effect of pushing $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ negative further than other PFTs for a given fraction $\frac{\partial \; ET}{\partial \; VPD} < 0$ and magnitude $\frac{\partial \; ET}{\partial \; VPD}(\overline{T,\ldots,VPD})$.
-GRA shows the opposite behavior; a relatively high $g1$  makes the function more linear, decreasing the magnitude of $-\overline{\frac{\partial \; ET}{\partial \; VPD}}$ for a given  [large] fraction $\frac{\partial \; ET}{\partial \; VPD} < 0$ and negative $\frac{\partial \; ET}{\partial \; VPD}(\overline{T,\ldots,VPD})$ (although aerodynamic conductance and Term 1 also probably have a role in this).
-Finally, low $uWUE$ of CSH pushes to toward by far the lowest values $\frac{\partial \; ET}{\partial \; VPD}$ (Figure \ref{term3}).
+GRA shows the opposite behavior; a relatively high $g1$  makes the function more linear, decreasing the magnitude of $-\overline{\frac{\partial \; ET}{\partial \; VPD}}$ for a given  [large] fraction $\frac{\partial \; ET}{\partial \; VPD} < 0$ and negative $\frac{\partial \; ET}{\partial \; VPD}(\overline{T,\ldots,VPD})$, However, aerodynamic conductance and scaling term also probably have a role in this.
+Finally, the low $uWUE$ of CSH pushes it toward by far the lowest values $\frac{\partial \; ET}{\partial \; VPD}$ (Figure \ref{term3}).
 Variability in $VPD$ accounts for the largest about of $ET$ variability for CSH.
-For the other PFTs, $VPD$ contributes less to $ET$ variability, but still represents about 15-20 \% of $R$'s contributions to ET variability.
+For the other PFTs, $VPD$ contributes less to $ET$ variability, but still represents about 15-20\explain{depending on how we take the derivative, this could be 30-40\%.} \% of $R$'s contributions to ET variability.
 
@@
 So far we have developed a theory for $\frac{\partial \; ET}{\partial \; VPD}$'s behavior.
-In particular, we determined a critical threshold $VPD_{crit}$ for $\frac{\partial \; ET}{\partial \; VPD}$.
-For $VPD < VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}  < 0$, and for $VPD > VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD} > 0$.
-This $VPD_{crit}$ is only a function of PFT.
-However, many assumptions, including constant uWUE, were made in deriving $VPD_{crit}$.
-So, we need to test if our theory and $VPD_{crit}$ hold up in the face of uncertainty.
+Most important to answering ``When does VPD drive or reduce ET?'' is that the sign of our derivative term is correct.
+So, we need to test if when we include uncertainty the observed derivitives match what our theory would say for the sign term.
+
+Figure \ref{agu_fig} shows the distribution of of the sign term when uncertainty is included, as compared to what the theory says it should be.
+Our goal for this paper was to capture the leading order behavior and guide intuition.
+For DBF and CSH our theory is successful towards this goal: the theory matches the leading order behavior of the function when uncertainty is included.
+Given the assumptions we've made, and the uncertainties of flux tower observations themselves, there will always be a large amount of noise.
+The plots for DBF and CSH look as one would expect if one just added noise on top of our theory.
+
+The theory does not match as well for CRO and GRA.
+Perhaps this should be expected.
+Our theory, for practical reasons, does not include important sources of variablility (particyularly between sites) for CRO and GRA.
+We do not account for seasonal changes in LAI, variability in crop height and surface roughness, differences in C3 vs C4 photosynthesis, or differences in irrigation practices between the sites.
+Given that the our thoery does not do as well for CRO and GRA, we will refrain from firm statements about how CRO and GRA respond to changes in VPD.
+
+ENF is a special case.
+The VPD dependence of the theory is not well captured by observations.
+However, the theory closely captures the intercept where $\frac{\partial \; ET}{\partial \; VPD} = 0$.
+Really resolving this feature is the key to answering our research question, so perhaps the theory is still useful for ENF for the purpose of addressing when VPD drives or reduces ET.
+It is also noteworthy that we have used published constants for $g1$, which determines the VPD dependence, from \cite{Lin_2015}.
+Had we fit our own constants we could have tweaked the theory to match the observations, but this was not our goal.
+
+As stated in the above discussion of ENF, we are mostly interested in the critical threshold $VPD_{crit}$ for which the sign of  $\frac{\partial \; ET}{\partial \; VPD}$ is dertmined.
+When $VPD < VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}  < 0$, and when $VPD > VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD} > 0$.
+Theoretically this $VPD_{crit}$ is only a function of PFT.
+We can look at how this value of $VPD_{crit}$ matches the data as different observed variables (including uncertainty) are allowed to vary.
 
@@
 Each column is a different quantity related to $\frac{\partial \; ET}{\partial \; VPD}$, and each row is a different PFT.
+Vertical black lines denote where our theory says $VPD_{crit}$ should be.
+In addition to seeing how the theory $VPD_{crit}$ matches observations, these plots are also useful for observing how the scaling term affects the magnitude of the derivative, if at all.
 
@@
 Through the columns of Figure \ref{real} we can see the impact of $\sigma$ and $g_a$ on the variability of $\frac{\partial \; ET}{\partial \; VPD}$.
-$g_a$'s scaling (included in columns 1 and 3) alters the magnitude considerably.
-$\sigma$ variability (included in columns 1 and 2) adds a lot of additional noise to the signal of $\frac{\partial \; ET}{\partial \; VPD}$, which is slightly undesirable given $\sigma$'s role in representing model and observational uncertainty.
+$g_a$'s scaling (included in columns 1 and 3) alters the magnitude considerably, particularly begtween observations.
+$\sigma$ variability (included in columns 1 and 2) adds a lot of additional noise to the signal of $\frac{\partial \; ET}{\partial \; VPD}$.
 However, the general story with the noise appears to match the cleaner signal when $\sigma$ is help constant and $VPD_{crit}$ is clearly visible.
-One exception is possibly with GRA, for which uncertainty represented in $\sigma$ is high and causes the full complexity plots (Columns 1 and 2) to not match well with $\sigma$ held fixed (Columns 3 and 4).
-
-For ENF and GRA $VPD_{crit}$ does not appear to be only a function of $\sigma$ (most observable in Column 4).
-It turns out that the site to site variability in $\gamma$ causes $VPD_{crit}$ to vary, which is not discussed in the previous section.
-The impact is observable in both ENF and GRA, but especially for ENF which has a larger $\frac{\partial^2 \; ET}{\partial^2 \; VPD}$ than the other PFTs.
+The exceptions are, as discussed above, for GRA and CRO, wher the full complexity plots (Columns 1 and 2) to not match well with $\sigma$ held fixed (Columns 3 and 4).
+
+While $g_a$ has a large impact on magnitudes in the scaling, temperature, through the $\Delta$ term, does not appear to have as much of an impact in the full observations.
+Therefore it appears that the absolute magnitude of the derivative can be mostly attributed to the sign term, and its direct scaling by aerodynamic conductance.
+Intuitively this makes sense; aerodynamic conductance will control how dominant balances at the land surface are communicated to the boundary layer.
 
@@
 We therefor have the most confidence in our conclusion based on Equation \ref{d_et} for PFTS CSH, DBF, and ENF, as the full complexity plots with uncertainty included closely match the story when $\sigma$ is held fixed.
-% **see somewhat preferred alterate figure \ref{real2} \explain{I think I like the alternate plot much  more as thinking in terms of T and RH is easier, and it makes the story easier to see at relatively low temperatures.
-However, I used the other plot because Fig 5 does not discuss things in terms of temperature, as this would make things more complicated (adding another dimension).
-I could just include both versions of figure 6 though (using figure 6a as a bridge to figure 6b)}.
+
 \begin{figure}[h]

### 2017-12-28 doc/paper/vpd_et_paper.tex
need to discuss vpd = 3.5 for pet in plots

@@
 
-Changes to vapor pressure deficit (VPD) alter the atmospheric demand for water from the land surface.
-Traditionally, atmospheric scientists and hydrometeorologists generally think that an increase in atmospheric demand induces an increases in evapotranspiration (ET) (citations?).
-This possible misconception developed in part due to the proliferation of studies examing potential ET (PET) rather than estimates of ET itself (citations?).
-In contrast, plant physiologists know that stomata have evolved to optimally regulate the exchange of water and carbon, and tend to close in response to increased atmospheric dryness \citep{Ball_1987, Leuning_1990, MEDLYN_2011}.
-Therefore, an increase (decrease) in VPD may not correspond to an increase (decrease) in ET because stomatal closure (opening) can cancel the effects of shifts to atmospheric demand.
+Answering the question ``When does vapor pressure deficit drive or reduce evapotranspiration?'' is motivated by two potenential but opposing perspectives on the matter.
+The first persepctive is that increases to vapor pressure deficit (VPD) increases atmospheric demand for water from the land surface, and this drives an increase in evapotranspiration (ET).
+This perspective is particularly relevant because potential evapotranspiration (PET), which is used in many drought indicies and hydrometeorological studies, only quantifies this atmospheric demand perspective.
+However, there is another possible perspective for vegetated surfaces.
+Plants' stomata have evolved to optimally regulate the exchange of water and carbon, and tend to close in response to increased atmospheric dryness \citep{Ball_1987, Leuning_1990, MEDLYN_2011}.
+Therefore, an increase in VPD may actually correspond to a decrease in ET because stomatal closure can cancel the effects of shifts to atmospheric de.
+So really a question of ``When does VPD drive or reduce ET?'' becomes a question of ``When does plant response or atmospheric demand dominate?.''
 
@@
 An increase in ET in reponse to atmospheric drying will reduce soil moisture, but contribute increased moisenting to the atmosperhere.
-Clearly, the sign and magnitude of land-surface responsedrives the co-evolution of the atmosphere and land-surface at many timescales, from diurna to interdecadal.
-
-We hypothesize that for most plant types a common resonse to increase in VPD will actually be a decrease in ET.
-The excpetion would be plants such as crops that are evolved (or bred) to prioritize gross primary production (GPP) over water conservation.
-However, for all otehr plants types, our hypothesis calls into question the validity of PET-based drought metrics deceloped by hydrometeorologists and used extensively in operations \citep[e.g.
-PDSI, P-PET,][]{2002, Alley_1984} \explain{more citations needed, including recent PET climate studies like Jack Scheff}.
-These metrics ignore the role of plants as gatekeepers for surface water loss to the atmosphere and have limited physical meaning for drought of vegetated land types.
-Additionally, plants evolved in arid climates should prioritize water conservation and we would expect a very negative ET response to increase in VPD.
-Therefore, vegetated locations most likely to experience droughts should show the strongest deviation between reality and a PET-based approximation.
+Understanding the one way response of the land surface to the atmosphere is a critical first step to disentangling land-atmosphere feedbacks.
+The sign and magnitude of land-surface response drives the co-evolution of the atmosphere and land-surface at many timescales, from diurnal to interdecadal.
+
+Before diving into math and data, we have some prior intuition for how the system might behave.
+First, whether VPD drive or reduces ET should be a function of plant type.
+Plants that evolved to conserve water (e.g.
+arid shrubs) will be more likely to reduce ET with increasing VPD, and plants that have evolved to care little about water (e.g.
+crops) will be more likely to increase ET with increasing VPD.
+Additionally, atmospheric conditions must matter as well.
+At the ecosystem scale there are limits to the strategies plants use to hold water back from the atmosphere.
+As atmospheric demand for water (VPD) increases, ecosystems will begin to reach their water conservation limits.
+At this stage any further increase in VPD will most likely drive an increase in ET, because there is little more the plants can do keep water back.
 
@@
 \label{lai_vpd_fig}
-\end{figure}
+\end{figure}nnn
 
@@
 Really resolving this feature is the key to answering our research question, so perhaps the theory is still useful for ENF for the purpose of addressing when VPD drives or reduces ET.
-It is also noteworthy that we have used published constants for $g1$, which determines the VPD dependence, from \cite{Lin_2015}.
-Had we fit our own constants we could have tweaked the theory to match the observations, but this was not our goal.
-
-As stated in the above discussion of ENF, we are mostly interested in the critical threshold $VPD_{crit}$ for which the sign of  $\frac{\partial \; ET}{\partial \; VPD}$ is dertmined.
+It is also noteworthy that we used published constants for $g1$, which determines the VPD dependence, from \cite{Lin_2015}.
+We could have fit our own constants to tweak the theory to match the observations, but this was not our goal.
+
+As stated in the above discussion of ENF, we are mostly interested in the critical threshold $VPD_{crit}$ for which the sign of  $\frac{\partial \; ET}{\partial \; VPD}$ is determined.
 When $VPD < VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}  < 0$, and when $VPD > VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD} > 0$.
@@
 
-The full observations generally confirm expectations from Section \ref{results}.
-CRO has the most positive values of $\frac{\partial \; ET}{\partial \; VPD}$, $\frac{\partial \; ET}{\partial \; VPD}$ is almost always negative for CSH, and response depends more with the environmental conditions for the other PFTs (especially ENF).
-Through the columns of Figure \ref{real} we can see the impact of $\sigma$ and $g_a$ on the variability of $\frac{\partial \; ET}{\partial \; VPD}$.
+The scatter plots generally confirm expectations from Figure \ref{agu_fig}.
+By comparing different columns of Figure \ref{real} we can see the impact of $\sigma$ and $g_a$ on the variability of $\frac{\partial \; ET}{\partial \; VPD}$.
 $g_a$'s scaling (included in columns 1 and 3) alters the magnitude considerably, particularly begtween observations.
-$\sigma$ variability (included in columns 1 and 2) adds a lot of additional noise to the signal of $\frac{\partial \; ET}{\partial \; VPD}$.
+$\sigma$ variability (included in columns 1 and 2) adds a lot of additional noise to the signal of $\frac{\partial \; ET}{\partial \; VPD}$, as observed in Figure \ref{agu_fig}.
 However, the general story with the noise appears to match the cleaner signal when $\sigma$ is help constant and $VPD_{crit}$ is clearly visible.
-The exceptions are, as discussed above, for GRA and CRO, wher the full complexity plots (Columns 1 and 2) to not match well with $\sigma$ held fixed (Columns 3 and 4).
+The exceptions are, as discussed above, for GRA and CRO, where the full complexity plots (Columns 1 and 2) to not match well with $\sigma$ held fixed (Columns 3 and 4).
 
@@
 Intuitively this makes sense; aerodynamic conductance will control how dominant balances at the land surface are communicated to the boundary layer.
-
-In general the full complexity plots of $\frac{\partial \; ET}{\partial \; VPD}$ match our expectations, even with the large sensitivity to $\sigma$ measures of uncertainty observed in Figure \ref{term3}.
-Our $\sigma$-based method of uncertainty propagation blurs the idealized expectations the most for GRA, and also has a considerable effect for CRO.
-We therefor have the most confidence in our conclusion based on Equation \ref{d_et} for PFTS CSH, DBF, and ENF, as the full complexity plots with uncertainty included closely match the story when $\sigma$ is held fixed.
+While $\Delta$ controlls the efficiency of latent fluxes in carying heat away from the surface, it appears this is a second order term, relative to $g_a$, for scaling $\frac{\partial \; ET}{\partial \; VPD}$.
+
+In general Figures \ref{real} and \ref{agu_fig} match our expectations, even with the large sensitivity to $\sigma$ seen in Figure \ref{term3}.
+Our $\sigma$-based method of uncertainty propagation blurs the idealized expectations the most for GRA and CRO.
+We therefor have the most confidence in our conclusion based on Equation \ref{d_et} for PFTS CSH, DBF, and ENF, as the sign of the derivative with uncertainty included matches (to leading order) the story when $\sigma$ is held fixed.
 

### 2017-12-28 doc/paper/vpd_et_paper.tex
wrote through conlcusions, now need to go back and just read throuhg (improving plots), then can send out to pierre.

@@
 
-The idealized representation of ET used here is successful in developing intuition for how ET responds to changes in $VPD$.
-This intuition will aid the community in interpreting observations and output from sophisticated full complexity climate models.
-
-The idealized framework leads to the following general conclusions:
-\begin{itemize}
-  \item Aerodynamic resistance plays an important role of scaling $\frac{\partial \; ET}{\partial \; VPD}$. This is a leading order effect for observing higher magnitude responses in DBF and ENF.
-  \item In general, CSH has the most negative (i.e.
-ET reduced) response to increases in $VPD$ (atmospheric drying).
-So CSH plants will almost always try and conserve water, effectively reducing ET with dry atmospheric perturbation.
-    \item Additionally for CSH, $VPD$ variability contributes the most to $ET$ variability.
-\item CRO has the most positive response (i.e. ET increased) in response to increases in $VPD$. This is consistent with CROs that may be evolved or bred to thrive in non-water-limited environments.
-\item The response is more a function of the environment for DBF, ENF, and GRA.
-Because as VPD increases the response is more likely to be positive, if RH is fixed then the response will be more likely to be positive at warmer T, or if T is fixed the response is more likely to be positive with decreasing RH.
-\item ENF has the strongest dependence on environmental conditions due to its small $g1$.
-\item Model and observational uncertainty is highest for GRA and CRO, so conclusions about those PFTs should be tempered.
-\item However, inclusion of uncertainty doesn't alter conclusions about DBF, ENF, and CSH.
-\end{itemize}
-
-The intuition developed using this framework can be used to understand how the land surface will respond and contribute to changes in the environment.
-Additionally, Equation \ref{et} gives provides an estimate of ET that requires no additional information beyond that required to calculate PET.
-Given that for all PFTs, with the exception of CRO, we found a high frequency and magnitude of negative $\frac{\partial \; ET}{\partial \; VPD}$, PET is a physically unrealisitic representation of ET for vegetated surfaces and PET-based drought metrics are not usefull.
-We advocate for drought metrics using Equation \ref{et} instead of PET.
+We derived a new form of Penman Monteith using $uWUE$ (\cite{Zhou_2015}) to remove implicit depedence on ET in the stomatal conductance term.
+With our new form of Penman Monteith we deveoped a theory for when an ecosystem will tend to reduce or increase ET with increasing VPD.
+The goal was to capture the leading order behavior of the stem to gain some intrinsic knowledge for its behaviro.
+This intuition can be used to disentangle land atmosphere feedbacks in more complicated scenariors, and will aid interpretation of obserations and sophisticated models.
+
+Our theory states that each plant type will have it's own critical VPD below which ET will decrease (plant response dominating) and above which ET will increase (atmospheric demand dominates).
+We tested our theory using data from FLUXNET 2015, and found that for DBF and CSH the theory succeeds in capturing the leading order behavor.
+For ENF, the thoery succeeds in capturing the leading order behavior for the sign of $\frac{\partial \; ET}{\partial \; VPD}$, but not for magnitude.
+In contrast, the model does not capture the behavior for CRO and GRA.
+In some ways we should have expected this, as our theory does not account for large variability we would expect for those plant types, including varying surface roughness, c3 vs c4 photosynthesis, and potential differences in irrigationa cross sites.
+Given these diffincies for CRO and GRA, we will withhold conclusions on their response to increasing VPD.
+
+For the forest sites (DBF, ENF) the observed environmental VPD approximately straddles the critical VPD, with about half the observations below the critical VPD and half above.
+However for CSH the environmental VPD never exceeds the critical VPD, and even when uncertainty is included the sign of the derivative is almost always negative (93\% of observations).
+
+While our theory predicts common occurence of positive $\frac{\partial \; ET}{\partial \; VPD}$ (atmospheric demand dominates), the plant response in attenuating ET from the land surface is still important even when $\frac{\partial \; ET}{\partial \; VPD} > 0$.
+The magnitude of the ``sign term'' of our theory is always far below what it would be if we only considered the atmospheric demand term.
+This means that any drought indicies or analysis using PET will not capture how the vegetated surface respond and react to increases in atmospheric demand.
+However, the new uWUE-version of Penman Monteith we derived could be used as a complement to PET in drought indices.
+Basically one could compare calculate and compare indices  when plant response is ignored (PET) or included (uWUE-Penman Monteith).
+
+This paper provides the initial intuition for how the land surface respondes to atmospheric drying.
+This intuition for the one-way behavior of  land surface resonse to atmospheric drying is critical first step to disentangling land-atmosphere feedbacks at various scales, from diurnal to seasonal and beyond.
 

### 2018-01-05 doc/paper/vpd_et_paper.tex
finished with my moring writing, need to revert src back to previous commit though for figures to be good

@@
 The first persepctive is that increases to vapor pressure deficit (VPD) increases atmospheric demand for water from the land surface, and this drives an increase in evapotranspiration (ET).
-This perspective is particularly relevant because potential evapotranspiration (PET), which is used in many drought indicies and hydrometeorological studies, only quantifies this atmospheric demand perspective.
+This perspective is particularly relevant because potential evapotranspiration (PET), which is used in many drought indicies and hydrometeorological studies, only quantifies atmospheric demand effects.
 However, there is another possible perspective for vegetated surfaces.
@@
 Quantifying the plant response to a perturbation to atmospheric VPD increases our understanding of land surface response to shifts in atmospheric conditions.
-If plant response reduces ET in response to atmospheric drying then soil moisture will be conserved.
+If plant response reduces ET in response to atmospheric drying then soil moisture will be conserved for later ET.
 An increase in ET in reponse to atmospheric drying will reduce soil moisture, but contribute increased moisenting to the atmosperhere.
@@
 Before diving into math and data, we have some prior intuition for how the system might behave.
-First, whether VPD drive or reduces ET should be a function of plant type.
+First, whether VPD drives or reduces ET should be a function of plant type.
 Plants that evolved to conserve water (e.g.
@@
 Each site's plant functional type (PFT) was classified using the International Geosphere-Biosphere Programme vegetation classification scheme (\cite{Loveland_1999}).
-Because $g_1$ coefficients \citep{Lin_2015} and uWUE were only both available for five plant functional types (PFTs - see Table \ref{pft}),  only 56 of the 77 sites were used.
+Because physical constats used in Section \ref{methods} are only published for five plant functional types (PFTs - see Table \ref{pft}),  only 56 of the 77 sites were used.
 Figure \ref{map_fig} \explain{map needs to be improved - it's a placeholder for now} presents each site and its plant functional type.
 
-The purpose of this study is examine ecosystem response to atmospheric drying during the growing season.
+The purpose of this study is to examine ecosystem response to atmospheric drying during the growing season.
 To accomplish this, we filter and quality control the data using a similiar procedure as \cite{Zhou_2015}:
@@
 \section{Methods}
+\label{methods}
 The Penman-Monteith equation (hereafter PM) estimates ET as a function of atmospheric and land-surface variables:
@@
 \includegraphics[width=20pc]{./fig01.pdf}
-\caption{Plant functional type and location of sites used in analysis. ***This is just a placeholder for now and needs to be improved i.e. with lat lon, better placement of continents, etc.)***}
+\caption{Plant functional type and location of sites used in analysis. ***\textbf{This is just  aplaceholder for now, I am currently fixing it***}}
 \label{map_fig}
@@
 Therefore, for our assumptions and framework to be reasonable $\sigma$ should be $O(1)$.
-Figure \ref{lai_fig} presents the histogram of calculated $\sigma$s with outliers (lowest and highest 5\% percent).
+Figure \ref{lai_fig} presents the histogram of calculated $\sigma$s with outliers removed (lowest and highest 5\% percent).
 All remaining $\sigma$ values are $O(1)$ which provides confidence in model framework.
@@
 \includegraphics[width=20pc]{./fig03.png}
-\caption{The joint distribution of $VPD$ and $\sigma$.
-$\sigma$ has only a weak dependence on $VPD$.
-***This plot could proabably benefit from a box plot of site specific correlations, because some sites do have stronger depednence than others.
-Note also Figs 3 and 2 can probably be combined because this figure shoes $\sigma$'s histogram.***}
+\caption{The joint distribution of $VPD$ and $\sigma$. $\sigma$ has only a weak dependence on $VPD$. \textbf{***note I am redoing/cleaning up this plot, ``lai'' should read sigma***}}
 \label{lai_vpd_fig}
-\end{figure}nnn
-
-Before calculating the sensitivity of ET to VPD, it is useful to consider the functional form of Equation \ref{d_et}.
+\end{figure}
+
+Before calculating the sensitivity of ET to VPD, we will consider the functional form of Equation \ref{d_et}.
 There are two main terms: a ``scaling'' term which modifies the mangnitude but not the sign:
 
-\[\frac{g_a \; P}{T(\Delta + \gamma)},\]
+\begin{equation}
+  \frac{g_a \; P}{T(\Delta + \gamma)},
+\end{equation}
 
@@
 
-\[\frac{c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right).\]
- All variables are positive, so the relative magnitude between the first term and the second term will deterimine whether ET increases or decreases with increasing VPD.
+\begin{equation}
+  \label{sign}
+  \frac{c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right).
+\end{equation}
+All variables are positive, so the relative magnitude between the first term and the second term in Equation \ref{sign} will deterimine whether ET increases or decreases with increasing VPD.
 If the second term is larger magnitude then plant response dominates, while if the first term is larger then atmospheric demand dominats.
@@
 
-Term 2 minus Term 3's role in determining the sign of the sensitivty of ET to VPD makes it crucuial for answering our question ``When does VPD drive or reduce ET?'' Exploring these terms more, $c_s$ and $\gamma$ variability is relatively less than $\sigma$ and $VPD$ variability, so variability within PFT will be solely determined by $\sigma$ and $VPD$.
-If we fix uncertainty $\sigma$ at 1, then Term 2 minus Term 3 is just a function of $VPD$.
+The sign term is crucuial for answering our question ``When does VPD drive or reduce ET?,'' so let's explore the terms within it more.
+The variability of $c_s$ and $\gamma$ is relatively less than $\sigma$ and $VPD$ variability, so we sign term variability within PFT will approximately determined by just $\sigma$ and $VPD$.
+If we fix uncertainty $\sigma$ at 1, then the sign term is approximately just a function of $VPD$.
 We can further dertmine a $VPD_{crit}$ where $\frac{\partial \; ET}{\partial \; VPD} = 0$:
@@
 Equation \ref{d_et_approx} aids interpretation of Figure \ref{term3}.
-Larger $uWUE$, and $g1$ shift the leading-order constant term ($\frac{1}{g_1}$) towards smaller values, and the sign term towards overall positive values.
-$uWUE$ and $g1$ are both water-use effiency type constants.
-Higher values corresponde to plants that are more willing to spend water on primary production and less evolved to conserve water.\explain{really need to think about this sentence more, tehre is more nuance and I actually thing g1 and uwue really partially cancel.} Figure \ref{term3} confirms our physical intuition: CROs are the least water conservative so have the smallest constant portion of Term 3, while CSH are the most water conservative and have the largest constant portion of Term 3.
+Larger $uWUE$ and $g1$ shift the leading-order constant term ($\frac{1}{g_1}$) towards smaller values, and the sign term towards overall positive values.
+% Higher values corresponde to plants that are more willing to spend water on primary production and less evolved to conserve water.\explain{really need to think about this sentence more, tehre is more nuance and I actually thing g1 and uwue really partially cancel.}
+Figure \ref{term3} confirms our physical intuition: CROs are the least water conservative so have the smallest constant portion of Term 3, while CSH are the most water conservative and have the largest constant portion of Term 3.
 For the VPD-dependent terms in Equation \ref{d_et_approx}, differences in $g1$ between PFTs exert a greater influence than difference $uWUE$, as the power of $g1$ increases.
 Increasing $uWUE$ and $g1$ decreases the VPD-dependence, but $g1$ has the bigger effect due to its increasing powers.
-ENF ($g1 = 74.31$) has by far the largest VPD dependence of response, while CRO ($g1 = 183.1$) has the smallest VPD dependence.\explain{could probably simplify and just say that g1 controls VPD dependence.}
+ENF ($g1 = 74.31$) has by far the largest VPD dependence of response, while CRO ($g1 = 183.1$) has the smallest VPD dependence.
 
@@
 \includegraphics[width=20pc]{./fig05.pdf}
-\caption{Sources of variability for sign term.
-Top: sign term as a function of VPD, with $\sigma$ held fixed at PFT averages.
+\caption{Sources of variability for the sign term.
+Top: sign term as a function of VPD, with $\sigma$ held fixed at 1.
 The observed range of VPD for each PFT is also shown below the x-axis.
@@
 
-Bottom: The location of the minima of ET, as a function of VPD and $\sigma$. Lines and stars denote the distribution of VPD and $\sigma$ next to each axis, following the same percentiles as above.}
+Bottom: The location of the minima of ET, as a function of VPD and $\sigma$.
+Lines and stars denote the distribution of VPD and $\sigma$ next to each axis, following the same percentiles as above.
+\textbf{***need to replot with correct sigma (normalzied to 1)***}}
 \label{term3}
@@
 Figure \ref{term3}b shows the location of the minima of ET, as a function of $\sigma$ and $VPD$.
-For any $\sigma$ or VPD less (more) than these curves, sign term will be negative (positive).
+For any $\sigma$ or VPD less (more) than these curves, the sign term will be negative (positive).
 It is clear that the portion of VPD observations below/above these curves will be a strong function of $\sigma$.
@@
 
-\subsection{Functional Form Scaling Term}
+\subsection{Functional Form of the Scaling Term}
 
 While the above discussion of the sign of $\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our research question, the magnitude of $\frac{\partial \; ET}{\partial \; VPD}$ will also impact statistics of $\frac{\partial \; ET}{\partial \; VPD}$ and the importance of $VPD$ variablility for overall $ET$ variability.
-So we will more closely examine the scaling Term 1: $\frac{P}{T} \propto \rho$, so this should vary little relative to aerodynamic conductance and $\Delta$.
-$\gamma$ should also be relatively constant, so Term 1 should be primarily a function of aerodynamic conductance and temperature (through the function $\Delta$).
+So we will more closely examine the scaling term: $\frac{P}{T} \propto \rho$, so this should vary little relative to aerodynamic conductance and $\Delta$.
+$\gamma$ should also be relatively constant, so the scaling term should be primarily a function of aerodynamic conductance and temperature (through the function $\Delta$).
 This makes sense, as aerodynamic conductance represents how efficently response at the surface is communicated to the atmosphere.
@@
 
-Figure \ref{scale_vary}A shows scaling term normalized by mean aerodynamic conductance (calculated for each plant functional type), and confirms that much of the relative variability of the scaling term is contained in the aerodynamic conductance variability.
-Generally, $T$ has less of a role.
-Additionally, the impact of $T$ on the relative variability increases with increasing aerodynamic conductance.
+Figure \ref{scale_vary}A shows scaling term normalized by mean aerodynamic conductance (calculated for each plant functional type), and confirms that much of the relative variability of the scaling term is caused by aerodynamic conductance variability.
+Generally, temperature (via $\Delta$) causes less relative variability.
+However, the impact of $T$ on the relative variability increases with increasing aerodynamic conductance.
 
@@
 Figure \ref{scale_vary}B shows scaling term evaluated with the mean aerodynamic conductance for each PFT, and at the range of observed temperatures for each PFT.
-As expected, for the tree PFTs (DBF, ENF) scaling term is much larger and the temperature dependence is much stronger.
+As expected, for the tree PFTs (DBF, ENF) the scaling term is much larger and the temperature dependence is much stronger.
 Systematic differences in observed temperatures also cause differences in the average magnitude of scaling term.
 For example, ENF experiences on average colder temperatures and is thus more likely to have a larger scaling term.
-Additionally, because the variability of aerodynamic conductance increases proportionally to the mean, the spread of scaling term due to aerodynamic conductance variability will be larger for the tree PFTs, although this is not shown for simplicity.
-To summarize, the variability of scaling term within each PFT will look like Figure \ref{scale_vary}A for each PFT, but the scale of the y-axis will increase or decrease according to mean aerodynamic conductance oberved in Figure \ref{scale_vary}B.
+Additionally, because the variability of aerodynamic conductance increases proportionally to the mean, the spread of the scaling term due to aerodynamic conductance variability will be larger for the tree PFTs, although this is not shown for simplicity.
+To summarize Figure \ref{scale_vary}, the variability of the scaling term within each PFT will look like Figure \ref{scale_vary}A for each PFT, but the scale of the y-axis will increase or decrease according to mean aerodynamic conductance oberved in Figure \ref{scale_vary}B.
 
@@
 $VPD$'s contribution to ET's variability ranges between 30 - 40 W m$^{-2}$ for all PFTs except for CSH, which is about 100 W m$^{-2}$.
-Another meaningful comparison is to $\frac{\partial \; ET}{\partial \; R} \cdot std(R)$, as net radiation is generally the driver of ET (cite joe berry here).
+Another meaningful comparison is to $\frac{\partial \; ET}{\partial \; R} \cdot std(R)$, as net radiation is generally the driver of ET (cite joe berry here)\explain{Find citation?}.
 For all PFTs except for CSH $VPD$ contributes between 30 - 40 \% of $R$'s contribution to variability.
 For CSH the portion is much larger, about 88 \%.
-$VPD$'s variability is certiantly a primary contributor to $ET$'s variability.
+$VPD$'s variability is certiantly a non-negligable contributor to $ET$'s variability.
 
 So far, idealized plots and statistics have illuminated the form of $\frac{\partial \; ET}{\partial \; VPD}$ and how it varies with PFT.
-Large mean uWUE shifts CRO and DBF towards positive $\frac{\partial \; ET}{\partial \; VPD}$.
-However, the strongly nonlinear function of $\frac{\partial \; ET}{\partial \; VPD}$ at $\frac{\partial \; ET}{\partial \; VPD} < 0$ pushes $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ negative for DBF (it does not do this for CRO because of CRO's high $g1$).
-ENF's low $g1$ value increases the dependence of $\frac{\partial \; ET}{\partial \; VPD}$ on $VPD$, and makes the function more strongly nonlinear.
+Large mean uWUE shifts CRO and DBF towards positive $\frac{\partial \; ET}{\partial \; VPD}$, but $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ is still negative for DBF due to its lower $g1$.
+ENF's low $g1$, even lower than DBF, increases the dependence of $\frac{\partial \; ET}{\partial \; VPD}$ on $VPD$, and makes the function very nonlinear.
 This has the side effect of pushing $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ negative further than other PFTs for a given fraction $\frac{\partial \; ET}{\partial \; VPD} < 0$ and magnitude $\frac{\partial \; ET}{\partial \; VPD}(\overline{T,\ldots,VPD})$.
-GRA shows the opposite behavior; a relatively high $g1$  makes the function more linear, decreasing the magnitude of $-\overline{\frac{\partial \; ET}{\partial \; VPD}}$ for a given  [large] fraction $\frac{\partial \; ET}{\partial \; VPD} < 0$ and negative $\frac{\partial \; ET}{\partial \; VPD}(\overline{T,\ldots,VPD})$, However, aerodynamic conductance and scaling term also probably have a role in this.
+GRA shows the opposite behavior; a relatively high $g1$  makes the function more linear, decreasing the magnitude of $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ for a given fraction $\frac{\partial \; ET}{\partial \; VPD} < 0$ and negative $\frac{\partial \; ET}{\partial \; VPD}(\overline{T,\ldots,VPD})$.
 Finally, the low $uWUE$ of CSH pushes it toward by far the lowest values $\frac{\partial \; ET}{\partial \; VPD}$ (Figure \ref{term3}).
 Variability in $VPD$ accounts for the largest about of $ET$ variability for CSH.
-For the other PFTs, $VPD$ contributes less to $ET$ variability, but still represents about 15-20\explain{depending on how we take the derivative, this could be 30-40\%.} \% of $R$'s contributions to ET variability.
+For the other PFTs, $VPD$ contributes less to $ET$ variability, but still represents about 30-40 \% of $R$'s contributions to ET variability.
 
@@
 \hline
-\multicolumn{2}{l}{$^{a}$Footnote text here.}
 
@@
 However, the theory closely captures the intercept where $\frac{\partial \; ET}{\partial \; VPD} = 0$.
-Really resolving this feature is the key to answering our research question, so perhaps the theory is still useful for ENF for the purpose of addressing when VPD drives or reduces ET.
+Resolving this feature is the key to answering our question, so perhaps the theory is still useful for ENF for the purpose of addressing when VPD drives or reduces ET.
 It is also noteworthy that we used published constants for $g1$, which determines the VPD dependence, from \cite{Lin_2015}.
-We could have fit our own constants to tweak the theory to match the observations, but this was not our goal.
+We could have fit our own constants to tweak the theory to match the observations, but our goal was to use a priori physical constants whenever possible..
 
@@
 
+A final comparison for Figure \ref{agu_fig} is to the value of the sign term for PET.
+For PET, the sign term is a constant $\frac{c_p}{R_{air}}\approx 3.5$.
+This is far more than any of the observed values in Figure \ref{agu_fig}, even when uncertainty is included.
+So, even when atmospheric demand dominates plant response, the plant response still has a large role in attenuating atmospheric demand.
+In order to understand and capture the behavior of land-surface reponse, representation of the plants must be included in some form.
+This comparison casts doubt on the usefulness of PET for interprenting land-atmosphere evolution, as is done often in hydrometeorlogical studies.
+
 Figure \ref{real} presents calculated $\frac{\partial \; ET}{\partial \; VPD}$ where, unless otherwise noted, all variables in Equation \ref{d_et} are allowed to vary, including uncertainty.
@@
 Vertical black lines denote where our theory says $VPD_{crit}$ should be.
+For any obervations to the left of this line $\frac{]aprial \; ET}{\partial \; VPD}$ should be negative, and any ovservations the righ right of the line should be postiive.
 In addition to seeing how the theory $VPD_{crit}$ matches observations, these plots are also useful for observing how the scaling term affects the magnitude of the derivative, if at all.
@@
 By comparing different columns of Figure \ref{real} we can see the impact of $\sigma$ and $g_a$ on the variability of $\frac{\partial \; ET}{\partial \; VPD}$.
-$g_a$'s scaling (included in columns 1 and 3) alters the magnitude considerably, particularly begtween observations.
+$g_a$'s scaling (included in columns 1 and 3) alters the magnitude considerably, particularly between PFTs.
 $\sigma$ variability (included in columns 1 and 2) adds a lot of additional noise to the signal of $\frac{\partial \; ET}{\partial \; VPD}$, as observed in Figure \ref{agu_fig}.
@@
 
-While $g_a$ has a large impact on magnitudes in the scaling, temperature, through the $\Delta$ term, does not appear to have as much of an impact in the full observations.
+While $g_a$ has a large impact on magnitudes in the scaling, temperature, through the $\Delta$ term, does not appear to have as much of an impact in the complete estimation of $\frac{\partial \; ET}{\partial \; VPD}$.
 Therefore it appears that the absolute magnitude of the derivative can be mostly attributed to the sign term, and its direct scaling by aerodynamic conductance.
@@
 Our $\sigma$-based method of uncertainty propagation blurs the idealized expectations the most for GRA and CRO.
-We therefor have the most confidence in our conclusion based on Equation \ref{d_et} for PFTS CSH, DBF, and ENF, as the sign of the derivative with uncertainty included matches (to leading order) the story when $\sigma$ is held fixed.
+We therefor have the most confidence in our conclusion based on Equation \ref{d_et} for PFTs CSH, DBF, and ENF, as the sign of the derivative with uncertainty included matches (to leading order) the story when $\sigma$ is held fixed.
 
@@
 \caption{Scatter plots of $\frac{\partial \; ET}{\partial \; VPD}$.
-Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; VPD}$, as labeled: Column 1 - $\frac{\partial \; ET}{\partial \; VPD}$; Column 2 - $\frac{\partial \; ET}{\partial \; VPD}$ normalized by $g_a$; Column 3 - $\frac{\partial \; ET}{\partial \; VPD}$ with $\sigma$ held fixed at PFT average; and Column 4 - $\frac{\partial \; ET}{\partial \; VPD}$ normalized by $g_a$ and with $\sigma$ held fixed.
+Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; VPD}$, as labeled: Column 1: $\frac{\partial \; ET}{\partial \; VPD}$; Column 2: $\frac{\partial \; ET}{\partial \; VPD}$ normalized by $g_a$; Column 3: $\frac{\partial \; ET}{\partial \; VPD}$ with $\sigma$ held fixed at 1; and Column 4: $\frac{\partial \; ET}{\partial \; VPD}$ normalized by $g_a$ and with $\sigma$ held fixed.
 For reference, lines corresponding to RH = 20\% and RH = 90 \% are drawn.
@@
 
-\section{Conclusions} \explain{also need to flesh this section out}
+\section{Conclusions}
 
@@
 
-Math coded inside display math mode \[ ...\]
- will not be numbered, e.g.,:
- \[ x^2=y^2 + z^2\]
-
- Math coded inside \begin{equation} and \end{equation} will
- be automatically numbered, e.g.,:
- \begin{equation}
- x^2=y^2 + z^2
- \end{equation}
-
-% To create multiline equations, use the
-% \begin{eqnarray} and \end{eqnarray} environment
-% as demonstrated below.
-\begin{eqnarray}
-  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
-        && + (y - y_{0}) \sin \Theta  \nonumber \\
-  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
-        && + (y - y_{0}) \cos \Theta.
-\end{eqnarray}
-
 %If you don't want an equation number, use the star form:

### 2018-01-05 src/analysis.py

@@ -35,6 +35,6 @@ _df = dfs['full']
 import seaborn as sns
 import matplotlib.pyplot as plt
 plt.figure()
-sns.distplot(_df.uwue)
-#sns.jointplot(_df.et_obs, _df.et, kind='hex')
+#sns.distplot(_df.uwue)
+sns.jointplot(_df.et_obs, _df.et, kind='hex')
 plt.savefig('%s/temp/garb1.png' % os.environ['PLOTS'])

### 2018-01-06 doc/paper/vpd_et_paper.tex
added agu fig, close to sending out to Pierre

@@
 % 11/23/2015
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
-%
-% This file includes commands and instructions
-% given in the order necessary to produce a final output that will
-% satisfy AGU requirements. 
-%
-% You may copy this file and give it your
-% article name, and enter your text.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% PLEASE DO NOT USE YOUR OWN MACROS
-% DSO NOT USE \newcommand, \renewcommand, or \def, etc.
-%
-% FOR FIGURES, DO NOT USE \psfrag or \subfigure.
-% DO NOT USE \psfrag or \subfigure commands.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Step 1: Set the \documentclass
-%
-% There are two options for article format:
-%
-% 1) PLEASE USE THE DRAFT OPTION TO SUBMIT YOUR PAPERS.
-% The draft option produces double spaced output.
-% 
-% 2) numberline will give you line numbers.
-
-%% To submit your paper:
 \documentclass[draft,linenumbers]{agujournal}
@@
 \drafttrue
-
-%% For final version.
-% \documentclass{agujournal}
-
-% Now, type in the journal name: \journalname{<Journal Name>}
-
-% ie, \journalname{Journal of Geophysical Research}
-%% Choose from this list of Journals:
-%
-% JGR-Atmospheres
-% JGR-Biogeosciences
-% JGR-Earth Surface
-% JGR-Oceans
-% JGR-Planets
-% JGR-Solid Earth
-% JGR-Space Physics
-% Global Biochemical Cycles
-% Geophysical Research Letters
-% Paleoceanography
-% Radio Science
-% Reviews of Geophysics
-% Tectonics
-% Space Weather
-% Water Resource Research
-% Geochemistry, Geophysics, Geosystems
-% Journal of Advances in Modeling Earth Systems (JAMES)
-% Earth's Future
-% Earth and Space Science
-%
-%
 
@@
 \sigma = - \frac{g_a \gamma c_s \sqrt{VPD} L_v P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p VPD\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})}
-    \label{et_sigma}
+    \label{sigma}
   \end{equation}
@@
 
-By construction, the variability in the $\sigma$ term (Equation \ref{lai}) contains all model and observational uncertainties.
+By construction, the variability in the $\sigma$ term (Equation \ref{sigma}) contains all model and observational uncertainties.
 For an observation that perfectly matches our model and constant uWUE assumption $\sigma$ will be one.
@@
 \includegraphics[width=20pc]{./fig02.png}
-\caption{Histogram of $\sigma$ values calculated for each site and time according to Equation \ref{lai}.
+\caption{Histogram of $\sigma$ values calculated for each site and time according to Equation \ref{sigma}.
 The lowest and highest 5\% are removed as outliers, as well as any values below 0.
@@
 \centering
+\includegraphics[width=\textwidth]{./agu_fig.pdf}
+\caption{Comparison of the sign term with moden uncertainty included (box plots) to the sign term as calculated with simplifying assumptions (blue line).}
+\label{agu_fig}
+\end{figure}
+
+\begin{figure}[h]
+\centering
 \includegraphics[width=\textwidth]{./fig06.png}
@@
 Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; VPD}$, as labeled: Column 1: $\frac{\partial \; ET}{\partial \; VPD}$; Column 2: $\frac{\partial \; ET}{\partial \; VPD}$ normalized by $g_a$; Column 3: $\frac{\partial \; ET}{\partial \; VPD}$ with $\sigma$ held fixed at 1; and Column 4: $\frac{\partial \; ET}{\partial \; VPD}$ normalized by $g_a$ and with $\sigma$ held fixed.
-For reference, lines corresponding to RH = 20\% and RH = 90 \% are drawn.
-Please note differences in the colorbar scale.
-***see alternate (or additional) plot below.***}
+Please note differences in the colorbar scale.}
 \label{real}
@@
 % AGU does not want a .bib or a .bbl file. Please copy in the contents of your .bbl file here.
-% \bibliography{references.bib}
+%\bibliography{references.bib}
 \input{vpd_et_paper.bbl}
 
-%% After you run BibTeX, Copy in the contents of the .bbl file here:
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Track Changes:
-% To add words, \added{<word added>}
-% To delete words, \deleted{<word deleted>}
-% To replace words, \replace{<word to be replaced>}{<replacement word>}
-% To explain why change was made: \explain{<explanation>} This will put
-% a comment into the right margin.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% At the end of the document, use \listofchanges, which will list the
-% changes and the page and line number where the change was made.
-
-% When final version, \listofchanges will not produce anything,
-% \added{<word or words>} word will be printed, \deleted{<word or words} will take away the word,
-% \replaced{<delete this word>}{<replace with this word>} will print only the replacement word.
-%  In the final version, \explain will not print anything.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%
 \listofchanges
-%%%
 
 \end{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Supporting Information
-%% (Optional) See AGUSuppInfoSamp.tex/pdf for requirements 
-%% for Supporting Information.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-More Information and Advice:
-
-%% ------------------------------------------------------------------------ %%
-%
-%  SECTION HEADS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Capitalize the first letter of each word (except for
-% prepositions, conjunctions, and articles that are
-% three or fewer letters).
-
-% AGU follows standard outline style; therefore, there cannot be a section 1 without
-% a section 2, or a section 2.3.1 without a section 2.3.2.
-% Please make sure your section numbers are balanced.
-% ---------------
-% Level 1 head
-%
-% Use the \section{} command to identify level 1 heads;
-% type the appropriate head wording between the curly
-% brackets, as shown below.
-%
-%An example:
-%\section{Level 1 Head: Introduction}
-%
-% ---------------
-% Level 2 head
-%
-% Use the \subsection{} command to identify level 2 heads.
-%An example:
-%\subsection{Level 2 Head}
-%
-% ---------------
-% Level 3 head
-%
-% Use the \subsubsection{} command to identify level 3 heads
-%An example:
-%\subsubsection{Level 3 Head}
-%
-%---------------
-% Level 4 head
-%
-% Use the \subsubsubsection{} command to identify level 3 heads
-% An example:
-%\subsubsubsection{Level 4 Head} An example.
-%
-%% ------------------------------------------------------------------------ %%
-%
-%  IN-TEXT LISTS
-%
-%% ------------------------------------------------------------------------ %%
-%
-% Do not use bulleted lists; enumerated lists are okay.
-% \begin{enumerate}
-% \item
-% \item
-% \item
-% \end{enumerate}
-%
-%% ------------------------------------------------------------------------ %%
-%
-%  EQUATIONS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Single-line equations are centered.
-% Equation arrays will appear left-aligned.
-
-%If you don't want an equation number, use the star form:
-%\begin{eqnarray*}...\end{eqnarray*}
-
-% Break each line at a sign of operation
-% (+, -, etc.) if possible, with the sign of operation
-% on the new line.
-
-% Indent second and subsequent lines to align with
-% the first character following the equal sign on the
-% first line.
-
-% Use an \hspace{} command to insert horizontal space
-% into your equation if necessary. Place an appropriate
-% unit of measure between the curly braces, e.g.
-% \hspace{1in}; you may have to experiment to achieve
-% the correct amount of space.
-
-%% ------------------------------------------------------------------------ %%
-%
-%  EQUATION NUMBERING: COUNTER
-%
-%% ------------------------------------------------------------------------ %%
-
-% You may change equation numbering by resetting
-% the equation counter or by explicitly numbering
-% an equation.
-
-% To explicitly number an equation, type \eqnum{}
-% (with the desired number between the brackets)
-% after the \begin{equation} or \begin{eqnarray}
-% command.  The \eqnum{} command will affect only
-% the equation it appears with; LaTeX will number
-% any equations appearing later in the manuscript
-% according to the equation counter.
-%
-
-% If you have a multiline equation that needs only
-% one equation number, use a \nonumber command in
-% front of the double backslashes (\\) as shown in
-% the multiline equation above.
-
-% If you are using line numbers, remember to surround
-% equations with \begin{linenomath*}...\end{linenomath*}
-
-%  To add line numbers to lines in equations:
-%  \begin{linenomath*}
-%  \begin{equation}
-%  \end{equation}
-%  \end{linenomath*}

### 2018-01-06 src/paper_figure_code/agu_fig.py

@@ -27,8 +27,8 @@ name_dict = {'CRO': 'Crops',\
 
 paren_string = r'$\left(\frac{ c_p}{R_{air}} '\
                r'- \frac{\gamma c_s }{1.6 \; R\; \sigma \; uWUE  }'\
-               r'\left( \frac{2 g_1 + \sqrt{D}}'\
-               r'{2 (g_1 + \sqrt{D})^2}\right)\right)$'
+               r'\left( \frac{2 g_1 + \sqrt{VPD}}'\
+               r'{2 (g_1 + \sqrt{VPD})^2}\right)\right)$'
 
 def plot_box(_df):
   """makes a box plot"""
@@ -68,18 +68,20 @@ def plot_box(_df):
   #         'b-', linewidth=1.0)
   #_df.groupby(pd.cut(_df.vpd, bins=20)).apply(add_box, ax=ax)
   #_df.boxplot(column='sign', by=pd.cut(_df.vpd, bins=20), ax=ax)
-  if (_df.pft.iloc[0] == 'GRA') | (_df.pft.iloc[0] == 'DBF'):
+  if (_df.pft.iloc[0] == 'ENF') | (_df.pft.iloc[0] == 'DBF')\
+     | (_df.pft.iloc[0] == 'CSH'):
     ax.set_ylabel(paren_string, fontsize=fontsize)
-  ax.set_xlabel('VPD (kPa)', fontsize=fontsize)
+  if (_df.pft.iloc[0] == 'CSH') | (_df.pft.iloc[0] == 'GRA'):
+    ax.set_xlabel('VPD (kPa)', fontsize=fontsize)
   ax.set_title(name_dict[_df.pft.iloc[0]], fontsize=fontsize+3)
   ax.plot(orig_xlim, [0., 0.], 'k--', linewidth=dashedlinewidth)
 
   ax.set_ylim(ylim)
 
   ax.set_xlim(orig_xlim)
-  if _df.pft.iloc[0] == 'ENF':
-    ax.text(1.95, 1.2, '*', horizontalalignment='center',\
-            verticalalignment='center', fontdict={'fontsize' : 40})
+  # if _df.pft.iloc[0] == 'ENF':
+  #   ax.text(1.95, 1.2, '*', horizontalalignment='center',\
+  #           verticalalignment='center', fontdict={'fontsize' : 40})
   plt.tight_layout()
[…]

### 2018-01-08 doc/paper/vpd_et_paper.tex
finished spell check

@@
 
-Answering the question ``When does vapor pressure deficit drive or reduce evapotranspiration?'' is motivated by two potenential but opposing perspectives on the matter.
-The first persepctive is that increases to vapor pressure deficit (VPD) increases atmospheric demand for water from the land surface, and this drives an increase in evapotranspiration (ET).
-This perspective is particularly relevant because potential evapotranspiration (PET), which is used in many drought indicies and hydrometeorological studies, only quantifies atmospheric demand effects.
+Answering the question ``When does vapor pressure deficit drive or reduce evapotranspiration?'' is motivated by two potential but opposing perspectives on the matter.
+The first perspective is that increases to vapor pressure deficit (VPD) increases atmospheric demand for water from the land surface, and this drives an increase in evapotranspiration (ET).
+This perspective is particularly relevant because potential evapotranspiration (PET), which is used in many drought indices and hydrometeorological studies, only quantifies atmospheric demand effects.
 However, there is another possible perspective for vegetated surfaces.
 Plants' stomata have evolved to optimally regulate the exchange of water and carbon, and tend to close in response to increased atmospheric dryness \citep{Ball_1987, Leuning_1990, MEDLYN_2011}.
-Therefore, an increase in VPD may actually correspond to a decrease in ET because stomatal closure can cancel the effects of shifts to atmospheric de.
+Therefore, an increase in VPD may actually correspond to a decrease in ET because stomatal closure can cancel the effects of shifts to atmospheric demand.
 So really a question of ``When does VPD drive or reduce ET?'' becomes a question of ``When does plant response or atmospheric demand dominate?.''
@@
 If plant response reduces ET in response to atmospheric drying then soil moisture will be conserved for later ET.
-An increase in ET in reponse to atmospheric drying will reduce soil moisture, but contribute increased moisenting to the atmosperhere.
+An increase in ET in response to atmospheric drying will reduce soil moisture, but contribute increased moistening to the atmosphere.
 Understanding the one way response of the land surface to the atmosphere is a critical first step to disentangling land-atmosphere feedbacks.
@@
 Each site's plant functional type (PFT) was classified using the International Geosphere-Biosphere Programme vegetation classification scheme (\cite{Loveland_1999}).
-Because physical constats used in Section \ref{methods} are only published for five plant functional types (PFTs - see Table \ref{pft}),  only 56 of the 77 sites were used.
+Because physical constants used in Section \ref{methods} are only published for five plant functional types (PFTs - see Table \ref{pft}),  only 56 of the 77 sites were used.
 Figure \ref{map_fig} \explain{map needs to be improved - it's a placeholder for now} presents each site and its plant functional type.
@@
 The purpose of this study is to examine ecosystem response to atmospheric drying during the growing season.
-To accomplish this, we filter and quality control the data using a similiar procedure as \cite{Zhou_2015}:
+To accomplish this, we filter and quality control the data using a similar procedure as \cite{Zhou_2015}:
 \begin{itemize}
-\item Only measured or highest (``good'') quality gapfilled data, according to qualtiy control flags, are used.
+\item Only measured or highest (``good'') quality gapfilled data, according to quality control flags, are used.
 \item To isolate the growing season, we only use days in which the average GPP exceeds 10\% of the observed 95th percentile of GPP for a given site.
-\item We remove days with rain and the day following to avoid issues with rain interveption and sensor satruation and high relative humidity (Medlyn et. al. 2017).
+\item We remove days with rain and the day following to avoid issues with rain interception and sensor saturation and high relative humidity (\cite{MEDLYN2016).
 \end{itemize}
@@
   \label{pft}
-\caption{Plant functional types, their abbreviation, Medlyn coefficient \citep[from ][]{Lin_2015}, and uWfUE \citep[from ][]{Zhou_2015}.
+\caption{Plant functional types, their abbreviation, Medlyn coefficient \citep[from ][]{Lin_2015}, and uWUE \citep[from ][]{Zhou_2015}.
 Note that units are converted such that the quantities fit into Equations 1-8 with the variables in Table 1.}
@@
 \includegraphics[width=20pc]{./fig01.pdf}
-\caption{Plant functional type and location of sites used in analysis. ***\textbf{This is just  aplaceholder for now, I am currently fixing it***}}
+\caption{Plant functional type and location of sites used in analysis. ***\textbf{This is just  a placeholder for now, I am currently fixing it***}}
 \label{map_fig}
@@
 Lines and stars denote the distribution of VPD and $\sigma$ next to each axis, following the same percentiles as above.
-\textbf{***need to replot with correct sigma (normalzied to 1)***}}
+\textbf{***need to replot with correct sigma (normalized to 1)***}}
 \label{term3}
@@
 Vertical black lines denote where our theory says $VPD_{crit}$ should be.
-For any obervations to the left of this line $\frac{]aprial \; ET}{\partial \; VPD}$ should be negative, and any ovservations the righ right of the line should be postiive.
+For any obervations to the left of this line $\frac{]\partial \; ET}{\partial \; VPD}$ should be negative, and any ovservations the righ right of the line should be postiive.
 In addition to seeing how the theory $VPD_{crit}$ matches observations, these plots are also useful for observing how the scaling term affects the magnitude of the derivative, if at all.
@@
 \includegraphics[width=\textwidth]{./agu_fig.pdf}
-\caption{Comparison of the sign term with moden uncertainty included (box plots) to the sign term as calculated with simplifying assumptions (blue line).}
+\caption{Comparison of the sign term with model uncertainty included (box plots) to the sign term as calculated with simplifying assumptions (blue line).}
 \label{agu_fig}

### 2018-01-09 doc/paper/vpd_et_paper.tex
small typo fixed, send off to Pierre and Changjie for ocmments

@@
 % \draftfalse
+\usepackage{makecell}
 \drafttrue
@@
 
-Quantifying the plant response to a perturbation to atmospheric VPD increases our understanding of land surface response to shifts in atmospheric conditions.
+Quantifying the plant response to a perturbation to atmospheric VPD increases our understanding of the land surface response to shifts in atmospheric conditions.
 If plant response reduces ET in response to atmospheric drying then soil moisture will be conserved for later ET.
@@
 crops) will be more likely to increase ET with increasing VPD.
-Additionally, atmospheric conditions must matter as well.
+Atmospheric conditions must matter as well.
 At the ecosystem scale there are limits to the strategies plants use to hold water back from the atmosphere.
@@
 \item To isolate the growing season, we only use days in which the average GPP exceeds 10\% of the observed 95th percentile of GPP for a given site.
-\item We remove days with rain and the day following to avoid issues with rain interception and sensor saturation and high relative humidity (\cite{MEDLYN2016).
+\item We remove days with rain and the day following to avoid issues with rain interception and sensor saturation and high relative humidity (\cite{MEDLYN_2011}).
 \end{itemize}
-Additionally, we restrict data to the daytime which is identified when downwelling shortwave radation is greater than 50 W m$^{-2}$ and sensible heat flux is greater than 5 W m$^{-2}$.
+Additionally, we restrict data to the daytime which is identified when downwelling shortwave radiation is greater than 50 W m$^{-2}$ and sensible heat flux is greater than 5 W m$^{-2}$.
 To further reduce the chance of sensor saturation at high relative humidity, we remove all time steps for which VPD is less than .01 kPa.
 Timesteps with negative observed GPP or ET are also removed.
-After these quality control proceedures, 332556 upscaled hourly observations remain.
+After these quality control procedures, 332556 upscaled hourly observations remain.
 
@@
  where variable definitions are given in Table 1.
-\citet{MEDLYN_2011} developed a model for $g_s$ by combining optimal photosynthesis theory with empirical approaches.
+\citet{MEDLYN_2011} developed a model for stomatal conductance ($g_s$) by combining optimal photosynthesis theory with empirical approaches.
 The result for leaf-scale stomatal resistance was:
@@
 
-Note that $uWUE$ is fit on the ecosystem scale in \citet{Zhou_2015} so GPP in \ref{uwue} is really $A\cdot \text{LAI}$.
+Note that $uWUE$ is fit on the ecosystem scale in \citet{Zhou_2015} so GPP in Equation \ref{uwue} is really $A\cdot \text{LAI}$.
 This leads to the cancelation of LAI in addition to uWUE in Equation \ref{medlyn}.
@@
 Given FLUXNET data described in Section \ref{data}, every term in Equation \ref{et} is known.
-However, our sampling of sites and our focus on the growing season may intoduce some deviations of $uWUE$ from those observed in \citet{Zhou_2015}.
+However, our sampling of sites at the global scale may introduce some deviations of $uWUE$ from those observed in \citet{Zhou_2015}.
 Also, we wish to include some measure of uncertainty in our analysis to guide if our many assumptions and simplifications are reasonable.
@@
 
-The variablility of $\sigma$ across sites and time will provides some measure of uncertainty in our model, assumptions, as well as the fluxnet observations themselves.
+The variability of $\sigma$ across sites and time will provides some measure of uncertainty in our model, assumptions, as well as the FLUXNET observations themselves.
 To correct for differences in sampling between \cite{Zhou_2015} and our data, we set uWUE such that $\overline{\sigma} = 1$.
-The variability of $\sigma$ then propogates through any uncertainty to our derivative of Equation \ref{et_sigma}:
+The variability of $\sigma$ then propagates through any uncertainty to our derivative of Equation \ref{et_sigma}:
 
@@
 
-With Equation \ref{d_et} we have provided an analytical franework for ecosystem reponse to atmospheric demand perturbations.
-There are a few subtelties to taking the derivative in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$) and $VPD$ are functionally related, so while taking the derivative we evaluate $\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET} {\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD} \Big|_{\text{RH fixed}} + \frac{\partial \; ET}{\partial \; RH} \frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$ fixed}}$\explain{Should I take the derivative this way?
-I think it's more mathematically correct, but it makes the explanation more complicated and only changes the solution by a factor of two.}.
+With Equation \ref{d_et} we have provided an analytical framework for ecosystem response to atmospheric demand perturbations.\explain{Should I take the derivative this way?
+I think it's more mathematically correct, but it makes the explanation more complicated and only changes the solution by a factor of two.} There are a few subtleties to taking the derivative in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$) and $VPD$ are functionally related, so while taking the derivative we evaluate $\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET} {\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD} \Big|_{\text{RH fixed}} + \frac{\partial \; ET}{\partial \; RH} \frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$ fixed}}$.
 $RH$ and $e_s$ are assumed to be approximately orthogonal.
@@
 
-One final comment on our derivation which will not be discussed further but is relevant for future analysis: if we  approximage $c_s$ at a global mean CO$_2$ concentration, then the RHS of Equation \ref{et} is fully defined using commonly available weather station data and the constants published in \citet{Zhou_2015, Lin_2015}.
-This then begs the question, why use PET for drought metrics in vegetated areas?
-It appears a much more physically realistic estimate of ET can be had with the same information reqiured to calculate PET.
+One final comment on our derivation which will not be discussed further but is relevant for future analysis: if we  approximate $c_s$ at a global mean CO$_2$ concentration, then the RHS of Equation \ref{et} is fully defined using commonly available weather station data and the constants published in \citet{Zhou_2015, Lin_2015}.
+This makes Equation \ref{et} a useful alternative to PET in drought indices and hydrometeorological analysis for vegetated surfaces.
+Equation \ref{et} better reflects the physics of water exchange at the land surface.
 
@@
 $R$  & net radiation at land surface minus ground heat flux & W m$^{-2}$   & - \\
-  $g_a$  & aerodynamic conductance & m s$^{-1}$  & \cite{Thom_1977}, \cite{Paulson_1970}, \cite{Beljaars_1991} \\
+  $g_a$  & aerodynamic conductance & m s$^{-1}$  & \makecell{\cite{Thom_1977} \\ \cite{Paulson_1970} \\ \cite{Beljaars_1991}} \\
   $\rho_a$  & air density & kg m$^{-3}$  & - \\
@@
   $VPD$  & vapor pressure deficit & Pa  & - \\
-  $\gamma$  & psychrometric constant & Pa K$^{-1}$   & - \\
+  $\gamma$  & psychometric constant & Pa K$^{-1}$   & - \\
   $g_s$  & stomatal conductance & m s$^{-1}$  & \cite{MEDLYN_2011} \\
@@
 Before calculating the sensitivity of ET to VPD, we will consider the functional form of Equation \ref{d_et}.
-There are two main terms: a ``scaling'' term which modifies the mangnitude but not the sign:
+There are two main terms: a ``scaling'' term which modifies the magnitude but not the sign:
 
@@
 
-and a ``sign'' term which determines whether the derivative is postive or negative:
+and a ``sign'' term which determines whether the derivative is positive or negative:
 
@@
 \end{equation}
-All variables are positive, so the relative magnitude between the first term and the second term in Equation \ref{sign} will deterimine whether ET increases or decreases with increasing VPD.
-If the second term is larger magnitude then plant response dominates, while if the first term is larger then atmospheric demand dominats.
+All variables are positive, so the relative magnitude between the first term and the second term in Equation \ref{sign} will determine whether ET increases or decreases with increasing VPD.
+If the second term is larger magnitude then plant response dominates, while if the first term is larger then atmospheric demand dominates.
 Given that the sign term is crucial to answering our question ``When does VPD drive or reduce ET?'', we will explore its functional form first.
@@
 \subsection{Functional Form of the Sign Term}
-
-The sign term is crucuial for answering our question ``When does VPD drive or reduce ET?,'' so let's explore the terms within it more.
-The variability of $c_s$ and $\gamma$ is relatively less than $\sigma$ and $VPD$ variability, so we sign term variability within PFT will approximately determined by just $\sigma$ and $VPD$.
+\label{sign_func}
+First, let's explore the variables within the sign term to gain better intuition for what determines when VPD drives vs.
+reduces ET.
+The variability of $c_s$ and $\gamma$ is relatively less than $\sigma$ and $VPD$ variability, so sign term variability within PFT will be approximately determined by just $\sigma$ and $VPD$.
 If we fix uncertainty $\sigma$ at 1, then the sign term is approximately just a function of $VPD$.
-We can further dertmine a $VPD_{crit}$ where $\frac{\partial \; ET}{\partial \; VPD} = 0$:
+We can further determine a $VPD_{crit}$ where $\frac{\partial \; ET}{\partial \; VPD} = 0$:
 
@@
 For reference, these values are also provided.
-For values of $VPD$ less than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be negative, and for values of $VPD$ greater than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be positive.}
+For values of $VPD$ less than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be negative, and for values of $VPD$ greater than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be positive.
+\textbf{**** this statistics are dated, need to update***}}
 \centering
@@
   \hline
-  PFT & $R_{air}$ & $c_s$ (ppm) & $\gamma$ & $\overline{\sigma}$ & $\overline{\sigma} \cdot uWUE$ & \textbf{$VPD_{crit}$ (Pa)} \\
+  PFT & $R_{air}$ & $c_s$ (ppm) & $\gamma$ &  uWUE    & \textbf{$VPD_{crit}$ (Pa)} \\
   \hline
-  CRO &  288.680920 & 372.567691& 65.351523& 0.684394&       2.602873&  \textbf{133.165438} \\
-  CSH &   289.067152& 381.593622& 67.613172& 0.997224&       2.175278& \textbf{4439.564212} \\
-  DBF &   288.624437& 377.449849& 63.421812& 0.881061&       2.746393&  \textbf{888.773243} \\
-  ENF &  288.183849& 377.676463& 61.559242& 1.217892&       4.015362&  \textbf{978.084845} \\
-  GRA &  288.425651& 377.264645& 61.598768& 0.850869&       2.281074& \textbf{1141.630778} \\
+  CRO &  288.680920 & 372.567691& 65.351523& 2.602873&  \textbf{133.165438} \\
+  CSH &   289.067152& 381.593622& 67.613172& 2.175278& \textbf{4439.564212} \\
+  DBF &   288.624437& 377.449849& 63.421812& 2.746393&  \textbf{888.773243} \\
+  ENF &  288.183849& 377.676463& 61.559242& 4.015362&  \textbf{978.084845} \\
+  GRA &  288.425651& 377.264645& 61.598768& 2.281074& \textbf{1141.630778} \\
 \hline
@@
 
-Figure \ref{term3} provides greater description for how the sign term varies with $VPD$, as a function of PFT.
+Figure \ref{term3} provides a greater description for how the sign term varies with $VPD$, as a function of PFT.
 Equation \ref{d_et_approx} aids interpretation of Figure \ref{term3}.
@@
 % Higher values corresponde to plants that are more willing to spend water on primary production and less evolved to conserve water.\explain{really need to think about this sentence more, tehre is more nuance and I actually thing g1 and uwue really partially cancel.}
-Figure \ref{term3} confirms our physical intuition: CROs are the least water conservative so have the smallest constant portion of Term 3, while CSH are the most water conservative and have the largest constant portion of Term 3.
-For the VPD-dependent terms in Equation \ref{d_et_approx}, differences in $g1$ between PFTs exert a greater influence than difference $uWUE$, as the power of $g1$ increases.
-Increasing $uWUE$ and $g1$ decreases the VPD-dependence, but $g1$ has the bigger effect due to its increasing powers.
+Figure \ref{term3} confirms our physical intuition: CROs are the least water conservative so have the smallest constant portion of Term 3 ($\frac{\gamma c_s }{1.6 \; R*\; \sigma \; \text{ uWUE }} \left(\frac{1}{g_1}\right)$), while CSH are the most water conservative and have the largest constant portion of Term 3.
+For the VPD-dependent terms in Equation \ref{d_et_approx}, differences in $g1$ between PFTs exert a greater influence than differences in $uWUE$.
+Increasing $uWUE$ and $g1$ decreases the VPD-dependence, but $g1$ has the bigger effect, particularly at high VPD when the series becomes disordered.
 ENF ($g1 = 74.31$) has by far the largest VPD dependence of response, while CRO ($g1 = 183.1$) has the smallest VPD dependence.
 
+A primary takeaway from Figure \ref{term3}a is that according to our theory for all PFTs except for crops there is frequent occurrence of a negative (plant dominating) ET response to increases in VPD.
+To better illustrate this, the ranges of observed environmental VPDs at the FLUXNET sites are plotted parallel to the x-axis.
+25th, 50th, and 75th percentiles are marked by stars, and the location of VPD$_{crit}$ is marked by a vertical line.
+For CSH (water conservative), VPD is always less than VPD$_{crit}$ (plant response dominates).
+For CRO, VPD is almost always more than VPD$_{crit}$.
+For forests and grass, whether plant response or atmospheric demand dominates will depend more on the environment, with about half of of observed VPD less than VPD$_{crit}$ (plant response dominates).
+It's also important to note that even when atmospheric demand dominates, the response is still far smaller than it would be for PET (atmospheric demand only).
+The sign term in this case would just be a constant ($\frac{c_p}{R_{air}} \approx 3.5$), which is far larger than any part of the curves for any PFT.
+This highlights the deficiencies of PET's ability to capture land response to changes in atmospheric dryness (VPD).
+However, up to this point we have assumed $\sigma$ (uncertainty) equal to one.
+We still need to test if inclusion of uncertainty changes the story.
 \begin{figure}[h]
@@
 Lines and stars denote the distribution of VPD and $\sigma$ next to each axis, following the same percentiles as above.
-\textbf{***need to replot with correct sigma (normalized to 1)***}}
+\textbf{***need to replot with correct sigma (normalized to 1), also correct uWUE label in first plot***}}
 \label{term3}
@@
 
-Figure \ref{term3}b shows the location of the minima of ET, as a function of $\sigma$ and $VPD$.
+Figure \ref{term3}b shows the location of VPD$_{crit}$, as a function of $\sigma$ (uncertainty) and $VPD$.
 For any $\sigma$ or VPD less (more) than these curves, the sign term will be negative (positive).
@@
 The fraction of positive $\frac{\partial \; ET}{\partial \; VPD}$ appears to be more even for ENF, GRA, and DBF, and we might expect a greater frequency of positive $\frac{\partial \; ET}{\partial \; VPD}$ for CRO.
+Section \ref{testing} further examines the consistency of our theory, given  the apparent sensitivity to uncertainty ($\sigma$).
 
@@
 
-While the above discussion of the sign of $\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our research question, the magnitude of $\frac{\partial \; ET}{\partial \; VPD}$ will also impact statistics of $\frac{\partial \; ET}{\partial \; VPD}$ and the importance of $VPD$ variablility for overall $ET$ variability.
+While the above discussion of the sign of $\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our research question, the magnitude of $\frac{\partial \; ET}{\partial \; VPD}$ will also impact statistics of $\frac{\partial \; ET}{\partial \; VPD}$ and the importance of $VPD$ variability for overall $ET$ variability.
 So we will more closely examine the scaling term: $\frac{P}{T} \propto \rho$, so this should vary little relative to aerodynamic conductance and $\Delta$.
 $\gamma$ should also be relatively constant, so the scaling term should be primarily a function of aerodynamic conductance and temperature (through the function $\Delta$).
-This makes sense, as aerodynamic conductance represents how efficently response at the surface is communicated to the atmosphere.
+This makes sense, as aerodynamic conductance represents how efficiently response at the surface is communicated to the atmosphere.
 As it increases, any plant response will be communicated more strongly to the atmosphere (and vice-versa).
 
-$\Delta$'s presence in the scaling term also matches physical intition.
+$\Delta$'s presence in the scaling term also matches physical intuition.
 Evaporative cooling will dampen the ability of the atmosphere to take more moisture, because $e_{s}$ decreases with decreasing temperature.
@@
 
-Figure \ref{scale_vary}A shows scaling term normalized by mean aerodynamic conductance (calculated for each plant functional type), and confirms that much of the relative variability of the scaling term is caused by aerodynamic conductance variability.
+Figure \ref{scale_vary}A shows the scaling term normalized by mean aerodynamic conductance (calculated for each plant functional type), and confirms that much of the relative variability of the scaling term is caused by aerodynamic conductance variability.
 Generally, temperature (via $\Delta$) causes less relative variability.
@@
 While the relative variability of scaling term is similar across PFT, the absolute value of scaling term varies strongly across PFT.
-Figure \ref{scale_vary}B shows scaling term evaluated with the mean aerodynamic conductance for each PFT, and at the range of observed temperatures for each PFT.
+Figure \ref{scale_vary}B shows scaling term evaluated at the mean aerodynamic conductance for each PFT, and at the range of observed temperatures for each PFT.
 As expected, for the tree PFTs (DBF, ENF) the scaling term is much larger and the temperature dependence is much stronger.
@@
 Additionally, because the variability of aerodynamic conductance increases proportionally to the mean, the spread of the scaling term due to aerodynamic conductance variability will be larger for the tree PFTs, although this is not shown for simplicity.
-To summarize Figure \ref{scale_vary}, the variability of the scaling term within each PFT will look like Figure \ref{scale_vary}A for each PFT, but the scale of the y-axis will increase or decrease according to mean aerodynamic conductance oberved in Figure \ref{scale_vary}B.
+To summarize Figure \ref{scale_vary}: the variability of the scaling term within each PFT will look like Figure \ref{scale_vary}A for each PFT, but the scale of the y-axis will increase or decrease according to mean aerodynamic conductance observed in Figure \ref{scale_vary}B.
 
@@
 The units of $\frac{\partial \; ET}{\partial \; VPD}$ make it difficult to interpret if $VPD$ is even a meaningful contributor to ET's variability.
-To better understand $VPD$'s contribution, we mulitply $\frac{\partial \; ET}{\partial \; VPD}$ with $VPD$'s standard deviation to define a (linearized) relative change in ET for variations in $VPD$ .
+To better understand $VPD$'s contribution, we multiply $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)$ with $VPD$'s standard deviation to define a (linearized) relative change in ET for variations in $VPD$ .
 $VPD$'s contribution to ET's variability ranges between 30 - 40 W m$^{-2}$ for all PFTs except for CSH, which is about 100 W m$^{-2}$.
@@
 For CSH the portion is much larger, about 88 \%.
-$VPD$'s variability is certiantly a non-negligable contributor to $ET$'s variability.
+$VPD$'s variability is certainly a non-negligible contributor to $ET$'s variability.
 
@@
 Finally, the low $uWUE$ of CSH pushes it toward by far the lowest values $\frac{\partial \; ET}{\partial \; VPD}$ (Figure \ref{term3}).
-Variability in $VPD$ accounts for the largest about of $ET$ variability for CSH.
+Variability in $VPD$ accounts for the largest amount of $ET$ variability for CSH.
 For the other PFTs, $VPD$ contributes less to $ET$ variability, but still represents about 30-40 \% of $R$'s contributions to ET variability.
@@
 \begin{table}
-\caption{Statistics of $\frac{\partial \; ET}{\partial \; VPD}$ as a function of PFT.}
+\caption{Statistics of $\frac{\partial \; ET}{\partial \; VPD}$ as a function of PFT. \textbf{***dated, need to update these statistics***}}
 \centering
@@
 \subsection{Testing the theory - does the data match the model?}
-
+\label{testing}
 So far we have developed a theory for $\frac{\partial \; ET}{\partial \; VPD}$'s behavior.
 Most important to answering ``When does VPD drive or reduce ET?'' is that the sign of our derivative term is correct.
-So, we need to test if when we include uncertainty the observed derivitives match what our theory would say for the sign term.
+So, we need to test if when we include uncertainty the observed derivatives match what our theory would say for the sign term.
 
@@
 Perhaps this should be expected.
-Our theory, for practical reasons, does not include important sources of variablility (particyularly between sites) for CRO and GRA.
+Our theory, for practical reasons, does not include important sources of variability (particularly between sites) for CRO and GRA.
 We do not account for seasonal changes in LAI, variability in crop height and surface roughness, differences in C3 vs C4 photosynthesis, or differences in irrigation practices between the sites.
-Given that the our thoery does not do as well for CRO and GRA, we will refrain from firm statements about how CRO and GRA respond to changes in VPD.
+Given that the our theory does not do as well for CRO and GRA, we will refrain from firm statements about how CRO and GRA respond to changes in VPD.
 
@@
 It is also noteworthy that we used published constants for $g1$, which determines the VPD dependence, from \cite{Lin_2015}.
-We could have fit our own constants to tweak the theory to match the observations, but our goal was to use a priori physical constants whenever possible..
+We could have fit our own constants to tweak the theory to match the observations, but our goal was to use a priori physical constants whenever possible.
 
+While the above discussion shows that our theory has some limitations when applied to real data, especially for crops and grasslands, another question could be, how does our theory's performance compare to often-used PET?
+Again, as in Section \ref{sign_func}, the PET response to changes in VPD will be constant at about 3.5.
+Even with uncertainty included, no values of the response approach the PET case.
+PET's status outside the error bars of our theory highlight the usefulness of our theory (and Equation \ref{et}) for estimating land response and moisture budgeting.
+
+\subsubsection{Testing VPD$_{crit}$}
 As stated in the above discussion of ENF, we are mostly interested in the critical threshold $VPD_{crit}$ for which the sign of  $\frac{\partial \; ET}{\partial \; VPD}$ is determined.
-When $VPD < VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}  < 0$, and when $VPD > VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD} > 0$.
+When $VPD < VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}  < 0$ (plant response dominates), and when $VPD > VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD} > 0$ (atmospheric demand dominates).
 Theoretically this $VPD_{crit}$ is only a function of PFT.
 We can look at how this value of $VPD_{crit}$ matches the data as different observed variables (including uncertainty) are allowed to vary.
-
-A final comparison for Figure \ref{agu_fig} is to the value of the sign term for PET.
-For PET, the sign term is a constant $\frac{c_p}{R_{air}}\approx 3.5$.
-This is far more than any of the observed values in Figure \ref{agu_fig}, even when uncertainty is included.
-So, even when atmospheric demand dominates plant response, the plant response still has a large role in attenuating atmospheric demand.
-In order to understand and capture the behavior of land-surface reponse, representation of the plants must be included in some form.
-This comparison casts doubt on the usefulness of PET for interprenting land-atmosphere evolution, as is done often in hydrometeorlogical studies.
 
@@
 Vertical black lines denote where our theory says $VPD_{crit}$ should be.
-For any obervations to the left of this line $\frac{]\partial \; ET}{\partial \; VPD}$ should be negative, and any ovservations the righ right of the line should be postiive.
+For any observations to the left of this line $\frac{\partial \; ET}{\partial \; VPD}$ should be negative (plant response dominates), and any observations the right of the line should be positive (atmospheric demand dominates).
 In addition to seeing how the theory $VPD_{crit}$ matches observations, these plots are also useful for observing how the scaling term affects the magnitude of the derivative, if at all.
@@
 Intuitively this makes sense; aerodynamic conductance will control how dominant balances at the land surface are communicated to the boundary layer.
-While $\Delta$ controlls the efficiency of latent fluxes in carying heat away from the surface, it appears this is a second order term, relative to $g_a$, for scaling $\frac{\partial \; ET}{\partial \; VPD}$.
+While $\Delta$ controls the efficiency of latent fluxes in carrying heat away from the surface, it appears this is a second order term, relative to $g_a$, for scaling $\frac{\partial \; ET}{\partial \; VPD}$.
 
@@
 \centering
-\includegraphics[width=\textwidth]{./fig06.png}
+\centerline{\includegraphics[width=1.4\textwidth]{./fig06.png}}
 \caption{Scatter plots of $\frac{\partial \; ET}{\partial \; VPD}$.
@@
 
-We derived a new form of Penman Monteith using $uWUE$ (\cite{Zhou_2015}) to remove implicit depedence on ET in the stomatal conductance term.
-With our new form of Penman Monteith we deveoped a theory for when an ecosystem will tend to reduce or increase ET with increasing VPD.
-The goal was to capture the leading order behavior of the stem to gain some intrinsic knowledge for its behaviro.
-This intuition can be used to disentangle land atmosphere feedbacks in more complicated scenariors, and will aid interpretation of obserations and sophisticated models.
+We derived a new form of Penman Monteith using $uWUE$ (\cite{Zhou_2015}) to remove the stomatal conductance term's implicit dependence on ET.
+With our new form of Penman Monteith we developed a theory for when an ecosystem will tend to reduce or increase ET with increasing VPD.
+The goal was to capture the leading order behavior of the system to gain some intrinsic knowledge for its behavior.
+This intuition can be used to disentangle land atmosphere feedbacks in more complicated scenarios, and will aid interpretation of observations and sophisticated models.
 
-Our theory states that each plant type will have it's own critical VPD below which ET will decrease (plant response dominating) and above which ET will increase (atmospheric demand dominates).
-We tested our theory using data from FLUXNET 2015, and found that for DBF and CSH the theory succeeds in capturing the leading order behavor.
-For ENF, the thoery succeeds in capturing the leading order behavior for the sign of $\frac{\partial \; ET}{\partial \; VPD}$, but not for magnitude.
+Our theory states that each plant type has a critical VPD below which ET will decrease (plant response dominating) and above which ET will increase (atmospheric demand dominates).
+We tested our theory using data from FLUXNET 2015, and found that for DBF and CSH the theory succeeds in capturing the leading order behavior.
+For ENF, the theory succeeds in capturing the leading order behavior for the sign of $\frac{\partial \; ET}{\partial \; VPD}$, but not for magnitude.
 In contrast, the model does not capture the behavior for CRO and GRA.
-In some ways we should have expected this, as our theory does not account for large variability we would expect for those plant types, including varying surface roughness, c3 vs c4 photosynthesis, and potential differences in irrigationa cross sites.
-Given these diffincies for CRO and GRA, we will withhold conclusions on their response to increasing VPD.
+In some ways we should have expected this, as our theory does not account for large sources of variability we would expect for those plant types, including varying surface roughness, c3 vs c4 photosynthesis, and potential differences in irrigation across sites.
+Given these deficiencies for CRO and GRA, we will withhold conclusions on their response to increasing VPD.
 
@@
 
-While our theory predicts common occurence of positive $\frac{\partial \; ET}{\partial \; VPD}$ (atmospheric demand dominates), the plant response in attenuating ET from the land surface is still important even when $\frac{\partial \; ET}{\partial \; VPD} > 0$.
-The magnitude of the ``sign term'' of our theory is always far below what it would be if we only considered the atmospheric demand term.
-This means that any drought indicies or analysis using PET will not capture how the vegetated surface respond and react to increases in atmospheric demand.
-However, the new uWUE-version of Penman Monteith we derived could be used as a complement to PET in drought indices.
-Basically one could compare calculate and compare indices  when plant response is ignored (PET) or included (uWUE-Penman Monteith).
+While our theory predicts common occurrence of positive $\frac{\partial \; ET}{\partial \; VPD}$ (atmospheric demand dominates), the plant response is still important even when $\frac{\partial \; ET}{\partial \; VPD} > 0$.
+The magnitude of the ``sign term'' of our theory is always far below what it would be if we only considered the atmospheric demand term (PET).
+This means that any drought indices or analysis using PET will not capture how the vegetated surface respond and react to increases in atmospheric demand.
+However, the new uWUE-version of Penman Monteith we derived (Equation \ref{et}) could be used as a complement to PET in drought indices.
+One could calculate and compare indices  when plant response is ignored (PET) or included (Equation \ref{et}).
 
-This paper provides the initial intuition for how the land surface respondes to atmospheric drying.
-This intuition for the one-way behavior of  land surface resonse to atmospheric drying is critical first step to disentangling land-atmosphere feedbacks at various scales, from diurnal to seasonal and beyond.
+This paper provides initial intuition for how the land surface responds to atmospheric drying.
+This intuition for the one-way behavior of  land surface response to atmospheric drying is critical first step to disentangling land-atmosphere feedbacks at various scales, from diurnal to seasonal and beyond.
 

### 2018-01-09 src/codebase/data_calc.py

@@ -150,6 +150,11 @@ def gpp_fixed_wrapper(_df, mean_df):
                                   lai=_df['lai_gppfixed'].mean()) # 1.0
   return _df
 
+def calc_sigma(_df):
+  """calcualtes a sgima for variabiility"""
+  _df['sigma'] = _df['uwue']/_df['uwue'].mean()
+  return _df
+
 def all_diagnostics(_df):
   """calcualtes all diagnostics"""
   _df['uwue'] = uwue(_df)
@@ -166,7 +171,10 @@ def all_diagnostics(_df):
   _df['et_pm_original'] = pm_et_orig(_df)
   _df['pet'] = pet(_df)
   # below is measure of uncertainty, comaprison to zhou et al
-  _df['sigma'] = _df['uwue']/_df['uwue_zhou']
+  # _df['sigma'] = _df['uwue']/_df['uwue_zhou']
+  # no, do comparison to mean to set mean sigma = 1
+  _df = _df.groupby('pft').apply(calc_sigma)
+  _df = _df.reset_index(drop=True)
   dfs = {'mean' : _df.groupby('pft').mean(),\
          'min' : _df.groupby('pft').min(),\
          'max' : _df.groupby('pft').max(),\

### 2018-01-17 doc/paper/vpd_et_paper_PG_v1.tex
added pierre's comments

--- 2018-01-09 doc/paper/vpd_et_paper.tex
@@
 
-\authors{A. Massmann\affil{1}, P. Gentine\affil{1}, C. Lin\affil{2}}
+\authors{A. Massmann\affil{1}, P. Gentine\affil{1}, C. Lin\affil{1,2}}
 
@@
 
-Answering the question ``When does vapor pressure deficit drive or reduce evapotranspiration?'' is motivated by two potential but opposing perspectives on the matter.
-The first perspective is that increases to vapor pressure deficit (VPD) increases atmospheric demand for water from the land surface, and this drives an increase in evapotranspiration (ET).
+Vapor pressure deficit (VPD)is expected to rise over continents in the future due to the combination of increased temperature and relative humidity drying (Byrne and O'Gorman 2013).
+In turn VPD modifies the atmospheric demand of evapotranspiration (ET) (Penman Monteith) but is also a stressors for stomata \citep{Leuning_1990, MEDLYN_2011}.
+Answering the question ``When does vapor pressure deficit drive or reduce evapotranspiration?'' is thus motivated by two potential but opposing perspectives on the matter.
+The first, hydrometeorological, perspective is that higher vapor pressure deficit  increases atmospheric demand for water from the land surface, and this drives an increase in evapotranspiration (ET).
 This perspective is particularly relevant because potential evapotranspiration (PET), which is used in many drought indices and hydrometeorological studies, only quantifies atmospheric demand effects.
-However, there is another possible perspective for vegetated surfaces.
+On the other hand, there is another perspective for vegetated surfaces.
 Plants' stomata have evolved to optimally regulate the exchange of water and carbon, and tend to close in response to increased atmospheric dryness \citep{Ball_1987, Leuning_1990, MEDLYN_2011}.
-Therefore, an increase in VPD may actually correspond to a decrease in ET because stomatal closure can cancel the effects of shifts to atmospheric demand.
-So really a question of ``When does VPD drive or reduce ET?'' becomes a question of ``When does plant response or atmospheric demand dominate?.''
+Therefore, an increase in VPD, in well watered soil conditions, may actually correspond to a decrease in ET because of stomatal closure.
+This decrease in ET would reduce and potentially cancel (in case of full closure) the effects of shifts to atmospheric demand.
+IN other words, the  question ``When does VPD drive or reduce ET?'' can be related to whether plant regulation or atmospheric demand dominate ET?.
 
-Quantifying the plant response to a perturbation to atmospheric VPD increases our understanding of the land surface response to shifts in atmospheric conditions.
-If plant response reduces ET in response to atmospheric drying then soil moisture will be conserved for later ET.
-An increase in ET in response to atmospheric drying will reduce soil moisture, but contribute increased moistening to the atmosphere.
-Understanding the one way response of the land surface to the atmosphere is a critical first step to disentangling land-atmosphere feedbacks.
-The sign and magnitude of land-surface response drives the co-evolution of the atmosphere and land-surface at many timescales, from diurnal to interdecadal.
+Adam: Merge those tow paragraphs, I only later noticed your paragraph below. Try to be as generic as possible when you are discussing the introduction.
 
-Before diving into math and data, we have some prior intuition for how the system might behave.
+If plant response reduces ET in response to atmospheric drying then soil moisture will be better conserved.
+This would seem a sensible evolutionary strategy to cope with aridity.
+If stomata were fully passive (similarly to soil pores.
+-cite Dani Or), increased atmospheric aridity would further reduce soil moisture.
+In turn, this would further increase aridity as low soil moisture levels increase the Bowen ratio, and in turn increase temperature and reduce atmospheric humidity (gentine et al.
+GRL 2016, Berg 2016).
+This however would not seem to be a sensible strategy for plants from an evolutionary standpoint.
+
 First, whether VPD drives or reduces ET should be a function of plant type.
@@
 
-In order to quantify plant response to perturbations to atmospheric demand for water, we apply a Penman-Monteith framework to eddy-covariance observations spanning various biomes and climates.
-Section 2 describes the data used, Section 3 derives the framework, Section 4 presents results, and Section 5 discusses conclusions.
+The objective of the present manuscript is thus to evaluate the VPD dependence of ET, in non-extreme soil drought conditions.
 The goal of this paper is to use reasonable approximations as a tool to increase intuition for plant response to atmospheric drying.
-This intuition will aid interpretation of observations and full complexity climate models.
+This intuition will aid interpretation of observations and full complexity models.
+In order to quantify plant response to perturbations in VPD, we apply a Penman-Monteith framework to derive theoretical response of ET to VPD.
+The model is then validated and tested at multiple eddy-covariance stations spanning various climates and plant functional types.
+Section 2 describes the data used.
+Section 3 derives the framework.
+Section 4 presents results and comparison to observations.
+Section 5 discusses conclusions.
 
@@
 \label{data}
-We use data from FLUXNET2015, including all sites with more than four years of data.
+We use data from the FLUXNET2015 database (ADAM: BE AS PRECISE AS POSSIBLE), including all sites with more than four years of data.
 Each site's plant functional type (PFT) was classified using the International Geosphere-Biosphere Programme vegetation classification scheme (\cite{Loveland_1999}).
@@
 \item To isolate the growing season, we only use days in which the average GPP exceeds 10\% of the observed 95th percentile of GPP for a given site.
-\item We remove days with rain and the day following to avoid issues with rain interception and sensor saturation and high relative humidity (\cite{MEDLYN_2011}).
+\item We remove days with rain and the day following to avoid issues with rain interception and sensor saturation at high relative humidity (\cite{MEDLYN_2011}).
 \end{itemize}
-Additionally, we restrict data to the daytime which is identified when downwelling shortwave radiation is greater than 50 W m$^{-2}$ and sensible heat flux is greater than 5 W m$^{-2}$.
+Additionally, we restrict data to the daytime, which is identified when downwelling shortwave radiation is greater than 50 W m$^{-2}$ and sensible heat flux is greater than 5 W m$^{-2}$.
 To further reduce the chance of sensor saturation at high relative humidity, we remove all time steps for which VPD is less than .01 kPa.
@@
 
+ADAM MENTION IF YOU ARE USING HOURLY OR SEMI-HOURLY DATA AND SAY THAT YOU ARE USING ALSO WEATHER STATION DATA.
+MENTION GPP WAS DERIVED FROM WHICH TECHNIQUE.
+CLOSURE?
+SEE CHANGJIE's DESCRIPTION OF DATA USED
+
 \section{Methods}
 \label{methods}
-The Penman-Monteith equation (hereafter PM) estimates ET as a function of atmospheric and land-surface variables:
+The Penman-Monteith equation (hereafter PM) REFERENCE estimates ET as a function of observable atmospheric variables and surface conductances:
 \begin{linenomath*}
@@
 \end{linenomath*}
-
 where variable definitions are given in Table 1.
-\citet{MEDLYN_2011} developed a model for stomatal conductance ($g_s$) by combining optimal photosynthesis theory with empirical approaches.
-The result for leaf-scale stomatal resistance was:
+\citet{MEDLYN_2011} developed a model for stomatal conductance ($g_s$) by combining optimal photosynthesis theory FARQUAHR, KATUL with empirical approaches, which describes the dependence of $g_s$ to VPD.
+The result for leaf-scale stomatal conductance was:
 
@@
 \end{linenomath*}
+This model has been shown to behave very well across PFTs, compared to other models (LIN AND MEDLYN NAT GEO).
 
-This can be adapted to an ecosystem-scale stomatal resistance by multiplying by leaf area index (LAI) and converting units to m s$^{-1}$:
+This can be adapted to the ecosystem scale by multiplying by leaf area index (LAI) and converting units to m s$^{-1}$:
 
@@
 
-While Equation 3 can be used in PM, it will make analytical work with the function intractable because $A$, net CO$_2$ assimilation, is closely functionally related to ET itself.
-To remove dependence of ET on $A$ we can use the semi-empirical results of \citet{Zhou_2015}.
-\citet{Zhou_2015} showed that:
+While Equation 3 can be used in PM (equation 1), it will make analytical work with the function intractable because $A$, net CO$_2$ assimilation, is functionally related to ET itself.
+To remove the dependence of ET on $A$ we can use the semi-empirical results of \citet{Zhou_2015}.
+\citet{Zhou_2015} showed that the underlying Water Use Efficiency $uWUE$:
 
@@
 is relatively constant across time and space (within plant functional type).
-If, following \citet{Lin_2015}, we approximate $g_0$ as $0$, we can use uWUE to remove $A$ from $g_s$ in a way that makes PM analytically tractable:
+If, following \citet{Lin_2015}, we approximate $g_0$ as $0$ (i.e.
+we neglect cuticular and epidermal losses - a reasonable assumption except in very dry conditions), we can use uWUE to remove $A$ from $g_s$ in a way that makes PM analytically tractable:
 
@@
 
-Note that $uWUE$ is fit on the ecosystem scale in \citet{Zhou_2015} so GPP in Equation \ref{uwue} is really $A\cdot \text{LAI}$.
+Note that $uWUE$ is fit on the ecosystem scale in \citet{Zhou_2015} so GPP in \ref{uwue} is really $A\cdot \text{LAI}$.
 This leads to the cancelation of LAI in addition to uWUE in Equation \ref{medlyn}.
-Plugging Equation 5 into Equation 1 and rearranging gives:
+Plugging Equation 5 into Equation 1 and rearranging gives a new expression for ET directly as a function of VPD:
 
@@
 However, our sampling of sites at the global scale may introduce some deviations of $uWUE$ from those observed in \citet{Zhou_2015}.
-Also, we wish to include some measure of uncertainty in our analysis to guide if our many assumptions and simplifications are reasonable.
+Also, we wish to include some measure of uncertainty in our analysis to check whether our assumptions and simplifications are reasonable.
 To account for both mean deviations of $uWUE$ and uncertainty, we will introduce an uncertainty parameter $\sigma$ modifying $uWUE$:
@@
 
-Now, from each FLUXNET observation we can calculate single value of $\sigma$:
-
-\begin{linenomath*}
-  \begin{equation}
-\sigma = - \frac{g_a \gamma c_s \sqrt{VPD} L_v P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p VPD\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})}
-    \label{sigma}
-  \end{equation}
-\end{linenomath*}
-
-The variability of $\sigma$ across sites and time will provides some measure of uncertainty in our model, assumptions, as well as the FLUXNET observations themselves.
+Now, from each FLUXNET observation we can evaluate $\sigma$ at each time step and thus we can evaluate departure from our theory (as a departure from unity).
+The variability of $\sigma$ across sites and time will provides a measure of uncertainty in our model, assumptions, as well as the FLUXNET observations themselves.
 To correct for differences in sampling between \cite{Zhou_2015} and our data, we set uWUE such that $\overline{\sigma} = 1$.
@@
 With Equation \ref{d_et} we have provided an analytical framework for ecosystem response to atmospheric demand perturbations.\explain{Should I take the derivative this way?
-I think it's more mathematically correct, but it makes the explanation more complicated and only changes the solution by a factor of two.} There are a few subtleties to taking the derivative in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$) and $VPD$ are functionally related, so while taking the derivative we evaluate $\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET} {\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD} \Big|_{\text{RH fixed}} + \frac{\partial \; ET}{\partial \; RH} \frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$ fixed}}$.
+I think it's more mathematically correct, but it makes the explanation more complicated and only changes the solution by a factor of two.
+ADAM I AM NOT SURE I UNDERSTNAD WHAT YOU MEAN HERE.
+USE THE RIGHT ONE!
+NOT ONE THAT IS SIMPLY PROPORTIONAL} There are a few subtleties to taking the derivative in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$) and $VPD$ are functionally related, so while taking the derivative we evaluate $\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET} {\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD} \Big|_{\text{RH fixed}} + \frac{\partial \; ET}{\partial \; RH} \frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$ fixed}}$.
 $RH$ and $e_s$ are assumed to be approximately orthogonal.
+IS THIS WHAT YOU ARE DOING?
+GIVE THE ANSWER YOU ARE USING
 
 The $VPD$ dependence in Equation \ref{d_et} is a little opaque.
-However, mean $VPD$ is 1062 Pa, so $\sqrt{VPD}$ is 32.6 Pa$^{1/2}$, which is much less than $g_1$ (with the exception of ENF; Table \ref{pft}).
-So a series expansion in the limit $\frac{\sqrt{VPD}}{g_1} \to 0$ gives an approximation which makes the functional form more clear: \explain{Should I even include the series expansion?}
+However, a typical value of $VPD$ is about 1kPa, so $\sqrt{VPD}$ is 33 Pa$^{1/2}$, which is much less than $g_1$ (with the exception of ENF; Table \ref{pft}).
+So a series expansion in the limit $\frac{\sqrt{VPD}}{g_1} \to 0$ gives an approximation which makes the functional form more clear: \explain{Should I even include the series expansion?
+NO}
 
@@
 
-One final comment on our derivation which will not be discussed further but is relevant for future analysis: if we  approximate $c_s$ at a global mean CO$_2$ concentration, then the RHS of Equation \ref{et} is fully defined using commonly available weather station data and the constants published in \citet{Zhou_2015, Lin_2015}.
+We note one final comment on our derivation which= is relevant for future analysis.
+If we  approximate $c_s$ at a global mean CO$_2$ concentration, then the RHS of Equation \ref{et} is fully defined using commonly available weather station data and the constants published in \citet{Zhou_2015, Lin_2015}.
 This makes Equation \ref{et} a useful alternative to PET in drought indices and hydrometeorological analysis for vegetated surfaces.
@@
   $\sigma$ & uncertainty parameter & -& - \\
-  $c_s$ & CO$_2$ concentration & $\mu$ mol CO$_2$ mol$^{-1}$ air& - \\
+  $c_s$ & surface CO$_2$ concentration & $\mu$ mol CO$_2$ mol$^{-1}$ air& - \\
 \hline
@@
 Figure \ref{lai_fig} presents the histogram of calculated $\sigma$s with outliers removed (lowest and highest 5\% percent).
-All remaining $\sigma$ values are $O(1)$ which provides confidence in model framework.
+All remaining $\sigma$ values are close to unity ($O(1)$) which provides confidence in model framework.
 
@@
 
-An additional concern is that the $\sigma$ term may in fact be some function of $VPD$, in which case the dependence would need to be accounted for when taking the derivative.
+An additional concern is that $\sigma$ may in fact be correlated with $VPD$, in which case the dependence would need to be accounted for when taking the derivative.
 Figure \ref{lai_vpd_fig} plots the joint distribution of $\sigma$ and VPD, and shows that $\sigma$ is very weakly a function of VPD.
@@
 Before calculating the sensitivity of ET to VPD, we will consider the functional form of Equation \ref{d_et}.
-There are two main terms: a ``scaling'' term which modifies the magnitude but not the sign:
+There are two main terms: a ``scaling'' term, which modifies the magnitude but not the sign of the derivative $\partial ET/\partial VPD$:
 
@@
 
-and a ``sign'' term which determines whether the derivative is positive or negative:
+and a ``sign'' term, which determines whether the derivative is positive i.e. atmospheric demand driven or negative i.e. physiologically controlled:
 
@@
 If the second term is larger magnitude then plant response dominates, while if the first term is larger then atmospheric demand dominates.
-Given that the sign term is crucial to answering our question ``When does VPD drive or reduce ET?'', we will explore its functional form first.
+The sign will determine whether ET is increasing or decreasing and we thus evaluate its functional form first.
+I AM OT A BIG FAN OF PLACING QUESTIONS IN THE TEXT, SOUNDS LIKE A TRAIN OF THOUGHT.ALSO TRY TO BE MORE FORMAL IN YOUR WRITING STYLE
 
@@
 \label{sign_func}
-First, let's explore the variables within the sign term to gain better intuition for what determines when VPD drives vs.
-reduces ET.
-The variability of $c_s$ and $\gamma$ is relatively less than $\sigma$ and $VPD$ variability, so sign term variability within PFT will be approximately determined by just $\sigma$ and $VPD$.
-If we fix uncertainty $\sigma$ at 1, then the sign term is approximately just a function of $VPD$.
-We can further determine a $VPD_{crit}$ where $\frac{\partial \; ET}{\partial \; VPD} = 0$:
+First, we explore the variables within the sign term to gain better intuition on the driver of either the increase or reduction of EF with VPD.
+$c_s$ and $\gamma$ are relatively constant to a good approximation so that the variability is dominated by $\sigma$ and $VPD$.
+$uWUE$ could vary with soil moisture but has been shown to be relatively constant except in very dry soil conditions ZHOU ET AL, LIN ET AL.
+We thus further assume that the soil is not too dry so that $uWUE$ can be approximated as a constant.
+This then means that the sign term only depends on VPD for a given PFT and is approximately just a function of $VPD$.
+We can further determine a critical threshold operating increase from decrease in ET, such that $VPD_{crit}$ where $\frac{\partial \; ET}{\partial \; VPD} = 0$:
 
@@
 For any values of $VPD$ less than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be negative, and for values of $VPD$ greater than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be positive.
-Intuitively this makes sense; as VPD increases ecosystems will start to run out of strategies to hold back water, and atmospheric demand will increase.
-Any further increases in atmospheric demand will then result in an increase in VPD.
+In other words, ecosystems can regulate and mitigate evaporative losses up to a VPD limit, above which atmospheric demand is just too high to be entirely compensated by stomatal regulation.
+We note however that even though ET increases again above the critical threshold, $VPD_{crit}$, ET is still much lower that potential evaporation as stomata are stills strongly regulating vapor fluxes.
+Even in the absence of soil pore evaporation, stomata do not shut down entirely at very high VPD and ET does not go to zero, because stomata are still slightly open to perform some photosynthesis (CITE BALL, MEDLYN ETC).
+In addition, upward xylem transport is necessary to maintain phloem transport and thus carbon allocation (MENCUCCINI).
+
+REMOVE THIS PARAGRAPH: Intuitively this makes sense TRY NOT TO WRITE LIKE THIS< BE MORE FORMAL; as VPD increases ecosystems will start to run out of strategies to hold back water, and atmospheric demand will increase.
+SOMETHING WAS NOT CONSISTENT IN THE SENTENCE HERE: Any further increases in atmospheric demand will then result in an increase in VPD.
 The extent to which a given ecosystem has evolved to conserve water will determine at what VPD atmospheric demand takes over.
@@
 \caption{Values of $VPD_{crit}$, where $\frac{\partial \; ET}{\partial \; VPD} = 0$, evaluated at PFT average values for $R_{air}$, $\sigma$, $\gamma$, and $c_s$.
-For reference, these values are also provided.
+For reference, these values are also provided CITE TABLE.
 For values of $VPD$ less than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be negative, and for values of $VPD$ greater than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be positive.
@@
 
-Figure \ref{term3} provides a greater description for how the sign term varies with $VPD$, as a function of PFT.
+Figure \ref{term3} provides a greater description for how the sign term varies with $VPD$, as a function of PFT I AM NOT A FAN OF SENTENCES LIKE: FGIGURE XX SHOWS BLABLABLA.
+RATHER SAY WHAT YOU SEE AND JUST PLACE FIGURE IN BRACKET.
+FRAME THIS DIFFERNTLY< SAY THAT THE OSBERVATIONS VALIDATE TEH CRITCAL THERESHOLD ANALYTICALLY FOUND.
 Equation \ref{d_et_approx} aids interpretation of Figure \ref{term3}.
@@
 % Higher values corresponde to plants that are more willing to spend water on primary production and less evolved to conserve water.\explain{really need to think about this sentence more, tehre is more nuance and I actually thing g1 and uwue really partially cancel.}
-Figure \ref{term3} confirms our physical intuition: CROs are the least water conservative so have the smallest constant portion of Term 3 ($\frac{\gamma c_s }{1.6 \; R*\; \sigma \; \text{ uWUE }} \left(\frac{1}{g_1}\right)$), while CSH are the most water conservative and have the largest constant portion of Term 3.
+Based on Figure \ref{term3}, CROs are the least water conservative so have the smallest constant portion of Term 3 ($\frac{\gamma c_s }{1.6 \; R*\; \sigma \; \text{ uWUE }} \left(\frac{1}{g_1}\right)$), while CSH are the most water conservative and have the largest constant portion of Term 3.
 For the VPD-dependent terms in Equation \ref{d_et_approx}, differences in $g1$ between PFTs exert a greater influence than differences in $uWUE$.
@@
 ENF ($g1 = 74.31$) has by far the largest VPD dependence of response, while CRO ($g1 = 183.1$) has the smallest VPD dependence.
+SAY WHAT IT MEANS IN LAYMAN's TERMS
 
 A primary takeaway from Figure \ref{term3}a is that according to our theory for all PFTs except for crops there is frequent occurrence of a negative (plant dominating) ET response to increases in VPD.
+This thus means that plants are able in most atmospheric conditions to redue ET in response to increased VPD and thus to reduce water losses.
 To better illustrate this, the ranges of observed environmental VPDs at the FLUXNET sites are plotted parallel to the x-axis.
-25th, 50th, and 75th percentiles are marked by stars, and the location of VPD$_{crit}$ is marked by a vertical line.
-For CSH (water conservative), VPD is always less than VPD$_{crit}$ (plant response dominates).
-For CRO, VPD is almost always more than VPD$_{crit}$.
-For forests and grass, whether plant response or atmospheric demand dominates will depend more on the environment, with about half of of observed VPD less than VPD$_{crit}$ (plant response dominates).
-It's also important to note that even when atmospheric demand dominates, the response is still far smaller than it would be for PET (atmospheric demand only).
+ALL OF THIS SHOULD BE IN CAPTION NOT IN MAIN TEXT: (25th, 50th, and 75th percentiles are marked by stars),
+The location of VPD$_{crit}$ is marked by a vertical line.
+For CSH, VPD is always less than VPD$_{crit}$ so that the plant response dominates and empathizes the water conservative strategy of those plants.
+For CRO on the other, VPD is almost always higher than VPD$_{crit}$, emphasizing that those plants are water intensive and were actually engineered for photosynthesis rather than water saving.
+For forests and grass, whether plant response or atmospheric demand dominates is more dependent on the environment, with about half of of observed VPD less than VPD$_{crit}$, i.e.
+in conditions where plant response dominates.
+It is also important to note that even when atmospheric demand dominates, the response is still far smaller than it would be for potential evaporation i.e.
+atmospheric demand only, emphasizing that there is still a strong regulation of evaporative flux by stomata and though the plant xylem.
 The sign term in this case would just be a constant ($\frac{c_p}{R_{air}} \approx 3.5$), which is far larger than any part of the curves for any PFT.
-This highlights the deficiencies of PET's ability to capture land response to changes in atmospheric dryness (VPD).
-However, up to this point we have assumed $\sigma$ (uncertainty) equal to one.
-We still need to test if inclusion of uncertainty changes the story.
+This highlights the deficiencies of PET's ability to capture land response to changes in atmospheric dryness (VPD) SEEMS TOO GENERIC BETTER EXPLAIN.
+However, we have assumed $\sigma$ (uncertainty) equal to one.
+We still need to test if inclusion of uncertainty could change our conclusions.
 \begin{figure}[h]
@@
 Section \ref{testing} further examines the consistency of our theory, given  the apparent sensitivity to uncertainty ($\sigma$).
+THIS PARAGRAPH DOES NOT SEEM CONCLUSIVE AND MORE A TRAIN OF THOUGHT.
+LAY DOWN EXACTLY WHAT YOU HAVE IN MIND.
 
@@
 
-While the above discussion of the sign of $\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our research question, the magnitude of $\frac{\partial \; ET}{\partial \; VPD}$ will also impact statistics of $\frac{\partial \; ET}{\partial \; VPD}$ and the importance of $VPD$ variability for overall $ET$ variability.
-So we will more closely examine the scaling term: $\frac{P}{T} \propto \rho$, so this should vary little relative to aerodynamic conductance and $\Delta$.
-$\gamma$ should also be relatively constant, so the scaling term should be primarily a function of aerodynamic conductance and temperature (through the function $\Delta$).
-This makes sense, as aerodynamic conductance represents how efficiently response at the surface is communicated to the atmosphere.
-As it increases, any plant response will be communicated more strongly to the atmosphere (and vice-versa).
+While the above discussion of the sign of $\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our research question, the magnitude of $\frac{\partial \; ET}{\partial \; VPD}$ will also the relative magnitude of the change $\frac{\partial \; ET}{\partial \; VPD}$ and the importance of $VPD$ variability for overall $ET$ variability.
+So we now more closely examine the scaling term: $\frac{P}{T} \propto \rho$.
+Air density $\rho$ scaling term varies little relative to aerodynamic conductance and $\Delta$.
+The psychrometric constant $\gamma$ is also relatively constant, so the scaling term should be primarily a function of aerodynamic conductance and temperature, through the slope of the Clauisu-Clapeyron relationship $\Delta$.
+This is as expected, given that the aerodynamic conductance represents the efficiency fo the transfer of surface anomalies to the atmosphere.
+As it increases I THINK YOU REALLY MEAN CONDUCATANCE HERE - BE CAREFUL, any plant response will be communicated more strongly to the atmosphere (and vice-versa).
+
+REMOVE ANY SNETENCES LIKE THIS MAKES SENSES ETC...
 
@@
 
-So far, idealized plots and statistics have illuminated the form of $\frac{\partial \; ET}{\partial \; VPD}$ and how it varies with PFT.
-Large mean uWUE shifts CRO and DBF towards positive $\frac{\partial \; ET}{\partial \; VPD}$, but $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ is still negative for DBF due to its lower $g1$.
-ENF's low $g1$, even lower than DBF, increases the dependence of $\frac{\partial \; ET}{\partial \; VPD}$ on $VPD$, and makes the function very nonlinear.
+Theoretical derivation has so far illuminated the dependence of $\frac{\partial \; ET}{\partial \; VPD}$ and how it varies across PFT.
+Large mean $uWUE$ shifts CRO and DBF towards positive $\frac{\partial \; ET}{\partial \; VPD}$ i.e.
+increasing ET at higher VPD, but $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ is still negative for DBF due to its lower $g1$ SAY PHISICALLY WHAT IT IS FOR THE READER TO UNDERSTAND.
+ENF's low $g1$, which is even lower than DBF, increases the dependence of $\frac{\partial \; ET}{\partial \; VPD}$ on $VPD$, and makes the function strongly nonlinear.
 This has the side effect of pushing $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ negative further than other PFTs for a given fraction $\frac{\partial \; ET}{\partial \; VPD} < 0$ and magnitude $\frac{\partial \; ET}{\partial \; VPD}(\overline{T,\ldots,VPD})$.
+SAY THIS IN WORDS AS WELL OR YOU WILL LSOE THE READER.
 GRA shows the opposite behavior; a relatively high $g1$  makes the function more linear, decreasing the magnitude of $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ for a given fraction $\frac{\partial \; ET}{\partial \; VPD} < 0$ and negative $\frac{\partial \; ET}{\partial \; VPD}(\overline{T,\ldots,VPD})$.
-Finally, the low $uWUE$ of CSH pushes it toward by far the lowest values $\frac{\partial \; ET}{\partial \; VPD}$ (Figure \ref{term3}).
+SAY IN WORDS AGAIN WHAT IT MEANS OTHERWISE PEOPLE WILL NOT UNDERSTAND OUR RESULTS.
+Finally, the low $uWUE$ of CSH pushes it toward the lowest values of $\frac{\partial \; ET}{\partial \; VPD}$ (Figure \ref{term3}) across all PFTs WHAT DOES IT MEAN AGAIN.
 Variability in $VPD$ accounts for the largest amount of $ET$ variability for CSH.
 For the other PFTs, $VPD$ contributes less to $ET$ variability, but still represents about 30-40 \% of $R$'s contributions to ET variability.
+WHAT IS THE SORUCE OF VARIBILITY OF THE REST OF ET I DON"T KNOW IF THAT IS RELEVANT FOR THE PRESENT PAPER.
 
@@
 
-\subsection{Testing the theory - does the data match the model?}
+\subsection{Validation of theory at eddy-covariance sites}
 \label{testing}
-So far we have developed a theory for $\frac{\partial \; ET}{\partial \; VPD}$'s behavior.
-Most important to answering ``When does VPD drive or reduce ET?'' is that the sign of our derivative term is correct.
-So, we need to test if when we include uncertainty the observed derivatives match what our theory would say for the sign term.
+So far we have developed a theory for $\frac{\partial \; ET}{\partial \; VPD}$'s behavior as a function of VPD and PFTs.
+We now turn to an direct evaluation against observations taken from a wide range of eddy-covariance datasets (SEE METHODS).
 
-Figure \ref{agu_fig} shows the distribution of of the sign term when uncertainty is included, as compared to what the theory says it should be.
-Our goal for this paper was to capture the leading order behavior and guide intuition.
-For DBF and CSH our theory is successful towards this goal: the theory matches the leading order behavior of the function when uncertainty is included.
-Given the assumptions we've made, and the uncertainties of flux tower observations themselves, there will always be a large amount of noise.
+Figure \ref{agu_fig} shows the distribution of of the sign term (when uncertainty is included), as compared to what the theory would predict.
+Our goal  was to capture the leading order behavior of the ET dependence on VPD.
+Given the assumptions we have made, and the uncertainties of flux tower observations themselves, we expect a large amount of noise when reprdrocuing the derivatives of ET.
+
+ALL OF THE REST SHOULDB EB EXPLAINED IN MUCH MORE DETAIL DESCRIBE EACH FIGURE AND THE RESULTS For DBF and CSH our theory is successful towards this goal: the theory matches the leading order behavior of the function when uncertainty is included.
 The plots for DBF and CSH look as one would expect if one just added noise on top of our theory.
@@
 It is also noteworthy that we used published constants for $g1$, which determines the VPD dependence, from \cite{Lin_2015}.
-We could have fit our own constants to tweak the theory to match the observations, but our goal was to use a priori physical constants whenever possible.
+We could have fit our own constants so that the theory would best match the observations, but our goal was to use a priori physical constants whenever possible.
 
-While the above discussion shows that our theory has some limitations when applied to real data, especially for crops and grasslands, another question could be, how does our theory's performance compare to often-used PET?
+While the above discussion shows that our theory has some limitations when applied to some sites, especially for crops and grasslands, another question could be, how does our theory's performance compare to often-used PET?
 Again, as in Section \ref{sign_func}, the PET response to changes in VPD will be constant at about 3.5.
-Even with uncertainty included, no values of the response approach the PET case.
+Even with uncertainty included, no values of the response approach the PET case YOU NEED TO ELABORATE WAY MORE AND GUDE THE READER THROUGH YOUR THOUGHTS.
 PET's status outside the error bars of our theory highlight the usefulness of our theory (and Equation \ref{et}) for estimating land response and moisture budgeting.
@@
 Figure \ref{real} presents calculated $\frac{\partial \; ET}{\partial \; VPD}$ where, unless otherwise noted, all variables in Equation \ref{d_et} are allowed to vary, including uncertainty.
-Each column is a different quantity related to $\frac{\partial \; ET}{\partial \; VPD}$, and each row is a different PFT.
+Each column represents a different quantity related to $\frac{\partial \; ET}{\partial \; VPD}$, and each row is a different PFT.
 Vertical black lines denote where our theory says $VPD_{crit}$ should be.
 For any observations to the left of this line $\frac{\partial \; ET}{\partial \; VPD}$ should be negative (plant response dominates), and any observations the right of the line should be positive (atmospheric demand dominates).
-In addition to seeing how the theory $VPD_{crit}$ matches observations, these plots are also useful for observing how the scaling term affects the magnitude of the derivative, if at all.
+In addition to seeing how our theoretically-derived VPD threshold, $VPD_{crit}$, matches observations, these analysis DO NOT SAY PLOT OR IT SOUNDS LIKE A MSC PROJECT are also useful for observing how the scaling term affects the magnitude of the derivative, if at all.
 
 The scatter plots generally confirm expectations from Figure \ref{agu_fig}.
-By comparing different columns of Figure \ref{real} we can see the impact of $\sigma$ and $g_a$ on the variability of $\frac{\partial \; ET}{\partial \; VPD}$.
+By comparing different columns of Figure \ref{real} we can evaluate the impact of $\sigma$ and $g_a$ on the variability of $\frac{\partial \; ET}{\partial \; VPD}$.
 $g_a$'s scaling (included in columns 1 and 3) alters the magnitude considerably, particularly between PFTs.
@@
 However, the general story with the noise appears to match the cleaner signal when $\sigma$ is help constant and $VPD_{crit}$ is clearly visible.
-The exceptions are, as discussed above, for GRA and CRO, where the full complexity plots (Columns 1 and 2) to not match well with $\sigma$ held fixed (Columns 3 and 4).
+The exceptions are, as discussed above, for GRA and CRO, where the full complexity plots (Columns 1 and 2) do not match well with $\sigma$ held fixed (Columns 3 and 4).
 
-While $g_a$ has a large impact on magnitudes in the scaling, temperature, through the $\Delta$ term, does not appear to have as much of an impact in the complete estimation of $\frac{\partial \; ET}{\partial \; VPD}$.
+While $g_a$ has a large impact on the magnitude of the scaling, temperature, through the derisive of the Clausius-Clapeyron, $\Delta$, term, does not appear to have as much of an impact in the complete estimation of $\frac{\partial \; ET}{\partial \; VPD}$.
 Therefore it appears that the absolute magnitude of the derivative can be mostly attributed to the sign term, and its direct scaling by aerodynamic conductance.
-Intuitively this makes sense; aerodynamic conductance will control how dominant balances at the land surface are communicated to the boundary layer.
+Intuitively this is reasonable; aerodynamic conductance will control how dominant balances at the land surface are communicated to the boundary layer.
 While $\Delta$ controls the efficiency of latent fluxes in carrying heat away from the surface, it appears this is a second order term, relative to $g_a$, for scaling $\frac{\partial \; ET}{\partial \; VPD}$.
@@
 Our $\sigma$-based method of uncertainty propagation blurs the idealized expectations the most for GRA and CRO.
-We therefor have the most confidence in our conclusion based on Equation \ref{d_et} for PFTs CSH, DBF, and ENF, as the sign of the derivative with uncertainty included matches (to leading order) the story when $\sigma$ is held fixed.
+We therefore have the most confidence in our conclusion based on Equation \ref{d_et} for PFTs CSH, DBF, and ENF, as the sign of the derivative with uncertainty included matches (to leading order) the story when $\sigma$ is held fixed.
 
@@
 % \end{figure}
+
+HERE DISCUSS LIMITATION OF THE METHODS TO BVERY DRY CONDITIONS AND THE FEEDBCAK TO THE ATMOSPHERE OF VERY DRY SOIL CONDITIONS (BERG ET AL 2016, GENTINE ET AL> 2016)
 

### 2018-01-17 doc/paper/vpd_et_paper_PG_v1.tex
done with wed mon's writing session

@@
 
-\section{Introduction}\explain{This section needs to be fleshed out, and I definitely need to add more citations}
+\section{Introduction}
 
-Vapor pressure deficit (VPD)is expected to rise over continents in the future due to the combination of increased temperature and relative humidity drying (Byrne and O'Gorman 2013).
-In turn VPD modifies the atmospheric demand of evapotranspiration (ET) (Penman Monteith) but is also a stressors for stomata \citep{Leuning_1990, MEDLYN_2011}.
+Vapor pressure deficit (VPD) is expected to rise over continents in the future due to the combination of increased temperature and relative humidity drying\explain{Is increased RH conclusive and universal?- I'm not so sure...
+is there another citation?
+I don't think Byrne 2013 shows that.} \citep{Byrne_2013}..
+In turn VPD modifies the atmospheric demand of evapotranspiration (ET) (Penman Monteith) but is also a stressor for stomata \citep{Leuning_1990, MEDLYN_2011}.
+
 Answering the question ``When does vapor pressure deficit drive or reduce evapotranspiration?'' is thus motivated by two potential but opposing perspectives on the matter.
@@
 This decrease in ET would reduce and potentially cancel (in case of full closure) the effects of shifts to atmospheric demand.
-IN other words, the  question ``When does VPD drive or reduce ET?'' can be related to whether plant regulation or atmospheric demand dominate ET?.
-
-Adam: Merge those tow paragraphs, I only later noticed your paragraph below. Try to be as generic as possible when you are discussing the introduction.
-
+In other words, the  question ``When does VPD drive or reduce ET?'' can be related to whether plant regulation or atmospheric demand dominate ET.
 If plant response reduces ET in response to atmospheric drying then soil moisture will be better conserved.
 This would seem a sensible evolutionary strategy to cope with aridity.
-If stomata were fully passive (similarly to soil pores.
--cite Dani Or), increased atmospheric aridity would further reduce soil moisture.
+If stomata were fully passive \citep [similar to soil pores, e.g.
+][]{Or_2013}, increased atmospheric aridity would further reduce soil moisture.
 In turn, this would further increase aridity as low soil moisture levels increase the Bowen ratio, and in turn increase temperature and reduce atmospheric humidity (gentine et al.
@@
 \label{data}
-We use data from the FLUXNET2015 database (ADAM: BE AS PRECISE AS POSSIBLE), including all sites with more than four years of data.
-Each site's plant functional type (PFT) was classified using the International Geosphere-Biosphere Programme vegetation classification scheme (\cite{Loveland_1999}).
-Because physical constants used in Section \ref{methods} are only published for five plant functional types (PFTs - see Table \ref{pft}),  only 56 of the 77 sites were used.
-Figure \ref{map_fig} \explain{map needs to be improved - it's a placeholder for now} presents each site and its plant functional type.
+We use both meterological and turbulent flux data from the FLUXNET2015 database, including all sites with more than four years of data\explain{Pierre says ``be as precise as possible,'' but not sure what he means}.
+Each site's plant functional type (PFT) was classified using the International Geosphere-Biosphere Programme vegetation classification scheme \citep{Loveland_1999}.
+The physical constants used in Section \ref{methods} are only published for five plant functional types (PFTs) crops (CRO), deciduous broadleaf forest (DBF), evergreen needleleaf forest (ENF), grass (GRA), and closed shrub (CSH) (see Table \ref{pft}).
+There are 56 sites with these plant functional types, and their location is shown in  Figure \ref{map_fig} \explain{map needs to be improved - it's a placeholder for now}.
 
@@
 \item To isolate the growing season, we only use days in which the average GPP exceeds 10\% of the observed 95th percentile of GPP for a given site.
+GPP is calculated using the nighttime respiration partitioning method.
 \item We remove days with rain and the day following to avoid issues with rain interception and sensor saturation at high relative humidity (\cite{MEDLYN_2011}).
@@
 To further reduce the chance of sensor saturation at high relative humidity, we remove all time steps for which VPD is less than .01 kPa.
-Timesteps with negative observed GPP or ET are also removed.
+Timesteps with negative observed GPP or ET are also removed, and we aggregate half hourly data to hourly averages to reduce noise.
 After these quality control procedures, 332556 upscaled hourly observations remain.
-
-ADAM MENTION IF YOU ARE USING HOURLY OR SEMI-HOURLY DATA AND SAY THAT YOU ARE USING ALSO WEATHER STATION DATA.
-MENTION GPP WAS DERIVED FROM WHICH TECHNIQUE.
-CLOSURE?
-SEE CHANGJIE's DESCRIPTION OF DATA USED
 
@@
 \label{methods}
-The Penman-Monteith equation (hereafter PM) REFERENCE estimates ET as a function of observable atmospheric variables and surface conductances:
+The Penman-Monteith equation (hereafter PM \explain{find pm ciatation in shuttlewroth}) estimates ET as a function of observable atmospheric variables and surface conductances:
 \begin{linenomath*}
@@
 where variable definitions are given in Table 1.
-\citet{MEDLYN_2011} developed a model for stomatal conductance ($g_s$) by combining optimal photosynthesis theory FARQUAHR, KATUL with empirical approaches, which describes the dependence of $g_s$ to VPD.
+\citet{MEDLYN_2011} developed a model for stomatal conductance ($g_s$) by combining optimal photosynthesis theory \citep{Farquhar_1980, Katul_2010} with empirical approaches, which describes the dependence of $g_s$ to VPD.
 The result for leaf-scale stomatal conductance was:
@@
 \end{linenomath*}
-This model has been shown to behave very well across PFTs, compared to other models (LIN AND MEDLYN NAT GEO).
+This model has been shown to behave very well across PFTs, compared to other models \citep{Lin_2015}.
 
@@
 Now, from each FLUXNET observation we can evaluate $\sigma$ at each time step and thus we can evaluate departure from our theory (as a departure from unity).
-The variability of $\sigma$ across sites and time will provides a measure of uncertainty in our model, assumptions, as well as the FLUXNET observations themselves.
+The variability of $\sigma$ across sites and time provides a measure of uncertainty in our model, assumptions, as well as the FLUXNET observations themselves.
 To correct for differences in sampling between \cite{Zhou_2015} and our data, we set uWUE such that $\overline{\sigma} = 1$.
@@
 
-With Equation \ref{d_et} we have provided an analytical framework for ecosystem response to atmospheric demand perturbations.\explain{Should I take the derivative this way?
-I think it's more mathematically correct, but it makes the explanation more complicated and only changes the solution by a factor of two.
-ADAM I AM NOT SURE I UNDERSTNAD WHAT YOU MEAN HERE.
-USE THE RIGHT ONE!
-NOT ONE THAT IS SIMPLY PROPORTIONAL} There are a few subtleties to taking the derivative in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$) and $VPD$ are functionally related, so while taking the derivative we evaluate $\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET} {\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD} \Big|_{\text{RH fixed}} + \frac{\partial \; ET}{\partial \; RH} \frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$ fixed}}$.
+With Equation \ref{d_et} we have provided an analytical framework for ecosystem response to atmospheric demand perturbations.
+There are a few subtleties to taking the derivative in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$) and $VPD$ are functionally related, so while taking the derivative we evaluate $\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET} {\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD} \Big|_{\text{RH fixed}} + \frac{\partial \; ET}{\partial \; RH} \frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$ fixed}}$.
 $RH$ and $e_s$ are assumed to be approximately orthogonal.
-IS THIS WHAT YOU ARE DOING?
-GIVE THE ANSWER YOU ARE USING
 
-The $VPD$ dependence in Equation \ref{d_et} is a little opaque.
-However, a typical value of $VPD$ is about 1kPa, so $\sqrt{VPD}$ is 33 Pa$^{1/2}$, which is much less than $g_1$ (with the exception of ENF; Table \ref{pft}).
-So a series expansion in the limit $\frac{\sqrt{VPD}}{g_1} \to 0$ gives an approximation which makes the functional form more clear: \explain{Should I even include the series expansion?
-NO}
-
-\begin{linenomath*}
-  \begin{equation}
-    \frac{\partial \;  ET}{\partial \; VPD} \approx \frac{g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_s }{1.6 \; R*\; \sigma \; \text{ uWUE }} \left( \frac{1}{g_{1}} - \frac{3 \sqrt{VPD}}{2 g_{1}^{2}} + \frac{2 \sqrt{VPD}^{2}}{g_{1}^{3}} - \frac{5 \sqrt{VPD}^{3}}{2 g_{1}^{4}} + \mathcal{O}\left(\left(\frac{\sqrt{VPD}}{g_1}\right)^{4}\right) \right) \right)
-    \label{d_et_approx}
-  \end{equation}
-\end{linenomath*}
-
-We note one final comment on our derivation which= is relevant for future analysis.
+We note one final comment on our derivation which is relevant for future analysis.
 If we  approximate $c_s$ at a global mean CO$_2$ concentration, then the RHS of Equation \ref{et} is fully defined using commonly available weather station data and the constants published in \citet{Zhou_2015, Lin_2015}.
@@
 
-% below show histogram of site specific spearmannr? think would be useful 
 \begin{figure}[h]
@@
 The sign will determine whether ET is increasing or decreasing and we thus evaluate its functional form first.
-I AM OT A BIG FAN OF PLACING QUESTIONS IN THE TEXT, SOUNDS LIKE A TRAIN OF THOUGHT.ALSO TRY TO BE MORE FORMAL IN YOUR WRITING STYLE
 
@@
 \label{sign_func}
-First, we explore the variables within the sign term to gain better intuition on the driver of either the increase or reduction of EF with VPD.
+First, we explore the variables within the sign term to gain better intuition on the driver of either the increase or reduction of ET with VPD.
 $c_s$ and $\gamma$ are relatively constant to a good approximation so that the variability is dominated by $\sigma$ and $VPD$.
-$uWUE$ could vary with soil moisture but has been shown to be relatively constant except in very dry soil conditions ZHOU ET AL, LIN ET AL.
+$uWUE$ could vary with soil moisture but has been shown to be relatively constant except in very dry soil conditions \citep{Zhou_2016, Lin_2015}.
 We thus further assume that the soil is not too dry so that $uWUE$ can be approximated as a constant.
@@
 We note however that even though ET increases again above the critical threshold, $VPD_{crit}$, ET is still much lower that potential evaporation as stomata are stills strongly regulating vapor fluxes.
-Even in the absence of soil pore evaporation, stomata do not shut down entirely at very high VPD and ET does not go to zero, because stomata are still slightly open to perform some photosynthesis (CITE BALL, MEDLYN ETC).
-In addition, upward xylem transport is necessary to maintain phloem transport and thus carbon allocation (MENCUCCINI).
-
-REMOVE THIS PARAGRAPH: Intuitively this makes sense TRY NOT TO WRITE LIKE THIS< BE MORE FORMAL; as VPD increases ecosystems will start to run out of strategies to hold back water, and atmospheric demand will increase.
-SOMETHING WAS NOT CONSISTENT IN THE SENTENCE HERE: Any further increases in atmospheric demand will then result in an increase in VPD.
-The extent to which a given ecosystem has evolved to conserve water will determine at what VPD atmospheric demand takes over.
+Even in the absence of soil pore evaporation, stomata do not shut down entirely at very high VPD and ET does not go to zero, because stomata are still slightly open to perform some photosynthesis \citep{Ball_1987, Leuning_1990, Medlyn_2011}.
+In addition, upward xylem transport is necessary to maintain phloem transport and thus carbon allocation (MENCUCCINI) \explain{Not sure citation  - Comparative criteria for models of the vasuclar transport systems of tall trees}.
 
@@
 
-Figure \ref{term3} provides a greater description for how the sign term varies with $VPD$, as a function of PFT I AM NOT A FAN OF SENTENCES LIKE: FGIGURE XX SHOWS BLABLABLA.
-RATHER SAY WHAT YOU SEE AND JUST PLACE FIGURE IN BRACKET.
-FRAME THIS DIFFERNTLY< SAY THAT THE OSBERVATIONS VALIDATE TEH CRITCAL THERESHOLD ANALYTICALLY FOUND.
-Equation \ref{d_et_approx} aids interpretation of Figure \ref{term3}.
-Larger $uWUE$ and $g1$ shift the leading-order constant term ($\frac{1}{g_1}$) towards smaller values, and the sign term towards overall positive values.
-% Higher values corresponde to plants that are more willing to spend water on primary production and less evolved to conserve water.\explain{really need to think about this sentence more, tehre is more nuance and I actually thing g1 and uwue really partially cancel.}
-Based on Figure \ref{term3}, CROs are the least water conservative so have the smallest constant portion of Term 3 ($\frac{\gamma c_s }{1.6 \; R*\; \sigma \; \text{ uWUE }} \left(\frac{1}{g_1}\right)$), while CSH are the most water conservative and have the largest constant portion of Term 3.
-For the VPD-dependent terms in Equation \ref{d_et_approx}, differences in $g1$ between PFTs exert a greater influence than differences in $uWUE$.
-Increasing $uWUE$ and $g1$ decreases the VPD-dependence, but $g1$ has the bigger effect, particularly at high VPD when the series becomes disordered.
-ENF ($g1 = 74.31$) has by far the largest VPD dependence of response, while CRO ($g1 = 183.1$) has the smallest VPD dependence.
-SAY WHAT IT MEANS IN LAYMAN's TERMS
+Differences in $uWUE$ and $g_1$ between PFTs alter the functional form of the sign term.
+Larger $uWUE$ and $g_1$ will shift the sign term towards overall positive values.
+$g_1$ additionally plays a primary role in determining dependnece on VPD: the smaller $g_1$, the greater the $VPD$ dependence for the PFT (Figure \ref{term3}).
+
+Based on Figure \ref{term3} and Table \ref{vpd_crit}, CROs are the least water conservative and have the most positive constant offset, while CSH are the most water conservative and have the most negative constant offset in the sign term.
+ENF ($g_1 = 74.31$) has by far the largest VPD dependence of response, while CRO ($g_1 = 183.1$) has the smallest VPD dependence.
+This means ENF's stomatal response evolved to change substantially depending on the envionmental dryness.
+This contasts\explain{think about this - what does it mean - go back to Franks and think about g1's relation to wue}
 
@@
 Large mean $uWUE$ shifts CRO and DBF towards positive $\frac{\partial \; ET}{\partial \; VPD}$ i.e.
-increasing ET at higher VPD, but $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ is still negative for DBF due to its lower $g1$ SAY PHISICALLY WHAT IT IS FOR THE READER TO UNDERSTAND.
-ENF's low $g1$, which is even lower than DBF, increases the dependence of $\frac{\partial \; ET}{\partial \; VPD}$ on $VPD$, and makes the function strongly nonlinear.
+increasing ET at higher VPD, but $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ is still negative for DBF due to its lower $g_1$ SAY PHISICALLY WHAT IT IS FOR THE READER TO UNDERSTAND.
+ENF's low $g_1$, which is even lower than DBF, increases the dependence of $\frac{\partial \; ET}{\partial \; VPD}$ on $VPD$, and makes the function strongly nonlinear.
 This has the side effect of pushing $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ negative further than other PFTs for a given fraction $\frac{\partial \; ET}{\partial \; VPD} < 0$ and magnitude $\frac{\partial \; ET}{\partial \; VPD}(\overline{T,\ldots,VPD})$.
 SAY THIS IN WORDS AS WELL OR YOU WILL LSOE THE READER.
-GRA shows the opposite behavior; a relatively high $g1$  makes the function more linear, decreasing the magnitude of $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ for a given fraction $\frac{\partial \; ET}{\partial \; VPD} < 0$ and negative $\frac{\partial \; ET}{\partial \; VPD}(\overline{T,\ldots,VPD})$.
+GRA shows the opposite behavior; a relatively high $g_1$  makes the function more linear, decreasing the magnitude of $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ for a given fraction $\frac{\partial \; ET}{\partial \; VPD} < 0$ and negative $\frac{\partial \; ET}{\partial \; VPD}(\overline{T,\ldots,VPD})$.
 SAY IN WORDS AGAIN WHAT IT MEANS OTHERWISE PEOPLE WILL NOT UNDERSTAND OUR RESULTS.
@@
 Resolving this feature is the key to answering our question, so perhaps the theory is still useful for ENF for the purpose of addressing when VPD drives or reduces ET.
-It is also noteworthy that we used published constants for $g1$, which determines the VPD dependence, from \cite{Lin_2015}.
+It is also noteworthy that we used published constants for $g_1$, which determines the VPD dependence, from \cite{Lin_2015}.
 We could have fit our own constants so that the theory would best match the observations, but our goal was to use a priori physical constants whenever possible.
@@
 %\bibliography{references.bib}
-\input{vpd_et_paper.bbl}
+\input{vpd_et_paper_PG_v1.bbl}
 

### 2018-01-19 doc/paper/vpd_et_paper_PG_v1.tex
good progress

@@
 This then means that the sign term only depends on VPD for a given PFT and is approximately just a function of $VPD$.
-We can further determine a critical threshold operating increase from decrease in ET, such that $VPD_{crit}$ where $\frac{\partial \; ET}{\partial \; VPD} = 0$:
+We can further determine a critical threshold separating an increase from decrease in ET, such that $VPD_{crit}$ where $\frac{\partial \; ET}{\partial \; VPD} = 0$:
 
@@
 ENF ($g_1 = 74.31$) has by far the largest VPD dependence of response, while CRO ($g_1 = 183.1$) has the smallest VPD dependence.
-This means ENF's stomatal response evolved to change substantially depending on the envionmental dryness.
-This contasts\explain{think about this - what does it mean - go back to Franks and think about g1's relation to wue}
+ENF are less willing to trade water for new growth, so the regulate water much more strongly depending on the dryness environment (VPD).
+CRO on the otherhand are more willing to trade water for new growth, so they regulate water relatively similiarly irregardless of the environment.
 
@@
 To better illustrate this, the ranges of observed environmental VPDs at the FLUXNET sites are plotted parallel to the x-axis.
-ALL OF THIS SHOULD BE IN CAPTION NOT IN MAIN TEXT: (25th, 50th, and 75th percentiles are marked by stars),
-The location of VPD$_{crit}$ is marked by a vertical line.
 For CSH, VPD is always less than VPD$_{crit}$ so that the plant response dominates and empathizes the water conservative strategy of those plants.
@@
 The sign term in this case would just be a constant ($\frac{c_p}{R_{air}} \approx 3.5$), which is far larger than any part of the curves for any PFT.
-This highlights the deficiencies of PET's ability to capture land response to changes in atmospheric dryness (VPD) SEEMS TOO GENERIC BETTER EXPLAIN.
+This highlights the deficiencies of PET's ability to capture land response to changes in atmospheric dryness (VPD).
+Plants are always regulating water exchange from the land surface, even when they reach the limits of they ability to do so.
+Therefore, the actual land surface reponse to a change in VPD will always me more negative than if the plants are not included in the physics (PET).
 However, we have assumed $\sigma$ (uncertainty) equal to one.
@@
 It is clear that the portion of VPD observations below/above these curves will be a strong function of $\sigma$.
-However, we can see some general trends.
-For CSH, $\frac{\partial \; ET}{\partial \; VPD}$ should be negative for the vast majority of observed $\sigma$ and VPD.
-The fraction of positive $\frac{\partial \; ET}{\partial \; VPD}$ appears to be more even for ENF, GRA, and DBF, and we might expect a greater frequency of positive $\frac{\partial \; ET}{\partial \; VPD}$ for CRO.
-Section \ref{testing} further examines the consistency of our theory, given  the apparent sensitivity to uncertainty ($\sigma$).
-THIS PARAGRAPH DOES NOT SEEM CONCLUSIVE AND MORE A TRAIN OF THOUGHT.
-LAY DOWN EXACTLY WHAT YOU HAVE IN MIND.
+This motivates a thorough examination of how inclusion of uncertainty ($\sigma$) might change the form of the sign term, and in particular the location of VPD$_{crit)$ for each of our plant types.
+In Secion \ref{testing} we examine if and how uncertainty changes the analysis presented in this section, for each PFT.
 
@@
 This is as expected, given that the aerodynamic conductance represents the efficiency fo the transfer of surface anomalies to the atmosphere.
-As it increases I THINK YOU REALLY MEAN CONDUCATANCE HERE - BE CAREFUL, any plant response will be communicated more strongly to the atmosphere (and vice-versa).
-
-REMOVE ANY SNETENCES LIKE THIS MAKES SENSES ETC...
+As aerodynamic conductace  increases, any plant response will be communicated more strongly to the atmosphere (and vice-versa).
 
@@
 Theoretical derivation has so far illuminated the dependence of $\frac{\partial \; ET}{\partial \; VPD}$ and how it varies across PFT.
-Large mean $uWUE$ shifts CRO and DBF towards positive $\frac{\partial \; ET}{\partial \; VPD}$ i.e.
-increasing ET at higher VPD, but $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ is still negative for DBF due to its lower $g_1$ SAY PHISICALLY WHAT IT IS FOR THE READER TO UNDERSTAND.
-ENF's low $g_1$, which is even lower than DBF, increases the dependence of $\frac{\partial \; ET}{\partial \; VPD}$ on $VPD$, and makes the function strongly nonlinear.
-This has the side effect of pushing $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ negative further than other PFTs for a given fraction $\frac{\partial \; ET}{\partial \; VPD} < 0$ and magnitude $\frac{\partial \; ET}{\partial \; VPD}(\overline{T,\ldots,VPD})$.
-SAY THIS IN WORDS AS WELL OR YOU WILL LSOE THE READER.
-GRA shows the opposite behavior; a relatively high $g_1$  makes the function more linear, decreasing the magnitude of $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ for a given fraction $\frac{\partial \; ET}{\partial \; VPD} < 0$ and negative $\frac{\partial \; ET}{\partial \; VPD}(\overline{T,\ldots,VPD})$.
-SAY IN WORDS AGAIN WHAT IT MEANS OTHERWISE PEOPLE WILL NOT UNDERSTAND OUR RESULTS.
-Finally, the low $uWUE$ of CSH pushes it toward the lowest values of $\frac{\partial \; ET}{\partial \; VPD}$ (Figure \ref{term3}) across all PFTs WHAT DOES IT MEAN AGAIN.
+Large mean $uWUE$ shifts CRO and DBF towards a tendancy to increase ET with VPD (positive $\frac{\partial \; ET}{\partial \; VPD}$), but a lower slope parameter ($g_1$) for DBF allows for decreaes in ET with VPD, at low environmental VPD.
+\explain{physically state how slope parameter is trade for water at leaf level, but uWUE is trade-ff between GPP and VPD at ecosystem-level}.
+ENF's very low slope parameter ($g_1$), which is even lower than for DBF, increases the dependence of ET response  ($\frac{\partial \; ET}{\partial \; VPD}$) on $VPD$, and makes the function strongly nonlinear.
+This has the effect of making mean ET response ($\overline{\frac{\partial \; ET}{\partial \; VPD}}$)  more negative for a given frequency of occurence of negative ET response ($\frac{\partial \; ET}{\partial \; VPD} < 0$), than for other PFT.
+GRA shows the opposite behavior; a relatively high slope parameter ($g_1$)  makes the function more linear, causing a more positive mean ET reponse ($\overline{\frac{\partial \; ET}{\partial \; VPD}}$) for a given given frequence of negative ET reponse  ($\frac{\partial \; ET}{\partial \; VPD} < 0$).
+Finally, CSH's very low $uWUE$ dominates the slope parameter's effects, causing ET to almost always decrees with increasing atmospheric dryness (VPD), even with uncertainty included.
 Variability in $VPD$ accounts for the largest amount of $ET$ variability for CSH.
 For the other PFTs, $VPD$ contributes less to $ET$ variability, but still represents about 30-40 \% of $R$'s contributions to ET variability.
-WHAT IS THE SORUCE OF VARIBILITY OF THE REST OF ET I DON"T KNOW IF THAT IS RELEVANT FOR THE PRESENT PAPER.
+\explain{think pierre is correct, could probably scrap this anecdote: WHAT IS THE SORUCE OF VARIBILITY OF THE REST OF ET I DON"T KNOW IF THAT IS RELEVANT FOR THE PRESENT PAPER.}
 
@@
 So far we have developed a theory for $\frac{\partial \; ET}{\partial \; VPD}$'s behavior as a function of VPD and PFTs.
-We now turn to an direct evaluation against observations taken from a wide range of eddy-covariance datasets (SEE METHODS).
+We now turn to an direct evaluation against observations taken from a wide range of eddy-covariance datasets (Section \ref{methods}).
 

### 2018-01-19 doc/paper/vpd_et_paper_PG_v1.tex
done with morn writing

@@
 
-ALL OF THE REST SHOULDB EB EXPLAINED IN MUCH MORE DETAIL DESCRIBE EACH FIGURE AND THE RESULTS For DBF and CSH our theory is successful towards this goal: the theory matches the leading order behavior of the function when uncertainty is included.
-The plots for DBF and CSH look as one would expect if one just added noise on top of our theory.
+For DBF our theory is successful towards this goal: the theory matches the leading order behavior of the function when uncertainty is included.
+The plots for DBF look as one would expect if one just added noise on top of our theory.
 
-The theory does not match as well for CRO and GRA.
+For CSH our theory is successful towards this goal: the theory matches the leading order behavior of the function when uncertainty is included.
+The plots for CSH look as one would expect if one just added noise on top of our theory.
+
+The theory does not match as well for CRO .
 Perhaps this should be expected.
-Our theory, for practical reasons, does not include important sources of variability (particularly between sites) for CRO and GRA.
+Our theory, for practical reasons, does not include important sources of variability (particularly between sites) for CRO.
 We do not account for seasonal changes in LAI, variability in crop height and surface roughness, differences in C3 vs C4 photosynthesis, or differences in irrigation practices between the sites.
-Given that the our theory does not do as well for CRO and GRA, we will refrain from firm statements about how CRO and GRA respond to changes in VPD.
+Given that the our theory does not do as well for CRO, we will refrain from firm statements about how CRO respond to changes in VPD.
+
+The theory does not match as well for GRA.
+Perhaps this should be expected.
+Our theory, for practical reasons, does not include important sources of variability (particularly between sites) for GRA.
+We do not account for seasonal changes in LAI, variability in crop height and surface roughness, differences in C3 vs C4 photosynthesis, or differences in irrigation practices between the sites.
+Given that the our theory does not do as well for GRA, we will refrain from firm statements about how  GRA respond to changes in VPD.
 
@@
 Again, as in Section \ref{sign_func}, the PET response to changes in VPD will be constant at about 3.5.
-Even with uncertainty included, no values of the response approach the PET case YOU NEED TO ELABORATE WAY MORE AND GUDE THE READER THROUGH YOUR THOUGHTS.
-PET's status outside the error bars of our theory highlight the usefulness of our theory (and Equation \ref{et}) for estimating land response and moisture budgeting.
+In Figure \ref{agu_fig} the ET reponse to shifts in dryness is always far below (more negative) 3.5 for observations, including uncertainty.
+So the PET reponse to atmospheric dryness is outside the range of all observable responses.
+This casts doubt on PET's usefulness for capturing the leading order behavior of water budgeting at the land surface.
+However our theory (Equation \ref{et}) could be used with published $uWUE$ and $g1$ as an alternative to PET to estimating land response and moisture budgeting.
 
@@
 For any observations to the left of this line $\frac{\partial \; ET}{\partial \; VPD}$ should be negative (plant response dominates), and any observations the right of the line should be positive (atmospheric demand dominates).
-In addition to seeing how our theoretically-derived VPD threshold, $VPD_{crit}$, matches observations, these analysis DO NOT SAY PLOT OR IT SOUNDS LIKE A MSC PROJECT are also useful for observing how the scaling term affects the magnitude of the derivative, if at all.
+In addition to seeing how our theoretically-derived VPD threshold, $VPD_{crit}$, matches observations, these analysis are also useful for observing how the scaling term affects the magnitude of the derivative, if at all.
 
-The scatter plots generally confirm expectations from Figure \ref{agu_fig}.
+The scatter figures generally confirm expectations from Figure \ref{agu_fig}.
 By comparing different columns of Figure \ref{real} we can evaluate the impact of $\sigma$ and $g_a$ on the variability of $\frac{\partial \; ET}{\partial \; VPD}$.
@@
 
-HERE DISCUSS LIMITATION OF THE METHODS TO BVERY DRY CONDITIONS AND THE FEEDBCAK TO THE ATMOSPHERE OF VERY DRY SOIL CONDITIONS (BERG ET AL 2016, GENTINE ET AL> 2016)
+\section{Limitations of theory}
+\explain{HERE DISCUSS LIMITATION OF THE METHODS TO BVERY DRY CONDITIONS AND THE FEEDBCAK TO THE ATMOSPHERE OF VERY DRY SOIL CONDITIONS (BERG ET AL 2016, GENTINE ET AL> 2016)}
 

### 2018-01-22 doc/paper/vpd_et_paper_PG_v1.tex
half way done with morn writing, takinga  break and will come back

@@
 It is clear that the portion of VPD observations below/above these curves will be a strong function of $\sigma$.
-This motivates a thorough examination of how inclusion of uncertainty ($\sigma$) might change the form of the sign term, and in particular the location of VPD$_{crit)$ for each of our plant types.
+This motivates a thorough examination of how inclusion of uncertainty ($\sigma$) might change the form of the sign term, and in particular the location of VPD$_{crit}$ for each of our plant types.
 In Secion \ref{testing} we examine if and how uncertainty changes the analysis presented in this section, for each PFT.
@@
 
-For DBF our theory is successful towards this goal: the theory matches the leading order behavior of the function when uncertainty is included.
-The plots for DBF look as one would expect if one just added noise on top of our theory.
+For DBF our theory is successful towards this goal: the theory matches the leading order behavior of the function when uncertainty is included, and the DBG figure looks as one would expect if one just added noise on top of our theory.
+The VPD dependence, given by the slope parameter ($g_1$), floows the median values of each box plot.
+Perhaps most importantly, the x-intercept, and thus VPD$_{crit}$, matches nearly exactly between the theory and the observations.
+Therefor the sign of the ET response to increases in VPD should be well matched, subject to the unavoidable constraints of noise, much of which may be from the observations themselves.
+The uncertainty is non negligable; there are bany observations in each box for which the the sign of observation is opposite the response predicted by the theory, but to leading order our theory matches the observations.
 
-For CSH our theory is successful towards this goal: the theory matches the leading order behavior of the function when uncertainty is included.
-The plots for CSH look as one would expect if one just added noise on top of our theory.
+While CSH has a much different functional form of the sign term than DBF, CSH observations also match our theory to leading order.
+Again, the VPD dependence mostly determined from the slope parameter ($g_1$) closely matches the medians in the observation bins.
+The VPD-independent, strongly negative response, which is determined by $uWUE$, is also captured.
+For CSH, there is extremely rare occurence of positive ET response with VPD, even with uncertainty, so the sign of the theory almost always matches the sign of the observations, even with all of the uncertainty.
 
-The theory does not match as well for CRO .
-Perhaps this should be expected.
-Our theory, for practical reasons, does not include important sources of variability (particularly between sites) for CRO.
-We do not account for seasonal changes in LAI, variability in crop height and surface roughness, differences in C3 vs C4 photosynthesis, or differences in irrigation practices between the sites.
-Given that the our theory does not do as well for CRO, we will refrain from firm statements about how CRO respond to changes in VPD.
+The theory does not match as well for CRO as for DBF and CSH.
+At low VPD the VPD dependence (due to $g_1$) is similiar to the leading order behavior of the observations, but the entire curve is shifted towards postiive values, due to some error in $uWUE$.
+Most strikingly, the observations show a tendency to move back towards negative values at high VPD, which given the functional form of VPD dependnce, is behavior that is impossible fo the theory to capture.
+This divergence between the theory and the observations for CRO could have been expected.
+For practical reasons, our theory does not include important sources of variability (particularly between sites) for CRO.
+We do not account for seasonal changes in LAI, variability in crop height and surface roughness, differences in C3 vs C4 photosynthesis, or differences in irrigation practices between the sites and/or years.
+This deficiencies would explain the inability of our theory to match obsercations.
+For example, a superpositions of sites with C3 plants (low VPD) and C4 plants (high VPD) would explain observed shift back towards negative ET reponse at high VPD when all sites are binned together, as in Figure \ref{agu_fig}.
 
-The theory does not match as well for GRA.
-Perhaps this should be expected.
-Our theory, for practical reasons, does not include important sources of variability (particularly between sites) for GRA.
-We do not account for seasonal changes in LAI, variability in crop height and surface roughness, differences in C3 vs C4 photosynthesis, or differences in irrigation practices between the sites.
-Given that the our theory does not do as well for GRA, we will refrain from firm statements about how  GRA respond to changes in VPD.
+The theory for GRA suffers from similiar limitations as for CRO.
+The differences between the theory and the observations are more pronounced for GRA, but are explainable using similiar logic as that used for CRO.
+GRA will also have differences between sites in C3 vs C4 photosythesis, and irrigation practices, as well as strong seasonal variability in crop height and LAI.
+For GRA, the observations show a very different VPD dependence of response than the theory, with a strong tendency back toewards a negative ET reponse at high VPD which our theory cannot account for.
+As with CRO, this could be explained by a superposition of sites with C3 vs C4 sites.
+Unfortunately, because detailed site-level data of C3 vs C4 plant type, as well as irrigation practes, was difficult to obtain, we could not bin GRO and GRA further by irrigation and photosyethesis stragtegy.
+However, we hypothesize that the theory would much more closel match obsercations with photosythesis and irrigation are accounted for.
 
-ENF is a special case.
-The VPD dependence of the theory is not well captured by observations.
-However, the theory closely captures the intercept where $\frac{\partial \; ET}{\partial \; VPD} = 0$.
-Resolving this feature is the key to answering our question, so perhaps the theory is still useful for ENF for the purpose of addressing when VPD drives or reduces ET.
-It is also noteworthy that we used published constants for $g_1$, which determines the VPD dependence, from \cite{Lin_2015}.
+The comparison between the theory and obsercations for ENF is unique from the rest of the PFTs.
+The VPD dependence of the theory, determined by the slope paramter, is not well captured by observations.
+However, the theory closely captures the x-intercept and VPD$_{crit}$.
+So, the theory still captures when the ET reponse will be positive or negative, but is less accurate with the magnitude of the reponse.
+Therefor the theory is still useful for addressing when VPD drives or reduces ET.
+Also, it's with noting that we used published constants for $g_1$ \citep[from ][]{Lin_2015}, which determines the VPD dependence.
 We could have fit our own constants so that the theory would best match the observations, but our goal was to use a priori physical constants whenever possible.
 
-While the above discussion shows that our theory has some limitations when applied to some sites, especially for crops and grasslands, another question could be, how does our theory's performance compare to often-used PET?
+While the above discussion shows that our theory has some limitations when applied to some sites, especially for crops and grasslands, another question could be, how would using PET, as many hydrometeorological studies and applications do, compare with our theory.
 Again, as in Section \ref{sign_func}, the PET response to changes in VPD will be constant at about 3.5.
-In Figure \ref{agu_fig} the ET reponse to shifts in dryness is always far below (more negative) 3.5 for observations, including uncertainty.
+In Figure \ref{agu_fig} the ET reponse to shifts in dryness is always far more negative than 3.5 for the observations, even with all uncertainty included.
 So the PET reponse to atmospheric dryness is outside the range of all observable responses.

### 2018-01-22 doc/paper/vpd_et_paper_PG_v1.tex
through methods, now onto results for reading thogh

@@
 
-%% ------------------------------------------------------------------------ %%
-%  Title
-% 
-% (A title should be specific, informative, and brief. Use
-% abbreviations only if they are defined in the abstract. Titles that
-% start with general keywords then specific terms are optimized in
-% searches)
-%
-%% ------------------------------------------------------------------------ %%
-
-% Example: \title{This is a test title}
-
 \title{When does vapor pressure deficit drive or reduce evapotranspiration?}
 
-%% ------------------------------------------------------------------------ %%
-%
-%  AUTHORS AND AFFILIATIONS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Authors are individuals who have significantly contributed to the
-% research and preparation of the article. Group authors are allowed, if
-% each author in the group is separately identified in an appendix.)
-
-% List authors by first name or initial followed by last name and
-% separated by commas. Use \affil{} to number affiliations, and
-% \thanks{} for author notes.  
-% Additional author notes should be indicated with \thanks{} (for
-% example, for current addresses).
-
-% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
-% Author\affil{3,4}\thanks{Also funded by Monsanto.}}
-
 \authors{A. Massmann\affil{1}, P. Gentine\affil{1}, C. Lin\affil{1,2}}
-
-% \affiliation{1}{First Affiliation}
-% \affiliation{2}{Second Affiliation}
-% \affiliation{3}{Third Affiliation}
-% \affiliation{4}{Fourth Affiliation}
 
@@
 
-  % (repeat as many times as is necessary)
-
-%% Corresponding Author:
-% Corresponding author mailing address and e-mail address:
-
-% (include name and email addresses of the corresponding author.  More
-% than one corresponding author is allowed in this LaTeX file and for
-% publication; but only one corresponding author is allowed in our
-% editorial system.)
-
-% Example: \correspondingauthor{First and Last Name}{email@address.edu}
-
 \correspondingauthor{Adam Massmann}{akm2203@columbia.edu}
-
-%% Keypoints, final entry on title page.
-
-% Example: 
-% \begin{keypoints}
-% \item	List up to three key points (at least one is required)
-% \item	Key Points summarize the main points and conclusions of the article
-% \item	Each must be 100 characters or less with no special characters or punctuation 
-% \end{keypoints}
-
-%  List up to three key points (at least one is required)
-%  Key Points summarize the main points and conclusions of the article
-%  Each must be 100 characters or less with no special characters or punctuation
 
@@
 
-%% ------------------------------------------------------------------------ %%
-%
-%  ABSTRACT
-%
-% A good abstract will begin with a short description of the problem
-% being addressed, briefly describe the new data or analyses, then
-% briefly states the main conclusion(s) and how they are supported and
-% uncertainties. 
-%% ------------------------------------------------------------------------ %%
-
-%% \begin{abstract} starts the second page
-
 \begin{abstract}
@@
 
-%% ------------------------------------------------------------------------ %%
-%
-%  TEXT
-%
-%% ------------------------------------------------------------------------ %%
-
-%%% Suggested section heads:
-% \section{Introduction}
-% 
-% The main text should start with an introduction. Except for short
-% manuscripts (such as comments and replies), the text should be divided
-% into sections, each with its own heading.
-
-% Headings should be sentence fragments and do not begin with a
-% lowercase letter or number. Examples of good headings are:
-
-% \section{Materials and Methods}
-% Here is text on Materials and Methods.
-%
-% \subsection{A descriptive heading about methods}
-% More about Methods.
-% 
-% \section{Data} (Or section title might be a descriptive heading about data)
-% 
-% \section{Results} (Or section title might be a descriptive heading about the
-% results)
-% 
-% \section{Conclusions}
-
 \section{Introduction}
 
-Vapor pressure deficit (VPD) is expected to rise over continents in the future due to the combination of increased temperature and relative humidity drying\explain{Is increased RH conclusive and universal?- I'm not so sure...
+Vapor pressure deficit (VPD) is expected to rise over continents in the future due to the combination of increased temperature and relative humidity drying\explain{Is decreased RH conclusive and universal?- I'm not so sure...
 is there another citation?
-I don't think Byrne 2013 shows that.} \citep{Byrne_2013}..
-In turn VPD modifies the atmospheric demand of evapotranspiration (ET) (Penman Monteith) but is also a stressor for stomata \citep{Leuning_1990, MEDLYN_2011}.
+I don't think Byrne 2013 shows that.} \citep{Byrne_2013}.
+Increases in VPD increase the atmospheric demand of evapotranspiration (ET) \citep{Monteith_1965} but also stress plant stomata \citep{Leuning_1990, MEDLYN_2011}.
 
-Answering the question ``When does vapor pressure deficit drive or reduce evapotranspiration?'' is thus motivated by two potential but opposing perspectives on the matter.
-The first, hydrometeorological, perspective is that higher vapor pressure deficit increases atmospheric demand for water from the land surface, and this drives an increase in evapotranspiration (ET).
+This leades to two potential but opposing perspectives on how ET adjusts to changes in VPD.
+The first, a hydrometeorological perspective, is that higher VPD increases atmospheric demand for water from the land surface, and this drives an increase in evapotranspiration (ET).
 This perspective is particularly relevant because potential evapotranspiration (PET), which is used in many drought indices and hydrometeorological studies, only quantifies atmospheric demand effects.
@@
 Plants' stomata have evolved to optimally regulate the exchange of water and carbon, and tend to close in response to increased atmospheric dryness \citep{Ball_1987, Leuning_1990, MEDLYN_2011}.
-Therefore, an increase in VPD, in well watered soil conditions, may actually correspond to a decrease in ET because of stomatal closure.
-This decrease in ET would reduce and potentially cancel (in case of full closure) the effects of shifts to atmospheric demand.
-In other words, the  question ``When does VPD drive or reduce ET?'' can be related to whether plant regulation or atmospheric demand dominate ET.
+Therefore, an increase in VPD, particularly in well watered soil conditions, may actually correspond to a decrease in ET because of stomatal closure.
+This decrease in ET would reduce and potentially cancel (in the case of full stomatal closure) the effects of shifts to atmospheric demand.
+In other words, the  question ``When does VPD drive or reduce ET?'' can be related to whether plant regulation or atmospheric demand dominates ET.
+
+The ET reponse to changes in VPD alters water partitioning between the soil and atmosphere.
 If plant response reduces ET in response to atmospheric drying then soil moisture will be better conserved.
@@
 
-First, whether VPD drives or reduces ET should be a function of plant type.
+We can use intuition about plant water conservation strategy to hypothesize about ET reponse to changes in VPD.
 Plants that evolved to conserve water (e.g.
@@
 Atmospheric conditions must matter as well.
-At the ecosystem scale there are limits to the strategies plants use to hold water back from the atmosphere.
+At the ecosystem scale there are limits to plant water conservation strategies.
 As atmospheric demand for water (VPD) increases, ecosystems will begin to reach their water conservation limits.
@@
 Each site's plant functional type (PFT) was classified using the International Geosphere-Biosphere Programme vegetation classification scheme \citep{Loveland_1999}.
-The physical constants used in Section \ref{methods} are only published for five plant functional types (PFTs) crops (CRO), deciduous broadleaf forest (DBF), evergreen needleleaf forest (ENF), grass (GRA), and closed shrub (CSH) (see Table \ref{pft}).
+The physical constants used in Section \ref{methods} are only published for five plant functional types (PFTs): crops (CRO), deciduous broadleaf forest (DBF), evergreen needleleaf forest (ENF), grass (GRA), and closed shrub (CSH) (see Table \ref{pft}).
 There are 56 sites with these plant functional types, and their location is shown in  Figure \ref{map_fig} \explain{map needs to be improved - it's a placeholder for now}.
@@
 \label{methods}
-The Penman-Monteith equation (hereafter PM \explain{find pm ciatation in shuttlewroth}) estimates ET as a function of observable atmospheric variables and surface conductances:
+The Penman-Monteith equation \citep [hereafter PM][]{Monteith_1965}} estimates ET as a function of observable atmospheric variables and surface conductances:
 \begin{linenomath*}
@@
 \end{linenomath*}
-where variable definitions are given in Table 1.
+where $\Delta$ is the change in saturation vapor pressure with temperature ($\frac{\partial \; e_s}{\partial \; T}$), $R$ is the net radiation minus ground heat flux, $g_a$ is aerodynamic conductance, $\rho_a$ is air density, $c_p$ is specific heat of air at constant pressure, $\gamma$ is the psychometric constant, and $g_s$ is the stomatal conudctance (Table 1).
 \citet{MEDLYN_2011} developed a model for stomatal conductance ($g_s$) by combining optimal photosynthesis theory \citep{Farquhar_1980, Katul_2010} with empirical approaches, which describes the dependence of $g_s$ to VPD.
@@
 \end{linenomath*}
-This model has been shown to behave very well across PFTs, compared to other models \citep{Lin_2015}.
+Where $g_1$ is a leaf-scale ``slope'' parameter, A is the net CO$_2$ assimilation rate, and $c_s$ is the atmospheric CO$_2$ concentration at the leaf surface.
+\cite{MEDLYN_2011} relate the slope parameter ($g_1$) to physical parameters as:
+\begin{linenomath*}
+  \label{slope}
+  \begin{equation}
+  g_1 = \sqrt{\frac{3 \, \Gamma \, \lambda}{1.6}},\footnote{Note the above expression has units of of (mmol mol$^{-1}$)$^{1/2}$, but this can be converted to Pa$^{1/2}$ using the ideal gas law.}
+  \end{equation}
+\end{linenomath*}
 
-This can be adapted to the ecosystem scale by multiplying by leaf area index (LAI) and converting units to m s$^{-1}$:
+where $\Gamma$ is the CO$_2$ compensation point for photosythesis (without dark respiration), and $\lamda$ is the marinal cost of leaf carbon.
+So, $g_1$ is a leaf-scale term reflecting the willingness of a leaf to trade water for carbon uptake and growth.
+The higher $g_1$, the more willing a leaf will be willing to release water in the exchange for water.
+
+The Medlyn model for stomatal conductance has been shown to behave very well across PFTs, compared to other models \citep{Lin_2015}, and can be adapted to the ecosystem scale by multiplying by leaf area index (LAI) and converting units to m s$^{-1}$:
 
@@
 
+With Equation \ref{et}, ET is explicitly a function of environmental variables and two plant-specific constats, the slope parameter ($g_1$), and $uWUE$.
+The slope parameter is a leaf-scale term related to the willingness of leaves to trade water for new growth.
+$uWUE$ is semi-empirical ecosystem-scale constant related to how WUE changes with VPD (specifically $\sqrt{VPD}$).
 Given FLUXNET data described in Section \ref{data}, every term in Equation \ref{et} is known.
@@
 
-%%
-%% Enter Figures and Tables near as possible to where they are first mentioned:
-%
-% DO NOT USE \psfrag or \subfigure commands.
-%
-% Figure captions go below the figure.
-% Table titles go above tables;  other caption information
-%  should be placed in last line of the table, using
-% \multicolumn2l{$^a$ This is a table note.}
-%
-%----------------
-% EXAMPLE FIGURE
-%
-% \begin{figure}[h]
-% \centering
-% when using pdflatex, use pdf file:
-% \includegraphics[width=20pc]{figsamp.pdf}
-%
-% when using dvips, use .eps file:
-% \includegraphics[width=20pc]{figsamp.eps}
-%
-% \caption{Short caption}
-% \label{figone}
-%  \end{figure}
-%
-% ---------------
-% EXAMPLE TABLE
-%
-% \begin{table}
-% \caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
-% \centering
-% \begin{tabular}{l c}
-% \hline
-%  Run  & Time (min)  \\
-% \hline
-%   $l1$  & 260   \\
-%   $l2$  & 300   \\
-%   $l3$  & 340   \\
-%   $h1$  & 270   \\
-%   $h2$  & 250   \\
-%   $h3$  & 380   \\
-%   $r1$  & 370   \\
-%   $r2$  & 390   \\
-% \hline
-% \multicolumn{2}{l}{$^{a}$Footnote text here.}
-% \end{tabular}
-% \end{table}
-
-%% SIDEWAYS FIGURE and TABLE 
-% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
-%
-% \begin{sidewaysfigure}
-% \includegraphics[width=20pc]{figsamp}
-% \caption{caption here}
-% \label{newfig}
-% \end{sidewaysfigure}
-% 
-%  \begin{sidewaystable}
-%  \caption{Caption here}
-% \label{tab:signif_gap_clos}
-%  \begin{tabular}{ccc}
-% one&two&three\\
-% four&five&six
-%  \end{tabular}
-%  \end{sidewaystable}
-
-%% If using numbered lines, please surround equations with \begin{linenomath*}...\end{linenomath*}
-%\begin{linenomath*}
-%\begin{equation}
-%y|{f} \sim g(m, \sigma),
-%\end{equation}
-%\end{linenomath*}
-
-%%% End of body of article
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Optional Appendix goes here
-%
-% The \appendix command resets counters and redefines section heads
-%
-% After typing \appendix
-%
-%\section{Here Is Appendix Title}
-% will show
-% A: Here Is Appendix Title
-%
-%\appendix
-%\section{Here is a sample appendix}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Optional Glossary, Notation or Acronym section goes here:
-%
-%%%%%%%%%%%%%%  
-% Glossary is only allowed in Reviews of Geophysics
-%  \begin{glossary}
-%  \term{Term}
-%   Term Definition here
-%  \term{Term}
-%   Term Definition here
-%  \term{Term}
-%   Term Definition here
-%  \end{glossary}
-
-%
-%%%%%%%%%%%%%%
-% Acronyms
-%   \begin{acronyms}
-%   \acro{Acronym}
-%   Definition here
-%   \acro{EMOS}
-%   Ensemble model output statistics 
-%   \acro{ECMWF}
-%   Centre for Medium-Range Weather Forecasts
-%   \end{acronyms}
-
-%
-%%%%%%%%%%%%%%
-% Notation 
-%   \begin{notation}
-%   \notation{$a+b$} Notation Definition here
-%   \notation{$e=mc^2$} 
-%   Equation in German-born physicist Albert Einstein's theory of special
-%  relativity that showed that the increased relativistic mass ($m$) of a
-%  body comes from the energy of motion of the body—that is, its kinetic
-%  energy ($E$)—divided by the speed of light squared ($c^2$).
-%   \end{notation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%  ACKNOWLEDGMENTS
-%
-% The acknowledgments must list:
-%
-% •	All funding sources related to this work from all authors
-%
-% •	Any real or perceived financial conflicts of interests for any
-%	author
-%
-% •	Other affiliations for any author that may be perceived as
-% 	having a conflict of interest with respect to the results of this
-% 	paper.
-%
-% •	A statement that indicates to the reader where the data
-% 	supporting the conclusions can be obtained (for example, in the
-% 	references, tables, supporting information, and other databases).
-%
-% It is also the appropriate place to thank colleagues and other contributors. 
-% AGU does not normally allow dedications.
-
 \acknowledgments
@@
 
-%% ------------------------------------------------------------------------ %%
-%% Citations
-
-% Please use ONLY \citet and \citep for reference citations.
-% DO NOT use other cite commands (e.g., \cite, \citeyear, \nocite, \citealp, etc.).
-
-%% Example \citet and \citep:
-%  ...as shown by \citet{Boug10}, \citet{Buiz07}, \citet{Fra10},
-%  \citet{Ghel00}, and \citet{Leit74}.
-
-%  ...as shown by \citep{Boug10}, \citep{Buiz07}, \citep{Fra10},
-%  \citep{Ghel00, Leit74}.
-
-%  ...has been shown \citep [e.g.,][]{Boug10,Buiz07,Fra10}.
-
-%%  REFERENCE LIST AND TEXT CITATIONS
-%
-% Either type in your references using
-%
-% \begin{thebibliography}{}
-% \bibitem[{\textit{Kobayashi et~al.}}(2003)]{R2013} Kobayashi, T.,
-% Tran, A.~H., Nishijo, H., Ono, T., and Matsumoto, G.  (2003).
-% Contribution of hippocampal place cell activity to learning and
-% formation of goal-directed navigation in rats. \textit{Neuroscience}
-% 117, 1025--1035.
-%
-% \bibitem{}
-% Text
-% \end{thebibliography}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Or, to use BibTeX:
-%
-% Follow these steps
-%
-% 1. Type in \bibliography{<name of your .bib file>} 
-%    Run LaTeX on your LaTeX file.
-%
-% 2. Run BiBTeX on your LaTeX file.
-%
-% 3. Open the new .bbl file containing the reference list and
-%   copy all the contents into your LaTeX file here.
-%
-% 4. Run LaTeX on your new file which will produce the citations.
-%
 % AGU does not want a .bib or a .bbl file. Please copy in the contents of your .bbl file here.

### 2018-01-23 doc/paper/vpd_et_paper_PG_v1.tex
done with this morn writing, drying to conceptualize wh g1 and uwue have some of the sme effects (and some different)

@@
 \label{methods}
-The Penman-Monteith equation \citep [hereafter PM][]{Monteith_1965}} estimates ET as a function of observable atmospheric variables and surface conductances:
+The Penman-Monteith equation \citep [hereafter PM][]{Monteith_1965} estimates ET as a function of observable atmospheric variables and surface conductances:
 \begin{linenomath*}
@@
 
-where $\Gamma$ is the CO$_2$ compensation point for photosythesis (without dark respiration), and $\lamda$ is the marinal cost of leaf carbon.
+where $\Gamma$ is the CO$_2$ compensation point for photosythesis (without dark respiration), and $\lambda$ is the marginal cost of leaf carbon.
 So, $g_1$ is a leaf-scale term reflecting the willingness of a leaf to trade water for carbon uptake and growth.
@@
 The slope parameter is a leaf-scale term related to the willingness of leaves to trade water for new growth.
-$uWUE$ is semi-empirical ecosystem-scale constant related to how WUE changes with VPD (specifically $\sqrt{VPD}$).
+$uWUE$ is semi-empirical ecosystem-scale constant related to how WUE changes with VPD (specifically $VPD^{-1/2}$).
 Given FLUXNET data described in Section \ref{data}, every term in Equation \ref{et} is known.
@@
 
-We note one final comment on our derivation which is relevant for future analysis.
+We note one final comment on our derivation which is relevant for future projects and analysis.
 If we approximate $c_s$ at a global mean CO$_2$ concentration, then the RHS of Equation \ref{et} is fully defined using commonly available weather station data and the constants published in \citet{Zhou_2015, Lin_2015}.
@@
 An additional concern is that $\sigma$ may in fact be correlated with $VPD$, in which case the dependence would need to be accounted for when taking the derivative.
-Figure \ref{lai_vpd_fig} plots the joint distribution of $\sigma$ and VPD, and shows that $\sigma$ is very weakly a function of VPD.
+However, the joint distribution of $\sigma$ and VPD shows that $\sigma$ is very weakly a function of VPD (Figure \ref{lai_vpd_fig}).
 Given this weak dependence, we argue that Equation \ref{d_et} is a valid approximation for ET response to $VPD$.
@@
 All variables are positive, so the relative magnitude between the first term and the second term in Equation \ref{sign} will determine whether ET increases or decreases with increasing VPD.
-If the second term is larger magnitude then plant response dominates, while if the first term is larger then atmospheric demand dominates.
-The sign will determine whether ET is increasing or decreasing and we thus evaluate its functional form first.
+If the second term is larger magnitude then plant response dominates and ET decreases with increasing VPD.
+However, if the first term is larger then atmospheric demand dominates and ET increases with increasing VPD.
+Given this sign
 
@@
 First, we explore the variables within the sign term to gain better intuition on the driver of either the increase or reduction of ET with VPD.
-$c_s$ and $\gamma$ are relatively constant to a good approximation so that the variability is dominated by $\sigma$ and $VPD$.
+CO$_2$ concentration ($c_s$) and the pschometric constant ($\gamma$) are relatively constant to a good approximation so that the variability is dominated by $\sigma$ and $VPD$.
 $uWUE$ could vary with soil moisture but has been shown to be relatively constant except in very dry soil conditions \citep{Zhou_2016, Lin_2015}.
-We thus further assume that the soil is not too dry so that $uWUE$ can be approximated as a constant.
+We thus further assume that the soil is not too dry so that $uWUE$ can be approximated as a constant (within PFT).
 This then means that the sign term only depends on VPD for a given PFT and is approximately just a function of $VPD$.
@@
 Values of $VPD_{crit}$ as a function of PFT are shown in Table \ref{vpd_crit}.
-For any values of $VPD$ less than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be negative, and for values of $VPD$ greater than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be positive.
+For any values of $VPD$ less than $VPD_{crit}$, ET will decrease with increasing VPD ($\frac{\partial \; ET}{\partial \; VPD} < 0$), and for values of $VPD$ greater than $VPD_{crit}$, ET will increase with increasing VPD ($\frac{\partial \; ET}{\partial \; VPD} > 0$).
 In other words, ecosystems can regulate and mitigate evaporative losses up to a VPD limit, above which atmospheric demand is just too high to be entirely compensated by stomatal regulation.
-We note however that even though ET increases again above the critical threshold, $VPD_{crit}$, ET is still much lower that potential evaporation as stomata are stills strongly regulating vapor fluxes.
+We note however that even though ET increases again above the critical threshold, $VPD_{crit}$, ET is still much lower that potential evaporation as stomata are still strongly regulating vapor fluxes.
 Even in the absence of soil pore evaporation, stomata do not shut down entirely at very high VPD and ET does not go to zero, because stomata are still slightly open to perform some photosynthesis \citep{Ball_1987, Leuning_1990, Medlyn_2011}.
@@
 
-Differences in $uWUE$ and $g_1$ between PFTs alter the functional form of the sign term.
-Larger $uWUE$ and $g_1$ will shift the sign term towards overall positive values.
-$g_1$ additionally plays a primary role in determining dependnece on VPD: the smaller $g_1$, the greater the $VPD$ dependence for the PFT (Figure \ref{term3}).
+Differences in the sign term between plant types is exclusively determined by $uWUE$ and the slope parameter ($g_1$).
+A larger $uWUE$ means that at the ecosystem scale $WUE$ is larger for a given VPD, so plants are more willing to keep stomata open for new growth despite high VPD.
+Similiarly, high slope parameters ($g_1$) means that at the leaf scale plants are more willing to trade water for growth.
+Both these effects (large $uWUE$, $g_1$) conspire to shift the sign term towards overall positive values for a given PFT.
+
+This effect of $uWUE$ and $g1$ on the sign term is most apparent by comparing CRO to CSH.
+CRO has a high slope parameter ($g_1 = 183.1 \, Pa^{1/2}$) and uWUE ($2.32$ -units-) compared to CSH ($g1 = 148.6 \, Pa^{1/2}$, $uWUE=2.11$ -units-).
+These differences in slope parameter cause opposite ET reponses to changse in VPD.
+ET almost always increases with increasing VPD for CRO, while ET almost always decreases with increasing VPD for CSH (Figure \ref{term3}).
+CRO, evolved or bred to prioritize GPP, are very willing to trade water for new growth, despite changes in VPD, while CSH are very unwilling to trade water for new growth.
+
+The slope paramter ($g_1$) is additionally a primary determinant of VPD  dependnece for the sign term shown in Figure \ref{term3}. Plants that are less willing
 

### 2018-01-24 doc/paper/vpd_et_paper_PG_v1.tex
at one interesting sid enote

@@
 
-The ET reponse to changes in VPD alters water partitioning between the soil and atmosphere.
+The ET response to changes in VPD alters water partitioning between the soil and atmosphere.
 If plant response reduces ET in response to atmospheric drying then soil moisture will be better conserved.
@@
 Differences in the sign term between plant types is exclusively determined by $uWUE$ and the slope parameter ($g_1$).
-A larger $uWUE$ means that at the ecosystem scale $WUE$ is larger for a given VPD, so plants are more willing to keep stomata open for new growth despite high VPD.
-Similiarly, high slope parameters ($g_1$) means that at the leaf scale plants are more willing to trade water for growth.
-Both these effects (large $uWUE$, $g_1$) conspire to shift the sign term towards overall positive values for a given PFT.
+A larger $uWUE$ causes means that at the ecosystem scale $WUE$ decreases more with increasing VPD, so plants lose more water per unit carbon production and are more likely to increase ET with increasing VPD.
+A tendenacy towards increasing ET reponse with increasing VPD can also be caused by high slope parameters ($g_1$), characteristic of plants that at the leaf scale plants are more willing to trade water for growth.
+Plants that give water more freely will be much more likely to increase ET with increasing VPD.
+Both the aforementioned effects (large $uWUE$, $g_1$) can amplify each other, and generally conspire to shift the sign term towards overall positive values for a given PFT.
 
@@
 CRO has a high slope parameter ($g_1 = 183.1 \, Pa^{1/2}$) and uWUE ($2.32$ -units-) compared to CSH ($g1 = 148.6 \, Pa^{1/2}$, $uWUE=2.11$ -units-).
-These differences in slope parameter cause opposite ET reponses to changse in VPD.
+These differences in slope parameter cause opposite ET reponses to changse in VPD between CRO and CSH.
 ET almost always increases with increasing VPD for CRO, while ET almost always decreases with increasing VPD for CSH (Figure \ref{term3}).
@@
 
-The slope paramter ($g_1$) is additionally a primary determinant of VPD  dependnece for the sign term shown in Figure \ref{term3}. Plants that are less willing
+The slope paramter ($g_1$) is additionally a primary determinant of VPD  dependnece for the sign term shown in Figure \ref{term3}.
+Plants that are less willing to trade water for new growth (small $g_1$) will have strong tendency to reduce ET with increasing VPD, and they will be very effective in reducing ET especially at low VPD.
+However, at very high VPD gradients in vapor pressure at the leaf scale will begin to dominate leaf-scale plant strategies (parameterized with $g_1$), so ET reponse will begin to asymtote towards ecosystem-scale values irregardless of leaf-scale strategy (and $g_1$).
+Therefore, plants with a low $g_1$ will have the largest VPD dependence of ET response because the difference in ET response at low VPD (leaf response dominates) and high VPD (VPD gradient dominates) is largest.
+This is apparent in the strong VPD dependece of ENF, which has the lowest slope parameter ($g_1=74.31$), in Figure \ref{term3}.
 
-Based on Figure \ref{term3} and Table \ref{vpd_crit}, CROs are the least water conservative and have the most positive constant offset, while CSH are the most water conservative and have the most negative constant offset in the sign term.
+To summarize Figure \ref{term3} and Table \ref{vpd_crit}, CROs are the least water conservative and have the strongest overall tendency to increase ET wirh increasing VPD, while CSH are the most water conservative and have the strongest tendency to decrease ET with increasing VPD.
 ENF ($g_1 = 74.31$) has by far the largest VPD dependence of response, while CRO ($g_1 = 183.1$) has the smallest VPD dependence.
-ENF are less willing to trade water for new growth, so the regulate water much more strongly depending on the dryness environment (VPD).
-CRO on the otherhand are more willing to trade water for new growth, so they regulate water relatively similiarly irregardless of the environment.
+ENF are less willing to trade water for new growth at the leaf scale, so the difference between the response at low VPD when leaf response domineates and high VPD when vapor pressure gradients dominate is much greater.
+CRO on the otherhand are more willing to trade water for new growth, so the leaf scale response is much smaller, and the difference in ET response between low VPD and high VPD is relatively low.
 
-A primary takeaway from Figure \ref{term3}a is that according to our theory for all PFTs except for crops there is frequent occurrence of a negative (plant dominating) ET response to increases in VPD.
-This thus means that plants are able in most atmospheric conditions to redue ET in response to increased VPD and thus to reduce water losses.
+Figure \ref{term3} show clearly that, accoridng to our thoery, for all PFTs except for crops there is frequent occurrence of a negative (plant dominating) ET response to increases in VPD.
+Therefore, plants are able in most atmospheric conditions to redue ET in response to increased VPD and thus to reduce water loss.
 To better illustrate this, the ranges of observed environmental VPDs at the FLUXNET sites are plotted parallel to the x-axis.
 For CSH, VPD is always less than VPD$_{crit}$ so that the plant response dominates and empathizes the water conservative strategy of those plants.
-For CRO on the other, VPD is almost always higher than VPD$_{crit}$, emphasizing that those plants are water intensive and were actually engineered for photosynthesis rather than water saving.
+For CRO on the otherhand, VPD is almost always higher than VPD$_{crit}$, emphasizing that those plants are water intensive and were actually engineered for photosynthesis rather than water saving.
 For forests and grass, whether plant response or atmospheric demand dominates is more dependent on the environment, with about half of of observed VPD less than VPD$_{crit}$, i.e.
@@
 Therefore, the actual land surface reponse to a change in VPD will always me more negative than if the plants are not included in the physics (PET).
-However, we have assumed $\sigma$ (uncertainty) equal to one.
-We still need to test if inclusion of uncertainty could change our conclusions.
 \begin{figure}[h]
@@
 Lines and stars denote the distribution of VPD and $\sigma$ next to each axis, following the same percentiles as above.
-\textbf{***need to replot with correct sigma (normalized to 1), also correct uWUE label in first plot***}}
+\textbf{***I think I'm going to remove the bottom plot: its discussion didn't fit in well with the text, and the box plots test inclusion of uncertainty***}}
 \label{term3}
@@
 
-Figure \ref{term3}b shows the location of VPD$_{crit}$, as a function of $\sigma$ (uncertainty) and $VPD$.
-For any $\sigma$ or VPD less (more) than these curves, the sign term will be negative (positive).
-It is clear that the portion of VPD observations below/above these curves will be a strong function of $\sigma$.
-This motivates a thorough examination of how inclusion of uncertainty ($\sigma$) might change the form of the sign term, and in particular the location of VPD$_{crit}$ for each of our plant types.
-In Secion \ref{testing} we examine if and how uncertainty changes the analysis presented in this section, for each PFT.
-
 \subsection{Functional Form of the Scaling Term}
 
-While the above discussion of the sign of $\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our research question, the magnitude of $\frac{\partial \; ET}{\partial \; VPD}$ will also the relative magnitude of the change $\frac{\partial \; ET}{\partial \; VPD}$ and the importance of $VPD$ variability for overall $ET$ variability.
-So we now more closely examine the scaling term: $\frac{P}{T} \propto \rho$.
-Air density $\rho$ scaling term varies little relative to aerodynamic conductance and $\Delta$.
-The psychrometric constant $\gamma$ is also relatively constant, so the scaling term should be primarily a function of aerodynamic conductance and temperature, through the slope of the Clauisu-Clapeyron relationship $\Delta$.
-This is as expected, given that the aerodynamic conductance represents the efficiency fo the transfer of surface anomalies to the atmosphere.
+While the above discussion of the sign of $\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our research question, understaing the overall magnitude of the ET reponse is important to the soil-plant-atmosphere water budgeting.
+So we now more closely examine the terms that affect how the sign term is scaled:
+
+\begin{equation}
+  \frac{g_a \; P}{T(\Delta + \gamma)},
+\end{equation}
+
+$\frac{P}{T}$ is an air-density term, which we would expect to vary relatively little compared to aerodynamic conductance and $\Delta$.
+The psychrometric constant ($\gamma$) is also relatively constant, so the scaling term should be primarily a function of aerodynamic conductance and temperature, through the slope of the Clauisu-Clapeyron relationship $\Delta$.
+This is as expected, given that the aerodynamic conductance represents the efficiency fo the transfer of the surface environment to the atmosphere.
 As aerodynamic conductace  increases, any plant response will be communicated more strongly to the atmosphere (and vice-versa).
@@
 Perhaps most importantly, the x-intercept, and thus VPD$_{crit}$, matches nearly exactly between the theory and the observations.
-Therefor the sign of the ET response to increases in VPD should be well matched, subject to the unavoidable constraints of noise, much of which may be from the observations themselves.
+Therefore the sign of the ET response to increases in VPD should be well matched, subject to the unavoidable constraints of noise, much of which may be from the observations themselves.
 The uncertainty is non negligable; there are bany observations in each box for which the the sign of observation is opposite the response predicted by the theory, but to leading order our theory matches the observations.
@@
 So, the theory still captures when the ET reponse will be positive or negative, but is less accurate with the magnitude of the reponse.
-Therefor the theory is still useful for addressing when VPD drives or reduces ET.
+Therefore the theory is still useful for addressing when VPD drives or reduces ET.
 Also, it's with noting that we used published constants for $g_1$ \citep[from ][]{Lin_2015}, which determines the VPD dependence.

### 2018-01-24 src/paper_figure_code/tables.py

@@ -14,29 +14,23 @@ def frequency(_df):
   return _df.d_et[_df.d_et < 0.].count() / _df.d_et.count()
 
 
-pft = df.groupby('site').apply(get_pft)
-mean = df.groupby('site').mean()
-mean['pft'] = pft
-std = df.groupby('site').std()
-#jacobian = site_analysis(mean)
-mean['d_et_bar'] = calc.d_et(mean)
-mean['d_et_bar_std'] = np.absolute(calc.d_et(mean))*std['vpd']
-mean['d_rn_bar'] = np.absolute(calc.d_et_d_r_net(mean))*std['r_net']
-mean['d_et_bar_norm_rn'] = mean['d_et_bar_std']\
-                           /mean['d_rn_bar']
+std = df.groupby('pft').std()
+mean_df['d_et_bar'] = d_et(mean_df)
+mean_df['d_et_bar_std'] = np.absolute(mean_df['d_et_bar'])*std['vpd']
+mean_df['d_et_bar_norm'] = mean_df['d_et_bar_std']\
+                           /std['et_obs']
 
-mean_df = mean.groupby('pft').mean()
 counts = df.groupby('pft').apply(frequency)
 
 print('\n mean d_et\n', mean_df.d_et)
 print('\n d_et(mean)\n', mean_df.d_et_bar)
 print('\n mean d_et *std vpd\nn', mean_df['d_et_bar_std'])
-print('\n mean d_et / rad\n', mean_df['d_et_bar_norm_rn'])
+print('\n mean d_et / rad\n', mean_df['d_et_bar_norm'])
 print('\n portion d_et < 0 \n',\
       counts)
 mean_df['counts'] = counts
 test = mean_df.loc[:, ['d_et', 'd_et_bar', 'd_et_bar_std',\
-                       'd_et_bar_norm_rn', 'counts']]
+                       'd_et_bar_norm', 'counts']]
 
 print(test)
 

### 2018-01-25 doc/paper/vpd_et_paper_PG_v1.tex
just have discussions section, conclusions, and thenr ead forward

@@
 $\frac{P}{T}$ is an air-density term, which we would expect to vary relatively little compared to aerodynamic conductance and $\Delta$.
-The psychrometric constant ($\gamma$) is also relatively constant, so the scaling term should be primarily a function of aerodynamic conductance and temperature, through the slope of the Clauisu-Clapeyron relationship $\Delta$.
-This is as expected, given that the aerodynamic conductance represents the efficiency fo the transfer of the surface environment to the atmosphere.
+The psychrometric constant ($\gamma$) is also relatively constant, so the scaling term should be primarily a function of aerodynamic conductance and temperature, through the Clauisu-Clapeyron relationship $\Delta$.
+This is as expected, given that the aerodynamic conductance represents the efficiency of exchange between the surface and the atmosphere.
 As aerodynamic conductace  increases, any plant response will be communicated more strongly to the atmosphere (and vice-versa).
@@
 The functional from of $\Delta$ will be the same across PFT, but the temperature range may vary slightly.
-In contrast, aerodynamic conductance will vary strongly with PFT due to the importance of surface roughness.
+In contrast, aerodynamic conductance will vary strongly with PFT due to the importance of surface roughness for aerodynamic conducatance.
 So most of the differences in scaling between PFT should be in the aerodynamic conductance term.
-One interesting side note is that the coefficient of variability for both aerodynamic conductance and the scaling term is relatively constant across PFT, suggesting that the influence of aerodynamic conductance on the relative (to the PFT mean) variability of the scaling term is approximately similar across PFT.
 
-Figure \ref{scale_vary}A shows the scaling term normalized by mean aerodynamic conductance (calculated for each plant functional type), and confirms that much of the relative variability of the scaling term is caused by aerodynamic conductance variability.
-Generally, temperature (via $\Delta$) causes less relative variability.
-However, the impact of $T$ on the relative variability increases with increasing aerodynamic conductance.
+The control of scaling term variability between PFTS by aerodynamic conductance is confirmed by the data (Figure \ref{scale_vary}).
+Differences between PFT are almost entiredly due to differences in aerodynamic conductance, rather than differences in observed temperature ranges.
+The Scaling term for the tree PFTs (DBF, ENF) is generally more than double the scaling terms for the other PFTs which have lower surface roughness and gererally smaller aerodynamic conductance.
 
-While the relative variability of scaling term is similar across PFT, the absolute value of scaling term varies strongly across PFT.
-Figure \ref{scale_vary}B shows scaling term evaluated at the mean aerodynamic conductance for each PFT, and at the range of observed temperatures for each PFT.
-As expected, for the tree PFTs (DBF, ENF) the scaling term is much larger and the temperature dependence is much stronger.
-Systematic differences in observed temperatures also cause differences in the average magnitude of scaling term.
-For example, ENF experiences on average colder temperatures and is thus more likely to have a larger scaling term.
-Additionally, because the variability of aerodynamic conductance increases proportionally to the mean, the spread of the scaling term due to aerodynamic conductance variability will be larger for the tree PFTs, although this is not shown for simplicity.
-To summarize Figure \ref{scale_vary}: the variability of the scaling term within each PFT will look like Figure \ref{scale_vary}A for each PFT, but the scale of the y-axis will increase or decrease according to mean aerodynamic conductance observed in Figure \ref{scale_vary}B.
+Within each PFT, scaling term variability is controlled both by environmental temperature and aerodynamic conductance variability (Figure \ref{scale_vary}).
+While the observed variability of aerodynamic conductance conrbutes more to scaling term variability than temperature, temperature's contributions are non-negliable.
+Specifically, the scaling term is generally larger at low temperatures when ET will be weakly damped by Clasius-Clapyeron, and the effect of aerodynamic conductance variability is stronger as it is less weakly damped by Clasius-Clapyeron.
+
+To summarize, variability between PFTs is mostly controlled by systematic differences in aerodynamic conducntace, due to differences in surface roughness, betwen each PFT.
+In contrast, variability within PFT is also controlled by temperature, through Clasius-Clapyeron.
+But, aerodynamic conductance variability generally impacts the scaling term more than temperature, even within PFT.
 
@@
 \centering
-\includegraphics[width=20pc]{./fig04.pdf}
-\caption{Primary sources of variability for scaling term.
-A) Variability within each PFT: scaling term normalized by mean $g_a$ for each PFT.
-B) Variability between each PFT: scaling term evaluated at mean $g_a$ for each PFT.
-Temperature range is 5-95th percentile for each PFT.
-Additionally, stars denote the location of the 25th, 50th, and 75th percentiles.}
+\includegraphics[width=20pc]{./fig04b.pdf}
+\caption{Primary sources of variability for the scaling term, as a function of PFT.
+Note the change in axis scale for each plot.
+The 5th-95th percentile range of temperature is plotted at the 5th, 25th, 50th, 75th, and 95th percentiles of aerodynamic conductance, as observed for each PFT's sites.}
 \label{scale_vary}
@@
 
-%%%%% replace below with idea:
-\subsection{Bulk statistics of $\frac{\partial \; ET}{\partial \; VPD}$}
+\subsection{Bulk statistics of ET response to VPD}
+\label{stats_sec}
+Previous sections only considered idealized scenarios without uncertainty ($\sigma = 1.0$).
+In this section we will consider bulk statistics, as a function of PFT, of ET reponse when $\sigma$ is allowed to vary and eddy-covariance derived uncertainty is included (Section \ref{methods}).
+These bulk statistics of ET reponse confirm our expectations based on the theory presented in Sections \ref{sign_term}-\ref{scale_term} (Table \ref{stats}.
+For all PFTs, mean ET reponse to increasing VPD is negative.
+However, ET response evaluated at the average of all variables (e.g.
+$\sigma$, $T$, $c_s$, $VPD$) is only negative for CSH and GRA.
+This difference in mean ET reponse as compared to the ET reponse at mean environmental conditions is due to the non-linear nature of the reponse, in which negative reponses are generally much larger magnitude than postive reponses (Figure \ref{sign}).
 
-Table 3 confirms our expectations for PFT behavior of $\frac{\partial \; ET}{\partial \; VPD}$.
-For all PFTs except for CRO, average $\frac{\partial \; ET}{\partial \; VPD}$ is less than zero.
-However, $\frac{\partial \; ET}{\partial \; VPD}$ evaluated at the average of all variables (e.g.
-$\sigma$, $T$, $c_s$, $VPD$) is only negative for CSH and GRA.
-And, DBF in addition to CRO experiences $\frac{\partial \; ET}{\partial \; VPD}$ < 0 less than half the time.
-These observations highlight the effect of the nonlinear function in Figure \ref{term3}: $\frac{\partial \; ET}{\partial \; VPD}$ has a much steeper slope when the function is negative, and is thus more likely to be large magnitude.
+Regarding the frequency of negative and postive ET response, observations of ENF, GRA and CSH are more frequently characterized by a negative ET reponse to increasing VPD, and are thus more frequently water conservative.
+DBF and CRO observations have greater frequency of postive ET reponse to increasing VPD, so are less frequently water conservative.
 
-The units of $\frac{\partial \; ET}{\partial \; VPD}$ make it difficult to interpret if $VPD$ is even a meaningful contributor to ET's variability.
-To better understand $VPD$'s contribution, we multiply $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)$ with $VPD$'s standard deviation to define a (linearized) relative change in ET for variations in $VPD$ .
-$VPD$'s contribution to ET's variability ranges between 30 - 40 W m$^{-2}$ for all PFTs except for CSH, which is about 100 W m$^{-2}$.
-Another meaningful comparison is to $\frac{\partial \; ET}{\partial \; R} \cdot std(R)$, as net radiation is generally the driver of ET (cite joe berry here)\explain{Find citation?}.
-For all PFTs except for CSH $VPD$ contributes between 30 - 40 \% of $R$'s contribution to variability.
-For CSH the portion is much larger, about 88 \%.
-$VPD$'s variability is certainly a non-negligible contributor to $ET$'s variability.
-
-Theoretical derivation has so far illuminated the dependence of $\frac{\partial \; ET}{\partial \; VPD}$ and how it varies across PFT.
-Large mean $uWUE$ shifts CRO and DBF towards a tendancy to increase ET with VPD (positive $\frac{\partial \; ET}{\partial \; VPD}$), but a lower slope parameter ($g_1$) for DBF allows for decreaes in ET with VPD, at low environmental VPD.
-\explain{physically state how slope parameter is trade for water at leaf level, but uWUE is trade-ff between GPP and VPD at ecosystem-level}.
-ENF's very low slope parameter ($g_1$), which is even lower than for DBF, increases the dependence of ET response  ($\frac{\partial \; ET}{\partial \; VPD}$) on $VPD$, and makes the function strongly nonlinear.
-This has the effect of making mean ET response ($\overline{\frac{\partial \; ET}{\partial \; VPD}}$)  more negative for a given frequency of occurence of negative ET response ($\frac{\partial \; ET}{\partial \; VPD} < 0$), than for other PFT.
-GRA shows the opposite behavior; a relatively high slope parameter ($g_1$)  makes the function more linear, causing a more positive mean ET reponse ($\overline{\frac{\partial \; ET}{\partial \; VPD}}$) for a given given frequence of negative ET reponse  ($\frac{\partial \; ET}{\partial \; VPD} < 0$).
-Finally, CSH's very low $uWUE$ dominates the slope parameter's effects, causing ET to almost always decrees with increasing atmospheric dryness (VPD), even with uncertainty included.
-Variability in $VPD$ accounts for the largest amount of $ET$ variability for CSH.
-For the other PFTs, $VPD$ contributes less to $ET$ variability, but still represents about 30-40 \% of $R$'s contributions to ET variability.
-\explain{think pierre is correct, could probably scrap this anecdote: WHAT IS THE SORUCE OF VARIBILITY OF THE REST OF ET I DON"T KNOW IF THAT IS RELEVANT FOR THE PRESENT PAPER.}
+In general the bulk statistics match our expectations based on previous idealzied and theoretical analysis, with the possible exception that inclusion of uncertainty shfits grass towards a slightly more negative reponse than the theory and Figure \ref{sign} would suggest.
+The bulk statistics motivate a more throrough examination of the structure of uncertainty and more sophsiticated validation of our theory's performance against observations.
 
 \begin{table}
+  \label{stats}
 \caption{Statistics of $\frac{\partial \; ET}{\partial \; VPD}$ as a function of PFT. \textbf{***dated, need to update these statistics***}}
@@
 \label{testing}
-So far we have developed a theory for $\frac{\partial \; ET}{\partial \; VPD}$'s behavior as a function of VPD and PFTs.
-We now turn to an direct evaluation against observations taken from a wide range of eddy-covariance datasets (Section \ref{methods}).
-
+We now turn to an direct evaluation of the functional forms presented in Section \ref{sign_term} against observations taken from a wide range of eddy-covariance datasets (Section \ref{methods}).
 Figure \ref{agu_fig} shows the distribution of of the sign term (when uncertainty is included), as compared to what the theory would predict.
@@
 For DBF our theory is successful towards this goal: the theory matches the leading order behavior of the function when uncertainty is included, and the DBG figure looks as one would expect if one just added noise on top of our theory.
-The VPD dependence, given by the slope parameter ($g_1$), floows the median values of each box plot.
+The VPD dependence, given by the slope parameter ($g_1$), follows the median values of each box plot.
 Perhaps most importantly, the x-intercept, and thus VPD$_{crit}$, matches nearly exactly between the theory and the observations.
 Therefore the sign of the ET response to increases in VPD should be well matched, subject to the unavoidable constraints of noise, much of which may be from the observations themselves.
-The uncertainty is non negligable; there are bany observations in each box for which the the sign of observation is opposite the response predicted by the theory, but to leading order our theory matches the observations.
+The uncertainty is non negligable; there are many observations in each box for which the the sign of observation is opposite the response predicted by the theory, but to leading order our theory matches the observations.
 
@@
 
-The theory does not match as well for CRO as for DBF and CSH.
-At low VPD the VPD dependence (due to $g_1$) is similiar to the leading order behavior of the observations, but the entire curve is shifted towards postiive values, due to some error in $uWUE$.
-Most strikingly, the observations show a tendency to move back towards negative values at high VPD, which given the functional form of VPD dependnce, is behavior that is impossible fo the theory to capture.
-This divergence between the theory and the observations for CRO could have been expected.
+As Section \ref{stats_sec} suggested, the theory does not match as well for CRO as for DBF and CSH.
+At low VPD the VPD dependence (due to $g_1$) is similiar to the leading order behavior of the observations, but the theory's curve is shifted towards postiive values relative to the observations, due to some error in $uWUE$.
+Most strikingly, the observations show a tendency to move back towards negative values at high VPD, which given the functional form of VPD dependnce, is behavior that is impossible for the theory to capture.
+This general postive bias the theory explains  the shift towards negative ET reponse observed in the bulk statistics of Section \ref{stats_sec}.
+The divergence between the theory and the observations for CRO could have been expected.
 For practical reasons, our theory does not include important sources of variability (particularly between sites) for CRO.
 We do not account for seasonal changes in LAI, variability in crop height and surface roughness, differences in C3 vs C4 photosynthesis, or differences in irrigation practices between the sites and/or years.
-This deficiencies would explain the inability of our theory to match obsercations.
-For example, a superpositions of sites with C3 plants (low VPD) and C4 plants (high VPD) would explain observed shift back towards negative ET reponse at high VPD when all sites are binned together, as in Figure \ref{agu_fig}.
+These deficiencies would explain the inability of our theory to match observations.
+For example, a superposition of sites with C3 plants (at low enviromental VPD) and C4 plants (at high environmental VPD) would explain observed shift back towards negative ET reponse at high VPD when all sites are binned together, as in Figure \ref{agu_fig}.
 
 The theory for GRA suffers from similiar limitations as for CRO.
-The differences between the theory and the observations are more pronounced for GRA, but are explainable using similiar logic as that used for CRO.
+GRA observations are characterized by a consistent trend back towards negative ET reponse at higher VPD, which the functional form of our theory is incapable of accounting for.
+As copared to CRO, the divergence between the theory and the observations is even greater for GRA, but these differences are explainable using similiar logic as presented for CRO.
 GRA will also have differences between sites in C3 vs C4 photosythesis, and irrigation practices, as well as strong seasonal variability in crop height and LAI.
-For GRA, the observations show a very different VPD dependence of response than the theory, with a strong tendency back toewards a negative ET reponse at high VPD which our theory cannot account for.
-As with CRO, this could be explained by a superposition of sites with C3 vs C4 sites.
+As with CRO, the trend back towards negative ET reponse at higher VPD could be explained by a superposition of sites with C3 vs C4 sites.
 Unfortunately, because detailed site-level data of C3 vs C4 plant type, as well as irrigation practes, was difficult to obtain, we could not bin GRO and GRA further by irrigation and photosyethesis stragtegy.
-However, we hypothesize that the theory would much more closel match obsercations with photosythesis and irrigation are accounted for.
+However, we hypothesize that the theory would validate against obsercations much better when photosythesis strategy and irrigation are accounted for.
 
-The comparison between the theory and obsercations for ENF is unique from the rest of the PFTs.
-The VPD dependence of the theory, determined by the slope paramter, is not well captured by observations.
+The comparison between the theory and observations for ENF is unique from the rest of the PFTs.
+The VPD dependence of the theory, determined by the slope paramter ($g_1$), is not well captured by observations.
 However, the theory closely captures the x-intercept and VPD$_{crit}$.
@@
 
-While the above discussion shows that our theory has some limitations when applied to some sites, especially for crops and grasslands, another question could be, how would using PET, as many hydrometeorological studies and applications do, compare with our theory.
+While the above discussion shows that our theory has some limitations when applied to some PFTs, especially for crops and grasslands, it does well for DBF and CSH PFTs, and captures the sign of the reponse conistently for ENF.
+It is useful to also compare our theory, the observed uncertainty in the reponse, and PET, as PET is ofted used as a proxy for ET in hydrometerological studies and drought indicies.
 Again, as in Section \ref{sign_func}, the PET response to changes in VPD will be constant at about 3.5.
@@
 This casts doubt on PET's usefulness for capturing the leading order behavior of water budgeting at the land surface.
-However our theory (Equation \ref{et}) could be used with published $uWUE$ and $g1$ as an alternative to PET to estimating land response and moisture budgeting.
+However our theory (Equation \ref{et}) could be used with published $uWUE$ and $g1$ as an alternative to PET to estimating land response and moisture budgeting, and would more accurately account for the response to atmospheric drying..
 
 \subsubsection{Testing VPD$_{crit}$}
-As stated in the above discussion of ENF, we are mostly interested in the critical threshold $VPD_{crit}$ for which the sign of  $\frac{\partial \; ET}{\partial \; VPD}$ is determined.
-When $VPD < VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}  < 0$ (plant response dominates), and when $VPD > VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD} > 0$ (atmospheric demand dominates).
+As stated in the above discussion of ENF, we are mostly interested in the critical threshold $VPD_{crit}$ for which the sign of the ET reponse to VPD ($\frac{\partial \; ET}{\partial \; VPD}$) is determined.
+When VPD is less than VPD$_{crit}$, ET reponse to VPD is negative ($\frac{\partial \; ET}{\partial \; VPD}  < 0$, plant response dominates), and when VPD is greater than VPD$_{crit}$, ET reponse to VPD is positive  ($\frac{\partial \; ET}{\partial \; VPD} > 0$ atmospheric demand dominates).
 Theoretically this $VPD_{crit}$ is only a function of PFT.
@@
 
-Figure \ref{real} presents calculated $\frac{\partial \; ET}{\partial \; VPD}$ where, unless otherwise noted, all variables in Equation \ref{d_et} are allowed to vary, including uncertainty.
+Figure \ref{real} presents calculated $\frac{\partial \; ET}{\partial \; VPD}$ where, unless otherwise noted, all variables in Equation \ref{d_et} are allowed to vary, including uncertainty ($\sigma$).
 Each column represents a different quantity related to $\frac{\partial \; ET}{\partial \; VPD}$, and each row is a different PFT.
@@
 
-The scatter figures generally confirm expectations from Figure \ref{agu_fig}.
-By comparing different columns of Figure \ref{real} we can evaluate the impact of $\sigma$ and $g_a$ on the variability of $\frac{\partial \; ET}{\partial \; VPD}$.
-$g_a$'s scaling (included in columns 1 and 3) alters the magnitude considerably, particularly between PFTs.
+Figure \ref{real} generally confirms expectations from Figure \ref{agu_fig}.
+By comparing different columns of Figure \ref{real} we can evaluate the impact of $\sigma$ and aerodynamic conductance ($g_a$) on the variability of ET reponse ($\frac{\partial \; ET}{\partial \; VPD}$).
+Aerodyamic conducntance's scaling (included in columns 1 and 3) alters the magnitude considerably, particularly between PFTs.
 $\sigma$ variability (included in columns 1 and 2) adds a lot of additional noise to the signal of $\frac{\partial \; ET}{\partial \; VPD}$, as observed in Figure \ref{agu_fig}.
@@
 
-While $g_a$ has a large impact on the magnitude of the scaling, temperature, through the derisive of the Clausius-Clapeyron, $\Delta$, term, does not appear to have as much of an impact in the complete estimation of $\frac{\partial \; ET}{\partial \; VPD}$.
+While $g_a$ has a large impact on the magnitude of the scaling, temperature, through Clausius-Clapeyron ($\Delta$), does not appear to have as much of an impact in the complete estimation of ET reponse ($\frac{\partial \; ET}{\partial \; VPD}$).
 Therefore it appears that the absolute magnitude of the derivative can be mostly attributed to the sign term, and its direct scaling by aerodynamic conductance.
 Intuitively this is reasonable; aerodynamic conductance will control how dominant balances at the land surface are communicated to the boundary layer.
-While $\Delta$ controls the efficiency of latent fluxes in carrying heat away from the surface, it appears this is a second order term, relative to $g_a$, for scaling $\frac{\partial \; ET}{\partial \; VPD}$.
+While $\Delta$ controls the efficiency of latent fluxes in carrying heat away from the surface, it appears this is a second order term, relative to $g_a$, for scaling ET reponse.
 
-In general Figures \ref{real} and \ref{agu_fig} match our expectations, even with the large sensitivity to $\sigma$ seen in Figure \ref{term3}.
+In general Figures \ref{real} and \ref{agu_fig} match our expectations.
 Our $\sigma$-based method of uncertainty propagation blurs the idealized expectations the most for GRA and CRO.

### 2018-01-25 doc/paper/vpd_et_paper_PG_v1.tex
done with formatting and spell check, now just ifnal read through

@@
 is there another citation?
-I don't think Byrne 2013 shows that.} \citep{Byrne_2013}.
+I'm not sure Byrne 2013 shows that.} \citep{Byrne_2013}.
 Increases in VPD increase the atmospheric demand of evapotranspiration (ET) \citep{Monteith_1965} but also stress plant stomata \citep{Leuning_1990, MEDLYN_2011}.
 
-This leades to two potential but opposing perspectives on how ET adjusts to changes in VPD.
+This leads to two potential but opposing perspectives on how ET adjusts to changes in VPD.
 The first, a hydrometeorological perspective, is that higher VPD increases atmospheric demand for water from the land surface, and this drives an increase in evapotranspiration (ET).
@@
 ][]{Or_2013}, increased atmospheric aridity would further reduce soil moisture.
-In turn, this would further increase aridity as low soil moisture levels increase the Bowen ratio, and in turn increase temperature and reduce atmospheric humidity (gentine et al.
-GRL 2016, Berg 2016).
+In turn, this would further increase aridity as low soil moisture levels increase the Bowen ratio, and in turn increase temperature and reduce atmospheric humidity \citep[gentine et al.
+GRL 2016][]{Berg_2016}.
 This however would not seem to be a sensible strategy for plants from an evolutionary standpoint.
 
-We can use intuition about plant water conservation strategy to hypothesize about ET reponse to changes in VPD.
+We can use intuition about plant water conservation strategy to hypothesize about ET response to changes in VPD.
 Plants that evolved to conserve water (e.g.
@@
 \label{data}
-We use both meterological and turbulent flux data from the FLUXNET2015 database, including all sites with more than four years of data\explain{Pierre says ``be as precise as possible,'' but not sure what he means}.
+We use both meteorological and turbulent flux data from the FLUXNET2015 database, including all sites with more than four years of data.
 Each site's plant functional type (PFT) was classified using the International Geosphere-Biosphere Programme vegetation classification scheme \citep{Loveland_1999}.
@@
 \end{linenomath*}
-where $\Delta$ is the change in saturation vapor pressure with temperature ($\frac{\partial \; e_s}{\partial \; T}$), $R$ is the net radiation minus ground heat flux, $g_a$ is aerodynamic conductance, $\rho_a$ is air density, $c_p$ is specific heat of air at constant pressure, $\gamma$ is the psychometric constant, and $g_s$ is the stomatal conudctance (Table 1).
+where $\Delta$ is the change in saturation vapor pressure with temperature ($\frac{\partial \; e_s}{\partial \; T}$), $R$ is the net radiation minus ground heat flux, $g_a$ is aerodynamic conductance, $\rho_a$ is air density, $c_p$ is specific heat of air at constant pressure, $\gamma$ is the psychometric constant, and $g_s$ is the stomatal conductance (Table 1).
 \citet{MEDLYN_2011} developed a model for stomatal conductance ($g_s$) by combining optimal photosynthesis theory \citep{Farquhar_1980, Katul_2010} with empirical approaches, which describes the dependence of $g_s$ to VPD.
@@
 
-where $\Gamma$ is the CO$_2$ compensation point for photosythesis (without dark respiration), and $\lambda$ is the marginal cost of leaf carbon.
+where $\Gamma$ is the CO$_2$ compensation point for photosynthesis (without dark respiration), and $\lambda$ is the marginal cost of leaf carbon.
 So, $g_1$ is a leaf-scale term reflecting the willingness of a leaf to trade water for carbon uptake and growth.
-The higher $g_1$, the more willing a leaf will be willing to release water in the exchange for water.
+The higher $g_1$, the more willing a leaf will be willing to open stomata and release water in the exchange for carbon.
 
@@
 Note that $uWUE$ is fit on the ecosystem scale in \citet{Zhou_2015} so GPP in \ref{uwue} is really $A\cdot \text{LAI}$.
-This leads to the cancelation of LAI in addition to uWUE in Equation \ref{medlyn}.
+This leads to the cancellation of LAI in addition to uWUE in Equation \ref{medlyn}.
 Plugging Equation 5 into Equation 1 and rearranging gives a new expression for ET directly as a function of VPD:
@@
 
-With Equation \ref{et}, ET is explicitly a function of environmental variables and two plant-specific constats, the slope parameter ($g_1$), and $uWUE$.
-The slope parameter is a leaf-scale term related to the willingness of leaves to trade water for new growth.
+With Equation \ref{et}, ET is explicitly a function of environmental variables and two plant-specific constants, the slope parameter ($g_1$), and $uWUE$.
+The slope parameter is a leaf-scale term related to the willingness of leaves to trade water for new growth and keep stomata open.
 $uWUE$ is semi-empirical ecosystem-scale constant related to how WUE changes with VPD (specifically $VPD^{-1/2}$).
@@
 
-Now, from each FLUXNET observation we can evaluate $\sigma$ at each time step and thus we can evaluate departure from our theory (as a departure from unity).
-The variability of $\sigma$ across sites and time provides a measure of uncertainty in our model, assumptions, as well as the FLUXNET observations themselves.
+Now, from each FLUXNET observation we can evaluate $\sigma$ at each time step:
+
+\begin{linenomath*}
+  \begin{equation}
+\sigma = - \frac{g_a \gamma c_s \sqrt{VPD} L_v P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p VPD\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})},
+    \label{sigma}
+  \end{equation}
+\end{linenomath*}
+
+and thus we can evaluate departure from our theory (as a departure of $\sigma$ from unity) The variability of $\sigma$ across sites and time provides a measure of uncertainty in our model, assumptions, as well as the FLUXNET observations themselves.
 To correct for differences in sampling between \cite{Zhou_2015} and our data, we set uWUE such that $\overline{\sigma} = 1$.
@@
 \subsection{Functional Form of the Sign Term}
-\label{sign_func}
+\label{sign_term}
 First, we explore the variables within the sign term to gain better intuition on the driver of either the increase or reduction of ET with VPD.
-CO$_2$ concentration ($c_s$) and the pschometric constant ($\gamma$) are relatively constant to a good approximation so that the variability is dominated by $\sigma$ and $VPD$.
+CO$_2$ concentration ($c_s$) and the psychometric constant ($\gamma$) are relatively constant to a good approximation so that the variability is dominated by $\sigma$ and $VPD$.
 $uWUE$ could vary with soil moisture but has been shown to be relatively constant except in very dry soil conditions \citep{Zhou_2016, Lin_2015}.
@@
 We note however that even though ET increases again above the critical threshold, $VPD_{crit}$, ET is still much lower that potential evaporation as stomata are still strongly regulating vapor fluxes.
-Even in the absence of soil pore evaporation, stomata do not shut down entirely at very high VPD and ET does not go to zero, because stomata are still slightly open to perform some photosynthesis \citep{Ball_1987, Leuning_1990, Medlyn_2011}.
+Even in the absence of soil pore evaporation, stomata do not shut down entirely at very high VPD and ET does not go to zero, because stomata are still slightly open to perform some photosynthesis \citep{Ball_1987, Leuning_1990, MEDLYN_2011}.
 In addition, upward xylem transport is necessary to maintain phloem transport and thus carbon allocation (MENCUCCINI) \explain{Not sure citation  - Comparative criteria for models of the vasuclar transport systems of tall trees}.
@@
 A larger $uWUE$ causes means that at the ecosystem scale $WUE$ decreases more with increasing VPD, so plants lose more water per unit carbon production and are more likely to increase ET with increasing VPD.
-A tendenacy towards increasing ET reponse with increasing VPD can also be caused by high slope parameters ($g_1$), characteristic of plants that at the leaf scale plants are more willing to trade water for growth.
+A tendency towards increasing ET response with increasing VPD can also be caused by high slope parameters ($g_1$), characteristic of plants that at the leaf scale plants are more willing to trade water for growth and open stomata.
 Plants that give water more freely will be much more likely to increase ET with increasing VPD.
@@
 CRO has a high slope parameter ($g_1 = 183.1 \, Pa^{1/2}$) and uWUE ($2.32$ -units-) compared to CSH ($g1 = 148.6 \, Pa^{1/2}$, $uWUE=2.11$ -units-).
-These differences in slope parameter cause opposite ET reponses to changse in VPD between CRO and CSH.
+These differences in slope parameter cause opposite ET responses to changes in VPD between CRO and CSH.
 ET almost always increases with increasing VPD for CRO, while ET almost always decreases with increasing VPD for CSH (Figure \ref{term3}).
@@
 
-The slope paramter ($g_1$) is additionally a primary determinant of VPD  dependnece for the sign term shown in Figure \ref{term3}.
+The slope parameter ($g_1$) is additionally a primary determinant of VPD  dependence for the sign term shown in Figure \ref{term3}.
 Plants that are less willing to trade water for new growth (small $g_1$) will have strong tendency to reduce ET with increasing VPD, and they will be very effective in reducing ET especially at low VPD.
-However, at very high VPD gradients in vapor pressure at the leaf scale will begin to dominate leaf-scale plant strategies (parameterized with $g_1$), so ET reponse will begin to asymtote towards ecosystem-scale values irregardless of leaf-scale strategy (and $g_1$).
+However, at very high VPD gradients in vapor pressure at the leaf scale will begin to dominate leaf-scale plant strategies of stomatal closure (parameterized with $g_1$), so ET response will begin to asymptote towards ecosystem-scale values irregardless of leaf-scale strategy and stomatal closure (through $g_1$).
 Therefore, plants with a low $g_1$ will have the largest VPD dependence of ET response because the difference in ET response at low VPD (leaf response dominates) and high VPD (VPD gradient dominates) is largest.
-This is apparent in the strong VPD dependece of ENF, which has the lowest slope parameter ($g_1=74.31$), in Figure \ref{term3}.
+This is apparent in the strong VPD dependence of ENF, which has the lowest slope parameter ($g_1=74.31$), in Figure \ref{term3}.
 
-To summarize Figure \ref{term3} and Table \ref{vpd_crit}, CROs are the least water conservative and have the strongest overall tendency to increase ET wirh increasing VPD, while CSH are the most water conservative and have the strongest tendency to decrease ET with increasing VPD.
+To summarize Figure \ref{term3} and Table \ref{vpd_crit}, CROs are the least water conservative and have the strongest overall tendency to increase ET with increasing VPD, while CSH are the most water conservative and have the strongest tendency to decrease ET with increasing VPD.
 ENF ($g_1 = 74.31$) has by far the largest VPD dependence of response, while CRO ($g_1 = 183.1$) has the smallest VPD dependence.
-ENF are less willing to trade water for new growth at the leaf scale, so the difference between the response at low VPD when leaf response domineates and high VPD when vapor pressure gradients dominate is much greater.
+ENF are less willing to trade water for new growth at the leaf scale, so the difference between the response at low VPD when leaf response dominates and high VPD when vapor pressure gradients dominate is much greater.
 CRO on the otherhand are more willing to trade water for new growth, so the leaf scale response is much smaller, and the difference in ET response between low VPD and high VPD is relatively low.
 
-Figure \ref{term3} show clearly that, accoridng to our thoery, for all PFTs except for crops there is frequent occurrence of a negative (plant dominating) ET response to increases in VPD.
-Therefore, plants are able in most atmospheric conditions to redue ET in response to increased VPD and thus to reduce water loss.
+Figure \ref{term3} show clearly that, according to our theory, for all PFTs except for crops there is frequent occurrence of a negative (plant dominating) ET response to increases in VPD.
+Therefore, plants are able in most atmospheric conditions to reduce ET in response to increased VPD and thus to reduce water loss.
 To better illustrate this, the ranges of observed environmental VPDs at the FLUXNET sites are plotted parallel to the x-axis.
@@
 Plants are always regulating water exchange from the land surface, even when they reach the limits of they ability to do so.
-Therefore, the actual land surface reponse to a change in VPD will always me more negative than if the plants are not included in the physics (PET).
+Therefore, the actual land surface response to a change in VPD will always me more negative than if the plants are not included in the physics (PET).
+
 \begin{figure}[h]
@@
 \subsection{Functional Form of the Scaling Term}
-
-While the above discussion of the sign of $\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our research question, understaing the overall magnitude of the ET reponse is important to the soil-plant-atmosphere water budgeting.
+\label{scale_term}
+While the above discussion of the sign of $\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our research question, understating the overall magnitude of the ET response is important to the soil-plant-atmosphere water budgeting.
 So we now more closely examine the terms that affect how the sign term is scaled:
@@
 $\frac{P}{T}$ is an air-density term, which we would expect to vary relatively little compared to aerodynamic conductance and $\Delta$.
-The psychrometric constant ($\gamma$) is also relatively constant, so the scaling term should be primarily a function of aerodynamic conductance and temperature, through the Clauisu-Clapeyron relationship $\Delta$.
+The psychometric constant ($\gamma$) is also relatively constant, so the scaling term should be primarily a function of aerodynamic conductance and temperature, through the Clausius-Clapeyron relationship $\Delta$.
 This is as expected, given that the aerodynamic conductance represents the efficiency of exchange between the surface and the atmosphere.
-As aerodynamic conductace  increases, any plant response will be communicated more strongly to the atmosphere (and vice-versa).
+As aerodynamic conductance  increases, any plant response will be communicated more strongly to the atmosphere (and vice-versa).
 
@@
 The functional from of $\Delta$ will be the same across PFT, but the temperature range may vary slightly.
-In contrast, aerodynamic conductance will vary strongly with PFT due to the importance of surface roughness for aerodynamic conducatance.
+In contrast, aerodynamic conductance will vary strongly with PFT due to the importance of surface roughness for aerodynamic conductance.
 So most of the differences in scaling between PFT should be in the aerodynamic conductance term.
@@
 The control of scaling term variability between PFTS by aerodynamic conductance is confirmed by the data (Figure \ref{scale_vary}).
-Differences between PFT are almost entiredly due to differences in aerodynamic conductance, rather than differences in observed temperature ranges.
-The Scaling term for the tree PFTs (DBF, ENF) is generally more than double the scaling terms for the other PFTs which have lower surface roughness and gererally smaller aerodynamic conductance.
+Differences between PFT are almost entirely due to differences in aerodynamic conductance, rather than differences in observed temperature ranges.
+The Scaling term for the tree PFTs (DBF, ENF) is generally more than double the scaling terms for the other PFTs which have lower surface roughness and generally smaller aerodynamic conductance.
 
 Within each PFT, scaling term variability is controlled both by environmental temperature and aerodynamic conductance variability (Figure \ref{scale_vary}).
-While the observed variability of aerodynamic conductance conrbutes more to scaling term variability than temperature, temperature's contributions are non-negliable.
-Specifically, the scaling term is generally larger at low temperatures when ET will be weakly damped by Clasius-Clapyeron, and the effect of aerodynamic conductance variability is stronger as it is less weakly damped by Clasius-Clapyeron.
+While the observed variability of aerodynamic conductance contributes more to scaling term variability than temperature, temperature's contributions are non-negligible.
+Specifically, the scaling term is generally larger at low temperatures when ET will be weakly damped by Clausius-Clapeyron, and the effect of aerodynamic conductance variability is stronger as it is less weakly damped by Clausius-Clapeyron.
 
-To summarize, variability between PFTs is mostly controlled by systematic differences in aerodynamic conducntace, due to differences in surface roughness, betwen each PFT.
-In contrast, variability within PFT is also controlled by temperature, through Clasius-Clapyeron.
+To summarize, variability between PFTs is mostly controlled by systematic differences in aerodynamic conductance, due to differences in surface roughness, between each PFT.
+In contrast, variability within PFT is also controlled by temperature, through Clausius-Clapeyron.
 But, aerodynamic conductance variability generally impacts the scaling term more than temperature, even within PFT.
@@
 \centering
-\includegraphics[width=20pc]{./fig04b.pdf}
+\includegraphics[width=\textwidth]{./fig04b.pdf}
 \caption{Primary sources of variability for the scaling term, as a function of PFT.
@@
 Previous sections only considered idealized scenarios without uncertainty ($\sigma = 1.0$).
-In this section we will consider bulk statistics, as a function of PFT, of ET reponse when $\sigma$ is allowed to vary and eddy-covariance derived uncertainty is included (Section \ref{methods}).
-These bulk statistics of ET reponse confirm our expectations based on the theory presented in Sections \ref{sign_term}-\ref{scale_term} (Table \ref{stats}.
-For all PFTs, mean ET reponse to increasing VPD is negative.
+In this section we will consider bulk statistics, as a function of PFT, of ET response when $\sigma$ is allowed to vary and eddy-covariance derived uncertainty is included (Section \ref{methods}).
+These bulk statistics of ET response confirm our expectations based on the theory presented in Sections \ref{sign_term}-\ref{scale_term} (Table \ref{stats}.
+For all PFTs, mean ET response to increasing VPD is negative.
 However, ET response evaluated at the average of all variables (e.g.
 $\sigma$, $T$, $c_s$, $VPD$) is only negative for CSH and GRA.
-This difference in mean ET reponse as compared to the ET reponse at mean environmental conditions is due to the non-linear nature of the reponse, in which negative reponses are generally much larger magnitude than postive reponses (Figure \ref{sign}).
+This difference in mean ET response as compared to the ET response at mean environmental conditions is due to the non-linear nature of the response, in which negative responses are generally much larger magnitude than positive responses (Figure \ref{sign}).
 
-Regarding the frequency of negative and postive ET response, observations of ENF, GRA and CSH are more frequently characterized by a negative ET reponse to increasing VPD, and are thus more frequently water conservative.
-DBF and CRO observations have greater frequency of postive ET reponse to increasing VPD, so are less frequently water conservative.
+Regarding the frequency of negative and positive ET response, observations of ENF, GRA and CSH are more frequently characterized by a negative ET response to increasing VPD, and are thus more frequently water conservative.
+DBF and CRO observations have greater frequency of positive ET response to increasing VPD, so are less frequently water conservative.
 
-In general the bulk statistics match our expectations based on previous idealzied and theoretical analysis, with the possible exception that inclusion of uncertainty shfits grass towards a slightly more negative reponse than the theory and Figure \ref{sign} would suggest.
-The bulk statistics motivate a more throrough examination of the structure of uncertainty and more sophsiticated validation of our theory's performance against observations.
+In general the bulk statistics match our expectations based on previous idealized and theoretical analysis, with the possible exception that inclusion of uncertainty shifts grass towards a slightly more negative response than the theory and Figure \ref{sign} would suggest.
+The bulk statistics motivate a more thorough examination of the structure of uncertainty and more sophisticated validation of our theory's performance against observations.
 
@@
 \label{testing}
-We now turn to an direct evaluation of the functional forms presented in Section \ref{sign_term} against observations taken from a wide range of eddy-covariance datasets (Section \ref{methods}).
+We now turn to an direct evaluation of the functional forms presented in Section \ref{sign_term} against
+
+t observations taken from a wide range of eddy-covariance datasets (Section \ref{methods}).
 Figure \ref{agu_fig} shows the distribution of of the sign term (when uncertainty is included), as compared to what the theory would predict.
 Our goal was to capture the leading order behavior of the ET dependence on VPD.
-Given the assumptions we have made, and the uncertainties of flux tower observations themselves, we expect a large amount of noise when reprdrocuing the derivatives of ET.
+Given the assumptions we have made, and the uncertainties of flux tower observations themselves, we expect a large amount of noise when reproducing the derivatives of ET.
 
-For DBF our theory is successful towards this goal: the theory matches the leading order behavior of the function when uncertainty is included, and the DBG figure looks as one would expect if one just added noise on top of our theory.
+For DBF our theory is successful towards this goal: the theory matches the leading order behavior of the function when uncertainty is included, and the DBF figure looks as one would expect if one just added noise on top of our theory.
 The VPD dependence, given by the slope parameter ($g_1$), follows the median values of each box plot.
@@
 Therefore the sign of the ET response to increases in VPD should be well matched, subject to the unavoidable constraints of noise, much of which may be from the observations themselves.
-The uncertainty is non negligable; there are many observations in each box for which the the sign of observation is opposite the response predicted by the theory, but to leading order our theory matches the observations.
+The uncertainty is non negligible; there are many observations in each box for which the the sign of observation is opposite the response predicted by the theory, but to leading order our theory matches the observations.
 
@@
 The VPD-independent, strongly negative response, which is determined by $uWUE$, is also captured.
-For CSH, there is extremely rare occurence of positive ET response with VPD, even with uncertainty, so the sign of the theory almost always matches the sign of the observations, even with all of the uncertainty.
+For CSH, there is extremely rare occurrence of positive ET response with VPD, even with uncertainty, so the sign of the theory almost always matches the sign of the observations, even with all of the uncertainty.
 
 As Section \ref{stats_sec} suggested, the theory does not match as well for CRO as for DBF and CSH.
-At low VPD the VPD dependence (due to $g_1$) is similiar to the leading order behavior of the observations, but the theory's curve is shifted towards postiive values relative to the observations, due to some error in $uWUE$.
-Most strikingly, the observations show a tendency to move back towards negative values at high VPD, which given the functional form of VPD dependnce, is behavior that is impossible for the theory to capture.
-This general postive bias the theory explains  the shift towards negative ET reponse observed in the bulk statistics of Section \ref{stats_sec}.
+At low VPD the VPD dependence (due to $g_1$) is similar to the leading order behavior of the observations, but the theory's curve is shifted towards positive values relative to the observations, due to some error in $uWUE$.
+Most strikingly, the observations show a tendency to move back towards negative values at high VPD, which given the functional form of VPD dependence, is behavior that is impossible for the theory to capture.
+This general positive bias the theory explains  the shift towards negative ET response observed in the bulk statistics of Section \ref{stats_sec}.
 The divergence between the theory and the observations for CRO could have been expected.
@@
 These deficiencies would explain the inability of our theory to match observations.
-For example, a superposition of sites with C3 plants (at low enviromental VPD) and C4 plants (at high environmental VPD) would explain observed shift back towards negative ET reponse at high VPD when all sites are binned together, as in Figure \ref{agu_fig}.
+For example, a superposition of sites with C3 plants (at low environmental VPD) and C4 plants (at high environmental VPD) would explain observed shift back towards negative ET response at high VPD when all sites are binned together, as in Figure \ref{agu_fig}.
 
-The theory for GRA suffers from similiar limitations as for CRO.
-GRA observations are characterized by a consistent trend back towards negative ET reponse at higher VPD, which the functional form of our theory is incapable of accounting for.
-As copared to CRO, the divergence between the theory and the observations is even greater for GRA, but these differences are explainable using similiar logic as presented for CRO.
-GRA will also have differences between sites in C3 vs C4 photosythesis, and irrigation practices, as well as strong seasonal variability in crop height and LAI.
-As with CRO, the trend back towards negative ET reponse at higher VPD could be explained by a superposition of sites with C3 vs C4 sites.
-Unfortunately, because detailed site-level data of C3 vs C4 plant type, as well as irrigation practes, was difficult to obtain, we could not bin GRO and GRA further by irrigation and photosyethesis stragtegy.
-However, we hypothesize that the theory would validate against obsercations much better when photosythesis strategy and irrigation are accounted for.
+The theory for GRA suffers from similar limitations as for CRO.
+GRA observations are characterized by a consistent trend back towards negative ET response at higher VPD, which the functional form of our theory is incapable of accounting for.
+As compared to CRO, the divergence between the theory and the observations is even greater for GRA, but these differences are explainable using similar logic as presented for CRO.
+GRA will also have differences between sites in C3 vs C4 photosynthesis, and irrigation practices, as well as strong seasonal variability in crop height and LAI.
+As with CRO, the trend back towards negative ET response at higher VPD could be explained by a superposition of sites with C3 vs C4 sites.
+Unfortunately, because detailed site-level data of C3 vs C4 plant type, as well as irrigation practices, was difficult to obtain, we could not bin CRO and GRA further by irrigation and photosynthesis strategy.
+However, we hypothesize that the theory would validate against observations much better when photosynthesis strategy and irrigation are accounted for.
 
 The comparison between the theory and observations for ENF is unique from the rest of the PFTs.
-The VPD dependence of the theory, determined by the slope paramter ($g_1$), is not well captured by observations.
+The VPD dependence of the theory, determined by the slope parameter ($g_1$), is not well captured by observations.
 However, the theory closely captures the x-intercept and VPD$_{crit}$.
-So, the theory still captures when the ET reponse will be positive or negative, but is less accurate with the magnitude of the reponse.
+So, the theory still captures when the ET response will be positive or negative, but is less accurate with the magnitude of the response.
 Therefore the theory is still useful for addressing when VPD drives or reduces ET.
@@
 
-While the above discussion shows that our theory has some limitations when applied to some PFTs, especially for crops and grasslands, it does well for DBF and CSH PFTs, and captures the sign of the reponse conistently for ENF.
-It is useful to also compare our theory, the observed uncertainty in the reponse, and PET, as PET is ofted used as a proxy for ET in hydrometerological studies and drought indicies.
-Again, as in Section \ref{sign_func}, the PET response to changes in VPD will be constant at about 3.5.
-In Figure \ref{agu_fig} the ET reponse to shifts in dryness is always far more negative than 3.5 for the observations, even with all uncertainty included.
-So the PET reponse to atmospheric dryness is outside the range of all observable responses.
+While the above discussion shows that our theory has some limitations when applied to some PFTs, especially for crops and grasslands, it does well for DBF and CSH PFTs, and captures the sign of the response consistently for ENF.
+It is useful to also compare our theory, the observed uncertainty in the response, and PET, as PET is often used as a proxy for ET in hydrometeorological studies and drought indices.
+Again, as in Section \ref{sign_term}, the PET response to changes in VPD will be constant at about 3.5.
+In Figure \ref{agu_fig} the ET response to shifts in dryness is always far more negative than 3.5 for the observations, even with all uncertainty included.
+So the PET response to atmospheric dryness is outside the range of all observable responses.
 This casts doubt on PET's usefulness for capturing the leading order behavior of water budgeting at the land surface.
@@
 \subsubsection{Testing VPD$_{crit}$}
-As stated in the above discussion of ENF, we are mostly interested in the critical threshold $VPD_{crit}$ for which the sign of the ET reponse to VPD ($\frac{\partial \; ET}{\partial \; VPD}$) is determined.
-When VPD is less than VPD$_{crit}$, ET reponse to VPD is negative ($\frac{\partial \; ET}{\partial \; VPD}  < 0$, plant response dominates), and when VPD is greater than VPD$_{crit}$, ET reponse to VPD is positive  ($\frac{\partial \; ET}{\partial \; VPD} > 0$ atmospheric demand dominates).
+As stated in the above discussion of ENF, we are mostly interested in the critical threshold $VPD_{crit}$ for which the sign of the ET response to VPD ($\frac{\partial \; ET}{\partial \; VPD}$) is determined.
+When VPD is less than VPD$_{crit}$, ET response to VPD is negative ($\frac{\partial \; ET}{\partial \; VPD}  < 0$, plant response dominates), and when VPD is greater than VPD$_{crit}$, ET response to VPD is positive  ($\frac{\partial \; ET}{\partial \; VPD} > 0$ atmospheric demand dominates).
 Theoretically this $VPD_{crit}$ is only a function of PFT.
@@
 Figure \ref{real} generally confirms expectations from Figure \ref{agu_fig}.
-By comparing different columns of Figure \ref{real} we can evaluate the impact of $\sigma$ and aerodynamic conductance ($g_a$) on the variability of ET reponse ($\frac{\partial \; ET}{\partial \; VPD}$).
-Aerodyamic conducntance's scaling (included in columns 1 and 3) alters the magnitude considerably, particularly between PFTs.
+By comparing different columns of Figure \ref{real} we can evaluate the impact of $\sigma$ and aerodynamic conductance ($g_a$) on the variability of ET response ($\frac{\partial \; ET}{\partial \; VPD}$).
+Aerodynamic conductance's scaling (included in columns 1 and 3) alters the magnitude considerably, particularly between PFTs.
 $\sigma$ variability (included in columns 1 and 2) adds a lot of additional noise to the signal of $\frac{\partial \; ET}{\partial \; VPD}$, as observed in Figure \ref{agu_fig}.
@@
 
-While $g_a$ has a large impact on the magnitude of the scaling, temperature, through Clausius-Clapeyron ($\Delta$), does not appear to have as much of an impact in the complete estimation of ET reponse ($\frac{\partial \; ET}{\partial \; VPD}$).
+While $g_a$ has a large impact on the magnitude of the scaling, temperature, through Clausius-Clapeyron ($\Delta$), does not appear to have as much of an impact in the complete estimation of ET response ($\frac{\partial \; ET}{\partial \; VPD}$).
 Therefore it appears that the absolute magnitude of the derivative can be mostly attributed to the sign term, and its direct scaling by aerodynamic conductance.
 Intuitively this is reasonable; aerodynamic conductance will control how dominant balances at the land surface are communicated to the boundary layer.
-While $\Delta$ controls the efficiency of latent fluxes in carrying heat away from the surface, it appears this is a second order term, relative to $g_a$, for scaling ET reponse.
+While $\Delta$ controls the efficiency of latent fluxes in carrying heat away from the surface, it appears this is a second order term, relative to $g_a$, for scaling ET response.
 
@@
 
-\section{Limitations of theory}
-\explain{HERE DISCUSS LIMITATION OF THE METHODS TO BVERY DRY CONDITIONS AND THE FEEDBCAK TO THE ATMOSPHERE OF VERY DRY SOIL CONDITIONS (BERG ET AL 2016, GENTINE ET AL> 2016)}
+\section{Limitations of theory - very dry soil moisture conditions}
+In formulating our theory with Penman Monteith, we implicitly do not account for very dry soil moisture conditions.
+For the majority of environmental conditions, we argue that soils provide enough water so that plants are not significantly stressed.
+We posit that the land surface will generally optimize to contain plants living in conditions which they evolved for.
+However, in extreme conditions and drought scenarios soil water content (SWC) could become the limiting factor for ET response to VPD, which our theory does not account for.
+Within our framework, any systematic bias due to the  failure to account for SWC's effects in very dry conditions should manifest itself in a functional relationship between $\sigma$, our uncertainty measure, and SWC, which is observed at the FLUXNET sites.
+Examining the relationship between $\sigma$ and SWC will test to what extent our theory breaks down in very dry soil moisture conditions.
+
+If we view $\sigma \cdot uWUE$ as, in addition to a measure of uncertainty, a short time-scale observation of uWUE, we would expect $\sigma \cdot uWUE$ to decrease at low SWC as water use becomes more inefficient.
+If $\sigma \cdot uWUE$ is a very strong function of SWC, then our should be conditioned more strongly on well watered soil conditions.
+If $\sigma \cdot uWUE$ is weakly a function of SWC, then our theory is more universal and independent of soil moisture conditions.
+
+Indeed, for all PFTs, there is some dependence of $\sigma \cdot uWUE$ on SWC, so there are some limitations to our theory, especially at low SWC (Figure \ref{swc}).
+The portion of observations for which our theory is biased by SWC-limitations varies by PFT, due to the nonlinear threshold at which soil moisture becomes a factor.
+For CRO and DBF, soil water content only matters for about the lowest 5\% of observations (each box is 5\% of observations in Figure \ref{swc}).
+And, CRO, relative to DBF, has a very weak dependence on soil moisture, which is reflective of the high likelihood that CRO sites are irrigated and not water stressed.
+For ENF, systematic SWC biases in $\sigma \cdot uWUE$ emerge in about the lowest 30\% of SWC conditions.
+Whiel for GRA, the relationship between $\sigma \cdot uWUE$ and SWC is more linear and affects a greater portion of observations, about 60\% of observations.
+Clearly, for GRA soil water frequently limits ET and its response, and our theory is limited.
+It is therefore no suprise that for GRA our theory tested worst againt the data (Section \ref{testing}); for all other PFTs occurence of soil mosture impacts was rare enough to not manifest itself in bulk statistics and plots\footnote{note that interpreting the soil moisture dependence for CSH is difficult as their are only two sites for that PFT.
+So systematic soil moisture relationshps are not apparent in Figure \ref{swc}, but instead blurred by the superposition of the two sites.
+We argue taht the soil moisture dependence is still there, but we don't have enough statistcs to comment on it}.
+
+The observed dependence of $\sigma \cdot uWUE$ on SWC for GRA would explain the deficiencies of our theory compared to the observations in Section \ref{testing}, specifically the trend back towards negative soil moisture at high VPD.
+Feedbacks between the land surface and the atmosphere, which our not accounted for due to our focus on the one-way response of the land surface to atmospheric conditions, would cause high VPD to be correlated with low SWC \citep[][Gentine et al.
+GRL 2016]{Berg_2016}\explain{Pierre, which publication were you thinking, your Evaporation from Weather Data one?}.
+So, at high VPD observations of low SWC are more likely, and this low SWC causes a lower water use efficiency (manifested by $\sigma \cdot uWUE$).
+The lower water use efficiency at low SWC/high VPD then leads to the observed downturn towards decreasing ET with increasing VPD, and the deviation between our theory (based on a constant uWUE assumption) and observations.
+
+To summarize, our theory is limited by its inability to account for soil water impacts on land surface response, and feedbacks between SWC and VPD.
+Fortunately, for most PFTs SWC's effect on ET is relatively rare (<30\% of observations) and does not manifest itself in the majority of observations and bulk statistics.
+However, for GRA SWC has a strong effect decreasing water use efficiency for the majority of the observations.
+The soil moisture effects explains the coefficients of our theory in Section \ref{testing}, when one considers feedbacks between the surface and the atmospheres, and coupling between SWC and VPD.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{./swc_boxplot.pdf}
+\caption{$\sigma \cot uWUE$ as a function of PFT and SWC. Each boxplot represents 5\% of all observations.}
+\label{swc}
+\end{figure}
 

### 2018-01-25 src/paper_figure_code/swc_box_plot.py

@@ -15,11 +15,14 @@ dashedlinewidth=1.4
 
 bluelinewidth=3.0
 
-def plot_box(_df, pft=False):
+def plot_box(_df, ax=None, pft=False):
   """makes a box plot"""
-  plt.close('all')
-  fig = plt.figure()
-  ax = fig.add_subplot(111)
+  if ax is None:
+    fig = plt.figure()
+    ax = fig.add_subplot(111)
+    savefig = True
+  else:
+    savefig = False
   # exclude the last 5 percent b/c usually large
   ptiles = np.arange(0.0, 91.0, 5.0)
   positions = []
@@ -40,7 +43,7 @@ def plot_box(_df, pft=False):
   orig_xlim = ax.get_xlim()
   #ax.plot(orig_xlim, [d_calc.CP/287.0, d_calc.CP/287.0], 'm--', linewidth=1.0)
   ylim = ax.get_ylim()#[-4.0, 4.0]
-  ax.set_ylabel('uWUE', fontsize=fontsize)
+  ax.set_ylabel('$\sigma \cdot$ uWUE', fontsize=fontsize)
   ax.set_xlabel('SWC', fontsize=fontsize)
   if pft:
     ax.set_title('PFT: %s'\
@@ -54,17 +57,32 @@ def plot_box(_df, pft=False):
   # ax.set_ylim(ylim)
 
   # ax.set_xlim(orig_xlim)
+  if savefig:
+    plt.tight_layout()
+    if pft:
+      plt.savefig('%s/climate_et/swc_box/pft/%s_box.png'\
+                  % (os.environ['PLOTS'], _df.pft.iloc[0]))
+    else:
+      plt.savefig('%s/climate_et/swc_box/%s_%s_box.png'\
+                  % (os.environ['PLOTS'], _df.pft.iloc[0], _df.site.iloc[0]))
+  return
+
+def scaling_wrapper(_df):
+  """wrapper that groups by pft and does scaling plot"""
+  fig = plt.figure()
+  fig.set_figheight(fig.get_figheight()*3)
+  fig.set_figwidth(fig.get_figwidth()*2)
+  for i, pft in enumerate(mean_df.index.values):
+    print(pft)
+    ax = fig.add_subplot(3, 2, i+1)
+    plot_box(_df.loc[(_df.pft==pft), :], ax=ax, pft=True)
   plt.tight_layout()
-  if pft:
[…]

### 2018-01-25 doc/paper/vpd_et_paper_PG_v1.tex
what I sent to Pierre

@@
 In turn, this would further increase aridity as low soil moisture levels increase the Bowen ratio, and in turn increase temperature and reduce atmospheric humidity \citep[gentine et al.
-GRL 2016][]{Berg_2016}.
+GRL 2016][]{Berg_2016}\explain{Pierre, which publication were you thinking, your Evaporation from Weather Data one?}.
 This however would not seem to be a sensible strategy for plants from an evolutionary standpoint.
@@
 \label{methods}
-The Penman-Monteith equation \citep [hereafter PM][]{Monteith_1965} estimates ET as a function of observable atmospheric variables and surface conductances:
+The Penman-Monteith equation \citep [hereafter PM,][]{Monteith_1965} estimates ET as a function of observable atmospheric variables and surface conductances:
 \begin{linenomath*}
@@
 \end{linenomath*}
-where $\Delta$ is the change in saturation vapor pressure with temperature ($\frac{\partial \; e_s}{\partial \; T}$), $R$ is the net radiation minus ground heat flux, $g_a$ is aerodynamic conductance, $\rho_a$ is air density, $c_p$ is specific heat of air at constant pressure, $\gamma$ is the psychometric constant, and $g_s$ is the stomatal conductance (Table 1).
+where $\Delta$ is the change in saturation vapor pressure with temperature, given by Clausius-Clapeyron ($\frac{d \; e_s}{d \; T}$), $R$ is the net radiation minus ground heat flux, $g_a$ is aerodynamic conductance, $\rho_a$ is air density, $c_p$ is specific heat of air at constant pressure, $\gamma$ is the psychometric constant, and $g_s$ is the stomatal conductance (Table 1).
 \citet{MEDLYN_2011} developed a model for stomatal conductance ($g_s$) by combining optimal photosynthesis theory \citep{Farquhar_1980, Katul_2010} with empirical approaches, which describes the dependence of $g_s$ to VPD.
@@
 
-While Equation 3 can be used in PM (equation 1), it will make analytical work with the function intractable because $A$, net CO$_2$ assimilation, is functionally related to ET itself.
-To remove the dependence of ET on $A$ we can use the semi-empirical results of \citet{Zhou_2015}.
-\citet{Zhou_2015} showed that the underlying Water Use Efficiency $uWUE$:
+While Equation 3 can be used in PM (Equation 1), it will make analytical work with the function intractable because $A$, net CO$_2$ assimilation, is functionally related to ET itself.
+To remove the dependence of ET on $A$ we can use the semi-empirical results of \citet{Zhou_2015}, who showed that the underlying Water Use Efficiency $uWUE$:
 
@@
 
-and thus we can evaluate departure from our theory (as a departure of $\sigma$ from unity) The variability of $\sigma$ across sites and time provides a measure of uncertainty in our model, assumptions, as well as the FLUXNET observations themselves.
+and thus we can evaluate departure from our theory (as a departure of $\sigma$ from unity).
+The variability of $\sigma$ across sites and time provides a measure of uncertainty in our model, assumptions, as well as the FLUXNET observations themselves.
 To correct for differences in sampling between \cite{Zhou_2015} and our data, we set uWUE such that $\overline{\sigma} = 1$.
@@
 There are a few subtleties to taking the derivative in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$) and $VPD$ are functionally related, so while taking the derivative we evaluate $\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET} {\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD} \Big|_{\text{RH fixed}} + \frac{\partial \; ET}{\partial \; RH} \frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$ fixed}}$.
-$RH$ and $e_s$ are assumed to be approximately orthogonal.
+$RH$ and $e_s$ are assumed to be approximately orthogonal, which is supported by the data.
 
 We note one final comment on our derivation which is relevant for future projects and analysis.
-If we approximate $c_s$ at a global mean CO$_2$ concentration, then the RHS of Equation \ref{et} is fully defined using commonly available weather station data and the constants published in \citet{Zhou_2015, Lin_2015}.
+If we approximate $c_s$ at a global mean CO$_2$ concentration, then the RHS of Equation \ref{et} is fully defined using commonly available weather station data and the constants published in \citet{Zhou_2015} and \citet{Lin_2015}.
 This makes Equation \ref{et} a useful alternative to PET in drought indices and hydrometeorological analysis for vegetated surfaces.
@@
 Before calculating the sensitivity of ET to VPD, we will consider the functional form of Equation \ref{d_et}.
-There are two main terms: a ``scaling'' term, which modifies the magnitude but not the sign of the derivative $\partial ET/\partial VPD$:
+There are two main terms: a ``scaling'' term, which modifies the magnitude but not the sign of the ET response to VPD  ($\frac{\partial \; ET}{\partial \; VPD}$):
 
@@
 
-and a ``sign'' term, which determines whether the derivative is positive i.e. atmospheric demand driven or negative i.e. physiologically controlled:
+and a ``sign'' term, which determines whether ET increases or decreases with VPD  (i.e. atmospheric demand driven or physiologically controlled):
 
@@
 However, if the first term is larger then atmospheric demand dominates and ET increases with increasing VPD.
-Given this sign
+Given that the sign term directly address the title of this manuscript, we will examine its functional form first.
 
@@
 CO$_2$ concentration ($c_s$) and the psychometric constant ($\gamma$) are relatively constant to a good approximation so that the variability is dominated by $\sigma$ and $VPD$.
-$uWUE$ could vary with soil moisture but has been shown to be relatively constant except in very dry soil conditions \citep{Zhou_2016, Lin_2015}.
+$uWUE$ could vary with soil moisture but has been shown to be relatively constant except in very dry soil conditions  (ZHOU ET AL, LIN ET AL)\explain{Pierre, which Zhou and Lin citations were you referring to that show that uWUE changes with very dry soil conditions?}.
 We thus further assume that the soil is not too dry so that $uWUE$ can be approximated as a constant (within PFT).
@@
 Even in the absence of soil pore evaporation, stomata do not shut down entirely at very high VPD and ET does not go to zero, because stomata are still slightly open to perform some photosynthesis \citep{Ball_1987, Leuning_1990, MEDLYN_2011}.
-In addition, upward xylem transport is necessary to maintain phloem transport and thus carbon allocation (MENCUCCINI) \explain{Not sure citation  - Comparative criteria for models of the vasuclar transport systems of tall trees}.
+In addition, upward xylem transport is necessary to maintain phloem transport and thus carbon allocation (MENCUCCINI) \explain{Not sure citation  - Comparative criteria for models of the vasuclar transport systems of tall trees?}.
 
 \begin{table}  
-  \label{vpd_crit}
 \caption{Values of $VPD_{crit}$, where $\frac{\partial \; ET}{\partial \; VPD} = 0$, evaluated at PFT average values for $R_{air}$, $\sigma$, $\gamma$, and $c_s$.
@@
 \end{tabular}
+\label{vpd_crit}
 \end{table}
@@
 Plants that are less willing to trade water for new growth (small $g_1$) will have strong tendency to reduce ET with increasing VPD, and they will be very effective in reducing ET especially at low VPD.
-However, at very high VPD gradients in vapor pressure at the leaf scale will begin to dominate leaf-scale plant strategies of stomatal closure (parameterized with $g_1$), so ET response will begin to asymptote towards ecosystem-scale values irregardless of leaf-scale strategy and stomatal closure (through $g_1$).
-Therefore, plants with a low $g_1$ will have the largest VPD dependence of ET response because the difference in ET response at low VPD (leaf response dominates) and high VPD (VPD gradient dominates) is largest.
+However, at very high VPD gradients in vapor pressure at the leaf scale will begin to dominate leaf-scale plant strategies of stomatal closure (parameterized with $g_1$), so ET response will begin to asymptote towards ecosystem-scale values irregardless of leaf-scale strategy and stomatal closure (parameterized with $g_1$).
+Therefore, plants with a low $g_1$ will have the largest VPD dependence of ET response because the difference in ET response at low VPD (leaf stomatal response dominates) and high VPD (VPD gradient dominates) is largest.
 This is apparent in the strong VPD dependence of ENF, which has the lowest slope parameter ($g_1=74.31$), in Figure \ref{term3}.
@@
 ENF are less willing to trade water for new growth at the leaf scale, so the difference between the response at low VPD when leaf response dominates and high VPD when vapor pressure gradients dominate is much greater.
-CRO on the otherhand are more willing to trade water for new growth, so the leaf scale response is much smaller, and the difference in ET response between low VPD and high VPD is relatively low.
+CRO on the otherhand are more willing to trade water for new growth, so the leaf scale response is much smaller, and the difference in ET response between low VPD and high VPD is relatively small.
 
@@
 To better illustrate this, the ranges of observed environmental VPDs at the FLUXNET sites are plotted parallel to the x-axis.
-For CSH, VPD is always less than VPD$_{crit}$ so that the plant response dominates and empathizes the water conservative strategy of those plants.
+For CSH, VPD is always less than VPD$_{crit}$ so that the plant response dominates, emphasizing the water conservative strategy of those plants.
 For CRO on the otherhand, VPD is almost always higher than VPD$_{crit}$, emphasizing that those plants are water intensive and were actually engineered for photosynthesis rather than water saving.
@@
 \textbf{***I think I'm going to remove the bottom plot: its discussion didn't fit in well with the text, and the box plots test inclusion of uncertainty***}}
-\label{term3}
+\label{sign_plot}
 \end{figure}
@@
 
-$\frac{P}{T}$ is an air-density term, which we would expect to vary relatively little compared to aerodynamic conductance and $\Delta$.
+$\frac{P}{T}$ is an air-density term, which we would expect to vary relatively little compared to aerodynamic conductance and Clausius-Clapeyron ($\Delta$).
 The psychometric constant ($\gamma$) is also relatively constant, so the scaling term should be primarily a function of aerodynamic conductance and temperature, through the Clausius-Clapeyron relationship $\Delta$.
@@
 
-The control of scaling term variability between PFTS by aerodynamic conductance is confirmed by the data (Figure \ref{scale_vary}).
+The control of scaling term variability between PFTs by aerodynamic conductance is confirmed by the data (Figure \ref{scale_vary}).
 Differences between PFT are almost entirely due to differences in aerodynamic conductance, rather than differences in observed temperature ranges.
-The Scaling term for the tree PFTs (DBF, ENF) is generally more than double the scaling terms for the other PFTs which have lower surface roughness and generally smaller aerodynamic conductance.
+The scaling term for the tree PFTs (DBF, ENF) is generally more than double the scaling terms for the other PFTs which have lower surface roughness and generally smaller aerodynamic conductance.
 
@@
 
-To summarize, variability between PFTs is mostly controlled by systematic differences in aerodynamic conductance, due to differences in surface roughness, between each PFT.
+To summarize, variability between PFTs is mostly controlled by systematic differences in aerodynamic conductance, due to differences in surface roughness between each PFT.
 In contrast, variability within PFT is also controlled by temperature, through Clausius-Clapeyron.
@@
 Previous sections only considered idealized scenarios without uncertainty ($\sigma = 1.0$).
-In this section we will consider bulk statistics, as a function of PFT, of ET response when $\sigma$ is allowed to vary and eddy-covariance derived uncertainty is included (Section \ref{methods}).
-These bulk statistics of ET response confirm our expectations based on the theory presented in Sections \ref{sign_term}-\ref{scale_term} (Table \ref{stats}.
+In this section we will consider bulk statistics of ET response, as a function of PFT, when $\sigma$ is allowed to vary and eddy-covariance derived uncertainty is included (Section \ref{methods}).
+These bulk statistics of ET response confirm our expectations based on the theory presented in Sections \ref{sign_term}-\ref{scale_term} (Table \ref{stats}).
 For all PFTs, mean ET response to increasing VPD is negative.
@@
 $\sigma$, $T$, $c_s$, $VPD$) is only negative for CSH and GRA.
-This difference in mean ET response as compared to the ET response at mean environmental conditions is due to the non-linear nature of the response, in which negative responses are generally much larger magnitude than positive responses (Figure \ref{sign}).
+This difference in mean ET response as compared to the ET response at mean environmental conditions is due to the non-linear nature of the response, in which negative responses are generally much larger magnitude than positive responses (Figure \ref{sign_plot}).
 
@@
 
-In general the bulk statistics match our expectations based on previous idealized and theoretical analysis, with the possible exception that inclusion of uncertainty shifts grass towards a slightly more negative response than the theory and Figure \ref{sign} would suggest.
+In general the bulk statistics match our expectations based on previous idealized and theoretical analysis, with the possible exception that inclusion of uncertainty shifts grass towards a slightly more negative response than the theory and Figure \ref{sign_plot} would suggest.
 The bulk statistics motivate a more thorough examination of the structure of uncertainty and more sophisticated validation of our theory's performance against observations.
@@
 \begin{table}
-  \label{stats}
 \caption{Statistics of $\frac{\partial \; ET}{\partial \; VPD}$ as a function of PFT. \textbf{***dated, need to update these statistics***}}
@@
 \hline
-
 \end{tabular}
+  \label{stats}
 \end{table}
@@
 \label{testing}
-We now turn to an direct evaluation of the functional forms presented in Section \ref{sign_term} against
-
-t observations taken from a wide range of eddy-covariance datasets (Section \ref{methods}).
+We now turn to an direct evaluation of the functional forms presented in Section \ref{sign_term} against observations taken from a wide range of eddy-covariance datasets (Section \ref{methods}).
 Figure \ref{agu_fig} shows the distribution of of the sign term (when uncertainty is included), as compared to what the theory would predict.
@@
 The VPD-independent, strongly negative response, which is determined by $uWUE$, is also captured.
-For CSH, there is extremely rare occurrence of positive ET response with VPD, even with uncertainty, so the sign of the theory almost always matches the sign of the observations, even with all of the uncertainty.
+For CSH, there is extremely rare occurrence of positive ET response with VPD, even with uncertainty, so the sign of the theory almost always matches the sign of the observations.
 
 As Section \ref{stats_sec} suggested, the theory does not match as well for CRO as for DBF and CSH.
-At low VPD the VPD dependence (due to $g_1$) is similar to the leading order behavior of the observations, but the theory's curve is shifted towards positive values relative to the observations, due to some error in $uWUE$.
+At low VPD the VPD dependence (due to $g_1$) is similar to the leading order behavior of the observations, but the theory's curve is shifted towards positive values relative to the observations, due to some bias in $uWUE$.
 Most strikingly, the observations show a tendency to move back towards negative values at high VPD, which given the functional form of VPD dependence, is behavior that is impossible for the theory to capture.
-This general positive bias the theory explains  the shift towards negative ET response observed in the bulk statistics of Section \ref{stats_sec}.
+This general positive bias of the theory explains the shift towards negative ET response observed in the bulk statistics of Section \ref{stats_sec}.
 The divergence between the theory and the observations for CRO could have been expected.
@@
 Also, it's with noting that we used published constants for $g_1$ \citep[from ][]{Lin_2015}, which determines the VPD dependence.
-We could have fit our own constants so that the theory would best match the observations, but our goal was to use a priori physical constants whenever possible.
+We could have fit our own constants so that the theory would best match the observations, but our goal was to use a priori results whenever possible.
 
 While the above discussion shows that our theory has some limitations when applied to some PFTs, especially for crops and grasslands, it does well for DBF and CSH PFTs, and captures the sign of the response consistently for ENF.
-It is useful to also compare our theory, the observed uncertainty in the response, and PET, as PET is often used as a proxy for ET in hydrometeorological studies and drought indices.
+It is useful to also compare our theory, the observed uncertainty in the ET response, and PET, as PET is often used as a proxy for ET in hydrometeorological studies and drought indices.
 Again, as in Section \ref{sign_term}, the PET response to changes in VPD will be constant at about 3.5.
@@
 
-In general Figures \ref{real} and \ref{agu_fig} match our expectations.
+In general Figures  \ref{agu_fig} and \ref{real} match our expectations.
 Our $\sigma$-based method of uncertainty propagation blurs the idealized expectations the most for GRA and CRO.
@@
 For the majority of environmental conditions, we argue that soils provide enough water so that plants are not significantly stressed.
-We posit that the land surface will generally optimize to contain plants living in conditions which they evolved for.
+We posit that ecosystems will generally optimize to host plants living in conditions which they evolved for.
 However, in extreme conditions and drought scenarios soil water content (SWC) could become the limiting factor for ET response to VPD, which our theory does not account for.
@@
 
-If we view $\sigma \cdot uWUE$ as, in addition to a measure of uncertainty, a short time-scale observation of uWUE, we would expect $\sigma \cdot uWUE$ to decrease at low SWC as water use becomes more inefficient.
-If $\sigma \cdot uWUE$ is a very strong function of SWC, then our should be conditioned more strongly on well watered soil conditions.
+If we view $\sigma \cdot uWUE$ as, in addition to a measure of uncertainty, a short time-scale observation of uWUE, we would expect $\sigma \cdot uWUE$ to decrease at low SWC (ZHOU ET AL).
+If $\sigma \cdot uWUE$ is a very strong function of SWC, then our theory should be conditioned more strongly on well watered soil conditions.
 If $\sigma \cdot uWUE$ is weakly a function of SWC, then our theory is more universal and independent of soil moisture conditions.
@@
 Indeed, for all PFTs, there is some dependence of $\sigma \cdot uWUE$ on SWC, so there are some limitations to our theory, especially at low SWC (Figure \ref{swc}).
-The portion of observations for which our theory is biased by SWC-limitations varies by PFT, due to the nonlinear threshold at which soil moisture becomes a factor.
+The portion of observations for which our theory is biased by SWC-limitations varies by PFT, due to the nonlinear threshold at which soil moisture availability limits plant function.
 For CRO and DBF, soil water content only matters for about the lowest 5\% of observations (each box is 5\% of observations in Figure \ref{swc}).
-And, CRO, relative to DBF, has a very weak dependence on soil moisture, which is reflective of the high likelihood that CRO sites are irrigated and not water stressed.
-For ENF, systematic SWC biases in $\sigma \cdot uWUE$ emerge in about the lowest 30\% of SWC conditions.
-Whiel for GRA, the relationship between $\sigma \cdot uWUE$ and SWC is more linear and affects a greater portion of observations, about 60\% of observations.
-Clearly, for GRA soil water frequently limits ET and its response, and our theory is limited.
-It is therefore no suprise that for GRA our theory tested worst againt the data (Section \ref{testing}); for all other PFTs occurence of soil mosture impacts was rare enough to not manifest itself in bulk statistics and plots\footnote{note that interpreting the soil moisture dependence for CSH is difficult as their are only two sites for that PFT.
+And, CRO, relative to DBF, has a very weak dependence on soil moisture, which is reflective of the high likelihood that CRO sites are optimally irrigated and not water stressed.
+For ENF, systematic SWC-induced biases in $\sigma \cdot uWUE$ emerge in about the lowest 30\% of SWC conditions.
+In crontast to ENF, CRO, and DBF, for GRA the relationship between $\sigma \cdot uWUE$ and SWC is more linear and affects a greater portion of observations; about 60\% of observations.
+Clearly, for GRA soil water frequently impacts plant function and alters ET response, and our theory is limited.
+It is therefore no suprise that for GRA our theory tested the worst againt the data (Section \ref{testing}); for all other PFTs occurence of soil mosture impacts was rare enough to not manifest itself in bulk statistics and plots\footnote{note that interpreting the soil moisture dependence for CSH is difficult as their are only two sites for that PFT.
 So systematic soil moisture relationshps are not apparent in Figure \ref{swc}, but instead blurred by the superposition of the two sites.
@@
 GRL 2016]{Berg_2016}\explain{Pierre, which publication were you thinking, your Evaporation from Weather Data one?}.
-So, at high VPD observations of low SWC are more likely, and this low SWC causes a lower water use efficiency (manifested by $\sigma \cdot uWUE$).
-The lower water use efficiency at low SWC/high VPD then leads to the observed downturn towards decreasing ET with increasing VPD, and the deviation between our theory (based on a constant uWUE assumption) and observations.
+So, at high VPD observations of low SWC are more likely, and this low SWC causes a lower uWUE (ZHOU ET AL).
+The lower uWUE at low SWC/high VPD then leads to the observed downturn towards decreasing ET with increasing VPD, and the deviation between our theory (based on a constant uWUE assumption) and observations.
 
@@
 However, for GRA SWC has a strong effect decreasing water use efficiency for the majority of the observations.
-The soil moisture effects explains the coefficients of our theory in Section \ref{testing}, when one considers feedbacks between the surface and the atmospheres, and coupling between SWC and VPD.
+Soil moisture effects explain the deficiencies of our theory in Section \ref{testing}, particularly for GRA.
 
@@
 In some ways we should have expected this, as our theory does not account for large sources of variability we would expect for those plant types, including varying surface roughness, c3 vs c4 photosynthesis, and potential differences in irrigation across sites.
+Also, our theory fails to account for soil water effects on plant function, which non-negligibly alters ET response for GRA.
 Given these deficiencies for CRO and GRA, we will withhold conclusions on their response to increasing VPD.

### 2018-01-30 src/paper_figure_code/fig04b.py
about to clean everything for final paper edits with pierre's comments

@@ -31,7 +31,6 @@ def plot_scaling(_df, ax, savefig=False):
     plt.savefig('../../doc/paper/fig04b.pdf')
   return
 
-
 def scaling_wrapper(_df):
   """wrapper that groups by pft and does scaling plot"""
   fig = plt.figure()

### 2018-01-30 doc/paper/vpd_et_paper.tex
done with wed mon's writing session
starting final edits

@@
 
-%% ------------------------------------------------------------------------ %%
-%  Title
-% 
-% (A title should be specific, informative, and brief. Use
-% abbreviations only if they are defined in the abstract. Titles that
-% start with general keywords then specific terms are optimized in
-% searches)
-%
-%% ------------------------------------------------------------------------ %%
-
-% Example: \title{This is a test title}
-
 \title{When does vapor pressure deficit drive or reduce evapotranspiration?}
 
-%% ------------------------------------------------------------------------ %%
-%
-%  AUTHORS AND AFFILIATIONS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Authors are individuals who have significantly contributed to the
-% research and preparation of the article. Group authors are allowed, if
-% each author in the group is separately identified in an appendix.)
-
-% List authors by first name or initial followed by last name and
-% separated by commas. Use \affil{} to number affiliations, and
-% \thanks{} for author notes.  
-% Additional author notes should be indicated with \thanks{} (for
-% example, for current addresses).
-
-% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
-% Author\affil{3,4}\thanks{Also funded by Monsanto.}}
-
-\authors{A. Massmann\affil{1}, P. Gentine\affil{1}, C. Lin\affil{2}}
-
-% \affiliation{1}{First Affiliation}
-% \affiliation{2}{Second Affiliation}
-% \affiliation{3}{Third Affiliation}
-% \affiliation{4}{Fourth Affiliation}
+\authors{A. Massmann\affil{1}, P. Gentine\affil{1}, C. Lin\affil{1,2}}
 
@@
 
-%% ------------------------------------------------------------------------ %%
-%
-%  TEXT
-%
-%% ------------------------------------------------------------------------ %%
+\section{Introduction}
 
-%%% Suggested section heads:
-% \section{Introduction}
-% 
-% The main text should start with an introduction. Except for short
-% manuscripts (such as comments and replies), the text should be divided
-% into sections, each with its own heading.
+Vapor pressure deficit (VPD) is expected to rise over continents in the future due to the combination of increased temperature and relative humidity drying \citep{Byrne_2013}.
+Increases in VPD increases the atmospheric demand of evapotranspiration (ET) \citep{Monteith_1965} but also stresses plant stomata \citep{Leuning_1990, MEDLYN_2011}.
 
-% Headings should be sentence fragments and do not begin with a
-% lowercase letter or number. Examples of good headings are:
+This conflicting response between increased demand and higher stomatal stress leads to two potential but opposing perspectives on how ET adjusts to changes in VPD.
+The first, a hydrometeorological perspective, is that higher VPD increases atmospheric demand for water from the land surface, and this drives an increase in evapotranspiration (ET).
+This perspective is particularly relevant because potential evapotranspiration (PET), which is used in many drought indices and hydrometeorological studies (CITATION NEEDED), typically only quantifies changes in atmospheric demand with limited surface regulation though ecosystem resistance regualtion (Swann et al.
+2016 FOR EXAMPLE).
+On the other hand, there is another perspective for vegetated surfaces.
+Plants' stomata have evolved to optimally regulate the exchange of water and carbon, and tend to partially close in response to increased atmospheric dryness I order to save \citep{Ball_1987, Leuning_1990, MEDLYN_2011}.
+Therefore, an increase in VPD, particularly in well-watered soil conditions, may actually correspond to a decrease in ET because of stomatal closure.
+This decrease in ET would reduce and potentially cancel, in the case of full stomatal closure, the effects of shifts to atmospheric demand.
+In other words, the  question ``When does VPD drive or reduce ET?'' can be related to whether plant regulation or atmospheric demand dominates ET.
 
-% \section{Materials and Methods}
-% Here is text on Materials and Methods.
-%
-% \subsection{A descriptive heading about methods}
-% More about Methods.
-% 
-% \section{Data} (Or section title might be a descriptive heading about data)
-% 
-% \section{Results} (Or section title might be a descriptive heading about the
-% results)
-% 
-% \section{Conclusions}
+The ET response to changes in VPD alters water partitioning between the soil and atmosphere.
+If plant response reduces ET in response to atmospheric drying then soil moisture will be better conserved.
+This would seem a sensible evolutionary strategy to cope with aridity.
+If stomata were fully passive \citep [similar to soil pores, e.g.
+][]{Or_2013}, increased atmospheric aridity would strongly reduce soil moisture (BERG GRL ON SOIL MOSITURE DIFFERNT BHEVAIOR IN TOP AND DEEP SOIL).
+In turn, this would further increase aridity as low soil moisture levels increase the Bowen ratio, and in turn increase temperature and atmospheric drying \citep[gentine et al.
+GRL 2016][]{Berg_2016}\explain{Pierre, which publication were you thinking, your Evaporation from Weather Data one?
+YES.
+Also cite papers on complementary relationship BRUTSAERT 1999, BOUCHET and MORTON ETC...
+all Salvucci and Gentine PNAS}.
+This however would not seem to be a sensible strategy for plants from an evolutionary standpoint.
 
-\section{Introduction}\explain{This section needs to be fleshed out, and I definitely need to add more citations}
+We can use intuition about plant water conservation strategy to hypothesize about ET response to changes in VPD.
+Plants that evolved to conserve water, such as arid shrubs, should be more likely to reduce ET with increasing VPD, and plants that have evolved or have been engineered to care little about water, such as crops, will be more likely to increase ET with increasing VPD.
+Atmospheric conditions must matter as well.
+At the ecosystem scale, there are limits to plant water conservation strategies.
+As atmospheric demand for water (VPD) increases further, ecosystems should begin to reach their water conservation limits and might not be able to entirely limit the ET flux to the atmosphere.
+At this stage any further increase in VPD will most likely drive a (limited) increase in ET, because there is little more the plants can do keep water back (REFOMRAULTE LAST PART OF SENSTENCE).
 
-Answering the question ``When does vapor pressure deficit drive or reduce evapotranspiration?'' is motivated by two potential but opposing perspectives on the matter.
-The first perspective is that increases to vapor pressure deficit (VPD) increases atmospheric demand for water from the land surface, and this drives an increase in evapotranspiration (ET).
-This perspective is particularly relevant because potential evapotranspiration (PET), which is used in many drought indices and hydrometeorological studies, only quantifies atmospheric demand effects.
-However, there is another possible perspective for vegetated surfaces.
-Plants' stomata have evolved to optimally regulate the exchange of water and carbon, and tend to close in response to increased atmospheric dryness \citep{Ball_1987, Leuning_1990, MEDLYN_2011}.
-Therefore, an increase in VPD may actually correspond to a decrease in ET because stomatal closure can cancel the effects of shifts to atmospheric demand.
-So really a question of ``When does VPD drive or reduce ET?'' becomes a question of ``When does plant response or atmospheric demand dominate?.''
-
-Quantifying the plant response to a perturbation to atmospheric VPD increases our understanding of the land surface response to shifts in atmospheric conditions.
-If plant response reduces ET in response to atmospheric drying then soil moisture will be conserved for later ET.
-An increase in ET in response to atmospheric drying will reduce soil moisture, but contribute increased moistening to the atmosphere.
-Understanding the one way response of the land surface to the atmosphere is a critical first step to disentangling land-atmosphere feedbacks.
-The sign and magnitude of land-surface response drives the co-evolution of the atmosphere and land-surface at many timescales, from diurnal to interdecadal.
-
-Before diving into math and data, we have some prior intuition for how the system might behave.
-First, whether VPD drives or reduces ET should be a function of plant type.
-Plants that evolved to conserve water (e.g.
-arid shrubs) will be more likely to reduce ET with increasing VPD, and plants that have evolved to care little about water (e.g.
-crops) will be more likely to increase ET with increasing VPD.
-Atmospheric conditions must matter as well.
-At the ecosystem scale there are limits to the strategies plants use to hold water back from the atmosphere.
-As atmospheric demand for water (VPD) increases, ecosystems will begin to reach their water conservation limits.
-At this stage any further increase in VPD will most likely drive an increase in ET, because there is little more the plants can do keep water back.
-
-In order to quantify plant response to perturbations to atmospheric demand for water, we apply a Penman-Monteith framework to eddy-covariance observations spanning various biomes and climates.
-Section 2 describes the data used, Section 3 derives the framework, Section 4 presents results, and Section 5 discusses conclusions.
-The goal of this paper is to use reasonable approximations as a tool to increase intuition for plant response to atmospheric drying.
-This intuition will aid interpretation of observations and full complexity climate models.
+The objective of the present manuscript is thus to evaluate the VPD dependence of ET, in non-extreme soil drought conditions.
+The goal of this paper is to use reasonable approximations as a tool to drive intuition for plant response to atmospheric drying.
+This intuition will aid interpretation of observations and full complexity models.
+In order to quantify plant response to perturbations in VPD, we apply a Penman-Monteith framework to derive theoretical response of ET to VPD.
+The model is then validated and tested at multiple eddy-covariance stations spanning various climates and plant functional types.
+Section 2 describes the data used.
+Section 3 derives the framework.
+Section 4 presents results and comparison to observations.
+Section 5 discusses conclusions.
 
@@
 \label{data}
-We use data from FLUXNET2015, including all sites with more than four years of data.
-Each site's plant functional type (PFT) was classified using the International Geosphere-Biosphere Programme vegetation classification scheme (\cite{Loveland_1999}).
-Because physical constants used in Section \ref{methods} are only published for five plant functional types (PFTs - see Table \ref{pft}),  only 56 of the 77 sites were used.
-Figure \ref{map_fig} \explain{map needs to be improved - it's a placeholder for now} presents each site and its plant functional type.
+We use both meteorological and eddy-covariance data from the FLUXNET2015 database (WEBSITE LINK), including all sites with more than four years of data.
+Each site's plant functional type (PFT) was classified using the International Geosphere-Biosphere Programme vegetation classification scheme \citep{Loveland_1999}.
+The physical constants used for water use efficiency in Section \ref{methods} are only published for five plant functional types (PFTs): crops (CRO), deciduous broadleaf forest (DBF), evergreen needleleaf forest (ENF), grass (GRA), and closed shrub (CSH) (see Table \ref{pft}).
+There are 56 sites with these plant functional types, and their location is shown in  Figure \ref{map_fig} \explain{map needs to be improved - it's a placeholder for now}.
 
-The purpose of this study is to examine ecosystem response to atmospheric drying during the growing season.
+The purpose of this study is to examine ecosystem response to atmospheric drying focusing on the growing season.
 To accomplish this, we filter and quality control the data using a similar procedure as \cite{Zhou_2015}:
@@
 \item Only measured or highest (``good'') quality gapfilled data, according to quality control flags, are used.
-\item To isolate the growing season, we only use days in which the average GPP exceeds 10\% of the observed 95th percentile of GPP for a given site.
-\item We remove days with rain and the day following to avoid issues with rain interception and sensor saturation and high relative humidity (\cite{MEDLYN_2011}).
+\item To isolate the growing season, we only use days in which the average Gross Primary Productivity (GPP) exceeds 10\% of the observed 95th percentile of GPP for a given site.
+GPP is calculated using the nighttime respiration partitioning method.
+\item We remove days with rain and the day following to avoid issues with rain interception and sensor saturation at high relative humidity (\cite{MEDLYN_2011}).
 \end{itemize}
-Additionally, we restrict data to the daytime which is identified when downwelling shortwave radiation is greater than 50 W m$^{-2}$ and sensible heat flux is greater than 5 W m$^{-2}$.
+Additionally, we restrict data to the daytime, which is identified when downwelling shortwave radiation is greater than 50 W m$^{-2}$ and sensible heat flux is greater than 5 W m$^{-2}$.
 To further reduce the chance of sensor saturation at high relative humidity, we remove all time steps for which VPD is less than .01 kPa.
-Timesteps with negative observed GPP or ET are also removed.
+Timesteps with negative observed GPP or ET are also removed, and we aggregate half hourly data to hourly averages to reduce noise.
 After these quality control procedures, 332556 upscaled hourly observations remain.
@@
 \label{methods}
-The Penman-Monteith equation (hereafter PM) estimates ET as a function of atmospheric and land-surface variables:
+The Penman-Monteith equation \citep [hereafter PM,][]{Monteith_1965} estimates ET as a function of observable atmospheric variables and surface conductances:
 \begin{linenomath*}
@@
 \end{linenomath*}
-
- where variable definitions are given in Table 1.
-\citet{MEDLYN_2011} developed a model for stomatal conductance ($g_s$) by combining optimal photosynthesis theory with empirical approaches.
-The result for leaf-scale stomatal resistance was:
+where $\Delta$ is the change in saturation vapor pressure with temperature, given by Clausius-Clapeyron ($\frac{d \; e_s}{d \; T}$), $R$ is the net radiation minus ground heat flux, $g_a$ is aerodynamic conductance, $\rho_a$ is air density, $c_p$ is specific heat of air at constant pressure, $\gamma$ is the psychometric constant, and $g_s$ is the stomatal conductance (Table 1).
+\citet{MEDLYN_2011} developed a model for stomatal conductance ($g_s$) by combining an optimal photosynthesis theory \citep{Farquhar_1980, Katul_2010} with an empirical approach, which describes the dependence of $g_s$ to VPD.
+This resulted in the following model for leaf-scale stomatal conductance was:
 
@@
 \end{linenomath*}
+Where $g_1$ is a leaf-scale ``slope'' parameter, A is the net CO$_2$ assimilation rate, and $c_s$ is the atmospheric CO$_2$ concentration at the leaf surface.
+\cite{MEDLYN_2011} relate the slope parameter ($g_1$) to physical parameters as:
+\begin{linenomath*}
+  \label{slope}
+  \begin{equation}
+  g_1 = \sqrt{\frac{3 \, \Gamma \, \lambda}{1.6}},\footnote{Note the above expression has units of of (mmol mol$^{-1}$)$^{1/2}$, but this can be converted to Pa$^{1/2}$ using the ideal gas law.}
+  \end{equation}
+\end{linenomath*}
 
-This can be adapted to an ecosystem-scale stomatal resistance by multiplying by leaf area index (LAI) and converting units to m s$^{-1}$:
+where $\Gamma$ is the CO$_2$ compensation point for photosynthesis (without dark respiration), and $\lambda$ is the marginal cost of leaf carbon.
+So, $g_1$ is a leaf-scale term reflecting the tradeoff of water for carbon uptake and growth.
+The higher $g_1$, the more open are stomata and the more they release water in exchange for carbon.
+
+The Medlyn model for stomatal conductance has been shown to behave very well across PFTs, compared to other models from the leaft to ecosystem scale \citep{Lin_2015, CITE ZHOU ET AL AS WELL}.
+It is adapted at the ecosystem scale by multiplying $g_{l-s}$ by leaf area index (LAI) and converting units to m s$^{-1}$:
 
@@
 
-While Equation 3 can be used in PM, it will make analytical work with the function intractable because $A$, net CO$_2$ assimilation, is closely functionally related to ET itself.
-To remove dependence of ET on $A$ we can use the semi-empirical results of \citet{Zhou_2015}.
-\citet{Zhou_2015} showed that:
+While Equation 3 can be used in PM (Equation 1), it will make analytical work with the function intractable because $A$, net CO$_2$ assimilation, is functionally related to ET itself.
+To remove the dependence of ET on $A$ we use a fundamental semi-empirical results of \citet{Zhou_2015}, who showed that the underlying Water Use Efficiency $uWUE$:
 
@@
 \end{linenomath*}
-is relatively constant across time and space (within plant functional type).
-If, following \citet{Lin_2015}, we approximate $g_0$ as $0$, we can use uWUE to remove $A$ from $g_s$ in a way that makes PM analytically tractable:
+is relatively constant across time and moisture conditions within a plant functional type (CITE ALSO CHAGJIE), in non-extreme drought conditions.
+If, following \citet{Lin_2015}, we approximate $g_0$ as $0$ (i.e.
+we neglect cuticular and epidermal losses - a reasonable assumption except in very dry conditions), we can use uWUE to remove the $A$ dependence from $g_s$ in a way that makes the PM equation analytically tractable:
 
@@
 
-Note that $uWUE$ is fit on the ecosystem scale in \citet{Zhou_2015} so GPP in Equation \ref{uwue} is really $A\cdot \text{LAI}$.
-This leads to the cancelation of LAI in addition to uWUE in Equation \ref{medlyn}.
-Plugging Equation 5 into Equation 1 and rearranging gives:
+Note that $uWUE$ is fit on the ecosystem scale as in \citet{Zhou_2015} so GPP in \ref{uwue} is really $A\cdot \text{LAI}$.
+This leads to the cancellation of LAI in addition to uWUE in Equation \ref{medlyn}.
+Plugging Equation 5 into Equation 1 and rearranging gives a new expression for PM ET directly as a function of VPD with no $A$ dependence:
 
@@
 
-Given FLUXNET data described in Section \ref{data}, every term in Equation \ref{et} is known.
-However, our sampling of sites at the global scale may introduce some deviations of $uWUE$ from those observed in \citet{Zhou_2015}.
-Also, we wish to include some measure of uncertainty in our analysis to guide if our many assumptions and simplifications are reasonable.
-To account for both mean deviations of $uWUE$ and uncertainty, we will introduce an uncertainty parameter $\sigma$ modifying $uWUE$:
+With Equation \ref{et}, ET is explicitly a function of environmental variables and two plant-specific constants, the slope parameter ($g_1$), and $uWUE$, both reflecting water conservation strategy.
+The slope parameter is a leaf-scale term related to the willingness of stomata to trade water for CO_2 and to keep stomata open.
+$uWUE$ is semi-empirical ecosystem-scale constant related to how WUE changes with VPD (specifically $VPD^{-1/2}$).
+In the eddy-covariance FLUXNET2015 data described in Section \ref{data}, every term in Equation \ref{et} is observed, except for the two parameters $g_1$, and $uWUE$, which have however been measured for the PFTs considered here (REF NEEDED HERE).
+However, eddy-covariance data are inherently noisy so that we include a measure of uncertainty in our analysis.
+To account for both mean deviations of $uWUE$ and uncertainty, we introduce an uncertainty parameter $\sigma$ modifying $uWUE$:
 
@@
 
-Now, from each FLUXNET observation we can calculate single value of $\sigma$:
+Now, from each FLUXNET2015 observation (i.e. for each semi-hourly data, at every time step) we can evaluate $\sigma$:
 
@@
   \begin{equation}
-\sigma = - \frac{g_a \gamma c_s \sqrt{VPD} L_v P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p VPD\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})}
+\sigma = - \frac{g_a \gamma c_s \sqrt{VPD} L_v P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p VPD\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})},
     \label{sigma}
@@
 
-The variability of $\sigma$ across sites and time will provides some measure of uncertainty in our model, assumptions, as well as the FLUXNET observations themselves.
+Thus, with this uncertainty analysis we can evaluate departure from our theory in observations, as a departure of $\sigma$ from unity.
+The variability of $\sigma$ across sites and time provides a measure of uncertainty in our model, assumptions, as well as in the FLUXNET201 observations themselves.
 To correct for differences in sampling between \cite{Zhou_2015} and our data, we set uWUE such that $\overline{\sigma} = 1$.
@@
 
-With Equation \ref{d_et} we have provided an analytical framework for ecosystem response to atmospheric demand perturbations.\explain{Should I take the derivative this way?
-I think it's more mathematically correct, but it makes the explanation more complicated and only changes the solution by a factor of two.} There are a few subtleties to taking the derivative in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$) and $VPD$ are functionally related, so while taking the derivative we evaluate $\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET} {\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD} \Big|_{\text{RH fixed}} + \frac{\partial \; ET}{\partial \; RH} \frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$ fixed}}$.
-$RH$ and $e_s$ are assumed to be approximately orthogonal.
+With Equation \ref{d_et} we have an analytical framework for ecosystem response to atmospheric demand perturbations, in non extreme drought conditions.
+There are a few subtleties to taking the derivative in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$) and $VPD$ are functionally related, so while taking the derivative we evaluate $\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET} {\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD} \Big|_{\text{RH fixed}} + \frac{\partial \; ET}{\partial \; RH} \frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$ fixed}}$.
+$RH$ and $e_s$ are assumed to be approximately independent, which is supported by the data (not shown).
 
-The $VPD$ dependence in Equation \ref{d_et} is a little opaque.
-However, mean $VPD$ is 1062 Pa, so $\sqrt{VPD}$ is 32.6 Pa$^{1/2}$, which is much less than $g_1$ (with the exception of ENF; Table \ref{pft}).
-So a series expansion in the limit $\frac{\sqrt{VPD}}{g_1} \to 0$ gives an approximation which makes the functional form more clear: \explain{Should I even include the series expansion?}
-
-\begin{linenomath*}
-  \begin{equation}
-    \frac{\partial \;  ET}{\partial \; VPD} \approx \frac{g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_s }{1.6 \; R*\; \sigma \; \text{ uWUE }} \left( \frac{1}{g_{1}} - \frac{3 \sqrt{VPD}}{2 g_{1}^{2}} + \frac{2 \sqrt{VPD}^{2}}{g_{1}^{3}} - \frac{5 \sqrt{VPD}^{3}}{2 g_{1}^{4}} + \mathcal{O}\left(\left(\frac{\sqrt{VPD}}{g_1}\right)^{4}\right) \right) \right)
-    \label{d_et_approx}
-  \end{equation}
-\end{linenomath*}
-
-One final comment on our derivation which will not be discussed further but is relevant for future analysis: if we  approximate $c_s$ at a global mean CO$_2$ concentration, then the RHS of Equation \ref{et} is fully defined using commonly available weather station data and the constants published in \citet{Zhou_2015, Lin_2015}.
+We note one final comment on our derivation which is relevant for drought indices.
+If we approximate $c_s$ at a global mean CO$_2$ concentration, then the RHS of Equation \ref{et} is fully defined using commonly available weather station data and the constants published in \citet{Zhou_2015} and \citet{Lin_2015}.
 This makes Equation \ref{et} a useful alternative to PET in drought indices and hydrometeorological analysis for vegetated surfaces.
-Equation \ref{et} better reflects the physics of water exchange at the land surface.
+Equation \ref{et} better reflects the physics of water exchange at the land surface and would only require fitting of $uWUE$ and $g_1$, and can also account for changes in CO$_2$ concentration, contrary to typical drought indices such as the Palmer Drought Severity Index (PDSI) (Swann et al.
+2016).
 
@@
 For an observation that perfectly matches our model and constant uWUE assumption $\sigma$ will be one.
-Therefore, for our assumptions and framework to be reasonable $\sigma$ should be $O(1)$.
+Therefore, for our assumptions and framework to be reasonable $\sigma$ should be close to 1.
 Figure \ref{lai_fig} presents the histogram of calculated $\sigma$s with outliers removed (lowest and highest 5\% percent).
-All remaining $\sigma$ values are $O(1)$ which provides confidence in model framework.
+All remaining $\sigma$ values are close to unity (i.e.
+$O(1)$), which provides further confidence in the model framework and in the data quality.
 
@@
 
-An additional concern is that the $\sigma$ term may in fact be some function of $VPD$, in which case the dependence would need to be accounted for when taking the derivative.
-Figure \ref{lai_vpd_fig} plots the joint distribution of $\sigma$ and VPD, and shows that $\sigma$ is very weakly a function of VPD.
+An additional concern is that $\sigma$ may in fact be correlated with $VPD$, in which case the dependence would need to be accounted for when taking the derivative.
+However, the joint distribution of $\sigma$ and VPD shows that $\sigma$ is very weakly a function of VPD (Figure \ref{lai_vpd_fig}).
 Given this weak dependence, we argue that Equation \ref{d_et} is a valid approximation for ET response to $VPD$.
 
-% below show histogram of site specific spearmannr? think would be useful 
 \begin{figure}[h]
@@
 Before calculating the sensitivity of ET to VPD, we will consider the functional form of Equation \ref{d_et}.
-There are two main terms: a ``scaling'' term which modifies the magnitude but not the sign:
+There are two main terms: a ``scaling'' term, which modifies the magnitude but not the sign of the ET response to VPD  ($\frac{\partial \; ET}{\partial \; VPD}$):
 
@@
 
-and a ``sign'' term which determines whether the derivative is positive or negative:
+and a ``sign'' term, which determines whether ET increases or decreases with VPD (i.e. atmospheric demand driven or physiologically controlled):
 
@@
 All variables are positive, so the relative magnitude between the first term and the second term in Equation \ref{sign} will determine whether ET increases or decreases with increasing VPD.
-If the second term is larger magnitude then plant response dominates, while if the first term is larger then atmospheric demand dominates.
-Given that the sign term is crucial to answering our question ``When does VPD drive or reduce ET?'', we will explore its functional form first.
+If the second term is larger then the plant control dominates and ET decreases with increasing VPD.
+However, if the first term is larger, then atmospheric demand dominates and ET increases with increasing VPD.
 
 \subsection{Functional Form of the Sign Term}
-\label{sign_func}
-First, let's explore the variables within the sign term to gain better intuition for what determines when VPD drives vs.
-reduces ET.
-The variability of $c_s$ and $\gamma$ is relatively less than $\sigma$ and $VPD$ variability, so sign term variability within PFT will be approximately determined by just $\sigma$ and $VPD$.
-If we fix uncertainty $\sigma$ at 1, then the sign term is approximately just a function of $VPD$.
-We can further determine a $VPD_{crit}$ where $\frac{\partial \; ET}{\partial \; VPD} = 0$:
+\label{sign_term}
+First, we explore the variables within the sign term to gain better intuition on the driver of either the increase or reduction of ET with VPD.
+CO$_2$ concentration ($c_s$) and the psychometric constant ($\gamma$) are relatively constant over the dataset considered here so that the variability is dominated by $\sigma$ and $VPD$.
+$uWUE$ could vary with soil moisture but has been shown to be relatively constant except in very dry soil conditions  (ZHOU ET AL, LIN ET AL)\explain{Pierre, which Zhou and Lin citations were you referring to that show that uWUE changes with very dry soil conditions?}.
+We thus further assume that the soil is not too dry so that $uWUE$ can be approximated as a constant (within PFT).
+This then means that the sign term only depends on VPD for a given PFT and is approximately just a function of $VPD$.
+We can further determine a critical threshold separating an increase from a decrease in ET, I.e.
+the threshold $VPD_{crit}$ such as the derivative vanishes $\frac{\partial \; ET}{\partial \; VPD} = 0$:
 
@@
 
-Values of $VPD_{crit}$ as a function of PFT are shown in Table \ref{vpd_crit}.
-For any values of $VPD$ less than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be negative, and for values of $VPD$ greater than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be positive.
-Intuitively this makes sense; as VPD increases ecosystems will start to run out of strategies to hold back water, and atmospheric demand will increase.
-Any further increases in atmospheric demand will then result in an increase in VPD.
-The extent to which a given ecosystem has evolved to conserve water will determine at what VPD atmospheric demand takes over.
+We note that $VPD_{crit}$ mostly depends on the PFT parameters $uWUE$ and $g_1$ but only very weakly on climate as most other parameters related to the environment are nearly constant.
+The value of $VPD_{crit}$ for each PFT is shown in Table \ref{vpd_crit}.
+For any values of $VPD$ less than $VPD_{crit}$, ET will thus decrease with increasing VPD ($\frac{\partial \; ET}{\partial \; VPD} < 0$), and for values of $VPD$ greater than $VPD_{crit}$, ET will increase with increasing VPD ($\frac{\partial \; ET}{\partial \; VPD} > 0$).
+In other words, ecosystems regulate and mitigate evaporative losses up to the VPD limit, $VPD_{crit}$, above which atmospheric demand is just too high to be entirely compensated by stomatal and ecosystem regulation.
+We note however that even though ET increases again above the critical threshold, $VPD_{crit}$, ET is still much lower that potential evaporation as stomata are still strongly regulating vapor fluxes tot he atmosphere.
+Indeed, even in the absence of soil pore evaporation, stomata do not shut down entirely at very high VPD so that ET does not go to zero, because stomata are still slightly open to perform some photosynthesis \citep{Ball_1987, Leuning_1990, MEDLYN_2011}.
+In addition, upward xylem transport is necessary to maintain phloem transport, as well as nutrient transport and thus carbon allocation (De Schepper, V., T.
+De Swaef, I.
+Bauweraerts, and K.
+Steppe (2013), Phloem transport: a review of mechanisms and controls, J Exp Bot, 64(16), 4839–4850, doi:10.1093/jxb/ert302.
+Nikinmaa, E., T.
+Hölttä, P.
+Hari, P.
+Kolari, A.
+Makela, S.
+Sevanto, and T.
+Vesala (2012), Assimilate transport in phloem sets conditions for leaf gas exchange, Plant Cell Environ, 36(3), 655–669, doi:10.1111/pce.12004.
+Ryan, M. G., and S. Asao (2014), Phloem transport in trees, Tree Physiol, 34(1), 1–4, doi:10.1093/treephys/tpt123.
+).
 
 \begin{table}  
-  \label{vpd_crit}
 \caption{Values of $VPD_{crit}$, where $\frac{\partial \; ET}{\partial \; VPD} = 0$, evaluated at PFT average values for $R_{air}$, $\sigma$, $\gamma$, and $c_s$.
-For reference, these values are also provided.
+For reference, these values are also provided CITE TABLE.
 For values of $VPD$ less than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be negative, and for values of $VPD$ greater than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be positive.
@@
 \end{tabular}
+\label{vpd_crit}
 \end{table}
 
-Figure \ref{term3} provides a greater description for how the sign term varies with $VPD$, as a function of PFT.
-Equation \ref{d_et_approx} aids interpretation of Figure \ref{term3}.
-Larger $uWUE$ and $g1$ shift the leading-order constant term ($\frac{1}{g_1}$) towards smaller values, and the sign term towards overall positive values.
-% Higher values corresponde to plants that are more willing to spend water on primary production and less evolved to conserve water.\explain{really need to think about this sentence more, tehre is more nuance and I actually thing g1 and uwue really partially cancel.}
-Figure \ref{term3} confirms our physical intuition: CROs are the least water conservative so have the smallest constant portion of Term 3 ($\frac{\gamma c_s }{1.6 \; R*\; \sigma \; \text{ uWUE }} \left(\frac{1}{g_1}\right)$), while CSH are the most water conservative and have the largest constant portion of Term 3.
-For the VPD-dependent terms in Equation \ref{d_et_approx}, differences in $g1$ between PFTs exert a greater influence than differences in $uWUE$.
-Increasing $uWUE$ and $g1$ decreases the VPD-dependence, but $g1$ has the bigger effect, particularly at high VPD when the series becomes disordered.
-ENF ($g1 = 74.31$) has by far the largest VPD dependence of response, while CRO ($g1 = 183.1$) has the smallest VPD dependence.
+Differences in $VPD_{crit}$ is exclusively determined by $uWUE$ and the slope parameter ($g_1$) related to the plant functional type.
+A larger underlying water use efficiency, $uWUE$, causes means that at the ecosystem scale $WUE$ decreases more with increasing VPD, so plants lose more water per unit carbon production and are more likely to increase ET with increasing VPD (REFLRMUALTE SENTENCE, THE RESULTS IS CONTER INUITTICE WE EDPECT THE OPPOSITE).
+A tendency towards increasing ET response with increasing VPD can also be caused by a high slope parameter ($g_1$), characteristic of plants that at the leaf scale are more willing to trade water for stomata opening.
+Plants that are less conservative will be thus be more likely to increase ET with increasing VPD.
+Both the aforementioned effects (large $uWUE$, $g_1$) can amplify each other, and generally conspire to shift the sign term towards a positive value for a given PFT.
 
-A primary takeaway from Figure \ref{term3}a is that according to our theory for all PFTs except for crops there is frequent occurrence of a negative (plant dominating) ET response to increases in VPD.
+This effect of $uWUE$ and $g1$ on the sign term is most apparent by comparing two extreme PFTs: water intensive CRO and water conservative CSH.
+CRO has a high slope parameter ($g_1 = 183.1 \, Pa^{1/2}$) compared to CSH ($g1 = 148.6 \, Pa^{1/2}$.
+It also has a slightly higher uWUE ($2.32$ -units-) compared to CSH ($uWUE=2.11$ -units-).
+MAYBE A DUMB QUESTION BUT CAN'T uWUE AND G1 BE ANALYTICALLY RELTED> I THINK AT LESAT YOU SHOULD MAKE CLEAR THAT UWUE IS AT THE ECOYSTEM SCALE (DID YOU FIT IT OR USED PUBLISHED DATA-I AM NOW CONFUSED).
+These differences in slope parameter cause opposite ET responses to changes in VPD between CRO and CSH.
+ET theoretically always increases with increasing VPD for the more water intensive CRO, while ET almost always decreases with increasing VPD for the more water conservative CSH (Figure \ref{term3}).
+CRO, evolved or were bred to prioritize GPP and yield and thus not water conservative so that they are very willing to trade water for photosynthesis and productivity, despite changes in VPD, while CSH are very unwilling to trade water for more photosynthesis.
+
+AS expected, the slope parameter ($g_1$) is a primary determinant of the VPD dependence for the sign term shown in Figure \ref{term3}.
+Plants that are more conservative (small $g_1$) will tend to reduce ET with increasing VPD, and will be very effective at reducing ET, especially at low VPD.
+However, at very high VPD, gradients in vapor pressure at the leaf scale will be so strong that they will begin to dominate leaf-scale plant strategies of stomatal (partial) closure (parameterized with $g_1$).
+As a result, ET response will begin to asymptote towards its ecosystem-scale values irregardless of leaf-scale strategy and stomatal regulation (parameterized with $g_1$).
+Therefore, plants with a low $g_1$ will have the largest VPD dependence of ET response because the difference in ET response at low VPD (leaf stomatal response dominates) and high VPD (VPD gradient dominates) is largest.
+This is apparent in the strong VPD dependence of ENF, which has the lowest slope parameter ($g_1=74.31$), in Figure \ref{term3}.
+
+To summarize our theoretical insights, Figure \ref{term3} and Table \ref{vpd_crit}, CROs are the least water conservative and have the strongest overall tendency to increase ET with increasing VPD, while CSH are the most water conservative and have the strongest tendency to decrease ET with increasing VPD.
+ENF ($g_1 = 74.31$) has by far the largest VPD response dependence, while CRO ($g_1 = 183.1$) has the smallest VPD dependence.
+ENF are less willing to trade water for GPP, so the difference between the response at low VPD when leaf response dominates and high VPD when vapor pressure gradients dominate is much greater.
+CROs on the other hand have a more aggressive water use strategy and are more willing to trade water for photosynthesis, so the leaf scale response is much smaller, and the difference in ET response between low VPD and high VPD is relatively small.
+
+Figure \ref{term3} clearly shows, according to our theory, that for all PFTs except for crops there is frequent occurrence of a negative (plant dominating) ET response to increases in VPD.
+Therefore, plants are able in most atmospheric conditions to reduce ET in response to increased VPD and thus to reduce water loss.
 To better illustrate this, the ranges of observed environmental VPDs at the FLUXNET sites are plotted parallel to the x-axis.
-25th, 50th, and 75th percentiles are marked by stars, and the location of VPD$_{crit}$ is marked by a vertical line.
-For CSH (water conservative), VPD is always less than VPD$_{crit}$ (plant response dominates).
-For CRO, VPD is almost always more than VPD$_{crit}$.
-For forests and grass, whether plant response or atmospheric demand dominates will depend more on the environment, with about half of of observed VPD less than VPD$_{crit}$ (plant response dominates).
-It's also important to note that even when atmospheric demand dominates, the response is still far smaller than it would be for PET (atmospheric demand only).
+For CSH, VPD is always less than VPD$_{crit}$ so that the plant response dominates in typical environmental conditions, emphasizing the water conservative strategy of those plants.
+For CRO on the other hand, VPD is almost always higher than VPD$_{crit}$, emphasizing that those plants operate with an aggressive water usage strategy, are water intensive and were actually engineered for photosynthesis rather than water saving.
+For forests and grass, whether plant response or atmospheric demand dominates is more dependent on environmental conditions, with about half of the observed VPD less than VPD$_{crit}$, i.e.
+in conditions where plant response dominates.
+It is also important to note that even when atmospheric demand dominates, ET response to VPD is still far smaller than it would be for potential evaporation $\partial PET/\partial VPD$, i.e.
+atmospheric demand only, emphasizing that there is still a strong regulation of evaporative flux by stomata and though the plant xylem.
 The sign term in this case would just be a constant ($\frac{c_p}{R_{air}} \approx 3.5$), which is far larger than any part of the curves for any PFT.
 This highlights the deficiencies of PET's ability to capture land response to changes in atmospheric dryness (VPD).
-However, up to this point we have assumed $\sigma$ (uncertainty) equal to one.
-We still need to test if inclusion of uncertainty changes the story.
+Plants are always regulating water exchange from the land surface, even when they reach the limits of they ability to do so.
+Therefore, the actual land surface response to a change in VPD will always me more negative than if the plants are not included in the physics (PET).
+
 \begin{figure}[h]
@@
 Lines and stars denote the distribution of VPD and $\sigma$ next to each axis, following the same percentiles as above.
-\textbf{***need to replot with correct sigma (normalized to 1), also correct uWUE label in first plot***}}
-\label{term3}
+\textbf{***I think I'm going to remove the bottom plot: its discussion didn't fit in well with the text, and the box plots test inclusion of uncertainty***}}
+\label{sign_plot}
 \end{figure}
 
-Figure \ref{term3}b shows the location of VPD$_{crit}$, as a function of $\sigma$ (uncertainty) and $VPD$.
-For any $\sigma$ or VPD less (more) than these curves, the sign term will be negative (positive).
-It is clear that the portion of VPD observations below/above these curves will be a strong function of $\sigma$.
-However, we can see some general trends.
-For CSH, $\frac{\partial \; ET}{\partial \; VPD}$ should be negative for the vast majority of observed $\sigma$ and VPD.
-The fraction of positive $\frac{\partial \; ET}{\partial \; VPD}$ appears to be more even for ENF, GRA, and DBF, and we might expect a greater frequency of positive $\frac{\partial \; ET}{\partial \; VPD}$ for CRO.
-Section \ref{testing} further examines the consistency of our theory, given  the apparent sensitivity to uncertainty ($\sigma$).
+\subsection{Functional Form of the Scaling Term}
+\label{scale_term}
+While the above discussion of the sign of $\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our question, understating the overall magnitude of the ET response is important to the soil-plant-atmosphere water budgeting.
+So we now more closely examine the terms that affect how the sign term is scaled:
 
-\subsection{Functional Form of the Scaling Term}
+\begin{equation}
+  \frac{g_a \; P}{T(\Delta + \gamma)},
+\end{equation}
 
-While the above discussion of the sign of $\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our research question, the magnitude of $\frac{\partial \; ET}{\partial \; VPD}$ will also impact statistics of $\frac{\partial \; ET}{\partial \; VPD}$ and the importance of $VPD$ variability for overall $ET$ variability.
-So we will more closely examine the scaling term: $\frac{P}{T} \propto \rho$, so this should vary little relative to aerodynamic conductance and $\Delta$.
-$\gamma$ should also be relatively constant, so the scaling term should be primarily a function of aerodynamic conductance and temperature (through the function $\Delta$).
-This makes sense, as aerodynamic conductance represents how efficiently response at the surface is communicated to the atmosphere.
-As it increases, any plant response will be communicated more strongly to the atmosphere (and vice-versa).
+$\frac{P}{T}$ is an air-density term, which we varies little compared to aerodynamic conductance and Clausius-Clapeyron ($\Delta$).
+The psychometric constant ($\gamma$) is also relatively constant, so the scaling term should be primarily a function of aerodynamic conductance and temperature, through the Clausius-Clapeyron relationship $\Delta$.
+This is as expected, given that the aerodynamic conductance represents the efficiency of exchange between the surface and the atmosphere.
+As aerodynamic conductance  increases, any plant response will be communicated more strongly to the atmosphere (and vice-versa).
 
 $\Delta$'s presence in the scaling term also matches physical intuition.
-Evaporative cooling will dampen the ability of the atmosphere to take more moisture, because $e_{s}$ decreases with decreasing temperature.
+Evaporative cooling will dampen the ability of the atmosphere to take more moisture (UNCLEAR SENTENCE REFORMULATE), because $e_{s}$ decreases with decreasing temperature.
 The decrease in $e_{s}$ is proportional to $\Delta$ ($\delta e_{s} = \Delta \delta T$).
@@
 The functional from of $\Delta$ will be the same across PFT, but the temperature range may vary slightly.
-In contrast, aerodynamic conductance will vary strongly with PFT due to the importance of surface roughness.
+In contrast, aerodynamic conductance will vary strongly with PFT due to the importance of surface roughness for aerodynamic conductance.
 So most of the differences in scaling between PFT should be in the aerodynamic conductance term.
-One interesting side note is that the coefficient of variability for both aerodynamic conductance and the scaling term is relatively constant across PFT, suggesting that the influence of aerodynamic conductance on the relative (to the PFT mean) variability of the scaling term is approximately similar across PFT.
 
-Figure \ref{scale_vary}A shows the scaling term normalized by mean aerodynamic conductance (calculated for each plant functional type), and confirms that much of the relative variability of the scaling term is caused by aerodynamic conductance variability.
-Generally, temperature (via $\Delta$) causes less relative variability.
-However, the impact of $T$ on the relative variability increases with increasing aerodynamic conductance.
+The control of the scaling term variability between PFTs by aerodynamic conductance is confirmed by the data (Figure \ref{scale_vary}).
+Differences between PFT are almost entirely due to differences in aerodynamic conductance, rather than differences in observed temperature ranges.
+The scaling term for the tree PFTs (DBF, ENF) is generally more than double the scaling terms for the other PFTs which have lower surface roughness and generally smaller aerodynamic conductance.
 
-While the relative variability of scaling term is similar across PFT, the absolute value of scaling term varies strongly across PFT.
-Figure \ref{scale_vary}B shows scaling term evaluated at the mean aerodynamic conductance for each PFT, and at the range of observed temperatures for each PFT.
-As expected, for the tree PFTs (DBF, ENF) the scaling term is much larger and the temperature dependence is much stronger.
-Systematic differences in observed temperatures also cause differences in the average magnitude of scaling term.
-For example, ENF experiences on average colder temperatures and is thus more likely to have a larger scaling term.
-Additionally, because the variability of aerodynamic conductance increases proportionally to the mean, the spread of the scaling term due to aerodynamic conductance variability will be larger for the tree PFTs, although this is not shown for simplicity.
-To summarize Figure \ref{scale_vary}: the variability of the scaling term within each PFT will look like Figure \ref{scale_vary}A for each PFT, but the scale of the y-axis will increase or decrease according to mean aerodynamic conductance observed in Figure \ref{scale_vary}B.
+Within each PFT, the scaling term variability is controlled both by environmental temperature and aerodynamic conductance variability (Figure \ref{scale_vary}).
+While the observed variability of the aerodynamic conductance contributes more to the scaling term variability than temperature, the temperature contribution is non-negligible.
+Specifically, the scaling term is generally larger at low temperatures when ET is weakly damped by Clausius-Clapeyron, and the effect of aerodynamic conductance variability is stronger as it is less weakly damped by Clausius-Clapeyron.
+
+To summarize, variability between PFTs is mostly controlled by systematic differences in aerodynamic conductance, due to differences in surface roughness between each PFT and wind conditions.
+In contrast, variability within PFT is also controlled by temperature, through Clausius-Clapeyron.
+But, aerodynamic conductance variability generally impacts the scaling term more than temperature, even within PFT.
 
@@
 \centering
-\includegraphics[width=20pc]{./fig04.pdf}
-\caption{Primary sources of variability for scaling term.
-A) Variability within each PFT: scaling term normalized by mean $g_a$ for each PFT.
-B) Variability between each PFT: scaling term evaluated at mean $g_a$ for each PFT.
-Temperature range is 5-95th percentile for each PFT.
-Additionally, stars denote the location of the 25th, 50th, and 75th percentiles.}
+\includegraphics[width=\textwidth]{./fig04b.pdf}
+\caption{Primary sources of variability for the scaling term, as a function of PFT.
+Note the change in axis scale for each plot.
+The 5th-95th percentile range of temperature is plotted at the 5th, 25th, 50th, 75th, and 95th percentiles of aerodynamic conductance, as observed for each PFT's sites.}
 \label{scale_vary}
@@
 
-%%%%% replace below with idea:
-\subsection{Bulk statistics of $\frac{\partial \; ET}{\partial \; VPD}$}
+\subsection{Bulk statistics of ET response to VPD}
+\label{stats_sec}
+FORULATE THAT DIFFERNTLY< SAY WE NOW TURN TO OBSERVATIONS> YOU HVAE ALREADY DISCSSED SIGAMA AND IRTS DEFINTION.
+MENTION IT WHILE HYOU DISCUSS THE RESULS BUT DO NOT MAKE IT THE MAIN FOCUS.
+I TRIED DOING IT
+In this section we consider direct observations of ET response with eddy-covariance data, while including uncertainty, with the $\sigma$ term (Section \ref{methods}).
+These observational results of ET response largely confirm our theoretical analysis, presented in Sections \ref{sign_term}-\ref{scale_term} (Table \ref{stats}).
+For all PFTs, mean ET response to increasing VPD is negative.
+However, ET response evaluated at the average of all variables (e.g.
+$\sigma$, $T$, $c_s$, $VPD$) is only negative for CSH and GRA.
+This difference in mean ET response as compared to the ET response at mean environmental conditions is due to the non-linear nature of the response, in which negative responses are generally much larger magnitude than positive responses (Figure \ref{sign_plot}).
+SAY IT IN WORDS AS WELL AND WHTEHR IT FITS PREVIOUS REULTS
 
-Table 3 confirms our expectations for PFT behavior of $\frac{\partial \; ET}{\partial \; VPD}$.
-For all PFTs except for CRO, average $\frac{\partial \; ET}{\partial \; VPD}$ is less than zero.
-However, $\frac{\partial \; ET}{\partial \; VPD}$ evaluated at the average of all variables (e.g.
-$\sigma$, $T$, $c_s$, $VPD$) is only negative for CSH and GRA.
-And, DBF in addition to CRO experiences $\frac{\partial \; ET}{\partial \; VPD}$ < 0 less than half the time.
-These observations highlight the effect of the nonlinear function in Figure \ref{term3}: $\frac{\partial \; ET}{\partial \; VPD}$ has a much steeper slope when the function is negative, and is thus more likely to be large magnitude.
+Regarding the frequency of negative and positive ET response, observations of ENF, GRA and CSH are more frequently characterized by a negative ET response to increasing VPD, and are thus more frequently water conservative TOO AVAGUE ESENTENCE REFORMUALTE.
+DBF and CRO observations have greater frequency of positive ET response to increasing VPD.
+i.e.
+water intensive strategies.
 
-The units of $\frac{\partial \; ET}{\partial \; VPD}$ make it difficult to interpret if $VPD$ is even a meaningful contributor to ET's variability.
-To better understand $VPD$'s contribution, we multiply $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)$ with $VPD$'s standard deviation to define a (linearized) relative change in ET for variations in $VPD$ .
-$VPD$'s contribution to ET's variability ranges between 30 - 40 W m$^{-2}$ for all PFTs except for CSH, which is about 100 W m$^{-2}$.
-Another meaningful comparison is to $\frac{\partial \; ET}{\partial \; R} \cdot std(R)$, as net radiation is generally the driver of ET (cite joe berry here)\explain{Find citation?}.
-For all PFTs except for CSH $VPD$ contributes between 30 - 40 \% of $R$'s contribution to variability.
-For CSH the portion is much larger, about 88 \%.
-$VPD$'s variability is certainly a non-negligible contributor to $ET$'s variability.
-
-So far, idealized plots and statistics have illuminated the form of $\frac{\partial \; ET}{\partial \; VPD}$ and how it varies with PFT.
-Large mean uWUE shifts CRO and DBF towards positive $\frac{\partial \; ET}{\partial \; VPD}$, but $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ is still negative for DBF due to its lower $g1$.
-ENF's low $g1$, even lower than DBF, increases the dependence of $\frac{\partial \; ET}{\partial \; VPD}$ on $VPD$, and makes the function very nonlinear.
-This has the side effect of pushing $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ negative further than other PFTs for a given fraction $\frac{\partial \; ET}{\partial \; VPD} < 0$ and magnitude $\frac{\partial \; ET}{\partial \; VPD}(\overline{T,\ldots,VPD})$.
-GRA shows the opposite behavior; a relatively high $g1$  makes the function more linear, decreasing the magnitude of $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ for a given fraction $\frac{\partial \; ET}{\partial \; VPD} < 0$ and negative $\frac{\partial \; ET}{\partial \; VPD}(\overline{T,\ldots,VPD})$.
-Finally, the low $uWUE$ of CSH pushes it toward by far the lowest values $\frac{\partial \; ET}{\partial \; VPD}$ (Figure \ref{term3}).
-Variability in $VPD$ accounts for the largest amount of $ET$ variability for CSH.
-For the other PFTs, $VPD$ contributes less to $ET$ variability, but still represents about 30-40 \% of $R$'s contributions to ET variability.
+In general the bulk statistics match our theoretical expectations well, with the caveat that inclusion of uncertainty shifts grass towards a slightly more negative response than the theory and Figure \ref{sign_plot} would suggest.
+The bulk statistics motivate a more thorough examination of the structure of uncertainty and more sophisticated validation of our theory's performance against observations.
 
@@
 \hline
-
 \end{tabular}
+  \label{stats}
 \end{table}
 
-\subsection{Testing the theory - does the data match the model?}
+\subsection{Validation of theory at eddy-covariance sites}
 \label{testing}
-So far we have developed a theory for $\frac{\partial \; ET}{\partial \; VPD}$'s behavior.
-Most important to answering ``When does VPD drive or reduce ET?'' is that the sign of our derivative term is correct.
-So, we need to test if when we include uncertainty the observed derivatives match what our theory would say for the sign term.
+MERGE WITH PREVIOUS PARAGRAPH
+We now turn to a direct evaluation of the functional forms presented in Section \ref{sign_term} against observations taken from a wide range of eddy-covariance datasets (Section \ref{methods}).
+Figure \ref{agu_fig} shows the distribution of of the sign term (when uncertainty is included), as compared to what the theory would predict.
+Our goal was to capture the leading order behavior of the ET dependence on VPD.
+Given the assumptions we have made, and the uncertainties of flux tower observations themselves, we expect a relatively large amount of noise when reproducing the derivatives of ET.
+However, the data largely reproduces our theoretical analysis.
 
-Figure \ref{agu_fig} shows the distribution of of the sign term when uncertainty is included, as compared to what the theory says it should be.
-Our goal for this paper was to capture the leading order behavior and guide intuition.
-For DBF and CSH our theory is successful towards this goal: the theory matches the leading order behavior of the function when uncertainty is included.
-Given the assumptions we've made, and the uncertainties of flux tower observations themselves, there will always be a large amount of noise.
-The plots for DBF and CSH look as one would expect if one just added noise on top of our theory.
+For DBF our theory is successful towards this goal: the theory matches the leading order behavior of the function when uncertainty is included, and the DBF response appears as expected: i.e.
+is similar to the theory with some additional noise.
+The VPD dependence, given by the slope parameter ($g_1$), follows the median values of each box plot.
+Perhaps most importantly, the x-intercept, and thus VPD$_{crit}$, matches nearly exactly between the theory and the observations.
+Therefore the sign of the ET response to increases in VPD should be well matched, subject to the unavoidable constraints of noise, much of which comes from the observations themselves.
+The uncertainty is non negligible; there are many observations in each box for which the the sign of observation is opposite the response predicted by the theory, but to leading order our theory matches the observations well.
 
-The theory does not match as well for CRO and GRA.
-Perhaps this should be expected.
-Our theory, for practical reasons, does not include important sources of variability (particularly between sites) for CRO and GRA.
-We do not account for seasonal changes in LAI, variability in crop height and surface roughness, differences in C3 vs C4 photosynthesis, or differences in irrigation practices between the sites.
-Given that the our theory does not do as well for CRO and GRA, we will refrain from firm statements about how CRO and GRA respond to changes in VPD.
+While CSH has a much different functional form of the sign term than DBF, CSH observations also match our theory to leading order.
+Again, the VPD dependence mostly determined from the slope parameter ($g_1$) closely matches the medians in the observation bins and the theoretical threshold is well reproduced.
+The VPD-independent, strongly negative response, which is determined by $uWUE$, is also captured.
+For CSH, there is extremely rare occurrence of positive ET response with VPD, even with uncertainty, so the sign of the observations almost always matches the sign of the theory.
 
-ENF is a special case.
-The VPD dependence of the theory is not well captured by observations.
-However, the theory closely captures the intercept where $\frac{\partial \; ET}{\partial \; VPD} = 0$.
-Resolving this feature is the key to answering our question, so perhaps the theory is still useful for ENF for the purpose of addressing when VPD drives or reduces ET.
-It is also noteworthy that we used published constants for $g1$, which determines the VPD dependence, from \cite{Lin_2015}.
-We could have fit our own constants to tweak the theory to match the observations, but our goal was to use a priori physical constants whenever possible.
+As Section \ref{stats_sec} suggested, the theory does not match as well for CRO as for DBF and CSH.
+At low VPD the VPD dependence (due to $g_1$) is similar to the leading order behavior of the observations, but the theory's curve is shifted towards more positive values relative to the observations, due to some bias in $uWUE$.
+Most strikingly, the observations show a tendency to move back towards negative values at high VPD, which given the functional form of VPD dependence, is behavior that is impossible for the theory to capture.
+This general positive bias of the theory explains the shift towards negative ET response observed in the bulk statistics of Section \ref{stats_sec}.
+The divergence between the theory and the observations for CRO could have been expected.
+For practical reasons, our theory does not include important sources of plant type heterogeneity within and across sites for CRO.
+In agricultural landscape, the footprint of an eddy-covariance tend to not only observe a single species but many differnent ones, with potentially varying water use strategies.
+We also do not account for seasonal changes in LAI, variability in crop height and surface roughness, differences in C3 vs C4 photosynthesis and water strategies, or differences in irrigation practices between the sites and years.
+These deficiencies should largely explain the inability of our theory to exactly match the observations in croplands.
+For example, a superposition of sites with C3 plants (at low environmental VPD) and C4 plants (at high environmental VPD) would explain observed shift back towards negative ET response at high VPD when all sites are binned together, as in Figure \ref{agu_fig}.
 
-While the above discussion shows that our theory has some limitations when applied to real data, especially for crops and grasslands, another question could be, how does our theory's performance compare to often-used PET?
-Again, as in Section \ref{sign_func}, the PET response to changes in VPD will be constant at about 3.5.
-Even with uncertainty included, no values of the response approach the PET case.
-PET's status outside the error bars of our theory highlight the usefulness of our theory (and Equation \ref{et}) for estimating land response and moisture budgeting.
+The theory for GRA suffers from similar limitations as for CRO.
+GRA observations are characterized by a consistent trend back towards negative ET response at higher VPD, which the functional form of our theory is incapable of accounting for.
+As compared to CRO, the divergence between the theory and the observations is even greater for GRA, but these differences are explainable by similar issues as for CRO.
+The GRA PFT tends to exhibit substantial within FPT difference, such as C3 vs C4 photosynthesis, as well as strong seasonal variability in height and LAI.
+As with CRO, the trend back towards negative ET response at higher VPD could be explained by a superposition of sites with C3 vs C4 sites.
+Unfortunately, because detailed site-level data of C3 vs C4 plant type is not available, we could not bin CRO and GRA further by irrigation and photosynthesis strategy.
+However, we hypothesize that the theory would validate against observations much better when photosynthesis pathways are accounted for.
+
+The comparison between the theory and observations for ENF is unique from the rest of the PFTs.
+The VPD dependence of the theory, determined by the slope parameter ($g_1$), is not well captured by observations.
+However, the theory closely captures the x-intercept and VPD$_{crit}$.
+So, the theory still captures when the ET response will be positive or negative, but is less accurate with the magnitude of the response.
+Therefore the theory is still useful for addressing when VPD drives or reduces ET.
+We note however that we used published constants for $g_1$ \citep[from ][]{Lin_2015}, which determines the VPD dependence and that deviations should be expected across sites.
+We could have fit our own constants so that the theory would best match the observations, but our goal was to use a priori results whenever possible in order to avoid overfitting and to have a theoretical approach as independent to the observations as possible for further confirmation of the validity of the approach.
+
+While the above discussion shows that our theory has some limitations when applied to some PFTs, especially for crops and grasslands, it does well for DBF and CSH PFTs, and captures the sign of the response consistently for ENF.
+It is useful to also compare our theory, the observed uncertainty in the ET response, and PET, as PET is often used as a proxy for ET in hydrometeorological studies and drought indices.
+Again, as in Section \ref{sign_term}, the PET sensitivity to VPD will be constant at about 3.5.
+In Figure \ref{agu_fig} the ET response to shifts in dryness is always far smaller than 3.5 in the observations, even with all uncertainty included.
+So the PET response to atmospheric dryness is outside the range of all observable responses.
+This casts doubt on PET's usefulness for capturing the leading order behavior of water budget and dryness response at the land surface, such as used in several drought indices.
+However our theory (Equation \ref{et}) could be used with published $uWUE$ and $g1$ data, as an alternative to PET to estimating land response and dryness, and would more accurately account for the response to atmospheric drying.
 
 \subsubsection{Testing VPD$_{crit}$}
-As stated in the above discussion of ENF, we are mostly interested in the critical threshold $VPD_{crit}$ for which the sign of  $\frac{\partial \; ET}{\partial \; VPD}$ is determined.
-When $VPD < VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}  < 0$ (plant response dominates), and when $VPD > VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD} > 0$ (atmospheric demand dominates).
-Theoretically this $VPD_{crit}$ is only a function of PFT.
-We can look at how this value of $VPD_{crit}$ matches the data as different observed variables (including uncertainty) are allowed to vary.
+As stated in the above discussion of ENF, we are mostly interested in the critical threshold $VPD_{crit}$ for which the sign of the ET response to VPD ($\frac{\partial \; ET}{\partial \; VPD}$) is determined.
+When VPD is less than VPD$_{crit}$, ET response to VPD is negative ($\frac{\partial \; ET}{\partial \; VPD}  < 0$), i.e.
+plant response dominates, and when VPD is greater than VPD$_{crit}$, ET response to VPD is positive  ($\frac{\partial \; ET}{\partial \; VPD} > 0$).
+i.e.
+atmospheric demand dominates.
+Theoretically, this $VPD_{crit}$ is mostly a function of PFT and not of climate.
+MAYBE MENTION THAT TEHR SHOULD BE STRONG SPECIES SELCETION BECAUSE OF THIS THRESHOLD
 
-Figure \ref{real} presents calculated $\frac{\partial \; ET}{\partial \; VPD}$ where, unless otherwise noted, all variables in Equation \ref{d_et} are allowed to vary, including uncertainty.
-Each column is a different quantity related to $\frac{\partial \; ET}{\partial \; VPD}$, and each row is a different PFT.
+Figure \ref{real} presents calculated $\frac{\partial \; ET}{\partial \; VPD}$ where, unless otherwise noted, all variables in Equation \ref{d_et} are allowed to vary, including uncertainty ($\sigma$).
+Each column represents a different quantity related to $\frac{\partial \; ET}{\partial \; VPD}$, and each row is a different PFT.
 Vertical black lines denote where our theory says $VPD_{crit}$ should be.
 For any observations to the left of this line $\frac{\partial \; ET}{\partial \; VPD}$ should be negative (plant response dominates), and any observations the right of the line should be positive (atmospheric demand dominates).
-In addition to seeing how the theory $VPD_{crit}$ matches observations, these plots are also useful for observing how the scaling term affects the magnitude of the derivative, if at all.
+In addition to seeing how our theoretically-derived VPD threshold, $VPD_{crit}$, matches observations, these analysis are also useful for observing how the scaling term affects the magnitude of the derivative, if at all.
 
-The scatter plots generally confirm expectations from Figure \ref{agu_fig}.
-By comparing different columns of Figure \ref{real} we can see the impact of $\sigma$ and $g_a$ on the variability of $\frac{\partial \; ET}{\partial \; VPD}$.
-$g_a$'s scaling (included in columns 1 and 3) alters the magnitude considerably, particularly between PFTs.
+Figure \ref{real} generally confirms expectations from Figure \ref{agu_fig}.
+By comparing different columns of Figure \ref{real} we can evaluate the impact of $\sigma$ and aerodynamic conductance ($g_a$) on the variability of ET response ($\frac{\partial \; ET}{\partial \; VPD}$).
+Aerodynamic conductance's scaling (included in columns 1 and 3) alters the magnitude considerably, particularly between PFTs.
 $\sigma$ variability (included in columns 1 and 2) adds a lot of additional noise to the signal of $\frac{\partial \; ET}{\partial \; VPD}$, as observed in Figure \ref{agu_fig}.
-However, the general story with the noise appears to match the cleaner signal when $\sigma$ is help constant and $VPD_{crit}$ is clearly visible.
-The exceptions are, as discussed above, for GRA and CRO, where the full complexity plots (Columns 1 and 2) to not match well with $\sigma$ held fixed (Columns 3 and 4).
+However, the general story with the noise appears to match the cleaner signal when $\sigma$ is held constant and $VPD_{crit}$ is clearly visible.
+The exceptions are, as discussed above, for GRA and CRO, where the full complexity plots (Columns 1 and 2) do not match well with $\sigma$ held fixed (Columns 3 and 4).
 
-While $g_a$ has a large impact on magnitudes in the scaling, temperature, through the $\Delta$ term, does not appear to have as much of an impact in the complete estimation of $\frac{\partial \; ET}{\partial \; VPD}$.
-Therefore it appears that the absolute magnitude of the derivative can be mostly attributed to the sign term, and its direct scaling by aerodynamic conductance.
-Intuitively this makes sense; aerodynamic conductance will control how dominant balances at the land surface are communicated to the boundary layer.
-While $\Delta$ controls the efficiency of latent fluxes in carrying heat away from the surface, it appears this is a second order term, relative to $g_a$, for scaling $\frac{\partial \; ET}{\partial \; VPD}$.
+While the stomatal conductance (MAKE SURE YOU SUE CONDUCTANE THROUGHOUT< YOU WERE USING RESITSANE SOMETIMES) s$g_a$ has a large impact on the magnitude of the scaling, temperature, through Clausius-Clapeyron ($\Delta$), does not appear to have as much of an impact in the complete estimation of ET response ($\frac{\partial \; ET}{\partial \; VPD}$).
+Therefore, it appears that the absolute magnitude of ($\frac{\partial \; ET}{\partial \; VPD}$ can be mostly attributed to the sign term, and its direct scaling by aerodynamic conductance.
+Intuitively this is reasonable; aerodynamic conductance will control how dominant balances at the land surface are communicated to the boundary layer.
+While $\Delta$ controls the efficiency of latent heat flux in removing heat from the surface, it appears this is a second order term, relative to $g_a$, for scaling ET response.
 
-In general Figures \ref{real} and \ref{agu_fig} match our expectations, even with the large sensitivity to $\sigma$ seen in Figure \ref{term3}.
+In general Figures  \ref{agu_fig} and \ref{real} match our expectations.
 Our $\sigma$-based method of uncertainty propagation blurs the idealized expectations the most for GRA and CRO.
-We therefor have the most confidence in our conclusion based on Equation \ref{d_et} for PFTs CSH, DBF, and ENF, as the sign of the derivative with uncertainty included matches (to leading order) the story when $\sigma$ is held fixed.
+We therefore have the most confidence in our conclusion based on Equation \ref{d_et} for PFTs CSH, DBF, and ENF, as the sign of the derivative with uncertainty included matches (to leading order) the story when $\sigma$ is held fixed.
 
@@
 
+\section{Limitations of theory - very dry soil moisture conditions}
+In formulating our theory with Penman Monteith, we implicitly did not account for very dry soil moisture conditions.
+For the majority of environmental conditions  observed at the eddy covariance sites used here, soil conditions were not extremely dry so that we could assume a constant $uWUE$ and $g_1$.
+We posit that ecosystems will generally optimize to host plants living in conditions which they evolved for.
+However, in extreme conditions and drought scenarios soil water content (SWC) could become the limiting factor for ET response to VPD, which our theory does not account for.
+IN addition low soil moisture conditions themselves increase PET and VPD through land-atmosphere feedback (Berg NAT CL CC, BOUCHET MRTON< BRUTSAERT< LINTENER AND GENTINE Budkyko and complementary relatnship).
+We first note that in our datasets those conditions very very infrequent (talk about pdf figures).
+Within our framework, any systematic bias due to the  failure to account for SWC's effects in very dry conditions should manifest itself in a functional relationship between $\sigma$, our uncertainty measure, and SWC, which is observed at the FLUXNET sites.
+Examining the relationship between $\sigma$ and SWC will test to what extent our theory breaks down in very dry soil moisture conditions.
+
+If we view $\sigma \cdot uWUE$ as, in addition to a measure of uncertainty, a short time-scale observation of uWUE, we would expect $\sigma \cdot uWUE$ to decrease at low SWC (ZHOU ET AL).
+If $\sigma \cdot uWUE$ is a very strong function of SWC, then our theory should be conditioned more strongly on well watered soil conditions.
+If $\sigma \cdot uWUE$ is weakly a function of SWC, then our theory is more universal and independent of soil moisture conditions.
+
+Indeed, for all PFTs, there is some slight dependence of $\sigma \cdot uWUE$ on SWC, especially at low SWC (Figure \ref{swc}).
+The portion of observations for which our theory is biased by SWC-limitations varies by PFT, due to the nonlinear threshold at which soil moisture availability limits plant function.
+For CRO and DBF, soil water content only matters for about the lowest 5\% of observations (each box is 5\% of observations in Figure \ref{swc}).
+And, CRO, relative to DBF, has a very weak dependence on soil moisture, which is reflective of the high likelihood that CRO sites are optimally irrigated and not water stressed.
+For ENF, systematic SWC-induced biases in $\sigma \cdot uWUE$ emerge in about the lowest 30\% of SWC conditions.
+In crontast to ENF, CRO, and DBF, for GRA the relationship between $\sigma \cdot uWUE$ and SWC is more linear and affects a greater portion of observations; about 60\% of observations.
+Clearly, for GRA soil water frequently impacts plant function and alters ET response, and our theory is limited.
+It is therefore not suprising that for GRA our theory tested the worst (ARE YOU SURE ABOUT THIS WORD) againt the data (Section \ref{testing}); for all other PFTs occurence of soil mosture impacts was rare enough to not manifest itself in bulk statistics and plots\footnote{note that interpreting the soil moisture dependence for CSH is difficult as their are only two sites for that PFT.
+So systematic soil moisture relationshps are not apparent in Figure \ref{swc}, but instead blurred by the superposition of the two sites.
+We argue taht the soil moisture dependence is still there, but we don't have enough statistcs to comment on it}.
+
+MERGE WITH THE SENTENCE I PLACED ABOVE WITH LAND-ATMOPHERE COUPLPINGThe observed dependence of $\sigma \cdot uWUE$ on SWC for GRA would explain the deficiencies of our theory compared to the observations in Section \ref{testing}, specifically the trend back towards negative soil moisture at high VPD.
+Feedbacks between the land surface and the atmosphere, which our not accounted for due to our focus on the one-way response of the land surface to atmospheric conditions, would cause high VPD to be correlated with low SWC \citep[][Gentine et al.
+GRL 2016]{Berg_2016}\explain{Pierre, which publication were you thinking, your Evaporation from Weather Data, YES?}.
+So, at high VPD observations of low SWC are more likely, and this low SWC causes a lower uWUE (ZHOU ET AL).
+The lower uWUE at low SWC/high VPD then leads to the observed downturn towards decreasing ET with increasing VPD, and the deviation between our theory (based on a constant uWUE assumption) and observations.
+
+To summarize, our theory is limited by its inability to account for soil water impacts on land surface response, and feedbacks between SWC and VPD.
+Fortunately, for most PFTs SWC's effect on ET is relatively rare (<30\% of observations) and does not manifest itself in the majority of observations and bulk statistics.
+However, for GRA SWC has a strong effect decreasing water use efficiency for the majority of the observations.
+Soil moisture effects explain the deficiencies of our theory in Section \ref{testing}, particularly for GRA.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{./swc_boxplot.pdf}
+\caption{$\sigma \cot uWUE$ as a function of PFT and SWC. Each boxplot represents 5\% of all observations.}
+\label{swc}
+\end{figure}
+
 \section{Conclusions}
 
-We derived a new form of Penman Monteith using $uWUE$ (\cite{Zhou_2015}) to remove the stomatal conductance term's implicit dependence on ET.
-With our new form of Penman Monteith we developed a theory for when an ecosystem will tend to reduce or increase ET with increasing VPD.
+We derived a new form of Penman Monteith using the concept of semi-empirical optimal stomatal regulation (CITE LIN AND MEDLYN) and near constant $uWUE$ (\cite{Zhou_2015}) to remove the stomatal conductance implicit dependence on GPP and ET.
+With our new form of Penman Monteith we developed a theory for when an ecosystem will tend to reduce or increase ET with increasing VPD, which we evaluated against a range of eddy-covariance data spanning different climates and plant functional types.
 The goal was to capture the leading order behavior of the system to gain some intrinsic knowledge for its behavior.
@@
 For ENF, the theory succeeds in capturing the leading order behavior for the sign of $\frac{\partial \; ET}{\partial \; VPD}$, but not for magnitude.
-In contrast, the model does not capture the behavior for CRO and GRA.
-In some ways we should have expected this, as our theory does not account for large sources of variability we would expect for those plant types, including varying surface roughness, c3 vs c4 photosynthesis, and potential differences in irrigation across sites.
+In contrast, the model does not capture the behavior for eddy-covariance CRO and GRA, which could have been expected, as our theory does not account for large sources of heterogeneity in those plant types, including varying surface roughness, c3 vs c4 photosynthesis, and potential differences in irrigation across sites.
+Also, our theory does not account for strong soil drying, which non-negligibly alters ET response for GRA.
 Given these deficiencies for CRO and GRA, we will withhold conclusions on their response to increasing VPD.
@@
 However for CSH the environmental VPD never exceeds the critical VPD, and even when uncertainty is included the sign of the derivative is almost always negative (93\% of observations).
+MAYBE HAVE A SIMPLE CONCLUDING STATEMETN REGARDING THE FACT THAT MOST PLANTS ARE CONSERVATIVE EXCEPT FOR CRO NA D THAT STOMATAL REGULATION STONGLY REGULATES ET. BASICALLY ANSWER THE TITEL QUESTION.
 
-While our theory predicts common occurrence of positive $\frac{\partial \; ET}{\partial \; VPD}$ (atmospheric demand dominates), the plant response is still important even when $\frac{\partial \; ET}{\partial \; VPD} > 0$.
+While our theory predicts common occurrence of positive $\frac{\partial \; ET}{\partial \; VPD}$ (atmospheric demand dominates), the plant regulation is still important even when $\frac{\partial \; ET}{\partial \; VPD} > 0$.
 The magnitude of the ``sign term'' of our theory is always far below what it would be if we only considered the atmospheric demand term (PET).
@@
 
-This paper provides initial intuition for how the land surface responds to atmospheric drying.
-This intuition for the one-way behavior of  land surface response to atmospheric drying is critical first step to disentangling land-atmosphere feedbacks at various scales, from diurnal to seasonal and beyond.
-
-%%
-%% Enter Figures and Tables near as possible to where they are first mentioned:
-%
-% DO NOT USE \psfrag or \subfigure commands.
-%
-% Figure captions go below the figure.
-% Table titles go above tables;  other caption information
-%  should be placed in last line of the table, using
-% \multicolumn2l{$^a$ This is a table note.}
-%
-%----------------
-% EXAMPLE FIGURE
-%
-% \begin{figure}[h]
-% \centering
-% when using pdflatex, use pdf file:
-% \includegraphics[width=20pc]{figsamp.pdf}
-%
-% when using dvips, use .eps file:
-% \includegraphics[width=20pc]{figsamp.eps}
-%
-% \caption{Short caption}
-% \label{figone}
-%  \end{figure}
-%
-% ---------------
-% EXAMPLE TABLE
-%
-% \begin{table}
-% \caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
-% \centering
-% \begin{tabular}{l c}
-% \hline
-%  Run  & Time (min)  \\
-% \hline
-%   $l1$  & 260   \\
-%   $l2$  & 300   \\
-%   $l3$  & 340   \\
-%   $h1$  & 270   \\
-%   $h2$  & 250   \\
-%   $h3$  & 380   \\
-%   $r1$  & 370   \\
-%   $r2$  & 390   \\
-% \hline
-% \multicolumn{2}{l}{$^{a}$Footnote text here.}
-% \end{tabular}
-% \end{table}
-
-%% SIDEWAYS FIGURE and TABLE 
-% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
-%
-% \begin{sidewaysfigure}
-% \includegraphics[width=20pc]{figsamp}
-% \caption{caption here}
-% \label{newfig}
-% \end{sidewaysfigure}
-% 
-%  \begin{sidewaystable}
-%  \caption{Caption here}
-% \label{tab:signif_gap_clos}
-%  \begin{tabular}{ccc}
-% one&two&three\\
-% four&five&six
-%  \end{tabular}
-%  \end{sidewaystable}
-
-%% If using numbered lines, please surround equations with \begin{linenomath*}...\end{linenomath*}
-%\begin{linenomath*}
-%\begin{equation}
-%y|{f} \sim g(m, \sigma),
-%\end{equation}
-%\end{linenomath*}
-
-%%% End of body of article
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Optional Appendix goes here
-%
-% The \appendix command resets counters and redefines section heads
-%
-% After typing \appendix
-%
-%\section{Here Is Appendix Title}
-% will show
-% A: Here Is Appendix Title
-%
-%\appendix
-%\section{Here is a sample appendix}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Optional Glossary, Notation or Acronym section goes here:
-%
-%%%%%%%%%%%%%%  
-% Glossary is only allowed in Reviews of Geophysics
-%  \begin{glossary}
-%  \term{Term}
-%   Term Definition here
-%  \term{Term}
-%   Term Definition here
-%  \term{Term}
-%   Term Definition here
-%  \end{glossary}
-
-%
-%%%%%%%%%%%%%%
-% Acronyms
-%   \begin{acronyms}
-%   \acro{Acronym}
-%   Definition here
-%   \acro{EMOS}
-%   Ensemble model output statistics 
-%   \acro{ECMWF}
-%   Centre for Medium-Range Weather Forecasts
-%   \end{acronyms}
-
-%
-%%%%%%%%%%%%%%
-% Notation 
-%   \begin{notation}
-%   \notation{$a+b$} Notation Definition here
-%   \notation{$e=mc^2$} 
-%   Equation in German-born physicist Albert Einstein's theory of special
-%  relativity that showed that the increased relativistic mass ($m$) of a
-%  body comes from the energy of motion of the body—that is, its kinetic
-%  energy ($E$)—divided by the speed of light squared ($c^2$).
-%   \end{notation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%  ACKNOWLEDGMENTS
-%
-% The acknowledgments must list:
-%
-% •	All funding sources related to this work from all authors
-%
-% •	Any real or perceived financial conflicts of interests for any
-%	author
-%
-% •	Other affiliations for any author that may be perceived as
-% 	having a conflict of interest with respect to the results of this
-% 	paper.
-%
-% •	A statement that indicates to the reader where the data
-% 	supporting the conclusions can be obtained (for example, in the
-% 	references, tables, supporting information, and other databases).
-%
-% It is also the appropriate place to thank colleagues and other contributors. 
-% AGU does not normally allow dedications.
+This paper provides novel intuition and quantitate assessment on how the land surface responds to atmospheric drying.
+This intuition for the response of the land surface response to atmospheric drying is critical first step to disentangling land-atmosphere feedbacks at various scales, from diurnal to seasonal and beyond, and to characterize ET response in a warmer, atmospherically drier, and enriched CO$_2$ world.
 
@@
 
-%% ------------------------------------------------------------------------ %%
-%% Citations
-
-% Please use ONLY \citet and \citep for reference citations.
-% DO NOT use other cite commands (e.g., \cite, \citeyear, \nocite, \citealp, etc.).
-
-%% Example \citet and \citep:
-%  ...as shown by \citet{Boug10}, \citet{Buiz07}, \citet{Fra10},
-%  \citet{Ghel00}, and \citet{Leit74}.
-
-%  ...as shown by \citep{Boug10}, \citep{Buiz07}, \citep{Fra10},
-%  \citep{Ghel00, Leit74}.
-
-%  ...has been shown \citep [e.g.,][]{Boug10,Buiz07,Fra10}.
-
-%%  REFERENCE LIST AND TEXT CITATIONS
-%
-% Either type in your references using
-%
-% \begin{thebibliography}{}
-% \bibitem[{\textit{Kobayashi et~al.}}(2003)]{R2013} Kobayashi, T.,
-% Tran, A.~H., Nishijo, H., Ono, T., and Matsumoto, G.  (2003).
-% Contribution of hippocampal place cell activity to learning and
-% formation of goal-directed navigation in rats. \textit{Neuroscience}
-% 117, 1025--1035.
-%
-% \bibitem{}
-% Text
-% \end{thebibliography}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Or, to use BibTeX:
-%
-% Follow these steps
-%
-% 1. Type in \bibliography{<name of your .bib file>} 
-%    Run LaTeX on your LaTeX file.
-%
-% 2. Run BiBTeX on your LaTeX file.
-%
-% 3. Open the new .bbl file containing the reference list and
-%   copy all the contents into your LaTeX file here.
-%
-% 4. Run LaTeX on your new file which will produce the citations.
-%
 % AGU does not want a .bib or a .bbl file. Please copy in the contents of your .bbl file here.
 %\bibliography{references.bib}
-\input{vpd_et_paper.bbl}
+\input{vpd_et_paper_PG_v1.bbl}
 

### 2018-02-12 doc/paper/vpd_et_paper.tex
back tow ork, jsut schanged a couple citations

@@
 The first, a hydrometeorological perspective, is that higher VPD increases atmospheric demand for water from the land surface, and this drives an increase in evapotranspiration (ET).
-This perspective is particularly relevant because potential evapotranspiration (PET), which is used in many drought indices and hydrometeorological studies (CITATION NEEDED), typically only quantifies changes in atmospheric demand with limited surface regulation though ecosystem resistance regualtion (Swann et al.
-2016 FOR EXAMPLE).
+This perspective is particularly relevant because potential evapotranspiration (PET), which is used in many drought indices and hydrometeorological studies \citep[e.g.,][]{Heim_2002, Scheff_2015}, typically only quantifies changes in atmospheric demand with limited surface regulation though ecosystem resistance regualtion \citep{Swann_2016}.
 On the other hand, there is another perspective for vegetated surfaces.
@@
 With Equation \ref{et}, ET is explicitly a function of environmental variables and two plant-specific constants, the slope parameter ($g_1$), and $uWUE$, both reflecting water conservation strategy.
-The slope parameter is a leaf-scale term related to the willingness of stomata to trade water for CO_2 and to keep stomata open.
+The slope parameter is a leaf-scale term related to the willingness of stomata to trade water for CO$_2$ and to keep stomata open.
 $uWUE$ is semi-empirical ecosystem-scale constant related to how WUE changes with VPD (specifically $VPD^{-1/2}$).
@@
 
-We note that $VPD_{crit}$ mostly depends on the PFT parameters $uWUE$ and $g_1$ but only very weakly on climate as most other parameters related to the environment are nearly constant.
+We note that $VPD_{crit}$ mostly depends on the PFT parameters $uWUE$ and $g_1$ but only vary weakly with climate as most other parameters related to the environment are nearly constant.
 The value of $VPD_{crit}$ for each PFT is shown in Table \ref{vpd_crit}.
@@
 % AGU does not want a .bib or a .bbl file. Please copy in the contents of your .bbl file here.
-%\bibliography{references.bib}
-\input{vpd_et_paper_PG_v1.bbl}
+\bibliography{references.bib}
+% \input{vpd_et_paper.bbl}
 

### 2018-02-12 doc/paper/vpd_et_paper.tex
good progress on revision, stuck at uwue statement

@@
 On the other hand, there is another perspective for vegetated surfaces.
-Plants' stomata have evolved to optimally regulate the exchange of water and carbon, and tend to partially close in response to increased atmospheric dryness I order to save \citep{Ball_1987, Leuning_1990, MEDLYN_2011}.
+Plants' stomata have evolved to optimally regulate the exchange of water and carbon, and tend to partially close in response to increased atmospheric dryness \citep{Ball_1987, Leuning_1990, MEDLYN_2011}.
 Therefore, an increase in VPD, particularly in well-watered soil conditions, may actually correspond to a decrease in ET because of stomatal closure.
@@
 If stomata were fully passive \citep [similar to soil pores, e.g.
-][]{Or_2013}, increased atmospheric aridity would strongly reduce soil moisture (BERG GRL ON SOIL MOSITURE DIFFERNT BHEVAIOR IN TOP AND DEEP SOIL).
-In turn, this would further increase aridity as low soil moisture levels increase the Bowen ratio, and in turn increase temperature and atmospheric drying \citep[gentine et al.
-GRL 2016][]{Berg_2016}\explain{Pierre, which publication were you thinking, your Evaporation from Weather Data one?
-YES.
-Also cite papers on complementary relationship BRUTSAERT 1999, BOUCHET and MORTON ETC...
-all Salvucci and Gentine PNAS}.
+][]{Or_2013}, increased atmospheric aridity would strongly reduce soil moisture \citep{Berg_2017}.
+In turn, this would further increase aridity as low soil moisture levels increase the Bowen ratio, and in turn increase temperature and atmospheric drying \citep[][]{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006, Salvucci_2013, Gentine_2016, Berg_2016}.
 This however would not seem to be a sensible strategy for plants from an evolutionary standpoint.
@@
 As atmospheric demand for water (VPD) increases further, ecosystems should begin to reach their water conservation limits and might not be able to entirely limit the ET flux to the atmosphere.
-At this stage any further increase in VPD will most likely drive a (limited) increase in ET, because there is little more the plants can do keep water back (REFOMRAULTE LAST PART OF SENSTENCE).
+At this stage any further increase in VPD will most likely drive a (limited) increase in ET, because the increase in atmospheric demand for water overwhelms the limited plant response to conserve water.
 
@@
 \label{data}
-We use both meteorological and eddy-covariance data from the FLUXNET2015 database (WEBSITE LINK), including all sites with more than four years of data.
+We use both meteorological and eddy-covariance data from the FLUXNET2015 database, including all sites with more than four years of data (data available at https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/).
 Each site's plant functional type (PFT) was classified using the International Geosphere-Biosphere Programme vegetation classification scheme \citep{Loveland_1999}.
@@
 So, $g_1$ is a leaf-scale term reflecting the tradeoff of water for carbon uptake and growth.
-The higher $g_1$, the more open are stomata and the more they release water in exchange for carbon.
+The higher $g_1$, the more open are the stomata and the more they release water in exchange for carbon.
 
-The Medlyn model for stomatal conductance has been shown to behave very well across PFTs, compared to other models from the leaft to ecosystem scale \citep{Lin_2015, CITE ZHOU ET AL AS WELL}.
+The Medlyn model for stomatal conductance has been shown to behave very well across PFTs, compared to other models from the leaf to ecosystem scale \citep[, CITE ZHOU ET AL AS WELL** ??][]{Lin_2015, Zhou_2015}.
 It is adapted at the ecosystem scale by multiplying $g_{l-s}$ by leaf area index (LAI) and converting units to m s$^{-1}$:
@@
 \end{linenomath*}
-is relatively constant across time and moisture conditions within a plant functional type (CITE ALSO CHAGJIE), in non-extreme drought conditions.
+is relatively constant across time and moisture conditions within a plant functional type (CITE ALSO CHAGJIE ** ??), in non-extreme drought conditions.
 If, following \citet{Lin_2015}, we approximate $g_0$ as $0$ (i.e.
@@
 $uWUE$ is semi-empirical ecosystem-scale constant related to how WUE changes with VPD (specifically $VPD^{-1/2}$).
-In the eddy-covariance FLUXNET2015 data described in Section \ref{data}, every term in Equation \ref{et} is observed, except for the two parameters $g_1$, and $uWUE$, which have however been measured for the PFTs considered here (REF NEEDED HERE).
+In the eddy-covariance FLUXNET2015 data described in Section \ref{data}, every term in Equation \ref{et} is observed, except for the two parameters $g_1$, and $uWUE$, which have however been measured for the PFTs considered here \citep{Zhou_2015, Lin_2015}.
 However, eddy-covariance data are inherently noisy so that we include a measure of uncertainty in our analysis.
@@
 This makes Equation \ref{et} a useful alternative to PET in drought indices and hydrometeorological analysis for vegetated surfaces.
-Equation \ref{et} better reflects the physics of water exchange at the land surface and would only require fitting of $uWUE$ and $g_1$, and can also account for changes in CO$_2$ concentration, contrary to typical drought indices such as the Palmer Drought Severity Index (PDSI) (Swann et al.
-2016).
+Equation \ref{et} better reflects the physics of water exchange at the land surface and would only require fitting of $uWUE$ and $g_1$, and can also account for changes in CO$_2$ concentration, contrary to typical drought indices such as the Palmer Drought Severity Index (PDSI) \citep{Swann_2016}.
 
@@
 CO$_2$ concentration ($c_s$) and the psychometric constant ($\gamma$) are relatively constant over the dataset considered here so that the variability is dominated by $\sigma$ and $VPD$.
-$uWUE$ could vary with soil moisture but has been shown to be relatively constant except in very dry soil conditions  (ZHOU ET AL, LIN ET AL)\explain{Pierre, which Zhou and Lin citations were you referring to that show that uWUE changes with very dry soil conditions?}.
+$uWUE$ could vary with soil moisture but has been shown to be relatively constant except in very dry soil conditions  (ZHOU ET AL, LIN ET AL ** ??)\explain{Pierre, which Zhou and Lin citations were you referring to that show that uWUE changes with very dry soil conditions?}.
 We thus further assume that the soil is not too dry so that $uWUE$ can be approximated as a constant (within PFT).
@@
 Indeed, even in the absence of soil pore evaporation, stomata do not shut down entirely at very high VPD so that ET does not go to zero, because stomata are still slightly open to perform some photosynthesis \citep{Ball_1987, Leuning_1990, MEDLYN_2011}.
-In addition, upward xylem transport is necessary to maintain phloem transport, as well as nutrient transport and thus carbon allocation (De Schepper, V., T.
-De Swaef, I.
-Bauweraerts, and K.
-Steppe (2013), Phloem transport: a review of mechanisms and controls, J Exp Bot, 64(16), 4839–4850, doi:10.1093/jxb/ert302.
-Nikinmaa, E., T.
-Hölttä, P.
-Hari, P.
-Kolari, A.
-Makela, S.
-Sevanto, and T.
-Vesala (2012), Assimilate transport in phloem sets conditions for leaf gas exchange, Plant Cell Environ, 36(3), 655–669, doi:10.1111/pce.12004.
-Ryan, M. G., and S. Asao (2014), Phloem transport in trees, Tree Physiol, 34(1), 1–4, doi:10.1093/treephys/tpt123.
-).
+In addition, upward xylem transport is necessary to maintain phloem transport, as well as nutrient transport and thus carbon allocation \citep{De_2013, Nikinmaa_2013, Ryan_2014}.
 
@@
 
-Differences in $VPD_{crit}$ is exclusively determined by $uWUE$ and the slope parameter ($g_1$) related to the plant functional type.
+Differences in $VPD_{crit}$ are exclusively determined by $uWUE$ and the slope parameter ($g_1$) related to the plant functional type.
 A larger underlying water use efficiency, $uWUE$, causes means that at the ecosystem scale $WUE$ decreases more with increasing VPD, so plants lose more water per unit carbon production and are more likely to increase ET with increasing VPD (REFLRMUALTE SENTENCE, THE RESULTS IS CONTER INUITTICE WE EDPECT THE OPPOSITE).

### 2018-02-13 doc/paper/vpd_et_paper.tex
some progress on pierre's comments, still thinking of delta role though

@@
 Differences in $VPD_{crit}$ are exclusively determined by $uWUE$ and the slope parameter ($g_1$) related to the plant functional type.
-A larger underlying water use efficiency, $uWUE$, causes means that at the ecosystem scale $WUE$ decreases more with increasing VPD, so plants lose more water per unit carbon production and are more likely to increase ET with increasing VPD (REFLRMUALTE SENTENCE, THE RESULTS IS CONTER INUITTICE WE EDPECT THE OPPOSITE).
+A larger $uWUE$ means a smaller $VPD_{crit}$, and an ET response to increases VPD that is more likely to be positive.
+At first glance this result is somewhat counterintuitive; we expect that plants with a higher water use effiency would be more water conservative.
+However, in reality $uWUE$ determines how $WUE$ changes with VPD:
+
+\[WUE = \frac{GPP}{ET} = \frac{uWUE}{\sqrt{VPD}}.\]
+\[\frac{\partial \; WUE}{\partial \; VPD} = -\frac{uWUE}{2 \; VPD^{3/2}}\]
+
+So, plants with a higher $uWUE$ will have a greater \textit{decrease} in ecosystem-scale $WUE$ in response to increases in VPD.
+This decrease in $WUE$ causes more water loss per unit carbon production, and explains the relationship between high $uWUE$ and high liklihood of increases of ET in response to increasing atmospheric drying (increases in VPD).
+
 A tendency towards increasing ET response with increasing VPD can also be caused by a high slope parameter ($g_1$), characteristic of plants that at the leaf scale are more willing to trade water for stomata opening.
@@
 It also has a slightly higher uWUE ($2.32$ -units-) compared to CSH ($uWUE=2.11$ -units-).
-MAYBE A DUMB QUESTION BUT CAN'T uWUE AND G1 BE ANALYTICALLY RELTED> I THINK AT LESAT YOU SHOULD MAKE CLEAR THAT UWUE IS AT THE ECOYSTEM SCALE (DID YOU FIT IT OR USED PUBLISHED DATA-I AM NOW CONFUSED).
+\explain{MAYBE A DUMB QUESTION BUT CAN'T uWUE AND G1 BE ANALYTICALLY RELTED> I THINK AT LESAT YOU SHOULD MAKE CLEAR THAT UWUE IS AT THE ECOYSTEM SCALE (DID YOU FIT IT OR USED PUBLISHED DATA-I AM NOW CONFUSED)}.
 These differences in slope parameter cause opposite ET responses to changes in VPD between CRO and CSH.
 ET theoretically always increases with increasing VPD for the more water intensive CRO, while ET almost always decreases with increasing VPD for the more water conservative CSH (Figure \ref{term3}).
-CRO, evolved or were bred to prioritize GPP and yield and thus not water conservative so that they are very willing to trade water for photosynthesis and productivity, despite changes in VPD, while CSH are very unwilling to trade water for more photosynthesis.
+CRO evolved or were bred to prioritize GPP and yield and are thus not water conservative.
+They are very willing to trade water for photosynthesis and productivity, despite changes in VPD, while CSH are very unwilling to trade water for more photosynthesis.
 
-AS expected, the slope parameter ($g_1$) is a primary determinant of the VPD dependence for the sign term shown in Figure \ref{term3}.
+As expected, the slope parameter ($g_1$) is a primary determinant of the VPD dependence for the sign term shown in Figure \ref{term3}.
 Plants that are more conservative (small $g_1$) will tend to reduce ET with increasing VPD, and will be very effective at reducing ET, especially at low VPD.
@@
 
-$\frac{P}{T}$ is an air-density term, which we varies little compared to aerodynamic conductance and Clausius-Clapeyron ($\Delta$).
+$\frac{P}{T}$ is an air-density term, which varies little compared to aerodynamic conductance and Clausius-Clapeyron ($\Delta$).
 The psychometric constant ($\gamma$) is also relatively constant, so the scaling term should be primarily a function of aerodynamic conductance and temperature, through the Clausius-Clapeyron relationship $\Delta$.

### 2018-02-14 doc/paper/vpd_et_paper.tex
done with pierre's comments, still ahve some q's (mostly with citations), and then just clean up, format to Glob Change Bio, and submit

@@
 $\Delta$'s presence in the scaling term also matches physical intuition.
-Evaporative cooling will dampen the ability of the atmosphere to take more moisture (UNCLEAR SENTENCE REFORMULATE), because $e_{s}$ decreases with decreasing temperature.
+Evaporative cooling will dampen the ability of the atmosphere to take more moisture (UNCLEAR SENTENCE REFORMULATE **), because $e_{s}$ decreases with decreasing temperature.
 The decrease in $e_{s}$ is proportional to $\Delta$ ($\delta e_{s} = \Delta \delta T$).
@@
 I TRIED DOING IT
-In this section we consider direct observations of ET response with eddy-covariance data, while including uncertainty, with the $\sigma$ term (Section \ref{methods}).
+
+In this section we consider direct observations of ET response with eddy-covariance data, while including uncertainty with the $\sigma$ term (Section \ref{methods}).
 These observational results of ET response largely confirm our theoretical analysis, presented in Sections \ref{sign_term}-\ref{scale_term} (Table \ref{stats}).
@@
 This difference in mean ET response as compared to the ET response at mean environmental conditions is due to the non-linear nature of the response, in which negative responses are generally much larger magnitude than positive responses (Figure \ref{sign_plot}).
-SAY IT IN WORDS AS WELL AND WHTEHR IT FITS PREVIOUS REULTS
+Therefore, both the mean ET response as well as the ET response at mean environmental conditions matches our expectations from the theory (Section \ref{sign_term}), with the exception that CRO observations are shifted more towards a negative ET response than we expect.
 
-Regarding the frequency of negative and positive ET response, observations of ENF, GRA and CSH are more frequently characterized by a negative ET response to increasing VPD, and are thus more frequently water conservative TOO AVAGUE ESENTENCE REFORMUALTE.
-DBF and CRO observations have greater frequency of positive ET response to increasing VPD.
-i.e.
-water intensive strategies.
+Regarding the frequency of negative and positive ET response, 53\% of ENF observations, 65\% of GRA observations, and 93\% of CSH observations are characterized by a negative ET response to increasing VPD.
+The response for these PFTs is thus more frequently physiologicaly controlled, and these PFTs will tend to conserve soil water for futre use.
+However DBF and CRO observations have greater frequency of positive ET response to increasing VPD (46\% and 49\% of observations, respectively), and are more intensive in their use of water.
 
-In general the bulk statistics match our theoretical expectations well, with the caveat that inclusion of uncertainty shifts grass towards a slightly more negative response than the theory and Figure \ref{sign_plot} would suggest.
+In general the bulk statistics match our theoretical expectations well, with the caveat that inclusion of uncertainty shifts crops towards a slightly more negative response than the theory and Figure \ref{sign_plot} would suggest.
 The bulk statistics motivate a more thorough examination of the structure of uncertainty and more sophisticated validation of our theory's performance against observations.
@@
 \label{testing}
-MERGE WITH PREVIOUS PARAGRAPH
-We now turn to a direct evaluation of the functional forms presented in Section \ref{sign_term} against observations taken from a wide range of eddy-covariance datasets (Section \ref{methods}).
-Figure \ref{agu_fig} shows the distribution of of the sign term (when uncertainty is included), as compared to what the theory would predict.
+While Section \ref{stats_sec} presented simple statistics of observed response, we now compare more sophisticated distributions of the observed response to our theory presented in Section \ref{sign_term}.
+Figure \ref{agu_fig} shows observed the distribution of of the sign term, as compared to what the theory would predict.
 Our goal was to capture the leading order behavior of the ET dependence on VPD.
@@
 Theoretically, this $VPD_{crit}$ is mostly a function of PFT and not of climate.
-MAYBE MENTION THAT TEHR SHOULD BE STRONG SPECIES SELCETION BECAUSE OF THIS THRESHOLD
+This $VPD_{crit}$, and how it compares with environmental VPD and water supply, will exert a strong control on species selection in competitive environments.
 
@@
 
-While the stomatal conductance (MAKE SURE YOU SUE CONDUCTANE THROUGHOUT< YOU WERE USING RESITSANE SOMETIMES) s$g_a$ has a large impact on the magnitude of the scaling, temperature, through Clausius-Clapeyron ($\Delta$), does not appear to have as much of an impact in the complete estimation of ET response ($\frac{\partial \; ET}{\partial \; VPD}$).
+While the stomatal conductance $g_a$ has a large impact on the magnitude of the scaling, temperature, through Clausius-Clapeyron ($\Delta$), does not appear to have as much of an impact in the complete estimation of ET response ($\frac{\partial \; ET}{\partial \; VPD}$).
 Therefore, it appears that the absolute magnitude of ($\frac{\partial \; ET}{\partial \; VPD}$ can be mostly attributed to the sign term, and its direct scaling by aerodynamic conductance.
@@
 However, in extreme conditions and drought scenarios soil water content (SWC) could become the limiting factor for ET response to VPD, which our theory does not account for.
-IN addition low soil moisture conditions themselves increase PET and VPD through land-atmosphere feedback (Berg NAT CL CC, BOUCHET MRTON< BRUTSAERT< LINTENER AND GENTINE Budkyko and complementary relatnship).
-We first note that in our datasets those conditions very very infrequent (talk about pdf figures).
+In addition, low soil moisture conditions themselves increase VPD through land-atmosphere feedback \citep[][]{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006, Salvucci_2013, Gentine_2016, Berg_2016}.
+
+We first note that in our datasets low SWC conditions are very infrequent (MAKE SEPERATE PDF OF SWC?).
 Within our framework, any systematic bias due to the  failure to account for SWC's effects in very dry conditions should manifest itself in a functional relationship between $\sigma$, our uncertainty measure, and SWC, which is observed at the FLUXNET sites.
@@
 
-If we view $\sigma \cdot uWUE$ as, in addition to a measure of uncertainty, a short time-scale observation of uWUE, we would expect $\sigma \cdot uWUE$ to decrease at low SWC (ZHOU ET AL).
+If we view $\sigma \cdot uWUE$ as, in addition to a measure of uncertainty, a short time-scale observation of uWUE, we would expect $\sigma \cdot uWUE$ to decrease at low SWC (CITE ZHOU ** which ZHOU shows low uWUE with low SWC).
 If $\sigma \cdot uWUE$ is a very strong function of SWC, then our theory should be conditioned more strongly on well watered soil conditions.
@@
 Clearly, for GRA soil water frequently impacts plant function and alters ET response, and our theory is limited.
-It is therefore not suprising that for GRA our theory tested the worst (ARE YOU SURE ABOUT THIS WORD) againt the data (Section \ref{testing}); for all other PFTs occurence of soil mosture impacts was rare enough to not manifest itself in bulk statistics and plots\footnote{note that interpreting the soil moisture dependence for CSH is difficult as their are only two sites for that PFT.
+It is therefore not suprising that our theory tested poorly againt the data for GRA, relative to ENF, DBF, and GRA (Section \ref{testing}).
+For all other PFTs occurence of soil mosture impacts was rare enough to not manifest itself in bulk statistics and plots\footnote{note that interpreting the soil moisture dependence for CSH is difficult as their are only two sites for that PFT.
 So systematic soil moisture relationshps are not apparent in Figure \ref{swc}, but instead blurred by the superposition of the two sites.
@@
 
-MERGE WITH THE SENTENCE I PLACED ABOVE WITH LAND-ATMOPHERE COUPLPINGThe observed dependence of $\sigma \cdot uWUE$ on SWC for GRA would explain the deficiencies of our theory compared to the observations in Section \ref{testing}, specifically the trend back towards negative soil moisture at high VPD.
-Feedbacks between the land surface and the atmosphere, which our not accounted for due to our focus on the one-way response of the land surface to atmospheric conditions, would cause high VPD to be correlated with low SWC \citep[][Gentine et al.
-GRL 2016]{Berg_2016}\explain{Pierre, which publication were you thinking, your Evaporation from Weather Data, YES?}.
-So, at high VPD observations of low SWC are more likely, and this low SWC causes a lower uWUE (ZHOU ET AL).
+The observed dependence of $\sigma \cdot uWUE$ on SWC for GRA would explain the deficiencies of our theory compared to the observations in Section \ref{testing}, specifically the trend back towards negative soil moisture at high VPD.
+Aforementioned feedbacks between the land surface and the atmosphere, which are not accounted for due to our focus on the one-way response of the land surface to atmospheric conditions, would cause high VPD to be correlated with low SWC \citep[][]{Getine_2016, Berg_2016}.
+So, at high VPD observations of low SWC are more likely, and this low SWC causes a lower uWUE (ZHOU ET AL **).
 The lower uWUE at low SWC/high VPD then leads to the observed downturn towards decreasing ET with increasing VPD, and the deviation between our theory (based on a constant uWUE assumption) and observations.
@@
 
-We derived a new form of Penman Monteith using the concept of semi-empirical optimal stomatal regulation (CITE LIN AND MEDLYN) and near constant $uWUE$ (\cite{Zhou_2015}) to remove the stomatal conductance implicit dependence on GPP and ET.
+We derived a new form of Penman Monteith using the concept of semi-empirical optimal stomatal regulation \citep{Lin_2015, MEDLYN_2011} and near constant $uWUE$ \citep{Zhou_2015} to remove the stomatal conductance implicit dependence on GPP and ET.
 With our new form of Penman Monteith we developed a theory for when an ecosystem will tend to reduce or increase ET with increasing VPD, which we evaluated against a range of eddy-covariance data spanning different climates and plant functional types.
@@
 For ENF, the theory succeeds in capturing the leading order behavior for the sign of $\frac{\partial \; ET}{\partial \; VPD}$, but not for magnitude.
-In contrast, the model does not capture the behavior for eddy-covariance CRO and GRA, which could have been expected, as our theory does not account for large sources of heterogeneity in those plant types, including varying surface roughness, c3 vs c4 photosynthesis, and potential differences in irrigation across sites.
+In contrast, the model does not capture the behavior for eddy-covariance observations of CRO and GRA, which could have been expected, as our theory does not account for large sources of heterogeneity in those plant types, including varying surface roughness, c3 vs c4 photosynthesis, and potential differences in irrigation across sites.
 Also, our theory does not account for strong soil drying, which non-negligibly alters ET response for GRA.
@@
 However for CSH the environmental VPD never exceeds the critical VPD, and even when uncertainty is included the sign of the derivative is almost always negative (93\% of observations).
-MAYBE HAVE A SIMPLE CONCLUDING STATEMETN REGARDING THE FACT THAT MOST PLANTS ARE CONSERVATIVE EXCEPT FOR CRO NA D THAT STOMATAL REGULATION STONGLY REGULATES ET. BASICALLY ANSWER THE TITEL QUESTION.
 
-While our theory predicts common occurrence of positive $\frac{\partial \; ET}{\partial \; VPD}$ (atmospheric demand dominates), the plant regulation is still important even when $\frac{\partial \; ET}{\partial \; VPD} > 0$.
+Generally speaking, our theory suggests that for a majority of enviornmental VPD conditions, plants will tend to conserve water and reduce ET with increasing VPD.
+Stomatal regulation and plant physiological response strongly regulates ET, but his regulation does vary by PFT.
+CROs are the least water conservative, while DBF, ENF, and GRA are progressively more water conservative, with CSH begin the most water conservative (most likely to reudce ET in response to increasing VPD).
+However, these results are subject to the caveat that our theory's performance is most uncertain for GRA and CRO.
+
+While our theory predicts occurrence of increasing ET with increasing VPD (atmospheric demand dominates), plant regulation is still important even when $\frac{\partial \; ET}{\partial \; VPD} > 0$.
 The magnitude of the ``sign term'' of our theory is always far below what it would be if we only considered the atmospheric demand term (PET).

### 2018-02-15 doc/paper/vpd_et_paper.tex
a few more commengts to do

@@
 
-The Medlyn model for stomatal conductance has been shown to behave very well across PFTs, compared to other models from the leaf to ecosystem scale \citep[, CITE ZHOU ET AL AS WELL** ??][]{Lin_2015, Zhou_2015}.
+The Medlyn model for stomatal conductance has been shown to behave very well across PFTs, compared to other models from the leaf to ecosystem scale \citep[ZHOU ET AL AS WELL** ??
+- Pierre says Zhou 2014][]{Lin_2015, Zhou_2015}.
 It is adapted at the ecosystem scale by multiplying $g_{l-s}$ by leaf area index (LAI) and converting units to m s$^{-1}$:

### 2018-02-16 doc/paper/vpd_et_paper.tex
starting conversion to glboal chagne biology
added gcb

@@
 
-\journalname{Agricultural and Forest Meteorology}
+\journalname{Global Change Biology}
 
@@
 
-Vapor pressure deficit (VPD) is expected to rise over continents in the future due to the combination of increased temperature and relative humidity drying \citep{Byrne_2013}.
+Vapor pressure deficit (VPD) is expected to rise over continents in the future due to the combination of increased temperature and relative humidity drying ** RH drying \citep{Byrne_2013 }.
 Increases in VPD increases the atmospheric demand of evapotranspiration (ET) \citep{Monteith_1965} but also stresses plant stomata \citep{Leuning_1990, MEDLYN_2011}.

### 2018-02-16 doc/paper/vpd_et_paper.tex
added journal styles, reset latex to old

@@
 % 11/23/2015
-\documentclass[draft,linenumbers]{agujournal}
+\documentclass[draft,linenumbers]{gcbjournal}
 % \draftfalse
@@
 
-Vapor pressure deficit (VPD) is expected to rise over continents in the future due to the combination of increased temperature and relative humidity drying ** RH drying \citep{Byrne_2013 }.
+Vapor pressure deficit (VPD) is expected to rise over continents in the future due to the combination of increased temperature and relative humidity drying ** RH drying \citep{Byrne_2013}.
 Increases in VPD increases the atmospheric demand of evapotranspiration (ET) \citep{Monteith_1965} but also stresses plant stomata \citep{Leuning_1990, MEDLYN_2011}.
@@
 
-\begin{figure}[h]
+\begin{figure}[!p]
 \centering
@@
 
-% \begin{figure}[h]
+% \begin{figure}[!p]
 % \centering
@@
 The observed dependence of $\sigma \cdot uWUE$ on SWC for GRA would explain the deficiencies of our theory compared to the observations in Section \ref{testing}, specifically the trend back towards negative soil moisture at high VPD.
-Aforementioned feedbacks between the land surface and the atmosphere, which are not accounted for due to our focus on the one-way response of the land surface to atmospheric conditions, would cause high VPD to be correlated with low SWC \citep[][]{Getine_2016, Berg_2016}.
+Aforementioned feedbacks between the land surface and the atmosphere, which are not accounted for due to our focus on the one-way response of the land surface to atmospheric conditions, would cause high VPD to be correlated with low SWC \citep[][]{Gentine_2016, Berg_2016}.
 So, at high VPD observations of low SWC are more likely, and this low SWC causes a lower uWUE (ZHOU ET AL **).
@@
 
-\listofchanges
+% \listofchanges
 

### 2018-02-19 doc/paper/vpd_et_paper.tex
got formatting up to snuff with GCB, now can really focus on finishin up the writing
okay, going ot do one last run of analytics to dot my i's and cross my t's, then will be done

@@
 \begin{keypoints}
-\item = enter point 1 here = 
+\item Add given Medlyn (nature response) and Zhou2014 uWUE should increase roughly proportional to atmospheric CO2 concentration,
+  so really the env VPD will be where plants will feel the effect of global warming stronges, b/c effect of CO2 cancels.
 \item = enter point 2 here = 
@@
 
-The Medlyn model for stomatal conductance has been shown to behave very well across PFTs, compared to other models from the leaf to ecosystem scale \citep[ZHOU ET AL AS WELL** ??
-- Pierre says Zhou 2014][]{Lin_2015, Zhou_2015}.
+The Medlyn model for stomatal conductance has been shown to behave very well across PFTs, compared to other models from the leaf to ecosystem scale \citep[][** still unsure about citing Zhou here]{Lin_2015, Zhou_2014}.
 It is adapted at the ecosystem scale by multiplying $g_{l-s}$ by leaf area index (LAI) and converting units to m s$^{-1}$:

### 2018-02-20 doc/paper/vpd_et_paper.tex
starting to clean up for publication

@@
 
-Vapor pressure deficit (VPD) is expected to rise over continents in the future due to the combination of increased temperature and relative humidity drying ** RH drying \citep{Byrne_2013}.
-Increases in VPD increases the atmospheric demand of evapotranspiration (ET) \citep{Monteith_1965} but also stresses plant stomata \citep{Leuning_1990, MEDLYN_2011}.
-
-This conflicting response between increased demand and higher stomatal stress leads to two potential but opposing perspectives on how ET adjusts to changes in VPD.
-The first, a hydrometeorological perspective, is that higher VPD increases atmospheric demand for water from the land surface, and this drives an increase in evapotranspiration (ET).
+Vapor pressure deficit (VPD) is expected to rise over continents in the future due to the combination of increased temperature and relative humidity drying ** RH drying?
+\citep{Byrne_2013}.
+Increases in VPD increase the atmospheric demand for evapotranspirated water \citep{Penman_1948, Monteith_1965}, but also stress plant stomata \citep{Leuning_1990, MEDLYN_2011}.
+These  conflicting responses, increased atmospheric demand and higher stomatal stress, leads to two possible responses of evapotraspiration (ET) to changes in VPD.
+The first, hydrometeorological perspective, is that higher VPD increases atmospheric demand for water from the land surface, and this drives an increase in evapotranspiration (ET).
 This perspective is particularly relevant because potential evapotranspiration (PET), which is used in many drought indices and hydrometeorological studies \citep[e.g.,][]{Heim_2002, Scheff_2015}, typically only quantifies changes in atmospheric demand with limited surface regulation though ecosystem resistance regualtion \citep{Swann_2016}.

### 2018-02-20 doc/paper/vpd_et_paper.tex
makes me think, should we set our own uWUE so we can include EBF?

@@
 \journalname{Global Change Biology}
+
+\newcommand{\appropto}{\mathrel{\vcenter{
+  \offinterlineskip\halign{\hfil$##$\cr
+    \propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}
 
@@
 Increases in VPD increase the atmospheric demand for evapotranspirated water \citep{Penman_1948, Monteith_1965}, but also stress plant stomata \citep{Leuning_1990, MEDLYN_2011}.
-These  conflicting responses, increased atmospheric demand and higher stomatal stress, leads to two possible responses of evapotraspiration (ET) to changes in VPD.
-The first, hydrometeorological perspective, is that higher VPD increases atmospheric demand for water from the land surface, and this drives an increase in evapotranspiration (ET).
+The conflicting effects of increased atmospheric demand and higher stomatal stress lead to two possible perspectives for how evapotraspiration (ET) respondes to shifts in VPD.
+The first, a hydrometeorological perspective, is that higher VPD increases atmospheric demand for water from the land surface, and this drives an increase in evapotranspiration (ET).
 This perspective is particularly relevant because potential evapotranspiration (PET), which is used in many drought indices and hydrometeorological studies \citep[e.g.,][]{Heim_2002, Scheff_2015}, typically only quantifies changes in atmospheric demand with limited surface regulation though ecosystem resistance regualtion \citep{Swann_2016}.
@@
 Plants' stomata have evolved to optimally regulate the exchange of water and carbon, and tend to partially close in response to increased atmospheric dryness \citep{Ball_1987, Leuning_1990, MEDLYN_2011}.
-Therefore, an increase in VPD, particularly in well-watered soil conditions, may actually correspond to a decrease in ET because of stomatal closure.
+This leads to a plant physiology perspective, in which an increase in VPD, particularly in well-watered soil conditions, may actually correspond to a decrease in ET because of stomatal closure.
 This decrease in ET would reduce and potentially cancel, in the case of full stomatal closure, the effects of shifts to atmospheric demand.
-In other words, the  question ``When does VPD drive or reduce ET?'' can be related to whether plant regulation or atmospheric demand dominates ET.
+In other words, the  question ``When does VPD drive or reduce ET?'' can be related to whether plant regulation or atmospheric demand dominates ET response.
 
@@
 At the ecosystem scale, there are limits to plant water conservation strategies.
-As atmospheric demand for water (VPD) increases further, ecosystems should begin to reach their water conservation limits and might not be able to entirely limit the ET flux to the atmosphere.
+As atmospheric demand for water (VPD) increases further, ecosystems should begin to reach their water conservation limits and might not be able to entirely limit ET flux to the atmosphere.
 At this stage any further increase in VPD will most likely drive a (limited) increase in ET, because the increase in atmospheric demand for water overwhelms the limited plant response to conserve water.
 
-The objective of the present manuscript is thus to evaluate the VPD dependence of ET, in non-extreme soil drought conditions.
+The objective of the present manuscript is to evaluate the VPD dependence of ET, in non-extreme soil drought conditions.
 The goal of this paper is to use reasonable approximations as a tool to drive intuition for plant response to atmospheric drying.
@@
 The model is then validated and tested at multiple eddy-covariance stations spanning various climates and plant functional types.
-Section 2 describes the data used.
-Section 3 derives the framework.
-Section 4 presents results and comparison to observations.
-Section 5 discusses conclusions.
+Section 2 describes the data used and derives the framework.
+Section 4 presents results and comparison to observations, and  Section 5 discusses the conclusions.
 
-\section{Data}
+\section{Materials and Methods}
+\subsection{Data}
 \label{data}
@@
 Each site's plant functional type (PFT) was classified using the International Geosphere-Biosphere Programme vegetation classification scheme \citep{Loveland_1999}.
-The physical constants used for water use efficiency in Section \ref{methods} are only published for five plant functional types (PFTs): crops (CRO), deciduous broadleaf forest (DBF), evergreen needleleaf forest (ENF), grass (GRA), and closed shrub (CSH) (see Table \ref{pft}).
+The physical constants used for water use efficiency in the Methods section (Section \ref{methods}) are only published for five plant functional types (PFTs): crops (CRO), deciduous broadleaf forest (DBF), evergreen needleleaf forest (ENF), grass (GRA), and closed shrub (CSH) (see Table \ref{pft}).
 There are 56 sites with these plant functional types, and their location is shown in  Figure \ref{map_fig} \explain{map needs to be improved - it's a placeholder for now}.
@@
 
-\section{Methods}
+\subsection{Methods}
 \label{methods}
@@
   \begin{equation}
+    \label{orig_pen}
     ET = \frac{\Delta R + g_a \rho_a c_p VPD}{\Delta + \gamma(1 + \frac{g_a}{g_s})},
@@
   \begin{equation}
-  g_1 = \sqrt{\frac{3 \, \Gamma \, \lambda}{1.6}},\footnote{Note the above expression has units of of (mmol mol$^{-1}$)$^{1/2}$, but this can be converted to Pa$^{1/2}$ using the ideal gas law.}
+  g_1 = \sqrt{\frac{3 \, \Gamma^* \, \lambda}{1.6}},\footnote{Note the above expression has units of of (mmol mol$^{-1}$)$^{1/2}$, but this can be converted to Pa$^{1/2}$ using the ideal gas law.}
   \end{equation}
@@
 
-where $\Gamma$ is the CO$_2$ compensation point for photosynthesis (without dark respiration), and $\lambda$ is the marginal cost of leaf carbon.
+where $\Gamma^*$ is the CO$_2$ compensation point for photosynthesis (without dark respiration), and $\lambda$ is the marginal water cost of leaf carbon.
 So, $g_1$ is a leaf-scale term reflecting the tradeoff of water for carbon uptake and growth.
-The higher $g_1$, the more open are the stomata and the more they release water in exchange for carbon.
+The higher $g_1$, the more open the stomata and the more they release water in exchange for carbon.
 
@@
 While Equation 3 can be used in PM (Equation 1), it will make analytical work with the function intractable because $A$, net CO$_2$ assimilation, is functionally related to ET itself.
-To remove the dependence of ET on $A$ we use a fundamental semi-empirical results of \citet{Zhou_2015}, who showed that the underlying Water Use Efficiency $uWUE$:
+To remove the dependence of ET on $A$ we use fundamental semi-empirical results of \citet{Zhou_2015}, who showed that the underlying Water Use Efficiency ($uWUE$):
 
@@
   \begin{equation}
+    uWUE = \frac{GPP \cdot \sqrt{VPD}}{ET}
     \label{uwue}
-uWUE = \frac{GPP \cdot \sqrt{VPD}}{ET}
   \end{equation}
@@
     g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{uWUE \; ET}{c_s \; \sqrt{VPD}}
+    \label{new_g_s}
   \end{equation}
@@
 
-Note that $uWUE$ is fit on the ecosystem scale as in \citet{Zhou_2015} so GPP in \ref{uwue} is really $A\cdot \text{LAI}$.
-This leads to the cancellation of LAI in addition to uWUE in Equation \ref{medlyn}.
-Plugging Equation 5 into Equation 1 and rearranging gives a new expression for PM ET directly as a function of VPD with no $A$ dependence:
+Note that $uWUE$ is fit on the ecosystem scale as in \citet{Zhou_2015} so GPP in Equation \ref{uwue} is really $A\cdot \text{LAI}$.
+This leads to the cancellation of LAI with the addition of uWUE in Equation \ref{medlyn}.
+Plugging Equation \ref{new_g_s} into Equation \ref{orig_pen} and rearranging gives a new expression for PM ET directly as a function of VPD with no $A$ dependence:
 
@@
 $uWUE$ is semi-empirical ecosystem-scale constant related to how WUE changes with VPD (specifically $VPD^{-1/2}$).
-In the eddy-covariance FLUXNET2015 data described in Section \ref{data}, every term in Equation \ref{et} is observed, except for the two parameters $g_1$, and $uWUE$, which have however been measured for the PFTs considered here \citep{Zhou_2015, Lin_2015}.
-However, eddy-covariance data are inherently noisy so that we include a measure of uncertainty in our analysis.
+Is is also roughly proportional to physical constants:
+
+\[uWUE \appropto \sqrt{\frac{c_s - \Gamma}{1.6 \lambda}},\]
+
+where $\Gamma$ is the CO$_2$ compensation point \citep{Zhou_2014}.
+So uWUE is related to atmospheric CO$_2$ concentration and compensation point, and is inversely proportional to the marginal water cost of leaf carbon.
+Therefore, generally speaking, plants that are very willing to trade water for new growth will have lower $uWUE$ and higher $g_1$, and will be less water conservative (more likely to increase ET with increasing VPD).
+
+Given the eddy-covariance FLUXNET2015 data described in Section \ref{data}, every term in Equation \ref{et} is observed, except for the two parameters $g_1$, and $uWUE$, which have however been measured for the PFTs considered here \citep{Zhou_2015, Lin_2015}.
+However, eddy-covariance data are inherently noisy so we include a measure of uncertainty in our analysis.
 To account for both mean deviations of $uWUE$ and uncertainty, we introduce an uncertainty parameter $\sigma$ modifying $uWUE$:
@@
 This then means that the sign term only depends on VPD for a given PFT and is approximately just a function of $VPD$.
-We can further determine a critical threshold separating an increase from a decrease in ET, I.e.
+We can further determine a critical threshold separating an increase from a decrease in ET, i.e.
 the threshold $VPD_{crit}$ such as the derivative vanishes $\frac{\partial \; ET}{\partial \; VPD} = 0$:
-
+\small
 \begin{linenomath*}
@@
 \end{linenomath*}
-
+\normalsize
 We note that $VPD_{crit}$ mostly depends on the PFT parameters $uWUE$ and $g_1$ but only vary weakly with climate as most other parameters related to the environment are nearly constant.
@@
 
-\section{Conclusions}
+\section{Discussion}
 

### 2018-02-21 doc/paper/vpd_et_paper.tex
cleaned up figures, also stopped tracking figs for paper b/c will generate those with a makefile in when I publish the repo

@@
 Each site's plant functional type (PFT) was classified using the International Geosphere-Biosphere Programme vegetation classification scheme \citep{Loveland_1999}.
-The physical constants used for water use efficiency in the Methods section (Section \ref{methods}) are only published for five plant functional types (PFTs): crops (CRO), deciduous broadleaf forest (DBF), evergreen needleleaf forest (ENF), grass (GRA), and closed shrub (CSH) (see Table \ref{pft}).
-There are 56 sites with these plant functional types, and their location is shown in  Figure \ref{map_fig} \explain{map needs to be improved - it's a placeholder for now}.
+The physical constants used in the stomatal conductance model in the Methods section (Section \ref{methods}) are only published for six plant functional types (PFTs): crops (CRO), deciduous broadleaf forest (DBF), evergreed broadleaf forests (EBF), evergreen needleleaf forest (ENF), grass (GRA), and closed shrub (CSH) (see Table \ref{pft}).
+There are 62 sites with these plant functional types, and their location is shown in  Figure \ref{map_fig} \explain{map needs to be improved - it's a placeholder for now}.
 
@@
 Timesteps with negative observed GPP or ET are also removed, and we aggregate half hourly data to hourly averages to reduce noise.
-After these quality control procedures, 332556 upscaled hourly observations remain.
+After these quality control procedures, 341788 upscaled hourly observations remain.
 

### 2018-02-21 src/paper_figure_code/agu_fig.py

@@ -19,13 +19,6 @@ PLOT_SERIES = False
 #below is to make slight changes for the talk
 PLOT_TALK = True
 
-name_dict = {'CRO': 'Crops',\
-             'DBF': 'Deciduous Forest',
-             'EBF': 'Evergreen Broadleaf Forest',
-             'ENF': 'Evergreen Needleleaf Forest',
-             'GRA': 'Grass',
-             'CSH': 'Shrub (closed)'}
-
 paren_string = r'$\left(\frac{ c_p}{R_{air}} '\
                r'- \frac{\gamma c_s }{1.6 \; R\; \sigma \; uWUE  }'\
                r'\left( \frac{2 g_1 + \sqrt{VPD}}'\
@@ -70,9 +63,9 @@ def plot_box(_df):
   #_df.groupby(pd.cut(_df.vpd, bins=20)).apply(add_box, ax=ax)
   #_df.boxplot(column='sign', by=pd.cut(_df.vpd, bins=20), ax=ax)
   if (_df.pft.iloc[0] == 'ENF') | (_df.pft.iloc[0] == 'DBF')\
-     | (_df.pft.iloc[0] == 'CSH'):
+     | (_df.pft.iloc[0] == 'CRO'):
     ax.set_ylabel(paren_string, fontsize=fontsize)
-  if (_df.pft.iloc[0] == 'CSH') | (_df.pft.iloc[0] == 'GRA'):
+  if (_df.pft.iloc[0] == 'CRO') | (_df.pft.iloc[0] == 'GRA'):
     ax.set_xlabel('VPD (kPa)', fontsize=fontsize)
   ax.set_title(name_dict[_df.pft.iloc[0]], fontsize=fontsize+3)
   ax.plot(orig_xlim, [0., 0.], 'k--', linewidth=dashedlinewidth)
@@ -91,12 +84,12 @@ def plot_box(_df):
 #plot_box(_df)
 df.groupby('pft').apply(plot_box)
 
-os.system('pdfjam %s/DBF_box.pdf %s/ENF_box.pdf %s/CSH_box.pdf '\
+os.system('pdfjam %s/DBF_box.pdf %s/ENF_box.pdf %s/CRO_box.pdf '\
           '--nup 1x3 --no-landscape --outfile %s/first-column.pdf'\
           % tuple(['../../doc/shared_figs']*4))
 os.system('pdfcrop --margins 10 %s %s'\
[…]

### 2018-02-21 src/paper_figure_code/fig03.py

@@ -6,6 +6,7 @@ from shared_functions import *
 
 plt.close('all')
 
+importlib.reload(plot_tools)
 ### FIGURE 3 ###
 # joint distribution of lai and swc
 df['sigma'] = df['sigma'].copy()
@@ -15,6 +16,9 @@ meta['xlim'] = (0., 4000.)
 meta['plot_type'] = '' #'simple'
 meta['x_var'] = 'vpd'
 meta['y_var'] = 'sigma'
+meta['x_label'] = 'VPD (Pa)'
+meta['y_label'] = '$\sigma$'
+meta['pft_plot'] = False
 plot_tools.scatter_wrapper(df, meta)
 os.system('cp %s/climate_et/scatters/vpd_sigma.png '\
           '../../doc/paper/fig03.png'\

### 2018-02-22 doc/paper/vpd_et_paper.tex
added makefile for creating paper, works great

@@
 There are 62 sites with these plant functional types, and their location is shown in  Figure \ref{map_fig} \explain{map needs to be improved - it's a placeholder for now}.
+
+\begin{figure}[!p]
+\centering
+\includegraphics[width=\textwidth]{./map.pdf}
+\caption{Plant functional type and location of sites used in analysis. ***\textbf{This is just  a placeholder for now, I am currently fixing it***}}
+\label{map_fig}
+ \end{figure}
 
@@
 
-\begin{figure}[!p]
-\centering
-\includegraphics[width=20pc]{./fig01.pdf}
-\caption{Plant functional type and location of sites used in analysis. ***\textbf{This is just  a placeholder for now, I am currently fixing it***}}
-\label{map_fig}
- \end{figure}
-
 \section{Results}
@@
 Therefore, for our assumptions and framework to be reasonable $\sigma$ should be close to 1.
-Figure \ref{lai_fig} presents the histogram of calculated $\sigma$s with outliers removed (lowest and highest 5\% percent).
+Figure \ref{joint_vpd_sigma} presents the histogram of calculated $\sigma$s with outliers removed (lowest and highest 5\% percent).
 All remaining $\sigma$ values are close to unity (i.e.
@@
 \centering
-\includegraphics[width=20pc]{./fig02.png}
-\caption{Histogram of $\sigma$ values calculated for each site and time according to Equation \ref{sigma}.
-The lowest and highest 5\% are removed as outliers, as well as any values below 0.
-The curve is normalized such that its area is 1.
-}
-\label{lai_fig}
+\includegraphics[width=\textwidth]{./joint_vpd_sigma.pdf}
+\caption{The joint distribution of $VPD$ and $\sigma$. $\sigma$ has only a weak dependence on $VPD$.}
+\label{joint_vpd_sigma}
 \end{figure}
@@
 An additional concern is that $\sigma$ may in fact be correlated with $VPD$, in which case the dependence would need to be accounted for when taking the derivative.
-However, the joint distribution of $\sigma$ and VPD shows that $\sigma$ is very weakly a function of VPD (Figure \ref{lai_vpd_fig}).
+However, the joint distribution of $\sigma$ and VPD shows that $\sigma$ is very weakly a function of VPD (Figure \ref{joint_vpd_sigma}).
 Given this weak dependence, we argue that Equation \ref{d_et} is a valid approximation for ET response to $VPD$.
-
-\begin{figure}[!p]
-\centering
-\includegraphics[width=20pc]{./fig03.png}
-\caption{The joint distribution of $VPD$ and $\sigma$. $\sigma$ has only a weak dependence on $VPD$. \textbf{***note I am redoing/cleaning up this plot, ``lai'' should read sigma***}}
-\label{lai_vpd_fig}
-\end{figure}
 
@@
 These differences in slope parameter cause opposite ET responses to changes in VPD between CRO and CSH.
-ET theoretically always increases with increasing VPD for the more water intensive CRO, while ET almost always decreases with increasing VPD for the more water conservative CSH (Figure \ref{term3}).
+ET theoretically always increases with increasing VPD for the more water intensive CRO, while ET almost always decreases with increasing VPD for the more water conservative CSH (Figure \ref{idealized_sign}).
 CRO evolved or were bred to prioritize GPP and yield and are thus not water conservative.
@@
 
-As expected, the slope parameter ($g_1$) is a primary determinant of the VPD dependence for the sign term shown in Figure \ref{term3}.
+\begin{figure}[!p]
+\centering
+\includegraphics[width=\textwidth]{./idealized_sign.pdf}
+\caption{Sources of variability for the sign term.
+Top: sign term as a function of VPD, with $\sigma$ held fixed at 1.
+The observed range of VPD for each PFT is also shown below the x-axis.
+Line extent corresponds to 5th and 95th percentiles, while stars denote the location of the 25th, 50th, and 75th percentiles.
+
+Bottom: The location of the minima of ET, as a function of VPD and $\sigma$.
+Lines and stars denote the distribution of VPD and $\sigma$ next to each axis, following the same percentiles as above.
+\textbf{***I think I'm going to remove the bottom plot: its discussion didn't fit in well with the text, and the box plots test inclusion of uncertainty***}}
+\label{idealized_sign}
+\end{figure}
+
+As expected, the slope parameter ($g_1$) is a primary determinant of the VPD dependence for the sign term shown in Figure \ref{idealized_sign}.
 Plants that are more conservative (small $g_1$) will tend to reduce ET with increasing VPD, and will be very effective at reducing ET, especially at low VPD.
@@
 Therefore, plants with a low $g_1$ will have the largest VPD dependence of ET response because the difference in ET response at low VPD (leaf stomatal response dominates) and high VPD (VPD gradient dominates) is largest.
-This is apparent in the strong VPD dependence of ENF, which has the lowest slope parameter ($g_1=74.31$), in Figure \ref{term3}.
+This is apparent in the strong VPD dependence of ENF, which has the lowest slope parameter ($g_1=74.31$), in Figure \ref{idealized_sign}.
 
-To summarize our theoretical insights, Figure \ref{term3} and Table \ref{vpd_crit}, CROs are the least water conservative and have the strongest overall tendency to increase ET with increasing VPD, while CSH are the most water conservative and have the strongest tendency to decrease ET with increasing VPD.
+To summarize our theoretical insights, Figure \ref{idealized_sign} and Table \ref{vpd_crit}, CROs are the least water conservative and have the strongest overall tendency to increase ET with increasing VPD, while CSH are the most water conservative and have the strongest tendency to decrease ET with increasing VPD.
 ENF ($g_1 = 74.31$) has by far the largest VPD response dependence, while CRO ($g_1 = 183.1$) has the smallest VPD dependence.
@@
 
-Figure \ref{term3} clearly shows, according to our theory, that for all PFTs except for crops there is frequent occurrence of a negative (plant dominating) ET response to increases in VPD.
+Figure \ref{idealized_sign} clearly shows, according to our theory, that for all PFTs except for crops there is frequent occurrence of a negative (plant dominating) ET response to increases in VPD.
 Therefore, plants are able in most atmospheric conditions to reduce ET in response to increased VPD and thus to reduce water loss.
@@
 Therefore, the actual land surface response to a change in VPD will always me more negative than if the plants are not included in the physics (PET).
-
-\begin{figure}[!p]
-\centering
-\includegraphics[width=20pc]{./fig05.pdf}
-\caption{Sources of variability for the sign term.
-Top: sign term as a function of VPD, with $\sigma$ held fixed at 1.
-The observed range of VPD for each PFT is also shown below the x-axis.
-Line extent corresponds to 5th and 95th percentiles, while stars denote the location of the 25th, 50th, and 75th percentiles.
-
-Bottom: The location of the minima of ET, as a function of VPD and $\sigma$.
-Lines and stars denote the distribution of VPD and $\sigma$ next to each axis, following the same percentiles as above.
-\textbf{***I think I'm going to remove the bottom plot: its discussion didn't fit in well with the text, and the box plots test inclusion of uncertainty***}}
-\label{sign_plot}
-\end{figure}
 
@@
 
+\begin{figure}[!p]
+\centering
+\includegraphics[width=\textwidth]{./idealized_scale.pdf}
+\caption{Primary sources of variability for the scaling term, as a function of PFT.
+Note the change in axis scale for each plot.
+The 5th-95th percentile range of temperature is plotted at the 5th, 25th, 50th, 75th, and 95th percentiles of aerodynamic conductance, as observed for each PFT's sites.}
+\label{scale_vary}
+\end{figure}
+
 Within each PFT, the scaling term variability is controlled both by environmental temperature and aerodynamic conductance variability (Figure \ref{scale_vary}).
@@
 But, aerodynamic conductance variability generally impacts the scaling term more than temperature, even within PFT.
-
-\begin{figure}[!p]
-\centering
-\includegraphics[width=\textwidth]{./fig04b.pdf}
-\caption{Primary sources of variability for the scaling term, as a function of PFT.
-Note the change in axis scale for each plot.
-The 5th-95th percentile range of temperature is plotted at the 5th, 25th, 50th, 75th, and 95th percentiles of aerodynamic conductance, as observed for each PFT's sites.}
-\label{scale_vary}
-\end{figure}
 
@@
 While Section \ref{stats_sec} presented simple statistics of observed response, we now compare more sophisticated distributions of the observed response to our theory presented in Section \ref{sign_term}.
-Figure \ref{agu_fig} shows observed the distribution of of the sign term, as compared to what the theory would predict.
+Figure \ref{test_sign} shows observed the distribution of of the sign term, as compared to what the theory would predict.
 Our goal was to capture the leading order behavior of the ET dependence on VPD.
@@
 However, the data largely reproduces our theoretical analysis.
+
+\begin{figure}[!p]
+\centering
+\includegraphics[width=\textwidth]{./test_sign.pdf}
+\caption{Comparison of the sign term with model uncertainty included (box plots) to the sign term as calculated with simplifying assumptions (blue line).}
+\label{test_sign}
+\end{figure}
 
@@
 These deficiencies should largely explain the inability of our theory to exactly match the observations in croplands.
-For example, a superposition of sites with C3 plants (at low environmental VPD) and C4 plants (at high environmental VPD) would explain observed shift back towards negative ET response at high VPD when all sites are binned together, as in Figure \ref{agu_fig}.
+For example, a superposition of sites with C3 plants (at low environmental VPD) and C4 plants (at high environmental VPD) would explain observed shift back towards negative ET response at high VPD when all sites are binned together, as in Figure \ref{test_sign}.
 
@@
 Again, as in Section \ref{sign_term}, the PET sensitivity to VPD will be constant at about 3.5.
-In Figure \ref{agu_fig} the ET response to shifts in dryness is always far smaller than 3.5 in the observations, even with all uncertainty included.
+In Figure \ref{test_sign} the ET response to shifts in dryness is always far smaller than 3.5 in the observations, even with all uncertainty included.
 So the PET response to atmospheric dryness is outside the range of all observable responses.
@@
 
-Figure \ref{real} presents calculated $\frac{\partial \; ET}{\partial \; VPD}$ where, unless otherwise noted, all variables in Equation \ref{d_et} are allowed to vary, including uncertainty ($\sigma$).
+Figure \ref{data_scatter} presents calculated $\frac{\partial \; ET}{\partial \; VPD}$ where, unless otherwise noted, all variables in Equation \ref{d_et} are allowed to vary, including uncertainty ($\sigma$).
 Each column represents a different quantity related to $\frac{\partial \; ET}{\partial \; VPD}$, and each row is a different PFT.
@@
 
-Figure \ref{real} generally confirms expectations from Figure \ref{agu_fig}.
-By comparing different columns of Figure \ref{real} we can evaluate the impact of $\sigma$ and aerodynamic conductance ($g_a$) on the variability of ET response ($\frac{\partial \; ET}{\partial \; VPD}$).
+\begin{figure}[!p]
+\centering
+\centerline{\includegraphics[width=1.4\textwidth]{./data_scatter.png}}
+\caption{Scatter plots of $\frac{\partial \; ET}{\partial \; VPD}$.
+Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; VPD}$, as labeled: Column 1: $\frac{\partial \; ET}{\partial \; VPD}$; Column 2: $\frac{\partial \; ET}{\partial \; VPD}$ normalized by $g_a$; Column 3: $\frac{\partial \; ET}{\partial \; VPD}$ with $\sigma$ held fixed at 1; and Column 4: $\frac{\partial \; ET}{\partial \; VPD}$ normalized by $g_a$ and with $\sigma$ held fixed.
+Please note differences in the colorbar scale.}
+\label{data_scatter}
+\end{figure}
+
+Figure \ref{data_scatter} generally confirms expectations from Figure \ref{test_sign}.
+By comparing different columns of Figure \ref{data_scatter} we can evaluate the impact of $\sigma$ and aerodynamic conductance ($g_a$) on the variability of ET response ($\frac{\partial \; ET}{\partial \; VPD}$).
 Aerodynamic conductance's scaling (included in columns 1 and 3) alters the magnitude considerably, particularly between PFTs.
-$\sigma$ variability (included in columns 1 and 2) adds a lot of additional noise to the signal of $\frac{\partial \; ET}{\partial \; VPD}$, as observed in Figure \ref{agu_fig}.
+$\sigma$ variability (included in columns 1 and 2) adds a lot of additional noise to the signal of $\frac{\partial \; ET}{\partial \; VPD}$, as observed in Figure \ref{test_sign}.
 However, the general story with the noise appears to match the cleaner signal when $\sigma$ is held constant and $VPD_{crit}$ is clearly visible.
@@
 
-In general Figures \ref{agu_fig} and \ref{real} match our expectations.
+In general Figures \ref{test_sign} and \ref{data_scatter} match our expectations.
 Our $\sigma$-based method of uncertainty propagation blurs the idealized expectations the most for GRA and CRO.
 We therefore have the most confidence in our conclusion based on Equation \ref{d_et} for PFTs CSH, DBF, and ENF, as the sign of the derivative with uncertainty included matches (to leading order) the story when $\sigma$ is held fixed.
-
-\begin{figure}[!p]
-\centering
-\includegraphics[width=\textwidth]{./agu_fig.pdf}
-\caption{Comparison of the sign term with model uncertainty included (box plots) to the sign term as calculated with simplifying assumptions (blue line).}
-\label{agu_fig}
-\end{figure}
-
-\begin{figure}[!p]
-\centering
-\centerline{\includegraphics[width=1.4\textwidth]{./fig06.png}}
-\caption{Scatter plots of $\frac{\partial \; ET}{\partial \; VPD}$.
-Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; VPD}$, as labeled: Column 1: $\frac{\partial \; ET}{\partial \; VPD}$; Column 2: $\frac{\partial \; ET}{\partial \; VPD}$ normalized by $g_a$; Column 3: $\frac{\partial \; ET}{\partial \; VPD}$ with $\sigma$ held fixed at 1; and Column 4: $\frac{\partial \; ET}{\partial \; VPD}$ normalized by $g_a$ and with $\sigma$ held fixed.
-Please note differences in the colorbar scale.}
-\label{real}
-\end{figure}
 
@@
 
-Indeed, for all PFTs, there is some slight dependence of $\sigma \cdot uWUE$ on SWC, especially at low SWC (Figure \ref{swc}).
+Indeed, for all PFTs, there is some slight dependence of $\sigma \cdot uWUE$ on SWC, especially at low SWC (Figure \ref{swc_boxplot}).
 The portion of observations for which our theory is biased by SWC-limitations varies by PFT, due to the nonlinear threshold at which soil moisture availability limits plant function.
-For CRO and DBF, soil water content only matters for about the lowest 5\% of observations (each box is 5\% of observations in Figure \ref{swc}).
+For CRO and DBF, soil water content only matters for about the lowest 5\% of observations (each box is 5\% of observations in Figure \ref{swc_boxplot}).
 And, CRO, relative to DBF, has a very weak dependence on soil moisture, which is reflective of the high likelihood that CRO sites are optimally irrigated and not water stressed.
@@
 For all other PFTs occurence of soil mosture impacts was rare enough to not manifest itself in bulk statistics and plots\footnote{note that interpreting the soil moisture dependence for CSH is difficult as their are only two sites for that PFT.
-So systematic soil moisture relationshps are not apparent in Figure \ref{swc}, but instead blurred by the superposition of the two sites.
+So systematic soil moisture relationshps are not apparent in Figure \ref{swc_boxplot}, but instead blurred by the superposition of the two sites.
 We argue taht the soil moisture dependence is still there, but we don't have enough statistcs to comment on it}.
+
+\begin{figure}[!p]
+\centering
+\includegraphics[width=\textwidth]{./swc_boxplot.pdf}
+\caption{$\sigma \cot uWUE$ as a function of PFT and SWC. Each boxplot represents 5\% of all observations.}
+\label{swc_boxplot}
+\end{figure}
 
@@
 Soil moisture effects explain the deficiencies of our theory in Section \ref{testing}, particularly for GRA.
-
-\begin{figure}[!p]
-\centering
-\includegraphics[width=\textwidth]{./swc_boxplot.pdf}
-\caption{$\sigma \cot uWUE$ as a function of PFT and SWC. Each boxplot represents 5\% of all observations.}
-\label{swc}
-\end{figure}
 

### 2018-02-26 doc/paper/vpd_et_paper.tex
fixed changjie's error where z0m=z0h,v, now is correct z0h,v=0.1*z0m
cleaned up some stuff

@@
 \label{data}
-We use both meteorological and eddy-covariance data from the FLUXNET2015 database, including all sites with more than four years of data (data available at https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/).
+We use both meteorological and eddy-covariance data from the FLUXNET2015 database,  all sites with at least four years of data (data available at https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/).
 Each site's plant functional type (PFT) was classified using the International Geosphere-Biosphere Programme vegetation classification scheme \citep{Loveland_1999}.
 The physical constants used in the stomatal conductance model in the Methods section (Section \ref{methods}) are only published for six plant functional types (PFTs): crops (CRO), deciduous broadleaf forest (DBF), evergreed broadleaf forests (EBF), evergreen needleleaf forest (ENF), grass (GRA), and closed shrub (CSH) (see Table \ref{pft}).
-There are 62 sites with these plant functional types, and their location is shown in  Figure \ref{map_fig} \explain{map needs to be improved - it's a placeholder for now}.
+There are 62 sites with these plant functional types, and their location is shown in  Figure \ref{map_fig}.
 
@@
 Timesteps with negative observed GPP or ET are also removed, and we aggregate half hourly data to hourly averages to reduce noise.
-After these quality control procedures, 341788 upscaled hourly observations remain.
+After these quality control procedures, 342461 upscaled hourly observations remain.
 

### 2018-02-26 src/paper_figure_code/data_scatter.py

@@ -30,11 +30,8 @@ def meshgrid_apply(_df, column='var', sample=False):
 
 def make_ax_plot(_ax, var, _df, meta):
   """makes an axis plot"""
-  # vmax = meta['vmax']
-  # vmin = -vmax
-  vmax = None
-  vmin = None
-
+  vmax = meta['vmax']
+  vmin = -vmax
   _df['var'] = var
   _df = _df.assign(x_cut=pd.cut(_df['vpd'], meta['nbins']),\
                    t_a_cut=pd.cut(_df['t_a'], meta['nbins']))
@@ -148,5 +145,9 @@ df.groupby('pft').apply(scatter_plot_paper, meta)
 os.system('convert -append %s/climate_et/paper_plots/scatter/*vpd.png '\
           '../../doc/paper/data_scatter%s.png'\
           % (os.environ['PLOTS'], meta['sample']))
+os.system('cp ../../doc/paper/data_scatter%s.png '\
+          '../../doc/paper/data_scatter%s.bak'
+          % (meta['sample'], meta['sample']))
+
 print('done with vpd, time was %f min' % ((time.time()-start)/60.))
 

### 2018-02-26 src/paper_figure_code/map.py

@@ -24,7 +24,7 @@ def make_map(_df):
     x, y = m(_ds.Longitude.values, _ds.Latitude.values)
     ax.scatter(x, y, s=12, label=pft, alpha=0.7,\
                marker=symbols[i], zorder=99999)
-  ax.set_title('FLUXNET2015 Sites with >4 Years Data')
+  ax.set_title('FLUXNET2015 Sites with $\geq4$ Years Data')
 
 
   plt.legend(loc='best')

### 2018-02-26 src/paper_figure_code/tables.py

@@ -19,7 +19,7 @@ mean_df['d_et_bar'] = d_et(mean_df)
 mean_df['d_et_bar_std'] = np.absolute(mean_df['d_et_bar'])*std['vpd']
 mean_df['d_et_d_rn_bar_std'] = mean_df['delta']/\
                                (mean_df['delta'] + mean_df['gamma'])\
-                               *std['r_n']
+                               *std['r_net']
 mean_df['d_et_bar_norm'] = mean_df['d_et_bar_std']\
                            /std['et_obs']
 mean_df['total_var'] = np.sqrt(mean_df.d_et_bar_std**2\

### 2018-02-27 doc/paper/vpd_et_paper.tex
think all figs and stuff in final shape, starting to write for submission

@@
 \affiliation{1}{Department of Earth and Environmental Engineering, Columbia University, New York, NY 10027}
-\affiliation{2}{Department of Hydraulic Engineering, Tsinghua University, Beijing, CN}
+\affiliation{2}{State Key Laboratory of Hydroscience and Engineering, Department of Hydraulic Engineering, Tsinghua University, Beijing, CN 100084}
 
@@
 \begin{keypoints}
-\item Add given Medlyn (nature response) and Zhou2014 uWUE should increase roughly proportional to atmospheric CO2 concentration,
-  so really the env VPD will be where plants will feel the effect of global warming stronges, b/c effect of CO2 cancels.
-\item = enter point 2 here = 
-\item = enter point 3 here = 
+\item one
+\item two
 \end{keypoints}
@@
 
-Vapor pressure deficit (VPD) is expected to rise over continents in the future due to the combination of increased temperature and relative humidity drying ** RH drying?
-\citep{Byrne_2013}.
+Vapor pressure deficit (VPD) is expected to rise over continents in the future due to the combination of increased temperature and, depending on region, decreased relative humidity \citep{Byrne_2013}.
 Increases in VPD increase the atmospheric demand for evapotranspirated water \citep{Penman_1948, Monteith_1965}, but also stress plant stomata \citep{Leuning_1990, MEDLYN_2011}.

### 2018-02-27 doc/paper/vpd_et_paper.tex
at methods

@@
 
-\authors{A. Massmann\affil{1}, P. Gentine\affil{1}, C. Lin\affil{1,2}}
+\authors{Adam Massmann\affil{1}, Pierre Gentine\affil{1}, Changjie Lin\affil{1,2}}
 
@@
 
-\correspondingauthor{Adam Massmann}{akm2203@columbia.edu}
+\correspondingauthor{Adam Massmann}{206-919-1364; akm2203@columbia.edu}
 
@@
 Increases in VPD increase the atmospheric demand for evapotranspirated water \citep{Penman_1948, Monteith_1965}, but also stress plant stomata \citep{Leuning_1990, MEDLYN_2011}.
-The conflicting effects of increased atmospheric demand and higher stomatal stress lead to two possible perspectives for how evapotraspiration (ET) respondes to shifts in VPD.
+
+The opposing effects of increased atmospheric demand and higher stomatal stress lead to two possible perspectives for how evapotraspiration (ET) respondes to shifts in VPD.
 The first, a hydrometeorological perspective, is that higher VPD increases atmospheric demand for water from the land surface, and this drives an increase in evapotranspiration (ET).
-This perspective is particularly relevant because potential evapotranspiration (PET), which is used in many drought indices and hydrometeorological studies \citep[e.g.,][]{Heim_2002, Scheff_2015}, typically only quantifies changes in atmospheric demand with limited surface regulation though ecosystem resistance regualtion \citep{Swann_2016}.
-On the other hand, there is another perspective for vegetated surfaces.
-Plants' stomata have evolved to optimally regulate the exchange of water and carbon, and tend to partially close in response to increased atmospheric dryness \citep{Ball_1987, Leuning_1990, MEDLYN_2011}.
+This perspective is particularly relevant because potential evapotranspiration (PET), which is used in many drought indices and hydrometeorological studies \citep[e.g.,][]{Heim_2002, Scheff_2015}, typically only quantifies changes in atmospheric demand and fails to account for ecosystem response \citep{Swann_2016}.
+In reality, plants' stomata have evolved to optimally regulate the exchange of water and carbon, and tend to partially close in response to increased atmospheric dryness \citep{Ball_1987, Leuning_1990, MEDLYN_2011}.
 This leads to a plant physiology perspective, in which an increase in VPD, particularly in well-watered soil conditions, may actually correspond to a decrease in ET because of stomatal closure.
@@
 At the ecosystem scale, there are limits to plant water conservation strategies.
-As atmospheric demand for water (VPD) increases further, ecosystems should begin to reach their water conservation limits and might not be able to entirely limit ET flux to the atmosphere.
+As atmospheric demand for water (VPD) increases, ecosystems should begin to reach their water conservation limits and might not be able to entirely limit ET flux to the atmosphere.
 At this stage any further increase in VPD will most likely drive a (limited) increase in ET, because the increase in atmospheric demand for water overwhelms the limited plant response to conserve water.
@@
 This intuition will aid interpretation of observations and full complexity models.
-In order to quantify plant response to perturbations in VPD, we apply a Penman-Monteith framework to derive theoretical response of ET to VPD.
+In order to quantify plant response to perturbations in VPD, we apply a Penman-Monteith framework to derive the theoretical response of ET to VPD.
 The model is then validated and tested at multiple eddy-covariance stations spanning various climates and plant functional types.
-Section 2 describes the data used and derives the framework.
-Section 4 presents results and comparison to observations, and  Section 5 discusses the conclusions.
 
 \section{Materials and Methods}
+\label{methods}
 \subsection{Data}
 \label{data}
-We use both meteorological and eddy-covariance data from the FLUXNET2015 database,  all sites with at least four years of data (data available at https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/).
+We use both meteorological and eddy-covariance data from the FLUXNET2015 database,  all sites with at least four years of data (data available at \sloppy https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/ \sloppy).
 Each site's plant functional type (PFT) was classified using the International Geosphere-Biosphere Programme vegetation classification scheme \citep{Loveland_1999}.
@@
 
-\begin{figure}[!p]
+\begin{figure}
 \centering
@@
 \end{itemize}
-Additionally, we restrict data to the daytime, which is identified when downwelling shortwave radiation is greater than 50 W m$^{-2}$ and sensible heat flux is greater than 5 W m$^{-2}$.
-To further reduce the chance of sensor saturation at high relative humidity, we remove all time steps for which VPD is less than .01 kPa.
-Timesteps with negative observed GPP or ET are also removed, and we aggregate half hourly data to hourly averages to reduce noise.
-After these quality control procedures, 342461 upscaled hourly observations remain.
+Additionally, as in \citet{Lin_2018}, we restrict data to the daytime, which is identified when downwelling shortwave radiation is greater than 50 W m$^{-2}$ and sensible heat flux is greater than 5 W m$^{-2}$.
+To reduce the chance of sensor saturation at high relative humidity, we remove all time steps for which VPD is less than .01 kPa, and to reduce errors at low windspeeds we remove all periods with wind magnitudes less than 0.5 m s$^{-1}$.
+Timesteps with negative observed GPP or ET are also removed, and we aggregate half hourly data to hourly averages to reduce noise \citep{Lin_2018}.
+After these quality control procedures, 335,939 upscaled hourly observations remain.
 
@@
 \label{methods}
-The Penman-Monteith equation \citep [hereafter PM,][]{Monteith_1965} estimates ET as a function of observable atmospheric variables and surface conductances:
+The Penman-Monteith equation \citep [hereafter PM,][]{Penman_1948, Monteith_1965} estimates ET as a function of observable atmospheric variables and surface conductances:
 \begin{linenomath*}
@@
 where $\Delta$ is the change in saturation vapor pressure with temperature, given by Clausius-Clapeyron ($\frac{d \; e_s}{d \; T}$), $R$ is the net radiation minus ground heat flux, $g_a$ is aerodynamic conductance, $\rho_a$ is air density, $c_p$ is specific heat of air at constant pressure, $\gamma$ is the psychometric constant, and $g_s$ is the stomatal conductance (Table 1).
+
+\begin{table}
+\caption{Definition of symbols and variables, with citation for calculation if applicable.}
+\centering
+\small
+\begin{tabular}{l c c c}
+\hline
+ Variable & Description & Units & Citation \\
+\hline
+$e_s$  & saturation vapor pressure & Pa  & - \\ 
+  $T$  & temperature  & K & - \\
+  $P$  & pressure & Pa  & - \\
+$\Delta$  & $\frac{\partial e_s}{\partial T}$ & Pa K$^{-1}$ & - \\
+$R$  & net radiation at land surface minus ground heat flux & W m$^{-2}$   & - \\
+  $g_a$  & aerodynamic conductance & m s$^{-1}$  & \makecell{\cite{Thom_1977} \\ \cite{Paulson_1970} \\ \cite{Beljaars_1991}} \\
+  $\rho_a$  & air density & kg m$^{-3}$  & - \\
+  $c_p$  & specific heat capacity of air at constant pressure & J K$^{-1}$ kg$^{-1}$ & - \\
+  $VPD$  & vapor pressure deficit & Pa  & - \\
+  $\gamma$  & psychometric constant & Pa K$^{-1}$   & - \\
+  $g_s$  & stomatal conductance & m s$^{-1}$  & \cite{MEDLYN_2011} \\
+  $g_{l-s}$  & leaf-scale stomatal conductance & mol m$^{-2}$ s$^{-1}$  & \cite{MEDLYN_2011} \\
+  $R*$ & universal gas constant & J mol$^{-1}$ K$^{-1}$ & - \\
+  $R_{air}$ & gas constant of air & J  K$^{-1}$ kg$^{-1}$ & - \\
+  $LAI$ & leaf area index & -& - \\
+  $\sigma$ & uncertainty parameter & -& - \\
+  $c_a$ & CO$_2$ concentration & $\mu$ mol CO$_2$ mol$^{-1}$ air& - \\
+  $\lambda$ & marginal water cost of leaf carbon & mol H$_2$O mol$^{-1}$ CO$_2$ & - \\
+  $\Gamma$ & CO$_2$ compensation point & - & - \\
+  $\Gamma^*$ & CO$_2$ compensation point without dark respiration & - & - \\
+
+\hline
+\end{tabular}
+\end{table}
+
 \citet{MEDLYN_2011} developed a model for stomatal conductance ($g_s$) by combining an optimal photosynthesis theory \citep{Farquhar_1980, Katul_2010} with an empirical approach, which describes the dependence of $g_s$ to VPD.
-This resulted in the following model for leaf-scale stomatal conductance was:
+This resulted in the following model for leaf-scale stomatal conductance:
 
@@
   \begin{equation}
-  g_{l-s} = g_0 + 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{A}{c_s}
+  g_{l-s} = g_0 + 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{A}{c_a}
   \end{equation}
 \end{linenomath*}
-Where $g_1$ is a leaf-scale ``slope'' parameter, A is the net CO$_2$ assimilation rate, and $c_s$ is the atmospheric CO$_2$ concentration at the leaf surface.
+Where $g_1$ is a leaf-scale ``slope'' parameter, A is the net CO$_2$ assimilation rate, and $c_a$ is the atmospheric CO$_2$ concentration at the leaf surface.
 \cite{MEDLYN_2011} relate the slope parameter ($g_1$) to physical parameters as:
@@
   \begin{equation}
-  g_1 = \sqrt{\frac{3 \, \Gamma^* \, \lambda}{1.6}},\footnote{Note the above expression has units of of (mmol mol$^{-1}$)$^{1/2}$, but this can be converted to Pa$^{1/2}$ using the ideal gas law.}
+  g_1 = \sqrt{\frac{3 \, \Gamma^* \, \lambda}{1.6}},\footnote{Note this expression has units of of (mmol mol$^{-1}$)$^{1/2}$, but this can be converted to Pa$^{1/2}$ using the ideal gas law.}
   \end{equation}
@@
 
-where $\Gamma^*$ is the CO$_2$ compensation point for photosynthesis (without dark respiration), and $\lambda$ is the marginal water cost of leaf carbon.
+where $\Gamma^*$ is the CO$_2$ compensation point for photosynthesis (without dark respiration), and $\lambda$ is the marginal water cost of leaf carbon ($\frac{\partial \; \text{transpiration}}{\partial \; A}$).
 So, $g_1$ is a leaf-scale term reflecting the tradeoff of water for carbon uptake and growth.
@@
 
-The Medlyn model for stomatal conductance has been shown to behave very well across PFTs, compared to other models from the leaf to ecosystem scale \citep[][** still unsure about citing Zhou here]{Lin_2015, Zhou_2014}.
-It is adapted at the ecosystem scale by multiplying $g_{l-s}$ by leaf area index (LAI) and converting units to m s$^{-1}$:
+The Medlyn model for stomatal conductance has been shown to behave very well across PFTs \citep[][]{Lin_2015}, and can be  adapted to the ecosystem scale by multiplying $g_{l-s}$ by leaf area index (LAI) and converting units to m s$^{-1}$ with the ideal gas law:
 
 \begin{linenomath*}
+  \begin{equation}
+    g_s = \text{LAI} \frac{R \,T}{P} \left( g_0 + 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{A}{c_a}\right)
       \label{medlyn}
-  \begin{equation}
-  g_s = \text{LAI} \frac{R \,T}{P} \left( g_0 + 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{A}{c_s}\right)
   \end{equation}
@@
 \end{linenomath*}
-is relatively constant across time and moisture conditions within a plant functional type (CITE ALSO CHAGJIE ** ??), in non-extreme drought conditions.
+is relatively constant across time and moisture conditions within a plant functional type, in non-extreme drought conditions.
 If, following \citet{Lin_2015}, we approximate $g_0$ as $0$ (i.e.
@@
   \begin{equation}
-    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{uWUE \; ET}{c_s \; \sqrt{VPD}}
+    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{uWUE \; ET}{c_a \; \sqrt{VPD}}
     \label{new_g_s}
@@
 
-Note that $uWUE$ is fit on the ecosystem scale as in \citet{Zhou_2015} so GPP in Equation \ref{uwue} is really $A\cdot \text{LAI}$.
-This leads to the cancellation of LAI with the addition of uWUE in Equation \ref{medlyn}.
-Plugging Equation \ref{new_g_s} into Equation \ref{orig_pen} and rearranging gives a new expression for PM ET directly as a function of VPD with no $A$ dependence:
+Note that $uWUE$ is fit on the ecosystem scale as in \citet{Zhou_2015}, so GPP in Equation \ref{uwue} is really $A \cdot \text{LAI}$.
+This leads to the cancellation of LAI in the step from Equation \ref{medlyn} to Equation \ref{new_g_s}.
+Plugging Equation \ref{new_g_s} into Equation \ref{orig_pen} and rearranging gives a new explicit expression for PM, in which dependence on $A$ is removed:
 
@@
   \begin{equation}
-    ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_s \sqrt{VPD} }{ R* \; 1.6\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
+    ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R* \; 1.6\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
     \label{et}
@@
 
-\[uWUE \appropto \sqrt{\frac{c_s - \Gamma}{1.6 \lambda}},\]
+\[uWUE \appropto \sqrt{\frac{c_a - \Gamma}{1.6 \lambda}},\]
 
@@
   \begin{equation}
-    ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_s \sqrt{VPD} }{ R* \; 1.6\; \sigma \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
+    ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R* \; 1.6\; \sigma \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
     \label{et_sigma}
@@
   \begin{equation}
-\sigma = - \frac{g_a \gamma c_s \sqrt{VPD} L_v P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p VPD\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})},
+\sigma = - \frac{g_a \gamma c_a \sqrt{VPD} L_v P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p VPD\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})},
     \label{sigma}
@@
   \begin{equation}
-    \frac{\partial \;  ET}{\partial \; VPD} = \frac{2\; g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_s }{1.6 \; R*\; \sigma \; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)
+    \frac{\partial \;  ET}{\partial \; VPD} = \frac{2\; g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_a }{1.6 \; R*\; \sigma \; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)
     \label{d_et}
@@
 We note one final comment on our derivation which is relevant for drought indices.
-If we approximate $c_s$ at a global mean CO$_2$ concentration, then the RHS of Equation \ref{et} is fully defined using commonly available weather station data and the constants published in \citet{Zhou_2015} and \citet{Lin_2015}.
+If we approximate $c_a$ at a global mean CO$_2$ concentration, then the RHS of Equation \ref{et} is fully defined using commonly available weather station data and the constants published in \citet{Zhou_2015} and \citet{Lin_2015}.
 This makes Equation \ref{et} a useful alternative to PET in drought indices and hydrometeorological analysis for vegetated surfaces.
 Equation \ref{et} better reflects the physics of water exchange at the land surface and would only require fitting of $uWUE$ and $g_1$, and can also account for changes in CO$_2$ concentration, contrary to typical drought indices such as the Palmer Drought Severity Index (PDSI) \citep{Swann_2016}.
-
-\begin{table}
-\caption{Definition of symbols and variables, with citation for calculation if applicable.}
-\centering
-\begin{tabular}{l c c c}
-\hline
- Variable & Description & Units & Citation \\
-\hline
-$e_s$  & saturation vapor pressure & Pa  & - \\ 
-$T$  & temperature  & K & - \\
-$\Delta$  & $\frac{\partial e_s}{\partial T}$ & Pa K$^{-1}$ & - \\
-$R$  & net radiation at land surface minus ground heat flux & W m$^{-2}$   & - \\
-  $g_a$  & aerodynamic conductance & m s$^{-1}$  & \makecell{\cite{Thom_1977} \\ \cite{Paulson_1970} \\ \cite{Beljaars_1991}} \\
-  $\rho_a$  & air density & kg m$^{-3}$  & - \\
-  $c_p$  & specific heat capacity of air at constant pressure & J K$^{-1}$ kg$^{-1}$ & - \\
-  $VPD$  & vapor pressure deficit & Pa  & - \\
-  $\gamma$  & psychometric constant & Pa K$^{-1}$   & - \\
-  $g_s$  & stomatal conductance & m s$^{-1}$  & \cite{MEDLYN_2011} \\
-  $g_{l-s}$  & leaf-scale stomatal conductance & mol m$^{-2}$ s$^{-1}$  & \cite{MEDLYN_2011} \\
-  $R*$ & universal gas constant & J mol$^{-1}$ K$^{-1}$ & - \\
-  $LAI$ & leaf area index & -& - \\
-  $\sigma$ & uncertainty parameter & -& - \\
-  $c_s$ & surface CO$_2$ concentration & $\mu$ mol CO$_2$ mol$^{-1}$ air& - \\
-\hline
-\multicolumn{2}{l}{$^{a}$Footnote text here.}
-\end{tabular}
-\end{table}
 
@@
   \label{sign}
-  \frac{c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right).
+  \frac{c_p}{R_{air}} - \frac{\gamma c_a }{1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right).
 \end{equation}
@@
 First, we explore the variables within the sign term to gain better intuition on the driver of either the increase or reduction of ET with VPD.
-CO$_2$ concentration ($c_s$) and the psychometric constant ($\gamma$) are relatively constant over the dataset considered here so that the variability is dominated by $\sigma$ and $VPD$.
+CO$_2$ concentration ($c_a$) and the psychometric constant ($\gamma$) are relatively constant over the dataset considered here so that the variability is dominated by $\sigma$ and $VPD$.
 $uWUE$ could vary with soil moisture but has been shown to be relatively constant except in very dry soil conditions  (ZHOU ET AL, LIN ET AL ** ??)\explain{Pierre, which Zhou and Lin citations were you referring to that show that uWUE changes with very dry soil conditions?}.
@@
   \begin{equation}
-VPD_{crit} = \frac{R_{air}}{4 c_p} \left( \frac{\gamma c_s}{1.6\; R \; \overline{\sigma} uWUE} + \sqrt{\frac{\gamma c_s}{1.6\; R \; \overline{\sigma} uWUE}\left( \frac{\gamma c_s}{1.6\; R \; \overline{\sigma} uWUE} + 8 g_1 \frac{c_p}{R_{air}}\right)} - 4 g_1 \frac{c_p}{R_{air}} \right)
+VPD_{crit} = \frac{R_{air}}{4 c_p} \left( \frac{\gamma c_a}{1.6\; R \; \overline{\sigma} uWUE} + \sqrt{\frac{\gamma c_a}{1.6\; R \; \overline{\sigma} uWUE}\left( \frac{\gamma c_a}{1.6\; R \; \overline{\sigma} uWUE} + 8 g_1 \frac{c_p}{R_{air}}\right)} - 4 g_1 \frac{c_p}{R_{air}} \right)
 \label{vpd_min_et}
@@
 \begin{table}  
-\caption{Values of $VPD_{crit}$, where $\frac{\partial \; ET}{\partial \; VPD} = 0$, evaluated at PFT average values for $R_{air}$, $\sigma$, $\gamma$, and $c_s$.
+\caption{Values of $VPD_{crit}$, where $\frac{\partial \; ET}{\partial \; VPD} = 0$, evaluated at PFT average values for $R_{air}$, $\sigma$, $\gamma$, and $c_a$.
 For reference, these values are also provided CITE TABLE.
@@
   \hline
-  PFT & $R_{air}$ & $c_s$ (ppm) & $\gamma$ &  uWUE    & \textbf{$VPD_{crit}$ (Pa)} \\
+  PFT & $R_{air}$ & $c_a$ (ppm) & $\gamma$ &  uWUE    & \textbf{$VPD_{crit}$ (Pa)} \\
   \hline
@@
 However, ET response evaluated at the average of all variables (e.g.
-$\sigma$, $T$, $c_s$, $VPD$) is only negative for CSH and GRA.
+$\sigma$, $T$, $c_a$, $VPD$) is only negative for CSH and GRA.
 This difference in mean ET response as compared to the ET response at mean environmental conditions is due to the non-linear nature of the response, in which negative responses are generally much larger magnitude than positive responses (Figure \ref{sign_plot}).
@@
 
-% \begin{figure}[!p]
+% \begin{figure}
 % \centering

### 2018-02-27 doc/paper/vpd_et_paper.tex
finished methods section

@@
 \label{data}
-We use both meteorological and eddy-covariance data from the FLUXNET2015 database,  all sites with at least four years of data (data available at \sloppy https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/ \sloppy).
+We use both meteorological and eddy-covariance data from the FLUXNET2015 database, including all sites with at least four years of data (data available at \sloppy https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/ \sloppy).
 Each site's plant functional type (PFT) was classified using the International Geosphere-Biosphere Programme vegetation classification scheme \citep{Loveland_1999}.
@@
 $R$  & net radiation at land surface minus ground heat flux & W m$^{-2}$   & - \\
-  $g_a$  & aerodynamic conductance & m s$^{-1}$  & \makecell{\cite{Thom_1977} \\ \cite{Paulson_1970} \\ \cite{Beljaars_1991}} \\
+  $g_a$  & aerodynamic conductance & m s$^{-1}$  & ** Shuttleworth \\
   $\rho_a$  & air density & kg m$^{-3}$  & - \\
@@
   $g_{l-s}$  & leaf-scale stomatal conductance & mol m$^{-2}$ s$^{-1}$  & \cite{MEDLYN_2011} \\
-  $R*$ & universal gas constant & J mol$^{-1}$ K$^{-1}$ & - \\
+  $R$ & universal gas constant & J mol$^{-1}$ K$^{-1}$ & - \\
   $R_{air}$ & gas constant of air & J  K$^{-1}$ kg$^{-1}$ & - \\
@@
 
-While Equation 3 can be used in PM (Equation 1), it will make analytical work with the function intractable because $A$, net CO$_2$ assimilation, is functionally related to ET itself.
+While Equation 3 can be used in PM (Equation 1), it will make analytical work with the function intractable because $A$, net CO$_2$ assimilation rate, is functionally related to ET itself.
 To remove the dependence of ET on $A$ we use fundamental semi-empirical results of \citet{Zhou_2015}, who showed that the underlying Water Use Efficiency ($uWUE$):
@@
   \begin{equation}
-    ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R* \; 1.6\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
+    ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R \; 1.6\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
     \label{et}
@@
 The slope parameter is a leaf-scale term related to the willingness of stomata to trade water for CO$_2$ and to keep stomata open.
-$uWUE$ is semi-empirical ecosystem-scale constant related to how WUE changes with VPD (specifically $VPD^{-1/2}$).
+$uWUE$ is a semi-empirical ecosystem-scale constant related to how WUE changes with VPD (specifically $VPD^{-1/2}$).
 Is is also roughly proportional to physical constants:
@@
 
-where $\Gamma$ is the CO$_2$ compensation point \citep{Zhou_2014}.
+where $\Gamma$ is the CO$_2$ compensation point \citep[Equation 5 in][]{Zhou_2014}.
 So uWUE is related to atmospheric CO$_2$ concentration and compensation point, and is inversely proportional to the marginal water cost of leaf carbon.
-Therefore, generally speaking, plants that are very willing to trade water for new growth will have lower $uWUE$ and higher $g_1$, and will be less water conservative (more likely to increase ET with increasing VPD).
 
-Given the eddy-covariance FLUXNET2015 data described in Section \ref{data}, every term in Equation \ref{et} is observed, except for the two parameters $g_1$, and $uWUE$, which have however been measured for the PFTs considered here \citep{Zhou_2015, Lin_2015}.
+Given the eddy-covariance FLUXNET2015 data described in Section \ref{data}, every term in our new version of PM (Equation \ref{et}) is observed, except for the two parameters $g_1$, and $uWUE$.
+$g_1$ has been measured and reported for the PFTs considered here \citep{Lin_2015, Franks_2017}, and we fit $uWUE$ by calculating its expectation, given the model and FLUXNET2015 data.
+
 However, eddy-covariance data are inherently noisy so we include a measure of uncertainty in our analysis.
-To account for both mean deviations of $uWUE$ and uncertainty, we introduce an uncertainty parameter $\sigma$ modifying $uWUE$:
+To account for observational error, as well as model uncertainty (e.g.
+temporal and spatial variations of $uWUE$), we introduce an uncertainty parameter $\sigma$ modifying $uWUE$:
 
@@
   \begin{equation}
-    ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R* \; 1.6\; \sigma \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
+    ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R \; 1.6\; \sigma \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
     \label{et_sigma}
@@
 
-Now, from each FLUXNET2015 observation (i.e. for each semi-hourly data, at every time step) we can evaluate $\sigma$:
+Now, from each FLUXNET2015 observation (i.e. for each hourly observation at every time step) we can evaluate $\sigma$:
 
@@
 The variability of $\sigma$ across sites and time provides a measure of uncertainty in our model, assumptions, as well as in the FLUXNET201 observations themselves.
-To correct for differences in sampling between \cite{Zhou_2015} and our data, we set uWUE such that $\overline{\sigma} = 1$.
 The variability of $\sigma$ then propagates through any uncertainty to our derivative of Equation \ref{et_sigma}:
@@
   \begin{equation}
-    \frac{\partial \;  ET}{\partial \; VPD} = \frac{2\; g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_a }{1.6 \; R*\; \sigma \; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)
+    \frac{\partial \;  ET}{\partial \; VPD} = \frac{2\; g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_a }{1.6 \; R\; \sigma \; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)
     \label{d_et}
@@
 
-With Equation \ref{d_et} we have an analytical framework for ecosystem response to atmospheric demand perturbations, in non extreme drought conditions.
+With Equation \ref{d_et} we have an analytical framework for ecosystem response to atmospheric demand perturbations, with an implicit assumption of non-extreme drought conditions.
 There are a few subtleties to taking the derivative in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$) and $VPD$ are functionally related, so while taking the derivative we evaluate $\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET} {\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD} \Big|_{\text{RH fixed}} + \frac{\partial \; ET}{\partial \; RH} \frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$ fixed}}$.

### 2018-02-27 doc/paper/vpd_et_paper.tex
finished sign theory section, good progress

@@
 \usepackage{makecell}
+\usepackage{multirow} %AM for tables
+\usepackage{booktabs}
 \drafttrue
@@
 
+\begin{table}
+  \caption{Plant functional types, their abbreviation, Medlyn coefficient (from \citet{Lin_2015}, via \citet{Franks_2017}) and caluclated uWUE.
+uWUE from \citet{Zhou_2015}, including the  observed standard deviation is shown for comparison.
+Note that uWUE from \citet{Zhou_2015} is calculated from a different set of sites, and that units are converted such that the quantities fit into Equations 1-8 with the variables in Table \ref{definitions}.}
+    \small
+\label{pft}
+\centering
+\begin{tabular}{l c c @{\qquad} c c}
+  \hline
+  \multirow{2}[3]{*}{Abbreviation} & \multirow{2}[3]{*}{PFT} & \multirow{2}[3]{*}{$g_1$ (Pa$^{0.5}$)} & \multicolumn{2}{c}{uWUE ($\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET])}  \\
+  \cmidrule{4-5}
+
+  & & & fitted & \citet{Zhou_2015} \\
+
+  \hline
+  DBF & deciduous broadleaf forest & 140.7 & 2.63 & 3.12 $\pm$ 0.52 \\
+  EBF & evergreen broadleaf forest & 130.3 & 2.56 &  N/A \\
+  ENF & evergreen needleleaf forest & 74.3 & 3.45 & 3.30 $\pm$ 0.91 \\
+  CRO & cropland & 183.1 & 2.27 & 3.80 $\pm$ 1.01 \\
+  CSH & closed shrub & 148.6 & 1.66 & 2.18 $\pm$ 0.44 \\
+  GRA & grassland (assumed C3) & 166.0 & 1.92 & 2.68 $\pm$ 0.61 \\
+\hline
+\end{tabular}
+\end{table}
+
 \begin{figure}
@@
     \label{orig_pen}
-    ET = \frac{\Delta R + g_a \rho_a c_p VPD}{\Delta + \gamma(1 + \frac{g_a}{g_s})},
+    ET = \frac{\Delta R_{net} + g_a \rho_a c_p VPD}{\Delta + \gamma(1 + \frac{g_a}{g_s})},
   \end{equation}
 \end{linenomath*}
-where $\Delta$ is the change in saturation vapor pressure with temperature, given by Clausius-Clapeyron ($\frac{d \; e_s}{d \; T}$), $R$ is the net radiation minus ground heat flux, $g_a$ is aerodynamic conductance, $\rho_a$ is air density, $c_p$ is specific heat of air at constant pressure, $\gamma$ is the psychometric constant, and $g_s$ is the stomatal conductance (Table 1).
+where $\Delta$ is the change in saturation vapor pressure with temperature, given by Clausius-Clapeyron ($\frac{d \; e_s}{d \; T}$), $R$ is the net radiation minus ground heat flux, $g_a$ is aerodynamic conductance, $\rho_a$ is air density, $c_p$ is specific heat of air at constant pressure, $\gamma$ is the psychometric constant, and $g_s$ is the stomatal conductance (Table \ref{definitions}).
 
 \begin{table}
+
   \caption{Definition of symbols and variables, with citation for calculation if applicable.}
+  \label{definitions}
 \centering
@@
 $\Delta$  & $\frac{\partial e_s}{\partial T}$ & Pa K$^{-1}$ & - \\
-$R$  & net radiation at land surface minus ground heat flux & W m$^{-2}$   & - \\
-  $g_a$  & aerodynamic conductance & m s$^{-1}$  & ** Shuttleworth \\
+$R_{net}$  & net radiation at land surface minus ground heat flux & W m$^{-2}$   & - \\
+  $g_a$  & aerodynamic conductance & m s$^{-1}$  & \citet{Shuttleworth_2012} \\
   $\rho_a$  & air density & kg m$^{-3}$  & - \\
@@
   $\gamma$  & psychometric constant & Pa K$^{-1}$   & - \\
-  $g_s$  & stomatal conductance & m s$^{-1}$  & \cite{MEDLYN_2011} \\
-  $g_{l-s}$  & leaf-scale stomatal conductance & mol m$^{-2}$ s$^{-1}$  & \cite{MEDLYN_2011} \\
+  $g_s$  & stomatal conductance & m s$^{-1}$  & \citet{MEDLYN_2011} \\
+  $g_{l-s}$  & leaf-scale stomatal conductance & mol m$^{-2}$ s$^{-1}$  & \citet{MEDLYN_2011} \\
   $R$ & universal gas constant & J mol$^{-1}$ K$^{-1}$ & - \\
@@
   $\Gamma^*$ & CO$_2$ compensation point without dark respiration & - & - \\
-
 \hline
@@
   \begin{equation}
-    ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R \; 1.6\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
+    ET = \frac{\Delta R_{net} + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R \; 1.6\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
     \label{et}
@@
   \begin{equation}
-    ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R \; 1.6\; \sigma \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
+    ET = \frac{\Delta R_{net} + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R \; 1.6\; \sigma \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
     \label{et_sigma}
@@
   \begin{equation}
-\sigma = - \frac{g_a \gamma c_a \sqrt{VPD} L_v P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p VPD\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})},
+\sigma = - \frac{g_a \gamma c_a \sqrt{VPD} L_v P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R_{net} - g_a \rho_a c_p VPD\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})},
     \label{sigma}
@@
 
-\begin{table}
-  \label{pft}
-\caption{Plant functional types, their abbreviation, Medlyn coefficient \citep[from ][]{Lin_2015}, and uWUE \citep[from ][]{Zhou_2015}.
-Note that units are converted such that the quantities fit into Equations 1-8 with the variables in Table 1.}
-\centering
-\begin{tabular}{l c c c}
-  \hline
-  Abbreviation & PFT & $g_1$ (Pa$^{0.5}$) & uWUE ($\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET])  \\
-  \hline
-  CRO & cropland & 183.1 & 3.80 \\
-  CSH & closed shrub & 148.6 & 2.18 \\
-  DBF & deciduous broadleaf forest & 140.7 & 3.12 \\
-  ENF & evergreen needleleaf forest & 74.3 & 3.30 \\
-  GRA & grassland (C3) & 166.0 & 2.68 \\
-\hline
-\multicolumn{2}{l}{$^{a}$Footnote text here.}
-\end{tabular}
-\end{table}
-
 \section{Results}
@@
 Therefore, for our assumptions and framework to be reasonable $\sigma$ should be close to 1.
-Figure \ref{joint_vpd_sigma} presents the histogram of calculated $\sigma$s with outliers removed (lowest and highest 5\% percent).
-All remaining $\sigma$ values are close to unity (i.e.
-$O(1)$), which provides further confidence in the model framework and in the data quality.
+An additional concern is that $\sigma$ may in fact be correlated with $VPD$, in which case the dependence would need to be accounted for when taking the derivative.
+Fortunately, there is a very weak dependence of $\sigma$ on VPD in their joint distrubution, and $\sigma$ is indeed close to unity i.e.
+$O(1)$ (Figure \ref{joint_vpd_sigma}).
+Given this weak dependence and the distribution of $\sigma$ we have confidence in our model framework and the data quality.
 
@@
 \includegraphics[width=\textwidth]{./joint_vpd_sigma.pdf}
-\caption{The joint distribution of $VPD$ and $\sigma$. $\sigma$ has only a weak dependence on $VPD$.}
+\caption{The joint distribution of $VPD$ and $\sigma$, with outliers removed (defined as lowest and highest 5\% of $\sigma$).
+$\sigma$ has only a weak dependence on $VPD$, and $\sigma$ is $O(1)$ for the bulk of thte observations.}
 \label{joint_vpd_sigma}
 \end{figure}
-
-An additional concern is that $\sigma$ may in fact be correlated with $VPD$, in which case the dependence would need to be accounted for when taking the derivative.
-However, the joint distribution of $\sigma$ and VPD shows that $\sigma$ is very weakly a function of VPD (Figure \ref{joint_vpd_sigma}).
-Given this weak dependence, we argue that Equation \ref{d_et} is a valid approximation for ET response to $VPD$.
 
@@
 \end{equation}
-All variables are positive, so the relative magnitude between the first term and the second term in Equation \ref{sign} will determine whether ET increases or decreases with increasing VPD.
+All variables are positive, so the relative magnitude between the first term and the second term in the sign term (Equation \ref{sign}) will determine whether ET increases or decreases with increasing VPD.
 If the second term is larger then the plant control dominates and ET decreases with increasing VPD.
@@
 CO$_2$ concentration ($c_a$) and the psychometric constant ($\gamma$) are relatively constant over the dataset considered here so that the variability is dominated by $\sigma$ and $VPD$.
-$uWUE$ could vary with soil moisture but has been shown to be relatively constant except in very dry soil conditions  (ZHOU ET AL, LIN ET AL ** ??)\explain{Pierre, which Zhou and Lin citations were you referring to that show that uWUE changes with very dry soil conditions?}.
-We thus further assume that the soil is not too dry so that $uWUE$ can be approximated as a constant (within PFT).
+$uWUE$ could vary with soil moisture but has been shown to be relatively constant \citep{Zhou_2015}.
 This then means that the sign term only depends on VPD for a given PFT and is approximately just a function of $VPD$.
@@
   \begin{equation}
-VPD_{crit} = \frac{R_{air}}{4 c_p} \left( \frac{\gamma c_a}{1.6\; R \; \overline{\sigma} uWUE} + \sqrt{\frac{\gamma c_a}{1.6\; R \; \overline{\sigma} uWUE}\left( \frac{\gamma c_a}{1.6\; R \; \overline{\sigma} uWUE} + 8 g_1 \frac{c_p}{R_{air}}\right)} - 4 g_1 \frac{c_p}{R_{air}} \right)
+VPD_{crit} = \frac{R_{air}}{4 c_p} \left( \frac{\gamma c_a}{1.6\; R \;  uWUE} + \sqrt{\frac{\gamma c_a}{1.6\; R \;  uWUE}\left( \frac{\gamma c_a}{1.6\; R \;  uWUE} + 8 g_1 \frac{c_p}{R_{air}}\right)} - 4 g_1 \frac{c_p}{R_{air}} \right)
 \label{vpd_min_et}
@@
 \begin{table}  
-\caption{Values of $VPD_{crit}$, where $\frac{\partial \; ET}{\partial \; VPD} = 0$, evaluated at PFT average values for $R_{air}$, $\sigma$, $\gamma$, and $c_a$.
-For reference, these values are also provided CITE TABLE.
-For values of $VPD$ less than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be negative, and for values of $VPD$ greater than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be positive.
-\textbf{**** this statistics are dated, need to update***}}
+\caption{Values of $VPD_{crit}$, where $\frac{\partial \; ET}{\partial \; VPD} = 0$, evaluated at PFT average values for $R_{air}$, $\gamma$, and $c_a$.
+PFT-specific constants ($g_1$, uWUE) are provided in Table \ref{pft}.
+For values of $VPD$ less than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be negative, and for values of $VPD$ greater than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be positive.}
 \centering
-\begin{tabular}{l c c c c c c}
+\begin{tabular}{l c c c c c}
   \hline
-  PFT & $R_{air}$ & $c_a$ (ppm) & $\gamma$ &  uWUE    & \textbf{$VPD_{crit}$ (Pa)} \\
+  PFT & $R_{air}$ & $c_a$ (ppm) & $\gamma$  & \textbf{$VPD_{crit}$ (Pa)} \\
   \hline
-  CRO &  288.680920 & 372.567691& 65.351523& 2.602873&  \textbf{133.165438} \\
-  CSH &   289.067152& 381.593622& 67.613172& 2.175278& \textbf{4439.564212} \\
-  DBF &   288.624437& 377.449849& 63.421812& 2.746393&  \textbf{888.773243} \\
-  ENF &  288.183849& 377.676463& 61.559242& 4.015362&  \textbf{978.084845} \\
-  GRA &  288.425651& 377.264645& 61.598768& 2.281074& \textbf{1141.630778} \\
+CRO & 288.6 & 376.0 & 65.2 &   \textbf{936.6} \\
+CSH & 289.0 & 383.6 & 67.5 & \textbf{12852.5} \\
+DBF & 288.6 & 379.5 & 63.6 &  \textbf{1266.6} \\
+EBF & 288.4 & 368.5 & 61.6 &  \textbf{1454.8} \\
+ENF & 288.28 & 379.8 & 61.0 &  \textbf{1853.7} \\
+GRA & 288.4 & 378.0 & 61.0 &  \textbf{3261.1} \\
 \hline
-\multicolumn{2}{l}{$^{a}$Footnote text here.}
 \end{tabular}
@@
 
-This effect of $uWUE$ and $g1$ on the sign term is most apparent by comparing two extreme PFTs: water intensive CRO and water conservative CSH.
-CRO has a high slope parameter ($g_1 = 183.1 \, Pa^{1/2}$) compared to CSH ($g1 = 148.6 \, Pa^{1/2}$.
-It also has a slightly higher uWUE ($2.32$ -units-) compared to CSH ($uWUE=2.11$ -units-).
-\explain{MAYBE A DUMB QUESTION BUT CAN'T uWUE AND G1 BE ANALYTICALLY RELTED> I THINK AT LESAT YOU SHOULD MAKE CLEAR THAT UWUE IS AT THE ECOYSTEM SCALE (DID YOU FIT IT OR USED PUBLISHED DATA-I AM NOW CONFUSED)}.
-These differences in slope parameter cause opposite ET responses to changes in VPD between CRO and CSH.
-ET theoretically always increases with increasing VPD for the more water intensive CRO, while ET almost always decreases with increasing VPD for the more water conservative CSH (Figure \ref{idealized_sign}).
+This effect of $uWUE$ and $g1$ on the sign term is most apparent by comparing two extreme PFTs: water intensive crops (CRO) and water conservative closed shrub (CSH).
+CRO has high slope parameters and uWUE ($g_1 = 183.1$ Pa$^{1/2}$; $2.27$ $\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET]) compared to CSH ($g1 = 148.6 \, Pa^{1/2}$, $uWUE=1.66$ $\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET]).
+These differences in PFT parameters cause opposite ET responses to changes in VPD between CRO and CSH.
+ET theoretically always decreases with increasing VPD for the more water conservative CSH, while ET will frequently  increases with increasing VPD for the more water intensive CRO (Figure \ref{idealized_sign}).
 CRO evolved or were bred to prioritize GPP and yield and are thus not water conservative.
@@
 Therefore, plants with a low $g_1$ will have the largest VPD dependence of ET response because the difference in ET response at low VPD (leaf stomatal response dominates) and high VPD (VPD gradient dominates) is largest.
-This is apparent in the strong VPD dependence of ENF, which has the lowest slope parameter ($g_1=74.31$), in Figure \ref{idealized_sign}.
+This is apparent in the strong VPD dependence of ENF, which has the lowest slope parameter ($g_1=74.31$ Pa$^{1/2}$) (Figure \ref{idealized_sign}).
 
-To summarize our theoretical insights, Figure \ref{idealized_sign} and Table \ref{vpd_crit}, CROs are the least water conservative and have the strongest overall tendency to increase ET with increasing VPD, while CSH are the most water conservative and have the strongest tendency to decrease ET with increasing VPD.
-ENF ($g_1 = 74.31$) has by far the largest VPD response dependence, while CRO ($g_1 = 183.1$) has the smallest VPD dependence.
-ENF are less willing to trade water for GPP, so the difference between the response at low VPD when leaf response dominates and high VPD when vapor pressure gradients dominate is much greater.
-CROs on the other hand have a more aggressive water use strategy and are more willing to trade water for photosynthesis, so the leaf scale response is much smaller, and the difference in ET response between low VPD and high VPD is relatively small.
-
+To summarize our theoretical insights (Figure \ref{idealized_sign} and Table \ref{vpd_crit}), CROs are the least water conservative and have the strongest overall tendency to increase ET with increasing VPD, while CSH are the most water conservative and have the strongest tendency to decrease ET with increasing VPD.
+ENF ($g_1 = 74.31$ Pa$^{1/2}$) has by far the largest VPD response dependence, while CRO ($g_1 = 183.1$ Pa$^{1/2}$) has the smallest VPD dependence.
 Figure \ref{idealized_sign} clearly shows, according to our theory, that for all PFTs except for crops there is frequent occurrence of a negative (plant dominating) ET response to increases in VPD.
@@
 To better illustrate this, the ranges of observed environmental VPDs at the FLUXNET sites are plotted parallel to the x-axis.
-For CSH, VPD is always less than VPD$_{crit}$ so that the plant response dominates in typical environmental conditions, emphasizing the water conservative strategy of those plants.
-For CRO on the other hand, VPD is almost always higher than VPD$_{crit}$, emphasizing that those plants operate with an aggressive water usage strategy, are water intensive and were actually engineered for photosynthesis rather than water saving.
-For forests and grass, whether plant response or atmospheric demand dominates is more dependent on environmental conditions, with about half of the observed VPD less than VPD$_{crit}$, i.e.
+For CSH, VPD is always less than VPD$_{crit}$ (off scale) so that the plant response dominates in typical environmental conditions, emphasizing the water conservative strategy of those plants.
+For CRO on the other hand, VPD is higher than VPD$_{crit}$ for more than 50\% of observations, emphasizing that those plants operate with an aggressive water usage strategy, are water intensive and were actually engineered for photosynthesis rather than water saving.
+For forests and grass,  more than half of the observed VPD are less than VPD$_{crit}$, i.e.
 in conditions where plant response dominates.
@@
 atmospheric demand only, emphasizing that there is still a strong regulation of evaporative flux by stomata and though the plant xylem.
-The sign term in this case would just be a constant ($\frac{c_p}{R_{air}} \approx 3.5$), which is far larger than any part of the curves for any PFT.
+The sign term in the PET case would just be a constant ($\frac{c_p}{R_{air}} \approx 3.5$), which is far larger than any part of the curves for any PFT.
 This highlights the deficiencies of PET's ability to capture land response to changes in atmospheric dryness (VPD).

### 2018-02-27 src/paper_figure_code/shared_functions.py

@@ -77,6 +77,8 @@ def custom_xlabel(_df, ax, label):
     ax.set_xlabel(label, fontsize=fontsize)
   return
 
+mean_df['vpd_crit'] = et_min_vpd(mean_df)
+
 # pft order for panel plots
 pft_order = ['DBF', 'EBF', 'ENF', 'CSH', 'CRO', 'GRA',]
 

### 2018-02-27 src/paper_figure_code/tables.py

@@ -14,6 +14,15 @@ def frequency(_df):
   return _df.d_et[_df.d_et < 0.].count() / _df.d_et.count()
 
 
+#### Table 1, pft tables
+print('\n\n******TABLE 1*****')
+print(mean_df.loc[:, ['g1', 'uwue', 'uwue_zhou', 'uwue_zhou_std']])
+
+
+#### Table 2, pft tables
+print('\n\n******TABLE 1*****')
+print(mean_df.loc[:, ['r_moist', 'c_a', 'gamma', 'vpd_crit']])
+
 std = df.groupby('pft').std()
 mean_df['d_et_bar'] = d_et(mean_df)
 mean_df['d_et_bar_std'] = np.absolute(mean_df['d_et_bar'])*std['vpd']
@@ -187,11 +196,11 @@ print(mean_df.uwue_zhou)
 # import codebase.plot_tools as plot_tools
 # df.groupby('pft').apply(plot_tools.histogram, meta)
 
-fig = plt.figure()
-ax = fig.add_subplot(111)
-for g1, uwue, label in zip(mean_df.g1, mean_df.uwue, mean_df.index):
-  ax.scatter(g1, uwue, label=label)
-ax.set_xlabel('g1')
-ax.set_ylabel('uwue')
-plt.legend(loc='best')
-plt.savefig('%s/climate_et/g1_uwue_relationship.png' % os.environ['PLOTS'])
+# fig = plt.figure()
+# ax = fig.add_subplot(111)
+# for g1, uwue, label in zip(mean_df.g1, mean_df.uwue, mean_df.index):
+#   ax.scatter(g1, uwue, label=label)
+# ax.set_xlabel('g1')
+# ax.set_ylabel('uwue')
+# plt.legend(loc='best')
+# plt.savefig('%s/climate_et/g1_uwue_relationship.png' % os.environ['PLOTS'])

### 2018-02-28 doc/paper/vpd_et_paper.tex
finished scale vary section

@@
 in conditions where plant response dominates.
-It is also important to note that even when atmospheric demand dominates, ET response to VPD is still far smaller than it would be for potential evaporation $\partial PET/\partial VPD$, i.e.
+It is also important to note that even when atmospheric demand dominates, ET response to VPD is still far more negative than it would be for potential evaporation $\partial PET/\partial VPD$, i.e.
 atmospheric demand only, emphasizing that there is still a strong regulation of evaporative flux by stomata and though the plant xylem.
@@
 Plants are always regulating water exchange from the land surface, even when they reach the limits of they ability to do so.
-Therefore, the actual land surface response to a change in VPD will always me more negative than if the plants are not included in the physics (PET).
 
@@
 \label{scale_term}
-While the above discussion of the sign of $\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our question, understating the overall magnitude of the ET response is important to the soil-plant-atmosphere water budgeting.
+While the above discussion of the sign of $\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our question of when ET reponse is negative or postive, understating the overall magnitude of the ET response is important to soil-plant-atmosphere water budgeting.
 So we now more closely examine the terms that affect how the sign term is scaled:
@@
 \begin{equation}
-  \frac{g_a \; P}{T(\Delta + \gamma)},
+  \frac{g_a \; P}{T(\Delta + \gamma)}:
 \end{equation}
@@
 $\Delta$'s presence in the scaling term also matches physical intuition.
-Evaporative cooling will dampen the ability of the atmosphere to take more moisture (UNCLEAR SENTENCE REFORMULATE **), because $e_{s}$ decreases with decreasing temperature.
-The decrease in $e_{s}$ is proportional to $\Delta$ ($\delta e_{s} = \Delta \delta T$).
-So as $\Delta$ increases, you will get a larger damping of ET due to evaporative cooling.
+$\Delta$ (and also the approxmately constant $\gamma$) control the efficiencies with which surface energy is converted to latent and sensible heat \citep{Monteith_1964}.
 The functional from of $\Delta$ will be the same across PFT, but the temperature range may vary slightly.
@@
 Differences between PFT are almost entirely due to differences in aerodynamic conductance, rather than differences in observed temperature ranges.
-The scaling term for the tree PFTs (DBF, ENF) is generally more than double the scaling terms for the other PFTs which have lower surface roughness and generally smaller aerodynamic conductance.
+The scaling term for the tree PFTs (DBF, EBF, ENF) is generally about double the scaling terms for the other PFTs which have lower surface roughness and generally smaller aerodynamic conductance.
 
@@
 While the observed variability of the aerodynamic conductance contributes more to the scaling term variability than temperature, the temperature contribution is non-negligible.
-Specifically, the scaling term is generally larger at low temperatures when ET is weakly damped by Clausius-Clapeyron, and the effect of aerodynamic conductance variability is stronger as it is less weakly damped by Clausius-Clapeyron.
+Specifically, the scaling term is generally larger at low temperatures when latent heat is relativel inefficient at moving energy away from the surface.
+This effect aplifies the role of aerodynamic conductance variability at low temperatures.
 
-To summarize, variability between PFTs is mostly controlled by systematic differences in aerodynamic conductance, due to differences in surface roughness between each PFT and wind conditions.
+To summarize, variability between PFTs is mostly controlled by systematic differences in aerodynamic conductance, due to differences in surface roughness between each PFT, and posibly to a lesser extent wind conditions.
 In contrast, variability within PFT is also controlled by temperature, through Clausius-Clapeyron.

### 2018-02-28 doc/paper/vpd_et_paper.tex
done with sign term test, on to the scatter plots which really need to be reworked

@@
 \label{scale_term}
-While the above discussion of the sign of $\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our question of when ET reponse is negative or postive, understating the overall magnitude of the ET response is important to soil-plant-atmosphere water budgeting.
+While the above discussion of the sign of $\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our question of when ET reponse increases or decreases with VPD, understating the overall magnitude of the ET response is important to soil-plant-atmosphere water budgeting.
 So we now more closely examine the terms that affect how the sign term is scaled:
@@
 $\Delta$'s presence in the scaling term also matches physical intuition.
-$\Delta$ (and also the approxmately constant $\gamma$) control the efficiencies with which surface energy is converted to latent and sensible heat \citep{Monteith_1964}.
-The functional from of $\Delta$ will be the same across PFT, but the temperature range may vary slightly.
+$\Delta$ (and also the approxmately constant $\gamma$) control the efficiencies with which surface energy is converted to latent and sensible heat \citep{Monteith_1965}.
+The functional from of $\Delta$ will be the same across PFTs, but the temperature range may vary slightly.
 In contrast, aerodynamic conductance will vary strongly with PFT due to the importance of surface roughness for aerodynamic conductance.
@@
 While the observed variability of the aerodynamic conductance contributes more to the scaling term variability than temperature, the temperature contribution is non-negligible.
-Specifically, the scaling term is generally larger at low temperatures when latent heat is relativel inefficient at moving energy away from the surface.
-This effect aplifies the role of aerodynamic conductance variability at low temperatures.
+Specifically, the scaling term is generally larger at low temperatures when latent heat is relatively inefficient at moving energy away from the surface.
+This effect amplifies the role of aerodynamic conductance variability at low temperatures.
 
@@
 \label{stats_sec}
-FORULATE THAT DIFFERNTLY< SAY WE NOW TURN TO OBSERVATIONS> YOU HVAE ALREADY DISCSSED SIGAMA AND IRTS DEFINTION.
-MENTION IT WHILE HYOU DISCUSS THE RESULS BUT DO NOT MAKE IT THE MAIN FOCUS.
-I TRIED DOING IT
 
 In this section we consider direct observations of ET response with eddy-covariance data, while including uncertainty with the $\sigma$ term (Section \ref{methods}).
-These observational results of ET response largely confirm our theoretical analysis, presented in Sections \ref{sign_term}-\ref{scale_term} (Table \ref{stats}).
+These observational results of ET response largely confirm our theoretical analysis, presented above (Table \ref{stats}).
 For all PFTs, mean ET response to increasing VPD is negative.
 However, ET response evaluated at the average of all variables (e.g.
-$\sigma$, $T$, $c_a$, $VPD$) is only negative for CSH and GRA.
-This difference in mean ET response as compared to the ET response at mean environmental conditions is due to the non-linear nature of the response, in which negative responses are generally much larger magnitude than positive responses (Figure \ref{sign_plot}).
+$\sigma$, $T$, $c_a$, $VPD$) is postive for CRO, and negative for all other PFTs.
+This difference in mean ET response as compared to the ET response at mean environmental conditions is due to the non-linear nature of the response, in which negative responses are generally larger magnitude than positive responses (Figure \ref{idealized_sign}).
 Therefore, both the mean ET response as well as the ET response at mean environmental conditions matches our expectations from the theory (Section \ref{sign_term}), with the exception that CRO observations are shifted more towards a negative ET response than we expect.
 
-Regarding the frequency of negative and positive ET response, 53\% of ENF observations, 65\% of GRA observations, and 93\% of CSH observations are characterized by a negative ET response to increasing VPD.
-The response for these PFTs is thus more frequently physiologicaly controlled, and these PFTs will tend to conserve soil water for futre use.
-However DBF and CRO observations have greater frequency of positive ET response to increasing VPD (46\% and 49\% of observations, respectively), and are more intensive in their use of water.
-
-In general the bulk statistics match our theoretical expectations well, with the caveat that inclusion of uncertainty shifts crops towards a slightly more negative response than the theory and Figure \ref{sign_plot} would suggest.
-The bulk statistics motivate a more thorough examination of the structure of uncertainty and more sophisticated validation of our theory's performance against observations.
-
 \begin{table}
-\caption{Statistics of $\frac{\partial \; ET}{\partial \; VPD}$ as a function of PFT. \textbf{***dated, need to update these statistics***}}
+\caption{Statistics of $\frac{\partial \; ET}{\partial \; VPD}$ as a function of PFT.}
 \centering
-\begin{tabular}{l c c c c c}
+\begin{tabular}{l c c c}
   \hline
-PFT & $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ & $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)$ & $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)*\text{std}(VPD)$ & $\frac{\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)*\text{std}(VPD)}{ \frac{\partial \; ET}{\partial \; R}\left(\overline{env}\right)*\text{std}(R)}$ & fraction $\frac{\partial \; ET}{\partial \; VPD} < 0.$ \\
+PFT & $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ & $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)$ & fraction $\frac{\partial \; ET}{\partial \; VPD} < 0.$ \\
   \hline
-CRO & 0.000853 & 0.026241 & 37.05 & 0.41 & 0.473311\\
-CSH & -0.108234 & -0.091526 & 101.72 & 0.88 & 0.931660\\
-DBF & -0.012727 & 0.013794 & 39.47 & 0.33 & 0.461674\\
-ENF & -0.034087 & 0.000706 & 33.22 & 0.30 & 0.534425\\
-GRA & -0.019637 & -0.000921 & 33.60 & 0.35 & 0.631735\\
+  CRO & -0.05 & 0.008 &  0.533 \\
+  DBF & -0.10 & -0.013 &  0.609 \\
+  EBF & -0.13 & -0.037 &  0.666 \\
+  GRA & -0.07 & -0.024 &  0.694 \\
+  ENF & -0.17 & -0.081 &  0.740 \\
+  CSH & -0.29 & -0.224 &  0.973 \\
   \hline
@@
 
+Regarding the frequency of negative and positive ET response, all PFTs exibit a  decreasing ET response with increasing VPD (physiologically controlled, water conservative response) for the majority of observations.
+When listed by increasing frequency of negative response, the PFTs are generally ordered as we would expect them to be, from most water intesive to most water conservative: CRO (53\%), DBF (61\%), EBF (67\%), GRA (69\%), ENF (74\%), and CSH (97\%).
+
+In general the bulk statistics match our theoretical expectations well, with the caveat that inclusion of uncertainty shifts crops towards a slightly more negative response than the theory and Figure \ref{idealized_sign} would suggest.
+The bulk statistics motivate a more thorough examination of the structure of uncertainty and more sophisticated validation of our theory's performance against observations.
+
 \subsection{Validation of theory at eddy-covariance sites}
 \label{testing}
-While Section \ref{stats_sec} presented simple statistics of observed response, we now compare more sophisticated distributions of the observed response to our theory presented in Section \ref{sign_term}.
-Figure \ref{test_sign} shows observed the distribution of of the sign term, as compared to what the theory would predict.
+We now compare more sophisticated distributions of the observed response to our simpified theory (Section \ref{sign_term}).
+Figure \ref{test_sign} shows the observed distribution of the sign term, as compared to what the theory would predict.
 Our goal was to capture the leading order behavior of the ET dependence on VPD.
@@
 
-For DBF our theory is successful towards this goal: the theory matches the leading order behavior of the function when uncertainty is included, and the DBF response appears as expected: i.e.
-is similar to the theory with some additional noise.
-The VPD dependence, given by the slope parameter ($g_1$), follows the median values of each box plot.
+For DBF our theory is successful towards this goal: the theory matches the leading order behavior of the function when uncertainty is included, and the observations match the theory with the addtion of noise.
+The VPD dependence, given by the slope parameter ($g_1$), follows the median values of each bin.
 Perhaps most importantly, the x-intercept, and thus VPD$_{crit}$, matches nearly exactly between the theory and the observations.
 Therefore the sign of the ET response to increases in VPD should be well matched, subject to the unavoidable constraints of noise, much of which comes from the observations themselves.
-The uncertainty is non negligible; there are many observations in each box for which the the sign of observation is opposite the response predicted by the theory, but to leading order our theory matches the observations well.
+The uncertainty is non negligible; there are many observations in each bin for which the the sign of observation is opposite the response predicted by the theory, but to leading order our theory matches the observations well.
 
@@
 The VPD-independent, strongly negative response, which is determined by $uWUE$, is also captured.
-For CSH, there is extremely rare occurrence of positive ET response with VPD, even with uncertainty, so the sign of the observations almost always matches the sign of the theory.
+For CSH, there is extremely rare occurrence of observed positive ET response with VPD ($\approx 3$\%), even with uncertainty, so the sign of the observations almost always matches the sign of the theory, which states that ET response should always be negative.
 
-As Section \ref{stats_sec} suggested, the theory does not match as well for CRO as for DBF and CSH.
-At low VPD the VPD dependence (due to $g_1$) is similar to the leading order behavior of the observations, but the theory's curve is shifted towards more positive values relative to the observations, due to some bias in $uWUE$.
+As Table \ref{stats} suggested, the theory does not match as well for CRO as for DBF and CSH.
+At low VPD the VPD dependence (due to $g_1$) is similar to the leading order behavior of the observations, but the theory's curve is shifted towards more positive values relative to the observations, likely due to some bias in $uWUE$.
 Most strikingly, the observations show a tendency to move back towards negative values at high VPD, which given the functional form of VPD dependence, is behavior that is impossible for the theory to capture.
-This general positive bias of the theory explains the shift towards negative ET response observed in the bulk statistics of Section \ref{stats_sec}.
+This general positive bias of the theory explains the shift towards negative ET response observed in the bulk statistics in Table \ref{stats}.
 The divergence between the theory and the observations for CRO could have been expected.
@@
 GRA observations are characterized by a consistent trend back towards negative ET response at higher VPD, which the functional form of our theory is incapable of accounting for.
-As compared to CRO, the divergence between the theory and the observations is even greater for GRA, but these differences are explainable by similar issues as for CRO.
-The GRA PFT tends to exhibit substantial within FPT difference, such as C3 vs C4 photosynthesis, as well as strong seasonal variability in height and LAI.
+As compared to CRO, the divergence between the theory and the observations is even greater for GRA, but these differences are explainable using similiar logic as we did for CRO.
+GRA also tends to exhibit substantial within-PFT variability, such as C3 vs C4 photosynthesis, as well as strong seasonal variability in height and LAI.
 As with CRO, the trend back towards negative ET response at higher VPD could be explained by a superposition of sites with C3 vs C4 sites.
@@
 
-The comparison between the theory and observations for ENF is unique from the rest of the PFTs.
-The VPD dependence of the theory, determined by the slope parameter ($g_1$), is not well captured by observations.
-However, the theory closely captures the x-intercept and VPD$_{crit}$.
-So, the theory still captures when the ET response will be positive or negative, but is less accurate with the magnitude of the response.
-Therefore the theory is still useful for addressing when VPD drives or reduces ET.
-We note however that we used published constants for $g_1$ \citep[from ][]{Lin_2015}, which determines the VPD dependence and that deviations should be expected across sites.
+The comparison between the theory and observations for ENF and EBF is unique from the rest of the PFTs.
+The VPD dependence of the theory, determined by the slope parameter ($g_1$), is not as well captured by observations.
+However, the theory still captures the general behavior and magnitude of the response, with some increase in error at low and/or high VPD.
+We note that we used published constants for $g_1$ \citep[from ][]{Lin_2015, Franks_2017}, which determines the VPD dependence, and that deviations should be expected across sites.
 We could have fit our own constants so that the theory would best match the observations, but our goal was to use a priori results whenever possible in order to avoid overfitting and to have a theoretical approach as independent to the observations as possible for further confirmation of the validity of the approach.
 
-While the above discussion shows that our theory has some limitations when applied to some PFTs, especially for crops and grasslands, it does well for DBF and CSH PFTs, and captures the sign of the response consistently for ENF.
+While the above discussion shows that our theory has some limitations when applied to some PFTs, especially for crops and grasslands, it does well for DBF and CSH PFTs, and captures the response at non-extreme VPD for ENF and EBF.
 It is useful to also compare our theory, the observed uncertainty in the ET response, and PET, as PET is often used as a proxy for ET in hydrometeorological studies and drought indices.
-Again, as in Section \ref{sign_term}, the PET sensitivity to VPD will be constant at about 3.5.
+Again, as in our discussion of the theoretical results, the PET sensitivity to VPD will be constant at about 3.5.
 In Figure \ref{test_sign} the ET response to shifts in dryness is always far smaller than 3.5 in the observations, even with all uncertainty included.
@@
 
-\subsubsection{Testing VPD$_{crit}$}
+\subsection{Testing VPD$_{crit}$}
 As stated in the above discussion of ENF, we are mostly interested in the critical threshold $VPD_{crit}$ for which the sign of the ET response to VPD ($\frac{\partial \; ET}{\partial \; VPD}$) is determined.

### 2018-02-28 src/paper_figure_code/tables.py

@@ -20,7 +20,7 @@ print(mean_df.loc[:, ['g1', 'uwue', 'uwue_zhou', 'uwue_zhou_std']])
 
 
 #### Table 2, pft tables
-print('\n\n******TABLE 1*****')
+print('\n\n******TABLE 3*****')
 print(mean_df.loc[:, ['r_moist', 'c_a', 'gamma', 'vpd_crit']])
 
 std = df.groupby('pft').std()
@@ -35,20 +35,16 @@ mean_df['total_var'] = np.sqrt(mean_df.d_et_bar_std**2\
                                +mean_df.d_et_d_rn_bar_std**2)
 mean_df['d_et_bar_norm_total'] = mean_df['d_et_bar_std']\
                            /mean_df['total_var']
-
+mean_df['et_std'] = std['et_obs']
 counts = df.groupby('pft').apply(frequency)
 
-print('\n mean d_et\n', mean_df.d_et)
-print('\n d_et(mean)\n', mean_df.d_et_bar)
-print('\n mean d_et *std vpd\nn', mean_df['d_et_bar_std'])
-print('\n mean d_et / rad\n', mean_df['d_et_bar_norm'])
-print('\n portion d_et < 0 \n',\
-      counts)
 mean_df['counts'] = counts
-test = mean_df.loc[:, ['d_et', 'd_et_bar', 'd_et_bar_std',\
-                       'd_et_bar_norm', 'd_et_bar_norm_total', 'counts']]
+# test = mean_df.loc[:, ['d_et', 'd_et_bar', 'd_et_bar_std',\
+#                        'd_et_bar_norm', 'counts']]
 
-print(test)
+
+print('\n\n******TABLE 4******')
+print(mean_df.loc[:, ['d_et', 'd_et_bar', 'counts']])
 
 
 print(mean_df.uwue)

### 2018-02-28 doc/paper/vpd_et_paper.tex
simplify scatter plot to just do derivative d_et/d_vpd?

@@
 
-\subsection{Testing VPD$_{crit}$}
+\subsection{Observed ET response to VPD}
 As stated in the above discussion of ENF, we are mostly interested in the critical threshold $VPD_{crit}$ for which the sign of the ET response to VPD ($\frac{\partial \; ET}{\partial \; VPD}$) is determined.

### 2018-02-28 src/paper_figure_code/data_scatter.py
only plot det/dvpd

@@ -7,6 +7,7 @@ import time
 
 # mean_df is defined up at this level
 local_fontsize = fontsize+3
+
 ### Figure 6 ###
 # note below really takes a long time
 def d_et_uwue_fixed(_df):
@@ -67,22 +68,21 @@ def make_ax_plot(_ax, var, _df, meta):
   z = np.ma.masked_array(grouped.values, mask=np.isnan(grouped.values))
   color = _ax.pcolormesh(_x, _y, z, cmap=meta['cmap'],\
                          vmin=vmin, vmax=vmax)
-  if str(_df['pft'].iloc[0]) == 'GRA':
-    _ax.set_xlabel('%s (Pa)' % 'vpd'.upper(), fontsize=local_fontsize)
-  if meta['title'] ==  r'$\frac{\partial \; ET}{\partial \, VPD}$':
-    _ax.set_ylabel('T (C)', fontsize=local_fontsize)
-  # _ax.set_title('%s: %s'\
-  #               % (str(_df['pft'].iloc[0]),\
-  #                  meta['title']), fontsize=local_fontsize+3)
+
+  custom_xlabel(_df, _ax, '%s (Pa)' % 'vpd'.upper())
+  custom_ylabel(_df, _ax, 'T (C)')
   _ax.set_title('%s'\
-                % (name_dict[_df['pft'].iloc[0]]), fontsize=local_fontsize+3)
-  _ax.yaxis.set_tick_params(labelsize=local_fontsize-3)
-  _ax.xaxis.set_tick_params(labelsize=local_fontsize-3)
+                % (name_dict[_df['pft'].iloc[0]]), fontsize=local_fontsize)
+  # _ax.yaxis.set_tick_params(labelsize=local_fontsize-3)
+  # _ax.xaxis.set_tick_params(labelsize=local_fontsize-3)
   cbar = plt.colorbar(color, ax=_ax)# , ax=_ax2)
-  cbar.set_label(meta['title'], fontsize=local_fontsize+3)
+  
+  if ((_df.pft.iloc[0] == 'EBF') | (_df.pft.iloc[0] == 'CSH')\
+      | (_df.pft.iloc[0] == 'GRA')):
+    cbar.set_label(meta['title'], fontsize=local_fontsize)
   y_lim = _ax.get_ylim()
-  vpd_crit = et_min_vpd(mean_df.loc[_df.pft.iloc[0], :])
-  _ax.plot(np.ones(2)*vpd_crit, np.array(y_lim), 'k-', linewidth=2.0)
[…]

### 2018-02-28 doc/paper/vpd_et_paper.tex
only plot det/dvpd
finished scatter section

--- 2018-02-28 doc/paper/vpd_et_paper.tex
@@
 
-\subsection{Testing VPD$_{crit}$}
-As stated in the above discussion of ENF, we are mostly interested in the critical threshold $VPD_{crit}$ for which the sign of the ET response to VPD ($\frac{\partial \; ET}{\partial \; VPD}$) is determined.
-When VPD is less than VPD$_{crit}$, ET response to VPD is negative ($\frac{\partial \; ET}{\partial \; VPD}  < 0$), i.e.
-plant response dominates, and when VPD is greater than VPD$_{crit}$, ET response to VPD is positive  ($\frac{\partial \; ET}{\partial \; VPD} > 0$).
-i.e.
-atmospheric demand dominates.
-Theoretically, this $VPD_{crit}$ is mostly a function of PFT and not of climate.
-This $VPD_{crit}$, and how it compares with environmental VPD and water supply, will exert a strong control on species selection in competitive environments.
+\subsection{Observed ET response to VPD}
 
-Figure \ref{data_scatter} presents calculated $\frac{\partial \; ET}{\partial \; VPD}$ where, unless otherwise noted, all variables in Equation \ref{d_et} are allowed to vary, including uncertainty ($\sigma$).
-Each column represents a different quantity related to $\frac{\partial \; ET}{\partial \; VPD}$, and each row is a different PFT.
-Vertical black lines denote where our theory says $VPD_{crit}$ should be.
-For any observations to the left of this line $\frac{\partial \; ET}{\partial \; VPD}$ should be negative (plant response dominates), and any observations the right of the line should be positive (atmospheric demand dominates).
-In addition to seeing how our theoretically-derived VPD threshold, $VPD_{crit}$, matches observations, these analysis are also useful for observing how the scaling term affects the magnitude of the derivative, if at all.
+Most of the results presented so far focus on the sign term, so now we turn to observations of ET response with the scaling term included (Figure \ref{data_scatter}).
+Until now, in the interest of developing leading order intuition for ET response, and to be conservative in our acknowledement of model and observational error, we've considered $\sigma$ variability to be a measure of uncertainty.
+An alternative viewpoint is that $\sigma \cdot uWUE$ represents spatial and temporal variability of $uWUE$, which may be expected within bounds \citep[see Table \ref{pft}, also ][]{Zhou_2015}.
+This is a less conservative view; some of the $\sigma \cdot uWUE$ variability will be due to model and observational error so by viewing $\sigma \cdot uWUE$ as \textit{real} variability we run the risk of mistaking noise for signal.
+However, the advantage of this viewpoint is that, from \citet{Zhou_2014}, we have very high confidence that $uWUE$ fit at the hourly timescale (as we do with $\sigma \cdot uWUE$) correctly captures the relationship between ET, GPP, and VPD, and that our form of PM introduces minimal error with its use of $uWUE$.
 
@@
 \centering
-\centerline{\includegraphics[width=1.4\textwidth]{./data_scatter.png}}
-\caption{Scatter plots of $\frac{\partial \; ET}{\partial \; VPD}$.
-Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; VPD}$, as labeled: Column 1: $\frac{\partial \; ET}{\partial \; VPD}$; Column 2: $\frac{\partial \; ET}{\partial \; VPD}$ normalized by $g_a$; Column 3: $\frac{\partial \; ET}{\partial \; VPD}$ with $\sigma$ held fixed at 1; and Column 4: $\frac{\partial \; ET}{\partial \; VPD}$ normalized by $g_a$ and with $\sigma$ held fixed.
-Please note differences in the colorbar scale.}
+\centerline{\includegraphics[width=\textwidth]{./data_scatter.png}}
+\caption{Scatter plots of observed $\frac{\partial \; ET}{\partial \; VPD}$, including $\sigma$ variability, as a function of PFT. Please note differences in the colorbar scale.}
 \label{data_scatter}
@@
 
-Figure \ref{data_scatter} generally confirms expectations from Figure \ref{test_sign}.
-By comparing different columns of Figure \ref{data_scatter} we can evaluate the impact of $\sigma$ and aerodynamic conductance ($g_a$) on the variability of ET response ($\frac{\partial \; ET}{\partial \; VPD}$).
-Aerodynamic conductance's scaling (included in columns 1 and 3) alters the magnitude considerably, particularly between PFTs.
-$\sigma$ variability (included in columns 1 and 2) adds a lot of additional noise to the signal of $\frac{\partial \; ET}{\partial \; VPD}$, as observed in Figure \ref{test_sign}.
-However, the general story with the noise appears to match the cleaner signal when $\sigma$ is held constant and $VPD_{crit}$ is clearly visible.
-The exceptions are, as discussed above, for GRA and CRO, where the full complexity plots (Columns 1 and 2) do not match well with $\sigma$ held fixed (Columns 3 and 4).
+Therefore, with the caveat that some of the signal presented in Figure \ref{data_scatter} may in fact be noise, we can interpret the observed distribution of ET response to VPD.
+In general, the observed response matches the intuitive theory.
+ET response to VPD shifts towards positive values as VPD increases (atmospheric demand dominating).
+CRO exhibit the highest occurence of postive ET response, and the obsevations confirm that CRO are the most water intensive.
+CSH are the most water conservative, with a strong negative response.
+DBF, EBF, ENF, and GRA are also water conservative, but show some occurence of positive ET response to VPD, particularly at higher observed VPD, as the theory predicts.
 
-While the stomatal conductance $g_a$ has a large impact on the magnitude of the scaling, temperature, through Clausius-Clapeyron ($\Delta$), does not appear to have as much of an impact in the complete estimation of ET response ($\frac{\partial \; ET}{\partial \; VPD}$).
-Therefore, it appears that the absolute magnitude of ($\frac{\partial \; ET}{\partial \; VPD}$ can be mostly attributed to the sign term, and its direct scaling by aerodynamic conductance.
+Figure \ref{data_scatter} also includes the impact of the scaling term.
+For a given VPD, the magnitude of ET response does not vary strongly with temperature, confirming that any impact of the scaling term on the magnitude of the response is primarily due to changes in aerodynamic conductance.
 Intuitively this is reasonable; aerodynamic conductance will control how dominant balances at the land surface are communicated to the boundary layer.
-While $\Delta$ controls the efficiency of latent heat flux in removing heat from the surface, it appears this is a second order term, relative to $g_a$, for scaling ET response.
+While $\Delta$ controls the efficiency of energy conversion to latent heat, it appears this is a second order term, relative to $g_a$, for scaling ET response.
 
-In general Figures \ref{test_sign} and \ref{data_scatter} match our expectations.
-Our $\sigma$-based method of uncertainty propagation blurs the idealized expectations the most for GRA and CRO.
-We therefore have the most confidence in our conclusion based on Equation \ref{d_et} for PFTs CSH, DBF, and ENF, as the sign of the derivative with uncertainty included matches (to leading order) the story when $\sigma$ is held fixed.
-
-% \begin{figure}
-% \centering
-% \includegraphics[width=\textwidth]{./fig06b.png}
-% \caption{****alternate Fig 06****  Scatter plots of $\frac{\partial \; ET}{\partial \; VPD}$.
-Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; VPD}$, as labeled.
-If I end up using this, I could also draw on the curve of $VPD_{crit}$ with $\overline{\frac{\text{LAI}}{\text{LAI$_{ref}$}}}$.
-}
-% \label{real2}
-% \end{figure}
+As in Figure \ref{test_sign}, Figure \ref{data_scatter} matches our expectation based on simplified theory.
+The sign term is most strongly scaled by $g_a$, and in general the sign term shifts towards positive values as VPD observations increase.
+The willingness with which a given plant type evolved to use water dictates the occurence of positve versus negative ET response.
+Water conservative plants are highly effective at mitigating the effects of atmospheric demand, and can store water for later use by reducing ET in response to increasing VPD.
 
@@
 So systematic soil moisture relationshps are not apparent in Figure \ref{swc_boxplot}, but instead blurred by the superposition of the two sites.
-We argue taht the soil moisture dependence is still there, but we don't have enough statistcs to comment on it}.
+We argue that the soil moisture dependence still exists, but commenting on the PFT-wide relationship is not currently possible.}.
 

### 2018-02-28 doc/paper/vpd_et_paper.tex
finished SWC section, onto discussion

@@
 
-\section{Limitations of theory - very dry soil moisture conditions}
+\subsection{Limitations of theory - very dry soil moisture conditions}
 In formulating our theory with Penman Monteith, we implicitly did not account for very dry soil moisture conditions.
@@
 
-We first note that in our datasets low SWC conditions are very infrequent (MAKE SEPERATE PDF OF SWC?).
+We first note that in the data used low SWC conditions are very infrequent.
 Within our framework, any systematic bias due to the  failure to account for SWC's effects in very dry conditions should manifest itself in a functional relationship between $\sigma$, our uncertainty measure, and SWC, which is observed at the FLUXNET sites.
@@
 
-If we view $\sigma \cdot uWUE$ as, in addition to a measure of uncertainty, a short time-scale observation of uWUE, we would expect $\sigma \cdot uWUE$ to decrease at low SWC (CITE ZHOU ** which ZHOU shows low uWUE with low SWC).
+If we again view $\sigma \cdot uWUE$ as, in addition to a measure of uncertainty, a short time-scale observation of uWUE, we would expect $\sigma \cdot uWUE$ to decrease at low SWC.
 If $\sigma \cdot uWUE$ is a very strong function of SWC, then our theory should be conditioned more strongly on well watered soil conditions.
@@
 And, CRO, relative to DBF, has a very weak dependence on soil moisture, which is reflective of the high likelihood that CRO sites are optimally irrigated and not water stressed.
-For ENF, systematic SWC-induced biases in $\sigma \cdot uWUE$ emerge in about the lowest 30\% of SWC conditions.
+For ENF and EBF, systematic SWC-induced biases in $\sigma \cdot uWUE$ emerge in about the lowest 30\% of SWC conditions.
 In crontast to ENF, CRO, and DBF, for GRA the relationship between $\sigma \cdot uWUE$ and SWC is more linear and affects a greater portion of observations; about 60\% of observations.
@@
 It is therefore not suprising that our theory tested poorly againt the data for GRA, relative to ENF, DBF, and GRA (Section \ref{testing}).
-For all other PFTs occurence of soil mosture impacts was rare enough to not manifest itself in bulk statistics and plots\footnote{note that interpreting the soil moisture dependence for CSH is difficult as their are only two sites for that PFT.
-So systematic soil moisture relationshps are not apparent in Figure \ref{swc_boxplot}, but instead blurred by the superposition of the two sites.
-We argue that the soil moisture dependence still exists, but commenting on the PFT-wide relationship is not currently possible.}.
+For all other PFTs occurence of soil mosture impacts was rare enough to not manifest itself in bulk statistics and plots.
+A possible exception is CSH, for which there are only two sites, which precludes comment on the exact PFT-wide relationship.
 
@@
 
-The observed dependence of $\sigma \cdot uWUE$ on SWC for GRA would explain the deficiencies of our theory compared to the observations in Section \ref{testing}, specifically the trend back towards negative soil moisture at high VPD.
+The observed dependence of $\sigma \cdot uWUE$ on SWC for GRA would explain the deficiencies of our theory compared to the observations in Figure \ref{idealized_sign}, specifically the trend back towards negative soil moisture at high VPD.
 Aforementioned feedbacks between the land surface and the atmosphere, which are not accounted for due to our focus on the one-way response of the land surface to atmospheric conditions, would cause high VPD to be correlated with low SWC \citep[][]{Gentine_2016, Berg_2016}.
-So, at high VPD observations of low SWC are more likely, and this low SWC causes a lower uWUE (ZHOU ET AL **).
+So, at high VPD observations of low SWC are more likely, and this low SWC causes a lower uWUE (Figure \ref{swc_boxplot}).
 The lower uWUE at low SWC/high VPD then leads to the observed downturn towards decreasing ET with increasing VPD, and the deviation between our theory (based on a constant uWUE assumption) and observations.

### 2018-02-28 doc/paper/vpd_et_paper.tex
wow, think I'm done. need to spell check and final read through

@@
 While Equation 3 can be used in PM (Equation 1), it will make analytical work with the function intractable because $A$, net CO$_2$ assimilation rate, is functionally related to ET itself.
-To remove the dependence of ET on $A$ we use fundamental semi-empirical results of \citet{Zhou_2015}, who showed that the underlying Water Use Efficiency ($uWUE$):
+To remove the dependence of ET on $A$ we use fundamental semi-empirical results of \citet{Zhou_2015}, who showed that the underlying Water Use Efficiency (uWUE):
 
@@
 
-Note that $uWUE$ is fit on the ecosystem scale as in \citet{Zhou_2015}, so GPP in Equation \ref{uwue} is really $A \cdot \text{LAI}$.
+Note that uWUE is fit on the ecosystem scale as in \citet{Zhou_2015}, so GPP in Equation \ref{uwue} is really $A \cdot \text{LAI}$.
 This leads to the cancellation of LAI in the step from Equation \ref{medlyn} to Equation \ref{new_g_s}.
@@
 
-With Equation \ref{et}, ET is explicitly a function of environmental variables and two plant-specific constants, the slope parameter ($g_1$), and $uWUE$, both reflecting water conservation strategy.
+With Equation \ref{et}, ET is explicitly a function of environmental variables and two plant-specific constants, the slope parameter ($g_1$), and uWUE, both reflecting water conservation strategy.
 The slope parameter is a leaf-scale term related to the willingness of stomata to trade water for CO$_2$ and to keep stomata open.
-$uWUE$ is a semi-empirical ecosystem-scale constant related to how WUE changes with VPD (specifically $VPD^{-1/2}$).
+uWUE is a semi-empirical ecosystem-scale constant related to how WUE changes with VPD (specifically $VPD^{-1/2}$).
 Is is also roughly proportional to physical constants:
@@
 
-Given the eddy-covariance FLUXNET2015 data described in Section \ref{data}, every term in our new version of PM (Equation \ref{et}) is observed, except for the two parameters $g_1$, and $uWUE$.
-$g_1$ has been measured and reported for the PFTs considered here \citep{Lin_2015, Franks_2017}, and we fit $uWUE$ by calculating its expectation, given the model and FLUXNET2015 data.
+Given the eddy-covariance FLUXNET2015 data described in Section \ref{data}, every term in our new version of PM (Equation \ref{et}) is observed, except for the two parameters $g_1$, and uWUE.
+$g_1$ has been measured and reported for the PFTs considered here \citep{Lin_2015, Franks_2017}, and we fit uWUE by calculating its expectation, given the model and FLUXNET2015 data.
 
@@
 To account for observational error, as well as model uncertainty (e.g.
-temporal and spatial variations of $uWUE$), we introduce an uncertainty parameter $\sigma$ modifying $uWUE$:
+temporal and spatial variations of uWUE), we introduce an uncertainty parameter $\sigma$ modifying uWUE:
 
@@
 This makes Equation \ref{et} a useful alternative to PET in drought indices and hydrometeorological analysis for vegetated surfaces.
-Equation \ref{et} better reflects the physics of water exchange at the land surface and would only require fitting of $uWUE$ and $g_1$, and can also account for changes in CO$_2$ concentration, contrary to typical drought indices such as the Palmer Drought Severity Index (PDSI) \citep{Swann_2016}.
+Equation \ref{et} better reflects the physics of water exchange at the land surface and would only require fitting of uWUE and $g_1$, and can also account for changes in CO$_2$ concentration, contrary to typical drought indices such as the Palmer Drought Severity Index (PDSI) \citep{Swann_2016}.
 
@@
 CO$_2$ concentration ($c_a$) and the psychometric constant ($\gamma$) are relatively constant over the dataset considered here so that the variability is dominated by $\sigma$ and $VPD$.
-$uWUE$ could vary with soil moisture but has been shown to be relatively constant \citep{Zhou_2015}.
+uWUE could vary with soil moisture but has been shown to be relatively constant \citep{Zhou_2015}.
 This then means that the sign term only depends on VPD for a given PFT and is approximately just a function of $VPD$.
@@
 \normalsize
-We note that $VPD_{crit}$ mostly depends on the PFT parameters $uWUE$ and $g_1$ but only vary weakly with climate as most other parameters related to the environment are nearly constant.
+We note that $VPD_{crit}$ mostly depends on the PFT parameters uWUE and $g_1$ but only vary weakly with climate as most other parameters related to the environment are nearly constant.
 The value of $VPD_{crit}$ for each PFT is shown in Table \ref{vpd_crit}.
@@
 
-Differences in $VPD_{crit}$ are exclusively determined by $uWUE$ and the slope parameter ($g_1$) related to the plant functional type.
-A larger $uWUE$ means a smaller $VPD_{crit}$, and an ET response to increases VPD that is more likely to be positive.
+Differences in $VPD_{crit}$ are exclusively determined by uWUE and the slope parameter ($g_1$) related to the plant functional type.
+A larger uWUE means a smaller $VPD_{crit}$, and an ET response to increases VPD that is more likely to be positive.
 At first glance this result is somewhat counterintuitive; we expect that plants with a higher water use effiency would be more water conservative.
-However, in reality $uWUE$ determines how $WUE$ changes with VPD:
+However, in reality uWUE determines how $WUE$ changes with VPD:
 
@@
 
-So, plants with a higher $uWUE$ will have a greater \textit{decrease} in ecosystem-scale $WUE$ in response to increases in VPD.
-This decrease in $WUE$ causes more water loss per unit carbon production, and explains the relationship between high $uWUE$ and high liklihood of increases of ET in response to increasing atmospheric drying (increases in VPD).
+So, plants with a higher uWUE will have a greater \textit{decrease} in ecosystem-scale $WUE$ in response to increases in VPD.
+This decrease in $WUE$ causes more water loss per unit carbon production, and explains the relationship between high uWUE and high liklihood of increases of ET in response to increasing atmospheric drying (increases in VPD).
 
@@
 Plants that are less conservative will be thus be more likely to increase ET with increasing VPD.
-Both the aforementioned effects (large $uWUE$, $g_1$) can amplify each other, and generally conspire to shift the sign term towards a positive value for a given PFT.
+Both the aforementioned effects (large uWUE, $g_1$) can amplify each other, and generally conspire to shift the sign term towards a positive value for a given PFT.
 
-This effect of $uWUE$ and $g1$ on the sign term is most apparent by comparing two extreme PFTs: water intensive crops (CRO) and water conservative closed shrub (CSH).
+This effect of uWUE and $g1$ on the sign term is most apparent by comparing two extreme PFTs: water intensive crops (CRO) and water conservative closed shrub (CSH).
 CRO has high slope parameters and uWUE ($g_1 = 183.1$ Pa$^{1/2}$; $2.27$ $\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET]) compared to CSH ($g1 = 148.6 \, Pa^{1/2}$, $uWUE=1.66$ $\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET]).
@@
 Again, the VPD dependence mostly determined from the slope parameter ($g_1$) closely matches the medians in the observation bins and the theoretical threshold is well reproduced.
-The VPD-independent, strongly negative response, which is determined by $uWUE$, is also captured.
+The VPD-independent, strongly negative response, which is determined by uWUE, is also captured.
 For CSH, there is extremely rare occurrence of observed positive ET response with VPD ($\approx 3$\%), even with uncertainty, so the sign of the observations almost always matches the sign of the theory, which states that ET response should always be negative.
@@
 As Table \ref{stats} suggested, the theory does not match as well for CRO as for DBF and CSH.
-At low VPD the VPD dependence (due to $g_1$) is similar to the leading order behavior of the observations, but the theory's curve is shifted towards more positive values relative to the observations, likely due to some bias in $uWUE$.
+At low VPD the VPD dependence (due to $g_1$) is similar to the leading order behavior of the observations, but the theory's curve is shifted towards more positive values relative to the observations, likely due to some bias in uWUE.
 Most strikingly, the observations show a tendency to move back towards negative values at high VPD, which given the functional form of VPD dependence, is behavior that is impossible for the theory to capture.
@@
 This casts doubt on PET's usefulness for capturing the leading order behavior of water budget and dryness response at the land surface, such as used in several drought indices.
-However our theory (Equation \ref{et}) could be used with published $uWUE$ and $g1$ data, as an alternative to PET to estimating land response and dryness, and would more accurately account for the response to atmospheric drying.
+However our theory (Equation \ref{et}) could be used with published uWUE and $g1$ data, as an alternative to PET to estimating land response and dryness, and would more accurately account for the response to atmospheric drying.
 
@@
 Until now, in the interest of developing leading order intuition for ET response, and to be conservative in our acknowledement of model and observational error, we've considered $\sigma$ variability to be a measure of uncertainty.
-An alternative viewpoint is that $\sigma \cdot uWUE$ represents spatial and temporal variability of $uWUE$, which may be expected within bounds \citep[see Table \ref{pft}, also ][]{Zhou_2015}.
+An alternative viewpoint is that $\sigma \cdot uWUE$ represents spatial and temporal variability of uWUE, which may be expected within bounds \citep[see Table \ref{pft}, also ][]{Zhou_2015}.
 This is a less conservative view; some of the $\sigma \cdot uWUE$ variability will be due to model and observational error so by viewing $\sigma \cdot uWUE$ as \textit{real} variability we run the risk of mistaking noise for signal.
-However, the advantage of this viewpoint is that, from \citet{Zhou_2014}, we have very high confidence that $uWUE$ fit at the hourly timescale (as we do with $\sigma \cdot uWUE$) correctly captures the relationship between ET, GPP, and VPD, and that our form of PM introduces minimal error with its use of $uWUE$.
+However, the advantage of this viewpoint is that, from \citet{Zhou_2014}, we have very high confidence that uWUE fit at the hourly timescale (as we do with $\sigma \cdot uWUE$) correctly captures the relationship between ET, GPP, and VPD, and that our form of PM introduces minimal error with its use of uWUE.
 
@@
 In formulating our theory with Penman Monteith, we implicitly did not account for very dry soil moisture conditions.
-For the majority of environmental conditions  observed at the eddy covariance sites used here, soil conditions were not extremely dry so that we could assume a constant $uWUE$ and $g_1$.
+For the majority of environmental conditions  observed at the eddy covariance sites used here, soil conditions were not extremely dry so that we could assume a constant uWUE and $g_1$.
 We posit that ecosystems will generally optimize to host plants living in conditions which they evolved for.
@@
 \includegraphics[width=\textwidth]{./swc_boxplot.pdf}
-\caption{$\sigma \cot uWUE$ as a function of PFT and SWC. Each boxplot represents 5\% of all observations.}
+\caption{$\sigma \cdot uWUE$ as a function of PFT and SWC. Each boxplot represents 5\% of all observations.}
 \label{swc_boxplot}
@@
 
-We derived a new form of Penman Monteith using the concept of semi-empirical optimal stomatal regulation \citep{Lin_2015, MEDLYN_2011} and near constant $uWUE$ \citep{Zhou_2015} to remove the stomatal conductance implicit dependence on GPP and ET.
+We derived a new form of Penman Monteith using the concept of semi-empirical optimal stomatal regulation \citep{Lin_2015, MEDLYN_2011} and near constant uWUE \citep{Zhou_2015} to remove the implicit dependence of stomatal conductance on GPP and ET.
 With our new form of Penman Monteith we developed a theory for when an ecosystem will tend to reduce or increase ET with increasing VPD, which we evaluated against a range of eddy-covariance data spanning different climates and plant functional types.
@@
 
-Our theory states that each plant type has a critical VPD below which ET will decrease (plant response dominating) and above which ET will increase (atmospheric demand dominates).
-We tested our theory using data from FLUXNET 2015, and found that for DBF and CSH the theory succeeds in capturing the leading order behavior.
-For ENF, the theory succeeds in capturing the leading order behavior for the sign of $\frac{\partial \; ET}{\partial \; VPD}$, but not for magnitude.
-In contrast, the model does not capture the behavior for eddy-covariance observations of CRO and GRA, which could have been expected, as our theory does not account for large sources of heterogeneity in those plant types, including varying surface roughness, c3 vs c4 photosynthesis, and potential differences in irrigation across sites.
-Also, our theory does not account for strong soil drying, which non-negligibly alters ET response for GRA.
-Given these deficiencies for CRO and GRA, we will withhold conclusions on their response to increasing VPD.
+Our theory suggests that for a majority of enviornmental conditions, plants will tend to conserve water and reduce ET with increasing VPD.
+Stomatal regulation and plant physiological response strongly regulate ET, and this regulation varies by PFT.
+CROs are the least water conservative, while DBF, EBF, ENF, GRA and CSH are progressively more water conservative (more likely to reudce ET in response to increasing VPD).
+Observations of ET response to VPD exhibit the same general behavior as the theory, with ET response becoming more postive (atmospheric demand dominating) as environmental VPD increases within a PFT, and more negative for PFTs that are adapted to arid conditions and prioritize water conservation over primary production.
 
-For the forest sites (DBF, ENF) the observed environmental VPD approximately straddles the critical VPD, with about half the observations below the critical VPD and half above.
-However for CSH the environmental VPD never exceeds the critical VPD, and even when uncertainty is included the sign of the derivative is almost always negative (93\% of observations).
-
-Generally speaking, our theory suggests that for a majority of enviornmental VPD conditions, plants will tend to conserve water and reduce ET with increasing VPD.
-Stomatal regulation and plant physiological response strongly regulates ET, but his regulation does vary by PFT.
-CROs are the least water conservative, while DBF, ENF, and GRA are progressively more water conservative, with CSH begin the most water conservative (most likely to reudce ET in response to increasing VPD).
-However, these results are subject to the caveat that our theory's performance is most uncertain for GRA and CRO.
-
-While our theory predicts occurrence of increasing ET with increasing VPD (atmospheric demand dominates), plant regulation is still important even when $\frac{\partial \; ET}{\partial \; VPD} > 0$.
-The magnitude of the ``sign term'' of our theory is always far below what it would be if we only considered the atmospheric demand term (PET).
-This means that any drought indices or analysis using PET will not capture how the vegetated surface respond and react to increases in atmospheric demand.
+The consequences of our theory our twofold.
+First, the ET response to VPD is always far more negative than it would be if only atmospheric demand is considered and plant phsiological reponse is ignored (as with PET).
+Therefore,  any drought indices or analysis using PET will not capture how  vegetated surfaces respond and react to increases in atmospheric demand.
 However, the new uWUE-version of Penman Monteith we derived (Equation \ref{et}) could be used as a complement to PET in drought indices.
@@
 
-This paper provides novel intuition and quantitate assessment on how the land surface responds to atmospheric drying.
-This intuition for the response of the land surface response to atmospheric drying is critical first step to disentangling land-atmosphere feedbacks at various scales, from diurnal to seasonal and beyond, and to characterize ET response in a warmer, atmospherically drier, and enriched CO$_2$ world.
+Lastly, our paper builds important intuition for how plants will respond to global change-induced shifts in VPD, which should increase in the future.
+Plant physiological responses to direct CO$_2$ effects \citep[e.g.,][]{Swann_2016} receives more attention than physiological response to indirect effects like increased VPD.
+Here, we provide broad PFT-focused results showing a likely decrease in ET in response to positive VPD perturbations (amtopsheric drying).
+Addionally, we derive an explicit conceptual framework for ET response as a function of two plant type parameters, $g_1$ and uWUE (e.g.
+Equation \ref{et}).
+This framework can be used to understand ET reponse to a range of environmental change, and forms a foundation for resolving varying respose to more nuanced consideration of plant type and climates.
+Any plant physiological heterogeneity that can be conceptualized with shifts in $g_1$ \citep[e.g.][]{Lin_2015} and/or uWUE \citep[e.g.][]{Zhou_2014} is represntable within our framework, which opens the door for a heirarchy of more sophisticated climate- and plant-specific analysis of ET sensitivity to environmental variables (including VPD).
+We argue that such simplified conceptual frameworks are critical tools for disentangling land-atmosphere feedbacks at various scales, from diurnal to seasonal and beyond, and to characterize ET response in a warmer, atmospherically drier, and enriched CO$_2$ world.
 

### 2018-02-28 doc/paper/vpd_et_paper.tex
done spell check

@@
 
-The opposing effects of increased atmospheric demand and higher stomatal stress lead to two possible perspectives for how evapotraspiration (ET) respondes to shifts in VPD.
+The opposing effects of increased atmospheric demand and higher stomatal stress lead to two possible perspectives for how evapotranspiration (ET) responds to shifts in VPD.
 The first, a hydrometeorological perspective, is that higher VPD increases atmospheric demand for water from the land surface, and this drives an increase in evapotranspiration (ET).
@@
 Each site's plant functional type (PFT) was classified using the International Geosphere-Biosphere Programme vegetation classification scheme \citep{Loveland_1999}.
-The physical constants used in the stomatal conductance model in the Methods section (Section \ref{methods}) are only published for six plant functional types (PFTs): crops (CRO), deciduous broadleaf forest (DBF), evergreed broadleaf forests (EBF), evergreen needleleaf forest (ENF), grass (GRA), and closed shrub (CSH) (see Table \ref{pft}).
+The physical constants used in the stomatal conductance model in the Methods section (Section \ref{methods}) are only published for six plant functional types (PFTs): crops (CRO), deciduous broadleaf forest (DBF), evergreen broadleaf forests (EBF), evergreen needleleaf forest (ENF), grass (GRA), and closed shrub (CSH) (see Table \ref{pft}).
 There are 62 sites with these plant functional types, and their location is shown in  Figure \ref{map_fig}.
@@
 \begin{table}
-  \caption{Plant functional types, their abbreviation, Medlyn coefficient (from \citet{Lin_2015}, via \citet{Franks_2017}) and caluclated uWUE.
+  \caption{Plant functional types, their abbreviation, Medlyn coefficient (from \citet{Lin_2015}, via \citet{Franks_2017}) and calculated uWUE.
 uWUE from \citet{Zhou_2015}, including the  observed standard deviation is shown for comparison.
@@
 A larger uWUE means a smaller $VPD_{crit}$, and an ET response to increases VPD that is more likely to be positive.
-At first glance this result is somewhat counterintuitive; we expect that plants with a higher water use effiency would be more water conservative.
+At first glance this result is somewhat counter-intuitive; we expect that plants with a higher water use efficiency would be more water conservative.
 However, in reality uWUE determines how $WUE$ changes with VPD:
@@
 So, plants with a higher uWUE will have a greater \textit{decrease} in ecosystem-scale $WUE$ in response to increases in VPD.
-This decrease in $WUE$ causes more water loss per unit carbon production, and explains the relationship between high uWUE and high liklihood of increases of ET in response to increasing atmospheric drying (increases in VPD).
+This decrease in $WUE$ causes more water loss per unit carbon production, and explains the relationship between high uWUE and high likelihood of increases of ET in response to increasing atmospheric drying (increases in VPD).
 
@@
 \label{scale_term}
-While the above discussion of the sign of $\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our question of when ET reponse increases or decreases with VPD, understating the overall magnitude of the ET response is important to soil-plant-atmosphere water budgeting.
+While the above discussion of the sign of $\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our question of when ET response increases or decreases with VPD, understating the overall magnitude of the ET response is important to soil-plant-atmosphere water budgeting.
 So we now more closely examine the terms that affect how the sign term is scaled:
@@
 $\Delta$'s presence in the scaling term also matches physical intuition.
-$\Delta$ (and also the approxmately constant $\gamma$) control the efficiencies with which surface energy is converted to latent and sensible heat \citep{Monteith_1965}.
+$\Delta$ (and also the approximately constant $\gamma$) control the efficiencies with which surface energy is converted to latent and sensible heat \citep{Monteith_1965}.
 The functional from of $\Delta$ will be the same across PFTs, but the temperature range may vary slightly.
@@
 
-To summarize, variability between PFTs is mostly controlled by systematic differences in aerodynamic conductance, due to differences in surface roughness between each PFT, and posibly to a lesser extent wind conditions.
+To summarize, variability between PFTs is mostly controlled by systematic differences in aerodynamic conductance, due to differences in surface roughness between each PFT, and possibly to a lesser extent wind conditions.
 In contrast, variability within PFT is also controlled by temperature, through Clausius-Clapeyron.
@@
 However, ET response evaluated at the average of all variables (e.g.
-$\sigma$, $T$, $c_a$, $VPD$) is postive for CRO, and negative for all other PFTs.
+$\sigma$, $T$, $c_a$, $VPD$) is positive for CRO, and negative for all other PFTs.
 This difference in mean ET response as compared to the ET response at mean environmental conditions is due to the non-linear nature of the response, in which negative responses are generally larger magnitude than positive responses (Figure \ref{idealized_sign}).
@@
 
-Regarding the frequency of negative and positive ET response, all PFTs exibit a  decreasing ET response with increasing VPD (physiologically controlled, water conservative response) for the majority of observations.
-When listed by increasing frequency of negative response, the PFTs are generally ordered as we would expect them to be, from most water intesive to most water conservative: CRO (53\%), DBF (61\%), EBF (67\%), GRA (69\%), ENF (74\%), and CSH (97\%).
+Regarding the frequency of negative and positive ET response, all PFTs exhibit a  decreasing ET response with increasing VPD (physiologically controlled, water conservative response) for the majority of observations.
+When listed by increasing frequency of negative response, the PFTs are generally ordered as we would expect them to be, from most water intensive to most water conservative: CRO (53\%), DBF (61\%), EBF (67\%), GRA (69\%), ENF (74\%), and CSH (97\%).
 
@@
 \label{testing}
-We now compare more sophisticated distributions of the observed response to our simpified theory (Section \ref{sign_term}).
+We now compare more sophisticated distributions of the observed response to our simplified theory (Section \ref{sign_term}).
 Figure \ref{test_sign} shows the observed distribution of the sign term, as compared to what the theory would predict.
@@
 
-For DBF our theory is successful towards this goal: the theory matches the leading order behavior of the function when uncertainty is included, and the observations match the theory with the addtion of noise.
+For DBF our theory is successful towards this goal: the theory matches the leading order behavior of the function when uncertainty is included, and the observations match the theory with the addition of noise.
 The VPD dependence, given by the slope parameter ($g_1$), follows the median values of each bin.
@@
 For practical reasons, our theory does not include important sources of plant type heterogeneity within and across sites for CRO.
-In agricultural landscape, the footprint of an eddy-covariance tend to not only observe a single species but many differnent ones, with potentially varying water use strategies.
+In agricultural landscape, the footprint of an eddy-covariance tend to not only observe a single species but many different ones, with potentially varying water use strategies.
 We also do not account for seasonal changes in LAI, variability in crop height and surface roughness, differences in C3 vs C4 photosynthesis and water strategies, or differences in irrigation practices between the sites and years.
@@
 GRA observations are characterized by a consistent trend back towards negative ET response at higher VPD, which the functional form of our theory is incapable of accounting for.
-As compared to CRO, the divergence between the theory and the observations is even greater for GRA, but these differences are explainable using similiar logic as we did for CRO.
+As compared to CRO, the divergence between the theory and the observations is even greater for GRA, but these differences are explainable using similar logic as we did for CRO.
 GRA also tends to exhibit substantial within-PFT variability, such as C3 vs C4 photosynthesis, as well as strong seasonal variability in height and LAI.
@@
 We note that we used published constants for $g_1$ \citep[from ][]{Lin_2015, Franks_2017}, which determines the VPD dependence, and that deviations should be expected across sites.
-We could have fit our own constants so that the theory would best match the observations, but our goal was to use a priori results whenever possible in order to avoid overfitting and to have a theoretical approach as independent to the observations as possible for further confirmation of the validity of the approach.
+We could have fit our own constants so that the theory would best match the observations, but our goal was to use a priori results whenever possible in order to avoid over-fitting and to have a theoretical approach as independent to the observations as possible for further confirmation of the validity of the approach.
 
@@
 Most of the results presented so far focus on the sign term, so now we turn to observations of ET response with the scaling term included (Figure \ref{data_scatter}).
-Until now, in the interest of developing leading order intuition for ET response, and to be conservative in our acknowledement of model and observational error, we've considered $\sigma$ variability to be a measure of uncertainty.
+Until now, in the interest of developing leading order intuition for ET response, and to be conservative in our acknowledgment of model and observational error, we've considered $\sigma$ variability to be a measure of uncertainty.
 An alternative viewpoint is that $\sigma \cdot uWUE$ represents spatial and temporal variability of uWUE, which may be expected within bounds \citep[see Table \ref{pft}, also ][]{Zhou_2015}.
@@
 ET response to VPD shifts towards positive values as VPD increases (atmospheric demand dominating).
-CRO exhibit the highest occurence of postive ET response, and the obsevations confirm that CRO are the most water intensive.
+CRO exhibit the highest occurrence of positive ET response, and the observations confirm that CRO are the most water intensive.
 CSH are the most water conservative, with a strong negative response.
-DBF, EBF, ENF, and GRA are also water conservative, but show some occurence of positive ET response to VPD, particularly at higher observed VPD, as the theory predicts.
+DBF, EBF, ENF, and GRA are also water conservative, but show some occurrence of positive ET response to VPD, particularly at higher observed VPD, as the theory predicts.
 
@@
 The sign term is most strongly scaled by $g_a$, and in general the sign term shifts towards positive values as VPD observations increase.
-The willingness with which a given plant type evolved to use water dictates the occurence of positve versus negative ET response.
+The willingness with which a given plant type evolved to use water dictates the occurrence of positive versus negative ET response.
 Water conservative plants are highly effective at mitigating the effects of atmospheric demand, and can store water for later use by reducing ET in response to increasing VPD.
@@
 
-Our theory suggests that for a majority of enviornmental conditions, plants will tend to conserve water and reduce ET with increasing VPD.
+Our theory suggests that for a majority of environmental conditions, plants will tend to conserve water and reduce ET with increasing VPD.
 Stomatal regulation and plant physiological response strongly regulate ET, and this regulation varies by PFT.
-CROs are the least water conservative, while DBF, EBF, ENF, GRA and CSH are progressively more water conservative (more likely to reudce ET in response to increasing VPD).
-Observations of ET response to VPD exhibit the same general behavior as the theory, with ET response becoming more postive (atmospheric demand dominating) as environmental VPD increases within a PFT, and more negative for PFTs that are adapted to arid conditions and prioritize water conservation over primary production.
+CROs are the least water conservative, while DBF, EBF, ENF, GRA and CSH are progressively more water conservative (more likely to reduce ET in response to increasing VPD).
+Observations of ET response to VPD exhibit the same general behavior as the theory, with ET response becoming more positive (atmospheric demand dominating) as environmental VPD increases within a PFT, and more negative for PFTs that are adapted to arid conditions and prioritize water conservation over primary production.
 
 The consequences of our theory our twofold.
-First, the ET response to VPD is always far more negative than it would be if only atmospheric demand is considered and plant phsiological reponse is ignored (as with PET).
+First, the ET response to VPD is always far more negative than it would be if only atmospheric demand is considered and plant physiological response is ignored (as with PET).
 Therefore,  any drought indices or analysis using PET will not capture how  vegetated surfaces respond and react to increases in atmospheric demand.
@@
 Plant physiological responses to direct CO$_2$ effects \citep[e.g.,][]{Swann_2016} receives more attention than physiological response to indirect effects like increased VPD.
-Here, we provide broad PFT-focused results showing a likely decrease in ET in response to positive VPD perturbations (amtopsheric drying).
-Addionally, we derive an explicit conceptual framework for ET response as a function of two plant type parameters, $g_1$ and uWUE (e.g.
+Here, we provide broad PFT-focused results showing a likely decrease in ET in response to positive VPD perturbations (atmospheric drying).
+Additionally, we derive an explicit conceptual framework for ET response as a function of two plant type parameters, $g_1$ and uWUE (e.g.
 Equation \ref{et}).
-This framework can be used to understand ET reponse to a range of environmental change, and forms a foundation for resolving varying respose to more nuanced consideration of plant type and climates.
-Any plant physiological heterogeneity that can be conceptualized with shifts in $g_1$ \citep[e.g.][]{Lin_2015} and/or uWUE \citep[e.g.][]{Zhou_2014} is represntable within our framework, which opens the door for a heirarchy of more sophisticated climate- and plant-specific analysis of ET sensitivity to environmental variables (including VPD).
+This framework can be used to understand ET response to a range of environmental change, and forms a foundation for resolving varying response to more nuanced consideration of plant type and climates.
+Any plant physiological heterogeneity that can be conceptualized with shifts in $g_1$ \citep[e.g.][]{Lin_2015} and/or uWUE \citep[e.g.][]{Zhou_2014} is representable within our framework, which opens the door for a hierarchy of more sophisticated climate- and plant-specific analysis of ET sensitivity to environmental variables (including VPD).
 We argue that such simplified conceptual frameworks are critical tools for disentangling land-atmosphere feedbacks at various scales, from diurnal to seasonal and beyond, and to characterize ET response in a warmer, atmospherically drier, and enriched CO$_2$ world.

### 2018-02-28 doc/paper/vpd_et_paper.tex
read through done, now just need to do checklist (figures, abstract, etc.)

@@
 This leads to a plant physiology perspective, in which an increase in VPD, particularly in well-watered soil conditions, may actually correspond to a decrease in ET because of stomatal closure.
-This decrease in ET would reduce and potentially cancel, in the case of full stomatal closure, the effects of shifts to atmospheric demand.
 In other words, the  question ``When does VPD drive or reduce ET?'' can be related to whether plant regulation or atmospheric demand dominates ET response.
@@
 The ET response to changes in VPD alters water partitioning between the soil and atmosphere.
-If plant response reduces ET in response to atmospheric drying then soil moisture will be better conserved.
+If plant response reduces ET with atmospheric drying then soil moisture will be better conserved.
 This would seem a sensible evolutionary strategy to cope with aridity.
@@
 ][]{Or_2013}, increased atmospheric aridity would strongly reduce soil moisture \citep{Berg_2017}.
-In turn, this would further increase aridity as low soil moisture levels increase the Bowen ratio, and in turn increase temperature and atmospheric drying \citep[][]{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006, Salvucci_2013, Gentine_2016, Berg_2016}.
+In turn, this would further increase aridity as low soil moisture levels increase the Bowen ratio, and cause increased temperature and atmospheric drying \citep[][]{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006, Salvucci_2013, Gentine_2016, Berg_2016}.
 This however would not seem to be a sensible strategy for plants from an evolutionary standpoint.
@@
 The objective of the present manuscript is to evaluate the VPD dependence of ET, in non-extreme soil drought conditions.
-The goal of this paper is to use reasonable approximations as a tool to drive intuition for plant response to atmospheric drying.
+The goal of this paper is to use reasonable approximations as a tool to develop intuition for plant response to atmospheric drying.
 This intuition will aid interpretation of observations and full complexity models.
@@
 
-The purpose of this study is to examine ecosystem response to atmospheric drying focusing on the growing season.
+The purpose of this study is to examine ecosystem response to atmospheric drying, focusing on the growing season.
 To accomplish this, we filter and quality control the data using a similar procedure as \cite{Zhou_2015}:
@@
 \end{linenomath*}
-where $\Delta$ is the change in saturation vapor pressure with temperature, given by Clausius-Clapeyron ($\frac{d \; e_s}{d \; T}$), $R$ is the net radiation minus ground heat flux, $g_a$ is aerodynamic conductance, $\rho_a$ is air density, $c_p$ is specific heat of air at constant pressure, $\gamma$ is the psychometric constant, and $g_s$ is the stomatal conductance (Table \ref{definitions}).
+where $\Delta$ is the change in saturation vapor pressure with temperature, given by Clausius-Clapeyron ($\frac{d \; e_s}{d \; T}$), $R_{net}$ is the net radiation minus ground heat flux, $g_a$ is aerodynamic conductance, $\rho_a$ is air density, $c_p$ is specific heat of air at constant pressure, $\gamma$ is the psychometric constant, and $g_s$ is the stomatal conductance (Table \ref{definitions}).
 
@@
 \end{linenomath*}
-is relatively constant across time and moisture conditions within a plant functional type, in non-extreme drought conditions.
+is relatively constant across time and moisture conditions within a plant functional type.
 If, following \citet{Lin_2015}, we approximate $g_0$ as $0$ (i.e.
@@
   \begin{equation}
-    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{uWUE \; ET}{c_a \; \sqrt{VPD}}
+    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{uWUE \; ET}{c_a \; \sqrt{VPD}}.
     \label{new_g_s}
@@
 
-Given the eddy-covariance FLUXNET2015 data described in Section \ref{data}, every term in our new version of PM (Equation \ref{et}) is observed, except for the two parameters $g_1$, and uWUE.
+Given eddy-covariance FLUXNET2015 data (Section \ref{data}), every term in our new version of PM (Equation \ref{et}) is observed, except for the two parameters $g_1$, and uWUE.
 $g_1$ has been measured and reported for the PFTs considered here \citep{Lin_2015, Franks_2017}, and we fit uWUE by calculating its expectation, given the model and FLUXNET2015 data.
@@
 Thus, with this uncertainty analysis we can evaluate departure from our theory in observations, as a departure of $\sigma$ from unity.
-The variability of $\sigma$ across sites and time provides a measure of uncertainty in our model, assumptions, as well as in the FLUXNET201 observations themselves.
+The variability of $\sigma$ across sites and time provides a measure of uncertainty in our model, assumptions, as well as in the FLUXNET2015 observations themselves.
 The variability of $\sigma$ then propagates through any uncertainty to our derivative of Equation \ref{et_sigma}:
@@
 By construction, the variability in the $\sigma$ term (Equation \ref{sigma}) contains all model and observational uncertainties.
-For an observation that perfectly matches our model and constant uWUE assumption $\sigma$ will be one.
+For an observation that perfectly matches our model and constant uWUE assumption, $\sigma$ will be one.
 Therefore, for our assumptions and framework to be reasonable $\sigma$ should be close to 1.
@@
 All variables are positive, so the relative magnitude between the first term and the second term in the sign term (Equation \ref{sign}) will determine whether ET increases or decreases with increasing VPD.
-If the second term is larger then the plant control dominates and ET decreases with increasing VPD.
+If the second term is larger then plant control dominates and ET decreases with increasing VPD.
 However, if the first term is larger, then atmospheric demand dominates and ET increases with increasing VPD.
@@
 We can further determine a critical threshold separating an increase from a decrease in ET, i.e.
-the threshold $VPD_{crit}$ such as the derivative vanishes $\frac{\partial \; ET}{\partial \; VPD} = 0$:
+the threshold $VPD_{crit}$ such that the derivative vanishes $\frac{\partial \; ET}{\partial \; VPD} = 0$:
 \small
@@
   \begin{equation}
-VPD_{crit} = \frac{R_{air}}{4 c_p} \left( \frac{\gamma c_a}{1.6\; R \;  uWUE} + \sqrt{\frac{\gamma c_a}{1.6\; R \;  uWUE}\left( \frac{\gamma c_a}{1.6\; R \;  uWUE} + 8 g_1 \frac{c_p}{R_{air}}\right)} - 4 g_1 \frac{c_p}{R_{air}} \right)
+VPD_{crit} = \frac{R_{air}}{4 c_p} \left( \frac{\gamma c_a}{1.6\; R \;  uWUE} + \sqrt{\frac{\gamma c_a}{1.6\; R \;  uWUE}\left( \frac{\gamma c_a}{1.6\; R \;  uWUE} + 8 g_1 \frac{c_p}{R_{air}}\right)} - 4 g_1 \frac{c_p}{R_{air}} \right),
 \label{vpd_min_et}
@@
 \normalsize
-We note that $VPD_{crit}$ mostly depends on the PFT parameters uWUE and $g_1$ but only vary weakly with climate as most other parameters related to the environment are nearly constant.
-The value of $VPD_{crit}$ for each PFT is shown in Table \ref{vpd_crit}.
+noting that $VPD_{crit}$ mostly depends on the PFT parameters uWUE and $g_1$, and only varies weakly with climate as most other parameters related to the environment are nearly constant.
+The calculated value of $VPD_{crit}$ for each PFT is shown in Table \ref{vpd_crit}.
 For any values of $VPD$ less than $VPD_{crit}$, ET will thus decrease with increasing VPD ($\frac{\partial \; ET}{\partial \; VPD} < 0$), and for values of $VPD$ greater than $VPD_{crit}$, ET will increase with increasing VPD ($\frac{\partial \; ET}{\partial \; VPD} > 0$).
 In other words, ecosystems regulate and mitigate evaporative losses up to the VPD limit, $VPD_{crit}$, above which atmospheric demand is just too high to be entirely compensated by stomatal and ecosystem regulation.
-We note however that even though ET increases again above the critical threshold, $VPD_{crit}$, ET is still much lower that potential evaporation as stomata are still strongly regulating vapor fluxes tot he atmosphere.
+We note however that even though ET increases again above the critical threshold, $VPD_{crit}$, ET is still much lower that potential evaporation as stomata are still strongly regulating vapor fluxes to the atmosphere.
 Indeed, even in the absence of soil pore evaporation, stomata do not shut down entirely at very high VPD so that ET does not go to zero, because stomata are still slightly open to perform some photosynthesis \citep{Ball_1987, Leuning_1990, MEDLYN_2011}.
@@
 
-A tendency towards increasing ET response with increasing VPD can also be caused by a high slope parameter ($g_1$), characteristic of plants that at the leaf scale are more willing to trade water for stomata opening.
+A tendency towards increasing ET response with increasing VPD can also be caused by a high slope parameter ($g_1$), characteristic of plants that at the leaf scale are more willing to trade water for access to atmospheric CO$_2$.
 Plants that are less conservative will be thus be more likely to increase ET with increasing VPD.
@@
 These differences in PFT parameters cause opposite ET responses to changes in VPD between CRO and CSH.
-ET theoretically always decreases with increasing VPD for the more water conservative CSH, while ET will frequently  increases with increasing VPD for the more water intensive CRO (Figure \ref{idealized_sign}).
+ET theoretically always decreases with increasing VPD for the more water conservative CSH, while ET frequently  increases with increasing VPD for the more water intensive CRO (Figure \ref{idealized_sign}).
 CRO evolved or were bred to prioritize GPP and yield and are thus not water conservative.
@@
 
-The control of the scaling term variability between PFTs by aerodynamic conductance is confirmed by the data (Figure \ref{scale_vary}).
+The control of the scaling term variability between PFTs by aerodynamic conductance is confirmed by data (Figure \ref{scale_vary}).
 Differences between PFT are almost entirely due to differences in aerodynamic conductance, rather than differences in observed temperature ranges.
@@
 We now compare more sophisticated distributions of the observed response to our simplified theory (Section \ref{sign_term}).
-Figure \ref{test_sign} shows the observed distribution of the sign term, as compared to what the theory would predict.
+The observed distribution of the sign term, as compared to what the theory would predict, is provided in Figure \ref{test_sign}.
 Our goal was to capture the leading order behavior of the ET dependence on VPD.
@@
 While CSH has a much different functional form of the sign term than DBF, CSH observations also match our theory to leading order.
-Again, the VPD dependence mostly determined from the slope parameter ($g_1$) closely matches the medians in the observation bins and the theoretical threshold is well reproduced.
+Again, the VPD dependence mostly determined from the slope parameter ($g_1$) closely matches the medians in the observation bins and the theoretical threshold is well reproduced (albeit off-scale).
 The VPD-independent, strongly negative response, which is determined by uWUE, is also captured.
@@
 For practical reasons, our theory does not include important sources of plant type heterogeneity within and across sites for CRO.
-In agricultural landscape, the footprint of an eddy-covariance tend to not only observe a single species but many different ones, with potentially varying water use strategies.
+In agricultural landscape, the footprint of an eddy-covariance towers tend to not only observe a single species but many different ones, with potentially varying water use strategies.
 We also do not account for seasonal changes in LAI, variability in crop height and surface roughness, differences in C3 vs C4 photosynthesis and water strategies, or differences in irrigation practices between the sites and years.
@@
 GRA also tends to exhibit substantial within-PFT variability, such as C3 vs C4 photosynthesis, as well as strong seasonal variability in height and LAI.
-As with CRO, the trend back towards negative ET response at higher VPD could be explained by a superposition of sites with C3 vs C4 sites.
+As with CRO, the trend back towards negative ET response at higher VPD could be explained by a superposition of sites with C3 vs C4 plants.
 Unfortunately, because detailed site-level data of C3 vs C4 plant type is not available, we could not bin CRO and GRA further by irrigation and photosynthesis strategy.
@@
 However, the theory still captures the general behavior and magnitude of the response, with some increase in error at low and/or high VPD.
-We note that we used published constants for $g_1$ \citep[from ][]{Lin_2015, Franks_2017}, which determines the VPD dependence, and that deviations should be expected across sites.
+We note that we used published constants for $g_1$ \citep[from ][]{Lin_2015, Franks_2017}, which determines the VPD dependence, and that deviations should be expected across sites \citep{Lin_2015}.
 We could have fit our own constants so that the theory would best match the observations, but our goal was to use a priori results whenever possible in order to avoid over-fitting and to have a theoretical approach as independent to the observations as possible for further confirmation of the validity of the approach.
@@
 So the PET response to atmospheric dryness is outside the range of all observable responses.
-This casts doubt on PET's usefulness for capturing the leading order behavior of water budget and dryness response at the land surface, such as used in several drought indices.
+This casts doubt on PET's usefulness for capturing the leading order behavior of water budget and dryness response at the land surface, as is done in several drought indices.
 However our theory (Equation \ref{et}) could be used with published uWUE and $g1$ data, as an alternative to PET to estimating land response and dryness, and would more accurately account for the response to atmospheric drying.
@@
 An alternative viewpoint is that $\sigma \cdot uWUE$ represents spatial and temporal variability of uWUE, which may be expected within bounds \citep[see Table \ref{pft}, also ][]{Zhou_2015}.
-This is a less conservative view; some of the $\sigma \cdot uWUE$ variability will be due to model and observational error so by viewing $\sigma \cdot uWUE$ as \textit{real} variability we run the risk of mistaking noise for signal.
+This is a less conservative view; some of the $\sigma \cdot uWUE$ variability will be due to model and observational error, so by viewing $\sigma \cdot uWUE$ as \textit{real} variability we run the risk of mistaking noise for signal.
 However, the advantage of this viewpoint is that, from \citet{Zhou_2014}, we have very high confidence that uWUE fit at the hourly timescale (as we do with $\sigma \cdot uWUE$) correctly captures the relationship between ET, GPP, and VPD, and that our form of PM introduces minimal error with its use of uWUE.
@@
 As in Figure \ref{test_sign}, Figure \ref{data_scatter} matches our expectation based on simplified theory.
-The sign term is most strongly scaled by $g_a$, and in general the sign term shifts towards positive values as VPD observations increase.
+The sign term is most strongly scaled by $g_a$, and in general the sign term shifts towards positive values as VPD increases.
 The willingness with which a given plant type evolved to use water dictates the occurrence of positive versus negative ET response.
@@
 Clearly, for GRA soil water frequently impacts plant function and alters ET response, and our theory is limited.
-It is therefore not suprising that our theory tested poorly againt the data for GRA, relative to ENF, DBF, and GRA (Section \ref{testing}).
+It is therefore not suprising that our theory tested poorly againt the data for GRA, relative to ENF, EBF, DBF, and CSH (Section \ref{testing}).
 For all other PFTs occurence of soil mosture impacts was rare enough to not manifest itself in bulk statistics and plots.
@@
 Equation \ref{et}).
-This framework can be used to understand ET response to a range of environmental change, and forms a foundation for resolving varying response to more nuanced consideration of plant type and climates.
+This framework can be used to understand ET response to a range of environmental change, and forms a foundation for resolving varying response to more nuanced consideration of plant type and climate.
 Any plant physiological heterogeneity that can be conceptualized with shifts in $g_1$ \citep[e.g.][]{Lin_2015} and/or uWUE \citep[e.g.][]{Zhou_2014} is representable within our framework, which opens the door for a hierarchy of more sophisticated climate- and plant-specific analysis of ET sensitivity to environmental variables (including VPD).
@@
 The FLUXNET eddy covariance data processing and harmonization was carried out by the European Fluxes Database Cluster, AmeriFlux Management Project, and Fluxdata project of FLUXNET, with the support of CDIAC and ICOS Ecosystem Thematic Center, and the OzFlux, ChinaFlux and AsiaFlux offices.
+This material is based upon work supported by the National Science Foundation Graduate Research Fellowship under Grant No.
+DGE 16-44869.
 
-% AGU does not want a .bib or a .bbl file. Please copy in the contents of your .bbl file here.
 \bibliography{references.bib}
-% \input{vpd_et_paper.bbl}
-
-% \listofchanges
 

### 2018-03-01 doc/paper/vpd_et_paper.tex
done with abstract

@@
 
-\title{When does vapor pressure deficit drive or reduce evapotranspiration?}
+\title{When does vapor pressure deficit drive or reduce evapotranspiration?\footnote{Primary Research Article}}
 
@@
 \begin{keypoints}
-\item one
-\item two
+\item hydrometeorology, evapotranspiration, aridity, vapor pressure deficit, Clausius-Clapeyron relation, plant physiology, stomatal response
 \end{keypoints}
@@
 \begin{abstract}
-= enter abstract here =
+  Increases to vapor pressure deficit (VPD) and atmospheric demand for water are expected with rising CO$_2$.
+While increased evapotranspiration (ET) in response to increased atmospheric demand seems intuitive, plants are capable of reducing ET in response to increased VPD by closing their stomata, in an effot to conserve water.
+Here we examine which effect dominates response to increasing VPD: atmospheric demand and increases in ET, or plant physiological response (stomata closure) and decreases in ET.
+We use Penman-Monteith, combined with semi-empirical optimal stomatal regulation theory and underlying water use efficiency, to develop a theoretical framework for understanding how ET responds to increases in VPD.
+  The theory suggests that for most environmental conditions and plant types, plant physiological response dominates and ET decreases with increasing VPD.
+Plants that are evolved or bred to prioritize primary production over water conservation (e.g.
+crops) exhibit a higher likelihood of atmospheric demand-driven response (ET increasing).
+However for forest, grass, and shrub plant types, ET more frequently decreases than increases with rising VPD.
+Therefore, we may expect reductions in ET due to CO$_2$-induced increases in VPD.
+This work serves as an example of the utility of our simplified framework for disentangling land-atmosphere feedbacks at multiple scales, including the characterization of ET response in an atmospherically drier, enriched CO$_2$ world.
+
 \end{abstract}

### 2018-03-01 doc/paper/vpd_et_paper.tex
paper ready to submit

@@
 \begin{abstract}
-  Increases to vapor pressure deficit (VPD) and atmospheric demand for water are expected with rising CO$_2$.
-While increased evapotranspiration (ET) in response to increased atmospheric demand seems intuitive, plants are capable of reducing ET in response to increased VPD by closing their stomata, in an effot to conserve water.
+  Increases to vapor pressure deficit (VPD) and atmospheric demand for water are expected with rising atmospheric CO$_2$.
+While increased evapotranspiration (ET) in response to increased atmospheric demand seems intuitive, plants are capable of reducing ET in response to increased VPD by closing their stomata, in an effort to conserve water.
 Here we examine which effect dominates response to increasing VPD: atmospheric demand and increases in ET, or plant physiological response (stomata closure) and decreases in ET.
@@
 \begin{table}
-  \caption{Plant functional types, their abbreviation, Medlyn coefficient (from \citet{Lin_2015}, via \citet{Franks_2017}) and calculated uWUE.
+  \caption{Plant functional types, their abbreviation, Medlyn coefficient (from \citet{Lin_2015}, via \citet{Franks_2017}), and calculated uWUE.
 uWUE from \citet{Zhou_2015}, including the  observed standard deviation is shown for comparison.
-Note that uWUE from \citet{Zhou_2015} is calculated from a different set of sites, and that units are converted such that the quantities fit into Equations 1-8 with the variables in Table \ref{definitions}.}
+Note that uWUE from \citet{Zhou_2015} is calculated from a different set of sites, and that units are converted such that the quantities work with Equations 1-8 with the variables defined Table \ref{definitions}.}
     \small
@@
 \includegraphics[width=\textwidth]{./map.pdf}
-\caption{Plant functional type and location of sites used in analysis. ***\textbf{This is just  a placeholder for now, I am currently fixing it***}}
+\caption{Plant functional type and location of FLUXNET2015 sites used in this analysis.}
 \label{map_fig}
@@
 \begin{table}
-
-  \caption{Definition of symbols and variables, with citation for calculation if applicable.}
+  \caption{Definition of symbols and variables, with citation for how values are calculated, if applicable.}
   \label{definitions}
@@
 \caption{The joint distribution of $VPD$ and $\sigma$, with outliers removed (defined as lowest and highest 5\% of $\sigma$).
-$\sigma$ has only a weak dependence on $VPD$, and $\sigma$ is $O(1)$ for the bulk of thte observations.}
+$\sigma$ exhibits a weak dependence on $VPD$, and $\sigma$ is $O(1)$ for the bulk of the observations.}
 \label{joint_vpd_sigma}
@@
 \includegraphics[width=\textwidth]{./idealized_sign.pdf}
-\caption{Sources of variability for the sign term.
-Top: sign term as a function of VPD, with $\sigma$ held fixed at 1.
-The observed range of VPD for each PFT is also shown below the x-axis.
-Line extent corresponds to 5th and 95th percentiles, while stars denote the location of the 25th, 50th, and 75th percentiles.
-
-Bottom: The location of the minima of ET, as a function of VPD and $\sigma$.
-Lines and stars denote the distribution of VPD and $\sigma$ next to each axis, following the same percentiles as above.
-\textbf{***I think I'm going to remove the bottom plot: its discussion didn't fit in well with the text, and the box plots test inclusion of uncertainty***}}
+\caption{The functional form of the sign term, with $\sigma$ held fixed at 1, and all terms except VPD set to PFT averages.
+For comparison, the observed range of VPD for each PFT is plotted below the x-axis.
+Stars denote 25th, 50th, and 75th percentiles, and the range of the line spans the 5th-95th percentiles of observed VPD.
+Vertical black lines denote the location of $VPD_{crit}$ for each PFT, with the exception of CSH, for which $VPD_{crit}$ is off-scale.}
 \label{idealized_sign}
@@
 \caption{Primary sources of variability for the scaling term, as a function of PFT.
-Note the change in axis scale for each plot.
-The 5th-95th percentile range of temperature is plotted at the 5th, 25th, 50th, 75th, and 95th percentiles of aerodynamic conductance, as observed for each PFT's sites.}
+The 5th-95th percentile range of temperature is plotted at the 5th, 25th, 50th, 75th, and 95th percentiles of aerodynamic conductance, as observed for each PFT.}
 \label{scale_vary}
@@
 \includegraphics[width=\textwidth]{./test_sign.pdf}
-\caption{Comparison of the sign term with model uncertainty included (box plots) to the sign term as calculated with simplifying assumptions (blue line).}
+\caption{Comparison of the sign term with model uncertainty included (box plots) to the sign term as calculated with simplifying assumptions (blue line, as in Figure \ref{idealized_sign}).
+Each box plot corresponds to 5\% of the data, and the 5-95\% range of VPD is plotted.}
 \label{test_sign}
@@
 \centerline{\includegraphics[width=\textwidth]{./data_scatter.png}}
-\caption{Scatter plots of observed $\frac{\partial \; ET}{\partial \; VPD}$, including $\sigma$ variability, as a function of PFT. Please note differences in the colorbar scale.}
+\caption{Scatter plots of observed $\frac{\partial \; ET}{\partial \; VPD}$, including $\sigma$ variability, as a function of PFT, temperature and VPD. Please note differences in the colorbar scale.}
 \label{data_scatter}
@@
 \includegraphics[width=\textwidth]{./swc_boxplot.pdf}
-\caption{$\sigma \cdot uWUE$ as a function of PFT and SWC. Each boxplot represents 5\% of all observations.}
+\caption{$\sigma \cdot uWUE$ as a function of PFT and SWC. Each box plot represents 5\% of all observations.}
 \label{swc_boxplot}
@@
 DGE 16-44869.
+All code and data used in this analysis, including those used to generate the figures and tables, is publicly available at https://github.com/massma/climate\_et.
 

### 2018-03-01 doc/paper/vpd_et_paper.tex
about to submit

@@
 
-\journalname{Global Change Biology}
+\journalname{Global Change Biology: When does VPD drive or reduce ET?}
 
@@
 \begin{keypoints}
-\item hydrometeorology, evapotranspiration, aridity, vapor pressure deficit, Clausius-Clapeyron relation, plant physiology, stomatal response
+\item ecohydrology, evapotranspiration, aridity, vapor pressure deficit, Clausius-Clapeyron relation, plant physiology, stomatal response
 \end{keypoints}

### 2018-03-19 doc/paper/vpd_et_paper.tex
rejected by GCB, need to do a better job communicating how my approach is new (enabled by medlyn and zhou). preparing for AFM submission now
reformulated g1/gs as ecosystem scale following Medlyn 2017. Fortunately (as expected) doesn't really change story, but will make ecosystem scale derivation much cleaner

@@
 % 11/23/2015
-\documentclass[draft,linenumbers]{gcbjournal}
+\documentclass[draft,linenumbers]{agujournal}
 % \draftfalse
@@
 
-\journalname{Global Change Biology: When does VPD drive or reduce ET?}
+\journalname{Agricultural and Forest Meteorology: When does VPD drive or reduce ET?}
 
@@
 
-\title{When does vapor pressure deficit drive or reduce evapotranspiration?\footnote{Primary Research Article}}
+\title{When does vapor pressure deficit drive or reduce evapotranspiration?}
 
-\authors{Adam Massmann\affil{1}, Pierre Gentine\affil{1}, Changjie Lin\affil{1,2}}
+\authors{Adam Massmann\affil{a}, Pierre Gentine\affil{a}, Changjie Lin\affil{a,b}}
 
-\affiliation{1}{Department of Earth and Environmental Engineering, Columbia University, New York, NY 10027}
-\affiliation{2}{State Key Laboratory of Hydroscience and Engineering, Department of Hydraulic Engineering, Tsinghua University, Beijing, CN 100084}
+\affiliation{a}{Department of Earth and Environmental Engineering, Columbia University, New York, NY 10027}
+\affiliation{b}{State Key Laboratory of Hydroscience and Engineering, Department of Hydraulic Engineering, Tsinghua University, Beijing, CN 100084}
 

### 2018-03-19 doc/paper/vpd_et_paper.tex
finished new introduction

@@
 This perspective is particularly relevant because potential evapotranspiration (PET), which is used in many drought indices and hydrometeorological studies \citep[e.g.,][]{Heim_2002, Scheff_2015}, typically only quantifies changes in atmospheric demand and fails to account for ecosystem response \citep{Swann_2016}.
-In reality, plants' stomata have evolved to optimally regulate the exchange of water and carbon, and tend to partially close in response to increased atmospheric dryness \citep{Ball_1987, Leuning_1990, MEDLYN_2011}.
+In reality, plants' stomata have evolved to optimally regulate the exchange of water and carbon, and tend to partially close in response to increased atmospheric dryness \citep{Farquhar_1978, Ball_1987, Leuning_1990, MEDLYN_2011}.
 This leads to a plant physiology perspective, in which an increase in VPD, particularly in well-watered soil conditions, may actually correspond to a decrease in ET because of stomatal closure.
@@
 This intuition will aid interpretation of observations and full complexity models.
-In order to quantify plant response to perturbations in VPD, we apply a Penman-Monteith framework to derive the theoretical response of ET to VPD.
-The model is then validated and tested at multiple eddy-covariance stations spanning various climates and plant functional types.
+In the past, similiar approaches were used to understand interactions between stomatal conductance and evapotranspiration \citep[e.g.,][]{Jarvis_1986, Mcnaughton_1991}.
+However, at the time researchers' understandng of the form of VPD's effect on plant physiology was limited, so they could not explore the sensitivity of ET to VPD, including VPD's effect on stomatal conductance and plant function.
+
+Recent results have drastically improved our understanding of VPD's impact on physiology.
+\citet{MEDLYN_2011} developed a model for leaf-scale stomatal conductance ($g_s$), including VPD response, by combining an optimal photosynthesis theory \citep{Farquhar_1980, Katul_2010} with an empirical approach, and extended this model to the ecosystem scale in \citet{Medlyn_2017}.
+Additionally, \citet{Zhou_2014} demonstrated that a new quantity $uWUE$ ($\frac{GPP\; \sqrt{vpd}}{ET}$) properly captures a constant relationship between GPP, ET, and VPD over a diurnal cycle.
+uWUE is also remarkably well conserved in the growing season across space and time, within a PFT \citep{Zhou_2015}.
+These results provide novell tools to examine how ET varies in response to VPD, while explicitly including VPD's effect on stoamtal conductance.
+In this manuscript, we derive the theoretical response of ET to VPD using a Penman Monteith framework.
+Our theory is validated and tested at multiple eddy-covariance stations spanning various climates and plant functional types.
 

### 2018-03-20 doc/paper/vpd_et_paper.tex
calling it a night, just have rewriting and address reviewer 1's comments

@@
 This intuition will aid interpretation of observations and full complexity models.
-In the past, similiar approaches were used to understand interactions between stomatal conductance and evapotranspiration \citep[e.g.,][]{Jarvis_1986, Mcnaughton_1991}.
+In the past, similiar approaches were used to understand interactions between stomatal conductance, evapotranspiration and the environment \citep[e.g.,][]{Jarvis_1986, Mcnaughton_1991}.
 However, at the time researchers' understandng of the form of VPD's effect on plant physiology was limited, so they could not explore the sensitivity of ET to VPD, including VPD's effect on stomatal conductance and plant function.
@@
 uWUE is also remarkably well conserved in the growing season across space and time, within a PFT \citep{Zhou_2015}.
-These results provide novell tools to examine how ET varies in response to VPD, while explicitly including VPD's effect on stoamtal conductance.
+These results provide novell tools to examine how ET varies in response to VPD, while explicitly including VPD's effect on stoamtal conductance for the first time.
 In this manuscript, we derive the theoretical response of ET to VPD using a Penman Monteith framework.

### 2018-03-20 doc/paper/vpd_et_paper.tex
some refactoring, also now with medlyn 2017 we can calcualte our own g1, which lets us do analysis on all pfts

@@
 Recent results have drastically improved our understanding of VPD's impact on physiology.
-\citet{MEDLYN_2011} developed a model for leaf-scale stomatal conductance ($g_s$), including VPD response, by combining an optimal photosynthesis theory \citep{Farquhar_1980, Katul_2010} with an empirical approach, and extended this model to the ecosystem scale in \citet{Medlyn_2017}.
-Additionally, \citet{Zhou_2014} demonstrated that a new quantity $uWUE$ ($\frac{GPP\; \sqrt{vpd}}{ET}$) properly captures a constant relationship between GPP, ET, and VPD over a diurnal cycle.
+\citet{MEDLYN_2011} developed a model for leaf-scale stomatal conductance ($g_s$), including VPD response, by combining an optimal photosynthesis theory \citep{Farquhar_1980, Katul_2010} with an empirical approach, and extended use of this model to the ecosystem scale in \citet{Medlyn_2017}.
+Additionally, \citet{Zhou_2014} demonstrated that a quantity $uWUE$ ($\frac{GPP\; \sqrt{vpd}}{ET}$) properly captures a constant relationship between GPP, ET, and VPD over a diurnal cycle.
 uWUE is also remarkably well conserved in the growing season across space and time, within a PFT \citep{Zhou_2015}.
 These results provide novell tools to examine how ET varies in response to VPD, while explicitly including VPD's effect on stoamtal conductance for the first time.
-In this manuscript, we derive the theoretical response of ET to VPD using a Penman Monteith framework.
+In this manuscript, we leverage these tools to derive the theoretical response of ET to VPD using a Penman Monteith framework.
 Our theory is validated and tested at multiple eddy-covariance stations spanning various climates and plant functional types.

### 2018-03-20 src/codebase/data_io.py

@@ -96,11 +96,11 @@ def load_file(filename):
     data_out['g1'] = WUE_MEDLYN.loc[pft, 'g1M']
     data_out['g1b'] = WUE_MEDLYN.loc[pft, 'g1b']
   except KeyError:
-    data_out['g1'] = np.nan
-    data_out['g1b'] = np.nan
+    data_out['g1'] = 1.0 #np.nan
+    data_out['g1b'] = 1.0 #np.nan
     print('error, no medlyn coeffieent for %s, pft: %s, setting to none'\
           % (filename, pft))
-    return None
+    #return None
   try:
     data_out['uwue_zhou'] = WUE.loc[pft, 'u_wue_yearly']
     data_out['uwue_zhou_std'] = WUE.loc[pft, 'u_wue_yearly_std']

### 2018-03-20 src/codebase/test_functions.py

@@ -35,7 +35,7 @@ def test_et_model(_df):
 def test_et_models(_df):
   """calcs histograms of c.v for all models"""
   plt.close('all')
-  for name in ['uwue', 'uwue_bb', 'iwue_bb', 'iwue']:
+  for name in ['uwue', 'iwue']: #['uwue', 'uwue_bb', 'iwue_bb', 'iwue']:
     plt.figure()
     sns.distplot(_df[name]/_df[name].mean())
     plt.xlabel("cv - %s" % name) #"(calc'ed uwue)/(zhou's uwue)")

### 2018-03-20 src/paper_figure_code/data_scatter.py

@@ -4,7 +4,7 @@ This script makes fig 6
 """
 from shared_functions import *
 import time
-
+mpl.rcParams.update(small_ax_params)
 # mean_df is defined up at this level
 local_fontsize = fontsize+3
 
@@ -76,9 +76,8 @@ def make_ax_plot(_ax, var, _df, meta):
   # _ax.yaxis.set_tick_params(labelsize=local_fontsize-3)
   # _ax.xaxis.set_tick_params(labelsize=local_fontsize-3)
   cbar = plt.colorbar(color, ax=_ax)# , ax=_ax2)
-  
-  if ((_df.pft.iloc[0] == 'EBF') | (_df.pft.iloc[0] == 'CSH')\
-      | (_df.pft.iloc[0] == 'GRA')):
+  if ((_df.pft.iloc[0] == 'MF') | (_df.pft.iloc[0] == 'WSA')\
+      | (_df.pft.iloc[0] == 'SAV')):
     cbar.set_label(meta['title'], fontsize=local_fontsize)
   y_lim = _ax.get_ylim()
   # vpd_crit = et_min_vpd(mean_df.loc[_df.pft.iloc[0], :])
@@ -90,7 +89,7 @@ def make_ax_plot(_ax, var, _df, meta):
   _ax.set_ylim(ylim)
   return
 
-def scatter_plot_paper(_df, meta):
+def scatter_plot_paper(_df, ax, meta):
   """
   creates scatter of derivatives wrt to VPD, assumes Delta(vpd) = 1.0 Pa
   """
@@ -103,7 +102,7 @@ def scatter_plot_paper(_df, meta):
 
   meta['vmax'] = np.nanmax(np.absolute([var.mean()+1.5*var.std(),\
                                           var.mean()-1.5*var.std()]))
-  make_ax_plot(meta['ax'], var, _df, meta)
+  make_ax_plot(ax, var, _df, meta)
 
   return
 
@@ -116,21 +115,7 @@ meta['sample'] = '' # 'sampled'
 # much faster if bins smaller, useful for testing
 meta['nbins'] = 1000
 
-def scatter_wrapper(_df):
-  """wrapper that groups by pft and does scaling plot"""
-  fig = plt.figure()
-  fig.set_figheight(fig.get_figheight()*3)
-  fig.set_figwidth(fig.get_figwidth()*2)
-  for i, pft in enumerate(pft_order):
-    print(pft)
-    meta['ax'] = fig.add_subplot(3, 2, i+1)
[…]

### 2018-03-20 src/paper_figure_code/map.py

@@ -4,7 +4,7 @@ This script makes fig 01 for the paper
 """
 from shared_functions import *
 
-symbols = ['o', 'v', '>', 's', '<', 'D']
+symbols = ['o', 'v', '>', 's', '<', 'D', 'X', 'P', '8']
 plt.close('all')
 def make_map(_df):
   """makes a map given a _df with Lat, Lon, and Cover"""

### 2018-03-20 src/paper_figure_code/shared_functions.py

@@ -18,6 +18,13 @@ import pickle
 
 mpl.use('Pdf')
 mpl.rcParams.update(mpl.rcParamsDefault)
+small_ax_params = mpl.rcParamsDefault
+# increase tick label size
+label_size = 14
+small_ax_params['xtick.labelsize'] = label_size
+small_ax_params['ytick.labelsize'] = label_size
+
+
 importlib.reload(util)
 importlib.reload(plot_tools)
 importlib.reload(d_calc)
@@ -45,7 +52,6 @@ def et_min_vpd(_df, uwue=None):
       sqrt_vpd = np.nan
   return sqrt_vpd**2
 
-
 def get_pft(_df):
   return _df['pft'].iloc[0]
 
@@ -53,13 +59,15 @@ def d_et(_df):
   """returns d et/d vpd given _df"""
   return d_calc.scaling(_df)*d_calc.sign(_df)
 
-
 name_dict = {'CRO': 'Crops (CRO)',\
              'DBF': 'Deciduous Forest (DBF)',
              'EBF': 'Evergreen Broadleaf Forest (EBF)',
              'ENF': 'Evergreen Needleleaf Forest (ENF)',
              'GRA': 'Grass (GRA)',
-             'CSH': 'Closed Shrub (CSH)'}
+             'CSH': 'Closed Shrub (CSH)',
+             'WSA': 'Woody Savannah (WSA)',
+             'MF': 'Mixed Forest (MF)',
+             'SAV': 'Savannah (SAV)'}
 
 fontsize=16
 
@@ -72,17 +80,30 @@ def custom_ylabel(_df, ax, label):
 
 def custom_xlabel(_df, ax, label):
   """labels x label for axis ordering"""
-  if (_df.pft.iloc[0] == 'CRO') | (_df.pft.iloc[0] == 'GRA'):
+  if (_df.pft.iloc[0] == 'CRO') | (_df.pft.iloc[0] == 'GRA')\
+     | (_df.pft.iloc[0] == 'SAV'):
     ax.set_xlabel(label, fontsize=fontsize)
   return
 
 mean_df['vpd_crit'] = et_min_vpd(mean_df)
 
 # pft order for panel plots
-pft_order = ['DBF', 'EBF', 'ENF', 'CSH', 'CRO', 'GRA',]
+pft_order = ['DBF', 'EBF', 'MF', 'ENF', 'CSH', 'WSA', 'CRO', 'GRA', 'SAV']
 
[…]

### 2018-03-20 src/paper_figure_code/tables.py

@@ -192,11 +192,11 @@ print(mean_df.uwue_zhou)
 # import codebase.plot_tools as plot_tools
 # df.groupby('pft').apply(plot_tools.histogram, meta)
 
-# fig = plt.figure()
-# ax = fig.add_subplot(111)
-# for g1, uwue, label in zip(mean_df.g1, mean_df.uwue, mean_df.index):
-#   ax.scatter(g1, uwue, label=label)
-# ax.set_xlabel('g1')
-# ax.set_ylabel('uwue')
-# plt.legend(loc='best')
-# plt.savefig('%s/climate_et/g1_uwue_relationship.png' % os.environ['PLOTS'])
+fig = plt.figure()
+ax = fig.add_subplot(111)
+for g1, uwue, label in zip(mean_df.g1, mean_df.uwue, mean_df.index):
+  ax.scatter(g1, uwue, label=label)
+ax.set_xlabel('g1')
+ax.set_ylabel('uwue')
+plt.legend(loc='best')
+plt.savefig('%s/climate_et/g1_uwue_relationship.png' % os.environ['PLOTS'])

### 2018-03-21 doc/paper/vpd_et_paper.tex
updated table 1 to write directly from code, should have done this long ago

@@
   \hline
-  DBF & deciduous broadleaf forest & 140.7 & 2.63 & 3.12 $\pm$ 0.52 \\
-  EBF & evergreen broadleaf forest & 130.3 & 2.56 &  N/A \\
-  ENF & evergreen needleleaf forest & 74.3 & 3.45 & 3.30 $\pm$ 0.91 \\
-  CRO & cropland & 183.1 & 2.27 & 3.80 $\pm$ 1.01 \\
-  CSH & closed shrub & 148.6 & 1.66 & 2.18 $\pm$ 0.44 \\
-  GRA & grassland (assumed C3) & 166.0 & 1.92 & 2.68 $\pm$ 0.61 \\
+\input{table_1}
   \hline

### 2018-03-21 src/paper_figure_code/tables.py

@@ -16,8 +16,27 @@ def frequency(_df):
 
 #### Table 1, pft tables
 print('\n\n******TABLE 1*****')
-print(mean_df.loc[:, ['g1', 'uwue', 'uwue_zhou', 'uwue_zhou_std']])
+mean_df['g1-dot-uwue'] = mean_df.g1*mean_df.uwue
+print(mean_df.loc[:, ['g1', 'uwue', 'uwue_zhou', 'uwue_zhou_std']])#,\
+                      #'g1-dot-uwue']])
+def write_table_1(_df):
+  """write output of table 1"""
+  fh = open("../../doc/paper/table_1.tex", "w")
+  for pft in _df.index:
+    row = _df.loc[pft, :]
+    if (row.uwue_zhou == 1.0) & (row.uwue_zhou_std == 1.0):
+      fh.write("%s & %s & %.2f & %.2f & N/A \\\\ \n"
+               % (pft, name_dict[pft][:-5], np.round(row.g1, 2),
+                  np.round(row.uwue, 2)))
+    else:
+      fh.write("%s & %s & %.2f & %.2f & %.2f $\pm$ %.2f \\\\ \n"
+               % (pft, name_dict[pft][:-5], np.round(row.g1, 2),
+                  np.round(row.uwue, 2), np.round(row.uwue_zhou, 2),
+                  np.round(row.uwue_zhou_std, 2)))
 
+  fh.close()
+
+write_table_1(mean_df)
 
 #### Table 2, pft tables
 print('\n\n******TABLE 3*****')
@@ -39,6 +58,8 @@ mean_df['et_std'] = std['et_obs']
 counts = df.groupby('pft').apply(frequency)
 
 mean_df['counts'] = counts
+
+
 # test = mean_df.loc[:, ['d_et', 'd_et_bar', 'd_et_bar_std',\
 #                        'd_et_bar_norm', 'counts']]
 

### 2018-03-21 src/paper_figure_code/test_sign.py

@@ -51,13 +51,13 @@ def plot_box(_df, ax):
   #\ uwue=mean_row.uwue_zhou),\
   ax.plot(vpd, d_calc.sign(mean_row, vpd=vpd),
           'b-', linewidth=bluelinewidth)
-  # ax.plot(vpd, d_calc.sign(mean_row, vpd=vpd, uwue=(mean_row.uwue_zhou\
-  #                                             +mean_row.uwue_zhou_std)),\
-  #         'b-', linewidth=1.0)
-  # ax.plot(vpd, d_calc.sign(mean_row, vpd=vpd,\
-  #                               uwue=(mean_row.uwue_zhou\
-  #                                     -mean_row.uwue_zhou_std)),\
-  #         'b-', linewidth=1.0)
+  ax.plot(vpd, d_calc.sign(mean_row, vpd=vpd, uwue=(mean_row.uwue_zhou\
+                                              +mean_row.uwue_zhou_std)),\
+          'b-', linewidth=1.0)
+  ax.plot(vpd, d_calc.sign(mean_row, vpd=vpd,\
+                                uwue=(mean_row.uwue_zhou\
+                                      -mean_row.uwue_zhou_std)),\
+          'b-', linewidth=1.0)
   #_df.groupby(pd.cut(_df.vpd, bins=20)).apply(add_box, ax=ax)
   #_df.boxplot(column='sign', by=pd.cut(_df.vpd, bins=20), ax=ax)
   custom_ylabel(_df, ax, paren_string)

### 2018-03-21 doc/paper/vpd_et_paper.tex
tables created

@@
 \citet{MEDLYN_2011} developed a model for leaf-scale stomatal conductance ($g_s$), including VPD response, by combining an optimal photosynthesis theory \citep{Farquhar_1980, Katul_2010} with an empirical approach, and extended use of this model to the ecosystem scale in \citet{Medlyn_2017}.
-Additionally, \citet{Zhou_2014} demonstrated that a quantity $uWUE$ ($\frac{GPP\; \sqrt{vpd}}{ET}$) properly captures a constant relationship between GPP, ET, and VPD over a diurnal cycle.
+Additionally, \citet{Zhou_2014} demonstrated that a quantity underlying water use efficiency $\left(uWUE = \frac{GPP\; \sqrt{VPD}}{ET}\right)$ properly captures a constant relationship between GPP, ET, and VPD over a diurnal cycle.
 uWUE is also remarkably well conserved in the growing season across space and time, within a PFT \citep{Zhou_2015}.
@@
   \hline
-\input{table_1}
+\input{pft_params}
   \hline
@@
   \hline
-CRO & 288.6 & 376.0 & 65.2 &   \textbf{936.6} \\
-CSH & 289.0 & 383.6 & 67.5 & \textbf{12852.5} \\
-DBF & 288.6 & 379.5 & 63.6 &  \textbf{1266.6} \\
-EBF & 288.4 & 368.5 & 61.6 &  \textbf{1454.8} \\
-ENF & 288.28 & 379.8 & 61.0 &  \textbf{1853.7} \\
-GRA & 288.4 & 378.0 & 61.0 &  \textbf{3261.1} \\
+  \input{vpd_crit.tex}
 \hline
@@
 \begin{figure}
-\centering
-\includegraphics[width=\textwidth]{./idealized_scale.pdf}
+%\centering
+\centerline{\includegraphics[width=1.4\textwidth]{./idealized_scale.pdf}}
 \caption{Primary sources of variability for the scaling term, as a function of PFT.
@@
   \hline
-  CRO & -0.05 & 0.008 &  0.533 \\
-  DBF & -0.10 & -0.013 &  0.609 \\
-  EBF & -0.13 & -0.037 &  0.666 \\
-  GRA & -0.07 & -0.024 &  0.694 \\
-  ENF & -0.17 & -0.081 &  0.740 \\
-  CSH & -0.29 & -0.224 &  0.973 \\
+  \input{stats.tex}
   \hline
@@
 \centering
-\includegraphics[width=\textwidth]{./test_sign.pdf}
+\centerline{\includegraphics[width=1.4\textwidth]{./test_sign.pdf}}
 \caption{Comparison of the sign term with model uncertainty included (box plots) to the sign term as calculated with simplifying assumptions (blue line, as in Figure \ref{idealized_sign}).
@@
 \centering
-\centerline{\includegraphics[width=\textwidth]{./data_scatter.png}}
+\centerline{\includegraphics[width=1.4\textwidth]{./data_scatter.png}}
 \caption{Scatter plots of observed $\frac{\partial \; ET}{\partial \; VPD}$, including $\sigma$ variability, as a function of PFT, temperature and VPD. Please note differences in the colorbar scale.}
@@
 \centering
-\includegraphics[width=\textwidth]{./swc_boxplot.pdf}
+\centerline{\includegraphics[width=1.4\textwidth]{./swc_boxplot.pdf}}
 \caption{$\sigma \cdot uWUE$ as a function of PFT and SWC. Each box plot represents 5\% of all observations.}

### 2018-03-21 src/paper_figure_code/data_scatter.py

@@ -100,8 +100,8 @@ def scatter_plot_paper(_df, ax, meta):
 
   var = _df['d_et']
 
-  meta['vmax'] = np.nanmax(np.absolute([var.mean()+1.5*var.std(),\
-                                          var.mean()-1.5*var.std()]))
+  meta['vmax'] = np.nanmax(np.absolute([var.mean()+1.0*var.std(),\
+                                          var.mean()-1.0*var.std()]))
   make_ax_plot(ax, var, _df, meta)
 
   return

### 2018-03-22 doc/paper/vpd_et_paper.tex
finished methods and materials, but un-proofread

@@
 In reality, plants' stomata have evolved to optimally regulate the exchange of water and carbon, and tend to partially close in response to increased atmospheric dryness \citep{Farquhar_1978, Ball_1987, Leuning_1990, MEDLYN_2011}.
-This leads to a plant physiology perspective, in which an increase in VPD, particularly in well-watered soil conditions, may actually correspond to a decrease in ET because of stomatal closure.
+This leads to a plant physiology perspective, in which an increase in VPD, particularly in well-watered soil conditions, may actually correspond to a decrease in ET because of stomatal closure \citep[e.g.][]{Rigden_2017}.
 In other words, the  question ``When does VPD drive or reduce ET?'' can be related to whether plant regulation or atmospheric demand dominates ET response.
@@
 We can use intuition about plant water conservation strategy to hypothesize about ET response to changes in VPD.
-Plants that evolved to conserve water, such as arid shrubs, should be more likely to reduce ET with increasing VPD, and plants that have evolved or have been engineered to care little about water, such as crops, will be more likely to increase ET with increasing VPD.
+Plants and ecosystems that evolved to conserve water, such as arid shrubs or savannah, should be more likely to reduce ET with increasing VPD, and plants that have evolved or have been engineered to care little about water, such as crops, will be more likely to increase ET with increasing VPD.
 Atmospheric conditions must matter as well.
@@
 
-The objective of the present manuscript is to evaluate the VPD dependence of ET, in non-extreme soil drought conditions.
-The goal of this paper is to use reasonable approximations as a tool to develop intuition for plant response to atmospheric drying.
-This intuition will aid interpretation of observations and full complexity models.
+The objective of the present manuscript is to use reasonable approximations as a tool to develop intuition for plant response to atmospheric drying and evaluate the VPD dependence of ET.
+This intuition will aid interpretation of observations, full complexity models, and facilitate the disentanglement of complex land-atmosphere feedbacks.
 In the past, similiar approaches were used to understand interactions between stomatal conductance, evapotranspiration and the environment \citep[e.g.,][]{Jarvis_1986, Mcnaughton_1991}.
@@
 \section{Materials and Methods}
-\label{methods}
 \subsection{Data}
 \label{data}
-We use both meteorological and eddy-covariance data from the FLUXNET2015 database, including all sites with at least four years of data (data available at \sloppy https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/ \sloppy).
-Each site's plant functional type (PFT) was classified using the International Geosphere-Biosphere Programme vegetation classification scheme \citep{Loveland_1999}.
-The physical constants used in the stomatal conductance model in the Methods section (Section \ref{methods}) are only published for six plant functional types (PFTs): crops (CRO), deciduous broadleaf forest (DBF), evergreen broadleaf forests (EBF), evergreen needleleaf forest (ENF), grass (GRA), and closed shrub (CSH) (see Table \ref{pft}).
-There are 62 sites with these plant functional types, and their location is shown in  Figure \ref{map_fig}.
-
-\begin{table}
-  \caption{Plant functional types, their abbreviation, Medlyn coefficient (from \citet{Lin_2015}, via \citet{Franks_2017}), and calculated uWUE.
-uWUE from \citet{Zhou_2015}, including the  observed standard deviation is shown for comparison.
-Note that uWUE from \citet{Zhou_2015} is calculated from a different set of sites, and that units are converted such that the quantities work with Equations 1-8 with the variables defined Table \ref{definitions}.}
-    \small
-\label{pft}
-\centering
-\begin{tabular}{l c c @{\qquad} c c}
-  \hline
-  \multirow{2}[3]{*}{Abbreviation} & \multirow{2}[3]{*}{PFT} & \multirow{2}[3]{*}{$g_1$ (Pa$^{0.5}$)} & \multicolumn{2}{c}{uWUE ($\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET])}  \\
-  \cmidrule{4-5}
-
-  & & & fitted & \citet{Zhou_2015} \\
-
-  \hline
-\input{pft_params}
-  \hline
-\end{tabular}
-\end{table}
+We use both meteorological and eddy-covariance data from the FLUXNET2015 database (data available at \sloppy https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/ \sloppy), including all sites with at least four years of data and observations of the variables described in the methods section (Section \ref{methods}).
+Seventy-three sites met these requirements, and were grouped into nine plant functional types (PFT) according to the International Geosphere-Biosphere Programme vegetation classification scheme \citep{Loveland_1999}: cropland (CRO), grass (GRA), deciduous broadleaf forest (DBF), evergreen broadlead forest (EBF), evergreen needleleaf forest (ENF), mixed forest (MF), closed shrub (CSH), savannah (SAV), and woody savannah (WSA) (site locations in Figure \ref{map_fig}).
 
@@
 Timesteps with negative observed GPP or ET are also removed, and we aggregate half hourly data to hourly averages to reduce noise \citep{Lin_2018}.
-After these quality control procedures, 335,939 upscaled hourly observations remain.
+After these quality control procedures, 434,765 upscaled hourly observations remain.
 
@@
   \begin{equation}
-  g_{l-s} = g_0 + 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{A}{c_a}
+  g_{s-leaf} = g_0 + 1.6 \left(1 + \frac{g_{1-leaf}}{\sqrt{VPD}}\right) \frac{A}{c_a}
   \end{equation}
 \end{linenomath*}
-Where $g_1$ is a leaf-scale ``slope'' parameter, A is the net CO$_2$ assimilation rate, and $c_a$ is the atmospheric CO$_2$ concentration at the leaf surface.
-\cite{MEDLYN_2011} relate the slope parameter ($g_1$) to physical parameters as:
+Where $g_{1-leaf}$ is a leaf-scale ``slope'' parameter, A is the net CO$_2$ assimilation rate, and $c_a$ is the atmospheric CO$_2$ concentration at the leaf surface.
+\cite{MEDLYN_2011} relate the slope parameter ($g_{1-leaf}$) to physical parameters as:
 \begin{linenomath*}
@@
   \begin{equation}
-  g_1 = \sqrt{\frac{3 \, \Gamma^* \, \lambda}{1.6}},\footnote{Note this expression has units of of (mmol mol$^{-1}$)$^{1/2}$, but this can be converted to Pa$^{1/2}$ using the ideal gas law.}
+  g_{1-leaf} = \sqrt{\frac{3 \, \Gamma^* \, \lambda}{1.6}},\footnote{Note this expression has units of of (mmol mol$^{-1}$)$^{1/2}$, but this can be converted to Pa$^{1/2}$ using the ideal gas law.}
   \end{equation}
@@
 where $\Gamma^*$ is the CO$_2$ compensation point for photosynthesis (without dark respiration), and $\lambda$ is the marginal water cost of leaf carbon ($\frac{\partial \; \text{transpiration}}{\partial \; A}$).
-So, $g_1$ is a leaf-scale term reflecting the tradeoff of water for carbon uptake and growth.
-The higher $g_1$, the more open the stomata and the more they release water in exchange for carbon.
+So, $g_{1-leaf}$ is a leaf-scale term reflecting the tradeoff of water for carbon uptake and growth.
+The higher $g_{1-leaf}$, the more open the stomata and the more they release water in exchange for carbon.
 
-The Medlyn model for stomatal conductance has been shown to behave very well across PFTs \citep[][]{Lin_2015}, and can be  adapted to the ecosystem scale by multiplying $g_{l-s}$ by leaf area index (LAI) and converting units to m s$^{-1}$ with the ideal gas law:
+The Medlyn model for stomatal conductance has been shown to behave very well across PFTs \citep[][]{Lin_2015}, and has been successfully adopted to ecosystem scale analysis in \citet{Medlyn_2017}.
+In units of m s$^{-1}$, the ecosystem scale stomatal conductance is:
 
@@
   \begin{equation}
-    g_s = \text{LAI} \frac{R \,T}{P} \left( g_0 + 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{A}{c_a}\right)
+    g_s = \frac{R \,T}{P} \; 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{GPP}{c_a},
       \label{medlyn}
@@
 
-While Equation 3 can be used in PM (Equation 1), it will make analytical work with the function intractable because $A$, net CO$_2$ assimilation rate, is functionally related to ET itself.
-To remove the dependence of ET on $A$ we use fundamental semi-empirical results of \citet{Zhou_2015}, who showed that the underlying Water Use Efficiency (uWUE):
+where GPP is the ecoysystem scale gross primary production, and $g_1$ is an ecosystem scale analogue to $g_{1-leaf}$.
+We solve for $g_1$ following \citet{Medlyn_2017} (Equation 5), and take the median $g_1$ value to be representative of each PFT (Table \ref{pft}).
+
+\begin{table}
+  \caption{Plant functional types, their abbreviation, calculated Medlyn coefficient, and calculated uWUE.
+uWUE from \citet{Zhou_2015}, including the  observed standard deviation is shown for comparison.
+Note that uWUE from \citet{Zhou_2015} is calculated from a different set of sites, and that units are converted such that the quantities work with Equations 1-8 and the variables defined Table \ref{definitions}.}
+    \small
+\label{pft}
+\centering
+\begin{tabular}{l c c @{\qquad} c c}
+  \hline
+  \multirow{2}[3]{*}{Abbreviation} & \multirow{2}[3]{*}{PFT} & \multirow{2}[3]{*}{$g_1$ (Pa$^{0.5}$)} & \multicolumn{2}{c}{uWUE ($\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET])}  \\
+  \cmidrule{4-5}
+
+  & & & fitted & \citet{Zhou_2015} \\
+
+  \hline
+\input{pft_params}
+  \hline
+\end{tabular}
+\end{table}
+
+While Equation 3 can be used in PM (Equation 1), it will make analytical work with the function intractable because $GPP$ is functionally related to ET itself.
+Additionally, a perturbation to VPD should induce a physiological plant response that will alter GPP and cause an indirect change in stomatal conductance, in addition to the direct effect of VPD in Equation \ref{medlyn}.
+Therefore, in order to derive the response of ET to VPD, we must account for the functional relationship between GPP, ET, and VPD and its effect on stomatal conductance.
+We can use aforementioned semi-empirical results of \citet{Zhou_2015} as a tool to approach this problem.
+\citet{Zhou_2015}, showed that the underlying Water Use Efficiency (uWUE):
 
@@
 \end{linenomath*}
-is relatively constant across time and moisture conditions within a plant functional type.
-If, following \citet{Lin_2015}, we approximate $g_0$ as $0$ (i.e.
-we neglect cuticular and epidermal losses - a reasonable assumption except in very dry conditions), we can use uWUE to remove the $A$ dependence from $g_s$ in a way that makes the PM equation analytically tractable:
+
+is relatively constant across time and moisture conditions within a plant functional type, and correctly captures a constant relationship between GPP, ET and VPD over a diurnal cycle \citep{Zhou_2014}.
+We can use uWUE to remove the $GPP$ dependence from $g_s$ in a way that makes the PM equation analytically tractable:
 
@@
 
-Note that uWUE is fit on the ecosystem scale as in \citet{Zhou_2015}, so GPP in Equation \ref{uwue} is really $A \cdot \text{LAI}$.
-This leads to the cancellation of LAI in the step from Equation \ref{medlyn} to Equation \ref{new_g_s}.
-Plugging Equation \ref{new_g_s} into Equation \ref{orig_pen} and rearranging gives a new explicit expression for PM, in which dependence on $A$ is removed:
+Plugging Equation \ref{new_g_s} into Equation \ref{orig_pen} and rearranging gives a new explicit expression for PM, in which dependence on $GPP$ is removed:
 
@@
 
-With Equation \ref{et}, ET is explicitly a function of environmental variables and two plant-specific constants, the slope parameter ($g_1$), and uWUE, both reflecting water conservation strategy.
-The slope parameter is a leaf-scale term related to the willingness of stomata to trade water for CO$_2$ and to keep stomata open.
+With Equation \ref{et} we have derived for the first time, using recent results \citep[][]{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017}, ET explicitly as function of environmental variables and two plant-specific constants, the slope parameter ($g_1$), and uWUE, both reflecting water conservation strategy.
+The slope parameter is related to the willingness of stomata to trade water for CO$_2$ and to keep stomata open.
 uWUE is a semi-empirical ecosystem-scale constant related to how WUE changes with VPD (specifically $VPD^{-1/2}$).
@@
 
-Given eddy-covariance FLUXNET2015 data (Section \ref{data}), every term in our new version of PM (Equation \ref{et}) is observed, except for the two parameters $g_1$, and uWUE.
-$g_1$ has been measured and reported for the PFTs considered here \citep{Lin_2015, Franks_2017}, and we fit uWUE by calculating its expectation, given the model and FLUXNET2015 data.
+Given eddy-covariance FLUXNET2015 data (Section \ref{data}), every term in our new version of PM (Equation \ref{et}) is observed, except for uWUE, which we fit by calculating its expectation, given the model and FLUXNET2015 data.
 
@@
 
-With Equation \ref{d_et} we have an analytical framework for ecosystem response to atmospheric demand perturbations, with an implicit assumption of non-extreme drought conditions.
+With Equation \ref{d_et} we have an analytical framework for ecosystem response to atmospheric demand perturbations, with an implicit assumption of non-extreme drought conditions associated with our constant $uWUE$ assumption (e.g.
+we might expect $uWUE$ to decrease in extreme drought).
 There are a few subtleties to taking the derivative in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$) and $VPD$ are functionally related, so while taking the derivative we evaluate $\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET} {\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD} \Big|_{\text{RH fixed}} + \frac{\partial \; ET}{\partial \; RH} \frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$ fixed}}$.
@@
 We note one final comment on our derivation which is relevant for drought indices.
-If we approximate $c_a$ at a global mean CO$_2$ concentration, then the RHS of Equation \ref{et} is fully defined using commonly available weather station data and the constants published in \citet{Zhou_2015} and \citet{Lin_2015}.
+If we approximate $c_a$ at a global mean CO$_2$ concentration, then the RHS of Equation \ref{et} is fully defined using commonly available weather station data and the constants published here.
 This makes Equation \ref{et} a useful alternative to PET in drought indices and hydrometeorological analysis for vegetated surfaces.
-Equation \ref{et} better reflects the physics of water exchange at the land surface and would only require fitting of uWUE and $g_1$, and can also account for changes in CO$_2$ concentration, contrary to typical drought indices such as the Palmer Drought Severity Index (PDSI) \citep{Swann_2016}.
+Equation \ref{et} better reflects the physics of water exchange at the land surface and would only require fitting of uWUE and $g_1$, and can also account for changes in CO$_2$ concentration (assuming some relationship between $\lambda$ and [CO$_2$]), contrary to typical drought indices such as the Palmer Drought Severity Index (PDSI) \citep{Swann_2016}.
 

### 2018-03-22 src/paper_figure_code/test_sign.py

@@ -51,13 +51,13 @@ def plot_box(_df, ax):
   #\ uwue=mean_row.uwue_zhou),\
   ax.plot(vpd, d_calc.sign(mean_row, vpd=vpd),
           'b-', linewidth=bluelinewidth)
-  ax.plot(vpd, d_calc.sign(mean_row, vpd=vpd, uwue=(mean_row.uwue_zhou\
-                                              +mean_row.uwue_zhou_std)),\
-          'b-', linewidth=1.0)
-  ax.plot(vpd, d_calc.sign(mean_row, vpd=vpd,\
-                                uwue=(mean_row.uwue_zhou\
-                                      -mean_row.uwue_zhou_std)),\
-          'b-', linewidth=1.0)
+  # ax.plot(vpd, d_calc.sign(mean_row, vpd=vpd, uwue=(mean_row.uwue_zhou\
+  #                                             +mean_row.uwue_zhou_std)),\
+  #         'b-', linewidth=1.0)
+  # ax.plot(vpd, d_calc.sign(mean_row, vpd=vpd,\
+  #                               uwue=(mean_row.uwue_zhou\
+  #                                     -mean_row.uwue_zhou_std)),\
+  #         'b-', linewidth=1.0)
   #_df.groupby(pd.cut(_df.vpd, bins=20)).apply(add_box, ax=ax)
   #_df.boxplot(column='sign', by=pd.cut(_df.vpd, bins=20), ax=ax)
   custom_ylabel(_df, ax, paren_string)

### 2018-03-26 doc/paper/vpd_et_paper.tex
finished idealized sign and scaling sections

@@
   \begin{equation}
-  g_{s-leaf} = g_0 + 1.6 \left(1 + \frac{g_{1-leaf}}{\sqrt{VPD}}\right) \frac{A}{c_a}
+  g_{s-leaf} = g_0 + 1.6 \left(1 + \frac{g_{1-leaf}}{\sqrt{VPD}}\right) \frac{A}{c_a},
   \end{equation}
 \end{linenomath*}
-Where $g_{1-leaf}$ is a leaf-scale ``slope'' parameter, A is the net CO$_2$ assimilation rate, and $c_a$ is the atmospheric CO$_2$ concentration at the leaf surface.
+where $g_{1-leaf}$ is a leaf-scale ``slope'' parameter, A is the net CO$_2$ assimilation rate, and $c_a$ is the atmospheric CO$_2$ concentration at the leaf surface.
 \cite{MEDLYN_2011} relate the slope parameter ($g_{1-leaf}$) to physical parameters as:
@@
 where $\Gamma^*$ is the CO$_2$ compensation point for photosynthesis (without dark respiration), and $\lambda$ is the marginal water cost of leaf carbon ($\frac{\partial \; \text{transpiration}}{\partial \; A}$).
-So, $g_{1-leaf}$ is a leaf-scale term reflecting the tradeoff of water for carbon uptake and growth.
+So, $g_{1-leaf}$ is a leaf-scale term reflecting the tradeoff of water for carbon uptake.
 The higher $g_{1-leaf}$, the more open the stomata and the more they release water in exchange for carbon.
@@
 Additionally, a perturbation to VPD should induce a physiological plant response that will alter GPP and cause an indirect change in stomatal conductance, in addition to the direct effect of VPD in Equation \ref{medlyn}.
-Therefore, in order to derive the response of ET to VPD, we must account for the functional relationship between GPP, ET, and VPD and its effect on stomatal conductance.
+Therefore, in order to derive the response of ET to VPD, we must account for the functional relationship between GPP, ET, and VPD, and its effect on stomatal conductance.
 We can use aforementioned semi-empirical results of \citet{Zhou_2015} as a tool to approach this problem.
@@
 uWUE is a semi-empirical ecosystem-scale constant related to how WUE changes with VPD (specifically $VPD^{-1/2}$).
-Is is also roughly proportional to physical constants:
+It is also roughly proportional to physical constants:
 
@@
 To account for observational error, as well as model uncertainty (e.g.
-temporal and spatial variations of uWUE), we introduce an uncertainty parameter $\sigma$ modifying uWUE:
+temporal and spatial variations of uWUE and $g_1$), we introduce an uncertainty parameter $\sigma$ modifying uWUE:
 
@@
 The calculated value of $VPD_{crit}$ for each PFT is shown in Table \ref{vpd_crit}.
-For any values of $VPD$ less than $VPD_{crit}$, ET will thus decrease with increasing VPD ($\frac{\partial \; ET}{\partial \; VPD} < 0$), and for values of $VPD$ greater than $VPD_{crit}$, ET will increase with increasing VPD ($\frac{\partial \; ET}{\partial \; VPD} > 0$).
+For any values of $VPD$ less than $VPD_{crit}$, ET will decrease with increasing VPD ($\frac{\partial \; ET}{\partial \; VPD} < 0$), and for values of $VPD$ greater than $VPD_{crit}$, ET will increase with increasing VPD ($\frac{\partial \; ET}{\partial \; VPD} > 0$).
 In other words, ecosystems regulate and mitigate evaporative losses up to the VPD limit, $VPD_{crit}$, above which atmospheric demand is just too high to be entirely compensated by stomatal and ecosystem regulation.
@@
 So, plants with a higher uWUE will have a greater \textit{decrease} in ecosystem-scale $WUE$ in response to increases in VPD.
-This decrease in $WUE$ causes more water loss per unit carbon production, and explains the relationship between high uWUE and high likelihood of increases of ET in response to increasing atmospheric drying (increases in VPD).
+This decrease in $WUE$ causes more water loss per unit carbon gain, and explains the relationship between high uWUE and high likelihood of increases of ET in response to increasing atmospheric drying (increases in VPD).
 
@@
 This effect of uWUE and $g1$ on the sign term is most apparent by comparing two extreme PFTs: water intensive crops (CRO) and water conservative closed shrub (CSH).
-CRO has high slope parameters and uWUE ($g_1 = 183.1$ Pa$^{1/2}$; $2.27$ $\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET]) compared to CSH ($g1 = 148.6 \, Pa^{1/2}$, $uWUE=1.66$ $\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET]).
+CRO has higher slope parameter and a slightyly higher uWUE ($g_1 = 140.7$ Pa$^{1/2}$; $2.85$ $\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET]) compared to CSH ($g1 = 75.1 \, Pa^{1/2}$, $uWUE=2.82$ $\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET]).
 These differences in PFT parameters cause opposite ET responses to changes in VPD between CRO and CSH.
@@
 Plants that are more conservative (small $g_1$) will tend to reduce ET with increasing VPD, and will be very effective at reducing ET, especially at low VPD.
-However, at very high VPD, gradients in vapor pressure at the leaf scale will be so strong that they will begin to dominate leaf-scale plant strategies of stomatal (partial) closure (parameterized with $g_1$).
-As a result, ET response will begin to asymptote towards its ecosystem-scale values irregardless of leaf-scale strategy and stomatal regulation (parameterized with $g_1$).
+However, at very high VPD, gradients in vapor pressure at the leaf scale will become very strong as stomata reach their limits of closure in response to VPD (parameterized with $g_1$).
+As a result, ET response will begin to asymptote towards a constant ecosystem-scale values as leaf-scale response to VPD asymptotes towards zero.
 Therefore, plants with a low $g_1$ will have the largest VPD dependence of ET response because the difference in ET response at low VPD (leaf stomatal response dominates) and high VPD (VPD gradient dominates) is largest.
-This is apparent in the strong VPD dependence of ENF, which has the lowest slope parameter ($g_1=74.31$ Pa$^{1/2}$) (Figure \ref{idealized_sign}).
+This is apparent in the strong VPD dependence of CSH, which has the lowest slope parameter ($g_1=75.1$ Pa$^{1/2}$) (Figure \ref{idealized_sign}).
 
-To summarize our theoretical insights (Figure \ref{idealized_sign} and Table \ref{vpd_crit}), CROs are the least water conservative and have the strongest overall tendency to increase ET with increasing VPD, while CSH are the most water conservative and have the strongest tendency to decrease ET with increasing VPD.
-ENF ($g_1 = 74.31$ Pa$^{1/2}$) has by far the largest VPD response dependence, while CRO ($g_1 = 183.1$ Pa$^{1/2}$) has the smallest VPD dependence.
+To summarize our theoretical insights (Figure \ref{idealized_sign} and Table \ref{vpd_crit}), CROs are the least water conservative and have the strongest overall tendency to increase ET with increasing VPD, while CSH are the most water conservative and have the strongest tendency to decrease ET with increasing VPD, as well as the strongest VPD dependence of response.
 Figure \ref{idealized_sign} clearly shows, according to our theory, that for all PFTs except for crops there is frequent occurrence of a negative (plant dominating) ET response to increases in VPD.
@@
 For CRO on the other hand, VPD is higher than VPD$_{crit}$ for more than 50\% of observations, emphasizing that those plants operate with an aggressive water usage strategy, are water intensive and were actually engineered for photosynthesis rather than water saving.
-For forests and grass,  more than half of the observed VPD are less than VPD$_{crit}$, i.e.
+For DBF, EBF, MF, GRA, and SAV  more than half of the observed VPD are less than VPD$_{crit}$, i.e.
 in conditions where plant response dominates.
-It is also important to note that even when atmospheric demand dominates, ET response to VPD is still far more negative than it would be for potential evaporation $\partial PET/\partial VPD$, i.e.
+SAV has a more water conservative response than the forest, grass, and crop plan types, but still respondes by increasing ET with increasing VPD for about a quarter of observations, due to the high aridity (VPD) of the SAV ecoclimate.
+It is also important to note that for all PFTs, even when atmospheric demand dominates, ET response to VPD is still far more negative than it would be for potential evaporation $\partial PET/\partial VPD$, i.e.
 atmospheric demand only, emphasizing that there is still a strong regulation of evaporative flux by stomata and though the plant xylem.
@@
 Differences between PFT are almost entirely due to differences in aerodynamic conductance, rather than differences in observed temperature ranges.
-The scaling term for the tree PFTs (DBF, EBF, ENF) is generally about double the scaling terms for the other PFTs which have lower surface roughness and generally smaller aerodynamic conductance.
+The scaling term for the tree PFTs (DBF, EBF, ENF, MF) is generally about double the scaling terms for other PFTs which have lower surface roughness and generally smaller aerodynamic conductance (GRA, CSH, CRO).
+The savannah (WSA, SAV) PFT's scaling is somewhere betwen GRA, CSH, and CRO, and DBF, EBF, ENF, and MF, due to higher variability and surface roughness.
 

### 2018-03-26 src/paper_figure_code/idealized_sign.py

@@ -45,7 +45,7 @@ def pft_leaf(_df, axs):
     p = axs[0].plot(vpd, d_calc.sign(mean_row, vpd=vpd, uwue=uwue),\
                     label="%s: $uWUE$=%4.2f, g1=%4.1f"\
                     % (_df.pft.iloc[0],\
-                       uwue,  _df.g1.iloc[0]))
+                       np.round(uwue, 2),  np.round(_df.g1.iloc[0], 1)))
 
   ptiles = np.array([_df.vpd.quantile(q=_p/100.)\
                      for _p in [25., 50., 75.]])

### 2018-03-26 src/paper_figure_code/shared_functions.py

@@ -23,7 +23,8 @@ small_ax_params = mpl.rcParamsDefault
 label_size = 14
 small_ax_params['xtick.labelsize'] = label_size
 small_ax_params['ytick.labelsize'] = label_size
-
+small_ax_fontsize = 19
+fontsize=16
 
 importlib.reload(util)
 importlib.reload(plot_tools)
@@ -69,16 +70,16 @@ name_dict = {'CRO': 'Crops (CRO)',\
              'MF': 'Mixed Forest (MF)',
              'SAV': 'Savannah (SAV)'}
 
-fontsize=16
 
-def custom_ylabel(_df, ax, label):
+
+def custom_ylabel(_df, ax, label, fontsize=fontsize):
   """labels y label for axis ordering"""
   if (_df.pft.iloc[0] == 'ENF') | (_df.pft.iloc[0] == 'DBF')\
      | (_df.pft.iloc[0] == 'CRO'):
     ax.set_ylabel(label, fontsize=fontsize)
   return
 
-def custom_xlabel(_df, ax, label):
+def custom_xlabel(_df, ax, label, fontsize=fontsize):
   """labels x label for axis ordering"""
   if (_df.pft.iloc[0] == 'CRO') | (_df.pft.iloc[0] == 'GRA')\
      | (_df.pft.iloc[0] == 'SAV'):

### 2018-03-26 doc/paper/vpd_et_paper.tex
done with everything but discussion

@@
 Regarding the frequency of negative and positive ET response, all PFTs exhibit a  decreasing ET response with increasing VPD (physiologically controlled, water conservative response) for the majority of observations.
-When listed by increasing frequency of negative response, the PFTs are generally ordered as we would expect them to be, from most water intensive to most water conservative: CRO (53\%), DBF (61\%), EBF (67\%), GRA (69\%), ENF (74\%), and CSH (97\%).
-
-In general the bulk statistics match our theoretical expectations well, with the caveat that inclusion of uncertainty shifts crops towards a slightly more negative response than the theory and Figure \ref{idealized_sign} would suggest.
+The more water conservative PFTs generally exhibit higher frequency of negative ET response, espcially when one factors in the distribution of environmental VPD (e.g.
+SAV and WSA grow in more arid climates, Figure \ref{idealized_sign}).
+In general the bulk statistics match our theoretical expectations well, with the caveat that inclusion of uncertainty shifts crops towards a slightly more negative response, and shifts many of the other PFTs, which still exhibit a high frequency of negative ET reponse, towards more frequent occurence of postive ET response than the theory and Figure \ref{idealized_sign} would suggest.
 The bulk statistics motivate a more thorough examination of the structure of uncertainty and more sophisticated validation of our theory's performance against observations.
@@
 
-For DBF our theory is successful towards this goal: the theory matches the leading order behavior of the function when uncertainty is included, and the observations match the theory with the addition of noise.
+This is particularly true for DBF and MF; the theory matches the leading order behavior of the function when uncertainty is included, and the observations match the theory with the addition of noise.
 The VPD dependence, given by the slope parameter ($g_1$), follows the median values of each bin.
@@
 
-While CSH has a much different functional form of the sign term than DBF, CSH observations also match our theory to leading order.
-Again, the VPD dependence mostly determined from the slope parameter ($g_1$) closely matches the medians in the observation bins and the theoretical threshold is well reproduced (albeit off-scale).
-The VPD-independent, strongly negative response, which is determined by uWUE, is also captured.
-For CSH, there is extremely rare occurrence of observed positive ET response with VPD ($\approx 3$\%), even with uncertainty, so the sign of the observations almost always matches the sign of the theory, which states that ET response should always be negative.
+While CSH has a much different functional form of the sign term than DBF and MF, CSH observations also match our theory to leading order, albiet with a bit more variability as a function of VPD.
+Again, the VPD dependence mostly determined from the slope parameter ($g_1$) closely matches the medians in the observation bins.
+The VPD-independent, strongly negative response is also captured.
+For CSH, there is rare occurrence of observed positive ET response with VPD ($\approx 6$\%), even with uncertainty, so the sign of the observations almost always matches the sign of the theory, which states that ET response should always be negative.
 
-As Table \ref{stats} suggested, the theory does not match as well for CRO as for DBF and CSH.
-At low VPD the VPD dependence (due to $g_1$) is similar to the leading order behavior of the observations, but the theory's curve is shifted towards more positive values relative to the observations, likely due to some bias in uWUE.
-Most strikingly, the observations show a tendency to move back towards negative values at high VPD, which given the functional form of VPD dependence, is behavior that is impossible for the theory to capture.
-This general positive bias of the theory explains the shift towards negative ET response observed in the bulk statistics in Table \ref{stats}.
-The divergence between the theory and the observations for CRO could have been expected.
-For practical reasons, our theory does not include important sources of plant type heterogeneity within and across sites for CRO.
-In agricultural landscape, the footprint of an eddy-covariance towers tend to not only observe a single species but many different ones, with potentially varying water use strategies.
-We also do not account for seasonal changes in LAI, variability in crop height and surface roughness, differences in C3 vs C4 photosynthesis and water strategies, or differences in irrigation practices between the sites and years.
-These deficiencies should largely explain the inability of our theory to exactly match the observations in croplands.
+Biases between the theory and observations are similiar for ENF, WSA, CRO and SAV.
+At low VPD the theory and the observations match well.
+However, at high VPD (upper 10-20\% of observations) the theory is biased slightly towards postive reponse as compared to the observations (Figure \ref{test_sign}).
+For ENF this slight bias could be explained by a negative bias in $g_1$.
+However, for WSA, CRO, and SAV the observations in the highest VPD bins exhibit a downturn towards more negative response, which cannot be captured by the functional form of the sign term (Figure \ref{idealized_sign}).
+Therefore, to explain the observations one of the varibles in the sign term must change at extreme VPD (upper 10-20\%).
+The most likely candidiate is $uWUE$, which we have assumed constant to meet our goal of developing intuition for leading order behavior, but might be expected to decrease at extremely low SWC, which should be correlated with high VPD.
+The theory for EBF exhibits similiar limiatations as for ENF, WSA, and SAV, except a larger portion of observations are biased negative as compared to the theory ($\approx$ 35\%, Figure \ref{test_sign}).
+However, in general, and specifically for non-extreme VPD, the theory matches the observations for the tree and savannah plant types.
+
+The theory for GRA suffers from similar, but much more severe, limitations as for CRO, WSA, and SAV.
+GRA observations are characterized by a consistent trend back towards negative ET response at higher VPD, which the functional form of our theory is incapable of accounting for.
+As compared to CRO, WSA and SAV, the divergence between the theory and the observations is far greater for GRA, biasing 40-50\% of observations at higher VPD.
+In addition for the potential for soil moisture to alter $uWUE$, there are other sources of plant heterogeneity specific to GRA (and some extent CRO) that may alter $uWUE$ (or $g_1$) or invalidate other assumptions made in the Methods section (Section \ref{methods}).
+We do not account for variability in plant height and surface roughness, or differences in C3 vs C4 photosynthesis and water strategies, which we might expect to vary substantially across sites, years, and season for GRA.
+These deficiencies could largely explain the inability of our theory to exactly match the observations in croplands.
 For example, a superposition of sites with C3 plants (at low environmental VPD) and C4 plants (at high environmental VPD) would explain observed shift back towards negative ET response at high VPD when all sites are binned together, as in Figure \ref{test_sign}.
+We hypothesize that the theory would validate against observations much better if these sources of varaibility were accounted for, at a cost of increased complexity and analytical opacity.
 
-The theory for GRA suffers from similar limitations as for CRO.
-GRA observations are characterized by a consistent trend back towards negative ET response at higher VPD, which the functional form of our theory is incapable of accounting for.
-As compared to CRO, the divergence between the theory and the observations is even greater for GRA, but these differences are explainable using similar logic as we did for CRO.
-GRA also tends to exhibit substantial within-PFT variability, such as C3 vs C4 photosynthesis, as well as strong seasonal variability in height and LAI.
-As with CRO, the trend back towards negative ET response at higher VPD could be explained by a superposition of sites with C3 vs C4 plants.
-Unfortunately, because detailed site-level data of C3 vs C4 plant type is not available, we could not bin CRO and GRA further by irrigation and photosynthesis strategy.
-However, we hypothesize that the theory would validate against observations much better when photosynthesis pathways are accounted for.
+While the above discussion shows that our theory has some limitations when applied to some PFTs, especially for grasslands, it does well for DBF, MF and CSH PFTs, and captures the response at non-extreme VPD for ENF, EBF, CRO, WSA, and SAV.
+In general, the leading order behavior observed in the data is captured by the theory.
+Departures between our theory and observations, specifically at high VPD, could be explained and conceptualized with shifts in $g_1$ and/or $uWUE$ due to site-spcific plant type variability (e.g.
+more arid-adapted ecosystems at more arid sites), or temporal variability for some enviornmental conditions (e.g.
+decreases in $uWUE$ at extremely low soil moisture).
+To focus on general behaviro and develop intuition for PFT-scale response, we have ignored these sources of variability in the present analysis.
+However, any site-specific or temporal plant functional variability that can be conceptualized with shifts to $g_1$ and $uWUE$ can be analyzed with our framework.
+This opens a door to a myriad future analysis in which plant behavior in anomalous conditions can be explained and analyzed using Equation \ref{et}.
 
-The comparison between the theory and observations for ENF and EBF is unique from the rest of the PFTs.
-The VPD dependence of the theory, determined by the slope parameter ($g_1$), is not as well captured by observations.
-However, the theory still captures the general behavior and magnitude of the response, with some increase in error at low and/or high VPD.
-We note that we used published constants for $g_1$ \citep[from ][]{Lin_2015, Franks_2017}, which determines the VPD dependence, and that deviations should be expected across sites \citep{Lin_2015}.
-We could have fit our own constants so that the theory would best match the observations, but our goal was to use a priori results whenever possible in order to avoid over-fitting and to have a theoretical approach as independent to the observations as possible for further confirmation of the validity of the approach.
-
-While the above discussion shows that our theory has some limitations when applied to some PFTs, especially for crops and grasslands, it does well for DBF and CSH PFTs, and captures the response at non-extreme VPD for ENF and EBF.
-It is useful to also compare our theory, the observed uncertainty in the ET response, and PET, as PET is often used as a proxy for ET in hydrometeorological studies and drought indices.
+% It is useful to also compare our theory, the observed uncertainty in the ET response, and PET, as PET is often used as a proxy for ET in hydrometeorological studies and drought indices.
 Again, as in our discussion of the theoretical results, the PET sensitivity to VPD will be constant at about 3.5.
@@
 CSH are the most water conservative, with a strong negative response.
-DBF, EBF, ENF, and GRA are also water conservative, but show some occurrence of positive ET response to VPD, particularly at higher observed VPD, as the theory predicts.
+DBF, EBF, ENF, MF, SAV and WSA are also water conservative, but show some occurrence of positive ET response to VPD, particularly at higher observed VPD, as the theory predicts.
+GRA, while generally water conservative, does not match the theory, with increasing frequency of \textit{negative} response at high VPD.
+This is as expected, given discussion in Section \ref{testing}.
 
@@
 The portion of observations for which our theory is biased by SWC-limitations varies by PFT, due to the nonlinear threshold at which soil moisture availability limits plant function.
-For CRO and DBF, soil water content only matters for about the lowest 5\% of observations (each box is 5\% of observations in Figure \ref{swc_boxplot}).
-And, CRO, relative to DBF, has a very weak dependence on soil moisture, which is reflective of the high likelihood that CRO sites are optimally irrigated and not water stressed.
-For ENF and EBF, systematic SWC-induced biases in $\sigma \cdot uWUE$ emerge in about the lowest 30\% of SWC conditions.
+For CRO, MF and DBF, soil water content only matters for about the lowest 5\% of observations (each box is 5\% of observations in Figure \ref{swc_boxplot}).
+And, CRO, relative to DBF, has a very weak dependence on soil moisture, which is reflective of the high likelihood that CRO sites are optimally irrigated and not water stressed, suggesting that observed depatures from the theory for CRO are due to other factors (i.e.
+photosythesis pathway) than soil moisture.
+For ENF, EBF, SAV, and WSA, systematic SWC-induced biases in $\sigma \cdot uWUE$ emerge in about the lowest 20-30\% of SWC conditions, although large variability in the SWC-$\sigma \cdot uWUE$ relationship hampers interpretation for WSA.
 In crontast to ENF, CRO, and DBF, for GRA the relationship between $\sigma \cdot uWUE$ and SWC is more linear and affects a greater portion of observations; about 60\% of observations.
-Clearly, for GRA soil water frequently impacts plant function and alters ET response, and our theory is limited.
-It is therefore not suprising that our theory tested poorly againt the data for GRA, relative to ENF, EBF, DBF, and CSH (Section \ref{testing}).
-For all other PFTs occurence of soil mosture impacts was rare enough to not manifest itself in bulk statistics and plots.
+Clearly, for GRA soil water frequently impacts plant function and alters ET response, and our theory is limited for the majority of encironmental condiitons.
+It is therefore not suprising that our theory tested poorly againt the data for GRA, relative to to the other PFTs (Section \ref{testing}).
+For all other PFTs occurence of soil mosture impacts was rare enough to not manifest itself in bulk statistics and figures.
 A possible exception is CSH, for which there are only two sites, which precludes comment on the exact PFT-wide relationship.
@@
 The lower uWUE at low SWC/high VPD then leads to the observed downturn towards decreasing ET with increasing VPD, and the deviation between our theory (based on a constant uWUE assumption) and observations.
+It is also perhaps not coincidental that the protion of $\sigma \cdot uWUE$ affected by low soil moisture observations is similiar to the portion of observations that to not match our theory at high VPD.
 
@@
 Soil moisture effects explain the deficiencies of our theory in Section \ref{testing}, particularly for GRA.
+By conceptualizing SWC effects as a change in $uWUE$ (or $g_1$ as well), it will be possible for future analysis to explore the importance of soil moisture on plant response to VPD.
 

### 2018-03-26 src/paper_figure_code/data_scatter.py

@@ -69,8 +69,8 @@ def make_ax_plot(_ax, var, _df, meta):
   color = _ax.pcolormesh(_x, _y, z, cmap=meta['cmap'],\
                          vmin=vmin, vmax=vmax)
 
-  custom_xlabel(_df, _ax, '%s (Pa)' % 'vpd'.upper())
-  custom_ylabel(_df, _ax, 'T (C)')
+  custom_xlabel(_df, _ax, '%s (Pa)' % 'vpd'.upper(), fontsize=small_ax_fontsize)
+  custom_ylabel(_df, _ax, 'T (C)', fontsize=small_ax_fontsize)
   _ax.set_title('%s'\
                 % (name_dict[_df['pft'].iloc[0]]), fontsize=local_fontsize)
   # _ax.yaxis.set_tick_params(labelsize=local_fontsize-3)
@@ -78,7 +78,7 @@ def make_ax_plot(_ax, var, _df, meta):
   cbar = plt.colorbar(color, ax=_ax)# , ax=_ax2)
   if ((_df.pft.iloc[0] == 'MF') | (_df.pft.iloc[0] == 'WSA')\
       | (_df.pft.iloc[0] == 'SAV')):
-    cbar.set_label(meta['title'], fontsize=local_fontsize)
+    cbar.set_label(meta['title'], fontsize=small_ax_fontsize)
   y_lim = _ax.get_ylim()
   # vpd_crit = et_min_vpd(mean_df.loc[_df.pft.iloc[0], :])
   # _ax.plot(np.ones(2)*vpd_crit, np.array(y_lim), 'k-', linewidth=2.0)

### 2018-03-26 src/paper_figure_code/swc_boxplot.py

@@ -2,6 +2,7 @@
 """
 This script makes box plot figures for code
 """
+from matplotlib.ticker import FuncFormatter
 from shared_functions import *
 import seaborn as sns
 import importlib
@@ -13,7 +14,8 @@ medianprops = {'linewidth' : 2.5}
 dashedlinewidth=1.4
 
 bluelinewidth=3.0
-
+xlim = {'WSA' : [0.00, 0.0225],
+        'SAV' : [0.002, 0.0073]}
 def plot_box(_df, ax=None, pft=False):
   """makes a box plot"""
   if ax is None:
@@ -51,9 +53,11 @@ def plot_box(_df, ax=None, pft=False):
     ax.set_title('Site: %s, PFT: %s'\
                  % (_df.site.iloc[0], _df.pft.iloc[0]),\
                  fontsize=fontsize+3)
-
+  if _df.pft.iloc[0] in xlim:
+    _xlims = xlim[_df.pft.iloc[0]]
+    ax.set_xlim(_xlims)
+    ax.xaxis.set_ticks(np.round(np.linspace(_xlims[0], _xlims[1], 4), 4))
   # ax.set_ylim(ylim)
-
   # ax.set_xlim(orig_xlim)
   if savefig:
     plt.tight_layout()

### 2018-03-26 doc/paper/vpd_et_paper.tex
done with discussion, need to add convex vs convave discussion

@@
 \centerline{\includegraphics[width=1.4\textwidth]{./swc_boxplot.pdf}}
-\caption{$\sigma \cdot uWUE$ as a function of PFT and SWC. Each box plot represents 5\% of all observations.}
+\caption{$\sigma \cdot uWUE$ as a function of PFT and SWC.
+Each box plot represents 5\% of all observations.
+Note that the highest 10\% of SWC observations were excluded from the WSA and SAV figures, to better resolve high variability in narrow low-SWC bins for those PFTs.}
 \label{swc_boxplot}
@@
 Soil moisture effects explain the deficiencies of our theory in Section \ref{testing}, particularly for GRA.
-By conceptualizing SWC effects as a change in $uWUE$ (or $g_1$ as well), it will be possible for future analysis to explore the importance of soil moisture on plant response to VPD.
+By conceptualizing SWC effects as a change in $uWUE$ (and/or $g_1$), it will be possible for future analysis to explore the importance of soil moisture on plant response to VPD.
 
@@
 Stomatal regulation and plant physiological response strongly regulate ET, and this regulation varies by PFT.
-CROs are the least water conservative, while DBF, EBF, ENF, GRA and CSH are progressively more water conservative (more likely to reduce ET in response to increasing VPD).
+CROs are the least water conservative, while DBF, EBF, SAV, GRA, MF, WSA, ENF and CSH are progressively more water conservative (more likely to reduce ET in response to increasing VPD).
+SAV and WSA exhibit positive ET response to VPD not necessarily because of poor water conservation stragegies relative to other PFTs, but because of greater occurence of high atmospheric demand (VPD) relative to other PFTs.
 Observations of ET response to VPD exhibit the same general behavior as the theory, with ET response becoming more positive (atmospheric demand dominating) as environmental VPD increases within a PFT, and more negative for PFTs that are adapted to arid conditions and prioritize water conservation over primary production.
@@
 
-Lastly, our paper builds important intuition for how plants will respond to global change-induced shifts in VPD, which should increase in the future.
+Lastly, our paper builds important intuition for how plants will respond to VPD pertubations, including expected increases in VPD induced by global change.
 Plant physiological responses to direct CO$_2$ effects \citep[e.g.,][]{Swann_2016} receives more attention than physiological response to indirect effects like increased VPD.
-Here, we provide broad PFT-focused results showing a likely decrease in ET in response to positive VPD perturbations (atmospheric drying).
-Additionally, we derive an explicit conceptual framework for ET response as a function of two plant type parameters, $g_1$ and uWUE (e.g.
-Equation \ref{et}).
-This framework can be used to understand ET response to a range of environmental change, and forms a foundation for resolving varying response to more nuanced consideration of plant type and climate.
-Any plant physiological heterogeneity that can be conceptualized with shifts in $g_1$ \citep[e.g.][]{Lin_2015} and/or uWUE \citep[e.g.][]{Zhou_2014} is representable within our framework, which opens the door for a hierarchy of more sophisticated climate- and plant-specific analysis of ET sensitivity to environmental variables (including VPD).
+Here, we provide broad PFT-focused results showing a likely decrease in ET in response to positive VPD perturbations (atmospheric drying), which is consistent with recent observational analysis \citep[e.g.,][]{Rigden_2017}.
+Feedbacks between the land and the atmosphere may alter the net response to a global VPD perturabation, but our focus on the one way plant response to a VPD perturbation in the atmopsheric boundary layer is an important first step to disentangling such feedbacks, both in observations and model simulations of the present and future.
+By removing Penman Monteith's dependence on implicit relationships between GPP, VPD, and ET, we allow for explicit, accuruate analysis of plant-VPD feedbacks in the atmospheric boundary layer (Equation \ref{et}).
+Our approach can be extended to examine varying plant response to more nuanced consideration of plant type and climate.
+Any plant physiological heterogeneity that can be conceptualized with shifts in $g_1$ \citep[e.g.][]{Lin_2015, Medlyn_2017} and/or uWUE \citep[e.g.][]{Zhou_2014} is representable within our framework, which opens the door for a hierarchy of more sophisticated climate- and plant-specific analysis of ET sensitivity to environmental variables (including VPD).
 We argue that such simplified conceptual frameworks are critical tools for disentangling land-atmosphere feedbacks at various scales, from diurnal to seasonal and beyond, and to characterize ET response in a warmer, atmospherically drier, and enriched CO$_2$ world.

### 2018-03-26 doc/paper/vpd_et_paper.tex
addeds econd derviative plot, should probably hash it to save ink, also need to write that section

@@
 Feedbacks between the land and the atmosphere may alter the net response to a global VPD perturabation, but our focus on the one way plant response to a VPD perturbation in the atmopsheric boundary layer is an important first step to disentangling such feedbacks, both in observations and model simulations of the present and future.
-By removing Penman Monteith's dependence on implicit relationships between GPP, VPD, and ET, we allow for explicit, accuruate analysis of plant-VPD feedbacks in the atmospheric boundary layer (Equation \ref{et}).
-Our approach can be extended to examine varying plant response to more nuanced consideration of plant type and climate.
+By removing Penman Monteith's dependence on implicit relationships between GPP, VPD, and ET, we allow for explicit, accuruate future analysis of plant-VPD feedbacks in the atmospheric boundary layer (Equation \ref{et}).
+Our approach can also be extended to examine varying plant response to more nuanced consideration of plant type and climate.
 Any plant physiological heterogeneity that can be conceptualized with shifts in $g_1$ \citep[e.g.][]{Lin_2015, Medlyn_2017} and/or uWUE \citep[e.g.][]{Zhou_2014} is representable within our framework, which opens the door for a hierarchy of more sophisticated climate- and plant-specific analysis of ET sensitivity to environmental variables (including VPD).

### 2018-03-26 src/paper_figure_code/shared_functions.py

@@ -72,17 +72,21 @@ name_dict = {'CRO': 'Crops (CRO)',\
 
 
 
-def custom_ylabel(_df, ax, label, fontsize=fontsize):
+def custom_ylabel(pft, ax, label, fontsize=fontsize):
   """labels y label for axis ordering"""
-  if (_df.pft.iloc[0] == 'ENF') | (_df.pft.iloc[0] == 'DBF')\
-     | (_df.pft.iloc[0] == 'CRO'):
+  if not isinstance(pft, str):
+    pft = pft.pft.iloc[0]
+  if (pft == 'ENF') | (pft == 'DBF')\
+     | (pft == 'CRO'):
     ax.set_ylabel(label, fontsize=fontsize)
   return
 
-def custom_xlabel(_df, ax, label, fontsize=fontsize):
+def custom_xlabel(pft, ax, label, fontsize=fontsize):
   """labels x label for axis ordering"""
-  if (_df.pft.iloc[0] == 'CRO') | (_df.pft.iloc[0] == 'GRA')\
-     | (_df.pft.iloc[0] == 'SAV'):
+  if not isinstance(pft, str):
+    pft = pft.pft.iloc[0]
+  if (pft == 'CRO') | (pft == 'GRA')\
+     | (pft == 'SAV'):
     ax.set_xlabel(label, fontsize=fontsize)
   return
 

### 2018-03-28 doc/paper/vpd_et_paper.tex
forgot concave plot

@@
 
+\subsection{Functional form of ET dependence on VPD and its relation to uWUE}
+
+Th theory described in Section \ref{sign_term} indicates that for a given $uWUE$ and $g_1$, the ET dependence on VPD should be concave upward.
+In other words, there should be some local minima in ET at a critical VPD$_{crit}$, assuming the scaling term (e.g.
+aerodynamic conductance and $\Delta$) is held fixed.
+This result warrants futher investigation, because to our knowledge no one has derived the theoretical ecosystem-scale relationship between ET and VPD while controlling for enivironmental conditiions.
+In particular, from personal communication, there is an apparent lack of consensus over whether the shape of the ET-VPD curve should be concave upward (our result) or concave downward.
+Given that understanding the ET-VPD relationship of the one-way plant response is absolutely fundamental for hypothesizing about any feedbacks between the land surface and the atmopshere, we analyze why our derived ET-VPD relationship is concave upward, particularly with respect to our use of $uWUE$ to make explicit the relationship betwen GPP, ET, and VPD.
+
 \section{Discussion}

### 2018-03-28 src/paper_figure_code/second_derivative.py

@@ -1,90 +1,0 @@
-import pandas as pd
-from sympy import *
-import pickle
-import os
-import matplotlib.pyplot as plt
-import numpy as np
-from shared_functions import *
-
-pft_order = ['CSH', 'MF', 'SAV', 'ENF',\
-             'EBF', 'WSA', 'DBF',  'CRO', 'GRA']
-
-
-def plot_second_derivative(row, ax, zero_func):
-  """
-  given a fucntion for second derivative, makes a plot
-  vpd on x axis wue power on y axis
-  """
-
-  _vpd = np.linspace(dfs['5'].loc[row.pft, 'vpd'],\
-                     dfs['95'].loc[row.pft, 'vpd'])
-  def temp_f(vpd):
-    return zero_func(row.g_a, row.p_a, row.t_a_k, d_calc.CP, vpd,\
-                   row.r_moist, row.gamma, row.delta, row.r_n,\
-                   row.c_a, d_calc.R_STAR, row.uwue, row.g1, 1.6)
-  _w_power = [temp_f(vpd) for vpd in _vpd]
-  ax.plot(_vpd, _w_power,\
-          label="%s"\
-          % row.pft)
-
-
-  return
-
-init_printing()
-# init rad
-r, delta = symbols('r Delta')
-# init aero
-g_a, p, t, cp, vpd, r_air = symbols('g_a P T c_p VPD R_air')
-# init plant
-gamma, c_s, r_star, uwue, g_1, onesix = symbols('gamma c_s R* uWUE g_1 1.6')
-
-# powers
-wue_power = symbols('n')
-
-et = (delta*r\
-      +g_a*p/t\
-       *(cp*vpd/r_air\
-         -gamma*c_s*vpd**wue_power\
-          /(r_star*onesix*uwue*(1+g_1/vpd**(1/2)))))\
-      /(delta + gamma)
-
-def sgn(expr):
-  """crudely returns the sign term given the d_et derivative"""
-  sign = expr*t*(delta+gamma)/(p*g_a)
-  neg = sign- cp/r_air
-  coll = collect(neg, (c_s*gamma/(1.6*r_star*uwue*onesix)))
-  return coll
-
-d_vpd = diff(et, vpd)
-sign = sgn(d_vpd)
-d2_vpd = diff(d_vpd, vpd)
-et
-vpd
-simple_second = simplify(d2_vpd)
-sign_terms = simple_second\
[…]

### 2018-03-29 doc/paper/vpd_et_paper.tex
refactored and back to writing

@@
 
-\subsection{Functional form of ET dependence on VPD and its relation to uWUE}
+\subsection{Functional form of ET dependence on VPD and its relation to VPD dependence}
 
@@
 In particular, from personal communication, there is an apparent lack of consensus over whether the shape of the ET-VPD curve should be concave upward (our result) or concave downward.
-Given that understanding the ET-VPD relationship of the one-way plant response is absolutely fundamental for hypothesizing about any feedbacks between the land surface and the atmopshere, we analyze why our derived ET-VPD relationship is concave upward, particularly with respect to our use of $uWUE$ to make explicit the relationship betwen GPP, ET, and VPD.
+Given that understanding the ET-VPD relationship of the one-way plant response is fundamental to hypothesizing about any feedbacks between the land surface and the atmopshere, we analyze why our derived ET-VPD relationship is concave upward, particularly with respect to the exponent of VPD depedence in $uWUE$ and the Medlyn unified stomatal conductace model.
+
+There is a theoretical basis for the square root VPD dependence in both the stomatal conducatance model and $uWUE$ based on the assumption that stomata behave to maximize carbon gain while minimizing water loss, which observations also generally support \citep{Lloyd_1991, MEDLYN_2011, Lin_2015, Zhou_2014, Zhou_2015, Medlyn_2017}.
+However, purely empiriacle results which fit the exponent of the VPD dependence have shown that it may vary from 1/2, suggesting that stomata, as well as ecosystem-scale quantities based on stomata theroy,  may not always function optimally \citep{Zhou_2015, Lin_2018}.
+Specifically with regards to $uWUE$, one would not expect that this ecosystem scale WUE quantity will responde to VPD exactly analgalously to stomata.
+Direct soil evaporation's contributions to ET should shift the exponential dependence of VPD to be greater than 1/2.
+In other words, direct soil evaporation causes WUE to decrease more strongly with increasing VPD, especially in well-watered conditions.
+\citet{Zhou_2015}'s results corroborate this: they found a mean imprically fit exponential VPD dependence of 0.55 istead of 0.5 for Ameriflux sites.
+Given that these recent results on the relatioship between VPD, GPP, and ET form the backbone of our analysis and are what allowed us to derive an explicit expression for ET for the first time (Equation \ref{et}), we will analyze if and how assumptions about the exonent of the VPD dependence impacts the shape of the ET-VPD dependence.
+By introducing $n$ and $m$ we can free our stomatal conductance model from assumptions about VPD dependence:
+\begin{linenomath*}
+  \begin{equation}
+    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{VPD^m}\right) \frac{*WUE \; ET}{c_a \; VPD^n},
+    \label{m_n}
+  \end{equation}
+\end{linenomath*}
+where:
+\[*WUE = \frac{GPP}{ET}VPD^n.\]
+To determine how the exponent $n$ and $m$ alter the shape of the ET-VPD dependence we find the roots of the second deriviative of ET, using Equation \ref{m_n} for stomatal conductance, with respect to VPD:
+\input{d2_solutions.tex}
 

### 2018-03-29 src/paper_figure_code/concave.py

@@ -6,25 +6,7 @@ import matplotlib.pyplot as plt
 import numpy as np
 from shared_functions import *
 
-pft_order = ['CSH', 'MF', 'SAV', 'ENF',\
-             'EBF', 'WSA', 'DBF',  'CRO', 'GRA']
-
 plt.close('all')
-def plot_second_derivative(row, ax, zero_func):
-  """
-  given a fucntion for second derivative, makes a plot
-  vpd on x axis wue power on y axis
-  """
-
-  # _vpd = np.linspace(dfs['5'].loc[row.pft, 'vpd'],\
-  #                    dfs['95'].loc[row.pft, 'vpd'])
-  _vpd = np.linspace(10., 4000.)
-  _w_power = [zero_func(vpd, row.g1) for vpd in _vpd]
-  print(np.nanmin(_w_power))
-  ax.plot(_vpd, _w_power,\
-          label="%s: g1: %5.1f"\
-          % (row.pft, np.round(row.g1, 1)))
-  return
 
 init_printing()
 # init rad
@@ -36,109 +18,6 @@ gamma, c_s, r_star, uwue, g_1, onesix = symbols('gamma c_s R* uWUE g_1 1.6')
 
 vpd, delta = symbols('VPD Delta')
 
-# powers
-wue_power = symbols('n')
-
-et_true = (delta*r\
-           +g_a*p/t\
-           *(cp*vpd/r_air\
-             -gamma*c_s*vpd**wue_power\
-             /(r_star*onesix*sigma*uwue*(1+g_1/vpd**(1/2)))))\
-             /(delta + gamma)
-et_pm = (delta*r + g_a*p/(t*r_air)*cp*vpd)/(delta+gamma*(1+g_a/g_s))
-et = symbols('ET')
-rhs = et_pm.subs(g_s, r_star*t/p*onesix*(1+ g_1/vpd**(1/2))\
-                 *sigma*uwue*et/(c_s*vpd**wue_power))
-
-# step 1
-lhs = et*(delta+gamma*(1+g_a/(r_star*t/p*onesix*(1+ g_1/vpd**(1/2))\
-                 *sigma*uwue*et/(c_s*vpd**wue_power))))
-rhs = rhs*(delta+gamma*(1+g_a/(r_star*t/p*onesix*(1+ g_1/vpd**(1/2))\
-                 *sigma*uwue*et/(c_s*vpd**wue_power))))
-
[…]

### 2018-04-01 doc/paper/vpd_et_paper.tex
filled formatted .tex, should have been doing this from the start

@@
 % \draftfalse
-\usepackage{makecell}
-\usepackage{multirow} %AM for tables
-\usepackage{booktabs}
-\drafttrue
+\usepackage{makecell} \usepackage{multirow} %AM for tables
+\usepackage{booktabs} \drafttrue
 
-\journalname{Agricultural and Forest Meteorology: When does VPD drive or reduce ET?}
+\journalname{Agricultural and Forest Meteorology: When does VPD drive
+  or reduce ET?}
 
@@
 
-\title{When does vapor pressure deficit drive or reduce evapotranspiration?}
+\title{When does vapor pressure deficit drive or reduce
+  evapotranspiration?}
 
-\authors{Adam Massmann\affil{a}, Pierre Gentine\affil{a}, Changjie Lin\affil{a,b}}
+\authors{Adam Massmann\affil{a}, Pierre Gentine\affil{a}, Changjie
+  Lin\affil{a,b}}
 
-\affiliation{a}{Department of Earth and Environmental Engineering, Columbia University, New York, NY 10027}
-\affiliation{b}{State Key Laboratory of Hydroscience and Engineering, Department of Hydraulic Engineering, Tsinghua University, Beijing, CN 100084}
+\affiliation{a}{Department of Earth and Environmental Engineering,
+  Columbia University, New York, NY 10027} \affiliation{b}{State Key
+  Laboratory of Hydroscience and Engineering, Department of Hydraulic
+  Engineering, Tsinghua University, Beijing, CN 100084}
 
-\correspondingauthor{Adam Massmann}{206-919-1364; akm2203@columbia.edu}
+\correspondingauthor{Adam Massmann}{206-919-1364;
+  akm2203@columbia.edu}
 
 \begin{keypoints}
-\item ecohydrology, evapotranspiration, aridity, vapor pressure deficit, Clausius-Clapeyron relation, plant physiology, stomatal response
+\item ecohydrology, evapotranspiration, aridity, vapor pressure
+  deficit, Clausius-Clapeyron relation, plant physiology, stomatal
+  response
 \end{keypoints}
@@
 \begin{abstract}
-  Increases to vapor pressure deficit (VPD) and atmospheric demand for water are expected with rising atmospheric CO$_2$.
-While increased evapotranspiration (ET) in response to increased atmospheric demand seems intuitive, plants are capable of reducing ET in response to increased VPD by closing their stomata, in an effort to conserve water.
-Here we examine which effect dominates response to increasing VPD: atmospheric demand and increases in ET, or plant physiological response (stomata closure) and decreases in ET.
-We use Penman-Monteith, combined with semi-empirical optimal stomatal regulation theory and underlying water use efficiency, to develop a theoretical framework for understanding how ET responds to increases in VPD.
-  The theory suggests that for most environmental conditions and plant types, plant physiological response dominates and ET decreases with increasing VPD.
-Plants that are evolved or bred to prioritize primary production over water conservation (e.g.
-crops) exhibit a higher likelihood of atmospheric demand-driven response (ET increasing).
-However for forest, grass, and shrub plant types, ET more frequently decreases than increases with rising VPD.
-Therefore, we may expect reductions in ET due to CO$_2$-induced increases in VPD.
-This work serves as an example of the utility of our simplified framework for disentangling land-atmosphere feedbacks at multiple scales, including the characterization of ET response in an atmospherically drier, enriched CO$_2$ world.
+  Increases to vapor pressure deficit (VPD) and atmospheric demand for
+  water are expected with rising atmospheric CO$_2$. While increased
+  evapotranspiration (ET) in response to increased atmospheric demand
+  seems intuitive, plants are capable of reducing ET in response to
+  increased VPD by closing their stomata, in an effort to conserve
+  water. Here we examine which effect dominates response to increasing
+  VPD: atmospheric demand and increases in ET, or plant physiological
+  response (stomata closure) and decreases in ET. We use
+  Penman-Monteith, combined with semi-empirical optimal stomatal
+  regulation theory and underlying water use efficiency, to develop a
+  theoretical framework for understanding how ET responds to increases
+  in VPD.
+
+  The theory suggests that for most environmental conditions and plant
+  types, plant physiological response dominates and ET decreases with
+  increasing VPD. Plants that are evolved or bred to prioritize
+  primary production over water conservation (e.g. crops) exhibit a
+  higher likelihood of atmospheric demand-driven response (ET
+  increasing). However for forest, grass, and shrub plant types, ET
+  more frequently decreases than increases with rising VPD. Therefore,
+  we may expect reductions in ET due to CO$_2$-induced increases in
+  VPD. This work serves as an example of the utility of our simplified
+  framework for disentangling land-atmosphere feedbacks at multiple
+  scales, including the characterization of ET response in an
+  atmospherically drier, enriched CO$_2$ world.
 
@@
 
-Vapor pressure deficit (VPD) is expected to rise over continents in the future due to the combination of increased temperature and, depending on region, decreased relative humidity \citep{Byrne_2013}.
-Increases in VPD increase the atmospheric demand for evapotranspirated water \citep{Penman_1948, Monteith_1965}, but also stress plant stomata \citep{Leuning_1990, MEDLYN_2011}.
+Vapor pressure deficit (VPD) is expected to rise over continents in
+the future due to the combination of increased temperature and,
+depending on region, decreased relative humidity
+\citep{Byrne_2013}. Increases in VPD increase the atmospheric demand
+for evapotranspirated water \citep{Penman_1948, Monteith_1965}, but
+also stress plant stomata \citep{Leuning_1990, MEDLYN_2011}.
 
-The opposing effects of increased atmospheric demand and higher stomatal stress lead to two possible perspectives for how evapotranspiration (ET) responds to shifts in VPD.
-The first, a hydrometeorological perspective, is that higher VPD increases atmospheric demand for water from the land surface, and this drives an increase in evapotranspiration (ET).
-This perspective is particularly relevant because potential evapotranspiration (PET), which is used in many drought indices and hydrometeorological studies \citep[e.g.,][]{Heim_2002, Scheff_2015}, typically only quantifies changes in atmospheric demand and fails to account for ecosystem response \citep{Swann_2016}.
-In reality, plants' stomata have evolved to optimally regulate the exchange of water and carbon, and tend to partially close in response to increased atmospheric dryness \citep{Farquhar_1978, Ball_1987, Leuning_1990, MEDLYN_2011}.
-This leads to a plant physiology perspective, in which an increase in VPD, particularly in well-watered soil conditions, may actually correspond to a decrease in ET because of stomatal closure \citep[e.g.][]{Rigden_2017}.
-In other words, the  question ``When does VPD drive or reduce ET?'' can be related to whether plant regulation or atmospheric demand dominates ET response.
+The opposing effects of increased atmospheric demand and higher
+stomatal stress lead to two possible perspectives for how
+evapotranspiration (ET) responds to shifts in VPD. The first, a
+hydrometeorological perspective, is that higher VPD increases
+atmospheric demand for water from the land surface, and this drives an
+increase in evapotranspiration (ET). This perspective is particularly
+relevant because potential evapotranspiration (PET), which is used in
+many drought indices and hydrometeorological studies
+\citep[e.g.,][]{Heim_2002, Scheff_2015}, typically only quantifies
+changes in atmospheric demand and fails to account for ecosystem
+response \citep{Swann_2016}. In reality, plants' stomata have evolved
+to optimally regulate the exchange of water and carbon, and tend to
+partially close in response to increased atmospheric dryness
+\citep{Farquhar_1978, Ball_1987, Leuning_1990, MEDLYN_2011}. This
+leads to a plant physiology perspective, in which an increase in VPD,
+particularly in well-watered soil conditions, may actually correspond
+to a decrease in ET because of stomatal closure
+\citep[e.g.][]{Rigden_2017}.  In other words, the question ``When does
+VPD drive or reduce ET?'' can be related to whether plant regulation
+or atmospheric demand dominates ET response.
 
-The ET response to changes in VPD alters water partitioning between the soil and atmosphere.
-If plant response reduces ET with atmospheric drying then soil moisture will be better conserved.
-This would seem a sensible evolutionary strategy to cope with aridity.
-If stomata were fully passive \citep [similar to soil pores, e.g.
-][]{Or_2013}, increased atmospheric aridity would strongly reduce soil moisture \citep{Berg_2017}.
-In turn, this would further increase aridity as low soil moisture levels increase the Bowen ratio, and cause increased temperature and atmospheric drying \citep[][]{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006, Salvucci_2013, Gentine_2016, Berg_2016}.
-This however would not seem to be a sensible strategy for plants from an evolutionary standpoint.
+The ET response to changes in VPD alters water partitioning between
+the soil and atmosphere. If plant response reduces ET with atmospheric
+drying then soil moisture will be better conserved. This would seem a
+sensible evolutionary strategy to cope with aridity. If stomata were
+fully passive \citep [similar to soil pores, e.g. ][]{Or_2013},
+increased atmospheric aridity would strongly reduce soil moisture
+\citep{Berg_2017}. In turn, this would further increase aridity as low
+soil moisture levels increase the Bowen ratio, and cause increased
+temperature and atmospheric drying \citep[][]{Bouchet_1963,
+  Morton_1965, Brutsaert_1999, Ozdogan_2006, Salvucci_2013,
+  Gentine_2016, Berg_2016}. This however would not seem to be a
+sensible strategy for plants from an evolutionary standpoint.
 
-We can use intuition about plant water conservation strategy to hypothesize about ET response to changes in VPD.
-Plants and ecosystems that evolved to conserve water, such as arid shrubs or savannah, should be more likely to reduce ET with increasing VPD, and plants that have evolved or have been engineered to care little about water, such as crops, will be more likely to increase ET with increasing VPD.
-Atmospheric conditions must matter as well.
-At the ecosystem scale, there are limits to plant water conservation strategies.
-As atmospheric demand for water (VPD) increases, ecosystems should begin to reach their water conservation limits and might not be able to entirely limit ET flux to the atmosphere.
-At this stage any further increase in VPD will most likely drive a (limited) increase in ET, because the increase in atmospheric demand for water overwhelms the limited plant response to conserve water.
+We can use intuition about plant water conservation strategy to
+hypothesize about ET response to changes in VPD. Plants and ecosystems
+that evolved to conserve water, such as arid shrubs or savannah,
+should be more likely to reduce ET with increasing VPD, and plants
+that have evolved or have been engineered to care little about water,
+such as crops, will be more likely to increase ET with increasing
+VPD. Atmospheric conditions must matter as well. At the ecosystem
+scale, there are limits to plant water conservation strategies. As
+atmospheric demand for water (VPD) increases, ecosystems should begin
+to reach their water conservation limits and might not be able to
+entirely limit ET flux to the atmosphere. At this stage any further
+increase in VPD will most likely drive a (limited) increase in ET,
+because the increase in atmospheric demand for water overwhelms the
+limited plant response to conserve water.
 
-The objective of the present manuscript is to use reasonable approximations as a tool to develop intuition for plant response to atmospheric drying and evaluate the VPD dependence of ET.
-This intuition will aid interpretation of observations, full complexity models, and facilitate the disentanglement of complex land-atmosphere feedbacks.
-In the past, similiar approaches were used to understand interactions between stomatal conductance, evapotranspiration and the environment \citep[e.g.,][]{Jarvis_1986, Mcnaughton_1991}.
-However, at the time researchers' understandng of the form of VPD's effect on plant physiology was limited, so they could not explore the sensitivity of ET to VPD, including VPD's effect on stomatal conductance and plant function.
+The objective of the present manuscript is to use reasonable
+approximations as a tool to develop intuition for plant response to
+atmospheric drying and evaluate the VPD dependence of ET. This
+intuition will aid interpretation of observations, full complexity
+models, and facilitate the disentanglement of complex land-atmosphere
+feedbacks. In the past, similiar approaches were used to understand
+interactions between stomatal conductance, evapotranspiration and the
+environment \citep[e.g.,][]{Jarvis_1986, Mcnaughton_1991}. However, at
+the time researchers' understandng of the form of VPD's effect on
+plant physiology was limited, so they could not explore the
+sensitivity of ET to VPD, including VPD's effect on stomatal
+conductance and plant function.
 
-Recent results have drastically improved our understanding of VPD's impact on physiology.
-\citet{MEDLYN_2011} developed a model for leaf-scale stomatal conductance ($g_s$), including VPD response, by combining an optimal photosynthesis theory \citep{Farquhar_1980, Katul_2010} with an empirical approach, and extended use of this model to the ecosystem scale in \citet{Medlyn_2017}.
-Additionally, \citet{Zhou_2014} demonstrated that a quantity underlying water use efficiency $\left(uWUE = \frac{GPP\; \sqrt{VPD}}{ET}\right)$ properly captures a constant relationship between GPP, ET, and VPD over a diurnal cycle.
-uWUE is also remarkably well conserved in the growing season across space and time, within a PFT \citep{Zhou_2015}.
-These results provide novell tools to examine how ET varies in response to VPD, while explicitly including VPD's effect on stoamtal conductance for the first time.
-In this manuscript, we leverage these tools to derive the theoretical response of ET to VPD using a Penman Monteith framework.
-Our theory is validated and tested at multiple eddy-covariance stations spanning various climates and plant functional types.
+Recent results have drastically improved our understanding of VPD's
+impact on physiology. \citet{MEDLYN_2011} developed a model for
+leaf-scale stomatal conductance ($g_s$), including VPD response, by
+combining an optimal photosynthesis theory \citep{Farquhar_1980,
+  Katul_2010} with an empirical approach, and extended use of this
+model to the ecosystem scale in \citet{Medlyn_2017}. Additionally,
+\citet{Zhou_2014} demonstrated that a quantity underlying water use
+efficiency $\left(uWUE = \frac{GPP\; \sqrt{VPD}}{ET}\right)$ properly
+captures a constant relationship between GPP, ET, and VPD over a
+diurnal cycle. uWUE is also remarkably well conserved in the growing
+season across space and time, within a PFT \citep{Zhou_2015}. These
+results provide novell tools to examine how ET varies in response to
+VPD, while explicitly including VPD's effect on stoamtal conductance
+for the first time. In this manuscript, we leverage these tools to
+derive the theoretical response of ET to VPD using a Penman Monteith
+framework. Our theory is validated and tested at multiple
+eddy-covariance stations spanning various climates and plant
+functional types.
 
@@
 \label{data}
-We use both meteorological and eddy-covariance data from the FLUXNET2015 database (data available at \sloppy https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/ \sloppy), including all sites with at least four years of data and observations of the variables described in the methods section (Section \ref{methods}).
-Seventy-three sites met these requirements, and were grouped into nine plant functional types (PFT) according to the International Geosphere-Biosphere Programme vegetation classification scheme \citep{Loveland_1999}: cropland (CRO), grass (GRA), deciduous broadleaf forest (DBF), evergreen broadlead forest (EBF), evergreen needleleaf forest (ENF), mixed forest (MF), closed shrub (CSH), savannah (SAV), and woody savannah (WSA) (site locations in Figure \ref{map_fig}).
+We use both meteorological and eddy-covariance data from the
+FLUXNET2015 database (data available at \sloppy
+https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/ \sloppy),
+including all sites with at least four years of data and observations
+of the variables described in the methods section (Section
+\ref{methods}). Seventy-three sites met these requirements, and were
+grouped into nine plant functional types (PFT) according to the
+International Geosphere-Biosphere Programme vegetation classification
+scheme \citep{Loveland_1999}: cropland (CRO), grass (GRA), deciduous
+broadleaf forest (DBF), evergreen broadlead forest (EBF), evergreen
+needleleaf forest (ENF), mixed forest (MF), closed shrub (CSH),
+savannah (SAV), and woody savannah (WSA) (site locations in Figure
+\ref{map_fig}).
 
 \begin{figure}
-\centering
-\includegraphics[width=\textwidth]{./map.pdf}
-\caption{Plant functional type and location of FLUXNET2015 sites used in this analysis.}
+  \centering \includegraphics[width=\textwidth]{./map.pdf}
+  \caption{Plant functional type and location of FLUXNET2015 sites
+    used in this analysis.}
   \label{map_fig}
@@
 
-The purpose of this study is to examine ecosystem response to atmospheric drying, focusing on the growing season.
-To accomplish this, we filter and quality control the data using a similar procedure as \cite{Zhou_2015}:
+The purpose of this study is to examine ecosystem response to
+atmospheric drying, focusing on the growing season. To accomplish
+this, we filter and quality control the data using a similar procedure
+as \cite{Zhou_2015}:
 \begin{itemize}
-\item Only measured or highest (``good'') quality gapfilled data, according to quality control flags, are used.
-\item To isolate the growing season, we only use days in which the average Gross Primary Productivity (GPP) exceeds 10\% of the observed 95th percentile of GPP for a given site.
-GPP is calculated using the nighttime respiration partitioning method.
-\item We remove days with rain and the day following to avoid issues with rain interception and sensor saturation at high relative humidity (\cite{MEDLYN_2011}).
+\item Only measured or highest (``good'') quality gapfilled data,
+  according to quality control flags, are used.
+\item To isolate the growing season, we only use days in which the
+  average Gross Primary Productivity (GPP) exceeds 10\% of the
+  observed 95th percentile of GPP for a given site. GPP is calculated
+  using the nighttime respiration partitioning method.
+\item We remove days with rain and the day following to avoid issues
+  with rain interception and sensor saturation at high relative
+  humidity (\cite{MEDLYN_2011}).
 \end{itemize}
-Additionally, as in \citet{Lin_2018}, we restrict data to the daytime, which is identified when downwelling shortwave radiation is greater than 50 W m$^{-2}$ and sensible heat flux is greater than 5 W m$^{-2}$.
-To reduce the chance of sensor saturation at high relative humidity, we remove all time steps for which VPD is less than .01 kPa, and to reduce errors at low windspeeds we remove all periods with wind magnitudes less than 0.5 m s$^{-1}$.
-Timesteps with negative observed GPP or ET are also removed, and we aggregate half hourly data to hourly averages to reduce noise \citep{Lin_2018}.
-After these quality control procedures, 434,765 upscaled hourly observations remain.
+Additionally, as in \citet{Lin_2018}, we restrict data to the daytime,
+which is identified when downwelling shortwave radiation is greater
+than 50 W m$^{-2}$ and sensible heat flux is greater than 5 W
+m$^{-2}$. To reduce the chance of sensor saturation at high relative
+humidity, we remove all time steps for which VPD is less than .01 kPa,
+and to reduce errors at low windspeeds we remove all periods with wind
+magnitudes less than 0.5 m s$^{-1}$. Timesteps with negative observed
+GPP or ET are also removed, and we aggregate half hourly data to
+hourly averages to reduce noise \citep{Lin_2018}. After these quality
+control procedures, 434,765 upscaled hourly observations remain.
 
@@
 \label{methods}
-The Penman-Monteith equation \citep [hereafter PM,][]{Penman_1948, Monteith_1965} estimates ET as a function of observable atmospheric variables and surface conductances:
+The Penman-Monteith equation \citep [hereafter PM,][]{Penman_1948,
+  Monteith_1965} estimates ET as a function of observable atmospheric
+variables and surface conductances:
 \begin{linenomath*}
@@
 \end{linenomath*}
-where $\Delta$ is the change in saturation vapor pressure with temperature, given by Clausius-Clapeyron ($\frac{d \; e_s}{d \; T}$), $R_{net}$ is the net radiation minus ground heat flux, $g_a$ is aerodynamic conductance, $\rho_a$ is air density, $c_p$ is specific heat of air at constant pressure, $\gamma$ is the psychometric constant, and $g_s$ is the stomatal conductance (Table \ref{definitions}).
+where $\Delta$ is the change in saturation vapor pressure with
+temperature, given by Clausius-Clapeyron ($\frac{d \; e_s}{d \; T}$),
+$R_{net}$ is the net radiation minus ground heat flux, $g_a$ is
+aerodynamic conductance, $\rho_a$ is air density, $c_p$ is specific
+heat of air at constant pressure, $\gamma$ is the psychometric
+constant, and $g_s$ is the stomatal conductance (Table
+\ref{definitions}).
 
 \begin{table}
-  \caption{Definition of symbols and variables, with citation for how values are calculated, if applicable.}
+  \caption{Definition of symbols and variables, with citation for how
+    values are calculated, if applicable.}
   \label{definitions}
-\centering
-\small
+  \centering \small
   \begin{tabular}{l c c c}
@@
 
-\citet{MEDLYN_2011} developed a model for stomatal conductance ($g_s$) by combining an optimal photosynthesis theory \citep{Farquhar_1980, Katul_2010} with an empirical approach, which describes the dependence of $g_s$ to VPD.
-This resulted in the following model for leaf-scale stomatal conductance:
+\citet{MEDLYN_2011} developed a model for stomatal conductance ($g_s$)
+by combining an optimal photosynthesis theory \citep{Farquhar_1980,
+  Katul_2010} with an empirical approach, which describes the
+dependence of $g_s$ to VPD. This resulted in the following model for
+leaf-scale stomatal conductance:
 
@@
 \end{linenomath*}
-where $g_{1-leaf}$ is a leaf-scale ``slope'' parameter, A is the net CO$_2$ assimilation rate, and $c_a$ is the atmospheric CO$_2$ concentration at the leaf surface.
-\cite{MEDLYN_2011} relate the slope parameter ($g_{1-leaf}$) to physical parameters as:
+where $g_{1-leaf}$ is a leaf-scale ``slope'' parameter, A is the net
+CO$_2$ assimilation rate, and $c_a$ is the atmospheric CO$_2$
+concentration at the leaf surface. \cite{MEDLYN_2011} relate the slope
+parameter ($g_{1-leaf}$) to physical parameters as:
 \begin{linenomath*}
@@
 
-where $\Gamma^*$ is the CO$_2$ compensation point for photosynthesis (without dark respiration), and $\lambda$ is the marginal water cost of leaf carbon ($\frac{\partial \; \text{transpiration}}{\partial \; A}$).
-So, $g_{1-leaf}$ is a leaf-scale term reflecting the tradeoff of water for carbon uptake.
-The higher $g_{1-leaf}$, the more open the stomata and the more they release water in exchange for carbon.
+where $\Gamma^*$ is the CO$_2$ compensation point for photosynthesis
+(without dark respiration), and $\lambda$ is the marginal water cost
+of leaf carbon
+($\frac{\partial \; \text{transpiration}}{\partial \; A}$). So,
+$g_{1-leaf}$ is a leaf-scale term reflecting the tradeoff of water for
+carbon uptake. The higher $g_{1-leaf}$, the more open the stomata and
+the more they release water in exchange for carbon.
 
-The Medlyn model for stomatal conductance has been shown to behave very well across PFTs \citep[][]{Lin_2015}, and has been successfully adopted to ecosystem scale analysis in \citet{Medlyn_2017}.
-In units of m s$^{-1}$, the ecosystem scale stomatal conductance is:
+The Medlyn model for stomatal conductance has been shown to behave
+very well across PFTs \citep[][]{Lin_2015}, and has been successfully
+adopted to ecosystem scale analysis in \citet{Medlyn_2017}. In units
+of m s$^{-1}$, the ecosystem scale stomatal conductance is:
 
@@
 
-where GPP is the ecoysystem scale gross primary production, and $g_1$ is an ecosystem scale analogue to $g_{1-leaf}$.
-We solve for $g_1$ following \citet{Medlyn_2017} (Equation 5), and take the median $g_1$ value to be representative of each PFT (Table \ref{pft}).
+where GPP is the ecoysystem scale gross primary production, and $g_1$
+is an ecosystem scale analogue to $g_{1-leaf}$. We solve for $g_1$
+following \citet{Medlyn_2017} (Equation 5), and take the median $g_1$
+value to be representative of each PFT (Table \ref{pft}).
 
 \begin{table}
-  \caption{Plant functional types, their abbreviation, calculated Medlyn coefficient, and calculated uWUE.
-uWUE from \citet{Zhou_2015}, including the  observed standard deviation is shown for comparison.
-Note that uWUE from \citet{Zhou_2015} is calculated from a different set of sites, and that units are converted such that the quantities work with Equations 1-8 and the variables defined Table \ref{definitions}.}
+  \caption{Plant functional types, their abbreviation, calculated
+    Medlyn coefficient, and calculated uWUE. uWUE from
+    \citet{Zhou_2015}, including the observed standard deviation is
+    shown for comparison. Note that uWUE from \citet{Zhou_2015} is
+    calculated from a different set of sites, and that units are
+    converted such that the quantities work with Equations 1-8 and the
+    variables defined Table \ref{definitions}.}
   \small
@@
 
-While Equation 3 can be used in PM (Equation 1), it will make analytical work with the function intractable because $GPP$ is functionally related to ET itself.
-Additionally, a perturbation to VPD should induce a physiological plant response that will alter GPP and cause an indirect change in stomatal conductance, in addition to the direct effect of VPD in Equation \ref{medlyn}.
-Therefore, in order to derive the response of ET to VPD, we must account for the functional relationship between GPP, ET, and VPD, and its effect on stomatal conductance.
-We can use aforementioned semi-empirical results of \citet{Zhou_2015} as a tool to approach this problem.
-\citet{Zhou_2015}, showed that the underlying Water Use Efficiency (uWUE):
+While Equation 3 can be used in PM (Equation 1), it will make
+analytical work with the function intractable because $GPP$ is
+functionally related to ET itself. Additionally, a perturbation to VPD
+should induce a physiological plant response that will alter GPP and
+cause an indirect change in stomatal conductance, in addition to the
+direct effect of VPD in Equation \ref{medlyn}. Therefore, in order to
+derive the response of ET to VPD, we must account for the functional
+relationship between GPP, ET, and VPD, and its effect on stomatal
+conductance. We can use aforementioned semi-empirical results of
+\citet{Zhou_2015} as a tool to approach this
+problem. \citet{Zhou_2015}, showed that the underlying Water Use
+Efficiency (uWUE):
 
@@
 
-is relatively constant across time and moisture conditions within a plant functional type, and correctly captures a constant relationship between GPP, ET and VPD over a diurnal cycle \citep{Zhou_2014}.
-We can use uWUE to remove the $GPP$ dependence from $g_s$ in a way that makes the PM equation analytically tractable:
+is relatively constant across time and moisture conditions within a
+plant functional type, and correctly captures a constant relationship
+between GPP, ET and VPD over a diurnal cycle \citep{Zhou_2014}. We can
+use uWUE to remove the $GPP$ dependence from $g_s$ in a way that makes
+the PM equation analytically tractable:
 
@@
 
-Plugging Equation \ref{new_g_s} into Equation \ref{orig_pen} and rearranging gives a new explicit expression for PM, in which dependence on $GPP$ is removed:
+Plugging Equation \ref{new_g_s} into Equation \ref{orig_pen} and
+rearranging gives a new explicit expression for PM, in which
+dependence on $GPP$ is removed:
 
@@
 
-With Equation \ref{et} we have derived for the first time, using recent results \citep[][]{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017}, ET explicitly as function of environmental variables and two plant-specific constants, the slope parameter ($g_1$), and uWUE, both reflecting water conservation strategy.
-The slope parameter is related to the willingness of stomata to trade water for CO$_2$ and to keep stomata open.
-uWUE is a semi-empirical ecosystem-scale constant related to how WUE changes with VPD (specifically $VPD^{-1/2}$).
-It is also roughly proportional to physical constants:
+With Equation \ref{et} we have derived for the first time, using
+recent results \citep[][]{MEDLYN_2011, Zhou_2014, Zhou_2015,
+  Medlyn_2017}, ET explicitly as function of environmental variables
+and two plant-specific constants, the slope parameter ($g_1$), and
+uWUE, both reflecting water conservation strategy. The slope parameter
+is related to the willingness of stomata to trade water for CO$_2$ and
+to keep stomata open. uWUE is a semi-empirical ecosystem-scale
+constant related to how WUE changes with VPD (specifically
+$VPD^{-1/2}$). It is also roughly proportional to physical constants:
 
@@
 
-where $\Gamma$ is the CO$_2$ compensation point \citep[Equation 5 in][]{Zhou_2014}.
-So uWUE is related to atmospheric CO$_2$ concentration and compensation point, and is inversely proportional to the marginal water cost of leaf carbon.
+where $\Gamma$ is the CO$_2$ compensation point \citep[Equation 5
+in][]{Zhou_2014}. So uWUE is related to atmospheric CO$_2$
+concentration and compensation point, and is inversely proportional to
+the marginal water cost of leaf carbon.
 
-Given eddy-covariance FLUXNET2015 data (Section \ref{data}), every term in our new version of PM (Equation \ref{et}) is observed, except for uWUE, which we fit by calculating its expectation, given the model and FLUXNET2015 data.
+Given eddy-covariance FLUXNET2015 data (Section \ref{data}), every
+term in our new version of PM (Equation \ref{et}) is observed, except
+for uWUE, which we fit by calculating its expectation, given the model
+and FLUXNET2015 data.
 
-However, eddy-covariance data are inherently noisy so we include a measure of uncertainty in our analysis.
-To account for observational error, as well as model uncertainty (e.g.
-temporal and spatial variations of uWUE and $g_1$), we introduce an uncertainty parameter $\sigma$ modifying uWUE:
+However, eddy-covariance data are inherently noisy so we include a
+measure of uncertainty in our analysis. To account for observational
+error, as well as model uncertainty (e.g. temporal and spatial
+variations of uWUE and $g_1$), we introduce an uncertainty parameter
+$\sigma$ modifying uWUE:
 
@@
 
-Now, from each FLUXNET2015 observation (i.e. for each hourly observation at every time step) we can evaluate $\sigma$:
+Now, from each FLUXNET2015 observation (i.e. for each hourly
+observation at every time step) we can evaluate $\sigma$:
 
@@
 
-Thus, with this uncertainty analysis we can evaluate departure from our theory in observations, as a departure of $\sigma$ from unity.
-The variability of $\sigma$ across sites and time provides a measure of uncertainty in our model, assumptions, as well as in the FLUXNET2015 observations themselves.
-The variability of $\sigma$ then propagates through any uncertainty to our derivative of Equation \ref{et_sigma}:
+Thus, with this uncertainty analysis we can evaluate departure from
+our theory in observations, as a departure of $\sigma$ from unity. The
+variability of $\sigma$ across sites and time provides a measure of
+uncertainty in our model, assumptions, as well as in the FLUXNET2015
+observations themselves. The variability of $\sigma$ then propagates
+through any uncertainty to our derivative of Equation \ref{et_sigma}:
 
@@
 
-With Equation \ref{d_et} we have an analytical framework for ecosystem response to atmospheric demand perturbations, with an implicit assumption of non-extreme drought conditions associated with our constant $uWUE$ assumption (e.g.
-we might expect $uWUE$ to decrease in extreme drought).
-There are a few subtleties to taking the derivative in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$) and $VPD$ are functionally related, so while taking the derivative we evaluate $\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET} {\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD} \Big|_{\text{RH fixed}} + \frac{\partial \; ET}{\partial \; RH} \frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$ fixed}}$.
-$RH$ and $e_s$ are assumed to be approximately independent, which is supported by the data (not shown).
+With Equation \ref{d_et} we have an analytical framework for ecosystem
+response to atmospheric demand perturbations, with an implicit
+assumption of non-extreme drought conditions associated with our
+constant $uWUE$ assumption (e.g. we might expect $uWUE$ to decrease in
+extreme drought). There are a few subtleties to taking the derivative
+in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$) and $VPD$ are
+functionally related, so while taking the derivative we evaluate
+$\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET}
+{\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD}
+\Big|_{\text{RH fixed}} + \frac{\partial \; ET}{\partial \; RH}
+\frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$
+    fixed}}$. $RH$ and $e_s$ are assumed to be approximately
+independent, which is supported by the data (not shown).
 
-We note one final comment on our derivation which is relevant for drought indices.
-If we approximate $c_a$ at a global mean CO$_2$ concentration, then the RHS of Equation \ref{et} is fully defined using commonly available weather station data and the constants published here.
-This makes Equation \ref{et} a useful alternative to PET in drought indices and hydrometeorological analysis for vegetated surfaces.
-Equation \ref{et} better reflects the physics of water exchange at the land surface and would only require fitting of uWUE and $g_1$, and can also account for changes in CO$_2$ concentration (assuming some relationship between $\lambda$ and [CO$_2$]), contrary to typical drought indices such as the Palmer Drought Severity Index (PDSI) \citep{Swann_2016}.
+We note one final comment on our derivation which is relevant for
+drought indices. If we approximate $c_a$ at a global mean CO$_2$
+concentration, then the RHS of Equation \ref{et} is fully defined
+using commonly available weather station data and the constants
+published here. This makes Equation \ref{et} a useful alternative to
+PET in drought indices and hydrometeorological analysis for vegetated
+surfaces. Equation \ref{et} better reflects the physics of water
+exchange at the land surface and would only require fitting of uWUE
+and $g_1$, and can also account for changes in CO$_2$ concentration
+(assuming some relationship between $\lambda$ and [CO$_2$]), contrary
+to typical drought indices such as the Palmer Drought Severity Index
+(PDSI) \citep{Swann_2016}.
 
@@
 
-By construction, the variability in the $\sigma$ term (Equation \ref{sigma}) contains all model and observational uncertainties.
-For an observation that perfectly matches our model and constant uWUE assumption, $\sigma$ will be one.
-Therefore, for our assumptions and framework to be reasonable $\sigma$ should be close to 1.
-An additional concern is that $\sigma$ may in fact be correlated with $VPD$, in which case the dependence would need to be accounted for when taking the derivative.
-Fortunately, there is a very weak dependence of $\sigma$ on VPD in their joint distrubution, and $\sigma$ is indeed close to unity i.e.
-$O(1)$ (Figure \ref{joint_vpd_sigma}).
-Given this weak dependence and the distribution of $\sigma$ we have confidence in our model framework and the data quality.
+By construction, the variability in the $\sigma$ term (Equation
+\ref{sigma}) contains all model and observational uncertainties. For
+an observation that perfectly matches our model and constant uWUE
+assumption, $\sigma$ will be one. Therefore, for our assumptions and
+framework to be reasonable $\sigma$ should be close to 1. An
+additional concern is that $\sigma$ may in fact be correlated with
+$VPD$, in which case the dependence would need to be accounted for
+when taking the derivative. Fortunately, there is a very weak
+dependence of $\sigma$ on VPD in their joint distrubution, and
+$\sigma$ is indeed close to unity i.e. $O(1)$ (Figure
+\ref{joint_vpd_sigma}). Given this weak dependence and the
+distribution of $\sigma$ we have confidence in our model framework and
+the data quality.
 
 \begin{figure}
-\centering
-\includegraphics[width=\textwidth]{./joint_vpd_sigma.pdf}
-\caption{The joint distribution of $VPD$ and $\sigma$, with outliers removed (defined as lowest and highest 5\% of $\sigma$).
-$\sigma$ exhibits a weak dependence on $VPD$, and $\sigma$ is $O(1)$ for the bulk of the observations.}
+  \centering \includegraphics[width=\textwidth]{./joint_vpd_sigma.pdf}
+  \caption{The joint distribution of $VPD$ and $\sigma$, with outliers
+    removed (defined as lowest and highest 5\% of $\sigma$). $\sigma$
+    exhibits a weak dependence on $VPD$, and $\sigma$ is $O(1)$ for
+    the bulk of the observations.}
   \label{joint_vpd_sigma}
@@
 
-Before calculating the sensitivity of ET to VPD, we will consider the functional form of Equation \ref{d_et}.
-There are two main terms: a ``scaling'' term, which modifies the magnitude but not the sign of the ET response to VPD  ($\frac{\partial \; ET}{\partial \; VPD}$):
+Before calculating the sensitivity of ET to VPD, we will consider the
+functional form of Equation \ref{d_et}. There are two main terms: a
+``scaling'' term, which modifies the magnitude but not the sign of the
+ET response to VPD ($\frac{\partial \; ET}{\partial \; VPD}$):
 
@@
 
-and a ``sign'' term, which determines whether ET increases or decreases with VPD (i.e. atmospheric demand driven or physiologically controlled):
+and a ``sign'' term, which determines whether ET increases or
+decreases with VPD (i.e. atmospheric demand driven or physiologically
+controlled):
 
@@
 \end{equation}
-All variables are positive, so the relative magnitude between the first term and the second term in the sign term (Equation \ref{sign}) will determine whether ET increases or decreases with increasing VPD.
-If the second term is larger then plant control dominates and ET decreases with increasing VPD.
-However, if the first term is larger, then atmospheric demand dominates and ET increases with increasing VPD.
+All variables are positive, so the relative magnitude between the
+first term and the second term in the sign term (Equation \ref{sign})
+will determine whether ET increases or decreases with increasing
+VPD. If the second term is larger then plant control dominates and ET
+decreases with increasing VPD. However, if the first term is larger,
+then atmospheric demand dominates and ET increases with increasing
+VPD.
 
@@
 \label{sign_term}
-First, we explore the variables within the sign term to gain better intuition on the driver of either the increase or reduction of ET with VPD.
-CO$_2$ concentration ($c_a$) and the psychometric constant ($\gamma$) are relatively constant over the dataset considered here so that the variability is dominated by $\sigma$ and $VPD$.
-uWUE could vary with soil moisture but has been shown to be relatively constant \citep{Zhou_2015}.
-This then means that the sign term only depends on VPD for a given PFT and is approximately just a function of $VPD$.
-We can further determine a critical threshold separating an increase from a decrease in ET, i.e.
-the threshold $VPD_{crit}$ such that the derivative vanishes $\frac{\partial \; ET}{\partial \; VPD} = 0$:
+First, we explore the variables within the sign term to gain better
+intuition on the driver of either the increase or reduction of ET with
+VPD. CO$_2$ concentration ($c_a$) and the psychometric constant
+($\gamma$) are relatively constant over the dataset considered here so
+that the variability is dominated by $\sigma$ and $VPD$. uWUE could
+vary with soil moisture but has been shown to be relatively constant
+\citep{Zhou_2015}. This then means that the sign term only depends on
+VPD for a given PFT and is approximately just a function of $VPD$. We
+can further determine a critical threshold separating an increase from
+a decrease in ET, i.e. the threshold $VPD_{crit}$ such that the
+derivative vanishes $\frac{\partial \; ET}{\partial \; VPD} = 0$:
 \small
@@
 \end{linenomath*}
-\normalsize
-noting that $VPD_{crit}$ mostly depends on the PFT parameters uWUE and $g_1$, and only varies weakly with climate as most other parameters related to the environment are nearly constant.
-The calculated value of $VPD_{crit}$ for each PFT is shown in Table \ref{vpd_crit}.
-For any values of $VPD$ less than $VPD_{crit}$, ET will decrease with increasing VPD ($\frac{\partial \; ET}{\partial \; VPD} < 0$), and for values of $VPD$ greater than $VPD_{crit}$, ET will increase with increasing VPD ($\frac{\partial \; ET}{\partial \; VPD} > 0$).
-In other words, ecosystems regulate and mitigate evaporative losses up to the VPD limit, $VPD_{crit}$, above which atmospheric demand is just too high to be entirely compensated by stomatal and ecosystem regulation.
-We note however that even though ET increases again above the critical threshold, $VPD_{crit}$, ET is still much lower that potential evaporation as stomata are still strongly regulating vapor fluxes to the atmosphere.
-Indeed, even in the absence of soil pore evaporation, stomata do not shut down entirely at very high VPD so that ET does not go to zero, because stomata are still slightly open to perform some photosynthesis \citep{Ball_1987, Leuning_1990, MEDLYN_2011}.
-In addition, upward xylem transport is necessary to maintain phloem transport, as well as nutrient transport and thus carbon allocation \citep{De_2013, Nikinmaa_2013, Ryan_2014}.
+\normalsize noting that $VPD_{crit}$ mostly depends on the PFT
+parameters uWUE and $g_1$, and only varies weakly with climate as most
+other parameters related to the environment are nearly constant. The
+calculated value of $VPD_{crit}$ for each PFT is shown in Table
+\ref{vpd_crit}. For any values of $VPD$ less than $VPD_{crit}$, ET
+will decrease with increasing VPD
+($\frac{\partial \; ET}{\partial \; VPD} < 0$), and for values of
+$VPD$ greater than $VPD_{crit}$, ET will increase with increasing VPD
+($\frac{\partial \; ET}{\partial \; VPD} > 0$). In other words,
+ecosystems regulate and mitigate evaporative losses up to the VPD
+limit, $VPD_{crit}$, above which atmospheric demand is just too high
+to be entirely compensated by stomatal and ecosystem regulation. We
+note however that even though ET increases again above the critical
+threshold, $VPD_{crit}$, ET is still much lower that potential
+evaporation as stomata are still strongly regulating vapor fluxes to
+the atmosphere. Indeed, even in the absence of soil pore evaporation,
+stomata do not shut down entirely at very high VPD so that ET does not
+go to zero, because stomata are still slightly open to perform some
+photosynthesis \citep{Ball_1987, Leuning_1990, MEDLYN_2011}. In
+addition, upward xylem transport is necessary to maintain phloem
+transport, as well as nutrient transport and thus carbon allocation
+\citep{De_2013, Nikinmaa_2013, Ryan_2014}.
 
 \begin{table}
-\caption{Values of $VPD_{crit}$, where $\frac{\partial \; ET}{\partial \; VPD} = 0$, evaluated at PFT average values for $R_{air}$, $\gamma$, and $c_a$.
-PFT-specific constants ($g_1$, uWUE) are provided in Table \ref{pft}.
-For values of $VPD$ less than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be negative, and for values of $VPD$ greater than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be positive.}
+  \caption{Values of $VPD_{crit}$, where
+    $\frac{\partial \; ET}{\partial \; VPD} = 0$, evaluated at PFT
+    average values for $R_{air}$, $\gamma$, and $c_a$. PFT-specific
+    constants ($g_1$, uWUE) are provided in Table \ref{pft}. For
+    values of $VPD$ less than $VPD_{crit}$,
+    $\frac{\partial \; ET}{\partial \; VPD}$ will be negative, and for
+    values of $VPD$ greater than $VPD_{crit}$,
+    $\frac{\partial \; ET}{\partial \; VPD}$ will be positive.}
   \centering
@@
 
-Differences in $VPD_{crit}$ are exclusively determined by uWUE and the slope parameter ($g_1$) related to the plant functional type.
-A larger uWUE means a smaller $VPD_{crit}$, and an ET response to increases VPD that is more likely to be positive.
-At first glance this result is somewhat counter-intuitive; we expect that plants with a higher water use efficiency would be more water conservative.
-However, in reality uWUE determines how $WUE$ changes with VPD:
+Differences in $VPD_{crit}$ are exclusively determined by uWUE and the
+slope parameter ($g_1$) related to the plant functional type. A larger
+uWUE means a smaller $VPD_{crit}$, and an ET response to increases VPD
+that is more likely to be positive. At first glance this result is
+somewhat counter-intuitive; we expect that plants with a higher water
+use efficiency would be more water conservative. However, in reality
+uWUE determines how $WUE$ changes with VPD:
 
 \[WUE = \frac{GPP}{ET} = \frac{uWUE}{\sqrt{VPD}}.\]
-\[\frac{\partial \; WUE}{\partial \; VPD} = -\frac{uWUE}{2 \; VPD^{3/2}}\]
+\[\frac{\partial \; WUE}{\partial \; VPD} = -\frac{uWUE}{2 \;
+    VPD^{3/2}}\]
 
-So, plants with a higher uWUE will have a greater \textit{decrease} in ecosystem-scale $WUE$ in response to increases in VPD.
-This decrease in $WUE$ causes more water loss per unit carbon gain, and explains the relationship between high uWUE and high likelihood of increases of ET in response to increasing atmospheric drying (increases in VPD).
+So, plants with a higher uWUE will have a greater \textit{decrease} in
+ecosystem-scale $WUE$ in response to increases in VPD. This decrease
+in $WUE$ causes more water loss per unit carbon gain, and explains the
+relationship between high uWUE and high likelihood of increases of ET
+in response to increasing atmospheric drying (increases in VPD).
 
-A tendency towards increasing ET response with increasing VPD can also be caused by a high slope parameter ($g_1$), characteristic of plants that at the leaf scale are more willing to trade water for access to atmospheric CO$_2$.
-Plants that are less conservative will be thus be more likely to increase ET with increasing VPD.
-Both the aforementioned effects (large uWUE, $g_1$) can amplify each other, and generally conspire to shift the sign term towards a positive value for a given PFT.
+A tendency towards increasing ET response with increasing VPD can also
+be caused by a high slope parameter ($g_1$), characteristic of plants
+that at the leaf scale are more willing to trade water for access to
+atmospheric CO$_2$. Plants that are less conservative will be thus be
+more likely to increase ET with increasing VPD. Both the
+aforementioned effects (large uWUE, $g_1$) can amplify each other, and
+generally conspire to shift the sign term towards a positive value for
+a given PFT.
 
-This effect of uWUE and $g1$ on the sign term is most apparent by comparing two extreme PFTs: water intensive crops (CRO) and water conservative closed shrub (CSH).
-CRO has higher slope parameter and a slightyly higher uWUE ($g_1 = 140.7$ Pa$^{1/2}$; $2.85$ $\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET]) compared to CSH ($g1 = 75.1 \, Pa^{1/2}$, $uWUE=2.82$ $\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET]).
-These differences in PFT parameters cause opposite ET responses to changes in VPD between CRO and CSH.
-ET theoretically always decreases with increasing VPD for the more water conservative CSH, while ET frequently  increases with increasing VPD for the more water intensive CRO (Figure \ref{idealized_sign}).
-CRO evolved or were bred to prioritize GPP and yield and are thus not water conservative.
-They are very willing to trade water for photosynthesis and productivity, despite changes in VPD, while CSH are very unwilling to trade water for more photosynthesis.
+This effect of uWUE and $g1$ on the sign term is most apparent by
+comparing two extreme PFTs: water intensive crops (CRO) and water
+conservative closed shrub (CSH). CRO has higher slope parameter and a
+slightyly higher uWUE ($g_1 = 140.7$ Pa$^{1/2}$; $2.85$ $\mu$-mol [C]
+Pa$^{0.5}$ J$^{-1}$ [ET]) compared to CSH ($g1 = 75.1 \, Pa^{1/2}$,
+$uWUE=2.82$ $\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET]). These differences
+in PFT parameters cause opposite ET responses to changes in VPD
+between CRO and CSH. ET theoretically always decreases with increasing
+VPD for the more water conservative CSH, while ET frequently increases
+with increasing VPD for the more water intensive CRO (Figure
+\ref{idealized_sign}). CRO evolved or were bred to prioritize GPP and
+yield and are thus not water conservative. They are very willing to
+trade water for photosynthesis and productivity, despite changes in
+VPD, while CSH are very unwilling to trade water for more
+photosynthesis.
 
 \begin{figure}
-\centering
-\includegraphics[width=\textwidth]{./idealized_sign.pdf}
-\caption{The functional form of the sign term, with $\sigma$ held fixed at 1, and all terms except VPD set to PFT averages.
-For comparison, the observed range of VPD for each PFT is plotted below the x-axis.
-Stars denote 25th, 50th, and 75th percentiles, and the range of the line spans the 5th-95th percentiles of observed VPD.
-Vertical black lines denote the location of $VPD_{crit}$ for each PFT, with the exception of CSH, for which $VPD_{crit}$ is off-scale.}
+  \centering \includegraphics[width=\textwidth]{./idealized_sign.pdf}
+  \caption{The functional form of the sign term, with $\sigma$ held
+    fixed at 1, and all terms except VPD set to PFT averages. For
+    comparison, the observed range of VPD for each PFT is plotted
+    below the x-axis. Stars denote 25th, 50th, and 75th percentiles,
+    and the range of the line spans the 5th-95th percentiles of
+    observed VPD. Vertical black lines denote the location of
+    $VPD_{crit}$ for each PFT, with the exception of CSH, for which
+    $VPD_{crit}$ is off-scale.}
   \label{idealized_sign}
@@
 
-As expected, the slope parameter ($g_1$) is a primary determinant of the VPD dependence for the sign term shown in Figure \ref{idealized_sign}.
-Plants that are more conservative (small $g_1$) will tend to reduce ET with increasing VPD, and will be very effective at reducing ET, especially at low VPD.
-However, at very high VPD, gradients in vapor pressure at the leaf scale will become very strong as stomata reach their limits of closure in response to VPD (parameterized with $g_1$).
-As a result, ET response will begin to asymptote towards a constant ecosystem-scale values as leaf-scale response to VPD asymptotes towards zero.
-Therefore, plants with a low $g_1$ will have the largest VPD dependence of ET response because the difference in ET response at low VPD (leaf stomatal response dominates) and high VPD (VPD gradient dominates) is largest.
-This is apparent in the strong VPD dependence of CSH, which has the lowest slope parameter ($g_1=75.1$ Pa$^{1/2}$) (Figure \ref{idealized_sign}).
+As expected, the slope parameter ($g_1$) is a primary determinant of
+the VPD dependence for the sign term shown in Figure
+\ref{idealized_sign}. Plants that are more conservative (small $g_1$)
+will tend to reduce ET with increasing VPD, and will be very effective
+at reducing ET, especially at low VPD. However, at very high VPD,
+gradients in vapor pressure at the leaf scale will become very strong
+as stomata reach their limits of closure in response to VPD
+(parameterized with $g_1$). As a result, ET response will begin to
+asymptote towards a constant ecosystem-scale values as leaf-scale
+response to VPD asymptotes towards zero.  Therefore, plants with a low
+$g_1$ will have the largest VPD dependence of ET response because the
+difference in ET response at low VPD (leaf stomatal response
+dominates) and high VPD (VPD gradient dominates) is largest. This is
+apparent in the strong VPD dependence of CSH, which has the lowest
+slope parameter ($g_1=75.1$ Pa$^{1/2}$) (Figure \ref{idealized_sign}).
 
-To summarize our theoretical insights (Figure \ref{idealized_sign} and Table \ref{vpd_crit}), CROs are the least water conservative and have the strongest overall tendency to increase ET with increasing VPD, while CSH are the most water conservative and have the strongest tendency to decrease ET with increasing VPD, as well as the strongest VPD dependence of response.
-Figure \ref{idealized_sign} clearly shows, according to our theory, that for all PFTs except for crops there is frequent occurrence of a negative (plant dominating) ET response to increases in VPD.
-Therefore, plants are able in most atmospheric conditions to reduce ET in response to increased VPD and thus to reduce water loss.
-To better illustrate this, the ranges of observed environmental VPDs at the FLUXNET sites are plotted parallel to the x-axis.
-For CSH, VPD is always less than VPD$_{crit}$ (off scale) so that the plant response dominates in typical environmental conditions, emphasizing the water conservative strategy of those plants.
-For CRO on the other hand, VPD is higher than VPD$_{crit}$ for more than 50\% of observations, emphasizing that those plants operate with an aggressive water usage strategy, are water intensive and were actually engineered for photosynthesis rather than water saving.
-For DBF, EBF, MF, GRA, and SAV  more than half of the observed VPD are less than VPD$_{crit}$, i.e.
-in conditions where plant response dominates.
-SAV has a more water conservative response than the forest, grass, and crop plan types, but still respondes by increasing ET with increasing VPD for about a quarter of observations, due to the high aridity (VPD) of the SAV ecoclimate.
-It is also important to note that for all PFTs, even when atmospheric demand dominates, ET response to VPD is still far more negative than it would be for potential evaporation $\partial PET/\partial VPD$, i.e.
-atmospheric demand only, emphasizing that there is still a strong regulation of evaporative flux by stomata and though the plant xylem.
-The sign term in the PET case would just be a constant ($\frac{c_p}{R_{air}} \approx 3.5$), which is far larger than any part of the curves for any PFT.
-This highlights the deficiencies of PET's ability to capture land response to changes in atmospheric dryness (VPD).
-Plants are always regulating water exchange from the land surface, even when they reach the limits of they ability to do so.
+To summarize our theoretical insights (Figure \ref{idealized_sign} and
+Table \ref{vpd_crit}), CROs are the least water conservative and have
+the strongest overall tendency to increase ET with increasing VPD,
+while CSH are the most water conservative and have the strongest
+tendency to decrease ET with increasing VPD, as well as the strongest
+VPD dependence of response. Figure \ref{idealized_sign} clearly shows,
+according to our theory, that for all PFTs except for crops there is
+frequent occurrence of a negative (plant dominating) ET response to
+increases in VPD. Therefore, plants are able in most atmospheric
+conditions to reduce ET in response to increased VPD and thus to
+reduce water loss. To better illustrate this, the ranges of observed
+environmental VPDs at the FLUXNET sites are plotted parallel to the
+x-axis. For CSH, VPD is always less than VPD$_{crit}$ (off scale) so
+that the plant response dominates in typical environmental conditions,
+emphasizing the water conservative strategy of those plants. For CRO
+on the other hand, VPD is higher than VPD$_{crit}$ for more than 50\%
+of observations, emphasizing that those plants operate with an
+aggressive water usage strategy, are water intensive and were actually
+engineered for photosynthesis rather than water saving. For DBF, EBF,
+MF, GRA, and SAV more than half of the observed VPD are less than
+VPD$_{crit}$, i.e. in conditions where plant response dominates. SAV
+has a more water conservative response than the forest, grass, and
+crop plan types, but still respondes by increasing ET with increasing
+VPD for about a quarter of observations, due to the high aridity (VPD)
+of the SAV ecoclimate. It is also important to note that for all PFTs,
+even when atmospheric demand dominates, ET response to VPD is still
+far more negative than it would be for potential evaporation
+$\partial PET/\partial VPD$, i.e. atmospheric demand only, emphasizing
+that there is still a strong regulation of evaporative flux by stomata
+and though the plant xylem. The sign term in the PET case would just
+be a constant ($\frac{c_p}{R_{air}} \approx 3.5$), which is far larger
+than any part of the curves for any PFT. This highlights the
+deficiencies of PET's ability to capture land response to changes in
+atmospheric dryness (VPD). Plants are always regulating water exchange
+from the land surface, even when they reach the limits of they ability
+to do so.
 
@@
 \label{scale_term}
-While the above discussion of the sign of $\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our question of when ET response increases or decreases with VPD, understating the overall magnitude of the ET response is important to soil-plant-atmosphere water budgeting.
-So we now more closely examine the terms that affect how the sign term is scaled:
+While the above discussion of the sign of
+$\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our
+question of when ET response increases or decreases with VPD,
+understating the overall magnitude of the ET response is important to
+soil-plant-atmosphere water budgeting. So we now more closely examine
+the terms that affect how the sign term is scaled:
 
@@
 
-$\frac{P}{T}$ is an air-density term, which varies little compared to aerodynamic conductance and Clausius-Clapeyron ($\Delta$).
-The psychometric constant ($\gamma$) is also relatively constant, so the scaling term should be primarily a function of aerodynamic conductance and temperature, through the Clausius-Clapeyron relationship $\Delta$.
-This is as expected, given that the aerodynamic conductance represents the efficiency of exchange between the surface and the atmosphere.
-As aerodynamic conductance  increases, any plant response will be communicated more strongly to the atmosphere (and vice-versa).
+$\frac{P}{T}$ is an air-density term, which varies little compared to
+aerodynamic conductance and Clausius-Clapeyron ($\Delta$). The
+psychometric constant ($\gamma$) is also relatively constant, so the
+scaling term should be primarily a function of aerodynamic conductance
+and temperature, through the Clausius-Clapeyron relationship
+$\Delta$. This is as expected, given that the aerodynamic conductance
+represents the efficiency of exchange between the surface and the
+atmosphere. As aerodynamic conductance increases, any plant response
+will be communicated more strongly to the atmosphere (and vice-versa).
 
-$\Delta$'s presence in the scaling term also matches physical intuition.
-$\Delta$ (and also the approximately constant $\gamma$) control the efficiencies with which surface energy is converted to latent and sensible heat \citep{Monteith_1965}.
-The functional from of $\Delta$ will be the same across PFTs, but the temperature range may vary slightly.
-In contrast, aerodynamic conductance will vary strongly with PFT due to the importance of surface roughness for aerodynamic conductance.
-So most of the differences in scaling between PFT should be in the aerodynamic conductance term.
+$\Delta$'s presence in the scaling term also matches physical
+intuition. $\Delta$ (and also the approximately constant $\gamma$)
+control the efficiencies with which surface energy is converted to
+latent and sensible heat \citep{Monteith_1965}. The functional from of
+$\Delta$ will be the same across PFTs, but the temperature range may
+vary slightly. In contrast, aerodynamic conductance will vary strongly
+with PFT due to the importance of surface roughness for aerodynamic
+conductance. So most of the differences in scaling between PFT should
+be in the aerodynamic conductance term.
 
-The control of the scaling term variability between PFTs by aerodynamic conductance is confirmed by data (Figure \ref{scale_vary}).
-Differences between PFT are almost entirely due to differences in aerodynamic conductance, rather than differences in observed temperature ranges.
-The scaling term for the tree PFTs (DBF, EBF, ENF, MF) is generally about double the scaling terms for other PFTs which have lower surface roughness and generally smaller aerodynamic conductance (GRA, CSH, CRO).
-The savannah (WSA, SAV) PFT's scaling is somewhere betwen GRA, CSH, and CRO, and DBF, EBF, ENF, and MF, due to higher variability and surface roughness.
+The control of the scaling term variability between PFTs by
+aerodynamic conductance is confirmed by data (Figure
+\ref{scale_vary}). Differences between PFT are almost entirely due to
+differences in aerodynamic conductance, rather than differences in
+observed temperature ranges. The scaling term for the tree PFTs (DBF,
+EBF, ENF, MF) is generally about double the scaling terms for other
+PFTs which have lower surface roughness and generally smaller
+aerodynamic conductance (GRA, CSH, CRO). The savannah (WSA, SAV) PFT's
+scaling is somewhere betwen GRA, CSH, and CRO, and DBF, EBF, ENF, and
+MF, due to higher variability and surface roughness.
 
 \begin{figure}
-%\centering
+  % \centering
   \centerline{\includegraphics[width=1.4\textwidth]{./idealized_scale.pdf}}
-\caption{Primary sources of variability for the scaling term, as a function of PFT.
-The 5th-95th percentile range of temperature is plotted at the 5th, 25th, 50th, 75th, and 95th percentiles of aerodynamic conductance, as observed for each PFT.}
+  \caption{Primary sources of variability for the scaling term, as a
+    function of PFT. The 5th-95th percentile range of temperature is
+    plotted at the 5th, 25th, 50th, 75th, and 95th percentiles of
+    aerodynamic conductance, as observed for each PFT.}
   \label{scale_vary}
@@
 
-Within each PFT, the scaling term variability is controlled both by environmental temperature and aerodynamic conductance variability (Figure \ref{scale_vary}).
-While the observed variability of the aerodynamic conductance contributes more to the scaling term variability than temperature, the temperature contribution is non-negligible.
-Specifically, the scaling term is generally larger at low temperatures when latent heat is relatively inefficient at moving energy away from the surface.
-This effect amplifies the role of aerodynamic conductance variability at low temperatures.
+Within each PFT, the scaling term variability is controlled both by
+environmental temperature and aerodynamic conductance variability
+(Figure \ref{scale_vary}). While the observed variability of the
+aerodynamic conductance contributes more to the scaling term
+variability than temperature, the temperature contribution is
+non-negligible. Specifically, the scaling term is generally larger at
+low temperatures when latent heat is relatively inefficient at moving
+energy away from the surface. This effect amplifies the role of
+aerodynamic conductance variability at low temperatures.
 
-To summarize, variability between PFTs is mostly controlled by systematic differences in aerodynamic conductance, due to differences in surface roughness between each PFT, and possibly to a lesser extent wind conditions.
-In contrast, variability within PFT is also controlled by temperature, through Clausius-Clapeyron.
-But, aerodynamic conductance variability generally impacts the scaling term more than temperature, even within PFT.
+To summarize, variability between PFTs is mostly controlled by
+systematic differences in aerodynamic conductance, due to differences
+in surface roughness between each PFT, and possibly to a lesser extent
+wind conditions. In contrast, variability within PFT is also
+controlled by temperature, through Clausius-Clapeyron. But,
+aerodynamic conductance variability generally impacts the scaling term
+more than temperature, even within PFT.
 
@@
 
-In this section we consider direct observations of ET response with eddy-covariance data, while including uncertainty with the $\sigma$ term (Section \ref{methods}).
-These observational results of ET response largely confirm our theoretical analysis, presented above (Table \ref{stats}).
-For all PFTs, mean ET response to increasing VPD is negative.
-However, ET response evaluated at the average of all variables (e.g.
-$\sigma$, $T$, $c_a$, $VPD$) is positive for CRO, and negative for all other PFTs.
-This difference in mean ET response as compared to the ET response at mean environmental conditions is due to the non-linear nature of the response, in which negative responses are generally larger magnitude than positive responses (Figure \ref{idealized_sign}).
-Therefore, both the mean ET response as well as the ET response at mean environmental conditions matches our expectations from the theory (Section \ref{sign_term}), with the exception that CRO observations are shifted more towards a negative ET response than we expect.
+In this section we consider direct observations of ET response with
+eddy-covariance data, while including uncertainty with the $\sigma$
+term (Section \ref{methods}). These observational results of ET
+response largely confirm our theoretical analysis, presented above
+(Table \ref{stats}). For all PFTs, mean ET response to increasing VPD
+is negative. However, ET response evaluated at the average of all
+variables (e.g. $\sigma$, $T$, $c_a$, $VPD$) is positive for CRO, and
+negative for all other PFTs. This difference in mean ET response as
+compared to the ET response at mean environmental conditions is due to
+the non-linear nature of the response, in which negative responses are
+generally larger magnitude than positive responses (Figure
+\ref{idealized_sign}). Therefore, both the mean ET response as well as
+the ET response at mean environmental conditions matches our
+expectations from the theory (Section \ref{sign_term}), with the
+exception that CRO observations are shifted more towards a negative ET
+response than we expect.
 
 \begin{table}
-\caption{Statistics of $\frac{\partial \; ET}{\partial \; VPD}$ as a function of PFT.}
+  \caption{Statistics of $\frac{\partial \; ET}{\partial \; VPD}$ as a
+    function of PFT.}
   \centering
@@
 
-Regarding the frequency of negative and positive ET response, all PFTs exhibit a  decreasing ET response with increasing VPD (physiologically controlled, water conservative response) for the majority of observations.
-The more water conservative PFTs generally exhibit higher frequency of negative ET response, espcially when one factors in the distribution of environmental VPD (e.g.
-SAV and WSA grow in more arid climates, Figure \ref{idealized_sign}).
-In general the bulk statistics match our theoretical expectations well, with the caveat that inclusion of uncertainty shifts crops towards a slightly more negative response, and shifts many of the other PFTs, which still exhibit a high frequency of negative ET reponse, towards more frequent occurence of postive ET response than the theory and Figure \ref{idealized_sign} would suggest.
-The bulk statistics motivate a more thorough examination of the structure of uncertainty and more sophisticated validation of our theory's performance against observations.
+Regarding the frequency of negative and positive ET response, all PFTs
+exhibit a decreasing ET response with increasing VPD (physiologically
+controlled, water conservative response) for the majority of
+observations. The more water conservative PFTs generally exhibit
+higher frequency of negative ET response, espcially when one factors
+in the distribution of environmental VPD (e.g. SAV and WSA grow in
+more arid climates, Figure \ref{idealized_sign}). In general the bulk
+statistics match our theoretical expectations well, with the caveat
+that inclusion of uncertainty shifts crops towards a slightly more
+negative response, and shifts many of the other PFTs, which still
+exhibit a high frequency of negative ET reponse, towards more frequent
+occurence of postive ET response than the theory and Figure
+\ref{idealized_sign} would suggest. The bulk statistics motivate a
+more thorough examination of the structure of uncertainty and more
+sophisticated validation of our theory's performance against
+observations.
 
@@
 \label{testing}
-We now compare more sophisticated distributions of the observed response to our simplified theory (Section \ref{sign_term}).
-The observed distribution of the sign term, as compared to what the theory would predict, is provided in Figure \ref{test_sign}.
-Our goal was to capture the leading order behavior of the ET dependence on VPD.
-Given the assumptions we have made, and the uncertainties of flux tower observations themselves, we expect a relatively large amount of noise when reproducing the derivatives of ET.
-However, the data largely reproduces our theoretical analysis.
+We now compare more sophisticated distributions of the observed
+response to our simplified theory (Section \ref{sign_term}). The
+observed distribution of the sign term, as compared to what the theory
+would predict, is provided in Figure \ref{test_sign}. Our goal was to
+capture the leading order behavior of the ET dependence on VPD. Given
+the assumptions we have made, and the uncertainties of flux tower
+observations themselves, we expect a relatively large amount of noise
+when reproducing the derivatives of ET. However, the data largely
+reproduces our theoretical analysis.
 
@@
   \centerline{\includegraphics[width=1.4\textwidth]{./test_sign.pdf}}
-\caption{Comparison of the sign term with model uncertainty included (box plots) to the sign term as calculated with simplifying assumptions (blue line, as in Figure \ref{idealized_sign}).
-Each box plot corresponds to 5\% of the data, and the 5-95\% range of VPD is plotted.}
+  \caption{Comparison of the sign term with model uncertainty included
+    (box plots) to the sign term as calculated with simplifying
+    assumptions (blue line, as in Figure \ref{idealized_sign}). Each
+    box plot corresponds to 5\% of the data, and the 5-95\% range of
+    VPD is plotted.}
   \label{test_sign}
@@
 
-This is particularly true for DBF and MF; the theory matches the leading order behavior of the function when uncertainty is included, and the observations match the theory with the addition of noise.
-The VPD dependence, given by the slope parameter ($g_1$), follows the median values of each bin.
-Perhaps most importantly, the x-intercept, and thus VPD$_{crit}$, matches nearly exactly between the theory and the observations.
-Therefore the sign of the ET response to increases in VPD should be well matched, subject to the unavoidable constraints of noise, much of which comes from the observations themselves.
-The uncertainty is non negligible; there are many observations in each bin for which the the sign of observation is opposite the response predicted by the theory, but to leading order our theory matches the observations well.
+This is particularly true for DBF and MF; the theory matches the
+leading order behavior of the function when uncertainty is included,
+and the observations match the theory with the addition of noise. The
+VPD dependence, given by the slope parameter ($g_1$), follows the
+median values of each bin. Perhaps most importantly, the x-intercept,
+and thus VPD$_{crit}$, matches nearly exactly between the theory and
+the observations. Therefore the sign of the ET response to increases
+in VPD should be well matched, subject to the unavoidable constraints
+of noise, much of which comes from the observations themselves. The
+uncertainty is non negligible; there are many observations in each bin
+for which the the sign of observation is opposite the response
+predicted by the theory, but to leading order our theory matches the
+observations well.
 
-While CSH has a much different functional form of the sign term than DBF and MF, CSH observations also match our theory to leading order, albiet with a bit more variability as a function of VPD.
-Again, the VPD dependence mostly determined from the slope parameter ($g_1$) closely matches the medians in the observation bins.
-The VPD-independent, strongly negative response is also captured.
-For CSH, there is rare occurrence of observed positive ET response with VPD ($\approx 6$\%), even with uncertainty, so the sign of the observations almost always matches the sign of the theory, which states that ET response should always be negative.
+While CSH has a much different functional form of the sign term than
+DBF and MF, CSH observations also match our theory to leading order,
+albiet with a bit more variability as a function of VPD. Again, the
+VPD dependence mostly determined from the slope parameter ($g_1$)
+closely matches the medians in the observation bins. The
+VPD-independent, strongly negative response is also captured. For CSH,
+there is rare occurrence of observed positive ET response with VPD
+($\approx 6$\%), even with uncertainty, so the sign of the
+observations almost always matches the sign of the theory, which
+states that ET response should always be negative.
 
-Biases between the theory and observations are similiar for ENF, WSA, CRO and SAV.
-At low VPD the theory and the observations match well.
-However, at high VPD (upper 10-20\% of observations) the theory is biased slightly towards postive reponse as compared to the observations (Figure \ref{test_sign}).
-For ENF this slight bias could be explained by a negative bias in $g_1$.
-However, for WSA, CRO, and SAV the observations in the highest VPD bins exhibit a downturn towards more negative response, which cannot be captured by the functional form of the sign term (Figure \ref{idealized_sign}).
-Therefore, to explain the observations one of the varibles in the sign term must change at extreme VPD (upper 10-20\%).
-The most likely candidiate is $uWUE$, which we have assumed constant to meet our goal of developing intuition for leading order behavior, but might be expected to decrease at extremely low SWC, which should be correlated with high VPD.
-The theory for EBF exhibits similiar limiatations as for ENF, WSA, and SAV, except a larger portion of observations are biased negative as compared to the theory ($\approx$ 35\%, Figure \ref{test_sign}).
-However, in general, and specifically for non-extreme VPD, the theory matches the observations for the tree and savannah plant types.
+Biases between the theory and observations are similiar for ENF, WSA,
+CRO and SAV. At low VPD the theory and the observations match
+well. However, at high VPD (upper 10-20\% of observations) the theory
+is biased slightly towards postive reponse as compared to the
+observations (Figure \ref{test_sign}). For ENF this slight bias could
+be explained by a negative bias in $g_1$. However, for WSA, CRO, and
+SAV the observations in the highest VPD bins exhibit a downturn
+towards more negative response, which cannot be captured by the
+functional form of the sign term (Figure
+\ref{idealized_sign}). Therefore, to explain the observations one of
+the varibles in the sign term must change at extreme VPD (upper
+10-20\%). The most likely candidiate is $uWUE$, which we have assumed
+constant to meet our goal of developing intuition for leading order
+behavior, but might be expected to decrease at extremely low SWC,
+which should be correlated with high VPD. The theory for EBF exhibits
+similiar limiatations as for ENF, WSA, and SAV, except a larger
+portion of observations are biased negative as compared to the theory
+($\approx$ 35\%, Figure \ref{test_sign}). However, in general, and
+specifically for non-extreme VPD, the theory matches the observations
+for the tree and savannah plant types.
 
-The theory for GRA suffers from similar, but much more severe, limitations as for CRO, WSA, and SAV.
-GRA observations are characterized by a consistent trend back towards negative ET response at higher VPD, which the functional form of our theory is incapable of accounting for.
-As compared to CRO, WSA and SAV, the divergence between the theory and the observations is far greater for GRA, biasing 40-50\% of observations at higher VPD.
-In addition for the potential for soil moisture to alter $uWUE$, there are other sources of plant heterogeneity specific to GRA (and some extent CRO) that may alter $uWUE$ (or $g_1$) or invalidate other assumptions made in the Methods section (Section \ref{methods}).
-We do not account for variability in plant height and surface roughness, or differences in C3 vs C4 photosynthesis and water strategies, which we might expect to vary substantially across sites, years, and season for GRA.
-These deficiencies could largely explain the inability of our theory to exactly match the observations in croplands.
-For example, a superposition of sites with C3 plants (at low environmental VPD) and C4 plants (at high environmental VPD) would explain observed shift back towards negative ET response at high VPD when all sites are binned together, as in Figure \ref{test_sign}.
-We hypothesize that the theory would validate against observations much better if these sources of varaibility were accounted for, at a cost of increased complexity and analytical opacity.
+The theory for GRA suffers from similar, but much more severe,
+limitations as for CRO, WSA, and SAV. GRA observations are
+characterized by a consistent trend back towards negative ET response
+at higher VPD, which the functional form of our theory is incapable of
+accounting for. As compared to CRO, WSA and SAV, the divergence
+between the theory and the observations is far greater for GRA,
+biasing 40-50\% of observations at higher VPD. In addition for the
+potential for soil moisture to alter $uWUE$, there are other sources
+of plant heterogeneity specific to GRA (and some extent CRO) that may
+alter $uWUE$ (or $g_1$) or invalidate other assumptions made in the
+Methods section (Section \ref{methods}). We do not account for
+variability in plant height and surface roughness, or differences in
+C3 vs C4 photosynthesis and water strategies, which we might expect to
+vary substantially across sites, years, and season for GRA. These
+deficiencies could largely explain the inability of our theory to
+exactly match the observations in croplands. For example, a
+superposition of sites with C3 plants (at low environmental VPD) and
+C4 plants (at high environmental VPD) would explain observed shift
+back towards negative ET response at high VPD when all sites are
+binned together, as in Figure \ref{test_sign}. We hypothesize that the
+theory would validate against observations much better if these
+sources of varaibility were accounted for, at a cost of increased
+complexity and analytical opacity.
 
-While the above discussion shows that our theory has some limitations when applied to some PFTs, especially for grasslands, it does well for DBF, MF and CSH PFTs, and captures the response at non-extreme VPD for ENF, EBF, CRO, WSA, and SAV.
-In general, the leading order behavior observed in the data is captured by the theory.
-Departures between our theory and observations, specifically at high VPD, could be explained and conceptualized with shifts in $g_1$ and/or $uWUE$ due to site-spcific plant type variability (e.g.
-more arid-adapted ecosystems at more arid sites), or temporal variability for some enviornmental conditions (e.g.
-decreases in $uWUE$ at extremely low soil moisture).
-To focus on general behaviro and develop intuition for PFT-scale response, we have ignored these sources of variability in the present analysis.
-However, any site-specific or temporal plant functional variability that can be conceptualized with shifts to $g_1$ and $uWUE$ can be analyzed with our framework.
-This opens a door to a myriad future analysis in which plant behavior in anomalous conditions can be explained and analyzed using Equation \ref{et}.
+While the above discussion shows that our theory has some limitations
+when applied to some PFTs, especially for grasslands, it does well for
+DBF, MF and CSH PFTs, and captures the response at non-extreme VPD for
+ENF, EBF, CRO, WSA, and SAV. In general, the leading order behavior
+observed in the data is captured by the theory. Departures between our
+theory and observations, specifically at high VPD, could be explained
+and conceptualized with shifts in $g_1$ and/or $uWUE$ due to
+site-spcific plant type variability (e.g. more arid-adapted ecosystems
+at more arid sites), or temporal variability for some enviornmental
+conditions (e.g. decreases in $uWUE$ at extremely low soil
+moisture). To focus on general behaviro and develop intuition for
+PFT-scale response, we have ignored these sources of variability in
+the present analysis. However, any site-specific or temporal plant
+functional variability that can be conceptualized with shifts to $g_1$
+and $uWUE$ can be analyzed with our framework. This opens a door to a
+myriad future analysis in which plant behavior in anomalous conditions
+can be explained and analyzed using Equation \ref{et}.
 
-% It is useful to also compare our theory, the observed uncertainty in the ET response, and PET, as PET is often used as a proxy for ET in hydrometeorological studies and drought indices.
-Again, as in our discussion of the theoretical results, the PET sensitivity to VPD will be constant at about 3.5.
-In Figure \ref{test_sign} the ET response to shifts in dryness is always far smaller than 3.5 in the observations, even with all uncertainty included.
-So the PET response to atmospheric dryness is outside the range of all observable responses.
-This casts doubt on PET's usefulness for capturing the leading order behavior of water budget and dryness response at the land surface, as is done in several drought indices.
-However our theory (Equation \ref{et}) could be used with published uWUE and $g1$ data, as an alternative to PET to estimating land response and dryness, and would more accurately account for the response to atmospheric drying.
+% It is useful to also compare our theory, the observed uncertainty in
+% the ET response, and PET, as PET is often used as a proxy for ET in
+% hydrometeorological studies and drought indices. Again, as in our
+% discussion of the theoretical results, the PET sensitivity to VPD
+% will be constant at about 3.5. In Figure \ref{test_sign} the ET
+% response to shifts in dryness is always far smaller than 3.5 in the
+% observations, even with all uncertainty included. So the PET
+% response to atmospheric dryness is outside the range of all
+% observable responses. This casts doubt on PET's usefulness for
+% capturing the leading order behavior of water budget and dryness
+% response at the land surface, as is done in several drought
+% indices. However our theory (Equation \ref{et}) could be used with
+% published uWUE and $g1$ data, as an alternative to PET to estimating
+% land response and dryness, and would more accurately account for the
+% response to atmospheric drying.
 
@@
 
-Most of the results presented so far focus on the sign term, so now we turn to observations of ET response with the scaling term included (Figure \ref{data_scatter}).
-Until now, in the interest of developing leading order intuition for ET response, and to be conservative in our acknowledgment of model and observational error, we've considered $\sigma$ variability to be a measure of uncertainty.
-An alternative viewpoint is that $\sigma \cdot uWUE$ represents spatial and temporal variability of uWUE, which may be expected within bounds \citep[see Table \ref{pft}, also ][]{Zhou_2015}.
-This is a less conservative view; some of the $\sigma \cdot uWUE$ variability will be due to model and observational error, so by viewing $\sigma \cdot uWUE$ as \textit{real} variability we run the risk of mistaking noise for signal.
-However, the advantage of this viewpoint is that, from \citet{Zhou_2014}, we have very high confidence that uWUE fit at the hourly timescale (as we do with $\sigma \cdot uWUE$) correctly captures the relationship between ET, GPP, and VPD, and that our form of PM introduces minimal error with its use of uWUE.
+Most of the results presented so far focus on the sign term, so now we
+turn to observations of ET response with the scaling term included
+(Figure \ref{data_scatter}). Until now, in the interest of developing
+leading order intuition for ET response, and to be conservative in our
+acknowledgment of model and observational error, we've considered
+$\sigma$ variability to be a measure of uncertainty. An alternative
+viewpoint is that $\sigma \cdot uWUE$ represents spatial and temporal
+variability of uWUE, which may be expected within bounds \citep[see
+Table \ref{pft}, also ][]{Zhou_2015}. This is a less conservative
+view; some of the $\sigma \cdot uWUE$ variability will be due to model
+and observational error, so by viewing $\sigma \cdot uWUE$ as
+\textit{real} variability we run the risk of mistaking noise for
+signal. However, the advantage of this viewpoint is that, from
+\citet{Zhou_2014}, we have very high confidence that uWUE fit at the
+hourly timescale (as we do with $\sigma \cdot uWUE$) correctly
+captures the relationship between ET, GPP, and VPD, and that our form
+of PM introduces minimal error with its use of uWUE.
 
@@
   \centerline{\includegraphics[width=1.4\textwidth]{./data_scatter.png}}
-\caption{Scatter plots of observed $\frac{\partial \; ET}{\partial \; VPD}$, including $\sigma$ variability, as a function of PFT, temperature and VPD. Please note differences in the colorbar scale.}
+  \caption{Scatter plots of observed
+    $\frac{\partial \; ET}{\partial \; VPD}$, including $\sigma$
+    variability, as a function of PFT, temperature and VPD. Please
+    note differences in the colorbar scale.}
   \label{data_scatter}
@@
 
-Therefore, with the caveat that some of the signal presented in Figure \ref{data_scatter} may in fact be noise, we can interpret the observed distribution of ET response to VPD.
-In general, the observed response matches the intuitive theory.
-ET response to VPD shifts towards positive values as VPD increases (atmospheric demand dominating).
-CRO exhibit the highest occurrence of positive ET response, and the observations confirm that CRO are the most water intensive.
-CSH are the most water conservative, with a strong negative response.
-DBF, EBF, ENF, MF, SAV and WSA are also water conservative, but show some occurrence of positive ET response to VPD, particularly at higher observed VPD, as the theory predicts.
-GRA, while generally water conservative, does not match the theory, with increasing frequency of \textit{negative} response at high VPD.
-This is as expected, given discussion in Section \ref{testing}.
+Therefore, with the caveat that some of the signal presented in Figure
+\ref{data_scatter} may in fact be noise, we can interpret the observed
+distribution of ET response to VPD. In general, the observed response
+matches the intuitive theory. ET response to VPD shifts towards
+positive values as VPD increases (atmospheric demand dominating). CRO
+exhibit the highest occurrence of positive ET response, and the
+observations confirm that CRO are the most water intensive. CSH are
+the most water conservative, with a strong negative response. DBF,
+EBF, ENF, MF, SAV and WSA are also water conservative, but show some
+occurrence of positive ET response to VPD, particularly at higher
+observed VPD, as the theory predicts. GRA, while generally water
+conservative, does not match the theory, with increasing frequency of
+\textit{negative} response at high VPD. This is as expected, given
+discussion in Section \ref{testing}.
 
-Figure \ref{data_scatter} also includes the impact of the scaling term.
-For a given VPD, the magnitude of ET response does not vary strongly with temperature, confirming that any impact of the scaling term on the magnitude of the response is primarily due to changes in aerodynamic conductance.
-Intuitively this is reasonable; aerodynamic conductance will control how dominant balances at the land surface are communicated to the boundary layer.
-While $\Delta$ controls the efficiency of energy conversion to latent heat, it appears this is a second order term, relative to $g_a$, for scaling ET response.
+Figure \ref{data_scatter} also includes the impact of the scaling
+term. For a given VPD, the magnitude of ET response does not vary
+strongly with temperature, confirming that any impact of the scaling
+term on the magnitude of the response is primarily due to changes in
+aerodynamic conductance. Intuitively this is reasonable; aerodynamic
+conductance will control how dominant balances at the land surface are
+communicated to the boundary layer. While $\Delta$ controls the
+efficiency of energy conversion to latent heat, it appears this is a
+second order term, relative to $g_a$, for scaling ET response.
 
-As in Figure \ref{test_sign}, Figure \ref{data_scatter} matches our expectation based on simplified theory.
-The sign term is most strongly scaled by $g_a$, and in general the sign term shifts towards positive values as VPD increases.
-The willingness with which a given plant type evolved to use water dictates the occurrence of positive versus negative ET response.
-Water conservative plants are highly effective at mitigating the effects of atmospheric demand, and can store water for later use by reducing ET in response to increasing VPD.
+As in Figure \ref{test_sign}, Figure \ref{data_scatter} matches our
+expectation based on simplified theory. The sign term is most strongly
+scaled by $g_a$, and in general the sign term shifts towards positive
+values as VPD increases. The willingness with which a given plant type
+evolved to use water dictates the occurrence of positive versus
+negative ET response. Water conservative plants are highly effective
+at mitigating the effects of atmospheric demand, and can store water
+for later use by reducing ET in response to increasing VPD.
 
 \subsection{Limitations of theory - very dry soil moisture conditions}
-In formulating our theory with Penman Monteith, we implicitly did not account for very dry soil moisture conditions.
-For the majority of environmental conditions  observed at the eddy covariance sites used here, soil conditions were not extremely dry so that we could assume a constant uWUE and $g_1$.
-We posit that ecosystems will generally optimize to host plants living in conditions which they evolved for.
-However, in extreme conditions and drought scenarios soil water content (SWC) could become the limiting factor for ET response to VPD, which our theory does not account for.
-In addition, low soil moisture conditions themselves increase VPD through land-atmosphere feedback \citep[][]{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006, Salvucci_2013, Gentine_2016, Berg_2016}.
+In formulating our theory with Penman Monteith, we implicitly did not
+account for very dry soil moisture conditions. For the majority of
+environmental conditions observed at the eddy covariance sites used
+here, soil conditions were not extremely dry so that we could assume a
+constant uWUE and $g_1$. We posit that ecosystems will generally
+optimize to host plants living in conditions which they evolved
+for. However, in extreme conditions and drought scenarios soil water
+content (SWC) could become the limiting factor for ET response to VPD,
+which our theory does not account for. In addition, low soil moisture
+conditions themselves increase VPD through land-atmosphere feedback
+\citep[][]{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006,
+  Salvucci_2013, Gentine_2016, Berg_2016}.
 
-We first note that in the data used low SWC conditions are very infrequent.
-Within our framework, any systematic bias due to the  failure to account for SWC's effects in very dry conditions should manifest itself in a functional relationship between $\sigma$, our uncertainty measure, and SWC, which is observed at the FLUXNET sites.
-Examining the relationship between $\sigma$ and SWC will test to what extent our theory breaks down in very dry soil moisture conditions.
+We first note that in the data used low SWC conditions are very
+infrequent. Within our framework, any systematic bias due to the
+failure to account for SWC's effects in very dry conditions should
+manifest itself in a functional relationship between $\sigma$, our
+uncertainty measure, and SWC, which is observed at the FLUXNET
+sites. Examining the relationship between $\sigma$ and SWC will test
+to what extent our theory breaks down in very dry soil moisture
+conditions.
 
-If we again view $\sigma \cdot uWUE$ as, in addition to a measure of uncertainty, a short time-scale observation of uWUE, we would expect $\sigma \cdot uWUE$ to decrease at low SWC.
-If $\sigma \cdot uWUE$ is a very strong function of SWC, then our theory should be conditioned more strongly on well watered soil conditions.
-If $\sigma \cdot uWUE$ is weakly a function of SWC, then our theory is more universal and independent of soil moisture conditions.
+If we again view $\sigma \cdot uWUE$ as, in addition to a measure of
+uncertainty, a short time-scale observation of uWUE, we would expect
+$\sigma \cdot uWUE$ to decrease at low SWC. If $\sigma \cdot uWUE$ is
+a very strong function of SWC, then our theory should be conditioned
+more strongly on well watered soil conditions. If $\sigma \cdot uWUE$
+is weakly a function of SWC, then our theory is more universal and
+independent of soil moisture conditions.
 
-Indeed, for all PFTs, there is some slight dependence of $\sigma \cdot uWUE$ on SWC, especially at low SWC (Figure \ref{swc_boxplot}).
-The portion of observations for which our theory is biased by SWC-limitations varies by PFT, due to the nonlinear threshold at which soil moisture availability limits plant function.
-For CRO, MF and DBF, soil water content only matters for about the lowest 5\% of observations (each box is 5\% of observations in Figure \ref{swc_boxplot}).
-And, CRO, relative to DBF, has a very weak dependence on soil moisture, which is reflective of the high likelihood that CRO sites are optimally irrigated and not water stressed, suggesting that observed depatures from the theory for CRO are due to other factors (i.e.
-photosythesis pathway) than soil moisture.
-For ENF, EBF, SAV, and WSA, systematic SWC-induced biases in $\sigma \cdot uWUE$ emerge in about the lowest 20-30\% of SWC conditions, although large variability in the SWC-$\sigma \cdot uWUE$ relationship hampers interpretation for WSA.
-In crontast to ENF, CRO, and DBF, for GRA the relationship between $\sigma \cdot uWUE$ and SWC is more linear and affects a greater portion of observations; about 60\% of observations.
-Clearly, for GRA soil water frequently impacts plant function and alters ET response, and our theory is limited for the majority of encironmental condiitons.
-It is therefore not suprising that our theory tested poorly againt the data for GRA, relative to to the other PFTs (Section \ref{testing}).
-For all other PFTs occurence of soil mosture impacts was rare enough to not manifest itself in bulk statistics and figures.
-A possible exception is CSH, for which there are only two sites, which precludes comment on the exact PFT-wide relationship.
+Indeed, for all PFTs, there is some slight dependence of
+$\sigma \cdot uWUE$ on SWC, especially at low SWC (Figure
+\ref{swc_boxplot}). The portion of observations for which our theory
+is biased by SWC-limitations varies by PFT, due to the nonlinear
+threshold at which soil moisture availability limits plant
+function. For CRO, MF and DBF, soil water content only matters for
+about the lowest 5\% of observations (each box is 5\% of observations
+in Figure \ref{swc_boxplot}). And, CRO, relative to DBF, has a very
+weak dependence on soil moisture, which is reflective of the high
+likelihood that CRO sites are optimally irrigated and not water
+stressed, suggesting that observed depatures from the theory for CRO
+are due to other factors (i.e. photosythesis pathway) than soil
+moisture. For ENF, EBF, SAV, and WSA, systematic SWC-induced biases in
+$\sigma \cdot uWUE$ emerge in about the lowest 20-30\% of SWC
+conditions, although large variability in the SWC-$\sigma \cdot uWUE$
+relationship hampers interpretation for WSA. In crontast to ENF, CRO,
+and DBF, for GRA the relationship between $\sigma \cdot uWUE$ and SWC
+is more linear and affects a greater portion of observations; about
+60\% of observations. Clearly, for GRA soil water frequently impacts
+plant function and alters ET response, and our theory is limited for
+the majority of encironmental condiitons. It is therefore not
+suprising that our theory tested poorly againt the data for GRA,
+relative to to the other PFTs (Section \ref{testing}). For all other
+PFTs occurence of soil mosture impacts was rare enough to not manifest
+itself in bulk statistics and figures. A possible exception is CSH,
+for which there are only two sites, which precludes comment on the
+exact PFT-wide relationship.
 
@@
   \centerline{\includegraphics[width=1.4\textwidth]{./swc_boxplot.pdf}}
-\caption{$\sigma \cdot uWUE$ as a function of PFT and SWC.
-Each box plot represents 5\% of all observations.
-Note that the highest 10\% of SWC observations were excluded from the WSA and SAV figures, to better resolve high variability in narrow low-SWC bins for those PFTs.}
+  \caption{$\sigma \cdot uWUE$ as a function of PFT and SWC. Each box
+    plot represents 5\% of all observations. Note that the highest
+    10\% of SWC observations were excluded from the WSA and SAV
+    figures, to better resolve high variability in narrow low-SWC bins
+    for those PFTs.}
   \label{swc_boxplot}
@@
 
-The observed dependence of $\sigma \cdot uWUE$ on SWC for GRA would explain the deficiencies of our theory compared to the observations in Figure \ref{idealized_sign}, specifically the trend back towards negative soil moisture at high VPD.
-Aforementioned feedbacks between the land surface and the atmosphere, which are not accounted for due to our focus on the one-way response of the land surface to atmospheric conditions, would cause high VPD to be correlated with low SWC \citep[][]{Gentine_2016, Berg_2016}.
-So, at high VPD observations of low SWC are more likely, and this low SWC causes a lower uWUE (Figure \ref{swc_boxplot}).
-The lower uWUE at low SWC/high VPD then leads to the observed downturn towards decreasing ET with increasing VPD, and the deviation between our theory (based on a constant uWUE assumption) and observations.
-It is also perhaps not coincidental that the protion of $\sigma \cdot uWUE$ affected by low soil moisture observations is similiar to the portion of observations that to not match our theory at high VPD.
+The observed dependence of $\sigma \cdot uWUE$ on SWC for GRA would
+explain the deficiencies of our theory compared to the observations in
+Figure \ref{idealized_sign}, specifically the trend back towards
+negative soil moisture at high VPD. Aforementioned feedbacks between
+the land surface and the atmosphere, which are not accounted for due
+to our focus on the one-way response of the land surface to
+atmospheric conditions, would cause high VPD to be correlated with low
+SWC \citep[][]{Gentine_2016, Berg_2016}. So, at high VPD observations
+of low SWC are more likely, and this low SWC causes a lower uWUE
+(Figure \ref{swc_boxplot}). The lower uWUE at low SWC/high VPD then
+leads to the observed downturn towards decreasing ET with increasing
+VPD, and the deviation between our theory (based on a constant uWUE
+assumption) and observations. It is also perhaps not coincidental that
+the protion of $\sigma \cdot uWUE$ affected by low soil moisture
+observations is similiar to the portion of observations that to not
+match our theory at high VPD.
 
-To summarize, our theory is limited by its inability to account for soil water impacts on land surface response, and feedbacks between SWC and VPD.
-Fortunately, for most PFTs SWC's effect on ET is relatively rare (<30\% of observations) and does not manifest itself in the majority of observations and bulk statistics.
-However, for GRA SWC has a strong effect decreasing water use efficiency for the majority of the observations.
-Soil moisture effects explain the deficiencies of our theory in Section \ref{testing}, particularly for GRA.
-By conceptualizing SWC effects as a change in $uWUE$ (and/or $g_1$), it will be possible for future analysis to explore the importance of soil moisture on plant response to VPD.
+To summarize, our theory is limited by its inability to account for
+soil water impacts on land surface response, and feedbacks between SWC
+and VPD. Fortunately, for most PFTs SWC's effect on ET is relatively
+rare (<30\% of observations) and does not manifest itself in the
+majority of observations and bulk statistics. However, for GRA SWC has
+a strong effect decreasing water use efficiency for the majority of
+the observations. Soil moisture effects explain the deficiencies of
+our theory in Section \ref{testing}, particularly for GRA. By
+conceptualizing SWC effects as a change in $uWUE$ (and/or $g_1$), it
+will be possible for future analysis to explore the importance of soil
+moisture on plant response to VPD.
 
-\subsection{Functional form of ET dependence on VPD and its relation to VPD dependence}
+\subsection{Functional form of ET dependence on VPD and its relation
+  to VPD dependence}
 
-Th theory described in Section \ref{sign_term} indicates that for a given $uWUE$ and $g_1$, the ET dependence on VPD should be concave upward.
-In other words, there should be some local minima in ET at a critical VPD$_{crit}$, assuming the scaling term (e.g.
-aerodynamic conductance and $\Delta$) is held fixed.
-This result warrants futher investigation, because to our knowledge no one has derived the theoretical ecosystem-scale relationship between ET and VPD while controlling for enivironmental conditiions.
-In particular, from personal communication, there is an apparent lack of consensus over whether the shape of the ET-VPD curve should be concave upward (our result) or concave downward.
-Given that understanding the ET-VPD relationship of the one-way plant response is fundamental to hypothesizing about any feedbacks between the land surface and the atmopshere, we analyze why our derived ET-VPD relationship is concave upward, particularly with respect to the exponent of VPD depedence in $uWUE$ and the Medlyn unified stomatal conductace model.
+Th theory described in Section \ref{sign_term} indicates that for a
+given $uWUE$ and $g_1$, the ET dependence on VPD should be concave
+upward. In other words, there should be some local minima in ET at a
+critical VPD$_{crit}$, assuming the scaling term (e.g. aerodynamic
+conductance and $\Delta$) is held fixed. This result warrants futher
+investigation, because to our knowledge no one has derived the
+theoretical ecosystem-scale relationship between ET and VPD while
+controlling for enivironmental conditiions. In particular, from
+personal communication, there is an apparent lack of consensus over
+whether the shape of the ET-VPD curve should be concave upward (our
+result) or concave downward. Given that understanding the ET-VPD
+relationship of the one-way plant response is fundamental to
+hypothesizing about any feedbacks between the land surface and the
+atmopshere, we analyze why our derived ET-VPD relationship is concave
+upward, particularly with respect to the exponent of VPD depedence in
+$uWUE$ and the Medlyn unified stomatal conductace model.
 
-There is a theoretical basis for the square root VPD dependence in both the stomatal conducatance model and $uWUE$ based on the assumption that stomata behave to maximize carbon gain while minimizing water loss, which observations also generally support \citep{Lloyd_1991, MEDLYN_2011, Lin_2015, Zhou_2014, Zhou_2015, Medlyn_2017}.
-However, purely empiriacle results which fit the exponent of the VPD dependence have shown that it may vary from 1/2, suggesting that stomata, as well as ecosystem-scale quantities based on stomata theroy,  may not always function optimally \citep{Zhou_2015, Lin_2018}.
-Specifically with regards to $uWUE$, one would not expect that this ecosystem scale WUE quantity will responde to VPD exactly analgalously to stomata.
-Direct soil evaporation's contributions to ET should shift the exponential dependence of VPD to be greater than 1/2.
-In other words, direct soil evaporation causes WUE to decrease more strongly with increasing VPD, especially in well-watered conditions.
-\citet{Zhou_2015}'s results corroborate this: they found a mean imprically fit exponential VPD dependence of 0.55 istead of 0.5 for Ameriflux sites.
-Given that these recent results on the relatioship between VPD, GPP, and ET form the backbone of our analysis and are what allowed us to derive an explicit expression for ET for the first time (Equation \ref{et}), we will analyze if and how assumptions about the exonent of the VPD dependence impacts the shape of the ET-VPD dependence.
-By introducing $n$ and $m$ we can free our stomatal conductance model from assumptions about VPD dependence:
+There is a theoretical basis for the square root VPD dependence in
+both the stomatal conducatance model and $uWUE$ based on the
+assumption that stomata behave to maximize carbon gain while
+minimizing water loss, which observations also generally support
+\citep{Lloyd_1991, MEDLYN_2011, Lin_2015, Zhou_2014, Zhou_2015,
+  Medlyn_2017}. However, purely empiriacle results which fit the
+exponent of the VPD dependence have shown that it may vary from 1/2,
+suggesting that stomata, as well as ecosystem-scale quantities based
+on stomata theroy, may not always function optimally \citep{Zhou_2015,
+  Lin_2018}. Specifically with regards to $uWUE$, one would not expect
+that this ecosystem scale WUE quantity will responde to VPD exactly
+analgalously to stomata. Direct soil evaporation's contributions to ET
+should shift the exponential dependence of VPD to be greater than
+1/2. In other words, direct soil evaporation causes WUE to decrease
+more strongly with increasing VPD, especially in well-watered
+conditions. \citet{Zhou_2015}'s results corroborate this: they found a
+mean imprically fit exponential VPD dependence of 0.55 istead of 0.5
+for Ameriflux sites. Given that these recent results on the
+relatioship between VPD, GPP, and ET form the backbone of our analysis
+and are what allowed us to derive an explicit expression for ET for
+the first time (Equation \ref{et}), we will analyze if and how
+assumptions about the exonent of the VPD dependence impacts the shape
+of the ET-VPD dependence. By introducing $n$ and $m$ we can free our
+stomatal conductance model from assumptions about VPD dependence:
 \begin{linenomath*}
@@
 where:
-\[*WUE = \frac{GPP}{ET}VPD^n.\]
-To determine how the exponent $n$ and $m$ alter the shape of the ET-VPD dependence we find the roots of the second deriviative of ET, using Equation \ref{m_n} for stomatal conductance, with respect to VPD:
-\input{d2_solutions.tex}
+\[*WUE = \frac{GPP}{ET}VPD^n.\] To determine how the exponent $n$ and
+$m$ alter the shape of the ET-VPD dependence we find the roots of the
+second deriviative of ET, using Equation \ref{m_n} for stomatal
+conductance, with respect to VPD: \input{d2_solutions.tex}
 
@@
 
-We derived a new form of Penman Monteith using the concept of semi-empirical optimal stomatal regulation \citep{Lin_2015, MEDLYN_2011} and near constant uWUE \citep{Zhou_2015} to remove the implicit dependence of stomatal conductance on GPP and ET.
-With our new form of Penman Monteith we developed a theory for when an ecosystem will tend to reduce or increase ET with increasing VPD, which we evaluated against a range of eddy-covariance data spanning different climates and plant functional types.
-The goal was to capture the leading order behavior of the system to gain some intrinsic knowledge for its behavior.
-This intuition can be used to disentangle land atmosphere feedbacks in more complicated scenarios, and will aid interpretation of observations and sophisticated models.
+We derived a new form of Penman Monteith using the concept of
+semi-empirical optimal stomatal regulation \citep{Lin_2015,
+  MEDLYN_2011} and near constant uWUE \citep{Zhou_2015} to remove the
+implicit dependence of stomatal conductance on GPP and ET. With our
+new form of Penman Monteith we developed a theory for when an
+ecosystem will tend to reduce or increase ET with increasing VPD,
+which we evaluated against a range of eddy-covariance data spanning
+different climates and plant functional types. The goal was to capture
+the leading order behavior of the system to gain some intrinsic
+knowledge for its behavior. This intuition can be used to disentangle
+land atmosphere feedbacks in more complicated scenarios, and will aid
+interpretation of observations and sophisticated models.
 
-Our theory suggests that for a majority of environmental conditions, plants will tend to conserve water and reduce ET with increasing VPD.
-Stomatal regulation and plant physiological response strongly regulate ET, and this regulation varies by PFT.
-CROs are the least water conservative, while DBF, EBF, SAV, GRA, MF, WSA, ENF and CSH are progressively more water conservative (more likely to reduce ET in response to increasing VPD).
-SAV and WSA exhibit positive ET response to VPD not necessarily because of poor water conservation stragegies relative to other PFTs, but because of greater occurence of high atmospheric demand (VPD) relative to other PFTs.
-Observations of ET response to VPD exhibit the same general behavior as the theory, with ET response becoming more positive (atmospheric demand dominating) as environmental VPD increases within a PFT, and more negative for PFTs that are adapted to arid conditions and prioritize water conservation over primary production.
+Our theory suggests that for a majority of environmental conditions,
+plants will tend to conserve water and reduce ET with increasing
+VPD. Stomatal regulation and plant physiological response strongly
+regulate ET, and this regulation varies by PFT. CROs are the least
+water conservative, while DBF, EBF, SAV, GRA, MF, WSA, ENF and CSH are
+progressively more water conservative (more likely to reduce ET in
+response to increasing VPD). SAV and WSA exhibit positive ET response
+to VPD not necessarily because of poor water conservation stragegies
+relative to other PFTs, but because of greater occurence of high
+atmospheric demand (VPD) relative to other PFTs. Observations of ET
+response to VPD exhibit the same general behavior as the theory, with
+ET response becoming more positive (atmospheric demand dominating) as
+environmental VPD increases within a PFT, and more negative for PFTs
+that are adapted to arid conditions and prioritize water conservation
+over primary production.
 
-The consequences of our theory our twofold.
-First, the ET response to VPD is always far more negative than it would be if only atmospheric demand is considered and plant physiological response is ignored (as with PET).
-Therefore,  any drought indices or analysis using PET will not capture how  vegetated surfaces respond and react to increases in atmospheric demand.
-However, the new uWUE-version of Penman Monteith we derived (Equation \ref{et}) could be used as a complement to PET in drought indices.
-One could calculate and compare indices  when plant response is ignored (PET) or included (Equation \ref{et}).
+The consequences of our theory our twofold. First, the ET response to
+VPD is always far more negative than it would be if only atmospheric
+demand is considered and plant physiological response is ignored (as
+with PET). Therefore, any drought indices or analysis using PET will
+not capture how vegetated surfaces respond and react to increases in
+atmospheric demand. However, the new uWUE-version of Penman Monteith
+we derived (Equation \ref{et}) could be used as a complement to PET in
+drought indices. One could calculate and compare indices when plant
+response is ignored (PET) or included (Equation \ref{et}).
 
-Lastly, our paper builds important intuition for how plants will respond to VPD pertubations, including expected increases in VPD induced by global change.
-Plant physiological responses to direct CO$_2$ effects \citep[e.g.,][]{Swann_2016} receives more attention than physiological response to indirect effects like increased VPD.
-Here, we provide broad PFT-focused results showing a likely decrease in ET in response to positive VPD perturbations (atmospheric drying), which is consistent with recent observational analysis \citep[e.g.,][]{Rigden_2017}.
-Feedbacks between the land and the atmosphere may alter the net response to a global VPD perturabation, but our focus on the one way plant response to a VPD perturbation in the atmopsheric boundary layer is an important first step to disentangling such feedbacks, both in observations and model simulations of the present and future.
-By removing Penman Monteith's dependence on implicit relationships between GPP, VPD, and ET, we allow for explicit, accuruate future analysis of plant-VPD feedbacks in the atmospheric boundary layer (Equation \ref{et}).
-Our approach can also be extended to examine varying plant response to more nuanced consideration of plant type and climate.
-Any plant physiological heterogeneity that can be conceptualized with shifts in $g_1$ \citep[e.g.][]{Lin_2015, Medlyn_2017} and/or uWUE \citep[e.g.][]{Zhou_2014} is representable within our framework, which opens the door for a hierarchy of more sophisticated climate- and plant-specific analysis of ET sensitivity to environmental variables (including VPD).
-We argue that such simplified conceptual frameworks are critical tools for disentangling land-atmosphere feedbacks at various scales, from diurnal to seasonal and beyond, and to characterize ET response in a warmer, atmospherically drier, and enriched CO$_2$ world.
+Lastly, our paper builds important intuition for how plants will
+respond to VPD pertubations, including expected increases in VPD
+induced by global change.  Plant physiological responses to direct
+CO$_2$ effects \citep[e.g.,][]{Swann_2016} receives more attention
+than physiological response to indirect effects like increased
+VPD. Here, we provide broad PFT-focused results showing a likely
+decrease in ET in response to positive VPD perturbations (atmospheric
+drying), which is consistent with recent observational analysis
+\citep[e.g.,][]{Rigden_2017}. Feedbacks between the land and the
+atmosphere may alter the net response to a global VPD perturabation,
+but our focus on the one way plant response to a VPD perturbation in
+the atmopsheric boundary layer is an important first step to
+disentangling such feedbacks, both in observations and model
+simulations of the present and future. By removing Penman Monteith's
+dependence on implicit relationships between GPP, VPD, and ET, we
+allow for explicit, accuruate future analysis of plant-VPD feedbacks
+in the atmospheric boundary layer (Equation \ref{et}). Our approach
+can also be extended to examine varying plant response to more nuanced
+consideration of plant type and climate. Any plant physiological
+heterogeneity that can be conceptualized with shifts in $g_1$
+\citep[e.g.][]{Lin_2015, Medlyn_2017} and/or uWUE
+\citep[e.g.][]{Zhou_2014} is representable within our framework, which
+opens the door for a hierarchy of more sophisticated climate- and
+plant-specific analysis of ET sensitivity to environmental variables
+(including VPD). We argue that such simplified conceptual frameworks
+are critical tools for disentangling land-atmosphere feedbacks at
+various scales, from diurnal to seasonal and beyond, and to
+characterize ET response in a warmer, atmospherically drier, and
+enriched CO$_2$ world.
 
-\acknowledgments
-This work used eddy covariance data acquired and shared by the FLUXNET community, including these networks: AmeriFlux, AfriFlux, AsiaFlux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont, ChinaFlux, Fluxnet-Canada, GreenGrass, ICOS, KoFlux, LBA, NECC, OzFlux-TERN, TCOS-Siberia, and USCCC.
-The ERA-Interim reanalysis data are provided by ECMWF and processed by LSCE.
-The FLUXNET eddy covariance data processing and harmonization was carried out by the European Fluxes Database Cluster, AmeriFlux Management Project, and Fluxdata project of FLUXNET, with the support of CDIAC and ICOS Ecosystem Thematic Center, and the OzFlux, ChinaFlux and AsiaFlux offices.
-This material is based upon work supported by the National Science Foundation Graduate Research Fellowship under Grant No.
-DGE 16-44869.
-All code and data used in this analysis, including those used to generate the figures and tables, is publicly available at https://github.com/massma/climate\_et.
+\acknowledgments This work used eddy covariance data acquired and
+shared by the FLUXNET community, including these networks: AmeriFlux,
+AfriFlux, AsiaFlux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont,
+ChinaFlux, Fluxnet-Canada, GreenGrass, ICOS, KoFlux, LBA, NECC,
+OzFlux-TERN, TCOS-Siberia, and USCCC. The ERA-Interim reanalysis data
+are provided by ECMWF and processed by LSCE. The FLUXNET eddy
+covariance data processing and harmonization was carried out by the
+European Fluxes Database Cluster, AmeriFlux Management Project, and
+Fluxdata project of FLUXNET, with the support of CDIAC and ICOS
+Ecosystem Thematic Center, and the OzFlux, ChinaFlux and AsiaFlux
+offices. This material is based upon work supported by the National
+Science Foundation Graduate Research Fellowship under Grant No. DGE
+16-44869. All code and data used in this analysis, including those
+used to generate the figures and tables, is publicly available at
+https://github.com/massma/climate\_et.
 

### 2018-04-03 doc/paper/vpd_et_paper.tex
finished funtional form section, and addressed reviewer and pierre's response.

@@
 The ET response to changes in VPD alters water partitioning between
-the soil and atmosphere. If plant response reduces ET with atmospheric
+the soil and atmosphere. If ecosystem plant response reduces ET with atmospheric
 drying then soil moisture will be better conserved. This would seem a
@@
   Gentine_2016, Berg_2016}. This however would not seem to be a
-sensible strategy for plants from an evolutionary standpoint.
+sensible strategy for plants from an evolutionary standpoint. One may
+argue that increases in ET with increasing VPD could increase the
+liklihood of precipitation \citep[e.g.,][]{Findell_2011}. However,
+increases in ET are not a gaurantee
+of increased likelhood of precipitation, which also dependes on the
+environemtnal conditions \citep[][]{Gentine_2013}. Futhermore,  any
+increases in precipitation are likely to be non-local, such that plants giving up
+water to the atmosphere are not gauranteed to reap the benefits of
+water returned to the atmosphere. Using this logic, from an ecoysystem
+evolutionary perspective, water stored in soil is worth much more than
+the \textit{chance} of water returned as precipitation.
 
@@
   with rain interception and sensor saturation at high relative
-  humidity (\cite{MEDLYN_2011}).
+  humidity (\cite{Medlyn_2017}).
 \end{itemize}
@@
     $\gamma$  & psychometric constant & Pa K$^{-1}$   & - \\
-    $g_s$  & stomatal conductance & m s$^{-1}$  & \citet{MEDLYN_2011} \\
-    $g_{l-s}$  & leaf-scale stomatal conductance & mol m$^{-2}$ s$^{-1}$  & \citet{MEDLYN_2011} \\
+    $g_{s-leaf}$  & leaf-scale stomatal conductance & m s$^{-1}$  & \citet{MEDLYN_2011} \\
+    $g_{s}$  &  stomatal conductance & m s$^{-1}$
+                                   & \citet{Medlyn_2017} \\
+    $g_{1-leaf}$  & leaf-scale slope parameter & Pa$^{0.5}$
+                                   & \citet{MEDLYN_2011} \\
+    $g_{1}$  & ecoysystem-scale slope parameter & Pa$^{0.5}$ & \citet{MEDLYN_2017} \\
     $R$ & universal gas constant & J mol$^{-1}$ K$^{-1}$ & - \\
     $R_{air}$ & gas constant of air & J  K$^{-1}$ kg$^{-1}$ & - \\
-    $LAI$ & leaf area index & -& - \\
     $\sigma$ & uncertainty parameter & -& - \\
@@
 \citet{MEDLYN_2011} developed a model for stomatal conductance ($g_s$)
-by combining an optimal photosynthesis theory \citep{Farquhar_1980,
-  Katul_2010} with an empirical approach, which describes the
+by combining an optimal photosynthesis theory \citep{Cowan_1977} with an empirical approach, which describes the
 dependence of $g_s$ to VPD. This resulted in the following model for
@@
   \begin{equation}
-    g_{s-leaf} = g_0 + 1.6 \left(1 + \frac{g_{1-leaf}}{\sqrt{VPD}}\right) \frac{A}{c_a},
+    g_{s-leaf} = g_0 + 1.6 \left(1 +
+      \frac{g_{1-leaf}}{\sqrt{VPD}}\right) \frac{A}{c_a},
+    \label{leaf_medlyn}
   \end{equation}
@@
 observations themselves. The variability of $\sigma$ then propagates
-through any uncertainty to our derivative of Equation \ref{et_sigma}:
+through any uncertainty to our partial derivative of Equation
+\ref{et_sigma} with respect to VPD:
 
@@
 With Equation \ref{d_et} we have an analytical framework for ecosystem
-response to atmospheric demand perturbations, with an implicit
-assumption of non-extreme drought conditions associated with our
-constant $uWUE$ assumption (e.g. we might expect $uWUE$ to decrease in
-extreme drought). There are a few subtleties to taking the derivative
-in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$) and $VPD$ are
-functionally related, so while taking the derivative we evaluate
+response to atmospheric demand perturbations with environmental
+conditions held fixed. There are a few subtleties to taking the
+derivative in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$)
+and $VPD$ are functionally related, so while taking the derivative we
+evaluate
 $\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET}
@@
 independent, which is supported by the data (not shown).
+
+This derivation relied either implicitly or explicitly on several
+assumptions. First, the underlying theory used to derive Equations
+\ref{medlyn}, \ref{leaf_medlyn} rely on a implicit assumptions that
+$ET=g_s \cdot VPD$; physically this implies that leaves are perfectly
+coupled to the atmosphere. In reality, for some conditions and plant
+types the leaves can become decoupled from the boundary layer
+\citep{De_2017, Medlyn_2017}. Given our focus on the growing season
+which is usually characterized by relatively high insolation, we would
+expect the surface to be well coupled. Additionally, by following
+\citet{Medlyn_2017}'s method for calcualting eco-system scale $g_1$,
+we have stronger confidence that surface coupling is represented
+within our framework, even if it is not represented in the underlying
+theory used to develop $g_s$. Additionally, the formulation of $uWUE$
+in \citet{Zhou_2014} and \citet{Zhou_2015} as well as our use of
+\citet{Medlyn_2017}'s stomatal conductance model assumes that that
+direct soil evaporation contributions to ET remian relatively
+small. This should be more true druing the growing season than other
+times; however some plant types allow for systematically larger
+contributions of ``E'' in ET, particularly those with less dense
+canopies and smaller aboslute amounts of transpiration. We therefore
+might expect that the theory will be most applicable to forest PFTs,
+which will be most strongly coupled to the boundary layer due to
+larger surface roughness, and will also generally have the highest
+ratios of transpiration to evaporation. Finally, for the goal of
+developing PFT-wide intuition, we assume that $g_1$ and $uWUE$ are
+constant in space and time. Both quantities have been shown to be well
+conserved over space and time within a PFT relative to inter-PFT
+variability. However, we would expect them to vary somehwat with
+environmental conditions including soil water content temperature, and
+with intra-PFT plant-specific characteristics like wood density for
+tree PFTs \citep{Lin_2015}. Given these strong assumptions made with
+the goal of understanding broad, leading order plant behaviors, we
+take the extra care of including and analyzing uncertainty (Section
+\ref{testing}) and examing sensitivity to the exponent of the VPD
+response derived using assumptions of $ET=g_s \cdot VPD$ (Section
+\ref{functional_form}).
 
@@
 ($\approx$ 35\%, Figure \ref{test_sign}). However, in general, and
-specifically for non-extreme VPD, the theory matches the observations
+specifically for non-extreme VPD (VPD < $\approx$70-90th percentiles), the theory matches the observations
 for the tree and savannah plant types.
@@
 when applied to some PFTs, especially for grasslands, it does well for
-DBF, MF and CSH PFTs, and captures the response at non-extreme VPD for
+DBF, MF and CSH PFTs, and captures the response at non-extreme VPD
+(VPD < $\approx$ 70th-90th percentiles) for
 ENF, EBF, CRO, WSA, and SAV. In general, the leading order behavior
@@
 \subsection{Limitations of theory - very dry soil moisture conditions}
+\label{swc_section}
 In formulating our theory with Penman Monteith, we implicitly did not
@@
 rare (<30\% of observations) and does not manifest itself in the
-majority of observations and bulk statistics. However, for GRA SWC has
+majority of observations and bulk statistics. This result is
+consistent with However, for GRA SWC has
 a strong effect decreasing water use efficiency for the majority of
@@
   to VPD dependence}
-
-Th theory described in Section \ref{sign_term} indicates that for a
+\label{functional_form}
+
+The theory described in Section \ref{sign_term} indicates that for a
 given $uWUE$ and $g_1$, the ET dependence on VPD should be concave
 upward. In other words, there should be some local minima in ET at a
-critical VPD$_{crit}$, assuming the scaling term (e.g. aerodynamic
-conductance and $\Delta$) is held fixed. This result warrants futher
+critical VPD$_{crit}$, assuming the scaling and plant terms  (e.g. aerodynamic
+conductance, $\Delta$, $g_1$ and $uWUE$) are held fixed. This result warrants futher
 investigation, because to our knowledge no one has derived the
@@
 \citep{Lloyd_1991, MEDLYN_2011, Lin_2015, Zhou_2014, Zhou_2015,
-  Medlyn_2017}. However, purely empiriacle results which fit the
-exponent of the VPD dependence have shown that it may vary from 1/2,
-suggesting that stomata, as well as ecosystem-scale quantities based
-on stomata theroy, may not always function optimally \citep{Zhou_2015,
-  Lin_2018}. Specifically with regards to $uWUE$, one would not expect
-that this ecosystem scale WUE quantity will responde to VPD exactly
-analgalously to stomata. Direct soil evaporation's contributions to ET
-should shift the exponential dependence of VPD to be greater than
-1/2. In other words, direct soil evaporation causes WUE to decrease
-more strongly with increasing VPD, especially in well-watered
-conditions. \citet{Zhou_2015}'s results corroborate this: they found a
-mean imprically fit exponential VPD dependence of 0.55 istead of 0.5
-for Ameriflux sites. Given that these recent results on the
-relatioship between VPD, GPP, and ET form the backbone of our analysis
-and are what allowed us to derive an explicit expression for ET for
-the first time (Equation \ref{et}), we will analyze if and how
-assumptions about the exonent of the VPD dependence impacts the shape
-of the ET-VPD dependence. By introducing $n$ and $m$ we can free our
-stomatal conductance model from assumptions about VPD dependence:
+  Medlyn_2017}. However, some purely empiriacle results which fit the
+exponent of the VPD dependence to data have shown that it may vary slightly
+from 1/2, suggesting that stomata, as well as ecosystem-scale
+quantities based on stomata theroy, may not always function optimally
+\citep{Zhou_2015, Lin_2018}. Specifically with regards to $uWUE$, one
+would not expect that this ecosystem scale WUE quantity will responde
+to VPD exactly analgalously to stomata. Direct soil evaporation's
+contributions to ET should shift the exponent of the VPD dependence,
+especially for conditions of low GPP when we would expect a
+systematically larger portion of direct soil evaporation contributions
+to ET, because we would also expect lower amounts of
+transpiration. \citet{Zhou_2015}'s results corroborate this: they
+found a mean imprically fit exponential VPD dependence of 0.55,
+varying slightly from the theoretically optimal value of of 0.5 for
+Ameriflux sites. Interpreting \citet{Lin_2018}'s results, which also
+show variance in the empiracle exponent of the VPD dependence of the
+stomatal conductance model, is more difficult as \citet{Lin_2018} does
+not handle GPP/A dependence of stomatal conductance in a directly
+analgoalous manner to the optimal theory in \citet{MEDLYN_2011} and
+\citet{Medlyn_2017}. Regardless, given that these recent results
+\citep{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017} on the
+relatioship between VPD, GPP, and ET \citep{MEDLYN_2011, Zhou_2014,
+  Zhou_2015, Medlyn_2017} form the backbone of our analysis and are
+what allowed us to derive an explicit ET expression for the first
+time (Equation \ref{et}), we will analyze if and how assumptions about
+the exponent of the VPD dependence impacts the shape of the ET-VPD
+dependence. By introducing $n$ and $m$ we can free our stomatal
+conductance model from assumptions about VPD dependence:
 \begin{linenomath*}
   \begin{equation}
-    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{VPD^m}\right) \frac{*WUE \; ET}{c_a \; VPD^n},
+    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_*}{VPD^m}\right) \frac{*WUE \; ET}{c_a \; VPD^n},
     \label{m_n}
@@
 where:
-\[*WUE = \frac{GPP}{ET}VPD^n.\] To determine how the exponent $n$ and
-$m$ alter the shape of the ET-VPD dependence we find the roots of the
-second deriviative of ET, using Equation \ref{m_n} for stomatal
-conductance, with respect to VPD: \input{d2_solutions.tex}
+\[*WUE = \frac{GPP}{ET}VPD^n,\] and $g_*$ is a generic slope parameter
+of units $VPD^m$. To determine how the exponent $n$ and $m$ alter the
+shape of the ET-VPD dependence we find the roots of the second
+deriviative of ET, using Equation \ref{m_n} for stomatal conductance
+($g_s$, with respect to VPD: \input{d2_solutions.tex}. With this
+result we've detemined a family of curves seperating concave up from
+concave down ET solutions (Figure \ref{concave}). These curves are
+only functions of the exponent of the VPD depedence and a quantity we
+call non-dimensional VPD ($VPD^m/g*$). Several important
+relations reveal themselves from Equation \ref{curves}:
+
+\begin{figure}
+  \centering
+  \centerline{\includegraphics[width=1\textwidth]{./concave.pdf}}
+  \caption{ Solutions corresponding to inflection points between
+    concave up and concave down ET-VPD curves (Equation \ref{curves})
+    for 3 specific scenarios. Solutions are defined in terms of a
+    non-dimensional VPD ($VPD^m/g_*$), but to aide physical
+    interpretation the horizontal axis is additionally provided interms of dimensionalized VPD assuming $m=1/2$ and
+    $g_*=110\; Pa^{1/2}$ (average of all PFT $g_1$). The vertical axis
+    has a different interpretation depending on the solution
+    curve. For the blue line ($m$ varying), it correspondes to $m$,
+    for the orange line ($n$ varying) it correspondes to $n$, and for
+    the green line it correspondes to the value of both $n$ and $m$
+    ($n=m$). Regions of the parameter space that corresponde to
+    concave up and concave down results are labeled.}
+  \label{concave}
+\end{figure}
+
+\begin{enumerate}
+  \item For optimal behavior (n, m = 1/2) the ET-VPD curve will be
+    concave up regardless of the magnitude of the plant constants
+    $g_1$ and $uWUE$. Therefore, the general concave up nature of our
+    results, given an assumption of optimal behavior, is insensitive
+    to plant type.
+  \item For all physically possible exponents of VPD dependence ($n,
+    m$), whether the solution is concave up or concave down does not
+    depend on $uWUE$.
+  \item In general, increasing the exponent of VPD dependence
+    increases the liklihood of a concave up result. Additionally, as
+    the exponent of VPD dependence increases from the optimum value of
+    1/2, whether the curve is concave upward or concave downward
+    becomes a function of the plant specific slope parameter $g_*$,
+    through non-dimensional VPD ($VPD^m/g_*$). Because the exponent of
+    the VPD dependencies is capable of altering the fundamental shape
+    of ET-VPD dependence, future research investment in understanding
+    the exact VPD depedence of stomatal conductance, and further
+    reconciliation of empiracle and theoretical stomatal and
+    ecosostyem behavior should be prioritized.
+\end{enumerate}
+However, while it is possible that in the future some other form of
+VPD dependence is derived, at present \cite{MEDLYN_2011} and
+\cite{Zhou_2014} firmly established n=m=1/2 as the most likely
+candidiate given current theory and empiracle data. Additionally, we
+argue that a concave up result matches physical intuition more than a
+concave down result. Plants must maintain nutrient and sugar transport
+through the phloem and xylem. To accomplish this, stomata must remain
+slightly open \citep{De_2013, Nikinmaa_2013, Ryan_2014}. Futhermore,
+even if complete stomatal closure were possible, cuticular water loss
+and [at the ecosystem-scale] direct soil evaporation are still sources
+of ET which increase with VPD, independent of stomatal
+closure. Therefore, in the limit as VPD becomes large and we assume
+plants are exercising all strategies to reduce ET, any futher increase
+in VPD should result in an increase in ET through cuticular water loss
+and/or direct soil evaporation. This inevitable transition from
+conditions where stomata responde strongly to VPD to conditions where
+stomata response is asymptoting towards full closure would cause a
+concave up ET-VPD curve, which is matched by the theory. In short,
+plant response becomes more limited as VPD increases, while
+atmospheric demand monotonically increases with VPD, leading to the
+intuitive result that atmospheric demand dominates plant response when
+atmospheric deman is high.
+
+This analysis allow us to understand the theoretical shape of the ET
+response to VPD with environmental conditions held fixed which would
+be difficult to disentagle purely from flux observations, given the
+relatively fast time scale of plant response and the non-stationarity
+of [solar forced] environmental conditions  over the relatively
+coarse (half hourly) flux estimates (which is required to obtain
+robust eddy-covariance statistics). This fundamental understanding of
+ET response to VPD with environmental conditoins held fixed can be
+built upon with future work examing how changes in VPD and enviornment
+conditions (e.g. soil water storage) feedback upon one another. In the
+soil water storage example, over very long time scales extremely high
+VPD perturbations coupled with no precipitation could result in
+decreases in soil water storage such that water becomes limiting. This
+could be represented by an extension of our framework in which uWUE is
+allowed to decrease with decreasing SWC, as observed in Secion
+\ref{swc_section}. Here, we focus our results by assuming constant
+PFT-wide conditions to build baseline intuition for ET-VPD
+depedence. For most PFTs, the theory with plant function held fixed
+matches the leading order behaviro of the observations where plant
+function varies (Section \ref{testing}).
 
@@
 Lastly, our paper builds important intuition for how plants will
-respond to VPD pertubations, including expected increases in VPD
+respond to VPD pertubations. We show that given optimal stomatal
+function and fixed environemntal conditions, the ET-VPD depednence is
+theoretically concave up, with ET increasing with increasing VPD as
+VPD increases past some critical value (Table \ref{vpd_crit}). However
+future research should focus on fully understanding the form of
+stomata VPD dependence, as this result is sensitive to the exponent of
+VPD dependence, which we currently believe is 1/2 \citep{MEDLYN_2011,
+  Zhou_2014}. Our results are applicable to understanding the impact
+of  expected increases in VPD
 induced by global change.  Plant physiological responses to direct
@@
 \citep[e.g.,][]{Rigden_2017}. Feedbacks between the land and the
-atmosphere may alter the net response to a global VPD perturabation,
+atmosphere may alter the net response to a long-timescale global VPD perturabation,
 but our focus on the one way plant response to a VPD perturbation in
@@
 consideration of plant type and climate. Any plant physiological
-heterogeneity that can be conceptualized with shifts in $g_1$
+heterogeneity or feedback that can be conceptualized with shifts in $g_1$
 \citep[e.g.][]{Lin_2015, Medlyn_2017} and/or uWUE
-\citep[e.g.][]{Zhou_2014} is representable within our framework, which
+\citep[e.g.][]{Zhou_2014} are representable within our framework, which
 opens the door for a hierarchy of more sophisticated climate- and

### 2018-04-03 src/paper_figure_code/concave.py

@@ -45,6 +45,7 @@ fh.write("\\begin{equation}\n")
 fh.write("\\frac{\\partial^2 \; ET}{\\partial \\; VPD^2} = 0 "\
          "\\quad \\forall \\quad")
 fh.write("%s%s%s" % ("\\frac{VPD^m}{g1} = ", latex(soln[0]), "\n"))
+fh.write("\\label{curves}\n")
 fh.write("\\end{equation}\n")
 fh.close()
 
@@ -75,9 +76,9 @@ ax2 = ax.twiny()
 for sub, solve_var, label in zip(substitutes, solve_vars, labels):
   subbed = gen_func(sub, solve_var, soln)
   ax = plot_curve(ax, subbed, label)
-ax.set_ylabel(r"VPD Exponent (m: $\frac{g_1}{VPD^m}$, "\
+ax.set_ylabel(r"VPD Exponent (m: $\frac{g_*}{VPD^m}$, "\
               r"n: $\frac{GPP}{ET}VPD^n$)", fontsize=single_ax_fontsize)
-ax.set_xlabel(r"VPD (Pa), assuming g1=110 Pa$^{0.5}$ and m=1/2",\
+ax.set_xlabel(r"VPD (Pa), assuming g$_*$=110 Pa$^{0.5}$ and m=1/2",\
               fontsize=single_ax_fontsize)
 ax.set_xlim([0., 4000.0])
 ax.set_ylim([0.5, 1.0])
@@ -85,7 +86,7 @@ ticks = ax.get_xticks()
 ax2.set_xticks(ticks)
 ax2.set_xticklabels(normalized(ticks))
 # ax2.set_xlabel(r"$\frac{VPD^m}{g_1}$         ", fontsize=single_ax_fontsize)
-ax2.set_xlabel(r"VPD$^m$/g$_1$", fontsize=single_ax_fontsize)
+ax2.set_xlabel(r"VPD$^m$/g$_*$", fontsize=single_ax_fontsize)
 ax.text(1050., 0.95, 'Concave Down', horizontalalignment='center',\
         verticalalignment='top', fontdict={'fontsize' : 22})
 ax.text(2750., 0.53, 'Concave Up', horizontalalignment='center',\
@@ -96,4 +97,18 @@ ax.legend(loc="lower right", bbox_to_anchor=(1.0, 0.6),
 plt.savefig("../../doc/paper/concave.pdf")
 
 
+
+plt.figure()
[…]

### 2018-04-03 doc/paper/vpd_et_paper.tex
readthrough done, now nee dto markup and send to Pierre:

@@
 \begin{abstract}
-  Increases to vapor pressure deficit (VPD) and atmospheric demand for
-  water are expected with rising atmospheric CO$_2$. While increased
-  evapotranspiration (ET) in response to increased atmospheric demand
-  seems intuitive, plants are capable of reducing ET in response to
-  increased VPD by closing their stomata, in an effort to conserve
-  water. Here we examine which effect dominates response to increasing
-  VPD: atmospheric demand and increases in ET, or plant physiological
-  response (stomata closure) and decreases in ET. We use
-  Penman-Monteith, combined with semi-empirical optimal stomatal
-  regulation theory and underlying water use efficiency, to develop a
-  theoretical framework for understanding how ET responds to increases
-  in VPD.
+  Increasing vapor pressure deficit increases atmospheric demand for
+  water, and vapor pressure deficit is expected to rise with
+  increasing CO$_2$. While increased evapotranspiration (ET) in
+  response to increased atmospheric demand seems intuitive, plants are
+  capable of reducing ET in response to increased VPD by closing their
+  stomata, in an effort to conserve water. Here we examine which
+  effect dominates response to increasing VPD: atmospheric demand and
+  increases in ET, or plant physiological response (stomata closure)
+  and decreases in ET. We use Penman-Monteith, combined with
+  semi-empirical optimal stomatal regulation theory and underlying
+  water use efficiency, to develop a theoretical framework for
+  understanding how ET responds to increases in VPD.
 
@@
   higher likelihood of atmospheric demand-driven response (ET
-  increasing). However for forest, grass, and shrub plant types, ET
-  more frequently decreases than increases with rising VPD. Therefore,
-  we may expect reductions in ET due to CO$_2$-induced increases in
-  VPD. This work serves as an example of the utility of our simplified
-  framework for disentangling land-atmosphere feedbacks at multiple
-  scales, including the characterization of ET response in an
-  atmospherically drier, enriched CO$_2$ world.
+  increasing). However for forest, grass, savannah, and shrub plant
+  types, ET more frequently decreases than increases with rising
+  VPD. This work serves as an example of the utility of
+  our simplified framework for disentangling land-atmosphere feedbacks
+  at multiple scales, including the characterization of ET response in
+  an atmospherically drier, enriched CO$_2$ world.
 
@@
 The ET response to changes in VPD alters water partitioning between
-the soil and atmosphere. If ecosystem plant response reduces ET with atmospheric
-drying then soil moisture will be better conserved. This would seem a
-sensible evolutionary strategy to cope with aridity. If stomata were
-fully passive \citep [similar to soil pores, e.g. ][]{Or_2013},
-increased atmospheric aridity would strongly reduce soil moisture
-\citep{Berg_2017}. In turn, this would further increase aridity as low
-soil moisture levels increase the Bowen ratio, and cause increased
-temperature and atmospheric drying \citep[][]{Bouchet_1963,
-  Morton_1965, Brutsaert_1999, Ozdogan_2006, Salvucci_2013,
-  Gentine_2016, Berg_2016}. This however would not seem to be a
-sensible strategy for plants from an evolutionary standpoint. One may
-argue that increases in ET with increasing VPD could increase the
-liklihood of precipitation \citep[e.g.,][]{Findell_2011}. However,
-increases in ET are not a gaurantee
-of increased likelhood of precipitation, which also dependes on the
-environemtnal conditions \citep[][]{Gentine_2013}. Futhermore,  any
-increases in precipitation are likely to be non-local, such that plants giving up
-water to the atmosphere are not gauranteed to reap the benefits of
-water returned to the atmosphere. Using this logic, from an ecoysystem
-evolutionary perspective, water stored in soil is worth much more than
-the \textit{chance} of water returned as precipitation.
+the soil and atmosphere. If ecosystem plant response reduces ET with
+atmospheric drying then soil moisture will be better conserved. This
+would seem a sensible evolutionary strategy to cope with aridity. If
+stomata were fully passive \citep [similar to soil pores,
+e.g. ][]{Or_2013}, increased atmospheric aridity would strongly reduce
+soil moisture \citep{Berg_2017}. In turn, this would further increase
+aridity as low soil moisture levels increase the Bowen ratio, and
+cause increased temperature and atmospheric drying
+\citep[][]{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006,
+  Salvucci_2013, Gentine_2016, Berg_2016}. This however would not seem
+to be a sensible strategy for plants from an evolutionary
+standpoint. As a counterpoint, one may argue that increases in ET with
+increasing VPD could increase the likelihood of precipitation
+\citep[e.g.,][]{Findell_2011}. However, increases in ET do not always
+guarantee an increased likelihood of precipitation, which depending on
+environmental conditions could cause a decrease in the likelihood
+of precipitation \citep[][]{Gentine_2013}. Furthermore, any increases
+in precipitation are likely to be non-local, such that plants giving
+up water to the atmosphere are not guaranteed to reap the benefits of
+water returned from the atmosphere to the soil. Using this logic, from
+an ecosystem evolutionary perspective, water stored in soil is worth
+much more than the chance of water returned as precipitation.
 
@@
 models, and facilitate the disentanglement of complex land-atmosphere
-feedbacks. In the past, similiar approaches were used to understand
+feedbacks. In the past, similar approaches were used to understand
 interactions between stomatal conductance, evapotranspiration and the
 environment \citep[e.g.,][]{Jarvis_1986, Mcnaughton_1991}. However, at
-the time researchers' understandng of the form of VPD's effect on
+the time researchers' understanding of the form of VPD's effect on
 plant physiology was limited, so they could not explore the
@@
 season across space and time, within a PFT \citep{Zhou_2015}. These
-results provide novell tools to examine how ET varies in response to
-VPD, while explicitly including VPD's effect on stoamtal conductance
+results provide novel tools to examine how ET varies in response to
+VPD, while explicitly including VPD's effect on stomatal conductance
 for the first time. In this manuscript, we leverage these tools to
@@
 scheme \citep{Loveland_1999}: cropland (CRO), grass (GRA), deciduous
-broadleaf forest (DBF), evergreen broadlead forest (EBF), evergreen
+broadleaf forest (DBF), evergreen broadleaf forest (EBF), evergreen
 needleleaf forest (ENF), mixed forest (MF), closed shrub (CSH),
@@
                                    & \citet{MEDLYN_2011} \\
-    $g_{1}$  & ecoysystem-scale slope parameter & Pa$^{0.5}$ & \citet{MEDLYN_2017} \\
+    $g_{1}$  & ecosystem-scale slope parameter & Pa$^{0.5}$ & \citet{MEDLYN_2017} \\
     $R$ & universal gas constant & J mol$^{-1}$ K$^{-1}$ & - \\
@@
 ($\frac{\partial \; \text{transpiration}}{\partial \; A}$). So,
-$g_{1-leaf}$ is a leaf-scale term reflecting the tradeoff of water for
+$g_{1-leaf}$ is a leaf-scale term reflecting the trade-off of water for
 carbon uptake. The higher $g_{1-leaf}$, the more open the stomata and
@@
 
-where GPP is the ecoysystem scale gross primary production, and $g_1$
+where GPP is the ecosystem scale gross primary production, and $g_1$
 is an ecosystem scale analogue to $g_{1-leaf}$. We solve for $g_1$
@@
 \citet{Zhou_2015} as a tool to approach this
-problem. \citet{Zhou_2015}, showed that the underlying Water Use
+problem. \citet{Zhou_2015}, showed that  underlying Water Use
 Efficiency (uWUE):
@@
 assumptions. First, the underlying theory used to derive Equations
-\ref{medlyn}, \ref{leaf_medlyn} rely on a implicit assumptions that
+\ref{leaf_medlyn} and \ref{medlyn} rely on an implicit assumption that
 $ET=g_s \cdot VPD$; physically this implies that leaves are perfectly
@@
 types the leaves can become decoupled from the boundary layer
-\citep{De_2017, Medlyn_2017}. Given our focus on the growing season
-which is usually characterized by relatively high insolation, we would
-expect the surface to be well coupled. Additionally, by following
-\citet{Medlyn_2017}'s method for calcualting eco-system scale $g_1$,
+\citep{De_2017, Medlyn_2017}. Given our focus on the growing season,
+which is usually characterized by relatively high insolation inducing
+instability and convective boundary layers, we would
+expect the surface to be generally well coupled. Additionally, by following
+\citet{Medlyn_2017}'s method for calculating ecosystem scale $g_1$
+using PM,
 we have stronger confidence that surface coupling is represented
 within our framework, even if it is not represented in the underlying
-theory used to develop $g_s$. Additionally, the formulation of $uWUE$
-in \citet{Zhou_2014} and \citet{Zhou_2015} as well as our use of
-\citet{Medlyn_2017}'s stomatal conductance model assumes that that
-direct soil evaporation contributions to ET remian relatively
-small. This should be more true druing the growing season than other
-times; however some plant types allow for systematically larger
+theory used to develop stomatal conductance ($g_s$). An additional
+assumption in the formulation of $uWUE$
+\citep{Zhou_2014, Zhou_2015} and
+\citet{Medlyn_2017}'s stomatal conductance model is that
+direct soil evaporation (E) contributions to ET remain
+small relative to transpiration (T). This should be more true during the growing season. However, some plant types allow for systematically larger
 contributions of ``E'' in ET, particularly those with less dense
-canopies and smaller aboslute amounts of transpiration. We therefore
+canopies and smaller absolute amounts of transpiration. We therefore
 might expect that the theory will be most applicable to forest PFTs,
@@
 conserved over space and time within a PFT relative to inter-PFT
-variability. However, we would expect them to vary somehwat with
-environmental conditions including soil water content temperature, and
+variability. However, we would expect them to vary somewhat with
+environmental conditions including soil water content, temperature, and
 with intra-PFT plant-specific characteristics like wood density for
@@
 take the extra care of including and analyzing uncertainty (Section
-\ref{testing}) and examing sensitivity to the exponent of the VPD
-response derived using assumptions of $ET=g_s \cdot VPD$ (Section
+\ref{testing}) and examining sensitivity to the exponent of the VPD
+response, which is derived using assumptions that $ET=g_s \cdot VPD$ (Section
 \ref{functional_form}).
@@
 to typical drought indices such as the Palmer Drought Severity Index
-(PDSI) \citep{Swann_2016}.
+(PDSI) \citep{Swann_2016}. Lastly, all code and data used in this
+analysis, including those used to generate the figures and tables, are
+publicly available at https://github.com/massma/climate\_et.
 
@@
 when taking the derivative. Fortunately, there is a very weak
-dependence of $\sigma$ on VPD in their joint distrubution, and
+dependence of $\sigma$ on VPD in their joint distribution, and
 $\sigma$ is indeed close to unity i.e. $O(1)$ (Figure
@@
 conservative closed shrub (CSH). CRO has higher slope parameter and a
-slightyly higher uWUE ($g_1 = 140.7$ Pa$^{1/2}$; $2.85$ $\mu$-mol [C]
+slightly higher uWUE ($g_1 = 140.7$ Pa$^{1/2}$; $2.85$ $\mu$-mol [C]
 Pa$^{0.5}$ J$^{-1}$ [ET]) compared to CSH ($g1 = 75.1 \, Pa^{1/2}$,
@@
 with increasing VPD for the more water intensive CRO (Figure
-\ref{idealized_sign}). CRO evolved or were bred to prioritize GPP and
+\ref{idealized_sign}). CROs evolved or were bred to prioritize GPP and
 yield and are thus not water conservative. They are very willing to
@@
 has a more water conservative response than the forest, grass, and
-crop plan types, but still respondes by increasing ET with increasing
+crop plan types, but still responds by increasing ET with increasing
 VPD for about a quarter of observations, due to the high aridity (VPD)
@@
 aerodynamic conductance (GRA, CSH, CRO). The savannah (WSA, SAV) PFT's
-scaling is somewhere betwen GRA, CSH, and CRO, and DBF, EBF, ENF, and
+scaling is somewhere between GRA, CSH, and CRO, and DBF, EBF, ENF, and
 MF, due to higher variability and surface roughness.
@@
 observations. The more water conservative PFTs generally exhibit
-higher frequency of negative ET response, espcially when one factors
+higher frequency of negative ET response, especially when one factors
 in the distribution of environmental VPD (e.g. SAV and WSA grow in
@@
 negative response, and shifts many of the other PFTs, which still
-exhibit a high frequency of negative ET reponse, towards more frequent
-occurence of postive ET response than the theory and Figure
+exhibit a high frequency of negative ET response, towards more frequent
+occurrence of positive ET response than the theory and Figure
 \ref{idealized_sign} would suggest. The bulk statistics motivate a
@@
 capture the leading order behavior of the ET dependence on VPD. Given
-the assumptions we have made, and the uncertainties of flux tower
+the assumptions we made, and the uncertainties of flux tower
 observations themselves, we expect a relatively large amount of noise
@@
 DBF and MF, CSH observations also match our theory to leading order,
-albiet with a bit more variability as a function of VPD. Again, the
+albeit with a bit more variability as a function of VPD. Again, the
 VPD dependence mostly determined from the slope parameter ($g_1$)
@@
 
-Biases between the theory and observations are similiar for ENF, WSA,
+Biases between the theory and observations are similar for ENF, WSA,
 CRO and SAV. At low VPD the theory and the observations match
 well. However, at high VPD (upper 10-20\% of observations) the theory
-is biased slightly towards postive reponse as compared to the
+is biased slightly towards positive response as compared to the
 observations (Figure \ref{test_sign}). For ENF this slight bias could
@@
 \ref{idealized_sign}). Therefore, to explain the observations one of
-the varibles in the sign term must change at extreme VPD (upper
-10-20\%). The most likely candidiate is $uWUE$, which we have assumed
+the variables in the sign term must change at extreme VPD (upper
+10-20\%). The most likely candidate is $uWUE$, which we have assumed
 constant to meet our goal of developing intuition for leading order
@@
 which should be correlated with high VPD. The theory for EBF exhibits
-similiar limiatations as for ENF, WSA, and SAV, except a larger
+similar limitations as for ENF, WSA, and SAV, except a larger
 portion of observations are biased negative as compared to the theory
@@
 superposition of sites with C3 plants (at low environmental VPD) and
-C4 plants (at high environmental VPD) would explain observed shift
+C4 plants (at high environmental VPD) would explain the observed shift
 back towards negative ET response at high VPD when all sites are
@@
 theory would validate against observations much better if these
-sources of varaibility were accounted for, at a cost of increased
+sources of variability were accounted for, at a cost of increased
 complexity and analytical opacity.
@@
 and conceptualized with shifts in $g_1$ and/or $uWUE$ due to
-site-spcific plant type variability (e.g. more arid-adapted ecosystems
-at more arid sites), or temporal variability for some enviornmental
+site-specific plant type variability (e.g. more arid-adapted ecosystems
+at more arid sites), or temporal variability for some environmental
 conditions (e.g. decreases in $uWUE$ at extremely low soil
-moisture). To focus on general behaviro and develop intuition for
+moisture). To focus on general behavior and develop intuition for
 PFT-scale response, we have ignored these sources of variability in
@@
 expectation based on simplified theory. The sign term is most strongly
-scaled by $g_a$, and in general the sign term shifts towards positive
-values as VPD increases. The willingness with which a given plant type
+scaled by $g_a$, and in general the occurrence of positive ET response
+increases as VPD increases. The willingness with which a given plant type
 evolved to use water dictates the occurrence of positive versus
-negative ET response. Water conservative plants are highly effective
+negative ET response. Water conservative ecosystems are highly effective
 at mitigating the effects of atmospheric demand, and can store water
@@
 likelihood that CRO sites are optimally irrigated and not water
-stressed, suggesting that observed depatures from the theory for CRO
-are due to other factors (i.e. photosythesis pathway) than soil
+stressed, suggesting that observed departures from the theory for CRO
+are due to other factors (i.e. photosynthesis pathway) than soil
 moisture. For ENF, EBF, SAV, and WSA, systematic SWC-induced biases in
@@
 conditions, although large variability in the SWC-$\sigma \cdot uWUE$
-relationship hampers interpretation for WSA. In crontast to ENF, CRO,
-and DBF, for GRA the relationship between $\sigma \cdot uWUE$ and SWC
-is more linear and affects a greater portion of observations; about
-60\% of observations. Clearly, for GRA soil water frequently impacts
-plant function and alters ET response, and our theory is limited for
-the majority of encironmental condiitons. It is therefore not
-suprising that our theory tested poorly againt the data for GRA,
-relative to to the other PFTs (Section \ref{testing}). For all other
-PFTs occurence of soil mosture impacts was rare enough to not manifest
-itself in bulk statistics and figures. A possible exception is CSH,
-for which there are only two sites, which precludes comment on the
+relationship hampers interpretation for WSA. CSH presents a special
+case: the limited number of sites (2) preclude comment on the
 exact PFT-wide relationship.
+
+In contrast to ENF, CRO, and DBF, for GRA the relationship between
+$\sigma \cdot uWUE$ and SWC is more linear and affects a greater
+portion of observations; about 60\% of observations. Clearly, for GRA
+soil water frequently impacts plant function and alters ET response,
+and our theory is limited for the majority of environmental
+conditions. It is therefore not surprising that our theory tested
+poorly against the data for GRA, relative to to the other PFTs
+(Section \ref{testing}). For all other PFTs occurrence of soil
+moisture impacts was rare enough to not manifest itself in bulk
+statistics and figures.
 
@@
 assumption) and observations. It is also perhaps not coincidental that
-the protion of $\sigma \cdot uWUE$ affected by low soil moisture
-observations is similiar to the portion of observations that to not
-match our theory at high VPD.
+the portion of $\sigma \cdot uWUE$ affected by low soil moisture
+observations is similar to the portion of observations that to not
+match our theory at high VPD. Indeed, this may suggest that for all PFTs
+except for GRA, coupling and feedbacks between SWC, VPD and plant
+function are relatively rare.
 
@@
 rare (<30\% of observations) and does not manifest itself in the
-majority of observations and bulk statistics. This result is
-consistent with However, for GRA SWC has
+majority of observations and bulk statistics. However, for GRA SWC has
 a strong effect decreasing water use efficiency for the majority of
@@
 will be possible for future analysis to explore the importance of soil
-moisture on plant response to VPD.
+moisture on plant response to VPD, and feedbacks between plant
+function, SWC, and VPD.
 
@@
 critical VPD$_{crit}$, assuming the scaling and plant terms  (e.g. aerodynamic
-conductance, $\Delta$, $g_1$ and $uWUE$) are held fixed. This result warrants futher
+conductance, $\Delta$, $g_1$ and $uWUE$) are held fixed. This result warrants further
 investigation, because to our knowledge no one has derived the
 theoretical ecosystem-scale relationship between ET and VPD while
-controlling for enivironmental conditiions. In particular, from
+controlling for environmental conditions. In particular, from
 personal communication, there is an apparent lack of consensus over
@@
 hypothesizing about any feedbacks between the land surface and the
-atmopshere, we analyze why our derived ET-VPD relationship is concave
-upward, particularly with respect to the exponent of VPD depedence in
-$uWUE$ and the Medlyn unified stomatal conductace model.
+atmosphere, we analyze why our derived ET-VPD relationship is concave
+upward, particularly with respect to the exponent of VPD dependence in
+$uWUE$ and the Medlyn unified stomatal conductance model.
 
 There is a theoretical basis for the square root VPD dependence in
-both the stomatal conducatance model and $uWUE$ based on the
-assumption that stomata behave to maximize carbon gain while
-minimizing water loss, which observations also generally support
-\citep{Lloyd_1991, MEDLYN_2011, Lin_2015, Zhou_2014, Zhou_2015,
-  Medlyn_2017}. However, some purely empiriacle results which fit the
-exponent of the VPD dependence to data have shown that it may vary slightly
-from 1/2, suggesting that stomata, as well as ecosystem-scale
-quantities based on stomata theroy, may not always function optimally
-\citep{Zhou_2015, Lin_2018}. Specifically with regards to $uWUE$, one
-would not expect that this ecosystem scale WUE quantity will responde
-to VPD exactly analgalously to stomata. Direct soil evaporation's
-contributions to ET should shift the exponent of the VPD dependence,
-especially for conditions of low GPP when we would expect a
-systematically larger portion of direct soil evaporation contributions
-to ET, because we would also expect lower amounts of
-transpiration. \citet{Zhou_2015}'s results corroborate this: they
-found a mean imprically fit exponential VPD dependence of 0.55,
-varying slightly from the theoretically optimal value of of 0.5 for
-Ameriflux sites. Interpreting \citet{Lin_2018}'s results, which also
-show variance in the empiracle exponent of the VPD dependence of the
-stomatal conductance model, is more difficult as \citet{Lin_2018} does
+both the stomatal conductance model and $uWUE$ based on the assumption
+that stomata behave to maximize carbon gain while minimizing water
+loss, which observations also generally support \citep{Lloyd_1991,
+  MEDLYN_2011, Lin_2015, Zhou_2014, Zhou_2015, Medlyn_2017}. However,
+some purely empirical results which fit the exponent of the VPD
+dependence to data have shown that it may vary slightly from 1/2,
+suggesting that stomata, as well as ecosystem-scale quantities based
+on stomata theory, may not always function optimally \citep{Zhou_2015,
+  Lin_2018}. Specifically with regards to $uWUE$, one would not expect
+that this ecosystem scale WUE quantity will respond to VPD exactly
+analogously to stomata. Direct soil evaporation's contributions to ET
+should shift the exponent of the VPD dependence, especially for
+conditions of low GPP when we would expect a systematically larger
+portion of direct soil evaporation contributions to ET, because we
+would also expect lower amounts of transpiration. \citet{Zhou_2015}'s
+results corroborate this: they found a mean empirically fit
+exponential VPD dependence of 0.55, varying slightly from the
+theoretically optimal value of of 0.5 for AmeriFlux
+sites. Interpreting \citet{Lin_2018}'s results, which also show
+variance in the empirical exponent of the VPD dependence of the
+stomatal conductance model, is more difficult as \citet{Lin_2018} do
 not handle GPP/A dependence of stomatal conductance in a directly
-analgoalous manner to the optimal theory in \citet{MEDLYN_2011} and
-\citet{Medlyn_2017}. Regardless, given that these recent results
-\citep{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017} on the
-relatioship between VPD, GPP, and ET \citep{MEDLYN_2011, Zhou_2014,
-  Zhou_2015, Medlyn_2017} form the backbone of our analysis and are
-what allowed us to derive an explicit ET expression for the first
-time (Equation \ref{et}), we will analyze if and how assumptions about
-the exponent of the VPD dependence impacts the shape of the ET-VPD
-dependence. By introducing $n$ and $m$ we can free our stomatal
+analogous manner to the optimal theory in \citet{MEDLYN_2011} and
+\citet{Medlyn_2017}. Regardless, given that these recent results on
+the relationship between VPD, GPP, and ET \citep{MEDLYN_2011,
+  Zhou_2014, Zhou_2015, Medlyn_2017} form the backbone of our analysis
+and are what allowed us to derive an explicit ET expression for the
+first time (Equation \ref{et}), we will analyze if and how assumptions
+about the exponent of the VPD dependence impacts the shape of the
+ET-VPD dependence. By introducing $n$ and $m$ we can free our stomatal
 conductance model from assumptions about VPD dependence:
@@
 shape of the ET-VPD dependence we find the roots of the second
-deriviative of ET, using Equation \ref{m_n} for stomatal conductance
-($g_s$, with respect to VPD: \input{d2_solutions.tex}. With this
-result we've detemined a family of curves seperating concave up from
+derivative of ET, using Equation \ref{m_n} for stomatal conductance
+($g_s$, with respect to VPD: \input{d2_solutions.tex} With this
+result we've defined the family of curves separating concave up from
 concave down ET solutions (Figure \ref{concave}). These curves are
-only functions of the exponent of the VPD depedence and a quantity we
+only functions of the exponent of the VPD dependence and a quantity we
 call non-dimensional VPD ($VPD^m/g*$). Several important
@@
     non-dimensional VPD ($VPD^m/g_*$), but to aide physical
-    interpretation the horizontal axis is additionally provided interms of dimensionalized VPD assuming $m=1/2$ and
+    interpretation the horizontal axis is additionally provided in terms of dimensionalized VPD assuming $m=1/2$ and
     $g_*=110\; Pa^{1/2}$ (average of all PFT $g_1$). The vertical axis
     has a different interpretation depending on the solution
-    curve. For the blue line ($m$ varying), it correspondes to $m$,
-    for the orange line ($n$ varying) it correspondes to $n$, and for
-    the green line it correspondes to the value of both $n$ and $m$
-    ($n=m$). Regions of the parameter space that corresponde to
+    curve. For the blue line ($m$ varying), it corresponds to $m$,
+    for the orange line ($n$ varying) it corresponds to $n$, and for
+    the green line it corresponds to the value of both $n$ and $m$
+    ($n=m$). Regions of the parameter space that correspond to
     concave up and concave down results are labeled.}
@@
   \item In general, increasing the exponent of VPD dependence
-    increases the liklihood of a concave up result. Additionally, as
+    increases the likelihood of a concave up result. Additionally, as
     the exponent of VPD dependence increases from the optimum value of
@@
     of ET-VPD dependence, future research investment in understanding
-    the exact VPD depedence of stomatal conductance, and further
-    reconciliation of empiracle and theoretical stomatal and
-    ecosostyem behavior should be prioritized.
+    the exact VPD dependence of stomatal conductance, and further
+    reconciliation of empirical and theoretical stomatal and
+    ecosystem behavior should be prioritized.
 \end{enumerate}
-However, while it is possible that in the future some other form of
+While it is possible that in the future some other form of
 VPD dependence is derived, at present \cite{MEDLYN_2011} and
 \cite{Zhou_2014} firmly established n=m=1/2 as the most likely
-candidiate given current theory and empiracle data. Additionally, we
+candidate given current theory and empirical data. Additionally, we
 argue that a concave up result matches physical intuition more than a
@@
 through the phloem and xylem. To accomplish this, stomata must remain
-slightly open \citep{De_2013, Nikinmaa_2013, Ryan_2014}. Futhermore,
+slightly open \citep{De_2013, Nikinmaa_2013, Ryan_2014}. Furthermore,
 even if complete stomatal closure were possible, cuticular water loss
@@
 closure. Therefore, in the limit as VPD becomes large and we assume
-plants are exercising all strategies to reduce ET, any futher increase
+plants are exercising all strategies to reduce ET, any further increase
 in VPD should result in an increase in ET through cuticular water loss
 and/or direct soil evaporation. This inevitable transition from
-conditions where stomata responde strongly to VPD to conditions where
+conditions when stomata respond strongly to VPD to conditions when
 stomata response is asymptoting towards full closure would cause a
@@
 intuitive result that atmospheric demand dominates plant response when
-atmospheric deman is high.
-
-This analysis allow us to understand the theoretical shape of the ET
-response to VPD with environmental conditions held fixed which would
-be difficult to disentagle purely from flux observations, given the
-relatively fast time scale of plant response and the non-stationarity
-of [solar forced] environmental conditions  over the relatively
-coarse (half hourly) flux estimates (which is required to obtain
-robust eddy-covariance statistics). This fundamental understanding of
-ET response to VPD with environmental conditoins held fixed can be
-built upon with future work examing how changes in VPD and enviornment
-conditions (e.g. soil water storage) feedback upon one another. In the
-soil water storage example, over very long time scales extremely high
-VPD perturbations coupled with no precipitation could result in
-decreases in soil water storage such that water becomes limiting. This
-could be represented by an extension of our framework in which uWUE is
-allowed to decrease with decreasing SWC, as observed in Secion
-\ref{swc_section}. Here, we focus our results by assuming constant
-PFT-wide conditions to build baseline intuition for ET-VPD
-depedence. For most PFTs, the theory with plant function held fixed
-matches the leading order behaviro of the observations where plant
-function varies (Section \ref{testing}).
+atmospheric demand is high.
+
+This analysis allows us to understand the theoretical shape of the ET
+response to VPD with environmental conditions held
+fixed. Accomplishing this with purely statistical
+methods applied to flux observations would be very difficult, given
+the relatively fast time scale of plant response and the
+non-stationarity of [solar forced] environmental conditions over the
+relatively coarse (half hourly) flux estimates (which is required to
+obtain robust eddy-covariance statistics). Our results on the shape of
+the ET-VPD curve with environmental conditions held
+fixed can be built upon with future work examining how changes in VPD
+and environmental conditions (e.g. soil water storage) feedback upon
+one another. In the soil water storage example, over very long time
+scales extremely high VPD perturbations coupled with no precipitation
+could result in decreases in soil water storage such that water
+becomes limiting. This could be represented by an extension of our
+framework in which uWUE is allowed to decrease with decreasing SWC, as
+observed in Section \ref{swc_section}. Here, we focus our results by
+assuming constant PFT-wide conditions to build baseline intuition for
+ET-VPD dependence. For most PFTs, the theory with plant function held
+fixed matches the leading order behavior of the observations where
+plant function varies (Section \ref{testing}).
 
@@
 response to increasing VPD). SAV and WSA exhibit positive ET response
-to VPD not necessarily because of poor water conservation stragegies
-relative to other PFTs, but because of greater occurence of high
+to VPD not necessarily because of poor water conservation strategies
+relative to other PFTs, but because of greater occurrence of high
 atmospheric demand (VPD) relative to other PFTs. Observations of ET
@@
 Lastly, our paper builds important intuition for how plants will
-respond to VPD pertubations. We show that given optimal stomatal
-function and fixed environemntal conditions, the ET-VPD depednence is
+respond to VPD perturbations. We show that given optimal stomatal
+function and fixed environmental conditions, the ET-VPD dependence is
 theoretically concave up, with ET increasing with increasing VPD as
@@
 future research should focus on fully understanding the form of
-stomata VPD dependence, as this result is sensitive to the exponent of
+stomatal VPD dependence, as this result is sensitive to the exponent of
 VPD dependence, which we currently believe is 1/2 \citep{MEDLYN_2011,
@@
 \citep[e.g.,][]{Rigden_2017}. Feedbacks between the land and the
-atmosphere may alter the net response to a long-timescale global VPD perturabation,
+atmosphere may alter the net response to a long-timescale global VPD perturbation,
 but our focus on the one way plant response to a VPD perturbation in
-the atmopsheric boundary layer is an important first step to
+the atmospheric boundary layer is an important first step to
 disentangling such feedbacks, both in observations and model
@@
 dependence on implicit relationships between GPP, VPD, and ET, we
-allow for explicit, accuruate future analysis of plant-VPD feedbacks
+allow for explicit, accurate future analysis of plant-VPD feedbacks
 in the atmospheric boundary layer (Equation \ref{et}). Our approach
-can also be extended to examine varying plant response to more nuanced
+can be extended to examine varying plant response to more nuanced
 consideration of plant type and climate. Any plant physiological
@@
 16-44869. All code and data used in this analysis, including those
-used to generate the figures and tables, is publicly available at
+used to generate the figures and tables, are publicly available at
 https://github.com/massma/climate\_et.

### 2018-04-03 src/paper_figure_code/concave.py

@@ -44,7 +44,7 @@ fh = open("../../doc/paper/d2_solutions.tex", "w")
 fh.write("\\begin{equation}\n")
 fh.write("\\frac{\\partial^2 \; ET}{\\partial \\; VPD^2} = 0 "\
          "\\quad \\forall \\quad")
-fh.write("%s%s%s" % ("\\frac{VPD^m}{g1} = ", latex(soln[0]), "\n"))
+fh.write("%s%s%s" % ("\\frac{VPD^m}{g1} = ", latex(soln[0]), ".\n"))
 fh.write("\\label{curves}\n")
 fh.write("\\end{equation}\n")
 fh.close()

### 2018-04-03 doc/paper/gcb_sub_vpd_et_paper.tex
successfully generated diff for Pierre

--- 2018-04-01 doc/paper/vpd_et_paper.tex
@@
 \documentclass[draft,linenumbers]{agujournal}
-% \draftfalse
-\usepackage{makecell} \usepackage{multirow} %AM for tables
+\draftfalse \usepackage{makecell} \usepackage{multirow} %AM for tables
 \usepackage{booktabs} \drafttrue
 
-\journalname{Agricultural and Forest Meteorology: When does VPD drive
-  or reduce ET?}
+\journalname{Global Change Biology: When does VPD drive or reduce ET?}
 
@@
 \title{When does vapor pressure deficit drive or reduce
-  evapotranspiration?}
-
-\authors{Adam Massmann\affil{a}, Pierre Gentine\affil{a}, Changjie
-  Lin\affil{a,b}}
-
-\affiliation{a}{Department of Earth and Environmental Engineering,
-  Columbia University, New York, NY 10027} \affiliation{b}{State Key
+evapotranspiration?\footnote{Primary Research Article}}
+
+\authors{Adam Massmann\affil{1}, Pierre Gentine\affil{1}, Changjie
+Lin\affil{1,2}}
+
+\affiliation{1}{Department of Earth and Environmental Engineering,
+Columbia University, New York, NY 10027} \affiliation{2}{State Key
 Laboratory of Hydroscience and Engineering, Department of Hydraulic
@@
 
-\begin{abstract}
-  Increases to vapor pressure deficit (VPD) and atmospheric demand for
-  water are expected with rising atmospheric CO$_2$. While increased
-  evapotranspiration (ET) in response to increased atmospheric demand
-  seems intuitive, plants are capable of reducing ET in response to
-  increased VPD by closing their stomata, in an effort to conserve
-  water. Here we examine which effect dominates response to increasing
-  VPD: atmospheric demand and increases in ET, or plant physiological
-  response (stomata closure) and decreases in ET. We use
-  Penman-Monteith, combined with semi-empirical optimal stomatal
+\begin{abstract} Increases to vapor pressure deficit (VPD) and
+atmospheric demand for water are expected with rising atmospheric
+CO$_2$. While increased evapotranspiration (ET) in response to
+increased atmospheric demand seems intuitive, plants are capable of
+reducing ET in response to increased VPD by closing their stomata, in
+an effort to conserve water. Here we examine which effect dominates
+response to increasing VPD: atmospheric demand and increases in ET, or
+plant physiological response (stomata closure) and decreases in ET. We
+use Penman-Monteith, combined with semi-empirical optimal stomatal
 regulation theory and underlying water use efficiency, to develop a
 theoretical framework for understanding how ET responds to increases
-  in VPD.
-
-  The theory suggests that for most environmental conditions and plant
-  types, plant physiological response dominates and ET decreases with
-  increasing VPD. Plants that are evolved or bred to prioritize
-  primary production over water conservation (e.g. crops) exhibit a
-  higher likelihood of atmospheric demand-driven response (ET
-  increasing). However for forest, grass, and shrub plant types, ET
-  more frequently decreases than increases with rising VPD. Therefore,
-  we may expect reductions in ET due to CO$_2$-induced increases in
-  VPD. This work serves as an example of the utility of our simplified
-  framework for disentangling land-atmosphere feedbacks at multiple
-  scales, including the characterization of ET response in an
-  atmospherically drier, enriched CO$_2$ world.
+in VPD.  The theory suggests that for most environmental conditions
+and plant types, plant physiological response dominates and ET
+decreases with increasing VPD. Plants that are evolved or bred to
+prioritize primary production over water conservation (e.g. crops)
+exhibit a higher likelihood of atmospheric demand-driven response (ET
+increasing). However for forest, grass, and shrub plant types, ET more
+frequently decreases than increases with rising VPD. Therefore, we may
+expect reductions in ET due to CO$_2$-induced increases in VPD. This
+work serves as an example of the utility of our simplified framework
+for disentangling land-atmosphere feedbacks at multiple scales,
+including the characterization of ET response in an atmospherically
+drier, enriched CO$_2$ world.
 
@@
 partially close in response to increased atmospheric dryness
-\citep{Farquhar_1978, Ball_1987, Leuning_1990, MEDLYN_2011}. This
-leads to a plant physiology perspective, in which an increase in VPD,
-particularly in well-watered soil conditions, may actually correspond
-to a decrease in ET because of stomatal closure
-\citep[e.g.][]{Rigden_2017}.  In other words, the question ``When does
-VPD drive or reduce ET?'' can be related to whether plant regulation
-or atmospheric demand dominates ET response.
+\citep{Ball_1987, Leuning_1990, MEDLYN_2011}. This leads to a plant
+physiology perspective, in which an increase in VPD, particularly in
+well-watered soil conditions, may actually correspond to a decrease in
+ET because of stomatal closure.  In other words, the question ``When
+does VPD drive or reduce ET?'' can be related to whether plant
+regulation or atmospheric demand dominates ET response.
 
@@
 Morton_1965, Brutsaert_1999, Ozdogan_2006, Salvucci_2013,
-  Gentine_2016, Berg_2016}. This however would not seem to be a
-sensible strategy for plants from an evolutionary standpoint.
+Gentine_2016, Berg_2016}. This however would not seem to be a sensible
+strategy for plants from an evolutionary standpoint.
 
 We can use intuition about plant water conservation strategy to
-hypothesize about ET response to changes in VPD. Plants and ecosystems
-that evolved to conserve water, such as arid shrubs or savannah,
-should be more likely to reduce ET with increasing VPD, and plants
-that have evolved or have been engineered to care little about water,
-such as crops, will be more likely to increase ET with increasing
-VPD. Atmospheric conditions must matter as well. At the ecosystem
-scale, there are limits to plant water conservation strategies. As
-atmospheric demand for water (VPD) increases, ecosystems should begin
-to reach their water conservation limits and might not be able to
-entirely limit ET flux to the atmosphere. At this stage any further
-increase in VPD will most likely drive a (limited) increase in ET,
-because the increase in atmospheric demand for water overwhelms the
-limited plant response to conserve water.
-
-The objective of the present manuscript is to use reasonable
-approximations as a tool to develop intuition for plant response to
-atmospheric drying and evaluate the VPD dependence of ET. This
-intuition will aid interpretation of observations, full complexity
-models, and facilitate the disentanglement of complex land-atmosphere
-feedbacks. In the past, similiar approaches were used to understand
-interactions between stomatal conductance, evapotranspiration and the
-environment \citep[e.g.,][]{Jarvis_1986, Mcnaughton_1991}. However, at
-the time researchers' understandng of the form of VPD's effect on
-plant physiology was limited, so they could not explore the
-sensitivity of ET to VPD, including VPD's effect on stomatal
-conductance and plant function.
-
-Recent results have drastically improved our understanding of VPD's
-impact on physiology. \citet{MEDLYN_2011} developed a model for
-leaf-scale stomatal conductance ($g_s$), including VPD response, by
-combining an optimal photosynthesis theory \citep{Farquhar_1980,
-  Katul_2010} with an empirical approach, and extended use of this
-model to the ecosystem scale in \citet{Medlyn_2017}. Additionally,
-\citet{Zhou_2014} demonstrated that a quantity underlying water use
-efficiency $\left(uWUE = \frac{GPP\; \sqrt{VPD}}{ET}\right)$ properly
-captures a constant relationship between GPP, ET, and VPD over a
-diurnal cycle. uWUE is also remarkably well conserved in the growing
-season across space and time, within a PFT \citep{Zhou_2015}. These
-results provide novell tools to examine how ET varies in response to
-VPD, while explicitly including VPD's effect on stoamtal conductance
-for the first time. In this manuscript, we leverage these tools to
-derive the theoretical response of ET to VPD using a Penman Monteith
-framework. Our theory is validated and tested at multiple
+hypothesize about ET response to changes in VPD. Plants that evolved
+to conserve water, such as arid shrubs, should be more likely to
+reduce ET with increasing VPD, and plants that have evolved or have
+been engineered to care little about water, such as crops, will be
+more likely to increase ET with increasing VPD. Atmospheric conditions
+must matter as well. At the ecosystem scale, there are limits to plant
+water conservation strategies. As atmospheric demand for water (VPD)
+increases, ecosystems should begin to reach their water conservation
+limits and might not be able to entirely limit ET flux to the
+atmosphere. At this stage any further increase in VPD will most likely
+drive a (limited) increase in ET, because the increase in atmospheric
+demand for water overwhelms the limited plant response to conserve
+water.
+
+The objective of the present manuscript is to evaluate the VPD
+dependence of ET, in non-extreme soil drought conditions. The goal of
+this paper is to use reasonable approximations as a tool to develop
+intuition for plant response to atmospheric drying. This intuition
+will aid interpretation of observations and full complexity models. In
+order to quantify plant response to perturbations in VPD, we apply a
+Penman-Monteith framework to derive the theoretical response of ET to
+VPD. The model is then validated and tested at multiple
 eddy-covariance stations spanning various climates and plant
@@
 \section{Materials and Methods}
+\label{methods}
 \subsection{Data}
-\label{data}
-We use both meteorological and eddy-covariance data from the
-FLUXNET2015 database (data available at \sloppy
-https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/ \sloppy),
-including all sites with at least four years of data and observations
-of the variables described in the methods section (Section
-\ref{methods}). Seventy-three sites met these requirements, and were
-grouped into nine plant functional types (PFT) according to the
+\label{data} We use both meteorological and eddy-covariance data from
+the FLUXNET2015 database, including all sites with at least four years
+of data (data available at \sloppy
+https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/ \sloppy). Each
+site's plant functional type (PFT) was classified using the
 International Geosphere-Biosphere Programme vegetation classification
-scheme \citep{Loveland_1999}: cropland (CRO), grass (GRA), deciduous
-broadleaf forest (DBF), evergreen broadlead forest (EBF), evergreen
-needleleaf forest (ENF), mixed forest (MF), closed shrub (CSH),
-savannah (SAV), and woody savannah (WSA) (site locations in Figure
-\ref{map_fig}).
-
-\begin{figure}
-  \centering \includegraphics[width=\textwidth]{./map.pdf}
-  \caption{Plant functional type and location of FLUXNET2015 sites
-    used in this analysis.}
+scheme \citep{Loveland_1999}. The physical constants used in the
+stomatal conductance model in the Methods section (Section
+\ref{methods}) are only published for six plant functional types
+(PFTs): crops (CRO), deciduous broadleaf forest (DBF), evergreen
+broadleaf forests (EBF), evergreen needleleaf forest (ENF), grass
+(GRA), and closed shrub (CSH) (see Table \ref{pft}). There are 62
+sites with these plant functional types, and their location is shown
+in Figure \ref{map_fig}.
+
+\begin{table}
+  \caption{Plant functional types, their abbreviation, Medlyn
+coefficient (from \citet{Lin_2015}, via \citet{Franks_2017}), and
+calculated uWUE. uWUE from \citet{Zhou_2015}, including the observed
+standard deviation is shown for comparison. Note that uWUE from
+\citet{Zhou_2015} is calculated from a different set of sites, and
+that units are converted such that the quantities work with Equations
+1-8 with the variables defined Table \ref{definitions}.}  \small
+\label{pft} \centering
+\begin{tabular}{l c c @{\qquad} c c} \hline
+\multirow{2}[3]{*}{Abbreviation} & \multirow{2}[3]{*}{PFT} &
+\multirow{2}[3]{*}{$g_1$ (Pa$^{0.5}$)} & \multicolumn{2}{c}{uWUE
+($\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET])} \\ \cmidrule{4-5}
+
+  & & & fitted & \citet{Zhou_2015} \\
+
+  \hline DBF & deciduous broadleaf forest & 140.7 & 2.63 & 3.12 $\pm$
+0.52 \\ EBF & evergreen broadleaf forest & 130.3 & 2.56 & N/A \\ ENF &
+evergreen needleleaf forest & 74.3 & 3.45 & 3.30 $\pm$ 0.91 \\ CRO &
+cropland & 183.1 & 2.27 & 3.80 $\pm$ 1.01 \\ CSH & closed shrub &
+148.6 & 1.66 & 2.18 $\pm$ 0.44 \\ GRA & grassland (assumed C3) & 166.0
+& 1.92 & 2.68 $\pm$ 0.61 \\ \hline
+\end{tabular}
+\end{table}
+
+\begin{figure} \centering
+\includegraphics[width=\textwidth]{./map.pdf}
+\caption{Plant functional type and location of FLUXNET2015 sites used
+in this analysis.}
 \label{map_fig}
@@
 \item To isolate the growing season, we only use days in which the
-  average Gross Primary Productivity (GPP) exceeds 10\% of the
-  observed 95th percentile of GPP for a given site. GPP is calculated
-  using the nighttime respiration partitioning method.
+average Gross Primary Productivity (GPP) exceeds 10\% of the observed
+95th percentile of GPP for a given site. GPP is calculated using the
+nighttime respiration partitioning method.
 \item We remove days with rain and the day following to avoid issues
-  with rain interception and sensor saturation at high relative
-  humidity (\cite{MEDLYN_2011}).
-\end{itemize}
-Additionally, as in \citet{Lin_2018}, we restrict data to the daytime,
-which is identified when downwelling shortwave radiation is greater
-than 50 W m$^{-2}$ and sensible heat flux is greater than 5 W
-m$^{-2}$. To reduce the chance of sensor saturation at high relative
-humidity, we remove all time steps for which VPD is less than .01 kPa,
-and to reduce errors at low windspeeds we remove all periods with wind
-magnitudes less than 0.5 m s$^{-1}$. Timesteps with negative observed
-GPP or ET are also removed, and we aggregate half hourly data to
-hourly averages to reduce noise \citep{Lin_2018}. After these quality
-control procedures, 434,765 upscaled hourly observations remain.
+with rain interception and sensor saturation at high relative humidity
+(\cite{MEDLYN_2011}).
+\end{itemize} Additionally, as in \citet{Lin_2018}, we restrict data
+to the daytime, which is identified when downwelling shortwave
+radiation is greater than 50 W m$^{-2}$ and sensible heat flux is
+greater than 5 W m$^{-2}$. To reduce the chance of sensor saturation
+at high relative humidity, we remove all time steps for which VPD is
+less than .01 kPa, and to reduce errors at low windspeeds we remove
+all periods with wind magnitudes less than 0.5 m s$^{-1}$. Timesteps
+with negative observed GPP or ET are also removed, and we aggregate
+half hourly data to hourly averages to reduce noise
+\citep{Lin_2018}. After these quality control procedures, 335,939
+upscaled hourly observations remain.
 
 \subsection{Methods}
-\label{methods}
-The Penman-Monteith equation \citep [hereafter PM,][]{Penman_1948,
-  Monteith_1965} estimates ET as a function of observable atmospheric
-variables and surface conductances:
+\label{methods} The Penman-Monteith equation \citep [hereafter
+PM,][]{Penman_1948, Monteith_1965} estimates ET as a function of
+observable atmospheric variables and surface conductances:
 \begin{linenomath*}
   \begin{equation}
-    \label{orig_pen}
-    ET = \frac{\Delta R_{net} + g_a \rho_a c_p VPD}{\Delta + \gamma(1 + \frac{g_a}{g_s})},
-  \end{equation}
-\end{linenomath*}
-where $\Delta$ is the change in saturation vapor pressure with
-temperature, given by Clausius-Clapeyron ($\frac{d \; e_s}{d \; T}$),
-$R_{net}$ is the net radiation minus ground heat flux, $g_a$ is
-aerodynamic conductance, $\rho_a$ is air density, $c_p$ is specific
-heat of air at constant pressure, $\gamma$ is the psychometric
-constant, and $g_s$ is the stomatal conductance (Table
+    \label{orig_pen} ET = \frac{\Delta R_{net} + g_a \rho_a c_p
+VPD}{\Delta + \gamma(1 + \frac{g_a}{g_s})},
+  \end{equation}
+\end{linenomath*} where $\Delta$ is the change in saturation vapor
+pressure with temperature, given by Clausius-Clapeyron ($\frac{d \;
+e_s}{d \; T}$), $R_{net}$ is the net radiation minus ground heat flux,
+$g_a$ is aerodynamic conductance, $\rho_a$ is air density, $c_p$ is
+specific heat of air at constant pressure, $\gamma$ is the
+psychometric constant, and $g_s$ is the stomatal conductance (Table
 \ref{definitions}).
@@
 values are calculated, if applicable.}
-  \label{definitions}
-  \centering \small
-  \begin{tabular}{l c c c}
-    \hline
-    Variable & Description & Units & Citation \\
-    \hline
-    $e_s$  & saturation vapor pressure & Pa  & - \\ 
-    $T$  & temperature  & K & - \\
-    $P$  & pressure & Pa  & - \\
-    $\Delta$  & $\frac{\partial e_s}{\partial T}$ & Pa K$^{-1}$ & - \\
-    $R_{net}$  & net radiation at land surface minus ground heat flux & W m$^{-2}$   & - \\
-    $g_a$  & aerodynamic conductance & m s$^{-1}$  & \citet{Shuttleworth_2012} \\
-    $\rho_a$  & air density & kg m$^{-3}$  & - \\
-    $c_p$  & specific heat capacity of air at constant pressure & J K$^{-1}$ kg$^{-1}$ & - \\
-    $VPD$  & vapor pressure deficit & Pa  & - \\
-    $\gamma$  & psychometric constant & Pa K$^{-1}$   & - \\
-    $g_s$  & stomatal conductance & m s$^{-1}$  & \citet{MEDLYN_2011} \\
-    $g_{l-s}$  & leaf-scale stomatal conductance & mol m$^{-2}$ s$^{-1}$  & \citet{MEDLYN_2011} \\
-    $R$ & universal gas constant & J mol$^{-1}$ K$^{-1}$ & - \\
-    $R_{air}$ & gas constant of air & J  K$^{-1}$ kg$^{-1}$ & - \\
-    $LAI$ & leaf area index & -& - \\
-    $\sigma$ & uncertainty parameter & -& - \\
-    $c_a$ & CO$_2$ concentration & $\mu$ mol CO$_2$ mol$^{-1}$ air& - \\
-    $\lambda$ & marginal water cost of leaf carbon & mol H$_2$O mol$^{-1}$ CO$_2$ & - \\
-    $\Gamma$ & CO$_2$ compensation point & - & - \\
-    $\Gamma^*$ & CO$_2$ compensation point without dark respiration & - & - \\
-    \hline
+  \label{definitions} \centering \small
+\begin{tabular}{l c c c} \hline Variable & Description & Units &
+Citation \\ \hline $e_s$ & saturation vapor pressure & Pa & - \\ $T$ &
+temperature & K & - \\ $P$ & pressure & Pa & - \\ $\Delta$ &
+$\frac{\partial e_s}{\partial T}$ & Pa K$^{-1}$ & - \\ $R_{net}$ & net
+radiation at land surface minus ground heat flux & W m$^{-2}$ & - \\
+$g_a$ & aerodynamic conductance & m s$^{-1}$ &
+\citet{Shuttleworth_2012} \\ $\rho_a$ & air density & kg m$^{-3}$ & -
+\\ $c_p$ & specific heat capacity of air at constant pressure & J
+K$^{-1}$ kg$^{-1}$ & - \\ $VPD$ & vapor pressure deficit & Pa & - \\
+$\gamma$ & psychometric constant & Pa K$^{-1}$ & - \\ $g_s$ & stomatal
+conductance & m s$^{-1}$ & \citet{MEDLYN_2011} \\ $g_{l-s}$ &
+leaf-scale stomatal conductance & mol m$^{-2}$ s$^{-1}$ &
+\citet{MEDLYN_2011} \\ $R$ & universal gas constant & J mol$^{-1}$
+K$^{-1}$ & - \\ $R_{air}$ & gas constant of air & J K$^{-1}$ kg$^{-1}$
+& - \\ $LAI$ & leaf area index & -& - \\ $\sigma$ & uncertainty
+parameter & -& - \\ $c_a$ & CO$_2$ concentration & $\mu$ mol CO$_2$
+mol$^{-1}$ air& - \\ $\lambda$ & marginal water cost of leaf carbon &
+mol H$_2$O mol$^{-1}$ CO$_2$ & - \\ $\Gamma$ & CO$_2$ compensation
+point & - & - \\ $\Gamma^*$ & CO$_2$ compensation point without dark
+respiration & - & - \\ \hline
 \end{tabular}
@@
 by combining an optimal photosynthesis theory \citep{Farquhar_1980,
-  Katul_2010} with an empirical approach, which describes the
-dependence of $g_s$ to VPD. This resulted in the following model for
-leaf-scale stomatal conductance:
+Katul_2010} with an empirical approach, which describes the dependence
+of $g_s$ to VPD. This resulted in the following model for leaf-scale
+stomatal conductance:
 
 \begin{linenomath*}
-  \begin{equation}
-    g_{s-leaf} = g_0 + 1.6 \left(1 + \frac{g_{1-leaf}}{\sqrt{VPD}}\right) \frac{A}{c_a},
-  \end{equation}
-\end{linenomath*}
-where $g_{1-leaf}$ is a leaf-scale ``slope'' parameter, A is the net
-CO$_2$ assimilation rate, and $c_a$ is the atmospheric CO$_2$
-concentration at the leaf surface. \cite{MEDLYN_2011} relate the slope
-parameter ($g_{1-leaf}$) to physical parameters as:
+  \begin{equation} g_{l-s} = g_0 + 1.6 \left(1 +
+\frac{g_1}{\sqrt{VPD}}\right) \frac{A}{c_a}
+  \end{equation}
+\end{linenomath*} Where $g_1$ is a leaf-scale ``slope'' parameter, A
+is the net CO$_2$ assimilation rate, and $c_a$ is the atmospheric
+CO$_2$ concentration at the leaf surface. \cite{MEDLYN_2011} relate
+the slope parameter ($g_1$) to physical parameters as:
 \begin{linenomath*}
   \label{slope}
-  \begin{equation}
-    g_{1-leaf} = \sqrt{\frac{3 \, \Gamma^* \, \lambda}{1.6}},\footnote{Note this expression has units of of (mmol mol$^{-1}$)$^{1/2}$, but this can be converted to Pa$^{1/2}$ using the ideal gas law.}
+  \begin{equation} g_1 = \sqrt{\frac{3 \, \Gamma^* \,
+\lambda}{1.6}},\footnote{Note this expression has units of of (mmol
+mol$^{-1}$)$^{1/2}$, but this can be converted to Pa$^{1/2}$ using the
+ideal gas law.}
   \end{equation}
@@
 (without dark respiration), and $\lambda$ is the marginal water cost
-of leaf carbon
-($\frac{\partial \; \text{transpiration}}{\partial \; A}$). So,
-$g_{1-leaf}$ is a leaf-scale term reflecting the tradeoff of water for
-carbon uptake. The higher $g_{1-leaf}$, the more open the stomata and
-the more they release water in exchange for carbon.
+of leaf carbon ($\frac{\partial \; \text{transpiration}}{\partial \;
+A}$). So, $g_1$ is a leaf-scale term reflecting the tradeoff of water
+for carbon uptake and growth. The higher $g_1$, the more open the
+stomata and the more they release water in exchange for carbon.
 
 The Medlyn model for stomatal conductance has been shown to behave
-very well across PFTs \citep[][]{Lin_2015}, and has been successfully
-adopted to ecosystem scale analysis in \citet{Medlyn_2017}. In units
-of m s$^{-1}$, the ecosystem scale stomatal conductance is:
+very well across PFTs \citep[][]{Lin_2015}, and can be adapted to the
+ecosystem scale by multiplying $g_{l-s}$ by leaf area index (LAI) and
+converting units to m s$^{-1}$ with the ideal gas law:
 
 \begin{linenomath*}
-  \begin{equation}
-    g_s = \frac{R \,T}{P} \; 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{GPP}{c_a},
+  \begin{equation} g_s = \text{LAI} \frac{R \,T}{P} \left( g_0 + 1.6
+\left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{A}{c_a}\right)
       \label{medlyn}
@@
 
-where GPP is the ecoysystem scale gross primary production, and $g_1$
-is an ecosystem scale analogue to $g_{1-leaf}$. We solve for $g_1$
-following \citet{Medlyn_2017} (Equation 5), and take the median $g_1$
-value to be representative of each PFT (Table \ref{pft}).
-
-\begin{table}
-  \caption{Plant functional types, their abbreviation, calculated
-    Medlyn coefficient, and calculated uWUE. uWUE from
-    \citet{Zhou_2015}, including the observed standard deviation is
-    shown for comparison. Note that uWUE from \citet{Zhou_2015} is
-    calculated from a different set of sites, and that units are
-    converted such that the quantities work with Equations 1-8 and the
-    variables defined Table \ref{definitions}.}
-  \small
-  \label{pft}
-  \centering
-  \begin{tabular}{l c c @{\qquad} c c}
-    \hline
-    \multirow{2}[3]{*}{Abbreviation} & \multirow{2}[3]{*}{PFT} & \multirow{2}[3]{*}{$g_1$ (Pa$^{0.5}$)} & \multicolumn{2}{c}{uWUE ($\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET])}  \\
-    \cmidrule{4-5}
-
-                                     & & & fitted & \citet{Zhou_2015} \\
-
-    \hline
-    \input{pft_params}
-    \hline
-  \end{tabular}
-\end{table}
-
 While Equation 3 can be used in PM (Equation 1), it will make
-analytical work with the function intractable because $GPP$ is
-functionally related to ET itself. Additionally, a perturbation to VPD
-should induce a physiological plant response that will alter GPP and
-cause an indirect change in stomatal conductance, in addition to the
-direct effect of VPD in Equation \ref{medlyn}. Therefore, in order to
-derive the response of ET to VPD, we must account for the functional
-relationship between GPP, ET, and VPD, and its effect on stomatal
-conductance. We can use aforementioned semi-empirical results of
-\citet{Zhou_2015} as a tool to approach this
-problem. \citet{Zhou_2015}, showed that the underlying Water Use
-Efficiency (uWUE):
+analytical work with the function intractable because $A$, net CO$_2$
+assimilation rate, is functionally related to ET itself. To remove the
+dependence of ET on $A$ we use fundamental semi-empirical results of
+\citet{Zhou_2015}, who showed that the underlying Water Use Efficiency
+(uWUE):
 
 \begin{linenomath*}
-  \begin{equation}
-    uWUE = \frac{GPP \cdot \sqrt{VPD}}{ET}
+  \begin{equation} uWUE = \frac{GPP \cdot \sqrt{VPD}}{ET}
     \label{uwue}
   \end{equation}
+\end{linenomath*} is relatively constant across time and moisture
+conditions within a plant functional type. If, following
+\citet{Lin_2015}, we approximate $g_0$ as $0$ (i.e. we neglect
+cuticular and epidermal losses - a reasonable assumption except in
+very dry conditions), we can use uWUE to remove the $A$ dependence
+from $g_s$ in a way that makes the PM equation analytically tractable:
+
+\begin{linenomath*}
+  \begin{equation} g_s = \frac{R \, T}{P} 1.6 \left(1 +
+\frac{g_1}{\sqrt{VPD}}\right) \frac{uWUE \; ET}{c_a \; \sqrt{VPD}}.
+    \label{new_g_s}
+  \end{equation}
 \end{linenomath*}
 
-is relatively constant across time and moisture conditions within a
-plant functional type, and correctly captures a constant relationship
-between GPP, ET and VPD over a diurnal cycle \citep{Zhou_2014}. We can
-use uWUE to remove the $GPP$ dependence from $g_s$ in a way that makes
-the PM equation analytically tractable:
+Note that uWUE is fit on the ecosystem scale as in \citet{Zhou_2015},
+so GPP in Equation \ref{uwue} is really $A \cdot \text{LAI}$. This
+leads to the cancellation of LAI in the step from Equation
+\ref{medlyn} to Equation \ref{new_g_s}. Plugging Equation
+\ref{new_g_s} into Equation \ref{orig_pen} and rearranging gives a new
+explicit expression for PM, in which dependence on $A$ is removed:
 
 \begin{linenomath*}
-  \begin{equation}
-    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{uWUE \; ET}{c_a \; \sqrt{VPD}}.
-    \label{new_g_s}
+  \begin{equation} ET = \frac{\Delta R_{net} + \frac{g_a\; P}{T}
+\left( \frac{ c_p VPD}{R_{air}} - \frac{\gamma c_a \sqrt{VPD} }{ R \;
+1.6\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta +
+\gamma}
+    \label{et}
   \end{equation}
@@
 
-Plugging Equation \ref{new_g_s} into Equation \ref{orig_pen} and
-rearranging gives a new explicit expression for PM, in which
-dependence on $GPP$ is removed:
-
-\begin{linenomath*}
-  \begin{equation}
-    ET = \frac{\Delta R_{net} + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R \; 1.6\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
-    \label{et}
-  \end{equation}
-\end{linenomath*}
-
-With Equation \ref{et} we have derived for the first time, using
-recent results \citep[][]{MEDLYN_2011, Zhou_2014, Zhou_2015,
-  Medlyn_2017}, ET explicitly as function of environmental variables
-and two plant-specific constants, the slope parameter ($g_1$), and
-uWUE, both reflecting water conservation strategy. The slope parameter
-is related to the willingness of stomata to trade water for CO$_2$ and
-to keep stomata open. uWUE is a semi-empirical ecosystem-scale
-constant related to how WUE changes with VPD (specifically
-$VPD^{-1/2}$). It is also roughly proportional to physical constants:
+With Equation \ref{et}, ET is explicitly a function of environmental
+variables and two plant-specific constants, the slope parameter
+($g_1$), and uWUE, both reflecting water conservation strategy. The
+slope parameter is a leaf-scale term related to the willingness of
+stomata to trade water for CO$_2$ and to keep stomata open. uWUE is a
+semi-empirical ecosystem-scale constant related to how WUE changes
+with VPD (specifically $VPD^{-1/2}$). Is is also roughly proportional
+to physical constants:
 
@@
 term in our new version of PM (Equation \ref{et}) is observed, except
-for uWUE, which we fit by calculating its expectation, given the model
-and FLUXNET2015 data.
+for the two parameters $g_1$, and uWUE. $g_1$ has been measured and
+reported for the PFTs considered here \citep{Lin_2015, Franks_2017},
+and we fit uWUE by calculating its expectation, given the model and
+FLUXNET2015 data.
 
@@
 error, as well as model uncertainty (e.g. temporal and spatial
-variations of uWUE and $g_1$), we introduce an uncertainty parameter
-$\sigma$ modifying uWUE:
+variations of uWUE), we introduce an uncertainty parameter $\sigma$
+modifying uWUE:
 
 \begin{linenomath*}
-  \begin{equation}
-    ET = \frac{\Delta R_{net} + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R \; 1.6\; \sigma \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
+  \begin{equation} ET = \frac{\Delta R_{net} + \frac{g_a\; P}{T}
+\left( \frac{ c_p VPD}{R_{air}} - \frac{\gamma c_a \sqrt{VPD} }{ R \;
+1.6\; \sigma \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{
+\Delta + \gamma}
     \label{et_sigma}
@@
 \begin{linenomath*}
-  \begin{equation}
-    \sigma = - \frac{g_a \gamma c_a \sqrt{VPD} L_v P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R_{net} - g_a \rho_a c_p VPD\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})},
+  \begin{equation} \sigma = - \frac{g_a \gamma c_a \sqrt{VPD} L_v P }{
+\left(\text{ ET } ( \Delta + \gamma) - \Delta R_{net} - g_a \rho_a c_p
+VPD\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})},
     \label{sigma}
@@
 \begin{linenomath*}
-  \begin{equation}
-    \frac{\partial \;  ET}{\partial \; VPD} = \frac{2\; g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_a }{1.6 \; R\; \sigma \; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)
+  \begin{equation} \frac{\partial \; ET}{\partial \; VPD} = \frac{2\;
+g_a \; P}{T(\Delta + \gamma)} \left(\frac{ c_p}{R_{air}} -
+\frac{\gamma c_a }{1.6 \; R\; \sigma \; \text{ uWUE }} \left( \frac{2
+g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)
     \label{d_et}
@@
 response to atmospheric demand perturbations, with an implicit
-assumption of non-extreme drought conditions associated with our
-constant $uWUE$ assumption (e.g. we might expect $uWUE$ to decrease in
-extreme drought). There are a few subtleties to taking the derivative
-in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$) and $VPD$ are
-functionally related, so while taking the derivative we evaluate
-$\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET}
-{\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD}
-\Big|_{\text{RH fixed}} + \frac{\partial \; ET}{\partial \; RH}
-\frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$
-    fixed}}$. $RH$ and $e_s$ are assumed to be approximately
-independent, which is supported by the data (not shown).
+assumption of non-extreme drought conditions. There are a few
+subtleties to taking the derivative in Equation \ref{d_et}: $\Delta$
+($\frac{d e_{s}}{d T}$) and $VPD$ are functionally related, so while
+taking the derivative we evaluate $\frac{\partial \; ET}{\partial \;
+VPD} = \frac{\partial \; ET} {\partial \; e_s} \frac{\partial \;
+e_s}{\partial \; VPD} \Big|_{\text{RH fixed}} + \frac{\partial \;
+ET}{\partial \; RH} \frac{\partial \; RH}{\partial \; VPD}
+\Big|_{\text{$e_s$ fixed}}$. $RH$ and $e_s$ are assumed to be
+approximately independent, which is supported by the data (not shown).
 
@@
 using commonly available weather station data and the constants
-published here. This makes Equation \ref{et} a useful alternative to
-PET in drought indices and hydrometeorological analysis for vegetated
-surfaces. Equation \ref{et} better reflects the physics of water
-exchange at the land surface and would only require fitting of uWUE
-and $g_1$, and can also account for changes in CO$_2$ concentration
-(assuming some relationship between $\lambda$ and [CO$_2$]), contrary
-to typical drought indices such as the Palmer Drought Severity Index
-(PDSI) \citep{Swann_2016}.
+published in \citet{Zhou_2015} and \citet{Lin_2015}. This makes
+Equation \ref{et} a useful alternative to PET in drought indices and
+hydrometeorological analysis for vegetated surfaces. Equation \ref{et}
+better reflects the physics of water exchange at the land surface and
+would only require fitting of uWUE and $g_1$, and can also account for
+changes in CO$_2$ concentration, contrary to typical drought indices
+such as the Palmer Drought Severity Index (PDSI) \citep{Swann_2016}.
 
@@
 
-\begin{figure}
-  \centering \includegraphics[width=\textwidth]{./joint_vpd_sigma.pdf}
+\begin{figure} \centering
+\includegraphics[width=\textwidth]{./joint_vpd_sigma.pdf}
 \caption{The joint distribution of $VPD$ and $\sigma$, with outliers
 removed (defined as lowest and highest 5\% of $\sigma$). $\sigma$
-    exhibits a weak dependence on $VPD$, and $\sigma$ is $O(1)$ for
-    the bulk of the observations.}
+exhibits a weak dependence on $VPD$, and $\sigma$ is $O(1)$ for the
+bulk of the observations.}
 \label{joint_vpd_sigma}
@@
 
-\begin{equation}
-  \frac{g_a \; P}{T(\Delta + \gamma)},
+\begin{equation} \frac{g_a \; P}{T(\Delta + \gamma)},
 \end{equation}
@@
 \begin{equation}
-  \label{sign}
-  \frac{c_p}{R_{air}} - \frac{\gamma c_a }{1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right).
-\end{equation}
-All variables are positive, so the relative magnitude between the
-first term and the second term in the sign term (Equation \ref{sign})
-will determine whether ET increases or decreases with increasing
-VPD. If the second term is larger then plant control dominates and ET
-decreases with increasing VPD. However, if the first term is larger,
-then atmospheric demand dominates and ET increases with increasing
-VPD.
+  \label{sign} \frac{c_p}{R_{air}} - \frac{\gamma c_a }{1.6 \; R\;
+\text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 +
+\sqrt{VPD})^2}\right).
+\end{equation} All variables are positive, so the relative magnitude
+between the first term and the second term in the sign term (Equation
+\ref{sign}) will determine whether ET increases or decreases with
+increasing VPD. If the second term is larger then plant control
+dominates and ET decreases with increasing VPD. However, if the first
+term is larger, then atmospheric demand dominates and ET increases
+with increasing VPD.
 
 \subsection{Functional Form of the Sign Term}
-\label{sign_term}
-First, we explore the variables within the sign term to gain better
-intuition on the driver of either the increase or reduction of ET with
-VPD. CO$_2$ concentration ($c_a$) and the psychometric constant
-($\gamma$) are relatively constant over the dataset considered here so
-that the variability is dominated by $\sigma$ and $VPD$. uWUE could
-vary with soil moisture but has been shown to be relatively constant
-\citep{Zhou_2015}. This then means that the sign term only depends on
-VPD for a given PFT and is approximately just a function of $VPD$. We
-can further determine a critical threshold separating an increase from
-a decrease in ET, i.e. the threshold $VPD_{crit}$ such that the
-derivative vanishes $\frac{\partial \; ET}{\partial \; VPD} = 0$:
-\small
+\label{sign_term} First, we explore the variables within the sign term
+to gain better intuition on the driver of either the increase or
+reduction of ET with VPD. CO$_2$ concentration ($c_a$) and the
+psychometric constant ($\gamma$) are relatively constant over the
+dataset considered here so that the variability is dominated by
+$\sigma$ and $VPD$. uWUE could vary with soil moisture but has been
+shown to be relatively constant \citep{Zhou_2015}. This then means
+that the sign term only depends on VPD for a given PFT and is
+approximately just a function of $VPD$. We can further determine a
+critical threshold separating an increase from a decrease in ET,
+i.e. the threshold $VPD_{crit}$ such that the derivative vanishes
+$\frac{\partial \; ET}{\partial \; VPD} = 0$: \small
 \begin{linenomath*}
-  \begin{equation}
-    VPD_{crit} = \frac{R_{air}}{4 c_p} \left( \frac{\gamma c_a}{1.6\; R \;  uWUE} + \sqrt{\frac{\gamma c_a}{1.6\; R \;  uWUE}\left( \frac{\gamma c_a}{1.6\; R \;  uWUE} + 8 g_1 \frac{c_p}{R_{air}}\right)} - 4 g_1 \frac{c_p}{R_{air}} \right),
+  \begin{equation} VPD_{crit} = \frac{R_{air}}{4 c_p} \left(
+\frac{\gamma c_a}{1.6\; R \; uWUE} + \sqrt{\frac{\gamma c_a}{1.6\; R
+\; uWUE}\left( \frac{\gamma c_a}{1.6\; R \; uWUE} + 8 g_1
+\frac{c_p}{R_{air}}\right)} - 4 g_1 \frac{c_p}{R_{air}} \right),
 \label{vpd_min_et}
   \end{equation}
-\end{linenomath*}
-\normalsize noting that $VPD_{crit}$ mostly depends on the PFT
-parameters uWUE and $g_1$, and only varies weakly with climate as most
-other parameters related to the environment are nearly constant. The
-calculated value of $VPD_{crit}$ for each PFT is shown in Table
-\ref{vpd_crit}. For any values of $VPD$ less than $VPD_{crit}$, ET
-will decrease with increasing VPD
+\end{linenomath*} \normalsize noting that $VPD_{crit}$ mostly depends
+on the PFT parameters uWUE and $g_1$, and only varies weakly with
+climate as most other parameters related to the environment are nearly
+constant. The calculated value of $VPD_{crit}$ for each PFT is shown
+in Table \ref{vpd_crit}. For any values of $VPD$ less than
+$VPD_{crit}$, ET will thus decrease with increasing VPD
 ($\frac{\partial \; ET}{\partial \; VPD} < 0$), and for values of
@@
 \begin{table}
-  \caption{Values of $VPD_{crit}$, where
-    $\frac{\partial \; ET}{\partial \; VPD} = 0$, evaluated at PFT
-    average values for $R_{air}$, $\gamma$, and $c_a$. PFT-specific
-    constants ($g_1$, uWUE) are provided in Table \ref{pft}. For
-    values of $VPD$ less than $VPD_{crit}$,
-    $\frac{\partial \; ET}{\partial \; VPD}$ will be negative, and for
-    values of $VPD$ greater than $VPD_{crit}$,
-    $\frac{\partial \; ET}{\partial \; VPD}$ will be positive.}
-  \centering
-  \begin{tabular}{l c c c c c}
-    \hline
-    PFT & $R_{air}$ & $c_a$ (ppm) & $\gamma$  & \textbf{$VPD_{crit}$ (Pa)} \\
-    \hline
-    \input{vpd_crit.tex}
+\caption{Values of $VPD_{crit}$, where $\frac{\partial \; ET}{\partial
+\; VPD} = 0$, evaluated at PFT average values for $R_{air}$, $\gamma$,
+and $c_a$. PFT-specific constants ($g_1$, uWUE) are provided in Table
+\ref{pft}. For values of $VPD$ less than $VPD_{crit}$, $\frac{\partial
+\; ET}{\partial \; VPD}$ will be negative, and for values of $VPD$
+greater than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$
+will be positive.}  \centering
+\begin{tabular}{l c c c c c} \hline PFT & $R_{air}$ & $c_a$ (ppm) &
+$\gamma$ & \textbf{$VPD_{crit}$ (Pa)} \\ \hline CRO & 288.6 & 376.0 &
+65.2 & \textbf{936.6} \\ CSH & 289.0 & 383.6 & 67.5 & \textbf{12852.5}
+\\ DBF & 288.6 & 379.5 & 63.6 & \textbf{1266.6} \\ EBF & 288.4 & 368.5
+& 61.6 & \textbf{1454.8} \\ ENF & 288.28 & 379.8 & 61.0 &
+\textbf{1853.7} \\ GRA & 288.4 & 378.0 & 61.0 & \textbf{3261.1} \\
 \hline
@@
 ecosystem-scale $WUE$ in response to increases in VPD. This decrease
-in $WUE$ causes more water loss per unit carbon gain, and explains the
-relationship between high uWUE and high likelihood of increases of ET
-in response to increasing atmospheric drying (increases in VPD).
+in $WUE$ causes more water loss per unit carbon production, and
+explains the relationship between high uWUE and high likelihood of
+increases of ET in response to increasing atmospheric drying
+(increases in VPD).
 
@@
 comparing two extreme PFTs: water intensive crops (CRO) and water
-conservative closed shrub (CSH). CRO has higher slope parameter and a
-slightyly higher uWUE ($g_1 = 140.7$ Pa$^{1/2}$; $2.85$ $\mu$-mol [C]
-Pa$^{0.5}$ J$^{-1}$ [ET]) compared to CSH ($g1 = 75.1 \, Pa^{1/2}$,
-$uWUE=2.82$ $\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET]). These differences
-in PFT parameters cause opposite ET responses to changes in VPD
-between CRO and CSH. ET theoretically always decreases with increasing
-VPD for the more water conservative CSH, while ET frequently increases
-with increasing VPD for the more water intensive CRO (Figure
+conservative closed shrub (CSH). CRO has high slope parameters and
+uWUE ($g_1 = 183.1$ Pa$^{1/2}$; $2.27$ $\mu$-mol [C] Pa$^{0.5}$
+J$^{-1}$ [ET]) compared to CSH ($g1 = 148.6 \, Pa^{1/2}$, $uWUE=1.66$
+$\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET]). These differences in PFT
+parameters cause opposite ET responses to changes in VPD between CRO
+and CSH. ET theoretically always decreases with increasing VPD for the
+more water conservative CSH, while ET frequently increases with
+increasing VPD for the more water intensive CRO (Figure
 \ref{idealized_sign}). CRO evolved or were bred to prioritize GPP and
@@
 
-\begin{figure}
-  \centering \includegraphics[width=\textwidth]{./idealized_sign.pdf}
+\begin{figure} \centering
+\includegraphics[width=\textwidth]{./idealized_sign.pdf}
 \caption{The functional form of the sign term, with $\sigma$ held
 fixed at 1, and all terms except VPD set to PFT averages. For
-    comparison, the observed range of VPD for each PFT is plotted
-    below the x-axis. Stars denote 25th, 50th, and 75th percentiles,
-    and the range of the line spans the 5th-95th percentiles of
-    observed VPD. Vertical black lines denote the location of
-    $VPD_{crit}$ for each PFT, with the exception of CSH, for which
-    $VPD_{crit}$ is off-scale.}
+comparison, the observed range of VPD for each PFT is plotted below
+the x-axis. Stars denote 25th, 50th, and 75th percentiles, and the
+range of the line spans the 5th-95th percentiles of observed
+VPD. Vertical black lines denote the location of $VPD_{crit}$ for each
+PFT, with the exception of CSH, for which $VPD_{crit}$ is off-scale.}
 \label{idealized_sign}
@@
 at reducing ET, especially at low VPD. However, at very high VPD,
-gradients in vapor pressure at the leaf scale will become very strong
-as stomata reach their limits of closure in response to VPD
-(parameterized with $g_1$). As a result, ET response will begin to
-asymptote towards a constant ecosystem-scale values as leaf-scale
-response to VPD asymptotes towards zero.  Therefore, plants with a low
-$g_1$ will have the largest VPD dependence of ET response because the
-difference in ET response at low VPD (leaf stomatal response
-dominates) and high VPD (VPD gradient dominates) is largest. This is
-apparent in the strong VPD dependence of CSH, which has the lowest
-slope parameter ($g_1=75.1$ Pa$^{1/2}$) (Figure \ref{idealized_sign}).
+gradients in vapor pressure at the leaf scale will be so strong that
+they will begin to dominate leaf-scale plant strategies of stomatal
+(partial) closure (parameterized with $g_1$). As a result, ET response
+will begin to asymptote towards its ecosystem-scale values
+irregardless of leaf-scale strategy and stomatal regulation
+(parameterized with $g_1$). Therefore, plants with a low $g_1$ will
+have the largest VPD dependence of ET response because the difference
+in ET response at low VPD (leaf stomatal response dominates) and high
+VPD (VPD gradient dominates) is largest. This is apparent in the
+strong VPD dependence of ENF, which has the lowest slope parameter
+($g_1=74.31$ Pa$^{1/2}$) (Figure \ref{idealized_sign}).
 
@@
 while CSH are the most water conservative and have the strongest
-tendency to decrease ET with increasing VPD, as well as the strongest
-VPD dependence of response. Figure \ref{idealized_sign} clearly shows,
-according to our theory, that for all PFTs except for crops there is
-frequent occurrence of a negative (plant dominating) ET response to
-increases in VPD. Therefore, plants are able in most atmospheric
-conditions to reduce ET in response to increased VPD and thus to
-reduce water loss. To better illustrate this, the ranges of observed
-environmental VPDs at the FLUXNET sites are plotted parallel to the
-x-axis. For CSH, VPD is always less than VPD$_{crit}$ (off scale) so
-that the plant response dominates in typical environmental conditions,
-emphasizing the water conservative strategy of those plants. For CRO
-on the other hand, VPD is higher than VPD$_{crit}$ for more than 50\%
-of observations, emphasizing that those plants operate with an
-aggressive water usage strategy, are water intensive and were actually
-engineered for photosynthesis rather than water saving. For DBF, EBF,
-MF, GRA, and SAV more than half of the observed VPD are less than
-VPD$_{crit}$, i.e. in conditions where plant response dominates. SAV
-has a more water conservative response than the forest, grass, and
-crop plan types, but still respondes by increasing ET with increasing
-VPD for about a quarter of observations, due to the high aridity (VPD)
-of the SAV ecoclimate. It is also important to note that for all PFTs,
-even when atmospheric demand dominates, ET response to VPD is still
-far more negative than it would be for potential evaporation
-$\partial PET/\partial VPD$, i.e. atmospheric demand only, emphasizing
-that there is still a strong regulation of evaporative flux by stomata
-and though the plant xylem. The sign term in the PET case would just
-be a constant ($\frac{c_p}{R_{air}} \approx 3.5$), which is far larger
-than any part of the curves for any PFT. This highlights the
-deficiencies of PET's ability to capture land response to changes in
-atmospheric dryness (VPD). Plants are always regulating water exchange
-from the land surface, even when they reach the limits of they ability
-to do so.
+tendency to decrease ET with increasing VPD.  ENF ($g_1 = 74.31$
+Pa$^{1/2}$) has by far the largest VPD response dependence, while CRO
+($g_1 = 183.1$ Pa$^{1/2}$) has the smallest VPD dependence. Figure
+\ref{idealized_sign} clearly shows, according to our theory, that for
+all PFTs except for crops there is frequent occurrence of a negative
+(plant dominating) ET response to increases in VPD. Therefore, plants
+are able in most atmospheric conditions to reduce ET in response to
+increased VPD and thus to reduce water loss. To better illustrate
+this, the ranges of observed environmental VPDs at the FLUXNET sites
+are plotted parallel to the x-axis. For CSH, VPD is always less than
+VPD$_{crit}$ (off scale) so that the plant response dominates in
+typical environmental conditions, emphasizing the water conservative
+strategy of those plants. For CRO on the other hand, VPD is higher
+than VPD$_{crit}$ for more than 50\% of observations, emphasizing that
+those plants operate with an aggressive water usage strategy, are
+water intensive and were actually engineered for photosynthesis rather
+than water saving. For forests and grass, more than half of the
+observed VPD are less than VPD$_{crit}$, i.e. in conditions where
+plant response dominates. It is also important to note that even when
+atmospheric demand dominates, ET response to VPD is still far more
+negative than it would be for potential evaporation $\partial
+PET/\partial VPD$, i.e. atmospheric demand only, emphasizing that
+there is still a strong regulation of evaporative flux by stomata and
+though the plant xylem. The sign term in the PET case would just be a
+constant ($\frac{c_p}{R_{air}} \approx 3.5$), which is far larger than
+any part of the curves for any PFT. This highlights the deficiencies
+of PET's ability to capture land response to changes in atmospheric
+dryness (VPD). Plants are always regulating water exchange from the
+land surface, even when they reach the limits of they ability to do
+so.
 
 \subsection{Functional Form of the Scaling Term}
-\label{scale_term}
-While the above discussion of the sign of
+\label{scale_term} While the above discussion of the sign of
 $\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our
@@
 
-\begin{equation}
-  \frac{g_a \; P}{T(\Delta + \gamma)}:
+\begin{equation} \frac{g_a \; P}{T(\Delta + \gamma)}:
 \end{equation}
@@
 observed temperature ranges. The scaling term for the tree PFTs (DBF,
-EBF, ENF, MF) is generally about double the scaling terms for other
+EBF, ENF) is generally about double the scaling terms for the other
 PFTs which have lower surface roughness and generally smaller
-aerodynamic conductance (GRA, CSH, CRO). The savannah (WSA, SAV) PFT's
-scaling is somewhere betwen GRA, CSH, and CRO, and DBF, EBF, ENF, and
-MF, due to higher variability and surface roughness.
-
-\begin{figure}
-  % \centering
-  \centerline{\includegraphics[width=1.4\textwidth]{./idealized_scale.pdf}}
+aerodynamic conductance.
+
+\begin{figure} \centering
+\includegraphics[width=\textwidth]{./idealized_scale.pdf}
 \caption{Primary sources of variability for the scaling term, as a
@@
 \caption{Statistics of $\frac{\partial \; ET}{\partial \; VPD}$ as a
-    function of PFT.}
-  \centering
-  \begin{tabular}{l c c c}
-    \hline
-    PFT & $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ & $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)$ & fraction $\frac{\partial \; ET}{\partial \; VPD} < 0.$ \\
-    \hline
-    \input{stats.tex}
-    \hline
+function of PFT.}  \centering
+\begin{tabular}{l c c c} \hline PFT & $\overline{\frac{\partial \;
+ET}{\partial \; VPD}}$ & $\frac{\partial \; ET}{\partial \;
+VPD}\left(\overline{env}\right)$ & fraction $\frac{\partial \;
+ET}{\partial \; VPD} < 0.$ \\ \hline CRO & -0.05 & 0.008 & 0.533 \\
+DBF & -0.10 & -0.013 & 0.609 \\ EBF & -0.13 & -0.037 & 0.666 \\ GRA &
+-0.07 & -0.024 & 0.694 \\ ENF & -0.17 & -0.081 & 0.740 \\ CSH & -0.29
+& -0.224 & 0.973 \\ \hline
 \end{tabular}
@@
 controlled, water conservative response) for the majority of
-observations. The more water conservative PFTs generally exhibit
-higher frequency of negative ET response, espcially when one factors
-in the distribution of environmental VPD (e.g. SAV and WSA grow in
-more arid climates, Figure \ref{idealized_sign}). In general the bulk
-statistics match our theoretical expectations well, with the caveat
-that inclusion of uncertainty shifts crops towards a slightly more
-negative response, and shifts many of the other PFTs, which still
-exhibit a high frequency of negative ET reponse, towards more frequent
-occurence of postive ET response than the theory and Figure
+observations. When listed by increasing frequency of negative
+response, the PFTs are generally ordered as we would expect them to
+be, from most water intensive to most water conservative: CRO (53\%),
+DBF (61\%), EBF (67\%), GRA (69\%), ENF (74\%), and CSH (97\%).
+
+In general the bulk statistics match our theoretical expectations
+well, with the caveat that inclusion of uncertainty shifts crops
+towards a slightly more negative response than the theory and Figure
 \ref{idealized_sign} would suggest. The bulk statistics motivate a
@@
 \subsection{Validation of theory at eddy-covariance sites}
-\label{testing}
-We now compare more sophisticated distributions of the observed
-response to our simplified theory (Section \ref{sign_term}). The
-observed distribution of the sign term, as compared to what the theory
-would predict, is provided in Figure \ref{test_sign}. Our goal was to
-capture the leading order behavior of the ET dependence on VPD. Given
-the assumptions we have made, and the uncertainties of flux tower
-observations themselves, we expect a relatively large amount of noise
-when reproducing the derivatives of ET. However, the data largely
-reproduces our theoretical analysis.
-
-\begin{figure}
-  \centering
-  \centerline{\includegraphics[width=1.4\textwidth]{./test_sign.pdf}}
+\label{testing} We now compare more sophisticated distributions of the
+observed response to our simplified theory (Section
+\ref{sign_term}). The observed distribution of the sign term, as
+compared to what the theory would predict, is provided in Figure
+\ref{test_sign}. Our goal was to capture the leading order behavior of
+the ET dependence on VPD. Given the assumptions we have made, and the
+uncertainties of flux tower observations themselves, we expect a
+relatively large amount of noise when reproducing the derivatives of
+ET. However, the data largely reproduces our theoretical analysis.
+
+\begin{figure} \centering
+\includegraphics[width=\textwidth]{./test_sign.pdf}
 \caption{Comparison of the sign term with model uncertainty included
 (box plots) to the sign term as calculated with simplifying
-    assumptions (blue line, as in Figure \ref{idealized_sign}). Each
-    box plot corresponds to 5\% of the data, and the 5-95\% range of
-    VPD is plotted.}
+assumptions (blue line, as in Figure \ref{idealized_sign}). Each box
+plot corresponds to 5\% of the data, and the 5-95\% range of VPD is
+plotted.}
 \label{test_sign}
@@
 
-This is particularly true for DBF and MF; the theory matches the
-leading order behavior of the function when uncertainty is included,
-and the observations match the theory with the addition of noise. The
-VPD dependence, given by the slope parameter ($g_1$), follows the
-median values of each bin. Perhaps most importantly, the x-intercept,
-and thus VPD$_{crit}$, matches nearly exactly between the theory and
-the observations. Therefore the sign of the ET response to increases
-in VPD should be well matched, subject to the unavoidable constraints
-of noise, much of which comes from the observations themselves. The
-uncertainty is non negligible; there are many observations in each bin
-for which the the sign of observation is opposite the response
-predicted by the theory, but to leading order our theory matches the
-observations well.
+For DBF our theory is successful towards this goal: the theory matches
+the leading order behavior of the function when uncertainty is
+included, and the observations match the theory with the addition of
+noise. The VPD dependence, given by the slope parameter ($g_1$),
+follows the median values of each bin. Perhaps most importantly, the
+x-intercept, and thus VPD$_{crit}$, matches nearly exactly between the
+theory and the observations. Therefore the sign of the ET response to
+increases in VPD should be well matched, subject to the unavoidable
+constraints of noise, much of which comes from the observations
+themselves. The uncertainty is non negligible; there are many
+observations in each bin for which the the sign of observation is
+opposite the response predicted by the theory, but to leading order
+our theory matches the observations well.
 
 While CSH has a much different functional form of the sign term than
-DBF and MF, CSH observations also match our theory to leading order,
-albiet with a bit more variability as a function of VPD. Again, the
-VPD dependence mostly determined from the slope parameter ($g_1$)
-closely matches the medians in the observation bins. The
-VPD-independent, strongly negative response is also captured. For CSH,
-there is rare occurrence of observed positive ET response with VPD
-($\approx 6$\%), even with uncertainty, so the sign of the
-observations almost always matches the sign of the theory, which
-states that ET response should always be negative.
-
-Biases between the theory and observations are similiar for ENF, WSA,
-CRO and SAV. At low VPD the theory and the observations match
-well. However, at high VPD (upper 10-20\% of observations) the theory
-is biased slightly towards postive reponse as compared to the
-observations (Figure \ref{test_sign}). For ENF this slight bias could
-be explained by a negative bias in $g_1$. However, for WSA, CRO, and
-SAV the observations in the highest VPD bins exhibit a downturn
-towards more negative response, which cannot be captured by the
-functional form of the sign term (Figure
-\ref{idealized_sign}). Therefore, to explain the observations one of
-the varibles in the sign term must change at extreme VPD (upper
-10-20\%). The most likely candidiate is $uWUE$, which we have assumed
-constant to meet our goal of developing intuition for leading order
-behavior, but might be expected to decrease at extremely low SWC,
-which should be correlated with high VPD. The theory for EBF exhibits
-similiar limiatations as for ENF, WSA, and SAV, except a larger
-portion of observations are biased negative as compared to the theory
-($\approx$ 35\%, Figure \ref{test_sign}). However, in general, and
-specifically for non-extreme VPD, the theory matches the observations
-for the tree and savannah plant types.
-
-The theory for GRA suffers from similar, but much more severe,
-limitations as for CRO, WSA, and SAV. GRA observations are
-characterized by a consistent trend back towards negative ET response
-at higher VPD, which the functional form of our theory is incapable of
-accounting for. As compared to CRO, WSA and SAV, the divergence
-between the theory and the observations is far greater for GRA,
-biasing 40-50\% of observations at higher VPD. In addition for the
-potential for soil moisture to alter $uWUE$, there are other sources
-of plant heterogeneity specific to GRA (and some extent CRO) that may
-alter $uWUE$ (or $g_1$) or invalidate other assumptions made in the
-Methods section (Section \ref{methods}). We do not account for
-variability in plant height and surface roughness, or differences in
-C3 vs C4 photosynthesis and water strategies, which we might expect to
-vary substantially across sites, years, and season for GRA. These
-deficiencies could largely explain the inability of our theory to
-exactly match the observations in croplands. For example, a
-superposition of sites with C3 plants (at low environmental VPD) and
-C4 plants (at high environmental VPD) would explain observed shift
-back towards negative ET response at high VPD when all sites are
-binned together, as in Figure \ref{test_sign}. We hypothesize that the
-theory would validate against observations much better if these
-sources of varaibility were accounted for, at a cost of increased
-complexity and analytical opacity.
+DBF, CSH observations also match our theory to leading order. Again,
+the VPD dependence mostly determined from the slope parameter ($g_1$)
+closely matches the medians in the observation bins and the
+theoretical threshold is well reproduced (albeit off-scale). The
+VPD-independent, strongly negative response, which is determined by
+uWUE, is also captured. For CSH, there is extremely rare occurrence of
+observed positive ET response with VPD ($\approx 3$\%), even with
+uncertainty, so the sign of the observations almost always matches the
+sign of the theory, which states that ET response should always be
+negative.
+
+As Table \ref{stats} suggested, the theory does not match as well for
+CRO as for DBF and CSH. At low VPD the VPD dependence (due to $g_1$)
+is similar to the leading order behavior of the observations, but the
+theory's curve is shifted towards more positive values relative to the
+observations, likely due to some bias in uWUE. Most strikingly, the
+observations show a tendency to move back towards negative values at
+high VPD, which given the functional form of VPD dependence, is
+behavior that is impossible for the theory to capture. This general
+positive bias of the theory explains the shift towards negative ET
+response observed in the bulk statistics in Table \ref{stats}. The
+divergence between the theory and the observations for CRO could have
+been expected. For practical reasons, our theory does not include
+important sources of plant type heterogeneity within and across sites
+for CRO. In agricultural landscape, the footprint of an
+eddy-covariance towers tend to not only observe a single species but
+many different ones, with potentially varying water use strategies. We
+also do not account for seasonal changes in LAI, variability in crop
+height and surface roughness, differences in C3 vs C4 photosynthesis
+and water strategies, or differences in irrigation practices between
+the sites and years. These deficiencies should largely explain the
+inability of our theory to exactly match the observations in
+croplands. For example, a superposition of sites with C3 plants (at
+low environmental VPD) and C4 plants (at high environmental VPD) would
+explain observed shift back towards negative ET response at high VPD
+when all sites are binned together, as in Figure \ref{test_sign}.
+
+The theory for GRA suffers from similar limitations as for CRO. GRA
+observations are characterized by a consistent trend back towards
+negative ET response at higher VPD, which the functional form of our
+theory is incapable of accounting for. As compared to CRO, the
+divergence between the theory and the observations is even greater for
+GRA, but these differences are explainable using similar logic as we
+did for CRO. GRA also tends to exhibit substantial within-PFT
+variability, such as C3 vs C4 photosynthesis, as well as strong
+seasonal variability in height and LAI. As with CRO, the trend back
+towards negative ET response at higher VPD could be explained by a
+superposition of sites with C3 vs C4 plants. Unfortunately, because
+detailed site-level data of C3 vs C4 plant type is not available, we
+could not bin CRO and GRA further by irrigation and photosynthesis
+strategy. However, we hypothesize that the theory would validate
+against observations much better when photosynthesis pathways are
+accounted for.
+
+The comparison between the theory and observations for ENF and EBF is
+unique from the rest of the PFTs. The VPD dependence of the theory,
+determined by the slope parameter ($g_1$), is not as well captured by
+observations. However, the theory still captures the general behavior
+and magnitude of the response, with some increase in error at low
+and/or high VPD.  We note that we used published constants for $g_1$
+\citep[from ][]{Lin_2015, Franks_2017}, which determines the VPD
+dependence, and that deviations should be expected across sites
+\citep{Lin_2015}. We could have fit our own constants so that the
+theory would best match the observations, but our goal was to use a
+priori results whenever possible in order to avoid over-fitting and to
+have a theoretical approach as independent to the observations as
+possible for further confirmation of the validity of the approach.
 
 While the above discussion shows that our theory has some limitations
-when applied to some PFTs, especially for grasslands, it does well for
-DBF, MF and CSH PFTs, and captures the response at non-extreme VPD for
-ENF, EBF, CRO, WSA, and SAV. In general, the leading order behavior
-observed in the data is captured by the theory. Departures between our
-theory and observations, specifically at high VPD, could be explained
-and conceptualized with shifts in $g_1$ and/or $uWUE$ due to
-site-spcific plant type variability (e.g. more arid-adapted ecosystems
-at more arid sites), or temporal variability for some enviornmental
-conditions (e.g. decreases in $uWUE$ at extremely low soil
-moisture). To focus on general behaviro and develop intuition for
-PFT-scale response, we have ignored these sources of variability in
-the present analysis. However, any site-specific or temporal plant
-functional variability that can be conceptualized with shifts to $g_1$
-and $uWUE$ can be analyzed with our framework. This opens a door to a
-myriad future analysis in which plant behavior in anomalous conditions
-can be explained and analyzed using Equation \ref{et}.
-
-% It is useful to also compare our theory, the observed uncertainty in
-% the ET response, and PET, as PET is often used as a proxy for ET in
-% hydrometeorological studies and drought indices. Again, as in our
-% discussion of the theoretical results, the PET sensitivity to VPD
-% will be constant at about 3.5. In Figure \ref{test_sign} the ET
-% response to shifts in dryness is always far smaller than 3.5 in the
-% observations, even with all uncertainty included. So the PET
-% response to atmospheric dryness is outside the range of all
-% observable responses. This casts doubt on PET's usefulness for
-% capturing the leading order behavior of water budget and dryness
-% response at the land surface, as is done in several drought
-% indices. However our theory (Equation \ref{et}) could be used with
-% published uWUE and $g1$ data, as an alternative to PET to estimating
-% land response and dryness, and would more accurately account for the
-% response to atmospheric drying.
+when applied to some PFTs, especially for crops and grasslands, it
+does well for DBF and CSH PFTs, and captures the response at
+non-extreme VPD for ENF and EBF. It is useful to also compare our
+theory, the observed uncertainty in the ET response, and PET, as PET
+is often used as a proxy for ET in hydrometeorological studies and
+drought indices. Again, as in our discussion of the theoretical
+results, the PET sensitivity to VPD will be constant at about 3.5. In
+Figure \ref{test_sign} the ET response to shifts in dryness is always
+far smaller than 3.5 in the observations, even with all uncertainty
+included. So the PET response to atmospheric dryness is outside the
+range of all observable responses. This casts doubt on PET's
+usefulness for capturing the leading order behavior of water budget
+and dryness response at the land surface, as is done in several
+drought indices. However our theory (Equation \ref{et}) could be used
+with published uWUE and $g1$ data, as an alternative to PET to
+estimating land response and dryness, and would more accurately
+account for the response to atmospheric drying.
 
@@
 
-\begin{figure}
-  \centering
-  \centerline{\includegraphics[width=1.4\textwidth]{./data_scatter.png}}
-  \caption{Scatter plots of observed
-    $\frac{\partial \; ET}{\partial \; VPD}$, including $\sigma$
-    variability, as a function of PFT, temperature and VPD. Please
-    note differences in the colorbar scale.}
+\begin{figure} \centering
+\centerline{\includegraphics[width=\textwidth]{./data_scatter.png}}
+\caption{Scatter plots of observed $\frac{\partial \; ET}{\partial \;
+VPD}$, including $\sigma$ variability, as a function of PFT,
+temperature and VPD. Please note differences in the colorbar scale.}
 \label{data_scatter}
@@
 the most water conservative, with a strong negative response. DBF,
-EBF, ENF, MF, SAV and WSA are also water conservative, but show some
+EBF, ENF, and GRA are also water conservative, but show some
 occurrence of positive ET response to VPD, particularly at higher
-observed VPD, as the theory predicts. GRA, while generally water
-conservative, does not match the theory, with increasing frequency of
-\textit{negative} response at high VPD. This is as expected, given
-discussion in Section \ref{testing}.
+observed VPD, as the theory predicts.
 
@@
 
-Indeed, for all PFTs, there is some slight dependence of
-$\sigma \cdot uWUE$ on SWC, especially at low SWC (Figure
-\ref{swc_boxplot}). The portion of observations for which our theory
-is biased by SWC-limitations varies by PFT, due to the nonlinear
-threshold at which soil moisture availability limits plant
-function. For CRO, MF and DBF, soil water content only matters for
-about the lowest 5\% of observations (each box is 5\% of observations
-in Figure \ref{swc_boxplot}). And, CRO, relative to DBF, has a very
-weak dependence on soil moisture, which is reflective of the high
+Indeed, for all PFTs, there is some slight dependence of $\sigma \cdot
+uWUE$ on SWC, especially at low SWC (Figure \ref{swc_boxplot}). The
+portion of observations for which our theory is biased by
+SWC-limitations varies by PFT, due to the nonlinear threshold at which
+soil moisture availability limits plant function. For CRO and DBF,
+soil water content only matters for about the lowest 5\% of
+observations (each box is 5\% of observations in Figure
+\ref{swc_boxplot}). And, CRO, relative to DBF, has a very weak
+dependence on soil moisture, which is reflective of the high
 likelihood that CRO sites are optimally irrigated and not water
-stressed, suggesting that observed depatures from the theory for CRO
-are due to other factors (i.e. photosythesis pathway) than soil
-moisture. For ENF, EBF, SAV, and WSA, systematic SWC-induced biases in
-$\sigma \cdot uWUE$ emerge in about the lowest 20-30\% of SWC
-conditions, although large variability in the SWC-$\sigma \cdot uWUE$
-relationship hampers interpretation for WSA. In crontast to ENF, CRO,
-and DBF, for GRA the relationship between $\sigma \cdot uWUE$ and SWC
-is more linear and affects a greater portion of observations; about
-60\% of observations. Clearly, for GRA soil water frequently impacts
-plant function and alters ET response, and our theory is limited for
-the majority of encironmental condiitons. It is therefore not
-suprising that our theory tested poorly againt the data for GRA,
-relative to to the other PFTs (Section \ref{testing}). For all other
-PFTs occurence of soil mosture impacts was rare enough to not manifest
-itself in bulk statistics and figures. A possible exception is CSH,
-for which there are only two sites, which precludes comment on the
-exact PFT-wide relationship.
-
-\begin{figure}
-  \centering
-  \centerline{\includegraphics[width=1.4\textwidth]{./swc_boxplot.pdf}}
+stressed. For ENF and EBF, systematic SWC-induced biases in $\sigma
+\cdot uWUE$ emerge in about the lowest 30\% of SWC conditions. In
+crontast to ENF, CRO, and DBF, for GRA the relationship between
+$\sigma \cdot uWUE$ and SWC is more linear and affects a greater
+portion of observations; about 60\% of observations. Clearly, for GRA
+soil water frequently impacts plant function and alters ET response,
+and our theory is limited. It is therefore not suprising that our
+theory tested poorly againt the data for GRA, relative to ENF, EBF,
+DBF, and CSH (Section \ref{testing}). For all other PFTs occurence of
+soil mosture impacts was rare enough to not manifest itself in bulk
+statistics and plots. A possible exception is CSH, for which there are
+only two sites, which precludes comment on the exact PFT-wide
+relationship.
+
+\begin{figure} \centering
+\includegraphics[width=\textwidth]{./swc_boxplot.pdf}
 \caption{$\sigma \cdot uWUE$ as a function of PFT and SWC. Each box
-    plot represents 5\% of all observations. Note that the highest
-    10\% of SWC observations were excluded from the WSA and SAV
-    figures, to better resolve high variability in narrow low-SWC bins
-    for those PFTs.}
+plot represents 5\% of all observations.}
 \label{swc_boxplot}
@@
 VPD, and the deviation between our theory (based on a constant uWUE
-assumption) and observations. It is also perhaps not coincidental that
-the protion of $\sigma \cdot uWUE$ affected by low soil moisture
-observations is similiar to the portion of observations that to not
-match our theory at high VPD.
+assumption) and observations.
 
@@
 the observations. Soil moisture effects explain the deficiencies of
-our theory in Section \ref{testing}, particularly for GRA. By
-conceptualizing SWC effects as a change in $uWUE$ (and/or $g_1$), it
-will be possible for future analysis to explore the importance of soil
-moisture on plant response to VPD.
-
-\subsection{Functional form of ET dependence on VPD and its relation
-  to VPD dependence}
-
-Th theory described in Section \ref{sign_term} indicates that for a
-given $uWUE$ and $g_1$, the ET dependence on VPD should be concave
-upward. In other words, there should be some local minima in ET at a
-critical VPD$_{crit}$, assuming the scaling term (e.g. aerodynamic
-conductance and $\Delta$) is held fixed. This result warrants futher
-investigation, because to our knowledge no one has derived the
-theoretical ecosystem-scale relationship between ET and VPD while
-controlling for enivironmental conditiions. In particular, from
-personal communication, there is an apparent lack of consensus over
-whether the shape of the ET-VPD curve should be concave upward (our
-result) or concave downward. Given that understanding the ET-VPD
-relationship of the one-way plant response is fundamental to
-hypothesizing about any feedbacks between the land surface and the
-atmopshere, we analyze why our derived ET-VPD relationship is concave
-upward, particularly with respect to the exponent of VPD depedence in
-$uWUE$ and the Medlyn unified stomatal conductace model.
-
-There is a theoretical basis for the square root VPD dependence in
-both the stomatal conducatance model and $uWUE$ based on the
-assumption that stomata behave to maximize carbon gain while
-minimizing water loss, which observations also generally support
-\citep{Lloyd_1991, MEDLYN_2011, Lin_2015, Zhou_2014, Zhou_2015,
-  Medlyn_2017}. However, purely empiriacle results which fit the
-exponent of the VPD dependence have shown that it may vary from 1/2,
-suggesting that stomata, as well as ecosystem-scale quantities based
-on stomata theroy, may not always function optimally \citep{Zhou_2015,
-  Lin_2018}. Specifically with regards to $uWUE$, one would not expect
-that this ecosystem scale WUE quantity will responde to VPD exactly
-analgalously to stomata. Direct soil evaporation's contributions to ET
-should shift the exponential dependence of VPD to be greater than
-1/2. In other words, direct soil evaporation causes WUE to decrease
-more strongly with increasing VPD, especially in well-watered
-conditions. \citet{Zhou_2015}'s results corroborate this: they found a
-mean imprically fit exponential VPD dependence of 0.55 istead of 0.5
-for Ameriflux sites. Given that these recent results on the
-relatioship between VPD, GPP, and ET form the backbone of our analysis
-and are what allowed us to derive an explicit expression for ET for
-the first time (Equation \ref{et}), we will analyze if and how
-assumptions about the exonent of the VPD dependence impacts the shape
-of the ET-VPD dependence. By introducing $n$ and $m$ we can free our
-stomatal conductance model from assumptions about VPD dependence:
-\begin{linenomath*}
-  \begin{equation}
-    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{VPD^m}\right) \frac{*WUE \; ET}{c_a \; VPD^n},
-    \label{m_n}
-  \end{equation}
-\end{linenomath*}
-where:
-\[*WUE = \frac{GPP}{ET}VPD^n.\] To determine how the exponent $n$ and
-$m$ alter the shape of the ET-VPD dependence we find the roots of the
-second deriviative of ET, using Equation \ref{m_n} for stomatal
-conductance, with respect to VPD: \input{d2_solutions.tex}
+our theory in Section \ref{testing}, particularly for GRA.
 
@@
 regulate ET, and this regulation varies by PFT. CROs are the least
-water conservative, while DBF, EBF, SAV, GRA, MF, WSA, ENF and CSH are
-progressively more water conservative (more likely to reduce ET in
-response to increasing VPD). SAV and WSA exhibit positive ET response
-to VPD not necessarily because of poor water conservation stragegies
-relative to other PFTs, but because of greater occurence of high
-atmospheric demand (VPD) relative to other PFTs. Observations of ET
-response to VPD exhibit the same general behavior as the theory, with
-ET response becoming more positive (atmospheric demand dominating) as
-environmental VPD increases within a PFT, and more negative for PFTs
-that are adapted to arid conditions and prioritize water conservation
-over primary production.
+water conservative, while DBF, EBF, ENF, GRA and CSH are progressively
+more water conservative (more likely to reduce ET in response to
+increasing VPD). Observations of ET response to VPD exhibit the same
+general behavior as the theory, with ET response becoming more
+positive (atmospheric demand dominating) as environmental VPD
+increases within a PFT, and more negative for PFTs that are adapted to
+arid conditions and prioritize water conservation over primary
+production.
 
@@
 Lastly, our paper builds important intuition for how plants will
-respond to VPD pertubations, including expected increases in VPD
-induced by global change.  Plant physiological responses to direct
-CO$_2$ effects \citep[e.g.,][]{Swann_2016} receives more attention
-than physiological response to indirect effects like increased
-VPD. Here, we provide broad PFT-focused results showing a likely
-decrease in ET in response to positive VPD perturbations (atmospheric
-drying), which is consistent with recent observational analysis
-\citep[e.g.,][]{Rigden_2017}. Feedbacks between the land and the
-atmosphere may alter the net response to a global VPD perturabation,
-but our focus on the one way plant response to a VPD perturbation in
-the atmopsheric boundary layer is an important first step to
-disentangling such feedbacks, both in observations and model
-simulations of the present and future. By removing Penman Monteith's
-dependence on implicit relationships between GPP, VPD, and ET, we
-allow for explicit, accuruate future analysis of plant-VPD feedbacks
-in the atmospheric boundary layer (Equation \ref{et}). Our approach
-can also be extended to examine varying plant response to more nuanced
-consideration of plant type and climate. Any plant physiological
-heterogeneity that can be conceptualized with shifts in $g_1$
-\citep[e.g.][]{Lin_2015, Medlyn_2017} and/or uWUE
-\citep[e.g.][]{Zhou_2014} is representable within our framework, which
-opens the door for a hierarchy of more sophisticated climate- and
-plant-specific analysis of ET sensitivity to environmental variables
-(including VPD). We argue that such simplified conceptual frameworks
-are critical tools for disentangling land-atmosphere feedbacks at
-various scales, from diurnal to seasonal and beyond, and to
-characterize ET response in a warmer, atmospherically drier, and
+respond to global change-induced shifts in VPD, which should increase
+in the future. Plant physiological responses to direct CO$_2$ effects
+\citep[e.g.,][]{Swann_2016} receives more attention than physiological
+response to indirect effects like increased VPD. Here, we provide
+broad PFT-focused results showing a likely decrease in ET in response
+to positive VPD perturbations (atmospheric drying). Additionally, we
+derive an explicit conceptual framework for ET response as a function
+of two plant type parameters, $g_1$ and uWUE (e.g. Equation
+\ref{et}). This framework can be used to understand ET response to a
+range of environmental change, and forms a foundation for resolving
+varying response to more nuanced consideration of plant type and
+climate. Any plant physiological heterogeneity that can be
+conceptualized with shifts in $g_1$ \citep[e.g.][]{Lin_2015} and/or
+uWUE \citep[e.g.][]{Zhou_2014} is representable within our framework,
+which opens the door for a hierarchy of more sophisticated climate-
+and plant-specific analysis of ET sensitivity to environmental
+variables (including VPD). We argue that such simplified conceptual
+frameworks are critical tools for disentangling land-atmosphere
+feedbacks at various scales, from diurnal to seasonal and beyond, and
+to characterize ET response in a warmer, atmospherically drier, and
 enriched CO$_2$ world.

### 2018-04-06 doc/paper/vpd_et_paper.tex
done with pierre's comments up to section 3.4

@@
   water, and vapor pressure deficit is expected to rise with
-  increasing CO$_2$. While increased evapotranspiration (ET) in
+  increasing greenhouse gases. While increased evapotranspiration (ET) in
   response to increased atmospheric demand seems intuitive, plants are
@@
 guarantee an increased likelihood of precipitation, which depending on
-environmental conditions could cause a decrease in the likelihood
-of precipitation \citep[][]{Gentine_2013}. Furthermore, any increases
-in precipitation are likely to be non-local, such that plants giving
-up water to the atmosphere are not guaranteed to reap the benefits of
-water returned from the atmosphere to the soil. Using this logic, from
-an ecosystem evolutionary perspective, water stored in soil is worth
-much more than the chance of water returned as precipitation.
+environmental conditions could cause a decrease in the likelihood of
+precipitation \citep[][]{Gentine_2013}. Furthermore, any increases in
+precipitation are likely to be non-local, such that plants giving up
+water to the atmosphere are not guaranteed to reap the benefits of
+water returned from the atmosphere to the soil. Using this subjective
+logic, from an ecosystem evolutionary perspective, water stored in
+soil seems to be worth much more than the chance of water returned as
+precipitation.
 
@@
 Recent results have drastically improved our understanding of VPD's
-impact on physiology. \citet{MEDLYN_2011} developed a model for
-leaf-scale stomatal conductance ($g_s$), including VPD response, by
-combining an optimal photosynthesis theory \citep{Farquhar_1980,
-  Katul_2010} with an empirical approach, and extended use of this
-model to the ecosystem scale in \citet{Medlyn_2017}. Additionally,
-\citet{Zhou_2014} demonstrated that a quantity underlying water use
-efficiency $\left(uWUE = \frac{GPP\; \sqrt{VPD}}{ET}\right)$ properly
-captures a constant relationship between GPP, ET, and VPD over a
-diurnal cycle. uWUE is also remarkably well conserved in the growing
-season across space and time, within a PFT \citep{Zhou_2015}. These
-results provide novel tools to examine how ET varies in response to
-VPD, while explicitly including VPD's effect on stomatal conductance
-for the first time. In this manuscript, we leverage these tools to
-derive the theoretical response of ET to VPD using a Penman Monteith
-framework. Our theory is validated and tested at multiple
-eddy-covariance stations spanning various climates and plant
-functional types.
+impact on physiology, especially at the leaf
+level. \citet{MEDLYN_2011} developed a model for leaf-scale stomatal
+conductance ($g_s$), including VPD response, by combining an optimal
+photosynthesis theory \citep{Farquhar_1980, Katul_2010} with an
+empirical approach, and extended use of this model to the ecosystem
+scale in \citet{Medlyn_2017}. Additionally, \citet{Zhou_2014}
+demonstrated that a quantity underlying water use efficiency
+$\left(uWUE = \frac{GPP\; \sqrt{VPD}}{ET}\right)$ properly captures a
+constant relationship between GPP, ET, and VPD over a diurnal cycle at
+the ecosystem scale. uWUE is also remarkably well conserved in the
+growing season across space and time, within a PFT
+\citep{Zhou_2015}. These results provide novel tools to examine how ET
+varies in response to VPD, while explicitly including VPD's effect on
+stomatal conductance for the first time. In this manuscript, we
+leverage these tools to derive the theoretical one-way response of ET
+to VPD with other environmental variables properly controlled for,
+i.e. we take the partial derivative of ET with respect to VPD. Our
+theory is validated and tested at multiple eddy-covariance stations
+spanning various climates and plant functional types.
 
@@
 following \citet{Medlyn_2017} (Equation 5), and take the median $g_1$
-value to be representative of each PFT (Table \ref{pft}).
+value to be representative of each PFT instead of the mean to avoid extra weighting of rare outliers (Table \ref{pft}).
 
@@
 plant functional type, and correctly captures a constant relationship
-between GPP, ET and VPD over a diurnal cycle \citep{Zhou_2014}. We can
-use uWUE to remove the $GPP$ dependence from $g_s$ in a way that makes
-the PM equation analytically tractable:
+between GPP, ET and VPD over a diurnal cycle \citep{Zhou_2014}. The
+theoretical derivation of the square root VPD dependence in $uWUE$
+leverages the same asumptions used in \cite{MEDLYN_2011} to derive the
+square-root VPD dependence of the stomatal conductace model (Equation
+\ref{medlyn}).  We can use uWUE to remove the $GPP$ dependence from
+$g_s$ in a way that makes the PM equation analytically tractable:
 
@@
 
-With Equation \ref{et} we have derived for the first time, using
-recent results \citep[][]{MEDLYN_2011, Zhou_2014, Zhou_2015,
-  Medlyn_2017}, ET explicitly as function of environmental variables
-and two plant-specific constants, the slope parameter ($g_1$), and
-uWUE, both reflecting water conservation strategy. The slope parameter
-is related to the willingness of stomata to trade water for CO$_2$ and
-to keep stomata open. uWUE is a semi-empirical ecosystem-scale
-constant related to how WUE changes with VPD (specifically
-$VPD^{-1/2}$). It is also roughly proportional to physical constants:
+By accounting for photosythesis changes in ecosystem conductace, with
+Equation \ref{et} we have derived for the first time, using recent
+results \citep[][]{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017}, ET
+explicitly as function of environmental variables and two
+plant-specific constants, the slope parameter ($g_1$), and uWUE, both
+reflecting water conservation strategy. The slope parameter is related
+to the willingness of stomata to trade water for CO$_2$ and to keep
+stomata open. uWUE is a semi-empirical ecosystem-scale constant
+related to how WUE changes with VPD (specifically $VPD^{-1/2}$). It is
+also roughly proportional to physical constants:
 
@@
 This derivation relied either implicitly or explicitly on several
-assumptions. First, the underlying theory used to derive Equations
-\ref{leaf_medlyn} and \ref{medlyn} rely on an implicit assumption that
-$ET=g_s \cdot VPD$; physically this implies that leaves are perfectly
+assumptions. First, we assume that VPD at the leaf surface is the same as VPD at measurement height; physically this implies that leaves are perfectly
 coupled to the atmosphere. In reality, for some conditions and plant
@@
 which is usually characterized by relatively high insolation inducing
-instability and convective boundary layers, we would
-expect the surface to be generally well coupled. Additionally, by following
-\citet{Medlyn_2017}'s method for calculating ecosystem scale $g_1$
-using PM,
-we have stronger confidence that surface coupling is represented
-within our framework, even if it is not represented in the underlying
-theory used to develop stomatal conductance ($g_s$). An additional
-assumption in the formulation of $uWUE$
-\citep{Zhou_2014, Zhou_2015} and
-\citet{Medlyn_2017}'s stomatal conductance model is that
-direct soil evaporation (E) contributions to ET remain
-small relative to transpiration (T). This should be more true during the growing season. However, some plant types allow for systematically larger
-contributions of ``E'' in ET, particularly those with less dense
-canopies and smaller absolute amounts of transpiration. We therefore
-might expect that the theory will be most applicable to forest PFTs,
-which will be most strongly coupled to the boundary layer due to
-larger surface roughness, and will also generally have the highest
-ratios of transpiration to evaporation. Finally, for the goal of
-developing PFT-wide intuition, we assume that $g_1$ and $uWUE$ are
-constant in space and time. Both quantities have been shown to be well
-conserved over space and time within a PFT relative to inter-PFT
-variability. However, we would expect them to vary somewhat with
-environmental conditions including soil water content, temperature, and
-with intra-PFT plant-specific characteristics like wood density for
-tree PFTs \citep{Lin_2015}. Given these strong assumptions made with
-the goal of understanding broad, leading order plant behaviors, we
-take the extra care of including and analyzing uncertainty (Section
-\ref{testing}) and examining sensitivity to the exponent of the VPD
-response, which is derived using assumptions that $ET=g_s \cdot VPD$ (Section
+instability and convective boundary layers, we would expect the
+surface to be generally well coupled. An
+additional assumption in the formulation of $uWUE$ \citep{Zhou_2014,
+  Zhou_2015} and \citet{Medlyn_2017}'s stomatal conductance model is
+that direct soil evaporation (E) contributions to ET remain small
+relative to transpiration (T). This should be more true during the
+growing season. The ratio of E to T may increase immediately after
+rainfall events due to high soil moisture and ponding, but VPD is
+generally low anyways during these times. However, some plant types
+allow for systematically larger contributions of E in ET,
+particularly those with sparse canopies and smaller relative amounts
+of transpiration. We therefore might expect that the theory will be
+most applicable to forest PFTs, which will be most strongly coupled to
+the boundary layer due to larger surface roughness, and will also
+generally have the highest ratios of transpiration to
+evaporation. Finally, for the goal of developing PFT-wide intuition,
+we assume that $g_1$ and $uWUE$ are constant in space and time. Both
+quantities have been shown to be well conserved over space and time
+within a PFT relative to inter-PFT variability. However, we would
+expect them to vary somewhat with environmental conditions including very low
+soil water content, temperature, and with intra-PFT plant-specific
+characteristics like wood density for tree PFTs
+\citep{Lin_2015}. Given these strong assumptions made with the goal of
+understanding broad, leading order plant behaviors, we take the extra
+care of including and analyzing uncertainty (Section \ref{testing})
+and examining sensitivity to the exponent of the VPD response (Section
 \ref{functional_form}).
@@
 to typical drought indices such as the Palmer Drought Severity Index
-(PDSI) \citep{Swann_2016}. Lastly, all code and data used in this
-analysis, including those used to generate the figures and tables, are
-publicly available at https://github.com/massma/climate\_et.
+(PDSI) \citep{Swann_2016, Lemordant_2016, Lemordant_2018}. Lastly, all
+code and data used in this analysis, including those used to generate
+the figures and tables, are publicly available at
+https://github.com/massma/climate\_et.
 
@@
 that inclusion of uncertainty shifts crops towards a slightly more
-negative response, and shifts many of the other PFTs, which still
-exhibit a high frequency of negative ET response, towards more frequent
+negative response to VPD, and shifts many of the other PFTs, which still
+exhibit a high frequency of negative ET response to VPD, towards more frequent
 occurrence of positive ET response than the theory and Figure

### 2018-04-06 doc/paper/vpd_et_paper.tex
pierre's comments done

@@
 constant to meet our goal of developing intuition for leading order
-behavior, but might be expected to decrease at extremely low SWC,
-which should be correlated with high VPD. The theory for EBF exhibits
-similar limitations as for ENF, WSA, and SAV, except a larger
-portion of observations are biased negative as compared to the theory
-($\approx$ 35\%, Figure \ref{test_sign}). However, in general, and
-specifically for non-extreme VPD (VPD < $\approx$70-90th percentiles), the theory matches the observations
-for the tree and savannah plant types.
+behavior, but might be expected to decrease at extremely low SWC, and should be correlated with high VPD in very dry environments
+through sensible heating increases \citep{Gentine_2016}. The theory
+for EBF exhibits similar limitations as for ENF, WSA, and SAV, except
+a larger portion of observations are biased negative as compared to
+the theory ($\approx$ 35\%, Figure \ref{test_sign}). However, in
+general, and specifically for non-extreme VPD (VPD < $\approx$70-90th
+percentiles), the theory matches the observations for the tree and
+savannah plant types well.
 
@@
 functional variability that can be conceptualized with shifts to $g_1$
-and $uWUE$ can be analyzed with our framework. This opens a door to a
-myriad future analysis in which plant behavior in anomalous conditions
+and $uWUE$ can be analyzed with our framework. This opens a door to future analyses in which plant behavior in anomalous conditions
 can be explained and analyzed using Equation \ref{et}.
@@
 observed VPD, as the theory predicts. GRA, while generally water
-conservative, does not match the theory, with increasing frequency of
+conservative, does not match the theory well, with increasing frequency of
 \textit{negative} response at high VPD. This is as expected, given
-discussion in Section \ref{testing}.
+previous discussion on how we might expect more inter-site and inter-year variability for GRA (Section \ref{testing}).
 
@@
 stressed, suggesting that observed departures from the theory for CRO
-are due to other factors (i.e. photosynthesis pathway) than soil
+are due to other factors (i.e. different photosynthesis pathways; C3/C4) than soil
 moisture. For ENF, EBF, SAV, and WSA, systematic SWC-induced biases in
@@
 the portion of $\sigma \cdot uWUE$ affected by low soil moisture
-observations is similar to the portion of observations that to not
+observations is similar to the portion of observations that do not
 match our theory at high VPD. Indeed, this may suggest that for all PFTs
@@
 given $uWUE$ and $g_1$, the ET dependence on VPD should be concave
-upward. In other words, there should be some local minima in ET at a
+upward, which is confirmed by eddy covariance data cross most PFTs. In other words, there should be some local minimum in ET at a
 critical VPD$_{crit}$, assuming the scaling and plant terms  (e.g. aerodynamic
@@
 theoretical ecosystem-scale relationship between ET and VPD while
-controlling for environmental conditions. In particular, from
+controlling for other environmental conditions. In particular, from
 personal communication, there is an apparent lack of consensus over
 whether the shape of the ET-VPD curve should be concave upward (our
-result) or concave downward. Given that understanding the ET-VPD
+result) or concave downward in the absence of dramatic water stress. Given that understanding the ET-VPD
 relationship of the one-way plant response is fundamental to
@@
   MEDLYN_2011, Lin_2015, Zhou_2014, Zhou_2015, Medlyn_2017}. However,
-some purely empirical results which fit the exponent of the VPD
+some purely empirical results that fit the exponent of the VPD
 dependence to data have shown that it may vary slightly from 1/2,
@@
 about the exponent of the VPD dependence impacts the shape of the
-ET-VPD dependence. By introducing $n$ and $m$ we can free our stomatal
+ET-VPD dependence.
+This analysis is also important to understand whether the choice of stomatal conductance model alters the fundamental behavior of the ET-VPD relatioships, as many commonly used models  utilize a VPD exponent other than the 1/2 suggested by optimal theory \citep[e.g.
+][ which uses an exponent of 1]{Leuning_1990}.
+
+By introducing $n$ and $m$ we can free our stomatal
 conductance model from assumptions about VPD dependence:
@@
 ($g_s$, with respect to VPD: \input{d2_solutions.tex} With this
-result we've defined the family of curves separating concave up from
+result we have defined the family of curves separating concave up from
 concave down ET solutions (Figure \ref{concave}). These curves are
@@
 atmospheric demand monotonically increases with VPD, leading to the
-intuitive result that atmospheric demand dominates plant response when
+ result that atmospheric demand dominates plant response when
 atmospheric demand is high.
@@
 induced by global change.  Plant physiological responses to direct
-CO$_2$ effects \citep[e.g.,][]{Swann_2016} receives more attention
+CO$_2$ effects \citep[e.g.,][]{Swann_2016, Lemordant_2018} receives more attention
 than physiological response to indirect effects like increased

### 2018-04-09 doc/paper/vpd_et_paper.tex
now just have final readthrough, formatting check, and write letter to editor
about to go through submission for AFM

@@
 % 11/23/2015
-\documentclass[draft,linenumbers]{agujournal}
+\documentclass[draft,linenumbers]{afmjournal}
 % \draftfalse
@@
 \item ecohydrology, evapotranspiration, aridity, vapor pressure
-  deficit, Clausius-Clapeyron relation, plant physiology, stomatal
-  response
+  deficit, plant physiology, stomatal response
 \end{keypoints}
@@
 conductance ($g_s$), including VPD response, by combining an optimal
-photosynthesis theory \citep{Farquhar_1980, Katul_2010} with an
+photosynthesis theory \citep{Cowan_1977} with an
 empirical approach, and extended use of this model to the ecosystem
@@
 to VPD with other environmental variables properly controlled for,
-i.e. we take the partial derivative of ET with respect to VPD. Our
-theory is validated and tested at multiple eddy-covariance stations
-spanning various climates and plant functional types.
+i.e. we develop a framework for evaluating the partial derivative of
+ET with respect to VPD. Our theory is validated and tested at multiple
+eddy-covariance stations spanning various climates and plant
+functional types.
 
@@
                                    & \citet{MEDLYN_2011} \\
-    $g_{1}$  & ecosystem-scale slope parameter & Pa$^{0.5}$ & \citet{MEDLYN_2017} \\
+    $g_{1}$  & ecosystem-scale slope parameter & Pa$^{0.5}$ & \citet{Medlyn_2017} \\
     $R$ & universal gas constant & J mol$^{-1}$ K$^{-1}$ & - \\
@@
 \ref{medlyn}).  We can use uWUE to remove the $GPP$ dependence from
-$g_s$ in a way that makes the PM equation analytically tractable:
+$g_s$ in a way that makes PM analytically tractable:
 
@@
 This derivation relied either implicitly or explicitly on several
-assumptions. First, we assume that VPD at the leaf surface is the same as VPD at measurement height; physically this implies that leaves are perfectly
-coupled to the atmosphere. In reality, for some conditions and plant
-types the leaves can become decoupled from the boundary layer
-\citep{De_2017, Medlyn_2017}. Given our focus on the growing season,
-which is usually characterized by relatively high insolation inducing
-instability and convective boundary layers, we would expect the
-surface to be generally well coupled. An
-additional assumption in the formulation of $uWUE$ \citep{Zhou_2014,
-  Zhou_2015} and \citet{Medlyn_2017}'s stomatal conductance model is
-that direct soil evaporation (E) contributions to ET remain small
-relative to transpiration (T). This should be more true during the
-growing season. The ratio of E to T may increase immediately after
-rainfall events due to high soil moisture and ponding, but VPD is
-generally low anyways during these times. However, some plant types
-allow for systematically larger contributions of E in ET,
-particularly those with sparse canopies and smaller relative amounts
-of transpiration. We therefore might expect that the theory will be
-most applicable to forest PFTs, which will be most strongly coupled to
-the boundary layer due to larger surface roughness, and will also
-generally have the highest ratios of transpiration to
-evaporation. Finally, for the goal of developing PFT-wide intuition,
-we assume that $g_1$ and $uWUE$ are constant in space and time. Both
-quantities have been shown to be well conserved over space and time
-within a PFT relative to inter-PFT variability. However, we would
-expect them to vary somewhat with environmental conditions including very low
-soil water content, temperature, and with intra-PFT plant-specific
-characteristics like wood density for tree PFTs
-\citep{Lin_2015}. Given these strong assumptions made with the goal of
-understanding broad, leading order plant behaviors, we take the extra
-care of including and analyzing uncertainty (Section \ref{testing})
-and examining sensitivity to the exponent of the VPD response (Section
-\ref{functional_form}).
+assumptions. First, we assume that VPD at the leaf surface is the same
+as VPD at measurement height; physically this implies that leaves are
+perfectly coupled to the atmosphere. In reality, for some conditions
+and plant types the leaves can become decoupled from the boundary
+layer \citep{De_2017, Medlyn_2017}. Given our focus on the growing
+season, which is usually characterized by relatively high insolation
+inducing instability and convective boundary layers, we would expect
+the surface to be generally well coupled. An additional assumption in
+the formulation of $uWUE$ \citep{Zhou_2014, Zhou_2015} and
+\citet{Medlyn_2017}'s stomatal conductance model is that direct soil
+evaporation (E) contributions to ET remain small relative to
+transpiration (T). This should be more true during the growing
+season. The ratio of E to T may increase immediately after rainfall
+events due to high soil moisture and ponding, but VPD is generally low
+anyways during these times. However, some plant types allow for
+systematically larger contributions of E in ET, particularly those
+with sparse canopies and smaller relative amounts of transpiration. We
+therefore might expect that the theory will be most applicable to
+forest PFTs, which will be most strongly coupled to the boundary layer
+due to larger surface roughness, and will also generally have the
+highest ratios of transpiration to evaporation. Finally, for the goal
+of developing PFT-wide intuition, we assume that $g_1$ and $uWUE$ are
+constant in space and time. Both quantities have been shown to be well
+conserved over space and time within a PFT relative to inter-PFT
+variability. However, we would expect them to vary somewhat with
+environmental conditions including very low soil water content,
+temperature, and with intra-PFT plant-specific characteristics like
+wood density for tree PFTs \citep{Lin_2015}. Given these strong
+assumptions made with the goal of understanding broad, leading order
+plant behaviors, we take the extra care of including and analyzing
+uncertainty (Section \ref{testing}) and examining sensitivity to the
+exponent of the VPD response (Section \ref{functional_form}).
 
@@
 note however that even though ET increases again above the critical
-threshold, $VPD_{crit}$, ET is still much lower that potential
+threshold, $VPD_{crit}$, ET is still much lower than potential
 evaporation as stomata are still strongly regulating vapor fluxes to
@@
     observed VPD. Vertical black lines denote the location of
-    $VPD_{crit}$ for each PFT, with the exception of CSH, for which
+    $VPD_{crit}$ for each PFT, with the exception of CSH and WSA, for which
     $VPD_{crit}$ is off-scale.}
@@
 environmental VPDs at the FLUXNET sites are plotted parallel to the
-x-axis. For CSH, VPD is always less than VPD$_{crit}$ (off scale) so
+x-axis. For CSH and WSA, VPD is always less than VPD$_{crit}$ (off scale) so
 that the plant response dominates in typical environmental conditions,
@@
 \begin{equation}
-  \frac{g_a \; P}{T(\Delta + \gamma)}:
+  \frac{g_a \; P}{T(\Delta + \gamma)}.
 \end{equation}
@@
 aerodynamic conductance variability generally impacts the scaling term
-more than temperature, even within PFT.
+more than temperature, even within PFTs.
 
@@
 constant to meet our goal of developing intuition for leading order
-behavior, but might be expected to decrease at extremely low SWC, and should be correlated with high VPD in very dry environments
+behavior, but might be expected to decrease at extremely low SWC. This extremely low SWC should be correlated with high VPD in very dry environments
 through sensible heating increases \citep{Gentine_2016}. The theory
@@
 
-We first note that in the data used low SWC conditions are very
-infrequent. Within our framework, any systematic bias due to the
+Within our framework, any systematic bias due to the
 failure to account for SWC's effects in very dry conditions should
@@
 $\sigma \cdot uWUE$ emerge in about the lowest 20-30\% of SWC
-conditions, although large variability in the SWC-$\sigma \cdot uWUE$
+conditions, although large variability in the SWC-$(\sigma \cdot uWUE)$
 relationship hampers interpretation for WSA. CSH presents a special
@@
 function are relatively rare.
+Future research will explore these relatively rare feedbacks in extreme conditions, which due to analytical intractablity will require more opaque numerical analysis of many more complex processes, including boundary layer growth and state and their relationship to surface layer coupling and free-tropospheric lapse rates and humidity.
 
@@
 analogously to stomata. Direct soil evaporation's contributions to ET
-should shift the exponent of the VPD dependence, especially for
+should shift the exponent of the VPD dependence, especially at
 conditions of low GPP when we would expect a systematically larger
 portion of direct soil evaporation contributions to ET, because we
-would also expect lower amounts of transpiration. \citet{Zhou_2015}'s
+would also expect lower amounts of transpiration at low GPP. \citet{Zhou_2015}'s
 results corroborate this: they found a mean empirically fit
@@
 
-\begin{enumerate}
+\begin{itemize}
   \item For optimal behavior (n, m = 1/2) the ET-VPD curve will be
@@
     ecosystem behavior should be prioritized.
-\end{enumerate}
+\end{itemize}
 While it is possible that in the future some other form of
@@
 function and fixed environmental conditions, the ET-VPD dependence is
-theoretically concave up, with ET increasing with increasing VPD as
+theoretically concave upward, with ET increasing with increasing VPD as
 VPD increases past some critical value (Table \ref{vpd_crit}). However

### 2018-04-09 doc/paper/afm_coverletter.tex
rough draft of coverletter, now need to shorten it up

\documentclass[10pt,stdletter,dateno]{newlfm}
\usepackage{kpfonts}

\widowpenalty=1000
\clubpenalty=1000

\newlfmP{headermarginskip=20pt}
\newlfmP{sigsize=50pt}
\newlfmP{dateskipafter=20pt}
\newlfmP{addrfromphone}
\newlfmP{addrfromemail}
\PhrPhone{Phone}
\PhrEmail{Email}

\addrfrom{%
    \today\\[10pt]
    Department of Earth and Environmental Engineering\\
    Columbia University\\
    500 West 120th Street, Mailcode: 4711\\
    New York, NY 10027
}
\phonefrom{206-919-1364}
\emailfrom{akm2203@columbia.edu}

\addrto{%
Dr. X. Lee\\
Editor-in-Chief\\
Agricultural and Forest Meteorology\\
Yale University
New Haven, CT}

\greetto{Dear Dr. Lee,}

\begin{document}
\begin{newlfm}

  We are pleased to submit our manuscript ``When does vapor pressure deficit drive or reduce evapotranspiration?'' for consideration for publication in the \textit{Journal of Agricultural and Forest Meteorology}. This research is particularly relevant now, given that vapor pressure deficit (VPD) is expected to increase as greenhouse gas concenration increases, but current research across ecology, climate, and meteorology exhibits a range of views on whether these increases in VPD drive increases in ET (atmospheric demand dominates response) or decreases in ET (plant response dominates).

  We show that for the majority of plant types and environmental conditions ET decreases with increasing VPD, suggesting that plant water conservation response to increasing VPD overwhelms the increase in atmospheric demand. However more water intensive ecosystems (e.g. crops) are more likely to allow increases to ET with increasing VPD than water conservative ecosystems (e.g. shrubs, savannah).

  While deriving these results, we developed a novel new analytical framework for calculating ET as a function of environmental variables. By leveraging recent results on the relatioship between ET, VPD and gross primary production [\textit{Zhou et al.}, 2014, 2015], for the first time we explicitly account for changes in photosythesis associated with changes in VPD. This approach reveals that a consequence of recently proposed optimal stomatal conductance theory [\textit{Medlyn et al.}, 2011, 2017] is that the ET-VPD curve is concave upward. That is, ET decreases with increasing VPD up until some ecosystem-specific VPD threshold, after which plant response cannot fully cancel increases in atmospheric demand (VPD) and further increases in VPD result in increases to ET. Furthermore, the fundamental shape of the ET-VPD curve could shift from concave up to concave down if one used pre-\textit{Medlyn et al.} 2011 stomatal conductance theory.

  A previous version of this manuscript was rejected from \textit{Global Change Biology}. The current version of the manuscript is greatly improved based on reviewers' comments; we applied recent results [\textit{Medlyn et al.} 2017] to extend our analysis to 3 more plant types, we added more clear explanation of our simplifying assumptions and their consequences for interpreting results, and we included a section showing how the concave up nature of the ET-VPD curve is a direct result of optimal stomatal conductance theory. This last addition was motivated by a reviewer's comment that the shape of the ET-VPD curve should be concave downward. The reviewer provided a generated figure instead of a citation to support this assertion, and we have not been able to locate a clear citation showing that the ecosystem-scale one-way plant response to a VPD perturbation, with environmenal conditions controlled for, should be concave downward. We have to assume the reviewer's figure is based on some other theory than the currently proposed state of the art optimal stomatal conductance theory. This interaction highlights for us the importance of publishing this manuscript, as experienced researchers in land-atmosphere feedbacks and coupling have apparent divergent [yet strong] opinions on the shape of the one-way ET-VPD curve, which is fundamental to disentangling much more complex feedbacks between the land surface and the atmosphere. We believe it is crucial for the community to see that the shape of the ET-VPD curve is dependent on cutting-edge stomatal conductance theory.

  Thank you very much for your time and consideration of this manuscript.

\closing{Sincerely,\\
%\fromsig{\includegraphics[scale=1]{signature.jpg}} \\
\fromname{Adam Massmann}
}

\end{newlfm}
\end{document}

### 2018-04-09 doc/paper/afm_coverletter.tex
cover letter finished

@@
 
-\newlfmP{headermarginskip=20pt}
-\newlfmP{sigsize=50pt}
-\newlfmP{dateskipafter=20pt}
+\newlfmP{headermarginskip=0pt}
+\newlfmP{topmarginskip=0pt}
+\newlfmP{unprtop=0pt}
+\newlfmP{headermarginsize=0pt}
+\newlfmP{bottommarginskip=20pt}
+
+\newlfmP{sigsize=0pt}
+\newlfmP{sigskipbefore=0pt}
+\newlfmP{sigskipafter=0pt}
+\newlfmP{dateskipbefore=0pt}
+\newlfmP{dateskipafter=0pt}
+
+% \newlfmP{greettoskipbefore=0pt}
+% \newlfmP{greettoskipafter=0pt}
+
+\newlfmP{addrtoskipbefore=18pt}
+\newlfmP{addrtoskipafter=18pt}
+
+\newlfmP{addrfromskipbefore=0pt}
+\newlfmP{addrfromskipafter=0pt}
+
+\newlfmP{closeskipbefore=0pt}
+\newlfmP{closeskipafter=0pt}
+
+\newlfmP{unprbottom=0pt}
+
 \newlfmP{addrfromphone}
@@
 
+\namefrom{Adam Massmann}
 \addrfrom{%
@@
 \greetto{Dear Dr. Lee,}
-
+\closeline{Sincerely,}
 \begin{document}
@@
   We are pleased to submit our manuscript ``When does vapor pressure deficit drive or reduce evapotranspiration?'' for consideration for publication in the \textit{Journal of Agricultural and Forest Meteorology}.
-This research is particularly relevant now, given that vapor pressure deficit (VPD) is expected to increase as greenhouse gas concenration increases, but current research across ecology, climate, and meteorology exhibits a range of views on whether these increases in VPD drive increases in ET (atmospheric demand dominates response) or decreases in ET (plant response dominates).
+This research is particularly relevant now, given that vapor pressure deficit (VPD) is expected to increase as greenhouse gas concentration increases, but current research across ecology, climate, and meteorology exhibits a range of views on whether these increases in VPD drive increases in ET (atmospheric demand dominates response) or decreases in ET (plant response dominates).
 
@@
   While deriving these results, we developed a novel new analytical framework for calculating ET as a function of environmental variables.
-By leveraging recent results on the relatioship between ET, VPD and gross primary production [\textit{Zhou et al.}, 2014, 2015], for the first time we explicitly account for changes in photosythesis associated with changes in VPD.
+By leveraging recent results on the relationship between ET, VPD and gross primary production [\textit{Zhou et al.}, 2014, 2015], for the first time we explicitly account for changes in photosynthesis associated with changes in VPD.
 This approach reveals that a consequence of recently proposed optimal stomatal conductance theory [\textit{Medlyn et al.}, 2011, 2017] is that the ET-VPD curve is concave upward.
@@
   A previous version of this manuscript was rejected from \textit{Global Change Biology}.
-The current version of the manuscript is greatly improved based on reviewers' comments; we applied recent results [\textit{Medlyn et al.} 2017] to extend our analysis to 3 more plant types, we added more clear explanation of our simplifying assumptions and their consequences for interpreting results, and we included a section showing how the concave up nature of the ET-VPD curve is a direct result of optimal stomatal conductance theory.
-This last addition was motivated by a reviewer's comment that the shape of the ET-VPD curve should be concave downward.
-The reviewer provided a generated figure instead of a citation to support this assertion, and we have not been able to locate a clear citation showing that the ecosystem-scale one-way plant response to a VPD perturbation, with environmenal conditions controlled for, should be concave downward.
-We have to assume the reviewer's figure is based on some other theory than the currently proposed state of the art optimal stomatal conductance theory.
-This interaction highlights for us the importance of publishing this manuscript, as experienced researchers in land-atmosphere feedbacks and coupling have apparent divergent [yet strong] opinions on the shape of the one-way ET-VPD curve, which is fundamental to disentangling much more complex feedbacks between the land surface and the atmosphere.
-We believe it is crucial for the community to see that the shape of the ET-VPD curve is dependent on cutting-edge stomatal conductance theory.
+The current version of the manuscript is greatly improved based on reviewers' comments; we applied recent results [\textit{Medlyn et al.} 2017] to extend our analysis to three more plant types, we added more clear explanation of our simplifying assumptions and their consequences for interpreting results, and we included a section showing how the concave up nature of the ET-VPD curve is a direct result of optimal stomatal conductance theory.
+This last addition was motivated by a reviewer's comment, supported by a generated figure rather than a citation, that the shape of the ET-VPD curve should be concave downward.
+This comment highlights for us the importance of publishing this manuscript, as experienced researchers in land-atmosphere feedbacks and coupling have apparent divergent [yet strong] opinions on the shape of the one-way ET-VPD curve, which is fundamental to disentangling much more complex feedbacks between the land surface and the atmosphere.
+We believe it is crucial for the community to see that the shape of the ET-VPD curve is dependent on use of cutting-edge stomatal conductance theory and is very much an active area of research.
 
@@
 
-\closing{Sincerely,\\
-%\fromsig{\includegraphics[scale=1]{signature.jpg}} \\
-\fromname{Adam Massmann}
-}
+% \closing{Sincerely,\\
+% %\fromsig{\includegraphics[scale=1]{signature.jpg}} \\
+% \fromname{Adam Massmann}
+% }
 

### 2018-04-09 doc/paper/afm_coverletter.tex
fixed axis albelling on DBF and improved cover letter w/ dave's comments

@@
   We are pleased to submit our manuscript ``When does vapor pressure deficit drive or reduce evapotranspiration?'' for consideration for publication in the \textit{Journal of Agricultural and Forest Meteorology}.
-This research is particularly relevant now, given that vapor pressure deficit (VPD) is expected to increase as greenhouse gas concentration increases, but current research across ecology, climate, and meteorology exhibits a range of views on whether these increases in VPD drive increases in ET (atmospheric demand dominates response) or decreases in ET (plant response dominates).
+This research is particularly relevant now, given that vapor pressure deficit (VPD) is expected to increase as greenhouse gas concentration increases, but current research across ecology, climate, and meteorology exhibits a range of views on whether these increases in VPD drive increases in ET (atmospheric demand dominates ET response) or decreases in ET (plant response dominates ET response).
 
-  We show that for the majority of plant types and environmental conditions ET decreases with increasing VPD, suggesting that plant water conservation response to increasing VPD overwhelms the increase in atmospheric demand.
+  We show that for the majority of plant types and environmental conditions, ET decreases with increasing VPD, suggesting that the plant water conservation response to increasing VPD overwhelms the increase in atmospheric demand.
 However more water intensive ecosystems (e.g.
@@
 
-  While deriving these results, we developed a novel new analytical framework for calculating ET as a function of environmental variables.
-By leveraging recent results on the relationship between ET, VPD and gross primary production [\textit{Zhou et al.}, 2014, 2015], for the first time we explicitly account for changes in photosynthesis associated with changes in VPD.
+  While deriving these results, we developed a novel analytical framework for calculating ET as a function of environmental variables.
+For the first time we explicitly account for changes in photosynthesis associated with changes in VPD by leveraging recent results on the relationship between ET, VPD and gross primary production [\textit{Zhou et al.}, 2014, 2015].
 This approach reveals that a consequence of recently proposed optimal stomatal conductance theory [\textit{Medlyn et al.}, 2011, 2017] is that the ET-VPD curve is concave upward.
@@
   A previous version of this manuscript was rejected from \textit{Global Change Biology}.
-The current version of the manuscript is greatly improved based on reviewers' comments; we applied recent results [\textit{Medlyn et al.} 2017] to extend our analysis to three more plant types, we added more clear explanation of our simplifying assumptions and their consequences for interpreting results, and we included a section showing how the concave up nature of the ET-VPD curve is a direct result of optimal stomatal conductance theory.
-This last addition was motivated by a reviewer's comment, supported by a generated figure rather than a citation, that the shape of the ET-VPD curve should be concave downward.
-This comment highlights for us the importance of publishing this manuscript, as experienced researchers in land-atmosphere feedbacks and coupling have apparent divergent [yet strong] opinions on the shape of the one-way ET-VPD curve, which is fundamental to disentangling much more complex feedbacks between the land surface and the atmosphere.
-We believe it is crucial for the community to see that the shape of the ET-VPD curve is dependent on use of cutting-edge stomatal conductance theory and is very much an active area of research.
+The current version of the manuscript is greatly improved based on reviewers' comments; we applied recent results [\textit{Medlyn et al.} 2017] to extend our analysis to three more plant types, we added additional discussion of our simplifying assumptions and their consequences for interpreting results, and we included a section showing how the concave up nature of the ET-VPD curve is a direct result of optimal stomatal conductance theory.
+This last addition was motivated by a reviewer's comment that the shape of the ET-VPD curve should be concave downward.
+This comment, which the reviewer supported via an unreferenced figure rather than a citation, highlights for us the importance of publishing this manuscript, as experienced researchers in land-atmosphere feedbacks and coupling maintain divergent [yet strong] opinions on the shape of the one-way ET-VPD curve, which is fundamental to disentangling more complex feedbacks between the land surface and the atmosphere.
+We believe it is crucial for the community to see that the nature of the ET-VPD curve is dependent on use of cutting-edge stomatal conductance theory and is very much an active area of research.
 

### 2018-04-09 src/paper_figure_code/data_scatter.py

@@ -68,7 +68,10 @@ def make_ax_plot(_ax, var, _df, meta):
   z = np.ma.masked_array(grouped.values, mask=np.isnan(grouped.values))
   color = _ax.pcolormesh(_x, _y, z, cmap=meta['cmap'],\
                          vmin=vmin, vmax=vmax)
-
+  if _df.pft.iloc[0] == 'DBF':
+    yticks = np.arange(10, 31, 5)
+    _ax.set_yticks(yticks)
+    _ax.set_yticklabels(['%2.0f' % y for y in yticks])
   custom_xlabel(_df, _ax, '%s (Pa)' % 'vpd'.upper(), fontsize=small_ax_fontsize)
   custom_ylabel(_df, _ax, 'T (C)', fontsize=small_ax_fontsize)
   _ax.set_title('%s'\

### 2018-05-14 doc/paper/vpd_et_paper.tex
rough draft of coverletter, now need to shorten it up
adjusting for arxiv submissions

@@
 % 11/23/2015
-\documentclass[draft,linenumbers]{afmjournal}
+% \documentclass[draft,linenumbers]{afmjournal}
+\documentclass{article}
 % \draftfalse
-\usepackage{makecell} \usepackage{multirow} %AM for tables
-\usepackage{booktabs} \drafttrue
-
-\journalname{Agricultural and Forest Meteorology: When does VPD drive
-  or reduce ET?}
+\usepackage{makecell}
+\usepackage{multirow} %AM for tables
+\usepackage{booktabs}
+% \drafttrue
+
+% \journalname{Agricultural and Forest Meteorology: When does VPD drive
+%   or reduce ET?}
 
@@
 
+\begin{abstract}
+  Increasing vapor pressure deficit increases atmospheric demand for  water, and vapor pressure deficit is expected to rise with  increasing greenhouse gases.
+While increased evapotranspiration (ET) in  response to increased atmospheric demand seems intuitive, plants are  capable of reducing ET in response to increased VPD by closing their  stomata, in an effort to conserve water.
+Here we examine which  effect dominates response to increasing VPD: atmospheric demand and  increases in ET, or plant physiological response (stomata closure)  and decreases in ET.
+We use Penman-Monteith, combined with  semi-empirical optimal stomatal regulation theory and underlying  water use efficiency, to develop a theoretical framework for  understanding how ET responds to increases in VPD.
+
+  The theory suggests that for most environmental conditions and plant  types, plant physiological response dominates and ET decreases with  increasing VPD.
+Plants that are evolved or bred to prioritize  primary production over water conservation (e.g.
+crops) exhibit a  higher likelihood of atmospheric demand-driven response (ET  increasing).
+However for forest, grass, savannah, and shrub plant  types, ET more frequently decreases than increases with rising  VPD.
+This work serves as an example of the utility of  our simplified framework for disentangling land-atmosphere feedbacks  at multiple scales, including the characterization of ET response in  an atmospherically drier, enriched CO$_2$ world.
+\end{abstract}
+
 \begin{keypoints}
@@
 \end{keypoints}
-
-\begin{abstract}
-  Increasing vapor pressure deficit increases atmospheric demand for
-  water, and vapor pressure deficit is expected to rise with
-  increasing greenhouse gases. While increased evapotranspiration (ET) in
-  response to increased atmospheric demand seems intuitive, plants are
-  capable of reducing ET in response to increased VPD by closing their
-  stomata, in an effort to conserve water. Here we examine which
-  effect dominates response to increasing VPD: atmospheric demand and
-  increases in ET, or plant physiological response (stomata closure)
-  and decreases in ET. We use Penman-Monteith, combined with
-  semi-empirical optimal stomatal regulation theory and underlying
-  water use efficiency, to develop a theoretical framework for
-  understanding how ET responds to increases in VPD.
-
-  The theory suggests that for most environmental conditions and plant
-  types, plant physiological response dominates and ET decreases with
-  increasing VPD. Plants that are evolved or bred to prioritize
-  primary production over water conservation (e.g. crops) exhibit a
-  higher likelihood of atmospheric demand-driven response (ET
-  increasing). However for forest, grass, savannah, and shrub plant
-  types, ET more frequently decreases than increases with rising
-  VPD. This work serves as an example of the utility of
-  our simplified framework for disentangling land-atmosphere feedbacks
-  at multiple scales, including the characterization of ET response in
-  an atmospherically drier, enriched CO$_2$ world.
-
-\end{abstract}
 
@@
 following \citet{Medlyn_2017} (Equation 5), and take the median $g_1$
-value to be representative of each PFT instead of the mean to avoid extra weighting of rare outliers (Table \ref{pft}).
+value to be representative of each PFT (Table \ref{pft}) instead of the mean to avoid extra weighting of rare outliers.
 
@@
     Medlyn coefficient, and calculated uWUE. uWUE from
-    \citet{Zhou_2015}, including the observed standard deviation is
+    \citet{Zhou_2015}, including the observed standard deviation, is
     shown for comparison. Note that uWUE from \citet{Zhou_2015} is
@@
 theoretical derivation of the square root VPD dependence in $uWUE$
-leverages the same asumptions used in \cite{MEDLYN_2011} to derive the
-square-root VPD dependence of the stomatal conductace model (Equation
+leverages the same assumptions used in \cite{MEDLYN_2011} to derive the
+square-root VPD dependence of the stomatal conductance model (Equation
 \ref{medlyn}).  We can use uWUE to remove the $GPP$ dependence from
@@
 
-By accounting for photosythesis changes in ecosystem conductace, with
+By accounting for photosynthesis changes in ecosystem conductance, with
 Equation \ref{et} we have derived for the first time, using recent
@@
 
-\section{Results}
+\section{Results and Discussion}
 \label{results}
@@
 term (Section \ref{methods}). These observational results of ET
-response largely confirm our theoretical analysis, presented above
-(Table \ref{stats}). For all PFTs, mean ET response to increasing VPD
+response (Table \ref{stats}) largely confirm our theoretical analysis, presented in the previous sections. For all PFTs, mean ET response to increasing VPD
 is negative. However, ET response evaluated at the average of all
@@
 
-As in Figure \ref{test_sign}, Figure \ref{data_scatter} matches our
+As with Figure \ref{test_sign}, Figure \ref{data_scatter} matches our
 expectation based on simplified theory. The sign term is most strongly
@@
     plot represents 5\% of all observations. Note that the highest
-    10\% of SWC observations were excluded from the WSA and SAV
-    figures, to better resolve high variability in narrow low-SWC bins
-    for those PFTs.}
+    10\% of SWC observations are excluded to better resolve variability in the much more narrow bins at lower SWC (e.g. SWC has long tails at high values).}
   \label{swc_boxplot}
@@
 function are relatively rare.
-Future research will explore these relatively rare feedbacks in extreme conditions, which due to analytical intractablity will require more opaque numerical analysis of many more complex processes, including boundary layer growth and state and their relationship to surface layer coupling and free-tropospheric lapse rates and humidity.
+Future research will explore these relatively rare feedbacks in extreme conditions, which due to analytical intractability will require more opaque numerical analysis of many more complex processes, including boundary layer growth and state and their relationship to surface layer coupling and free-tropospheric lapse rates and humidity.
 
@@
 rare (<30\% of observations) and does not manifest itself in the
-majority of observations and bulk statistics. However, for GRA SWC has
-a strong effect decreasing water use efficiency for the majority of
+majority of observations and bulk statistics. However for GRA, SWC decreases water use efficiency for the majority of
 the observations. Soil moisture effects explain the deficiencies of
@@
 given $uWUE$ and $g_1$, the ET dependence on VPD should be concave
-upward, which is confirmed by eddy covariance data cross most PFTs. In other words, there should be some local minimum in ET at a
+upward, which is confirmed by eddy covariance data across most PFTs. In other words, there should be some local minimum in ET at a
 critical VPD$_{crit}$, assuming the scaling and plant terms  (e.g. aerodynamic
@@
 ET-VPD dependence.
-This analysis is also important to understand whether the choice of stomatal conductance model alters the fundamental behavior of the ET-VPD relatioships, as many commonly used models  utilize a VPD exponent other than the 1/2 suggested by optimal theory \citep[e.g.
+This analysis is also important to understand whether the choice of stomatal conductance model alters the fundamental behavior of the ET-VPD relationships, as many commonly used models  utilize a VPD exponent other than the 1/2 suggested by optimal theory \citep[e.g.
 ][ which uses an exponent of 1]{Leuning_1990}.
@@
 derivative of ET, using Equation \ref{m_n} for stomatal conductance
-($g_s$, with respect to VPD: \input{d2_solutions.tex} With this
+($g_s$), with respect to VPD: \input{d2_solutions.tex} With this
 result we have defined the family of curves separating concave up from
@@
     concave up and concave down ET-VPD curves (Equation \ref{curves})
-    for 3 specific scenarios. Solutions are defined in terms of a
+    for three specific scenarios. Solutions are defined in terms of a
     non-dimensional VPD ($VPD^m/g_*$), but to aide physical
@@
 
-\section{Discussion}
+\section{Conclusions}
 
@@
 
+\cite{Zhou_2016}
+
 \bibliography{references.bib}

### 2018-05-14 doc/paper/vpd_et_paper.tex
cleaned up some references, ready to proofread and then post to arxiv

@@
-% 11/23/2015
-% \documentclass[draft,linenumbers]{afmjournal}
-\documentclass{article}
-% \draftfalse
+\input{def}
 \usepackage{makecell}
@@
 \usepackage{booktabs}
+\usepackage{authblk}
+\usepackage{textcomp}
+
+\bibliographystyle{agufull08}
+\RequirePackage{natbib}
+\RequirePackage{graphicx}
+
+\def\acknowledgments{\vskip12pt\noindent{\bf
+Acknowledgments\vrule depth 6pt
+width0pt\relax}\\*\noindent\ignorespaces}
+
 % \drafttrue
@@
 
-\authors{Adam Massmann\affil{a}, Pierre Gentine\affil{a}, Changjie
-  Lin\affil{a,b}}
-
-\affiliation{a}{Department of Earth and Environmental Engineering,
-  Columbia University, New York, NY 10027} \affiliation{b}{State Key
-  Laboratory of Hydroscience and Engineering, Department of Hydraulic
+\author[1]{Adam Massmann\thanks{Corresponding author: akm2203@columbia.edu}}
+\author[1]{Pierre Gentine}
+\author[1,2]{Changjie Lin}
+
+\affil[1]{Department of Earth and Environmental Engineering, Columbia University, New York, NY 10027}
+\affil[2]{State Key Laboratory of Hydroscience and Engineering, Department of Hydraulic
   Engineering, Tsinghua University, Beijing, CN 100084}
 
-\correspondingauthor{Adam Massmann}{206-919-1364;
-  akm2203@columbia.edu}
+\maketitle
 
@@
 
-\begin{keypoints}
-\item ecohydrology, evapotranspiration, aridity, vapor pressure
-  deficit, plant physiology, stomatal response
-\end{keypoints}
+\onehalfspacing
 
@@
 \begin{figure}
-  \centering \includegraphics[width=\textwidth]{./map.pdf}
+  \centering \includegraphics{./map.pdf}
   \caption{Plant functional type and location of FLUXNET2015 sites
@@
 variables and surface conductances:
-\begin{linenomath*}
+% \begin{linenomath*}
   \begin{equation}
@@
   \end{equation}
-\end{linenomath*}
+% \end{linenomath*}
 where $\Delta$ is the change in saturation vapor pressure with
@@
   \label{definitions}
-  \centering \small
+  \centering \footnotesize
   \begin{tabular}{l c c c}
@@
 
-\begin{linenomath*}
+% \begin{linenomath*}
   \begin{equation}
@@
   \end{equation}
-\end{linenomath*}
+% \end{linenomath*}
 where $g_{1-leaf}$ is a leaf-scale ``slope'' parameter, A is the net
@@
 parameter ($g_{1-leaf}$) to physical parameters as:
-\begin{linenomath*}
+% \begin{linenomath*}
   \label{slope}
@@
   \end{equation}
-\end{linenomath*}
+% \end{linenomath*}
 
@@
 \begin{figure}
-  \centering \includegraphics[width=\textwidth]{./joint_vpd_sigma.pdf}
+  \centering \includegraphics{./joint_vpd_sigma.pdf}
   \caption{The joint distribution of $VPD$ and $\sigma$, with outliers
@@
 \small
-\begin{linenomath*}
+% \begin{linenomath*}
   \begin{equation}
@@
   \end{equation}
-\end{linenomath*}
+% \end{linenomath*}
 \normalsize noting that $VPD_{crit}$ mostly depends on the PFT
@@
 \begin{figure}
-  \centering \includegraphics[width=\textwidth]{./idealized_sign.pdf}
+  \centering \includegraphics{./idealized_sign.pdf}
   \caption{The functional form of the sign term, with $\sigma$ held
@@
   % \centering
-  \centerline{\includegraphics[width=1.4\textwidth]{./idealized_scale.pdf}}
+  \centerline{\includegraphics[width=1.2\textwidth]{./idealized_scale.pdf}}
   \caption{Primary sources of variability for the scaling term, as a
@@
   \centering
-  \centerline{\includegraphics[width=1.4\textwidth]{./test_sign.pdf}}
+  \centerline{\includegraphics[width=1.2\textwidth]{./test_sign.pdf}}
   \caption{Comparison of the sign term with model uncertainty included
@@
   \centering
-  \centerline{\includegraphics[width=1.4\textwidth]{./data_scatter.png}}
+  \centerline{\includegraphics[width=1.2\textwidth]{./data_scatter.png}}
   \caption{Scatter plots of observed
@@
   \centering
-  \centerline{\includegraphics[width=1.4\textwidth]{./swc_boxplot.pdf}}
+  \centerline{\includegraphics[width=1.2\textwidth]{./swc_boxplot.pdf}}
   \caption{$\sigma \cdot uWUE$ as a function of PFT and SWC. Each box
@@
 conductance model from assumptions about VPD dependence:
-\begin{linenomath*}
+% \begin{linenomath*}
   \begin{equation}
@@
   \end{equation}
-\end{linenomath*}
+% \end{linenomath*}
 where:
@@
 shared by the FLUXNET community, including these networks: AmeriFlux,
-AfriFlux, AsiaFlux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont,
+AfriFlux, AsiaFlknux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont,
 ChinaFlux, Fluxnet-Canada, GreenGrass, ICOS, KoFlux, LBA, NECC,
@@
 
-\cite{Zhou_2016}
-
 \bibliography{references.bib}

### 2018-05-14 doc/paper/vpd_et_paper.tex
some tweaking of figure and positions for arkiv submission
arkiv submission

@@
-\input{def}
-\usepackage{makecell}
-\usepackage{multirow} %AM for tables
-\usepackage{booktabs}
-\usepackage{authblk}
-\usepackage{textcomp}
-
-\bibliographystyle{agufull08}
-\RequirePackage{natbib}
-\RequirePackage{graphicx}
-
-\def\acknowledgments{\vskip12pt\noindent{\bf
-Acknowledgments\vrule depth 6pt
-width0pt\relax}\\*\noindent\ignorespaces}
-
-% \drafttrue
-
-% \journalname{Agricultural and Forest Meteorology: When does VPD drive
-%   or reduce ET?}
-
-\newcommand{\appropto}{\mathrel{\vcenter{
-      \offinterlineskip\halign{\hfil$##$\cr
-        \propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}
+\documentclass[12pt]{article}
+\input{def.tex}
 
@@
 FLUXNET2015 database (data available at \sloppy
-https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/ \sloppy),
+\url{https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/} \sloppy),
 including all sites with at least four years of data and observations
@@
 
-\begin{figure}
-  \centering \includegraphics{./map.pdf}
+\begin{figure}[h]
+  \centering \includegraphics[trim={0 3cm 0 3cm}, clip]{./map.pdf}
   \caption{Plant functional type and location of FLUXNET2015 sites
@@
     \hline
-    \input{pft_params}
+    \input{pft_params.tex}
     \hline
@@
 the figures and tables, are publicly available at
-https://github.com/massma/climate\_et.
+\url{https://github.com/massma/climate\_et}.
+%\hyperref[https://github.com/massma/climate\_et]{https://github.com/massma/climate_et}.
 
@@
 
-\begin{figure}
-  \centering \includegraphics{./joint_vpd_sigma.pdf}
+\begin{figure}[h]
+  \centering\includegraphics[width=0.75\textwidth]{./joint_vpd_sigma.pdf}
   \caption{The joint distribution of $VPD$ and $\sigma$, with outliers
@@
 the theory ($\approx$ 35\%, Figure \ref{test_sign}). However, in
-general, and specifically for non-extreme VPD (VPD < $\approx$70-90th
+general, and specifically for non-extreme VPD (VPD  $< \approx$70-90th
 percentiles), the theory matches the observations for the tree and
@@
 DBF, MF and CSH PFTs, and captures the response at non-extreme VPD
-(VPD < $\approx$ 70th-90th percentiles) for
+(VPD $< \approx$ 70th-90th percentiles) for
 ENF, EBF, CRO, WSA, and SAV. In general, the leading order behavior
@@
 and VPD. Fortunately, for most PFTs SWC's effect on ET is relatively
-rare (<30\% of observations) and does not manifest itself in the
+rare ($<$30\% of observations) and does not manifest itself in the
 majority of observations and bulk statistics. However for GRA, SWC decreases water use efficiency for the majority of
@@
   \centering
-  \centerline{\includegraphics[width=1\textwidth]{./concave.pdf}}
+  \centerline{\includegraphics[width=0.75\textwidth]{./concave.pdf}}
   \caption{ Solutions corresponding to inflection points between
@@
 used to generate the figures and tables, are publicly available at
-https://github.com/massma/climate\_et.
+\url{https://github.com/massma/climate\_et}.
+%\hyperref[https://github.com/massma/climate\_et]{https://github.com/massma/climate_et}.
 

### 2018-05-25 doc/paper/body.tex
refactored tex seperate body, removed unknown external dependency on my util package in python modules, and set the stage for removal of tier 2 sites and addition of citation table <credit Trevor Keenan for catch>

--- 2018-05-14 doc/paper/vpd_et_paper.tex
@@
-\documentclass[12pt]{article}
-\input{def.tex}
-
-\begin{document}
-
-\title{When does vapor pressure deficit drive or reduce
-  evapotranspiration?}
-
-\author[1]{Adam Massmann\thanks{Corresponding author: akm2203@columbia.edu}}
-\author[1]{Pierre Gentine}
-\author[1,2]{Changjie Lin}
-
-\affil[1]{Department of Earth and Environmental Engineering, Columbia University, New York, NY 10027}
-\affil[2]{State Key Laboratory of Hydroscience and Engineering, Department of Hydraulic
-  Engineering, Tsinghua University, Beijing, CN 100084}
-
-\maketitle
-
-\begin{abstract}
-  Increasing vapor pressure deficit increases atmospheric demand for water, and vapor pressure deficit is expected to rise with increasing greenhouse gases.
-While increased evapotranspiration (ET) in response to increased atmospheric demand seems intuitive, plants are capable of reducing ET in response to increased VPD by closing their stomata, in an effort to conserve water.
-Here we examine which effect dominates response to increasing VPD: atmospheric demand and increases in ET, or plant physiological response (stomata closure) and decreases in ET.
-We use Penman-Monteith, combined with semi-empirical optimal stomatal regulation theory and underlying water use efficiency, to develop a theoretical framework for understanding how ET responds to increases in VPD.
-
- The theory suggests that for most environmental conditions and plant types, plant physiological response dominates and ET decreases with increasing VPD.
-Plants that are evolved or bred to prioritize primary production over water conservation (e.g.
-crops) exhibit a higher likelihood of atmospheric demand-driven response (ET increasing).
-However for forest, grass, savannah, and shrub plant types, ET more frequently decreases than increases with rising VPD.
-This work serves as an example of the utility of our simplified framework for disentangling land-atmosphere feedbacks at multiple scales, including the characterization of ET response in an atmospherically drier, enriched CO$_2$ world.
-\end{abstract}
-
-\onehalfspacing
-
 \section{Introduction}
@@
 needleleaf forest (ENF), mixed forest (MF), closed shrub (CSH),
-savannah (SAV), and woody savannah (WSA) (site locations in Figure
-\ref{map_fig}).
+savannah (SAV), and woody savannah (WSA) (site locations and citations in Figure
+\ref{map_fig} and Table \ref{flux_sites}).
 
@@
 \end{figure}
+
+\begin{table}
+  \caption{Metadata and citations for flux sites used in this analysis.
+All data are gathered from \url{www.fluxdata.org} using tools available at \url{https://github.com/trevorkeenan/FLUXNET_citations}.}
+  \label{flux_sites}
+  \label{definitions}
+  \centering \footnotesize
+  \begin{tabular}{l l l l l l l}
+    \hline
+    Site name & PFT & Lat & Lon & Clim$^{a}$ & Period & References \\
+    \hline
+    \input{flux_sites.tex}
+    \hline
+    \multicolumn{7}{l}{$^{a}$K{\"o}ppen Climate classification.}
+    \hline
+\end{table}
 
@@
 
-\acknowledgments This work used eddy covariance data acquired and
-shared by the FLUXNET community, including these networks: AmeriFlux,
-AfriFlux, AsiaFlknux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont,
+\acknowledgments This work used eddy covariance data acquired and shared
+by the FLUXNET community, including these networks: AmeriFlux, AfriFlux,
+AsiaFlknux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont,
 ChinaFlux, Fluxnet-Canada, GreenGrass, ICOS, KoFlux, LBA, NECC,
 OzFlux-TERN, TCOS-Siberia, and USCCC. The ERA-Interim reanalysis data
-are provided by ECMWF and processed by LSCE. The FLUXNET eddy
-covariance data processing and harmonization was carried out by the
-European Fluxes Database Cluster, AmeriFlux Management Project, and
-Fluxdata project of FLUXNET, with the support of CDIAC and ICOS
-Ecosystem Thematic Center, and the OzFlux, ChinaFlux and AsiaFlux
-offices. This material is based upon work supported by the National
-Science Foundation Graduate Research Fellowship under Grant No. DGE
-16-44869. All code and data used in this analysis, including those
-used to generate the figures and tables, are publicly available at
+are provided by ECMWF and processed by LSCE. The FLUXNET eddy covariance
+data processing and harmonization was carried out by the European Fluxes
+Database Cluster, AmeriFlux Management Project, and Fluxdata project of
+FLUXNET, with the support of CDIAC and ICOS Ecosystem Thematic Center,
+and the OzFlux, ChinaFlux and AsiaFlux offices. We would like to think
+Dr. Trevor Keenan for providing tools for citation of the FLUXNET2015
+dataset, available at
+\url{https://github.com/trevorkeenan/FLUXNET_citations}. This material is
+based upon work supported by the National Science Foundation Graduate
+Research Fellowship under Grant No. DGE 16-44869. All code and data used
+in this analysis, including those used to generate the figures and
+tables, are publicly available at
 \url{https://github.com/massma/climate\_et}.

### 2018-05-25 doc/paper/vpd_et_paper_afm.tex
AFM resubmission
refactored tex seperate body, removed unknown external dependency on my util package in python modules, and set the stage for removal of tier 2 sites and addition of citation table <credit Trevor Keenan for catch>

--- 2018-05-14 doc/paper/vpd_et_paper.tex
@@
 % 11/23/2015
-% \documentclass[draft,linenumbers]{afmjournal}
-\documentclass{article}
+\documentclass[draft,linenumbers]{afmjournal}
 % \draftfalse
-\usepackage{makecell}
-\usepackage{multirow} %AM for tables
-\usepackage{booktabs}
-% \drafttrue
+\usepackage{makecell} \usepackage{multirow} %AM for tables
+\usepackage{booktabs} \drafttrue
 
-% \journalname{Agricultural and Forest Meteorology: When does VPD drive
-%   or reduce ET?}
+\journalname{Agricultural and Forest Meteorology: When does VPD drive
+  or reduce ET?}
 
@@
 
-\section{Introduction}
-
-Vapor pressure deficit (VPD) is expected to rise over continents in
-the future due to the combination of increased temperature and,
-depending on region, decreased relative humidity
-\citep{Byrne_2013}. Increases in VPD increase the atmospheric demand
-for evapotranspirated water \citep{Penman_1948, Monteith_1965}, but
-also stress plant stomata \citep{Leuning_1990, MEDLYN_2011}.
-
-The opposing effects of increased atmospheric demand and higher
-stomatal stress lead to two possible perspectives for how
-evapotranspiration (ET) responds to shifts in VPD. The first, a
-hydrometeorological perspective, is that higher VPD increases
-atmospheric demand for water from the land surface, and this drives an
-increase in evapotranspiration (ET). This perspective is particularly
-relevant because potential evapotranspiration (PET), which is used in
-many drought indices and hydrometeorological studies
-\citep[e.g.,][]{Heim_2002, Scheff_2015}, typically only quantifies
-changes in atmospheric demand and fails to account for ecosystem
-response \citep{Swann_2016}. In reality, plants' stomata have evolved
-to optimally regulate the exchange of water and carbon, and tend to
-partially close in response to increased atmospheric dryness
-\citep{Farquhar_1978, Ball_1987, Leuning_1990, MEDLYN_2011}. This
-leads to a plant physiology perspective, in which an increase in VPD,
-particularly in well-watered soil conditions, may actually correspond
-to a decrease in ET because of stomatal closure
-\citep[e.g.][]{Rigden_2017}.  In other words, the question ``When does
-VPD drive or reduce ET?'' can be related to whether plant regulation
-or atmospheric demand dominates ET response.
-
-The ET response to changes in VPD alters water partitioning between
-the soil and atmosphere. If ecosystem plant response reduces ET with
-atmospheric drying then soil moisture will be better conserved. This
-would seem a sensible evolutionary strategy to cope with aridity. If
-stomata were fully passive \citep [similar to soil pores,
-e.g. ][]{Or_2013}, increased atmospheric aridity would strongly reduce
-soil moisture \citep{Berg_2017}. In turn, this would further increase
-aridity as low soil moisture levels increase the Bowen ratio, and
-cause increased temperature and atmospheric drying
-\citep[][]{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006,
-  Salvucci_2013, Gentine_2016, Berg_2016}. This however would not seem
-to be a sensible strategy for plants from an evolutionary
-standpoint. As a counterpoint, one may argue that increases in ET with
-increasing VPD could increase the likelihood of precipitation
-\citep[e.g.,][]{Findell_2011}. However, increases in ET do not always
-guarantee an increased likelihood of precipitation, which depending on
-environmental conditions could cause a decrease in the likelihood of
-precipitation \citep[][]{Gentine_2013}. Furthermore, any increases in
-precipitation are likely to be non-local, such that plants giving up
-water to the atmosphere are not guaranteed to reap the benefits of
-water returned from the atmosphere to the soil. Using this subjective
-logic, from an ecosystem evolutionary perspective, water stored in
-soil seems to be worth much more than the chance of water returned as
-precipitation.
-
-We can use intuition about plant water conservation strategy to
-hypothesize about ET response to changes in VPD. Plants and ecosystems
-that evolved to conserve water, such as arid shrubs or savannah,
-should be more likely to reduce ET with increasing VPD, and plants
-that have evolved or have been engineered to care little about water,
-such as crops, will be more likely to increase ET with increasing
-VPD. Atmospheric conditions must matter as well. At the ecosystem
-scale, there are limits to plant water conservation strategies. As
-atmospheric demand for water (VPD) increases, ecosystems should begin
-to reach their water conservation limits and might not be able to
-entirely limit ET flux to the atmosphere. At this stage any further
-increase in VPD will most likely drive a (limited) increase in ET,
-because the increase in atmospheric demand for water overwhelms the
-limited plant response to conserve water.
-
-The objective of the present manuscript is to use reasonable
-approximations as a tool to develop intuition for plant response to
-atmospheric drying and evaluate the VPD dependence of ET. This
-intuition will aid interpretation of observations, full complexity
-models, and facilitate the disentanglement of complex land-atmosphere
-feedbacks. In the past, similar approaches were used to understand
-interactions between stomatal conductance, evapotranspiration and the
-environment \citep[e.g.,][]{Jarvis_1986, Mcnaughton_1991}. However, at
-the time researchers' understanding of the form of VPD's effect on
-plant physiology was limited, so they could not explore the
-sensitivity of ET to VPD, including VPD's effect on stomatal
-conductance and plant function.
-
-Recent results have drastically improved our understanding of VPD's
-impact on physiology, especially at the leaf
-level. \citet{MEDLYN_2011} developed a model for leaf-scale stomatal
-conductance ($g_s$), including VPD response, by combining an optimal
-photosynthesis theory \citep{Cowan_1977} with an
-empirical approach, and extended use of this model to the ecosystem
-scale in \citet{Medlyn_2017}. Additionally, \citet{Zhou_2014}
-demonstrated that a quantity underlying water use efficiency
-$\left(uWUE = \frac{GPP\; \sqrt{VPD}}{ET}\right)$ properly captures a
-constant relationship between GPP, ET, and VPD over a diurnal cycle at
-the ecosystem scale. uWUE is also remarkably well conserved in the
-growing season across space and time, within a PFT
-\citep{Zhou_2015}. These results provide novel tools to examine how ET
-varies in response to VPD, while explicitly including VPD's effect on
-stomatal conductance for the first time. In this manuscript, we
-leverage these tools to derive the theoretical one-way response of ET
-to VPD with other environmental variables properly controlled for,
-i.e. we develop a framework for evaluating the partial derivative of
-ET with respect to VPD. Our theory is validated and tested at multiple
-eddy-covariance stations spanning various climates and plant
-functional types.
-
-\section{Materials and Methods}
-\subsection{Data}
-\label{data}
-We use both meteorological and eddy-covariance data from the
-FLUXNET2015 database (data available at \sloppy
-https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/ \sloppy),
-including all sites with at least four years of data and observations
-of the variables described in the methods section (Section
-\ref{methods}). Seventy-three sites met these requirements, and were
-grouped into nine plant functional types (PFT) according to the
-International Geosphere-Biosphere Programme vegetation classification
-scheme \citep{Loveland_1999}: cropland (CRO), grass (GRA), deciduous
-broadleaf forest (DBF), evergreen broadleaf forest (EBF), evergreen
-needleleaf forest (ENF), mixed forest (MF), closed shrub (CSH),
-savannah (SAV), and woody savannah (WSA) (site locations in Figure
-\ref{map_fig}).
-
-\begin{figure}
-  \centering \includegraphics[width=\textwidth]{./map.pdf}
-  \caption{Plant functional type and location of FLUXNET2015 sites
-    used in this analysis.}
-  \label{map_fig}
-\end{figure}
-
-The purpose of this study is to examine ecosystem response to
-atmospheric drying, focusing on the growing season. To accomplish
-this, we filter and quality control the data using a similar procedure
-as \cite{Zhou_2015}:
-\begin{itemize}
-\item Only measured or highest (``good'') quality gapfilled data,
-  according to quality control flags, are used.
-\item To isolate the growing season, we only use days in which the
-  average Gross Primary Productivity (GPP) exceeds 10\% of the
-  observed 95th percentile of GPP for a given site. GPP is calculated
-  using the nighttime respiration partitioning method.
-\item We remove days with rain and the day following to avoid issues
-  with rain interception and sensor saturation at high relative
-  humidity (\cite{Medlyn_2017}).
-\end{itemize}
-Additionally, as in \citet{Lin_2018}, we restrict data to the daytime,
-which is identified when downwelling shortwave radiation is greater
-than 50 W m$^{-2}$ and sensible heat flux is greater than 5 W
-m$^{-2}$. To reduce the chance of sensor saturation at high relative
-humidity, we remove all time steps for which VPD is less than .01 kPa,
-and to reduce errors at low windspeeds we remove all periods with wind
-magnitudes less than 0.5 m s$^{-1}$. Timesteps with negative observed
-GPP or ET are also removed, and we aggregate half hourly data to
-hourly averages to reduce noise \citep{Lin_2018}. After these quality
-control procedures, 434,765 upscaled hourly observations remain.
-
-\subsection{Methods}
-\label{methods}
-The Penman-Monteith equation \citep [hereafter PM,][]{Penman_1948,
-  Monteith_1965} estimates ET as a function of observable atmospheric
-variables and surface conductances:
-\begin{linenomath*}
-  \begin{equation}
-    \label{orig_pen}
-    ET = \frac{\Delta R_{net} + g_a \rho_a c_p VPD}{\Delta + \gamma(1 + \frac{g_a}{g_s})},
-  \end{equation}
-\end{linenomath*}
-where $\Delta$ is the change in saturation vapor pressure with
-temperature, given by Clausius-Clapeyron ($\frac{d \; e_s}{d \; T}$),
-$R_{net}$ is the net radiation minus ground heat flux, $g_a$ is
-aerodynamic conductance, $\rho_a$ is air density, $c_p$ is specific
-heat of air at constant pressure, $\gamma$ is the psychometric
-constant, and $g_s$ is the stomatal conductance (Table
-\ref{definitions}).
-
-\begin{table}
-  \caption{Definition of symbols and variables, with citation for how
-    values are calculated, if applicable.}
-  \label{definitions}
-  \centering \small
-  \begin{tabular}{l c c c}
-    \hline
-    Variable & Description & Units & Citation \\
-    \hline
-    $e_s$  & saturation vapor pressure & Pa  & - \\ 
-    $T$  & temperature  & K & - \\
-    $P$  & pressure & Pa  & - \\
-    $\Delta$  & $\frac{\partial e_s}{\partial T}$ & Pa K$^{-1}$ & - \\
-    $R_{net}$  & net radiation at land surface minus ground heat flux & W m$^{-2}$   & - \\
-    $g_a$  & aerodynamic conductance & m s$^{-1}$  & \citet{Shuttleworth_2012} \\
-    $\rho_a$  & air density & kg m$^{-3}$  & - \\
-    $c_p$  & specific heat capacity of air at constant pressure & J K$^{-1}$ kg$^{-1}$ & - \\
-    $VPD$  & vapor pressure deficit & Pa  & - \\
-    $\gamma$  & psychometric constant & Pa K$^{-1}$   & - \\
-    $g_{s-leaf}$  & leaf-scale stomatal conductance & m s$^{-1}$  & \citet{MEDLYN_2011} \\
-    $g_{s}$  &  stomatal conductance & m s$^{-1}$
-                                   & \citet{Medlyn_2017} \\
-    $g_{1-leaf}$  & leaf-scale slope parameter & Pa$^{0.5}$
-                                   & \citet{MEDLYN_2011} \\
-    $g_{1}$  & ecosystem-scale slope parameter & Pa$^{0.5}$ & \citet{Medlyn_2017} \\
-    $R$ & universal gas constant & J mol$^{-1}$ K$^{-1}$ & - \\
-    $R_{air}$ & gas constant of air & J  K$^{-1}$ kg$^{-1}$ & - \\
-    $\sigma$ & uncertainty parameter & -& - \\
-    $c_a$ & CO$_2$ concentration & $\mu$ mol CO$_2$ mol$^{-1}$ air& - \\
-    $\lambda$ & marginal water cost of leaf carbon & mol H$_2$O mol$^{-1}$ CO$_2$ & - \\
-    $\Gamma$ & CO$_2$ compensation point & - & - \\
-    $\Gamma^*$ & CO$_2$ compensation point without dark respiration & - & - \\
-    \hline
-  \end{tabular}
-\end{table}
-
-\citet{MEDLYN_2011} developed a model for stomatal conductance ($g_s$)
-by combining an optimal photosynthesis theory \citep{Cowan_1977} with an empirical approach, which describes the
-dependence of $g_s$ to VPD. This resulted in the following model for
-leaf-scale stomatal conductance:
-
-\begin{linenomath*}
-  \begin{equation}
-    g_{s-leaf} = g_0 + 1.6 \left(1 +
-      \frac{g_{1-leaf}}{\sqrt{VPD}}\right) \frac{A}{c_a},
-    \label{leaf_medlyn}
-  \end{equation}
-\end{linenomath*}
-where $g_{1-leaf}$ is a leaf-scale ``slope'' parameter, A is the net
-CO$_2$ assimilation rate, and $c_a$ is the atmospheric CO$_2$
-concentration at the leaf surface. \cite{MEDLYN_2011} relate the slope
-parameter ($g_{1-leaf}$) to physical parameters as:
-\begin{linenomath*}
-  \label{slope}
-  \begin{equation}
-    g_{1-leaf} = \sqrt{\frac{3 \, \Gamma^* \, \lambda}{1.6}},\footnote{Note this expression has units of of (mmol mol$^{-1}$)$^{1/2}$, but this can be converted to Pa$^{1/2}$ using the ideal gas law.}
-  \end{equation}
-\end{linenomath*}
-
-where $\Gamma^*$ is the CO$_2$ compensation point for photosynthesis
-(without dark respiration), and $\lambda$ is the marginal water cost
-of leaf carbon
-($\frac{\partial \; \text{transpiration}}{\partial \; A}$). So,
-$g_{1-leaf}$ is a leaf-scale term reflecting the trade-off of water for
-carbon uptake. The higher $g_{1-leaf}$, the more open the stomata and
-the more they release water in exchange for carbon.
-
-The Medlyn model for stomatal conductance has been shown to behave
-very well across PFTs \citep[][]{Lin_2015}, and has been successfully
-adopted to ecosystem scale analysis in \citet{Medlyn_2017}. In units
-of m s$^{-1}$, the ecosystem scale stomatal conductance is:
-
-\begin{linenomath*}
-  \begin{equation}
-    g_s = \frac{R \,T}{P} \; 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{GPP}{c_a},
-    \label{medlyn}
-  \end{equation}
-\end{linenomath*}
-
-where GPP is the ecosystem scale gross primary production, and $g_1$
-is an ecosystem scale analogue to $g_{1-leaf}$. We solve for $g_1$
-following \citet{Medlyn_2017} (Equation 5), and take the median $g_1$
-value to be representative of each PFT (Table \ref{pft}) instead of the mean to avoid extra weighting of rare outliers.
-
-\begin{table}
-  \caption{Plant functional types, their abbreviation, calculated
-    Medlyn coefficient, and calculated uWUE. uWUE from
-    \citet{Zhou_2015}, including the observed standard deviation, is
-    shown for comparison. Note that uWUE from \citet{Zhou_2015} is
-    calculated from a different set of sites, and that units are
-    converted such that the quantities work with Equations 1-8 and the
-    variables defined Table \ref{definitions}.}
-  \small
-  \label{pft}
-  \centering
-  \begin{tabular}{l c c @{\qquad} c c}
-    \hline
-    \multirow{2}[3]{*}{Abbreviation} & \multirow{2}[3]{*}{PFT} & \multirow{2}[3]{*}{$g_1$ (Pa$^{0.5}$)} & \multicolumn{2}{c}{uWUE ($\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET])}  \\
-    \cmidrule{4-5}
-
-                                     & & & fitted & \citet{Zhou_2015} \\
-
-    \hline
-    \input{pft_params}
-    \hline
-  \end{tabular}
-\end{table}
-
-While Equation 3 can be used in PM (Equation 1), it will make
-analytical work with the function intractable because $GPP$ is
-functionally related to ET itself. Additionally, a perturbation to VPD
-should induce a physiological plant response that will alter GPP and
-cause an indirect change in stomatal conductance, in addition to the
-direct effect of VPD in Equation \ref{medlyn}. Therefore, in order to
-derive the response of ET to VPD, we must account for the functional
-relationship between GPP, ET, and VPD, and its effect on stomatal
-conductance. We can use aforementioned semi-empirical results of
-\citet{Zhou_2015} as a tool to approach this
-problem. \citet{Zhou_2015}, showed that  underlying Water Use
-Efficiency (uWUE):
-
-\begin{linenomath*}
-  \begin{equation}
-    uWUE = \frac{GPP \cdot \sqrt{VPD}}{ET}
-    \label{uwue}
-  \end{equation}
-\end{linenomath*}
-
-is relatively constant across time and moisture conditions within a
-plant functional type, and correctly captures a constant relationship
-between GPP, ET and VPD over a diurnal cycle \citep{Zhou_2014}. The
-theoretical derivation of the square root VPD dependence in $uWUE$
-leverages the same assumptions used in \cite{MEDLYN_2011} to derive the
-square-root VPD dependence of the stomatal conductance model (Equation
-\ref{medlyn}).  We can use uWUE to remove the $GPP$ dependence from
-$g_s$ in a way that makes PM analytically tractable:
-
-\begin{linenomath*}
-  \begin{equation}
-    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{uWUE \; ET}{c_a \; \sqrt{VPD}}.
-    \label{new_g_s}
-  \end{equation}
-\end{linenomath*}
-
-Plugging Equation \ref{new_g_s} into Equation \ref{orig_pen} and
-rearranging gives a new explicit expression for PM, in which
-dependence on $GPP$ is removed:
-
-\begin{linenomath*}
-  \begin{equation}
-    ET = \frac{\Delta R_{net} + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R \; 1.6\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
-    \label{et}
-  \end{equation}
-\end{linenomath*}
-
-By accounting for photosynthesis changes in ecosystem conductance, with
-Equation \ref{et} we have derived for the first time, using recent
-results \citep[][]{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017}, ET
-explicitly as function of environmental variables and two
-plant-specific constants, the slope parameter ($g_1$), and uWUE, both
-reflecting water conservation strategy. The slope parameter is related
-to the willingness of stomata to trade water for CO$_2$ and to keep
-stomata open. uWUE is a semi-empirical ecosystem-scale constant
-related to how WUE changes with VPD (specifically $VPD^{-1/2}$). It is
-also roughly proportional to physical constants:
-
-\[uWUE \appropto \sqrt{\frac{c_a - \Gamma}{1.6 \lambda}},\]
-
-where $\Gamma$ is the CO$_2$ compensation point \citep[Equation 5
-in][]{Zhou_2014}. So uWUE is related to atmospheric CO$_2$
-concentration and compensation point, and is inversely proportional to
-the marginal water cost of leaf carbon.
-
-Given eddy-covariance FLUXNET2015 data (Section \ref{data}), every
-term in our new version of PM (Equation \ref{et}) is observed, except
-for uWUE, which we fit by calculating its expectation, given the model
-and FLUXNET2015 data.
-
-However, eddy-covariance data are inherently noisy so we include a
-measure of uncertainty in our analysis. To account for observational
-error, as well as model uncertainty (e.g. temporal and spatial
-variations of uWUE and $g_1$), we introduce an uncertainty parameter
-$\sigma$ modifying uWUE:
-
-\begin{linenomath*}
-  \begin{equation}
-    ET = \frac{\Delta R_{net} + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R \; 1.6\; \sigma \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
-    \label{et_sigma}
-  \end{equation}
-\end{linenomath*}
-
-Now, from each FLUXNET2015 observation (i.e. for each hourly
-observation at every time step) we can evaluate $\sigma$:
-
-\begin{linenomath*}
-  \begin{equation}
-    \sigma = - \frac{g_a \gamma c_a \sqrt{VPD} L_v P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R_{net} - g_a \rho_a c_p VPD\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})},
-    \label{sigma}
-  \end{equation}
-\end{linenomath*}
-
-Thus, with this uncertainty analysis we can evaluate departure from
-our theory in observations, as a departure of $\sigma$ from unity. The
-variability of $\sigma$ across sites and time provides a measure of
-uncertainty in our model, assumptions, as well as in the FLUXNET2015
-observations themselves. The variability of $\sigma$ then propagates
-through any uncertainty to our partial derivative of Equation
-\ref{et_sigma} with respect to VPD:
-
-\begin{linenomath*}
-  \begin{equation}
-    \frac{\partial \;  ET}{\partial \; VPD} = \frac{2\; g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_a }{1.6 \; R\; \sigma \; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)
-    \label{d_et}
-  \end{equation}
-\end{linenomath*}
-
-With Equation \ref{d_et} we have an analytical framework for ecosystem
-response to atmospheric demand perturbations with environmental
-conditions held fixed. There are a few subtleties to taking the
-derivative in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$)
-and $VPD$ are functionally related, so while taking the derivative we
-evaluate
-$\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET}
-{\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD}
-\Big|_{\text{RH fixed}} + \frac{\partial \; ET}{\partial \; RH}
-\frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$
-    fixed}}$. $RH$ and $e_s$ are assumed to be approximately
-independent, which is supported by the data (not shown).
-
-This derivation relied either implicitly or explicitly on several
-assumptions. First, we assume that VPD at the leaf surface is the same
-as VPD at measurement height; physically this implies that leaves are
-perfectly coupled to the atmosphere. In reality, for some conditions
-and plant types the leaves can become decoupled from the boundary
-layer \citep{De_2017, Medlyn_2017}. Given our focus on the growing
-season, which is usually characterized by relatively high insolation
-inducing instability and convective boundary layers, we would expect
-the surface to be generally well coupled. An additional assumption in
-the formulation of $uWUE$ \citep{Zhou_2014, Zhou_2015} and
-\citet{Medlyn_2017}'s stomatal conductance model is that direct soil
-evaporation (E) contributions to ET remain small relative to
-transpiration (T). This should be more true during the growing
-season. The ratio of E to T may increase immediately after rainfall
-events due to high soil moisture and ponding, but VPD is generally low
-anyways during these times. However, some plant types allow for
-systematically larger contributions of E in ET, particularly those
-with sparse canopies and smaller relative amounts of transpiration. We
-therefore might expect that the theory will be most applicable to
-forest PFTs, which will be most strongly coupled to the boundary layer
-due to larger surface roughness, and will also generally have the
-highest ratios of transpiration to evaporation. Finally, for the goal
-of developing PFT-wide intuition, we assume that $g_1$ and $uWUE$ are
-constant in space and time. Both quantities have been shown to be well
-conserved over space and time within a PFT relative to inter-PFT
-variability. However, we would expect them to vary somewhat with
-environmental conditions including very low soil water content,
-temperature, and with intra-PFT plant-specific characteristics like
-wood density for tree PFTs \citep{Lin_2015}. Given these strong
-assumptions made with the goal of understanding broad, leading order
-plant behaviors, we take the extra care of including and analyzing
-uncertainty (Section \ref{testing}) and examining sensitivity to the
-exponent of the VPD response (Section \ref{functional_form}).
-
-We note one final comment on our derivation which is relevant for
-drought indices. If we approximate $c_a$ at a global mean CO$_2$
-concentration, then the RHS of Equation \ref{et} is fully defined
-using commonly available weather station data and the constants
-published here. This makes Equation \ref{et} a useful alternative to
-PET in drought indices and hydrometeorological analysis for vegetated
-surfaces. Equation \ref{et} better reflects the physics of water
-exchange at the land surface and would only require fitting of uWUE
-and $g_1$, and can also account for changes in CO$_2$ concentration
-(assuming some relationship between $\lambda$ and [CO$_2$]), contrary
-to typical drought indices such as the Palmer Drought Severity Index
-(PDSI) \citep{Swann_2016, Lemordant_2016, Lemordant_2018}. Lastly, all
-code and data used in this analysis, including those used to generate
-the figures and tables, are publicly available at
-https://github.com/massma/climate\_et.
-
-\section{Results and Discussion}
-\label{results}
-
-By construction, the variability in the $\sigma$ term (Equation
-\ref{sigma}) contains all model and observational uncertainties. For
-an observation that perfectly matches our model and constant uWUE
-assumption, $\sigma$ will be one. Therefore, for our assumptions and
-framework to be reasonable $\sigma$ should be close to 1. An
-additional concern is that $\sigma$ may in fact be correlated with
-$VPD$, in which case the dependence would need to be accounted for
-when taking the derivative. Fortunately, there is a very weak
-dependence of $\sigma$ on VPD in their joint distribution, and
-$\sigma$ is indeed close to unity i.e. $O(1)$ (Figure
-\ref{joint_vpd_sigma}). Given this weak dependence and the
-distribution of $\sigma$ we have confidence in our model framework and
-the data quality.
-
-\begin{figure}
-  \centering \includegraphics[width=\textwidth]{./joint_vpd_sigma.pdf}
-  \caption{The joint distribution of $VPD$ and $\sigma$, with outliers
-    removed (defined as lowest and highest 5\% of $\sigma$). $\sigma$
-    exhibits a weak dependence on $VPD$, and $\sigma$ is $O(1)$ for
-    the bulk of the observations.}
-  \label{joint_vpd_sigma}
-\end{figure}
-
-Before calculating the sensitivity of ET to VPD, we will consider the
-functional form of Equation \ref{d_et}. There are two main terms: a
-``scaling'' term, which modifies the magnitude but not the sign of the
-ET response to VPD ($\frac{\partial \; ET}{\partial \; VPD}$):
-
-\begin{equation}
-  \frac{g_a \; P}{T(\Delta + \gamma)},
-\end{equation}
-
-and a ``sign'' term, which determines whether ET increases or
-decreases with VPD (i.e. atmospheric demand driven or physiologically
-controlled):
-
-\begin{equation}
-  \label{sign}
-  \frac{c_p}{R_{air}} - \frac{\gamma c_a }{1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right).
-\end{equation}
-All variables are positive, so the relative magnitude between the
-first term and the second term in the sign term (Equation \ref{sign})
-will determine whether ET increases or decreases with increasing
-VPD. If the second term is larger then plant control dominates and ET
-decreases with increasing VPD. However, if the first term is larger,
-then atmospheric demand dominates and ET increases with increasing
-VPD.
-
-\subsection{Functional Form of the Sign Term}
-\label{sign_term}
-First, we explore the variables within the sign term to gain better
-intuition on the driver of either the increase or reduction of ET with
-VPD. CO$_2$ concentration ($c_a$) and the psychometric constant
-($\gamma$) are relatively constant over the dataset considered here so
-that the variability is dominated by $\sigma$ and $VPD$. uWUE could
-vary with soil moisture but has been shown to be relatively constant
-\citep{Zhou_2015}. This then means that the sign term only depends on
-VPD for a given PFT and is approximately just a function of $VPD$. We
-can further determine a critical threshold separating an increase from
-a decrease in ET, i.e. the threshold $VPD_{crit}$ such that the
-derivative vanishes $\frac{\partial \; ET}{\partial \; VPD} = 0$:
-\small
-\begin{linenomath*}
-  \begin{equation}
-    VPD_{crit} = \frac{R_{air}}{4 c_p} \left( \frac{\gamma c_a}{1.6\; R \;  uWUE} + \sqrt{\frac{\gamma c_a}{1.6\; R \;  uWUE}\left( \frac{\gamma c_a}{1.6\; R \;  uWUE} + 8 g_1 \frac{c_p}{R_{air}}\right)} - 4 g_1 \frac{c_p}{R_{air}} \right),
-    \label{vpd_min_et}
-  \end{equation}
-\end{linenomath*}
-\normalsize noting that $VPD_{crit}$ mostly depends on the PFT
-parameters uWUE and $g_1$, and only varies weakly with climate as most
-other parameters related to the environment are nearly constant. The
-calculated value of $VPD_{crit}$ for each PFT is shown in Table
-\ref{vpd_crit}. For any values of $VPD$ less than $VPD_{crit}$, ET
-will decrease with increasing VPD
-($\frac{\partial \; ET}{\partial \; VPD} < 0$), and for values of
-$VPD$ greater than $VPD_{crit}$, ET will increase with increasing VPD
-($\frac{\partial \; ET}{\partial \; VPD} > 0$). In other words,
-ecosystems regulate and mitigate evaporative losses up to the VPD
-limit, $VPD_{crit}$, above which atmospheric demand is just too high
-to be entirely compensated by stomatal and ecosystem regulation. We
-note however that even though ET increases again above the critical
-threshold, $VPD_{crit}$, ET is still much lower than potential
-evaporation as stomata are still strongly regulating vapor fluxes to
-the atmosphere. Indeed, even in the absence of soil pore evaporation,
-stomata do not shut down entirely at very high VPD so that ET does not
-go to zero, because stomata are still slightly open to perform some
-photosynthesis \citep{Ball_1987, Leuning_1990, MEDLYN_2011}. In
-addition, upward xylem transport is necessary to maintain phloem
-transport, as well as nutrient transport and thus carbon allocation
-\citep{De_2013, Nikinmaa_2013, Ryan_2014}.
-
-\begin{table}
-  \caption{Values of $VPD_{crit}$, where
-    $\frac{\partial \; ET}{\partial \; VPD} = 0$, evaluated at PFT
-    average values for $R_{air}$, $\gamma$, and $c_a$. PFT-specific
-    constants ($g_1$, uWUE) are provided in Table \ref{pft}. For
-    values of $VPD$ less than $VPD_{crit}$,
-    $\frac{\partial \; ET}{\partial \; VPD}$ will be negative, and for
-    values of $VPD$ greater than $VPD_{crit}$,
-    $\frac{\partial \; ET}{\partial \; VPD}$ will be positive.}
-  \centering
-  \begin{tabular}{l c c c c c}
-    \hline
-    PFT & $R_{air}$ & $c_a$ (ppm) & $\gamma$  & \textbf{$VPD_{crit}$ (Pa)} \\
-    \hline
-    \input{vpd_crit.tex}
-    \hline
-  \end{tabular}
-  \label{vpd_crit}
-\end{table}
-
-Differences in $VPD_{crit}$ are exclusively determined by uWUE and the
-slope parameter ($g_1$) related to the plant functional type. A larger
-uWUE means a smaller $VPD_{crit}$, and an ET response to increases VPD
-that is more likely to be positive. At first glance this result is
-somewhat counter-intuitive; we expect that plants with a higher water
-use efficiency would be more water conservative. However, in reality
-uWUE determines how $WUE$ changes with VPD:
-
-\[WUE = \frac{GPP}{ET} = \frac{uWUE}{\sqrt{VPD}}.\]
-\[\frac{\partial \; WUE}{\partial \; VPD} = -\frac{uWUE}{2 \;
-    VPD^{3/2}}\]
-
-So, plants with a higher uWUE will have a greater \textit{decrease} in
-ecosystem-scale $WUE$ in response to increases in VPD. This decrease
-in $WUE$ causes more water loss per unit carbon gain, and explains the
-relationship between high uWUE and high likelihood of increases of ET
-in response to increasing atmospheric drying (increases in VPD).
-
-A tendency towards increasing ET response with increasing VPD can also
-be caused by a high slope parameter ($g_1$), characteristic of plants
-that at the leaf scale are more willing to trade water for access to
-atmospheric CO$_2$. Plants that are less conservative will be thus be
-more likely to increase ET with increasing VPD. Both the
-aforementioned effects (large uWUE, $g_1$) can amplify each other, and
-generally conspire to shift the sign term towards a positive value for
-a given PFT.
-
-This effect of uWUE and $g1$ on the sign term is most apparent by
-comparing two extreme PFTs: water intensive crops (CRO) and water
-conservative closed shrub (CSH). CRO has higher slope parameter and a
-slightly higher uWUE ($g_1 = 140.7$ Pa$^{1/2}$; $2.85$ $\mu$-mol [C]
-Pa$^{0.5}$ J$^{-1}$ [ET]) compared to CSH ($g1 = 75.1 \, Pa^{1/2}$,
-$uWUE=2.82$ $\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET]). These differences
-in PFT parameters cause opposite ET responses to changes in VPD
-between CRO and CSH. ET theoretically always decreases with increasing
-VPD for the more water conservative CSH, while ET frequently increases
-with increasing VPD for the more water intensive CRO (Figure
-\ref{idealized_sign}). CROs evolved or were bred to prioritize GPP and
-yield and are thus not water conservative. They are very willing to
-trade water for photosynthesis and productivity, despite changes in
-VPD, while CSH are very unwilling to trade water for more
-photosynthesis.
-
-\begin{figure}
-  \centering \includegraphics[width=\textwidth]{./idealized_sign.pdf}
-  \caption{The functional form of the sign term, with $\sigma$ held
-    fixed at 1, and all terms except VPD set to PFT averages. For
-    comparison, the observed range of VPD for each PFT is plotted
-    below the x-axis. Stars denote 25th, 50th, and 75th percentiles,
-    and the range of the line spans the 5th-95th percentiles of
-    observed VPD. Vertical black lines denote the location of
-    $VPD_{crit}$ for each PFT, with the exception of CSH and WSA, for which
-    $VPD_{crit}$ is off-scale.}
-  \label{idealized_sign}
-\end{figure}
-
-As expected, the slope parameter ($g_1$) is a primary determinant of
-the VPD dependence for the sign term shown in Figure
-\ref{idealized_sign}. Plants that are more conservative (small $g_1$)
-will tend to reduce ET with increasing VPD, and will be very effective
-at reducing ET, especially at low VPD. However, at very high VPD,
-gradients in vapor pressure at the leaf scale will become very strong
-as stomata reach their limits of closure in response to VPD
-(parameterized with $g_1$). As a result, ET response will begin to
-asymptote towards a constant ecosystem-scale values as leaf-scale
-response to VPD asymptotes towards zero.  Therefore, plants with a low
-$g_1$ will have the largest VPD dependence of ET response because the
-difference in ET response at low VPD (leaf stomatal response
-dominates) and high VPD (VPD gradient dominates) is largest. This is
-apparent in the strong VPD dependence of CSH, which has the lowest
-slope parameter ($g_1=75.1$ Pa$^{1/2}$) (Figure \ref{idealized_sign}).
-
-To summarize our theoretical insights (Figure \ref{idealized_sign} and
-Table \ref{vpd_crit}), CROs are the least water conservative and have
-the strongest overall tendency to increase ET with increasing VPD,
-while CSH are the most water conservative and have the strongest
-tendency to decrease ET with increasing VPD, as well as the strongest
-VPD dependence of response. Figure \ref{idealized_sign} clearly shows,
-according to our theory, that for all PFTs except for crops there is
-frequent occurrence of a negative (plant dominating) ET response to
-increases in VPD. Therefore, plants are able in most atmospheric
-conditions to reduce ET in response to increased VPD and thus to
-reduce water loss. To better illustrate this, the ranges of observed
-environmental VPDs at the FLUXNET sites are plotted parallel to the
-x-axis. For CSH and WSA, VPD is always less than VPD$_{crit}$ (off scale) so
-that the plant response dominates in typical environmental conditions,
-emphasizing the water conservative strategy of those plants. For CRO
-on the other hand, VPD is higher than VPD$_{crit}$ for more than 50\%
-of observations, emphasizing that those plants operate with an
-aggressive water usage strategy, are water intensive and were actually
-engineered for photosynthesis rather than water saving. For DBF, EBF,
-MF, GRA, and SAV more than half of the observed VPD are less than
-VPD$_{crit}$, i.e. in conditions where plant response dominates. SAV
-has a more water conservative response than the forest, grass, and
-crop plan types, but still responds by increasing ET with increasing
-VPD for about a quarter of observations, due to the high aridity (VPD)
-of the SAV ecoclimate. It is also important to note that for all PFTs,
-even when atmospheric demand dominates, ET response to VPD is still
-far more negative than it would be for potential evaporation
-$\partial PET/\partial VPD$, i.e. atmospheric demand only, emphasizing
-that there is still a strong regulation of evaporative flux by stomata
-and though the plant xylem. The sign term in the PET case would just
-be a constant ($\frac{c_p}{R_{air}} \approx 3.5$), which is far larger
-than any part of the curves for any PFT. This highlights the
-deficiencies of PET's ability to capture land response to changes in
-atmospheric dryness (VPD). Plants are always regulating water exchange
-from the land surface, even when they reach the limits of they ability
-to do so.
-
-\subsection{Functional Form of the Scaling Term}
-\label{scale_term}
-While the above discussion of the sign of
-$\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our
-question of when ET response increases or decreases with VPD,
-understating the overall magnitude of the ET response is important to
-soil-plant-atmosphere water budgeting. So we now more closely examine
-the terms that affect how the sign term is scaled:
-
-\begin{equation}
-  \frac{g_a \; P}{T(\Delta + \gamma)}.
-\end{equation}
-
-$\frac{P}{T}$ is an air-density term, which varies little compared to
-aerodynamic conductance and Clausius-Clapeyron ($\Delta$). The
-psychometric constant ($\gamma$) is also relatively constant, so the
-scaling term should be primarily a function of aerodynamic conductance
-and temperature, through the Clausius-Clapeyron relationship
-$\Delta$. This is as expected, given that the aerodynamic conductance
-represents the efficiency of exchange between the surface and the
-atmosphere. As aerodynamic conductance increases, any plant response
-will be communicated more strongly to the atmosphere (and vice-versa).
-
-$\Delta$'s presence in the scaling term also matches physical
-intuition. $\Delta$ (and also the approximately constant $\gamma$)
-control the efficiencies with which surface energy is converted to
-latent and sensible heat \citep{Monteith_1965}. The functional from of
-$\Delta$ will be the same across PFTs, but the temperature range may
-vary slightly. In contrast, aerodynamic conductance will vary strongly
-with PFT due to the importance of surface roughness for aerodynamic
-conductance. So most of the differences in scaling between PFT should
-be in the aerodynamic conductance term.
-
-The control of the scaling term variability between PFTs by
-aerodynamic conductance is confirmed by data (Figure
-\ref{scale_vary}). Differences between PFT are almost entirely due to
-differences in aerodynamic conductance, rather than differences in
-observed temperature ranges. The scaling term for the tree PFTs (DBF,
-EBF, ENF, MF) is generally about double the scaling terms for other
-PFTs which have lower surface roughness and generally smaller
-aerodynamic conductance (GRA, CSH, CRO). The savannah (WSA, SAV) PFT's
-scaling is somewhere between GRA, CSH, and CRO, and DBF, EBF, ENF, and
-MF, due to higher variability and surface roughness.
-
-\begin{figure}
-  % \centering
-  \centerline{\includegraphics[width=1.4\textwidth]{./idealized_scale.pdf}}
-  \caption{Primary sources of variability for the scaling term, as a
-    function of PFT. The 5th-95th percentile range of temperature is
-    plotted at the 5th, 25th, 50th, 75th, and 95th percentiles of
-    aerodynamic conductance, as observed for each PFT.}
-  \label{scale_vary}
-\end{figure}
-
-Within each PFT, the scaling term variability is controlled both by
-environmental temperature and aerodynamic conductance variability
-(Figure \ref{scale_vary}). While the observed variability of the
-aerodynamic conductance contributes more to the scaling term
-variability than temperature, the temperature contribution is
-non-negligible. Specifically, the scaling term is generally larger at
-low temperatures when latent heat is relatively inefficient at moving
-energy away from the surface. This effect amplifies the role of
-aerodynamic conductance variability at low temperatures.
-
-To summarize, variability between PFTs is mostly controlled by
-systematic differences in aerodynamic conductance, due to differences
-in surface roughness between each PFT, and possibly to a lesser extent
-wind conditions. In contrast, variability within PFT is also
-controlled by temperature, through Clausius-Clapeyron. But,
-aerodynamic conductance variability generally impacts the scaling term
-more than temperature, even within PFTs.
-
-\subsection{Bulk statistics of ET response to VPD}
-\label{stats_sec}
-
-In this section we consider direct observations of ET response with
-eddy-covariance data, while including uncertainty with the $\sigma$
-term (Section \ref{methods}). These observational results of ET
-response (Table \ref{stats}) largely confirm our theoretical analysis, presented in the previous sections. For all PFTs, mean ET response to increasing VPD
-is negative. However, ET response evaluated at the average of all
-variables (e.g. $\sigma$, $T$, $c_a$, $VPD$) is positive for CRO, and
-negative for all other PFTs. This difference in mean ET response as
-compared to the ET response at mean environmental conditions is due to
-the non-linear nature of the response, in which negative responses are
-generally larger magnitude than positive responses (Figure
-\ref{idealized_sign}). Therefore, both the mean ET response as well as
-the ET response at mean environmental conditions matches our
-expectations from the theory (Section \ref{sign_term}), with the
-exception that CRO observations are shifted more towards a negative ET
-response than we expect.
-
-\begin{table}
-  \caption{Statistics of $\frac{\partial \; ET}{\partial \; VPD}$ as a
-    function of PFT.}
-  \centering
-  \begin{tabular}{l c c c}
-    \hline
-    PFT & $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ & $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)$ & fraction $\frac{\partial \; ET}{\partial \; VPD} < 0.$ \\
-    \hline
-    \input{stats.tex}
-    \hline
-  \end{tabular}
-  \label{stats}
-\end{table}
-
-Regarding the frequency of negative and positive ET response, all PFTs
-exhibit a decreasing ET response with increasing VPD (physiologically
-controlled, water conservative response) for the majority of
-observations. The more water conservative PFTs generally exhibit
-higher frequency of negative ET response, especially when one factors
-in the distribution of environmental VPD (e.g. SAV and WSA grow in
-more arid climates, Figure \ref{idealized_sign}). In general the bulk
-statistics match our theoretical expectations well, with the caveat
-that inclusion of uncertainty shifts crops towards a slightly more
-negative response to VPD, and shifts many of the other PFTs, which still
-exhibit a high frequency of negative ET response to VPD, towards more frequent
-occurrence of positive ET response than the theory and Figure
-\ref{idealized_sign} would suggest. The bulk statistics motivate a
-more thorough examination of the structure of uncertainty and more
-sophisticated validation of our theory's performance against
-observations.
-
-\subsection{Validation of theory at eddy-covariance sites}
-\label{testing}
-We now compare more sophisticated distributions of the observed
-response to our simplified theory (Section \ref{sign_term}). The
-observed distribution of the sign term, as compared to what the theory
-would predict, is provided in Figure \ref{test_sign}. Our goal was to
-capture the leading order behavior of the ET dependence on VPD. Given
-the assumptions we made, and the uncertainties of flux tower
-observations themselves, we expect a relatively large amount of noise
-when reproducing the derivatives of ET. However, the data largely
-reproduces our theoretical analysis.
-
-\begin{figure}
-  \centering
-  \centerline{\includegraphics[width=1.4\textwidth]{./test_sign.pdf}}
-  \caption{Comparison of the sign term with model uncertainty included
-    (box plots) to the sign term as calculated with simplifying
-    assumptions (blue line, as in Figure \ref{idealized_sign}). Each
-    box plot corresponds to 5\% of the data, and the 5-95\% range of
-    VPD is plotted.}
-  \label{test_sign}
-\end{figure}
-
-This is particularly true for DBF and MF; the theory matches the
-leading order behavior of the function when uncertainty is included,
-and the observations match the theory with the addition of noise. The
-VPD dependence, given by the slope parameter ($g_1$), follows the
-median values of each bin. Perhaps most importantly, the x-intercept,
-and thus VPD$_{crit}$, matches nearly exactly between the theory and
-the observations. Therefore the sign of the ET response to increases
-in VPD should be well matched, subject to the unavoidable constraints
-of noise, much of which comes from the observations themselves. The
-uncertainty is non negligible; there are many observations in each bin
-for which the the sign of observation is opposite the response
-predicted by the theory, but to leading order our theory matches the
-observations well.
-
-While CSH has a much different functional form of the sign term than
-DBF and MF, CSH observations also match our theory to leading order,
-albeit with a bit more variability as a function of VPD. Again, the
-VPD dependence mostly determined from the slope parameter ($g_1$)
-closely matches the medians in the observation bins. The
-VPD-independent, strongly negative response is also captured. For CSH,
-there is rare occurrence of observed positive ET response with VPD
-($\approx 6$\%), even with uncertainty, so the sign of the
-observations almost always matches the sign of the theory, which
-states that ET response should always be negative.
-
-Biases between the theory and observations are similar for ENF, WSA,
-CRO and SAV. At low VPD the theory and the observations match
-well. However, at high VPD (upper 10-20\% of observations) the theory
-is biased slightly towards positive response as compared to the
-observations (Figure \ref{test_sign}). For ENF this slight bias could
-be explained by a negative bias in $g_1$. However, for WSA, CRO, and
-SAV the observations in the highest VPD bins exhibit a downturn
-towards more negative response, which cannot be captured by the
-functional form of the sign term (Figure
-\ref{idealized_sign}). Therefore, to explain the observations one of
-the variables in the sign term must change at extreme VPD (upper
-10-20\%). The most likely candidate is $uWUE$, which we have assumed
-constant to meet our goal of developing intuition for leading order
-behavior, but might be expected to decrease at extremely low SWC. This extremely low SWC should be correlated with high VPD in very dry environments
-through sensible heating increases \citep{Gentine_2016}. The theory
-for EBF exhibits similar limitations as for ENF, WSA, and SAV, except
-a larger portion of observations are biased negative as compared to
-the theory ($\approx$ 35\%, Figure \ref{test_sign}). However, in
-general, and specifically for non-extreme VPD (VPD < $\approx$70-90th
-percentiles), the theory matches the observations for the tree and
-savannah plant types well.
-
-The theory for GRA suffers from similar, but much more severe,
-limitations as for CRO, WSA, and SAV. GRA observations are
-characterized by a consistent trend back towards negative ET response
-at higher VPD, which the functional form of our theory is incapable of
-accounting for. As compared to CRO, WSA and SAV, the divergence
-between the theory and the observations is far greater for GRA,
-biasing 40-50\% of observations at higher VPD. In addition for the
-potential for soil moisture to alter $uWUE$, there are other sources
-of plant heterogeneity specific to GRA (and some extent CRO) that may
-alter $uWUE$ (or $g_1$) or invalidate other assumptions made in the
-Methods section (Section \ref{methods}). We do not account for
-variability in plant height and surface roughness, or differences in
-C3 vs C4 photosynthesis and water strategies, which we might expect to
-vary substantially across sites, years, and season for GRA. These
-deficiencies could largely explain the inability of our theory to
-exactly match the observations in croplands. For example, a
-superposition of sites with C3 plants (at low environmental VPD) and
-C4 plants (at high environmental VPD) would explain the observed shift
-back towards negative ET response at high VPD when all sites are
-binned together, as in Figure \ref{test_sign}. We hypothesize that the
-theory would validate against observations much better if these
-sources of variability were accounted for, at a cost of increased
-complexity and analytical opacity.
-
-While the above discussion shows that our theory has some limitations
-when applied to some PFTs, especially for grasslands, it does well for
-DBF, MF and CSH PFTs, and captures the response at non-extreme VPD
-(VPD < $\approx$ 70th-90th percentiles) for
-ENF, EBF, CRO, WSA, and SAV. In general, the leading order behavior
-observed in the data is captured by the theory. Departures between our
-theory and observations, specifically at high VPD, could be explained
-and conceptualized with shifts in $g_1$ and/or $uWUE$ due to
-site-specific plant type variability (e.g. more arid-adapted ecosystems
-at more arid sites), or temporal variability for some environmental
-conditions (e.g. decreases in $uWUE$ at extremely low soil
-moisture). To focus on general behavior and develop intuition for
-PFT-scale response, we have ignored these sources of variability in
-the present analysis. However, any site-specific or temporal plant
-functional variability that can be conceptualized with shifts to $g_1$
-and $uWUE$ can be analyzed with our framework. This opens a door to future analyses in which plant behavior in anomalous conditions
-can be explained and analyzed using Equation \ref{et}.
-
-% It is useful to also compare our theory, the observed uncertainty in
-% the ET response, and PET, as PET is often used as a proxy for ET in
-% hydrometeorological studies and drought indices. Again, as in our
-% discussion of the theoretical results, the PET sensitivity to VPD
-% will be constant at about 3.5. In Figure \ref{test_sign} the ET
-% response to shifts in dryness is always far smaller than 3.5 in the
-% observations, even with all uncertainty included. So the PET
-% response to atmospheric dryness is outside the range of all
-% observable responses. This casts doubt on PET's usefulness for
-% capturing the leading order behavior of water budget and dryness
-% response at the land surface, as is done in several drought
-% indices. However our theory (Equation \ref{et}) could be used with
-% published uWUE and $g1$ data, as an alternative to PET to estimating
-% land response and dryness, and would more accurately account for the
-% response to atmospheric drying.
-
-\subsection{Observed ET response to VPD}
-
-Most of the results presented so far focus on the sign term, so now we
-turn to observations of ET response with the scaling term included
-(Figure \ref{data_scatter}). Until now, in the interest of developing
-leading order intuition for ET response, and to be conservative in our
-acknowledgment of model and observational error, we've considered
-$\sigma$ variability to be a measure of uncertainty. An alternative
-viewpoint is that $\sigma \cdot uWUE$ represents spatial and temporal
-variability of uWUE, which may be expected within bounds \citep[see
-Table \ref{pft}, also ][]{Zhou_2015}. This is a less conservative
-view; some of the $\sigma \cdot uWUE$ variability will be due to model
-and observational error, so by viewing $\sigma \cdot uWUE$ as
-\textit{real} variability we run the risk of mistaking noise for
-signal. However, the advantage of this viewpoint is that, from
-\citet{Zhou_2014}, we have very high confidence that uWUE fit at the
-hourly timescale (as we do with $\sigma \cdot uWUE$) correctly
-captures the relationship between ET, GPP, and VPD, and that our form
-of PM introduces minimal error with its use of uWUE.
-
-\begin{figure}
-  \centering
-  \centerline{\includegraphics[width=1.4\textwidth]{./data_scatter.png}}
-  \caption{Scatter plots of observed
-    $\frac{\partial \; ET}{\partial \; VPD}$, including $\sigma$
-    variability, as a function of PFT, temperature and VPD. Please
-    note differences in the colorbar scale.}
-  \label{data_scatter}
-\end{figure}
-
-Therefore, with the caveat that some of the signal presented in Figure
-\ref{data_scatter} may in fact be noise, we can interpret the observed
-distribution of ET response to VPD. In general, the observed response
-matches the intuitive theory. ET response to VPD shifts towards
-positive values as VPD increases (atmospheric demand dominating). CRO
-exhibit the highest occurrence of positive ET response, and the
-observations confirm that CRO are the most water intensive. CSH are
-the most water conservative, with a strong negative response. DBF,
-EBF, ENF, MF, SAV and WSA are also water conservative, but show some
-occurrence of positive ET response to VPD, particularly at higher
-observed VPD, as the theory predicts. GRA, while generally water
-conservative, does not match the theory well, with increasing frequency of
-\textit{negative} response at high VPD. This is as expected, given
-previous discussion on how we might expect more inter-site and inter-year variability for GRA (Section \ref{testing}).
-
-Figure \ref{data_scatter} also includes the impact of the scaling
-term. For a given VPD, the magnitude of ET response does not vary
-strongly with temperature, confirming that any impact of the scaling
-term on the magnitude of the response is primarily due to changes in
-aerodynamic conductance. Intuitively this is reasonable; aerodynamic
-conductance will control how dominant balances at the land surface are
-communicated to the boundary layer. While $\Delta$ controls the
-efficiency of energy conversion to latent heat, it appears this is a
-second order term, relative to $g_a$, for scaling ET response.
-
-As with Figure \ref{test_sign}, Figure \ref{data_scatter} matches our
-expectation based on simplified theory. The sign term is most strongly
-scaled by $g_a$, and in general the occurrence of positive ET response
-increases as VPD increases. The willingness with which a given plant type
-evolved to use water dictates the occurrence of positive versus
-negative ET response. Water conservative ecosystems are highly effective
-at mitigating the effects of atmospheric demand, and can store water
-for later use by reducing ET in response to increasing VPD.
-
-\subsection{Limitations of theory - very dry soil moisture conditions}
-\label{swc_section}
-In formulating our theory with Penman Monteith, we implicitly did not
-account for very dry soil moisture conditions. For the majority of
-environmental conditions observed at the eddy covariance sites used
-here, soil conditions were not extremely dry so that we could assume a
-constant uWUE and $g_1$. We posit that ecosystems will generally
-optimize to host plants living in conditions which they evolved
-for. However, in extreme conditions and drought scenarios soil water
-content (SWC) could become the limiting factor for ET response to VPD,
-which our theory does not account for. In addition, low soil moisture
-conditions themselves increase VPD through land-atmosphere feedback
-\citep[][]{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006,
-  Salvucci_2013, Gentine_2016, Berg_2016}.
-
-Within our framework, any systematic bias due to the
-failure to account for SWC's effects in very dry conditions should
-manifest itself in a functional relationship between $\sigma$, our
-uncertainty measure, and SWC, which is observed at the FLUXNET
-sites. Examining the relationship between $\sigma$ and SWC will test
-to what extent our theory breaks down in very dry soil moisture
-conditions.
-
-If we again view $\sigma \cdot uWUE$ as, in addition to a measure of
-uncertainty, a short time-scale observation of uWUE, we would expect
-$\sigma \cdot uWUE$ to decrease at low SWC. If $\sigma \cdot uWUE$ is
-a very strong function of SWC, then our theory should be conditioned
-more strongly on well watered soil conditions. If $\sigma \cdot uWUE$
-is weakly a function of SWC, then our theory is more universal and
-independent of soil moisture conditions.
-
-Indeed, for all PFTs, there is some slight dependence of
-$\sigma \cdot uWUE$ on SWC, especially at low SWC (Figure
-\ref{swc_boxplot}). The portion of observations for which our theory
-is biased by SWC-limitations varies by PFT, due to the nonlinear
-threshold at which soil moisture availability limits plant
-function. For CRO, MF and DBF, soil water content only matters for
-about the lowest 5\% of observations (each box is 5\% of observations
-in Figure \ref{swc_boxplot}). And, CRO, relative to DBF, has a very
-weak dependence on soil moisture, which is reflective of the high
-likelihood that CRO sites are optimally irrigated and not water
-stressed, suggesting that observed departures from the theory for CRO
-are due to other factors (i.e. different photosynthesis pathways; C3/C4) than soil
-moisture. For ENF, EBF, SAV, and WSA, systematic SWC-induced biases in
-$\sigma \cdot uWUE$ emerge in about the lowest 20-30\% of SWC
-conditions, although large variability in the SWC-$(\sigma \cdot uWUE)$
-relationship hampers interpretation for WSA. CSH presents a special
-case: the limited number of sites (2) preclude comment on the
-exact PFT-wide relationship.
-
-In contrast to ENF, CRO, and DBF, for GRA the relationship between
-$\sigma \cdot uWUE$ and SWC is more linear and affects a greater
-portion of observations; about 60\% of observations. Clearly, for GRA
-soil water frequently impacts plant function and alters ET response,
-and our theory is limited for the majority of environmental
-conditions. It is therefore not surprising that our theory tested
-poorly against the data for GRA, relative to to the other PFTs
-(Section \ref{testing}). For all other PFTs occurrence of soil
-moisture impacts was rare enough to not manifest itself in bulk
-statistics and figures.
-
-\begin{figure}
-  \centering
-  \centerline{\includegraphics[width=1.4\textwidth]{./swc_boxplot.pdf}}
-  \caption{$\sigma \cdot uWUE$ as a function of PFT and SWC. Each box
-    plot represents 5\% of all observations. Note that the highest
-    10\% of SWC observations are excluded to better resolve variability in the much more narrow bins at lower SWC (e.g. SWC has long tails at high values).}
-  \label{swc_boxplot}
-\end{figure}
-
-The observed dependence of $\sigma \cdot uWUE$ on SWC for GRA would
-explain the deficiencies of our theory compared to the observations in
-Figure \ref{idealized_sign}, specifically the trend back towards
-negative soil moisture at high VPD. Aforementioned feedbacks between
-the land surface and the atmosphere, which are not accounted for due
-to our focus on the one-way response of the land surface to
-atmospheric conditions, would cause high VPD to be correlated with low
-SWC \citep[][]{Gentine_2016, Berg_2016}. So, at high VPD observations
-of low SWC are more likely, and this low SWC causes a lower uWUE
-(Figure \ref{swc_boxplot}). The lower uWUE at low SWC/high VPD then
-leads to the observed downturn towards decreasing ET with increasing
-VPD, and the deviation between our theory (based on a constant uWUE
-assumption) and observations. It is also perhaps not coincidental that
-the portion of $\sigma \cdot uWUE$ affected by low soil moisture
-observations is similar to the portion of observations that do not
-match our theory at high VPD. Indeed, this may suggest that for all PFTs
-except for GRA, coupling and feedbacks between SWC, VPD and plant
-function are relatively rare.
-Future research will explore these relatively rare feedbacks in extreme conditions, which due to analytical intractability will require more opaque numerical analysis of many more complex processes, including boundary layer growth and state and their relationship to surface layer coupling and free-tropospheric lapse rates and humidity.
-
-To summarize, our theory is limited by its inability to account for
-soil water impacts on land surface response, and feedbacks between SWC
-and VPD. Fortunately, for most PFTs SWC's effect on ET is relatively
-rare (<30\% of observations) and does not manifest itself in the
-majority of observations and bulk statistics. However for GRA, SWC decreases water use efficiency for the majority of
-the observations. Soil moisture effects explain the deficiencies of
-our theory in Section \ref{testing}, particularly for GRA. By
-conceptualizing SWC effects as a change in $uWUE$ (and/or $g_1$), it
-will be possible for future analysis to explore the importance of soil
-moisture on plant response to VPD, and feedbacks between plant
-function, SWC, and VPD.
-
-\subsection{Functional form of ET dependence on VPD and its relation
-  to VPD dependence}
-\label{functional_form}
-
-The theory described in Section \ref{sign_term} indicates that for a
-given $uWUE$ and $g_1$, the ET dependence on VPD should be concave
-upward, which is confirmed by eddy covariance data across most PFTs. In other words, there should be some local minimum in ET at a
-critical VPD$_{crit}$, assuming the scaling and plant terms  (e.g. aerodynamic
-conductance, $\Delta$, $g_1$ and $uWUE$) are held fixed. This result warrants further
-investigation, because to our knowledge no one has derived the
-theoretical ecosystem-scale relationship between ET and VPD while
-controlling for other environmental conditions. In particular, from
-personal communication, there is an apparent lack of consensus over
-whether the shape of the ET-VPD curve should be concave upward (our
-result) or concave downward in the absence of dramatic water stress. Given that understanding the ET-VPD
-relationship of the one-way plant response is fundamental to
-hypothesizing about any feedbacks between the land surface and the
-atmosphere, we analyze why our derived ET-VPD relationship is concave
-upward, particularly with respect to the exponent of VPD dependence in
-$uWUE$ and the Medlyn unified stomatal conductance model.
-
-There is a theoretical basis for the square root VPD dependence in
-both the stomatal conductance model and $uWUE$ based on the assumption
-that stomata behave to maximize carbon gain while minimizing water
-loss, which observations also generally support \citep{Lloyd_1991,
-  MEDLYN_2011, Lin_2015, Zhou_2014, Zhou_2015, Medlyn_2017}. However,
-some purely empirical results that fit the exponent of the VPD
-dependence to data have shown that it may vary slightly from 1/2,
-suggesting that stomata, as well as ecosystem-scale quantities based
-on stomata theory, may not always function optimally \citep{Zhou_2015,
-  Lin_2018}. Specifically with regards to $uWUE$, one would not expect
-that this ecosystem scale WUE quantity will respond to VPD exactly
-analogously to stomata. Direct soil evaporation's contributions to ET
-should shift the exponent of the VPD dependence, especially at
-conditions of low GPP when we would expect a systematically larger
-portion of direct soil evaporation contributions to ET, because we
-would also expect lower amounts of transpiration at low GPP. \citet{Zhou_2015}'s
-results corroborate this: they found a mean empirically fit
-exponential VPD dependence of 0.55, varying slightly from the
-theoretically optimal value of of 0.5 for AmeriFlux
-sites. Interpreting \citet{Lin_2018}'s results, which also show
-variance in the empirical exponent of the VPD dependence of the
-stomatal conductance model, is more difficult as \citet{Lin_2018} do
-not handle GPP/A dependence of stomatal conductance in a directly
-analogous manner to the optimal theory in \citet{MEDLYN_2011} and
-\citet{Medlyn_2017}. Regardless, given that these recent results on
-the relationship between VPD, GPP, and ET \citep{MEDLYN_2011,
-  Zhou_2014, Zhou_2015, Medlyn_2017} form the backbone of our analysis
-and are what allowed us to derive an explicit ET expression for the
-first time (Equation \ref{et}), we will analyze if and how assumptions
-about the exponent of the VPD dependence impacts the shape of the
-ET-VPD dependence.
-This analysis is also important to understand whether the choice of stomatal conductance model alters the fundamental behavior of the ET-VPD relationships, as many commonly used models  utilize a VPD exponent other than the 1/2 suggested by optimal theory \citep[e.g.
-][ which uses an exponent of 1]{Leuning_1990}.
-
-By introducing $n$ and $m$ we can free our stomatal
-conductance model from assumptions about VPD dependence:
-\begin{linenomath*}
-  \begin{equation}
-    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_*}{VPD^m}\right) \frac{*WUE \; ET}{c_a \; VPD^n},
-    \label{m_n}
-  \end{equation}
-\end{linenomath*}
-where:
-\[*WUE = \frac{GPP}{ET}VPD^n,\] and $g_*$ is a generic slope parameter
-of units $VPD^m$. To determine how the exponent $n$ and $m$ alter the
-shape of the ET-VPD dependence we find the roots of the second
-derivative of ET, using Equation \ref{m_n} for stomatal conductance
-($g_s$), with respect to VPD: \input{d2_solutions.tex} With this
-result we have defined the family of curves separating concave up from
-concave down ET solutions (Figure \ref{concave}). These curves are
-only functions of the exponent of the VPD dependence and a quantity we
-call non-dimensional VPD ($VPD^m/g*$). Several important
-relations reveal themselves from Equation \ref{curves}:
-
-\begin{figure}
-  \centering
-  \centerline{\includegraphics[width=1\textwidth]{./concave.pdf}}
-  \caption{ Solutions corresponding to inflection points between
-    concave up and concave down ET-VPD curves (Equation \ref{curves})
-    for three specific scenarios. Solutions are defined in terms of a
-    non-dimensional VPD ($VPD^m/g_*$), but to aide physical
-    interpretation the horizontal axis is additionally provided in terms of dimensionalized VPD assuming $m=1/2$ and
-    $g_*=110\; Pa^{1/2}$ (average of all PFT $g_1$). The vertical axis
-    has a different interpretation depending on the solution
-    curve. For the blue line ($m$ varying), it corresponds to $m$,
-    for the orange line ($n$ varying) it corresponds to $n$, and for
-    the green line it corresponds to the value of both $n$ and $m$
-    ($n=m$). Regions of the parameter space that correspond to
-    concave up and concave down results are labeled.}
-  \label{concave}
-\end{figure}
-
-\begin{itemize}
-  \item For optimal behavior (n, m = 1/2) the ET-VPD curve will be
-    concave up regardless of the magnitude of the plant constants
-    $g_1$ and $uWUE$. Therefore, the general concave up nature of our
-    results, given an assumption of optimal behavior, is insensitive
-    to plant type.
-  \item For all physically possible exponents of VPD dependence ($n,
-    m$), whether the solution is concave up or concave down does not
-    depend on $uWUE$.
-  \item In general, increasing the exponent of VPD dependence
-    increases the likelihood of a concave up result. Additionally, as
-    the exponent of VPD dependence increases from the optimum value of
-    1/2, whether the curve is concave upward or concave downward
-    becomes a function of the plant specific slope parameter $g_*$,
-    through non-dimensional VPD ($VPD^m/g_*$). Because the exponent of
-    the VPD dependencies is capable of altering the fundamental shape
-    of ET-VPD dependence, future research investment in understanding
-    the exact VPD dependence of stomatal conductance, and further
-    reconciliation of empirical and theoretical stomatal and
-    ecosystem behavior should be prioritized.
-\end{itemize}
-While it is possible that in the future some other form of
-VPD dependence is derived, at present \cite{MEDLYN_2011} and
-\cite{Zhou_2014} firmly established n=m=1/2 as the most likely
-candidate given current theory and empirical data. Additionally, we
-argue that a concave up result matches physical intuition more than a
-concave down result. Plants must maintain nutrient and sugar transport
-through the phloem and xylem. To accomplish this, stomata must remain
-slightly open \citep{De_2013, Nikinmaa_2013, Ryan_2014}. Furthermore,
-even if complete stomatal closure were possible, cuticular water loss
-and [at the ecosystem-scale] direct soil evaporation are still sources
-of ET which increase with VPD, independent of stomatal
-closure. Therefore, in the limit as VPD becomes large and we assume
-plants are exercising all strategies to reduce ET, any further increase
-in VPD should result in an increase in ET through cuticular water loss
-and/or direct soil evaporation. This inevitable transition from
-conditions when stomata respond strongly to VPD to conditions when
-stomata response is asymptoting towards full closure would cause a
-concave up ET-VPD curve, which is matched by the theory. In short,
-plant response becomes more limited as VPD increases, while
-atmospheric demand monotonically increases with VPD, leading to the
- result that atmospheric demand dominates plant response when
-atmospheric demand is high.
-
-This analysis allows us to understand the theoretical shape of the ET
-response to VPD with environmental conditions held
-fixed. Accomplishing this with purely statistical
-methods applied to flux observations would be very difficult, given
-the relatively fast time scale of plant response and the
-non-stationarity of [solar forced] environmental conditions over the
-relatively coarse (half hourly) flux estimates (which is required to
-obtain robust eddy-covariance statistics). Our results on the shape of
-the ET-VPD curve with environmental conditions held
-fixed can be built upon with future work examining how changes in VPD
-and environmental conditions (e.g. soil water storage) feedback upon
-one another. In the soil water storage example, over very long time
-scales extremely high VPD perturbations coupled with no precipitation
-could result in decreases in soil water storage such that water
-becomes limiting. This could be represented by an extension of our
-framework in which uWUE is allowed to decrease with decreasing SWC, as
-observed in Section \ref{swc_section}. Here, we focus our results by
-assuming constant PFT-wide conditions to build baseline intuition for
-ET-VPD dependence. For most PFTs, the theory with plant function held
-fixed matches the leading order behavior of the observations where
-plant function varies (Section \ref{testing}).
-
-\section{Conclusions}
-
-We derived a new form of Penman Monteith using the concept of
-semi-empirical optimal stomatal regulation \citep{Lin_2015,
-  MEDLYN_2011} and near constant uWUE \citep{Zhou_2015} to remove the
-implicit dependence of stomatal conductance on GPP and ET. With our
-new form of Penman Monteith we developed a theory for when an
-ecosystem will tend to reduce or increase ET with increasing VPD,
-which we evaluated against a range of eddy-covariance data spanning
-different climates and plant functional types. The goal was to capture
-the leading order behavior of the system to gain some intrinsic
-knowledge for its behavior. This intuition can be used to disentangle
-land atmosphere feedbacks in more complicated scenarios, and will aid
-interpretation of observations and sophisticated models.
-
-Our theory suggests that for a majority of environmental conditions,
-plants will tend to conserve water and reduce ET with increasing
-VPD. Stomatal regulation and plant physiological response strongly
-regulate ET, and this regulation varies by PFT. CROs are the least
-water conservative, while DBF, EBF, SAV, GRA, MF, WSA, ENF and CSH are
-progressively more water conservative (more likely to reduce ET in
-response to increasing VPD). SAV and WSA exhibit positive ET response
-to VPD not necessarily because of poor water conservation strategies
-relative to other PFTs, but because of greater occurrence of high
-atmospheric demand (VPD) relative to other PFTs. Observations of ET
-response to VPD exhibit the same general behavior as the theory, with
-ET response becoming more positive (atmospheric demand dominating) as
-environmental VPD increases within a PFT, and more negative for PFTs
-that are adapted to arid conditions and prioritize water conservation
-over primary production.
-
-The consequences of our theory our twofold. First, the ET response to
-VPD is always far more negative than it would be if only atmospheric
-demand is considered and plant physiological response is ignored (as
-with PET). Therefore, any drought indices or analysis using PET will
-not capture how vegetated surfaces respond and react to increases in
-atmospheric demand. However, the new uWUE-version of Penman Monteith
-we derived (Equation \ref{et}) could be used as a complement to PET in
-drought indices. One could calculate and compare indices when plant
-response is ignored (PET) or included (Equation \ref{et}).
-
-Lastly, our paper builds important intuition for how plants will
-respond to VPD perturbations. We show that given optimal stomatal
-function and fixed environmental conditions, the ET-VPD dependence is
-theoretically concave upward, with ET increasing with increasing VPD as
-VPD increases past some critical value (Table \ref{vpd_crit}). However
-future research should focus on fully understanding the form of
-stomatal VPD dependence, as this result is sensitive to the exponent of
-VPD dependence, which we currently believe is 1/2 \citep{MEDLYN_2011,
-  Zhou_2014}. Our results are applicable to understanding the impact
-of  expected increases in VPD
-induced by global change.  Plant physiological responses to direct
-CO$_2$ effects \citep[e.g.,][]{Swann_2016, Lemordant_2018} receives more attention
-than physiological response to indirect effects like increased
-VPD. Here, we provide broad PFT-focused results showing a likely
-decrease in ET in response to positive VPD perturbations (atmospheric
-drying), which is consistent with recent observational analysis
-\citep[e.g.,][]{Rigden_2017}. Feedbacks between the land and the
-atmosphere may alter the net response to a long-timescale global VPD perturbation,
-but our focus on the one way plant response to a VPD perturbation in
-the atmospheric boundary layer is an important first step to
-disentangling such feedbacks, both in observations and model
-simulations of the present and future. By removing Penman Monteith's
-dependence on implicit relationships between GPP, VPD, and ET, we
-allow for explicit, accurate future analysis of plant-VPD feedbacks
-in the atmospheric boundary layer (Equation \ref{et}). Our approach
-can be extended to examine varying plant response to more nuanced
-consideration of plant type and climate. Any plant physiological
-heterogeneity or feedback that can be conceptualized with shifts in $g_1$
-\citep[e.g.][]{Lin_2015, Medlyn_2017} and/or uWUE
-\citep[e.g.][]{Zhou_2014} are representable within our framework, which
-opens the door for a hierarchy of more sophisticated climate- and
-plant-specific analysis of ET sensitivity to environmental variables
-(including VPD). We argue that such simplified conceptual frameworks
-are critical tools for disentangling land-atmosphere feedbacks at
-various scales, from diurnal to seasonal and beyond, and to
-characterize ET response in a warmer, atmospherically drier, and
-enriched CO$_2$ world.
-
-\acknowledgments This work used eddy covariance data acquired and
-shared by the FLUXNET community, including these networks: AmeriFlux,
-AfriFlux, AsiaFlux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont,
-ChinaFlux, Fluxnet-Canada, GreenGrass, ICOS, KoFlux, LBA, NECC,
-OzFlux-TERN, TCOS-Siberia, and USCCC. The ERA-Interim reanalysis data
-are provided by ECMWF and processed by LSCE. The FLUXNET eddy
-covariance data processing and harmonization was carried out by the
-European Fluxes Database Cluster, AmeriFlux Management Project, and
-Fluxdata project of FLUXNET, with the support of CDIAC and ICOS
-Ecosystem Thematic Center, and the OzFlux, ChinaFlux and AsiaFlux
-offices. This material is based upon work supported by the National
-Science Foundation Graduate Research Fellowship under Grant No. DGE
-16-44869. All code and data used in this analysis, including those
-used to generate the figures and tables, are publicly available at
-https://github.com/massma/climate\_et.
-
-\cite{Zhou_2016}
-
-\bibliography{references.bib}
-
-\end{document}
+\input{body.tex}

### 2018-05-25 doc/paper/vpd_et_paper_arxiv.tex
refactored tex seperate body, removed unknown external dependency on my util package in python modules, and set the stage for removal of tier 2 sites and addition of citation table <credit Trevor Keenan for catch>

--- 2018-05-14 doc/paper/vpd_et_paper.tex
@@
 
-\section{Introduction}
-
-Vapor pressure deficit (VPD) is expected to rise over continents in
-the future due to the combination of increased temperature and,
-depending on region, decreased relative humidity
-\citep{Byrne_2013}. Increases in VPD increase the atmospheric demand
-for evapotranspirated water \citep{Penman_1948, Monteith_1965}, but
-also stress plant stomata \citep{Leuning_1990, MEDLYN_2011}.
-
-The opposing effects of increased atmospheric demand and higher
-stomatal stress lead to two possible perspectives for how
-evapotranspiration (ET) responds to shifts in VPD. The first, a
-hydrometeorological perspective, is that higher VPD increases
-atmospheric demand for water from the land surface, and this drives an
-increase in evapotranspiration (ET). This perspective is particularly
-relevant because potential evapotranspiration (PET), which is used in
-many drought indices and hydrometeorological studies
-\citep[e.g.,][]{Heim_2002, Scheff_2015}, typically only quantifies
-changes in atmospheric demand and fails to account for ecosystem
-response \citep{Swann_2016}. In reality, plants' stomata have evolved
-to optimally regulate the exchange of water and carbon, and tend to
-partially close in response to increased atmospheric dryness
-\citep{Farquhar_1978, Ball_1987, Leuning_1990, MEDLYN_2011}. This
-leads to a plant physiology perspective, in which an increase in VPD,
-particularly in well-watered soil conditions, may actually correspond
-to a decrease in ET because of stomatal closure
-\citep[e.g.][]{Rigden_2017}.  In other words, the question ``When does
-VPD drive or reduce ET?'' can be related to whether plant regulation
-or atmospheric demand dominates ET response.
-
-The ET response to changes in VPD alters water partitioning between
-the soil and atmosphere. If ecosystem plant response reduces ET with
-atmospheric drying then soil moisture will be better conserved. This
-would seem a sensible evolutionary strategy to cope with aridity. If
-stomata were fully passive \citep [similar to soil pores,
-e.g. ][]{Or_2013}, increased atmospheric aridity would strongly reduce
-soil moisture \citep{Berg_2017}. In turn, this would further increase
-aridity as low soil moisture levels increase the Bowen ratio, and
-cause increased temperature and atmospheric drying
-\citep[][]{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006,
-  Salvucci_2013, Gentine_2016, Berg_2016}. This however would not seem
-to be a sensible strategy for plants from an evolutionary
-standpoint. As a counterpoint, one may argue that increases in ET with
-increasing VPD could increase the likelihood of precipitation
-\citep[e.g.,][]{Findell_2011}. However, increases in ET do not always
-guarantee an increased likelihood of precipitation, which depending on
-environmental conditions could cause a decrease in the likelihood of
-precipitation \citep[][]{Gentine_2013}. Furthermore, any increases in
-precipitation are likely to be non-local, such that plants giving up
-water to the atmosphere are not guaranteed to reap the benefits of
-water returned from the atmosphere to the soil. Using this subjective
-logic, from an ecosystem evolutionary perspective, water stored in
-soil seems to be worth much more than the chance of water returned as
-precipitation.
-
-We can use intuition about plant water conservation strategy to
-hypothesize about ET response to changes in VPD. Plants and ecosystems
-that evolved to conserve water, such as arid shrubs or savannah,
-should be more likely to reduce ET with increasing VPD, and plants
-that have evolved or have been engineered to care little about water,
-such as crops, will be more likely to increase ET with increasing
-VPD. Atmospheric conditions must matter as well. At the ecosystem
-scale, there are limits to plant water conservation strategies. As
-atmospheric demand for water (VPD) increases, ecosystems should begin
-to reach their water conservation limits and might not be able to
-entirely limit ET flux to the atmosphere. At this stage any further
-increase in VPD will most likely drive a (limited) increase in ET,
-because the increase in atmospheric demand for water overwhelms the
-limited plant response to conserve water.
-
-The objective of the present manuscript is to use reasonable
-approximations as a tool to develop intuition for plant response to
-atmospheric drying and evaluate the VPD dependence of ET. This
-intuition will aid interpretation of observations, full complexity
-models, and facilitate the disentanglement of complex land-atmosphere
-feedbacks. In the past, similar approaches were used to understand
-interactions between stomatal conductance, evapotranspiration and the
-environment \citep[e.g.,][]{Jarvis_1986, Mcnaughton_1991}. However, at
-the time researchers' understanding of the form of VPD's effect on
-plant physiology was limited, so they could not explore the
-sensitivity of ET to VPD, including VPD's effect on stomatal
-conductance and plant function.
-
-Recent results have drastically improved our understanding of VPD's
-impact on physiology, especially at the leaf
-level. \citet{MEDLYN_2011} developed a model for leaf-scale stomatal
-conductance ($g_s$), including VPD response, by combining an optimal
-photosynthesis theory \citep{Cowan_1977} with an
-empirical approach, and extended use of this model to the ecosystem
-scale in \citet{Medlyn_2017}. Additionally, \citet{Zhou_2014}
-demonstrated that a quantity underlying water use efficiency
-$\left(uWUE = \frac{GPP\; \sqrt{VPD}}{ET}\right)$ properly captures a
-constant relationship between GPP, ET, and VPD over a diurnal cycle at
-the ecosystem scale. uWUE is also remarkably well conserved in the
-growing season across space and time, within a PFT
-\citep{Zhou_2015}. These results provide novel tools to examine how ET
-varies in response to VPD, while explicitly including VPD's effect on
-stomatal conductance for the first time. In this manuscript, we
-leverage these tools to derive the theoretical one-way response of ET
-to VPD with other environmental variables properly controlled for,
-i.e. we develop a framework for evaluating the partial derivative of
-ET with respect to VPD. Our theory is validated and tested at multiple
-eddy-covariance stations spanning various climates and plant
-functional types.
-
-\section{Materials and Methods}
-\subsection{Data}
-\label{data}
-We use both meteorological and eddy-covariance data from the
-FLUXNET2015 database (data available at \sloppy
-\url{https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/} \sloppy),
-including all sites with at least four years of data and observations
-of the variables described in the methods section (Section
-\ref{methods}). Seventy-three sites met these requirements, and were
-grouped into nine plant functional types (PFT) according to the
-International Geosphere-Biosphere Programme vegetation classification
-scheme \citep{Loveland_1999}: cropland (CRO), grass (GRA), deciduous
-broadleaf forest (DBF), evergreen broadleaf forest (EBF), evergreen
-needleleaf forest (ENF), mixed forest (MF), closed shrub (CSH),
-savannah (SAV), and woody savannah (WSA) (site locations in Figure
-\ref{map_fig}).
-
-\begin{figure}[h]
-  \centering \includegraphics[trim={0 3cm 0 3cm}, clip]{./map.pdf}
-  \caption{Plant functional type and location of FLUXNET2015 sites
-    used in this analysis.}
-  \label{map_fig}
-\end{figure}
-
-The purpose of this study is to examine ecosystem response to
-atmospheric drying, focusing on the growing season. To accomplish
-this, we filter and quality control the data using a similar procedure
-as \cite{Zhou_2015}:
-\begin{itemize}
-\item Only measured or highest (``good'') quality gapfilled data,
-  according to quality control flags, are used.
-\item To isolate the growing season, we only use days in which the
-  average Gross Primary Productivity (GPP) exceeds 10\% of the
-  observed 95th percentile of GPP for a given site. GPP is calculated
-  using the nighttime respiration partitioning method.
-\item We remove days with rain and the day following to avoid issues
-  with rain interception and sensor saturation at high relative
-  humidity (\cite{Medlyn_2017}).
-\end{itemize}
-Additionally, as in \citet{Lin_2018}, we restrict data to the daytime,
-which is identified when downwelling shortwave radiation is greater
-than 50 W m$^{-2}$ and sensible heat flux is greater than 5 W
-m$^{-2}$. To reduce the chance of sensor saturation at high relative
-humidity, we remove all time steps for which VPD is less than .01 kPa,
-and to reduce errors at low windspeeds we remove all periods with wind
-magnitudes less than 0.5 m s$^{-1}$. Timesteps with negative observed
-GPP or ET are also removed, and we aggregate half hourly data to
-hourly averages to reduce noise \citep{Lin_2018}. After these quality
-control procedures, 434,765 upscaled hourly observations remain.
-
-\subsection{Methods}
-\label{methods}
-The Penman-Monteith equation \citep [hereafter PM,][]{Penman_1948,
-  Monteith_1965} estimates ET as a function of observable atmospheric
-variables and surface conductances:
-% \begin{linenomath*}
-  \begin{equation}
-    \label{orig_pen}
-    ET = \frac{\Delta R_{net} + g_a \rho_a c_p VPD}{\Delta + \gamma(1 + \frac{g_a}{g_s})},
-  \end{equation}
-% \end{linenomath*}
-where $\Delta$ is the change in saturation vapor pressure with
-temperature, given by Clausius-Clapeyron ($\frac{d \; e_s}{d \; T}$),
-$R_{net}$ is the net radiation minus ground heat flux, $g_a$ is
-aerodynamic conductance, $\rho_a$ is air density, $c_p$ is specific
-heat of air at constant pressure, $\gamma$ is the psychometric
-constant, and $g_s$ is the stomatal conductance (Table
-\ref{definitions}).
-
-\begin{table}
-  \caption{Definition of symbols and variables, with citation for how
-    values are calculated, if applicable.}
-  \label{definitions}
-  \centering \footnotesize
-  \begin{tabular}{l c c c}
-    \hline
-    Variable & Description & Units & Citation \\
-    \hline
-    $e_s$  & saturation vapor pressure & Pa  & - \\ 
-    $T$  & temperature  & K & - \\
-    $P$  & pressure & Pa  & - \\
-    $\Delta$  & $\frac{\partial e_s}{\partial T}$ & Pa K$^{-1}$ & - \\
-    $R_{net}$  & net radiation at land surface minus ground heat flux & W m$^{-2}$   & - \\
-    $g_a$  & aerodynamic conductance & m s$^{-1}$  & \citet{Shuttleworth_2012} \\
-    $\rho_a$  & air density & kg m$^{-3}$  & - \\
-    $c_p$  & specific heat capacity of air at constant pressure & J K$^{-1}$ kg$^{-1}$ & - \\
-    $VPD$  & vapor pressure deficit & Pa  & - \\
-    $\gamma$  & psychometric constant & Pa K$^{-1}$   & - \\
-    $g_{s-leaf}$  & leaf-scale stomatal conductance & m s$^{-1}$  & \citet{MEDLYN_2011} \\
-    $g_{s}$  &  stomatal conductance & m s$^{-1}$
-                                   & \citet{Medlyn_2017} \\
-    $g_{1-leaf}$  & leaf-scale slope parameter & Pa$^{0.5}$
-                                   & \citet{MEDLYN_2011} \\
-    $g_{1}$  & ecosystem-scale slope parameter & Pa$^{0.5}$ & \citet{Medlyn_2017} \\
-    $R$ & universal gas constant & J mol$^{-1}$ K$^{-1}$ & - \\
-    $R_{air}$ & gas constant of air & J  K$^{-1}$ kg$^{-1}$ & - \\
-    $\sigma$ & uncertainty parameter & -& - \\
-    $c_a$ & CO$_2$ concentration & $\mu$ mol CO$_2$ mol$^{-1}$ air& - \\
-    $\lambda$ & marginal water cost of leaf carbon & mol H$_2$O mol$^{-1}$ CO$_2$ & - \\
-    $\Gamma$ & CO$_2$ compensation point & - & - \\
-    $\Gamma^*$ & CO$_2$ compensation point without dark respiration & - & - \\
-    \hline
-  \end{tabular}
-\end{table}
-
-\citet{MEDLYN_2011} developed a model for stomatal conductance ($g_s$)
-by combining an optimal photosynthesis theory \citep{Cowan_1977} with an empirical approach, which describes the
-dependence of $g_s$ to VPD. This resulted in the following model for
-leaf-scale stomatal conductance:
-
-% \begin{linenomath*}
-  \begin{equation}
-    g_{s-leaf} = g_0 + 1.6 \left(1 +
-      \frac{g_{1-leaf}}{\sqrt{VPD}}\right) \frac{A}{c_a},
-    \label{leaf_medlyn}
-  \end{equation}
-% \end{linenomath*}
-where $g_{1-leaf}$ is a leaf-scale ``slope'' parameter, A is the net
-CO$_2$ assimilation rate, and $c_a$ is the atmospheric CO$_2$
-concentration at the leaf surface. \cite{MEDLYN_2011} relate the slope
-parameter ($g_{1-leaf}$) to physical parameters as:
-% \begin{linenomath*}
-  \label{slope}
-  \begin{equation}
-    g_{1-leaf} = \sqrt{\frac{3 \, \Gamma^* \, \lambda}{1.6}},\footnote{Note this expression has units of of (mmol mol$^{-1}$)$^{1/2}$, but this can be converted to Pa$^{1/2}$ using the ideal gas law.}
-  \end{equation}
-% \end{linenomath*}
-
-where $\Gamma^*$ is the CO$_2$ compensation point for photosynthesis
-(without dark respiration), and $\lambda$ is the marginal water cost
-of leaf carbon
-($\frac{\partial \; \text{transpiration}}{\partial \; A}$). So,
-$g_{1-leaf}$ is a leaf-scale term reflecting the trade-off of water for
-carbon uptake. The higher $g_{1-leaf}$, the more open the stomata and
-the more they release water in exchange for carbon.
-
-The Medlyn model for stomatal conductance has been shown to behave
-very well across PFTs \citep[][]{Lin_2015}, and has been successfully
-adopted to ecosystem scale analysis in \citet{Medlyn_2017}. In units
-of m s$^{-1}$, the ecosystem scale stomatal conductance is:
-
-% \begin{linenomath*}
-  \begin{equation}
-    g_s = \frac{R \,T}{P} \; 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{GPP}{c_a},
-    \label{medlyn}
-  \end{equation}
-% \end{linenomath*}
-
-where GPP is the ecosystem scale gross primary production, and $g_1$
-is an ecosystem scale analogue to $g_{1-leaf}$. We solve for $g_1$
-following \citet{Medlyn_2017} (Equation 5), and take the median $g_1$
-value to be representative of each PFT (Table \ref{pft}) instead of the mean to avoid extra weighting of rare outliers.
-
-\begin{table}
-  \caption{Plant functional types, their abbreviation, calculated
-    Medlyn coefficient, and calculated uWUE. uWUE from
-    \citet{Zhou_2015}, including the observed standard deviation, is
-    shown for comparison. Note that uWUE from \citet{Zhou_2015} is
-    calculated from a different set of sites, and that units are
-    converted such that the quantities work with Equations 1-8 and the
-    variables defined Table \ref{definitions}.}
-  \small
-  \label{pft}
-  \centering
-  \begin{tabular}{l c c @{\qquad} c c}
-    \hline
-    \multirow{2}[3]{*}{Abbreviation} & \multirow{2}[3]{*}{PFT} & \multirow{2}[3]{*}{$g_1$ (Pa$^{0.5}$)} & \multicolumn{2}{c}{uWUE ($\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET])}  \\
-    \cmidrule{4-5}
-
-                                     & & & fitted & \citet{Zhou_2015} \\
-
-    \hline
-    \input{pft_params.tex}
-    \hline
-  \end{tabular}
-\end{table}
-
-While Equation 3 can be used in PM (Equation 1), it will make
-analytical work with the function intractable because $GPP$ is
-functionally related to ET itself. Additionally, a perturbation to VPD
-should induce a physiological plant response that will alter GPP and
-cause an indirect change in stomatal conductance, in addition to the
-direct effect of VPD in Equation \ref{medlyn}. Therefore, in order to
-derive the response of ET to VPD, we must account for the functional
-relationship between GPP, ET, and VPD, and its effect on stomatal
-conductance. We can use aforementioned semi-empirical results of
-\citet{Zhou_2015} as a tool to approach this
-problem. \citet{Zhou_2015}, showed that  underlying Water Use
-Efficiency (uWUE):
-
-% \begin{linenomath*}
-  \begin{equation}
-    uWUE = \frac{GPP \cdot \sqrt{VPD}}{ET}
-    \label{uwue}
-  \end{equation}
-% \end{linenomath*}
-
-is relatively constant across time and moisture conditions within a
-plant functional type, and correctly captures a constant relationship
-between GPP, ET and VPD over a diurnal cycle \citep{Zhou_2014}. The
-theoretical derivation of the square root VPD dependence in $uWUE$
-leverages the same assumptions used in \cite{MEDLYN_2011} to derive the
-square-root VPD dependence of the stomatal conductance model (Equation
-\ref{medlyn}).  We can use uWUE to remove the $GPP$ dependence from
-$g_s$ in a way that makes PM analytically tractable:
-
-% \begin{linenomath*}
-  \begin{equation}
-    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{uWUE \; ET}{c_a \; \sqrt{VPD}}.
-    \label{new_g_s}
-  \end{equation}
-% \end{linenomath*}
-
-Plugging Equation \ref{new_g_s} into Equation \ref{orig_pen} and
-rearranging gives a new explicit expression for PM, in which
-dependence on $GPP$ is removed:
-
-% \begin{linenomath*}
-  \begin{equation}
-    ET = \frac{\Delta R_{net} + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R \; 1.6\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
-    \label{et}
-  \end{equation}
-% \end{linenomath*}
-
-By accounting for photosynthesis changes in ecosystem conductance, with
-Equation \ref{et} we have derived for the first time, using recent
-results \citep[][]{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017}, ET
-explicitly as function of environmental variables and two
-plant-specific constants, the slope parameter ($g_1$), and uWUE, both
-reflecting water conservation strategy. The slope parameter is related
-to the willingness of stomata to trade water for CO$_2$ and to keep
-stomata open. uWUE is a semi-empirical ecosystem-scale constant
-related to how WUE changes with VPD (specifically $VPD^{-1/2}$). It is
-also roughly proportional to physical constants:
-
-\[uWUE \appropto \sqrt{\frac{c_a - \Gamma}{1.6 \lambda}},\]
-
-where $\Gamma$ is the CO$_2$ compensation point \citep[Equation 5
-in][]{Zhou_2014}. So uWUE is related to atmospheric CO$_2$
-concentration and compensation point, and is inversely proportional to
-the marginal water cost of leaf carbon.
-
-Given eddy-covariance FLUXNET2015 data (Section \ref{data}), every
-term in our new version of PM (Equation \ref{et}) is observed, except
-for uWUE, which we fit by calculating its expectation, given the model
-and FLUXNET2015 data.
-
-However, eddy-covariance data are inherently noisy so we include a
-measure of uncertainty in our analysis. To account for observational
-error, as well as model uncertainty (e.g. temporal and spatial
-variations of uWUE and $g_1$), we introduce an uncertainty parameter
-$\sigma$ modifying uWUE:
-
-% \begin{linenomath*}
-  \begin{equation}
-    ET = \frac{\Delta R_{net} + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R \; 1.6\; \sigma \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
-    \label{et_sigma}
-  \end{equation}
-% \end{linenomath*}
-
-Now, from each FLUXNET2015 observation (i.e. for each hourly
-observation at every time step) we can evaluate $\sigma$:
-
-% \begin{linenomath*}
-  \begin{equation}
-    \sigma = - \frac{g_a \gamma c_a \sqrt{VPD} L_v P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R_{net} - g_a \rho_a c_p VPD\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})},
-    \label{sigma}
-  \end{equation}
-% \end{linenomath*}
-
-Thus, with this uncertainty analysis we can evaluate departure from
-our theory in observations, as a departure of $\sigma$ from unity. The
-variability of $\sigma$ across sites and time provides a measure of
-uncertainty in our model, assumptions, as well as in the FLUXNET2015
-observations themselves. The variability of $\sigma$ then propagates
-through any uncertainty to our partial derivative of Equation
-\ref{et_sigma} with respect to VPD:
-
-% \begin{linenomath*}
-  \begin{equation}
-    \frac{\partial \;  ET}{\partial \; VPD} = \frac{2\; g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_a }{1.6 \; R\; \sigma \; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)
-    \label{d_et}
-  \end{equation}
-% \end{linenomath*}
-
-With Equation \ref{d_et} we have an analytical framework for ecosystem
-response to atmospheric demand perturbations with environmental
-conditions held fixed. There are a few subtleties to taking the
-derivative in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$)
-and $VPD$ are functionally related, so while taking the derivative we
-evaluate
-$\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET}
-{\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD}
-\Big|_{\text{RH fixed}} + \frac{\partial \; ET}{\partial \; RH}
-\frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$
-    fixed}}$. $RH$ and $e_s$ are assumed to be approximately
-independent, which is supported by the data (not shown).
-
-This derivation relied either implicitly or explicitly on several
-assumptions. First, we assume that VPD at the leaf surface is the same
-as VPD at measurement height; physically this implies that leaves are
-perfectly coupled to the atmosphere. In reality, for some conditions
-and plant types the leaves can become decoupled from the boundary
-layer \citep{De_2017, Medlyn_2017}. Given our focus on the growing
-season, which is usually characterized by relatively high insolation
-inducing instability and convective boundary layers, we would expect
-the surface to be generally well coupled. An additional assumption in
-the formulation of $uWUE$ \citep{Zhou_2014, Zhou_2015} and
-\citet{Medlyn_2017}'s stomatal conductance model is that direct soil
-evaporation (E) contributions to ET remain small relative to
-transpiration (T). This should be more true during the growing
-season. The ratio of E to T may increase immediately after rainfall
-events due to high soil moisture and ponding, but VPD is generally low
-anyways during these times. However, some plant types allow for
-systematically larger contributions of E in ET, particularly those
-with sparse canopies and smaller relative amounts of transpiration. We
-therefore might expect that the theory will be most applicable to
-forest PFTs, which will be most strongly coupled to the boundary layer
-due to larger surface roughness, and will also generally have the
-highest ratios of transpiration to evaporation. Finally, for the goal
-of developing PFT-wide intuition, we assume that $g_1$ and $uWUE$ are
-constant in space and time. Both quantities have been shown to be well
-conserved over space and time within a PFT relative to inter-PFT
-variability. However, we would expect them to vary somewhat with
-environmental conditions including very low soil water content,
-temperature, and with intra-PFT plant-specific characteristics like
-wood density for tree PFTs \citep{Lin_2015}. Given these strong
-assumptions made with the goal of understanding broad, leading order
-plant behaviors, we take the extra care of including and analyzing
-uncertainty (Section \ref{testing}) and examining sensitivity to the
-exponent of the VPD response (Section \ref{functional_form}).
-
-We note one final comment on our derivation which is relevant for
-drought indices. If we approximate $c_a$ at a global mean CO$_2$
-concentration, then the RHS of Equation \ref{et} is fully defined
-using commonly available weather station data and the constants
-published here. This makes Equation \ref{et} a useful alternative to
-PET in drought indices and hydrometeorological analysis for vegetated
-surfaces. Equation \ref{et} better reflects the physics of water
-exchange at the land surface and would only require fitting of uWUE
-and $g_1$, and can also account for changes in CO$_2$ concentration
-(assuming some relationship between $\lambda$ and [CO$_2$]), contrary
-to typical drought indices such as the Palmer Drought Severity Index
-(PDSI) \citep{Swann_2016, Lemordant_2016, Lemordant_2018}. Lastly, all
-code and data used in this analysis, including those used to generate
-the figures and tables, are publicly available at
-\url{https://github.com/massma/climate\_et}.
-%\hyperref[https://github.com/massma/climate\_et]{https://github.com/massma/climate_et}.
-
-\section{Results and Discussion}
-\label{results}
-
-By construction, the variability in the $\sigma$ term (Equation
-\ref{sigma}) contains all model and observational uncertainties. For
-an observation that perfectly matches our model and constant uWUE
-assumption, $\sigma$ will be one. Therefore, for our assumptions and
-framework to be reasonable $\sigma$ should be close to 1. An
-additional concern is that $\sigma$ may in fact be correlated with
-$VPD$, in which case the dependence would need to be accounted for
-when taking the derivative. Fortunately, there is a very weak
-dependence of $\sigma$ on VPD in their joint distribution, and
-$\sigma$ is indeed close to unity i.e. $O(1)$ (Figure
-\ref{joint_vpd_sigma}). Given this weak dependence and the
-distribution of $\sigma$ we have confidence in our model framework and
-the data quality.
-
-\begin{figure}[h]
-  \centering\includegraphics[width=0.75\textwidth]{./joint_vpd_sigma.pdf}
-  \caption{The joint distribution of $VPD$ and $\sigma$, with outliers
-    removed (defined as lowest and highest 5\% of $\sigma$). $\sigma$
-    exhibits a weak dependence on $VPD$, and $\sigma$ is $O(1)$ for
-    the bulk of the observations.}
-  \label{joint_vpd_sigma}
-\end{figure}
-
-Before calculating the sensitivity of ET to VPD, we will consider the
-functional form of Equation \ref{d_et}. There are two main terms: a
-``scaling'' term, which modifies the magnitude but not the sign of the
-ET response to VPD ($\frac{\partial \; ET}{\partial \; VPD}$):
-
-\begin{equation}
-  \frac{g_a \; P}{T(\Delta + \gamma)},
-\end{equation}
-
-and a ``sign'' term, which determines whether ET increases or
-decreases with VPD (i.e. atmospheric demand driven or physiologically
-controlled):
-
-\begin{equation}
-  \label{sign}
-  \frac{c_p}{R_{air}} - \frac{\gamma c_a }{1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right).
-\end{equation}
-All variables are positive, so the relative magnitude between the
-first term and the second term in the sign term (Equation \ref{sign})
-will determine whether ET increases or decreases with increasing
-VPD. If the second term is larger then plant control dominates and ET
-decreases with increasing VPD. However, if the first term is larger,
-then atmospheric demand dominates and ET increases with increasing
-VPD.
-
-\subsection{Functional Form of the Sign Term}
-\label{sign_term}
-First, we explore the variables within the sign term to gain better
-intuition on the driver of either the increase or reduction of ET with
-VPD. CO$_2$ concentration ($c_a$) and the psychometric constant
-($\gamma$) are relatively constant over the dataset considered here so
-that the variability is dominated by $\sigma$ and $VPD$. uWUE could
-vary with soil moisture but has been shown to be relatively constant
-\citep{Zhou_2015}. This then means that the sign term only depends on
-VPD for a given PFT and is approximately just a function of $VPD$. We
-can further determine a critical threshold separating an increase from
-a decrease in ET, i.e. the threshold $VPD_{crit}$ such that the
-derivative vanishes $\frac{\partial \; ET}{\partial \; VPD} = 0$:
-\small
-% \begin{linenomath*}
-  \begin{equation}
-    VPD_{crit} = \frac{R_{air}}{4 c_p} \left( \frac{\gamma c_a}{1.6\; R \;  uWUE} + \sqrt{\frac{\gamma c_a}{1.6\; R \;  uWUE}\left( \frac{\gamma c_a}{1.6\; R \;  uWUE} + 8 g_1 \frac{c_p}{R_{air}}\right)} - 4 g_1 \frac{c_p}{R_{air}} \right),
-    \label{vpd_min_et}
-  \end{equation}
-% \end{linenomath*}
-\normalsize noting that $VPD_{crit}$ mostly depends on the PFT
-parameters uWUE and $g_1$, and only varies weakly with climate as most
-other parameters related to the environment are nearly constant. The
-calculated value of $VPD_{crit}$ for each PFT is shown in Table
-\ref{vpd_crit}. For any values of $VPD$ less than $VPD_{crit}$, ET
-will decrease with increasing VPD
-($\frac{\partial \; ET}{\partial \; VPD} < 0$), and for values of
-$VPD$ greater than $VPD_{crit}$, ET will increase with increasing VPD
-($\frac{\partial \; ET}{\partial \; VPD} > 0$). In other words,
-ecosystems regulate and mitigate evaporative losses up to the VPD
-limit, $VPD_{crit}$, above which atmospheric demand is just too high
-to be entirely compensated by stomatal and ecosystem regulation. We
-note however that even though ET increases again above the critical
-threshold, $VPD_{crit}$, ET is still much lower than potential
-evaporation as stomata are still strongly regulating vapor fluxes to
-the atmosphere. Indeed, even in the absence of soil pore evaporation,
-stomata do not shut down entirely at very high VPD so that ET does not
-go to zero, because stomata are still slightly open to perform some
-photosynthesis \citep{Ball_1987, Leuning_1990, MEDLYN_2011}. In
-addition, upward xylem transport is necessary to maintain phloem
-transport, as well as nutrient transport and thus carbon allocation
-\citep{De_2013, Nikinmaa_2013, Ryan_2014}.
-
-\begin{table}
-  \caption{Values of $VPD_{crit}$, where
-    $\frac{\partial \; ET}{\partial \; VPD} = 0$, evaluated at PFT
-    average values for $R_{air}$, $\gamma$, and $c_a$. PFT-specific
-    constants ($g_1$, uWUE) are provided in Table \ref{pft}. For
-    values of $VPD$ less than $VPD_{crit}$,
-    $\frac{\partial \; ET}{\partial \; VPD}$ will be negative, and for
-    values of $VPD$ greater than $VPD_{crit}$,
-    $\frac{\partial \; ET}{\partial \; VPD}$ will be positive.}
-  \centering
-  \begin{tabular}{l c c c c c}
-    \hline
-    PFT & $R_{air}$ & $c_a$ (ppm) & $\gamma$  & \textbf{$VPD_{crit}$ (Pa)} \\
-    \hline
-    \input{vpd_crit.tex}
-    \hline
-  \end{tabular}
-  \label{vpd_crit}
-\end{table}
-
-Differences in $VPD_{crit}$ are exclusively determined by uWUE and the
-slope parameter ($g_1$) related to the plant functional type. A larger
-uWUE means a smaller $VPD_{crit}$, and an ET response to increases VPD
-that is more likely to be positive. At first glance this result is
-somewhat counter-intuitive; we expect that plants with a higher water
-use efficiency would be more water conservative. However, in reality
-uWUE determines how $WUE$ changes with VPD:
-
-\[WUE = \frac{GPP}{ET} = \frac{uWUE}{\sqrt{VPD}}.\]
-\[\frac{\partial \; WUE}{\partial \; VPD} = -\frac{uWUE}{2 \;
-    VPD^{3/2}}\]
-
-So, plants with a higher uWUE will have a greater \textit{decrease} in
-ecosystem-scale $WUE$ in response to increases in VPD. This decrease
-in $WUE$ causes more water loss per unit carbon gain, and explains the
-relationship between high uWUE and high likelihood of increases of ET
-in response to increasing atmospheric drying (increases in VPD).
-
-A tendency towards increasing ET response with increasing VPD can also
-be caused by a high slope parameter ($g_1$), characteristic of plants
-that at the leaf scale are more willing to trade water for access to
-atmospheric CO$_2$. Plants that are less conservative will be thus be
-more likely to increase ET with increasing VPD. Both the
-aforementioned effects (large uWUE, $g_1$) can amplify each other, and
-generally conspire to shift the sign term towards a positive value for
-a given PFT.
-
-This effect of uWUE and $g1$ on the sign term is most apparent by
-comparing two extreme PFTs: water intensive crops (CRO) and water
-conservative closed shrub (CSH). CRO has higher slope parameter and a
-slightly higher uWUE ($g_1 = 140.7$ Pa$^{1/2}$; $2.85$ $\mu$-mol [C]
-Pa$^{0.5}$ J$^{-1}$ [ET]) compared to CSH ($g1 = 75.1 \, Pa^{1/2}$,
-$uWUE=2.82$ $\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET]). These differences
-in PFT parameters cause opposite ET responses to changes in VPD
-between CRO and CSH. ET theoretically always decreases with increasing
-VPD for the more water conservative CSH, while ET frequently increases
-with increasing VPD for the more water intensive CRO (Figure
-\ref{idealized_sign}). CROs evolved or were bred to prioritize GPP and
-yield and are thus not water conservative. They are very willing to
-trade water for photosynthesis and productivity, despite changes in
-VPD, while CSH are very unwilling to trade water for more
-photosynthesis.
-
-\begin{figure}
-  \centering \includegraphics{./idealized_sign.pdf}
-  \caption{The functional form of the sign term, with $\sigma$ held
-    fixed at 1, and all terms except VPD set to PFT averages. For
-    comparison, the observed range of VPD for each PFT is plotted
-    below the x-axis. Stars denote 25th, 50th, and 75th percentiles,
-    and the range of the line spans the 5th-95th percentiles of
-    observed VPD. Vertical black lines denote the location of
-    $VPD_{crit}$ for each PFT, with the exception of CSH and WSA, for which
-    $VPD_{crit}$ is off-scale.}
-  \label{idealized_sign}
-\end{figure}
-
-As expected, the slope parameter ($g_1$) is a primary determinant of
-the VPD dependence for the sign term shown in Figure
-\ref{idealized_sign}. Plants that are more conservative (small $g_1$)
-will tend to reduce ET with increasing VPD, and will be very effective
-at reducing ET, especially at low VPD. However, at very high VPD,
-gradients in vapor pressure at the leaf scale will become very strong
-as stomata reach their limits of closure in response to VPD
-(parameterized with $g_1$). As a result, ET response will begin to
-asymptote towards a constant ecosystem-scale values as leaf-scale
-response to VPD asymptotes towards zero.  Therefore, plants with a low
-$g_1$ will have the largest VPD dependence of ET response because the
-difference in ET response at low VPD (leaf stomatal response
-dominates) and high VPD (VPD gradient dominates) is largest. This is
-apparent in the strong VPD dependence of CSH, which has the lowest
-slope parameter ($g_1=75.1$ Pa$^{1/2}$) (Figure \ref{idealized_sign}).
-
-To summarize our theoretical insights (Figure \ref{idealized_sign} and
-Table \ref{vpd_crit}), CROs are the least water conservative and have
-the strongest overall tendency to increase ET with increasing VPD,
-while CSH are the most water conservative and have the strongest
-tendency to decrease ET with increasing VPD, as well as the strongest
-VPD dependence of response. Figure \ref{idealized_sign} clearly shows,
-according to our theory, that for all PFTs except for crops there is
-frequent occurrence of a negative (plant dominating) ET response to
-increases in VPD. Therefore, plants are able in most atmospheric
-conditions to reduce ET in response to increased VPD and thus to
-reduce water loss. To better illustrate this, the ranges of observed
-environmental VPDs at the FLUXNET sites are plotted parallel to the
-x-axis. For CSH and WSA, VPD is always less than VPD$_{crit}$ (off scale) so
-that the plant response dominates in typical environmental conditions,
-emphasizing the water conservative strategy of those plants. For CRO
-on the other hand, VPD is higher than VPD$_{crit}$ for more than 50\%
-of observations, emphasizing that those plants operate with an
-aggressive water usage strategy, are water intensive and were actually
-engineered for photosynthesis rather than water saving. For DBF, EBF,
-MF, GRA, and SAV more than half of the observed VPD are less than
-VPD$_{crit}$, i.e. in conditions where plant response dominates. SAV
-has a more water conservative response than the forest, grass, and
-crop plan types, but still responds by increasing ET with increasing
-VPD for about a quarter of observations, due to the high aridity (VPD)
-of the SAV ecoclimate. It is also important to note that for all PFTs,
-even when atmospheric demand dominates, ET response to VPD is still
-far more negative than it would be for potential evaporation
-$\partial PET/\partial VPD$, i.e. atmospheric demand only, emphasizing
-that there is still a strong regulation of evaporative flux by stomata
-and though the plant xylem. The sign term in the PET case would just
-be a constant ($\frac{c_p}{R_{air}} \approx 3.5$), which is far larger
-than any part of the curves for any PFT. This highlights the
-deficiencies of PET's ability to capture land response to changes in
-atmospheric dryness (VPD). Plants are always regulating water exchange
-from the land surface, even when they reach the limits of they ability
-to do so.
-
-\subsection{Functional Form of the Scaling Term}
-\label{scale_term}
-While the above discussion of the sign of
-$\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our
-question of when ET response increases or decreases with VPD,
-understating the overall magnitude of the ET response is important to
-soil-plant-atmosphere water budgeting. So we now more closely examine
-the terms that affect how the sign term is scaled:
-
-\begin{equation}
-  \frac{g_a \; P}{T(\Delta + \gamma)}.
-\end{equation}
-
-$\frac{P}{T}$ is an air-density term, which varies little compared to
-aerodynamic conductance and Clausius-Clapeyron ($\Delta$). The
-psychometric constant ($\gamma$) is also relatively constant, so the
-scaling term should be primarily a function of aerodynamic conductance
-and temperature, through the Clausius-Clapeyron relationship
-$\Delta$. This is as expected, given that the aerodynamic conductance
-represents the efficiency of exchange between the surface and the
-atmosphere. As aerodynamic conductance increases, any plant response
-will be communicated more strongly to the atmosphere (and vice-versa).
-
-$\Delta$'s presence in the scaling term also matches physical
-intuition. $\Delta$ (and also the approximately constant $\gamma$)
-control the efficiencies with which surface energy is converted to
-latent and sensible heat \citep{Monteith_1965}. The functional from of
-$\Delta$ will be the same across PFTs, but the temperature range may
-vary slightly. In contrast, aerodynamic conductance will vary strongly
-with PFT due to the importance of surface roughness for aerodynamic
-conductance. So most of the differences in scaling between PFT should
-be in the aerodynamic conductance term.
-
-The control of the scaling term variability between PFTs by
-aerodynamic conductance is confirmed by data (Figure
-\ref{scale_vary}). Differences between PFT are almost entirely due to
-differences in aerodynamic conductance, rather than differences in
-observed temperature ranges. The scaling term for the tree PFTs (DBF,
-EBF, ENF, MF) is generally about double the scaling terms for other
-PFTs which have lower surface roughness and generally smaller
-aerodynamic conductance (GRA, CSH, CRO). The savannah (WSA, SAV) PFT's
-scaling is somewhere between GRA, CSH, and CRO, and DBF, EBF, ENF, and
-MF, due to higher variability and surface roughness.
-
-\begin{figure}
-  % \centering
-  \centerline{\includegraphics[width=1.2\textwidth]{./idealized_scale.pdf}}
-  \caption{Primary sources of variability for the scaling term, as a
-    function of PFT. The 5th-95th percentile range of temperature is
-    plotted at the 5th, 25th, 50th, 75th, and 95th percentiles of
-    aerodynamic conductance, as observed for each PFT.}
-  \label{scale_vary}
-\end{figure}
-
-Within each PFT, the scaling term variability is controlled both by
-environmental temperature and aerodynamic conductance variability
-(Figure \ref{scale_vary}). While the observed variability of the
-aerodynamic conductance contributes more to the scaling term
-variability than temperature, the temperature contribution is
-non-negligible. Specifically, the scaling term is generally larger at
-low temperatures when latent heat is relatively inefficient at moving
-energy away from the surface. This effect amplifies the role of
-aerodynamic conductance variability at low temperatures.
-
-To summarize, variability between PFTs is mostly controlled by
-systematic differences in aerodynamic conductance, due to differences
-in surface roughness between each PFT, and possibly to a lesser extent
-wind conditions. In contrast, variability within PFT is also
-controlled by temperature, through Clausius-Clapeyron. But,
-aerodynamic conductance variability generally impacts the scaling term
-more than temperature, even within PFTs.
-
-\subsection{Bulk statistics of ET response to VPD}
-\label{stats_sec}
-
-In this section we consider direct observations of ET response with
-eddy-covariance data, while including uncertainty with the $\sigma$
-term (Section \ref{methods}). These observational results of ET
-response (Table \ref{stats}) largely confirm our theoretical analysis, presented in the previous sections. For all PFTs, mean ET response to increasing VPD
-is negative. However, ET response evaluated at the average of all
-variables (e.g. $\sigma$, $T$, $c_a$, $VPD$) is positive for CRO, and
-negative for all other PFTs. This difference in mean ET response as
-compared to the ET response at mean environmental conditions is due to
-the non-linear nature of the response, in which negative responses are
-generally larger magnitude than positive responses (Figure
-\ref{idealized_sign}). Therefore, both the mean ET response as well as
-the ET response at mean environmental conditions matches our
-expectations from the theory (Section \ref{sign_term}), with the
-exception that CRO observations are shifted more towards a negative ET
-response than we expect.
-
-\begin{table}
-  \caption{Statistics of $\frac{\partial \; ET}{\partial \; VPD}$ as a
-    function of PFT.}
-  \centering
-  \begin{tabular}{l c c c}
-    \hline
-    PFT & $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ & $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)$ & fraction $\frac{\partial \; ET}{\partial \; VPD} < 0.$ \\
-    \hline
-    \input{stats.tex}
-    \hline
-  \end{tabular}
-  \label{stats}
-\end{table}
-
-Regarding the frequency of negative and positive ET response, all PFTs
-exhibit a decreasing ET response with increasing VPD (physiologically
-controlled, water conservative response) for the majority of
-observations. The more water conservative PFTs generally exhibit
-higher frequency of negative ET response, especially when one factors
-in the distribution of environmental VPD (e.g. SAV and WSA grow in
-more arid climates, Figure \ref{idealized_sign}). In general the bulk
-statistics match our theoretical expectations well, with the caveat
-that inclusion of uncertainty shifts crops towards a slightly more
-negative response to VPD, and shifts many of the other PFTs, which still
-exhibit a high frequency of negative ET response to VPD, towards more frequent
-occurrence of positive ET response than the theory and Figure
-\ref{idealized_sign} would suggest. The bulk statistics motivate a
-more thorough examination of the structure of uncertainty and more
-sophisticated validation of our theory's performance against
-observations.
-
-\subsection{Validation of theory at eddy-covariance sites}
-\label{testing}
-We now compare more sophisticated distributions of the observed
-response to our simplified theory (Section \ref{sign_term}). The
-observed distribution of the sign term, as compared to what the theory
-would predict, is provided in Figure \ref{test_sign}. Our goal was to
-capture the leading order behavior of the ET dependence on VPD. Given
-the assumptions we made, and the uncertainties of flux tower
-observations themselves, we expect a relatively large amount of noise
-when reproducing the derivatives of ET. However, the data largely
-reproduces our theoretical analysis.
-
-\begin{figure}
-  \centering
-  \centerline{\includegraphics[width=1.2\textwidth]{./test_sign.pdf}}
-  \caption{Comparison of the sign term with model uncertainty included
-    (box plots) to the sign term as calculated with simplifying
-    assumptions (blue line, as in Figure \ref{idealized_sign}). Each
-    box plot corresponds to 5\% of the data, and the 5-95\% range of
-    VPD is plotted.}
-  \label{test_sign}
-\end{figure}
-
-This is particularly true for DBF and MF; the theory matches the
-leading order behavior of the function when uncertainty is included,
-and the observations match the theory with the addition of noise. The
-VPD dependence, given by the slope parameter ($g_1$), follows the
-median values of each bin. Perhaps most importantly, the x-intercept,
-and thus VPD$_{crit}$, matches nearly exactly between the theory and
-the observations. Therefore the sign of the ET response to increases
-in VPD should be well matched, subject to the unavoidable constraints
-of noise, much of which comes from the observations themselves. The
-uncertainty is non negligible; there are many observations in each bin
-for which the the sign of observation is opposite the response
-predicted by the theory, but to leading order our theory matches the
-observations well.
-
-While CSH has a much different functional form of the sign term than
-DBF and MF, CSH observations also match our theory to leading order,
-albeit with a bit more variability as a function of VPD. Again, the
-VPD dependence mostly determined from the slope parameter ($g_1$)
-closely matches the medians in the observation bins. The
-VPD-independent, strongly negative response is also captured. For CSH,
-there is rare occurrence of observed positive ET response with VPD
-($\approx 6$\%), even with uncertainty, so the sign of the
-observations almost always matches the sign of the theory, which
-states that ET response should always be negative.
-
-Biases between the theory and observations are similar for ENF, WSA,
-CRO and SAV. At low VPD the theory and the observations match
-well. However, at high VPD (upper 10-20\% of observations) the theory
-is biased slightly towards positive response as compared to the
-observations (Figure \ref{test_sign}). For ENF this slight bias could
-be explained by a negative bias in $g_1$. However, for WSA, CRO, and
-SAV the observations in the highest VPD bins exhibit a downturn
-towards more negative response, which cannot be captured by the
-functional form of the sign term (Figure
-\ref{idealized_sign}). Therefore, to explain the observations one of
-the variables in the sign term must change at extreme VPD (upper
-10-20\%). The most likely candidate is $uWUE$, which we have assumed
-constant to meet our goal of developing intuition for leading order
-behavior, but might be expected to decrease at extremely low SWC. This extremely low SWC should be correlated with high VPD in very dry environments
-through sensible heating increases \citep{Gentine_2016}. The theory
-for EBF exhibits similar limitations as for ENF, WSA, and SAV, except
-a larger portion of observations are biased negative as compared to
-the theory ($\approx$ 35\%, Figure \ref{test_sign}). However, in
-general, and specifically for non-extreme VPD (VPD  $< \approx$70-90th
-percentiles), the theory matches the observations for the tree and
-savannah plant types well.
-
-The theory for GRA suffers from similar, but much more severe,
-limitations as for CRO, WSA, and SAV. GRA observations are
-characterized by a consistent trend back towards negative ET response
-at higher VPD, which the functional form of our theory is incapable of
-accounting for. As compared to CRO, WSA and SAV, the divergence
-between the theory and the observations is far greater for GRA,
-biasing 40-50\% of observations at higher VPD. In addition for the
-potential for soil moisture to alter $uWUE$, there are other sources
-of plant heterogeneity specific to GRA (and some extent CRO) that may
-alter $uWUE$ (or $g_1$) or invalidate other assumptions made in the
-Methods section (Section \ref{methods}). We do not account for
-variability in plant height and surface roughness, or differences in
-C3 vs C4 photosynthesis and water strategies, which we might expect to
-vary substantially across sites, years, and season for GRA. These
-deficiencies could largely explain the inability of our theory to
-exactly match the observations in croplands. For example, a
-superposition of sites with C3 plants (at low environmental VPD) and
-C4 plants (at high environmental VPD) would explain the observed shift
-back towards negative ET response at high VPD when all sites are
-binned together, as in Figure \ref{test_sign}. We hypothesize that the
-theory would validate against observations much better if these
-sources of variability were accounted for, at a cost of increased
-complexity and analytical opacity.
-
-While the above discussion shows that our theory has some limitations
-when applied to some PFTs, especially for grasslands, it does well for
-DBF, MF and CSH PFTs, and captures the response at non-extreme VPD
-(VPD $< \approx$ 70th-90th percentiles) for
-ENF, EBF, CRO, WSA, and SAV. In general, the leading order behavior
-observed in the data is captured by the theory. Departures between our
-theory and observations, specifically at high VPD, could be explained
-and conceptualized with shifts in $g_1$ and/or $uWUE$ due to
-site-specific plant type variability (e.g. more arid-adapted ecosystems
-at more arid sites), or temporal variability for some environmental
-conditions (e.g. decreases in $uWUE$ at extremely low soil
-moisture). To focus on general behavior and develop intuition for
-PFT-scale response, we have ignored these sources of variability in
-the present analysis. However, any site-specific or temporal plant
-functional variability that can be conceptualized with shifts to $g_1$
-and $uWUE$ can be analyzed with our framework. This opens a door to future analyses in which plant behavior in anomalous conditions
-can be explained and analyzed using Equation \ref{et}.
-
-% It is useful to also compare our theory, the observed uncertainty in
-% the ET response, and PET, as PET is often used as a proxy for ET in
-% hydrometeorological studies and drought indices. Again, as in our
-% discussion of the theoretical results, the PET sensitivity to VPD
-% will be constant at about 3.5. In Figure \ref{test_sign} the ET
-% response to shifts in dryness is always far smaller than 3.5 in the
-% observations, even with all uncertainty included. So the PET
-% response to atmospheric dryness is outside the range of all
-% observable responses. This casts doubt on PET's usefulness for
-% capturing the leading order behavior of water budget and dryness
-% response at the land surface, as is done in several drought
-% indices. However our theory (Equation \ref{et}) could be used with
-% published uWUE and $g1$ data, as an alternative to PET to estimating
-% land response and dryness, and would more accurately account for the
-% response to atmospheric drying.
-
-\subsection{Observed ET response to VPD}
-
-Most of the results presented so far focus on the sign term, so now we
-turn to observations of ET response with the scaling term included
-(Figure \ref{data_scatter}). Until now, in the interest of developing
-leading order intuition for ET response, and to be conservative in our
-acknowledgment of model and observational error, we've considered
-$\sigma$ variability to be a measure of uncertainty. An alternative
-viewpoint is that $\sigma \cdot uWUE$ represents spatial and temporal
-variability of uWUE, which may be expected within bounds \citep[see
-Table \ref{pft}, also ][]{Zhou_2015}. This is a less conservative
-view; some of the $\sigma \cdot uWUE$ variability will be due to model
-and observational error, so by viewing $\sigma \cdot uWUE$ as
-\textit{real} variability we run the risk of mistaking noise for
-signal. However, the advantage of this viewpoint is that, from
-\citet{Zhou_2014}, we have very high confidence that uWUE fit at the
-hourly timescale (as we do with $\sigma \cdot uWUE$) correctly
-captures the relationship between ET, GPP, and VPD, and that our form
-of PM introduces minimal error with its use of uWUE.
-
-\begin{figure}
-  \centering
-  \centerline{\includegraphics[width=1.2\textwidth]{./data_scatter.png}}
-  \caption{Scatter plots of observed
-    $\frac{\partial \; ET}{\partial \; VPD}$, including $\sigma$
-    variability, as a function of PFT, temperature and VPD. Please
-    note differences in the colorbar scale.}
-  \label{data_scatter}
-\end{figure}
-
-Therefore, with the caveat that some of the signal presented in Figure
-\ref{data_scatter} may in fact be noise, we can interpret the observed
-distribution of ET response to VPD. In general, the observed response
-matches the intuitive theory. ET response to VPD shifts towards
-positive values as VPD increases (atmospheric demand dominating). CRO
-exhibit the highest occurrence of positive ET response, and the
-observations confirm that CRO are the most water intensive. CSH are
-the most water conservative, with a strong negative response. DBF,
-EBF, ENF, MF, SAV and WSA are also water conservative, but show some
-occurrence of positive ET response to VPD, particularly at higher
-observed VPD, as the theory predicts. GRA, while generally water
-conservative, does not match the theory well, with increasing frequency of
-\textit{negative} response at high VPD. This is as expected, given
-previous discussion on how we might expect more inter-site and inter-year variability for GRA (Section \ref{testing}).
-
-Figure \ref{data_scatter} also includes the impact of the scaling
-term. For a given VPD, the magnitude of ET response does not vary
-strongly with temperature, confirming that any impact of the scaling
-term on the magnitude of the response is primarily due to changes in
-aerodynamic conductance. Intuitively this is reasonable; aerodynamic
-conductance will control how dominant balances at the land surface are
-communicated to the boundary layer. While $\Delta$ controls the
-efficiency of energy conversion to latent heat, it appears this is a
-second order term, relative to $g_a$, for scaling ET response.
-
-As with Figure \ref{test_sign}, Figure \ref{data_scatter} matches our
-expectation based on simplified theory. The sign term is most strongly
-scaled by $g_a$, and in general the occurrence of positive ET response
-increases as VPD increases. The willingness with which a given plant type
-evolved to use water dictates the occurrence of positive versus
-negative ET response. Water conservative ecosystems are highly effective
-at mitigating the effects of atmospheric demand, and can store water
-for later use by reducing ET in response to increasing VPD.
-
-\subsection{Limitations of theory - very dry soil moisture conditions}
-\label{swc_section}
-In formulating our theory with Penman Monteith, we implicitly did not
-account for very dry soil moisture conditions. For the majority of
-environmental conditions observed at the eddy covariance sites used
-here, soil conditions were not extremely dry so that we could assume a
-constant uWUE and $g_1$. We posit that ecosystems will generally
-optimize to host plants living in conditions which they evolved
-for. However, in extreme conditions and drought scenarios soil water
-content (SWC) could become the limiting factor for ET response to VPD,
-which our theory does not account for. In addition, low soil moisture
-conditions themselves increase VPD through land-atmosphere feedback
-\citep[][]{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006,
-  Salvucci_2013, Gentine_2016, Berg_2016}.
-
-Within our framework, any systematic bias due to the
-failure to account for SWC's effects in very dry conditions should
-manifest itself in a functional relationship between $\sigma$, our
-uncertainty measure, and SWC, which is observed at the FLUXNET
-sites. Examining the relationship between $\sigma$ and SWC will test
-to what extent our theory breaks down in very dry soil moisture
-conditions.
-
-If we again view $\sigma \cdot uWUE$ as, in addition to a measure of
-uncertainty, a short time-scale observation of uWUE, we would expect
-$\sigma \cdot uWUE$ to decrease at low SWC. If $\sigma \cdot uWUE$ is
-a very strong function of SWC, then our theory should be conditioned
-more strongly on well watered soil conditions. If $\sigma \cdot uWUE$
-is weakly a function of SWC, then our theory is more universal and
-independent of soil moisture conditions.
-
-Indeed, for all PFTs, there is some slight dependence of
-$\sigma \cdot uWUE$ on SWC, especially at low SWC (Figure
-\ref{swc_boxplot}). The portion of observations for which our theory
-is biased by SWC-limitations varies by PFT, due to the nonlinear
-threshold at which soil moisture availability limits plant
-function. For CRO, MF and DBF, soil water content only matters for
-about the lowest 5\% of observations (each box is 5\% of observations
-in Figure \ref{swc_boxplot}). And, CRO, relative to DBF, has a very
-weak dependence on soil moisture, which is reflective of the high
-likelihood that CRO sites are optimally irrigated and not water
-stressed, suggesting that observed departures from the theory for CRO
-are due to other factors (i.e. different photosynthesis pathways; C3/C4) than soil
-moisture. For ENF, EBF, SAV, and WSA, systematic SWC-induced biases in
-$\sigma \cdot uWUE$ emerge in about the lowest 20-30\% of SWC
-conditions, although large variability in the SWC-$(\sigma \cdot uWUE)$
-relationship hampers interpretation for WSA. CSH presents a special
-case: the limited number of sites (2) preclude comment on the
-exact PFT-wide relationship.
-
-In contrast to ENF, CRO, and DBF, for GRA the relationship between
-$\sigma \cdot uWUE$ and SWC is more linear and affects a greater
-portion of observations; about 60\% of observations. Clearly, for GRA
-soil water frequently impacts plant function and alters ET response,
-and our theory is limited for the majority of environmental
-conditions. It is therefore not surprising that our theory tested
-poorly against the data for GRA, relative to to the other PFTs
-(Section \ref{testing}). For all other PFTs occurrence of soil
-moisture impacts was rare enough to not manifest itself in bulk
-statistics and figures.
-
-\begin{figure}
-  \centering
-  \centerline{\includegraphics[width=1.2\textwidth]{./swc_boxplot.pdf}}
-  \caption{$\sigma \cdot uWUE$ as a function of PFT and SWC. Each box
-    plot represents 5\% of all observations. Note that the highest
-    10\% of SWC observations are excluded to better resolve variability in the much more narrow bins at lower SWC (e.g. SWC has long tails at high values).}
-  \label{swc_boxplot}
-\end{figure}
-
-The observed dependence of $\sigma \cdot uWUE$ on SWC for GRA would
-explain the deficiencies of our theory compared to the observations in
-Figure \ref{idealized_sign}, specifically the trend back towards
-negative soil moisture at high VPD. Aforementioned feedbacks between
-the land surface and the atmosphere, which are not accounted for due
-to our focus on the one-way response of the land surface to
-atmospheric conditions, would cause high VPD to be correlated with low
-SWC \citep[][]{Gentine_2016, Berg_2016}. So, at high VPD observations
-of low SWC are more likely, and this low SWC causes a lower uWUE
-(Figure \ref{swc_boxplot}). The lower uWUE at low SWC/high VPD then
-leads to the observed downturn towards decreasing ET with increasing
-VPD, and the deviation between our theory (based on a constant uWUE
-assumption) and observations. It is also perhaps not coincidental that
-the portion of $\sigma \cdot uWUE$ affected by low soil moisture
-observations is similar to the portion of observations that do not
-match our theory at high VPD. Indeed, this may suggest that for all PFTs
-except for GRA, coupling and feedbacks between SWC, VPD and plant
-function are relatively rare.
-Future research will explore these relatively rare feedbacks in extreme conditions, which due to analytical intractability will require more opaque numerical analysis of many more complex processes, including boundary layer growth and state and their relationship to surface layer coupling and free-tropospheric lapse rates and humidity.
-
-To summarize, our theory is limited by its inability to account for
-soil water impacts on land surface response, and feedbacks between SWC
-and VPD. Fortunately, for most PFTs SWC's effect on ET is relatively
-rare ($<$30\% of observations) and does not manifest itself in the
-majority of observations and bulk statistics. However for GRA, SWC decreases water use efficiency for the majority of
-the observations. Soil moisture effects explain the deficiencies of
-our theory in Section \ref{testing}, particularly for GRA. By
-conceptualizing SWC effects as a change in $uWUE$ (and/or $g_1$), it
-will be possible for future analysis to explore the importance of soil
-moisture on plant response to VPD, and feedbacks between plant
-function, SWC, and VPD.
-
-\subsection{Functional form of ET dependence on VPD and its relation
-  to VPD dependence}
-\label{functional_form}
-
-The theory described in Section \ref{sign_term} indicates that for a
-given $uWUE$ and $g_1$, the ET dependence on VPD should be concave
-upward, which is confirmed by eddy covariance data across most PFTs. In other words, there should be some local minimum in ET at a
-critical VPD$_{crit}$, assuming the scaling and plant terms  (e.g. aerodynamic
-conductance, $\Delta$, $g_1$ and $uWUE$) are held fixed. This result warrants further
-investigation, because to our knowledge no one has derived the
-theoretical ecosystem-scale relationship between ET and VPD while
-controlling for other environmental conditions. In particular, from
-personal communication, there is an apparent lack of consensus over
-whether the shape of the ET-VPD curve should be concave upward (our
-result) or concave downward in the absence of dramatic water stress. Given that understanding the ET-VPD
-relationship of the one-way plant response is fundamental to
-hypothesizing about any feedbacks between the land surface and the
-atmosphere, we analyze why our derived ET-VPD relationship is concave
-upward, particularly with respect to the exponent of VPD dependence in
-$uWUE$ and the Medlyn unified stomatal conductance model.
-
-There is a theoretical basis for the square root VPD dependence in
-both the stomatal conductance model and $uWUE$ based on the assumption
-that stomata behave to maximize carbon gain while minimizing water
-loss, which observations also generally support \citep{Lloyd_1991,
-  MEDLYN_2011, Lin_2015, Zhou_2014, Zhou_2015, Medlyn_2017}. However,
-some purely empirical results that fit the exponent of the VPD
-dependence to data have shown that it may vary slightly from 1/2,
-suggesting that stomata, as well as ecosystem-scale quantities based
-on stomata theory, may not always function optimally \citep{Zhou_2015,
-  Lin_2018}. Specifically with regards to $uWUE$, one would not expect
-that this ecosystem scale WUE quantity will respond to VPD exactly
-analogously to stomata. Direct soil evaporation's contributions to ET
-should shift the exponent of the VPD dependence, especially at
-conditions of low GPP when we would expect a systematically larger
-portion of direct soil evaporation contributions to ET, because we
-would also expect lower amounts of transpiration at low GPP. \citet{Zhou_2015}'s
-results corroborate this: they found a mean empirically fit
-exponential VPD dependence of 0.55, varying slightly from the
-theoretically optimal value of of 0.5 for AmeriFlux
-sites. Interpreting \citet{Lin_2018}'s results, which also show
-variance in the empirical exponent of the VPD dependence of the
-stomatal conductance model, is more difficult as \citet{Lin_2018} do
-not handle GPP/A dependence of stomatal conductance in a directly
-analogous manner to the optimal theory in \citet{MEDLYN_2011} and
-\citet{Medlyn_2017}. Regardless, given that these recent results on
-the relationship between VPD, GPP, and ET \citep{MEDLYN_2011,
-  Zhou_2014, Zhou_2015, Medlyn_2017} form the backbone of our analysis
-and are what allowed us to derive an explicit ET expression for the
-first time (Equation \ref{et}), we will analyze if and how assumptions
-about the exponent of the VPD dependence impacts the shape of the
-ET-VPD dependence.
-This analysis is also important to understand whether the choice of stomatal conductance model alters the fundamental behavior of the ET-VPD relationships, as many commonly used models  utilize a VPD exponent other than the 1/2 suggested by optimal theory \citep[e.g.
-][ which uses an exponent of 1]{Leuning_1990}.
-
-By introducing $n$ and $m$ we can free our stomatal
-conductance model from assumptions about VPD dependence:
-% \begin{linenomath*}
-  \begin{equation}
-    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_*}{VPD^m}\right) \frac{*WUE \; ET}{c_a \; VPD^n},
-    \label{m_n}
-  \end{equation}
-% \end{linenomath*}
-where:
-\[*WUE = \frac{GPP}{ET}VPD^n,\] and $g_*$ is a generic slope parameter
-of units $VPD^m$. To determine how the exponent $n$ and $m$ alter the
-shape of the ET-VPD dependence we find the roots of the second
-derivative of ET, using Equation \ref{m_n} for stomatal conductance
-($g_s$), with respect to VPD: \input{d2_solutions.tex} With this
-result we have defined the family of curves separating concave up from
-concave down ET solutions (Figure \ref{concave}). These curves are
-only functions of the exponent of the VPD dependence and a quantity we
-call non-dimensional VPD ($VPD^m/g*$). Several important
-relations reveal themselves from Equation \ref{curves}:
-
-\begin{figure}
-  \centering
-  \centerline{\includegraphics[width=0.75\textwidth]{./concave.pdf}}
-  \caption{ Solutions corresponding to inflection points between
-    concave up and concave down ET-VPD curves (Equation \ref{curves})
-    for three specific scenarios. Solutions are defined in terms of a
-    non-dimensional VPD ($VPD^m/g_*$), but to aide physical
-    interpretation the horizontal axis is additionally provided in terms of dimensionalized VPD assuming $m=1/2$ and
-    $g_*=110\; Pa^{1/2}$ (average of all PFT $g_1$). The vertical axis
-    has a different interpretation depending on the solution
-    curve. For the blue line ($m$ varying), it corresponds to $m$,
-    for the orange line ($n$ varying) it corresponds to $n$, and for
-    the green line it corresponds to the value of both $n$ and $m$
-    ($n=m$). Regions of the parameter space that correspond to
-    concave up and concave down results are labeled.}
-  \label{concave}
-\end{figure}
-
-\begin{itemize}
-  \item For optimal behavior (n, m = 1/2) the ET-VPD curve will be
-    concave up regardless of the magnitude of the plant constants
-    $g_1$ and $uWUE$. Therefore, the general concave up nature of our
-    results, given an assumption of optimal behavior, is insensitive
-    to plant type.
-  \item For all physically possible exponents of VPD dependence ($n,
-    m$), whether the solution is concave up or concave down does not
-    depend on $uWUE$.
-  \item In general, increasing the exponent of VPD dependence
-    increases the likelihood of a concave up result. Additionally, as
-    the exponent of VPD dependence increases from the optimum value of
-    1/2, whether the curve is concave upward or concave downward
-    becomes a function of the plant specific slope parameter $g_*$,
-    through non-dimensional VPD ($VPD^m/g_*$). Because the exponent of
-    the VPD dependencies is capable of altering the fundamental shape
-    of ET-VPD dependence, future research investment in understanding
-    the exact VPD dependence of stomatal conductance, and further
-    reconciliation of empirical and theoretical stomatal and
-    ecosystem behavior should be prioritized.
-\end{itemize}
-While it is possible that in the future some other form of
-VPD dependence is derived, at present \cite{MEDLYN_2011} and
-\cite{Zhou_2014} firmly established n=m=1/2 as the most likely
-candidate given current theory and empirical data. Additionally, we
-argue that a concave up result matches physical intuition more than a
-concave down result. Plants must maintain nutrient and sugar transport
-through the phloem and xylem. To accomplish this, stomata must remain
-slightly open \citep{De_2013, Nikinmaa_2013, Ryan_2014}. Furthermore,
-even if complete stomatal closure were possible, cuticular water loss
-and [at the ecosystem-scale] direct soil evaporation are still sources
-of ET which increase with VPD, independent of stomatal
-closure. Therefore, in the limit as VPD becomes large and we assume
-plants are exercising all strategies to reduce ET, any further increase
-in VPD should result in an increase in ET through cuticular water loss
-and/or direct soil evaporation. This inevitable transition from
-conditions when stomata respond strongly to VPD to conditions when
-stomata response is asymptoting towards full closure would cause a
-concave up ET-VPD curve, which is matched by the theory. In short,
-plant response becomes more limited as VPD increases, while
-atmospheric demand monotonically increases with VPD, leading to the
- result that atmospheric demand dominates plant response when
-atmospheric demand is high.
-
-This analysis allows us to understand the theoretical shape of the ET
-response to VPD with environmental conditions held
-fixed. Accomplishing this with purely statistical
-methods applied to flux observations would be very difficult, given
-the relatively fast time scale of plant response and the
-non-stationarity of [solar forced] environmental conditions over the
-relatively coarse (half hourly) flux estimates (which is required to
-obtain robust eddy-covariance statistics). Our results on the shape of
-the ET-VPD curve with environmental conditions held
-fixed can be built upon with future work examining how changes in VPD
-and environmental conditions (e.g. soil water storage) feedback upon
-one another. In the soil water storage example, over very long time
-scales extremely high VPD perturbations coupled with no precipitation
-could result in decreases in soil water storage such that water
-becomes limiting. This could be represented by an extension of our
-framework in which uWUE is allowed to decrease with decreasing SWC, as
-observed in Section \ref{swc_section}. Here, we focus our results by
-assuming constant PFT-wide conditions to build baseline intuition for
-ET-VPD dependence. For most PFTs, the theory with plant function held
-fixed matches the leading order behavior of the observations where
-plant function varies (Section \ref{testing}).
-
-\section{Conclusions}
-
-We derived a new form of Penman Monteith using the concept of
-semi-empirical optimal stomatal regulation \citep{Lin_2015,
-  MEDLYN_2011} and near constant uWUE \citep{Zhou_2015} to remove the
-implicit dependence of stomatal conductance on GPP and ET. With our
-new form of Penman Monteith we developed a theory for when an
-ecosystem will tend to reduce or increase ET with increasing VPD,
-which we evaluated against a range of eddy-covariance data spanning
-different climates and plant functional types. The goal was to capture
-the leading order behavior of the system to gain some intrinsic
-knowledge for its behavior. This intuition can be used to disentangle
-land atmosphere feedbacks in more complicated scenarios, and will aid
-interpretation of observations and sophisticated models.
-
-Our theory suggests that for a majority of environmental conditions,
-plants will tend to conserve water and reduce ET with increasing
-VPD. Stomatal regulation and plant physiological response strongly
-regulate ET, and this regulation varies by PFT. CROs are the least
-water conservative, while DBF, EBF, SAV, GRA, MF, WSA, ENF and CSH are
-progressively more water conservative (more likely to reduce ET in
-response to increasing VPD). SAV and WSA exhibit positive ET response
-to VPD not necessarily because of poor water conservation strategies
-relative to other PFTs, but because of greater occurrence of high
-atmospheric demand (VPD) relative to other PFTs. Observations of ET
-response to VPD exhibit the same general behavior as the theory, with
-ET response becoming more positive (atmospheric demand dominating) as
-environmental VPD increases within a PFT, and more negative for PFTs
-that are adapted to arid conditions and prioritize water conservation
-over primary production.
-
-The consequences of our theory our twofold. First, the ET response to
-VPD is always far more negative than it would be if only atmospheric
-demand is considered and plant physiological response is ignored (as
-with PET). Therefore, any drought indices or analysis using PET will
-not capture how vegetated surfaces respond and react to increases in
-atmospheric demand. However, the new uWUE-version of Penman Monteith
-we derived (Equation \ref{et}) could be used as a complement to PET in
-drought indices. One could calculate and compare indices when plant
-response is ignored (PET) or included (Equation \ref{et}).
-
-Lastly, our paper builds important intuition for how plants will
-respond to VPD perturbations. We show that given optimal stomatal
-function and fixed environmental conditions, the ET-VPD dependence is
-theoretically concave upward, with ET increasing with increasing VPD as
-VPD increases past some critical value (Table \ref{vpd_crit}). However
-future research should focus on fully understanding the form of
-stomatal VPD dependence, as this result is sensitive to the exponent of
-VPD dependence, which we currently believe is 1/2 \citep{MEDLYN_2011,
-  Zhou_2014}. Our results are applicable to understanding the impact
-of  expected increases in VPD
-induced by global change.  Plant physiological responses to direct
-CO$_2$ effects \citep[e.g.,][]{Swann_2016, Lemordant_2018} receives more attention
-than physiological response to indirect effects like increased
-VPD. Here, we provide broad PFT-focused results showing a likely
-decrease in ET in response to positive VPD perturbations (atmospheric
-drying), which is consistent with recent observational analysis
-\citep[e.g.,][]{Rigden_2017}. Feedbacks between the land and the
-atmosphere may alter the net response to a long-timescale global VPD perturbation,
-but our focus on the one way plant response to a VPD perturbation in
-the atmospheric boundary layer is an important first step to
-disentangling such feedbacks, both in observations and model
-simulations of the present and future. By removing Penman Monteith's
-dependence on implicit relationships between GPP, VPD, and ET, we
-allow for explicit, accurate future analysis of plant-VPD feedbacks
-in the atmospheric boundary layer (Equation \ref{et}). Our approach
-can be extended to examine varying plant response to more nuanced
-consideration of plant type and climate. Any plant physiological
-heterogeneity or feedback that can be conceptualized with shifts in $g_1$
-\citep[e.g.][]{Lin_2015, Medlyn_2017} and/or uWUE
-\citep[e.g.][]{Zhou_2014} are representable within our framework, which
-opens the door for a hierarchy of more sophisticated climate- and
-plant-specific analysis of ET sensitivity to environmental variables
-(including VPD). We argue that such simplified conceptual frameworks
-are critical tools for disentangling land-atmosphere feedbacks at
-various scales, from diurnal to seasonal and beyond, and to
-characterize ET response in a warmer, atmospherically drier, and
-enriched CO$_2$ world.
-
-\acknowledgments This work used eddy covariance data acquired and
-shared by the FLUXNET community, including these networks: AmeriFlux,
-AfriFlux, AsiaFlknux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont,
-ChinaFlux, Fluxnet-Canada, GreenGrass, ICOS, KoFlux, LBA, NECC,
-OzFlux-TERN, TCOS-Siberia, and USCCC. The ERA-Interim reanalysis data
-are provided by ECMWF and processed by LSCE. The FLUXNET eddy
-covariance data processing and harmonization was carried out by the
-European Fluxes Database Cluster, AmeriFlux Management Project, and
-Fluxdata project of FLUXNET, with the support of CDIAC and ICOS
-Ecosystem Thematic Center, and the OzFlux, ChinaFlux and AsiaFlux
-offices. This material is based upon work supported by the National
-Science Foundation Graduate Research Fellowship under Grant No. DGE
-16-44869. All code and data used in this analysis, including those
-used to generate the figures and tables, are publicly available at
-\url{https://github.com/massma/climate\_et}.
-%\hyperref[https://github.com/massma/climate\_et]{https://github.com/massma/climate_et}.
-
-\bibliography{references.bib}
-
-\end{document}
+\input{body.tex}

### 2018-05-25 src/paper_figure_code/map.py

@@ -3,6 +3,8 @@
 This script makes fig 01 for the paper
 """
 from shared_functions import *
+from mpl_toolkits.basemap import Basemap
+
 
 symbols = ['o', 'v', '>', 's', '<', 'D', 'X', 'P', '8']
 plt.close('all')
@@ -29,14 +31,12 @@ def make_map(_df):
 
   plt.legend(loc='best')
   # plt.tight_layout()
-  util.test_savefig('../../doc/paper/map.pdf')
+  plot_tools.test_savefig('../../doc/paper/map.pdf')
   return
 
 ### FIGURE 1 ###
 #make a map fig 1
 # get metadata
-sites_used = df.loc[:, 'site'].drop_duplicates()
-
 site_list = pd.read_csv('%s/changjie/fluxnet_algorithm/'\
                    'Site_list_(canopy_height).csv' % (os.environ['DATA']))
 site_list.index = site_list.Site

### 2018-05-25 src/paper_figure_code/shared_functions.py

@@ -2,21 +2,19 @@
 """
 This script makes all figs for the paper
 """
+import matplotlib as mpl
+mpl.use('Pdf')
 import os
 import importlib
 import pandas as pd
 import numpy as np
-import matplotlib as mpl
 import matplotlib.pyplot as plt
-from mpl_toolkits.basemap import Basemap
 from mpl_toolkits.axes_grid1 import make_axes_locatable
 import codebase.plot_tools as plot_tools
-import util
 import codebase.data_calc as d_calc
 import codebase.data_prep as d_prep
 import pickle
 
-mpl.use('Pdf')
 mpl.rcParams.update(mpl.rcParamsDefault)
 small_ax_params = mpl.rcParamsDefault
 # increase tick label size
@@ -28,11 +26,10 @@ fontsize=16
 single_ax_fontsize = 14
 
 
-importlib.reload(util)
 importlib.reload(plot_tools)
 importlib.reload(d_calc)
 
-# grab sites actually used in analysis
+
 with open('%s/changjie/diagnosed_data.pkl' % os.environ['DATA'],\
           mode='rb') as file:
   dfs = pickle.load(file)
@@ -114,3 +111,7 @@ def panel_wrapper(_df, function, name, args=()):
   plt.tight_layout()
   plt.savefig('../../doc/paper/%s' % name)
   return
+
+# grab sites actually used in analysis
+sites_used = df.loc[:, 'site'].drop_duplicates()
+

### 2018-05-25 doc/paper/body.tex
figs, data and citations now match data policy, just need to proofread body and then can submit v2 to arxiv

@@
 savannah (SAV), and woody savannah (WSA) (site locations and citations in Figure
-\ref{map_fig} and Table \ref{flux_sites}).
+\ref{map_fig} and Appendex \ref{flux_sites}).
 
@@
 \end{figure}
-
-\begin{table}
-  \caption{Metadata and citations for flux sites used in this analysis.
-All data are gathered from \url{www.fluxdata.org} using tools available at \url{https://github.com/trevorkeenan/FLUXNET_citations}.}
-  \label{flux_sites}
-  \label{definitions}
-  \centering \footnotesize
-  \begin{tabular}{l l l l l l l}
-    \hline
-    Site name & PFT & Lat & Lon & Clim$^{a}$ & Period & References \\
-    \hline
-    \input{flux_sites.tex}
-    \hline
-    \multicolumn{7}{l}{$^{a}$K{\"o}ppen Climate classification.}
-    \hline
-\end{table}
 
@@
 
+\begin{appendices}
+  \section{FLUXNET2015 sites}\label{flux_sites}
+  % below taken adapted from Trevor Keenan's FLUXNET citations
+  % at github.com/trevorkeenan/FLUXNET_citations
+  Table A1: Metadata and citations for flux sites used in this analysis.
+All data are gathered from \url{www.fluxdata.org}, and citations are aggregated using tools available at \url{https://github.com/trevorkeenan/FLUXNET_citations}.
+  \begin{longtable}{l l l l l l l}
+    \hline 
+    \textbf{Site} & 
+    \textbf{PFT} & 
+    \textbf{Lat} & 
+    \textbf{Lon} & 
+    \textbf{Clim\textsuperscript{1}} & 
+    \textbf{Period} & 
+    \textbf{References} \\
+    [0.5ex] 
+    \hline 
+    \input{flux_sites.tex}
+    [1ex]
+    \hline
+  \end{longtable}
+
+  \textsuperscript{1}  K{\"o}ppen Climate classification.
+\end{appendices}
+
 \bibliography{references.bib}

### 2018-05-26 doc/paper/body.tex
can't get spellcheck to work in emacs tex mode

@@
 \label{data}
-We use both meteorological and eddy-covariance data from the
-FLUXNET2015 database (data available at \sloppy
+We use both meteorological and eddy-covariance data from the FLUXNET2015
+database (data available at \sloppy
 \url{https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/} \sloppy),
-including all sites with at least four years of data and observations
-of the variables described in the methods section (Section
-\ref{methods}). Seventy-three sites met these requirements, and were
-grouped into nine plant functional types (PFT) according to the
-International Geosphere-Biosphere Programme vegetation classification
-scheme \citep{Loveland_1999}: cropland (CRO), grass (GRA), deciduous
-broadleaf forest (DBF), evergreen broadleaf forest (EBF), evergreen
-needleleaf forest (ENF), mixed forest (MF), closed shrub (CSH),
-savannah (SAV), and woody savannah (WSA) (site locations and citations in Figure
-\ref{map_fig} and Appendex \ref{flux_sites}).
+including all Tier 1 sites with at least four years of data and
+observations of the variables described in the methods section (Section
+\ref{methods}). Sixty-six sites met these requirements, and were grouped
+into nine plant functional types (PFT) according to the International
+Geosphere-Biosphere Programme vegetation classification scheme
+\citep{Loveland_1999}: cropland (CRO), grass (GRA), deciduous broadleaf
+forest (DBF), evergreen broadleaf forest (EBF), evergreen needleleaf
+forest (ENF), mixed forest (MF), closed shrub (CSH), savannah (SAV), and
+woody savannah (WSA) (site locations and citations are in Figure
+\ref{map_fig} and Appendex \ref{flux_sites}, respectively).
 
@@
 hourly averages to reduce noise \citep{Lin_2018}. After these quality
-control procedures, 434,765 upscaled hourly observations remain.
+control procedures, 400,983 upscaled hourly observations remain.
 
@@
 FLUXNET, with the support of CDIAC and ICOS Ecosystem Thematic Center,
-and the OzFlux, ChinaFlux and AsiaFlux offices. We would like to think
+and the OzFlux, ChinaFlux and AsiaFlux offices. We would like to thank
 Dr. Trevor Keenan for providing tools for citation of the FLUXNET2015
@@
   \end{longtable}
-
   \textsuperscript{1}  K{\"o}ppen Climate classification.

### 2018-05-27 doc/paper/body.tex
proofread, about to submit to arxiv

@@
 alter $uWUE$ (or $g_1$) or invalidate other assumptions made in the
-Methods section (Section \ref{methods}). We do not account for
+methods section (Section \ref{methods}). We do not account for
 variability in plant height and surface roughness, or differences in
@@
 
-Indeed, for all PFTs, there is some slight dependence of
-$\sigma \cdot uWUE$ on SWC, especially at low SWC (Figure
-\ref{swc_boxplot}). The portion of observations for which our theory
-is biased by SWC-limitations varies by PFT, due to the nonlinear
-threshold at which soil moisture availability limits plant
-function. For CRO, MF and DBF, soil water content only matters for
-about the lowest 5\% of observations (each box is 5\% of observations
-in Figure \ref{swc_boxplot}). And, CRO, relative to DBF, has a very
-weak dependence on soil moisture, which is reflective of the high
-likelihood that CRO sites are optimally irrigated and not water
-stressed, suggesting that observed departures from the theory for CRO
-are due to other factors (i.e. different photosynthesis pathways; C3/C4) than soil
-moisture. For ENF, EBF, SAV, and WSA, systematic SWC-induced biases in
-$\sigma \cdot uWUE$ emerge in about the lowest 20-30\% of SWC
-conditions, although large variability in the SWC-$(\sigma \cdot uWUE)$
-relationship hampers interpretation for WSA. CSH presents a special
-case: the limited number of sites (2) preclude comment on the
-exact PFT-wide relationship.
+Indeed, for all PFTs, there is some slight dependence of $\sigma \cdot
+uWUE$ on SWC, especially at low SWC (Figure \ref{swc_boxplot}). The
+portion of observations for which our theory is biased by
+SWC-limitations varies by PFT, due to the nonlinear threshold at which
+soil moisture availability limits plant function. For CRO, MF and DBF,
+soil water content only matters for about the lowest 5\% of observations
+(each box is 5\% of observations in Figure \ref{swc_boxplot}). And, CRO,
+relative to DBF, has a very weak dependence on soil moisture, which is
+reflective of the high likelihood that CRO sites are optimally irrigated
+and not water stressed, suggesting that observed departures from the
+theory for CRO are due to other factors (e.g. different photosynthesis
+pathways; C3/C4) than soil moisture. For ENF, EBF, SAV, and WSA,
+systematic SWC-induced biases in $\sigma \cdot uWUE$ emerge in about the
+lowest 20-30\% of SWC conditions, although large variability in the
+SWC-$(\sigma \cdot uWUE)$ relationship hampers interpretation for
+WSA. CSH presents a special case: the limited number of sites (2)
+preclude comment on the exact PFT-wide relationship.
 

### 2018-06-25 doc/paper/abstract.tex
refactored to have independent, single, abstract file.

--- 2018-05-25 doc/paper/vpd_et_paper_arxiv.tex
@@
-\documentclass[12pt]{article}
-\input{def.tex}
+Increasing vapor pressure deficit (VPD) increases atmospheric demand for
+water, and vapor pressure deficit is expected to rise with increasing
+greenhouse gases. While increased evapotranspiration (ET) in response to
+increased atmospheric demand seems intuitive, plants are capable of
+reducing ET in response to increased VPD by closing their stomata, in an
+effort to conserve water. Here we examine which effect dominates
+response to increasing VPD: atmospheric demand and increases in ET, or
+plant physiological response (stomata closure) and decreases in ET. We
+use Penman-Monteith, combined with semi-empirical optimal stomatal
+regulation theory and underlying water use efficiency, to develop a
+theoretical framework for understanding how ET responds to increases in
+VPD.
 
-\begin{document}
-
-\title{When does vapor pressure deficit drive or reduce
-  evapotranspiration?}
-
-\author[1]{Adam Massmann\thanks{Corresponding author: akm2203@columbia.edu}}
-\author[1]{Pierre Gentine}
-\author[1,2]{Changjie Lin}
-
-\affil[1]{Department of Earth and Environmental Engineering, Columbia University, New York, NY 10027}
-\affil[2]{State Key Laboratory of Hydroscience and Engineering, Department of Hydraulic
-  Engineering, Tsinghua University, Beijing, CN 100084}
-
-\maketitle
-
-\begin{abstract}
-  Increasing vapor pressure deficit increases atmospheric demand for water, and vapor pressure deficit is expected to rise with increasing greenhouse gases.
-While increased evapotranspiration (ET) in response to increased atmospheric demand seems intuitive, plants are capable of reducing ET in response to increased VPD by closing their stomata, in an effort to conserve water.
-Here we examine which effect dominates response to increasing VPD: atmospheric demand and increases in ET, or plant physiological response (stomata closure) and decreases in ET.
-We use Penman-Monteith, combined with semi-empirical optimal stomatal regulation theory and underlying water use efficiency, to develop a theoretical framework for understanding how ET responds to increases in VPD.
-
- The theory suggests that for most environmental conditions and plant types, plant physiological response dominates and ET decreases with increasing VPD.
-Plants that are evolved or bred to prioritize primary production over water conservation (e.g.
-crops) exhibit a higher likelihood of atmospheric demand-driven response (ET increasing).
-However for forest, grass, savannah, and shrub plant types, ET more frequently decreases than increases with rising VPD.
-This work serves as an example of the utility of our simplified framework for disentangling land-atmosphere feedbacks at multiple scales, including the characterization of ET response in an atmospherically drier, enriched CO$_2$ world.
-\end{abstract}
-
-\onehalfspacing
-
-\input{body.tex}
+The theory suggests that for most environmental conditions and plant
+types, plant physiological response dominates and ET decreases with
+increasing VPD. Plants that are evolved or bred to prioritize primary
+production over water conservation (e.g. crops) exhibit a higher
+likelihood of atmospheric demand-driven response (ET
+increasing). However for forest, grass, savannah, and shrub plant types,
+ET more frequently decreases than increases with rising VPD. This work
+serves as an example of the utility of our simplified framework for
+disentangling land-atmosphere feedbacks at multiple scales, including
+the characterization of ET response in an atmospherically drier,
+enriched CO$_2$ world.

### 2018-06-25 doc/paper/vpd_et_paper_afm.tex
figs, data and citations now match data policy, just need to proofread body and then can submit v2 to arxiv
refactored to have independent, single, abstract file.

@@
 \usepackage{makecell} \usepackage{multirow} %AM for tables
+\usepackage{hyperref}
+\usepackage[title]{appendix}
 \usepackage{booktabs} \drafttrue
+\usepackage{longtable}
 
@@
 \begin{abstract}
-  Increasing vapor pressure deficit increases atmospheric demand for  water, and vapor pressure deficit is expected to rise with  increasing greenhouse gases.
-While increased evapotranspiration (ET) in  response to increased atmospheric demand seems intuitive, plants are  capable of reducing ET in response to increased VPD by closing their  stomata, in an effort to conserve water.
-Here we examine which  effect dominates response to increasing VPD: atmospheric demand and  increases in ET, or plant physiological response (stomata closure)  and decreases in ET.
-We use Penman-Monteith, combined with  semi-empirical optimal stomatal regulation theory and underlying  water use efficiency, to develop a theoretical framework for  understanding how ET responds to increases in VPD.
-
-  The theory suggests that for most environmental conditions and plant  types, plant physiological response dominates and ET decreases with  increasing VPD.
-Plants that are evolved or bred to prioritize  primary production over water conservation (e.g.
-crops) exhibit a  higher likelihood of atmospheric demand-driven response (ET  increasing).
-However for forest, grass, savannah, and shrub plant  types, ET more frequently decreases than increases with rising  VPD.
-This work serves as an example of the utility of  our simplified framework for disentangling land-atmosphere feedbacks  at multiple scales, including the characterization of ET response in  an atmospherically drier, enriched CO$_2$ world.
+\input{abstract.tex}
 \end{abstract}

### 2018-10-30 src/codebase/data_io.py
starting hess submission

@@ -84,7 +84,7 @@ def tier1_test(filename):
     return True
   else:
     return False
-  
+
 def load_file(filename):
   """loads up a matlab file from changjie"""
   data = io.loadmat(filename)

### 2018-10-30 doc/paper/body.tex
starting hess submission
preparing hess submission

@@
-\section{Introduction}
-
-Vapor pressure deficit (VPD) is expected to rise over continents in
-the future due to the combination of increased temperature and,
-depending on region, decreased relative humidity
-\citep{Byrne_2013}. Increases in VPD increase the atmospheric demand
-for evapotranspirated water \citep{Penman_1948, Monteith_1965}, but
-also stress plant stomata \citep{Leuning_1990, MEDLYN_2011}.
-
-The opposing effects of increased atmospheric demand and higher
-stomatal stress lead to two possible perspectives for how
-evapotranspiration (ET) responds to shifts in VPD. The first, a
-hydrometeorological perspective, is that higher VPD increases
-atmospheric demand for water from the land surface, and this drives an
-increase in evapotranspiration (ET). This perspective is particularly
-relevant because potential evapotranspiration (PET), which is used in
-many drought indices and hydrometeorological studies
-\citep[e.g.,][]{Heim_2002, Scheff_2015}, typically only quantifies
-changes in atmospheric demand and fails to account for ecosystem
-response \citep{Swann_2016}. In reality, plants' stomata have evolved
-to optimally regulate the exchange of water and carbon, and tend to
-partially close in response to increased atmospheric dryness
-\citep{Farquhar_1978, Ball_1987, Leuning_1990, MEDLYN_2011}. This
-leads to a plant physiology perspective, in which an increase in VPD,
-particularly in well-watered soil conditions, may actually correspond
-to a decrease in ET because of stomatal closure
-\citep[e.g.][]{Rigden_2017}.  In other words, the question ``When does
-VPD drive or reduce ET?'' can be related to whether plant regulation
-or atmospheric demand dominates ET response.
-
-The ET response to changes in VPD alters water partitioning between
-the soil and atmosphere. If ecosystem plant response reduces ET with
-atmospheric drying then soil moisture will be better conserved. This
-would seem a sensible evolutionary strategy to cope with aridity. If
-stomata were fully passive \citep [similar to soil pores,
-e.g. ][]{Or_2013}, increased atmospheric aridity would strongly reduce
-soil moisture \citep{Berg_2017}. In turn, this would further increase
-aridity as low soil moisture levels increase the Bowen ratio, and
-cause increased temperature and atmospheric drying
-\citep[][]{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006,
-  Salvucci_2013, Gentine_2016, Berg_2016}. This however would not seem
-to be a sensible strategy for plants from an evolutionary
-standpoint. As a counterpoint, one may argue that increases in ET with
-increasing VPD could increase the likelihood of precipitation
-\citep[e.g.,][]{Findell_2011}. However, increases in ET do not always
-guarantee an increased likelihood of precipitation, which depending on
-environmental conditions could cause a decrease in the likelihood of
-precipitation \citep[][]{Gentine_2013}. Furthermore, any increases in
-precipitation are likely to be non-local, such that plants giving up
-water to the atmosphere are not guaranteed to reap the benefits of
-water returned from the atmosphere to the soil. Using this subjective
-logic, from an ecosystem evolutionary perspective, water stored in
-soil seems to be worth much more than the chance of water returned as
-precipitation.
-
-We can use intuition about plant water conservation strategy to
-hypothesize about ET response to changes in VPD. Plants and ecosystems
-that evolved to conserve water, such as arid shrubs or savannah,
-should be more likely to reduce ET with increasing VPD, and plants
-that have evolved or have been engineered to care little about water,
-such as crops, will be more likely to increase ET with increasing
-VPD. Atmospheric conditions must matter as well. At the ecosystem
-scale, there are limits to plant water conservation strategies. As
-atmospheric demand for water (VPD) increases, ecosystems should begin
-to reach their water conservation limits and might not be able to
-entirely limit ET flux to the atmosphere. At this stage any further
-increase in VPD will most likely drive a (limited) increase in ET,
-because the increase in atmospheric demand for water overwhelms the
-limited plant response to conserve water.
-
-The objective of the present manuscript is to use reasonable
-approximations as a tool to develop intuition for plant response to
-atmospheric drying and evaluate the VPD dependence of ET. This
-intuition will aid interpretation of observations, full complexity
-models, and facilitate the disentanglement of complex land-atmosphere
-feedbacks. In the past, similar approaches were used to understand
-interactions between stomatal conductance, evapotranspiration and the
-environment \citep[e.g.,][]{Jarvis_1986, Mcnaughton_1991}. However, at
-the time researchers' understanding of the form of VPD's effect on
-plant physiology was limited, so they could not explore the
-sensitivity of ET to VPD, including VPD's effect on stomatal
-conductance and plant function.
-
-Recent results have drastically improved our understanding of VPD's
-impact on physiology, especially at the leaf
-level. \citet{MEDLYN_2011} developed a model for leaf-scale stomatal
-conductance ($g_s$), including VPD response, by combining an optimal
-photosynthesis theory \citep{Cowan_1977} with an
-empirical approach, and extended use of this model to the ecosystem
-scale in \citet{Medlyn_2017}. Additionally, \citet{Zhou_2014}
-demonstrated that a quantity underlying water use efficiency
-$\left(uWUE = \frac{GPP\; \sqrt{VPD}}{ET}\right)$ properly captures a
-constant relationship between GPP, ET, and VPD over a diurnal cycle at
-the ecosystem scale. uWUE is also remarkably well conserved in the
-growing season across space and time, within a PFT
-\citep{Zhou_2015}. These results provide novel tools to examine how ET
-varies in response to VPD, while explicitly including VPD's effect on
-stomatal conductance for the first time. In this manuscript, we
-leverage these tools to derive the theoretical one-way response of ET
-to VPD with other environmental variables properly controlled for,
-i.e. we develop a framework for evaluating the partial derivative of
-ET with respect to VPD. Our theory is validated and tested at multiple
-eddy-covariance stations spanning various climates and plant
-functional types.
-
 \section{Materials and Methods}
@@
 
-\section{Conclusions}
-
-We derived a new form of Penman Monteith using the concept of
-semi-empirical optimal stomatal regulation \citep{Lin_2015,
-  MEDLYN_2011} and near constant uWUE \citep{Zhou_2015} to remove the
-implicit dependence of stomatal conductance on GPP and ET. With our
-new form of Penman Monteith we developed a theory for when an
-ecosystem will tend to reduce or increase ET with increasing VPD,
-which we evaluated against a range of eddy-covariance data spanning
-different climates and plant functional types. The goal was to capture
-the leading order behavior of the system to gain some intrinsic
-knowledge for its behavior. This intuition can be used to disentangle
-land atmosphere feedbacks in more complicated scenarios, and will aid
-interpretation of observations and sophisticated models.
-
-Our theory suggests that for a majority of environmental conditions,
-plants will tend to conserve water and reduce ET with increasing
-VPD. Stomatal regulation and plant physiological response strongly
-regulate ET, and this regulation varies by PFT. CROs are the least
-water conservative, while DBF, EBF, SAV, GRA, MF, WSA, ENF and CSH are
-progressively more water conservative (more likely to reduce ET in
-response to increasing VPD). SAV and WSA exhibit positive ET response
-to VPD not necessarily because of poor water conservation strategies
-relative to other PFTs, but because of greater occurrence of high
-atmospheric demand (VPD) relative to other PFTs. Observations of ET
-response to VPD exhibit the same general behavior as the theory, with
-ET response becoming more positive (atmospheric demand dominating) as
-environmental VPD increases within a PFT, and more negative for PFTs
-that are adapted to arid conditions and prioritize water conservation
-over primary production.
-
-The consequences of our theory our twofold. First, the ET response to
-VPD is always far more negative than it would be if only atmospheric
-demand is considered and plant physiological response is ignored (as
-with PET). Therefore, any drought indices or analysis using PET will
-not capture how vegetated surfaces respond and react to increases in
-atmospheric demand. However, the new uWUE-version of Penman Monteith
-we derived (Equation \ref{et}) could be used as a complement to PET in
-drought indices. One could calculate and compare indices when plant
-response is ignored (PET) or included (Equation \ref{et}).
-
-Lastly, our paper builds important intuition for how plants will
-respond to VPD perturbations. We show that given optimal stomatal
-function and fixed environmental conditions, the ET-VPD dependence is
-theoretically concave upward, with ET increasing with increasing VPD as
-VPD increases past some critical value (Table \ref{vpd_crit}). However
-future research should focus on fully understanding the form of
-stomatal VPD dependence, as this result is sensitive to the exponent of
-VPD dependence, which we currently believe is 1/2 \citep{MEDLYN_2011,
-  Zhou_2014}. Our results are applicable to understanding the impact
-of  expected increases in VPD
-induced by global change.  Plant physiological responses to direct
-CO$_2$ effects \citep[e.g.,][]{Swann_2016, Lemordant_2018} receives more attention
-than physiological response to indirect effects like increased
-VPD. Here, we provide broad PFT-focused results showing a likely
-decrease in ET in response to positive VPD perturbations (atmospheric
-drying), which is consistent with recent observational analysis
-\citep[e.g.,][]{Rigden_2017}. Feedbacks between the land and the
-atmosphere may alter the net response to a long-timescale global VPD perturbation,
-but our focus on the one way plant response to a VPD perturbation in
-the atmospheric boundary layer is an important first step to
-disentangling such feedbacks, both in observations and model
-simulations of the present and future. By removing Penman Monteith's
-dependence on implicit relationships between GPP, VPD, and ET, we
-allow for explicit, accurate future analysis of plant-VPD feedbacks
-in the atmospheric boundary layer (Equation \ref{et}). Our approach
-can be extended to examine varying plant response to more nuanced
-consideration of plant type and climate. Any plant physiological
-heterogeneity or feedback that can be conceptualized with shifts in $g_1$
-\citep[e.g.][]{Lin_2015, Medlyn_2017} and/or uWUE
-\citep[e.g.][]{Zhou_2014} are representable within our framework, which
-opens the door for a hierarchy of more sophisticated climate- and
-plant-specific analysis of ET sensitivity to environmental variables
-(including VPD). We argue that such simplified conceptual frameworks
-are critical tools for disentangling land-atmosphere feedbacks at
-various scales, from diurnal to seasonal and beyond, and to
-characterize ET response in a warmer, atmospherically drier, and
-enriched CO$_2$ world.
-
 \acknowledgments This work used eddy covariance data acquired and shared

### 2018-10-30 doc/paper/conclusions.tex
preparing hess submission

--- 2018-05-14 doc/paper/vpd_et_paper.tex
@@
-\documentclass[12pt]{article}
-\input{def.tex}
-
-\begin{document}
-
-\title{When does vapor pressure deficit drive or reduce
-  evapotranspiration?}
-
-\author[1]{Adam Massmann\thanks{Corresponding author: akm2203@columbia.edu}}
-\author[1]{Pierre Gentine}
-\author[1,2]{Changjie Lin}
-
-\affil[1]{Department of Earth and Environmental Engineering, Columbia University, New York, NY 10027}
-\affil[2]{State Key Laboratory of Hydroscience and Engineering, Department of Hydraulic
-  Engineering, Tsinghua University, Beijing, CN 100084}
-
-\maketitle
-
-\begin{abstract}
-  Increasing vapor pressure deficit increases atmospheric demand for water, and vapor pressure deficit is expected to rise with increasing greenhouse gases.
-While increased evapotranspiration (ET) in response to increased atmospheric demand seems intuitive, plants are capable of reducing ET in response to increased VPD by closing their stomata, in an effort to conserve water.
-Here we examine which effect dominates response to increasing VPD: atmospheric demand and increases in ET, or plant physiological response (stomata closure) and decreases in ET.
-We use Penman-Monteith, combined with semi-empirical optimal stomatal regulation theory and underlying water use efficiency, to develop a theoretical framework for understanding how ET responds to increases in VPD.
-
- The theory suggests that for most environmental conditions and plant types, plant physiological response dominates and ET decreases with increasing VPD.
-Plants that are evolved or bred to prioritize primary production over water conservation (e.g.
-crops) exhibit a higher likelihood of atmospheric demand-driven response (ET increasing).
-However for forest, grass, savannah, and shrub plant types, ET more frequently decreases than increases with rising VPD.
-This work serves as an example of the utility of our simplified framework for disentangling land-atmosphere feedbacks at multiple scales, including the characterization of ET response in an atmospherically drier, enriched CO$_2$ world.
-\end{abstract}
-
-\onehalfspacing
-
-\section{Introduction}
-
-Vapor pressure deficit (VPD) is expected to rise over continents in
-the future due to the combination of increased temperature and,
-depending on region, decreased relative humidity
-\citep{Byrne_2013}. Increases in VPD increase the atmospheric demand
-for evapotranspirated water \citep{Penman_1948, Monteith_1965}, but
-also stress plant stomata \citep{Leuning_1990, MEDLYN_2011}.
-
-The opposing effects of increased atmospheric demand and higher
-stomatal stress lead to two possible perspectives for how
-evapotranspiration (ET) responds to shifts in VPD. The first, a
-hydrometeorological perspective, is that higher VPD increases
-atmospheric demand for water from the land surface, and this drives an
-increase in evapotranspiration (ET). This perspective is particularly
-relevant because potential evapotranspiration (PET), which is used in
-many drought indices and hydrometeorological studies
-\citep[e.g.,][]{Heim_2002, Scheff_2015}, typically only quantifies
-changes in atmospheric demand and fails to account for ecosystem
-response \citep{Swann_2016}. In reality, plants' stomata have evolved
-to optimally regulate the exchange of water and carbon, and tend to
-partially close in response to increased atmospheric dryness
-\citep{Farquhar_1978, Ball_1987, Leuning_1990, MEDLYN_2011}. This
-leads to a plant physiology perspective, in which an increase in VPD,
-particularly in well-watered soil conditions, may actually correspond
-to a decrease in ET because of stomatal closure
-\citep[e.g.][]{Rigden_2017}.  In other words, the question ``When does
-VPD drive or reduce ET?'' can be related to whether plant regulation
-or atmospheric demand dominates ET response.
-
-The ET response to changes in VPD alters water partitioning between
-the soil and atmosphere. If ecosystem plant response reduces ET with
-atmospheric drying then soil moisture will be better conserved. This
-would seem a sensible evolutionary strategy to cope with aridity. If
-stomata were fully passive \citep [similar to soil pores,
-e.g. ][]{Or_2013}, increased atmospheric aridity would strongly reduce
-soil moisture \citep{Berg_2017}. In turn, this would further increase
-aridity as low soil moisture levels increase the Bowen ratio, and
-cause increased temperature and atmospheric drying
-\citep[][]{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006,
-  Salvucci_2013, Gentine_2016, Berg_2016}. This however would not seem
-to be a sensible strategy for plants from an evolutionary
-standpoint. As a counterpoint, one may argue that increases in ET with
-increasing VPD could increase the likelihood of precipitation
-\citep[e.g.,][]{Findell_2011}. However, increases in ET do not always
-guarantee an increased likelihood of precipitation, which depending on
-environmental conditions could cause a decrease in the likelihood of
-precipitation \citep[][]{Gentine_2013}. Furthermore, any increases in
-precipitation are likely to be non-local, such that plants giving up
-water to the atmosphere are not guaranteed to reap the benefits of
-water returned from the atmosphere to the soil. Using this subjective
-logic, from an ecosystem evolutionary perspective, water stored in
-soil seems to be worth much more than the chance of water returned as
-precipitation.
-
-We can use intuition about plant water conservation strategy to
-hypothesize about ET response to changes in VPD. Plants and ecosystems
-that evolved to conserve water, such as arid shrubs or savannah,
-should be more likely to reduce ET with increasing VPD, and plants
-that have evolved or have been engineered to care little about water,
-such as crops, will be more likely to increase ET with increasing
-VPD. Atmospheric conditions must matter as well. At the ecosystem
-scale, there are limits to plant water conservation strategies. As
-atmospheric demand for water (VPD) increases, ecosystems should begin
-to reach their water conservation limits and might not be able to
-entirely limit ET flux to the atmosphere. At this stage any further
-increase in VPD will most likely drive a (limited) increase in ET,
-because the increase in atmospheric demand for water overwhelms the
-limited plant response to conserve water.
-
-The objective of the present manuscript is to use reasonable
-approximations as a tool to develop intuition for plant response to
-atmospheric drying and evaluate the VPD dependence of ET. This
-intuition will aid interpretation of observations, full complexity
-models, and facilitate the disentanglement of complex land-atmosphere
-feedbacks. In the past, similar approaches were used to understand
-interactions between stomatal conductance, evapotranspiration and the
-environment \citep[e.g.,][]{Jarvis_1986, Mcnaughton_1991}. However, at
-the time researchers' understanding of the form of VPD's effect on
-plant physiology was limited, so they could not explore the
-sensitivity of ET to VPD, including VPD's effect on stomatal
-conductance and plant function.
-
-Recent results have drastically improved our understanding of VPD's
-impact on physiology, especially at the leaf
-level. \citet{MEDLYN_2011} developed a model for leaf-scale stomatal
-conductance ($g_s$), including VPD response, by combining an optimal
-photosynthesis theory \citep{Cowan_1977} with an
-empirical approach, and extended use of this model to the ecosystem
-scale in \citet{Medlyn_2017}. Additionally, \citet{Zhou_2014}
-demonstrated that a quantity underlying water use efficiency
-$\left(uWUE = \frac{GPP\; \sqrt{VPD}}{ET}\right)$ properly captures a
-constant relationship between GPP, ET, and VPD over a diurnal cycle at
-the ecosystem scale. uWUE is also remarkably well conserved in the
-growing season across space and time, within a PFT
-\citep{Zhou_2015}. These results provide novel tools to examine how ET
-varies in response to VPD, while explicitly including VPD's effect on
-stomatal conductance for the first time. In this manuscript, we
-leverage these tools to derive the theoretical one-way response of ET
-to VPD with other environmental variables properly controlled for,
-i.e. we develop a framework for evaluating the partial derivative of
-ET with respect to VPD. Our theory is validated and tested at multiple
-eddy-covariance stations spanning various climates and plant
-functional types.
-
-\section{Materials and Methods}
-\subsection{Data}
-\label{data}
-We use both meteorological and eddy-covariance data from the
-FLUXNET2015 database (data available at \sloppy
-\url{https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/} \sloppy),
-including all sites with at least four years of data and observations
-of the variables described in the methods section (Section
-\ref{methods}). Seventy-three sites met these requirements, and were
-grouped into nine plant functional types (PFT) according to the
-International Geosphere-Biosphere Programme vegetation classification
-scheme \citep{Loveland_1999}: cropland (CRO), grass (GRA), deciduous
-broadleaf forest (DBF), evergreen broadleaf forest (EBF), evergreen
-needleleaf forest (ENF), mixed forest (MF), closed shrub (CSH),
-savannah (SAV), and woody savannah (WSA) (site locations in Figure
-\ref{map_fig}).
-
-\begin{figure}[h]
-  \centering \includegraphics[trim={0 3cm 0 3cm}, clip]{./map.pdf}
-  \caption{Plant functional type and location of FLUXNET2015 sites
-    used in this analysis.}
-  \label{map_fig}
-\end{figure}
-
-The purpose of this study is to examine ecosystem response to
-atmospheric drying, focusing on the growing season. To accomplish
-this, we filter and quality control the data using a similar procedure
-as \cite{Zhou_2015}:
-\begin{itemize}
-\item Only measured or highest (``good'') quality gapfilled data,
-  according to quality control flags, are used.
-\item To isolate the growing season, we only use days in which the
-  average Gross Primary Productivity (GPP) exceeds 10\% of the
-  observed 95th percentile of GPP for a given site. GPP is calculated
-  using the nighttime respiration partitioning method.
-\item We remove days with rain and the day following to avoid issues
-  with rain interception and sensor saturation at high relative
-  humidity (\cite{Medlyn_2017}).
-\end{itemize}
-Additionally, as in \citet{Lin_2018}, we restrict data to the daytime,
-which is identified when downwelling shortwave radiation is greater
-than 50 W m$^{-2}$ and sensible heat flux is greater than 5 W
-m$^{-2}$. To reduce the chance of sensor saturation at high relative
-humidity, we remove all time steps for which VPD is less than .01 kPa,
-and to reduce errors at low windspeeds we remove all periods with wind
-magnitudes less than 0.5 m s$^{-1}$. Timesteps with negative observed
-GPP or ET are also removed, and we aggregate half hourly data to
-hourly averages to reduce noise \citep{Lin_2018}. After these quality
-control procedures, 434,765 upscaled hourly observations remain.
-
-\subsection{Methods}
-\label{methods}
-The Penman-Monteith equation \citep [hereafter PM,][]{Penman_1948,
-  Monteith_1965} estimates ET as a function of observable atmospheric
-variables and surface conductances:
-% \begin{linenomath*}
-  \begin{equation}
-    \label{orig_pen}
-    ET = \frac{\Delta R_{net} + g_a \rho_a c_p VPD}{\Delta + \gamma(1 + \frac{g_a}{g_s})},
-  \end{equation}
-% \end{linenomath*}
-where $\Delta$ is the change in saturation vapor pressure with
-temperature, given by Clausius-Clapeyron ($\frac{d \; e_s}{d \; T}$),
-$R_{net}$ is the net radiation minus ground heat flux, $g_a$ is
-aerodynamic conductance, $\rho_a$ is air density, $c_p$ is specific
-heat of air at constant pressure, $\gamma$ is the psychometric
-constant, and $g_s$ is the stomatal conductance (Table
-\ref{definitions}).
-
-\begin{table}
-  \caption{Definition of symbols and variables, with citation for how
-    values are calculated, if applicable.}
-  \label{definitions}
-  \centering \footnotesize
-  \begin{tabular}{l c c c}
-    \hline
-    Variable & Description & Units & Citation \\
-    \hline
-    $e_s$  & saturation vapor pressure & Pa  & - \\ 
-    $T$  & temperature  & K & - \\
-    $P$  & pressure & Pa  & - \\
-    $\Delta$  & $\frac{\partial e_s}{\partial T}$ & Pa K$^{-1}$ & - \\
-    $R_{net}$  & net radiation at land surface minus ground heat flux & W m$^{-2}$   & - \\
-    $g_a$  & aerodynamic conductance & m s$^{-1}$  & \citet{Shuttleworth_2012} \\
-    $\rho_a$  & air density & kg m$^{-3}$  & - \\
-    $c_p$  & specific heat capacity of air at constant pressure & J K$^{-1}$ kg$^{-1}$ & - \\
-    $VPD$  & vapor pressure deficit & Pa  & - \\
-    $\gamma$  & psychometric constant & Pa K$^{-1}$   & - \\
-    $g_{s-leaf}$  & leaf-scale stomatal conductance & m s$^{-1}$  & \citet{MEDLYN_2011} \\
-    $g_{s}$  &  stomatal conductance & m s$^{-1}$
-                                   & \citet{Medlyn_2017} \\
-    $g_{1-leaf}$  & leaf-scale slope parameter & Pa$^{0.5}$
-                                   & \citet{MEDLYN_2011} \\
-    $g_{1}$  & ecosystem-scale slope parameter & Pa$^{0.5}$ & \citet{Medlyn_2017} \\
-    $R$ & universal gas constant & J mol$^{-1}$ K$^{-1}$ & - \\
-    $R_{air}$ & gas constant of air & J  K$^{-1}$ kg$^{-1}$ & - \\
-    $\sigma$ & uncertainty parameter & -& - \\
-    $c_a$ & CO$_2$ concentration & $\mu$ mol CO$_2$ mol$^{-1}$ air& - \\
-    $\lambda$ & marginal water cost of leaf carbon & mol H$_2$O mol$^{-1}$ CO$_2$ & - \\
-    $\Gamma$ & CO$_2$ compensation point & - & - \\
-    $\Gamma^*$ & CO$_2$ compensation point without dark respiration & - & - \\
-    \hline
-  \end{tabular}
-\end{table}
-
-\citet{MEDLYN_2011} developed a model for stomatal conductance ($g_s$)
-by combining an optimal photosynthesis theory \citep{Cowan_1977} with an empirical approach, which describes the
-dependence of $g_s$ to VPD. This resulted in the following model for
-leaf-scale stomatal conductance:
-
-% \begin{linenomath*}
-  \begin{equation}
-    g_{s-leaf} = g_0 + 1.6 \left(1 +
-      \frac{g_{1-leaf}}{\sqrt{VPD}}\right) \frac{A}{c_a},
-    \label{leaf_medlyn}
-  \end{equation}
-% \end{linenomath*}
-where $g_{1-leaf}$ is a leaf-scale ``slope'' parameter, A is the net
-CO$_2$ assimilation rate, and $c_a$ is the atmospheric CO$_2$
-concentration at the leaf surface. \cite{MEDLYN_2011} relate the slope
-parameter ($g_{1-leaf}$) to physical parameters as:
-% \begin{linenomath*}
-  \label{slope}
-  \begin{equation}
-    g_{1-leaf} = \sqrt{\frac{3 \, \Gamma^* \, \lambda}{1.6}},\footnote{Note this expression has units of of (mmol mol$^{-1}$)$^{1/2}$, but this can be converted to Pa$^{1/2}$ using the ideal gas law.}
-  \end{equation}
-% \end{linenomath*}
-
-where $\Gamma^*$ is the CO$_2$ compensation point for photosynthesis
-(without dark respiration), and $\lambda$ is the marginal water cost
-of leaf carbon
-($\frac{\partial \; \text{transpiration}}{\partial \; A}$). So,
-$g_{1-leaf}$ is a leaf-scale term reflecting the trade-off of water for
-carbon uptake. The higher $g_{1-leaf}$, the more open the stomata and
-the more they release water in exchange for carbon.
-
-The Medlyn model for stomatal conductance has been shown to behave
-very well across PFTs \citep[][]{Lin_2015}, and has been successfully
-adopted to ecosystem scale analysis in \citet{Medlyn_2017}. In units
-of m s$^{-1}$, the ecosystem scale stomatal conductance is:
-
-% \begin{linenomath*}
-  \begin{equation}
-    g_s = \frac{R \,T}{P} \; 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{GPP}{c_a},
-    \label{medlyn}
-  \end{equation}
-% \end{linenomath*}
-
-where GPP is the ecosystem scale gross primary production, and $g_1$
-is an ecosystem scale analogue to $g_{1-leaf}$. We solve for $g_1$
-following \citet{Medlyn_2017} (Equation 5), and take the median $g_1$
-value to be representative of each PFT (Table \ref{pft}) instead of the mean to avoid extra weighting of rare outliers.
-
-\begin{table}
-  \caption{Plant functional types, their abbreviation, calculated
-    Medlyn coefficient, and calculated uWUE. uWUE from
-    \citet{Zhou_2015}, including the observed standard deviation, is
-    shown for comparison. Note that uWUE from \citet{Zhou_2015} is
-    calculated from a different set of sites, and that units are
-    converted such that the quantities work with Equations 1-8 and the
-    variables defined Table \ref{definitions}.}
-  \small
-  \label{pft}
-  \centering
-  \begin{tabular}{l c c @{\qquad} c c}
-    \hline
-    \multirow{2}[3]{*}{Abbreviation} & \multirow{2}[3]{*}{PFT} & \multirow{2}[3]{*}{$g_1$ (Pa$^{0.5}$)} & \multicolumn{2}{c}{uWUE ($\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET])}  \\
-    \cmidrule{4-5}
-
-                                     & & & fitted & \citet{Zhou_2015} \\
-
-    \hline
-    \input{pft_params.tex}
-    \hline
-  \end{tabular}
-\end{table}
-
-While Equation 3 can be used in PM (Equation 1), it will make
-analytical work with the function intractable because $GPP$ is
-functionally related to ET itself. Additionally, a perturbation to VPD
-should induce a physiological plant response that will alter GPP and
-cause an indirect change in stomatal conductance, in addition to the
-direct effect of VPD in Equation \ref{medlyn}. Therefore, in order to
-derive the response of ET to VPD, we must account for the functional
-relationship between GPP, ET, and VPD, and its effect on stomatal
-conductance. We can use aforementioned semi-empirical results of
-\citet{Zhou_2015} as a tool to approach this
-problem. \citet{Zhou_2015}, showed that  underlying Water Use
-Efficiency (uWUE):
-
-% \begin{linenomath*}
-  \begin{equation}
-    uWUE = \frac{GPP \cdot \sqrt{VPD}}{ET}
-    \label{uwue}
-  \end{equation}
-% \end{linenomath*}
-
-is relatively constant across time and moisture conditions within a
-plant functional type, and correctly captures a constant relationship
-between GPP, ET and VPD over a diurnal cycle \citep{Zhou_2014}. The
-theoretical derivation of the square root VPD dependence in $uWUE$
-leverages the same assumptions used in \cite{MEDLYN_2011} to derive the
-square-root VPD dependence of the stomatal conductance model (Equation
-\ref{medlyn}).  We can use uWUE to remove the $GPP$ dependence from
-$g_s$ in a way that makes PM analytically tractable:
-
-% \begin{linenomath*}
-  \begin{equation}
-    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{uWUE \; ET}{c_a \; \sqrt{VPD}}.
-    \label{new_g_s}
-  \end{equation}
-% \end{linenomath*}
-
-Plugging Equation \ref{new_g_s} into Equation \ref{orig_pen} and
-rearranging gives a new explicit expression for PM, in which
-dependence on $GPP$ is removed:
-
-% \begin{linenomath*}
-  \begin{equation}
-    ET = \frac{\Delta R_{net} + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R \; 1.6\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
-    \label{et}
-  \end{equation}
-% \end{linenomath*}
-
-By accounting for photosynthesis changes in ecosystem conductance, with
-Equation \ref{et} we have derived for the first time, using recent
-results \citep[][]{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017}, ET
-explicitly as function of environmental variables and two
-plant-specific constants, the slope parameter ($g_1$), and uWUE, both
-reflecting water conservation strategy. The slope parameter is related
-to the willingness of stomata to trade water for CO$_2$ and to keep
-stomata open. uWUE is a semi-empirical ecosystem-scale constant
-related to how WUE changes with VPD (specifically $VPD^{-1/2}$). It is
-also roughly proportional to physical constants:
-
-\[uWUE \appropto \sqrt{\frac{c_a - \Gamma}{1.6 \lambda}},\]
-
-where $\Gamma$ is the CO$_2$ compensation point \citep[Equation 5
-in][]{Zhou_2014}. So uWUE is related to atmospheric CO$_2$
-concentration and compensation point, and is inversely proportional to
-the marginal water cost of leaf carbon.
-
-Given eddy-covariance FLUXNET2015 data (Section \ref{data}), every
-term in our new version of PM (Equation \ref{et}) is observed, except
-for uWUE, which we fit by calculating its expectation, given the model
-and FLUXNET2015 data.
-
-However, eddy-covariance data are inherently noisy so we include a
-measure of uncertainty in our analysis. To account for observational
-error, as well as model uncertainty (e.g. temporal and spatial
-variations of uWUE and $g_1$), we introduce an uncertainty parameter
-$\sigma$ modifying uWUE:
-
-% \begin{linenomath*}
-  \begin{equation}
-    ET = \frac{\Delta R_{net} + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R \; 1.6\; \sigma \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
-    \label{et_sigma}
-  \end{equation}
-% \end{linenomath*}
-
-Now, from each FLUXNET2015 observation (i.e. for each hourly
-observation at every time step) we can evaluate $\sigma$:
-
-% \begin{linenomath*}
-  \begin{equation}
-    \sigma = - \frac{g_a \gamma c_a \sqrt{VPD} L_v P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R_{net} - g_a \rho_a c_p VPD\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})},
-    \label{sigma}
-  \end{equation}
-% \end{linenomath*}
-
-Thus, with this uncertainty analysis we can evaluate departure from
-our theory in observations, as a departure of $\sigma$ from unity. The
-variability of $\sigma$ across sites and time provides a measure of
-uncertainty in our model, assumptions, as well as in the FLUXNET2015
-observations themselves. The variability of $\sigma$ then propagates
-through any uncertainty to our partial derivative of Equation
-\ref{et_sigma} with respect to VPD:
-
-% \begin{linenomath*}
-  \begin{equation}
-    \frac{\partial \;  ET}{\partial \; VPD} = \frac{2\; g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_a }{1.6 \; R\; \sigma \; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)
-    \label{d_et}
-  \end{equation}
-% \end{linenomath*}
-
-With Equation \ref{d_et} we have an analytical framework for ecosystem
-response to atmospheric demand perturbations with environmental
-conditions held fixed. There are a few subtleties to taking the
-derivative in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$)
-and $VPD$ are functionally related, so while taking the derivative we
-evaluate
-$\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET}
-{\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD}
-\Big|_{\text{RH fixed}} + \frac{\partial \; ET}{\partial \; RH}
-\frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$
-    fixed}}$. $RH$ and $e_s$ are assumed to be approximately
-independent, which is supported by the data (not shown).
-
-This derivation relied either implicitly or explicitly on several
-assumptions. First, we assume that VPD at the leaf surface is the same
-as VPD at measurement height; physically this implies that leaves are
-perfectly coupled to the atmosphere. In reality, for some conditions
-and plant types the leaves can become decoupled from the boundary
-layer \citep{De_2017, Medlyn_2017}. Given our focus on the growing
-season, which is usually characterized by relatively high insolation
-inducing instability and convective boundary layers, we would expect
-the surface to be generally well coupled. An additional assumption in
-the formulation of $uWUE$ \citep{Zhou_2014, Zhou_2015} and
-\citet{Medlyn_2017}'s stomatal conductance model is that direct soil
-evaporation (E) contributions to ET remain small relative to
-transpiration (T). This should be more true during the growing
-season. The ratio of E to T may increase immediately after rainfall
-events due to high soil moisture and ponding, but VPD is generally low
-anyways during these times. However, some plant types allow for
-systematically larger contributions of E in ET, particularly those
-with sparse canopies and smaller relative amounts of transpiration. We
-therefore might expect that the theory will be most applicable to
-forest PFTs, which will be most strongly coupled to the boundary layer
-due to larger surface roughness, and will also generally have the
-highest ratios of transpiration to evaporation. Finally, for the goal
-of developing PFT-wide intuition, we assume that $g_1$ and $uWUE$ are
-constant in space and time. Both quantities have been shown to be well
-conserved over space and time within a PFT relative to inter-PFT
-variability. However, we would expect them to vary somewhat with
-environmental conditions including very low soil water content,
-temperature, and with intra-PFT plant-specific characteristics like
-wood density for tree PFTs \citep{Lin_2015}. Given these strong
-assumptions made with the goal of understanding broad, leading order
-plant behaviors, we take the extra care of including and analyzing
-uncertainty (Section \ref{testing}) and examining sensitivity to the
-exponent of the VPD response (Section \ref{functional_form}).
-
-We note one final comment on our derivation which is relevant for
-drought indices. If we approximate $c_a$ at a global mean CO$_2$
-concentration, then the RHS of Equation \ref{et} is fully defined
-using commonly available weather station data and the constants
-published here. This makes Equation \ref{et} a useful alternative to
-PET in drought indices and hydrometeorological analysis for vegetated
-surfaces. Equation \ref{et} better reflects the physics of water
-exchange at the land surface and would only require fitting of uWUE
-and $g_1$, and can also account for changes in CO$_2$ concentration
-(assuming some relationship between $\lambda$ and [CO$_2$]), contrary
-to typical drought indices such as the Palmer Drought Severity Index
-(PDSI) \citep{Swann_2016, Lemordant_2016, Lemordant_2018}. Lastly, all
-code and data used in this analysis, including those used to generate
-the figures and tables, are publicly available at
-\url{https://github.com/massma/climate\_et}.
-%\hyperref[https://github.com/massma/climate\_et]{https://github.com/massma/climate_et}.
-
-\section{Results and Discussion}
-\label{results}
-
-By construction, the variability in the $\sigma$ term (Equation
-\ref{sigma}) contains all model and observational uncertainties. For
-an observation that perfectly matches our model and constant uWUE
-assumption, $\sigma$ will be one. Therefore, for our assumptions and
-framework to be reasonable $\sigma$ should be close to 1. An
-additional concern is that $\sigma$ may in fact be correlated with
-$VPD$, in which case the dependence would need to be accounted for
-when taking the derivative. Fortunately, there is a very weak
-dependence of $\sigma$ on VPD in their joint distribution, and
-$\sigma$ is indeed close to unity i.e. $O(1)$ (Figure
-\ref{joint_vpd_sigma}). Given this weak dependence and the
-distribution of $\sigma$ we have confidence in our model framework and
-the data quality.
-
-\begin{figure}[h]
-  \centering\includegraphics[width=0.75\textwidth]{./joint_vpd_sigma.pdf}
-  \caption{The joint distribution of $VPD$ and $\sigma$, with outliers
-    removed (defined as lowest and highest 5\% of $\sigma$). $\sigma$
-    exhibits a weak dependence on $VPD$, and $\sigma$ is $O(1)$ for
-    the bulk of the observations.}
-  \label{joint_vpd_sigma}
-\end{figure}
-
-Before calculating the sensitivity of ET to VPD, we will consider the
-functional form of Equation \ref{d_et}. There are two main terms: a
-``scaling'' term, which modifies the magnitude but not the sign of the
-ET response to VPD ($\frac{\partial \; ET}{\partial \; VPD}$):
-
-\begin{equation}
-  \frac{g_a \; P}{T(\Delta + \gamma)},
-\end{equation}
-
-and a ``sign'' term, which determines whether ET increases or
-decreases with VPD (i.e. atmospheric demand driven or physiologically
-controlled):
-
-\begin{equation}
-  \label{sign}
-  \frac{c_p}{R_{air}} - \frac{\gamma c_a }{1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right).
-\end{equation}
-All variables are positive, so the relative magnitude between the
-first term and the second term in the sign term (Equation \ref{sign})
-will determine whether ET increases or decreases with increasing
-VPD. If the second term is larger then plant control dominates and ET
-decreases with increasing VPD. However, if the first term is larger,
-then atmospheric demand dominates and ET increases with increasing
-VPD.
-
-\subsection{Functional Form of the Sign Term}
-\label{sign_term}
-First, we explore the variables within the sign term to gain better
-intuition on the driver of either the increase or reduction of ET with
-VPD. CO$_2$ concentration ($c_a$) and the psychometric constant
-($\gamma$) are relatively constant over the dataset considered here so
-that the variability is dominated by $\sigma$ and $VPD$. uWUE could
-vary with soil moisture but has been shown to be relatively constant
-\citep{Zhou_2015}. This then means that the sign term only depends on
-VPD for a given PFT and is approximately just a function of $VPD$. We
-can further determine a critical threshold separating an increase from
-a decrease in ET, i.e. the threshold $VPD_{crit}$ such that the
-derivative vanishes $\frac{\partial \; ET}{\partial \; VPD} = 0$:
-\small
-% \begin{linenomath*}
-  \begin{equation}
-    VPD_{crit} = \frac{R_{air}}{4 c_p} \left( \frac{\gamma c_a}{1.6\; R \;  uWUE} + \sqrt{\frac{\gamma c_a}{1.6\; R \;  uWUE}\left( \frac{\gamma c_a}{1.6\; R \;  uWUE} + 8 g_1 \frac{c_p}{R_{air}}\right)} - 4 g_1 \frac{c_p}{R_{air}} \right),
-    \label{vpd_min_et}
-  \end{equation}
-% \end{linenomath*}
-\normalsize noting that $VPD_{crit}$ mostly depends on the PFT
-parameters uWUE and $g_1$, and only varies weakly with climate as most
-other parameters related to the environment are nearly constant. The
-calculated value of $VPD_{crit}$ for each PFT is shown in Table
-\ref{vpd_crit}. For any values of $VPD$ less than $VPD_{crit}$, ET
-will decrease with increasing VPD
-($\frac{\partial \; ET}{\partial \; VPD} < 0$), and for values of
-$VPD$ greater than $VPD_{crit}$, ET will increase with increasing VPD
-($\frac{\partial \; ET}{\partial \; VPD} > 0$). In other words,
-ecosystems regulate and mitigate evaporative losses up to the VPD
-limit, $VPD_{crit}$, above which atmospheric demand is just too high
-to be entirely compensated by stomatal and ecosystem regulation. We
-note however that even though ET increases again above the critical
-threshold, $VPD_{crit}$, ET is still much lower than potential
-evaporation as stomata are still strongly regulating vapor fluxes to
-the atmosphere. Indeed, even in the absence of soil pore evaporation,
-stomata do not shut down entirely at very high VPD so that ET does not
-go to zero, because stomata are still slightly open to perform some
-photosynthesis \citep{Ball_1987, Leuning_1990, MEDLYN_2011}. In
-addition, upward xylem transport is necessary to maintain phloem
-transport, as well as nutrient transport and thus carbon allocation
-\citep{De_2013, Nikinmaa_2013, Ryan_2014}.
-
-\begin{table}
-  \caption{Values of $VPD_{crit}$, where
-    $\frac{\partial \; ET}{\partial \; VPD} = 0$, evaluated at PFT
-    average values for $R_{air}$, $\gamma$, and $c_a$. PFT-specific
-    constants ($g_1$, uWUE) are provided in Table \ref{pft}. For
-    values of $VPD$ less than $VPD_{crit}$,
-    $\frac{\partial \; ET}{\partial \; VPD}$ will be negative, and for
-    values of $VPD$ greater than $VPD_{crit}$,
-    $\frac{\partial \; ET}{\partial \; VPD}$ will be positive.}
-  \centering
-  \begin{tabular}{l c c c c c}
-    \hline
-    PFT & $R_{air}$ & $c_a$ (ppm) & $\gamma$  & \textbf{$VPD_{crit}$ (Pa)} \\
-    \hline
-    \input{vpd_crit.tex}
-    \hline
-  \end{tabular}
-  \label{vpd_crit}
-\end{table}
-
-Differences in $VPD_{crit}$ are exclusively determined by uWUE and the
-slope parameter ($g_1$) related to the plant functional type. A larger
-uWUE means a smaller $VPD_{crit}$, and an ET response to increases VPD
-that is more likely to be positive. At first glance this result is
-somewhat counter-intuitive; we expect that plants with a higher water
-use efficiency would be more water conservative. However, in reality
-uWUE determines how $WUE$ changes with VPD:
-
-\[WUE = \frac{GPP}{ET} = \frac{uWUE}{\sqrt{VPD}}.\]
-\[\frac{\partial \; WUE}{\partial \; VPD} = -\frac{uWUE}{2 \;
-    VPD^{3/2}}\]
-
-So, plants with a higher uWUE will have a greater \textit{decrease} in
-ecosystem-scale $WUE$ in response to increases in VPD. This decrease
-in $WUE$ causes more water loss per unit carbon gain, and explains the
-relationship between high uWUE and high likelihood of increases of ET
-in response to increasing atmospheric drying (increases in VPD).
-
-A tendency towards increasing ET response with increasing VPD can also
-be caused by a high slope parameter ($g_1$), characteristic of plants
-that at the leaf scale are more willing to trade water for access to
-atmospheric CO$_2$. Plants that are less conservative will be thus be
-more likely to increase ET with increasing VPD. Both the
-aforementioned effects (large uWUE, $g_1$) can amplify each other, and
-generally conspire to shift the sign term towards a positive value for
-a given PFT.
-
-This effect of uWUE and $g1$ on the sign term is most apparent by
-comparing two extreme PFTs: water intensive crops (CRO) and water
-conservative closed shrub (CSH). CRO has higher slope parameter and a
-slightly higher uWUE ($g_1 = 140.7$ Pa$^{1/2}$; $2.85$ $\mu$-mol [C]
-Pa$^{0.5}$ J$^{-1}$ [ET]) compared to CSH ($g1 = 75.1 \, Pa^{1/2}$,
-$uWUE=2.82$ $\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET]). These differences
-in PFT parameters cause opposite ET responses to changes in VPD
-between CRO and CSH. ET theoretically always decreases with increasing
-VPD for the more water conservative CSH, while ET frequently increases
-with increasing VPD for the more water intensive CRO (Figure
-\ref{idealized_sign}). CROs evolved or were bred to prioritize GPP and
-yield and are thus not water conservative. They are very willing to
-trade water for photosynthesis and productivity, despite changes in
-VPD, while CSH are very unwilling to trade water for more
-photosynthesis.
-
-\begin{figure}
-  \centering \includegraphics{./idealized_sign.pdf}
-  \caption{The functional form of the sign term, with $\sigma$ held
-    fixed at 1, and all terms except VPD set to PFT averages. For
-    comparison, the observed range of VPD for each PFT is plotted
-    below the x-axis. Stars denote 25th, 50th, and 75th percentiles,
-    and the range of the line spans the 5th-95th percentiles of
-    observed VPD. Vertical black lines denote the location of
-    $VPD_{crit}$ for each PFT, with the exception of CSH and WSA, for which
-    $VPD_{crit}$ is off-scale.}
-  \label{idealized_sign}
-\end{figure}
-
-As expected, the slope parameter ($g_1$) is a primary determinant of
-the VPD dependence for the sign term shown in Figure
-\ref{idealized_sign}. Plants that are more conservative (small $g_1$)
-will tend to reduce ET with increasing VPD, and will be very effective
-at reducing ET, especially at low VPD. However, at very high VPD,
-gradients in vapor pressure at the leaf scale will become very strong
-as stomata reach their limits of closure in response to VPD
-(parameterized with $g_1$). As a result, ET response will begin to
-asymptote towards a constant ecosystem-scale values as leaf-scale
-response to VPD asymptotes towards zero.  Therefore, plants with a low
-$g_1$ will have the largest VPD dependence of ET response because the
-difference in ET response at low VPD (leaf stomatal response
-dominates) and high VPD (VPD gradient dominates) is largest. This is
-apparent in the strong VPD dependence of CSH, which has the lowest
-slope parameter ($g_1=75.1$ Pa$^{1/2}$) (Figure \ref{idealized_sign}).
-
-To summarize our theoretical insights (Figure \ref{idealized_sign} and
-Table \ref{vpd_crit}), CROs are the least water conservative and have
-the strongest overall tendency to increase ET with increasing VPD,
-while CSH are the most water conservative and have the strongest
-tendency to decrease ET with increasing VPD, as well as the strongest
-VPD dependence of response. Figure \ref{idealized_sign} clearly shows,
-according to our theory, that for all PFTs except for crops there is
-frequent occurrence of a negative (plant dominating) ET response to
-increases in VPD. Therefore, plants are able in most atmospheric
-conditions to reduce ET in response to increased VPD and thus to
-reduce water loss. To better illustrate this, the ranges of observed
-environmental VPDs at the FLUXNET sites are plotted parallel to the
-x-axis. For CSH and WSA, VPD is always less than VPD$_{crit}$ (off scale) so
-that the plant response dominates in typical environmental conditions,
-emphasizing the water conservative strategy of those plants. For CRO
-on the other hand, VPD is higher than VPD$_{crit}$ for more than 50\%
-of observations, emphasizing that those plants operate with an
-aggressive water usage strategy, are water intensive and were actually
-engineered for photosynthesis rather than water saving. For DBF, EBF,
-MF, GRA, and SAV more than half of the observed VPD are less than
-VPD$_{crit}$, i.e. in conditions where plant response dominates. SAV
-has a more water conservative response than the forest, grass, and
-crop plan types, but still responds by increasing ET with increasing
-VPD for about a quarter of observations, due to the high aridity (VPD)
-of the SAV ecoclimate. It is also important to note that for all PFTs,
-even when atmospheric demand dominates, ET response to VPD is still
-far more negative than it would be for potential evaporation
-$\partial PET/\partial VPD$, i.e. atmospheric demand only, emphasizing
-that there is still a strong regulation of evaporative flux by stomata
-and though the plant xylem. The sign term in the PET case would just
-be a constant ($\frac{c_p}{R_{air}} \approx 3.5$), which is far larger
-than any part of the curves for any PFT. This highlights the
-deficiencies of PET's ability to capture land response to changes in
-atmospheric dryness (VPD). Plants are always regulating water exchange
-from the land surface, even when they reach the limits of they ability
-to do so.
-
-\subsection{Functional Form of the Scaling Term}
-\label{scale_term}
-While the above discussion of the sign of
-$\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our
-question of when ET response increases or decreases with VPD,
-understating the overall magnitude of the ET response is important to
-soil-plant-atmosphere water budgeting. So we now more closely examine
-the terms that affect how the sign term is scaled:
-
-\begin{equation}
-  \frac{g_a \; P}{T(\Delta + \gamma)}.
-\end{equation}
-
-$\frac{P}{T}$ is an air-density term, which varies little compared to
-aerodynamic conductance and Clausius-Clapeyron ($\Delta$). The
-psychometric constant ($\gamma$) is also relatively constant, so the
-scaling term should be primarily a function of aerodynamic conductance
-and temperature, through the Clausius-Clapeyron relationship
-$\Delta$. This is as expected, given that the aerodynamic conductance
-represents the efficiency of exchange between the surface and the
-atmosphere. As aerodynamic conductance increases, any plant response
-will be communicated more strongly to the atmosphere (and vice-versa).
-
-$\Delta$'s presence in the scaling term also matches physical
-intuition. $\Delta$ (and also the approximately constant $\gamma$)
-control the efficiencies with which surface energy is converted to
-latent and sensible heat \citep{Monteith_1965}. The functional from of
-$\Delta$ will be the same across PFTs, but the temperature range may
-vary slightly. In contrast, aerodynamic conductance will vary strongly
-with PFT due to the importance of surface roughness for aerodynamic
-conductance. So most of the differences in scaling between PFT should
-be in the aerodynamic conductance term.
-
-The control of the scaling term variability between PFTs by
-aerodynamic conductance is confirmed by data (Figure
-\ref{scale_vary}). Differences between PFT are almost entirely due to
-differences in aerodynamic conductance, rather than differences in
-observed temperature ranges. The scaling term for the tree PFTs (DBF,
-EBF, ENF, MF) is generally about double the scaling terms for other
-PFTs which have lower surface roughness and generally smaller
-aerodynamic conductance (GRA, CSH, CRO). The savannah (WSA, SAV) PFT's
-scaling is somewhere between GRA, CSH, and CRO, and DBF, EBF, ENF, and
-MF, due to higher variability and surface roughness.
-
-\begin{figure}
-  % \centering
-  \centerline{\includegraphics[width=1.2\textwidth]{./idealized_scale.pdf}}
-  \caption{Primary sources of variability for the scaling term, as a
-    function of PFT. The 5th-95th percentile range of temperature is
-    plotted at the 5th, 25th, 50th, 75th, and 95th percentiles of
-    aerodynamic conductance, as observed for each PFT.}
-  \label{scale_vary}
-\end{figure}
-
-Within each PFT, the scaling term variability is controlled both by
-environmental temperature and aerodynamic conductance variability
-(Figure \ref{scale_vary}). While the observed variability of the
-aerodynamic conductance contributes more to the scaling term
-variability than temperature, the temperature contribution is
-non-negligible. Specifically, the scaling term is generally larger at
-low temperatures when latent heat is relatively inefficient at moving
-energy away from the surface. This effect amplifies the role of
-aerodynamic conductance variability at low temperatures.
-
-To summarize, variability between PFTs is mostly controlled by
-systematic differences in aerodynamic conductance, due to differences
-in surface roughness between each PFT, and possibly to a lesser extent
-wind conditions. In contrast, variability within PFT is also
-controlled by temperature, through Clausius-Clapeyron. But,
-aerodynamic conductance variability generally impacts the scaling term
-more than temperature, even within PFTs.
-
-\subsection{Bulk statistics of ET response to VPD}
-\label{stats_sec}
-
-In this section we consider direct observations of ET response with
-eddy-covariance data, while including uncertainty with the $\sigma$
-term (Section \ref{methods}). These observational results of ET
-response (Table \ref{stats}) largely confirm our theoretical analysis, presented in the previous sections. For all PFTs, mean ET response to increasing VPD
-is negative. However, ET response evaluated at the average of all
-variables (e.g. $\sigma$, $T$, $c_a$, $VPD$) is positive for CRO, and
-negative for all other PFTs. This difference in mean ET response as
-compared to the ET response at mean environmental conditions is due to
-the non-linear nature of the response, in which negative responses are
-generally larger magnitude than positive responses (Figure
-\ref{idealized_sign}). Therefore, both the mean ET response as well as
-the ET response at mean environmental conditions matches our
-expectations from the theory (Section \ref{sign_term}), with the
-exception that CRO observations are shifted more towards a negative ET
-response than we expect.
-
-\begin{table}
-  \caption{Statistics of $\frac{\partial \; ET}{\partial \; VPD}$ as a
-    function of PFT.}
-  \centering
-  \begin{tabular}{l c c c}
-    \hline
-    PFT & $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ & $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)$ & fraction $\frac{\partial \; ET}{\partial \; VPD} < 0.$ \\
-    \hline
-    \input{stats.tex}
-    \hline
-  \end{tabular}
-  \label{stats}
-\end{table}
-
-Regarding the frequency of negative and positive ET response, all PFTs
-exhibit a decreasing ET response with increasing VPD (physiologically
-controlled, water conservative response) for the majority of
-observations. The more water conservative PFTs generally exhibit
-higher frequency of negative ET response, especially when one factors
-in the distribution of environmental VPD (e.g. SAV and WSA grow in
-more arid climates, Figure \ref{idealized_sign}). In general the bulk
-statistics match our theoretical expectations well, with the caveat
-that inclusion of uncertainty shifts crops towards a slightly more
-negative response to VPD, and shifts many of the other PFTs, which still
-exhibit a high frequency of negative ET response to VPD, towards more frequent
-occurrence of positive ET response than the theory and Figure
-\ref{idealized_sign} would suggest. The bulk statistics motivate a
-more thorough examination of the structure of uncertainty and more
-sophisticated validation of our theory's performance against
-observations.
-
-\subsection{Validation of theory at eddy-covariance sites}
-\label{testing}
-We now compare more sophisticated distributions of the observed
-response to our simplified theory (Section \ref{sign_term}). The
-observed distribution of the sign term, as compared to what the theory
-would predict, is provided in Figure \ref{test_sign}. Our goal was to
-capture the leading order behavior of the ET dependence on VPD. Given
-the assumptions we made, and the uncertainties of flux tower
-observations themselves, we expect a relatively large amount of noise
-when reproducing the derivatives of ET. However, the data largely
-reproduces our theoretical analysis.
-
-\begin{figure}
-  \centering
-  \centerline{\includegraphics[width=1.2\textwidth]{./test_sign.pdf}}
-  \caption{Comparison of the sign term with model uncertainty included
-    (box plots) to the sign term as calculated with simplifying
-    assumptions (blue line, as in Figure \ref{idealized_sign}). Each
-    box plot corresponds to 5\% of the data, and the 5-95\% range of
-    VPD is plotted.}
-  \label{test_sign}
-\end{figure}
-
-This is particularly true for DBF and MF; the theory matches the
-leading order behavior of the function when uncertainty is included,
-and the observations match the theory with the addition of noise. The
-VPD dependence, given by the slope parameter ($g_1$), follows the
-median values of each bin. Perhaps most importantly, the x-intercept,
-and thus VPD$_{crit}$, matches nearly exactly between the theory and
-the observations. Therefore the sign of the ET response to increases
-in VPD should be well matched, subject to the unavoidable constraints
-of noise, much of which comes from the observations themselves. The
-uncertainty is non negligible; there are many observations in each bin
-for which the the sign of observation is opposite the response
-predicted by the theory, but to leading order our theory matches the
-observations well.
-
-While CSH has a much different functional form of the sign term than
-DBF and MF, CSH observations also match our theory to leading order,
-albeit with a bit more variability as a function of VPD. Again, the
-VPD dependence mostly determined from the slope parameter ($g_1$)
-closely matches the medians in the observation bins. The
-VPD-independent, strongly negative response is also captured. For CSH,
-there is rare occurrence of observed positive ET response with VPD
-($\approx 6$\%), even with uncertainty, so the sign of the
-observations almost always matches the sign of the theory, which
-states that ET response should always be negative.
-
-Biases between the theory and observations are similar for ENF, WSA,
-CRO and SAV. At low VPD the theory and the observations match
-well. However, at high VPD (upper 10-20\% of observations) the theory
-is biased slightly towards positive response as compared to the
-observations (Figure \ref{test_sign}). For ENF this slight bias could
-be explained by a negative bias in $g_1$. However, for WSA, CRO, and
-SAV the observations in the highest VPD bins exhibit a downturn
-towards more negative response, which cannot be captured by the
-functional form of the sign term (Figure
-\ref{idealized_sign}). Therefore, to explain the observations one of
-the variables in the sign term must change at extreme VPD (upper
-10-20\%). The most likely candidate is $uWUE$, which we have assumed
-constant to meet our goal of developing intuition for leading order
-behavior, but might be expected to decrease at extremely low SWC. This extremely low SWC should be correlated with high VPD in very dry environments
-through sensible heating increases \citep{Gentine_2016}. The theory
-for EBF exhibits similar limitations as for ENF, WSA, and SAV, except
-a larger portion of observations are biased negative as compared to
-the theory ($\approx$ 35\%, Figure \ref{test_sign}). However, in
-general, and specifically for non-extreme VPD (VPD  $< \approx$70-90th
-percentiles), the theory matches the observations for the tree and
-savannah plant types well.
-
-The theory for GRA suffers from similar, but much more severe,
-limitations as for CRO, WSA, and SAV. GRA observations are
-characterized by a consistent trend back towards negative ET response
-at higher VPD, which the functional form of our theory is incapable of
-accounting for. As compared to CRO, WSA and SAV, the divergence
-between the theory and the observations is far greater for GRA,
-biasing 40-50\% of observations at higher VPD. In addition for the
-potential for soil moisture to alter $uWUE$, there are other sources
-of plant heterogeneity specific to GRA (and some extent CRO) that may
-alter $uWUE$ (or $g_1$) or invalidate other assumptions made in the
-Methods section (Section \ref{methods}). We do not account for
-variability in plant height and surface roughness, or differences in
-C3 vs C4 photosynthesis and water strategies, which we might expect to
-vary substantially across sites, years, and season for GRA. These
-deficiencies could largely explain the inability of our theory to
-exactly match the observations in croplands. For example, a
-superposition of sites with C3 plants (at low environmental VPD) and
-C4 plants (at high environmental VPD) would explain the observed shift
-back towards negative ET response at high VPD when all sites are
-binned together, as in Figure \ref{test_sign}. We hypothesize that the
-theory would validate against observations much better if these
-sources of variability were accounted for, at a cost of increased
-complexity and analytical opacity.
-
-While the above discussion shows that our theory has some limitations
-when applied to some PFTs, especially for grasslands, it does well for
-DBF, MF and CSH PFTs, and captures the response at non-extreme VPD
-(VPD $< \approx$ 70th-90th percentiles) for
-ENF, EBF, CRO, WSA, and SAV. In general, the leading order behavior
-observed in the data is captured by the theory. Departures between our
-theory and observations, specifically at high VPD, could be explained
-and conceptualized with shifts in $g_1$ and/or $uWUE$ due to
-site-specific plant type variability (e.g. more arid-adapted ecosystems
-at more arid sites), or temporal variability for some environmental
-conditions (e.g. decreases in $uWUE$ at extremely low soil
-moisture). To focus on general behavior and develop intuition for
-PFT-scale response, we have ignored these sources of variability in
-the present analysis. However, any site-specific or temporal plant
-functional variability that can be conceptualized with shifts to $g_1$
-and $uWUE$ can be analyzed with our framework. This opens a door to future analyses in which plant behavior in anomalous conditions
-can be explained and analyzed using Equation \ref{et}.
-
-% It is useful to also compare our theory, the observed uncertainty in
-% the ET response, and PET, as PET is often used as a proxy for ET in
-% hydrometeorological studies and drought indices. Again, as in our
-% discussion of the theoretical results, the PET sensitivity to VPD
-% will be constant at about 3.5. In Figure \ref{test_sign} the ET
-% response to shifts in dryness is always far smaller than 3.5 in the
-% observations, even with all uncertainty included. So the PET
-% response to atmospheric dryness is outside the range of all
-% observable responses. This casts doubt on PET's usefulness for
-% capturing the leading order behavior of water budget and dryness
-% response at the land surface, as is done in several drought
-% indices. However our theory (Equation \ref{et}) could be used with
-% published uWUE and $g1$ data, as an alternative to PET to estimating
-% land response and dryness, and would more accurately account for the
-% response to atmospheric drying.
-
-\subsection{Observed ET response to VPD}
-
-Most of the results presented so far focus on the sign term, so now we
-turn to observations of ET response with the scaling term included
-(Figure \ref{data_scatter}). Until now, in the interest of developing
-leading order intuition for ET response, and to be conservative in our
-acknowledgment of model and observational error, we've considered
-$\sigma$ variability to be a measure of uncertainty. An alternative
-viewpoint is that $\sigma \cdot uWUE$ represents spatial and temporal
-variability of uWUE, which may be expected within bounds \citep[see
-Table \ref{pft}, also ][]{Zhou_2015}. This is a less conservative
-view; some of the $\sigma \cdot uWUE$ variability will be due to model
-and observational error, so by viewing $\sigma \cdot uWUE$ as
-\textit{real} variability we run the risk of mistaking noise for
-signal. However, the advantage of this viewpoint is that, from
-\citet{Zhou_2014}, we have very high confidence that uWUE fit at the
-hourly timescale (as we do with $\sigma \cdot uWUE$) correctly
-captures the relationship between ET, GPP, and VPD, and that our form
-of PM introduces minimal error with its use of uWUE.
-
-\begin{figure}
-  \centering
-  \centerline{\includegraphics[width=1.2\textwidth]{./data_scatter.png}}
-  \caption{Scatter plots of observed
-    $\frac{\partial \; ET}{\partial \; VPD}$, including $\sigma$
-    variability, as a function of PFT, temperature and VPD. Please
-    note differences in the colorbar scale.}
-  \label{data_scatter}
-\end{figure}
-
-Therefore, with the caveat that some of the signal presented in Figure
-\ref{data_scatter} may in fact be noise, we can interpret the observed
-distribution of ET response to VPD. In general, the observed response
-matches the intuitive theory. ET response to VPD shifts towards
-positive values as VPD increases (atmospheric demand dominating). CRO
-exhibit the highest occurrence of positive ET response, and the
-observations confirm that CRO are the most water intensive. CSH are
-the most water conservative, with a strong negative response. DBF,
-EBF, ENF, MF, SAV and WSA are also water conservative, but show some
-occurrence of positive ET response to VPD, particularly at higher
-observed VPD, as the theory predicts. GRA, while generally water
-conservative, does not match the theory well, with increasing frequency of
-\textit{negative} response at high VPD. This is as expected, given
-previous discussion on how we might expect more inter-site and inter-year variability for GRA (Section \ref{testing}).
-
-Figure \ref{data_scatter} also includes the impact of the scaling
-term. For a given VPD, the magnitude of ET response does not vary
-strongly with temperature, confirming that any impact of the scaling
-term on the magnitude of the response is primarily due to changes in
-aerodynamic conductance. Intuitively this is reasonable; aerodynamic
-conductance will control how dominant balances at the land surface are
-communicated to the boundary layer. While $\Delta$ controls the
-efficiency of energy conversion to latent heat, it appears this is a
-second order term, relative to $g_a$, for scaling ET response.
-
-As with Figure \ref{test_sign}, Figure \ref{data_scatter} matches our
-expectation based on simplified theory. The sign term is most strongly
-scaled by $g_a$, and in general the occurrence of positive ET response
-increases as VPD increases. The willingness with which a given plant type
-evolved to use water dictates the occurrence of positive versus
-negative ET response. Water conservative ecosystems are highly effective
-at mitigating the effects of atmospheric demand, and can store water
-for later use by reducing ET in response to increasing VPD.
-
-\subsection{Limitations of theory - very dry soil moisture conditions}
-\label{swc_section}
-In formulating our theory with Penman Monteith, we implicitly did not
-account for very dry soil moisture conditions. For the majority of
-environmental conditions observed at the eddy covariance sites used
-here, soil conditions were not extremely dry so that we could assume a
-constant uWUE and $g_1$. We posit that ecosystems will generally
-optimize to host plants living in conditions which they evolved
-for. However, in extreme conditions and drought scenarios soil water
-content (SWC) could become the limiting factor for ET response to VPD,
-which our theory does not account for. In addition, low soil moisture
-conditions themselves increase VPD through land-atmosphere feedback
-\citep[][]{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006,
-  Salvucci_2013, Gentine_2016, Berg_2016}.
-
-Within our framework, any systematic bias due to the
-failure to account for SWC's effects in very dry conditions should
-manifest itself in a functional relationship between $\sigma$, our
-uncertainty measure, and SWC, which is observed at the FLUXNET
-sites. Examining the relationship between $\sigma$ and SWC will test
-to what extent our theory breaks down in very dry soil moisture
-conditions.
-
-If we again view $\sigma \cdot uWUE$ as, in addition to a measure of
-uncertainty, a short time-scale observation of uWUE, we would expect
-$\sigma \cdot uWUE$ to decrease at low SWC. If $\sigma \cdot uWUE$ is
-a very strong function of SWC, then our theory should be conditioned
-more strongly on well watered soil conditions. If $\sigma \cdot uWUE$
-is weakly a function of SWC, then our theory is more universal and
-independent of soil moisture conditions.
-
-Indeed, for all PFTs, there is some slight dependence of
-$\sigma \cdot uWUE$ on SWC, especially at low SWC (Figure
-\ref{swc_boxplot}). The portion of observations for which our theory
-is biased by SWC-limitations varies by PFT, due to the nonlinear
-threshold at which soil moisture availability limits plant
-function. For CRO, MF and DBF, soil water content only matters for
-about the lowest 5\% of observations (each box is 5\% of observations
-in Figure \ref{swc_boxplot}). And, CRO, relative to DBF, has a very
-weak dependence on soil moisture, which is reflective of the high
-likelihood that CRO sites are optimally irrigated and not water
-stressed, suggesting that observed departures from the theory for CRO
-are due to other factors (i.e. different photosynthesis pathways; C3/C4) than soil
-moisture. For ENF, EBF, SAV, and WSA, systematic SWC-induced biases in
-$\sigma \cdot uWUE$ emerge in about the lowest 20-30\% of SWC
-conditions, although large variability in the SWC-$(\sigma \cdot uWUE)$
-relationship hampers interpretation for WSA. CSH presents a special
-case: the limited number of sites (2) preclude comment on the
-exact PFT-wide relationship.
-
-In contrast to ENF, CRO, and DBF, for GRA the relationship between
-$\sigma \cdot uWUE$ and SWC is more linear and affects a greater
-portion of observations; about 60\% of observations. Clearly, for GRA
-soil water frequently impacts plant function and alters ET response,
-and our theory is limited for the majority of environmental
-conditions. It is therefore not surprising that our theory tested
-poorly against the data for GRA, relative to to the other PFTs
-(Section \ref{testing}). For all other PFTs occurrence of soil
-moisture impacts was rare enough to not manifest itself in bulk
-statistics and figures.
-
-\begin{figure}
-  \centering
-  \centerline{\includegraphics[width=1.2\textwidth]{./swc_boxplot.pdf}}
-  \caption{$\sigma \cdot uWUE$ as a function of PFT and SWC. Each box
-    plot represents 5\% of all observations. Note that the highest
-    10\% of SWC observations are excluded to better resolve variability in the much more narrow bins at lower SWC (e.g. SWC has long tails at high values).}
-  \label{swc_boxplot}
-\end{figure}
-
-The observed dependence of $\sigma \cdot uWUE$ on SWC for GRA would
-explain the deficiencies of our theory compared to the observations in
-Figure \ref{idealized_sign}, specifically the trend back towards
-negative soil moisture at high VPD. Aforementioned feedbacks between
-the land surface and the atmosphere, which are not accounted for due
-to our focus on the one-way response of the land surface to
-atmospheric conditions, would cause high VPD to be correlated with low
-SWC \citep[][]{Gentine_2016, Berg_2016}. So, at high VPD observations
-of low SWC are more likely, and this low SWC causes a lower uWUE
-(Figure \ref{swc_boxplot}). The lower uWUE at low SWC/high VPD then
-leads to the observed downturn towards decreasing ET with increasing
-VPD, and the deviation between our theory (based on a constant uWUE
-assumption) and observations. It is also perhaps not coincidental that
-the portion of $\sigma \cdot uWUE$ affected by low soil moisture
-observations is similar to the portion of observations that do not
-match our theory at high VPD. Indeed, this may suggest that for all PFTs
-except for GRA, coupling and feedbacks between SWC, VPD and plant
-function are relatively rare.
-Future research will explore these relatively rare feedbacks in extreme conditions, which due to analytical intractability will require more opaque numerical analysis of many more complex processes, including boundary layer growth and state and their relationship to surface layer coupling and free-tropospheric lapse rates and humidity.
-
-To summarize, our theory is limited by its inability to account for
-soil water impacts on land surface response, and feedbacks between SWC
-and VPD. Fortunately, for most PFTs SWC's effect on ET is relatively
-rare ($<$30\% of observations) and does not manifest itself in the
-majority of observations and bulk statistics. However for GRA, SWC decreases water use efficiency for the majority of
-the observations. Soil moisture effects explain the deficiencies of
-our theory in Section \ref{testing}, particularly for GRA. By
-conceptualizing SWC effects as a change in $uWUE$ (and/or $g_1$), it
-will be possible for future analysis to explore the importance of soil
-moisture on plant response to VPD, and feedbacks between plant
-function, SWC, and VPD.
-
-\subsection{Functional form of ET dependence on VPD and its relation
-  to VPD dependence}
-\label{functional_form}
-
-The theory described in Section \ref{sign_term} indicates that for a
-given $uWUE$ and $g_1$, the ET dependence on VPD should be concave
-upward, which is confirmed by eddy covariance data across most PFTs. In other words, there should be some local minimum in ET at a
-critical VPD$_{crit}$, assuming the scaling and plant terms  (e.g. aerodynamic
-conductance, $\Delta$, $g_1$ and $uWUE$) are held fixed. This result warrants further
-investigation, because to our knowledge no one has derived the
-theoretical ecosystem-scale relationship between ET and VPD while
-controlling for other environmental conditions. In particular, from
-personal communication, there is an apparent lack of consensus over
-whether the shape of the ET-VPD curve should be concave upward (our
-result) or concave downward in the absence of dramatic water stress. Given that understanding the ET-VPD
-relationship of the one-way plant response is fundamental to
-hypothesizing about any feedbacks between the land surface and the
-atmosphere, we analyze why our derived ET-VPD relationship is concave
-upward, particularly with respect to the exponent of VPD dependence in
-$uWUE$ and the Medlyn unified stomatal conductance model.
-
-There is a theoretical basis for the square root VPD dependence in
-both the stomatal conductance model and $uWUE$ based on the assumption
-that stomata behave to maximize carbon gain while minimizing water
-loss, which observations also generally support \citep{Lloyd_1991,
-  MEDLYN_2011, Lin_2015, Zhou_2014, Zhou_2015, Medlyn_2017}. However,
-some purely empirical results that fit the exponent of the VPD
-dependence to data have shown that it may vary slightly from 1/2,
-suggesting that stomata, as well as ecosystem-scale quantities based
-on stomata theory, may not always function optimally \citep{Zhou_2015,
-  Lin_2018}. Specifically with regards to $uWUE$, one would not expect
-that this ecosystem scale WUE quantity will respond to VPD exactly
-analogously to stomata. Direct soil evaporation's contributions to ET
-should shift the exponent of the VPD dependence, especially at
-conditions of low GPP when we would expect a systematically larger
-portion of direct soil evaporation contributions to ET, because we
-would also expect lower amounts of transpiration at low GPP. \citet{Zhou_2015}'s
-results corroborate this: they found a mean empirically fit
-exponential VPD dependence of 0.55, varying slightly from the
-theoretically optimal value of of 0.5 for AmeriFlux
-sites. Interpreting \citet{Lin_2018}'s results, which also show
-variance in the empirical exponent of the VPD dependence of the
-stomatal conductance model, is more difficult as \citet{Lin_2018} do
-not handle GPP/A dependence of stomatal conductance in a directly
-analogous manner to the optimal theory in \citet{MEDLYN_2011} and
-\citet{Medlyn_2017}. Regardless, given that these recent results on
-the relationship between VPD, GPP, and ET \citep{MEDLYN_2011,
-  Zhou_2014, Zhou_2015, Medlyn_2017} form the backbone of our analysis
-and are what allowed us to derive an explicit ET expression for the
-first time (Equation \ref{et}), we will analyze if and how assumptions
-about the exponent of the VPD dependence impacts the shape of the
-ET-VPD dependence.
-This analysis is also important to understand whether the choice of stomatal conductance model alters the fundamental behavior of the ET-VPD relationships, as many commonly used models  utilize a VPD exponent other than the 1/2 suggested by optimal theory \citep[e.g.
-][ which uses an exponent of 1]{Leuning_1990}.
-
-By introducing $n$ and $m$ we can free our stomatal
-conductance model from assumptions about VPD dependence:
-% \begin{linenomath*}
-  \begin{equation}
-    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_*}{VPD^m}\right) \frac{*WUE \; ET}{c_a \; VPD^n},
-    \label{m_n}
-  \end{equation}
-% \end{linenomath*}
-where:
-\[*WUE = \frac{GPP}{ET}VPD^n,\] and $g_*$ is a generic slope parameter
-of units $VPD^m$. To determine how the exponent $n$ and $m$ alter the
-shape of the ET-VPD dependence we find the roots of the second
-derivative of ET, using Equation \ref{m_n} for stomatal conductance
-($g_s$), with respect to VPD: \input{d2_solutions.tex} With this
-result we have defined the family of curves separating concave up from
-concave down ET solutions (Figure \ref{concave}). These curves are
-only functions of the exponent of the VPD dependence and a quantity we
-call non-dimensional VPD ($VPD^m/g*$). Several important
-relations reveal themselves from Equation \ref{curves}:
-
-\begin{figure}
-  \centering
-  \centerline{\includegraphics[width=0.75\textwidth]{./concave.pdf}}
-  \caption{ Solutions corresponding to inflection points between
-    concave up and concave down ET-VPD curves (Equation \ref{curves})
-    for three specific scenarios. Solutions are defined in terms of a
-    non-dimensional VPD ($VPD^m/g_*$), but to aide physical
-    interpretation the horizontal axis is additionally provided in terms of dimensionalized VPD assuming $m=1/2$ and
-    $g_*=110\; Pa^{1/2}$ (average of all PFT $g_1$). The vertical axis
-    has a different interpretation depending on the solution
-    curve. For the blue line ($m$ varying), it corresponds to $m$,
-    for the orange line ($n$ varying) it corresponds to $n$, and for
-    the green line it corresponds to the value of both $n$ and $m$
-    ($n=m$). Regions of the parameter space that correspond to
-    concave up and concave down results are labeled.}
-  \label{concave}
-\end{figure}
-
-\begin{itemize}
-  \item For optimal behavior (n, m = 1/2) the ET-VPD curve will be
-    concave up regardless of the magnitude of the plant constants
-    $g_1$ and $uWUE$. Therefore, the general concave up nature of our
-    results, given an assumption of optimal behavior, is insensitive
-    to plant type.
-  \item For all physically possible exponents of VPD dependence ($n,
-    m$), whether the solution is concave up or concave down does not
-    depend on $uWUE$.
-  \item In general, increasing the exponent of VPD dependence
-    increases the likelihood of a concave up result. Additionally, as
-    the exponent of VPD dependence increases from the optimum value of
-    1/2, whether the curve is concave upward or concave downward
-    becomes a function of the plant specific slope parameter $g_*$,
-    through non-dimensional VPD ($VPD^m/g_*$). Because the exponent of
-    the VPD dependencies is capable of altering the fundamental shape
-    of ET-VPD dependence, future research investment in understanding
-    the exact VPD dependence of stomatal conductance, and further
-    reconciliation of empirical and theoretical stomatal and
-    ecosystem behavior should be prioritized.
-\end{itemize}
-While it is possible that in the future some other form of
-VPD dependence is derived, at present \cite{MEDLYN_2011} and
-\cite{Zhou_2014} firmly established n=m=1/2 as the most likely
-candidate given current theory and empirical data. Additionally, we
-argue that a concave up result matches physical intuition more than a
-concave down result. Plants must maintain nutrient and sugar transport
-through the phloem and xylem. To accomplish this, stomata must remain
-slightly open \citep{De_2013, Nikinmaa_2013, Ryan_2014}. Furthermore,
-even if complete stomatal closure were possible, cuticular water loss
-and [at the ecosystem-scale] direct soil evaporation are still sources
-of ET which increase with VPD, independent of stomatal
-closure. Therefore, in the limit as VPD becomes large and we assume
-plants are exercising all strategies to reduce ET, any further increase
-in VPD should result in an increase in ET through cuticular water loss
-and/or direct soil evaporation. This inevitable transition from
-conditions when stomata respond strongly to VPD to conditions when
-stomata response is asymptoting towards full closure would cause a
-concave up ET-VPD curve, which is matched by the theory. In short,
-plant response becomes more limited as VPD increases, while
-atmospheric demand monotonically increases with VPD, leading to the
- result that atmospheric demand dominates plant response when
-atmospheric demand is high.
-
-This analysis allows us to understand the theoretical shape of the ET
-response to VPD with environmental conditions held
-fixed. Accomplishing this with purely statistical
-methods applied to flux observations would be very difficult, given
-the relatively fast time scale of plant response and the
-non-stationarity of [solar forced] environmental conditions over the
-relatively coarse (half hourly) flux estimates (which is required to
-obtain robust eddy-covariance statistics). Our results on the shape of
-the ET-VPD curve with environmental conditions held
-fixed can be built upon with future work examining how changes in VPD
-and environmental conditions (e.g. soil water storage) feedback upon
-one another. In the soil water storage example, over very long time
-scales extremely high VPD perturbations coupled with no precipitation
-could result in decreases in soil water storage such that water
-becomes limiting. This could be represented by an extension of our
-framework in which uWUE is allowed to decrease with decreasing SWC, as
-observed in Section \ref{swc_section}. Here, we focus our results by
-assuming constant PFT-wide conditions to build baseline intuition for
-ET-VPD dependence. For most PFTs, the theory with plant function held
-fixed matches the leading order behavior of the observations where
-plant function varies (Section \ref{testing}).
-
-\section{Conclusions}
-
 We derived a new form of Penman Monteith using the concept of
@@
 enriched CO$_2$ world.
-
-\acknowledgments This work used eddy covariance data acquired and
-shared by the FLUXNET community, including these networks: AmeriFlux,
-AfriFlux, AsiaFlknux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont,
-ChinaFlux, Fluxnet-Canada, GreenGrass, ICOS, KoFlux, LBA, NECC,
-OzFlux-TERN, TCOS-Siberia, and USCCC. The ERA-Interim reanalysis data
-are provided by ECMWF and processed by LSCE. The FLUXNET eddy
-covariance data processing and harmonization was carried out by the
-European Fluxes Database Cluster, AmeriFlux Management Project, and
-Fluxdata project of FLUXNET, with the support of CDIAC and ICOS
-Ecosystem Thematic Center, and the OzFlux, ChinaFlux and AsiaFlux
-offices. This material is based upon work supported by the National
-Science Foundation Graduate Research Fellowship under Grant No. DGE
-16-44869. All code and data used in this analysis, including those
-used to generate the figures and tables, are publicly available at
-\url{https://github.com/massma/climate\_et}.
-%\hyperref[https://github.com/massma/climate\_et]{https://github.com/massma/climate_et}.
-
-\bibliography{references.bib}
-
-\end{document}

### 2018-10-30 doc/paper/intro.tex
preparing hess submission

--- 2018-05-14 doc/paper/vpd_et_paper.tex
@@
-\documentclass[12pt]{article}
-\input{def.tex}
-
-\begin{document}
-
-\title{When does vapor pressure deficit drive or reduce
-  evapotranspiration?}
-
-\author[1]{Adam Massmann\thanks{Corresponding author: akm2203@columbia.edu}}
-\author[1]{Pierre Gentine}
-\author[1,2]{Changjie Lin}
-
-\affil[1]{Department of Earth and Environmental Engineering, Columbia University, New York, NY 10027}
-\affil[2]{State Key Laboratory of Hydroscience and Engineering, Department of Hydraulic
-  Engineering, Tsinghua University, Beijing, CN 100084}
-
-\maketitle
-
-\begin{abstract}
-  Increasing vapor pressure deficit increases atmospheric demand for water, and vapor pressure deficit is expected to rise with increasing greenhouse gases.
-While increased evapotranspiration (ET) in response to increased atmospheric demand seems intuitive, plants are capable of reducing ET in response to increased VPD by closing their stomata, in an effort to conserve water.
-Here we examine which effect dominates response to increasing VPD: atmospheric demand and increases in ET, or plant physiological response (stomata closure) and decreases in ET.
-We use Penman-Monteith, combined with semi-empirical optimal stomatal regulation theory and underlying water use efficiency, to develop a theoretical framework for understanding how ET responds to increases in VPD.
-
- The theory suggests that for most environmental conditions and plant types, plant physiological response dominates and ET decreases with increasing VPD.
-Plants that are evolved or bred to prioritize primary production over water conservation (e.g.
-crops) exhibit a higher likelihood of atmospheric demand-driven response (ET increasing).
-However for forest, grass, savannah, and shrub plant types, ET more frequently decreases than increases with rising VPD.
-This work serves as an example of the utility of our simplified framework for disentangling land-atmosphere feedbacks at multiple scales, including the characterization of ET response in an atmospherically drier, enriched CO$_2$ world.
-\end{abstract}
-
-\onehalfspacing
-
-\section{Introduction}
-
 Vapor pressure deficit (VPD) is expected to rise over continents in
@@
 functional types.
-
-\section{Materials and Methods}
-\subsection{Data}
-\label{data}
-We use both meteorological and eddy-covariance data from the
-FLUXNET2015 database (data available at \sloppy
-\url{https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/} \sloppy),
-including all sites with at least four years of data and observations
-of the variables described in the methods section (Section
-\ref{methods}). Seventy-three sites met these requirements, and were
-grouped into nine plant functional types (PFT) according to the
-International Geosphere-Biosphere Programme vegetation classification
-scheme \citep{Loveland_1999}: cropland (CRO), grass (GRA), deciduous
-broadleaf forest (DBF), evergreen broadleaf forest (EBF), evergreen
-needleleaf forest (ENF), mixed forest (MF), closed shrub (CSH),
-savannah (SAV), and woody savannah (WSA) (site locations in Figure
-\ref{map_fig}).
-
-\begin{figure}[h]
-  \centering \includegraphics[trim={0 3cm 0 3cm}, clip]{./map.pdf}
-  \caption{Plant functional type and location of FLUXNET2015 sites
-    used in this analysis.}
-  \label{map_fig}
-\end{figure}
-
-The purpose of this study is to examine ecosystem response to
-atmospheric drying, focusing on the growing season. To accomplish
-this, we filter and quality control the data using a similar procedure
-as \cite{Zhou_2015}:
-\begin{itemize}
-\item Only measured or highest (``good'') quality gapfilled data,
-  according to quality control flags, are used.
-\item To isolate the growing season, we only use days in which the
-  average Gross Primary Productivity (GPP) exceeds 10\% of the
-  observed 95th percentile of GPP for a given site. GPP is calculated
-  using the nighttime respiration partitioning method.
-\item We remove days with rain and the day following to avoid issues
-  with rain interception and sensor saturation at high relative
-  humidity (\cite{Medlyn_2017}).
-\end{itemize}
-Additionally, as in \citet{Lin_2018}, we restrict data to the daytime,
-which is identified when downwelling shortwave radiation is greater
-than 50 W m$^{-2}$ and sensible heat flux is greater than 5 W
-m$^{-2}$. To reduce the chance of sensor saturation at high relative
-humidity, we remove all time steps for which VPD is less than .01 kPa,
-and to reduce errors at low windspeeds we remove all periods with wind
-magnitudes less than 0.5 m s$^{-1}$. Timesteps with negative observed
-GPP or ET are also removed, and we aggregate half hourly data to
-hourly averages to reduce noise \citep{Lin_2018}. After these quality
-control procedures, 434,765 upscaled hourly observations remain.
-
-\subsection{Methods}
-\label{methods}
-The Penman-Monteith equation \citep [hereafter PM,][]{Penman_1948,
-  Monteith_1965} estimates ET as a function of observable atmospheric
-variables and surface conductances:
-% \begin{linenomath*}
-  \begin{equation}
-    \label{orig_pen}
-    ET = \frac{\Delta R_{net} + g_a \rho_a c_p VPD}{\Delta + \gamma(1 + \frac{g_a}{g_s})},
-  \end{equation}
-% \end{linenomath*}
-where $\Delta$ is the change in saturation vapor pressure with
-temperature, given by Clausius-Clapeyron ($\frac{d \; e_s}{d \; T}$),
-$R_{net}$ is the net radiation minus ground heat flux, $g_a$ is
-aerodynamic conductance, $\rho_a$ is air density, $c_p$ is specific
-heat of air at constant pressure, $\gamma$ is the psychometric
-constant, and $g_s$ is the stomatal conductance (Table
-\ref{definitions}).
-
-\begin{table}
-  \caption{Definition of symbols and variables, with citation for how
-    values are calculated, if applicable.}
-  \label{definitions}
-  \centering \footnotesize
-  \begin{tabular}{l c c c}
-    \hline
-    Variable & Description & Units & Citation \\
-    \hline
-    $e_s$  & saturation vapor pressure & Pa  & - \\ 
-    $T$  & temperature  & K & - \\
-    $P$  & pressure & Pa  & - \\
-    $\Delta$  & $\frac{\partial e_s}{\partial T}$ & Pa K$^{-1}$ & - \\
-    $R_{net}$  & net radiation at land surface minus ground heat flux & W m$^{-2}$   & - \\
-    $g_a$  & aerodynamic conductance & m s$^{-1}$  & \citet{Shuttleworth_2012} \\
-    $\rho_a$  & air density & kg m$^{-3}$  & - \\
-    $c_p$  & specific heat capacity of air at constant pressure & J K$^{-1}$ kg$^{-1}$ & - \\
-    $VPD$  & vapor pressure deficit & Pa  & - \\
-    $\gamma$  & psychometric constant & Pa K$^{-1}$   & - \\
-    $g_{s-leaf}$  & leaf-scale stomatal conductance & m s$^{-1}$  & \citet{MEDLYN_2011} \\
-    $g_{s}$  &  stomatal conductance & m s$^{-1}$
-                                   & \citet{Medlyn_2017} \\
-    $g_{1-leaf}$  & leaf-scale slope parameter & Pa$^{0.5}$
-                                   & \citet{MEDLYN_2011} \\
-    $g_{1}$  & ecosystem-scale slope parameter & Pa$^{0.5}$ & \citet{Medlyn_2017} \\
-    $R$ & universal gas constant & J mol$^{-1}$ K$^{-1}$ & - \\
-    $R_{air}$ & gas constant of air & J  K$^{-1}$ kg$^{-1}$ & - \\
-    $\sigma$ & uncertainty parameter & -& - \\
-    $c_a$ & CO$_2$ concentration & $\mu$ mol CO$_2$ mol$^{-1}$ air& - \\
-    $\lambda$ & marginal water cost of leaf carbon & mol H$_2$O mol$^{-1}$ CO$_2$ & - \\
-    $\Gamma$ & CO$_2$ compensation point & - & - \\
-    $\Gamma^*$ & CO$_2$ compensation point without dark respiration & - & - \\
-    \hline
-  \end{tabular}
-\end{table}
-
-\citet{MEDLYN_2011} developed a model for stomatal conductance ($g_s$)
-by combining an optimal photosynthesis theory \citep{Cowan_1977} with an empirical approach, which describes the
-dependence of $g_s$ to VPD. This resulted in the following model for
-leaf-scale stomatal conductance:
-
-% \begin{linenomath*}
-  \begin{equation}
-    g_{s-leaf} = g_0 + 1.6 \left(1 +
-      \frac{g_{1-leaf}}{\sqrt{VPD}}\right) \frac{A}{c_a},
-    \label{leaf_medlyn}
-  \end{equation}
-% \end{linenomath*}
-where $g_{1-leaf}$ is a leaf-scale ``slope'' parameter, A is the net
-CO$_2$ assimilation rate, and $c_a$ is the atmospheric CO$_2$
-concentration at the leaf surface. \cite{MEDLYN_2011} relate the slope
-parameter ($g_{1-leaf}$) to physical parameters as:
-% \begin{linenomath*}
-  \label{slope}
-  \begin{equation}
-    g_{1-leaf} = \sqrt{\frac{3 \, \Gamma^* \, \lambda}{1.6}},\footnote{Note this expression has units of of (mmol mol$^{-1}$)$^{1/2}$, but this can be converted to Pa$^{1/2}$ using the ideal gas law.}
-  \end{equation}
-% \end{linenomath*}
-
-where $\Gamma^*$ is the CO$_2$ compensation point for photosynthesis
-(without dark respiration), and $\lambda$ is the marginal water cost
-of leaf carbon
-($\frac{\partial \; \text{transpiration}}{\partial \; A}$). So,
-$g_{1-leaf}$ is a leaf-scale term reflecting the trade-off of water for
-carbon uptake. The higher $g_{1-leaf}$, the more open the stomata and
-the more they release water in exchange for carbon.
-
-The Medlyn model for stomatal conductance has been shown to behave
-very well across PFTs \citep[][]{Lin_2015}, and has been successfully
-adopted to ecosystem scale analysis in \citet{Medlyn_2017}. In units
-of m s$^{-1}$, the ecosystem scale stomatal conductance is:
-
-% \begin{linenomath*}
-  \begin{equation}
-    g_s = \frac{R \,T}{P} \; 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{GPP}{c_a},
-    \label{medlyn}
-  \end{equation}
-% \end{linenomath*}
-
-where GPP is the ecosystem scale gross primary production, and $g_1$
-is an ecosystem scale analogue to $g_{1-leaf}$. We solve for $g_1$
-following \citet{Medlyn_2017} (Equation 5), and take the median $g_1$
-value to be representative of each PFT (Table \ref{pft}) instead of the mean to avoid extra weighting of rare outliers.
-
-\begin{table}
-  \caption{Plant functional types, their abbreviation, calculated
-    Medlyn coefficient, and calculated uWUE. uWUE from
-    \citet{Zhou_2015}, including the observed standard deviation, is
-    shown for comparison. Note that uWUE from \citet{Zhou_2015} is
-    calculated from a different set of sites, and that units are
-    converted such that the quantities work with Equations 1-8 and the
-    variables defined Table \ref{definitions}.}
-  \small
-  \label{pft}
-  \centering
-  \begin{tabular}{l c c @{\qquad} c c}
-    \hline
-    \multirow{2}[3]{*}{Abbreviation} & \multirow{2}[3]{*}{PFT} & \multirow{2}[3]{*}{$g_1$ (Pa$^{0.5}$)} & \multicolumn{2}{c}{uWUE ($\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET])}  \\
-    \cmidrule{4-5}
-
-                                     & & & fitted & \citet{Zhou_2015} \\
-
-    \hline
-    \input{pft_params.tex}
-    \hline
-  \end{tabular}
-\end{table}
-
-While Equation 3 can be used in PM (Equation 1), it will make
-analytical work with the function intractable because $GPP$ is
-functionally related to ET itself. Additionally, a perturbation to VPD
-should induce a physiological plant response that will alter GPP and
-cause an indirect change in stomatal conductance, in addition to the
-direct effect of VPD in Equation \ref{medlyn}. Therefore, in order to
-derive the response of ET to VPD, we must account for the functional
-relationship between GPP, ET, and VPD, and its effect on stomatal
-conductance. We can use aforementioned semi-empirical results of
-\citet{Zhou_2015} as a tool to approach this
-problem. \citet{Zhou_2015}, showed that  underlying Water Use
-Efficiency (uWUE):
-
-% \begin{linenomath*}
-  \begin{equation}
-    uWUE = \frac{GPP \cdot \sqrt{VPD}}{ET}
-    \label{uwue}
-  \end{equation}
-% \end{linenomath*}
-
-is relatively constant across time and moisture conditions within a
-plant functional type, and correctly captures a constant relationship
-between GPP, ET and VPD over a diurnal cycle \citep{Zhou_2014}. The
-theoretical derivation of the square root VPD dependence in $uWUE$
-leverages the same assumptions used in \cite{MEDLYN_2011} to derive the
-square-root VPD dependence of the stomatal conductance model (Equation
-\ref{medlyn}).  We can use uWUE to remove the $GPP$ dependence from
-$g_s$ in a way that makes PM analytically tractable:
-
-% \begin{linenomath*}
-  \begin{equation}
-    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{uWUE \; ET}{c_a \; \sqrt{VPD}}.
-    \label{new_g_s}
-  \end{equation}
-% \end{linenomath*}
-
-Plugging Equation \ref{new_g_s} into Equation \ref{orig_pen} and
-rearranging gives a new explicit expression for PM, in which
-dependence on $GPP$ is removed:
-
-% \begin{linenomath*}
-  \begin{equation}
-    ET = \frac{\Delta R_{net} + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R \; 1.6\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
-    \label{et}
-  \end{equation}
-% \end{linenomath*}
-
-By accounting for photosynthesis changes in ecosystem conductance, with
-Equation \ref{et} we have derived for the first time, using recent
-results \citep[][]{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017}, ET
-explicitly as function of environmental variables and two
-plant-specific constants, the slope parameter ($g_1$), and uWUE, both
-reflecting water conservation strategy. The slope parameter is related
-to the willingness of stomata to trade water for CO$_2$ and to keep
-stomata open. uWUE is a semi-empirical ecosystem-scale constant
-related to how WUE changes with VPD (specifically $VPD^{-1/2}$). It is
-also roughly proportional to physical constants:
-
-\[uWUE \appropto \sqrt{\frac{c_a - \Gamma}{1.6 \lambda}},\]
-
-where $\Gamma$ is the CO$_2$ compensation point \citep[Equation 5
-in][]{Zhou_2014}. So uWUE is related to atmospheric CO$_2$
-concentration and compensation point, and is inversely proportional to
-the marginal water cost of leaf carbon.
-
-Given eddy-covariance FLUXNET2015 data (Section \ref{data}), every
-term in our new version of PM (Equation \ref{et}) is observed, except
-for uWUE, which we fit by calculating its expectation, given the model
-and FLUXNET2015 data.
-
-However, eddy-covariance data are inherently noisy so we include a
-measure of uncertainty in our analysis. To account for observational
-error, as well as model uncertainty (e.g. temporal and spatial
-variations of uWUE and $g_1$), we introduce an uncertainty parameter
-$\sigma$ modifying uWUE:
-
-% \begin{linenomath*}
-  \begin{equation}
-    ET = \frac{\Delta R_{net} + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R \; 1.6\; \sigma \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
-    \label{et_sigma}
-  \end{equation}
-% \end{linenomath*}
-
-Now, from each FLUXNET2015 observation (i.e. for each hourly
-observation at every time step) we can evaluate $\sigma$:
-
-% \begin{linenomath*}
-  \begin{equation}
-    \sigma = - \frac{g_a \gamma c_a \sqrt{VPD} L_v P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R_{net} - g_a \rho_a c_p VPD\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})},
-    \label{sigma}
-  \end{equation}
-% \end{linenomath*}
-
-Thus, with this uncertainty analysis we can evaluate departure from
-our theory in observations, as a departure of $\sigma$ from unity. The
-variability of $\sigma$ across sites and time provides a measure of
-uncertainty in our model, assumptions, as well as in the FLUXNET2015
-observations themselves. The variability of $\sigma$ then propagates
-through any uncertainty to our partial derivative of Equation
-\ref{et_sigma} with respect to VPD:
-
-% \begin{linenomath*}
-  \begin{equation}
-    \frac{\partial \;  ET}{\partial \; VPD} = \frac{2\; g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_a }{1.6 \; R\; \sigma \; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)
-    \label{d_et}
-  \end{equation}
-% \end{linenomath*}
-
-With Equation \ref{d_et} we have an analytical framework for ecosystem
-response to atmospheric demand perturbations with environmental
-conditions held fixed. There are a few subtleties to taking the
-derivative in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$)
-and $VPD$ are functionally related, so while taking the derivative we
-evaluate
-$\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET}
-{\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD}
-\Big|_{\text{RH fixed}} + \frac{\partial \; ET}{\partial \; RH}
-\frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$
-    fixed}}$. $RH$ and $e_s$ are assumed to be approximately
-independent, which is supported by the data (not shown).
-
-This derivation relied either implicitly or explicitly on several
-assumptions. First, we assume that VPD at the leaf surface is the same
-as VPD at measurement height; physically this implies that leaves are
-perfectly coupled to the atmosphere. In reality, for some conditions
-and plant types the leaves can become decoupled from the boundary
-layer \citep{De_2017, Medlyn_2017}. Given our focus on the growing
-season, which is usually characterized by relatively high insolation
-inducing instability and convective boundary layers, we would expect
-the surface to be generally well coupled. An additional assumption in
-the formulation of $uWUE$ \citep{Zhou_2014, Zhou_2015} and
-\citet{Medlyn_2017}'s stomatal conductance model is that direct soil
-evaporation (E) contributions to ET remain small relative to
-transpiration (T). This should be more true during the growing
-season. The ratio of E to T may increase immediately after rainfall
-events due to high soil moisture and ponding, but VPD is generally low
-anyways during these times. However, some plant types allow for
-systematically larger contributions of E in ET, particularly those
-with sparse canopies and smaller relative amounts of transpiration. We
-therefore might expect that the theory will be most applicable to
-forest PFTs, which will be most strongly coupled to the boundary layer
-due to larger surface roughness, and will also generally have the
-highest ratios of transpiration to evaporation. Finally, for the goal
-of developing PFT-wide intuition, we assume that $g_1$ and $uWUE$ are
-constant in space and time. Both quantities have been shown to be well
-conserved over space and time within a PFT relative to inter-PFT
-variability. However, we would expect them to vary somewhat with
-environmental conditions including very low soil water content,
-temperature, and with intra-PFT plant-specific characteristics like
-wood density for tree PFTs \citep{Lin_2015}. Given these strong
-assumptions made with the goal of understanding broad, leading order
-plant behaviors, we take the extra care of including and analyzing
-uncertainty (Section \ref{testing}) and examining sensitivity to the
-exponent of the VPD response (Section \ref{functional_form}).
-
-We note one final comment on our derivation which is relevant for
-drought indices. If we approximate $c_a$ at a global mean CO$_2$
-concentration, then the RHS of Equation \ref{et} is fully defined
-using commonly available weather station data and the constants
-published here. This makes Equation \ref{et} a useful alternative to
-PET in drought indices and hydrometeorological analysis for vegetated
-surfaces. Equation \ref{et} better reflects the physics of water
-exchange at the land surface and would only require fitting of uWUE
-and $g_1$, and can also account for changes in CO$_2$ concentration
-(assuming some relationship between $\lambda$ and [CO$_2$]), contrary
-to typical drought indices such as the Palmer Drought Severity Index
-(PDSI) \citep{Swann_2016, Lemordant_2016, Lemordant_2018}. Lastly, all
-code and data used in this analysis, including those used to generate
-the figures and tables, are publicly available at
-\url{https://github.com/massma/climate\_et}.
-%\hyperref[https://github.com/massma/climate\_et]{https://github.com/massma/climate_et}.
-
-\section{Results and Discussion}
-\label{results}
-
-By construction, the variability in the $\sigma$ term (Equation
-\ref{sigma}) contains all model and observational uncertainties. For
-an observation that perfectly matches our model and constant uWUE
-assumption, $\sigma$ will be one. Therefore, for our assumptions and
-framework to be reasonable $\sigma$ should be close to 1. An
-additional concern is that $\sigma$ may in fact be correlated with
-$VPD$, in which case the dependence would need to be accounted for
-when taking the derivative. Fortunately, there is a very weak
-dependence of $\sigma$ on VPD in their joint distribution, and
-$\sigma$ is indeed close to unity i.e. $O(1)$ (Figure
-\ref{joint_vpd_sigma}). Given this weak dependence and the
-distribution of $\sigma$ we have confidence in our model framework and
-the data quality.
-
-\begin{figure}[h]
-  \centering\includegraphics[width=0.75\textwidth]{./joint_vpd_sigma.pdf}
-  \caption{The joint distribution of $VPD$ and $\sigma$, with outliers
-    removed (defined as lowest and highest 5\% of $\sigma$). $\sigma$
-    exhibits a weak dependence on $VPD$, and $\sigma$ is $O(1)$ for
-    the bulk of the observations.}
-  \label{joint_vpd_sigma}
-\end{figure}
-
-Before calculating the sensitivity of ET to VPD, we will consider the
-functional form of Equation \ref{d_et}. There are two main terms: a
-``scaling'' term, which modifies the magnitude but not the sign of the
-ET response to VPD ($\frac{\partial \; ET}{\partial \; VPD}$):
-
-\begin{equation}
-  \frac{g_a \; P}{T(\Delta + \gamma)},
-\end{equation}
-
-and a ``sign'' term, which determines whether ET increases or
-decreases with VPD (i.e. atmospheric demand driven or physiologically
-controlled):
-
-\begin{equation}
-  \label{sign}
-  \frac{c_p}{R_{air}} - \frac{\gamma c_a }{1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right).
-\end{equation}
-All variables are positive, so the relative magnitude between the
-first term and the second term in the sign term (Equation \ref{sign})
-will determine whether ET increases or decreases with increasing
-VPD. If the second term is larger then plant control dominates and ET
-decreases with increasing VPD. However, if the first term is larger,
-then atmospheric demand dominates and ET increases with increasing
-VPD.
-
-\subsection{Functional Form of the Sign Term}
-\label{sign_term}
-First, we explore the variables within the sign term to gain better
-intuition on the driver of either the increase or reduction of ET with
-VPD. CO$_2$ concentration ($c_a$) and the psychometric constant
-($\gamma$) are relatively constant over the dataset considered here so
-that the variability is dominated by $\sigma$ and $VPD$. uWUE could
-vary with soil moisture but has been shown to be relatively constant
-\citep{Zhou_2015}. This then means that the sign term only depends on
-VPD for a given PFT and is approximately just a function of $VPD$. We
-can further determine a critical threshold separating an increase from
-a decrease in ET, i.e. the threshold $VPD_{crit}$ such that the
-derivative vanishes $\frac{\partial \; ET}{\partial \; VPD} = 0$:
-\small
-% \begin{linenomath*}
-  \begin{equation}
-    VPD_{crit} = \frac{R_{air}}{4 c_p} \left( \frac{\gamma c_a}{1.6\; R \;  uWUE} + \sqrt{\frac{\gamma c_a}{1.6\; R \;  uWUE}\left( \frac{\gamma c_a}{1.6\; R \;  uWUE} + 8 g_1 \frac{c_p}{R_{air}}\right)} - 4 g_1 \frac{c_p}{R_{air}} \right),
-    \label{vpd_min_et}
-  \end{equation}
-% \end{linenomath*}
-\normalsize noting that $VPD_{crit}$ mostly depends on the PFT
-parameters uWUE and $g_1$, and only varies weakly with climate as most
-other parameters related to the environment are nearly constant. The
-calculated value of $VPD_{crit}$ for each PFT is shown in Table
-\ref{vpd_crit}. For any values of $VPD$ less than $VPD_{crit}$, ET
-will decrease with increasing VPD
-($\frac{\partial \; ET}{\partial \; VPD} < 0$), and for values of
-$VPD$ greater than $VPD_{crit}$, ET will increase with increasing VPD
-($\frac{\partial \; ET}{\partial \; VPD} > 0$). In other words,
-ecosystems regulate and mitigate evaporative losses up to the VPD
-limit, $VPD_{crit}$, above which atmospheric demand is just too high
-to be entirely compensated by stomatal and ecosystem regulation. We
-note however that even though ET increases again above the critical
-threshold, $VPD_{crit}$, ET is still much lower than potential
-evaporation as stomata are still strongly regulating vapor fluxes to
-the atmosphere. Indeed, even in the absence of soil pore evaporation,
-stomata do not shut down entirely at very high VPD so that ET does not
-go to zero, because stomata are still slightly open to perform some
-photosynthesis \citep{Ball_1987, Leuning_1990, MEDLYN_2011}. In
-addition, upward xylem transport is necessary to maintain phloem
-transport, as well as nutrient transport and thus carbon allocation
-\citep{De_2013, Nikinmaa_2013, Ryan_2014}.
-
-\begin{table}
-  \caption{Values of $VPD_{crit}$, where
-    $\frac{\partial \; ET}{\partial \; VPD} = 0$, evaluated at PFT
-    average values for $R_{air}$, $\gamma$, and $c_a$. PFT-specific
-    constants ($g_1$, uWUE) are provided in Table \ref{pft}. For
-    values of $VPD$ less than $VPD_{crit}$,
-    $\frac{\partial \; ET}{\partial \; VPD}$ will be negative, and for
-    values of $VPD$ greater than $VPD_{crit}$,
-    $\frac{\partial \; ET}{\partial \; VPD}$ will be positive.}
-  \centering
-  \begin{tabular}{l c c c c c}
-    \hline
-    PFT & $R_{air}$ & $c_a$ (ppm) & $\gamma$  & \textbf{$VPD_{crit}$ (Pa)} \\
-    \hline
-    \input{vpd_crit.tex}
-    \hline
-  \end{tabular}
-  \label{vpd_crit}
-\end{table}
-
-Differences in $VPD_{crit}$ are exclusively determined by uWUE and the
-slope parameter ($g_1$) related to the plant functional type. A larger
-uWUE means a smaller $VPD_{crit}$, and an ET response to increases VPD
-that is more likely to be positive. At first glance this result is
-somewhat counter-intuitive; we expect that plants with a higher water
-use efficiency would be more water conservative. However, in reality
-uWUE determines how $WUE$ changes with VPD:
-
-\[WUE = \frac{GPP}{ET} = \frac{uWUE}{\sqrt{VPD}}.\]
-\[\frac{\partial \; WUE}{\partial \; VPD} = -\frac{uWUE}{2 \;
-    VPD^{3/2}}\]
-
-So, plants with a higher uWUE will have a greater \textit{decrease} in
-ecosystem-scale $WUE$ in response to increases in VPD. This decrease
-in $WUE$ causes more water loss per unit carbon gain, and explains the
-relationship between high uWUE and high likelihood of increases of ET
-in response to increasing atmospheric drying (increases in VPD).
-
-A tendency towards increasing ET response with increasing VPD can also
-be caused by a high slope parameter ($g_1$), characteristic of plants
-that at the leaf scale are more willing to trade water for access to
-atmospheric CO$_2$. Plants that are less conservative will be thus be
-more likely to increase ET with increasing VPD. Both the
-aforementioned effects (large uWUE, $g_1$) can amplify each other, and
-generally conspire to shift the sign term towards a positive value for
-a given PFT.
-
-This effect of uWUE and $g1$ on the sign term is most apparent by
-comparing two extreme PFTs: water intensive crops (CRO) and water
-conservative closed shrub (CSH). CRO has higher slope parameter and a
-slightly higher uWUE ($g_1 = 140.7$ Pa$^{1/2}$; $2.85$ $\mu$-mol [C]
-Pa$^{0.5}$ J$^{-1}$ [ET]) compared to CSH ($g1 = 75.1 \, Pa^{1/2}$,
-$uWUE=2.82$ $\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET]). These differences
-in PFT parameters cause opposite ET responses to changes in VPD
-between CRO and CSH. ET theoretically always decreases with increasing
-VPD for the more water conservative CSH, while ET frequently increases
-with increasing VPD for the more water intensive CRO (Figure
-\ref{idealized_sign}). CROs evolved or were bred to prioritize GPP and
-yield and are thus not water conservative. They are very willing to
-trade water for photosynthesis and productivity, despite changes in
-VPD, while CSH are very unwilling to trade water for more
-photosynthesis.
-
-\begin{figure}
-  \centering \includegraphics{./idealized_sign.pdf}
-  \caption{The functional form of the sign term, with $\sigma$ held
-    fixed at 1, and all terms except VPD set to PFT averages. For
-    comparison, the observed range of VPD for each PFT is plotted
-    below the x-axis. Stars denote 25th, 50th, and 75th percentiles,
-    and the range of the line spans the 5th-95th percentiles of
-    observed VPD. Vertical black lines denote the location of
-    $VPD_{crit}$ for each PFT, with the exception of CSH and WSA, for which
-    $VPD_{crit}$ is off-scale.}
-  \label{idealized_sign}
-\end{figure}
-
-As expected, the slope parameter ($g_1$) is a primary determinant of
-the VPD dependence for the sign term shown in Figure
-\ref{idealized_sign}. Plants that are more conservative (small $g_1$)
-will tend to reduce ET with increasing VPD, and will be very effective
-at reducing ET, especially at low VPD. However, at very high VPD,
-gradients in vapor pressure at the leaf scale will become very strong
-as stomata reach their limits of closure in response to VPD
-(parameterized with $g_1$). As a result, ET response will begin to
-asymptote towards a constant ecosystem-scale values as leaf-scale
-response to VPD asymptotes towards zero.  Therefore, plants with a low
-$g_1$ will have the largest VPD dependence of ET response because the
-difference in ET response at low VPD (leaf stomatal response
-dominates) and high VPD (VPD gradient dominates) is largest. This is
-apparent in the strong VPD dependence of CSH, which has the lowest
-slope parameter ($g_1=75.1$ Pa$^{1/2}$) (Figure \ref{idealized_sign}).
-
-To summarize our theoretical insights (Figure \ref{idealized_sign} and
-Table \ref{vpd_crit}), CROs are the least water conservative and have
-the strongest overall tendency to increase ET with increasing VPD,
-while CSH are the most water conservative and have the strongest
-tendency to decrease ET with increasing VPD, as well as the strongest
-VPD dependence of response. Figure \ref{idealized_sign} clearly shows,
-according to our theory, that for all PFTs except for crops there is
-frequent occurrence of a negative (plant dominating) ET response to
-increases in VPD. Therefore, plants are able in most atmospheric
-conditions to reduce ET in response to increased VPD and thus to
-reduce water loss. To better illustrate this, the ranges of observed
-environmental VPDs at the FLUXNET sites are plotted parallel to the
-x-axis. For CSH and WSA, VPD is always less than VPD$_{crit}$ (off scale) so
-that the plant response dominates in typical environmental conditions,
-emphasizing the water conservative strategy of those plants. For CRO
-on the other hand, VPD is higher than VPD$_{crit}$ for more than 50\%
-of observations, emphasizing that those plants operate with an
-aggressive water usage strategy, are water intensive and were actually
-engineered for photosynthesis rather than water saving. For DBF, EBF,
-MF, GRA, and SAV more than half of the observed VPD are less than
-VPD$_{crit}$, i.e. in conditions where plant response dominates. SAV
-has a more water conservative response than the forest, grass, and
-crop plan types, but still responds by increasing ET with increasing
-VPD for about a quarter of observations, due to the high aridity (VPD)
-of the SAV ecoclimate. It is also important to note that for all PFTs,
-even when atmospheric demand dominates, ET response to VPD is still
-far more negative than it would be for potential evaporation
-$\partial PET/\partial VPD$, i.e. atmospheric demand only, emphasizing
-that there is still a strong regulation of evaporative flux by stomata
-and though the plant xylem. The sign term in the PET case would just
-be a constant ($\frac{c_p}{R_{air}} \approx 3.5$), which is far larger
-than any part of the curves for any PFT. This highlights the
-deficiencies of PET's ability to capture land response to changes in
-atmospheric dryness (VPD). Plants are always regulating water exchange
-from the land surface, even when they reach the limits of they ability
-to do so.
-
-\subsection{Functional Form of the Scaling Term}
-\label{scale_term}
-While the above discussion of the sign of
-$\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our
-question of when ET response increases or decreases with VPD,
-understating the overall magnitude of the ET response is important to
-soil-plant-atmosphere water budgeting. So we now more closely examine
-the terms that affect how the sign term is scaled:
-
-\begin{equation}
-  \frac{g_a \; P}{T(\Delta + \gamma)}.
-\end{equation}
-
-$\frac{P}{T}$ is an air-density term, which varies little compared to
-aerodynamic conductance and Clausius-Clapeyron ($\Delta$). The
-psychometric constant ($\gamma$) is also relatively constant, so the
-scaling term should be primarily a function of aerodynamic conductance
-and temperature, through the Clausius-Clapeyron relationship
-$\Delta$. This is as expected, given that the aerodynamic conductance
-represents the efficiency of exchange between the surface and the
-atmosphere. As aerodynamic conductance increases, any plant response
-will be communicated more strongly to the atmosphere (and vice-versa).
-
-$\Delta$'s presence in the scaling term also matches physical
-intuition. $\Delta$ (and also the approximately constant $\gamma$)
-control the efficiencies with which surface energy is converted to
-latent and sensible heat \citep{Monteith_1965}. The functional from of
-$\Delta$ will be the same across PFTs, but the temperature range may
-vary slightly. In contrast, aerodynamic conductance will vary strongly
-with PFT due to the importance of surface roughness for aerodynamic
-conductance. So most of the differences in scaling between PFT should
-be in the aerodynamic conductance term.
-
-The control of the scaling term variability between PFTs by
-aerodynamic conductance is confirmed by data (Figure
-\ref{scale_vary}). Differences between PFT are almost entirely due to
-differences in aerodynamic conductance, rather than differences in
-observed temperature ranges. The scaling term for the tree PFTs (DBF,
-EBF, ENF, MF) is generally about double the scaling terms for other
-PFTs which have lower surface roughness and generally smaller
-aerodynamic conductance (GRA, CSH, CRO). The savannah (WSA, SAV) PFT's
-scaling is somewhere between GRA, CSH, and CRO, and DBF, EBF, ENF, and
-MF, due to higher variability and surface roughness.
-
-\begin{figure}
-  % \centering
-  \centerline{\includegraphics[width=1.2\textwidth]{./idealized_scale.pdf}}
-  \caption{Primary sources of variability for the scaling term, as a
-    function of PFT. The 5th-95th percentile range of temperature is
-    plotted at the 5th, 25th, 50th, 75th, and 95th percentiles of
-    aerodynamic conductance, as observed for each PFT.}
-  \label{scale_vary}
-\end{figure}
-
-Within each PFT, the scaling term variability is controlled both by
-environmental temperature and aerodynamic conductance variability
-(Figure \ref{scale_vary}). While the observed variability of the
-aerodynamic conductance contributes more to the scaling term
-variability than temperature, the temperature contribution is
-non-negligible. Specifically, the scaling term is generally larger at
-low temperatures when latent heat is relatively inefficient at moving
-energy away from the surface. This effect amplifies the role of
-aerodynamic conductance variability at low temperatures.
-
-To summarize, variability between PFTs is mostly controlled by
-systematic differences in aerodynamic conductance, due to differences
-in surface roughness between each PFT, and possibly to a lesser extent
-wind conditions. In contrast, variability within PFT is also
-controlled by temperature, through Clausius-Clapeyron. But,
-aerodynamic conductance variability generally impacts the scaling term
-more than temperature, even within PFTs.
-
-\subsection{Bulk statistics of ET response to VPD}
-\label{stats_sec}
-
-In this section we consider direct observations of ET response with
-eddy-covariance data, while including uncertainty with the $\sigma$
-term (Section \ref{methods}). These observational results of ET
-response (Table \ref{stats}) largely confirm our theoretical analysis, presented in the previous sections. For all PFTs, mean ET response to increasing VPD
-is negative. However, ET response evaluated at the average of all
-variables (e.g. $\sigma$, $T$, $c_a$, $VPD$) is positive for CRO, and
-negative for all other PFTs. This difference in mean ET response as
-compared to the ET response at mean environmental conditions is due to
-the non-linear nature of the response, in which negative responses are
-generally larger magnitude than positive responses (Figure
-\ref{idealized_sign}). Therefore, both the mean ET response as well as
-the ET response at mean environmental conditions matches our
-expectations from the theory (Section \ref{sign_term}), with the
-exception that CRO observations are shifted more towards a negative ET
-response than we expect.
-
-\begin{table}
-  \caption{Statistics of $\frac{\partial \; ET}{\partial \; VPD}$ as a
-    function of PFT.}
-  \centering
-  \begin{tabular}{l c c c}
-    \hline
-    PFT & $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ & $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)$ & fraction $\frac{\partial \; ET}{\partial \; VPD} < 0.$ \\
-    \hline
-    \input{stats.tex}
-    \hline
-  \end{tabular}
-  \label{stats}
-\end{table}
-
-Regarding the frequency of negative and positive ET response, all PFTs
-exhibit a decreasing ET response with increasing VPD (physiologically
-controlled, water conservative response) for the majority of
-observations. The more water conservative PFTs generally exhibit
-higher frequency of negative ET response, especially when one factors
-in the distribution of environmental VPD (e.g. SAV and WSA grow in
-more arid climates, Figure \ref{idealized_sign}). In general the bulk
-statistics match our theoretical expectations well, with the caveat
-that inclusion of uncertainty shifts crops towards a slightly more
-negative response to VPD, and shifts many of the other PFTs, which still
-exhibit a high frequency of negative ET response to VPD, towards more frequent
-occurrence of positive ET response than the theory and Figure
-\ref{idealized_sign} would suggest. The bulk statistics motivate a
-more thorough examination of the structure of uncertainty and more
-sophisticated validation of our theory's performance against
-observations.
-
-\subsection{Validation of theory at eddy-covariance sites}
-\label{testing}
-We now compare more sophisticated distributions of the observed
-response to our simplified theory (Section \ref{sign_term}). The
-observed distribution of the sign term, as compared to what the theory
-would predict, is provided in Figure \ref{test_sign}. Our goal was to
-capture the leading order behavior of the ET dependence on VPD. Given
-the assumptions we made, and the uncertainties of flux tower
-observations themselves, we expect a relatively large amount of noise
-when reproducing the derivatives of ET. However, the data largely
-reproduces our theoretical analysis.
-
-\begin{figure}
-  \centering
-  \centerline{\includegraphics[width=1.2\textwidth]{./test_sign.pdf}}
-  \caption{Comparison of the sign term with model uncertainty included
-    (box plots) to the sign term as calculated with simplifying
-    assumptions (blue line, as in Figure \ref{idealized_sign}). Each
-    box plot corresponds to 5\% of the data, and the 5-95\% range of
-    VPD is plotted.}
-  \label{test_sign}
-\end{figure}
-
-This is particularly true for DBF and MF; the theory matches the
-leading order behavior of the function when uncertainty is included,
-and the observations match the theory with the addition of noise. The
-VPD dependence, given by the slope parameter ($g_1$), follows the
-median values of each bin. Perhaps most importantly, the x-intercept,
-and thus VPD$_{crit}$, matches nearly exactly between the theory and
-the observations. Therefore the sign of the ET response to increases
-in VPD should be well matched, subject to the unavoidable constraints
-of noise, much of which comes from the observations themselves. The
-uncertainty is non negligible; there are many observations in each bin
-for which the the sign of observation is opposite the response
-predicted by the theory, but to leading order our theory matches the
-observations well.
-
-While CSH has a much different functional form of the sign term than
-DBF and MF, CSH observations also match our theory to leading order,
-albeit with a bit more variability as a function of VPD. Again, the
-VPD dependence mostly determined from the slope parameter ($g_1$)
-closely matches the medians in the observation bins. The
-VPD-independent, strongly negative response is also captured. For CSH,
-there is rare occurrence of observed positive ET response with VPD
-($\approx 6$\%), even with uncertainty, so the sign of the
-observations almost always matches the sign of the theory, which
-states that ET response should always be negative.
-
-Biases between the theory and observations are similar for ENF, WSA,
-CRO and SAV. At low VPD the theory and the observations match
-well. However, at high VPD (upper 10-20\% of observations) the theory
-is biased slightly towards positive response as compared to the
-observations (Figure \ref{test_sign}). For ENF this slight bias could
-be explained by a negative bias in $g_1$. However, for WSA, CRO, and
-SAV the observations in the highest VPD bins exhibit a downturn
-towards more negative response, which cannot be captured by the
-functional form of the sign term (Figure
-\ref{idealized_sign}). Therefore, to explain the observations one of
-the variables in the sign term must change at extreme VPD (upper
-10-20\%). The most likely candidate is $uWUE$, which we have assumed
-constant to meet our goal of developing intuition for leading order
-behavior, but might be expected to decrease at extremely low SWC. This extremely low SWC should be correlated with high VPD in very dry environments
-through sensible heating increases \citep{Gentine_2016}. The theory
-for EBF exhibits similar limitations as for ENF, WSA, and SAV, except
-a larger portion of observations are biased negative as compared to
-the theory ($\approx$ 35\%, Figure \ref{test_sign}). However, in
-general, and specifically for non-extreme VPD (VPD  $< \approx$70-90th
-percentiles), the theory matches the observations for the tree and
-savannah plant types well.
-
-The theory for GRA suffers from similar, but much more severe,
-limitations as for CRO, WSA, and SAV. GRA observations are
-characterized by a consistent trend back towards negative ET response
-at higher VPD, which the functional form of our theory is incapable of
-accounting for. As compared to CRO, WSA and SAV, the divergence
-between the theory and the observations is far greater for GRA,
-biasing 40-50\% of observations at higher VPD. In addition for the
-potential for soil moisture to alter $uWUE$, there are other sources
-of plant heterogeneity specific to GRA (and some extent CRO) that may
-alter $uWUE$ (or $g_1$) or invalidate other assumptions made in the
-Methods section (Section \ref{methods}). We do not account for
-variability in plant height and surface roughness, or differences in
-C3 vs C4 photosynthesis and water strategies, which we might expect to
-vary substantially across sites, years, and season for GRA. These
-deficiencies could largely explain the inability of our theory to
-exactly match the observations in croplands. For example, a
-superposition of sites with C3 plants (at low environmental VPD) and
-C4 plants (at high environmental VPD) would explain the observed shift
-back towards negative ET response at high VPD when all sites are
-binned together, as in Figure \ref{test_sign}. We hypothesize that the
-theory would validate against observations much better if these
-sources of variability were accounted for, at a cost of increased
-complexity and analytical opacity.
-
-While the above discussion shows that our theory has some limitations
-when applied to some PFTs, especially for grasslands, it does well for
-DBF, MF and CSH PFTs, and captures the response at non-extreme VPD
-(VPD $< \approx$ 70th-90th percentiles) for
-ENF, EBF, CRO, WSA, and SAV. In general, the leading order behavior
-observed in the data is captured by the theory. Departures between our
-theory and observations, specifically at high VPD, could be explained
-and conceptualized with shifts in $g_1$ and/or $uWUE$ due to
-site-specific plant type variability (e.g. more arid-adapted ecosystems
-at more arid sites), or temporal variability for some environmental
-conditions (e.g. decreases in $uWUE$ at extremely low soil
-moisture). To focus on general behavior and develop intuition for
-PFT-scale response, we have ignored these sources of variability in
-the present analysis. However, any site-specific or temporal plant
-functional variability that can be conceptualized with shifts to $g_1$
-and $uWUE$ can be analyzed with our framework. This opens a door to future analyses in which plant behavior in anomalous conditions
-can be explained and analyzed using Equation \ref{et}.
-
-% It is useful to also compare our theory, the observed uncertainty in
-% the ET response, and PET, as PET is often used as a proxy for ET in
-% hydrometeorological studies and drought indices. Again, as in our
-% discussion of the theoretical results, the PET sensitivity to VPD
-% will be constant at about 3.5. In Figure \ref{test_sign} the ET
-% response to shifts in dryness is always far smaller than 3.5 in the
-% observations, even with all uncertainty included. So the PET
-% response to atmospheric dryness is outside the range of all
-% observable responses. This casts doubt on PET's usefulness for
-% capturing the leading order behavior of water budget and dryness
-% response at the land surface, as is done in several drought
-% indices. However our theory (Equation \ref{et}) could be used with
-% published uWUE and $g1$ data, as an alternative to PET to estimating
-% land response and dryness, and would more accurately account for the
-% response to atmospheric drying.
-
-\subsection{Observed ET response to VPD}
-
-Most of the results presented so far focus on the sign term, so now we
-turn to observations of ET response with the scaling term included
-(Figure \ref{data_scatter}). Until now, in the interest of developing
-leading order intuition for ET response, and to be conservative in our
-acknowledgment of model and observational error, we've considered
-$\sigma$ variability to be a measure of uncertainty. An alternative
-viewpoint is that $\sigma \cdot uWUE$ represents spatial and temporal
-variability of uWUE, which may be expected within bounds \citep[see
-Table \ref{pft}, also ][]{Zhou_2015}. This is a less conservative
-view; some of the $\sigma \cdot uWUE$ variability will be due to model
-and observational error, so by viewing $\sigma \cdot uWUE$ as
-\textit{real} variability we run the risk of mistaking noise for
-signal. However, the advantage of this viewpoint is that, from
-\citet{Zhou_2014}, we have very high confidence that uWUE fit at the
-hourly timescale (as we do with $\sigma \cdot uWUE$) correctly
-captures the relationship between ET, GPP, and VPD, and that our form
-of PM introduces minimal error with its use of uWUE.
-
-\begin{figure}
-  \centering
-  \centerline{\includegraphics[width=1.2\textwidth]{./data_scatter.png}}
-  \caption{Scatter plots of observed
-    $\frac{\partial \; ET}{\partial \; VPD}$, including $\sigma$
-    variability, as a function of PFT, temperature and VPD. Please
-    note differences in the colorbar scale.}
-  \label{data_scatter}
-\end{figure}
-
-Therefore, with the caveat that some of the signal presented in Figure
-\ref{data_scatter} may in fact be noise, we can interpret the observed
-distribution of ET response to VPD. In general, the observed response
-matches the intuitive theory. ET response to VPD shifts towards
-positive values as VPD increases (atmospheric demand dominating). CRO
-exhibit the highest occurrence of positive ET response, and the
-observations confirm that CRO are the most water intensive. CSH are
-the most water conservative, with a strong negative response. DBF,
-EBF, ENF, MF, SAV and WSA are also water conservative, but show some
-occurrence of positive ET response to VPD, particularly at higher
-observed VPD, as the theory predicts. GRA, while generally water
-conservative, does not match the theory well, with increasing frequency of
-\textit{negative} response at high VPD. This is as expected, given
-previous discussion on how we might expect more inter-site and inter-year variability for GRA (Section \ref{testing}).
-
-Figure \ref{data_scatter} also includes the impact of the scaling
-term. For a given VPD, the magnitude of ET response does not vary
-strongly with temperature, confirming that any impact of the scaling
-term on the magnitude of the response is primarily due to changes in
-aerodynamic conductance. Intuitively this is reasonable; aerodynamic
-conductance will control how dominant balances at the land surface are
-communicated to the boundary layer. While $\Delta$ controls the
-efficiency of energy conversion to latent heat, it appears this is a
-second order term, relative to $g_a$, for scaling ET response.
-
-As with Figure \ref{test_sign}, Figure \ref{data_scatter} matches our
-expectation based on simplified theory. The sign term is most strongly
-scaled by $g_a$, and in general the occurrence of positive ET response
-increases as VPD increases. The willingness with which a given plant type
-evolved to use water dictates the occurrence of positive versus
-negative ET response. Water conservative ecosystems are highly effective
-at mitigating the effects of atmospheric demand, and can store water
-for later use by reducing ET in response to increasing VPD.
-
-\subsection{Limitations of theory - very dry soil moisture conditions}
-\label{swc_section}
-In formulating our theory with Penman Monteith, we implicitly did not
-account for very dry soil moisture conditions. For the majority of
-environmental conditions observed at the eddy covariance sites used
-here, soil conditions were not extremely dry so that we could assume a
-constant uWUE and $g_1$. We posit that ecosystems will generally
-optimize to host plants living in conditions which they evolved
-for. However, in extreme conditions and drought scenarios soil water
-content (SWC) could become the limiting factor for ET response to VPD,
-which our theory does not account for. In addition, low soil moisture
-conditions themselves increase VPD through land-atmosphere feedback
-\citep[][]{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006,
-  Salvucci_2013, Gentine_2016, Berg_2016}.
-
-Within our framework, any systematic bias due to the
-failure to account for SWC's effects in very dry conditions should
-manifest itself in a functional relationship between $\sigma$, our
-uncertainty measure, and SWC, which is observed at the FLUXNET
-sites. Examining the relationship between $\sigma$ and SWC will test
-to what extent our theory breaks down in very dry soil moisture
-conditions.
-
-If we again view $\sigma \cdot uWUE$ as, in addition to a measure of
-uncertainty, a short time-scale observation of uWUE, we would expect
-$\sigma \cdot uWUE$ to decrease at low SWC. If $\sigma \cdot uWUE$ is
-a very strong function of SWC, then our theory should be conditioned
-more strongly on well watered soil conditions. If $\sigma \cdot uWUE$
-is weakly a function of SWC, then our theory is more universal and
-independent of soil moisture conditions.
-
-Indeed, for all PFTs, there is some slight dependence of
-$\sigma \cdot uWUE$ on SWC, especially at low SWC (Figure
-\ref{swc_boxplot}). The portion of observations for which our theory
-is biased by SWC-limitations varies by PFT, due to the nonlinear
-threshold at which soil moisture availability limits plant
-function. For CRO, MF and DBF, soil water content only matters for
-about the lowest 5\% of observations (each box is 5\% of observations
-in Figure \ref{swc_boxplot}). And, CRO, relative to DBF, has a very
-weak dependence on soil moisture, which is reflective of the high
-likelihood that CRO sites are optimally irrigated and not water
-stressed, suggesting that observed departures from the theory for CRO
-are due to other factors (i.e. different photosynthesis pathways; C3/C4) than soil
-moisture. For ENF, EBF, SAV, and WSA, systematic SWC-induced biases in
-$\sigma \cdot uWUE$ emerge in about the lowest 20-30\% of SWC
-conditions, although large variability in the SWC-$(\sigma \cdot uWUE)$
-relationship hampers interpretation for WSA. CSH presents a special
-case: the limited number of sites (2) preclude comment on the
-exact PFT-wide relationship.
-
-In contrast to ENF, CRO, and DBF, for GRA the relationship between
-$\sigma \cdot uWUE$ and SWC is more linear and affects a greater
-portion of observations; about 60\% of observations. Clearly, for GRA
-soil water frequently impacts plant function and alters ET response,
-and our theory is limited for the majority of environmental
-conditions. It is therefore not surprising that our theory tested
-poorly against the data for GRA, relative to to the other PFTs
-(Section \ref{testing}). For all other PFTs occurrence of soil
-moisture impacts was rare enough to not manifest itself in bulk
-statistics and figures.
-
-\begin{figure}
-  \centering
-  \centerline{\includegraphics[width=1.2\textwidth]{./swc_boxplot.pdf}}
-  \caption{$\sigma \cdot uWUE$ as a function of PFT and SWC. Each box
-    plot represents 5\% of all observations. Note that the highest
-    10\% of SWC observations are excluded to better resolve variability in the much more narrow bins at lower SWC (e.g. SWC has long tails at high values).}
-  \label{swc_boxplot}
-\end{figure}
-
-The observed dependence of $\sigma \cdot uWUE$ on SWC for GRA would
-explain the deficiencies of our theory compared to the observations in
-Figure \ref{idealized_sign}, specifically the trend back towards
-negative soil moisture at high VPD. Aforementioned feedbacks between
-the land surface and the atmosphere, which are not accounted for due
-to our focus on the one-way response of the land surface to
-atmospheric conditions, would cause high VPD to be correlated with low
-SWC \citep[][]{Gentine_2016, Berg_2016}. So, at high VPD observations
-of low SWC are more likely, and this low SWC causes a lower uWUE
-(Figure \ref{swc_boxplot}). The lower uWUE at low SWC/high VPD then
-leads to the observed downturn towards decreasing ET with increasing
-VPD, and the deviation between our theory (based on a constant uWUE
-assumption) and observations. It is also perhaps not coincidental that
-the portion of $\sigma \cdot uWUE$ affected by low soil moisture
-observations is similar to the portion of observations that do not
-match our theory at high VPD. Indeed, this may suggest that for all PFTs
-except for GRA, coupling and feedbacks between SWC, VPD and plant
-function are relatively rare.
-Future research will explore these relatively rare feedbacks in extreme conditions, which due to analytical intractability will require more opaque numerical analysis of many more complex processes, including boundary layer growth and state and their relationship to surface layer coupling and free-tropospheric lapse rates and humidity.
-
-To summarize, our theory is limited by its inability to account for
-soil water impacts on land surface response, and feedbacks between SWC
-and VPD. Fortunately, for most PFTs SWC's effect on ET is relatively
-rare ($<$30\% of observations) and does not manifest itself in the
-majority of observations and bulk statistics. However for GRA, SWC decreases water use efficiency for the majority of
-the observations. Soil moisture effects explain the deficiencies of
-our theory in Section \ref{testing}, particularly for GRA. By
-conceptualizing SWC effects as a change in $uWUE$ (and/or $g_1$), it
-will be possible for future analysis to explore the importance of soil
-moisture on plant response to VPD, and feedbacks between plant
-function, SWC, and VPD.
-
-\subsection{Functional form of ET dependence on VPD and its relation
-  to VPD dependence}
-\label{functional_form}
-
-The theory described in Section \ref{sign_term} indicates that for a
-given $uWUE$ and $g_1$, the ET dependence on VPD should be concave
-upward, which is confirmed by eddy covariance data across most PFTs. In other words, there should be some local minimum in ET at a
-critical VPD$_{crit}$, assuming the scaling and plant terms  (e.g. aerodynamic
-conductance, $\Delta$, $g_1$ and $uWUE$) are held fixed. This result warrants further
-investigation, because to our knowledge no one has derived the
-theoretical ecosystem-scale relationship between ET and VPD while
-controlling for other environmental conditions. In particular, from
-personal communication, there is an apparent lack of consensus over
-whether the shape of the ET-VPD curve should be concave upward (our
-result) or concave downward in the absence of dramatic water stress. Given that understanding the ET-VPD
-relationship of the one-way plant response is fundamental to
-hypothesizing about any feedbacks between the land surface and the
-atmosphere, we analyze why our derived ET-VPD relationship is concave
-upward, particularly with respect to the exponent of VPD dependence in
-$uWUE$ and the Medlyn unified stomatal conductance model.
-
-There is a theoretical basis for the square root VPD dependence in
-both the stomatal conductance model and $uWUE$ based on the assumption
-that stomata behave to maximize carbon gain while minimizing water
-loss, which observations also generally support \citep{Lloyd_1991,
-  MEDLYN_2011, Lin_2015, Zhou_2014, Zhou_2015, Medlyn_2017}. However,
-some purely empirical results that fit the exponent of the VPD
-dependence to data have shown that it may vary slightly from 1/2,
-suggesting that stomata, as well as ecosystem-scale quantities based
-on stomata theory, may not always function optimally \citep{Zhou_2015,
-  Lin_2018}. Specifically with regards to $uWUE$, one would not expect
-that this ecosystem scale WUE quantity will respond to VPD exactly
-analogously to stomata. Direct soil evaporation's contributions to ET
-should shift the exponent of the VPD dependence, especially at
-conditions of low GPP when we would expect a systematically larger
-portion of direct soil evaporation contributions to ET, because we
-would also expect lower amounts of transpiration at low GPP. \citet{Zhou_2015}'s
-results corroborate this: they found a mean empirically fit
-exponential VPD dependence of 0.55, varying slightly from the
-theoretically optimal value of of 0.5 for AmeriFlux
-sites. Interpreting \citet{Lin_2018}'s results, which also show
-variance in the empirical exponent of the VPD dependence of the
-stomatal conductance model, is more difficult as \citet{Lin_2018} do
-not handle GPP/A dependence of stomatal conductance in a directly
-analogous manner to the optimal theory in \citet{MEDLYN_2011} and
-\citet{Medlyn_2017}. Regardless, given that these recent results on
-the relationship between VPD, GPP, and ET \citep{MEDLYN_2011,
-  Zhou_2014, Zhou_2015, Medlyn_2017} form the backbone of our analysis
-and are what allowed us to derive an explicit ET expression for the
-first time (Equation \ref{et}), we will analyze if and how assumptions
-about the exponent of the VPD dependence impacts the shape of the
-ET-VPD dependence.
-This analysis is also important to understand whether the choice of stomatal conductance model alters the fundamental behavior of the ET-VPD relationships, as many commonly used models  utilize a VPD exponent other than the 1/2 suggested by optimal theory \citep[e.g.
-][ which uses an exponent of 1]{Leuning_1990}.
-
-By introducing $n$ and $m$ we can free our stomatal
-conductance model from assumptions about VPD dependence:
-% \begin{linenomath*}
-  \begin{equation}
-    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_*}{VPD^m}\right) \frac{*WUE \; ET}{c_a \; VPD^n},
-    \label{m_n}
-  \end{equation}
-% \end{linenomath*}
-where:
-\[*WUE = \frac{GPP}{ET}VPD^n,\] and $g_*$ is a generic slope parameter
-of units $VPD^m$. To determine how the exponent $n$ and $m$ alter the
-shape of the ET-VPD dependence we find the roots of the second
-derivative of ET, using Equation \ref{m_n} for stomatal conductance
-($g_s$), with respect to VPD: \input{d2_solutions.tex} With this
-result we have defined the family of curves separating concave up from
-concave down ET solutions (Figure \ref{concave}). These curves are
-only functions of the exponent of the VPD dependence and a quantity we
-call non-dimensional VPD ($VPD^m/g*$). Several important
-relations reveal themselves from Equation \ref{curves}:
-
-\begin{figure}
-  \centering
-  \centerline{\includegraphics[width=0.75\textwidth]{./concave.pdf}}
-  \caption{ Solutions corresponding to inflection points between
-    concave up and concave down ET-VPD curves (Equation \ref{curves})
-    for three specific scenarios. Solutions are defined in terms of a
-    non-dimensional VPD ($VPD^m/g_*$), but to aide physical
-    interpretation the horizontal axis is additionally provided in terms of dimensionalized VPD assuming $m=1/2$ and
-    $g_*=110\; Pa^{1/2}$ (average of all PFT $g_1$). The vertical axis
-    has a different interpretation depending on the solution
-    curve. For the blue line ($m$ varying), it corresponds to $m$,
-    for the orange line ($n$ varying) it corresponds to $n$, and for
-    the green line it corresponds to the value of both $n$ and $m$
-    ($n=m$). Regions of the parameter space that correspond to
-    concave up and concave down results are labeled.}
-  \label{concave}
-\end{figure}
-
-\begin{itemize}
-  \item For optimal behavior (n, m = 1/2) the ET-VPD curve will be
-    concave up regardless of the magnitude of the plant constants
-    $g_1$ and $uWUE$. Therefore, the general concave up nature of our
-    results, given an assumption of optimal behavior, is insensitive
-    to plant type.
-  \item For all physically possible exponents of VPD dependence ($n,
-    m$), whether the solution is concave up or concave down does not
-    depend on $uWUE$.
-  \item In general, increasing the exponent of VPD dependence
-    increases the likelihood of a concave up result. Additionally, as
-    the exponent of VPD dependence increases from the optimum value of
-    1/2, whether the curve is concave upward or concave downward
-    becomes a function of the plant specific slope parameter $g_*$,
-    through non-dimensional VPD ($VPD^m/g_*$). Because the exponent of
-    the VPD dependencies is capable of altering the fundamental shape
-    of ET-VPD dependence, future research investment in understanding
-    the exact VPD dependence of stomatal conductance, and further
-    reconciliation of empirical and theoretical stomatal and
-    ecosystem behavior should be prioritized.
-\end{itemize}
-While it is possible that in the future some other form of
-VPD dependence is derived, at present \cite{MEDLYN_2011} and
-\cite{Zhou_2014} firmly established n=m=1/2 as the most likely
-candidate given current theory and empirical data. Additionally, we
-argue that a concave up result matches physical intuition more than a
-concave down result. Plants must maintain nutrient and sugar transport
-through the phloem and xylem. To accomplish this, stomata must remain
-slightly open \citep{De_2013, Nikinmaa_2013, Ryan_2014}. Furthermore,
-even if complete stomatal closure were possible, cuticular water loss
-and [at the ecosystem-scale] direct soil evaporation are still sources
-of ET which increase with VPD, independent of stomatal
-closure. Therefore, in the limit as VPD becomes large and we assume
-plants are exercising all strategies to reduce ET, any further increase
-in VPD should result in an increase in ET through cuticular water loss
-and/or direct soil evaporation. This inevitable transition from
-conditions when stomata respond strongly to VPD to conditions when
-stomata response is asymptoting towards full closure would cause a
-concave up ET-VPD curve, which is matched by the theory. In short,
-plant response becomes more limited as VPD increases, while
-atmospheric demand monotonically increases with VPD, leading to the
- result that atmospheric demand dominates plant response when
-atmospheric demand is high.
-
-This analysis allows us to understand the theoretical shape of the ET
-response to VPD with environmental conditions held
-fixed. Accomplishing this with purely statistical
-methods applied to flux observations would be very difficult, given
-the relatively fast time scale of plant response and the
-non-stationarity of [solar forced] environmental conditions over the
-relatively coarse (half hourly) flux estimates (which is required to
-obtain robust eddy-covariance statistics). Our results on the shape of
-the ET-VPD curve with environmental conditions held
-fixed can be built upon with future work examining how changes in VPD
-and environmental conditions (e.g. soil water storage) feedback upon
-one another. In the soil water storage example, over very long time
-scales extremely high VPD perturbations coupled with no precipitation
-could result in decreases in soil water storage such that water
-becomes limiting. This could be represented by an extension of our
-framework in which uWUE is allowed to decrease with decreasing SWC, as
-observed in Section \ref{swc_section}. Here, we focus our results by
-assuming constant PFT-wide conditions to build baseline intuition for
-ET-VPD dependence. For most PFTs, the theory with plant function held
-fixed matches the leading order behavior of the observations where
-plant function varies (Section \ref{testing}).
-
-\section{Conclusions}
-
-We derived a new form of Penman Monteith using the concept of
-semi-empirical optimal stomatal regulation \citep{Lin_2015,
-  MEDLYN_2011} and near constant uWUE \citep{Zhou_2015} to remove the
-implicit dependence of stomatal conductance on GPP and ET. With our
-new form of Penman Monteith we developed a theory for when an
-ecosystem will tend to reduce or increase ET with increasing VPD,
-which we evaluated against a range of eddy-covariance data spanning
-different climates and plant functional types. The goal was to capture
-the leading order behavior of the system to gain some intrinsic
-knowledge for its behavior. This intuition can be used to disentangle
-land atmosphere feedbacks in more complicated scenarios, and will aid
-interpretation of observations and sophisticated models.
-
-Our theory suggests that for a majority of environmental conditions,
-plants will tend to conserve water and reduce ET with increasing
-VPD. Stomatal regulation and plant physiological response strongly
-regulate ET, and this regulation varies by PFT. CROs are the least
-water conservative, while DBF, EBF, SAV, GRA, MF, WSA, ENF and CSH are
-progressively more water conservative (more likely to reduce ET in
-response to increasing VPD). SAV and WSA exhibit positive ET response
-to VPD not necessarily because of poor water conservation strategies
-relative to other PFTs, but because of greater occurrence of high
-atmospheric demand (VPD) relative to other PFTs. Observations of ET
-response to VPD exhibit the same general behavior as the theory, with
-ET response becoming more positive (atmospheric demand dominating) as
-environmental VPD increases within a PFT, and more negative for PFTs
-that are adapted to arid conditions and prioritize water conservation
-over primary production.
-
-The consequences of our theory our twofold. First, the ET response to
-VPD is always far more negative than it would be if only atmospheric
-demand is considered and plant physiological response is ignored (as
-with PET). Therefore, any drought indices or analysis using PET will
-not capture how vegetated surfaces respond and react to increases in
-atmospheric demand. However, the new uWUE-version of Penman Monteith
-we derived (Equation \ref{et}) could be used as a complement to PET in
-drought indices. One could calculate and compare indices when plant
-response is ignored (PET) or included (Equation \ref{et}).
-
-Lastly, our paper builds important intuition for how plants will
-respond to VPD perturbations. We show that given optimal stomatal
-function and fixed environmental conditions, the ET-VPD dependence is
-theoretically concave upward, with ET increasing with increasing VPD as
-VPD increases past some critical value (Table \ref{vpd_crit}). However
-future research should focus on fully understanding the form of
-stomatal VPD dependence, as this result is sensitive to the exponent of
-VPD dependence, which we currently believe is 1/2 \citep{MEDLYN_2011,
-  Zhou_2014}. Our results are applicable to understanding the impact
-of  expected increases in VPD
-induced by global change.  Plant physiological responses to direct
-CO$_2$ effects \citep[e.g.,][]{Swann_2016, Lemordant_2018} receives more attention
-than physiological response to indirect effects like increased
-VPD. Here, we provide broad PFT-focused results showing a likely
-decrease in ET in response to positive VPD perturbations (atmospheric
-drying), which is consistent with recent observational analysis
-\citep[e.g.,][]{Rigden_2017}. Feedbacks between the land and the
-atmosphere may alter the net response to a long-timescale global VPD perturbation,
-but our focus on the one way plant response to a VPD perturbation in
-the atmospheric boundary layer is an important first step to
-disentangling such feedbacks, both in observations and model
-simulations of the present and future. By removing Penman Monteith's
-dependence on implicit relationships between GPP, VPD, and ET, we
-allow for explicit, accurate future analysis of plant-VPD feedbacks
-in the atmospheric boundary layer (Equation \ref{et}). Our approach
-can be extended to examine varying plant response to more nuanced
-consideration of plant type and climate. Any plant physiological
-heterogeneity or feedback that can be conceptualized with shifts in $g_1$
-\citep[e.g.][]{Lin_2015, Medlyn_2017} and/or uWUE
-\citep[e.g.][]{Zhou_2014} are representable within our framework, which
-opens the door for a hierarchy of more sophisticated climate- and
-plant-specific analysis of ET sensitivity to environmental variables
-(including VPD). We argue that such simplified conceptual frameworks
-are critical tools for disentangling land-atmosphere feedbacks at
-various scales, from diurnal to seasonal and beyond, and to
-characterize ET response in a warmer, atmospherically drier, and
-enriched CO$_2$ world.
-
-\acknowledgments This work used eddy covariance data acquired and
-shared by the FLUXNET community, including these networks: AmeriFlux,
-AfriFlux, AsiaFlknux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont,
-ChinaFlux, Fluxnet-Canada, GreenGrass, ICOS, KoFlux, LBA, NECC,
-OzFlux-TERN, TCOS-Siberia, and USCCC. The ERA-Interim reanalysis data
-are provided by ECMWF and processed by LSCE. The FLUXNET eddy
-covariance data processing and harmonization was carried out by the
-European Fluxes Database Cluster, AmeriFlux Management Project, and
-Fluxdata project of FLUXNET, with the support of CDIAC and ICOS
-Ecosystem Thematic Center, and the OzFlux, ChinaFlux and AsiaFlux
-offices. This material is based upon work supported by the National
-Science Foundation Graduate Research Fellowship under Grant No. DGE
-16-44869. All code and data used in this analysis, including those
-used to generate the figures and tables, are publicly available at
-\url{https://github.com/massma/climate\_et}.
-%\hyperref[https://github.com/massma/climate\_et]{https://github.com/massma/climate_et}.
-
-\bibliography{references.bib}
-
-\end{document}

### 2018-10-30 doc/paper/body.tex
syling for hess

@@
 plant function varies (Section \ref{testing}).
-
-\acknowledgments This work used eddy covariance data acquired and shared
-by the FLUXNET community, including these networks: AmeriFlux, AfriFlux,
-AsiaFlknux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont,
-ChinaFlux, Fluxnet-Canada, GreenGrass, ICOS, KoFlux, LBA, NECC,
-OzFlux-TERN, TCOS-Siberia, and USCCC. The ERA-Interim reanalysis data
-are provided by ECMWF and processed by LSCE. The FLUXNET eddy covariance
-data processing and harmonization was carried out by the European Fluxes
-Database Cluster, AmeriFlux Management Project, and Fluxdata project of
-FLUXNET, with the support of CDIAC and ICOS Ecosystem Thematic Center,
-and the OzFlux, ChinaFlux and AsiaFlux offices. We would like to thank
-Dr. Trevor Keenan for providing tools for citation of the FLUXNET2015
-dataset, available at
-\url{https://github.com/trevorkeenan/FLUXNET_citations}. This material is
-based upon work supported by the National Science Foundation Graduate
-Research Fellowship under Grant No. DGE 16-44869. All code and data used
-in this analysis, including those used to generate the figures and
-tables, are publicly available at
-\url{https://github.com/massma/climate\_et}.
-%\hyperref[https://github.com/massma/climate\_et]{https://github.com/massma/climate_et}.
-
-\begin{appendices}
-  \section{FLUXNET2015 sites}\label{flux_sites}
-  % below taken adapted from Trevor Keenan's FLUXNET citations
-  % at github.com/trevorkeenan/FLUXNET_citations
-  Table A1: Metadata and citations for flux sites used in this analysis.
-All data are gathered from \url{www.fluxdata.org}, and citations are aggregated using tools available at \url{https://github.com/trevorkeenan/FLUXNET_citations}.
-  \begin{longtable}{l l l l l l l}
-    \hline
-    \textbf{Site} &
-    \textbf{PFT} &
-    \textbf{Lat} &
-    \textbf{Lon} &
-    \textbf{Clim\textsuperscript{1}} &
-    \textbf{Period} &
-    \textbf{References} \\
-    [0.5ex]
-    \hline
-    \input{flux_sites.tex}
-    [1ex]
-    \hline
-  \end{longtable}
-  \textsuperscript{1}  K{\"o}ppen Climate classification.
-\end{appendices}
-
-\bibliography{references.bib}
-
-\end{document}

### 2018-10-30 doc/paper/abstract.tex
need to test reoproducibility

@@
 serves as an example of the utility of our simplified framework for
-disentangling land-atmosphere feedbacks at multiple scales, including
+disentangling land-atmosphere feedbacks, including
 the characterization of ET response in an atmospherically drier,

### 2018-10-30 doc/paper/body.tex
need to test reoproducibility

@@
 
-Thus, with this uncertainty analysis we can evaluate departure from
+So, with this uncertainty analysis we can evaluate departure from
 our theory in observations, as a departure of $\sigma$ from unity. The
@@
 observations themselves. The variability of $\sigma$ then propagates
-through any uncertainty to our partial derivative of Equation
+ any uncertainty through to our partial derivative of Equation
 \ref{et_sigma} with respect to VPD:
@@
 using commonly available weather station data and the constants
-published here. This makes Equation \ref{et} a useful alternative to
+published here. This makes Equation \ref{et} useful in addition to
 PET in drought indices and hydrometeorological analysis for vegetated
@@
 and $g_1$, and can also account for changes in CO$_2$ concentration
-(assuming some relationship between $\lambda$ and [CO$_2$]), contrary
-to typical drought indices such as the Palmer Drought Severity Index
+(assuming some relationship between $\lambda$ and [CO$_2$]), which is missing
+in other drought indices such as the Palmer Drought Severity Index
 (PDSI) \citep{Swann_2016, Lemordant_2016, Lemordant_2018}. Lastly, all
@@
 % \end{linenomath*}
-\normalsize noting that $VPD_{crit}$ mostly depends on the PFT
-parameters uWUE and $g_1$, and only varies weakly with climate as most
-other parameters related to the environment are nearly constant. The
+  \normalsize
+
+  noting that $VPD_{crit}$ mostly depends on the PFT parameters uWUE
+  and $g_1$, and only varies weakly with climate as most other
+  parameters related to the environment are nearly constant. The
   calculated value of $VPD_{crit}$ for each PFT is shown in Table
@@
   ($\frac{\partial \; ET}{\partial \; VPD} < 0$), and for values of
-$VPD$ greater than $VPD_{crit}$, ET will increase with increasing VPD
-($\frac{\partial \; ET}{\partial \; VPD} > 0$). In other words,
+  $VPD$ greater than $VPD_{crit}$, ET will increase with increasing
+  VPD ($\frac{\partial \; ET}{\partial \; VPD} > 0$). In other words,
   ecosystems regulate and mitigate evaporative losses up to the VPD
@@
   evaporation as stomata are still strongly regulating vapor fluxes to
-the atmosphere. Indeed, even in the absence of soil pore evaporation,
-stomata do not shut down entirely at very high VPD so that ET does not
-go to zero, because stomata are still slightly open to perform some
-photosynthesis \citep{Ball_1987, Leuning_1990, MEDLYN_2011}. In
-addition, upward xylem transport is necessary to maintain phloem
-transport, as well as nutrient transport and thus carbon allocation
-\citep{De_2013, Nikinmaa_2013, Ryan_2014}.
+  the atmosphere. However, even in the absence of soil pore
+  evaporation ET cannot go completely to zero at high VPD, because
+  stomata are still slightly open to perform some photosynthesis
+  \citep{Ball_1987, Leuning_1990, MEDLYN_2011}. In addition, upward
+  xylem transport is necessary to maintain phloem transport, as well
+  as nutrient transport and thus carbon allocation \citep{De_2013,
+    Nikinmaa_2013, Ryan_2014}.
 
@@
 environmental VPDs at the FLUXNET sites are plotted parallel to the
-x-axis. For CSH and WSA, VPD is always less than VPD$_{crit}$ (off scale) so
-that the plant response dominates in typical environmental conditions,
-emphasizing the water conservative strategy of those plants. For CRO
-on the other hand, VPD is higher than VPD$_{crit}$ for more than 50\%
-of observations, emphasizing that those plants operate with an
-aggressive water usage strategy, are water intensive and were actually
-engineered for photosynthesis rather than water saving. For DBF, EBF,
-MF, GRA, and SAV more than half of the observed VPD are less than
-VPD$_{crit}$, i.e. in conditions where plant response dominates. SAV
-has a more water conservative response than the forest, grass, and
-crop plan types, but still responds by increasing ET with increasing
-VPD for about a quarter of observations, due to the high aridity (VPD)
-of the SAV ecoclimate. It is also important to note that for all PFTs,
-even when atmospheric demand dominates, ET response to VPD is still
-far more negative than it would be for potential evaporation
-$\partial PET/\partial VPD$, i.e. atmospheric demand only, emphasizing
-that there is still a strong regulation of evaporative flux by stomata
-and though the plant xylem. The sign term in the PET case would just
-be a constant ($\frac{c_p}{R_{air}} \approx 3.5$), which is far larger
-than any part of the curves for any PFT. This highlights the
-deficiencies of PET's ability to capture land response to changes in
-atmospheric dryness (VPD). Plants are always regulating water exchange
-from the land surface, even when they reach the limits of they ability
-to do so.
+x-axis. For CSH and WSA, VPD is always less than VPD$_{crit}$ (off
+scale) so that the plant response dominates in typical environmental
+conditions, emphasizing the water conservative strategy of those
+plants. For CRO on the other hand, VPD is higher than VPD$_{crit}$ for
+more than 50\% of observations, emphasizing that those plants operate
+with an aggressive water usage strategy, are water intensive and were
+actually engineered for photosynthesis rather than water saving. For
+DBF, EBF, MF, GRA, and SAV more than half of the observed VPD are less
+than VPD$_{crit}$, i.e. in conditions where plant response
+dominates. SAV has a more water conservative response than the forest,
+grass, and crop plan types, but still responds by increasing ET with
+increasing VPD for about a quarter of observations, due to the high
+aridity (VPD) of the SAV ecoclimate. It is also important to note that
+for all PFTs, even when atmospheric demand dominates, ET response to
+VPD is still far more negative than it would be for potential
+evaporation $\partial PET/\partial VPD$, i.e. atmospheric demand only,
+emphasizing that there is still a strong regulation of evaporative
+flux by stomata and though the plant xylem. The sign term in the PET
+case would just be a constant ($\frac{c_p}{R_{air}} \approx 3.5$),
+which is far larger than any part of the curves for any PFT. Plants
+are always regulating water exchange from the land surface, even when
+they reach the limits of they ability to do so.
 
@@
   \item In general, increasing the exponent of VPD dependence
-    increases the likelihood of a concave up result. Additionally, as
+    increases the likelihood of a concave down result. Additionally, as
     the exponent of VPD dependence increases from the optimum value of

### 2018-10-30 doc/paper/conclusions.tex
need to test reoproducibility

@@
 
-The consequences of our theory our twofold. First, the ET response to
-VPD is always far more negative than it would be if only atmospheric
-demand is considered and plant physiological response is ignored (as
-with PET). Therefore, any drought indices or analysis using PET will
-not capture how vegetated surfaces respond and react to increases in
-atmospheric demand. However, the new uWUE-version of Penman Monteith
-we derived (Equation \ref{et}) could be used as a complement to PET in
-drought indices. One could calculate and compare indices when plant
-response is ignored (PET) or included (Equation \ref{et}).
-
-Lastly, our paper builds important intuition for how plants will
+Our paper builds important intuition for how plants will
 respond to VPD perturbations. We show that given optimal stomatal

### 2018-10-30 doc/paper/intro.tex
need to test reoproducibility

@@
 atmospheric drying then soil moisture will be better conserved. This
-would seem a sensible evolutionary strategy to cope with aridity. If
-stomata were fully passive \citep [similar to soil pores,
-e.g. ][]{Or_2013}, increased atmospheric aridity would strongly reduce
-soil moisture \citep{Berg_2017}. In turn, this would further increase
-aridity as low soil moisture levels increase the Bowen ratio, and
-cause increased temperature and atmospheric drying
+represents a sensible evolutionary strategy to cope with aridity: save
+water for periods when atmospheric demand for water is relatively low,
+and atmospheric carbon can be accessed with a relatively smaller cost
+in water loss. If instead stomata were fully passive \citep [similar
+to soil pores, e.g. ][]{Or_2013}, increased atmospheric aridity would
+strongly reduce soil moisture \citep{Berg_2017}. This could further
+increase aridity as low soil moisture levels increases the Bowen
+ratio, leading to increased temperature and atmospheric drying
 \citep[][]{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006,
-  Salvucci_2013, Gentine_2016, Berg_2016}. This however would not seem
-to be a sensible strategy for plants from an evolutionary
-standpoint. As a counterpoint, one may argue that increases in ET with
-increasing VPD could increase the likelihood of precipitation
+  Salvucci_2013, Gentine_2016, Berg_2016}. Therefore, passive
+regulation and a lack of soil moisture conservation does not seem to
+be a sensible strategy for plants from an evolutionary standpoint.
+
+As a counterpoint, one may argue that increases in ET with increasing
+VPD could increase the likelihood of precipitation
 \citep[e.g.,][]{Findell_2011}. However, increases in ET do not always
@@
 conductance ($g_s$), including VPD response, by combining an optimal
-photosynthesis theory \citep{Cowan_1977} with an
-empirical approach, and extended use of this model to the ecosystem
-scale in \citet{Medlyn_2017}. Additionally, \citet{Zhou_2014}
-demonstrated that a quantity underlying water use efficiency
+photosynthesis theory \citep{Cowan_1977} with an empirical approach,
+and extended use of this model to the ecosystem scale in
+\citet{Medlyn_2017}. Additionally, \citet{Zhou_2014} demonstrated that
+a quantity underlying water use efficiency
 $\left(uWUE = \frac{GPP\; \sqrt{VPD}}{ET}\right)$ properly captures a
@@
 growing season across space and time, within a PFT
-\citep{Zhou_2015}. These results provide novel tools to examine how ET
-varies in response to VPD, while explicitly including VPD's effect on
-stomatal conductance for the first time. In this manuscript, we
-leverage these tools to derive the theoretical one-way response of ET
-to VPD with other environmental variables properly controlled for,
-i.e. we develop a framework for evaluating the partial derivative of
-ET with respect to VPD. Our theory is validated and tested at multiple
-eddy-covariance stations spanning various climates and plant
-functional types.
+\citep{Zhou_2015}. While stomatal conductance parameterizations and
+uWUE greatly simplify complex plant physiological processes, they
+still capture leading order ecosystem behavior for vegetated
+surfaces \citep{Medlyn_2017, Zhou_2014}, and are novel tools to
+transparently understand and develop intuition for complex
+ecohydrologic response to environmental forcing.
+
+In this manuscript,
+we leverage uWUE and recent developments in stomatal conductance
+parameterizations \citep{MEDLYN_2011} to derive the theoretical
+one-way response of ET to VPD with other environmental variables
+properly controlled for, i.e. we develop a framework for evaluating
+the partial derivative of ET with respect to VPD. For the first time,
+we explicitly include VPD's full effect on stomatal conductance,
+including its leading order impact on photosynthesis. Our theory is
+validated and tested at multiple eddy-covariance stations spanning
+various climates and plant functional types.

### 2018-10-30 doc/paper/ms.tex
forgot to add archive base file (ms.tex), updated readme

--- 2018-05-14 doc/paper/vpd_et_paper.tex
@@
+% this is used for arkiv submissions
 \documentclass[12pt]{article}
@@
 \maketitle
-
 \begin{abstract}
-  Increasing vapor pressure deficit increases atmospheric demand for water, and vapor pressure deficit is expected to rise with increasing greenhouse gases.
-While increased evapotranspiration (ET) in response to increased atmospheric demand seems intuitive, plants are capable of reducing ET in response to increased VPD by closing their stomata, in an effort to conserve water.
-Here we examine which effect dominates response to increasing VPD: atmospheric demand and increases in ET, or plant physiological response (stomata closure) and decreases in ET.
-We use Penman-Monteith, combined with semi-empirical optimal stomatal regulation theory and underlying water use efficiency, to develop a theoretical framework for understanding how ET responds to increases in VPD.
-
- The theory suggests that for most environmental conditions and plant types, plant physiological response dominates and ET decreases with increasing VPD.
-Plants that are evolved or bred to prioritize primary production over water conservation (e.g.
-crops) exhibit a higher likelihood of atmospheric demand-driven response (ET increasing).
-However for forest, grass, savannah, and shrub plant types, ET more frequently decreases than increases with rising VPD.
-This work serves as an example of the utility of our simplified framework for disentangling land-atmosphere feedbacks at multiple scales, including the characterization of ET response in an atmospherically drier, enriched CO$_2$ world.
+  \input{abstract}
 \end{abstract}
@@
 
-Vapor pressure deficit (VPD) is expected to rise over continents in
-the future due to the combination of increased temperature and,
-depending on region, decreased relative humidity
-\citep{Byrne_2013}. Increases in VPD increase the atmospheric demand
-for evapotranspirated water \citep{Penman_1948, Monteith_1965}, but
-also stress plant stomata \citep{Leuning_1990, MEDLYN_2011}.
+\input{intro.tex}
 
-The opposing effects of increased atmospheric demand and higher
-stomatal stress lead to two possible perspectives for how
-evapotranspiration (ET) responds to shifts in VPD. The first, a
-hydrometeorological perspective, is that higher VPD increases
-atmospheric demand for water from the land surface, and this drives an
-increase in evapotranspiration (ET). This perspective is particularly
-relevant because potential evapotranspiration (PET), which is used in
-many drought indices and hydrometeorological studies
-\citep[e.g.,][]{Heim_2002, Scheff_2015}, typically only quantifies
-changes in atmospheric demand and fails to account for ecosystem
-response \citep{Swann_2016}. In reality, plants' stomata have evolved
-to optimally regulate the exchange of water and carbon, and tend to
-partially close in response to increased atmospheric dryness
-\citep{Farquhar_1978, Ball_1987, Leuning_1990, MEDLYN_2011}. This
-leads to a plant physiology perspective, in which an increase in VPD,
-particularly in well-watered soil conditions, may actually correspond
-to a decrease in ET because of stomatal closure
-\citep[e.g.][]{Rigden_2017}.  In other words, the question ``When does
-VPD drive or reduce ET?'' can be related to whether plant regulation
-or atmospheric demand dominates ET response.
-
-The ET response to changes in VPD alters water partitioning between
-the soil and atmosphere. If ecosystem plant response reduces ET with
-atmospheric drying then soil moisture will be better conserved. This
-would seem a sensible evolutionary strategy to cope with aridity. If
-stomata were fully passive \citep [similar to soil pores,
-e.g. ][]{Or_2013}, increased atmospheric aridity would strongly reduce
-soil moisture \citep{Berg_2017}. In turn, this would further increase
-aridity as low soil moisture levels increase the Bowen ratio, and
-cause increased temperature and atmospheric drying
-\citep[][]{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006,
-  Salvucci_2013, Gentine_2016, Berg_2016}. This however would not seem
-to be a sensible strategy for plants from an evolutionary
-standpoint. As a counterpoint, one may argue that increases in ET with
-increasing VPD could increase the likelihood of precipitation
-\citep[e.g.,][]{Findell_2011}. However, increases in ET do not always
-guarantee an increased likelihood of precipitation, which depending on
-environmental conditions could cause a decrease in the likelihood of
-precipitation \citep[][]{Gentine_2013}. Furthermore, any increases in
-precipitation are likely to be non-local, such that plants giving up
-water to the atmosphere are not guaranteed to reap the benefits of
-water returned from the atmosphere to the soil. Using this subjective
-logic, from an ecosystem evolutionary perspective, water stored in
-soil seems to be worth much more than the chance of water returned as
-precipitation.
-
-We can use intuition about plant water conservation strategy to
-hypothesize about ET response to changes in VPD. Plants and ecosystems
-that evolved to conserve water, such as arid shrubs or savannah,
-should be more likely to reduce ET with increasing VPD, and plants
-that have evolved or have been engineered to care little about water,
-such as crops, will be more likely to increase ET with increasing
-VPD. Atmospheric conditions must matter as well. At the ecosystem
-scale, there are limits to plant water conservation strategies. As
-atmospheric demand for water (VPD) increases, ecosystems should begin
-to reach their water conservation limits and might not be able to
-entirely limit ET flux to the atmosphere. At this stage any further
-increase in VPD will most likely drive a (limited) increase in ET,
-because the increase in atmospheric demand for water overwhelms the
-limited plant response to conserve water.
-
-The objective of the present manuscript is to use reasonable
-approximations as a tool to develop intuition for plant response to
-atmospheric drying and evaluate the VPD dependence of ET. This
-intuition will aid interpretation of observations, full complexity
-models, and facilitate the disentanglement of complex land-atmosphere
-feedbacks. In the past, similar approaches were used to understand
-interactions between stomatal conductance, evapotranspiration and the
-environment \citep[e.g.,][]{Jarvis_1986, Mcnaughton_1991}. However, at
-the time researchers' understanding of the form of VPD's effect on
-plant physiology was limited, so they could not explore the
-sensitivity of ET to VPD, including VPD's effect on stomatal
-conductance and plant function.
-
-Recent results have drastically improved our understanding of VPD's
-impact on physiology, especially at the leaf
-level. \citet{MEDLYN_2011} developed a model for leaf-scale stomatal
-conductance ($g_s$), including VPD response, by combining an optimal
-photosynthesis theory \citep{Cowan_1977} with an
-empirical approach, and extended use of this model to the ecosystem
-scale in \citet{Medlyn_2017}. Additionally, \citet{Zhou_2014}
-demonstrated that a quantity underlying water use efficiency
-$\left(uWUE = \frac{GPP\; \sqrt{VPD}}{ET}\right)$ properly captures a
-constant relationship between GPP, ET, and VPD over a diurnal cycle at
-the ecosystem scale. uWUE is also remarkably well conserved in the
-growing season across space and time, within a PFT
-\citep{Zhou_2015}. These results provide novel tools to examine how ET
-varies in response to VPD, while explicitly including VPD's effect on
-stomatal conductance for the first time. In this manuscript, we
-leverage these tools to derive the theoretical one-way response of ET
-to VPD with other environmental variables properly controlled for,
-i.e. we develop a framework for evaluating the partial derivative of
-ET with respect to VPD. Our theory is validated and tested at multiple
-eddy-covariance stations spanning various climates and plant
-functional types.
-
-\section{Materials and Methods}
-\subsection{Data}
-\label{data}
-We use both meteorological and eddy-covariance data from the
-FLUXNET2015 database (data available at \sloppy
-\url{https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/} \sloppy),
-including all sites with at least four years of data and observations
-of the variables described in the methods section (Section
-\ref{methods}). Seventy-three sites met these requirements, and were
-grouped into nine plant functional types (PFT) according to the
-International Geosphere-Biosphere Programme vegetation classification
-scheme \citep{Loveland_1999}: cropland (CRO), grass (GRA), deciduous
-broadleaf forest (DBF), evergreen broadleaf forest (EBF), evergreen
-needleleaf forest (ENF), mixed forest (MF), closed shrub (CSH),
-savannah (SAV), and woody savannah (WSA) (site locations in Figure
-\ref{map_fig}).
-
-\begin{figure}[h]
-  \centering \includegraphics[trim={0 3cm 0 3cm}, clip]{./map.pdf}
-  \caption{Plant functional type and location of FLUXNET2015 sites
-    used in this analysis.}
-  \label{map_fig}
-\end{figure}
-
-The purpose of this study is to examine ecosystem response to
-atmospheric drying, focusing on the growing season. To accomplish
-this, we filter and quality control the data using a similar procedure
-as \cite{Zhou_2015}:
-\begin{itemize}
-\item Only measured or highest (``good'') quality gapfilled data,
-  according to quality control flags, are used.
-\item To isolate the growing season, we only use days in which the
-  average Gross Primary Productivity (GPP) exceeds 10\% of the
-  observed 95th percentile of GPP for a given site. GPP is calculated
-  using the nighttime respiration partitioning method.
-\item We remove days with rain and the day following to avoid issues
-  with rain interception and sensor saturation at high relative
-  humidity (\cite{Medlyn_2017}).
-\end{itemize}
-Additionally, as in \citet{Lin_2018}, we restrict data to the daytime,
-which is identified when downwelling shortwave radiation is greater
-than 50 W m$^{-2}$ and sensible heat flux is greater than 5 W
-m$^{-2}$. To reduce the chance of sensor saturation at high relative
-humidity, we remove all time steps for which VPD is less than .01 kPa,
-and to reduce errors at low windspeeds we remove all periods with wind
-magnitudes less than 0.5 m s$^{-1}$. Timesteps with negative observed
-GPP or ET are also removed, and we aggregate half hourly data to
-hourly averages to reduce noise \citep{Lin_2018}. After these quality
-control procedures, 434,765 upscaled hourly observations remain.
-
-\subsection{Methods}
-\label{methods}
-The Penman-Monteith equation \citep [hereafter PM,][]{Penman_1948,
-  Monteith_1965} estimates ET as a function of observable atmospheric
-variables and surface conductances:
-% \begin{linenomath*}
-  \begin{equation}
-    \label{orig_pen}
-    ET = \frac{\Delta R_{net} + g_a \rho_a c_p VPD}{\Delta + \gamma(1 + \frac{g_a}{g_s})},
-  \end{equation}
-% \end{linenomath*}
-where $\Delta$ is the change in saturation vapor pressure with
-temperature, given by Clausius-Clapeyron ($\frac{d \; e_s}{d \; T}$),
-$R_{net}$ is the net radiation minus ground heat flux, $g_a$ is
-aerodynamic conductance, $\rho_a$ is air density, $c_p$ is specific
-heat of air at constant pressure, $\gamma$ is the psychometric
-constant, and $g_s$ is the stomatal conductance (Table
-\ref{definitions}).
-
-\begin{table}
-  \caption{Definition of symbols and variables, with citation for how
-    values are calculated, if applicable.}
-  \label{definitions}
-  \centering \footnotesize
-  \begin{tabular}{l c c c}
-    \hline
-    Variable & Description & Units & Citation \\
-    \hline
-    $e_s$  & saturation vapor pressure & Pa  & - \\ 
-    $T$  & temperature  & K & - \\
-    $P$  & pressure & Pa  & - \\
-    $\Delta$  & $\frac{\partial e_s}{\partial T}$ & Pa K$^{-1}$ & - \\
-    $R_{net}$  & net radiation at land surface minus ground heat flux & W m$^{-2}$   & - \\
-    $g_a$  & aerodynamic conductance & m s$^{-1}$  & \citet{Shuttleworth_2012} \\
-    $\rho_a$  & air density & kg m$^{-3}$  & - \\
-    $c_p$  & specific heat capacity of air at constant pressure & J K$^{-1}$ kg$^{-1}$ & - \\
-    $VPD$  & vapor pressure deficit & Pa  & - \\
-    $\gamma$  & psychometric constant & Pa K$^{-1}$   & - \\
-    $g_{s-leaf}$  & leaf-scale stomatal conductance & m s$^{-1}$  & \citet{MEDLYN_2011} \\
-    $g_{s}$  &  stomatal conductance & m s$^{-1}$
-                                   & \citet{Medlyn_2017} \\
-    $g_{1-leaf}$  & leaf-scale slope parameter & Pa$^{0.5}$
-                                   & \citet{MEDLYN_2011} \\
-    $g_{1}$  & ecosystem-scale slope parameter & Pa$^{0.5}$ & \citet{Medlyn_2017} \\
-    $R$ & universal gas constant & J mol$^{-1}$ K$^{-1}$ & - \\
-    $R_{air}$ & gas constant of air & J  K$^{-1}$ kg$^{-1}$ & - \\
-    $\sigma$ & uncertainty parameter & -& - \\
-    $c_a$ & CO$_2$ concentration & $\mu$ mol CO$_2$ mol$^{-1}$ air& - \\
-    $\lambda$ & marginal water cost of leaf carbon & mol H$_2$O mol$^{-1}$ CO$_2$ & - \\
-    $\Gamma$ & CO$_2$ compensation point & - & - \\
-    $\Gamma^*$ & CO$_2$ compensation point without dark respiration & - & - \\
-    \hline
-  \end{tabular}
-\end{table}
-
-\citet{MEDLYN_2011} developed a model for stomatal conductance ($g_s$)
-by combining an optimal photosynthesis theory \citep{Cowan_1977} with an empirical approach, which describes the
-dependence of $g_s$ to VPD. This resulted in the following model for
-leaf-scale stomatal conductance:
-
-% \begin{linenomath*}
-  \begin{equation}
-    g_{s-leaf} = g_0 + 1.6 \left(1 +
-      \frac{g_{1-leaf}}{\sqrt{VPD}}\right) \frac{A}{c_a},
-    \label{leaf_medlyn}
-  \end{equation}
-% \end{linenomath*}
-where $g_{1-leaf}$ is a leaf-scale ``slope'' parameter, A is the net
-CO$_2$ assimilation rate, and $c_a$ is the atmospheric CO$_2$
-concentration at the leaf surface. \cite{MEDLYN_2011} relate the slope
-parameter ($g_{1-leaf}$) to physical parameters as:
-% \begin{linenomath*}
-  \label{slope}
-  \begin{equation}
-    g_{1-leaf} = \sqrt{\frac{3 \, \Gamma^* \, \lambda}{1.6}},\footnote{Note this expression has units of of (mmol mol$^{-1}$)$^{1/2}$, but this can be converted to Pa$^{1/2}$ using the ideal gas law.}
-  \end{equation}
-% \end{linenomath*}
-
-where $\Gamma^*$ is the CO$_2$ compensation point for photosynthesis
-(without dark respiration), and $\lambda$ is the marginal water cost
-of leaf carbon
-($\frac{\partial \; \text{transpiration}}{\partial \; A}$). So,
-$g_{1-leaf}$ is a leaf-scale term reflecting the trade-off of water for
-carbon uptake. The higher $g_{1-leaf}$, the more open the stomata and
-the more they release water in exchange for carbon.
-
-The Medlyn model for stomatal conductance has been shown to behave
-very well across PFTs \citep[][]{Lin_2015}, and has been successfully
-adopted to ecosystem scale analysis in \citet{Medlyn_2017}. In units
-of m s$^{-1}$, the ecosystem scale stomatal conductance is:
-
-% \begin{linenomath*}
-  \begin{equation}
-    g_s = \frac{R \,T}{P} \; 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{GPP}{c_a},
-    \label{medlyn}
-  \end{equation}
-% \end{linenomath*}
-
-where GPP is the ecosystem scale gross primary production, and $g_1$
-is an ecosystem scale analogue to $g_{1-leaf}$. We solve for $g_1$
-following \citet{Medlyn_2017} (Equation 5), and take the median $g_1$
-value to be representative of each PFT (Table \ref{pft}) instead of the mean to avoid extra weighting of rare outliers.
-
-\begin{table}
-  \caption{Plant functional types, their abbreviation, calculated
-    Medlyn coefficient, and calculated uWUE. uWUE from
-    \citet{Zhou_2015}, including the observed standard deviation, is
-    shown for comparison. Note that uWUE from \citet{Zhou_2015} is
-    calculated from a different set of sites, and that units are
-    converted such that the quantities work with Equations 1-8 and the
-    variables defined Table \ref{definitions}.}
-  \small
-  \label{pft}
-  \centering
-  \begin{tabular}{l c c @{\qquad} c c}
-    \hline
-    \multirow{2}[3]{*}{Abbreviation} & \multirow{2}[3]{*}{PFT} & \multirow{2}[3]{*}{$g_1$ (Pa$^{0.5}$)} & \multicolumn{2}{c}{uWUE ($\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET])}  \\
-    \cmidrule{4-5}
-
-                                     & & & fitted & \citet{Zhou_2015} \\
-
-    \hline
-    \input{pft_params.tex}
-    \hline
-  \end{tabular}
-\end{table}
-
-While Equation 3 can be used in PM (Equation 1), it will make
-analytical work with the function intractable because $GPP$ is
-functionally related to ET itself. Additionally, a perturbation to VPD
-should induce a physiological plant response that will alter GPP and
-cause an indirect change in stomatal conductance, in addition to the
-direct effect of VPD in Equation \ref{medlyn}. Therefore, in order to
-derive the response of ET to VPD, we must account for the functional
-relationship between GPP, ET, and VPD, and its effect on stomatal
-conductance. We can use aforementioned semi-empirical results of
-\citet{Zhou_2015} as a tool to approach this
-problem. \citet{Zhou_2015}, showed that  underlying Water Use
-Efficiency (uWUE):
-
-% \begin{linenomath*}
-  \begin{equation}
-    uWUE = \frac{GPP \cdot \sqrt{VPD}}{ET}
-    \label{uwue}
-  \end{equation}
-% \end{linenomath*}
-
-is relatively constant across time and moisture conditions within a
-plant functional type, and correctly captures a constant relationship
-between GPP, ET and VPD over a diurnal cycle \citep{Zhou_2014}. The
-theoretical derivation of the square root VPD dependence in $uWUE$
-leverages the same assumptions used in \cite{MEDLYN_2011} to derive the
-square-root VPD dependence of the stomatal conductance model (Equation
-\ref{medlyn}).  We can use uWUE to remove the $GPP$ dependence from
-$g_s$ in a way that makes PM analytically tractable:
-
-% \begin{linenomath*}
-  \begin{equation}
-    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{uWUE \; ET}{c_a \; \sqrt{VPD}}.
-    \label{new_g_s}
-  \end{equation}
-% \end{linenomath*}
-
-Plugging Equation \ref{new_g_s} into Equation \ref{orig_pen} and
-rearranging gives a new explicit expression for PM, in which
-dependence on $GPP$ is removed:
-
-% \begin{linenomath*}
-  \begin{equation}
-    ET = \frac{\Delta R_{net} + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R \; 1.6\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
-    \label{et}
-  \end{equation}
-% \end{linenomath*}
-
-By accounting for photosynthesis changes in ecosystem conductance, with
-Equation \ref{et} we have derived for the first time, using recent
-results \citep[][]{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017}, ET
-explicitly as function of environmental variables and two
-plant-specific constants, the slope parameter ($g_1$), and uWUE, both
-reflecting water conservation strategy. The slope parameter is related
-to the willingness of stomata to trade water for CO$_2$ and to keep
-stomata open. uWUE is a semi-empirical ecosystem-scale constant
-related to how WUE changes with VPD (specifically $VPD^{-1/2}$). It is
-also roughly proportional to physical constants:
-
-\[uWUE \appropto \sqrt{\frac{c_a - \Gamma}{1.6 \lambda}},\]
-
-where $\Gamma$ is the CO$_2$ compensation point \citep[Equation 5
-in][]{Zhou_2014}. So uWUE is related to atmospheric CO$_2$
-concentration and compensation point, and is inversely proportional to
-the marginal water cost of leaf carbon.
-
-Given eddy-covariance FLUXNET2015 data (Section \ref{data}), every
-term in our new version of PM (Equation \ref{et}) is observed, except
-for uWUE, which we fit by calculating its expectation, given the model
-and FLUXNET2015 data.
-
-However, eddy-covariance data are inherently noisy so we include a
-measure of uncertainty in our analysis. To account for observational
-error, as well as model uncertainty (e.g. temporal and spatial
-variations of uWUE and $g_1$), we introduce an uncertainty parameter
-$\sigma$ modifying uWUE:
-
-% \begin{linenomath*}
-  \begin{equation}
-    ET = \frac{\Delta R_{net} + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R \; 1.6\; \sigma \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
-    \label{et_sigma}
-  \end{equation}
-% \end{linenomath*}
-
-Now, from each FLUXNET2015 observation (i.e. for each hourly
-observation at every time step) we can evaluate $\sigma$:
-
-% \begin{linenomath*}
-  \begin{equation}
-    \sigma = - \frac{g_a \gamma c_a \sqrt{VPD} L_v P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R_{net} - g_a \rho_a c_p VPD\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})},
-    \label{sigma}
-  \end{equation}
-% \end{linenomath*}
-
-Thus, with this uncertainty analysis we can evaluate departure from
-our theory in observations, as a departure of $\sigma$ from unity. The
-variability of $\sigma$ across sites and time provides a measure of
-uncertainty in our model, assumptions, as well as in the FLUXNET2015
-observations themselves. The variability of $\sigma$ then propagates
-through any uncertainty to our partial derivative of Equation
-\ref{et_sigma} with respect to VPD:
-
-% \begin{linenomath*}
-  \begin{equation}
-    \frac{\partial \;  ET}{\partial \; VPD} = \frac{2\; g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_a }{1.6 \; R\; \sigma \; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)
-    \label{d_et}
-  \end{equation}
-% \end{linenomath*}
-
-With Equation \ref{d_et} we have an analytical framework for ecosystem
-response to atmospheric demand perturbations with environmental
-conditions held fixed. There are a few subtleties to taking the
-derivative in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$)
-and $VPD$ are functionally related, so while taking the derivative we
-evaluate
-$\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET}
-{\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD}
-\Big|_{\text{RH fixed}} + \frac{\partial \; ET}{\partial \; RH}
-\frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$
-    fixed}}$. $RH$ and $e_s$ are assumed to be approximately
-independent, which is supported by the data (not shown).
-
-This derivation relied either implicitly or explicitly on several
-assumptions. First, we assume that VPD at the leaf surface is the same
-as VPD at measurement height; physically this implies that leaves are
-perfectly coupled to the atmosphere. In reality, for some conditions
-and plant types the leaves can become decoupled from the boundary
-layer \citep{De_2017, Medlyn_2017}. Given our focus on the growing
-season, which is usually characterized by relatively high insolation
-inducing instability and convective boundary layers, we would expect
-the surface to be generally well coupled. An additional assumption in
-the formulation of $uWUE$ \citep{Zhou_2014, Zhou_2015} and
-\citet{Medlyn_2017}'s stomatal conductance model is that direct soil
-evaporation (E) contributions to ET remain small relative to
-transpiration (T). This should be more true during the growing
-season. The ratio of E to T may increase immediately after rainfall
-events due to high soil moisture and ponding, but VPD is generally low
-anyways during these times. However, some plant types allow for
-systematically larger contributions of E in ET, particularly those
-with sparse canopies and smaller relative amounts of transpiration. We
-therefore might expect that the theory will be most applicable to
-forest PFTs, which will be most strongly coupled to the boundary layer
-due to larger surface roughness, and will also generally have the
-highest ratios of transpiration to evaporation. Finally, for the goal
-of developing PFT-wide intuition, we assume that $g_1$ and $uWUE$ are
-constant in space and time. Both quantities have been shown to be well
-conserved over space and time within a PFT relative to inter-PFT
-variability. However, we would expect them to vary somewhat with
-environmental conditions including very low soil water content,
-temperature, and with intra-PFT plant-specific characteristics like
-wood density for tree PFTs \citep{Lin_2015}. Given these strong
-assumptions made with the goal of understanding broad, leading order
-plant behaviors, we take the extra care of including and analyzing
-uncertainty (Section \ref{testing}) and examining sensitivity to the
-exponent of the VPD response (Section \ref{functional_form}).
-
-We note one final comment on our derivation which is relevant for
-drought indices. If we approximate $c_a$ at a global mean CO$_2$
-concentration, then the RHS of Equation \ref{et} is fully defined
-using commonly available weather station data and the constants
-published here. This makes Equation \ref{et} a useful alternative to
-PET in drought indices and hydrometeorological analysis for vegetated
-surfaces. Equation \ref{et} better reflects the physics of water
-exchange at the land surface and would only require fitting of uWUE
-and $g_1$, and can also account for changes in CO$_2$ concentration
-(assuming some relationship between $\lambda$ and [CO$_2$]), contrary
-to typical drought indices such as the Palmer Drought Severity Index
-(PDSI) \citep{Swann_2016, Lemordant_2016, Lemordant_2018}. Lastly, all
-code and data used in this analysis, including those used to generate
-the figures and tables, are publicly available at
-\url{https://github.com/massma/climate\_et}.
-%\hyperref[https://github.com/massma/climate\_et]{https://github.com/massma/climate_et}.
-
-\section{Results and Discussion}
-\label{results}
-
-By construction, the variability in the $\sigma$ term (Equation
-\ref{sigma}) contains all model and observational uncertainties. For
-an observation that perfectly matches our model and constant uWUE
-assumption, $\sigma$ will be one. Therefore, for our assumptions and
-framework to be reasonable $\sigma$ should be close to 1. An
-additional concern is that $\sigma$ may in fact be correlated with
-$VPD$, in which case the dependence would need to be accounted for
-when taking the derivative. Fortunately, there is a very weak
-dependence of $\sigma$ on VPD in their joint distribution, and
-$\sigma$ is indeed close to unity i.e. $O(1)$ (Figure
-\ref{joint_vpd_sigma}). Given this weak dependence and the
-distribution of $\sigma$ we have confidence in our model framework and
-the data quality.
-
-\begin{figure}[h]
-  \centering\includegraphics[width=0.75\textwidth]{./joint_vpd_sigma.pdf}
-  \caption{The joint distribution of $VPD$ and $\sigma$, with outliers
-    removed (defined as lowest and highest 5\% of $\sigma$). $\sigma$
-    exhibits a weak dependence on $VPD$, and $\sigma$ is $O(1)$ for
-    the bulk of the observations.}
-  \label{joint_vpd_sigma}
-\end{figure}
-
-Before calculating the sensitivity of ET to VPD, we will consider the
-functional form of Equation \ref{d_et}. There are two main terms: a
-``scaling'' term, which modifies the magnitude but not the sign of the
-ET response to VPD ($\frac{\partial \; ET}{\partial \; VPD}$):
-
-\begin{equation}
-  \frac{g_a \; P}{T(\Delta + \gamma)},
-\end{equation}
-
-and a ``sign'' term, which determines whether ET increases or
-decreases with VPD (i.e. atmospheric demand driven or physiologically
-controlled):
-
-\begin{equation}
-  \label{sign}
-  \frac{c_p}{R_{air}} - \frac{\gamma c_a }{1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right).
-\end{equation}
-All variables are positive, so the relative magnitude between the
-first term and the second term in the sign term (Equation \ref{sign})
-will determine whether ET increases or decreases with increasing
-VPD. If the second term is larger then plant control dominates and ET
-decreases with increasing VPD. However, if the first term is larger,
-then atmospheric demand dominates and ET increases with increasing
-VPD.
-
-\subsection{Functional Form of the Sign Term}
-\label{sign_term}
-First, we explore the variables within the sign term to gain better
-intuition on the driver of either the increase or reduction of ET with
-VPD. CO$_2$ concentration ($c_a$) and the psychometric constant
-($\gamma$) are relatively constant over the dataset considered here so
-that the variability is dominated by $\sigma$ and $VPD$. uWUE could
-vary with soil moisture but has been shown to be relatively constant
-\citep{Zhou_2015}. This then means that the sign term only depends on
-VPD for a given PFT and is approximately just a function of $VPD$. We
-can further determine a critical threshold separating an increase from
-a decrease in ET, i.e. the threshold $VPD_{crit}$ such that the
-derivative vanishes $\frac{\partial \; ET}{\partial \; VPD} = 0$:
-\small
-% \begin{linenomath*}
-  \begin{equation}
-    VPD_{crit} = \frac{R_{air}}{4 c_p} \left( \frac{\gamma c_a}{1.6\; R \;  uWUE} + \sqrt{\frac{\gamma c_a}{1.6\; R \;  uWUE}\left( \frac{\gamma c_a}{1.6\; R \;  uWUE} + 8 g_1 \frac{c_p}{R_{air}}\right)} - 4 g_1 \frac{c_p}{R_{air}} \right),
-    \label{vpd_min_et}
-  \end{equation}
-% \end{linenomath*}
-\normalsize noting that $VPD_{crit}$ mostly depends on the PFT
-parameters uWUE and $g_1$, and only varies weakly with climate as most
-other parameters related to the environment are nearly constant. The
-calculated value of $VPD_{crit}$ for each PFT is shown in Table
-\ref{vpd_crit}. For any values of $VPD$ less than $VPD_{crit}$, ET
-will decrease with increasing VPD
-($\frac{\partial \; ET}{\partial \; VPD} < 0$), and for values of
-$VPD$ greater than $VPD_{crit}$, ET will increase with increasing VPD
-($\frac{\partial \; ET}{\partial \; VPD} > 0$). In other words,
-ecosystems regulate and mitigate evaporative losses up to the VPD
-limit, $VPD_{crit}$, above which atmospheric demand is just too high
-to be entirely compensated by stomatal and ecosystem regulation. We
-note however that even though ET increases again above the critical
-threshold, $VPD_{crit}$, ET is still much lower than potential
-evaporation as stomata are still strongly regulating vapor fluxes to
-the atmosphere. Indeed, even in the absence of soil pore evaporation,
-stomata do not shut down entirely at very high VPD so that ET does not
-go to zero, because stomata are still slightly open to perform some
-photosynthesis \citep{Ball_1987, Leuning_1990, MEDLYN_2011}. In
-addition, upward xylem transport is necessary to maintain phloem
-transport, as well as nutrient transport and thus carbon allocation
-\citep{De_2013, Nikinmaa_2013, Ryan_2014}.
-
-\begin{table}
-  \caption{Values of $VPD_{crit}$, where
-    $\frac{\partial \; ET}{\partial \; VPD} = 0$, evaluated at PFT
-    average values for $R_{air}$, $\gamma$, and $c_a$. PFT-specific
-    constants ($g_1$, uWUE) are provided in Table \ref{pft}. For
-    values of $VPD$ less than $VPD_{crit}$,
-    $\frac{\partial \; ET}{\partial \; VPD}$ will be negative, and for
-    values of $VPD$ greater than $VPD_{crit}$,
-    $\frac{\partial \; ET}{\partial \; VPD}$ will be positive.}
-  \centering
-  \begin{tabular}{l c c c c c}
-    \hline
-    PFT & $R_{air}$ & $c_a$ (ppm) & $\gamma$  & \textbf{$VPD_{crit}$ (Pa)} \\
-    \hline
-    \input{vpd_crit.tex}
-    \hline
-  \end{tabular}
-  \label{vpd_crit}
-\end{table}
-
-Differences in $VPD_{crit}$ are exclusively determined by uWUE and the
-slope parameter ($g_1$) related to the plant functional type. A larger
-uWUE means a smaller $VPD_{crit}$, and an ET response to increases VPD
-that is more likely to be positive. At first glance this result is
-somewhat counter-intuitive; we expect that plants with a higher water
-use efficiency would be more water conservative. However, in reality
-uWUE determines how $WUE$ changes with VPD:
-
-\[WUE = \frac{GPP}{ET} = \frac{uWUE}{\sqrt{VPD}}.\]
-\[\frac{\partial \; WUE}{\partial \; VPD} = -\frac{uWUE}{2 \;
-    VPD^{3/2}}\]
-
-So, plants with a higher uWUE will have a greater \textit{decrease} in
-ecosystem-scale $WUE$ in response to increases in VPD. This decrease
-in $WUE$ causes more water loss per unit carbon gain, and explains the
-relationship between high uWUE and high likelihood of increases of ET
-in response to increasing atmospheric drying (increases in VPD).
-
-A tendency towards increasing ET response with increasing VPD can also
-be caused by a high slope parameter ($g_1$), characteristic of plants
-that at the leaf scale are more willing to trade water for access to
-atmospheric CO$_2$. Plants that are less conservative will be thus be
-more likely to increase ET with increasing VPD. Both the
-aforementioned effects (large uWUE, $g_1$) can amplify each other, and
-generally conspire to shift the sign term towards a positive value for
-a given PFT.
-
-This effect of uWUE and $g1$ on the sign term is most apparent by
-comparing two extreme PFTs: water intensive crops (CRO) and water
-conservative closed shrub (CSH). CRO has higher slope parameter and a
-slightly higher uWUE ($g_1 = 140.7$ Pa$^{1/2}$; $2.85$ $\mu$-mol [C]
-Pa$^{0.5}$ J$^{-1}$ [ET]) compared to CSH ($g1 = 75.1 \, Pa^{1/2}$,
-$uWUE=2.82$ $\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET]). These differences
-in PFT parameters cause opposite ET responses to changes in VPD
-between CRO and CSH. ET theoretically always decreases with increasing
-VPD for the more water conservative CSH, while ET frequently increases
-with increasing VPD for the more water intensive CRO (Figure
-\ref{idealized_sign}). CROs evolved or were bred to prioritize GPP and
-yield and are thus not water conservative. They are very willing to
-trade water for photosynthesis and productivity, despite changes in
-VPD, while CSH are very unwilling to trade water for more
-photosynthesis.
-
-\begin{figure}
-  \centering \includegraphics{./idealized_sign.pdf}
-  \caption{The functional form of the sign term, with $\sigma$ held
-    fixed at 1, and all terms except VPD set to PFT averages. For
-    comparison, the observed range of VPD for each PFT is plotted
-    below the x-axis. Stars denote 25th, 50th, and 75th percentiles,
-    and the range of the line spans the 5th-95th percentiles of
-    observed VPD. Vertical black lines denote the location of
-    $VPD_{crit}$ for each PFT, with the exception of CSH and WSA, for which
-    $VPD_{crit}$ is off-scale.}
-  \label{idealized_sign}
-\end{figure}
-
-As expected, the slope parameter ($g_1$) is a primary determinant of
-the VPD dependence for the sign term shown in Figure
-\ref{idealized_sign}. Plants that are more conservative (small $g_1$)
-will tend to reduce ET with increasing VPD, and will be very effective
-at reducing ET, especially at low VPD. However, at very high VPD,
-gradients in vapor pressure at the leaf scale will become very strong
-as stomata reach their limits of closure in response to VPD
-(parameterized with $g_1$). As a result, ET response will begin to
-asymptote towards a constant ecosystem-scale values as leaf-scale
-response to VPD asymptotes towards zero.  Therefore, plants with a low
-$g_1$ will have the largest VPD dependence of ET response because the
-difference in ET response at low VPD (leaf stomatal response
-dominates) and high VPD (VPD gradient dominates) is largest. This is
-apparent in the strong VPD dependence of CSH, which has the lowest
-slope parameter ($g_1=75.1$ Pa$^{1/2}$) (Figure \ref{idealized_sign}).
-
-To summarize our theoretical insights (Figure \ref{idealized_sign} and
-Table \ref{vpd_crit}), CROs are the least water conservative and have
-the strongest overall tendency to increase ET with increasing VPD,
-while CSH are the most water conservative and have the strongest
-tendency to decrease ET with increasing VPD, as well as the strongest
-VPD dependence of response. Figure \ref{idealized_sign} clearly shows,
-according to our theory, that for all PFTs except for crops there is
-frequent occurrence of a negative (plant dominating) ET response to
-increases in VPD. Therefore, plants are able in most atmospheric
-conditions to reduce ET in response to increased VPD and thus to
-reduce water loss. To better illustrate this, the ranges of observed
-environmental VPDs at the FLUXNET sites are plotted parallel to the
-x-axis. For CSH and WSA, VPD is always less than VPD$_{crit}$ (off scale) so
-that the plant response dominates in typical environmental conditions,
-emphasizing the water conservative strategy of those plants. For CRO
-on the other hand, VPD is higher than VPD$_{crit}$ for more than 50\%
-of observations, emphasizing that those plants operate with an
-aggressive water usage strategy, are water intensive and were actually
-engineered for photosynthesis rather than water saving. For DBF, EBF,
-MF, GRA, and SAV more than half of the observed VPD are less than
-VPD$_{crit}$, i.e. in conditions where plant response dominates. SAV
-has a more water conservative response than the forest, grass, and
-crop plan types, but still responds by increasing ET with increasing
-VPD for about a quarter of observations, due to the high aridity (VPD)
-of the SAV ecoclimate. It is also important to note that for all PFTs,
-even when atmospheric demand dominates, ET response to VPD is still
-far more negative than it would be for potential evaporation
-$\partial PET/\partial VPD$, i.e. atmospheric demand only, emphasizing
-that there is still a strong regulation of evaporative flux by stomata
-and though the plant xylem. The sign term in the PET case would just
-be a constant ($\frac{c_p}{R_{air}} \approx 3.5$), which is far larger
-than any part of the curves for any PFT. This highlights the
-deficiencies of PET's ability to capture land response to changes in
-atmospheric dryness (VPD). Plants are always regulating water exchange
-from the land surface, even when they reach the limits of they ability
-to do so.
-
-\subsection{Functional Form of the Scaling Term}
-\label{scale_term}
-While the above discussion of the sign of
-$\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our
-question of when ET response increases or decreases with VPD,
-understating the overall magnitude of the ET response is important to
-soil-plant-atmosphere water budgeting. So we now more closely examine
-the terms that affect how the sign term is scaled:
-
-\begin{equation}
-  \frac{g_a \; P}{T(\Delta + \gamma)}.
-\end{equation}
-
-$\frac{P}{T}$ is an air-density term, which varies little compared to
-aerodynamic conductance and Clausius-Clapeyron ($\Delta$). The
-psychometric constant ($\gamma$) is also relatively constant, so the
-scaling term should be primarily a function of aerodynamic conductance
-and temperature, through the Clausius-Clapeyron relationship
-$\Delta$. This is as expected, given that the aerodynamic conductance
-represents the efficiency of exchange between the surface and the
-atmosphere. As aerodynamic conductance increases, any plant response
-will be communicated more strongly to the atmosphere (and vice-versa).
-
-$\Delta$'s presence in the scaling term also matches physical
-intuition. $\Delta$ (and also the approximately constant $\gamma$)
-control the efficiencies with which surface energy is converted to
-latent and sensible heat \citep{Monteith_1965}. The functional from of
-$\Delta$ will be the same across PFTs, but the temperature range may
-vary slightly. In contrast, aerodynamic conductance will vary strongly
-with PFT due to the importance of surface roughness for aerodynamic
-conductance. So most of the differences in scaling between PFT should
-be in the aerodynamic conductance term.
-
-The control of the scaling term variability between PFTs by
-aerodynamic conductance is confirmed by data (Figure
-\ref{scale_vary}). Differences between PFT are almost entirely due to
-differences in aerodynamic conductance, rather than differences in
-observed temperature ranges. The scaling term for the tree PFTs (DBF,
-EBF, ENF, MF) is generally about double the scaling terms for other
-PFTs which have lower surface roughness and generally smaller
-aerodynamic conductance (GRA, CSH, CRO). The savannah (WSA, SAV) PFT's
-scaling is somewhere between GRA, CSH, and CRO, and DBF, EBF, ENF, and
-MF, due to higher variability and surface roughness.
-
-\begin{figure}
-  % \centering
-  \centerline{\includegraphics[width=1.2\textwidth]{./idealized_scale.pdf}}
-  \caption{Primary sources of variability for the scaling term, as a
-    function of PFT. The 5th-95th percentile range of temperature is
-    plotted at the 5th, 25th, 50th, 75th, and 95th percentiles of
-    aerodynamic conductance, as observed for each PFT.}
-  \label{scale_vary}
-\end{figure}
-
-Within each PFT, the scaling term variability is controlled both by
-environmental temperature and aerodynamic conductance variability
-(Figure \ref{scale_vary}). While the observed variability of the
-aerodynamic conductance contributes more to the scaling term
-variability than temperature, the temperature contribution is
-non-negligible. Specifically, the scaling term is generally larger at
-low temperatures when latent heat is relatively inefficient at moving
-energy away from the surface. This effect amplifies the role of
-aerodynamic conductance variability at low temperatures.
-
-To summarize, variability between PFTs is mostly controlled by
-systematic differences in aerodynamic conductance, due to differences
-in surface roughness between each PFT, and possibly to a lesser extent
-wind conditions. In contrast, variability within PFT is also
-controlled by temperature, through Clausius-Clapeyron. But,
-aerodynamic conductance variability generally impacts the scaling term
-more than temperature, even within PFTs.
-
-\subsection{Bulk statistics of ET response to VPD}
-\label{stats_sec}
-
-In this section we consider direct observations of ET response with
-eddy-covariance data, while including uncertainty with the $\sigma$
-term (Section \ref{methods}). These observational results of ET
-response (Table \ref{stats}) largely confirm our theoretical analysis, presented in the previous sections. For all PFTs, mean ET response to increasing VPD
-is negative. However, ET response evaluated at the average of all
-variables (e.g. $\sigma$, $T$, $c_a$, $VPD$) is positive for CRO, and
-negative for all other PFTs. This difference in mean ET response as
-compared to the ET response at mean environmental conditions is due to
-the non-linear nature of the response, in which negative responses are
-generally larger magnitude than positive responses (Figure
-\ref{idealized_sign}). Therefore, both the mean ET response as well as
-the ET response at mean environmental conditions matches our
-expectations from the theory (Section \ref{sign_term}), with the
-exception that CRO observations are shifted more towards a negative ET
-response than we expect.
-
-\begin{table}
-  \caption{Statistics of $\frac{\partial \; ET}{\partial \; VPD}$ as a
-    function of PFT.}
-  \centering
-  \begin{tabular}{l c c c}
-    \hline
-    PFT & $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ & $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)$ & fraction $\frac{\partial \; ET}{\partial \; VPD} < 0.$ \\
-    \hline
-    \input{stats.tex}
-    \hline
-  \end{tabular}
-  \label{stats}
-\end{table}
-
-Regarding the frequency of negative and positive ET response, all PFTs
-exhibit a decreasing ET response with increasing VPD (physiologically
-controlled, water conservative response) for the majority of
-observations. The more water conservative PFTs generally exhibit
-higher frequency of negative ET response, especially when one factors
-in the distribution of environmental VPD (e.g. SAV and WSA grow in
-more arid climates, Figure \ref{idealized_sign}). In general the bulk
-statistics match our theoretical expectations well, with the caveat
-that inclusion of uncertainty shifts crops towards a slightly more
-negative response to VPD, and shifts many of the other PFTs, which still
-exhibit a high frequency of negative ET response to VPD, towards more frequent
-occurrence of positive ET response than the theory and Figure
-\ref{idealized_sign} would suggest. The bulk statistics motivate a
-more thorough examination of the structure of uncertainty and more
-sophisticated validation of our theory's performance against
-observations.
-
-\subsection{Validation of theory at eddy-covariance sites}
-\label{testing}
-We now compare more sophisticated distributions of the observed
-response to our simplified theory (Section \ref{sign_term}). The
-observed distribution of the sign term, as compared to what the theory
-would predict, is provided in Figure \ref{test_sign}. Our goal was to
-capture the leading order behavior of the ET dependence on VPD. Given
-the assumptions we made, and the uncertainties of flux tower
-observations themselves, we expect a relatively large amount of noise
-when reproducing the derivatives of ET. However, the data largely
-reproduces our theoretical analysis.
-
-\begin{figure}
-  \centering
-  \centerline{\includegraphics[width=1.2\textwidth]{./test_sign.pdf}}
-  \caption{Comparison of the sign term with model uncertainty included
-    (box plots) to the sign term as calculated with simplifying
-    assumptions (blue line, as in Figure \ref{idealized_sign}). Each
-    box plot corresponds to 5\% of the data, and the 5-95\% range of
-    VPD is plotted.}
-  \label{test_sign}
-\end{figure}
-
-This is particularly true for DBF and MF; the theory matches the
-leading order behavior of the function when uncertainty is included,
-and the observations match the theory with the addition of noise. The
-VPD dependence, given by the slope parameter ($g_1$), follows the
-median values of each bin. Perhaps most importantly, the x-intercept,
-and thus VPD$_{crit}$, matches nearly exactly between the theory and
-the observations. Therefore the sign of the ET response to increases
-in VPD should be well matched, subject to the unavoidable constraints
-of noise, much of which comes from the observations themselves. The
-uncertainty is non negligible; there are many observations in each bin
-for which the the sign of observation is opposite the response
-predicted by the theory, but to leading order our theory matches the
-observations well.
-
-While CSH has a much different functional form of the sign term than
-DBF and MF, CSH observations also match our theory to leading order,
-albeit with a bit more variability as a function of VPD. Again, the
-VPD dependence mostly determined from the slope parameter ($g_1$)
-closely matches the medians in the observation bins. The
-VPD-independent, strongly negative response is also captured. For CSH,
-there is rare occurrence of observed positive ET response with VPD
-($\approx 6$\%), even with uncertainty, so the sign of the
-observations almost always matches the sign of the theory, which
-states that ET response should always be negative.
-
-Biases between the theory and observations are similar for ENF, WSA,
-CRO and SAV. At low VPD the theory and the observations match
-well. However, at high VPD (upper 10-20\% of observations) the theory
-is biased slightly towards positive response as compared to the
-observations (Figure \ref{test_sign}). For ENF this slight bias could
-be explained by a negative bias in $g_1$. However, for WSA, CRO, and
-SAV the observations in the highest VPD bins exhibit a downturn
-towards more negative response, which cannot be captured by the
-functional form of the sign term (Figure
-\ref{idealized_sign}). Therefore, to explain the observations one of
-the variables in the sign term must change at extreme VPD (upper
-10-20\%). The most likely candidate is $uWUE$, which we have assumed
-constant to meet our goal of developing intuition for leading order
-behavior, but might be expected to decrease at extremely low SWC. This extremely low SWC should be correlated with high VPD in very dry environments
-through sensible heating increases \citep{Gentine_2016}. The theory
-for EBF exhibits similar limitations as for ENF, WSA, and SAV, except
-a larger portion of observations are biased negative as compared to
-the theory ($\approx$ 35\%, Figure \ref{test_sign}). However, in
-general, and specifically for non-extreme VPD (VPD  $< \approx$70-90th
-percentiles), the theory matches the observations for the tree and
-savannah plant types well.
-
-The theory for GRA suffers from similar, but much more severe,
-limitations as for CRO, WSA, and SAV. GRA observations are
-characterized by a consistent trend back towards negative ET response
-at higher VPD, which the functional form of our theory is incapable of
-accounting for. As compared to CRO, WSA and SAV, the divergence
-between the theory and the observations is far greater for GRA,
-biasing 40-50\% of observations at higher VPD. In addition for the
-potential for soil moisture to alter $uWUE$, there are other sources
-of plant heterogeneity specific to GRA (and some extent CRO) that may
-alter $uWUE$ (or $g_1$) or invalidate other assumptions made in the
-Methods section (Section \ref{methods}). We do not account for
-variability in plant height and surface roughness, or differences in
-C3 vs C4 photosynthesis and water strategies, which we might expect to
-vary substantially across sites, years, and season for GRA. These
-deficiencies could largely explain the inability of our theory to
-exactly match the observations in croplands. For example, a
-superposition of sites with C3 plants (at low environmental VPD) and
-C4 plants (at high environmental VPD) would explain the observed shift
-back towards negative ET response at high VPD when all sites are
-binned together, as in Figure \ref{test_sign}. We hypothesize that the
-theory would validate against observations much better if these
-sources of variability were accounted for, at a cost of increased
-complexity and analytical opacity.
-
-While the above discussion shows that our theory has some limitations
-when applied to some PFTs, especially for grasslands, it does well for
-DBF, MF and CSH PFTs, and captures the response at non-extreme VPD
-(VPD $< \approx$ 70th-90th percentiles) for
-ENF, EBF, CRO, WSA, and SAV. In general, the leading order behavior
-observed in the data is captured by the theory. Departures between our
-theory and observations, specifically at high VPD, could be explained
-and conceptualized with shifts in $g_1$ and/or $uWUE$ due to
-site-specific plant type variability (e.g. more arid-adapted ecosystems
-at more arid sites), or temporal variability for some environmental
-conditions (e.g. decreases in $uWUE$ at extremely low soil
-moisture). To focus on general behavior and develop intuition for
-PFT-scale response, we have ignored these sources of variability in
-the present analysis. However, any site-specific or temporal plant
-functional variability that can be conceptualized with shifts to $g_1$
-and $uWUE$ can be analyzed with our framework. This opens a door to future analyses in which plant behavior in anomalous conditions
-can be explained and analyzed using Equation \ref{et}.
-
-% It is useful to also compare our theory, the observed uncertainty in
-% the ET response, and PET, as PET is often used as a proxy for ET in
-% hydrometeorological studies and drought indices. Again, as in our
-% discussion of the theoretical results, the PET sensitivity to VPD
-% will be constant at about 3.5. In Figure \ref{test_sign} the ET
-% response to shifts in dryness is always far smaller than 3.5 in the
-% observations, even with all uncertainty included. So the PET
-% response to atmospheric dryness is outside the range of all
-% observable responses. This casts doubt on PET's usefulness for
-% capturing the leading order behavior of water budget and dryness
-% response at the land surface, as is done in several drought
-% indices. However our theory (Equation \ref{et}) could be used with
-% published uWUE and $g1$ data, as an alternative to PET to estimating
-% land response and dryness, and would more accurately account for the
-% response to atmospheric drying.
-
-\subsection{Observed ET response to VPD}
-
-Most of the results presented so far focus on the sign term, so now we
-turn to observations of ET response with the scaling term included
-(Figure \ref{data_scatter}). Until now, in the interest of developing
-leading order intuition for ET response, and to be conservative in our
-acknowledgment of model and observational error, we've considered
-$\sigma$ variability to be a measure of uncertainty. An alternative
-viewpoint is that $\sigma \cdot uWUE$ represents spatial and temporal
-variability of uWUE, which may be expected within bounds \citep[see
-Table \ref{pft}, also ][]{Zhou_2015}. This is a less conservative
-view; some of the $\sigma \cdot uWUE$ variability will be due to model
-and observational error, so by viewing $\sigma \cdot uWUE$ as
-\textit{real} variability we run the risk of mistaking noise for
-signal. However, the advantage of this viewpoint is that, from
-\citet{Zhou_2014}, we have very high confidence that uWUE fit at the
-hourly timescale (as we do with $\sigma \cdot uWUE$) correctly
-captures the relationship between ET, GPP, and VPD, and that our form
-of PM introduces minimal error with its use of uWUE.
-
-\begin{figure}
-  \centering
-  \centerline{\includegraphics[width=1.2\textwidth]{./data_scatter.png}}
-  \caption{Scatter plots of observed
-    $\frac{\partial \; ET}{\partial \; VPD}$, including $\sigma$
-    variability, as a function of PFT, temperature and VPD. Please
-    note differences in the colorbar scale.}
-  \label{data_scatter}
-\end{figure}
-
-Therefore, with the caveat that some of the signal presented in Figure
-\ref{data_scatter} may in fact be noise, we can interpret the observed
-distribution of ET response to VPD. In general, the observed response
-matches the intuitive theory. ET response to VPD shifts towards
-positive values as VPD increases (atmospheric demand dominating). CRO
-exhibit the highest occurrence of positive ET response, and the
-observations confirm that CRO are the most water intensive. CSH are
-the most water conservative, with a strong negative response. DBF,
-EBF, ENF, MF, SAV and WSA are also water conservative, but show some
-occurrence of positive ET response to VPD, particularly at higher
-observed VPD, as the theory predicts. GRA, while generally water
-conservative, does not match the theory well, with increasing frequency of
-\textit{negative} response at high VPD. This is as expected, given
-previous discussion on how we might expect more inter-site and inter-year variability for GRA (Section \ref{testing}).
-
-Figure \ref{data_scatter} also includes the impact of the scaling
-term. For a given VPD, the magnitude of ET response does not vary
-strongly with temperature, confirming that any impact of the scaling
-term on the magnitude of the response is primarily due to changes in
-aerodynamic conductance. Intuitively this is reasonable; aerodynamic
-conductance will control how dominant balances at the land surface are
-communicated to the boundary layer. While $\Delta$ controls the
-efficiency of energy conversion to latent heat, it appears this is a
-second order term, relative to $g_a$, for scaling ET response.
-
-As with Figure \ref{test_sign}, Figure \ref{data_scatter} matches our
-expectation based on simplified theory. The sign term is most strongly
-scaled by $g_a$, and in general the occurrence of positive ET response
-increases as VPD increases. The willingness with which a given plant type
-evolved to use water dictates the occurrence of positive versus
-negative ET response. Water conservative ecosystems are highly effective
-at mitigating the effects of atmospheric demand, and can store water
-for later use by reducing ET in response to increasing VPD.
-
-\subsection{Limitations of theory - very dry soil moisture conditions}
-\label{swc_section}
-In formulating our theory with Penman Monteith, we implicitly did not
-account for very dry soil moisture conditions. For the majority of
-environmental conditions observed at the eddy covariance sites used
-here, soil conditions were not extremely dry so that we could assume a
-constant uWUE and $g_1$. We posit that ecosystems will generally
-optimize to host plants living in conditions which they evolved
-for. However, in extreme conditions and drought scenarios soil water
-content (SWC) could become the limiting factor for ET response to VPD,
-which our theory does not account for. In addition, low soil moisture
-conditions themselves increase VPD through land-atmosphere feedback
-\citep[][]{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006,
-  Salvucci_2013, Gentine_2016, Berg_2016}.
-
-Within our framework, any systematic bias due to the
-failure to account for SWC's effects in very dry conditions should
-manifest itself in a functional relationship between $\sigma$, our
-uncertainty measure, and SWC, which is observed at the FLUXNET
-sites. Examining the relationship between $\sigma$ and SWC will test
-to what extent our theory breaks down in very dry soil moisture
-conditions.
-
-If we again view $\sigma \cdot uWUE$ as, in addition to a measure of
-uncertainty, a short time-scale observation of uWUE, we would expect
-$\sigma \cdot uWUE$ to decrease at low SWC. If $\sigma \cdot uWUE$ is
-a very strong function of SWC, then our theory should be conditioned
-more strongly on well watered soil conditions. If $\sigma \cdot uWUE$
-is weakly a function of SWC, then our theory is more universal and
-independent of soil moisture conditions.
-
-Indeed, for all PFTs, there is some slight dependence of
-$\sigma \cdot uWUE$ on SWC, especially at low SWC (Figure
-\ref{swc_boxplot}). The portion of observations for which our theory
-is biased by SWC-limitations varies by PFT, due to the nonlinear
-threshold at which soil moisture availability limits plant
-function. For CRO, MF and DBF, soil water content only matters for
-about the lowest 5\% of observations (each box is 5\% of observations
-in Figure \ref{swc_boxplot}). And, CRO, relative to DBF, has a very
-weak dependence on soil moisture, which is reflective of the high
-likelihood that CRO sites are optimally irrigated and not water
-stressed, suggesting that observed departures from the theory for CRO
-are due to other factors (i.e. different photosynthesis pathways; C3/C4) than soil
-moisture. For ENF, EBF, SAV, and WSA, systematic SWC-induced biases in
-$\sigma \cdot uWUE$ emerge in about the lowest 20-30\% of SWC
-conditions, although large variability in the SWC-$(\sigma \cdot uWUE)$
-relationship hampers interpretation for WSA. CSH presents a special
-case: the limited number of sites (2) preclude comment on the
-exact PFT-wide relationship.
-
-In contrast to ENF, CRO, and DBF, for GRA the relationship between
-$\sigma \cdot uWUE$ and SWC is more linear and affects a greater
-portion of observations; about 60\% of observations. Clearly, for GRA
-soil water frequently impacts plant function and alters ET response,
-and our theory is limited for the majority of environmental
-conditions. It is therefore not surprising that our theory tested
-poorly against the data for GRA, relative to to the other PFTs
-(Section \ref{testing}). For all other PFTs occurrence of soil
-moisture impacts was rare enough to not manifest itself in bulk
-statistics and figures.
-
-\begin{figure}
-  \centering
-  \centerline{\includegraphics[width=1.2\textwidth]{./swc_boxplot.pdf}}
-  \caption{$\sigma \cdot uWUE$ as a function of PFT and SWC. Each box
-    plot represents 5\% of all observations. Note that the highest
-    10\% of SWC observations are excluded to better resolve variability in the much more narrow bins at lower SWC (e.g. SWC has long tails at high values).}
-  \label{swc_boxplot}
-\end{figure}
-
-The observed dependence of $\sigma \cdot uWUE$ on SWC for GRA would
-explain the deficiencies of our theory compared to the observations in
-Figure \ref{idealized_sign}, specifically the trend back towards
-negative soil moisture at high VPD. Aforementioned feedbacks between
-the land surface and the atmosphere, which are not accounted for due
-to our focus on the one-way response of the land surface to
-atmospheric conditions, would cause high VPD to be correlated with low
-SWC \citep[][]{Gentine_2016, Berg_2016}. So, at high VPD observations
-of low SWC are more likely, and this low SWC causes a lower uWUE
-(Figure \ref{swc_boxplot}). The lower uWUE at low SWC/high VPD then
-leads to the observed downturn towards decreasing ET with increasing
-VPD, and the deviation between our theory (based on a constant uWUE
-assumption) and observations. It is also perhaps not coincidental that
-the portion of $\sigma \cdot uWUE$ affected by low soil moisture
-observations is similar to the portion of observations that do not
-match our theory at high VPD. Indeed, this may suggest that for all PFTs
-except for GRA, coupling and feedbacks between SWC, VPD and plant
-function are relatively rare.
-Future research will explore these relatively rare feedbacks in extreme conditions, which due to analytical intractability will require more opaque numerical analysis of many more complex processes, including boundary layer growth and state and their relationship to surface layer coupling and free-tropospheric lapse rates and humidity.
-
-To summarize, our theory is limited by its inability to account for
-soil water impacts on land surface response, and feedbacks between SWC
-and VPD. Fortunately, for most PFTs SWC's effect on ET is relatively
-rare ($<$30\% of observations) and does not manifest itself in the
-majority of observations and bulk statistics. However for GRA, SWC decreases water use efficiency for the majority of
-the observations. Soil moisture effects explain the deficiencies of
-our theory in Section \ref{testing}, particularly for GRA. By
-conceptualizing SWC effects as a change in $uWUE$ (and/or $g_1$), it
-will be possible for future analysis to explore the importance of soil
-moisture on plant response to VPD, and feedbacks between plant
-function, SWC, and VPD.
-
-\subsection{Functional form of ET dependence on VPD and its relation
-  to VPD dependence}
-\label{functional_form}
-
-The theory described in Section \ref{sign_term} indicates that for a
-given $uWUE$ and $g_1$, the ET dependence on VPD should be concave
-upward, which is confirmed by eddy covariance data across most PFTs. In other words, there should be some local minimum in ET at a
-critical VPD$_{crit}$, assuming the scaling and plant terms  (e.g. aerodynamic
-conductance, $\Delta$, $g_1$ and $uWUE$) are held fixed. This result warrants further
-investigation, because to our knowledge no one has derived the
-theoretical ecosystem-scale relationship between ET and VPD while
-controlling for other environmental conditions. In particular, from
-personal communication, there is an apparent lack of consensus over
-whether the shape of the ET-VPD curve should be concave upward (our
-result) or concave downward in the absence of dramatic water stress. Given that understanding the ET-VPD
-relationship of the one-way plant response is fundamental to
-hypothesizing about any feedbacks between the land surface and the
-atmosphere, we analyze why our derived ET-VPD relationship is concave
-upward, particularly with respect to the exponent of VPD dependence in
-$uWUE$ and the Medlyn unified stomatal conductance model.
-
-There is a theoretical basis for the square root VPD dependence in
-both the stomatal conductance model and $uWUE$ based on the assumption
-that stomata behave to maximize carbon gain while minimizing water
-loss, which observations also generally support \citep{Lloyd_1991,
-  MEDLYN_2011, Lin_2015, Zhou_2014, Zhou_2015, Medlyn_2017}. However,
-some purely empirical results that fit the exponent of the VPD
-dependence to data have shown that it may vary slightly from 1/2,
-suggesting that stomata, as well as ecosystem-scale quantities based
-on stomata theory, may not always function optimally \citep{Zhou_2015,
-  Lin_2018}. Specifically with regards to $uWUE$, one would not expect
-that this ecosystem scale WUE quantity will respond to VPD exactly
-analogously to stomata. Direct soil evaporation's contributions to ET
-should shift the exponent of the VPD dependence, especially at
-conditions of low GPP when we would expect a systematically larger
-portion of direct soil evaporation contributions to ET, because we
-would also expect lower amounts of transpiration at low GPP. \citet{Zhou_2015}'s
-results corroborate this: they found a mean empirically fit
-exponential VPD dependence of 0.55, varying slightly from the
-theoretically optimal value of of 0.5 for AmeriFlux
-sites. Interpreting \citet{Lin_2018}'s results, which also show
-variance in the empirical exponent of the VPD dependence of the
-stomatal conductance model, is more difficult as \citet{Lin_2018} do
-not handle GPP/A dependence of stomatal conductance in a directly
-analogous manner to the optimal theory in \citet{MEDLYN_2011} and
-\citet{Medlyn_2017}. Regardless, given that these recent results on
-the relationship between VPD, GPP, and ET \citep{MEDLYN_2011,
-  Zhou_2014, Zhou_2015, Medlyn_2017} form the backbone of our analysis
-and are what allowed us to derive an explicit ET expression for the
-first time (Equation \ref{et}), we will analyze if and how assumptions
-about the exponent of the VPD dependence impacts the shape of the
-ET-VPD dependence.
-This analysis is also important to understand whether the choice of stomatal conductance model alters the fundamental behavior of the ET-VPD relationships, as many commonly used models  utilize a VPD exponent other than the 1/2 suggested by optimal theory \citep[e.g.
-][ which uses an exponent of 1]{Leuning_1990}.
-
-By introducing $n$ and $m$ we can free our stomatal
-conductance model from assumptions about VPD dependence:
-% \begin{linenomath*}
-  \begin{equation}
-    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_*}{VPD^m}\right) \frac{*WUE \; ET}{c_a \; VPD^n},
-    \label{m_n}
-  \end{equation}
-% \end{linenomath*}
-where:
-\[*WUE = \frac{GPP}{ET}VPD^n,\] and $g_*$ is a generic slope parameter
-of units $VPD^m$. To determine how the exponent $n$ and $m$ alter the
-shape of the ET-VPD dependence we find the roots of the second
-derivative of ET, using Equation \ref{m_n} for stomatal conductance
-($g_s$), with respect to VPD: \input{d2_solutions.tex} With this
-result we have defined the family of curves separating concave up from
-concave down ET solutions (Figure \ref{concave}). These curves are
-only functions of the exponent of the VPD dependence and a quantity we
-call non-dimensional VPD ($VPD^m/g*$). Several important
-relations reveal themselves from Equation \ref{curves}:
-
-\begin{figure}
-  \centering
-  \centerline{\includegraphics[width=0.75\textwidth]{./concave.pdf}}
-  \caption{ Solutions corresponding to inflection points between
-    concave up and concave down ET-VPD curves (Equation \ref{curves})
-    for three specific scenarios. Solutions are defined in terms of a
-    non-dimensional VPD ($VPD^m/g_*$), but to aide physical
-    interpretation the horizontal axis is additionally provided in terms of dimensionalized VPD assuming $m=1/2$ and
-    $g_*=110\; Pa^{1/2}$ (average of all PFT $g_1$). The vertical axis
-    has a different interpretation depending on the solution
-    curve. For the blue line ($m$ varying), it corresponds to $m$,
-    for the orange line ($n$ varying) it corresponds to $n$, and for
-    the green line it corresponds to the value of both $n$ and $m$
-    ($n=m$). Regions of the parameter space that correspond to
-    concave up and concave down results are labeled.}
-  \label{concave}
-\end{figure}
-
-\begin{itemize}
-  \item For optimal behavior (n, m = 1/2) the ET-VPD curve will be
-    concave up regardless of the magnitude of the plant constants
-    $g_1$ and $uWUE$. Therefore, the general concave up nature of our
-    results, given an assumption of optimal behavior, is insensitive
-    to plant type.
-  \item For all physically possible exponents of VPD dependence ($n,
-    m$), whether the solution is concave up or concave down does not
-    depend on $uWUE$.
-  \item In general, increasing the exponent of VPD dependence
-    increases the likelihood of a concave up result. Additionally, as
-    the exponent of VPD dependence increases from the optimum value of
-    1/2, whether the curve is concave upward or concave downward
-    becomes a function of the plant specific slope parameter $g_*$,
-    through non-dimensional VPD ($VPD^m/g_*$). Because the exponent of
-    the VPD dependencies is capable of altering the fundamental shape
-    of ET-VPD dependence, future research investment in understanding
-    the exact VPD dependence of stomatal conductance, and further
-    reconciliation of empirical and theoretical stomatal and
-    ecosystem behavior should be prioritized.
-\end{itemize}
-While it is possible that in the future some other form of
-VPD dependence is derived, at present \cite{MEDLYN_2011} and
-\cite{Zhou_2014} firmly established n=m=1/2 as the most likely
-candidate given current theory and empirical data. Additionally, we
-argue that a concave up result matches physical intuition more than a
-concave down result. Plants must maintain nutrient and sugar transport
-through the phloem and xylem. To accomplish this, stomata must remain
-slightly open \citep{De_2013, Nikinmaa_2013, Ryan_2014}. Furthermore,
-even if complete stomatal closure were possible, cuticular water loss
-and [at the ecosystem-scale] direct soil evaporation are still sources
-of ET which increase with VPD, independent of stomatal
-closure. Therefore, in the limit as VPD becomes large and we assume
-plants are exercising all strategies to reduce ET, any further increase
-in VPD should result in an increase in ET through cuticular water loss
-and/or direct soil evaporation. This inevitable transition from
-conditions when stomata respond strongly to VPD to conditions when
-stomata response is asymptoting towards full closure would cause a
-concave up ET-VPD curve, which is matched by the theory. In short,
-plant response becomes more limited as VPD increases, while
-atmospheric demand monotonically increases with VPD, leading to the
- result that atmospheric demand dominates plant response when
-atmospheric demand is high.
-
-This analysis allows us to understand the theoretical shape of the ET
-response to VPD with environmental conditions held
-fixed. Accomplishing this with purely statistical
-methods applied to flux observations would be very difficult, given
-the relatively fast time scale of plant response and the
-non-stationarity of [solar forced] environmental conditions over the
-relatively coarse (half hourly) flux estimates (which is required to
-obtain robust eddy-covariance statistics). Our results on the shape of
-the ET-VPD curve with environmental conditions held
-fixed can be built upon with future work examining how changes in VPD
-and environmental conditions (e.g. soil water storage) feedback upon
-one another. In the soil water storage example, over very long time
-scales extremely high VPD perturbations coupled with no precipitation
-could result in decreases in soil water storage such that water
-becomes limiting. This could be represented by an extension of our
-framework in which uWUE is allowed to decrease with decreasing SWC, as
-observed in Section \ref{swc_section}. Here, we focus our results by
-assuming constant PFT-wide conditions to build baseline intuition for
-ET-VPD dependence. For most PFTs, the theory with plant function held
-fixed matches the leading order behavior of the observations where
-plant function varies (Section \ref{testing}).
+\input{body.tex}
 
@@
 
-We derived a new form of Penman Monteith using the concept of
-semi-empirical optimal stomatal regulation \citep{Lin_2015,
-  MEDLYN_2011} and near constant uWUE \citep{Zhou_2015} to remove the
-implicit dependence of stomatal conductance on GPP and ET. With our
-new form of Penman Monteith we developed a theory for when an
-ecosystem will tend to reduce or increase ET with increasing VPD,
-which we evaluated against a range of eddy-covariance data spanning
-different climates and plant functional types. The goal was to capture
-the leading order behavior of the system to gain some intrinsic
-knowledge for its behavior. This intuition can be used to disentangle
-land atmosphere feedbacks in more complicated scenarios, and will aid
-interpretation of observations and sophisticated models.
+\input{conclusions.tex}
 
-Our theory suggests that for a majority of environmental conditions,
-plants will tend to conserve water and reduce ET with increasing
-VPD. Stomatal regulation and plant physiological response strongly
-regulate ET, and this regulation varies by PFT. CROs are the least
-water conservative, while DBF, EBF, SAV, GRA, MF, WSA, ENF and CSH are
-progressively more water conservative (more likely to reduce ET in
-response to increasing VPD). SAV and WSA exhibit positive ET response
-to VPD not necessarily because of poor water conservation strategies
-relative to other PFTs, but because of greater occurrence of high
-atmospheric demand (VPD) relative to other PFTs. Observations of ET
-response to VPD exhibit the same general behavior as the theory, with
-ET response becoming more positive (atmospheric demand dominating) as
-environmental VPD increases within a PFT, and more negative for PFTs
-that are adapted to arid conditions and prioritize water conservation
-over primary production.
-
-The consequences of our theory our twofold. First, the ET response to
-VPD is always far more negative than it would be if only atmospheric
-demand is considered and plant physiological response is ignored (as
-with PET). Therefore, any drought indices or analysis using PET will
-not capture how vegetated surfaces respond and react to increases in
-atmospheric demand. However, the new uWUE-version of Penman Monteith
-we derived (Equation \ref{et}) could be used as a complement to PET in
-drought indices. One could calculate and compare indices when plant
-response is ignored (PET) or included (Equation \ref{et}).
-
-Lastly, our paper builds important intuition for how plants will
-respond to VPD perturbations. We show that given optimal stomatal
-function and fixed environmental conditions, the ET-VPD dependence is
-theoretically concave upward, with ET increasing with increasing VPD as
-VPD increases past some critical value (Table \ref{vpd_crit}). However
-future research should focus on fully understanding the form of
-stomatal VPD dependence, as this result is sensitive to the exponent of
-VPD dependence, which we currently believe is 1/2 \citep{MEDLYN_2011,
-  Zhou_2014}. Our results are applicable to understanding the impact
-of  expected increases in VPD
-induced by global change.  Plant physiological responses to direct
-CO$_2$ effects \citep[e.g.,][]{Swann_2016, Lemordant_2018} receives more attention
-than physiological response to indirect effects like increased
-VPD. Here, we provide broad PFT-focused results showing a likely
-decrease in ET in response to positive VPD perturbations (atmospheric
-drying), which is consistent with recent observational analysis
-\citep[e.g.,][]{Rigden_2017}. Feedbacks between the land and the
-atmosphere may alter the net response to a long-timescale global VPD perturbation,
-but our focus on the one way plant response to a VPD perturbation in
-the atmospheric boundary layer is an important first step to
-disentangling such feedbacks, both in observations and model
-simulations of the present and future. By removing Penman Monteith's
-dependence on implicit relationships between GPP, VPD, and ET, we
-allow for explicit, accurate future analysis of plant-VPD feedbacks
-in the atmospheric boundary layer (Equation \ref{et}). Our approach
-can be extended to examine varying plant response to more nuanced
-consideration of plant type and climate. Any plant physiological
-heterogeneity or feedback that can be conceptualized with shifts in $g_1$
-\citep[e.g.][]{Lin_2015, Medlyn_2017} and/or uWUE
-\citep[e.g.][]{Zhou_2014} are representable within our framework, which
-opens the door for a hierarchy of more sophisticated climate- and
-plant-specific analysis of ET sensitivity to environmental variables
-(including VPD). We argue that such simplified conceptual frameworks
-are critical tools for disentangling land-atmosphere feedbacks at
-various scales, from diurnal to seasonal and beyond, and to
-characterize ET response in a warmer, atmospherically drier, and
-enriched CO$_2$ world.
-
-\acknowledgments This work used eddy covariance data acquired and
-shared by the FLUXNET community, including these networks: AmeriFlux,
-AfriFlux, AsiaFlknux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont,
+\acknowledgments This work used eddy covariance data acquired and shared
+by the FLUXNET community, including these networks: AmeriFlux, AfriFlux,
+AsiaFlux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont,
 ChinaFlux, Fluxnet-Canada, GreenGrass, ICOS, KoFlux, LBA, NECC,
 OzFlux-TERN, TCOS-Siberia, and USCCC. The ERA-Interim reanalysis data
-are provided by ECMWF and processed by LSCE. The FLUXNET eddy
-covariance data processing and harmonization was carried out by the
-European Fluxes Database Cluster, AmeriFlux Management Project, and
-Fluxdata project of FLUXNET, with the support of CDIAC and ICOS
-Ecosystem Thematic Center, and the OzFlux, ChinaFlux and AsiaFlux
-offices. This material is based upon work supported by the National
-Science Foundation Graduate Research Fellowship under Grant No. DGE
-16-44869. All code and data used in this analysis, including those
-used to generate the figures and tables, are publicly available at
-\url{https://github.com/massma/climate\_et}.
-%\hyperref[https://github.com/massma/climate\_et]{https://github.com/massma/climate_et}.
+are provided by ECMWF and processed by LSCE. The FLUXNET eddy covariance
+data processing and harmonization was carried out by the European Fluxes
+Database Cluster, AmeriFlux Management Project, and Fluxdata project of
+FLUXNET, with the support of CDIAC and ICOS Ecosystem Thematic Center,
+and the OzFlux, ChinaFlux and AsiaFlux offices. We would like to thank
+Dr. Trevor Keenan for providing tools for citation of the FLUXNET2015
+dataset, available at
+\url{https://github.com/trevorkeenan/FLUXNET_citations}. This material is
+based upon work supported by the National Science Foundation Graduate
+Research Fellowship under Grant No. DGE 16-44869. All code and data used
+in this analysis, including those used to generate the figures and
+tables, are publicly available at \url{https://github.com/massma/climate\_et}
+
+\begin{appendices}
+  \section{FLUXNET2015 sites}\label{flux_sites}
+  % below taken adapted from Trevor Keenan's FLUXNET citations
+  % at github.com/trevorkeenan/FLUXNET_citations
+  Table A1: Metadata and citations for flux sites used in this analysis.
+All data are gathered from \url{www.fluxdata.org}, and citations are aggregated using tools available at \url{https://github.com/trevorkeenan/FLUXNET_citations}.
+  \begin{longtable}{l l l l l l l}
+    \hline
+    \textbf{Site} &
+    \textbf{PFT} &
+    \textbf{Lat} &
+    \textbf{Lon} &
+    \textbf{Clim\textsuperscript{1}} &
+    \textbf{Period} &
+    \textbf{References} \\
+    [0.5ex]
+    \hline
+    \input{flux_sites.tex}
+    [1ex]
+    \hline
+  \end{longtable}
+  \textsuperscript{1}  K{\"o}ppen Climate classification.
+\end{appendices}
 

### 2018-10-31 doc/paper/vpd_et_paper_hess.tex
the tex file might be important, gotta be a bettter way to test reproducibility that this

--- 2018-10-30 doc/paper/ms.tex
@@
-% this is used for arkiv submissions
-\documentclass[12pt]{article}
-\input{def.tex}
+\documentclass[hess, manuscript]{copernicus}
+
+\usepackage{makecell} \usepackage{multirow} %AM for tables
+\usepackage{hyperref}
+\usepackage[title]{appendix}
+\usepackage{booktabs}
+\usepackage{longtable}
+
+\newcommand{\appropto}{\mathrel{\vcenter{
+      \offinterlineskip\halign{\hfil$##$\cr
+        \propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}
 
@@
   evapotranspiration?}
+\Author[1]{Adam}{Massmann}
+\Author[1]{Pierre}{Gentine}
+\Author[1,2]{Changjie}{Lin}
 
-\author[1]{Adam Massmann\thanks{Corresponding author: akm2203@columbia.edu}}
-\author[1]{Pierre Gentine}
-\author[1,2]{Changjie Lin}
-
-\affil[1]{Department of Earth and Environmental Engineering, Columbia University, New York, NY 10027}
+\affil[1]{Department of Earth and Environmental Engineering,
+  Columbia University, New York, NY 10027}
 \affil[2]{State Key Laboratory of Hydroscience and Engineering, Department of Hydraulic
@@
 
+\runningtitle{When does VPD drive or reduce ET?}
+
+\runningauthor{A. Massmann}
+
+\correspondence{Adam Massmann (akm2203@columbia.edu)}
+
+\firstpage{1}
+
 \maketitle
+
 \begin{abstract}
-  \input{abstract}
+\input{abstract.tex}
 \end{abstract}
 
-\onehalfspacing
-
-\section{Introduction}
-
+\introduction
 \input{intro.tex}
@@
 
-\section{Conclusions}
-
+\conclusions
 \input{conclusions.tex}
 
-\acknowledgments This work used eddy covariance data acquired and shared
-by the FLUXNET community, including these networks: AmeriFlux, AfriFlux,
-AsiaFlux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont,
-ChinaFlux, Fluxnet-Canada, GreenGrass, ICOS, KoFlux, LBA, NECC,
-OzFlux-TERN, TCOS-Siberia, and USCCC. The ERA-Interim reanalysis data
-are provided by ECMWF and processed by LSCE. The FLUXNET eddy covariance
-data processing and harmonization was carried out by the European Fluxes
-Database Cluster, AmeriFlux Management Project, and Fluxdata project of
-FLUXNET, with the support of CDIAC and ICOS Ecosystem Thematic Center,
-and the OzFlux, ChinaFlux and AsiaFlux offices. We would like to thank
-Dr. Trevor Keenan for providing tools for citation of the FLUXNET2015
-dataset, available at
-\url{https://github.com/trevorkeenan/FLUXNET_citations}. This material is
-based upon work supported by the National Science Foundation Graduate
-Research Fellowship under Grant No. DGE 16-44869. All code and data used
+\codedataavailability{All code and data used
 in this analysis, including those used to generate the figures and
-tables, are publicly available at \url{https://github.com/massma/climate\_et}
+tables, are publicly available at
+\url{https://github.com/massma/climate\_et}}
 
-\begin{appendices}
+\appendix
+
 \section{FLUXNET2015 sites}\label{flux_sites}
+
+\appendixtables   %% needs to be added in front of appendix tables
+
 % below taken adapted from Trevor Keenan's FLUXNET citations
@@
 \textsuperscript{1}  K{\"o}ppen Climate classification.
-\end{appendices}
 
+\authorcontribution{AM: formal analysis; methodology; software;
+  validation; visualization; writing - original draft. PG:
+  conceptualization; methodology; supervision; writing - review \&
+  editing. CL: data curation; writing - review \& editing.}
+
+\competinginterests{We declare no competing interests.}
+
+\begin{acknowledgements}
+This work used eddy covariance data acquired and shared
+by the FLUXNET community, including these networks: AmeriFlux, AfriFlux,
+AsiaFlux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont,
+ChinaFlux, Fluxnet-Canada, GreenGrass, ICOS, KoFlux, LBA, NECC,
+OzFlux-TERN, TCOS-Siberia, and USCCC. The ERA-Interim reanalysis data
+are provided by ECMWF and processed by LSCE. The FLUXNET eddy covariance
+data processing and harmonization was carried out by the European Fluxes
+Database Cluster, AmeriFlux Management Project, and Fluxdata project of
+FLUXNET, with the support of CDIAC and ICOS Ecosystem Thematic Center,
+and the OzFlux, ChinaFlux and AsiaFlux offices. We would like to thank
+Dr. Trevor Keenan for providing tools for citation of the FLUXNET2015
+dataset, available at
+\url{https://github.com/trevorkeenan/FLUXNET_citations}. This material is
+based upon work supported by the National Science Foundation Graduate
+Research Fellowship under Grant No. DGE 16-44869.
+%\hyperref[https://github.com/massma/climate\_et]{https://github.com/massma/climate_et}.
+\end{acknowledgements}
+
+\bibliographystyle{copernicus}
 \bibliography{references.bib}

### 2018-10-31 doc/paper/body.tex
about to spell check

@@
 including all Tier 1 sites with at least four years of data and
-observations of the variables described in the methods section (Section
+observations of the variables described in the methods section (Sect.
 \ref{methods}). Sixty-six sites met these requirements, and were grouped
@@
 forest (ENF), mixed forest (MF), closed shrub (CSH), savannah (SAV), and
-woody savannah (WSA) (site locations and citations are in Figure
+woody savannah (WSA) (site locations and citations are in Fig.
 \ref{map_fig} and Appendex \ref{flux_sites}, respectively).
@@
 is an ecosystem scale analogue to $g_{1-leaf}$. We solve for $g_1$
-following \citet{Medlyn_2017} (Equation 5), and take the median $g_1$
+following \citet{Medlyn_2017} (Eq. (5)), and take the median $g_1$
 value to be representative of each PFT (Table \ref{pft}) instead of the mean to avoid extra weighting of rare outliers.
@@
 
-While Equation 3 can be used in PM (Equation 1), it will make
+While Eq. (3) can be used in PM (Eq. (1)), it will make
 analytical work with the function intractable because $GPP$ is
@@
 cause an indirect change in stomatal conductance, in addition to the
-direct effect of VPD in Equation \ref{medlyn}. Therefore, in order to
+direct effect of VPD in Eq. (\ref{medlyn}). Therefore, in order to
 derive the response of ET to VPD, we must account for the functional
@@
 leverages the same assumptions used in \cite{MEDLYN_2011} to derive the
-square-root VPD dependence of the stomatal conductance model (Equation
-\ref{medlyn}).  We can use uWUE to remove the $GPP$ dependence from
+square-root VPD dependence of the stomatal conductance model (Eq.
+(\ref{medlyn})).  We can use uWUE to remove the $GPP$ dependence from
 $g_s$ in a way that makes PM analytically tractable:
@@
 
-Plugging Equation \ref{new_g_s} into Equation \ref{orig_pen} and
+Plugging Eq. (\ref{new_g_s}) into Eq. (\ref{orig_pen}) and
 rearranging gives a new explicit expression for PM, in which
@@
 By accounting for photosynthesis changes in ecosystem conductance, with
-Equation \ref{et} we have derived for the first time, using recent
+Eq. (\ref{et}) we have derived for the first time, using recent
 results \citep[][]{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017}, ET
@@
 
-where $\Gamma$ is the CO$_2$ compensation point \citep[Equation 5
+where $\Gamma$ is the CO$_2$ compensation point \citep[Eq. (5)
 in][]{Zhou_2014}. So uWUE is related to atmospheric CO$_2$
@@
 
-Given eddy-covariance FLUXNET2015 data (Section \ref{data}), every
-term in our new version of PM (Equation \ref{et}) is observed, except
+Given eddy-covariance FLUXNET2015 data (Sect. \ref{data}), every
+term in our new version of PM (Eq. (\ref{et})) is observed, except
 for uWUE, which we fit by calculating its expectation, given the model
@@
 observations themselves. The variability of $\sigma$ then propagates
- any uncertainty through to our partial derivative of Equation
-\ref{et_sigma} with respect to VPD:
+ any uncertainty through to our partial derivative of Eq.
+(\ref{et_sigma}) with respect to VPD:
 
@@
 
-With Equation \ref{d_et} we have an analytical framework for ecosystem
+With Eq. (\ref{d_et}) we have an analytical framework for ecosystem
 response to atmospheric demand perturbations with environmental
 conditions held fixed. There are a few subtleties to taking the
-derivative in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$)
+derivative in Eq. (\ref{d_et}): $\Delta$ ($\frac{d e_{s}}{d T}$)
 and $VPD$ are functionally related, so while taking the derivative we
@@
 plant behaviors, we take the extra care of including and analyzing
-uncertainty (Section \ref{testing}) and examining sensitivity to the
-exponent of the VPD response (Section \ref{functional_form}).
+uncertainty (Sect. \ref{testing}) and examining sensitivity to the
+exponent of the VPD response (Sect. \ref{functional_form}).
 
@@
 drought indices. If we approximate $c_a$ at a global mean CO$_2$
-concentration, then the RHS of Equation \ref{et} is fully defined
+concentration, then the RHS of Eq. (\ref{et}) is fully defined
 using commonly available weather station data and the constants
-published here. This makes Equation \ref{et} useful in addition to
+published here. This makes Eq. (\ref{et}) useful in addition to
 PET in drought indices and hydrometeorological analysis for vegetated
-surfaces. Equation \ref{et} better reflects the physics of water
+surfaces. Equation (\ref{et}) better reflects the physics of water
 exchange at the land surface and would only require fitting of uWUE
@@
 
-By construction, the variability in the $\sigma$ term (Equation
-\ref{sigma}) contains all model and observational uncertainties. For
+By construction, the variability in the $\sigma$ term (Eq.
+(\ref{sigma})) contains all model and observational uncertainties. For
 an observation that perfectly matches our model and constant uWUE
@@
 dependence of $\sigma$ on VPD in their joint distribution, and
-$\sigma$ is indeed close to unity i.e. $O(1)$ (Figure
+$\sigma$ is indeed close to unity i.e. $O(1)$ (Fig.
 \ref{joint_vpd_sigma}). Given this weak dependence and the
@@
 Before calculating the sensitivity of ET to VPD, we will consider the
-functional form of Equation \ref{d_et}. There are two main terms: a
+functional form of Eq. (\ref{d_et}). There are two main terms: a
 ``scaling'' term, which modifies the magnitude but not the sign of the
@@
 All variables are positive, so the relative magnitude between the
-first term and the second term in the sign term (Equation \ref{sign})
+first term and the second term in the sign term (Eq. (\ref{sign}))
 will determine whether ET increases or decreases with increasing
@@
 VPD for the more water conservative CSH, while ET frequently increases
-with increasing VPD for the more water intensive CRO (Figure
+with increasing VPD for the more water intensive CRO (Fig.
 \ref{idealized_sign}). CROs evolved or were bred to prioritize GPP and
@@
 As expected, the slope parameter ($g_1$) is a primary determinant of
-the VPD dependence for the sign term shown in Figure
+the VPD dependence for the sign term shown in Fig.
 \ref{idealized_sign}. Plants that are more conservative (small $g_1$)
@@
 apparent in the strong VPD dependence of CSH, which has the lowest
-slope parameter ($g_1=75.1$ Pa$^{1/2}$) (Figure \ref{idealized_sign}).
-
-To summarize our theoretical insights (Figure \ref{idealized_sign} and
+slope parameter ($g_1=75.1$ Pa$^{1/2}$) (Fig. \ref{idealized_sign}).
+
+To summarize our theoretical insights (Fig. \ref{idealized_sign} and
 Table \ref{vpd_crit}), CROs are the least water conservative and have
@@
 tendency to decrease ET with increasing VPD, as well as the strongest
-VPD dependence of response. Figure \ref{idealized_sign} clearly shows,
+VPD dependence of response. Fig. \ref{idealized_sign} clearly shows,
 according to our theory, that for all PFTs except for crops there is
@@
 The control of the scaling term variability between PFTs by
-aerodynamic conductance is confirmed by data (Figure
+aerodynamic conductance is confirmed by data (Fig.
 \ref{scale_vary}). Differences between PFT are almost entirely due to
@@
 environmental temperature and aerodynamic conductance variability
-(Figure \ref{scale_vary}). While the observed variability of the
+(Fig. \ref{scale_vary}). While the observed variability of the
 aerodynamic conductance contributes more to the scaling term
@@
 eddy-covariance data, while including uncertainty with the $\sigma$
-term (Section \ref{methods}). These observational results of ET
+term (Sect. \ref{methods}). These observational results of ET
 response (Table \ref{stats}) largely confirm our theoretical analysis, presented in the previous sections. For all PFTs, mean ET response to increasing VPD
@@
 the non-linear nature of the response, in which negative responses are
-generally larger magnitude than positive responses (Figure
+generally larger magnitude than positive responses (Fig.
 \ref{idealized_sign}). Therefore, both the mean ET response as well as
 the ET response at mean environmental conditions matches our
-expectations from the theory (Section \ref{sign_term}), with the
+expectations from the theory (Sect. \ref{sign_term}), with the
 exception that CRO observations are shifted more towards a negative ET
@@
 in the distribution of environmental VPD (e.g. SAV and WSA grow in
-more arid climates, Figure \ref{idealized_sign}). In general the bulk
+more arid climates, Fig. \ref{idealized_sign}). In general the bulk
 statistics match our theoretical expectations well, with the caveat
@@
 exhibit a high frequency of negative ET response to VPD, towards more frequent
-occurrence of positive ET response than the theory and Figure
+occurrence of positive ET response than the theory and Fig.
 \ref{idealized_sign} would suggest. The bulk statistics motivate a
@@
 We now compare more sophisticated distributions of the observed
-response to our simplified theory (Section \ref{sign_term}). The
+response to our simplified theory (Sect. \ref{sign_term}). The
 observed distribution of the sign term, as compared to what the theory
-would predict, is provided in Figure \ref{test_sign}. Our goal was to
+would predict, is provided in Fig. \ref{test_sign}. Our goal was to
 capture the leading order behavior of the ET dependence on VPD. Given
@@
     (box plots) to the sign term as calculated with simplifying
-    assumptions (blue line, as in Figure \ref{idealized_sign}). Each
+    assumptions (blue line, as in Fig. \ref{idealized_sign}). Each
     box plot corresponds to 5\% of the data, and the 5-95\% range of
@@
 is biased slightly towards positive response as compared to the
-observations (Figure \ref{test_sign}). For ENF this slight bias could
+observations (Fig. \ref{test_sign}). For ENF this slight bias could
 be explained by a negative bias in $g_1$. However, for WSA, CRO, and
@@
 towards more negative response, which cannot be captured by the
-functional form of the sign term (Figure
+functional form of the sign term (Fig.
 \ref{idealized_sign}). Therefore, to explain the observations one of
@@
 a larger portion of observations are biased negative as compared to
-the theory ($\approx$ 35\%, Figure \ref{test_sign}). However, in
+the theory ($\approx$ 35\%, Fig. \ref{test_sign}). However, in
 general, and specifically for non-extreme VPD (VPD  $< \approx$70-90th
@@
 alter $uWUE$ (or $g_1$) or invalidate other assumptions made in the
-methods section (Section \ref{methods}). We do not account for
+methods section (Sect. \ref{methods}). We do not account for
 variability in plant height and surface roughness, or differences in
@@
 back towards negative ET response at high VPD when all sites are
-binned together, as in Figure \ref{test_sign}. We hypothesize that the
+binned together, as in Fig. \ref{test_sign}. We hypothesize that the
 theory would validate against observations much better if these
@@
 and $uWUE$ can be analyzed with our framework. This opens a door to future analyses in which plant behavior in anomalous conditions
-can be explained and analyzed using Equation \ref{et}.
-
-% It is useful to also compare our theory, the observed uncertainty in
-% the ET response, and PET, as PET is often used as a proxy for ET in
-% hydrometeorological studies and drought indices. Again, as in our
-% discussion of the theoretical results, the PET sensitivity to VPD
-% will be constant at about 3.5. In Figure \ref{test_sign} the ET
-% response to shifts in dryness is always far smaller than 3.5 in the
-% observations, even with all uncertainty included. So the PET
-% response to atmospheric dryness is outside the range of all
-% observable responses. This casts doubt on PET's usefulness for
-% capturing the leading order behavior of water budget and dryness
-% response at the land surface, as is done in several drought
-% indices. However our theory (Equation \ref{et}) could be used with
-% published uWUE and $g1$ data, as an alternative to PET to estimating
-% land response and dryness, and would more accurately account for the
-% response to atmospheric drying.
+can be explained and analyzed using Eq. (\ref{et}).
 
@@
 turn to observations of ET response with the scaling term included
-(Figure \ref{data_scatter}). Until now, in the interest of developing
+(Fig. \ref{data_scatter}). Until now, in the interest of developing
 leading order intuition for ET response, and to be conservative in our
@@
 
-Therefore, with the caveat that some of the signal presented in Figure
+Therefore, with the caveat that some of the signal presented in Fig.
 \ref{data_scatter} may in fact be noise, we can interpret the observed
@@
 \textit{negative} response at high VPD. This is as expected, given
-previous discussion on how we might expect more inter-site and inter-year variability for GRA (Section \ref{testing}).
-
-Figure \ref{data_scatter} also includes the impact of the scaling
+previous discussion on how we might expect more inter-site and inter-year variability for GRA (Sect. \ref{testing}).
+
+Fig. \ref{data_scatter} also includes the impact of the scaling
 term. For a given VPD, the magnitude of ET response does not vary
@@
 
-As with Figure \ref{test_sign}, Figure \ref{data_scatter} matches our
+As with Fig. \ref{test_sign}, Fig. \ref{data_scatter} matches our
 expectation based on simplified theory. The sign term is most strongly
@@
 Indeed, for all PFTs, there is some slight dependence of $\sigma \cdot
-uWUE$ on SWC, especially at low SWC (Figure \ref{swc_boxplot}). The
+uWUE$ on SWC, especially at low SWC (Fig. \ref{swc_boxplot}). The
 portion of observations for which our theory is biased by
@@
 soil water content only matters for about the lowest 5\% of observations
-(each box is 5\% of observations in Figure \ref{swc_boxplot}). And, CRO,
+(each box is 5\% of observations in Fig. \ref{swc_boxplot}). And, CRO,
 relative to DBF, has a very weak dependence on soil moisture, which is
@@
 poorly against the data for GRA, relative to to the other PFTs
-(Section \ref{testing}). For all other PFTs occurrence of soil
+(Sect. \ref{testing}). For all other PFTs occurrence of soil
 moisture impacts was rare enough to not manifest itself in bulk
@@
 explain the deficiencies of our theory compared to the observations in
-Figure \ref{idealized_sign}, specifically the trend back towards
+Fig. \ref{idealized_sign}, specifically the trend back towards
 negative soil moisture at high VPD. Aforementioned feedbacks between
@@
 of low SWC are more likely, and this low SWC causes a lower uWUE
-(Figure \ref{swc_boxplot}). The lower uWUE at low SWC/high VPD then
+(Fig. \ref{swc_boxplot}). The lower uWUE at low SWC/high VPD then
 leads to the observed downturn towards decreasing ET with increasing
@@
 the observations. Soil moisture effects explain the deficiencies of
-our theory in Section \ref{testing}, particularly for GRA. By
+our theory in Sect. \ref{testing}, particularly for GRA. By
 conceptualizing SWC effects as a change in $uWUE$ (and/or $g_1$), it
@@
 
-The theory described in Section \ref{sign_term} indicates that for a
+The theory described in Sect. \ref{sign_term} indicates that for a
 given $uWUE$ and $g_1$, the ET dependence on VPD should be concave
@@
 and are what allowed us to derive an explicit ET expression for the
-first time (Equation \ref{et}), we will analyze if and how assumptions
+first time (Eq. (\ref{et})), we will analyze if and how assumptions
 about the exponent of the VPD dependence impacts the shape of the
@@
 shape of the ET-VPD dependence we find the roots of the second
-derivative of ET, using Equation \ref{m_n} for stomatal conductance
+derivative of ET, using Eq. (\ref{m_n}) for stomatal conductance
 ($g_s$), with respect to VPD: \input{d2_solutions.tex} With this
 result we have defined the family of curves separating concave up from
-concave down ET solutions (Figure \ref{concave}). These curves are
+concave down ET solutions (Fig. \ref{concave}). These curves are
 only functions of the exponent of the VPD dependence and a quantity we
 call non-dimensional VPD ($VPD^m/g*$). Several important
-relations reveal themselves from Equation \ref{curves}:
+relations reveal themselves from Eq. (\ref{curves}):
 
@@
   \caption{ Solutions corresponding to inflection points between
-    concave up and concave down ET-VPD curves (Equation \ref{curves})
+    concave up and concave down ET-VPD curves (Eq. (\ref{curves}))
     for three specific scenarios. Solutions are defined in terms of a
@@
 framework in which uWUE is allowed to decrease with decreasing SWC, as
-observed in Section \ref{swc_section}. Here, we focus our results by
+observed in Sect. \ref{swc_section}. Here, we focus our results by
 assuming constant PFT-wide conditions to build baseline intuition for
@@
 fixed matches the leading order behavior of the observations where
-plant function varies (Section \ref{testing}).
+plant function varies (Sect. \ref{testing}).

### 2018-10-31 doc/paper/conclusions.tex
about to spell check

@@
 allow for explicit, accurate future analysis of plant-VPD feedbacks
-in the atmospheric boundary layer (Equation \ref{et}). Our approach
+in the atmospheric boundary layer (Eq. (\ref{et})). Our approach
 can be extended to examine varying plant response to more nuanced

### 2018-10-31 doc/paper/body.tex
spellcheck
ready to submit

@@
 woody savannah (WSA) (site locations and citations are in Fig.
-\ref{map_fig} and Appendex \ref{flux_sites}, respectively).
+\ref{map_fig} and Appendix \ref{flux_sites}, respectively).
 
@@
 
-\subsection{Limitations of theory - very dry soil moisture conditions}
+\subsection{Limitations of theory: very dry soil moisture conditions}
 \label{swc_section}
@@
 explain the deficiencies of our theory compared to the observations in
-Fig. \ref{idealized_sign}, specifically the trend back towards
-negative soil moisture at high VPD. Aforementioned feedbacks between
+Fig. \ref{test_sign}, specifically the trend back towards
+negative ET response at high VPD. Aforementioned feedbacks between
 the land surface and the atmosphere, which are not accounted for due
@@
 \subsection{Functional form of ET dependence on VPD and its relation
-  to VPD dependence}
+  to the VPD exponent}
 \label{functional_form}
@@
   \begin{equation}
-    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_*}{VPD^m}\right) \frac{*WUE \; ET}{c_a \; VPD^n},
+    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g*}{VPD^m}\right) \frac{*WUE \; ET}{c_a \; VPD^n},
     \label{m_n}
@@
 where:
-\[*WUE = \frac{GPP}{ET}VPD^n,\] and $g_*$ is a generic slope parameter
+\[*WUE = \frac{GPP}{ET}VPD^n,\] and $g*$ is a generic slope parameter
 of units $VPD^m$. To determine how the exponent $n$ and $m$ alter the

### 2018-10-31 doc/paper/conclusions.tex
ready to submit

@@
 
-Our paper builds important intuition for how plants will
-respond to VPD perturbations. We show that given optimal stomatal
-function and fixed environmental conditions, the ET-VPD dependence is
-theoretically concave upward, with ET increasing with increasing VPD as
-VPD increases past some critical value (Table \ref{vpd_crit}). However
+Our paper builds important intuition for how plants respond to VPD
+perturbations. We show that given optimal stomatal function and fixed
+environmental conditions, the ET-VPD dependence is theoretically
+concave upward, with ET increasing with increasing VPD as VPD
+increases past some critical value (Table \ref{vpd_crit}). However
 future research should focus on fully understanding the form of
-stomatal VPD dependence, as this result is sensitive to the exponent of
-VPD dependence, which we currently believe is 1/2 \citep{MEDLYN_2011,
-  Zhou_2014}. Our results are applicable to understanding the impact
+stomatal VPD dependence, as this result is sensitive to the exponent
+of VPD dependence, which we currently believe is 1/2
+\citep{MEDLYN_2011, Zhou_2014}. Indeed, this sensitivity to the exponent of
+VPD dependence is an important result itself: land surface models,
+including those used in earth system models for climate forecasts,
+employ different assumptions about the exponent of VPD dependence in
+stomatal conductance \citep[e.g.,][]{Ball_1987, Leuning_1990,
+MEDLYN_2011}, and these assumptions can fundamentally change the
+relationship between ET and VPD from one that is concave upward (local
+minimum in ET) to one that is concave downward (local maximum in ET).
+
+Our results are also applicable to understanding the impact
 of  expected increases in VPD

### 2018-10-31 doc/paper/intro.tex
ready to submit

@@
 surfaces \citep{Medlyn_2017, Zhou_2014}, and are novel tools to
-transparently understand and develop intuition for complex
-ecohydrologic response to environmental forcing.
+transparently develop intuition for the behavior of complex, multiscale
+ecohydrologic systems.
 

### 2018-10-31 src/paper_figure_code/concave.py

@@ -14,7 +14,7 @@ r, sigma = symbols('r sigma')
 # init aero
 g_a, p, t, cp, r_air, x, g_s = symbols('g_a P T c_p R_air x g_s')
 # init plant
-gamma, c_s, r_star, uwue, g_1, onesix = symbols('gamma c_s R* uWUE g_* 1.6')
+gamma, c_s, r_star, uwue, g_1, onesix = symbols('gamma c_s R* uWUE g* 1.6')
 
 vpd, delta = symbols('VPD Delta')
 
@@ -44,7 +44,7 @@ fh = open("../../doc/paper/d2_solutions.tex", "w")
 fh.write("\\begin{equation}\n")
 fh.write("\\frac{\\partial^2 \; ET}{\\partial \\; VPD^2} = 0 "\
          "\\quad \\forall \\quad")
-fh.write("%s%s%s" % ("\\frac{VPD^m}{g1} = ", latex(soln[0]), ".\n"))
+fh.write("%s%s%s" % ("\\frac{VPD^m}{g*} = ", latex(soln[0]), ".\n"))
 fh.write("\\label{curves}\n")
 fh.write("\\end{equation}\n")
 fh.close()

### 2018-12-17 doc/paper/hess-revision/review-1.tex
rough draft of RC1 response

\RequirePackage{natbib}
\bibliographystyle{agufull08}

Thank you very much for an excellent and very thoughtful
review. Addressing your comments will greatly improve the
manuscript. We agree with your comments, and think the manuscript
could use substantial reframing and rewording to clarify how we are
answering our research question, and how the answer may vary with
climate and environmental factors.

Just a quick general comment before addressing your specific
comments. By including a lot of content on PFT-focused analysis we
failed to communicate the major goal and scope of our project: we were
trying to characterize the response of ET to VPD, with all other
environment variables held fixed. To accomplish this, we needed to
formulate an explicit function of ET in terms of environemntal
variabiles and parameters, where any parameters can be approximated as
constant with regards to some [arbitrary] VPD perturbation
scenario. We will elaborate more below on why we used PFT-focused
scenarios for much of our analysis. However, the goal of our
manuscript was much more simple and fundamental: at a given place or
time, if you introduce a perturbation to VPD, what is the immediate ET
response (postive or negative)? Answering this question does not
require the much stronger assertion that any parameters must be
invariant within a given PFT.

\section{Compounding uncertainties}

We agree that the uncertainties are large both within a PFT and across
a PFT. Looking back, we believe that some of our language
communicating \cite{Lin_2018}, \cite{Medlyn_2017}, and
\cite{Zhou_2014}'s results was misleading, and this was exacerbated by
our focus on PFT-analysis. We will remove that language from the
manuscript, and add language to better communicate our results as
reflections of the considerable within-PFT uncertainty.

I think comparing the rankings of given PFT values for g$_1$ and uWUE
can be misleading, given the magnitude of the uncertainties
involved. Most of our calculated values for uWUE are within one
standard deviation of \cite{Zhou_2015}'s results, but these deviations
can result in some changes in the ordering of PFTs from high to low
uWUE. The bigger problem in our eyes is that we made a mistake with
some language suggesting within intra-PFT variability is less than
inter-PFT variability, but clearly this is not the case. We will
remove this language. These misinterpretation should not have made the
manuscript; we actually included \cite{Zhou_2015}'s results in Table 2
to be transparent about the within-PFT uncertainties. g$_1$ values
also exhibit considerable uncertainty, and again, while the relative
rankings of these values may change, they are all within observed
ranges in \cite{Medlyn_2017}. We did not include these in Table 2
because \cite{Medlyn_2017} does not provide the numerical values;
however, we think it would be useful to provide estimates of these
ranges form \cite{Medlyn_2017}'s figures in our Table 2.

Regarding how to interpret g$_1$ and uWUE: we think the most best way
is with established physical relations from \cite{MEDLYN_2011}
(Equation 3 in the manuscript) and \cite{Zhou_2014} (Pg. 7, line 10 in
the manuscript), with the knowledge that there is uncertainty
involved. The quantities in these relationships all have some
intrinsic physical meaning, but can vary substantially within a
PFT. Again, we need to alter our language to better reflect this
intra-PFT variability.

However, \cite{Zhou_2014} did establish a constant uWUE approximation
as a good approximation for capturing the relationship between GPP, ET
and VPD at a given place and time, so it is still a very useful and
robust approximation for answering our research question (see
discussion earlier in this comment). Additionally, a constant g$_1$
approximation is used in many earth system models, so while it
introduces considerable uncertainty, it also makes our framework
useful for interpreting modelled vegetation response in ESMs and GCMs.

We also agree with the comment that the best way to think of this is
probabalistically. This is the most robust approach to dealing with
approximations - we introduce randomness to our variables and
paramters to account for all of the physics that are not explicitly
accounted for, as well as observational uncertainty. However, as far
as I know we still have not developed a general, robust, and efficient
arithmetic for random variables. We could try and adapt a Bayesian
model to this problem, but given the large amount of arithmetic
invloved, fully incorporating a Bayesian represntation to every
variable in the analysis would be a very hard problem, and a
significant research project in its own right. I think a good
compromise is to add language directing the reader to interpret our
results more probabalistically, which we have already presented
probablistically in the figures. For example, in Figure 5 the range
of values in each plot represents a range of possible responses in the
sign term. Some portion of this could be due to oveservational
uncertainty, and others could be due to physical intra-PFT variations
inducing changes in uWUE and/or g$_1$. It's also worth noting that our
intra-PFT variability is greater than the inter-PFT
variability. Again, we need to add language highlighting this, and its
consistency with previous results. We included this uncertainty
explicitly to be transparent.

\subsection{Attribution to physiological responses}

We do not want our analysis to be interpretted as an assertion that
g$_1$ and uWUE are attributes of PFT only. We expect them to vary both
within a PFT and across a PFT. Our primaty goal in using them was to
develop an explict expression of ET as a function of environmental
variables, and use this to assess the response to a change in VPD. We
focused our analysis using PFTs because this is how it was framed in
previous studies in our field, and specificically the studies used in
our derivation \citep{Zhou_2014, Zhou_2015, Medlyn_2017}. We were
thinking the PFT-focused analysis could be useful, especially given
that climate models generally hold plant paramters fixed with respect
to PFT, so long as we were transparent about the large uncertainties
and problems with this approach (see Figure 5).

We agree with you, and think stating that any quantity is fixed within
a PFT is hard to believe. Phenotypic variation and adaptation within a
given specicies can be considerable (i.e. effecting $\lambda$, g$_1$,
and uWUE), so it would be hard to say that anything would be constant
within a PFT made up of many different diverse species. Both
phenotypic variation as well as the species distirbution and dominant
PFT at a given location will all be strongly optimized in response to
climate. In this sense, the distribution and evolution of local
climate is a strong control on the local ecosystem structure and
physiology of that ecosystem.

What we need to communicate better is that a given ecosystem's state
at a place or time controlls its response to a VPD perturbation. We
are making an approximation that we can parameterize the effect of the
ecoystems' state on VPD response with g$_1$ and uWUE. These quantities
can also vary due to soil moisture condition; the approximation we
need to make to answer for our research question is that they are
fixed with respect to a VPD perturbation. In this sense, we do not
view the two statements: ``plants that are evolved to bred to
prioritize primary production over water conservation (e.g., crops)
exhibit a higher likelihood of atmospheric demand-driven response''
and ``ecosystem types are responding in this way because they have, on
the whole, been subjected to less soil water limitation (due to the
non-negligible effects of irrigation?)'' as mutually exclusive. In
fact, we view them as consistent with a view that crops and thier
physiology (parameterized by g$_1$ and uWUE) exist in a given
environment because they have not been subjected to soil water
limitation, and they have a given response to VPD because of their
physiology (which is a direct effect of the environment). In this way,
climate and land surface state are causes of the VPD response through
their effect on plant physiology (parameterized by g$_1$ and uWUE).

When writing this manuscript, there was definitely some internal
tension and debate about how to frame the analysis and results. We
could either focus on PFT-oriented results as previous literature has
done (e.g. holding plant physiology fixed within a given PFT), or
allow plant physiology terms to vary through time and space and look
at the distribution of ET response to VPD (see Section 3.5). After
re-examining the manuscript both after some time away from the
problem and in light of your comments I think a strong argument could
be made that we made the wrong choice with respect to this focus. It
may have made more sense to focus our analysis on the ET response more
generally across space and time as ecosystem-scale plant physiology
varies in response to climate and soil moisture.

\section{Next Steps}

In order to fix the manuscript and our communication of the answer to
the question ``When does VPD drive or reduce ET?'', we see a few
potential paths that we will consider between now and the final
response after the discussion period. If you (or anyone else) has any
opinions or comments in the meantime, we would appreciate the insight
and feedback.

\begin{enumerate}
\item We include the discussion presented in this review and response,
  and include language explicitly stating that the purpose of the PFT
  analysis is to provide connections to other PFT-constant analyses
  and models (e.g. ESMs and GCMs). We will add extensive language on
  the uncertainty and weaknesses of this approach, and reframe our
  conclusions to reflect his uncertainty.
\item We instead alter our analysis to look at how VPD response
  varies with climate-induced variations and general plan
  physiological variability instead of just PFT variability that
  poorly captures all of the observed variability in
  response. Sections 3.1 - 3.2 would be replaced by analysis direcly
  relating ET response to general climate and plant physiological
  terms, rather than PFT-mean analysis. For example, the ``scaling
  term'' analysis would be presented in terms of generic changes in
  plant height and temperature, and the ``sign term'' analysis would
  be framed by generic changes to g$_1$ and uWUE, as informed by
  previous literature.
\item This is the most extreme option in terms of modifying the
  manuscript. We significantly alter the manuscript and instead focus
  on just the general shape of the ET-VPD curve (with environmental
  variables held fixed). This was not discussed in the review, but one
  of our most noteworthy results is on the general shape of the ET-VPD
  curve being concave up, given an assumption of a square-root VPD
  dependence of plant response. This result is independent of any
  assumptions of uWUE and g$_1$, and to our knowledge it is first
  derived curve of ecosystem response to VPD (see Section 3.7). It
  also highlights the importance of discerning the exact exponent of
  VPD dependence, as this alters the fundamental nature and shape of
  the curve. Essentially, the manuscript would become just our
  motivation and derivation, followed by Section 3.7 exploring the
  consequences of the derivation for the shape of the VPD curve. This
  alteration of the manuscript would likely result in more of a
  technical note type paper, and we are hesitant to do this because we
  think there is still a lot of useful information obtained by tying
  our results to real-world scenarios (as in Options 1 and 2 above).
\end{enumerate}

Here we present some final minor comments and concerns on the
technical details of our proposed changes to the manuscript, which are
relevant to representing uncertainty and spatiotemporal variabiliy in
uWUE \textit{and} g$_1$. In the original manuscript we used a single
$\sigma$ term to represent this variability. We did this because there
was changes in uWUE and g$_1$ induce a very similiar change in the ET
solution, which made solving for independent $\sigma_{uWUE}$ and
$\sigma_{g1}$ terms intractable, and represnting variability in both
g$_1$ and uWUE difficult. We decided to hold g$_1$ fixed within a PFT
for a two reasons: 1) ESMs and GCMs generally hold g$_1$ fixed, and 2)
letting uWUE vary seemed more appropraite to represent specifically
soil water variations on stomatal conductance, as $uWUE$ modifies
stomatal conductance analgalously to how soil moisture factors modify
a maximum stomatal conductance in land surface models (e.g. it is a
multiplicative factor on the entire stomatal conductance
term). Because uWUE and g$_1$ induce similar changes in ET, at least
qualitatively we think that having a single $\sigma$ can at least
represent some of the variability in both $uWUE$ and g$_1$. However,
formally there is no explicit variability in the g$_1$ term. The
discussion in this review rightly points out that based on previous
results we do expect some variability in g$_1$ as well. So, an
unresolved question is how important is it to the analysis and its
interpretation to include an explicit g$_1$ variability term in
addition to the exisiting $\sigma$ term. I'm including these comments
to hopefully stimulate some discussion. I could imagine some timescale
based approaches where we might be able to account for both the g$_1$
and uWUE variability, for example by making assumptions over what time
scale each quantity is fixed, and fitting based on that (e.g. g$_1$ is
fixed for a given season, and uWUE is fixed for a given day). This
approach might allow a tractable solution, and also could help us
filter error and observational noise from ``true''  plant
physiological and climatic variability in uWUE and g$_1$. The cost of
all of this is a significant increase in the analysis and content of
the paper, as well as some increased opacity to the methods.

Thank you again for the thoughtfull review. We hope it stimulates
further discussion.

\bibliography{references.bib}

### 2018-12-17 doc/paper/hess-revision/review-1.tex
what I sent to Pierre and Changjie for edits

@@
 manuscript. We agree with your comments, and think the manuscript
-could use substantial reframing and rewording to clarify how we are
+could use substantial re-framing and rewording to clarify how we are
 answering our research question, and how the answer may vary with
@@
 comments. By including a lot of content on PFT-focused analysis we
-failed to communicate the major goal and scope of our project: we were
+failed to communicate the major goal and scope of our project: we are
 trying to characterize the response of ET to VPD, with all other
-environment variables held fixed. To accomplish this, we needed to
-formulate an explicit function of ET in terms of environemntal
-variabiles and parameters, where any parameters can be approximated as
+environment variables held fixed. To accomplish this, we need to
+formulate an explicit function of ET in terms of environmental
+variables and parameters, where any parameters can be approximated as
 constant with regards to some [arbitrary] VPD perturbation
@@
 time, if you introduce a perturbation to VPD, what is the immediate ET
-response (postive or negative)? Answering this question does not
+response (e.g. positive or negative)? Answering this question does not
 require the much stronger assertion that any parameters must be
@@
 
-I think comparing the rankings of given PFT values for g$_1$ and uWUE
+We think comparing the rankings of given PFT values for g$_1$ and uWUE
 can be misleading, given the magnitude of the uncertainties
@@
 inter-PFT variability, but clearly this is not the case. We will
-remove this language. These misinterpretation should not have made the
+remove this language. This misinterpretation should not have made the
 manuscript; we actually included \cite{Zhou_2015}'s results in Table 2
@@
 however, we think it would be useful to provide estimates of these
-ranges form \cite{Medlyn_2017}'s figures in our Table 2.
+ranges from \cite{Medlyn_2017}'s figures in our Table 2.
 
@@
 introduces considerable uncertainty, it also makes our framework
-useful for interpreting modelled vegetation response in ESMs and GCMs.
+useful for interpreting modeled vegetation response in ESMs and GCMs.
 
 We also agree with the comment that the best way to think of this is
-probabalistically. This is the most robust approach to dealing with
+probabilistically. This is the most robust approach to dealing with
 approximations - we introduce randomness to our variables and
-paramters to account for all of the physics that are not explicitly
+parameters to account for all of the physics that are not explicitly
 accounted for, as well as observational uncertainty. However, as far
-as I know we still have not developed a general, robust, and efficient
+as we know we still have not developed a general, robust, and efficient
 arithmetic for random variables. We could try and adapt a Bayesian
 model to this problem, but given the large amount of arithmetic
-invloved, fully incorporating a Bayesian represntation to every
+involved, fully incorporating a Bayesian representation to every
 variable in the analysis would be a very hard problem, and a
-significant research project in its own right. I think a good
+significant research project in its own right. We think a good
 compromise is to add language directing the reader to interpret our
-results more probabalistically, which we have already presented
-probablistically in the figures. For example, in Figure 5 the range
-of values in each plot represents a range of possible responses in the
-sign term. Some portion of this could be due to oveservational
-uncertainty, and others could be due to physical intra-PFT variations
-inducing changes in uWUE and/or g$_1$. It's also worth noting that our
-intra-PFT variability is greater than the inter-PFT
-variability. Again, we need to add language highlighting this, and its
-consistency with previous results. We included this uncertainty
-explicitly to be transparent.
+results more probabilistically, which we have already presented
+probabilistically in the figures. For example, in Figure 5 the range of
+values in each plot represents a range of possible responses in the
+sign term. Within this figure, it's worth noting that the intra-PFT
+variability is greater than the inter-PFT variability, which is
+consistent with some of the previous results you highlight. Again, we
+need to add language highlighting this, and its consistency with
+previous results. We included this variability and uncertainty
+explicitly to be transparent. A more difficult question is how much of
+this variability is due to observational and model error, and how much
+of it is due to climate and plant physiological variability (see
+Section 3.5).
 
@@
 
-We do not want our analysis to be interpretted as an assertion that
+We do not want our analysis to be interpreted as an assertion that
 g$_1$ and uWUE are attributes of PFT only. We expect them to vary both
-within a PFT and across a PFT. Our primaty goal in using them was to
-develop an explict expression of ET as a function of environmental
-variables, and use this to assess the response to a change in VPD. We
-focused our analysis using PFTs because this is how it was framed in
-previous studies in our field, and specificically the studies used in
-our derivation \citep{Zhou_2014, Zhou_2015, Medlyn_2017}. We were
-thinking the PFT-focused analysis could be useful, especially given
-that climate models generally hold plant paramters fixed with respect
-to PFT, so long as we were transparent about the large uncertainties
-and problems with this approach (see Figure 5).
+within a PFT and across a PFT. Our primary goal in using g$_1$ and
+uWUE was to develop an explicit expression of ET as a function of
+environmental variables, and use this to assess the response to a
+change in VPD. We focused our analysis using PFTs because this is how
+it was framed in previous studies in the field, and specifically the
+studies used in our derivation \citep{Zhou_2014, Zhou_2015,
+  Medlyn_2017}. We were originally thinking the PFT-focused analysis
+could be useful, especially given that climate models generally hold
+plant parameters fixed with respect to PFT, so long as we were
+transparent about the large uncertainties and problems with this
+approach (see Figure 5).
 
-We agree with you, and think stating that any quantity is fixed within
-a PFT is hard to believe. Phenotypic variation and adaptation within a
-given specicies can be considerable (i.e. effecting $\lambda$, g$_1$,
-and uWUE), so it would be hard to say that anything would be constant
+We agree with you that stating that any quantity is fixed within a PFT
+is hard to believe. Phenotypic variation and adaptation within a given
+\textit{species} can be considerable (i.e. effecting $\lambda$, g$_1$, and
+uWUE), so it would be hard to say that anything would be constant
 within a PFT made up of many different diverse species. Both
-phenotypic variation as well as the species distirbution and dominant
+phenotypic variation as well as the species distribution and dominant
 PFT at a given location will all be strongly optimized in response to
@@
 climate is a strong control on the local ecosystem structure and
-physiology of that ecosystem.
+physiology of a given ecosystem.
 
 What we need to communicate better is that a given ecosystem's state
-at a place or time controlls its response to a VPD perturbation. We
-are making an approximation that we can parameterize the effect of the
-ecoystems' state on VPD response with g$_1$ and uWUE. These quantities
+at a place or time controls its response to a VPD perturbation. We are
+making an approximation that we can parameterize the effect of the
+ecoystem's state on VPD response with g$_1$ and uWUE. These quantities
 can also vary due to soil moisture condition; the approximation we
-need to make to answer for our research question is that they are
-fixed with respect to a VPD perturbation. In this sense, we do not
-view the two statements: ``plants that are evolved to bred to
-prioritize primary production over water conservation (e.g., crops)
-exhibit a higher likelihood of atmospheric demand-driven response''
-and ``ecosystem types are responding in this way because they have, on
-the whole, been subjected to less soil water limitation (due to the
+need to make to answer our research question is that they are fixed
+with respect to a VPD perturbation. In this sense, we do not view the
+two statements: ``plants that are evolved to bred to prioritize
+primary production over water conservation (e.g., crops) exhibit a
+higher likelihood of atmospheric demand-driven response'' and
+``ecosystem types are responding in this way because they have, on the
+whole, been subjected to less soil water limitation (due to the
 non-negligible effects of irrigation?)'' as mutually exclusive. In
-fact, we view them as consistent with a view that crops and thier
-physiology (parameterized by g$_1$ and uWUE) exist in a given
-environment because they have not been subjected to soil water
-limitation, and they have a given response to VPD because of their
-physiology (which is a direct effect of the environment). In this way,
-climate and land surface state are causes of the VPD response through
-their effect on plant physiology (parameterized by g$_1$ and uWUE).
+fact, we view them as consistent with a view that crops and their
+physiology (parameterized by g$_1$ and uWUE) exist at a given time and
+place because they have not been subjected to soil water limitation,
+and they have a given response to VPD because of their physiology
+(which is a direct effect of the environment). In this way, climate
+and land surface state are causes of the VPD response both directly
+and through their effect on plant physiology (parameterized by g$_1$
+and uWUE).
 
@@
 re-examining the manuscript both after some time away from the
-problem and in light of your comments I think a strong argument could
+problem and in light of your comments we think a strong argument could
 be made that we made the wrong choice with respect to this focus. It
@@
   conclusions to reflect his uncertainty.
-\item We instead alter our analysis to look at how VPD response
-  varies with climate-induced variations and general plan
-  physiological variability instead of just PFT variability that
-  poorly captures all of the observed variability in
-  response. Sections 3.1 - 3.2 would be replaced by analysis direcly
-  relating ET response to general climate and plant physiological
-  terms, rather than PFT-mean analysis. For example, the ``scaling
-  term'' analysis would be presented in terms of generic changes in
-  plant height and temperature, and the ``sign term'' analysis would
-  be framed by generic changes to g$_1$ and uWUE, as informed by
-  previous literature.
+\item We instead alter our analysis to look at how VPD response varies
+  with climate and general plant physiological variability, instead of
+  focussing only on PFT analysis that poorly captures all of the observed
+  variability in ET response. Sections 3.1 - 3.2 would be replaced by
+  analysis direcly relating ET response to general climate and plant
+  physiological terms, rather than PFT-mean analysis. For example, the
+  ``scaling term'' analysis would be presented in terms of generic
+  changes in plant height and temperature, and the ``sign term''
+  analysis would be framed by generic changes to g$_1$ and uWUE, as
+  informed by previous literature.
 \item This is the most extreme option in terms of modifying the
@@
   curve being concave up, given an assumption of a square-root VPD
-  dependence of plant response. This result is independent of any
-  assumptions of uWUE and g$_1$, and to our knowledge it is first
-  derived curve of ecosystem response to VPD (see Section 3.7). It
-  also highlights the importance of discerning the exact exponent of
-  VPD dependence, as this alters the fundamental nature and shape of
-  the curve. Essentially, the manuscript would become just our
-  motivation and derivation, followed by Section 3.7 exploring the
-  consequences of the derivation for the shape of the VPD curve. This
-  alteration of the manuscript would likely result in more of a
-  technical note type paper, and we are hesitant to do this because we
-  think there is still a lot of useful information obtained by tying
-  our results to real-world scenarios (as in Options 1 and 2 above).
+  dependence of ET. This result is independent of any assumptions of
+  uWUE and g$_1$, and to our knowledge it is first derived curve of
+  ecosystem response to VPD (see Section 3.7). It also highlights the
+  importance of discerning the exact exponent of VPD dependence, as
+  this alters the fundamental nature and shape of the
+  curve. Essentially, the manuscript would become just our motivation
+  and derivation, followed by Section 3.7 exploring the consequences
+  of the derivation for the shape of the VPD curve. This alteration of
+  the manuscript would likely result in more of a technical note-type
+  paper, and we are hesitant to do this because we think there is
+  still a lot of useful information obtained by tying our results to
+  real-world scenarios (as in Options 1 and 2 above).
 \end{enumerate}
@@
 technical details of our proposed changes to the manuscript, which are
-relevant to representing uncertainty and spatiotemporal variabiliy in
+relevant to representing uncertainty and spatiotemporal variability in
 uWUE \textit{and} g$_1$. In the original manuscript we used a single
-$\sigma$ term to represent this variability. We did this because there
-was changes in uWUE and g$_1$ induce a very similiar change in the ET
+$\sigma$ term to represent this variability. We did this because
+changes in uWUE and g$_1$ induce a very similar change in the ET
 solution, which made solving for independent $\sigma_{uWUE}$ and
-$\sigma_{g1}$ terms intractable, and represnting variability in both
-g$_1$ and uWUE difficult. We decided to hold g$_1$ fixed within a PFT
-for a two reasons: 1) ESMs and GCMs generally hold g$_1$ fixed, and 2)
-letting uWUE vary seemed more appropraite to represent specifically
-soil water variations on stomatal conductance, as $uWUE$ modifies
-stomatal conductance analgalously to how soil moisture factors modify
-a maximum stomatal conductance in land surface models (e.g. it is a
-multiplicative factor on the entire stomatal conductance
-term). Because uWUE and g$_1$ induce similar changes in ET, at least
-qualitatively we think that having a single $\sigma$ can at least
-represent some of the variability in both $uWUE$ and g$_1$. However,
-formally there is no explicit variability in the g$_1$ term. The
-discussion in this review rightly points out that based on previous
-results we do expect some variability in g$_1$ as well. So, an
-unresolved question is how important is it to the analysis and its
-interpretation to include an explicit g$_1$ variability term in
-addition to the exisiting $\sigma$ term. I'm including these comments
-to hopefully stimulate some discussion. I could imagine some timescale
-based approaches where we might be able to account for both the g$_1$
-and uWUE variability, for example by making assumptions over what time
-scale each quantity is fixed, and fitting based on that (e.g. g$_1$ is
-fixed for a given season, and uWUE is fixed for a given day). This
-approach might allow a tractable solution, and also could help us
-filter error and observational noise from ``true''  plant
-physiological and climatic variability in uWUE and g$_1$. The cost of
-all of this is a significant increase in the analysis and content of
-the paper, as well as some increased opacity to the methods.
+$\sigma_{g1}$ terms intractable at a given time and place, and
+representing variability in both g$_1$ and uWUE difficult. We decided
+to hold g$_1$ fixed within a PFT for a two reasons: 1) ESMs and GCMs
+generally hold g$_1$ fixed, and 2) letting uWUE vary seemed more
+appropriate to represent specifically soil water variations on
+stomatal conductance, as $uWUE$ modifies stomatal conductance
+analogously to how soil moisture factors modify a maximum stomatal
+conductance in land surface models (e.g. it is a multiplicative factor
+on the entire stomatal conductance term). Because uWUE and g$_1$
+induce similar changes in ET, at least qualitatively we think that
+having a single $\sigma$ can represent some of the variability in both
+$uWUE$ and g$_1$. However, formally there is no explicit variability
+in the g$_1$ term. The discussion in this review rightly points out
+that based on previous results we do expect some variability in g$_1$
+as well. So, an unresolved question is how important is it to the
+analysis and its interpretation to include an explicit g$_1$
+variability term in addition to the existing $\sigma$ term. We are
+including these comments to hopefully stimulate some discussion. We
+could imagine some timescale based approaches where we might be able
+to account for both the g$_1$ and uWUE variability, for example by
+making assumptions over what time scale each quantity is fixed, and
+fitting based on that (e.g. g$_1$ is fixed for a given season, and
+uWUE is fixed for a given day). This approach might allow a tractable
+solution, and also could help us filter error and observational noise
+from ``true'' plant physiological and climatic variability in uWUE and
+g$_1$. The cost of all of this is a significant increase in the
+analysis and content of the paper, as well as some increased opacity
+to the methods. Comments are welcome.
 
-Thank you again for the thoughtfull review. We hope it stimulates
+Thank you again for the thoughtful review. We hope it stimulates
 further discussion.

### 2018-12-17 doc/paper/hess-revision/review-1-pierre-comments.tex
added pierre's revision comments

--- 2018-12-17 doc/paper/hess-revision/review-1.tex
@@
 Just a quick general comment before addressing your specific
-comments. By including a lot of content on PFT-focused analysis we
-failed to communicate the major goal and scope of our project: we are
+comments. By including a PFT-focused analysis we did not fully communicate the major goal and scope of our project: we are
 trying to characterize the response of ET to VPD, with all other
@@
 We agree that the uncertainties are large both within a PFT and across
-a PFT. Looking back, we believe that some of our language
+PFTs. Looking back, we believe that some of our language
 communicating \cite{Lin_2018}, \cite{Medlyn_2017}, and
@@
 remove this language. This misinterpretation should not have made the
-manuscript; we actually included \cite{Zhou_2015}'s results in Table 2
+manuscript SOMETHING MISSING HERE; we actually included \cite{Zhou_2015}'s results in Table 2
 to be transparent about the within-PFT uncertainties. g$_1$ values
@@
 
-Regarding how to interpret g$_1$ and uWUE: we think the most best way
+Regarding how to interpret g$_1$ and uWUE: we think the best way
 is with established physical relations from \cite{MEDLYN_2011}
@@
 approximation is used in many earth system models, so while it
-introduces considerable uncertainty, it also makes our framework
+introduces uncertainty, it also makes our framework
 useful for interpreting modeled vegetation response in ESMs and GCMs.
@@
 can also vary due to soil moisture condition; the approximation we
-need to make to answer our research question is that they are fixed
+need to answer our research question is that they are fixed
 with respect to a VPD perturbation. In this sense, we do not view the
@@
 
-When writing this manuscript, there was definitely some internal
-tension and debate about how to frame the analysis and results. We
+When writing this manuscript, there was definitely some discussion as to how to best frame the analysis and results. We
 could either focus on PFT-oriented results as previous literature has
@@
 
-In order to fix the manuscript and our communication of the answer to
+In order to improve the manuscript and our communication of the answer to
 the question ``When does VPD drive or reduce ET?'', we see a few
@@
   variability in ET response. Sections 3.1 - 3.2 would be replaced by
-  analysis direcly relating ET response to general climate and plant
+  analysis directly relating ET response to general climate and plant
   physiological terms, rather than PFT-mean analysis. For example, the

### 2018-12-18 doc/paper/hess-revision/review-1.tex
better formatting on comment

@@
+\documentclass[12pt]{article}
+% \usepackage{amsmath}
+% \usepackage{amssymb}
+% \usepackage{bm}
+% \usepackage{graphicx}
+% \usepackage{epstopdf}
+% \usepackage{xcolor}
+% \DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `basename #1 .tif`.png}
+% \usepackage{color}
+% \pagestyle{plain}
+
+% \usepackage[
+% 	backend=biber,
+% 	url=false,
+% 	isbn=false,
+% 	doi=false,
+%         eprint=false,
+%         bibstyle=authoryear,
+%         style=authoryear,
+%         % style=numeric-comp,
+% 	% uniquename=false,
+% 	% uniquelist=false,
+% 	% firstinits=true,
+%  	natbib=true,
+% 	sorting=none,
+%         ]{biblatex}
+
+% \addbibresource{./references.bib}
+
 \RequirePackage{natbib}
 \bibliographystyle{agufull08}
+
+% Bib aliases
+\makeatletter
+\def\@citex[#1]#2{\leavevmode
+  \let\@citea\@empty
+  \@cite{\@for\@citeb:=#2\do
+    {\@citea\def\@citea{,\penalty\@m\ }%
+\edef\magic##1{\let##1\expandafter\noexpand\csname bibalias@\@citeb\endcsname}%
+\magic\tmp \ifx\tmp\relax\else \let\@citeb\tmp\fi
+     \edef\@citeb{\expandafter\@firstofone\@citeb\@empty}%
+     \if@filesw\immediate\write\@auxout{\string\citation{\@citeb}}\fi
+     \@ifundefined{b@\@citeb}{\hbox{\reset@font\bfseries ?}%
+       \G@refundefinedtrue
+       \@latex@warning
+         {Citation `\@citeb' on page \thepage \space undefined}}%
+       {\@cite@ofmt{\csname b@\@citeb\endcsname}}}}{#1}}
+\def\bibalias#1#2{\expandafter\def\csname bibalias@#1\endcsname{#2}}
+\makeatother
+
+\title{Response to RC1 of ``When does vapor pressure deficit drive or
+  reduce evapotranspiration?''}
+
+\author{Adam Massmann, Pierre Gentine, and Changjie Lin}
+
+\date{December 17th, 2018}
+
+\begin{document}
 
@@
 Just a quick general comment before addressing your specific
-comments. By including a lot of content on PFT-focused analysis we
-failed to communicate the major goal and scope of our project: we are
-trying to characterize the response of ET to VPD, with all other
-environment variables held fixed. To accomplish this, we need to
-formulate an explicit function of ET in terms of environmental
-variables and parameters, where any parameters can be approximated as
-constant with regards to some [arbitrary] VPD perturbation
-scenario. We will elaborate more below on why we used PFT-focused
-scenarios for much of our analysis. However, the goal of our
-manuscript was much more simple and fundamental: at a given place or
-time, if you introduce a perturbation to VPD, what is the immediate ET
-response (e.g. positive or negative)? Answering this question does not
-require the much stronger assertion that any parameters must be
-invariant within a given PFT.
+comments. By including a PFT-focused analysis we did not fully
+communicate the major goal and scope of our project: we are trying to
+characterize the response of ET to VPD, with all other environment
+variables held fixed. To accomplish this, we need to formulate an
+explicit function of ET in terms of environmental variables and
+parameters, where any parameters can be approximated as constant with
+regards to some [arbitrary] VPD perturbation scenario. We will
+elaborate more below on why we used PFT-focused scenarios for much of
+our analysis. However, the goal of our manuscript was much more simple
+and fundamental: at a given place or time, if you introduce a
+perturbation to VPD, what is the immediate ET response (e.g. positive
+or negative)? Answering this question does not require the much
+stronger assertion that any parameters must be invariant within a
+given PFT.
 
@@
 inter-PFT variability, but clearly this is not the case. We will
-remove this language. This misinterpretation should not have made the
-manuscript; we actually included \cite{Zhou_2015}'s results in Table 2
-to be transparent about the within-PFT uncertainties. g$_1$ values
-also exhibit considerable uncertainty, and again, while the relative
-rankings of these values may change, they are all within observed
-ranges in \cite{Medlyn_2017}. We did not include these in Table 2
-because \cite{Medlyn_2017} does not provide the numerical values;
-however, we think it would be useful to provide estimates of these
-ranges from \cite{Medlyn_2017}'s figures in our Table 2.
+remove this language. This misinterpretation should not have been in
+the manuscript and actually contradicts other manuscript content; for
+example, we included \cite{Zhou_2015}'s results on uWUE in Table 2
+explicitly to be transparent about the within-PFT uncertainties. g$_1$
+values also exhibit considerable uncertainty, and again, while the
+relative rankings of these values may change, they are all within
+observed ranges in \cite{Medlyn_2017}. We did not include these in
+Table 2 because \cite{Medlyn_2017} does not provide the numerical
+values; however, we think it would be useful to provide estimates of
+these ranges from \cite{Medlyn_2017}'s figures in our Table 2.
 
-Regarding how to interpret g$_1$ and uWUE: we think the most best way
+Regarding how to interpret g$_1$ and uWUE: we think the best way
 is with established physical relations from \cite{MEDLYN_2011}
@@
 approximation is used in many earth system models, so while it
-introduces considerable uncertainty, it also makes our framework
-useful for interpreting modeled vegetation response in ESMs and GCMs.
+introduces uncertainty, it also makes our framework useful for
+interpreting modeled vegetation response in ESMs and GCMs.
 
@@
 it was framed in previous studies in the field, and specifically the
-studies used in our derivation \citep{Zhou_2014, Zhou_2015,
+studies used in our derivation \cite{Zhou_2014, Zhou_2015,
   Medlyn_2017}. We were originally thinking the PFT-focused analysis
@@
 making an approximation that we can parameterize the effect of the
-ecoystem's state on VPD response with g$_1$ and uWUE. These quantities
+ecosystem's state on VPD response with g$_1$ and uWUE. These quantities
 can also vary due to soil moisture condition; the approximation we
-need to make to answer our research question is that they are fixed
+need to answer our research question is that they are fixed
 with respect to a VPD perturbation. In this sense, we do not view the
@@
 When writing this manuscript, there was definitely some internal
-tension and debate about how to frame the analysis and results. We
-could either focus on PFT-oriented results as previous literature has
-done (e.g. holding plant physiology fixed within a given PFT), or
-allow plant physiology terms to vary through time and space and look
-at the distribution of ET response to VPD (see Section 3.5). After
-re-examining the manuscript both after some time away from the
-problem and in light of your comments we think a strong argument could
-be made that we made the wrong choice with respect to this focus. It
-may have made more sense to focus our analysis on the ET response more
-generally across space and time as ecosystem-scale plant physiology
-varies in response to climate and soil moisture.
+tension and debate about how to best frame the analysis and
+results. We could either focus on PFT-oriented results as previous
+literature has done (e.g. holding plant physiology fixed within a
+given PFT), or allow plant physiology terms to vary through time and
+space and look at the distribution of ET response to VPD (see Section
+3.5). After re-examining the manuscript both after some time away from
+the problem and in light of your comments we think a strong argument
+could be made that we made the wrong choice with respect to this
+focus. It may have made more sense to focus our analysis on the ET
+response more generally across space and time as ecosystem-scale plant
+physiology varies in response to climate and soil moisture.
 
@@
 
+% \printbibliography
 \bibliography{references.bib}
+
+\end{document}

### 2018-12-18 doc/paper/hess-revision/review-1.tex
RC1 reponse submitted to HESS

@@
 communicating \cite{Lin_2018}, \cite{Medlyn_2017}, and
-\cite{Zhou_2014}'s results was misleading, and this was exacerbated by
+\cite{Zhou_2015}'s results was misleading, and this was exacerbated by
 our focus on PFT-analysis. We will remove that language from the
@@
 robust approximation for answering our research question (see
-discussion earlier in this comment). Additionally, a constant g$_1$
+discussion in the introduction of this comment). Additionally, a constant g$_1$
 approximation is used in many earth system models, so while it
@@
 
-\subsection{Attribution to physiological responses}
+\section{Attribution to physiological responses}
 
@@
 it was framed in previous studies in the field, and specifically the
-studies used in our derivation \cite{Zhou_2014, Zhou_2015,
+studies used in our derivation \citep{Zhou_2014, Zhou_2015,
   Medlyn_2017}. We were originally thinking the PFT-focused analysis
@@
 is hard to believe. Phenotypic variation and adaptation within a given
-\textit{species} can be considerable (i.e. effecting $\lambda$, g$_1$, and
-uWUE), so it would be hard to say that anything would be constant
+\textit{species} can be considerable (i.e. effecting $\lambda$, g$_1$,
+and uWUE), so it would be hard to say that anything would be constant
 within a PFT made up of many different diverse species. Both
@@
 climate. In this sense, the distribution and evolution of local
-climate is a strong control on the local ecosystem structure and
-physiology of a given ecosystem.
+climate is a strong control on the local structure and physiology of a
+given ecosystem.
 
@@
 
-\begin{enumerate}
-\item We include the discussion presented in this review and response,
-  and include language explicitly stating that the purpose of the PFT
-  analysis is to provide connections to other PFT-constant analyses
-  and models (e.g. ESMs and GCMs). We will add extensive language on
-  the uncertainty and weaknesses of this approach, and reframe our
-  conclusions to reflect his uncertainty.
-\item We instead alter our analysis to look at how VPD response varies
-  with climate and general plant physiological variability, instead of
-  focussing only on PFT analysis that poorly captures all of the observed
-  variability in ET response. Sections 3.1 - 3.2 would be replaced by
-  analysis directly relating ET response to general climate and plant
-  physiological terms, rather than PFT-mean analysis. For example, the
-  ``scaling term'' analysis would be presented in terms of generic
-  changes in plant height and temperature, and the ``sign term''
-  analysis would be framed by generic changes to g$_1$ and uWUE, as
-  informed by previous literature.
-\item This is the most extreme option in terms of modifying the
-  manuscript. We significantly alter the manuscript and instead focus
-  on just the general shape of the ET-VPD curve (with environmental
-  variables held fixed). This was not discussed in the review, but one
-  of our most noteworthy results is on the general shape of the ET-VPD
-  curve being concave up, given an assumption of a square-root VPD
-  dependence of ET. This result is independent of any assumptions of
-  uWUE and g$_1$, and to our knowledge it is first derived curve of
-  ecosystem response to VPD (see Section 3.7). It also highlights the
-  importance of discerning the exact exponent of VPD dependence, as
-  this alters the fundamental nature and shape of the
-  curve. Essentially, the manuscript would become just our motivation
-  and derivation, followed by Section 3.7 exploring the consequences
-  of the derivation for the shape of the VPD curve. This alteration of
-  the manuscript would likely result in more of a technical note-type
-  paper, and we are hesitant to do this because we think there is
-  still a lot of useful information obtained by tying our results to
-  real-world scenarios (as in Options 1 and 2 above).
-\end{enumerate}
+\begin{itemize}
+\item \textbf{Option 1:} We include the discussion presented in this
+  review and response, and include language explicitly stating that
+  the purpose of the PFT analysis is to provide connections to other
+  PFT-constant analyses and models (e.g. ESMs and GCMs). We will add
+  extensive language on the uncertainty and weaknesses of this
+  approach, and reframe our conclusions to reflect his uncertainty.
+\item \textbf{Option 2:} We instead alter our analysis to look at how
+  VPD response varies with climate and general plant physiological
+  variability, instead of focusing only on PFT analysis that poorly
+  captures all of the observed variability in ET response. Sections
+  3.1 - 3.2 would be replaced by analysis directly relating ET
+  response to general climate and plant physiological terms, rather
+  than PFT-mean analysis. For example, the ``scaling term'' analysis
+  would be presented in terms of generic changes in plant height and
+  temperature, and the ``sign term'' analysis would be framed by
+  generic changes to g$_1$ and uWUE, as informed by previous
+  literature.
+\item \textbf{Option 3:} This is the most extreme option in terms of
+  modifying the manuscript. We significantly alter the manuscript and
+  instead focus on just the general shape of the ET-VPD curve (with
+  environmental variables held fixed). This was not discussed in the
+  review, but one of our most noteworthy results is on the general
+  shape of the ET-VPD curve being concave up, given an assumption of a
+  square-root VPD dependence of ET. This result is independent of any
+  assumptions of uWUE and g$_1$, and to our knowledge it is first
+  derived curve of ecosystem response to VPD (see Section 3.7). It
+  also highlights the importance of discerning the exact exponent of
+  VPD dependence, as this alters the fundamental nature and shape of
+  the curve. Essentially, the manuscript would become just our
+  motivation and derivation, followed by Section 3.7 exploring the
+  consequences of the derivation for the shape of the VPD curve. This
+  alteration of the manuscript would likely result in more of a
+  technical note-type paper, and we are hesitant to do this because we
+  think there is still a lot of useful information obtained by tying
+  our results to real-world scenarios (as in Options 1 and 2 above).
+\end{itemize}
 
@@
 appropriate to represent specifically soil water variations on
-stomatal conductance, as $uWUE$ modifies stomatal conductance
+stomatal conductance, as uWUE modifies stomatal conductance
 analogously to how soil moisture factors modify a maximum stomatal
@@
 variability term in addition to the existing $\sigma$ term. We are
-including these comments to hopefully stimulate some discussion. We
-could imagine some timescale based approaches where we might be able
-to account for both the g$_1$ and uWUE variability, for example by
-making assumptions over what time scale each quantity is fixed, and
-fitting based on that (e.g. g$_1$ is fixed for a given season, and
-uWUE is fixed for a given day). This approach might allow a tractable
-solution, and also could help us filter error and observational noise
-from ``true'' plant physiological and climatic variability in uWUE and
-g$_1$. The cost of all of this is a significant increase in the
-analysis and content of the paper, as well as some increased opacity
-to the methods. Comments are welcome.
+including these comments to hopefully stimulate some discussion on the
+importance of explicitly representing g$_1$ variability, given the
+difficulties of doing so within our framework. However, we could
+imagine some timescale based approaches where we might be able to
+account for both the g$_1$ and uWUE variability, for example by making
+assumptions over what time scale each quantity is fixed, and fitting
+based on that (e.g. g$_1$ is fixed for a given season, and uWUE is
+fixed for a given day). This approach might allow a tractable
+solution, and also could help us filter model error and observational
+noise from ``true'' plant physiological and climatic variability in
+uWUE and g$_1$. The cost of all of this is a significant increase in
+the analysis and content of the paper, as well as some increased
+opacity to the methods. Comments are welcome.
 
@@
 % \printbibliography
-\bibliography{references.bib}
+\bibliography{references}
 

### 2018-12-18 doc/paper/hess-revision/review-2.tex
progress on review 2

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{xcolor}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `basename #1 .tif`.png}
\usepackage{color}
\pagestyle{plain}

\RequirePackage{natbib}
\bibliographystyle{agufull08}

% Bib aliases
\makeatletter
\def\@citex[#1]#2{\leavevmode
  \let\@citea\@empty
  \@cite{\@for\@citeb:=#2\do
    {\@citea\def\@citea{,\penalty\@m\ }%
\edef\magic##1{\let##1\expandafter\noexpand\csname bibalias@\@citeb\endcsname}%
\magic\tmp \ifx\tmp\relax\else \let\@citeb\tmp\fi
     \edef\@citeb{\expandafter\@firstofone\@citeb\@empty}%
     \if@filesw\immediate\write\@auxout{\string\citation{\@citeb}}\fi
     \@ifundefined{b@\@citeb}{\hbox{\reset@font\bfseries ?}%
       \G@refundefinedtrue
       \@latex@warning
         {Citation `\@citeb' on page \thepage \space undefined}}%
       {\@cite@ofmt{\csname b@\@citeb\endcsname}}}}{#1}}
\def\bibalias#1#2{\expandafter\def\csname bibalias@#1\endcsname{#2}}
\makeatother

\begin{document}

Thank you for a very thoughtful review; as in our response to RC1, we
feel very lucky to have received such quality reviews.

We read through your major and minor comments, and do not anticipate
issues reconciling your minor comments for the final response after
the discussion period. We will focus out response on your major
comments, with the goal of opening the door to further discussion.

Also please note our response to RC1, as some of the discussion there
is relelvant to this response; particularly discussion on PFT-scale
analysis \url{https://editor.copernicus.org/index.php/hess-2018-553-AC1.pdf?_mdl=msover_md&_jrl=13&_lcm=oc108lcm109w&_acm=get_comm_file&_ms=72556&c=153745&salt=17166479111051651323}.

\section{Major comment (i)}

One issue with analyses attempting to diagnose the ET response to VPD
with pure data (i.e. using binned boxplots, as you suggest), is that
ET varies strongly with both aerodynamic resistance and
radiation. Simply binning by VPD or by site includes too many
confounding factors to really diagnose what the attirbutable response
is to VPD. This issue is really what motivated our analysis and makes
it unique; by building an analytical framework based on confirmed
results in the literatre (uWUE, Medlyn model, Penman-Monteith), we are
able to formally deduce what the actual ET response is, with other
environmental factors held fixed. This ``other environmental factors
held fixed'' component is what is, in our mind, nearly impossible to
deduce from pure data, because the environment is always changing and
direct analogues are either impossible to find or result in too small
of a sample size to form a meaningful conclusion. However, while we
cannot assess our formulation of the derivative with respect to VPD,
we can assess the actual model of ET, which includes all
approximations we introudced (specifically uWUE). As a part of our
analysis test suite we generated a figure of this comparison and
reproduce it here (Figure \ref{rmse}). Apologies this is not a
publication quality figure, but we wanted to get this response out
sooner rather than later to stimulate discussion, especially as we are
already behind because of the AGU Fall Meeting. This comparison is
fair in the sense that by using uWUE we introduce an extra free
parameter, so in the original PM model we introduce an extra free
parameter modifying g$_s$ and fit it also for each given PFT. The
introduction of the uWUE approximation does degrade the quality of the
ET model relative to the original PM model. However, we expect this,
as using uWUE is a simplification over using GPP directly, and may
break down in extreme, limiting cases like VPD $\to$ 0. As
a purpose-built model for assessing the ET response, we think the uWUE
approach is justified. Essentially we are saying that to the degree we
trust arithmetic, uWUE, Penman-Monteith, and the Medlyn stomatal
conductance model, our evaluation of the derivative is robust.

\begin{figure}[h]
  \centering \includegraphics{./rmse_bias.png}
  \caption{Mean bias and root mean squared error (RMSE) for three
    different ET estimates, relative to FLUXNET-2015 observations:
    original (original Penman-Monteith formulation, e.g. Equation 1 in
    the manuscript, with Equation 4 providing g$_s$; uWUE (Equation 8
    in the manuscript); iWUE (as in Equation 8, but if we had used
    iWUE to remove GPP dependence rather than uWUE).}
  \label{rmse}
\end{figure}

However, the discussion so far assumes that the variables in our model
(Equation 8) do not change directly with VPD and we are free to
evaluate the derivative. You bring up a very good example that we do
not explictly consider: a change in VPD will induce a change in
surface temperature, which will alter the net radiation term. We will
definitely add language addressing this oversight. Ideally, we would
incorporate that effect directly in Equation 8. We briefly tried to do
this in time for this response, but difficulties deriving a diagnostic
equation for temperature sabatoged the effort, and we do not want to
delay the response further. We will continue to work on this for the
final response phase. In the meantime, I think we can make some
hand-wavy arguments in the interest of discussion as to what the
effect would be. In the case where the aerodynamic and physiological
ET response (which the manuscript does consider) increases with VPD,
we would expect a decrease in surface temperature. This would induce
an increase in net radiation, as OLR and ground heat flux
decrease. This increase in net radation would induce a further
increase in ET response. Granted, the net radiation perturbation would
damp the decreased temperature response to the
aerodynamic/physiological terms, but all of the signs of the change
would stay the same. The converse is true for a decrease in ET with
respect to VPD: surface temperature increases, OLR and ground heat
flux increase, net radiation decreases, ET decreases. In this way, we
believe the effects of VPD on radiation do not change the nature of
the sign of the response, but could amplify the magnitude. Comments on
this are incouraged, and we will work to include a more quantitative or
rigorous analysis in the final response. Sometimes it is easy to trick
onself with these types of qualitative arguments, particularly with
complicated systems, so hopefully we have not presented faulty logic
in our rush to make this response public.

\end{document}

### 2018-12-19 doc/paper/hess-revision/review-2.tex
done with comment (ii)

@@
 
+\graphicspath{../../../etc/plots/climate_et/test_plots}
 \RequirePackage{natbib}
@@
   \centering \includegraphics{./rmse_bias.png}
-  \caption{Mean bias and root mean squared error (RMSE) for three
-    different ET estimates, relative to FLUXNET-2015 observations:
-    original (original Penman-Monteith formulation, e.g. Equation 1 in
-    the manuscript, with Equation 4 providing g$_s$; uWUE (Equation 8
-    in the manuscript); iWUE (as in Equation 8, but if we had used
-    iWUE to remove GPP dependence rather than uWUE).}
+  \caption{Mean bias (absolute value) and root mean squared error
+    (RMSE) for three different ET estimates, relative to FLUXNET-2015
+    observations: original (original Penman-Monteith formulation,
+    e.g. Equation 1 in the manuscript, with Equation 4 providing
+    g$_s$; uWUE (Equation 8 in the manuscript); iWUE (as in Equation
+    8, but if we had used iWUE to remove GPP dependence rather than
+    uWUE).}
   \label{rmse}
@@
 (Equation 8) do not change directly with VPD and we are free to
-evaluate the derivative. You bring up a very good example that we do
-not explictly consider: a change in VPD will induce a change in
-surface temperature, which will alter the net radiation term. We will
+evaluate the derivative. You bring up a very good example that we did
+not consider: a change in VPD will induce a change in net radiation
+through surface tmperature that PM does not consider. We will
 definitely add language addressing this oversight. Ideally, we would
@@
 delay the response further. We will continue to work on this for the
-final response phase. In the meantime, I think we can make some
-hand-wavy arguments in the interest of discussion as to what the
+final response phase. In the meantime, we think we can make some
+logic-based arguments in the interest of discussion as to what the
 effect would be. In the case where the aerodynamic and physiological
@@
 the sign of the response, but could amplify the magnitude. Comments on
-this are incouraged, and we will work to include a more quantitative or
-rigorous analysis in the final response. Sometimes it is easy to trick
-onself with these types of qualitative arguments, particularly with
-complicated systems, so hopefully we have not presented faulty logic
-in our rush to make this response public.
+this are incouraged, and we will work to include a more quantitative
+and rigorous analysis in the final response. Sometimes it is easy to
+trick onself with these types of qualitative arguments, particularly
+with complicated systems, so hopefully we have not presented faulty
+logic in our rush to make this response public.
+
+\section{Major comment (ii)}
+
+Thank you for having an open mind about the concave up ET-VPD curve
+result and providing clear explanations of concerns about this result
+based on literature. Given the apparent controversy, we will
+definitely add more language connecting previous results, particularly
+at the leaf scale, with the presented results. Here we will informally
+elaborate in the interest of discussion.
+
+One of the primary differences between ecosystem scale assessments of
+ET (specifically using Penman Monteith) and experimental and modeling
+results of leaf scale response to VPD is that by taking an energy
+balance approach Penman Monteith accounts for changes in $e_{sat}(T)$
+at the surface induced by ET. To demonstrate this, compare Figure
+\ref{leaf} and Figure \ref{PM}. In Figure \ref{leaf} we apply a naive
+leaf scale approach to calculate ecosystem ET without the energy
+balance consideration introduced by PM. In this figure, the curve
+shows a concave downward shape, in direct analogy to leaf scale
+T. Once the effects of evaporation's thermodyanmics are included in
+the energy balance, the curve's shape is no longer concave
+downward. How these conceptual results translate to leaf scale
+experiments depends on the experimental setup
+\citep[e.g.,]{Rawson1988, Turner1984, Mott2013}; however generally
+chamber based leaf scale experiments do not preserve the energy
+balance relationships we expect for a free surface in a natural
+environment (sometimes intentionally by design). Additionally, we
+would like to note that a concave up result is not necesarily
+inconsistent with the statement ``transpiration at the leaf scale is
+mostly positive'' with VPD. Especially for plants in well watered
+environments like crops, the manuscript's ecosystem scale shape
+(Figure \ref{uWUE}) would not be out of place on the figures in
+\cite{Rawson1977} and \cite{Mott2007} for the range of VPD considered
+(Figure \ref{uWUE} compared to Fig. 2 in \cite{Rawson1977}, also
+Figure 7 in \cite{Mott2007}). However, the change in relative
+magnitudes are not similiar, and this effect can be traced partially
+to the effect of decreasing surface $e_{sat}(T)$ in response to
+increasing evapotranspiration. While not direcly considered here (or
+in the manuscript) we aslo think it is important to consider how plant
+physiology in the natural environment could vary from plants grown in
+well-watered conditions in experiments. It is not inconceivable that
+phenotypic variability and adaptation in response to water limitation
+could result in plants with different responses to those grown in a
+lab.
+
+The previous paragraph discussed the change in frame of reference
+induced by using PM to assess ecoystem scale ET rather than directly
+applying leaf scale logic in controlled experiments. However, our
+uWUE-based PM framework additionally accounts for changes in
+photosythesis induced by VPD. Discussion in \cite{Damour2010} (``The
+issue of co-regulation of g$_s$ and A$_{net}$'') provides some
+interesting background both on the general weaknesses of current
+stomatal theory, and specifically on the possible importance of resolving the
+effects of water stress on $A_{net}$ in stomatal conudctance. Given
+that the current state of the art in physically-derived theory has not
+converged on the proper way to account for changes in $A_{net}$ and
+$g_s$, we believe using the validated uWUE empiracle results
+\citep[e.g. Figure 3]{Zhou_2014} is a good approach to develop an
+estimate of ET response to VPD accounting for changes in
+photosythesis. This method of using an empricle result in the absence
+of a complete physical theory has been used often in science
+(e.g. turbulence) to increase our knowlege about a given systems'
+behavior, and we think it is well motivated here. However, that said
+it is important to add language and consideration on how introducing
+this empricism could alter interpretation of our results. In
+particular, while uWUE was developed using leaf scale photosythesis
+theory, it was validated with observations of a fully coupled
+land-atmosphere system. It captures a relatively consistent observed
+relationship between GPP, ET, and VPD as they vary. There is a chance
+that some of this relatioship could be due to feedbacks between the
+land and the atmosphere, even though the relationship was derived by
+leaf scale theory without feedbacks (e.g., if the assumptions behind
+the uWUE theory introduced ``real'' errors that were compensated for
+by feedbacks or ecosystem-scale physics in the observations).  Were
+this the case, then we introduced a conceptual impurity to the
+statement ``we are examining the one way response of ET to VPD.''
+But, given the theoretical motivations for uWUE we think the portion
+the relationship related to land-atmosphere feedbacks are likely
+small, if they exist at all, and regardless this approach is still an
+improvement over ignoring the documented effect of water stress on net
+photosythesis.
+
+Given the continued gaps in our ability to represent stomatal behavior
+\citep(Damour2010}, particulary for ecosystems in natural
+environments, we think it is hard to rule out our results. In the end,
+it comes down to the degree to which we trust our three tools: the
+Medlyn model, Penman-Monteith, and uWUE. To us these offered the most
+compete analytical approach to the problem at present. However, it's
+important to emphasize that we are not saying this is certainly
+correct, and to do so would be irresponsible given the continued
+development of stomatal theory and how it translates to the ecoystem
+scale. What is important is to recognize that something as fundamental
+as the shape of the ET-VPD curve is still very much an area of open
+research, and sensitive to the representation of plant physiology (of
+which there are many possibilities) in a given model or framework. We
+thank the reviewer for having an open mind and giving us a chance to
+elaborate; others in the community have rejected this result off-hand
+(or even ridiculed it), which reinforces for us the importance of
+demonstrating current uncertainty in VPD-ET relationships.
+
+Just a quick couple of final notes:
+
+We think that we miscommunicated with our (mis)use of ``plant physiology''
+throughout the manuscript, and will alter our language in the final
+review phase to be more precise. Many times when we refer to ``plant
+physiology'' we mean any effect (direct or indirect) of an ecosystem
+response, e.g. any behaviour that would not be observed over a wet or
+bare soil surface.
 

### 2018-12-20 doc/paper/hess-revision/review-2.tex
response 2: what I sent to pierre changjie for edits

--- 2018-12-18 doc/paper/hess-revision/review-2.tex
@@
 \usepackage{xcolor}
+\usepackage{hyperref}
 \DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `basename #1 .tif`.png}
@@
 \pagestyle{plain}
+
+\graphicspath{{../../../etc/plots/climate_et/test_plots/}}
 
@@
 Also please note our response to RC1, as some of the discussion there
-is relelvant to this response; particularly discussion on PFT-scale
-analysis \url{https://editor.copernicus.org/index.php/hess-2018-553-AC1.pdf?_mdl=msover_md&_jrl=13&_lcm=oc108lcm109w&_acm=get_comm_file&_ms=72556&c=153745&salt=17166479111051651323}.
+is relevant to this response; particularly discussion on PFT-scale
+analysis and variability of g$_1$ and uWUE
+(\href{https://editor.copernicus.org/index.php/hess-2018-553-AC1.pdf?_mdl=msover_md&_jrl=13&_lcm=oc108lcm109w&_acm=get_comm_file&_ms=72556&c=153745&salt=17166479111051651323}{RC1
+Response}).
 
 \section{Major comment (i)}
-
+\label{i}
 One issue with analyses attempting to diagnose the ET response to VPD
-with pure data (i.e. using binned boxplots, as you suggest), is that
+with pure data (e.g. using binned boxplots, as you suggest), is that
 ET varies strongly with both aerodynamic resistance and
 radiation. Simply binning by VPD or by site includes too many
-confounding factors to really diagnose what the attirbutable response
+confounding factors to really diagnose what the attributable response
 is to VPD. This issue is really what motivated our analysis and makes
 it unique; by building an analytical framework based on confirmed
-results in the literatre (uWUE, Medlyn model, Penman-Monteith), we are
+results in the literature (uWUE, Medlyn model, Penman-Monteith [PM]), we are
 able to formally deduce what the actual ET response is, with other
 environmental factors held fixed. This ``other environmental factors
-held fixed'' component is what is, in our mind, nearly impossible to
+held fixed'' component is, in our mind, nearly impossible to
 deduce from pure data, because the environment is always changing and
@@
 we can assess the actual model of ET, which includes all
-approximations we introudced (specifically uWUE). As a part of our
+approximations we introduced (specifically uWUE). As a part of our
 analysis test suite we generated a figure of this comparison and
@@
 
-\begin{figure}[h]
-  \centering \includegraphics{./rmse_bias.png}
-  \caption{Mean bias and root mean squared error (RMSE) for three
-    different ET estimates, relative to FLUXNET-2015 observations:
-    original (original Penman-Monteith formulation, e.g. Equation 1 in
-    the manuscript, with Equation 4 providing g$_s$; uWUE (Equation 8
-    in the manuscript); iWUE (as in Equation 8, but if we had used
-    iWUE to remove GPP dependence rather than uWUE).}
+\begin{figure}
+  \centering \includegraphics[width=0.75\textwidth]{rmse_bias.png}
+  \caption{Mean bias (absolute value) and root mean squared error
+    (RMSE) for three different ET estimates, relative to FLUXNET-2015
+    observations: original (original Penman-Monteith formulation,
+    e.g. Equation 1 in the manuscript, with Equation 4 providing
+    g$_s$); uWUE (Equation 8 in the manuscript); iWUE (as in Equation
+    8, but if we had used iWUE to remove GPP dependence rather than
+    uWUE).}
   \label{rmse}
@@
 (Equation 8) do not change directly with VPD and we are free to
-evaluate the derivative. You bring up a very good example that we do
-not explictly consider: a change in VPD will induce a change in
-surface temperature, which will alter the net radiation term. We will
+evaluate the derivative. You bring up a very good example that we did
+not consider: a change in VPD will induce a change in net radiation
+through surface temperature that PM does not consider. We will
 definitely add language addressing this oversight. Ideally, we would
@@
 this in time for this response, but difficulties deriving a diagnostic
-equation for temperature sabatoged the effort, and we do not want to
+equation for temperature sabotaged the effort, and we do not want to
 delay the response further. We will continue to work on this for the
-final response phase. In the meantime, I think we can make some
-hand-wavy arguments in the interest of discussion as to what the
+final response phase. In the meantime, we think we can make some
+logic-based arguments in the interest of discussion as to what the
 effect would be. In the case where the aerodynamic and physiological
@@
 an increase in net radiation, as OLR and ground heat flux
-decrease. This increase in net radation would induce a further
+decrease. This increase in net radiation would induce a further
 increase in ET response. Granted, the net radiation perturbation would
@@
 the sign of the response, but could amplify the magnitude. Comments on
-this are incouraged, and we will work to include a more quantitative or
-rigorous analysis in the final response. Sometimes it is easy to trick
-onself with these types of qualitative arguments, particularly with
-complicated systems, so hopefully we have not presented faulty logic
-in our rush to make this response public.
+this are encouraged, and we will work to include a more quantitative
+and rigorous analysis in the final response. Sometimes it is easy to
+trick oneself with these types of qualitative arguments, particularly
+with complicated systems, so hopefully we have not presented faulty
+logic in our rush to make this response public.
+
+\section{Major comment (ii)}
+\label{ii}
+Thank you for having an open mind about the concave up ET-VPD curve
+result and providing clear explanations of concerns about this result
+based on literature. Given the apparent controversy, we will
+definitely add more language connecting previous results, particularly
+at the leaf scale, with the presented results. Here we will informally
+elaborate in the interest of discussion.
+
+One of the primary differences between ecosystem scale assessments of
+ET (specifically using Penman Monteith) and experimental and modeling
+results of leaf scale response to VPD is that by taking an energy
+balance approach Penman Monteith accounts for changes in $e_{sat}(T)$
+at the surface induced by ET. To demonstrate this, compare Figure
+\ref{leaf} and Figure \ref{PM} (``Original Penman Monteith''
+Curve). In Figure \ref{leaf} we apply a conceptual leaf scale
+approach to calculate ecosystem ET without the energy balance
+consideration introduced by PM. In this figure, the curve shows a
+concave downward shape, in direct analogy to leaf scale T. Once the
+effects of evaporation's thermodynamics are included in the energy
+balance, the curve's shape is no longer concave downward (Figure
+\ref{PM}, ``Original Penman Monteith'' curve). How these conceptual
+results translate to leaf scale experiments depends on the
+experimental setup \citep[e.g.,][]{Rawson1977, Turner1984, Mott2013};
+however generally chamber-based leaf scale experiments do not preserve
+the energy balance relationships we expect for a surface in a
+natural environment (sometimes intentionally by design). Additionally,
+we would like to note that a concave up result is not necessarily
+inconsistent with the statement ``transpiration at the leaf scale is
+mostly positive'' with VPD. Especially for plants in well watered
+environments like crops, the manuscript's ecosystem scale shape
+(Figure \ref{PM}) would not be out of place on the figures in
+\cite{Rawson1977} and \cite{Mott2007} for the range of VPD considered
+(Figure \ref{PM} compared to Fig. 2 in \cite{Rawson1977}, also
+Figure 7 in \cite{Mott2007}). However, the change in relative
+magnitudes are not similar, and this effect can be traced partially to
+the effect of decreasing surface $e_{sat}(T)$ in response to
+increasing evapotranspiration. While not directly considered here (or
+in the manuscript) we also think it is important to consider how plant
+physiology in the natural environment could vary from plants grown in
+well-watered conditions in experiments. It is not inconceivable that
+phenotypic variability and adaptation in response to water limitation
+could result in plants with different responses to those grown in a
+lab. We presented crops in our figures because they are the most likely
+to well watered and analogous with laboratory results; however for
+other PFTs the concavity of the VPD-ET curves are more pronounced at
+higher VPD as compared to Figure \ref{PM}.
+
+\begin{figure}
+  \centering \includegraphics[width=1.0\textwidth]{CRO_et_vpd_curve_leaf.png}
+  \caption{The conceptual relationship between ET with VPD using a
+    model analogous to leaf scale models and experiments, evaluated
+    with the study's median ecosystem scale resistances for the crop
+    PFT. This estimation of ET is by definition not physically
+    representative, but the figure is intended as a conceptual
+    description of how leaf-scale theory's ET-VPD relationship deviates
+    from energy balance ecosystem theory (Fig. \ref{PM}). This figure is
+    demonstrative, and not intended to be of publication quality.}
+  \label{leaf}
+\end{figure}
+
+\begin{figure}
+  \centering \includegraphics[width=1.0\textwidth]{CRO_et_vpd_curve_with_pm.png}
+  \caption{The conceptual relationship between ET and VPD using the
+    uWUE derived version of Penman Monteith (Equation 8 in the manuscript)
+    and the original Penman Monteith formulation (Equation 1 in the
+    manuscript), evaluated at the study's median conditions for the
+    crop PFT. uWUE Penman Monteith includes the effects of VPD on
+    photosynthesis. This figure is demonstrative, and not intended to be of
+    publication quality.}
+  \label{PM}
+\end{figure}
+
+The previous paragraph discussed the change in frame of reference
+induced by using PM (and energy balance) to assess ecosystem scale ET
+rather than directly applying leaf scale logic in controlled
+experiments. However, our uWUE-based PM framework additionally
+accounts for changes in photosynthesis induced by VPD (Figure
+\ref{PM}). Discussion in \cite{Damour2010} (``The issue of
+co-regulation of g$_s$ and A$_{net}$'') provides some interesting
+background both on the general weaknesses of current stomatal theory,
+and specifically on the possible importance of resolving the effects
+of water stress on $A_{net}$ in stomatal conductance. Given that the
+current state of the art in physically-derived theory has not
+converged on the proper way to account for changes in $A_{net}$ and
+$g_s$ \citep{Damour2010}, we believe using the validated uWUE
+empiracle results \citep[e.g. Figure 3,][]{Zhou_2014} is a good
+approach to develop an estimate of ET response to VPD accounting for
+changes in photosynthesis. This method of using an empirical result in
+the absence of a complete physical theory has been used often in
+science (e.g. turbulence) to increase our knowlege about a given
+systems' behavior, and we think it is well motivated here. However, it
+is important to add language on how introducing this
+empricism could alter interpretation of our results. In particular,
+while uWUE was developed using leaf scale photosythesis theory, it was
+validated with observations of a fully coupled land-atmosphere
+system. It captures a relatively consistent observed relationship
+between GPP, ET, and VPD as they vary. There is a chance that some of
+this relationship could be due to feedbacks between the land and the
+atmosphere, even though the relationship was derived by leaf scale
+theory without feedbacks (e.g., if the assumptions behind the uWUE
+theory introduced ``real'' errors that were compensated for by
+feedbacks or ecosystem-scale physics in the observations).  Were this
+the case, then we introduced a conceptual impurity to the statement
+``we are examining the one way response of ET to VPD.''  But, given
+the theoretical motivations for uWUE we think the portion of the
+relationship related to land-atmosphere feedbacks are likely small, if
+they exist at all, and regardless this approach is still an
+improvement over ignoring the documented effect of water stress on net
+photosynthesis.
+
+Given the continued gaps in our ability to represent stomatal behavior
+\citep{Damour2010}, particularly for ecosystems in natural
+environments, we think it is hard to rule out our results. In the end,
+it comes down to the degree to which we trust our three tools: the
+Medlyn model, Penman-Monteith, and uWUE. To us these offered the most
+compete analytical approach to the problem at present, and are
+supported in the literature. However, it's important to emphasize that
+we are not saying this is certainly correct, and to do so would be
+irresponsible given the continued development of stomatal theory and
+how it translates to the ecosystem scale. What is important is to
+recognize that something as fundamental as the shape of the ET-VPD
+curve is still very much an area of open research, and sensitive to
+the representation of plant physiology (of which there are many
+possibilities) in a given model or framework. We thank the reviewer
+for having an open mind and giving us a chance to elaborate; others in
+the community have rejected this result off-hand, which reinforces for
+us the importance of demonstrating current uncertainty in VPD-ET
+relationships.
+
+\vspace{0.2in}
+Just a quick final note:
+
+We think that we miscommunicated with our (mis)use of ``plant physiology''
+throughout the manuscript, and will alter our language in the final
+response phase to be more precise. Many times when we refer to ``plant
+physiology'' we mean any effect (direct or indirect) of an ecosystem
+response, e.g. any behavior that would not be observed over a wet or
+bare soil surface.
+
+\section{Major comment (iii)}
+\label{iii}
+Yes, through our manipulation of PM we have replaced stomatal
+conductance with two parameters (g$_1$ and uWUE). However, given that
+any physically reasonable representation of stomatal conductance will
+include more parameters (usually at least some sort of VPD-related
+parameter [g$_1$], and a model for photosynthesis with more
+parameters), we think we are actually reducing the number of unknowns
+in our model for ET. We agree with you that in reality g$_1$ and
+uWUE are spatially variable, and we elaborate in detail in our
+response to Reviewer 1
+(\href{https://editor.copernicus.org/index.php/hess-2018-553-AC1.pdf?_mdl=msover_md&_jrl=13&_lcm=oc108lcm109w&_acm=get_comm_file&_ms=72556&c=153745&salt=1660920846993335212}{RC1
+  Response})
+on why and how it might be better to frame our results more in terms
+of this variability. Also, see our comments in Section \ref{i} on the
+effect of VPD on R$_n$ and the suitability of a binning approach to
+VPD.
+
+We also agree with you that uWUE includes a representation of ecosystem
+response. However, we would argue that this response dependence of
+stomatal conductance is not introduced by uWUE, but is actually an
+issue that exists inherently in any stomatal conductance theory that
+includes a photosynthesis term: photosynthesis depends on the
+physiological response and environmental conditions. Using uWUE is our
+(approximate) solution to this problem, and it works because we are
+able to algebraically manipulate PM to isolate ET after introducing
+uWUE. We do recognize that using uWUE introduces some uncertainty, but
+as previously discussed in this response we believe it is a useful tool
+for representing photosynthesis dependence of stomatal conductance in an
+analytical framework.
+
+\section{Major comment (iv)}
+
+We agree on both counts: use of ``leading order'' is misguided and
+confusing in this context. Also, better separation of theory and
+empirical results would help communication/understanding.
+\vspace{0.3in}
+
+Thank you again for such a thoughtful review. Including these insights
+will really improve the manuscript in the final response phase. In our
+response to Reviewer 1
+(\href{https://editor.copernicus.org/index.php/hess-2018-553-AC1.pdf?_mdl=msover_md&_jrl=13&_lcm=oc108lcm109w&_acm=get_comm_file&_ms=72556&c=153745&salt=1660920846993335212}{RC1
+Response})
+we also laid out some alternatives to our approach for the final
+review phase not explicitly discussed here. If you (or anyone else)
+has any opinions on that please weigh in.
+
+\bibliography{references}
 

### 2018-12-20 src/codebase/test_functions.py

@@ -87,15 +87,15 @@ def plot_et_curve(mean_row, min_row, max_row, index, _df):
   # print('for pft %s, time to subset was %f s, and nobs was %d'\
   #     % (index, (end-start), _df.shape[0]))
   # plt.scatter(_df.vpd, _df.et_obs, s=4.0, c='m')
-  plt.plot(vpd, et, label='uWUE PM')
-  plt.plot(vpd, et_orig, label='Mean GPP')
-  plt.plot(vpd, et_iwue, label='iWUE PM')
+  plt.plot(vpd, et, label='uWUE Penman Monteith')
+  plt.plot(vpd, et_orig, label='Original Penman Monteith')
+  # plt.plot(vpd, et_iwue, label='iWUE PM')
   # plt.plot(vpd, et_bb, label='uWUE BB')
   # plt.plot(vpd, et_iwue_bb, label='iWUE BB')
   plt.legend(loc='best')
   plt.title(index)
-  plt.xlabel('vpd (Pa')
-  plt.ylabel('et (w/m2)')
+  plt.xlabel('VPD (Pa)')
+  plt.ylabel('ET (W/m2)')
   plt.savefig('%s/%s_et_vpd_curve_with_pm.png' % (PLOTDIR, index))
   return
 
@@ -108,8 +108,8 @@ def plot_et_curve_onlyuWUE(mean_row, min_row, max_row, index, _df):
   plt.plot(vpd, et, label='uWUE PM')
   plt.legend(loc='best')
   plt.title(index)
-  plt.xlabel('vpd (Pa')
-  plt.ylabel('et (w/m2)')
+  plt.xlabel('VPD (Pa)')
+  plt.ylabel('ET (W/m2)')
   plt.savefig('%s/%s_et_vpd_curve_uWUE.png' % (PLOTDIR, index))
   return
 
@@ -121,8 +121,8 @@ def plot_et_curve_onlyPM(mean_row, min_row, max_row, index, _df):
   plt.plot(vpd, et_orig, label='Original Penman Monteith')
   plt.legend(loc='best')
   plt.title(index)
-  plt.xlabel('vpd (Pa')
-  plt.ylabel('et (w/m2)')
+  plt.xlabel('VPD (Pa)')
+  plt.ylabel('ET (W/m2)')
   plt.savefig('%s/%s_et_vpd_curve_PM.png' % (PLOTDIR, index))
   return
 
[…]

### 2018-12-20 doc/paper/hess-revision/review-3.tex
response 3, sent off to pierre and changjie

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{xcolor}
\usepackage{hyperref}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `basename #1 .tif`.png}
\usepackage{color}
\pagestyle{plain}

\graphicspath{{../../../etc/plots/climate_et/test_plots/}}

\RequirePackage{natbib}
\bibliographystyle{agufull08}

\begin{document}

Thank you for a very thoughtful review. This third review adds further
to our feeling that we are extremely lucky to have received such
quality reviews.

We agree with your comments and will respond with some general views
to open the door for further discussion in this discussion phase. We
look forward to reconciling all reviews with a revised manuscript and
formal response in the final response phase. Also of relevance to this
review is our \href{https://editor.copernicus.org/index.php/hess-2018-553-AC1.pdf?_mdl=msover_md&_jrl=13&_lcm=oc108lcm109w&_acm=get_comm_file&_ms=72556&c=153745&salt=17166479111051651323}{
  response to RC1}, which we will refer to as needed to avoid
repeating content.

\section{Soil moisture, uWUE and g$_1$}
Environmental conditions, particularly soil moisture, will effect both
uWUE and g$_1$ as you point out. Not accounting for this variability
by using a PFT-focused analysis is problematic. We discuss this in
\href{https://editor.copernicus.org/index.php/hess-2018-553-AC1.pdf?_mdl=msover_md&_jrl=13&_lcm=oc108lcm109w&_acm=get_comm_file&_ms=72556&c=153745&salt=17166479111051651323}{our
  response to RC1}, specifically Sections 2 and 3, and think a strong
argument could be made that we should have framed our analysis in
terms of more general uWUE and g$_1$ variability instead of focusing
primarily on PFT variability. If we let uWUE and g$_1$ vary in space
and time, then we include a measure of soil moisture through its
effect on uWUE and g$_1$. In this case, interpretation of our results
relies on a less restrictive approximation that soil moisture
conditions remain relatively fixed with respect to the VPD
perturbation, which is consistent with the general interpretation of
our results (e.g. we evaluate the ET response to VPD with
other environmental conditions held fixed).

There are also issues with only including uncertainty in uWUE. We
discuss reasons for this approach in our response to
\href{https://editor.copernicus.org/index.php/hess-2018-553-AC1.pdf?_mdl=msover_md&_jrl=13&_lcm=oc108lcm109w&_acm=get_comm_file&_ms=72556&c=153745&salt=17166479111051651323}{RC1}
(specifically the final paragraph), and we think there may be some
tractable approaches to representing uncertainty and variability in
both uWUE and g$_1$ for the final manuscript.

\section{Methods/Introduction}
Adding the clarifications you suggest for the methods section will
greatly improve the manuscript, as well as more thorough discussion in
the introduction about what is known about the transpiration response to
VPD. This is also discussed peripherally in our response to
\href{https://editor.copernicus.org/index.php/hess-2018-553-RC2.pdf?_mdl=msover_md&_jrl=13&_lcm=oc108lcm109w&_acm=get_comm_file&_ms=72556&c=153384&salt=74148443849227084}{RC2}.

\section{Results}
Regarding the presentation of the results, we agree that in its current
form the manuscript can be disorienting. Just discussing the sign of
the derivative, and then discussing the magnitude (while including
what the current manuscript refers to as the ``sign term'') seems to
be a much cleaner approach, possibly also differentiating between
environmental controls versus plant/canopy controls.

In our response to RC1 we discussed a few revision approaches
regarding the results, some of them including significant changes,
which should help both with acknowledging g$_1$ and uWUE variability
and streamlining the results. One example (``Option 2'') is to remove
the PFT focus from the first portion of the results (Section 3.1-3.2;
this could also go in Methods as you suggest), and instead focus
generally on how changes in the environment (e.g. wind speed,
temperature) and land surface terms (e.g. g$_1$, uWUE, canopy height)
influence the ET response. Using a more sophisticated representation
of uncertainty we could then present the distribution of
observation-informed ET responses as a function of the model terms
(g$_a$, uWUE, g$_1$, etc.), and finally tie all this to either PFTs or
specific sites with maps of statistics from the distribution of model
parameters and ET response. While this could add a lot of analysis and
content to the paper, we think in the end it could streamline the
results and their interpretation. Basically we would start with the
idealized ET response as a function of parameters (no uncertainty), to
the observed ET response as a function of parameters (with
uncertainty), to how both the parameters and ET response (with
uncertainty) map to specific sites, years and/or PFTs. Thoughts on
this are welcome.

We also acknowledge that we need to be more precise about some of our
language (e.g. ``leading order behavior'', ``bit more variability''),
and thank you for the other minor comments as well; fixing these will
improve the manuscript.

\end{document}

### 2018-12-23 doc/paper/hess-revision/review-2.tex
going to submit response 2

@@
 issues reconciling your minor comments for the final response after
-the discussion period. We will focus out response on your major
+the discussion period. We will focus our response on your major
 comments, with the goal of opening the door to further discussion.
@@
 with pure data (e.g. using binned boxplots, as you suggest), is that
-ET varies strongly with both aerodynamic resistance and
+ET varies strongly with both aerodynamic conductance and
 radiation. Simply binning by VPD or by site includes too many
@@
 it unique; by building an analytical framework based on confirmed
-results in the literature (uWUE, Medlyn model, Penman-Monteith [PM]), we are
-able to formally deduce what the actual ET response is, with other
-environmental factors held fixed. This ``other environmental factors
-held fixed'' component is, in our mind, nearly impossible to
+results in the literature (uWUE, Medlyn model, Penman-Monteith [PM]),
+we are able to formally deduce what the actual ET response is, with
+other environmental factors held fixed. This ``other environmental
+factors held fixed'' component is, in our mind, nearly impossible to
 deduce from pure data, because the environment is always changing and
@@
 as using uWUE is a simplification over using GPP directly, and may
-break down in extreme, limiting cases like VPD $\to$ 0. As
-a purpose-built model for assessing the ET response, we think the uWUE
-approach is justified. Essentially we are saying that to the degree we
-trust arithmetic, uWUE, Penman-Monteith, and the Medlyn stomatal
-conductance model, our evaluation of the derivative is robust.
+break down in extreme, limiting cases like when VPD, ET or GPP go to
+0. Additionally, we expect errors introduced by assuming that uWUE is
+fixed within a given PFT. These errors may be reduced if we account
+for spatiotemporal variation in uWUE, as discussed further in our
+response to RC1.  As a purpose-built model for assessing the ET
+response, we think the uWUE approach is justified. Essentially we are
+saying that to the degree we trust arithmetic, uWUE, Penman-Monteith,
+and the Medlyn stomatal conductance model, our evaluation of the
+derivative is robust.
 
@@
 increase in ET response. Granted, the net radiation perturbation would
-damp the decreased temperature response to the
+damp the negative temperature response to the
 aerodynamic/physiological terms, but all of the signs of the change
@@
 \label{ii}
+
 Thank you for having an open mind about the concave up ET-VPD curve
@@
 results of leaf scale response to VPD is that by taking an energy
-balance approach Penman Monteith accounts for changes in $e_{sat}(T)$
-at the surface induced by ET. To demonstrate this, compare Figure
-\ref{leaf} and Figure \ref{PM} (``Original Penman Monteith''
-Curve). In Figure \ref{leaf} we apply a conceptual leaf scale
-approach to calculate ecosystem ET without the energy balance
-consideration introduced by PM. In this figure, the curve shows a
-concave downward shape, in direct analogy to leaf scale T. Once the
-effects of evaporation's thermodynamics are included in the energy
-balance, the curve's shape is no longer concave downward (Figure
-\ref{PM}, ``Original Penman Monteith'' curve). How these conceptual
-results translate to leaf scale experiments depends on the
-experimental setup \citep[e.g.,][]{Rawson1977, Turner1984, Mott2013};
-however generally chamber-based leaf scale experiments do not preserve
-the energy balance relationships we expect for a surface in a
-natural environment (sometimes intentionally by design). Additionally,
-we would like to note that a concave up result is not necessarily
-inconsistent with the statement ``transpiration at the leaf scale is
-mostly positive'' with VPD. Especially for plants in well watered
-environments like crops, the manuscript's ecosystem scale shape
-(Figure \ref{PM}) would not be out of place on the figures in
-\cite{Rawson1977} and \cite{Mott2007} for the range of VPD considered
-(Figure \ref{PM} compared to Fig. 2 in \cite{Rawson1977}, also
-Figure 7 in \cite{Mott2007}). However, the change in relative
-magnitudes are not similar, and this effect can be traced partially to
-the effect of decreasing surface $e_{sat}(T)$ in response to
-increasing evapotranspiration. While not directly considered here (or
-in the manuscript) we also think it is important to consider how plant
-physiology in the natural environment could vary from plants grown in
-well-watered conditions in experiments. It is not inconceivable that
-phenotypic variability and adaptation in response to water limitation
-could result in plants with different responses to those grown in a
-lab. We presented crops in our figures because they are the most likely
-to well watered and analogous with laboratory results; however for
-other PFTs the concavity of the VPD-ET curves are more pronounced at
-higher VPD as compared to Figure \ref{PM}.
+balance approach Penman Monteith consistently accounts for the effects
+of the energy cost of evaporating water from a surface (e.g., changes
+in e$_{sat}(T)$). To demonstrate this, compare Figure \ref{leaf} and
+Figure \ref{PM} (``Original Penman Monteith'' Curve). In Figure
+\ref{leaf} we apply a conceptual leaf scale approach to calculate
+ecosystem ET without the energy balance consideration introduced by
+PM. In this figure, the curve shows a concave downward shape, in
+direct analogy to leaf scale T. Once the effects of evaporation's
+thermodynamics are included in the energy balance, the curve's shape
+is no longer concave downward (Figure \ref{PM}, ``Original Penman
+Monteith'' curve). How these conceptual results translate to leaf
+scale experiments depends on the experimental setup
+\citep[e.g.,][]{Rawson1977, Turner1984, Mott2013}; however generally
+chamber-based leaf scale experiments do not preserve the energy
+balance relationships we expect for a surface in a natural environment
+(sometimes intentionally by design). Additionally, we would like to
+note that a concave up result is not necessarily inconsistent with the
+statement ``transpiration at the leaf scale is mostly positive'' with
+VPD. Especially for plants in well watered environments like crops,
+the manuscript's ecosystem scale shape (Figure \ref{PM}) would not be
+out of place on the figures in \cite{Rawson1977} and \cite{Mott2007}
+for the range of VPD considered (Figure \ref{PM} compared to Fig. 2 in
+\cite{Rawson1977}, also Figure 7 in \cite{Mott2007}). However, the
+change in relative magnitudes are not similar, and this effect can be
+traced partially to the effect of decreasing surface $e_{sat}(T)$ in
+response to increasing evapotranspiration. While not directly
+considered here (or in the manuscript) we also think it is important
+to consider how plant physiologic response in the natural environment
+could vary from plants grown in well-watered conditions in
+experiments. It is not inconceivable that phenotypic variability and
+adaptation in response to water limitation could result in plants with
+different responses to those grown in a lab. We presented crops in our
+figures because they are the most likely to well watered and analogous
+with laboratory results; however for other PFTs the concavity of the
+VPD-ET curves is more pronounced at higher VPD as compared to Figure
+\ref{PM}.
 
@@
 induced by using PM (and energy balance) to assess ecosystem scale ET
-rather than directly applying leaf scale logic in controlled
+rather than directly applying leaf-scale logic from controlled
 experiments. However, our uWUE-based PM framework additionally
@@
 $g_s$ \citep{Damour2010}, we believe using the validated uWUE
-empiracle results \citep[e.g. Figure 3,][]{Zhou_2014} is a good
+empirical results \citep[e.g. Figure 3,][]{Zhou_2014} is a good
 approach to develop an estimate of ET response to VPD accounting for
@@
 the absence of a complete physical theory has been used often in
-science (e.g. turbulence) to increase our knowlege about a given
-systems' behavior, and we think it is well motivated here. However, it
-is important to add language on how introducing this
-empricism could alter interpretation of our results. In particular,
-while uWUE was developed using leaf scale photosythesis theory, it was
-validated with observations of a fully coupled land-atmosphere
-system. It captures a relatively consistent observed relationship
-between GPP, ET, and VPD as they vary. There is a chance that some of
-this relationship could be due to feedbacks between the land and the
+science (e.g. turbulence) to increase our knowledge about a given
+system's behavior, and we think it is well motivated here. However, it
+is important to add language on how introducing this empiricism could
+alter interpretation of our results. In particular, while uWUE was
+developed using leaf scale photosynthesis theory, it was validated
+with observations of a fully coupled land-atmosphere system. It
+captures a relatively consistent observed relationship between GPP,
+ET, and VPD as they vary. There is a chance that some of this
+relationship could be due to feedbacks between the land and the
 atmosphere, even though the relationship was derived by leaf scale
@@
 theory introduced ``real'' errors that were compensated for by
-feedbacks or ecosystem-scale physics in the observations).  Were this
-the case, then we introduced a conceptual impurity to the statement
-``we are examining the one way response of ET to VPD.''  But, given
-the theoretical motivations for uWUE we think the portion of the
-relationship related to land-atmosphere feedbacks are likely small, if
-they exist at all, and regardless this approach is still an
+feedbacks or ecosystem-scale processes in the observations).  Were
+this the case, then we introduced a conceptual impurity to the
+statement ``we are examining the one way response of ET to VPD.''
+But, given the theoretical motivations for uWUE we think the portion
+of the relationship related to land-atmosphere feedbacks are likely
+small, and regardless we argue that this approach is still an
 improvement over ignoring the documented effect of water stress on net
@@
 
-We also agree with you that uWUE includes a representation of ecosystem
-response. However, we would argue that this response dependence of
-stomatal conductance is not introduced by uWUE, but is actually an
-issue that exists inherently in any stomatal conductance theory that
-includes a photosynthesis term: photosynthesis depends on the
-physiological response and environmental conditions. Using uWUE is our
-(approximate) solution to this problem, and it works because we are
-able to algebraically manipulate PM to isolate ET after introducing
-uWUE. We do recognize that using uWUE introduces some uncertainty, but
-as previously discussed in this response we believe it is a useful tool
-for representing photosynthesis dependence of stomatal conductance in an
-analytical framework.
+We also agree with you that uWUE includes a representation of
+ecosystem response. However, we would argue that this response
+dependence of stomatal conductance is not introduced by uWUE, but is
+actually an issue that exists inherently in any stomatal conductance
+theory that includes a photosynthesis term: photosynthesis depends on
+the physiological response and environmental conditions. Using uWUE is
+our (approximate) solution to this problem, and it works because we
+are able to algebraically manipulate PM to isolate ET after
+introducing uWUE. We do recognize that using uWUE introduces some
+uncertainty, but as previously discussed in this response we believe
+it is a useful tool for representing the photosynthesis dependence of
+stomatal conductance in an analytical framework, with the specific
+goal of examining VPD sensitivity.
 

### 2018-12-23 doc/paper/hess-revision/review-3.tex
done with responses, all posted

@@
 to open the door for further discussion in this discussion phase. We
-look forward to reconciling all reviews with a revised manuscript and
+look forward to reconciling all reviews with a
 formal response in the final response phase. Also of relevance to this
@@
 \section{Methods/Introduction}
-Adding the clarifications you suggest for the methods section will
-greatly improve the manuscript, as well as more thorough discussion in
-the introduction about what is known about the transpiration response to
-VPD. This is also discussed peripherally in our response to
-\href{https://editor.copernicus.org/index.php/hess-2018-553-RC2.pdf?_mdl=msover_md&_jrl=13&_lcm=oc108lcm109w&_acm=get_comm_file&_ms=72556&c=153384&salt=74148443849227084}{RC2}.
+A more thorough discussion in the introduction about what is known
+about the transpiration response to VPD is definitely needed. This is
+also discussed peripherally in our response to
+\href{https://editor.copernicus.org/index.php/hess-2018-553-AC2.pdf?_mdl=msover_md&_jrl=13&_lcm=oc108lcm109w&_acm=get_comm_file&_ms=72556&c=154058&salt=19427645101542423504}{RC2}.
+The clarifications you
+suggest for the methods section will also improve the manuscript.
 
@@
 temperature) and land surface terms (e.g. g$_1$, uWUE, canopy height)
-influence the ET response. Using a more sophisticated representation
+influence the ET response. By using a more sophisticated representation
 of uncertainty we could then present the distribution of

### 2019-02-15 doc/paper/intro.tex
intro rough done, onto body

@@
 The objective of the present manuscript is to use reasonable
-approximations as a tool to develop intuition for plant response to
-atmospheric drying and evaluate the VPD dependence of ET. This
-intuition will aid interpretation of observations, full complexity
-models, and facilitate the disentanglement of complex land-atmosphere
-feedbacks. In the past, similar approaches were used to understand
-interactions between stomatal conductance, evapotranspiration and the
-environment \citep[e.g.,][]{Jarvis_1986, Mcnaughton_1991}. However, at
-the time researchers' understanding of the form of VPD's effect on
-plant physiology was limited, so they could not explore the
-sensitivity of ET to VPD, including VPD's effect on stomatal
-conductance and plant function.
+approximations established in prior research as a tool to develop a
+framework for understanding plant response to atmospheric drying and
+evaluating the VPD dependence of ET. This framework will aid
+interpretation of observations, full complexity models, and facilitate
+the disentanglement of complex land-atmosphere feedbacks. In the past,
+similar approaches were used to understand interactions between
+stomatal conductance, evapotranspiration and the environment
+\citep[e.g.,][]{Jarvis_1986, Mcnaughton_1991}. However, at the time
+researchers' understanding of the form of VPD's effect on plant
+physiology was limited, so they could not explore the sensitivity of
+ET to VPD, including VPD's effect on stomatal conductance and plant
+function.
 
@@
 \citet{Medlyn_2017}. Additionally, \citet{Zhou_2014} demonstrated that
-a quantity underlying water use efficiency
-$\left(uWUE = \frac{GPP\; \sqrt{VPD}}{ET}\right)$ properly captures a
-constant relationship between GPP, ET, and VPD over a diurnal cycle at
-the ecosystem scale. uWUE is also remarkably well conserved in the
-growing season across space and time, within a PFT
-\citep{Zhou_2015}. While stomatal conductance parameterizations and
-uWUE greatly simplify complex plant physiological processes, they
-still capture leading order ecosystem behavior for vegetated
-surfaces \citep{Medlyn_2017, Zhou_2014}, and are novel tools to
-transparently develop intuition for the behavior of complex, multiscale
-ecohydrologic systems.
+a quantity underlying water use efficiency $\left(uWUE = \frac{GPP\;
+\sqrt{VPD}}{ET}\right)$ properly captures a constant relationship
+between GPP, ET, and VPD over a diurnal cycle at the ecosystem
+scale. uWUE is also relatively well conserved in the growing season
+across space and time, within a PFT \citep{Zhou_2015}. While stomatal
+conductance parameterizations and uWUE greatly simplify complex plant
+physiological processes, they still capture ecosystem behavior for
+vegetated surfaces \citep{Medlyn_2017, Zhou_2014}, and are novel tools
+to transparently develop intuition for the behavior of complex,
+multiscale ecohydrologic systems.
 
-In this manuscript,
-we leverage uWUE and recent developments in stomatal conductance
-parameterizations \citep{MEDLYN_2011} to derive the theoretical
-one-way response of ET to VPD with other environmental variables
-properly controlled for, i.e. we develop a framework for evaluating
-the partial derivative of ET with respect to VPD. For the first time,
-we explicitly include VPD's full effect on stomatal conductance,
-including its leading order impact on photosynthesis. Our theory is
-validated and tested at multiple eddy-covariance stations spanning
-various climates and plant functional types.
+In this manuscript, we leverage uWUE and recent developments in
+stomatal conductance parameterizations \citep{MEDLYN_2011} to derive
+the theoretical one-way response of ET to VPD with other environmental
+variables properly controlled for, i.e. we develop a framework for
+evaluating the partial derivative of ET with respect to VPD. For the
+first time, we explicitly include VPD's full effect on stomatal
+conductance, including its impact on photosynthesis. We explore the
+range of possible ET responses to VPD, given parameters previously
+established in peer reviewed literature. Additionally, we show the
+sensitivity of the ET-VPD relationship to model and framework choice,
+highlighting the importance of: 1) future research on stomatal
+conductance and ecosystem scaling, and 2) thoughtful selection of
+photosynthesis and stomatal conductance model in more sophisticated
+land surface and earth system models.

### 2019-02-16 doc/paper/body.tex
now done with methods, except need to write out experiment table

@@
-\section{Materials and Methods}
-\subsection{Data}
-\label{data}
-We use both meteorological and eddy-covariance data from the FLUXNET2015
-database (data available at \sloppy
-\url{https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/} \sloppy),
-including all Tier 1 sites with at least four years of data and
-observations of the variables described in the methods section (Sect.
-\ref{methods}). Sixty-six sites met these requirements, and were grouped
-into nine plant functional types (PFT) according to the International
-Geosphere-Biosphere Programme vegetation classification scheme
-\citep{Loveland_1999}: cropland (CRO), grass (GRA), deciduous broadleaf
-forest (DBF), evergreen broadleaf forest (EBF), evergreen needleleaf
-forest (ENF), mixed forest (MF), closed shrub (CSH), savannah (SAV), and
-woody savannah (WSA) (site locations and citations are in Fig.
-\ref{map_fig} and Appendix \ref{flux_sites}, respectively).
-
-\begin{figure}[h]
-  \centering \includegraphics[trim={0 3cm 0 3cm}, clip]{./map.pdf}
-  \caption{Plant functional type and location of FLUXNET2015 sites
-    used in this analysis.}
-  \label{map_fig}
-\end{figure}
-
-The purpose of this study is to examine ecosystem response to
-atmospheric drying, focusing on the growing season. To accomplish
-this, we filter and quality control the data using a similar procedure
-as \cite{Zhou_2015}:
-\begin{itemize}
-\item Only measured or highest (``good'') quality gapfilled data,
-  according to quality control flags, are used.
-\item To isolate the growing season, we only use days in which the
-  average Gross Primary Productivity (GPP) exceeds 10\% of the
-  observed 95th percentile of GPP for a given site. GPP is calculated
-  using the nighttime respiration partitioning method.
-\item We remove days with rain and the day following to avoid issues
-  with rain interception and sensor saturation at high relative
-  humidity (\cite{Medlyn_2017}).
-\end{itemize}
-Additionally, as in \citet{Lin_2018}, we restrict data to the daytime,
-which is identified when downwelling shortwave radiation is greater
-than 50 W m$^{-2}$ and sensible heat flux is greater than 5 W
-m$^{-2}$. To reduce the chance of sensor saturation at high relative
-humidity, we remove all time steps for which VPD is less than .01 kPa,
-and to reduce errors at low windspeeds we remove all periods with wind
-magnitudes less than 0.5 m s$^{-1}$. Timesteps with negative observed
-GPP or ET are also removed, and we aggregate half hourly data to
-hourly averages to reduce noise \citep{Lin_2018}. After these quality
-control procedures, 400,983 upscaled hourly observations remain.
-
-\subsection{Methods}
+\section{Methods}
 \label{methods}
@@
   following \citet{Medlyn_2017} (Eq. (5)), and take the median $g_1$
-value to be representative of each PFT (Table \ref{pft}) instead of the mean to avoid extra weighting of rare outliers.
+  value to be representative of each PFT (Table \ref{pft}) instead of
+  the mean to avoid extra weighting of rare outliers.
 
@@
 
-Given eddy-covariance FLUXNET2015 data (Sect. \ref{data}), every
-term in our new version of PM (Eq. (\ref{et})) is observed, except
-for uWUE, which we fit by calculating its expectation, given the model
-and FLUXNET2015 data.
-
-However, eddy-covariance data are inherently noisy so we include a
-measure of uncertainty in our analysis. To account for observational
-error, as well as model uncertainty (e.g. temporal and spatial
-variations of uWUE and $g_1$), we introduce an uncertainty parameter
-$\sigma$ modifying uWUE:
+With Equation \ref{eq} we can take the partial derivative with respect
+to VPD:
 
@@
   \begin{equation}
-    ET = \frac{\Delta R_{net} + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R \; 1.6\; \sigma \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
-    \label{et_sigma}
+    \frac{\partial \;  ET}{\partial \; VPD} = \frac{2\; g_a \;
+      P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} -
+      \frac{\gamma c_a }{1.6 \; R\; \; \text{ uWUE }} \left(
+        \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right)
+    \right),
+    \label{d_et}
   \end{equation}
@@
 
-Now, from each FLUXNET2015 observation (i.e. for each hourly
-observation at every time step) we can evaluate $\sigma$:
-
-% \begin{linenomath*}
-  \begin{equation}
-    \sigma = - \frac{g_a \gamma c_a \sqrt{VPD} L_v P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R_{net} - g_a \rho_a c_p VPD\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})},
-    \label{sigma}
-  \end{equation}
-% \end{linenomath*}
-
-So, with this uncertainty analysis we can evaluate departure from
-our theory in observations, as a departure of $\sigma$ from unity. The
-variability of $\sigma$ across sites and time provides a measure of
-uncertainty in our model, assumptions, as well as in the FLUXNET2015
-observations themselves. The variability of $\sigma$ then propagates
- any uncertainty through to our partial derivative of Eq.
-(\ref{et_sigma}) with respect to VPD:
-
-% \begin{linenomath*}
-  \begin{equation}
-    \frac{\partial \;  ET}{\partial \; VPD} = \frac{2\; g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_a }{1.6 \; R\; \sigma \; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)
-    \label{d_et}
-  \end{equation}
-% \end{linenomath*}
-
-With Eq. (\ref{d_et}) we have an analytical framework for ecosystem
-response to atmospheric demand perturbations with environmental
-conditions held fixed. There are a few subtleties to taking the
-derivative in Eq. (\ref{d_et}): $\Delta$ ($\frac{d e_{s}}{d T}$)
-and $VPD$ are functionally related, so while taking the derivative we
-evaluate
+  deriving an analytical framework for ecosystem response to
+  atmospheric demand perturbations with environmental conditions held
+  fixed. There are a few subtleties to taking the derivative in
+  Eq. (\ref{d_et}): $\Delta$ ($\frac{d e_{s}}{d T}$) and $VPD$ are
+  functionally related, so while taking the derivative we evaluate
   $\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET}
@@
       fixed}}$. $RH$ and $e_s$ are assumed to be approximately
-independent, which is supported by the data (not shown).
+  independent, which is supported by data (not shown).
 
   This derivation relied either implicitly or explicitly on several
-assumptions. First, we assume that VPD at the leaf surface is the same
-as VPD at measurement height; physically this implies that leaves are
-perfectly coupled to the atmosphere. In reality, for some conditions
-and plant types the leaves can become decoupled from the boundary
-layer \citep{De_2017, Medlyn_2017}. Given our focus on the growing
-season, which is usually characterized by relatively high insolation
-inducing instability and convective boundary layers, we would expect
-the surface to be generally well coupled. An additional assumption in
-the formulation of $uWUE$ \citep{Zhou_2014, Zhou_2015} and
-\citet{Medlyn_2017}'s stomatal conductance model is that direct soil
-evaporation (E) contributions to ET remain small relative to
-transpiration (T). This should be more true during the growing
-season. The ratio of E to T may increase immediately after rainfall
-events due to high soil moisture and ponding, but VPD is generally low
-anyways during these times. However, some plant types allow for
-systematically larger contributions of E in ET, particularly those
-with sparse canopies and smaller relative amounts of transpiration. We
-therefore might expect that the theory will be most applicable to
-forest PFTs, which will be most strongly coupled to the boundary layer
-due to larger surface roughness, and will also generally have the
-highest ratios of transpiration to evaporation. Finally, for the goal
-of developing PFT-wide intuition, we assume that $g_1$ and $uWUE$ are
-constant in space and time. Both quantities have been shown to be well
-conserved over space and time within a PFT relative to inter-PFT
-variability. However, we would expect them to vary somewhat with
-environmental conditions including very low soil water content,
-temperature, and with intra-PFT plant-specific characteristics like
-wood density for tree PFTs \citep{Lin_2015}. Given these strong
-assumptions made with the goal of understanding broad, leading order
-plant behaviors, we take the extra care of including and analyzing
-uncertainty (Sect. \ref{testing}) and examining sensitivity to the
-exponent of the VPD response (Sect. \ref{functional_form}).
-
-We note one final comment on our derivation which is relevant for
-drought indices. If we approximate $c_a$ at a global mean CO$_2$
-concentration, then the RHS of Eq. (\ref{et}) is fully defined
-using commonly available weather station data and the constants
-published here. This makes Eq. (\ref{et}) useful in addition to
-PET in drought indices and hydrometeorological analysis for vegetated
-surfaces. Equation (\ref{et}) better reflects the physics of water
-exchange at the land surface and would only require fitting of uWUE
-and $g_1$, and can also account for changes in CO$_2$ concentration
-(assuming some relationship between $\lambda$ and [CO$_2$]), which is missing
-in other drought indices such as the Palmer Drought Severity Index
-(PDSI) \citep{Swann_2016, Lemordant_2016, Lemordant_2018}. Lastly, all
-code and data used in this analysis, including those used to generate
-the figures and tables, are publicly available at
-\url{https://github.com/massma/climate\_et}.
+  assumptions. First, we assume that VPD at the leaf surface is the
+  same as VPD at measurement height; physically this implies that
+  leaves are perfectly coupled to the atmosphere. In reality, for some
+  conditions and plant types the leaves can become decoupled from the
+  boundary layer \citep{De_2017, Medlyn_2017}. Therefor, our
+  derivation will be most applicable in times like the growing season,
+  when relatively high insolation induces instability and convective
+  boundary layers, and we would expect the surface to be generally
+  well coupled. An additional assumption in the formulation of $uWUE$
+  \citep{Zhou_2014, Zhou_2015} and \citet{Medlyn_2017}'s stomatal
+  conductance model is that direct soil evaporation (E) contributions
+  to ET remain small relative to transpiration (T). This should be
+  more true during the growing season. The ratio of E to T may
+  increase immediately after rainfall events due to high soil moisture
+  and ponding, but VPD is generally low anyways during these
+  times. However, some plant types allow for systematically larger
+  contributions of E in ET, particularly those with sparse canopies
+  and smaller relative amounts of transpiration. We therefore might
+  expect that the theory will be most applicable to forest PFTs, which
+  will be most strongly coupled to the boundary layer due to larger
+  surface roughness, and will also generally have the highest ratios
+  of transpiration to evaporation. Finally, we assume that $g_1$ and
+  $uWUE$ are constant with respect to the conceptual VPD
+  perturbation. Both quantities have been shown to be relatively
+  constant with respect to changes in
+  % TODO: make this a general discussion on partial derivatives and
+  % what things are constant/how to interpret?
+  VPD \cite{Franks_2017, Zhou_2014}. These parameters will however
+  vary with plant species and characteristics \citep[e.g. wood density,
+  ]{Lin_2015}, as well as environmental conditions including soil
+  water content and temperature.
+  % TODO: write exact parameter ranges we use, and introduce
+  % scaling/sign term here?
+  % \item
+
+To account for both this variability in plant parameters and the
+environmental, we will systematically analyze how the ET reponse to
+VPD (Equation \ref{d_et}) varies. Equation \ref{d_et} includes a
+``sign'' term  that determines the sign of the response in addition to
+magnitude:
+
+\begin{equation}
+  \label{sign}
+  \frac{c_p}{R_{air}} - \frac{\gamma c_a }{1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right),
+\end{equation}
+
+and a ``scaling'' multiplying the ``sign'' term:
+
+\begin{equation}
+  \frac{g_a \; P}{T(\Delta + \gamma)}.
+\end{equation}
+
+In the ``sign term'' most of quantities are relatively constant,
+except for VPD and the plant parameters g$_1$ and uWUE. In the scaling
+term, most of the terms are relatively constant with the exception of
+aerodynamic conductance ($g_a$) and temperature (especially its effect
+on $\Delta$). To determine the range of probable ET reponses to VPD we
+will systematically vary this parameters according to Table
+\ref{param_values}. Environmental parameters ($g_a$, $T$) are varied
+according to ranges we would expect for growing season conditions
+across a range of plant types (canopy height is a determinant $g_a$),
+and plant parameters (g$_1$, uWUE) are varied according to ranges
+established in previous literature \citep{Zhou_2015,
+  Medlyn_2017}. Using this previous literature we can connect the
+effect of varying plant parameters to specific plant types and
+characteristics. Lastly, all code and data used in this analysis,
+including those used to generate the figures and tables, are publicly
+available at \url{https://github.com/massma/climate\_et}.
 %\hyperref[https://github.com/massma/climate\_et]{https://github.com/massma/climate_et}.
@@
 
-\begin{equation}
-  \frac{g_a \; P}{T(\Delta + \gamma)},
-\end{equation}
-
 and a ``sign'' term, which determines whether ET increases or
@@
 
-\begin{equation}
-  \label{sign}
-  \frac{c_p}{R_{air}} - \frac{\gamma c_a }{1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right).
-\end{equation}
 All variables are positive, so the relative magnitude between the

### 2019-02-17 doc/paper/body.tex
tables look good but untested within latex

@@
     converted such that the quantities work with Equations 1-8 and the
-    variables defined Table \ref{definitions}.}
+    variables defined Table \ref{definitions}. }
   \small
@@
   \centering
-  \begin{tabular}{l c c @{\qquad} c c}
-    \hline
-    \multirow{2}[3]{*}{Abbreviation} & \multirow{2}[3]{*}{PFT} & \multirow{2}[3]{*}{$g_1$ (Pa$^{0.5}$)} & \multicolumn{2}{c}{uWUE ($\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET])}  \\
-    \cmidrule{4-5}
-
-                                     & & & fitted & \citet{Zhou_2015} \\
+  \begin{tabular}{l c c c}
+    \hline
+    Symbol & min (~15th percentile) & med (~50th percentile ) & max
+                                                                (~15th
+                                                                percentile)
+    & citation
 
@@
 will systematically vary this parameters according to Table
-\ref{param_values}. Environmental parameters ($g_a$, $T$) are varied
-according to ranges we would expect for growing season conditions
-across a range of plant types (canopy height is a determinant $g_a$),
-and plant parameters (g$_1$, uWUE) are varied according to ranges
-established in previous literature \citep{Zhou_2015,
-  Medlyn_2017}. Using this previous literature we can connect the
-effect of varying plant parameters to specific plant types and
-characteristics. Lastly, all code and data used in this analysis,
-including those used to generate the figures and tables, are publicly
-available at \url{https://github.com/massma/climate\_et}.
+\ref{param_varying}, while all other parametrs are held fixed (Table
+\ref{param_fixed}). Physical variables ($g_a$, $T$) and plant
+physiological paramters (g$_1$, uWUE) are varied according to
+literature-based expectations for a range of growing season conditions
+and plant types \citep{Zhou_2015, Medlyn_2017}. Using this previous
+literature we can connect the effect of varying plant parameters to
+specific plant types and characteristics. Lastly, all code and data
+used in this analysis, including those used to generate the figures
+and tables, are publicly available at \url{https://github.com/massma/climate\_et}.
 %\hyperref[https://github.com/massma/climate\_et]{https://github.com/massma/climate_et}.
+
+\begin{table}
+  \caption{Variable quantities in the ET response to VPD. Each value
+    is varied to determine the effect of a range of expected plant and
+    environmental conditions on ET response to VPD. A citation is
+    provided in cases where the quantites are directly derived from
+    previous literature. Conceptual percentiles
+    for interpreting each quantity are noted.}
+  \label{param_varying}
+  \centering
+  \begin{tabular}{l c c c c c}
+    \hline
+    Symbol & units [units in citation] & min (~15th percentile) & med (~50th percentile) & max
+                                                                                            (~15th
+                                                                                            percentile)
+    & citation  \\
+    \hline
+    \input{param_varying.tex}
+    \hline
+  \end{tabular}
+\end{table}
+
+\begin{table}
+  \caption{Quantities that are fixed in the ET response to VPD
+    (relative to those in Table \ref{param_varying}).}
+  \label{param_varying}
+  \centering
+  \begin{tabular}{l c c}
+    \hline
+    Symbol & units & value \\
+    \hline
+    \input{param_fixed.tex}
+    \hline
+  \end{tabular}
+\end{table}
 

### 2019-02-17 src/paper_figure_code/fully_idealized_figure.py

@@ -31,15 +31,17 @@ uwue_zhou = [6.99, 9.52, 12.05]
 uwues = list(map(d_io.uWUE_converter, uwue_zhou))
 ts = [10.0, 20.0, 30.0]
 gas = [0.015, 0.035, 0.055]
-p_default =
-
+p_d = 100000.0
+gamma_d = 64.5
+r_moist_d = 288.0
+c_a_d = 400.0
 
 def get_scaling_quantiles(_df):
   out = d_calc.scaling(_df)
   return list(map((lambda p: out.quantile(q=p)), [0.15, 0.5, 0.85]))
 
 def idealized_scaling(t, g):
-  _df = {'t_a_k': t + 273.15, 'p_a': 100000.0, 'gamma': 64.5}
+  _df = {'t_a_k': t + 273.15, 'p_a': p_d, 'gamma': gamma_d}
   return d_calc.scaling(_df, t_a=t, g_a=g)
 
 def get_scaling_quantiles_idealized():
@@ -49,7 +51,7 @@ def get_scaling_quantiles_idealized():
   return x
 
 def idealized_sign(g1, uwue):
-  _df = {'gamma': 64.6, 'r_moist' : 288.0, "c_a"  : 400.0}
+  _df = {'gamma': gamma_d, 'r_moist' : r_moist_d, "c_a"  : c_a_d}
   return d_calc.sign(_df, g1=g1, uwue=uwue, vpd=vpd)
 
 
@@ -138,3 +140,37 @@ for p in [0.15, 0.5, 0.85]:
   print("\n*****Percentile: %5.2f" % p)
   for var in ["p_a", "g_a","t_a","gamma","delta"]:
     print("%s: %f" % (var, df[var].quantile(q=p)))
+
+
+# write tables
+def write_varying():
+  """write output of varying plant params"""
+  fh = open("../../doc/paper/param_varying.tex", "w")
+  for symbol, unit, citation, values, citevals \
+      in zip(["g$_1$", "uWUE", "T", "g$_a$"],
+             ["Pa$^{1/2}$ [kPa$^{1/2}$]", "$\\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET] [g [C] hPa$^{1/2}$ kg$^{-1}$ [H$_2$O]]", "K", "m/s"],
+             ["Fig. 2, 7; \\citet{Medlyn_2014}", "Table 4; \\citet{Zhou_2015}", "-", "-"],
+             [g1s, uwues, ts, gas],
[…]

### 2019-02-17 doc/paper/body.tex
results done, onto discussion

@@
   the mean to avoid extra weighting of rare outliers.
-
-\begin{table}
-  \caption{Plant functional types, their abbreviation, calculated
-    Medlyn coefficient, and calculated uWUE. uWUE from
-    \citet{Zhou_2015}, including the observed standard deviation, is
-    shown for comparison. Note that uWUE from \citet{Zhou_2015} is
-    calculated from a different set of sites, and that units are
-    converted such that the quantities work with Equations 1-8 and the
-    variables defined Table \ref{definitions}. }
-  \small
-  \label{pft}
-  \centering
-  \begin{tabular}{l c c c}
-    \hline
-    Symbol & min (~15th percentile) & med (~50th percentile ) & max
-                                                                (~15th
-                                                                percentile)
-    & citation
-
-    \hline
-    \input{pft_params.tex}
-    \hline
-  \end{tabular}
-\end{table}
 
@@
   To account for both this variability in plant parameters and the
-environmental, we will systematically analyze how the ET reponse to
+  environment, we will systematically analyze how the ET reponse to
   VPD (Equation \ref{d_et}) varies. Equation \ref{d_et} includes a
-``sign'' term  that determines the sign of the response in addition to
-magnitude:
+  ``sign'' term that determines the sign of the response in addition
+  to magnitude:
 
@@
 aerodynamic conductance ($g_a$) and temperature (especially its effect
-on $\Delta$). To determine the range of probable ET reponses to VPD we
-will systematically vary this parameters according to Table
+through $\Delta$). To determine the range of probable ET reponses to
+VPD we will systematically vary this parameters according to Table
 \ref{param_varying}, while all other parametrs are held fixed (Table
@@
 literature we can connect the effect of varying plant parameters to
-specific plant types and characteristics. Lastly, all code and data
-used in this analysis, including those used to generate the figures
-and tables, are publicly available at \url{https://github.com/massma/climate\_et}.
+specific plant types and characteristics.
+
+All code and data used in this analysis, including those used to
+generate the figures and tables, are publicly available at
+\url{https://github.com/massma/climate\_et}.
 %\hyperref[https://github.com/massma/climate\_et]{https://github.com/massma/climate_et}.
@@
 
-\section{Results and Discussion}
+\section{Results}
 \label{results}
 
-By construction, the variability in the $\sigma$ term (Eq.
-(\ref{sigma})) contains all model and observational uncertainties. For
-an observation that perfectly matches our model and constant uWUE
-assumption, $\sigma$ will be one. Therefore, for our assumptions and
-framework to be reasonable $\sigma$ should be close to 1. An
-additional concern is that $\sigma$ may in fact be correlated with
-$VPD$, in which case the dependence would need to be accounted for
-when taking the derivative. Fortunately, there is a very weak
-dependence of $\sigma$ on VPD in their joint distribution, and
-$\sigma$ is indeed close to unity i.e. $O(1)$ (Fig.
-\ref{joint_vpd_sigma}). Given this weak dependence and the
-distribution of $\sigma$ we have confidence in our model framework and
-the data quality.
+By varying four parameters (g$_a$, T, uWUE, g$_1$) at three different
+values (Table \ref{param_varying}) we generate nine different values
+for the scaling term, nine different curves (as a function of VPD) for
+the sign term, and 81 different curves for the ET response to VPD
+($\frac{\partial \; ET}{\partial \; VPD}$). Examining a subset of nine
+of the curves for the ET response to VPD, defined by the minimum, median
+and maximum of for each of the scaling and sign term, it is clear
+that the range of responses includes those where ET almost always
+decreases with increasing VPD (water conservative), ET almost always
+increases with increasing VPD (water intensive), and intermediate
+values where ET will increase or decrease with VPD depending on the
+atmospheric demand for water (Figure \ref{full}).
 
-\begin{figure}[h]
-  \centering\includegraphics[width=0.75\textwidth]{./joint_vpd_sigma.pdf}
-  \caption{The joint distribution of $VPD$ and $\sigma$, with outliers
-    removed (defined as lowest and highest 5\% of $\sigma$). $\sigma$
-    exhibits a weak dependence on $VPD$, and $\sigma$ is $O(1)$ for
-    the bulk of the observations.}
-  \label{joint_vpd_sigma}
+\begin{figure}
+  \centering \includegraphics{./fully_idealized.pdf}
+  \caption{The functional form of the $\frac{\partial \; ET}{\partial
+      \; VPD}$ for minimum, median and maximum values of both the sign
+    term and the scaling term.}
+  \label{full}
 \end{figure}
 
-Before calculating the sensitivity of ET to VPD, we will consider the
-functional form of Eq. (\ref{d_et}). There are two main terms: a
-``scaling'' term, which modifies the magnitude but not the sign of the
-ET response to VPD ($\frac{\partial \; ET}{\partial \; VPD}$):
+Examining the sign term independent of the scaling term indicates the
+dependence of water conservative versus water intensive response on
+the main plant parameters, g$_1$ and uWUE (Figure \ref{sign}). Higher
+g$_1$ and uWUE shift the curve towards increasing ET responses with
+VPD (water intensive), and with increasing g$_1$ also leading to a
+higher non-linear VPD dependence of the response.
 
-and a ``sign'' term, which determines whether ET increases or
-decreases with VPD (i.e. atmospheric demand driven or physiologically
-controlled):
+\begin{figure}
+  \centering \includegraphics{./fully_idealized)sign.pdf}
+  \caption{The functional form of the $\frac{\partial \; ET}{\partial
+      \; VPD}$ evaluated at the median value of the scaling term, for
+    varying values of g$_1$ and uWUE as given in Table \ref{param_varying}.}
+  \label{sign}
+\end{figure}
 
-All variables are positive, so the relative magnitude between the
-first term and the second term in the sign term (Eq. (\ref{sign}))
-will determine whether ET increases or decreases with increasing
-VPD. If the second term is larger then plant control dominates and ET
-decreases with increasing VPD. However, if the first term is larger,
-then atmospheric demand dominates and ET increases with increasing
-VPD.
+Examining the scaling term independent of the sign term shows how
+aerodynamic conductance (communication between the atmosphere and the
+surface) and temperature (through $\Delta$ controls the efficiency of
+energy conversion to latent heat), amplify or supress the plant
+physiological response represented in the sign term. Both lower
+temperatures and higher aerodynamic conductance lead to amplified ET
+response to VPD, with the variabilty of aerodynamic conudcance
+resulting in a slightly higher variability of ET response relative to
+temperature, which is still an important factor.
 
-\subsection{Functional Form of the Sign Term}
-\label{sign_term}
-First, we explore the variables within the sign term to gain better
-intuition on the driver of either the increase or reduction of ET with
-VPD. CO$_2$ concentration ($c_a$) and the psychometric constant
-($\gamma$) are relatively constant over the dataset considered here so
-that the variability is dominated by $\sigma$ and $VPD$. uWUE could
-vary with soil moisture but has been shown to be relatively constant
-\citep{Zhou_2015}. This then means that the sign term only depends on
-VPD for a given PFT and is approximately just a function of $VPD$. We
-can further determine a critical threshold separating an increase from
-a decrease in ET, i.e. the threshold $VPD_{crit}$ such that the
-derivative vanishes $\frac{\partial \; ET}{\partial \; VPD} = 0$:
+\section{Discussion}
+\label{discussion}
+
 \small
@@
   \normalsize
-
-  noting that $VPD_{crit}$ mostly depends on the PFT parameters uWUE
-  and $g_1$, and only varies weakly with climate as most other
-  parameters related to the environment are nearly constant. The
-  calculated value of $VPD_{crit}$ for each PFT is shown in Table
-  \ref{vpd_crit}. For any values of $VPD$ less than $VPD_{crit}$, ET
-  will decrease with increasing VPD
-  ($\frac{\partial \; ET}{\partial \; VPD} < 0$), and for values of
-  $VPD$ greater than $VPD_{crit}$, ET will increase with increasing
-  VPD ($\frac{\partial \; ET}{\partial \; VPD} > 0$). In other words,
-  ecosystems regulate and mitigate evaporative losses up to the VPD
-  limit, $VPD_{crit}$, above which atmospheric demand is just too high
-  to be entirely compensated by stomatal and ecosystem regulation. We
-  note however that even though ET increases again above the critical
-  threshold, $VPD_{crit}$, ET is still much lower than potential
-  evaporation as stomata are still strongly regulating vapor fluxes to
-  the atmosphere. However, even in the absence of soil pore
-  evaporation ET cannot go completely to zero at high VPD, because
-  stomata are still slightly open to perform some photosynthesis
-  \citep{Ball_1987, Leuning_1990, MEDLYN_2011}. In addition, upward
-  xylem transport is necessary to maintain phloem transport, as well
-  as nutrient transport and thus carbon allocation \citep{De_2013,
-    Nikinmaa_2013, Ryan_2014}.
-
-\begin{table}
-  \caption{Values of $VPD_{crit}$, where
-    $\frac{\partial \; ET}{\partial \; VPD} = 0$, evaluated at PFT
-    average values for $R_{air}$, $\gamma$, and $c_a$. PFT-specific
-    constants ($g_1$, uWUE) are provided in Table \ref{pft}. For
-    values of $VPD$ less than $VPD_{crit}$,
-    $\frac{\partial \; ET}{\partial \; VPD}$ will be negative, and for
-    values of $VPD$ greater than $VPD_{crit}$,
-    $\frac{\partial \; ET}{\partial \; VPD}$ will be positive.}
-  \centering
-  \begin{tabular}{l c c c c c}
-    \hline
-    PFT & $R_{air}$ & $c_a$ (ppm) & $\gamma$  & \textbf{$VPD_{crit}$ (Pa)} \\
-    \hline
-    \input{vpd_crit.tex}
-    \hline
-  \end{tabular}
-  \label{vpd_crit}
-\end{table}
 
@@
 
-\subsection{Bulk statistics of ET response to VPD}
-\label{stats_sec}
-
-In this section we consider direct observations of ET response with
-eddy-covariance data, while including uncertainty with the $\sigma$
-term (Sect. \ref{methods}). These observational results of ET
-response (Table \ref{stats}) largely confirm our theoretical analysis, presented in the previous sections. For all PFTs, mean ET response to increasing VPD
-is negative. However, ET response evaluated at the average of all
-variables (e.g. $\sigma$, $T$, $c_a$, $VPD$) is positive for CRO, and
-negative for all other PFTs. This difference in mean ET response as
-compared to the ET response at mean environmental conditions is due to
-the non-linear nature of the response, in which negative responses are
-generally larger magnitude than positive responses (Fig.
-\ref{idealized_sign}). Therefore, both the mean ET response as well as
-the ET response at mean environmental conditions matches our
-expectations from the theory (Sect. \ref{sign_term}), with the
-exception that CRO observations are shifted more towards a negative ET
-response than we expect.
-
-\begin{table}
-  \caption{Statistics of $\frac{\partial \; ET}{\partial \; VPD}$ as a
-    function of PFT.}
-  \centering
-  \begin{tabular}{l c c c}
-    \hline
-    PFT & $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ & $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)$ & fraction $\frac{\partial \; ET}{\partial \; VPD} < 0.$ \\
-    \hline
-    \input{stats.tex}
-    \hline
-  \end{tabular}
-  \label{stats}
-\end{table}
-
-Regarding the frequency of negative and positive ET response, all PFTs
-exhibit a decreasing ET response with increasing VPD (physiologically
-controlled, water conservative response) for the majority of
-observations. The more water conservative PFTs generally exhibit
-higher frequency of negative ET response, especially when one factors
-in the distribution of environmental VPD (e.g. SAV and WSA grow in
-more arid climates, Fig. \ref{idealized_sign}). In general the bulk
-statistics match our theoretical expectations well, with the caveat
-that inclusion of uncertainty shifts crops towards a slightly more
-negative response to VPD, and shifts many of the other PFTs, which still
-exhibit a high frequency of negative ET response to VPD, towards more frequent
-occurrence of positive ET response than the theory and Fig.
-\ref{idealized_sign} would suggest. The bulk statistics motivate a
-more thorough examination of the structure of uncertainty and more
-sophisticated validation of our theory's performance against
-observations.
-
-\subsection{Validation of theory at eddy-covariance sites}
-\label{testing}
-We now compare more sophisticated distributions of the observed
-response to our simplified theory (Sect. \ref{sign_term}). The
-observed distribution of the sign term, as compared to what the theory
-would predict, is provided in Fig. \ref{test_sign}. Our goal was to
-capture the leading order behavior of the ET dependence on VPD. Given
-the assumptions we made, and the uncertainties of flux tower
-observations themselves, we expect a relatively large amount of noise
-when reproducing the derivatives of ET. However, the data largely
-reproduces our theoretical analysis.
-
-\begin{figure}
-  \centering
-  \centerline{\includegraphics[width=1.2\textwidth]{./test_sign.pdf}}
-  \caption{Comparison of the sign term with model uncertainty included
-    (box plots) to the sign term as calculated with simplifying
-    assumptions (blue line, as in Fig. \ref{idealized_sign}). Each
-    box plot corresponds to 5\% of the data, and the 5-95\% range of
-    VPD is plotted.}
-  \label{test_sign}
-\end{figure}
-
-This is particularly true for DBF and MF; the theory matches the
-leading order behavior of the function when uncertainty is included,
-and the observations match the theory with the addition of noise. The
-VPD dependence, given by the slope parameter ($g_1$), follows the
-median values of each bin. Perhaps most importantly, the x-intercept,
-and thus VPD$_{crit}$, matches nearly exactly between the theory and
-the observations. Therefore the sign of the ET response to increases
-in VPD should be well matched, subject to the unavoidable constraints
-of noise, much of which comes from the observations themselves. The
-uncertainty is non negligible; there are many observations in each bin
-for which the the sign of observation is opposite the response
-predicted by the theory, but to leading order our theory matches the
-observations well.
-
-While CSH has a much different functional form of the sign term than
-DBF and MF, CSH observations also match our theory to leading order,
-albeit with a bit more variability as a function of VPD. Again, the
-VPD dependence mostly determined from the slope parameter ($g_1$)
-closely matches the medians in the observation bins. The
-VPD-independent, strongly negative response is also captured. For CSH,
-there is rare occurrence of observed positive ET response with VPD
-($\approx 6$\%), even with uncertainty, so the sign of the
-observations almost always matches the sign of the theory, which
-states that ET response should always be negative.
-
-Biases between the theory and observations are similar for ENF, WSA,
-CRO and SAV. At low VPD the theory and the observations match
-well. However, at high VPD (upper 10-20\% of observations) the theory
-is biased slightly towards positive response as compared to the
-observations (Fig. \ref{test_sign}). For ENF this slight bias could
-be explained by a negative bias in $g_1$. However, for WSA, CRO, and
-SAV the observations in the highest VPD bins exhibit a downturn
-towards more negative response, which cannot be captured by the
-functional form of the sign term (Fig.
-\ref{idealized_sign}). Therefore, to explain the observations one of
-the variables in the sign term must change at extreme VPD (upper
-10-20\%). The most likely candidate is $uWUE$, which we have assumed
-constant to meet our goal of developing intuition for leading order
-behavior, but might be expected to decrease at extremely low SWC. This extremely low SWC should be correlated with high VPD in very dry environments
-through sensible heating increases \citep{Gentine_2016}. The theory
-for EBF exhibits similar limitations as for ENF, WSA, and SAV, except
-a larger portion of observations are biased negative as compared to
-the theory ($\approx$ 35\%, Fig. \ref{test_sign}). However, in
-general, and specifically for non-extreme VPD (VPD  $< \approx$70-90th
-percentiles), the theory matches the observations for the tree and
-savannah plant types well.
-
-The theory for GRA suffers from similar, but much more severe,
-limitations as for CRO, WSA, and SAV. GRA observations are
-characterized by a consistent trend back towards negative ET response
-at higher VPD, which the functional form of our theory is incapable of
-accounting for. As compared to CRO, WSA and SAV, the divergence
-between the theory and the observations is far greater for GRA,
-biasing 40-50\% of observations at higher VPD. In addition for the
-potential for soil moisture to alter $uWUE$, there are other sources
-of plant heterogeneity specific to GRA (and some extent CRO) that may
-alter $uWUE$ (or $g_1$) or invalidate other assumptions made in the
-methods section (Sect. \ref{methods}). We do not account for
-variability in plant height and surface roughness, or differences in
-C3 vs C4 photosynthesis and water strategies, which we might expect to
-vary substantially across sites, years, and season for GRA. These
-deficiencies could largely explain the inability of our theory to
-exactly match the observations in croplands. For example, a
-superposition of sites with C3 plants (at low environmental VPD) and
-C4 plants (at high environmental VPD) would explain the observed shift
-back towards negative ET response at high VPD when all sites are
-binned together, as in Fig. \ref{test_sign}. We hypothesize that the
-theory would validate against observations much better if these
-sources of variability were accounted for, at a cost of increased
-complexity and analytical opacity.
-
-While the above discussion shows that our theory has some limitations
-when applied to some PFTs, especially for grasslands, it does well for
-DBF, MF and CSH PFTs, and captures the response at non-extreme VPD
-(VPD $< \approx$ 70th-90th percentiles) for
-ENF, EBF, CRO, WSA, and SAV. In general, the leading order behavior
-observed in the data is captured by the theory. Departures between our
-theory and observations, specifically at high VPD, could be explained
-and conceptualized with shifts in $g_1$ and/or $uWUE$ due to
-site-specific plant type variability (e.g. more arid-adapted ecosystems
-at more arid sites), or temporal variability for some environmental
-conditions (e.g. decreases in $uWUE$ at extremely low soil
-moisture). To focus on general behavior and develop intuition for
-PFT-scale response, we have ignored these sources of variability in
-the present analysis. However, any site-specific or temporal plant
-functional variability that can be conceptualized with shifts to $g_1$
-and $uWUE$ can be analyzed with our framework. This opens a door to future analyses in which plant behavior in anomalous conditions
-can be explained and analyzed using Eq. (\ref{et}).
-
-\subsection{Observed ET response to VPD}
-
-Most of the results presented so far focus on the sign term, so now we
-turn to observations of ET response with the scaling term included
-(Fig. \ref{data_scatter}). Until now, in the interest of developing
-leading order intuition for ET response, and to be conservative in our
-acknowledgment of model and observational error, we've considered
-$\sigma$ variability to be a measure of uncertainty. An alternative
-viewpoint is that $\sigma \cdot uWUE$ represents spatial and temporal
-variability of uWUE, which may be expected within bounds \citep[see
-Table \ref{pft}, also ][]{Zhou_2015}. This is a less conservative
-view; some of the $\sigma \cdot uWUE$ variability will be due to model
-and observational error, so by viewing $\sigma \cdot uWUE$ as
-\textit{real} variability we run the risk of mistaking noise for
-signal. However, the advantage of this viewpoint is that, from
-\citet{Zhou_2014}, we have very high confidence that uWUE fit at the
-hourly timescale (as we do with $\sigma \cdot uWUE$) correctly
-captures the relationship between ET, GPP, and VPD, and that our form
-of PM introduces minimal error with its use of uWUE.
-
-\begin{figure}
-  \centering
-  \centerline{\includegraphics[width=1.2\textwidth]{./data_scatter.png}}
-  \caption{Scatter plots of observed
-    $\frac{\partial \; ET}{\partial \; VPD}$, including $\sigma$
-    variability, as a function of PFT, temperature and VPD. Please
-    note differences in the colorbar scale.}
-  \label{data_scatter}
-\end{figure}
-
-Therefore, with the caveat that some of the signal presented in Fig.
-\ref{data_scatter} may in fact be noise, we can interpret the observed
-distribution of ET response to VPD. In general, the observed response
-matches the intuitive theory. ET response to VPD shifts towards
-positive values as VPD increases (atmospheric demand dominating). CRO
-exhibit the highest occurrence of positive ET response, and the
-observations confirm that CRO are the most water intensive. CSH are
-the most water conservative, with a strong negative response. DBF,
-EBF, ENF, MF, SAV and WSA are also water conservative, but show some
-occurrence of positive ET response to VPD, particularly at higher
-observed VPD, as the theory predicts. GRA, while generally water
-conservative, does not match the theory well, with increasing frequency of
-\textit{negative} response at high VPD. This is as expected, given
-previous discussion on how we might expect more inter-site and inter-year variability for GRA (Sect. \ref{testing}).
-
-Fig. \ref{data_scatter} also includes the impact of the scaling
-term. For a given VPD, the magnitude of ET response does not vary
-strongly with temperature, confirming that any impact of the scaling
-term on the magnitude of the response is primarily due to changes in
-aerodynamic conductance. Intuitively this is reasonable; aerodynamic
-conductance will control how dominant balances at the land surface are
-communicated to the boundary layer. While $\Delta$ controls the
-efficiency of energy conversion to latent heat, it appears this is a
-second order term, relative to $g_a$, for scaling ET response.
-
-As with Fig. \ref{test_sign}, Fig. \ref{data_scatter} matches our
-expectation based on simplified theory. The sign term is most strongly
-scaled by $g_a$, and in general the occurrence of positive ET response
-increases as VPD increases. The willingness with which a given plant type
-evolved to use water dictates the occurrence of positive versus
-negative ET response. Water conservative ecosystems are highly effective
-at mitigating the effects of atmospheric demand, and can store water
-for later use by reducing ET in response to increasing VPD.
-
-\subsection{Limitations of theory: very dry soil moisture conditions}
-\label{swc_section}
-In formulating our theory with Penman Monteith, we implicitly did not
-account for very dry soil moisture conditions. For the majority of
-environmental conditions observed at the eddy covariance sites used
-here, soil conditions were not extremely dry so that we could assume a
-constant uWUE and $g_1$. We posit that ecosystems will generally
-optimize to host plants living in conditions which they evolved
-for. However, in extreme conditions and drought scenarios soil water
-content (SWC) could become the limiting factor for ET response to VPD,
-which our theory does not account for. In addition, low soil moisture
-conditions themselves increase VPD through land-atmosphere feedback
-\citep[][]{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006,
-  Salvucci_2013, Gentine_2016, Berg_2016}.
-
-Within our framework, any systematic bias due to the
-failure to account for SWC's effects in very dry conditions should
-manifest itself in a functional relationship between $\sigma$, our
-uncertainty measure, and SWC, which is observed at the FLUXNET
-sites. Examining the relationship between $\sigma$ and SWC will test
-to what extent our theory breaks down in very dry soil moisture
-conditions.
-
-If we again view $\sigma \cdot uWUE$ as, in addition to a measure of
-uncertainty, a short time-scale observation of uWUE, we would expect
-$\sigma \cdot uWUE$ to decrease at low SWC. If $\sigma \cdot uWUE$ is
-a very strong function of SWC, then our theory should be conditioned
-more strongly on well watered soil conditions. If $\sigma \cdot uWUE$
-is weakly a function of SWC, then our theory is more universal and
-independent of soil moisture conditions.
-
-Indeed, for all PFTs, there is some slight dependence of $\sigma \cdot
-uWUE$ on SWC, especially at low SWC (Fig. \ref{swc_boxplot}). The
-portion of observations for which our theory is biased by
-SWC-limitations varies by PFT, due to the nonlinear threshold at which
-soil moisture availability limits plant function. For CRO, MF and DBF,
-soil water content only matters for about the lowest 5\% of observations
-(each box is 5\% of observations in Fig. \ref{swc_boxplot}). And, CRO,
-relative to DBF, has a very weak dependence on soil moisture, which is
-reflective of the high likelihood that CRO sites are optimally irrigated
-and not water stressed, suggesting that observed departures from the
-theory for CRO are due to other factors (e.g. different photosynthesis
-pathways; C3/C4) than soil moisture. For ENF, EBF, SAV, and WSA,
-systematic SWC-induced biases in $\sigma \cdot uWUE$ emerge in about the
-lowest 20-30\% of SWC conditions, although large variability in the
-SWC-$(\sigma \cdot uWUE)$ relationship hampers interpretation for
-WSA. CSH presents a special case: the limited number of sites (2)
-preclude comment on the exact PFT-wide relationship.
-
-In contrast to ENF, CRO, and DBF, for GRA the relationship between
-$\sigma \cdot uWUE$ and SWC is more linear and affects a greater
-portion of observations; about 60\% of observations. Clearly, for GRA
-soil water frequently impacts plant function and alters ET response,
-and our theory is limited for the majority of environmental
-conditions. It is therefore not surprising that our theory tested
-poorly against the data for GRA, relative to to the other PFTs
-(Sect. \ref{testing}). For all other PFTs occurrence of soil
-moisture impacts was rare enough to not manifest itself in bulk
-statistics and figures.
-
-\begin{figure}
-  \centering
-  \centerline{\includegraphics[width=1.2\textwidth]{./swc_boxplot.pdf}}
-  \caption{$\sigma \cdot uWUE$ as a function of PFT and SWC. Each box
-    plot represents 5\% of all observations. Note that the highest
-    10\% of SWC observations are excluded to better resolve variability in the much more narrow bins at lower SWC (e.g. SWC has long tails at high values).}
-  \label{swc_boxplot}
-\end{figure}
-
-The observed dependence of $\sigma \cdot uWUE$ on SWC for GRA would
-explain the deficiencies of our theory compared to the observations in
-Fig. \ref{test_sign}, specifically the trend back towards
-negative ET response at high VPD. Aforementioned feedbacks between
-the land surface and the atmosphere, which are not accounted for due
-to our focus on the one-way response of the land surface to
-atmospheric conditions, would cause high VPD to be correlated with low
-SWC \citep[][]{Gentine_2016, Berg_2016}. So, at high VPD observations
-of low SWC are more likely, and this low SWC causes a lower uWUE
-(Fig. \ref{swc_boxplot}). The lower uWUE at low SWC/high VPD then
-leads to the observed downturn towards decreasing ET with increasing
-VPD, and the deviation between our theory (based on a constant uWUE
-assumption) and observations. It is also perhaps not coincidental that
-the portion of $\sigma \cdot uWUE$ affected by low soil moisture
-observations is similar to the portion of observations that do not
-match our theory at high VPD. Indeed, this may suggest that for all PFTs
-except for GRA, coupling and feedbacks between SWC, VPD and plant
-function are relatively rare.
-Future research will explore these relatively rare feedbacks in extreme conditions, which due to analytical intractability will require more opaque numerical analysis of many more complex processes, including boundary layer growth and state and their relationship to surface layer coupling and free-tropospheric lapse rates and humidity.
-
-To summarize, our theory is limited by its inability to account for
-soil water impacts on land surface response, and feedbacks between SWC
-and VPD. Fortunately, for most PFTs SWC's effect on ET is relatively
-rare ($<$30\% of observations) and does not manifest itself in the
-majority of observations and bulk statistics. However for GRA, SWC decreases water use efficiency for the majority of
-the observations. Soil moisture effects explain the deficiencies of
-our theory in Sect. \ref{testing}, particularly for GRA. By
-conceptualizing SWC effects as a change in $uWUE$ (and/or $g_1$), it
-will be possible for future analysis to explore the importance of soil
-moisture on plant response to VPD, and feedbacks between plant
-function, SWC, and VPD.
-
 \subsection{Functional form of ET dependence on VPD and its relation

### 2019-02-17 doc/paper/body.tex
done with extremely rough draft
tomorrow morning need to compile, proof read, and write abstract.

make sure to consult notes of ultimate vision for revisions.

@@
 
-\small
-% \begin{linenomath*}
-  \begin{equation}
-    VPD_{crit} = \frac{R_{air}}{4 c_p} \left( \frac{\gamma c_a}{1.6\; R \;  uWUE} + \sqrt{\frac{\gamma c_a}{1.6\; R \;  uWUE}\left( \frac{\gamma c_a}{1.6\; R \;  uWUE} + 8 g_1 \frac{c_p}{R_{air}}\right)} - 4 g_1 \frac{c_p}{R_{air}} \right),
-    \label{vpd_min_et}
-  \end{equation}
-% \end{linenomath*}
-  \normalsize
+Interpreting both the derivation and results of this manuscript hinges
+on an understanding and appreciation for the usefulness of partial
+dervatives for understanding the behavior of complex systems. Our
+results are a conceptual tool for answering the question, ``What is
+the ET response to VPD with all other quantities held fixed?'' This is
+useful, and we would argue is critical, for understanding complex
+responses when many quantities are varying symotaneously. Through
+partial derivatives, we can approach answering questions like ``What
+is the relative contribution of VPD in some observed response given
+that X Y and Z were also varying?'' In this example, X Y and Z could
+be, for example, radiation, soil moisture, and temperature. Equation
+\ref{d_et} explicitly provides an estimate of the ET response to VPD,
+given assumptions about ecosystem plant characteristics through the
+parameters g$_1$ and uWUE. Equation \ref{d_et} is generic to external
+environmental factors and the timescale over which a VPD response
+acts, and has a straightforward conceptual meaning. It provides the ET
+response, given other quantities and parameters held fixed. So, while
+the plant parameters g$_1$ and uWUE may vary with plant species and
+environmental conditions like soil moisture, we can still assess the
+ET response to VPD for a given plant species, or for a given soil
+moisture condition.
 
-Differences in $VPD_{crit}$ are exclusively determined by uWUE and the
-slope parameter ($g_1$) related to the plant functional type. A larger
-uWUE means a smaller $VPD_{crit}$, and an ET response to increases VPD
-that is more likely to be positive. At first glance this result is
-somewhat counter-intuitive; we expect that plants with a higher water
-use efficiency would be more water conservative. However, in reality
-uWUE determines how $WUE$ changes with VPD:
+However, given that g$_1$ and uWUE can vary, and that they both have a
+large impact on the ET response (Figure \ref{full}, \ref{sign}) it is
+useful to rexamine how these quantities vary in previous literature,
+framed by our results on ET responses. In \citet{Zhou_2015} most plant
+types have uWUE similiar to our median value, where whether ET
+response is postive (water intensive) or negative (water conservative)
+depends on VPD. However, crops, which we expect to prioritize carbon
+uptake over water conservation, have a higher uWUE closer to our
+maximum value. This higher uWUE results in a higher liklhood for an
+increasing ET response to VPD, which intuitively amkes sense given
+taht we expect crops ro keep stoamta open for access to carbon, at the
+cost of water loss. Shrubs, and to a lesser extent grass, gave an uWUE
+closer to the minumum uWUE. For these plant types, we would then
+expect a higher liklhood of a water conservatve response; that is, a
+decrease of ET in response to VPD. It is important to note that, while
+\citet{Zhou_2015} did not examine the roll of soil moisture in within
+plant-type variability of uWUE, we might expect some variation,
+especially in extreme cases when soil water becomes a limiting factor.
 
-\[WUE = \frac{GPP}{ET} = \frac{uWUE}{\sqrt{VPD}}.\]
-\[\frac{\partial \; WUE}{\partial \; VPD} = -\frac{uWUE}{2 \;
-    VPD^{3/2}}\]
+In \citet{Medlyn_2017} there is not as robust of a relationship
+between plant type and g$_1$ as there was for uWUE in
+\citet{Zhou_2015}. Given that there is a robust relatiohsip between
+g$_1$ and plant type in \citet{Lin_2015}, the ambiguos results in
+\citet{Medlyn_2017} could be due to noise in ecosystem scale
+observations, the consequences of imposing a model structure in some
+of the g$_1$ observations, or real ecosystem-scale within-plant type
+variability. These differences between leaf-scale behavior and
+ecosystem scale behaviro highlight the importance of understanding how
+and why slope parameters vary at the ecosystem scale, as argued by
+\citet{Medlyn_2017}. However, if we assume that there is some real
+analogy between leaf scale results in \citet{Lin_2015} and ecosystem
+scale behavior, and that some of the ambiguity in \citet{Medlyn_2017}
+is due to model and/or observational error, then we can make some
+connections between g$_1$-induced ET response variability and plant
+characteristics, which are subject to the strong caveat that we still
+do not fully understand the g$_1$'s utility at the ecosystem
+scale. Extrapolating \citet{Lin_2015}'s results (e.g. Fig. 2) to the
+ecosystem scale would reveal the following relationships between g$_1$
+and plant types:
 
-So, plants with a higher uWUE will have a greater \textit{decrease} in
-ecosystem-scale $WUE$ in response to increases in VPD. This decrease
-in $WUE$ causes more water loss per unit carbon gain, and explains the
-relationship between high uWUE and high likelihood of increases of ET
-in response to increasing atmospheric drying (increases in VPD).
+\begin{itemize}
+  \item C3 plants would have a generally higher g$_1$ than C4 plants,
+  \item Crops would have a larger g$_1$ than shrubs, grass, and
+    angiosperm trees, which have a higher g$_1$ than gymnosperm
+    trees.
+  \item Tropical and temperate climates would generally host plants
+    with a higher g$_1$ than artic and boreal climates.
+\end{itemize}
+For all of the above relationships, a higher g$_1$ means a higher
+liklihood of a postive ET response to increasing VPD (water
+intensive), and a smaller VPD dependence of the ET response.
 
-A tendency towards increasing ET response with increasing VPD can also
-be caused by a high slope parameter ($g_1$), characteristic of plants
-that at the leaf scale are more willing to trade water for access to
-atmospheric CO$_2$. Plants that are less conservative will be thus be
-more likely to increase ET with increasing VPD. Both the
-aforementioned effects (large uWUE, $g_1$) can amplify each other, and
-generally conspire to shift the sign term towards a positive value for
-a given PFT.
+Our theoretical results, when framed by previous literature tying
+parametets to plant types and climate, highlight the variability of ET
+response to VPD as a function of plant characteristics. Whether an
+ecosystem increases or decreases ET in response to VPD will vary
+depending on vegetation and climate. Generalizations about ET response
+to VPD therefore require thoughtful consideration of ecosystem
+characteristics; statements such as ``ET increases with warming due to
+increases in VPD'' may be false depending on the water conservation
+strategy of a given ecosystem **CITE SOME EXAMPLES OF THIS LOGIC FROM
+GLOBAL WARMING LIT**.
 
-This effect of uWUE and $g1$ on the sign term is most apparent by
-comparing two extreme PFTs: water intensive crops (CRO) and water
-conservative closed shrub (CSH). CRO has higher slope parameter and a
-slightly higher uWUE ($g_1 = 140.7$ Pa$^{1/2}$; $2.85$ $\mu$-mol [C]
-Pa$^{0.5}$ J$^{-1}$ [ET]) compared to CSH ($g1 = 75.1 \, Pa^{1/2}$,
-$uWUE=2.82$ $\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET]). These differences
-in PFT parameters cause opposite ET responses to changes in VPD
-between CRO and CSH. ET theoretically always decreases with increasing
-VPD for the more water conservative CSH, while ET frequently increases
-with increasing VPD for the more water intensive CRO (Fig.
-\ref{idealized_sign}). CROs evolved or were bred to prioritize GPP and
-yield and are thus not water conservative. They are very willing to
-trade water for photosynthesis and productivity, despite changes in
-VPD, while CSH are very unwilling to trade water for more
-photosynthesis.
+Additionally, the sensitivity of ET reponse to plant parameter
+highlights the importance of understanding and developing stomatal
+conductance models and parameters for the ecosystem scale. In
+particular, why and how g$_1$ behavior at the leaf and ecosystem scale
+are not directly analgalous much be understood. While the derivation
+here was explicit about the assumptions of a constant g$_1$ term with
+respect to the VPD perturbation, many sophsiticated land surface
+models and earth system models employing stomatal conductance models
+make similiar, and even stronger assumptions about the stationarity of
+VPD slope terms. In particular, many models will assume a constant
+g$_1$ term within a given plant functional type, and research derived
+from these models will not acknowledge limitations of this assumption,
+or difficulty in the literature with quantifying an ecosystem scale
+slope paramter \citep{Medlyn_2017}. Given that something so
+fundamental as the ET resposne to VPD varies strongly with g$_1$, we
+must invenst more in understanding g$_1$'s behavir at the ecosystem
+scale, and the efficacy of models using g$_1$'s for ecosystem-scale
+fluxes.
 
-\begin{figure}
-  \centering \includegraphics{./idealized_sign.pdf}
-  \caption{The functional form of the sign term, with $\sigma$ held
-    fixed at 1, and all terms except VPD set to PFT averages. For
-    comparison, the observed range of VPD for each PFT is plotted
-    below the x-axis. Stars denote 25th, 50th, and 75th percentiles,
-    and the range of the line spans the 5th-95th percentiles of
-    observed VPD. Vertical black lines denote the location of
-    $VPD_{crit}$ for each PFT, with the exception of CSH and WSA, for which
-    $VPD_{crit}$ is off-scale.}
-  \label{idealized_sign}
-\end{figure}
-
-As expected, the slope parameter ($g_1$) is a primary determinant of
-the VPD dependence for the sign term shown in Fig.
-\ref{idealized_sign}. Plants that are more conservative (small $g_1$)
-will tend to reduce ET with increasing VPD, and will be very effective
-at reducing ET, especially at low VPD. However, at very high VPD,
-gradients in vapor pressure at the leaf scale will become very strong
-as stomata reach their limits of closure in response to VPD
-(parameterized with $g_1$). As a result, ET response will begin to
-asymptote towards a constant ecosystem-scale values as leaf-scale
-response to VPD asymptotes towards zero.  Therefore, plants with a low
-$g_1$ will have the largest VPD dependence of ET response because the
-difference in ET response at low VPD (leaf stomatal response
-dominates) and high VPD (VPD gradient dominates) is largest. This is
-apparent in the strong VPD dependence of CSH, which has the lowest
-slope parameter ($g_1=75.1$ Pa$^{1/2}$) (Fig. \ref{idealized_sign}).
-
-To summarize our theoretical insights (Fig. \ref{idealized_sign} and
-Table \ref{vpd_crit}), CROs are the least water conservative and have
-the strongest overall tendency to increase ET with increasing VPD,
-while CSH are the most water conservative and have the strongest
-tendency to decrease ET with increasing VPD, as well as the strongest
-VPD dependence of response. Fig. \ref{idealized_sign} clearly shows,
-according to our theory, that for all PFTs except for crops there is
-frequent occurrence of a negative (plant dominating) ET response to
-increases in VPD. Therefore, plants are able in most atmospheric
-conditions to reduce ET in response to increased VPD and thus to
-reduce water loss. To better illustrate this, the ranges of observed
-environmental VPDs at the FLUXNET sites are plotted parallel to the
-x-axis. For CSH and WSA, VPD is always less than VPD$_{crit}$ (off
-scale) so that the plant response dominates in typical environmental
-conditions, emphasizing the water conservative strategy of those
-plants. For CRO on the other hand, VPD is higher than VPD$_{crit}$ for
-more than 50\% of observations, emphasizing that those plants operate
-with an aggressive water usage strategy, are water intensive and were
-actually engineered for photosynthesis rather than water saving. For
-DBF, EBF, MF, GRA, and SAV more than half of the observed VPD are less
-than VPD$_{crit}$, i.e. in conditions where plant response
-dominates. SAV has a more water conservative response than the forest,
-grass, and crop plan types, but still responds by increasing ET with
-increasing VPD for about a quarter of observations, due to the high
-aridity (VPD) of the SAV ecoclimate. It is also important to note that
-for all PFTs, even when atmospheric demand dominates, ET response to
-VPD is still far more negative than it would be for potential
-evaporation $\partial PET/\partial VPD$, i.e. atmospheric demand only,
-emphasizing that there is still a strong regulation of evaporative
-flux by stomata and though the plant xylem. The sign term in the PET
-case would just be a constant ($\frac{c_p}{R_{air}} \approx 3.5$),
-which is far larger than any part of the curves for any PFT. Plants
-are always regulating water exchange from the land surface, even when
-they reach the limits of they ability to do so.
-
-\subsection{Functional Form of the Scaling Term}
-\label{scale_term}
-While the above discussion of the sign of
-$\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our
-question of when ET response increases or decreases with VPD,
-understating the overall magnitude of the ET response is important to
-soil-plant-atmosphere water budgeting. So we now more closely examine
-the terms that affect how the sign term is scaled:
-
-\begin{equation}
-  \frac{g_a \; P}{T(\Delta + \gamma)}.
-\end{equation}
-
-$\frac{P}{T}$ is an air-density term, which varies little compared to
-aerodynamic conductance and Clausius-Clapeyron ($\Delta$). The
-psychometric constant ($\gamma$) is also relatively constant, so the
-scaling term should be primarily a function of aerodynamic conductance
-and temperature, through the Clausius-Clapeyron relationship
-$\Delta$. This is as expected, given that the aerodynamic conductance
-represents the efficiency of exchange between the surface and the
-atmosphere. As aerodynamic conductance increases, any plant response
-will be communicated more strongly to the atmosphere (and vice-versa).
-
-$\Delta$'s presence in the scaling term also matches physical
-intuition. $\Delta$ (and also the approximately constant $\gamma$)
-control the efficiencies with which surface energy is converted to
-latent and sensible heat \citep{Monteith_1965}. The functional from of
-$\Delta$ will be the same across PFTs, but the temperature range may
-vary slightly. In contrast, aerodynamic conductance will vary strongly
-with PFT due to the importance of surface roughness for aerodynamic
-conductance. So most of the differences in scaling between PFT should
-be in the aerodynamic conductance term.
-
-The control of the scaling term variability between PFTs by
-aerodynamic conductance is confirmed by data (Fig.
-\ref{scale_vary}). Differences between PFT are almost entirely due to
-differences in aerodynamic conductance, rather than differences in
-observed temperature ranges. The scaling term for the tree PFTs (DBF,
-EBF, ENF, MF) is generally about double the scaling terms for other
-PFTs which have lower surface roughness and generally smaller
-aerodynamic conductance (GRA, CSH, CRO). The savannah (WSA, SAV) PFT's
-scaling is somewhere between GRA, CSH, and CRO, and DBF, EBF, ENF, and
-MF, due to higher variability and surface roughness.
-
-\begin{figure}
-  % \centering
-  \centerline{\includegraphics[width=1.2\textwidth]{./idealized_scale.pdf}}
-  \caption{Primary sources of variability for the scaling term, as a
-    function of PFT. The 5th-95th percentile range of temperature is
-    plotted at the 5th, 25th, 50th, 75th, and 95th percentiles of
-    aerodynamic conductance, as observed for each PFT.}
-  \label{scale_vary}
-\end{figure}
-
-Within each PFT, the scaling term variability is controlled both by
-environmental temperature and aerodynamic conductance variability
-(Fig. \ref{scale_vary}). While the observed variability of the
-aerodynamic conductance contributes more to the scaling term
-variability than temperature, the temperature contribution is
-non-negligible. Specifically, the scaling term is generally larger at
-low temperatures when latent heat is relatively inefficient at moving
-energy away from the surface. This effect amplifies the role of
-aerodynamic conductance variability at low temperatures.
-
-To summarize, variability between PFTs is mostly controlled by
-systematic differences in aerodynamic conductance, due to differences
-in surface roughness between each PFT, and possibly to a lesser extent
-wind conditions. In contrast, variability within PFT is also
-controlled by temperature, through Clausius-Clapeyron. But,
-aerodynamic conductance variability generally impacts the scaling term
-more than temperature, even within PFTs.
-
+So far we have discussed the role of different parameters in the ET
+response. However, the form of the ET response is imposed by our
+choice of stomatal conductance model. We now examine how the model
+choice, and speicifcally the exponent of the VPD dependence in teh
+stomatal conductance model, can impact the general form of the ET-VPD
+relationship. This is relevant as many models include an
+exponententiated VPD term \citep[e.g.,][]{Leuning_1990}, but the value
+of the exponent can vary. It is useful to understand how generally the
+choice of model imposes the functional form of the ET-VPD
+relationship, given its fundamental importance for land atmosphere
+interaction.
 \subsection{Functional form of ET dependence on VPD and its relation
@@
 \end{itemize}
+
 While it is possible that in the future some other form of
@@
 response to VPD with environmental conditions held
-fixed. Accomplishing this with purely statistical
-methods applied to flux observations would be very difficult, given
-the relatively fast time scale of plant response and the
-non-stationarity of [solar forced] environmental conditions over the
-relatively coarse (half hourly) flux estimates (which is required to
-obtain robust eddy-covariance statistics). Our results on the shape of
-the ET-VPD curve with environmental conditions held
-fixed can be built upon with future work examining how changes in VPD
-and environmental conditions (e.g. soil water storage) feedback upon
-one another. In the soil water storage example, over very long time
-scales extremely high VPD perturbations coupled with no precipitation
-could result in decreases in soil water storage such that water
-becomes limiting. This could be represented by an extension of our
-framework in which uWUE is allowed to decrease with decreasing SWC, as
-observed in Sect. \ref{swc_section}. Here, we focus our results by
-assuming constant PFT-wide conditions to build baseline intuition for
-ET-VPD dependence. For most PFTs, the theory with plant function held
-fixed matches the leading order behavior of the observations where
-plant function varies (Sect. \ref{testing}).
+fixed. Accomplishing this with purely statistical methods applied to
+flux observations would be very difficult, given the relatively fast
+time scale of plant response and the non-stationarity of [solar
+forced] environmental conditions over the relatively coarse (half
+hourly) flux estimates (which is required to obtain robust
+eddy-covariance statistics). Our results on the shape of the ET-VPD
+curve with environmental conditions held fixed can be built upon with
+future work examining how changes in VPD and environmental conditions
+(e.g. soil water storage) feedback upon one another. In the soil water
+storage example, over very long time scales extremely high VPD
+perturbations coupled with no precipitation could result in decreases
+in soil water storage such that water becomes limiting. This could be
+represented by an extension of our framework in which uWUE is allowed
+to decrease with decreasing SWC. Here, we focus our results by
+framing theme with previous literature to  build baseline intuition for
+ET-VPD dependence.

### 2019-02-17 doc/paper/conclusions.tex
done with extremely rough draft
tomorrow morning need to compile, proof read, and write abstract.

make sure to consult notes of ultimate vision for revisions.

--- 2018-10-31 doc/paper/conclusions.tex
@@
 ecosystem will tend to reduce or increase ET with increasing VPD,
-which we evaluated against a range of eddy-covariance data spanning
-different climates and plant functional types. The goal was to capture
-the leading order behavior of the system to gain some intrinsic
-knowledge for its behavior. This intuition can be used to disentangle
-land atmosphere feedbacks in more complicated scenarios, and will aid
-interpretation of observations and sophisticated models.
+which we framed using previous literature exploring the relatiohsip
+between plant parameters, plant types, and climate. The goal was to
+understand the range of likely ET responses to VPD and develop some
+intuition for how the ET response may vary with plant types and
+climate. This intuition can be used to disentangle land atmosphere
+feedbacks in more complicated scenarios, and will aid interpretation
+of observations and more sophisticated models.
 
-Our theory suggests that for a majority of environmental conditions,
-plants will tend to conserve water and reduce ET with increasing
-VPD. Stomatal regulation and plant physiological response strongly
-regulate ET, and this regulation varies by PFT. CROs are the least
-water conservative, while DBF, EBF, SAV, GRA, MF, WSA, ENF and CSH are
-progressively more water conservative (more likely to reduce ET in
-response to increasing VPD). SAV and WSA exhibit positive ET response
-to VPD not necessarily because of poor water conservation strategies
-relative to other PFTs, but because of greater occurrence of high
-atmospheric demand (VPD) relative to other PFTs. Observations of ET
-response to VPD exhibit the same general behavior as the theory, with
-ET response becoming more positive (atmospheric demand dominating) as
-environmental VPD increases within a PFT, and more negative for PFTs
-that are adapted to arid conditions and prioritize water conservation
-over primary production.
+ET response to VPD can vary from strongly water conservative (ET
+always decreasing in response to increasing VPD), to strongly water
+intensive (ET always increasing in response to VPD), which is
+idicative of the large observed diversity of stomatal regulation and
+plany physiological strategy. Higher uWUE and g$_1$ values make the ET
+response more likely to be positive, while decreasing temperature and
+increasing g$_a$ amplify the magnitue of the response. Previous
+literature \citep{Zhou_2015, Lin_2015} suggests that crops, through
+association with higher g$_1$ and uWUE, are more likely to exhibit a
+postiive ET response to VPD. Shrubs (lower uWUE), C4 plants (lower
+g$_1$), gymnosperm trees (lower g$_1$), and plants in a artic and
+boreal climates (lower g$_1$ are more likely to exhibit a negative ET
+response to VPD. However interpretation of g$_1$-induced variability
+is partially muddied by ambiguity in upscaling leaf-scale estimations
+of g$_1$ to the ecosystem scale \citep{Medlyn_2017}.
 
-Our paper builds important intuition for how plants will
-respond to VPD perturbations. We show that given optimal stomatal
-function and fixed environmental conditions, the ET-VPD dependence is
-theoretically concave upward, with ET increasing with increasing VPD as
-VPD increases past some critical value (Table \ref{vpd_crit}). However
-future research should focus on fully understanding the form of
-stomatal VPD dependence, as this result is sensitive to the exponent of
-VPD dependence, which we currently believe is 1/2 \citep{MEDLYN_2011,
-  Zhou_2014}. Our results are applicable to understanding the impact
-of  expected increases in VPD
-induced by global change.  Plant physiological responses to direct
-CO$_2$ effects \citep[e.g.,][]{Swann_2016, Lemordant_2018} receives more attention
+Our paper builds important intuition for how plants respond to VPD
+perturbations. We show that given optimal stomatal function and fixed
+environmental conditions, the ET-VPD dependence is theoretically
+concave upward, with ET increasing with increasing VPD as VPD
+increases past some critical value **Add VPD_crit plot to
+results??**. However future research should focus on fully
+understanding the form of stomatal VPD dependence, as this result is
+sensitive to the exponent of VPD dependence, which we currently
+believe is 1/2 \citep{MEDLYN_2011, Zhou_2014}. Indeed, this
+sensitivity to the exponent of VPD dependence is an important result
+itself: land surface models, including those used in earth system
+models for climate forecasts, employ different assumptions about the
+exponent of VPD dependence in stomatal conductance
+\citep[e.g.,][]{Ball_1987, Leuning_1990, MEDLYN_2011}, and these
+assumptions can fundamentally change the relationship between ET and
+VPD from one that is concave upward (local minimum in ET) to one that
+is concave downward (local maximum in ET).
+
+Our results are also applicable to understanding the impact of
+expected increases in VPD induced by global change.  Plant
+physiological responses to direct CO$_2$ effects
+\citep[e.g.,][]{Swann_2016, Lemordant_2018} receives more attention
 than physiological response to indirect effects like increased
-VPD. Here, we provide broad PFT-focused results showing a likely
-decrease in ET in response to positive VPD perturbations (atmospheric
-drying), which is consistent with recent observational analysis
-\citep[e.g.,][]{Rigden_2017}. Feedbacks between the land and the
-atmosphere may alter the net response to a long-timescale global VPD perturbation,
-but our focus on the one way plant response to a VPD perturbation in
-the atmospheric boundary layer is an important first step to
-disentangling such feedbacks, both in observations and model
-simulations of the present and future. By removing Penman Monteith's
-dependence on implicit relationships between GPP, VPD, and ET, we
-allow for explicit, accurate future analysis of plant-VPD feedbacks
-in the atmospheric boundary layer (Eq. (\ref{et})). Our approach
-can be extended to examine varying plant response to more nuanced
-consideration of plant type and climate. Any plant physiological
-heterogeneity or feedback that can be conceptualized with shifts in $g_1$
-\citep[e.g.][]{Lin_2015, Medlyn_2017} and/or uWUE
-\citep[e.g.][]{Zhou_2014} are representable within our framework, which
-opens the door for a hierarchy of more sophisticated climate- and
-plant-specific analysis of ET sensitivity to environmental variables
-(including VPD). We argue that such simplified conceptual frameworks
-are critical tools for disentangling land-atmosphere feedbacks at
-various scales, from diurnal to seasonal and beyond, and to
-characterize ET response in a warmer, atmospherically drier, and
-enriched CO$_2$ world.
+VPD. Here, we provide a framework for understanding ET response to ET
+using a simplfied model fo two plant parameters. Feedbacks between the
+land and the atmosphere may alter the net response to a long-timescale
+global VPD perturbation, but our focus on the one way plant response
+to a VPD perturbation in the atmospheric boundary layer is an
+important first step to disentangling such feedbacks, both in
+observations and model simulations of the present and future. By
+removing Penman Monteith's dependence on implicit relationships
+between GPP, VPD, and ET, we allow for explicit, accurate future
+analysis of plant-VPD feedbacks in the atmospheric boundary layer
+(Eq. (\ref{full})). Our approach can be extended to examine varying
+plant response to more nuanced consideration of plant type and
+climate. Any plant physiological heterogeneity or feedback that can be
+conceptualized with shifts in $g_1$ \citep[e.g.][]{Lin_2015,
+Medlyn_2017} and/or uWUE \citep[e.g.][]{Zhou_2014} are representable
+within our framework, which opens the door for a hierarchy of more
+sophisticated climate- and plant-specific analysis of ET sensitivity
+to environmental variables (including VPD). We argue that such
+simplified conceptual frameworks are critical tools for disentangling
+land-atmosphere feedbacks at various scales, from diurnal to seasonal
+and beyond, and to characterize ET response in a warmer,
+atmospherically drier, and enriched CO$_2$ world.

### 2019-02-17 doc/paper/abstract.tex
spell check done and makefile updated
now proofread to send off to Pierre and Changjie

@@
-Increasing vapor pressure deficit (VPD) increases atmospheric demand for
-water, and vapor pressure deficit is expected to rise with increasing
-greenhouse gases. While increased evapotranspiration (ET) in response to
-increased atmospheric demand seems intuitive, plants are capable of
-reducing ET in response to increased VPD by closing their stomata, in an
-effort to conserve water. Here we examine which effect dominates
-response to increasing VPD: atmospheric demand and increases in ET, or
-plant physiological response (stomata closure) and decreases in ET. We
-use Penman-Monteith, combined with semi-empirical optimal stomatal
-regulation theory and underlying water use efficiency, to develop a
-theoretical framework for understanding how ET responds to increases in
-VPD.
+Increasing vapor pressure deficit (VPD) increases atmospheric demand
+for water, and vapor pressure deficit is expected to rise with
+increasing greenhouse gases. While increased evapotranspiration (ET)
+in response to increased atmospheric demand seems intuitive, plants
+are capable of reducing ET in response to increased VPD by closing
+their stomata, in an effort to conserve water. Here we examine which
+effect dominates response to increasing VPD: atmospheric demand and
+increases in ET, or plant physiological response (stomata closure) and
+decreases in ET. We use Penman-Monteith, combined with semi-empirical
+optimal stomatal regulation theory and underlying water use
+efficiency, to develop a theoretical framework for understanding how
+ET responds to increases in VPD. The theoretical ET response to VPD
+over a range of reasonable environmental conditions and plant
+characteristics varies from a strong decrease in ET in response to
+increasing VPD (water conservative) to a strong increase in ET in
+response to VPD (water intensive), highlighting the diversity of plant
+water regulation strategies.
 
-The theory suggests that for most environmental conditions and plant
-types, plant physiological response dominates and ET decreases with
-increasing VPD. Plants that are evolved or bred to prioritize primary
-production over water conservation (e.g. crops) exhibit a higher
-likelihood of atmospheric demand-driven response (ET
-increasing). However for forest, grass, savannah, and shrub plant types,
-ET more frequently decreases than increases with rising VPD. This work
-serves as an example of the utility of our simplified framework for
-disentangling land-atmosphere feedbacks, including
-the characterization of ET response in an atmospherically drier,
-enriched CO$_2$ world.
+The ET response varies due to: 1) climate, with tropical and
+temperate climates more likely to exhibit a positive ET response than
+boreal and arctic climates; 2) photosynthesis strategy, with C3 plants
+more likely to exhibit a positive ET response than C4 plants; and 3)
+due to plant type, with crops more likely to exhibit a positive ET
+response, and shrubs and gymniosperm trees more likely to exhibit a
+negative ET response. These results, derived from previous literature
+connecting plant parameters to plant and climate characteristics,
+highlight the utility of our simplified framework for understanding
+complex land atmosphere systems in terms of idealized scenarios.

### 2019-02-17 doc/paper/body.tex
spell check done and makefile updated
now proofread to send off to Pierre and Changjie

@@
   To account for both this variability in plant parameters and the
-  environment, we will systematically analyze how the ET reponse to
+  environment, we will systematically analyze how the ET response to
   VPD (Equation \ref{d_et}) varies. Equation \ref{d_et} includes a
@@
 aerodynamic conductance ($g_a$) and temperature (especially its effect
-through $\Delta$). To determine the range of probable ET reponses to
+through $\Delta$). To determine the range of probable ET responses to
 VPD we will systematically vary this parameters according to Table
-\ref{param_varying}, while all other parametrs are held fixed (Table
+\ref{param_varying}, while all other parameters are held fixed (Table
 \ref{param_fixed}). Physical variables ($g_a$, $T$) and plant
-physiological paramters (g$_1$, uWUE) are varied according to
+physiological parameters (g$_1$, uWUE) are varied according to
 literature-based expectations for a range of growing season conditions
@@
     environmental conditions on ET response to VPD. A citation is
-    provided in cases where the quantites are directly derived from
+    provided in cases where the quantities are directly derived from
     previous literature. Conceptual percentiles
@@
 \begin{figure}
-  \centering \includegraphics{./fully_idealized)sign.pdf}
+  \centering \includegraphics{./fully_idealized_sign.pdf}
   \caption{The functional form of the $\frac{\partial \; ET}{\partial
@@
 surface) and temperature (through $\Delta$ controls the efficiency of
-energy conversion to latent heat), amplify or supress the plant
+energy conversion to latent heat), amplify or suppress the plant
 physiological response represented in the sign term. Both lower
 temperatures and higher aerodynamic conductance lead to amplified ET
-response to VPD, with the variabilty of aerodynamic conudcance
+response to VPD, with the variability of aerodynamic conductance
 resulting in a slightly higher variability of ET response relative to
@@
 on an understanding and appreciation for the usefulness of partial
-dervatives for understanding the behavior of complex systems. Our
+derivatives for understanding the behavior of complex systems. Our
 results are a conceptual tool for answering the question, ``What is
@@
 useful, and we would argue is critical, for understanding complex
-responses when many quantities are varying symotaneously. Through
+responses when many quantities are varying simultaneously. Through
 partial derivatives, we can approach answering questions like ``What
@@
 large impact on the ET response (Figure \ref{full}, \ref{sign}) it is
-useful to rexamine how these quantities vary in previous literature,
+useful to examine how these quantities vary in previous literature,
 framed by our results on ET responses. In \citet{Zhou_2015} most plant
-types have uWUE similiar to our median value, where whether ET
-response is postive (water intensive) or negative (water conservative)
+types have uWUE similar to our median value, where whether ET
+response is positive (water intensive) or negative (water conservative)
 depends on VPD. However, crops, which we expect to prioritize carbon
 uptake over water conservation, have a higher uWUE closer to our
-maximum value. This higher uWUE results in a higher liklhood for an
-increasing ET response to VPD, which intuitively amkes sense given
-taht we expect crops ro keep stoamta open for access to carbon, at the
+maximum value. This higher uWUE results in a higher likelihood for an
+increasing ET response to VPD, which intuitively makes sense given
+that we expect crops to keep stomata open for access to carbon, at the
 cost of water loss. Shrubs, and to a lesser extent grass, gave an uWUE
-closer to the minumum uWUE. For these plant types, we would then
-expect a higher liklhood of a water conservatve response; that is, a
+closer to the minimum uWUE. For these plant types, we would then
+expect a higher likelihood of a water conservative response; that is, a
 decrease of ET in response to VPD. It is important to note that, while
@@
 between plant type and g$_1$ as there was for uWUE in
-\citet{Zhou_2015}. Given that there is a robust relatiohsip between
-g$_1$ and plant type in \citet{Lin_2015}, the ambiguos results in
+\citet{Zhou_2015}. Given that there is a robust relationship between
+g$_1$ and plant type in \citet{Lin_2015}, the ambiguous results in
 \citet{Medlyn_2017} could be due to noise in ecosystem scale
@@
 variability. These differences between leaf-scale behavior and
-ecosystem scale behaviro highlight the importance of understanding how
+ecosystem scale behavior highlight the importance of understanding how
 and why slope parameters vary at the ecosystem scale, as argued by
@@
   \item Tropical and temperate climates would generally host plants
-    with a higher g$_1$ than artic and boreal climates.
+    with a higher g$_1$ than arctic and boreal climates.
 \end{itemize}
 For all of the above relationships, a higher g$_1$ means a higher
-liklihood of a postive ET response to increasing VPD (water
+likelihood of a positive ET response to increasing VPD (water
 intensive), and a smaller VPD dependence of the ET response.
@@
 Our theoretical results, when framed by previous literature tying
-parametets to plant types and climate, highlight the variability of ET
+parameters to plant types and climate, highlight the variability of ET
 response to VPD as a function of plant characteristics. Whether an
@@
 
-Additionally, the sensitivity of ET reponse to plant parameter
+Additionally, the sensitivity of ET response to plant parameter
 highlights the importance of understanding and developing stomatal
@@
 particular, why and how g$_1$ behavior at the leaf and ecosystem scale
-are not directly analgalous much be understood. While the derivation
+are not directly analogous much be understood. While the derivation
 here was explicit about the assumptions of a constant g$_1$ term with
-respect to the VPD perturbation, many sophsiticated land surface
+respect to the VPD perturbation, many sophisticated land surface
 models and earth system models employing stomatal conductance models
-make similiar, and even stronger assumptions about the stationarity of
+make similar, and even stronger assumptions about the stationarity of
 VPD slope terms. In particular, many models will assume a constant
@@
 or difficulty in the literature with quantifying an ecosystem scale
-slope paramter \citep{Medlyn_2017}. Given that something so
-fundamental as the ET resposne to VPD varies strongly with g$_1$, we
-must invenst more in understanding g$_1$'s behavir at the ecosystem
+slope parameter \citep{Medlyn_2017}. Given that something so
+fundamental as the ET response to VPD varies strongly with g$_1$, we
+must invest more in understanding g$_1$'s behavior at the ecosystem
 scale, and the efficacy of models using g$_1$'s for ecosystem-scale
@@
 choice of stomatal conductance model. We now examine how the model
-choice, and speicifcally the exponent of the VPD dependence in teh
+choice, and specifically the exponent of the VPD dependence in the
 stomatal conductance model, can impact the general form of the ET-VPD
 relationship. This is relevant as many models include an
-exponententiated VPD term \citep[e.g.,][]{Leuning_1990}, but the value
+exponentiated VPD term \citep[e.g.,][]{Leuning_1990}, but the value
 of the exponent can vary. It is useful to understand how generally the

### 2019-02-17 doc/paper/conclusions.tex
spell check done and makefile updated
now proofread to send off to Pierre and Changjie

@@
 ecosystem will tend to reduce or increase ET with increasing VPD,
-which we framed using previous literature exploring the relatiohsip
+which we framed using previous literature exploring the relationship
 between plant parameters, plant types, and climate. The goal was to
@@
 intensive (ET always increasing in response to VPD), which is
-idicative of the large observed diversity of stomatal regulation and
-plany physiological strategy. Higher uWUE and g$_1$ values make the ET
+indicative of the large observed diversity of stomatal regulation and
+plant physiological strategy. Higher uWUE and g$_1$ values make the ET
 response more likely to be positive, while decreasing temperature and
-increasing g$_a$ amplify the magnitue of the response. Previous
+increasing g$_a$ amplify the magnitude of the response. Previous
 literature \citep{Zhou_2015, Lin_2015} suggests that crops, through
 association with higher g$_1$ and uWUE, are more likely to exhibit a
-postiive ET response to VPD. Shrubs (lower uWUE), C4 plants (lower
-g$_1$), gymnosperm trees (lower g$_1$), and plants in a artic and
+positive ET response to VPD. Shrubs (lower uWUE), C4 plants (lower
+g$_1$), gymnosperm trees (lower g$_1$), and plants in a arctic and
 boreal climates (lower g$_1$ are more likely to exhibit a negative ET
@@
 concave upward, with ET increasing with increasing VPD as VPD
-increases past some critical value **Add VPD_crit plot to
+increases past some critical value **Add VPD crit plot to
 results??**. However future research should focus on fully
@@
 VPD. Here, we provide a framework for understanding ET response to ET
-using a simplfied model fo two plant parameters. Feedbacks between the
+using a simplified model of two plant parameters. Feedbacks between the
 land and the atmosphere may alter the net response to a long-timescale

### 2019-02-27 doc/paper/abstract.tex
about to start responding to pierre's comments

@@
 
-The ET response varies due to: 1) climate, with tropical and
-temperate climates more likely to exhibit a positive ET response than
-boreal and arctic climates; 2) photosynthesis strategy, with C3 plants
-more likely to exhibit a positive ET response than C4 plants; and 3)
-due to plant type, with crops more likely to exhibit a positive ET
-response, and shrubs and gymniosperm trees more likely to exhibit a
-negative ET response. These results, derived from previous literature
-connecting plant parameters to plant and climate characteristics,
-highlight the utility of our simplified framework for understanding
-complex land atmosphere systems in terms of idealized scenarios.
+The ET response varies due to: 1) climate, with tropical and temperate
+climates more likely to exhibit a positive ET response to increasing
+VPD than boreal and arctic climates; 2) photosynthesis strategy, with
+C3 plants more likely to exhibit a positive ET response than C4
+plants; and 3) due to plant type, with crops more likely to exhibit a
+positive ET response, and shrubs and gymniosperm trees more likely to
+exhibit a negative ET response. These results, derived from previous
+literature connecting plant parameters to plant and climate
+characteristics, highlight the utility of our simplified framework for
+understanding complex land atmosphere systems in terms of idealized
+scenarios.

### 2019-02-27 doc/paper/body.tex
about to start responding to pierre's comments

@@
 
-The Medlyn model for stomatal conductance has been shown to behave
-very well across PFTs \citep[][]{Lin_2015}, and has been successfully
-adopted to ecosystem scale analysis in \citet{Medlyn_2017}. In units
-of m s$^{-1}$, the ecosystem scale stomatal conductance is:
+The Medlyn model adopted to ecosystem scale analysis in
+\citet{Medlyn_2017}. In units of m s$^{-1}$, the ecosystem scale
+stomatal conductance is:
 
@@
   where GPP is the ecosystem scale gross primary production, and $g_1$
-  is an ecosystem scale analogue to $g_{1-leaf}$. We solve for $g_1$
-  following \citet{Medlyn_2017} (Eq. (5)), and take the median $g_1$
-  value to be representative of each PFT (Table \ref{pft}) instead of
-  the mean to avoid extra weighting of rare outliers.
+  is an ecosystem scale analogue to $g_{1-leaf}$.
 
-While Eq. (3) can be used in PM (Eq. (1)), it will make
+While Eq. (\ref{medlyn}) can be used in PM (Eq. (1)), it will make
 analytical work with the function intractable because $GPP$ is
@@
   is relatively constant across time and moisture conditions within a
-plant functional type, and correctly captures a constant relationship
-between GPP, ET and VPD over a diurnal cycle \citep{Zhou_2014}. The
-theoretical derivation of the square root VPD dependence in $uWUE$
-leverages the same assumptions used in \cite{MEDLYN_2011} to derive the
-square-root VPD dependence of the stomatal conductance model (Eq.
-(\ref{medlyn})).  We can use uWUE to remove the $GPP$ dependence from
-$g_s$ in a way that makes PM analytically tractable:
+  plant functional type, and correctly captures a constant
+  relationship between GPP, ET and VPD over a diurnal cycle during the
+  growing season \citep{Zhou_2014}. The theoretical derivation of the
+  square root VPD dependence in $uWUE$ leverages the same assumptions
+  used in \cite{MEDLYN_2011} to derive the square-root VPD dependence
+  of the stomatal conductance model (Eq.  (\ref{medlyn})).  We can use
+  uWUE to remove the $GPP$ dependence from $g_s$ in a way that makes
+  PM analytically tractable:
 
@@
 
-With Equation \ref{eq} we can take the partial derivative with respect
+With Equation \ref{et} we can take the partial derivative with respect
 to VPD:
@@
   conditions and plant types the leaves can become decoupled from the
-  boundary layer \citep{De_2017, Medlyn_2017}. Therefor, our
-  derivation will be most applicable in times like the growing season,
-  when relatively high insolation induces instability and convective
-  boundary layers, and we would expect the surface to be generally
-  well coupled. An additional assumption in the formulation of $uWUE$
-  \citep{Zhou_2014, Zhou_2015} and \citet{Medlyn_2017}'s stomatal
-  conductance model is that direct soil evaporation (E) contributions
-  to ET remain small relative to transpiration (T). This should be
-  more true during the growing season. The ratio of E to T may
-  increase immediately after rainfall events due to high soil moisture
-  and ponding, but VPD is generally low anyways during these
-  times. However, some plant types allow for systematically larger
-  contributions of E in ET, particularly those with sparse canopies
-  and smaller relative amounts of transpiration. We therefore might
-  expect that the theory will be most applicable to forest PFTs, which
-  will be most strongly coupled to the boundary layer due to larger
-  surface roughness, and will also generally have the highest ratios
-  of transpiration to evaporation. Finally, we assume that $g_1$ and
-  $uWUE$ are constant with respect to the conceptual VPD
-  perturbation. Both quantities have been shown to be relatively
-  constant with respect to changes in
+  boundary layer \citep{De_2017, Medlyn_2017}. Therefore, our
+  derivation will be most applicable in times like the growing season
+  (when we also expect uWUE to be most valid), when relatively high
+  insolation induces instability and convective boundary layers, and
+  we would expect the surface to be generally well coupled. An
+  additional assumption in the formulation of $uWUE$ \citep{Zhou_2014,
+    Zhou_2015} and \citet{Medlyn_2017}'s stomatal conductance model is
+  that direct soil evaporation (E) contributions to ET remain small
+  relative to transpiration (T). This should be more true during the
+  growing season. The ratio of E to T may increase immediately after
+  rainfall events due to high soil moisture and ponding, but VPD is
+  generally low anyways during these times. However, some plant types
+  allow for systematically larger contributions of E in ET,
+  particularly those with sparse canopies and smaller relative amounts
+  of transpiration. We therefore might expect that the theory will be
+  most applicable to forest PFTs, which will be most strongly coupled
+  to the boundary layer due to larger surface roughness, and will also
+  generally have the highest ratios of transpiration to
+  evaporation. Finally, we assume that $g_1$ and $uWUE$ are constant
+  with respect to the conceptual VPD perturbation. Both quantities
+  have been shown to be relatively constant with respect to changes in
   % TODO: make this a general discussion on partial derivatives and
   % what things are constant/how to interpret?
-  VPD \cite{Franks_2017, Zhou_2014}. These parameters will however
+  VPD \citep{Franks_2017, Zhou_2014}. These parameters will however
   vary with plant species and characteristics \citep[e.g. wood density,
-  ]{Lin_2015}, as well as environmental conditions including soil
+  ][]{Lin_2015}, as well as environmental conditions including soil
   water content and temperature.
@@
   % \item
-
   To account for both this variability in plant parameters and the
@@
 
-and a ``scaling'' multiplying the ``sign'' term:
+and a ``scaling'' term multiplying the ``sign'' term:
 
@@
 
-In the ``sign term'' most of quantities are relatively constant,
+In the ``sign term'' most of the quantities are relatively constant,
 except for VPD and the plant parameters g$_1$ and uWUE. In the scaling
@@
 through $\Delta$). To determine the range of probable ET responses to
-VPD we will systematically vary this parameters according to Table
+VPD we will systematically vary these parameters according to Table
 \ref{param_varying}, while all other parameters are held fixed (Table
@@
     provided in cases where the quantities are directly derived from
-    previous literature. Conceptual percentiles
-    for interpreting each quantity are noted.}
+    previous literature. Conceptually, min. values are extracted from
+    literature to correspond to approximately the 15th percentile of
+    observed conditions during the growing season, med. values
+    correspond to approximately the 50th percentile, and max. values
+    correspond to approximately the 85th percentile.}
   \label{param_varying}
@@
     \hline
-    Symbol & units [units in citation] & min (~15th percentile) & med (~50th percentile) & max
-                                                                                            (~15th
-                                                                                            percentile)
+    Symbol & units [units in citation] & min  & med & max
     & citation  \\
@@
     (relative to those in Table \ref{param_varying}).}
-  \label{param_varying}
+  \label{param_fixed}
   \centering
@@
   \centering \includegraphics{./fully_idealized.pdf}
-  \caption{The functional form of the $\frac{\partial \; ET}{\partial
+  \caption{The functional form of $\frac{\partial \; ET}{\partial
       \; VPD}$ for minimum, median and maximum values of both the sign
@@
 
-Examining the sign term independent of the scaling term indicates the
-dependence of water conservative versus water intensive response on
-the main plant parameters, g$_1$ and uWUE (Figure \ref{sign}). Higher
-g$_1$ and uWUE shift the curve towards increasing ET responses with
-VPD (water intensive), and with increasing g$_1$ also leading to a
-higher non-linear VPD dependence of the response.
+Examining the sign term independent of the scaling term illuminates
+the role of the two plant parameters, g$_1$ and uWUE, in determining
+the degree to which the response is water conservative
+($\frac{\partial \; ET}{\partial \; VPD} < 0$) or water intensive
+($\frac{\partial \; ET}{\partial \; VPD} > 0$) (Figure \ref{sign})
+. Higher g$_1$ and uWUE shift the curve towards increasing ET
+responses with VPD (water intensive), and smaller g$_1$ values lead to
+a higher non-linear VPD dependence of the response.
 
@@
   \centering \includegraphics{./fully_idealized_sign.pdf}
-  \caption{The functional form of the $\frac{\partial \; ET}{\partial
+  \caption{The functional form of $\frac{\partial \; ET}{\partial
       \; VPD}$ evaluated at the median value of the scaling term, for
@@
 
+\begin{figure}
+  \centering \includegraphics{./fully_idealized_scaling.pdf}
+  \caption{The functional form of $\frac{\partial \; ET}{\partial
+      \; VPD}$ evaluated at the median value of the sign term, for
+    varying values of g$_a$ and T as given in Table \ref{param_varying}.}
+  \label{scaling}
+\end{figure}
+
 Examining the scaling term independent of the sign term shows how
 aerodynamic conductance (communication between the atmosphere and the
-surface) and temperature (through $\Delta$ controls the efficiency of
-energy conversion to latent heat), amplify or suppress the plant
-physiological response represented in the sign term. Both lower
+surface) and temperature (controls the efficiency of
+energy conversion to latent heat through $\Delta$), amplify or suppress the plant
+physiological response represented in the sign term (Figure \ref{scaing}). Both lower
 temperatures and higher aerodynamic conductance lead to amplified ET
@@
 resulting in a slightly higher variability of ET response relative to
-temperature, which is still an important factor.
+temperature.
 
@@
 useful, and we would argue is critical, for understanding complex
-responses when many quantities are varying simultaneously. Through
-partial derivatives, we can approach answering questions like ``What
-is the relative contribution of VPD in some observed response given
-that X Y and Z were also varying?'' In this example, X Y and Z could
-be, for example, radiation, soil moisture, and temperature. Equation
-\ref{d_et} explicitly provides an estimate of the ET response to VPD,
-given assumptions about ecosystem plant characteristics through the
-parameters g$_1$ and uWUE. Equation \ref{d_et} is generic to external
+responses when many quantities are varying simultaneously. If we
+cannot understand the contribution of one term to observed
+variability, independent of other confounding factors, there is little
+hope of disentangling and understanding the relative role of many
+quantities in determining observed variability.  Equation \ref{d_et}
+explicitly provides an estimate of the ET response to VPD, given
+assumptions about ecosystem plant characteristics through the
+parameters g$_1$ and uWUE. Eq. (\ref{d_et}) is generic to external
 environmental factors and the timescale over which a VPD response
 acts, and has a straightforward conceptual meaning. It provides the ET
-response, given other quantities and parameters held fixed. So, while
-the plant parameters g$_1$ and uWUE may vary with plant species and
+response, given other quantities and parameters held fixed, subject to
+the assumptions outlined in Sect. \ref{methods}. So, while the plant
+parameters g$_1$ and uWUE may vary with plant species and
 environmental conditions like soil moisture, we can still assess the
@@
 
-However, given that g$_1$ and uWUE can vary, and that they both have a
-large impact on the ET response (Figure \ref{full}, \ref{sign}) it is
+However, given that g$_1$ and uWUE can vary, and they both have a
+large impact on the ET response (Figure \ref{full}, \ref{sign}), it is
 useful to examine how these quantities vary in previous literature,
-framed by our results on ET responses. In \citet{Zhou_2015} most plant
-types have uWUE similar to our median value, where whether ET
-response is positive (water intensive) or negative (water conservative)
-depends on VPD. However, crops, which we expect to prioritize carbon
-uptake over water conservation, have a higher uWUE closer to our
-maximum value. This higher uWUE results in a higher likelihood for an
+framed by our results on $\frac{\partial \; ET}{\partial \; VPD}$. In
+\citet{Zhou_2015} most plant types have uWUE similar to our median
+value, under which the sign of the ET response depends on
+VPD. However, crops, which we expect to prioritize carbon uptake over
+water conservation, have a higher uWUE closer to our maximum
+value. This higher uWUE results in a higher likelihood for an
 increasing ET response to VPD, which intuitively makes sense given
 that we expect crops to keep stomata open for access to carbon, at the
-cost of water loss. Shrubs, and to a lesser extent grass, gave an uWUE
-closer to the minimum uWUE. For these plant types, we would then
-expect a higher likelihood of a water conservative response; that is, a
-decrease of ET in response to VPD. It is important to note that, while
-\citet{Zhou_2015} did not examine the roll of soil moisture in within
-plant-type variability of uWUE, we might expect some variation,
-especially in extreme cases when soil water becomes a limiting factor.
+cost of increased water loss during high VPD. Shrubs, and to a lesser
+extent grass, have an uWUE closer to the minimum uWUE. For these plant
+types, we would then expect a higher likelihood of a water
+conservative response; that is, a decrease of ET in response to
+VPD. It is important to note that, while \citet{Zhou_2015} did not
+examine the role of soil moisture for within plant-type variability of
+uWUE, we might expect some variation, especially in extreme cases when
+soil water becomes a limiting factor.
 
@@
 observations, the consequences of imposing a model structure in some
-of the g$_1$ observations, or real ecosystem-scale within-plant type
+of their g$_1$ esimations, or real ecosystem-scale within-plant type
 variability. These differences between leaf-scale behavior and
 ecosystem scale behavior highlight the importance of understanding how
-and why slope parameters vary at the ecosystem scale, as argued by
-\citet{Medlyn_2017}. However, if we assume that there is some real
-analogy between leaf scale results in \citet{Lin_2015} and ecosystem
-scale behavior, and that some of the ambiguity in \citet{Medlyn_2017}
-is due to model and/or observational error, then we can make some
-connections between g$_1$-induced ET response variability and plant
-characteristics, which are subject to the strong caveat that we still
-do not fully understand the g$_1$'s utility at the ecosystem
-scale. Extrapolating \citet{Lin_2015}'s results (e.g. Fig. 2) to the
-ecosystem scale would reveal the following relationships between g$_1$
-and plant types:
+and why slope parameters show increased variability at the ecosystem
+scale, as argued by \citet{Medlyn_2017}. However, if we assume that
+there is some real analogy between leaf scale results in
+\citet{Lin_2015} and ecosystem scale behavior, and that some of the
+ambiguity in \citet{Medlyn_2017} is due to model and/or observational
+error, then we can make some connections between g$_1$-induced ET
+response variability and plant characteristics, which are subject to
+the strong caveat that we still do not fully understand g$_1$'s
+utility at the ecosystem scale. Extrapolating \citet{Lin_2015}'s
+results (e.g. Fig. 2) to the ecosystem scale would reveal the
+following relationships between g$_1$ and plant types:
 
@@
 
-Additionally, the sensitivity of ET response to plant parameter
+Additionally, the sensitivity of ET response to plant parameters
 highlights the importance of understanding and developing stomatal
@@
 are not directly analogous much be understood. While the derivation
-here was explicit about the assumptions of a constant g$_1$ term with
+here was explicit about the assumption of a constant g$_1$ term with
 respect to the VPD perturbation, many sophisticated land surface
@@
 make similar, and even stronger assumptions about the stationarity of
-VPD slope terms. In particular, many models will assume a constant
-g$_1$ term within a given plant functional type, and research derived
-from these models will not acknowledge limitations of this assumption,
-or difficulty in the literature with quantifying an ecosystem scale
-slope parameter \citep{Medlyn_2017}. Given that something so
-fundamental as the ET response to VPD varies strongly with g$_1$, we
-must invest more in understanding g$_1$'s behavior at the ecosystem
-scale, and the efficacy of models using g$_1$'s for ecosystem-scale
-fluxes.
+VPD slope terms. Many models will assume a constant g$_1$ term within
+a given plant functional type, and research derived from these models
+often does not acknowledge limitations of this assumption, or the
+difficulty in the literature with quantifying an ecosystem scale slope
+parameter \citep{Medlyn_2017}. Given that something so fundamental as
+the ET response to VPD varies strongly with g$_1$, we must invest more
+in understanding g$_1$'s behavior at the ecosystem scale, and the
+efficacy of models using g$_1$'s for ecosystem-scale
+fluxes. Fortunately, our framework, and the novel general approach of
+using robust ecosystem-scale ratios (uWUE) to remove the implicit
+dependence of stomatal conductance on ET, is flexible enough to be
+applied to any future-derived ecosystem-scale stomatal conductance
+model that contains a dependence on GPP.
 
-So far we have discussed the role of different parameters in the ET
-response. However, the form of the ET response is imposed by our
-choice of stomatal conductance model. We now examine how the model
-choice, and specifically the exponent of the VPD dependence in the
-stomatal conductance model, can impact the general form of the ET-VPD
-relationship. This is relevant as many models include an
-exponentiated VPD term \citep[e.g.,][]{Leuning_1990}, but the value
-of the exponent can vary. It is useful to understand how generally the
+So far we have discussed the role of different parameters in
+modulating the ET response. However, the form of the ET response is
+imposed by our choice of stomatal conductance model. We now examine
+how the model choice, and specifically the exponent of the VPD
+dependence in the stomatal conductance model, can impact the general
+form of the ET-VPD relationship. This is relevant as many models
+include an exponentiated VPD term \citep[e.g.,][]{Leuning_1990}, but
+the value of the exponent can vary. It is useful to understand how the
 choice of model imposes the functional form of the ET-VPD
@@
 interaction.
+
 \subsection{Functional form of ET dependence on VPD and its relation
@@
 
-The theory described in Sect. \ref{sign_term} indicates that for a
+The results presented in Sect. \ref{results} indicates that for a
 given $uWUE$ and $g_1$, the ET dependence on VPD should be concave
-upward, which is confirmed by eddy covariance data across most PFTs. In other words, there should be some local minimum in ET at a
+upward. In other words, there should be some local minimum in ET at a
 critical VPD$_{crit}$, assuming the scaling and plant terms  (e.g. aerodynamic

### 2019-02-27 doc/paper/conclusions.tex
about to start responding to pierre's comments

@@
 intensive (ET always increasing in response to VPD), which is
-indicative of the large observed diversity of stomatal regulation and
-plant physiological strategy. Higher uWUE and g$_1$ values make the ET
-response more likely to be positive, while decreasing temperature and
-increasing g$_a$ amplify the magnitude of the response. Previous
-literature \citep{Zhou_2015, Lin_2015} suggests that crops, through
-association with higher g$_1$ and uWUE, are more likely to exhibit a
-positive ET response to VPD. Shrubs (lower uWUE), C4 plants (lower
-g$_1$), gymnosperm trees (lower g$_1$), and plants in a arctic and
-boreal climates (lower g$_1$ are more likely to exhibit a negative ET
+indicative of the diversity of possible plant water conservation
+strategies. Higher uWUE and g$_1$ values increase the likelihood of a
+postive ET response, while decreasing temperature and increasing
+g$_a$ amplify the magnitude of the response. Previous literature
+\citep{Zhou_2015, Lin_2015} suggests that crops, through association
+with higher g$_1$ and uWUE, are more likely to exhibit a positive ET
+response to VPD. Shrubs (lower uWUE), C4 plants (lower g$_1$),
+gymnosperm trees (lower g$_1$), and plants in arctic and boreal
+climates (lower g$_1$) are more likely to exhibit a negative ET
 response to VPD. However interpretation of g$_1$-induced variability
@@
 analysis of plant-VPD feedbacks in the atmospheric boundary layer
-(Eq. (\ref{full})). Our approach can be extended to examine varying
+(Eq. (\ref{et})). Our approach can be extended to examine varying
 plant response to more nuanced consideration of plant type and

### 2019-02-27 doc/paper/intro.tex
about to start responding to pierre's comments

@@
 atmospheric demand for water from the land surface, and this drives an
-increase in evapotranspiration (ET). This perspective is particularly
-relevant because potential evapotranspiration (PET), which is used in
-many drought indices and hydrometeorological studies
-\citep[e.g.,][]{Heim_2002, Scheff_2015}, typically only quantifies
-changes in atmospheric demand and fails to account for ecosystem
-response \citep{Swann_2016}. In reality, plants' stomata have evolved
-to optimally regulate the exchange of water and carbon, and tend to
-partially close in response to increased atmospheric dryness
+increase in evapotranspiration (ET). However, plants' stomata have
+evolved to optimally regulate the exchange of water and carbon, and
+tend to partially close in response to increased atmospheric dryness
 \citep{Farquhar_1978, Ball_1987, Leuning_1990, MEDLYN_2011}. This
-leads to a plant physiology perspective, in which an increase in VPD,
-particularly in well-watered soil conditions, may actually correspond
-to a decrease in ET because of stomatal closure
-\citep[e.g.][]{Rigden_2017}.  In other words, the question ``When does
-VPD drive or reduce ET?'' can be related to whether plant regulation
-or atmospheric demand dominates ET response.
+leads to a plant physiology perspective, in which an increase in VPD
+may actually correspond to a decrease in ET because of stomatal
+closure \citep[e.g.][]{Rigden_2017}.  In other words, the question
+``When does VPD drive or reduce ET?'' can be related to whether plant
+regulation or atmospheric demand dominates ET response.
 
@@
 
-As a counterpoint, one may argue that increases in ET with increasing
-VPD could increase the likelihood of precipitation
-\citep[e.g.,][]{Findell_2011}. However, increases in ET do not always
-guarantee an increased likelihood of precipitation, which depending on
-environmental conditions could cause a decrease in the likelihood of
-precipitation \citep[][]{Gentine_2013}. Furthermore, any increases in
-precipitation are likely to be non-local, such that plants giving up
-water to the atmosphere are not guaranteed to reap the benefits of
-water returned from the atmosphere to the soil. Using this subjective
-logic, from an ecosystem evolutionary perspective, water stored in
-soil seems to be worth much more than the chance of water returned as
-precipitation.
-
 We can use intuition about plant water conservation strategy to
 hypothesize about ET response to changes in VPD. Plants and ecosystems
-that evolved to conserve water, such as arid shrubs or savannah,
-should be more likely to reduce ET with increasing VPD, and plants
-that have evolved or have been engineered to care little about water,
-such as crops, will be more likely to increase ET with increasing
-VPD. Atmospheric conditions must matter as well. At the ecosystem
-scale, there are limits to plant water conservation strategies. As
-atmospheric demand for water (VPD) increases, ecosystems should begin
-to reach their water conservation limits and might not be able to
-entirely limit ET flux to the atmosphere. At this stage any further
-increase in VPD will most likely drive a (limited) increase in ET,
-because the increase in atmospheric demand for water overwhelms the
-limited plant response to conserve water.
+that evolved to conserve water, such as arid shrubs, should be more
+likely to reduce ET with increasing VPD, and plants that have evolved
+or have been engineered to care little about water, such as crops,
+will be more likely to increase ET with increasing VPD. Atmospheric
+conditions must matter as well. At the ecosystem scale, there are
+limits to plant water conservation strategies. As atmospheric demand
+for water (VPD) increases, ecosystems should begin to reach their
+water conservation limits and might not be able to entirely limit ET
+flux to the atmosphere. At this stage any further increase in VPD will
+most likely drive a (limited) increase in ET, because the increase in
+atmospheric demand for water overwhelms the limited plant response to
+conserve water.
 
@@
 scale. uWUE is also relatively well conserved in the growing season
-across space and time, within a PFT \citep{Zhou_2015}. While stomatal
-conductance parameterizations and uWUE greatly simplify complex plant
-physiological processes, they still capture ecosystem behavior for
-vegetated surfaces \citep{Medlyn_2017, Zhou_2014}, and are novel tools
-to transparently develop intuition for the behavior of complex,
-multiscale ecohydrologic systems.
+across space and time, within a plant functional type (PFT)
+\citep{Zhou_2015}. While stomatal conductance parameterizations and
+uWUE greatly simplify complex plant physiological processes, they
+still capture ecosystem behavior for vegetated surfaces
+\citep{Medlyn_2017, Zhou_2014}, and are novel tools to transparently
+develop intuition for the behavior of complex, multiscale
+ecohydrologic systems.
 
@@
 highlighting the importance of: 1) future research on stomatal
-conductance and ecosystem scaling, and 2) thoughtful selection of
-photosynthesis and stomatal conductance model in more sophisticated
-land surface and earth system models.
+conductance and ecosystem scale modeling, and 2) thoughtful selection
+of photosynthesis and stomatal conductance models in more
+sophisticated land surface and earth system models.

### 2019-02-27 doc/paper/intro.tex
progress through intro on pierres comments
also set up site specific SWC box plots

@@
 for evapotranspirated water \citep{Penman_1948, Monteith_1965}, but
-also stress plant stomata \citep{Leuning_1990, MEDLYN_2011}.
+also stress plant stomata \citep{Leuning_1990,
+MEDLYN_2011}. Understanding the net evaportanspiration (ET) response
+to these two opposing effects of changes in VPD is crucial for
+assessing the impact of environmental perturbations to VPD on the
+water cycle.
 
 The opposing effects of increased atmospheric demand and higher
-stomatal stress lead to two possible perspectives for how
-evapotranspiration (ET) responds to shifts in VPD. The first, a
-hydrometeorological perspective, is that higher VPD increases
-atmospheric demand for water from the land surface, and this drives an
-increase in evapotranspiration (ET). However, plants' stomata have
-evolved to optimally regulate the exchange of water and carbon, and
-tend to partially close in response to increased atmospheric dryness
-\citep{Farquhar_1978, Ball_1987, Leuning_1990, MEDLYN_2011}. This
+stomatal stress lead to two possible perspectives for how ET responds
+to shifts in VPD. The first, a hydrometeorological perspective, is
+that higher VPD increases atmospheric demand for water from the land
+surface, and this drives an increase in ET
+\citep{Penman_1948}. However, plants' stomata have evolved to
+optimally regulate the exchange of water and carbon, and tend to
+partially close in response to increased atmospheric dryness
+\citep{Farquhar_1978, Ball_1987, Leuning_1990, Katul_2009, MEDLYN_2011}. This
 leads to a plant physiology perspective, in which an increase in VPD
@@
 ``When does VPD drive or reduce ET?'' can be related to whether plant
-regulation or atmospheric demand dominates ET response.
+regulation or atmospheric demand dominates the ET response.
 
@@
 \citep[][]{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006,
-  Salvucci_2013, Gentine_2016, Berg_2016}. Therefore, passive
+Salvucci_2013, Gentine_2016, Berg_2016, Zhou_2019}. Therefore, passive
 regulation and a lack of soil moisture conservation does not seem to
 be a sensible strategy for plants from an evolutionary standpoint.
+
+** paragraph on soil moisture dependenec? **
 
@@
 likely to reduce ET with increasing VPD, and plants that have evolved
-or have been engineered to care little about water, such as crops,
-will be more likely to increase ET with increasing VPD. Atmospheric
-conditions must matter as well. At the ecosystem scale, there are
-limits to plant water conservation strategies. As atmospheric demand
-for water (VPD) increases, ecosystems should begin to reach their
-water conservation limits and might not be able to entirely limit ET
-flux to the atmosphere. At this stage any further increase in VPD will
-most likely drive a (limited) increase in ET, because the increase in
-atmospheric demand for water overwhelms the limited plant response to
-conserve water.
+or have been engineered to prioritize carbon gain over water
+conservation, such as crops, will be more likely to increase ET with
+increasing VPD. Atmospheric conditions must matter as well. At the
+ecosystem scale, there are limits to plant water conservation
+strategies. As atmospheric demand for water (VPD) increases,
+ecosystems may begin to reach their water conservation limits and
+might not be able to entirely limit ET flux to the atmosphere. At this
+stage any further increase in VPD will most likely drive a (limited)
+increase in ET, because the increase in atmospheric demand for water
+overwhelms the limited plant response to conserve water.
 
@@
 framework for understanding plant response to atmospheric drying and
-evaluating the VPD dependence of ET. This framework will aid
-interpretation of observations, full complexity models, and facilitate
-the disentanglement of complex land-atmosphere feedbacks. In the past,
-similar approaches were used to understand interactions between
-stomatal conductance, evapotranspiration and the environment
-\citep[e.g.,][]{Jarvis_1986, Mcnaughton_1991}. However, at the time
-researchers' understanding of the form of VPD's effect on plant
-physiology was limited, so they could not explore the sensitivity of
-ET to VPD, including VPD's effect on stomatal conductance and plant
-function.
+evaluating the VPD dependence of ET while keeping other variables
+fixed. This framework will aid interpretation of observations, full
+complexity models, and facilitate the disentanglement of complex
+land-atmosphere feedbacks. In particular, we can understand the impact
+of climate change, given the expected increase of VPD.  In the past,
+similar simplified approaches were used to understand interactions
+between stomatal conductance, evapotranspiration and the environment
+\citep[e.g.,][]{Jarvis_1984, Jarvis_1986, Mcnaughton_1991}. However,
+at the time researchers' understanding of the form of VPD's effect on
+plant physiology was limited, so they could not systematically explore
+the sensitivity of ET to VPD, including VPD's effect on stomatal
+conductance and plant function.
 
@@
 conductance ($g_s$), including VPD response, by combining an optimal
-photosynthesis theory \citep{Cowan_1977} with an empirical approach,
-and extended use of this model to the ecosystem scale in
+photosynthesis theory \citep{Cowan_1977, Katul_2009} with an empirical
+approach, and extended use of this model to the ecosystem scale in
 \citet{Medlyn_2017}. Additionally, \citet{Zhou_2014} demonstrated that
@@
 across space and time, within a plant functional type (PFT)
-\citep{Zhou_2015}. While stomatal conductance parameterizations and
-uWUE greatly simplify complex plant physiological processes, they
-still capture ecosystem behavior for vegetated surfaces
-\citep{Medlyn_2017, Zhou_2014}, and are novel tools to transparently
-develop intuition for the behavior of complex, multiscale
-ecohydrologic systems.
+\citep{Zhou_2015}. ** here talk about difference between leaf level
+and ecosystem scale, also connect uWUE to g1 ** While stomatal
+conductance parameterizations and uWUE greatly simplify complex plant
+physiological processes, they therefore still capture ecosystem
+behavior for vegetated surfaces \citep{Medlyn_2017, Zhou_2014}, and
+are novel tools to transparently develop intuition for the behavior of
+complex land-atmosphere systems.
 
@@
 variables properly controlled for, i.e. we develop a framework for
-evaluating the partial derivative of ET with respect to VPD. For the
-first time, we explicitly include VPD's full effect on stomatal
-conductance, including its impact on photosynthesis. We explore the
-range of possible ET responses to VPD, given parameters previously
-established in peer reviewed literature. Additionally, we show the
-sensitivity of the ET-VPD relationship to model and framework choice,
-highlighting the importance of: 1) future research on stomatal
-conductance and ecosystem scale modeling, and 2) thoughtful selection
-of photosynthesis and stomatal conductance models in more
-sophisticated land surface and earth system models.
+evaluating the partial derivative of ET with respect to VPD. We
+explicitly include VPD's full effect on stomatal conductance,
+including its impact on photosynthesis. We explore the range of
+possible ET responses to VPD, given parameters previously established
+in peer reviewed literature. Additionally, we show the sensitivity of
+the ET-VPD relationship to model and framework choice, highlighting
+the importance of: 1) future research on stomatal conductance and
+ecosystem scale modeling, and 2) thoughtful selection of
+photosynthesis and stomatal conductance models in more sophisticated
+land surface and earth system models. Finally, we discuss the caveats
+of our approach and the sensitivity of results to assumptions about
+stomatal conductance.

### 2019-02-27 src/paper_figure_code/swc_boxplot.py

@@ -40,7 +40,7 @@ def plot_box(_df, ax=None, pft=False):
     labels.append('%4.0f' % positions[-1])
   ax.boxplot(x=data, positions=positions,\
              sym=(''), widths=widths, manage_xticks=False,\
-             whis=[5, 95], boxprops=boxprops, medianprops=medianprops)#sym=('')
+             whis=[10, 90], boxprops=boxprops, medianprops=medianprops)#sym=('')
   print('make hist for %s' % _df.pft.iloc[0])
   orig_xlim = ax.get_xlim()
   #ax.plot(orig_xlim, [d_calc.CP/287.0, d_calc.CP/287.0], 'm--', linewidth=1.0)
@@ -51,8 +51,8 @@ def plot_box(_df, ax=None, pft=False):
     ax.set_title(name_dict[_df.pft.iloc[0]],\
                  fontsize=fontsize+3)
   else:
-    ax.set_title('Site: %s, PFT: %s'\
-                 % (_df.site.iloc[0], _df.pft.iloc[0]),\
+    ax.set_title('Site: %s, PFT: %s, Nobs: %d'\
+                 % (_df.site.iloc[0], _df.pft.iloc[0], _df.shape[0]),\
                  fontsize=fontsize+3)
   # if _df.pft.iloc[0] in xlim:
   #   _xlims = xlim[_df.pft.iloc[0]]
@@ -63,17 +63,18 @@ def plot_box(_df, ax=None, pft=False):
   if savefig:
     plt.tight_layout()
     if pft:
-      plt.savefig('%s/climate_et/swc_box/pft/%s_box.png'\
+      safesavefig('%s/climate_et/swc_box/pft/%s_box.png'\
                   % (os.environ['PLOTS'], _df.pft.iloc[0]))
     else:
-      plt.savefig('%s/climate_et/swc_box/%s_%s_box.png'\
+      safesavefig('%s/climate_et/swc_box/%s_%s_box.png'\
                   % (os.environ['PLOTS'], _df.pft.iloc[0], _df.site.iloc[0]))
   return
 
-
+df_uwue_obs = df.copy()
+df_uwue_obs['uwue'] = df_uwue_obs['gpp_obs']/df_uwue_obs['et_obs']*np.sqrt(df_uwue_obs['vpd'])
 #_df = df.iloc[:1000, :].copy()
 #plot_box(_df)
-#df.groupby('site').apply(plot_box)
[…]

### 2019-02-27 doc/paper/body.tex
kidn of done with methods, need to decide where to put paragraph on soil moisture

@@
 constant, and $g_s$ is the stomatal conductance (Table
-\ref{definitions}).
+\ref{definitions}). The issue with PM is that $g_s$ is a function of
+carbon uptake, which is has a strong functionally relation to ET
+through stomatal function. Therefore, when PM is formulated in terms
+of $g_s$, it is an implicit funciton of ET as a function of itself
+rather than an explicit function of ET. Here we will derive a new form
+of PM in which ET is an explicit function of plant parameters and
+environmental conditions, and use it to assess the ecosystem scale
+response to VPD (by taking a partial derivative).
 
@@
 of leaf carbon
-($\frac{\partial \; \text{transpiration}}{\partial \; A}$). So,
+($\frac{\partial \; \text{transpiration}}{\partial \; A}$)
+\citep{Farquhar_1980, Katul_2009}. So,
 $g_{1-leaf}$ is a leaf-scale term reflecting the trade-off of water for
@@
   analytical work with the function intractable because $GPP$ is
-functionally related to ET itself. Additionally, a perturbation to VPD
-should induce a physiological plant response that will alter GPP and
-cause an indirect change in stomatal conductance, in addition to the
-direct effect of VPD in Eq. (\ref{medlyn}). Therefore, in order to
-derive the response of ET to VPD, we must account for the functional
-relationship between GPP, ET, and VPD, and its effect on stomatal
-conductance. We can use aforementioned semi-empirical results of
-\citet{Zhou_2015} as a tool to approach this
+  functionally related to ET itself. Additionally, a perturbation to
+  VPD should induce a physiological plant response that will alter GPP
+  and cause an indirect change in stomatal conductance, in addition to
+  the direct effect of VPD in Eq. (\ref{medlyn}). Therefore, in order
+  to derive the response of ET to VPD, we must account for the
+  functional relationship between GPP, ET, and VPD, and its effect on
+  stomatal conductance. We can use aforementioned semi-empirical
+  results of \citet{Zhou_2015}, which were inspired optimal
+  photosynthesis theory, as a tool to approach this
   problem. \citet{Zhou_2015}, showed that underlying Water Use
@@
 
-  is relatively constant across time and moisture conditions within a
-  plant functional type, and correctly captures a constant
-  relationship between GPP, ET and VPD over a diurnal cycle during the
-  growing season \citep{Zhou_2014}. The theoretical derivation of the
-  square root VPD dependence in $uWUE$ leverages the same assumptions
-  used in \cite{MEDLYN_2011} to derive the square-root VPD dependence
-  of the stomatal conductance model (Eq.  (\ref{medlyn})).  We can use
-  uWUE to remove the $GPP$ dependence from $g_s$ in a way that makes
-  PM analytically tractable:
+  is relatively constant across time within a plant functional type,
+  and correctly captures a constant relationship between GPP, ET and
+  VPD over a diurnal cycle during the growing season
+  \citep{Zhou_2014}. The theoretical derivation of the square root VPD
+  dependence in $uWUE$ leverages the same assumptions used in
+  \cite{MEDLYN_2011} to derive the square-root VPD dependence of the
+  stomatal conductance model (Eq.  (\ref{medlyn})).  We can use uWUE
+  to remove the $GPP$ dependence from $g_s$ in a way that makes PM
+  analytically tractable:
 
@@
 
-By accounting for photosynthesis changes in ecosystem conductance, with
-Eq. (\ref{et}) we have derived for the first time, using recent
-results \citep[][]{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017}, ET
-explicitly as function of environmental variables and two
-plant-specific constants, the slope parameter ($g_1$), and uWUE, both
-reflecting water conservation strategy. The slope parameter is related
-to the willingness of stomata to trade water for CO$_2$ and to keep
-stomata open. uWUE is a semi-empirical ecosystem-scale constant
-related to how WUE changes with VPD (specifically $VPD^{-1/2}$). It is
-also roughly proportional to physical constants:
+  By accounting for photosynthesis changes in ecosystem conductance,
+  with Eq. (\ref{et}) we have used recent results
+  \citep[][]{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017} to derive
+  ET explicitly as function of environmental variables and two
+  plant-specific constants, the slope parameter ($g_1$), and uWUE. For
+  the first time we have removed the implicit dependence of ET on
+  itself through the stomatal conductance term, and we have also
+  replaced the added complexity of a stomatal conductance reduction
+  factor and a photosythesis model with a signle paramter (uWUE). Both
+  the plant paramters reflect water conservation strategy.The
+  slope parameter is related to the willingness of stomata to trade
+  water for CO$_2$ and to keep stomata open (carbon cost in terhms of
+  water)). uWUE is a semi-empirical ecosystem-scale constant related
+  to how WUE changes with VPD (specifically $VPD^{-1/2}$). It is also
+  roughly proportional to physical constants:
 
@@
 
-With Equation \ref{et} we can take the partial derivative with respect
-to VPD:
+With Equation \ref{et} we can take the partial derivative of ET with respect
+to VPD to understand whether VPD drives or reduces ecosystem-scale ET:
 
@@
   deriving an analytical framework for ecosystem response to
-  atmospheric demand perturbations with environmental conditions held
+  atmospheric demand with environmental conditions held
   fixed. There are a few subtleties to taking the derivative in
@@
   that direct soil evaporation (E) contributions to ET remain small
-  relative to transpiration (T). This should be more true during the
-  growing season. The ratio of E to T may increase immediately after
-  rainfall events due to high soil moisture and ponding, but VPD is
-  generally low anyways during these times. However, some plant types
-  allow for systematically larger contributions of E in ET,
-  particularly those with sparse canopies and smaller relative amounts
-  of transpiration. We therefore might expect that the theory will be
-  most applicable to forest PFTs, which will be most strongly coupled
-  to the boundary layer due to larger surface roughness, and will also
-  generally have the highest ratios of transpiration to
-  evaporation. Finally, we assume that $g_1$ and $uWUE$ are constant
-  with respect to the conceptual VPD perturbation. Both quantities
-  have been shown to be relatively constant with respect to changes in
+  relative to transpiration (T). Again, this should be more true
+  during the growing season. The ratio of E to T may increase
+  immediately after rainfall events due to high soil moisture and
+  ponding, but VPD is generally low anyways during these
+  times. However, some plant types allow for systematically larger
+  contributions of E in ET, particularly those with sparse canopies
+  and smaller relative amounts of transpiration. We therefore might
+  expect that the theory will be most applicable to forest PFTs, which
+  will be most strongly coupled to the boundary layer due to larger
+  surface roughness, and will also generally have the highest ratios
+  of transpiration to evaporation. Finally, we assume that $g_1$ and
+  $uWUE$ are constant with respect to the conceptual VPD
+  perturbation. Both quantities have been shown to be relatively
+  constant with respect to changes in
   % TODO: make this a general discussion on partial derivatives and
@@
   to magnitude:
+
+  ** comment on temperature and acclimation??? **
+  ** impact of T is much smaller than VPD? **
+  ** paragraph on soil moisture **
 

### 2019-02-27 doc/paper/body.tex
need to find good citiations for ESM/LSM claim, email dan?

@@
 the sign term, and 81 different curves for the ET response to VPD
-($\frac{\partial \; ET}{\partial \; VPD}$). Examining a subset of nine
-of the curves for the ET response to VPD, defined by the minimum, median
-and maximum of for each of the scaling and sign term, it is clear
-that the range of responses includes those where ET almost always
-decreases with increasing VPD (water conservative), ET almost always
-increases with increasing VPD (water intensive), and intermediate
-values where ET will increase or decrease with VPD depending on the
-atmospheric demand for water (Figure \ref{full}).
+($\frac{\partial \; ET}{\partial \; VPD}$). To aid visualization we
+can examine a subset of nine of these 81 curves, defined by the
+minimum, median and maximum values for both the scaling and sign term
+(Figure \ref{full}). The range of ET responses to VPD vary from those
+where ET almost always decreases with increasing VPD (water
+conservative), to those where ET almost always increases with
+increasing VPD (water intensive). Additionally, for some parameters
+whether ET will increase or decrease with increasing VPD depends on
+atmospheric demand (Figure \ref{full}).
 
@@
 
+\begin{figure}
+  \centering \includegraphics{./fully_idealized_sign.pdf}
+  \caption{The functional form of $\frac{\partial \; ET}{\partial
+      \; VPD}$ evaluated at the median value of the scaling term, for
+    varying values of g$_1$ and uWUE as given in Table \ref{param_varying}.}
+  \label{sign}
+\end{figure}
+
 Examining the sign term independent of the scaling term illuminates
@@
 responses with VPD (water intensive), and smaller g$_1$ values lead to
-a higher non-linear VPD dependence of the response.
-
-\begin{figure}
-  \centering \includegraphics{./fully_idealized_sign.pdf}
-  \caption{The functional form of $\frac{\partial \; ET}{\partial
-      \; VPD}$ evaluated at the median value of the scaling term, for
-    varying values of g$_1$ and uWUE as given in Table \ref{param_varying}.}
-  \label{sign}
-\end{figure}
+a higher non-linear VPD dependence of the response. g$_1$ and uWUE
+should generally co-vary with $\lambda$. As $\lambda$ increases, g$_1$
+should increase and uWUE should decrease. ** ellaborate **
 
@@
 on an understanding and appreciation for the usefulness of partial
-derivatives for understanding the behavior of complex systems. Our
-results are a conceptual tool for answering the question, ``What is
-the ET response to VPD with all other quantities held fixed?'' This is
-useful, and we would argue is critical, for understanding complex
-responses when many quantities are varying simultaneously. If we
-cannot understand the contribution of one term to observed
-variability, independent of other confounding factors, there is little
-hope of disentangling and understanding the relative role of many
-quantities in determining observed variability.  Equation \ref{d_et}
-explicitly provides an estimate of the ET response to VPD, given
-assumptions about ecosystem plant characteristics through the
-parameters g$_1$ and uWUE. Eq. (\ref{d_et}) is generic to external
-environmental factors and the timescale over which a VPD response
-acts, and has a straightforward conceptual meaning. It provides the ET
+derivatives for understanding the behavior of complex systems in a
+simplifed framework. Our results are a conceptual tool for answering
+the question, ``What is the ET response to VPD with all other
+quantities held fixed?'' This is useful, and we would argue is
+critical, for understanding complex responses when many quantities are
+varying simultaneously \citep{Zhou_2019}. If we cannot understand the
+contribution of one term to observed variability, independent of other
+confounding factors, there is little hope of disentangling and
+understanding the relative role of many covarying quantities in
+determining observed variability.  Equation \ref{d_et} explicitly
+provides an estimate of the ET response to VPD, given assumptions
+about ecosystem plant characteristics through the parameters g$_1$ and
+uWUE. Eq. (\ref{d_et}) is generic to external environmental factors
+and the timescale over which a VPD response acts, and has a
+straightforward conceptual interpretation. It provides the ET
 response, given other quantities and parameters held fixed, subject to
@@
 ET response to VPD for a given plant species, or for a given soil
-moisture condition.
+moisture condition. ** talk about uWUE-SM, with supplementary here **?
 
@@
 VPD. However, crops, which we expect to prioritize carbon uptake over
-water conservation, have a higher uWUE closer to our maximum
+water conservation, have a higher uWUE value closer to our maximum
 value. This higher uWUE results in a higher likelihood for an
-increasing ET response to VPD, which intuitively makes sense given
-that we expect crops to keep stomata open for access to carbon, at the
-cost of increased water loss during high VPD. Shrubs, and to a lesser
-extent grass, have an uWUE closer to the minimum uWUE. For these plant
-types, we would then expect a higher likelihood of a water
-conservative response; that is, a decrease of ET in response to
-VPD. It is important to note that, while \citet{Zhou_2015} did not
-examine the role of soil moisture for within plant-type variability of
-uWUE, we might expect some variation, especially in extreme cases when
-soil water becomes a limiting factor.
+increasing ET response to VPD, which matches intuition given that we
+expect crops to keep stomata open for access to carbon, at the cost of
+increased water loss during high VPD. Shrubs, and to a lesser extent
+grass, have an uWUE closer to the minimum uWUE. For these plant types,
+we would then expect a decrease of ET in response to VPD. It is
+important to note that, while \citet{Zhou_2015} did not examine the
+role of soil moisture for within plant-type variability of uWUE, we
+might expect some variation, especially in extreme cases when soil
+water becomes a limiting factor. ** talk about uWUE-SM, with supplementary here **?
 
-In \citet{Medlyn_2017} there is not as robust of a relationship
-between plant type and g$_1$ as there was for uWUE in
-\citet{Zhou_2015}. Given that there is a robust relationship between
-g$_1$ and plant type in \citet{Lin_2015}, the ambiguous results in
-\citet{Medlyn_2017} could be due to noise in ecosystem scale
-observations, the consequences of imposing a model structure in some
-of their g$_1$ esimations, or real ecosystem-scale within-plant type
-variability. These differences between leaf-scale behavior and
-ecosystem scale behavior highlight the importance of understanding how
-and why slope parameters show increased variability at the ecosystem
-scale, as argued by \citet{Medlyn_2017}. However, if we assume that
-there is some real analogy between leaf scale results in
-\citet{Lin_2015} and ecosystem scale behavior, and that some of the
-ambiguity in \citet{Medlyn_2017} is due to model and/or observational
-error, then we can make some connections between g$_1$-induced ET
-response variability and plant characteristics, which are subject to
-the strong caveat that we still do not fully understand g$_1$'s
-utility at the ecosystem scale. Extrapolating \citet{Lin_2015}'s
+\citet{Medlyn_2017}'s results did not display the robust relationship
+between plant type and and plant parameter (g$_1$) that was
+demontrated by \citet{Zhou_2015} for uWUE. Given that there is a
+robust relationship between g$_1$ and plant type in \citet{Lin_2015},
+and that \citet{Zhou_2015}'s ecosystem-scale results rely on the same
+underlying theory as \citet{Medlyn_2017}'s results, the ambiguous
+results in \citet{Medlyn_2017} could be due to noise in ecosystem
+scale observations, the consequences of imposing a model structure in
+some of their g$_1$ esimations, or real ecosystem-scale within-plant
+type variability (e.g. multiple species impacting the relationship
+with the eddy-covariance footprint). These differences between
+leaf-scale behavior and ecosystem scale behavior highlight the
+importance of understanding how and why slope parameters show
+increased variability at the ecosystem scale, as argued by
+\citet{Medlyn_2017}. However, if we assume that there is some real
+analogy between leaf scale results in \citet{Lin_2015} and ecosystem
+scale behavior, and that some of the ambiguity in \citet{Medlyn_2017}
+is due to model and/or observational error, then we can g$_1$-induced
+ET response variability and plant characteristics, which are subject
+to the strong caveat that we still do not fully understand g$_1$
+behavior at the ecosystem scale. Extrapolating \citet{Lin_2015}'s
 results (e.g. Fig. 2) to the ecosystem scale would reveal the
@@
     trees.
-  \item Tropical and temperate climates would generally host plants
-    with a higher g$_1$ than arctic and boreal climates.
+  \item Tropical and temperate climates would generally be
+    characterized by plants with a higher g$_1$ than arctic and boreal
+    climates.
 \end{itemize}
@@
 
-Our theoretical results, when framed by previous literature tying
-parameters to plant types and climate, highlight the variability of ET
-response to VPD as a function of plant characteristics. Whether an
+Our theoretical results highlight the variability of ET
+response to VPD as a function of both climate  and plant
+characteristics, especially water usage strategy. Whether an
 ecosystem increases or decreases ET in response to VPD will vary
@@
 increases in VPD'' may be false depending on the water conservation
-strategy of a given ecosystem **CITE SOME EXAMPLES OF THIS LOGIC FROM
-GLOBAL WARMING LIT**.
+strategy of a given ecosystem \citep{Lemordant_2018}.
 
@@
 conductance models and parameters for the ecosystem scale. In
-particular, why and how g$_1$ behavior at the leaf and ecosystem scale
-are not directly analogous much be understood. While the derivation
+particular, why and how g$_1$ behavior at the leaf scale is not
+analagous to the ecosystem scale must be understood. While the derivation
 here was explicit about the assumption of a constant g$_1$ term with
@@
 efficacy of models using g$_1$'s for ecosystem-scale
-fluxes. Fortunately, our framework, and the novel general approach of
+fluxes. Fortunately, our framework, and the novel approach of
 using robust ecosystem-scale ratios (uWUE) to remove the implicit

### 2019-02-28 doc/paper/body.tex
done with pierre's comments, still need to put in soil moisture discussion and supplementary materials somewhere

@@
 particular, why and how g$_1$ behavior at the leaf scale is not
-analagous to the ecosystem scale must be understood. While the derivation
-here was explicit about the assumption of a constant g$_1$ term with
-respect to the VPD perturbation, many sophisticated land surface
-models and earth system models employing stomatal conductance models
-make similar, and even stronger assumptions about the stationarity of
-VPD slope terms. Many models will assume a constant g$_1$ term within
-a given plant functional type, and research derived from these models
-often does not acknowledge limitations of this assumption, or the
-difficulty in the literature with quantifying an ecosystem scale slope
-parameter \citep{Medlyn_2017}. Given that something so fundamental as
-the ET response to VPD varies strongly with g$_1$, we must invest more
-in understanding g$_1$'s behavior at the ecosystem scale, and the
-efficacy of models using g$_1$'s for ecosystem-scale
-fluxes. Fortunately, our framework, and the novel approach of
-using robust ecosystem-scale ratios (uWUE) to remove the implicit
-dependence of stomatal conductance on ET, is flexible enough to be
-applied to any future-derived ecosystem-scale stomatal conductance
-model that contains a dependence on GPP.
+analagous to the ecosystem scale must be understood. While the
+derivation here was explicit about the assumption of a constant g$_1$
+term with respect to the VPD perturbation, many sophisticated land
+surface models and earth system models employing stomatal conductance
+models make similar, and even stronger assumptions about the
+stationarity of VPD slope terms \citep{Niu_2011, Franks_2017,
+  Rogers_2017, Lawrence_2019}. Many models will assume a constant
+g$_1$ term within a given plant functional type, and research derived
+from these models often does not acknowledge limitations of this
+assumption, or the difficulty in the literature with quantifying an
+ecosystem scale slope parameter \citep{Medlyn_2017}. Given that
+something so fundamental as the ET response to VPD varies strongly
+with g$_1$, we must invest more in understanding g$_1$'s behavior at
+the ecosystem scale, and the efficacy of models using g$_1$'s for
+ecosystem-scale fluxes. Fortunately, our framework, and the novel
+approach of using robust ecosystem-scale ratios (uWUE) to remove the
+implicit dependence of stomatal conductance on ET, is flexible enough
+to be applied to any stomatal conductance model that contains a
+dependence on GPP.
 
@@
 imposed by our choice of stomatal conductance model. We now examine
-how the model choice, and specifically the exponent of the VPD
-dependence in the stomatal conductance model, can impact the general
-form of the ET-VPD relationship. This is relevant as many models
-include an exponentiated VPD term \citep[e.g.,][]{Leuning_1990}, but
-the value of the exponent can vary. It is useful to understand how the
-choice of model imposes the functional form of the ET-VPD
-relationship, given its fundamental importance for land atmosphere
-interaction.
+how the stomatal model choice, and specifically the exponent of the
+VPD dependence in the stomatal conductance model, can impact the
+general form of the ET-VPD relationship. There is a theoretical basis
+for the square root VPD dependence in both the stomatal conductance
+model and $uWUE$ based on the assumption that stomata behave to
+maximize carbon gain while minimizing water loss, which observations
+also generally support \citep{Lloyd_1991, MEDLYN_2011, Lin_2015,
+  Zhou_2014, Zhou_2015, Medlyn_2017}. However, some purely empirical
+results that fit the exponent of the VPD dependence to data have shown
+that it may vary slightly from 1/2, suggesting that stomata, as well
+as ecosystem-scale quantities based on stomata theory, may not always
+function optimally \citep{Zhou_2015, Lin_2018}. Specifically with
+regards to $uWUE$, one would not expect that this ecosystem scale WUE
+quantity will respond to VPD exactly analogously to stomata. Direct
+soil evaporation's contributions to ET should shift the exponent of
+the VPD dependence, especially at conditions of low GPP when we would
+expect a systematically larger portion of direct soil evaporation
+contributions to ET, because we would also expect lower amounts of
+transpiration at low GPP. \citet{Zhou_2015}'s results corroborate
+this: they found a mean empirically fit exponential VPD dependence of
+0.55, varying slightly from the theoretically optimal value of of 0.5
+for AmeriFlux sites. \citet{Lin_2018}'s results also show variance in
+the empirical exponent of the VPD dependence of the stomatal
+conductance model, but interpretation of this variance is more
+difficult as \citet{Lin_2018} do not handle GPP/A dependence of
+stomatal conductance in a directly analogous manner to the optimal
+theory in \citet{MEDLYN_2011} and \citet{Medlyn_2017}. Regardless,
+given that these recent results on the relationship between VPD, GPP,
+and ET \citep{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017} form the
+backbone of our analysis and are what allowed us to derive an explicit
+ET expression for the first time (Eq. (\ref{et})), we will analyze if
+and how assumptions about the exponent of the VPD dependence impacts
+the shape of the ET-VPD dependence. This analysis is also important to
+understand whether the choice of stomatal conductance model alters the
+fundamental behavior of the ET-VPD relationships, as many commonly
+used models utilize a VPD exponent other than the 1/2 suggested by
+optimal theory \citep[e.g. ][ which uses an exponent of
+1]{Leuning_1990}. It is useful to understand how the choice of model
+imposes the functional form of the ET-VPD relationship, given its
+fundamental importance for land-atmosphere interactions.
 
@@
 upward. In other words, there should be some local minimum in ET at a
-critical VPD$_{crit}$, assuming the scaling and plant terms  (e.g. aerodynamic
-conductance, $\Delta$, $g_1$ and $uWUE$) are held fixed. This result warrants further
-investigation, because to our knowledge no one has derived the
-theoretical ecosystem-scale relationship between ET and VPD while
-controlling for other environmental conditions. In particular, from
-personal communication, there is an apparent lack of consensus over
-whether the shape of the ET-VPD curve should be concave upward (our
-result) or concave downward in the absence of dramatic water stress. Given that understanding the ET-VPD
-relationship of the one-way plant response is fundamental to
-hypothesizing about any feedbacks between the land surface and the
-atmosphere, we analyze why our derived ET-VPD relationship is concave
-upward, particularly with respect to the exponent of VPD dependence in
-$uWUE$ and the Medlyn unified stomatal conductance model.
+critical VPD$_{crit}$, assuming the scaling and plant terms
+(e.g. aerodynamic conductance, $\Delta$, $g_1$ and $uWUE$) are held
+fixed. This result warrants further investigation, because to our
+knowledge no earlier work has derived the theoretical ecosystem-scale
+relationship between ET and VPD while controlling for other
+environmental conditions. In particular there is an apparent lack of
+consensus over whether the shape of the ET-VPD curve should be concave
+upward (our result) or concave downward in the absence of dramatic
+water stress. Given that understanding the ET-VPD relationship of the
+one-way plant response is fundamental to hypothesizing about any
+feedbacks between the land surface and the atmosphere, we analyze why
+our derived ET-VPD relationship is concave upward, particularly with
+respect to the exponent of VPD dependence in $uWUE$ and the Medlyn
+unified stomatal conductance model, as other models and empiracle
+results have suggested exponents varying from 1/2 \citep{Leuning_1990,
+  Zhou_2015, Lin_2018}.
 
-There is a theoretical basis for the square root VPD dependence in
-both the stomatal conductance model and $uWUE$ based on the assumption
-that stomata behave to maximize carbon gain while minimizing water
-loss, which observations also generally support \citep{Lloyd_1991,
-  MEDLYN_2011, Lin_2015, Zhou_2014, Zhou_2015, Medlyn_2017}. However,
-some purely empirical results that fit the exponent of the VPD
-dependence to data have shown that it may vary slightly from 1/2,
-suggesting that stomata, as well as ecosystem-scale quantities based
-on stomata theory, may not always function optimally \citep{Zhou_2015,
-  Lin_2018}. Specifically with regards to $uWUE$, one would not expect
-that this ecosystem scale WUE quantity will respond to VPD exactly
-analogously to stomata. Direct soil evaporation's contributions to ET
-should shift the exponent of the VPD dependence, especially at
-conditions of low GPP when we would expect a systematically larger
-portion of direct soil evaporation contributions to ET, because we
-would also expect lower amounts of transpiration at low GPP. \citet{Zhou_2015}'s
-results corroborate this: they found a mean empirically fit
-exponential VPD dependence of 0.55, varying slightly from the
-theoretically optimal value of of 0.5 for AmeriFlux
-sites. Interpreting \citet{Lin_2018}'s results, which also show
-variance in the empirical exponent of the VPD dependence of the
-stomatal conductance model, is more difficult as \citet{Lin_2018} do
-not handle GPP/A dependence of stomatal conductance in a directly
-analogous manner to the optimal theory in \citet{MEDLYN_2011} and
-\citet{Medlyn_2017}. Regardless, given that these recent results on
-the relationship between VPD, GPP, and ET \citep{MEDLYN_2011,
-  Zhou_2014, Zhou_2015, Medlyn_2017} form the backbone of our analysis
-and are what allowed us to derive an explicit ET expression for the
-first time (Eq. (\ref{et})), we will analyze if and how assumptions
-about the exponent of the VPD dependence impacts the shape of the
-ET-VPD dependence.
-This analysis is also important to understand whether the choice of stomatal conductance model alters the fundamental behavior of the ET-VPD relationships, as many commonly used models  utilize a VPD exponent other than the 1/2 suggested by optimal theory \citep[e.g.
-][ which uses an exponent of 1]{Leuning_1990}.
-
-By introducing $n$ and $m$ we can free our stomatal
-conductance model from assumptions about VPD dependence:
+By introducing $n$, the exponent of VPD in uWUE, and $m$, the exponent
+of VPD in the stomatal conductance model, we can free our theory from
+assumptions about VPD dependence:
 % \begin{linenomath*}
@@
 of ET which increase with VPD, independent of stomatal
-closure. Therefore, in the limit as VPD becomes large and we assume
+closure ** cite katul??? **. Therefore, in the limit as VPD becomes large and we assume
 plants are exercising all strategies to reduce ET, any further increase
@@
 stomata response is asymptoting towards full closure would cause a
-concave up ET-VPD curve, which is matched by the theory. In short,
+concave up ET-VPD curve, which is matched by our theory. In short,
 plant response becomes more limited as VPD increases, while
@@
 represented by an extension of our framework in which uWUE is allowed
-to decrease with decreasing SWC. Here, we focus our results by
-framing theme with previous literature to  build baseline intuition for
-ET-VPD dependence.
+to change with changing SWC. Here, we focus our results by framing
+theme with previous literature to build baseline intuition for ET-VPD
+dependence.

### 2019-02-28 doc/paper/hess-revision/final-response.tex
finished rough of response 1

--- 2018-12-18 doc/paper/hess-revision/review-1.tex
@@
 
-\title{Response to RC1 of ``When does vapor pressure deficit drive or
-  reduce evapotranspiration?''}
+\title{Final Response: When does vapor pressure deficit drive or
+  reduce evapotranspiration?}
 
@@
 
-\section{Compounding uncertainties}
+\section{Reviewer 1}
 
-We agree that the uncertainties are large both within a PFT and across
-PFTs. Looking back, we believe that some of our language
-communicating \cite{Lin_2018}, \cite{Medlyn_2017}, and
-\cite{Zhou_2014}'s results was misleading, and this was exacerbated by
-our focus on PFT-analysis. We will remove that language from the
-manuscript, and add language to better communicate our results as
-reflections of the considerable within-PFT uncertainty.
+Thank you for a very thoughtful review. You identified some critical
+flaws in our methods and presentation, and addressing these flaws
+have, in our opinion, greatly improved the strength and clariy of the
+manuscript.
 
-We think comparing the rankings of given PFT values for g$_1$ and uWUE
-can be misleading, given the magnitude of the uncertainties
-involved. Most of our calculated values for uWUE are within one
-standard deviation of \cite{Zhou_2015}'s results, but these deviations
-can result in some changes in the ordering of PFTs from high to low
-uWUE. The bigger problem in our eyes is that we made a mistake with
-some language suggesting within intra-PFT variability is less than
-inter-PFT variability, but clearly this is not the case. We will
-remove this language. This misinterpretation should not have been in
-the manuscript and actually contradicts other manuscript content; for
-example, we included \cite{Zhou_2015}'s results on uWUE in Table 2
-explicitly to be transparent about the within-PFT uncertainties. g$_1$
-values also exhibit considerable uncertainty, and again, while the
-relative rankings of these values may change, they are all within
-observed ranges in \cite{Medlyn_2017}. We did not include these in
-Table 2 because \cite{Medlyn_2017} does not provide the numerical
-values; however, we think it would be useful to provide estimates of
-these ranges from \cite{Medlyn_2017}'s figures in our Table 2.
+\subsection{Compounding uncertainties}
 
-Regarding how to interpret g$_1$ and uWUE: we think the best way
-is with established physical relations from \cite{MEDLYN_2011}
-(Equation 3 in the manuscript) and \cite{Zhou_2014} (Pg. 7, line 10 in
-the manuscript), with the knowledge that there is uncertainty
-involved. The quantities in these relationships all have some
-intrinsic physical meaning, but can vary substantially within a
-PFT. Again, we need to alter our language to better reflect this
-intra-PFT variability.
+As stated in our response during the open discussion period, we agree
+with you that there are problems prescribing any parameters as fixed
+within a given PFT, as we would expect relatively large variability
+within a PFT. While this type of analysis was motivated by examples in
+previous literature, we think it is misleading in this case to present
+results where these parameters are suggested to be fixed. These issues
+were exacerbated by the manner in which we fit our parameters: it was
+unclear given our methods how model and observational uncertainty,
+which can be considerable, are partitioned into our PFT-wide estimates
+of g1 and uWUE. Also, as you point out, there is already previous
+literature on the values of g1 and uWUE, so in many ways by fitting
+out own values we are repeating previous work, but with defects given
+our handling of observational and model error. Also, much of our
+observational work, in addition to its flaws, decreased the clarity of
+our manuscript and distracted from the focus, which a flexible
+derivation, based completely on previous literature, of ET explicitly
+as a function of enviornmental variables and plant parameters.
 
-However, \cite{Zhou_2014} did establish a constant uWUE approximation
-as a good approximation for capturing the relationship between GPP, ET
-and VPD at a given place and time, so it is still a very useful and
-robust approximation for answering our research question (see
-discussion earlier in this comment). Additionally, a constant g$_1$
-approximation is used in many earth system models, so while it
-introduces uncertainty, it also makes our framework useful for
-interpreting modeled vegetation response in ESMs and GCMs.
-
-We also agree with the comment that the best way to think of this is
-probabilistically. This is the most robust approach to dealing with
-approximations - we introduce randomness to our variables and
-parameters to account for all of the physics that are not explicitly
-accounted for, as well as observational uncertainty. However, as far
-as we know we still have not developed a general, robust, and efficient
-arithmetic for random variables. We could try and adapt a Bayesian
-model to this problem, but given the large amount of arithmetic
-involved, fully incorporating a Bayesian representation to every
-variable in the analysis would be a very hard problem, and a
-significant research project in its own right. We think a good
-compromise is to add language directing the reader to interpret our
-results more probabilistically, which we have already presented
-probabilistically in the figures. For example, in Figure 5 the range of
-values in each plot represents a range of possible responses in the
-sign term. Within this figure, it's worth noting that the intra-PFT
-variability is greater than the inter-PFT variability, which is
-consistent with some of the previous results you highlight. Again, we
-need to add language highlighting this, and its consistency with
-previous results. We included this variability and uncertainty
-explicitly to be transparent. A more difficult question is how much of
-this variability is due to observational and model error, and how much
-of it is due to climate and plant physiological variability (see
-Section 3.5).
+These realizations led us to make significant alterations to our
+manuscript. We no longer use observations or fit parameters
+ourselves. Instead we frame our theoretical derivation in terms of
+parameters established in the literature. Using this literature, we
+can then make statements about the general liklihood of a positive or
+negative ET response to increasing VPD in terms of climate, plant
+type, and photosynthetic pathway. This seemed a more natural
+delineation of scope: our paper presents a derivation tying plant
+parameters to an ET response to VPD, while previous literature
+establishes relationships between plant parameters and plant
+type/climate.
 
@@
 
-We do not want our analysis to be interpreted as an assertion that
-g$_1$ and uWUE are attributes of PFT only. We expect them to vary both
-within a PFT and across a PFT. Our primary goal in using g$_1$ and
-uWUE was to develop an explicit expression of ET as a function of
-environmental variables, and use this to assess the response to a
-change in VPD. We focused our analysis using PFTs because this is how
-it was framed in previous studies in the field, and specifically the
-studies used in our derivation \cite{Zhou_2014, Zhou_2015,
-  Medlyn_2017}. We were originally thinking the PFT-focused analysis
-could be useful, especially given that climate models generally hold
-plant parameters fixed with respect to PFT, so long as we were
-transparent about the large uncertainties and problems with this
-approach (see Figure 5).
+In our open discussion response we detailed our agreement with you that g1 and
+uWUE should vary within a given PFT. Removing this PFT-focused
+analysis should prevent any misinterpretation that we are making
+claims that these paremeters should be fixed within a given PFT. For
+convenience, we also reproduce a particlarly relevant passage from our
+orignal response:
 
-We agree with you that stating that any quantity is fixed within a PFT
-is hard to believe. Phenotypic variation and adaptation within a given
-\textit{species} can be considerable (i.e. effecting $\lambda$, g$_1$, and
-uWUE), so it would be hard to say that anything would be constant
-within a PFT made up of many different diverse species. Both
-phenotypic variation as well as the species distribution and dominant
-PFT at a given location will all be strongly optimized in response to
-climate. In this sense, the distribution and evolution of local
-climate is a strong control on the local ecosystem structure and
-physiology of a given ecosystem.
-
-What we need to communicate better is that a given ecosystem's state
+``What we need to communicate better is that a given ecosystem's state
 at a place or time controls its response to a VPD perturbation. We are
@@
 and through their effect on plant physiology (parameterized by g$_1$
-and uWUE).
+and uWUE).''
 
-When writing this manuscript, there was definitely some internal
-tension and debate about how to best frame the analysis and
-results. We could either focus on PFT-oriented results as previous
-literature has done (e.g. holding plant physiology fixed within a
-given PFT), or allow plant physiology terms to vary through time and
-space and look at the distribution of ET response to VPD (see Section
-3.5). After re-examining the manuscript both after some time away from
-the problem and in light of your comments we think a strong argument
-could be made that we made the wrong choice with respect to this
-focus. It may have made more sense to focus our analysis on the ET
-response more generally across space and time as ecosystem-scale plant
-physiology varies in response to climate and soil moisture.
+In addition we added language to the manuscript that our results
+should be used by a reader/scientist under the assumption that plant
+parameters are fixed with respect to the VPD perturbation
+considered. Previous literature \citep{Zhou_2015, Lin_2015} suggests
+that these parameters might also be approximately constant across a
+relatively broader range of plant types and environmental conditions,
+but our manucript avoids claims about this costancy other than the
+importance of conceptually considering these parameters to be constant
+when applying our theory to a problem of interest.
 
-\section{Next Steps}
-
-In order to improve the manuscript and our communication of the answer to
-the question ``When does VPD drive or reduce ET?'', we see a few
-potential paths that we will consider between now and the final
-response after the discussion period. If you (or anyone else) has any
-opinions or comments in the meantime, we would appreciate the insight
-and feedback.
-
-\begin{enumerate}
-\item We include the discussion presented in this review and response,
-  and include language explicitly stating that the purpose of the PFT
-  analysis is to provide connections to other PFT-constant analyses
-  and models (e.g. ESMs and GCMs). We will add extensive language on
-  the uncertainty and weaknesses of this approach, and reframe our
-  conclusions to reflect his uncertainty.
-\item We instead alter our analysis to look at how VPD response varies
-  with climate and general plant physiological variability, instead of
-  focussing only on PFT analysis that poorly captures all of the observed
-  variability in ET response. Sections 3.1 - 3.2 would be replaced by
-  analysis directly relating ET response to general climate and plant
-  physiological terms, rather than PFT-mean analysis. For example, the
-  ``scaling term'' analysis would be presented in terms of generic
-  changes in plant height and temperature, and the ``sign term''
-  analysis would be framed by generic changes to g$_1$ and uWUE, as
-  informed by previous literature.
-\item This is the most extreme option in terms of modifying the
-  manuscript. We significantly alter the manuscript and instead focus
-  on just the general shape of the ET-VPD curve (with environmental
-  variables held fixed). This was not discussed in the review, but one
-  of our most noteworthy results is on the general shape of the ET-VPD
-  curve being concave up, given an assumption of a square-root VPD
-  dependence of ET. This result is independent of any assumptions of
-  uWUE and g$_1$, and to our knowledge it is first derived curve of
-  ecosystem response to VPD (see Section 3.7). It also highlights the
-  importance of discerning the exact exponent of VPD dependence, as
-  this alters the fundamental nature and shape of the
-  curve. Essentially, the manuscript would become just our motivation
-  and derivation, followed by Section 3.7 exploring the consequences
-  of the derivation for the shape of the VPD curve. This alteration of
-  the manuscript would likely result in more of a technical note-type
-  paper, and we are hesitant to do this because we think there is
-  still a lot of useful information obtained by tying our results to
-  real-world scenarios (as in Options 1 and 2 above).
-\end{enumerate}
-
-Here we present some final minor comments and concerns on the
-technical details of our proposed changes to the manuscript, which are
-relevant to representing uncertainty and spatiotemporal variability in
-uWUE \textit{and} g$_1$. In the original manuscript we used a single
-$\sigma$ term to represent this variability. We did this because
-changes in uWUE and g$_1$ induce a very similar change in the ET
-solution, which made solving for independent $\sigma_{uWUE}$ and
-$\sigma_{g1}$ terms intractable at a given time and place, and
-representing variability in both g$_1$ and uWUE difficult. We decided
-to hold g$_1$ fixed within a PFT for a two reasons: 1) ESMs and GCMs
-generally hold g$_1$ fixed, and 2) letting uWUE vary seemed more
-appropriate to represent specifically soil water variations on
-stomatal conductance, as $uWUE$ modifies stomatal conductance
-analogously to how soil moisture factors modify a maximum stomatal
-conductance in land surface models (e.g. it is a multiplicative factor
-on the entire stomatal conductance term). Because uWUE and g$_1$
-induce similar changes in ET, at least qualitatively we think that
-having a single $\sigma$ can represent some of the variability in both
-$uWUE$ and g$_1$. However, formally there is no explicit variability
-in the g$_1$ term. The discussion in this review rightly points out
-that based on previous results we do expect some variability in g$_1$
-as well. So, an unresolved question is how important is it to the
-analysis and its interpretation to include an explicit g$_1$
-variability term in addition to the existing $\sigma$ term. We are
-including these comments to hopefully stimulate some discussion. We
-could imagine some timescale based approaches where we might be able
-to account for both the g$_1$ and uWUE variability, for example by
-making assumptions over what time scale each quantity is fixed, and
-fitting based on that (e.g. g$_1$ is fixed for a given season, and
-uWUE is fixed for a given day). This approach might allow a tractable
-solution, and also could help us filter error and observational noise
-from ``true'' plant physiological and climatic variability in uWUE and
-g$_1$. The cost of all of this is a significant increase in the
-analysis and content of the paper, as well as some increased opacity
-to the methods. Comments are welcome.
-
-Thank you again for the thoughtful review. We hope it stimulates
-further discussion.
+The issue of soil moisture dependence is of particular importance,
+however. If our plant parameters are more stationary with respect to
+soil moisture, then our theory can be applied to a broader range of
+VPD scenarios, including observed compound events between high VPD and
+low soil moisture \citep{Zhou_2019}. To help the reader assess the
+soil moisture dependence of uWUE (and partially by extension
+$\lambda$), we provided supplementaty material showing the
+distribution of uWUE with SWC for each of 66 FLUXNET sites. The
+functional relationship between uWUE and SWC varies, with a mix of
+sites showing strong and weak SWC-dependence. For all sites the ratio
+of signal to noise is very low, an unfortunate consequence of taking a
+ratio of two highly uncertain eddy-covariance derived fluxes. In
+general we find this analysis inconclusive. Our paper does not rely on
+assumptions about uWUE's functional relationship to soil moisture so
+we do not include these analysis in the body of the text, but these
+types of assumptions can make our theory more useful to the
+practitioner so we provide them for their interpretation and
+motivation for future research.
 
 % \printbibliography
-\bibliography{references.bib}
+\bibliography{references}
 

### 2019-02-28 doc/paper/body.tex
done with reviewer 2 comments, onto reviewer 3

@@
 % \begin{linenomath*}
+  \begin{equation}
+    g_{1-leaf} = \sqrt{\frac{3 \, \Gamma^* \,
+        \lambda}{1.6}},\footnote{Note this expression has units of of
+      (mmol mol$^{-1}$)$^{1/2}$, but this can be converted to
+      Pa$^{1/2}$ using the ideal gas law.}
     \label{slope}
-  \begin{equation}
-    g_{1-leaf} = \sqrt{\frac{3 \, \Gamma^* \, \lambda}{1.6}},\footnote{Note this expression has units of of (mmol mol$^{-1}$)$^{1/2}$, but this can be converted to Pa$^{1/2}$ using the ideal gas law.}
   \end{equation}
@@
   constant with respect to changes in
-  % TODO: make this a general discussion on partial derivatives and
-  % what things are constant/how to interpret?
   VPD \citep{Franks_2017, Zhou_2014}. These parameters will however
@@
   ][]{Lin_2015}, as well as environmental conditions including soil
-  water content and temperature.
-  % TODO: write exact parameter ranges we use, and introduce
-  % scaling/sign term here?
-  % \item
-  To account for both this variability in plant parameters and the
+  water content \citep{Manzoni2013} and temperature.
+
+  As Eq. (\ref{d_et}) is a partial derivative, it's utility is as a
+  conceptual model for the change in ET in response to VPD with all
+  other variables held fixed. This concept is a useful tool for
+  identifying the effect of VPD on ET through atmospheric demand and
+  plant response, and allows a practitioner to disentangle complicated
+  feedbacks when many quantities co-vary. However users of
+  Eq. (\ref{d_et}) should take care that their interpretation matches
+  the assumptions inherent in a partial derivative. For example,
+  results will only be as valid as the assumption that g$_1$ and uWUE
+  (and by extension $\lambda$) are fixed with respect to the user's
+  conceptual change in VPD. Fortunately, these quantities are
+  relatively constant for many under many scenarios
+  \citep{Franks_2017}. Care must also be taken with possible indirect
+  effects associated with a change in VPD: for example, a change in ET
+  induced by a change in VPD can cause a change in surface
+  temperature, which would drive a change in net radiation. These
+  types of indirect effects and feedbacks are not considered in
+  Eq. (\ref{d_et}): temperature (a variable) is mathematically fixed.
+
+  To account for both the variability in plant parameters and the
   environment, we will systematically analyze how the ET response to
-  VPD (Equation \ref{d_et}) varies. Equation \ref{d_et} includes a
+  VPD (Eq. (\ref{d_et})) varies. Equation (\ref{d_et}) includes a
   ``sign'' term that determines the sign of the response in addition
@@
 aerodynamic conductance (communication between the atmosphere and the
-surface) and temperature (controls the efficiency of
-energy conversion to latent heat through $\Delta$), amplify or suppress the plant
-physiological response represented in the sign term (Figure \ref{scaing}). Both lower
-temperatures and higher aerodynamic conductance lead to amplified ET
-response to VPD, with the variability of aerodynamic conductance
-resulting in a slightly higher variability of ET response relative to
-temperature.
+surface) and temperature (controls the efficiency of energy conversion
+to latent heat through $\Delta$), amplify or suppress the plant
+response represented in the sign term (Figure \ref{scaing}). Both
+lower temperatures and higher aerodynamic conductance lead to
+amplified ET response to VPD, with the variability of aerodynamic
+conductance resulting in a slightly higher variability of ET response
+relative to temperature.
 
@@
     depend on $uWUE$.
-  \item In general, increasing the exponent of VPD dependence
-    increases the likelihood of a concave down result. Additionally, as
-    the exponent of VPD dependence increases from the optimum value of
-    1/2, whether the curve is concave upward or concave downward
-    becomes a function of the plant specific slope parameter $g_*$,
-    through non-dimensional VPD ($VPD^m/g_*$). Because the exponent of
-    the VPD dependencies is capable of altering the fundamental shape
-    of ET-VPD dependence, future research investment in understanding
-    the exact VPD dependence of stomatal conductance, and further
-    reconciliation of empirical and theoretical stomatal and
-    ecosystem behavior should be prioritized.
+  \item In general, increasing the exponent of VPD dependence for
+    either $uWUE*$ or $g_*$ increases the likelihood of a concave down
+    result. Additionally, as the exponent of VPD dependence increases
+    from the optimum value of 1/2, whether the curve is concave upward
+    or concave downward becomes a function of the plant specific slope
+    parameter $g_*$, through non-dimensional VPD
+    ($VPD^m/g_*$). Because the exponent of the VPD dependencies is
+    capable of altering the fundamental shape of ET-VPD dependence,
+    future research investment in understanding the exact VPD
+    dependence of stomatal conductance, and further reconciliation of
+    empirical and theoretical stomatal and ecosystem behavior should
+    be prioritized.
 \end{itemize}

### 2019-02-28 doc/paper/hess-revision/final-response.tex
done with reviewer 2 comments, onto reviewer 3

@@
 flaws in our methods and presentation, and addressing these flaws
-have, in our opinion, greatly improved the strength and clariy of the
+have, in our opinion, greatly improved the strength and clarity of the
 manuscript.
@@
 \subsection{Compounding uncertainties}
-
-As stated in our response during the open discussion period, we agree
+\label{pft}
+As stated in
+\href{https://editor.copernicus.org/index.php/hess-2018-553-AC1.pdf}{our
+response} during the open discussion period, we agree
 with you that there are problems prescribing any parameters as fixed
@@
 
-In our open discussion response we detailed our agreement with you that g1 and
-uWUE should vary within a given PFT. Removing this PFT-focused
-analysis should prevent any misinterpretation that we are making
-claims that these paremeters should be fixed within a given PFT. For
-convenience, we also reproduce a particlarly relevant passage from our
-orignal response:
+In our open discussion
+\href{https://editor.copernicus.org/index.php/hess-2018-553-AC1.pdf}{response}
+we detailed our agreement with you that g1 and uWUE should vary within
+a given PFT. Removing this PFT-focused analysis should prevent any
+misinterpretation that we are making claims that these paremeters
+should be fixed within a given PFT. For convenience, we also reproduce
+a particlarly relevant passage from our orignal response:
 
@@
 
+\section{Reviewer 2}
+
+Thank you again for a high quality review. The discussion it
+inspired improved the strength of the manuscript.
+
+\subsection{(i)}
+\label{(i)}
+Please see our
+\href{https://www.hydrol-earth-syst-sci-discuss.net/hess-2018-553/hess-2018-553-AC2-supplement.pdf}{previous
+  response} for a full discussion of the barriers to assessing the ET
+response to VPD using pure data. For convience, we includ an exerpt
+here:
+
+``One issue with analyses attempting to diagnose the ET response to
+VPD with pure data (e.g. using binned boxplots, as you suggest), is
+that ET varies strongly with both aerodynamic conductance and
+radiation. Simply binning by VPD or by site includes too many
+confounding factors to really diagnose what the attributable response
+is to VPD. This issue is really what motivated our analysis and makes
+it unique; by building an analytical framework based on confirmed
+results in the literature (uWUE, Medlyn model, Penman-Monteith [PM]),
+we are able to formally deduce what the actual ET response is, with
+other environmental factors held fixed. This ``other environmental
+factors held fixed'' component is, in our mind, nearly impossible to
+deduce from pure data, because the environment is always changing and
+direct analogues are either impossible to find or result in too small
+of a sample size to form a meaningful conclusion... Essentially we are
+saying that to the degree we trust arithmetic, uWUE, Penman-Monteith,
+and the Medlyn stomatal conductance model, our evaluation of the
+derivative is robust.''
+
+You also mention possible net radiation dependence on VPD, which we
+address in the minor comments.
+
+\subsection{(ii)}
+
+As in our
+\href{https://www.hydrol-earth-syst-sci-discuss.net/hess-2018-553/hess-2018-553-AC2-supplement.pdf}{previous
+  response}, we agree that we need more discussion connecting and
+contrasting leaf level results with our ecosystems scale approach. We
+have summarized the discussion from our reviewer response and added it
+to the introduction in lines ** **.
+
+\subsection{(iii)}
+
+We agree that are PFT-based formulation was too easy to be
+misintrepreted as an assertion that g1 and uWUE are constant and never
+vary in time or space. As described in Section \ref{pft}, we have
+removed the observational and PFT-focused component of the analysis,
+and instead focus exclusvively on using previously published constants
+(and their range) to interpret results. We would also like to
+reiterate that we do not view our use of uWUE as an introduction of an
+extra parameter. Our use of uWUE replaces a photosynthesis model,
+which will generally have many more parameters. Using uWUE is instead
+a simplifying approximation that relies on assumptions (based on
+optimal theory), but one that certainly reduces the total number of
+parameters of a more traditional estimate of ET.
+
+\subsection{(iv)}
+
+Yes, we should not have used ``leading-order'' language. We removed
+that confusing and inappropraite language. Also, by removing the
+empirical results and focusing exclusively theory framed by observational
+results from the literature, we have clarified the manuscript along
+lines you suggest (e.g. better seperation of theory vs empirical
+data).
+
+\subsection{Minor Comments}
+
+\begin{itemize}
+\item P 1 L 17: We removed ``stress'' language.
+\item P 1 L 22-24: Discussion of PET was removed to simplify
+  manuscript (it seemed less relevant). Otherwise the Roderick
+  citation is a good one, thanks.
+\item P 2 L 17: Discussion of precipitation removed to simplify
+  manuscript.
+\item P 2 L 33-35: You're right, and this was speculation; removed
+  this language and replaced with: ``However these researchers did not
+  explore explicitly the sensitivity of ET to VPD, including VPD's
+  effect on stomatal conductance and plant function.''
+\item P 3 L 2: Added citation.
+\item P 3 L 9: Replaced novel with useful.
+\item P 5 L 4: Yes it is related to water use efficiency, but we think
+  that could add more confusion given the ``water use efficiency''
+  language in uWUE; by expressing it in terms of $\lambda$ hopefully
+  it is clear to the reader that g$_1$ is related to water use
+  efficiency.
+\item P 6 L 4: Yes, that equation was mislabeled. Should be fixed now.
+\item P 7 L 14: We no longer fit our own values, so this is no
+  longer relevant.
+\item P 7 Eq. (10): Yes, a change in ET/VPD would induce indrect
+  effects like changes in temperature and (be extension) net
+  radiation. We added discussion guiding the reader on how to avoid
+  misuse partial derivatives, because it seems like readers may not
+  the sufficient understanding, and included explicit discussion of
+  net radiation as an example: ``Care must also be taken with possible
+  indirect effects associated with a change in VPD: for example, a
+  change in ET induced by a change in VPD can cause a change in
+  surface temperature, which would drive a change in net
+  radiation. These types of indirect effects and feedbacks are not
+  considered in Eq. (\ref{d_et}): temperature (a variable) is
+  mathematically fixed.''
+\item P 10 LL 2 and LL 5: Yes, ``physiological controls'' can be
+  misintrepreted and doesn't tell the full story. This language is
+  removed.
+\item P 10 Eq. (12): Yes there should have been a $\sigma$, but
+  $\sigma$ is now removed from the manuscript.
+\item P 10. L 11: Yes, both g$_1$ and uWUE can vary with soil moisture
+  as both are dependent on $\lambda$. This section is however removed
+  from the updated manuscript, but we also added language reflecting
+  this ``These parameters will however vary with plant species and
+  characteristics \citep[e.g. wood density, ][]{Lin_2015}, as well as
+  environmental conditions including soil water content
+  \citep{Manzoni2013} and temperature.''
+\item P 10 Equation (13): $\sigma$ should not have been removed, but
+  it is now no longer in the manuscript.
+\item P 10 L 29-30 ** think on this, gets back to $\lambda$ and how it
+  changes with aridity **
+\item P 11 L 14-17 : this discussion removed from new version of
+  paper.
+\item P 13 L 7-9: this discussion removed from new version of paper.
+\item Figure 4: removed from paper.
+\item P 16 L 16: Yes this was somewhat circular and limited: it only
+  allowed us to test if/how uWUE, given a prescribed model form
+  and the data, would systematically vary with VPD within a PFT. There
+  were definitely some issues with our PFT-focused methods and how we
+  handled uncertainty, which let us to remove this portion of the
+  analysis (see Section \ref{pft}).
+\item P 16, L 15: Agreed, and thankfully this whole analysis is now
+  removed from the paper.
+\item P 18 L 32 / P 21 L 1-3: These sections have been
+  removed. However, we do cannot completely follow the feedback logic
+  you are using; can you ellaborate? Also we provide some citations
+  suggesting that uWUE and g1 can be \textit{relatively} constant
+  \citep{Lin_2005, Franks_2017, Zhou_2015}, so would be interested in
+  citations showing that they are not constant for the majority fo
+  conditions. Please note we also added some plots of the relationship
+  between soil moisture and uWUE in our supplementary material to aid
+  the reader, but view this as an active research as well.
+\item P 23 L 6: This section removed.
+\item P 24 L 24: Added clarifying language about the exponent.
+  ``In general, increasing the exponent of VPD dependence for
+  either $uWUE*$ or $g_*$ increases the likelihood ``
+\item P 26 L 2-4: Please see discussion in Section \ref{(i)}.
+\item P 27 Appendix A: removed use of FLUXNET data in this manuscript.
+\end{itemize}
+
 % \printbibliography

### 2019-02-28 doc/paper/intro.tex
done with reviewer 2 comments, onto reviewer 3

@@
 for evapotranspirated water \citep{Penman_1948, Monteith_1965}, but
-also stress plant stomata \citep{Leuning_1990,
-MEDLYN_2011}. Understanding the net evaportanspiration (ET) response
+also reduce stomatal conductance through stomatal closure \citep{Leuning_1990, MEDLYN_2011} by
+p \citep{Rawson1977, Mott2007,
+Darmour2010}. Understanding the net evaportanspiration (ET) response
 to these two opposing effects of changes in VPD is crucial for
@@
 complexity models, and facilitate the disentanglement of complex
-land-atmosphere feedbacks. In particular, we can understand the impact
-of climate change, given the expected increase of VPD.  In the past,
-similar simplified approaches were used to understand interactions
-between stomatal conductance, evapotranspiration and the environment
-\citep[e.g.,][]{Jarvis_1984, Jarvis_1986, Mcnaughton_1991}. However,
-at the time researchers' understanding of the form of VPD's effect on
-plant physiology was limited, so they could not systematically explore
-the sensitivity of ET to VPD, including VPD's effect on stomatal
-conductance and plant function.
+land-atmosphere feedbacks. In particular, it facilitates increased
+understading of the impact of climate change, given the expected
+increase of VPD.  In the past, similar simplified approaches were used
+to understand interactions between stomatal conductance,
+evapotranspiration and the environment \citep[e.g.,][]{Jarvis_1984,
+Jarvis_1986, Mcnaughton_1991}. However these researchers did not
+explore explicitly the sensitivity of ET to VPD, including VPD's
+effect on stomatal conductance and plant function.
 
@@
 conductance ($g_s$), including VPD response, by combining an optimal
-photosynthesis theory \citep{Cowan_1977, Katul_2009} with an empirical
+photosynthesis theory \citep{Cowan_1977, Katul_2009, Katul_2010} with an empirical
 approach, and extended use of this model to the ecosystem scale in
@@
 across space and time, within a plant functional type (PFT)
-\citep{Zhou_2015}. ** here talk about difference between leaf level
-and ecosystem scale, also connect uWUE to g1 ** While stomatal
-conductance parameterizations and uWUE greatly simplify complex plant
-physiological processes, they therefore still capture ecosystem
-behavior for vegetated surfaces \citep{Medlyn_2017, Zhou_2014}, and
-are novel tools to transparently develop intuition for the behavior of
-complex land-atmosphere systems.
+\citep{Zhou_2015}.  While stomatal conductance parameterizations and
+uWUE greatly simplify complex plant physiological processes, they
+therefore still capture ecosystem behavior for vegetated surfaces
+\citep{Medlyn_2017, Zhou_2014}, and are useful tools to transparently
+develop intuition for the behavior of complex land-atmosphere systems.
 
 In this manuscript, we leverage uWUE and recent developments in
-stomatal conductance parameterizations \citep{MEDLYN_2011} to derive
-the theoretical one-way response of ET to VPD with other environmental
-variables properly controlled for, i.e. we develop a framework for
-evaluating the partial derivative of ET with respect to VPD. We
-explicitly include VPD's full effect on stomatal conductance,
-including its impact on photosynthesis. We explore the range of
-possible ET responses to VPD, given parameters previously established
-in peer reviewed literature. Additionally, we show the sensitivity of
-the ET-VPD relationship to model and framework choice, highlighting
-the importance of: 1) future research on stomatal conductance and
-ecosystem scale modeling, and 2) thoughtful selection of
-photosynthesis and stomatal conductance models in more sophisticated
-land surface and earth system models. Finally, we discuss the caveats
-of our approach and the sensitivity of results to assumptions about
-stomatal conductance.
+stomatal conductance parameterizations \citep{MEDLYN_2011} with a
+Penman-Monteith framework \citep [hereafter PM,][]{Penman_1948,
+Monteith_1965} to derive the theoretical one-way response of ET to VPD
+with other environmental variables properly controlled for, i.e. we
+develop a framework for evaluating the partial derivative of ET with
+respect to VPD. We explicitly include VPD's full effect on stomatal
+conductance, including its impact on photosynthesis. The use of an
+energy balance framework (PM) allows us to include in our analysis the
+effects of the energy cost of evaporating water from a surface, which
+is an important factor in the natural environment, compared to
+prescribed in situ envirnmental conditions. This is relevant because
+previous research focusing on the leaf scale \citep{Rawson1977,
+Turner1984, Oren1999, Darmour2010, Mott2013} does not consistently or
+analogously include the energetic constraints on evapotranspiration
+under which ecosystems operate. The leaf-scale results agree that
+stomatal conductance decreases in response to VPD \citep{Oren199,
+Damour2010}, which we expect to be true for an ecosystem as
+well. However leaf scale results also indicate that transpiration
+usually increases with increasing VPD in a concave downward shape
+\citep[e.g.,][]{Rawson1977, Turner1983, Mott2013}, which may not be
+true for ET at the ecosystem scale once energetic constraints on ET
+are included in the analysis. Our approach allows us to estimate the
+expected ET response to VPD at the ecosystem scale, including the
+effects of surface energy balance constraints, and assess how
+ecosystem ET response to VPD deviates from previous leaf scale
+analyses. It is indeed key to disconnect the impact of VPD from other
+variables as VPD is known to increase dramatically with future climate
+** (Novick JGR 2017) **, yet other environmental conditions such as
+soil moisture and radiation will not increase as
+dramatically. Nonetheless previous studies have typically been
+hampered by the covariations between different variables
+\citep[e.g. radiation and VPD or soil moisture and VPD, ][]{Lin_2018,
+Zhou_2019}, which limits the interpretability and understanding of the
+response of ecosystem-scale ET to VPD.
+
+This manuscript presents the range of possible ecosystem-scale ET
+responses to VPD, given parameters previously established in peer
+reviewed literature. Additionally, we explore the sensitivity of the
+ET-VPD relationship to stomatal model and framework choice,
+highlighting the importance of: 1) future research on stomatal
+conductance and ecosystem scale modeling, and 2) thoughtful selection
+of photosynthesis and stomatal conductance models in more
+sophisticated land surface and earth system models.

### 2019-02-28 doc/paper/body.tex
done with reviewer 3, now just edit paper make it nice and then response to editor

@@
     $\gamma$  & psychometric constant & Pa K$^{-1}$   & - \\
+    $g_{0}$  & leaf-scale mimimum stomatal conductance & m s$^{-1}$  & \citet{MEDLYN_2011} \\
+
     $g_{s-leaf}$  & leaf-scale stomatal conductance & m s$^{-1}$  & \citet{MEDLYN_2011} \\
@@
     $c_a$ & CO$_2$ concentration & $\mu$ mol CO$_2$ mol$^{-1}$ air& - \\
-    $\lambda$ & marginal water cost of leaf carbon & mol H$_2$O mol$^{-1}$ CO$_2$ & - \\
+    $\lambda = \frac{\partial \; transpiration}{\partial\; CO_2\; assimilation}$ & marginal water cost of leaf carbon & mol H$_2$O mol$^{-1}$ CO$_2$ & - \\
     $\Gamma$ & CO$_2$ compensation point & - & - \\
@@
     Zhou_2015} and \citet{Medlyn_2017}'s stomatal conductance model is
-  that direct soil evaporation (E) contributions to ET remain small
-  relative to transpiration (T). Again, this should be more true
-  during the growing season. The ratio of E to T may increase
-  immediately after rainfall events due to high soil moisture and
-  ponding, but VPD is generally low anyways during these
+  that direct soil evaporation contributions to ET remain small
+  relative to transpiration. Again, this should be more true during
+  the growing season. The ratio of evaporation to transpiration may
+  increase immediately after rainfall events due to high soil moisture
+  and ponding, but VPD is generally low anyways during these
   times. However, some plant types allow for systematically larger
-  contributions of E in ET, particularly those with sparse canopies
-  and smaller relative amounts of transpiration. We therefore might
-  expect that the theory will be most applicable to forest PFTs, which
-  will be most strongly coupled to the boundary layer due to larger
-  surface roughness, and will also generally have the highest ratios
-  of transpiration to evaporation. Finally, we assume that $g_1$ and
-  $uWUE$ are constant with respect to the conceptual VPD
+  contributions of evaporation in ET, particularly those with sparse
+  canopies and smaller relative amounts of transpiration. We therefore
+  might expect that the theory will be most applicable to forest PFTs,
+  which will be most strongly coupled to the boundary layer due to
+  larger surface roughness, and will also generally have the highest
+  ratios of transpiration to evaporation. Finally, we assume that
+  $g_1$ and $uWUE$ are constant with respect to the conceptual VPD
   perturbation. Both quantities have been shown to be relatively
-  constant with respect to changes in
-  VPD \citep{Franks_2017, Zhou_2014}. These parameters will however
-  vary with plant species and characteristics \citep[e.g. wood density,
-  ][]{Lin_2015}, as well as environmental conditions including soil
-  water content \citep{Manzoni2013} and temperature.
+  constant with respect to changes in VPD \citep{Franks_2017,
+    Zhou_2014}. These parameters will however vary with plant species
+  and characteristics \citep[e.g. wood density, ][]{Lin_2015}, as well
+  as environmental conditions including soil water content
+  \citep{Manzoni2013} and temperature.
 

### 2019-02-28 doc/paper/hess-revision/final-response.tex
done with reviewer 3, now just edit paper make it nice and then response to editor

@@
 \subsection{(ii)}
-
+\label{leaf}
 As in our
@@
 reiterate that we do not view our use of uWUE as an introduction of an
-extra parameter. Our use of uWUE replaces a photosynthesis model,
-which will generally have many more parameters. Using uWUE is instead
-a simplifying approximation that relies on assumptions (based on
-optimal theory), but one that certainly reduces the total number of
-parameters of a more traditional estimate of ET.
+extra parameter. Our use of uWUE replaces a photosynthesis model and
+stomatal conductance reduction factor, which will generally have many
+more parameters. Using uWUE is instead a simplifying approximation
+that relies on assumptions (based on optimal theory), but one that
+certainly reduces the total number of parameters used in a more
+traditional estimate of ET.
 
@@
   misuse partial derivatives, because it seems like readers may not
-  the sufficient understanding, and included explicit discussion of
+  have sufficient understanding, and included explicit discussion of
   net radiation as an example: ``Care must also be taken with possible
@@
   suggesting that uWUE and g1 can be \textit{relatively} constant
-  \citep{Lin_2005, Franks_2017, Zhou_2015}, so would be interested in
-  citations showing that they are not constant for the majority fo
-  conditions. Please note we also added some plots of the relationship
-  between soil moisture and uWUE in our supplementary material to aid
-  the reader, but view this as an active research as well.
+  \citep{Lin_2005, Franks_2017, Zhou_2015}, so we would be interested
+  in citations showing that they are not constant for the majority fo
+  conditions. We also added some plots of the relationship between
+  soil moisture and uWUE in our supplementary material to aid the
+  reader, but view those as inconclusive and an active research as
+  well.
 \item P 23 L 6: This section removed.
+\item P 23 LL 16-21 : we added language about the leaf-scale derived
+  shape of the ET-VPD curve in the introduction. We prefer this
+  section in the discussion, as we view it as a discussion of the
+  implication of our results (which is the derivation itself), and how
+  it is sensitive to stomatal model choice.
 \item P 24 L 24: Added clarifying language about the exponent.
@@
 
+\section{Reviewer 3}
+
+Thank you for a very high quality review.
+
+\subsection{Consequences of $\lambda$ in g$_1$ and uWUE}
+
+\subsubsection{Soil moisture and $\lambda$}
+
+Yes, $\lambda$, and by extension uWUE and g$_1$ can vary with
+environental conditions, including soil moisture. In the worst case,
+the consequences of this dependency for scientists applying our theory
+is that they must conceptually consider the change in ET with respect
+to VPD only for scenarios in which soil moisture is fixed. We believe
+this more restricted scenario still has utility for disentangling land
+atmosphere feedbacks, and understanding observations and more
+sophisticated models. However, \cite{Zhou_2015}'s results do suggest
+that uWUE (and by extension $\lambda$) may be somewhat constant with
+respect to soil moisture. The degree to which it is constant
+influences the degree to which a reader can relax their interpretation
+to one in which soil moisture is not fixed with respect to VPD. We
+attempted to shed some more light on this issue by plotting the
+distribution of uWUE with SWC for 66 fluxnet sites, but the signal to
+noise ratio was too low and when there was a signal of a relationship,
+it was inconsistent across sites. We did include this in supplemental
+material to aid the reader and draw attention to uWUE (and $\lambda$)
+- SWC relationships at the ecosystem scale as an important area for
+future research.
+
+\subsubsection{Interpretation issues with the use of the uncertainty
+  parameter}
+
+We agree that forcing all uncertainty (and real variability) into a
+single parameter made succesful interpretation our results extremely
+difficult. The general issues with this formulation lead us to remove
+the observational component of the paper, to focus on the derivation
+using plant parameters previously established in obsevational studies
+in the literature \citep{Zhou_2015, Medlyn_2017}. This simplification
+and focus also [hopefully] clarifies the manuscript and the
+presntation of results.
+
+\subsection{General comments}
+
+\begin{itemize}
+  \item \textbf{Symbol definition}: thank you for pointing out these mistakes
+    and/or potential sources of confusion. We have gone through the
+    manuscript and attempted to clarify/fix these issues.
+  \item \textbf{Dependence on $\lambda$}: we did present both uWUE and g$_1$ in
+    terms of $\lambda$ in our original manuscript, but we added
+    language drawing the readers attention to this, and elaborating
+    its importance for interpreting how they co-vary.
+  \item \textbf{Presentation of leaf-level stomatal conductace}: we thought
+    this might help the reader and draw attention that it the theory
+    originated at the leaf scale. However we agree with you that it
+    seems unnecesary given we do not use it, and makes the manuscript
+    less clear. We removed it.
+  \item \textbf{g$_a$ calculation}: in the newer idealized version of
+    the manuscript we do not calculate g$_a$, but represent a range of
+    reasonable values suggested by data and the literature
+    \citep[e.g.,][]{Raupach_1995}.
+  \item \textbf{Fitting of parameters with FLUXNET data}: as
+    mentioned, we removed this part of the analysis.
+\end{itemize}
+
+\subsection{Presentation of results}
+
+We agree that the results are more clear when all figures are in terms
+of $\frac{\partial ET}{\partial VPD}$. All figures are now presented
+in terms of $\frac{\partial ET}{\partial VPD}$, including those where
+only parameters in the sign term are varied.
+
+We also agree that our original manuscript was confusing and
+disjointed. Hopefully the signficiant changes to the structure of the
+manuscript have clarified, streamlined, and simplifed the
+results. Regarding Section 3.7, we moved it into a new ``discussion''
+section. While it could be placed in the methods, we view it as more
+of a discussion on the consequences of using a given stomatal
+conductance theory, and think it is hopefully easier for the reader to
+digest after seeing the concave up nature of our results.
+
+Yes, we did misuse ``leading order'' language, which we removed. Also many
+of the subjective interpretations of quantiative results have been
+removed with the removal of the observational component.
+
+\subsection{Minor comments}
+
+\begin{itemize}
+  \item \textbf{Introduction}: We added more discussion of previous
+    work on transpiration response to VPD, as elaborated on in Section
+    \ref{leaf}. We also introduce \cite{Novick_2016}'s results, and
+    added langauge recognizing there are more details to plant strategy
+    and response to VPD than our simplified logic in the introduction.
+  \item \textbf{P 9 results equations}: These were moved to methods.
+  \item \textbf{P 13, L32, Temperature effect}: Yes, the direct effect is
+    included in our idealized analysis (although because of CC the
+    effect through $\Delta$ should be larger than the ``direct''
+    effect).
+  \item \textbf{P 15 L 15 ``between''}: removed.
+\end{itemize}
+
 % \printbibliography

### 2019-02-28 doc/paper/intro.tex
done with reviewer 3, now just edit paper make it nice and then response to editor

@@
 regulation and a lack of soil moisture conservation does not seem to
-be a sensible strategy for plants from an evolutionary standpoint.
-
-** paragraph on soil moisture dependenec? **
+be a sensible strategy for plants from an evolutionary
+standpoint. This simplified logic explains generally why plants
+evolved to respod to VPD, but also excludes many possible details and
+special cases (e.g. plant to plant ineractionand highly specialized
+photosythesis strategies like Crassulacean acid metabolism
+photosntheisis).
 
@@
 develop a framework for evaluating the partial derivative of ET with
-respect to VPD. We explicitly include VPD's full effect on stomatal
+respect to VPD. It is indeed key to disconnect the impact of VPD from
+other variables as VPD is known to increase dramatically with future
+climate and limit ET more than soil moisture
+\citep{Novick_2016}. Nonetheless previous studies have typically been
+hampered by the covariations between different variables
+\citep[e.g. soil moisture and VPD, ][]{Lin_2018, Zhou_2019}, which
+limits the interpretability and understanding of the response of
+ecosystem-scale ET to VPD. We are able to disentangle the effect of
+VPD on ET because we explictly include VPD's full effect on stomatal
 conductance, including its impact on photosynthesis. The use of an
@@
 under which ecosystems operate. The leaf-scale results agree that
-stomatal conductance decreases in response to VPD \citep{Oren199,
+stomatal conductance decreases in response to VPD \citep{Oren1990,
 Damour2010}, which we expect to be true for an ecosystem as
@@
 ecosystem ET response to VPD deviates from previous leaf scale
-analyses. It is indeed key to disconnect the impact of VPD from other
-variables as VPD is known to increase dramatically with future climate
-** (Novick JGR 2017) **, yet other environmental conditions such as
-soil moisture and radiation will not increase as
-dramatically. Nonetheless previous studies have typically been
-hampered by the covariations between different variables
-\citep[e.g. radiation and VPD or soil moisture and VPD, ][]{Lin_2018,
-Zhou_2019}, which limits the interpretability and understanding of the
-response of ecosystem-scale ET to VPD.
+analyses.
 

### 2019-03-01 doc/paper/hess-revision/final-response.tex
added more supplementary material and it compiles

@@
 \documentclass[12pt]{article}
-% \usepackage{amsmath}
-% \usepackage{amssymb}
-% \usepackage{bm}
-% \usepackage{graphicx}
-% \usepackage{epstopdf}
-% \usepackage{xcolor}
-% \DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `basename #1 .tif`.png}
-% \usepackage{color}
-% \pagestyle{plain}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{bm}
+\usepackage{graphicx}
+\usepackage{epstopdf}
+\usepackage{xcolor}
+\usepackage{hyperref}
+\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `basename #1 .tif`.png}
+\usepackage{color}
+\pagestyle{plain}
 
@@
 
-\date{December 17th, 2018}
+\date{February 28th, 2019}
 
@@
 soil moisture dependence of uWUE (and partially by extension
-$\lambda$), we provided supplementaty material showing the
-distribution of uWUE with SWC for each of 66 FLUXNET sites. The
-functional relationship between uWUE and SWC varies, with a mix of
-sites showing strong and weak SWC-dependence. For all sites the ratio
-of signal to noise is very low, an unfortunate consequence of taking a
-ratio of two highly uncertain eddy-covariance derived fluxes. In
-general we find this analysis inconclusive. Our paper does not rely on
-assumptions about uWUE's functional relationship to soil moisture so
-we do not include these analysis in the body of the text, but these
-types of assumptions can make our theory more useful to the
-practitioner so we provide them for their interpretation and
-motivation for future research.
+$\lambda$), we provide figures in supplementary material showing the
+distribution of uWUE with SWC for each of 66 FLUXNET sites from the
+FLUXNET-2015 database. The functional relationship between uWUE and
+SWC varies, with a mix of sites showing strong and weak
+SWC-dependence. For all sites the ratio of signal to noise is very
+low, an unfortunate consequence of taking a ratio of two highly
+uncertain eddy-covariance derived fluxes. In general we find this
+analysis inconclusive. The paper does not rely on assumptions about
+uWUE's functional relationship with soil moisture so we do not include
+the figures in the body of the manuscript. But given that constant
+uWUE g$_1$ assumptions can make our theory more useful to the reader
+we provide the figures for their interpretation, and motivation for
+future research.
 
@@
 that uWUE (and by extension $\lambda$) may be somewhat constant with
-respect to soil moisture. The degree to which it is constant
-influences the degree to which a reader can relax their interpretation
-to one in which soil moisture is not fixed with respect to VPD. We
-attempted to shed some more light on this issue by plotting the
-distribution of uWUE with SWC for 66 fluxnet sites, but the signal to
-noise ratio was too low and when there was a signal of a relationship,
-it was inconsistent across sites. We did include this in supplemental
-material to aid the reader and draw attention to uWUE (and $\lambda$)
-- SWC relationships at the ecosystem scale as an important area for
-future research.
+across a range of conditions (including implicitly soil moisture). The
+degree to which it is constant influences the degree to which a reader
+can relax their interpretation to one in which soil moisture is not
+fixed with respect to VPD. We attempted to shed some more light on
+this issue by plotting the distribution of uWUE with SWC for 66
+fluxnet sites, but the signal to noise ratio was too low and when
+there was a signal of a relationship, it was inconsistent across
+sites. We did include this in supplemental material to aid the reader
+and draw attention to uWUE (and $\lambda$) - SWC relationships at the
+ecosystem scale as an important area for future research.
 

### 2019-03-01 doc/paper/hess-supplement.tex
added more supplementary material and it compiles

--- 2019-02-28 doc/paper/hess-revision/final-response.tex
@@
 \documentclass[12pt]{article}
-% \usepackage{amsmath}
-% \usepackage{amssymb}
-% \usepackage{bm}
-% \usepackage{graphicx}
-% \usepackage{epstopdf}
-% \usepackage{xcolor}
-% \DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `basename #1 .tif`.png}
-% \usepackage{color}
-% \pagestyle{plain}
-
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{bm}
+\usepackage{graphicx}
+\usepackage{epstopdf}
+\usepackage{xcolor}
+\usepackage{hyperref}
+\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `basename #1 .tif`.png}
+\usepackage{color}
+\pagestyle{plain}
+\usepackage{longtable}
 % \usepackage[
@@
 \bibliographystyle{agufull08}
+\maxdeadcycles=200
 
@@
 
-\title{Final Response: When does vapor pressure deficit drive or
+\title{Supplementary material: When does vapor pressure deficit drive or
   reduce evapotranspiration?}
@@
 
-Thank you very much for an excellent and very thoughtful
-review. Addressing your comments will greatly improve the
-manuscript. We agree with your comments, and think the manuscript
-could use substantial re-framing and rewording to clarify how we are
-answering our research question, and how the answer may vary with
-climate and environmental factors.
+\maketitle
 
-Just a quick general comment before addressing your specific
-comments. By including a PFT-focused analysis we did not fully
-communicate the major goal and scope of our project: we are trying to
-characterize the response of ET to VPD, with all other environment
-variables held fixed. To accomplish this, we need to formulate an
-explicit function of ET in terms of environmental variables and
-parameters, where any parameters can be approximated as constant with
-regards to some [arbitrary] VPD perturbation scenario. We will
-elaborate more below on why we used PFT-focused scenarios for much of
-our analysis. However, the goal of our manuscript was much more simple
-and fundamental: at a given place or time, if you introduce a
-perturbation to VPD, what is the immediate ET response (e.g. positive
-or negative)? Answering this question does not require the much
-stronger assertion that any parameters must be invariant within a
-given PFT.
+The manuscript analyzes the partial derivative of ET with respect to
+VPD. This assumes that all other quantites remain fixed, including the
+plant parameters g$_1$ and uWUE (and by extension, $\lambda$). In
+reality, these parameters may vary with environmental conditions, and
+specifically soil moisture. However because soil moisture
+\textit{only} enters the partial derivative directly through these
+plant terms, if the plant parameters are weak functions of soil
+moisture then the theory can be directly applied to a broader range of
+conceptual VPD scenarios, including observed compound events between
+high VPD and low soil moisture \citep{Zhou_2019}. To help the reader
+assess the soil moisture dependence of uWUE (and partially by
+extension $\lambda$ and g$_1$), we provide here figures showing the
+distribution of uWUE with SWC for each of 66 FLUXNET sites from the
+FLUXNET-2015 database. The functional relationship between uWUE and
+SWC varies, with a mix of sites showing strong and weak
+SWC-dependence. For all sites the ratio of signal to noise is very
+low, an unfortunate consequence of taking a ratio of two highly
+uncertain eddy-covariance derived fluxes. In general we find this
+analysis inconclusive.
 
-\section{Reviewer 1}
+The paper does not rely on assumptions about uWUE's functional
+relationship with soil moisture so we do not include the figures in
+the body of the manuscript. But given that constant uWUE g$_1$
+assumptions can make our theory more useful to the reader we provide
+the figures for their interpretation, and motivation for future
+research.
 
-Thank you for a very thoughtful review. You identified some critical
-flaws in our methods and presentation, and addressing these flaws
-have, in our opinion, greatly improved the strength and clarity of the
-manuscript.
+\input{supp-figs.tex}
 
-\subsection{Compounding uncertainties}
-\label{pft}
-As stated in
-\href{https://editor.copernicus.org/index.php/hess-2018-553-AC1.pdf}{our
-response} during the open discussion period, we agree
-with you that there are problems prescribing any parameters as fixed
-within a given PFT, as we would expect relatively large variability
-within a PFT. While this type of analysis was motivated by examples in
-previous literature, we think it is misleading in this case to present
-results where these parameters are suggested to be fixed. These issues
-were exacerbated by the manner in which we fit our parameters: it was
-unclear given our methods how model and observational uncertainty,
-which can be considerable, are partitioned into our PFT-wide estimates
-of g1 and uWUE. Also, as you point out, there is already previous
-literature on the values of g1 and uWUE, so in many ways by fitting
-out own values we are repeating previous work, but with defects given
-our handling of observational and model error. Also, much of our
-observational work, in addition to its flaws, decreased the clarity of
-our manuscript and distracted from the focus, which a flexible
-derivation, based completely on previous literature, of ET explicitly
-as a function of enviornmental variables and plant parameters.
+\section{FLUXNET2015 sites}
+  % below taken adapted from Trevor Keenan's FLUXNET citations
+  % at github.com/trevorkeenan/FLUXNET_citations
+  Table A1: Metadata and citations for flux sites used in this analysis.
+All data are gathered from \url{www.fluxdata.org}, and citations are aggregated using tools available at \url{https://github.com/trevorkeenan/FLUXNET_citations}.
+  \begin{longtable}{l l l l l l l}
+    \hline
+    \textbf{Site} &
+    \textbf{PFT} &
+    \textbf{Lat} &
+    \textbf{Lon} &
+    \textbf{Clim\textsuperscript{1}} &
+    \textbf{Period} &
+    \textbf{References} \\
+    [0.5ex]
+    \hline
+    \input{flux_sites.tex}
+    [1ex]
+    \hline
+  \end{longtable}
+\textsuperscript{1} K{\"o}ppen Climate classification.
 
-These realizations led us to make significant alterations to our
-manuscript. We no longer use observations or fit parameters
-ourselves. Instead we frame our theoretical derivation in terms of
-parameters established in the literature. Using this literature, we
-can then make statements about the general liklihood of a positive or
-negative ET response to increasing VPD in terms of climate, plant
-type, and photosynthetic pathway. This seemed a more natural
-delineation of scope: our paper presents a derivation tying plant
-parameters to an ET response to VPD, while previous literature
-establishes relationships between plant parameters and plant
-type/climate.
-
-\subsection{Attribution to physiological responses}
-
-In our open discussion
-\href{https://editor.copernicus.org/index.php/hess-2018-553-AC1.pdf}{response}
-we detailed our agreement with you that g1 and uWUE should vary within
-a given PFT. Removing this PFT-focused analysis should prevent any
-misinterpretation that we are making claims that these paremeters
-should be fixed within a given PFT. For convenience, we also reproduce
-a particlarly relevant passage from our orignal response:
-
-``What we need to communicate better is that a given ecosystem's state
-at a place or time controls its response to a VPD perturbation. We are
-making an approximation that we can parameterize the effect of the
-ecosystem's state on VPD response with g$_1$ and uWUE. These quantities
-can also vary due to soil moisture condition; the approximation we
-need to answer our research question is that they are fixed
-with respect to a VPD perturbation. In this sense, we do not view the
-two statements: ``plants that are evolved to bred to prioritize
-primary production over water conservation (e.g., crops) exhibit a
-higher likelihood of atmospheric demand-driven response'' and
-``ecosystem types are responding in this way because they have, on the
-whole, been subjected to less soil water limitation (due to the
-non-negligible effects of irrigation?)'' as mutually exclusive. In
-fact, we view them as consistent with a view that crops and their
-physiology (parameterized by g$_1$ and uWUE) exist at a given time and
-place because they have not been subjected to soil water limitation,
-and they have a given response to VPD because of their physiology
-(which is a direct effect of the environment). In this way, climate
-and land surface state are causes of the VPD response both directly
-and through their effect on plant physiology (parameterized by g$_1$
-and uWUE).''
-
-In addition we added language to the manuscript that our results
-should be used by a reader/scientist under the assumption that plant
-parameters are fixed with respect to the VPD perturbation
-considered. Previous literature \citep{Zhou_2015, Lin_2015} suggests
-that these parameters might also be approximately constant across a
-relatively broader range of plant types and environmental conditions,
-but our manucript avoids claims about this costancy other than the
-importance of conceptually considering these parameters to be constant
-when applying our theory to a problem of interest.
-
-The issue of soil moisture dependence is of particular importance,
-however. If our plant parameters are more stationary with respect to
-soil moisture, then our theory can be applied to a broader range of
-VPD scenarios, including observed compound events between high VPD and
-low soil moisture \citep{Zhou_2019}. To help the reader assess the
-soil moisture dependence of uWUE (and partially by extension
-$\lambda$), we provided supplementaty material showing the
-distribution of uWUE with SWC for each of 66 FLUXNET sites. The
-functional relationship between uWUE and SWC varies, with a mix of
-sites showing strong and weak SWC-dependence. For all sites the ratio
-of signal to noise is very low, an unfortunate consequence of taking a
-ratio of two highly uncertain eddy-covariance derived fluxes. In
-general we find this analysis inconclusive. Our paper does not rely on
-assumptions about uWUE's functional relationship to soil moisture so
-we do not include these analysis in the body of the text, but these
-types of assumptions can make our theory more useful to the
-practitioner so we provide them for their interpretation and
-motivation for future research.
-
-\section{Reviewer 2}
-
-Thank you again for a high quality review. The discussion it
-inspired improved the strength of the manuscript.
-
-\subsection{(i)}
-\label{(i)}
-Please see our
-\href{https://www.hydrol-earth-syst-sci-discuss.net/hess-2018-553/hess-2018-553-AC2-supplement.pdf}{previous
-  response} for a full discussion of the barriers to assessing the ET
-response to VPD using pure data. For convience, we includ an exerpt
-here:
-
-``One issue with analyses attempting to diagnose the ET response to
-VPD with pure data (e.g. using binned boxplots, as you suggest), is
-that ET varies strongly with both aerodynamic conductance and
-radiation. Simply binning by VPD or by site includes too many
-confounding factors to really diagnose what the attributable response
-is to VPD. This issue is really what motivated our analysis and makes
-it unique; by building an analytical framework based on confirmed
-results in the literature (uWUE, Medlyn model, Penman-Monteith [PM]),
-we are able to formally deduce what the actual ET response is, with
-other environmental factors held fixed. This ``other environmental
-factors held fixed'' component is, in our mind, nearly impossible to
-deduce from pure data, because the environment is always changing and
-direct analogues are either impossible to find or result in too small
-of a sample size to form a meaningful conclusion... Essentially we are
-saying that to the degree we trust arithmetic, uWUE, Penman-Monteith,
-and the Medlyn stomatal conductance model, our evaluation of the
-derivative is robust.''
-
-You also mention possible net radiation dependence on VPD, which we
-address in the minor comments.
-
-\subsection{(ii)}
-
-As in our
-\href{https://www.hydrol-earth-syst-sci-discuss.net/hess-2018-553/hess-2018-553-AC2-supplement.pdf}{previous
-  response}, we agree that we need more discussion connecting and
-contrasting leaf level results with our ecosystems scale approach. We
-have summarized the discussion from our reviewer response and added it
-to the introduction in lines ** **.
-
-\subsection{(iii)}
-
-We agree that are PFT-based formulation was too easy to be
-misintrepreted as an assertion that g1 and uWUE are constant and never
-vary in time or space. As described in Section \ref{pft}, we have
-removed the observational and PFT-focused component of the analysis,
-and instead focus exclusvively on using previously published constants
-(and their range) to interpret results. We would also like to
-reiterate that we do not view our use of uWUE as an introduction of an
-extra parameter. Our use of uWUE replaces a photosynthesis model,
-which will generally have many more parameters. Using uWUE is instead
-a simplifying approximation that relies on assumptions (based on
-optimal theory), but one that certainly reduces the total number of
-parameters of a more traditional estimate of ET.
-
-\subsection{(iv)}
-
-Yes, we should not have used ``leading-order'' language. We removed
-that confusing and inappropraite language. Also, by removing the
-empirical results and focusing exclusively theory framed by observational
-results from the literature, we have clarified the manuscript along
-lines you suggest (e.g. better seperation of theory vs empirical
-data).
-
-\subsection{Minor Comments}
-
-\begin{itemize}
-\item P 1 L 17: We removed ``stress'' language.
-\item P 1 L 22-24: Discussion of PET was removed to simplify
-  manuscript (it seemed less relevant). Otherwise the Roderick
-  citation is a good one, thanks.
-\item P 2 L 17: Discussion of precipitation removed to simplify
-  manuscript.
-\item P 2 L 33-35: You're right, and this was speculation; removed
-  this language and replaced with: ``However these researchers did not
-  explore explicitly the sensitivity of ET to VPD, including VPD's
-  effect on stomatal conductance and plant function.''
-\item P 3 L 2: Added citation.
-\item P 3 L 9: Replaced novel with useful.
-\item P 5 L 4: Yes it is related to water use efficiency, but we think
-  that could add more confusion given the ``water use efficiency''
-  language in uWUE; by expressing it in terms of $\lambda$ hopefully
-  it is clear to the reader that g$_1$ is related to water use
-  efficiency.
-\item P 6 L 4: Yes, that equation was mislabeled. Should be fixed now.
-\item P 7 L 14: We no longer fit our own values, so this is no
-  longer relevant.
-\item P 7 Eq. (10): Yes, a change in ET/VPD would induce indrect
-  effects like changes in temperature and (be extension) net
-  radiation. We added discussion guiding the reader on how to avoid
-  misuse partial derivatives, because it seems like readers may not
-  the sufficient understanding, and included explicit discussion of
-  net radiation as an example: ``Care must also be taken with possible
-  indirect effects associated with a change in VPD: for example, a
-  change in ET induced by a change in VPD can cause a change in
-  surface temperature, which would drive a change in net
-  radiation. These types of indirect effects and feedbacks are not
-  considered in Eq. (\ref{d_et}): temperature (a variable) is
-  mathematically fixed.''
-\item P 10 LL 2 and LL 5: Yes, ``physiological controls'' can be
-  misintrepreted and doesn't tell the full story. This language is
-  removed.
-\item P 10 Eq. (12): Yes there should have been a $\sigma$, but
-  $\sigma$ is now removed from the manuscript.
-\item P 10. L 11: Yes, both g$_1$ and uWUE can vary with soil moisture
-  as both are dependent on $\lambda$. This section is however removed
-  from the updated manuscript, but we also added language reflecting
-  this ``These parameters will however vary with plant species and
-  characteristics \citep[e.g. wood density, ][]{Lin_2015}, as well as
-  environmental conditions including soil water content
-  \citep{Manzoni2013} and temperature.''
-\item P 10 Equation (13): $\sigma$ should not have been removed, but
-  it is now no longer in the manuscript.
-\item P 10 L 29-30 ** think on this, gets back to $\lambda$ and how it
-  changes with aridity **
-\item P 11 L 14-17 : this discussion removed from new version of
-  paper.
-\item P 13 L 7-9: this discussion removed from new version of paper.
-\item Figure 4: removed from paper.
-\item P 16 L 16: Yes this was somewhat circular and limited: it only
-  allowed us to test if/how uWUE, given a prescribed model form
-  and the data, would systematically vary with VPD within a PFT. There
-  were definitely some issues with our PFT-focused methods and how we
-  handled uncertainty, which let us to remove this portion of the
-  analysis (see Section \ref{pft}).
-\item P 16, L 15: Agreed, and thankfully this whole analysis is now
-  removed from the paper.
-\item P 18 L 32 / P 21 L 1-3: These sections have been
-  removed. However, we do cannot completely follow the feedback logic
-  you are using; can you ellaborate? Also we provide some citations
-  suggesting that uWUE and g1 can be \textit{relatively} constant
-  \citep{Lin_2005, Franks_2017, Zhou_2015}, so would be interested in
-  citations showing that they are not constant for the majority fo
-  conditions. Please note we also added some plots of the relationship
-  between soil moisture and uWUE in our supplementary material to aid
-  the reader, but view this as an active research as well.
-\item P 23 L 6: This section removed.
-\item P 24 L 24: Added clarifying language about the exponent.
-  ``In general, increasing the exponent of VPD dependence for
-  either $uWUE*$ or $g_*$ increases the likelihood ``
-\item P 26 L 2-4: Please see discussion in Section \ref{(i)}.
-\item P 27 Appendix A: removed use of FLUXNET data in this manuscript.
-\end{itemize}
-
-% \printbibliography
 \bibliography{references}

### 2019-03-01 doc/paper/hess-revision/final-response.tex
everything compiles, now we can just read through

@@
 \begin{document}
-
-Thank you very much for an excellent and very thoughtful
-review. Addressing your comments will greatly improve the
-manuscript. We agree with your comments, and think the manuscript
-could use substantial re-framing and rewording to clarify how we are
-answering our research question, and how the answer may vary with
-climate and environmental factors.
-
-Just a quick general comment before addressing your specific
-comments. By including a PFT-focused analysis we did not fully
-communicate the major goal and scope of our project: we are trying to
-characterize the response of ET to VPD, with all other environment
-variables held fixed. To accomplish this, we need to formulate an
-explicit function of ET in terms of environmental variables and
-parameters, where any parameters can be approximated as constant with
-regards to some [arbitrary] VPD perturbation scenario. We will
-elaborate more below on why we used PFT-focused scenarios for much of
-our analysis. However, the goal of our manuscript was much more simple
-and fundamental: at a given place or time, if you introduce a
-perturbation to VPD, what is the immediate ET response (e.g. positive
-or negative)? Answering this question does not require the much
-stronger assertion that any parameters must be invariant within a
-given PFT.
 

### 2019-03-01 doc/paper/intro.tex
starting to read through paper

@@
 for evapotranspirated water \citep{Penman_1948, Monteith_1965}, but
-also reduce stomatal conductance through stomatal closure \citep{Leuning_1990, MEDLYN_2011} by
-p \citep{Rawson1977, Mott2007,
-Darmour2010}. Understanding the net evaportanspiration (ET) response
+also reduce stomatal conductance through stomatal closure
+\citep{Rawson1977, Leuning_1990, Mott2007,
+Darmour2010, MEDLYN_2011}. Understanding the net evaportanspiration (ET) response
 to these two opposing effects of changes in VPD is crucial for

### 2019-03-01 doc/paper/abstract.tex
progress through intro on pierres comments
also set up site specific SWC box plots
abstract done

@@
 understanding complex land atmosphere systems in terms of idealized
-scenarios.
+scenarios in which evapotranspiration responds to VPD only. This
+response is otherwise challenging to assess in an environment where
+many processes co-evolve together.

### 2019-03-01 doc/paper/intro.tex
abstract done
done with intro

@@
 also reduce stomatal conductance through stomatal closure
-\citep{Rawson1977, Leuning_1990, Mott2007,
-Darmour2010, MEDLYN_2011}. Understanding the net evaportanspiration (ET) response
+\citep{Rawson1977, Leuning_1990, Mott2007, Damour2010,
+MEDLYN_2011}. Understanding the net evapotranspiration (ET) response
 to these two opposing effects of changes in VPD is crucial for
-assessing the impact of environmental perturbations to VPD on the
-water cycle.
+assessing the impact of environmental VPD perturbations on the water
+cycle.
 
 The opposing effects of increased atmospheric demand and higher
-stomatal stress lead to two possible perspectives for how ET responds
+stomatal closure lead to two possible perspectives for how ET responds
 to shifts in VPD. The first, a hydrometeorological perspective, is
@@
 partially close in response to increased atmospheric dryness
-\citep{Farquhar_1978, Ball_1987, Leuning_1990, Katul_2009, MEDLYN_2011}. This
-leads to a plant physiology perspective, in which an increase in VPD
-may actually correspond to a decrease in ET because of stomatal
-closure \citep[e.g.][]{Rigden_2017}.  In other words, the question
-``When does VPD drive or reduce ET?'' can be related to whether plant
-regulation or atmospheric demand dominates the ET response.
+\citep{Farquhar_1978, Ball_1987, Leuning_1990, Katul_2009,
+MEDLYN_2011}. This leads to a plant physiology perspective, in which
+an increase in VPD may actually correspond to a decrease in ET because
+of stomatal closure \citep[e.g.][]{Rigden_2017}.  In other words, the
+question ``When does VPD drive or reduce ET?'' can be related to
+whether plant regulation or atmospheric demand dominates the ET
+response.
 
@@
 standpoint. This simplified logic explains generally why plants
-evolved to respod to VPD, but also excludes many possible details and
-special cases (e.g. plant to plant ineractionand highly specialized
-photosythesis strategies like Crassulacean acid metabolism
-photosntheisis).
+evolved to respond to VPD, but also excludes many details and special
+cases (e.g. plant to plant interaction and highly specialized
+photosynthesis strategies like Crassulacean acid metabolism
+photosynthesis).
 
@@
 approximations established in prior research as a tool to develop a
-framework for understanding plant response to atmospheric drying and
-evaluating the VPD dependence of ET while keeping other variables
-fixed. This framework will aid interpretation of observations, full
-complexity models, and facilitate the disentanglement of complex
-land-atmosphere feedbacks. In particular, it facilitates increased
-understading of the impact of climate change, given the expected
-increase of VPD.  In the past, similar simplified approaches were used
-to understand interactions between stomatal conductance,
+framework for understanding plant responses to atmospheric drying and
+the the VPD dependence of ET while keeping other variables fixed. This
+framework will aid interpretation of observations, full complexity
+models, and facilitate the disentanglement of complex land-atmosphere
+feedbacks. In particular, our approach has applications for
+understanding climate change impacts, given expected increases in VPD
+with rising temperature.  In the past, similar simplified approaches
+were used to understand interactions between stomatal conductance,
 evapotranspiration and the environment \citep[e.g.,][]{Jarvis_1984,
@@
 
-Recent results have drastically improved our understanding of VPD's
-impact on physiology, especially at the leaf
-level. \citet{MEDLYN_2011} developed a model for leaf-scale stomatal
-conductance ($g_s$), including VPD response, by combining an optimal
-photosynthesis theory \citep{Cowan_1977, Katul_2009, Katul_2010} with an empirical
-approach, and extended use of this model to the ecosystem scale in
+Approaching the problem of ET response to VPD is aided by recent
+results drastically improving our understanding of VPD's impact on
+physiology, especially at the leaf level. \citet{MEDLYN_2011}
+developed a model for leaf-scale stomatal conductance ($g_s$),
+including VPD response, by combining an optimal photosynthesis theory
+\citep{Cowan_1977, Katul_2009, Katul_2010} with an empirical approach,
+and extended use of this model to the ecosystem scale in
 \citet{Medlyn_2017}. Additionally, \citet{Zhou_2014} demonstrated that
@@
 uWUE greatly simplify complex plant physiological processes, they
-therefore still capture ecosystem behavior for vegetated surfaces
+ still capture ecosystem behavior for vegetated surfaces
 \citep{Medlyn_2017, Zhou_2014}, and are useful tools to transparently
@@
 develop a framework for evaluating the partial derivative of ET with
-respect to VPD. It is indeed key to disconnect the impact of VPD from
+respect to VPD. It is useful to disconnect the impact of VPD from
 other variables as VPD is known to increase dramatically with future
-climate and limit ET more than soil moisture
-\citep{Novick_2016}. Nonetheless previous studies have typically been
-hampered by the covariations between different variables
-\citep[e.g. soil moisture and VPD, ][]{Lin_2018, Zhou_2019}, which
-limits the interpretability and understanding of the response of
-ecosystem-scale ET to VPD. We are able to disentangle the effect of
-VPD on ET because we explictly include VPD's full effect on stomatal
-conductance, including its impact on photosynthesis. The use of an
-energy balance framework (PM) allows us to include in our analysis the
-effects of the energy cost of evaporating water from a surface, which
-is an important factor in the natural environment, compared to
-prescribed in situ envirnmental conditions. This is relevant because
-previous research focusing on the leaf scale \citep{Rawson1977,
-Turner1984, Oren1999, Darmour2010, Mott2013} does not consistently or
-analogously include the energetic constraints on evapotranspiration
-under which ecosystems operate. The leaf-scale results agree that
-stomatal conductance decreases in response to VPD \citep{Oren1990,
-Damour2010}, which we expect to be true for an ecosystem as
-well. However leaf scale results also indicate that transpiration
-usually increases with increasing VPD in a concave downward shape
-\citep[e.g.,][]{Rawson1977, Turner1983, Mott2013}, which may not be
-true for ET at the ecosystem scale once energetic constraints on ET
-are included in the analysis.  Our approach allows us to estimate the
-expected ET response to VPD at the ecosystem scale, including the
-effects of surface energy balance constraints, and assess how
-ecosystem ET response to VPD deviates from previous leaf scale
-analyses.
+climate and limit ET more than soil moisture \citep{Novick_2016}, but
+disentangling VPD effects from soil moisture effects has been
+difficult in previous research, given their co-variability
+\citep[][]{Lin_2018, Zhou_2019}.  We are
+able to disentangle the effect of VPD on ET because we explicitly
+include VPD's full effect on stomatal conductance, including its
+impact on photosynthesis. The use of an energy balance framework (PM)
+allows us to include in our analysis the effects of the energy cost of
+evaporating water from a surface, which is an important factor in the
+natural environment, compared to prescribed in situ environmental
+conditions. This is relevant because previous research focusing on the
+leaf scale \citep{Rawson1977, Turner1984, Oren1999, Damour2010,
+Mott2013} does not consistently or analogously include the energetic
+constraints on evapotranspiration under which ecosystems operate. The
+leaf-scale results agree that stomatal conductance decreases in
+response to VPD \citep{Oren1999, Damour2010}, which we expect to be
+true for an ecosystem as well. However leaf scale results also
+indicate that transpiration usually increases with increasing VPD in a
+concave downward shape \citep[e.g.,][]{Rawson1977, Turner1984,
+Mott2013}, which may not be true for ET at the ecosystem scale once
+energetic constraints on ET are included in the analysis.  Our
+approach allows us to estimate the expected ET response to VPD at the
+ecosystem scale, including the effects of surface energy balance
+constraints, and assess how ecosystem ET response to VPD deviates from
+previous leaf scale analyses.
 

### 2019-03-01 doc/paper/body.tex
done through methods, onto results

@@
   where $\Delta$ is the change in saturation vapor pressure with
-temperature, given by Clausius-Clapeyron ($\frac{d \; e_s}{d \; T}$),
-$R_{net}$ is the net radiation minus ground heat flux, $g_a$ is
-aerodynamic conductance, $\rho_a$ is air density, $c_p$ is specific
-heat of air at constant pressure, $\gamma$ is the psychometric
-constant, and $g_s$ is the stomatal conductance (Table
-\ref{definitions}). The issue with PM is that $g_s$ is a function of
-carbon uptake, which is has a strong functionally relation to ET
-through stomatal function. Therefore, when PM is formulated in terms
-of $g_s$, it is an implicit funciton of ET as a function of itself
-rather than an explicit function of ET. Here we will derive a new form
-of PM in which ET is an explicit function of plant parameters and
-environmental conditions, and use it to assess the ecosystem scale
-response to VPD (by taking a partial derivative).
+  temperature, given by Clausius-Clapeyron
+  ($\frac{d \; e_s}{d \; T}$), $R_{net}$ is the net radiation minus
+  ground heat flux, $g_a$ is aerodynamic conductance, $\rho_a$ is air
+  density, $c_p$ is specific heat of air at constant pressure,
+  $\gamma$ is the psychometric constant, and $g_s$ is the stomatal
+  conductance (Table \ref{definitions}). The issue with PM is that
+  $g_s$ is a function of carbon uptake, which is has a strong
+  functionally relation to ET through stomatal function. Therefore,
+  when PM is formulated in terms of $g_s$, it is an implicit function
+  of ET itself rather than an explicit function of ET. Here we will
+  derive a new form of PM in which ET is an explicit function of plant
+  parameters and environmental conditions, and use it to assess the
+  ecosystem scale response to VPD (by taking a partial derivative).
 
@@
     $\gamma$  & psychometric constant & Pa K$^{-1}$   & - \\
-    $g_{0}$  & leaf-scale mimimum stomatal conductance & m s$^{-1}$  & \citet{MEDLYN_2011} \\
-
-    $g_{s-leaf}$  & leaf-scale stomatal conductance & m s$^{-1}$  & \citet{MEDLYN_2011} \\
     $g_{s}$  &  stomatal conductance & m s$^{-1}$
                                    & \citet{Medlyn_2017} \\
-    $g_{1-leaf}$  & leaf-scale slope parameter & Pa$^{0.5}$
-                                   & \citet{MEDLYN_2011} \\
     $g_{1}$  & ecosystem-scale slope parameter & Pa$^{0.5}$ & \citet{Medlyn_2017} \\
@@
     $\Gamma$ & CO$_2$ compensation point & - & - \\
-    $\Gamma^*$ & CO$_2$ compensation point without dark respiration & - & - \\
+    $\Gamma^*$ & CO$_2$ compensation point without dark respiration &
+                                                                      - & - \\
+    GPP & gross primary production & $\mu$-mol C s$^{-1}$ m$^{-2}$ & -
+    \\
+    ET & evapotranspiration & W m$^{-2}$ & - \\
+   uWUE & underlying water use efficiency & $\mu$-mol C Pa$^{0.5}$
+                                            J$^{-1}$ ET  &
+                                                           \citet{Zhou_2015} \\
     \hline
@@
 \citet{MEDLYN_2011} developed a model for stomatal conductance ($g_s$)
-by combining an optimal photosynthesis theory \citep{Cowan_1977} with an empirical approach, which describes the
-dependence of $g_s$ to VPD. This resulted in the following model for
-leaf-scale stomatal conductance:
+by combining an optimal photosynthesis theory \citep{Cowan_1977} with
+an empirical approach, which describes the dependence of $g_s$ to
+VPD. They also extended this model to the ecosystem scale
+\citep{Medlyn_2017}:
 
@@
   \begin{equation}
-    g_{s-leaf} = g_0 + 1.6 \left(1 +
-      \frac{g_{1-leaf}}{\sqrt{VPD}}\right) \frac{A}{c_a},
-    \label{leaf_medlyn}
+    g_s = \frac{R \,T}{P} \; 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{GPP}{c_a},
+    \label{medlyn}
   \end{equation}
 % \end{linenomath*}
-where $g_{1-leaf}$ is a leaf-scale ``slope'' parameter, A is the net
-CO$_2$ assimilation rate, and $c_a$ is the atmospheric CO$_2$
-concentration at the leaf surface. \cite{MEDLYN_2011} relate the slope
-parameter ($g_{1-leaf}$) to physical parameters as:
+
+  where $g_{1}$ is a VPD ``slope'' parameter, GPP is the ecosystem
+  scale gross primary production, and $c_a$ is the atmospheric CO$_2$
+  concentration at the canopy. \cite{MEDLYN_2011} relate the slope
+  parameter ($g_{1}$) to physical parameters as:
 % \begin{linenomath*}
+
   \begin{equation}
-    g_{1-leaf} = \sqrt{\frac{3 \, \Gamma^* \,
+    g_{1}  \propto \sqrt{\frac{3 \, \Gamma^* \,
         \lambda}{1.6}},\footnote{Note this expression has units of of
@@
 of leaf carbon
-($\frac{\partial \; \text{transpiration}}{\partial \; A}$)
+($\frac{\partial \; \text{transpiration}}{\partial \; CO_2 assimilation}$)
 \citep{Farquhar_1980, Katul_2009}. So,
-$g_{1-leaf}$ is a leaf-scale term reflecting the trade-off of water for
-carbon uptake. The higher $g_{1-leaf}$, the more open the stomata and
+$g_{1}$ is a leaf-scale term reflecting the trade-off of water for
+carbon uptake. The higher $g_{1}$, the more open the stomata and
 the more they release water in exchange for carbon.
 
-The Medlyn model adopted to ecosystem scale analysis in
-\citet{Medlyn_2017}. In units of m s$^{-1}$, the ecosystem scale
-stomatal conductance is:
-
-% \begin{linenomath*}
-  \begin{equation}
-    g_s = \frac{R \,T}{P} \; 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{GPP}{c_a},
-    \label{medlyn}
-  \end{equation}
-% \end{linenomath*}
-
-  where GPP is the ecosystem scale gross primary production, and $g_1$
-  is an ecosystem scale analogue to $g_{1-leaf}$.
-
 While Eq. (\ref{medlyn}) can be used in PM (Eq. (1)), it will make
-  analytical work with the function intractable because $GPP$ is
-  functionally related to ET itself. Additionally, a perturbation to
-  VPD should induce a physiological plant response that will alter GPP
-  and cause an indirect change in stomatal conductance, in addition to
-  the direct effect of VPD in Eq. (\ref{medlyn}). Therefore, in order
-  to derive the response of ET to VPD, we must account for the
+analytical work with the function intractable because GPP is
+functionally related to ET itself. Additionally, a perturbation to VPD
+should induce a physiological plant response that will alter GPP and
+cause an indirect change in stomatal conductance, in addition to the
+direct effect of VPD in Eq. (\ref{medlyn}). Therefore, in order to
+derive the full plant response of ET to VPD, we must account for the
 functional relationship between GPP, ET, and VPD, and its effect on
-  stomatal conductance. We can use aforementioned semi-empirical
-  results of \citet{Zhou_2015}, which were inspired optimal
-  photosynthesis theory, as a tool to approach this
-  problem. \citet{Zhou_2015}, showed that underlying Water Use
-  Efficiency (uWUE):
+stomatal conductance. We can use aforementioned semi-empirical results
+of \citet{Zhou_2015}, which were inspired optimal photosynthesis
+theory, as a tool to approach this problem. \citet{Zhou_2015}, showed
+that underlying Water Use Efficiency (uWUE):
 
@@
   \citep{Zhou_2014}. The theoretical derivation of the square root VPD
-  dependence in $uWUE$ leverages the same assumptions used in
+  dependence in uWUE leverages the same assumptions used in
   \cite{MEDLYN_2011} to derive the square-root VPD dependence of the
@@
 concentration and compensation point, and is inversely proportional to
-the marginal water cost of leaf carbon.
+the marginal water cost of leaf carbon ($\lambda$). This relationship
+with $\lambda$ ($\propto \lambda^{-1/2}$) is important as it is the inverse of g$_1$'s
+relationship with $\lambda$ ($\propto \sqrt{\lambda}$). So, uWUE
+should increase as $\lambda$ decreases, and g$_1$ should decrease as
+$\lambda$ decreases.
 
@@
 
-  deriving an analytical framework for ecosystem response to
+  providing analytical framework for ecosystem response to
   atmospheric demand with environmental conditions held
@@
   we would expect the surface to be generally well coupled. An
-  additional assumption in the formulation of $uWUE$ \citep{Zhou_2014,
+  additional assumption in the formulation of uWUE \citep{Zhou_2014,
     Zhou_2015} and \citet{Medlyn_2017}'s stomatal conductance model is
@@
   ratios of transpiration to evaporation. Finally, we assume that
-  $g_1$ and $uWUE$ are constant with respect to the conceptual VPD
+  $g_1$ and uWUE are constant with respect to the conceptual VPD
   perturbation. Both quantities have been shown to be relatively
@@
   and characteristics \citep[e.g. wood density, ][]{Lin_2015}, as well
-  as environmental conditions including soil water content
-  \citep{Manzoni2013} and temperature.
+  as environmental conditions including soil water content and
+  temperature \citep{Lin_2015, Manzoni2013}. Exploring possible soil
+  moisture (in)dependnce of the plant parameters (g$_1$ and uWUE) is
+  particularly interesting because soil moisture only enters the
+  partial derivative directly through these plant prameters. If the
+  plant parameters are weak functions of soil moisture then the theory
+  can be directly applied to a broader range of conceptual VPD
+  scenarios, including observed compound events between high VPD and
+  low soil moisture \citep{Zhou_2019}. Supplementary material for this
+  manuscript explores the soil moisture dependence of the plant
+  parameters, but excessive noise and inconsistencies preclude
+  conclusions about the nature of any dependence. We provide it in
+  case it is of interest to the reader, and to motivate future
+  research.
 
-  As Eq. (\ref{d_et}) is a partial derivative, it's utility is as a
+   Because Eq. (\ref{d_et}) is a partial derivative, it's utility is as a
    conceptual model for the change in ET in response to VPD with all
-  other variables held fixed. This concept is a useful tool for
+   other variables held fixed. This provides is a useful tool for
    identifying the effect of VPD on ET through atmospheric demand and
-  plant response, and allows a practitioner to disentangle complicated
-  feedbacks when many quantities co-vary. However users of
-  Eq. (\ref{d_et}) should take care that their interpretation matches
-  the assumptions inherent in a partial derivative. For example,
-  results will only be as valid as the assumption that g$_1$ and uWUE
-  (and by extension $\lambda$) are fixed with respect to the user's
-  conceptual change in VPD. Fortunately, these quantities are
-  relatively constant for many under many scenarios
-  \citep{Franks_2017}. Care must also be taken with possible indirect
-  effects associated with a change in VPD: for example, a change in ET
-  induced by a change in VPD can cause a change in surface
-  temperature, which would drive a change in net radiation. These
-  types of indirect effects and feedbacks are not considered in
-  Eq. (\ref{d_et}): temperature (a variable) is mathematically fixed.
+   plant response, and allows a practitioner to disentangle
+   complicated feedbacks when many quantities co-vary. However users
+   of Eq. (\ref{d_et}) should take care that their interpretation
+   matches the assumptions inherent in a partial derivative. For
+   example, results will only be as valid as the assumption that g$_1$
+   and uWUE (and by extension $\lambda$) are fixed with respect to the
+   user's conceptual change in VPD. Care must also be taken with
+   possible indirect effects associated with a change in VPD: for
+   example, a change in ET induced by a change in VPD can cause a
+   change in surface temperature, which would drive a change in net
+   radiation. These types of indirect effects and feedbacks are not
+   considered in Eq. (\ref{d_et}): temperature (a variable) is
+   mathematically fixed.
+
+\subsection{Framing Eq. \ref{d_et} with previous research }
 
@@
   to magnitude:
-
-  ** comment on temperature and acclimation??? **
-  ** impact of T is much smaller than VPD? **
-  ** paragraph on soil moisture **
 
@@
 for the square root VPD dependence in both the stomatal conductance
-model and $uWUE$ based on the assumption that stomata behave to
+model and uWUE based on the assumption that stomata behave to
 maximize carbon gain while minimizing water loss, which observations
@@
 function optimally \citep{Zhou_2015, Lin_2018}. Specifically with
-regards to $uWUE$, one would not expect that this ecosystem scale WUE
+regards to uWUE, one would not expect that this ecosystem scale WUE
 quantity will respond to VPD exactly analogously to stomata. Direct
@@
 The results presented in Sect. \ref{results} indicates that for a
-given $uWUE$ and $g_1$, the ET dependence on VPD should be concave
+given uWUE and $g_1$, the ET dependence on VPD should be concave
 upward. In other words, there should be some local minimum in ET at a
 critical VPD$_{crit}$, assuming the scaling and plant terms
-(e.g. aerodynamic conductance, $\Delta$, $g_1$ and $uWUE$) are held
+(e.g. aerodynamic conductance, $\Delta$, $g_1$ and uWUE) are held
 fixed. This result warrants further investigation, because to our
@@
 our derived ET-VPD relationship is concave upward, particularly with
-respect to the exponent of VPD dependence in $uWUE$ and the Medlyn
+respect to the exponent of VPD dependence in uWUE and the Medlyn
 unified stomatal conductance model, as other models and empiracle
@@
     concave up regardless of the magnitude of the plant constants
-    $g_1$ and $uWUE$. Therefore, the general concave up nature of our
+    $g_1$ and uWUE. Therefore, the general concave up nature of our
     results, given an assumption of optimal behavior, is insensitive
@@
     m$), whether the solution is concave up or concave down does not
-    depend on $uWUE$.
+    depend on uWUE.
   \item In general, increasing the exponent of VPD dependence for

### 2019-03-01 doc/paper/hess-supplement.tex
done through methods, onto results

@@
 
+\section{Description of data}
+\subsection{Data}
+\label{data}
+We use both meteorological and eddy-covariance data from the FLUXNET2015
+database (data available at \sloppy
+\url{https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/} \sloppy),
+including all Tier 1 sites with at least four years of data and
+observations of the variables described in the methods section (Sect.
+\ref{methods}). Sixty-six sites met these requirements, and were grouped
+into nine plant functional types (PFT) according to the International
+Geosphere-Biosphere Programme vegetation classification scheme
+\citep{Loveland_1999}: cropland (CRO), grass (GRA), deciduous broadleaf
+forest (DBF), evergreen broadleaf forest (EBF), evergreen needleleaf
+forest (ENF), mixed forest (MF), closed shrub (CSH), savannah (SAV), and
+woody savannah (WSA) (site locations and citations are in Fig.
+\ref{map_fig} and Appendix \ref{flux_sites}, respectively).
+
+\begin{figure}[h]
+  \centering \includegraphics[trim={0 3cm 0 3cm}, clip]{./map.pdf}
+  \caption{Plant functional type and location of FLUXNET2015 sites
+    used in this analysis.}
+  \label{map_fig}
+\end{figure}
+We filter and quality control the FLUXNET-2015 data using a similar
+procedure as \cite{Zhou_2015}:
+
+\begin{itemize}
+\item Only measured or highest (``good'') quality gapfilled data,
+  according to quality control flags, are used.
+\item To isolate the growing season, we only use days in which the
+  average Gross Primary Productivity (GPP) exceeds 10\% of the
+  observed 95th percentile of GPP for a given site. GPP is calculated
+  using the nighttime respiration partitioning method.
+\item We remove days with rain and the day following to avoid issues
+  with rain interception and sensor saturation at high relative
+  humidity (\cite{Medlyn_2017}).
+\end{itemize}
+Additionally, as in \citet{Lin_2018}, we restrict data to the daytime,
+which is identified when downwelling shortwave radiation is greater
+than 50 W m$^{-2}$ and sensible heat flux is greater than 5 W
+m$^{-2}$. To reduce the chance of sensor saturation at high relative
+humidity, we remove all time steps for which VPD is less than .01 kPa,
+and to reduce errors at low windspeeds we remove all periods with wind
+magnitudes less than 0.5 m s$^{-1}$. Timesteps with negative observed
+GPP or ET are also removed, and we aggregate half hourly data to
+hourly averages to reduce noise \citep{Lin_2018}. After these quality
+control procedures, 400,983 upscaled hourly observations remain.
+
 \input{supp-figs.tex}

### 2019-03-01 doc/paper/body.tex
finished body, onto conclusions

@@
   replaced the added complexity of a stomatal conductance reduction
-  factor and a photosythesis model with a signle paramter (uWUE). Both
-  the plant paramters reflect water conservation strategy.The
+  factor and a photosynthesis model with a single parameter (uWUE). Both
+  the plant parameters reflect water conservation strategy.The
   slope parameter is related to the willingness of stomata to trade
-  water for CO$_2$ and to keep stomata open (carbon cost in terhms of
+  water for CO$_2$ and to keep stomata open (carbon cost in terms of
   water)). uWUE is a semi-empirical ecosystem-scale constant related
@@
 
-With Equation \ref{et} we can take the partial derivative of ET with respect
+With Eq. (\ref{et}) we can take the partial derivative of ET with respect
 to VPD to understand whether VPD drives or reduces ecosystem-scale ET:
@@
   temperature \citep{Lin_2015, Manzoni2013}. Exploring possible soil
-  moisture (in)dependnce of the plant parameters (g$_1$ and uWUE) is
+  moisture (in)dependence of the plant parameters (g$_1$ and uWUE) is
   particularly interesting because soil moisture only enters the
-  partial derivative directly through these plant prameters. If the
+  partial derivative directly through these plant parameters. If the
   plant parameters are weak functions of soil moisture then the theory
@@
   environment, we will systematically analyze how the ET response to
-  VPD (Eq. (\ref{d_et})) varies. Equation (\ref{d_et}) includes a
+  VPD (Eq. (\ref{d_et})) varies. Eq. (\ref{d_et}) includes a
   ``sign'' term that determines the sign of the response in addition
@@
 responses with VPD (water intensive), and smaller g$_1$ values lead to
-a higher non-linear VPD dependence of the response. g$_1$ and uWUE
-should generally co-vary with $\lambda$. As $\lambda$ increases, g$_1$
-should increase and uWUE should decrease. ** ellaborate **
+a higher non-linear VPD dependence of the response. However care
+should be exercised when interpreting the range of ET responses,
+because g$_1$ and uWUE should generally be anti-correlated due to
+their dependencies on $\lambda$. As $\lambda$ increases, g$_1$ should
+increase and uWUE should decrease. Because high and low values of uWUE
+and g$_1$ were determined independently from each other in
+previous literature, plant types with anomalously high or low values
+for \textit{both} g$_1$ and uWUE should be relatively rare.
 
@@
 to latent heat through $\Delta$), amplify or suppress the plant
-response represented in the sign term (Figure \ref{scaing}). Both
+response represented in the sign term (Figure \ref{scaling}). Both
 lower temperatures and higher aerodynamic conductance lead to
@@
 derivatives for understanding the behavior of complex systems in a
-simplifed framework. Our results are a conceptual tool for answering
+simplified framework. Our results are a conceptual tool for answering
 the question, ``What is the ET response to VPD with all other
@@
 confounding factors, there is little hope of disentangling and
-understanding the relative role of many covarying quantities in
-determining observed variability.  Equation \ref{d_et} explicitly
+understanding the relative role of many co-varying quantities in
+determining observed variability.  Eq. (\ref{d_et}) explicitly
 provides an estimate of the ET response to VPD, given assumptions
@@
 environmental conditions like soil moisture, we can still assess the
-ET response to VPD for a given plant species, or for a given soil
-moisture condition. ** talk about uWUE-SM, with supplementary here **?
+ET response to VPD for a given ecosystem state, or for a given soil
+moisture condition.
 
@@
 might expect some variation, especially in extreme cases when soil
-water becomes a limiting factor. ** talk about uWUE-SM, with supplementary here **?
+water becomes a limiting factor (see supplementary material).
 
@@
 between plant type and and plant parameter (g$_1$) that was
-demontrated by \citet{Zhou_2015} for uWUE. Given that there is a
+demonstrated by \citet{Zhou_2015} for uWUE. Given that there is a
 robust relationship between g$_1$ and plant type in \citet{Lin_2015},
@@
 scale observations, the consequences of imposing a model structure in
-some of their g$_1$ esimations, or real ecosystem-scale within-plant
+some of their g$_1$ estimations, or real ecosystem-scale within-plant
 type variability (e.g. multiple species impacting the relationship
-with the eddy-covariance footprint). These differences between
+within the eddy-covariance footprint). These differences between
 leaf-scale behavior and ecosystem scale behavior highlight the
@@
 scale behavior, and that some of the ambiguity in \citet{Medlyn_2017}
-is due to model and/or observational error, then we can g$_1$-induced
-ET response variability and plant characteristics, which are subject
+is due to model and/or observational error, then we can use the
+relationships between g$_1$ and plant characteristics defined by \citet{Lin_2015} to
+frame the ET response to VPD. This is subject
 to the strong caveat that we still do not fully understand g$_1$
@@
 particular, why and how g$_1$ behavior at the leaf scale is not
-analagous to the ecosystem scale must be understood. While the
+analogous to the ecosystem scale must be understood. While the
 derivation here was explicit about the assumption of a constant g$_1$
@@
 surface models and earth system models employing stomatal conductance
-models make similar, and even stronger assumptions about the
-stationarity of VPD slope terms \citep{Niu_2011, Franks_2017,
-  Rogers_2017, Lawrence_2019}. Many models will assume a constant
-g$_1$ term within a given plant functional type, and research derived
-from these models often does not acknowledge limitations of this
-assumption, or the difficulty in the literature with quantifying an
-ecosystem scale slope parameter \citep{Medlyn_2017}. Given that
-something so fundamental as the ET response to VPD varies strongly
-with g$_1$, we must invest more in understanding g$_1$'s behavior at
-the ecosystem scale, and the efficacy of models using g$_1$'s for
-ecosystem-scale fluxes. Fortunately, our framework, and the novel
-approach of using robust ecosystem-scale ratios (uWUE) to remove the
-implicit dependence of stomatal conductance on ET, is flexible enough
-to be applied to any stomatal conductance model that contains a
-dependence on GPP.
+models make similar assumptions about the stationarity of VPD slope
+terms \citep{Niu_2011, Franks_2017, Rogers_2017, Lawrence_2019}. Many
+models will assume a constant VPD slope term (e.g. g$_1$) term within
+a given plant functional type, and research derived from these models
+often does not acknowledge limitations of this assumption, or the
+difficulty in the literature with quantifying an ecosystem scale slope
+parameter \citep{Medlyn_2017}. Given that something so fundamental as
+the ET response to VPD varies strongly with g$_1$, we must invest more
+in understanding g$_1$'s behavior at the ecosystem scale, and the
+efficacy of models using constant slope terms for ecosystem-scale
+fluxes. Fortunately, our framework, and the new approach of using
+robust ecosystem-scale ratios (uWUE) to remove the implicit dependence
+of stomatal conductance on ET, is flexible enough to be applied to any
+stomatal conductance model that contains a dependence on GPP.
 
@@
 for the square root VPD dependence in both the stomatal conductance
-model and uWUE based on the assumption that stomata behave to
-maximize carbon gain while minimizing water loss, which observations
-also generally support \citep{Lloyd_1991, MEDLYN_2011, Lin_2015,
-  Zhou_2014, Zhou_2015, Medlyn_2017}. However, some purely empirical
-results that fit the exponent of the VPD dependence to data have shown
-that it may vary slightly from 1/2, suggesting that stomata, as well
-as ecosystem-scale quantities based on stomata theory, may not always
+model and uWUE based on the assumption that stomata behave to maximize
+carbon gain while minimizing water loss, which observations also
+generally support \citep{Lloyd_1991, MEDLYN_2011, Lin_2015, Zhou_2014,
+  Zhou_2015, Medlyn_2017}. However, some purely empirical results that
+fit the exponent of the VPD dependence to data have shown that it may
+vary slightly from 1/2, suggesting that stomata, as well as
+ecosystem-scale quantities based on stomata theory, may not always
 function optimally \citep{Zhou_2015, Lin_2018}. Specifically with
@@
 soil evaporation's contributions to ET should shift the exponent of
-the VPD dependence, especially at conditions of low GPP when we would
-expect a systematically larger portion of direct soil evaporation
-contributions to ET, because we would also expect lower amounts of
-transpiration at low GPP. \citet{Zhou_2015}'s results corroborate
-this: they found a mean empirically fit exponential VPD dependence of
-0.55, varying slightly from the theoretically optimal value of of 0.5
-for AmeriFlux sites. \citet{Lin_2018}'s results also show variance in
-the empirical exponent of the VPD dependence of the stomatal
-conductance model, but interpretation of this variance is more
-difficult as \citet{Lin_2018} do not handle GPP/A dependence of
-stomatal conductance in a directly analogous manner to the optimal
-theory in \citet{MEDLYN_2011} and \citet{Medlyn_2017}. Regardless,
-given that these recent results on the relationship between VPD, GPP,
-and ET \citep{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017} form the
+the VPD dependence. \citet{Zhou_2015}'s results corroborate this: they
+found a mean empirically fit exponential VPD dependence of 0.55,
+varying slightly from the theoretically optimal value of of 0.5 for
+AmeriFlux sites. \citet{Lin_2018}'s results also show variance in the
+empirical exponent of the VPD dependence in the stomatal conductance
+model, but interpretation of this variance is more difficult as
+\citet{Lin_2018} do not handle the GPP dependence of stomatal
+conductance in a directly analogous manner to the optimal theory in
+\citet{MEDLYN_2011} and \citet{Medlyn_2017}. Regardless, given that
+these recent results on the relationship between VPD, GPP, and ET
+\citep{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017} form the
 backbone of our analysis and are what allowed us to derive an explicit
 ET expression for the first time (Eq. (\ref{et})), we will analyze if
-and how assumptions about the exponent of the VPD dependence impacts
+and how assumptions about the exponent of the VPD dependence impact
 the shape of the ET-VPD dependence. This analysis is also important to
@@
 knowledge no earlier work has derived the theoretical ecosystem-scale
-relationship between ET and VPD while controlling for other
-environmental conditions. In particular there is an apparent lack of
-consensus over whether the shape of the ET-VPD curve should be concave
-upward (our result) or concave downward in the absence of dramatic
-water stress. Given that understanding the ET-VPD relationship of the
-one-way plant response is fundamental to hypothesizing about any
-feedbacks between the land surface and the atmosphere, we analyze why
-our derived ET-VPD relationship is concave upward, particularly with
-respect to the exponent of VPD dependence in uWUE and the Medlyn
-unified stomatal conductance model, as other models and empiracle
-results have suggested exponents varying from 1/2 \citep{Leuning_1990,
-  Zhou_2015, Lin_2018}.
+relationship between ET and VPD in an energy balance framework, while
+controlling for other environmental conditions. In particular there is
+an apparent lack of consensus over whether the shape of the ET-VPD
+curve should be concave upward (our result) or concave downward in the
+absence of dramatic water stress. Given that understanding the ET-VPD
+relationship of the one-way plant response is fundamental to
+hypothesizing about any feedbacks between the land surface and the
+atmosphere, we analyze why our derived ET-VPD relationship is concave
+upward, particularly with respect to the exponent of VPD dependence in
+uWUE and the Medlyn unified stomatal conductance model, as other
+models and empirical results have suggested exponents varying from 1/2
+\citep{Leuning_1990, Zhou_2015, Lin_2018}.
 
@@
 derivative of ET, using Eq. (\ref{m_n}) for stomatal conductance
-($g_s$), with respect to VPD: \input{d2_solutions.tex} With this
+($g_s$), with respect to VPD:
+
+\input{d2_solutions.tex}
+
+With this
 result we have defined the family of curves separating concave up from
@@
 
-While it is possible that in the future some other form of
-VPD dependence is derived, at present \cite{MEDLYN_2011} and
-\cite{Zhou_2014} firmly established n=m=1/2 as the most likely
-candidate given current theory and empirical data. Additionally, we
-argue that a concave up result matches physical intuition more than a
-concave down result. Plants must maintain nutrient and sugar transport
-through the phloem and xylem. To accomplish this, stomata must remain
-slightly open \citep{De_2013, Nikinmaa_2013, Ryan_2014}. Furthermore,
-even if complete stomatal closure were possible, cuticular water loss
-and [at the ecosystem-scale] direct soil evaporation are still sources
-of ET which increase with VPD, independent of stomatal
-closure ** cite katul??? **. Therefore, in the limit as VPD becomes large and we assume
-plants are exercising all strategies to reduce ET, any further increase
-in VPD should result in an increase in ET through cuticular water loss
-and/or direct soil evaporation. This inevitable transition from
-conditions when stomata respond strongly to VPD to conditions when
-stomata response is asymptoting towards full closure would cause a
-concave up ET-VPD curve, which is matched by our theory. In short,
-plant response becomes more limited as VPD increases, while
-atmospheric demand monotonically increases with VPD, leading to the
- result that atmospheric demand dominates plant response when
-atmospheric demand is high.
+While it is possible that in the future some other form of VPD
+dependence is derived, at present \cite{MEDLYN_2011} and
+\cite{Zhou_2014} established n=m=1/2 as the most likely candidate
+given current theory and empirical data. Additionally, we argue that a
+concave up result matches physical intuition more than a concave down
+result. Plants must maintain nutrient and sugar transport through the
+phloem and xylem. To accomplish this, stomata must remain slightly
+open \citep{De_2013, Nikinmaa_2013, Ryan_2014}. Furthermore, even if
+complete stomatal closure were possible, cuticular water loss and [at
+the ecosystem-scale] direct soil evaporation are still sources of ET
+which increase with VPD, independent of stomatal closure. Therefore,
+in the limit as VPD becomes large and we assume plants are exercising
+all strategies to reduce ET, any further increase in VPD should result
+in an increase in ET through cuticular water loss and/or direct soil
+evaporation. This inevitable transition from conditions when stomata
+respond strongly to VPD to conditions when stomata response is
+asymptoting towards full closure would cause a concave up ET-VPD
+curve, which is matched by our theory. In short, plant response
+becomes more limited as VPD increases, while atmospheric demand
+monotonically increases with VPD, leading to the result that
+atmospheric demand dominates plant response when atmospheric demand is
+high.
 
@@
 in soil water storage such that water becomes limiting. This could be
-represented by an extension of our framework in which uWUE is allowed
-to change with changing SWC. Here, we focus our results by framing
-theme with previous literature to build baseline intuition for ET-VPD
-dependence.
+represented by an extension of our framework in which $\lambda$ is
+allowed to decrease with decreasing soil water, increasing uWUE and
+decreasing g$_1$. Here, we focus our results by framing them with
+previous literature to build baseline intuition for ET-VPD dependence.

### 2019-03-01 doc/paper/conclusions.tex
done with pierre's comments, still need to put in soil moisture discussion and supplementary materials somewhere
done with conclusions

@@
 between plant parameters, plant types, and climate. The goal was to
-understand the range of likely ET responses to VPD and develop some
+understand the range of possible ET responses to VPD and develop some
 intuition for how the ET response may vary with plant types and
@@
 ET response to VPD can vary from strongly water conservative (ET
-always decreasing in response to increasing VPD), to strongly water
-intensive (ET always increasing in response to VPD), which is
-indicative of the diversity of possible plant water conservation
-strategies. Higher uWUE and g$_1$ values increase the likelihood of a
-postive ET response, while decreasing temperature and increasing
-g$_a$ amplify the magnitude of the response. Previous literature
-\citep{Zhou_2015, Lin_2015} suggests that crops, through association
-with higher g$_1$ and uWUE, are more likely to exhibit a positive ET
-response to VPD. Shrubs (lower uWUE), C4 plants (lower g$_1$),
-gymnosperm trees (lower g$_1$), and plants in arctic and boreal
-climates (lower g$_1$) are more likely to exhibit a negative ET
-response to VPD. However interpretation of g$_1$-induced variability
-is partially muddied by ambiguity in upscaling leaf-scale estimations
-of g$_1$ to the ecosystem scale \citep{Medlyn_2017}.
+decreasing in response to increasing VPD), to strongly water intensive
+(ET increasing in response to VPD), which is indicative of the
+diversity of possible plant water conservation strategies. Higher uWUE
+and g$_1$ values increase the likelihood of a positive ET response,
+while decreasing temperature and increasing g$_a$ amplify the
+magnitude of the response. Previous literature \citep{Zhou_2015,
+Lin_2015} suggests that crops, through association with higher g$_1$
+and uWUE, are more likely to exhibit a positive ET response to
+VPD. Shrubs (lower uWUE), C4 plants (lower g$_1$), gymnosperm trees
+(lower g$_1$), and plants in arctic and boreal climates (lower g$_1$)
+are more likely to exhibit a negative ET response to VPD. However
+interpretation of g$_1$-induced variability is partially muddied by
+ambiguity in applying leaf-scale estimations of g$_1$ to the ecosystem
+scale \citep{Medlyn_2017}.
 
-Our paper builds important intuition for how plants respond to VPD
+Our paper builds intuition for how plants respond to VPD
 perturbations. We show that given optimal stomatal function and fixed
@@
 concave upward, with ET increasing with increasing VPD as VPD
-increases past some critical value **Add VPD crit plot to
-results??**. However future research should focus on fully
-understanding the form of stomatal VPD dependence, as this result is
-sensitive to the exponent of VPD dependence, which we currently
-believe is 1/2 \citep{MEDLYN_2011, Zhou_2014}. Indeed, this
+increases past some critical value where $\frac{\partial \;
+ET}{\partial \; VPD} = 0$. However future research should focus on
+fully understanding the functional form of VPD dependence, as this
+concave up result is sensitive to the exponent of VPD dependence,
+which we currently believe is 1/2 for both uWUE and the stomatal
+conductance model \citep{MEDLYN_2011, Zhou_2014}. Indeed, this
 sensitivity to the exponent of VPD dependence is an important result
@@
 than physiological response to indirect effects like increased
-VPD. Here, we provide a framework for understanding ET response to ET
+VPD \citep{Novick_2016}. Here, we provide a framework for understanding ET response to VPD
 using a simplified model of two plant parameters. Feedbacks between the
@@
 within our framework, which opens the door for a hierarchy of more
-sophisticated climate- and plant-specific analysis of ET sensitivity
+sophisticated climate- and plant-specific analyses of ET sensitivity
 to environmental variables (including VPD). We argue that such

### 2019-03-01 doc/paper/hess-supplement.tex
done with supplement onto response

@@
 \usepackage{longtable}
-% \usepackage[
-% 	backend=biber,
-% 	url=false,
-% 	isbn=false,
-% 	doi=false,
-%         eprint=false,
-%         bibstyle=authoryear,
-%         style=authoryear,
-%         % style=numeric-comp,
-% 	% uniquename=false,
-% 	% uniquelist=false,
-% 	% firstinits=true,
-%  	natbib=true,
-% 	sorting=none,
-%         ]{biblatex}
-
-% \addbibresource{./references.bib}
+\usepackage{textcomp}
+\graphicspath{{../../../etc/plots/climate_et/test_plots/}}
 
@@
 \bibliographystyle{agufull08}
-\maxdeadcycles=200
 
@@
 \label{data}
-We use both meteorological and eddy-covariance data from the FLUXNET2015
-database (data available at \sloppy
+We use both meteorological and eddy-covariance data from the
+FLUXNET2015 database (data available at \sloppy
 \url{https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/} \sloppy),
-including all Tier 1 sites with at least four years of data and
-observations of the variables described in the methods section (Sect.
-\ref{methods}). Sixty-six sites met these requirements, and were grouped
-into nine plant functional types (PFT) according to the International
+including all Tier 1 sites with at least four years of data. Sixty-six
+sites met these requirements, and were grouped into nine plant
+functional types (PFT) according to the International
 Geosphere-Biosphere Programme vegetation classification scheme
-\citep{Loveland_1999}: cropland (CRO), grass (GRA), deciduous broadleaf
-forest (DBF), evergreen broadleaf forest (EBF), evergreen needleleaf
-forest (ENF), mixed forest (MF), closed shrub (CSH), savannah (SAV), and
-woody savannah (WSA) (site locations and citations are in Fig.
-\ref{map_fig} and Appendix \ref{flux_sites}, respectively).
+\citep{Loveland_1999}: cropland (CRO), grass (GRA), deciduous
+broadleaf forest (DBF), evergreen broadleaf forest (EBF), evergreen
+needleleaf forest (ENF), mixed forest (MF), closed shrub (CSH),
+savannah (SAV), and woody savannah (WSA).
 

### 2019-03-01 src/paper_figure_code/swc_boxplot.py

@@ -54,15 +54,16 @@ def plot_box(_df, ax=None, pft=False):
   orig_xlim = ax.get_xlim()
   #ax.plot(orig_xlim, [d_calc.CP/287.0, d_calc.CP/287.0], 'm--', linewidth=1.0)
   ylim = ax.get_ylim()#[-4.0, 4.0]
-  custom_xlabel(_df, ax, 'Volumetric SWC', fontsize=small_ax_fontsize-4)
-  custom_ylabel(_df, ax, 'uWUE', fontsize=small_ax_fontsize-4)
+  ax.set_xlabel('Volumetric SWC')
+  ax.set_ylabel('uWUE')
+  # custom_xlabel(_df, ax, 'Volumetric SWC', fontsize=small_ax_fontsize-4)
+  # custom_ylabel(_df, ax, , fontsize=small_ax_fontsize-4)
   if pft:
     ax.set_title(name_dict[_df.pft.iloc[0]],\
                  fontsize=fontsize+3)
   else:
-    ax.set_title('Site: %s, PFT: %s, b# obs: %d'\
-                 % (_df.site.iloc[0], _df.pft.iloc[0], _df.shape[0]),\
-                 fontsize=fontsize+3)
+    ax.set_title('Site: %s, PFT: %s, # obs: %d'\
+                 % (_df.site.iloc[0], _df.pft.iloc[0], _df.shape[0]))
   # if _df.pft.iloc[0] in xlim:
   #   _xlims = xlim[_df.pft.iloc[0]]
   #   ax.set_xlim(_xlims)

### 2019-03-01 doc/paper/hess-revision/final-response.tex
done with supplement onto response
finished, what I submitted. need to refactor tomorrow and add makdefile

@@
 \pagestyle{plain}
-
-% \usepackage[
-% 	backend=biber,
-% 	url=false,
-% 	isbn=false,
-% 	doi=false,
-%         eprint=false,
-%         bibstyle=authoryear,
-%         style=authoryear,
-%         % style=numeric-comp,
-% 	% uniquename=false,
-% 	% uniquelist=false,
-% 	% firstinits=true,
-%  	natbib=true,
-% 	sorting=none,
-%         ]{biblatex}
-
-% \addbibresource{./references.bib}
+\usepackage{longtable}
+\usepackage{textcomp}
+\graphicspath{{../../../etc/plots/climate_et/test_plots/}}
 
@@
 
+\section{To the editor}
+
+Thank you for facilitating a rewarding review process. In addressing
+the reviewers' comments we made significant alterations to the
+manuscript, bordering on a complete rewrite. We believe this was
+warranted to improve the clarity and address some methodological flaws
+identified by the reviewers.
+
+Relevant to your comments to us, we added a lot of language clarifying
+that with a partial derivative, environmental conditions and parameters
+should be considered fixed with respect to a (conceptual) change in
+VPD. Your comments suggest that such a conceptual scenario does not
+have scientific utility unless it is explicitly observed in the
+observations (e.g. we see observe only VPD changes and nothing
+else). On this point I think we disagree, as we believe such idealized
+and conceptual approaches still have tremendous utility disentangling
+complicated observations in which many variables are co-varying
+together, and represent a path towards deeper understanding for how
+extremely complicated systems behave. Understanding how one quantity
+induces a change in another, with all other quantities held fixed, is
+for us a crucial first step towards quantifying environmental drivers,
+causal relationships, and feedback pathways. A serious issue with the
+original version of the manuscript was that we believe we did a poor
+job communicating both the conceptual consequences and utility of a
+partial derivative approach. We hope that the language we added in
+this version of the manuscript more strongly communicates our approach
+and how it is intended to be used.
+
+We also hope that the manuscript is not misleading. Part (most?)  of
+the problem may have been afore mentioned deficiencies in our
+communication in the original version. We took a simpler, idealized
+approach with very explicit assumptions in large part because we value
+conceptual clarity and ease of interpretation. With improved language
+and organization in the updated manuscript version hopefully the
+reader feels the same way. The real benefit of this conceptual clarity
+is that it enables the reader to make a decision about whether the
+tool or result is useful or relevant to them. We think this has
+tremendous value, and find that many published papers have limited
+utility explicitly because they do not provide this value (it often
+feels impossible to identify all the assumptions some results rely
+on). Our reasons for mentioning ESM/GCMs was not to provide evidence
+that it is scientifically justified to use a constant slope parameter
+or $\lambda$, but to 1) draw attention of an application of our
+framework for interpreting GCMs/ESMs, and more importantly 2) draw
+attention to a difference in approaches: while we have made an
+assumption very explicit (increasing the interpretability of our
+results), many published results (based on ESMs) are inherently opaque
+because of the same assumption made implicitly and buried deep in
+documentation (and even more complicated by multiple such assumptions
+interacting non-linearly). Our worry is that there exists a
+possibility that the more explicit one makes assumptions, the lower
+the likelihood that results will be accepted and disseminated. We hope
+the relationship is the opposite. To us, a result we may not agree
+with, but for which we understand the assumptions behind it, is far
+more valuable than a result we agree with but for which we cannot
+decipher the underlying assumptions. However, hopefully improved
+communication and organization in the updated version of the
+manuscript mitigates a lot of the the concerns about misleading
+readers.
+
+Thank you again for your comments and facilitating such a high quality
+review process. We look forward to further comments and feedback, and
+greatly appreciate the significant time everyone has spent improving
+our manuscript.
+
+\section{To all reviewers}
+
+Thank you for all for very thoughtful reviews. You identified some
+critical flaws in our methods and presentation, as well as key points
+to elaborate upon. Addressing these flaws have, in our opinion,
+greatly improved the strength and clarity of the manuscript.
+
 \section{Reviewer 1}
-
-Thank you for a very thoughtful review. You identified some critical
-flaws in our methods and presentation, and addressing these flaws
-have, in our opinion, greatly improved the strength and clarity of the
-manuscript.
 
@@
 derivation, based completely on previous literature, of ET explicitly
-as a function of enviornmental variables and plant parameters.
+as a function of environmental variables and plant parameters.
 
@@
 parameters established in the literature. Using this literature, we
-can then make statements about the general liklihood of a positive or
+can then make statements about the general likelihood of a positive or
 negative ET response to increasing VPD in terms of climate, plant
@@
 a given PFT. Removing this PFT-focused analysis should prevent any
-misinterpretation that we are making claims that these paremeters
+misinterpretation that we are making claims that these parameters
 should be fixed within a given PFT. For convenience, we also reproduce
-a particlarly relevant passage from our orignal response:
+a particularly relevant passage from our original response:
+
+\medskip
 
@@
 
-In addition we added language to the manuscript that our results
-should be used by a reader/scientist under the assumption that plant
-parameters are fixed with respect to the VPD perturbation
-considered. Previous literature \citep{Zhou_2015, Lin_2015} suggests
-that these parameters might also be approximately constant across a
-relatively broader range of plant types and environmental conditions,
-but our manucript avoids claims about this costancy other than the
-importance of conceptually considering these parameters to be constant
-when applying our theory to a problem of interest.
+\medskip
+
+In addition we added language to the manuscript stating that our
+results should be used by a reader/scientist under the assumption that
+plant parameters are fixed with respect to the VPD perturbation
+considered.
 
@@
 the figures in the body of the manuscript. But given that constant
-uWUE g$_1$ assumptions can make our theory more useful to the reader
-we provide the figures for their interpretation, and motivation for
-future research.
+uWUE and g$_1$ assumptions can make our theory more useful to the
+reader we provide the figures for their interpretation, and motivation
+for future research.
 
 \section{Reviewer 2}
-
-Thank you again for a high quality review. The discussion it
-inspired improved the strength of the manuscript.
-
-\subsection{(i)}
+\subsection{(i)  ET Response to VPD diagnosed from data}
 \label{(i)}
@@
   response} for a full discussion of the barriers to assessing the ET
-response to VPD using pure data. For convience, we includ an exerpt
+response to VPD using pure data. For convenience, we inclined an excerpt
 here:
+
+\medskip
 
@@
 
+\medskip
+
 You also mention possible net radiation dependence on VPD, which we
@@
 
-\subsection{(ii)}
+\subsection{(ii) Leaf scale discussion}
 \label{leaf}
@@
 have summarized the discussion from our reviewer response and added it
-to the introduction in lines ** **.
+to the introduction.
 
-\subsection{(iii)}
+\subsection{(iii) Additional uWUE parameter}
 
 We agree that are PFT-based formulation was too easy to be
-misintrepreted as an assertion that g1 and uWUE are constant and never
+misinterpreted as an assertion that g1 and uWUE are constant and never
 vary in time or space. As described in Section \ref{pft}, we have
 removed the observational and PFT-focused component of the analysis,
-and instead focus exclusvively on using previously published constants
+and instead focus exclusively on using previously published constants
 (and their range) to interpret results. We would also like to
@@
 
-\subsection{(iv)}
+\subsection{(iv) Bad language}
 
 Yes, we should not have used ``leading-order'' language. We removed
-that confusing and inappropraite language. Also, by removing the
+that confusing and inappropriate language. Also, by removing the
 empirical results and focusing exclusively theory framed by observational
 results from the literature, we have clarified the manuscript along
-lines you suggest (e.g. better seperation of theory vs empirical
+lines you suggest (e.g. better separation of theory vs empirical
 data).
@@
   longer relevant.
-\item P 7 Eq. (10): Yes, a change in ET/VPD would induce indrect
+\item P 7 Eq. (10): Yes, a change in ET/VPD would induce indirect
   effects like changes in temperature and (be extension) net
@@
   radiation. These types of indirect effects and feedbacks are not
-  considered in Eq. (\ref{d_et}): temperature (a variable) is
+  considered in Eq. (6): temperature (a variable) is
   mathematically fixed.''
 \item P 10 LL 2 and LL 5: Yes, ``physiological controls'' can be
-  misintrepreted and doesn't tell the full story. This language is
+  misinterpreted and doesn't tell the full story. This language is
   removed.
@@
   from the updated manuscript, but we also added language reflecting
-  this ``These parameters will however vary with plant species and
-  characteristics \citep[e.g. wood density, ][]{Lin_2015}, as well as
-  environmental conditions including soil water content
-  \citep{Manzoni2013} and temperature.''
+  this ``These parameters will however vary with plant species
+  and characteristics \citep[e.g. wood density, ][]{Lin_2015}, as well
+  as environmental conditions including soil water content and
+  temperature \citep{Lin_2015, Manzoni2013}.''
 \item P 10 Equation (13): $\sigma$ should not have been removed, but
   it is now no longer in the manuscript.
-\item P 10 L 29-30 ** think on this, gets back to $\lambda$ and how it
-  changes with aridity **
+\item P 10 L 29-30 Yes, we added more language on both parameters, and
+  how they are related through $\lambda$.
 \item P 11 L 14-17 : this discussion removed from new version of
@@
   were definitely some issues with our PFT-focused methods and how we
-  handled uncertainty, which let us to remove this portion of the
+  handled uncertainty, which led us to remove this portion of the
   analysis (see Section \ref{pft}).
@@
   removed. However, we do cannot completely follow the feedback logic
-  you are using; can you ellaborate? Also we provide some citations
-  suggesting that uWUE and g1 can be \textit{relatively} constant
-  \citep{Lin_2005, Franks_2017, Zhou_2015}, so we would be interested
-  in citations showing that they are not constant for the majority fo
-  conditions. We also added some plots of the relationship between
-  soil moisture and uWUE in our supplementary material to aid the
-  reader, but view those as inconclusive and an active research as
-  well.
+  you are using; can you elaborate? We also added some plots of the
+  relationship between soil moisture and uWUE in our supplementary
+  material to aid the reader, but view those as inconclusive and an
+  are of active research as well. Finally, in general the updated  manuscript is
+  much better about acknowledging that uWUE and g$_1$ can vary through
+  time, but should be fixed with respect to a VPD perturbation.
 \item P 23 L 6: This section removed.
-\item P 23 LL 16-21 : we added language about the leaf-scale derived
-  shape of the ET-VPD curve in the introduction. We prefer this
-  section in the discussion, as we view it as a discussion of the
-  implication of our results (which is the derivation itself), and how
-  it is sensitive to stomatal model choice.
+\item P 23 LL 16-21 : W added language about the leaf-scale derived
+  shape of the ET-VPD curve in the introduction. We prefer the section
+  on the ET-VPD shape in the discussion, as we view it as a discussion
+  of the implication of our results (which is the derivation itself),
+  and how it is sensitive to stomatal model choice.
 \item P 24 L 24: Added clarifying language about the exponent.
@@
 
-Thank you for a very high quality review.
-
 \subsection{Consequences of $\lambda$ in g$_1$ and uWUE}
@@
 Yes, $\lambda$, and by extension uWUE and g$_1$ can vary with
-environental conditions, including soil moisture. In the worst case,
+environmental conditions, including soil moisture. In the worst case,
 the consequences of this dependency for scientists applying our theory
@@
 atmosphere feedbacks, and understanding observations and more
-sophisticated models. However, \cite{Zhou_2015}'s results do suggest
-that uWUE (and by extension $\lambda$) may be somewhat constant with
-across a range of conditions (including implicitly soil moisture). The
-degree to which it is constant influences the degree to which a reader
-can relax their interpretation to one in which soil moisture is not
-fixed with respect to VPD. We attempted to shed some more light on
-this issue by plotting the distribution of uWUE with SWC for 66
-fluxnet sites, but the signal to noise ratio was too low and when
-there was a signal of a relationship, it was inconsistent across
+sophisticated models. However, \cite{Zhou_2015}'s results do faintly
+suggest that uWUE (and by extension $\lambda$) may be somewhat
+constant with across a range of conditions (including implicitly soil
+moisture). The degree to which it is constant influences the degree to
+which a reader can relax their interpretation to one in which soil
+moisture is not fixed with respect to VPD. We attempted to shed some
+more light on this issue by plotting the distribution of uWUE with SWC
+for 66 FLUXNET sites, but the signal to noise ratio was too low and
+when there was a signal of a relationship, it was inconsistent across
 sites. We did include this in supplemental material to aid the reader
@@
 We agree that forcing all uncertainty (and real variability) into a
-single parameter made succesful interpretation our results extremely
-difficult. The general issues with this formulation lead us to remove
+single parameter made successful interpretation our results extremely
+difficult. The general issues with this formulation led us to remove
 the observational component of the paper, to focus on the derivation
-using plant parameters previously established in obsevational studies
+using plant parameters previously established in observational studies
 in the literature \citep{Zhou_2015, Medlyn_2017}. This simplification
 and focus also [hopefully] clarifies the manuscript and the
-presntation of results.
+presentation of results.
 
@@
     its importance for interpreting how they co-vary.
-  \item \textbf{Presentation of leaf-level stomatal conductace}: we thought
-    this might help the reader and draw attention that it the theory
+  \item \textbf{Presentation of leaf-level stomatal conductance}: we
+    thought this might help the reader and emphasize that the theory
     originated at the leaf scale. However we agree with you that it
-    seems unnecesary given we do not use it, and makes the manuscript
+    seems unnecessary given we do not use it, and makes the manuscript
     less clear. We removed it.
@@
 We also agree that our original manuscript was confusing and
-disjointed. Hopefully the signficiant changes to the structure of the
-manuscript have clarified, streamlined, and simplifed the
+disjointed. Hopefully the significant changes to the structure of the
+manuscript have clarified, streamlined, and simplified the
 results. Regarding Section 3.7, we moved it into a new ``discussion''
@@
 Yes, we did misuse ``leading order'' language, which we removed. Also many
-of the subjective interpretations of quantiative results have been
+of the subjective interpretations of quantitative results have been
 removed with the removal of the observational component.
@@
     \ref{leaf}. We also introduce \cite{Novick_2016}'s results, and
-    added langauge recognizing there are more details to plant strategy
+    added language recognizing there are more details to plant strategy
     and response to VPD than our simplified logic in the introduction.
@@
 % \printbibliography
-\bibliography{references}
+\bibliography{final-references}
 

### 2019-06-11 doc/paper/agu-james-submission.tex
starting to work on james submission for pierre

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
%
% This file includes commands and instructions
% given in the order necessary to produce a final output that will
% satisfy AGU requirements, including customized APA reference formatting.
%
% You may copy this file and give it your
% article name, and enter your text.
%
%
% Step 1: Set the \documentclass
%
%

%% To submit your paper:
\documentclass[draft]{agujournal2019}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers
%%%%%%%
% As of 2018 we recommend use of the TrackChanges package to mark revisions.
% The trackchanges package adds five new LaTeX commands:
%
%  \note[editor]{The note}
%  \annote[editor]{Text to annotate}{The note}
%  \add[editor]{Text to add}
%  \remove[editor]{Text to remove}
%  \change[editor]{Text to remove}{Text to add}
%
% complete documentation is here: http://trackchanges.sourceforge.net/
%%%%%%%

\draftfalse

%% Enter journal name below.
%% Choose from this list of Journals:
%
% JGR: Atmospheres
% JGR: Biogeosciences
% JGR: Earth Surface
% JGR: Oceans
% JGR: Planets
% JGR: Solid Earth
% JGR: Space Physics
% Global Biogeochemical Cycles
% Geophysical Research Letters
% Paleoceanography and Paleoclimatology
% Radio Science
% Reviews of Geophysics
% Tectonics
% Space Weather
% Water Resources Research
% Geochemistry, Geophysics, Geosystems
% Journal of Advances in Modeling Earth Systems (JAMES)
% Earth's Future
% Earth and Space Science
% Geohealth
%
% ie, \journalname{Water Resources Research}

\journalname{Enter journal name here}

\begin{document}

%% ------------------------------------------------------------------------ %%
%  Title
%
% (A title should be specific, informative, and brief. Use
% abbreviations only if they are defined in the abstract. Titles that
% start with general keywords then specific terms are optimized in
% searches)
%
%% ------------------------------------------------------------------------ %%

% Example: \title{This is a test title}

\title{=enter title here=}

%% ------------------------------------------------------------------------ %%
%
%  AUTHORS AND AFFILIATIONS
%
%% ------------------------------------------------------------------------ %%

% Authors are individuals who have significantly contributed to the
% research and preparation of the article. Group authors are allowed, if
% each author in the group is separately identified in an appendix.)

% List authors by first name or initial followed by last name and
% separated by commas. Use \affil{} to number affiliations, and
% \thanks{} for author notes.
% Additional author notes should be indicated with \thanks{} (for
% example, for current addresses).

% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
% Author\affil{3,4}\thanks{Also funded by Monsanto.}}

\authors{=list all authors here=}

% \affiliation{1}{First Affiliation}
% \affiliation{2}{Second Affiliation}
% \affiliation{3}{Third Affiliation}
% \affiliation{4}{Fourth Affiliation}

\affiliation{=number=}{=Affiliation Address=}
%(repeat as many times as is necessary)

%% Corresponding Author:
% Corresponding author mailing address and e-mail address:

% (include name and email addresses of the corresponding author.  More
% than one corresponding author is allowed in this LaTeX file and for
% publication; but only one corresponding author is allowed in our
% editorial system.)

% Example: \correspondingauthor{First and Last Name}{email@address.edu}

\correspondingauthor{=name=}{=email address=}

%% Keypoints, final entry on title page.

%  List up to three key points (at least one is required)
%  Key Points summarize the main points and conclusions of the article
%  Each must be 100 characters or less with no special characters or punctuation and must be complete sentences

% Example:
% \begin{keypoints}
% \item	List up to three key points (at least one is required)
% \item	Key Points summarize the main points and conclusions of the article
% \item	Each must be 100 characters or less with no special characters or punctuation and must be complete sentences
% \end{keypoints}

\begin{keypoints}
\item enter point 1 here
\item enter point 2 here
\item enter point 3 here
\end{keypoints}

%% ------------------------------------------------------------------------ %%
%
%  ABSTRACT and PLAIN LANGUAGE SUMMARY
%
% A good Abstract will begin with a short description of the problem
% being addressed, briefly describe the new data or analyses, then
% briefly states the main conclusion(s) and how they are supported and
% uncertainties.

% The Plain Language Summary should be written for a broad audience,
% including journalists and the science-interested public, that will not have 
% a background in your field.
%
% A Plain Language Summary is required in GRL, JGR: Planets, JGR: Biogeosciences,
% JGR: Oceans, G-Cubed, Reviews of Geophysics, and JAMES.
% see http://sharingscience.agu.org/creating-plain-language-summary/)
%
%% ------------------------------------------------------------------------ %%

%% \begin{abstract} starts the second page

\begin{abstract}
[ enter your Abstract here ]
\end{abstract}

\section*{Plain Language Summary}
[ enter your Plain Language Summary here or delete this section]

%% ------------------------------------------------------------------------ %%
%
%  TEXT
%
%% ------------------------------------------------------------------------ %%

%%% Suggested section heads:
% \section{Introduction}
%
% The main text should start with an introduction. Except for short
% manuscripts (such as comments and replies), the text should be divided
% into sections, each with its own heading.

% Headings should be sentence fragments and do not begin with a
% lowercase letter or number. Examples of good headings are:

% \section{Materials and Methods}
% Here is text on Materials and Methods.
%
% \subsection{A descriptive heading about methods}
% More about Methods.
%
% \section{Data} (Or section title might be a descriptive heading about data)
%
% \section{Results} (Or section title might be a descriptive heading about the
% results)
%
% \section{Conclusions}

\section{= enter section title =}
%Text here ===>>>

%%

%  Numbered lines in equations:
%  To add line numbers to lines in equations,
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}

%% Enter Figures and Tables near as possible to where they are first mentioned:
%
% DO NOT USE \psfrag or \subfigure commands.
%
% Figure captions go below the figure.
% Table titles go above tables;  other caption information
%  should be placed in last line of the table, using
% \multicolumn2l{$^a$ This is a table note.}
%
%----------------
% EXAMPLE FIGURES
%
% \begin{figure}
% \includegraphics{example.png}
% \caption{caption}
% \end{figure}
%
% Giving latex a width will help it to scale the figure properly. A simple trick is to use \textwidth. Try this if large figures run off the side of the page.
% \begin{figure}
% \noindent\includegraphics[width=\textwidth]{anothersample.png}
%\caption{caption}
%\label{pngfiguresample}
%\end{figure}
%
%
% If you get an error about an unknown bounding box, try specifying the width and height of the figure with the natwidth and natheight options. This is common when trying to add a PDF figure without pdflatex.
% \begin{figure}
% \noindent\includegraphics[natwidth=800px,natheight=600px]{samplefigure.pdf}
%\caption{caption}
%\label{pdffiguresample}
%\end{figure}
%
%
% PDFLatex does not seem to be able to process EPS figures. You may want to try the epstopdf package.
%

%
% ---------------
% EXAMPLE TABLE
%
% \begin{table}
% \caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
% \centering
% \begin{tabular}{l c}
% \hline
%  Run  & Time (min)  \\
% \hline
%   $l1$  & 260   \\
%   $l2$  & 300   \\
%   $l3$  & 340   \\
%   $h1$  & 270   \\
%   $h2$  & 250   \\
%   $h3$  & 380   \\
%   $r1$  & 370   \\
%   $r2$  & 390   \\
% \hline
% \multicolumn{2}{l}{$^{a}$Footnote text here.}
% \end{tabular}
% \end{table}

%% SIDEWAYS FIGURE and TABLE
% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
%
% \begin{sidewaysfigure}
% \includegraphics[width=20pc]{figsamp}
% \caption{caption here}
% \label{newfig}
% \end{sidewaysfigure}
%
%  \begin{sidewaystable}
%  \caption{Caption here}
% \label{tab:signif_gap_clos}
%  \begin{tabular}{ccc}
% one&two&three\\
% four&five&six
%  \end{tabular}
%  \end{sidewaystable}

%% If using numbered lines, please surround equations with \begin{linenomath*}...\end{linenomath*}
%\begin{linenomath*}
%\begin{equation}
%y|{f} \sim g(m, \sigma),
%\end{equation}
%\end{linenomath*}

%%% End of body of article

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional Appendix goes here
%
% The \appendix command resets counters and redefines section heads
%
% After typing \appendix
%
%\section{Here Is Appendix Title}
% will show
% A: Here Is Appendix Title
%
%\appendix
%\section{Here is a sample appendix}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Optional Glossary, Notation or Acronym section goes here:
%
%%%%%%%%%%%%%%
% Glossary is only allowed in Reviews of Geophysics
%  \begin{glossary}
%  \term{Term}
%   Term Definition here
%  \term{Term}
%   Term Definition here
%  \term{Term}
%   Term Definition here
%  \end{glossary}

%
%%%%%%%%%%%%%%
% Acronyms
%   \begin{acronyms}
%   \acro{Acronym}
%   Definition here
%   \acro{EMOS}
%   Ensemble model output statistics
%   \acro{ECMWF}
%   Centre for Medium-Range Weather Forecasts
%   \end{acronyms}

%
%%%%%%%%%%%%%%
% Notation
%   \begin{notation}
%   \notation{$a+b$} Notation Definition here
%   \notation{$e=mc^2$}
%   Equation in German-born physicist Albert Einstein's theory of special
%  relativity that showed that the increased relativistic mass ($m$) of a
%  body comes from the energy of motion of the body—that is, its kinetic
%  energy ($E$)—divided by the speed of light squared ($c^2$).
%   \end{notation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  ACKNOWLEDGMENTS
%
% The acknowledgments must list:
%
% >>>>	A statement that indicates to the reader where the data
% 	supporting the conclusions can be obtained (for example, in the
% 	references, tables, supporting information, and other databases).
%
% 	All funding sources related to this work from all authors
%
% 	Any real or perceived financial conflicts of interests for any
%	author
%
% 	Other affiliations for any author that may be perceived as
% 	having a conflict of interest with respect to the results of this
% 	paper.
%
%
% It is also the appropriate place to thank colleagues and other contributors.
% AGU does not normally allow dedications.

\acknowledgments
Enter acknowledgments, including your data availability statement, here.

%% ------------------------------------------------------------------------ %%
%% References and Citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% \bibliography{<name of your .bib file>} don't specify the file extension
%
% don't specify bibliographystyle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliography{ enter your bibtex bibliography filename here }

%Reference citation instructions and examples:
%
% Please use ONLY \cite and \citeA for reference citations.
% \cite for parenthetical references
% ...as shown in recent studies (Simpson et al., 2019)
% \citeA for in-text citations
% ...Simpson et al. (2019) have shown...
%
%
%...as shown by \citeA{jskilby}.
%...as shown by \citeA{lewin76}, \citeA{carson86}, \citeA{bartoldy02}, and \citeA{rinaldi03}.
%...has been shown \cite{jskilbye}.
%...has been shown \cite{lewin76,carson86,bartoldy02,rinaldi03}.
%... \cite <i.e.>[]{lewin76,carson86,bartoldy02,rinaldi03}.
%...has been shown by \cite <e.g.,>[and others]{lewin76}.
%
% apacite uses < > for prenotes and [ ] for postnotes
% DO NOT use other cite commands (e.g., \citet, \citep, \citeyear, \nocite, \citealp, etc.).
%

\end{document}

More Information and Advice:

%% ------------------------------------------------------------------------ %%
%
%  SECTION HEADS
%
%% ------------------------------------------------------------------------ %%

% Capitalize the first letter of each word (except for
% prepositions, conjunctions, and articles that are
% three or fewer letters).

% AGU follows standard outline style; therefore, there cannot be a section 1 without
% a section 2, or a section 2.3.1 without a section 2.3.2.
% Please make sure your section numbers are balanced.
% ---------------
% Level 1 head
%
% Use the \section{} command to identify level 1 heads;
% type the appropriate head wording between the curly
% brackets, as shown below.
%
%An example:
%\section{Level 1 Head: Introduction}
%
% ---------------
% Level 2 head
%
% Use the \subsection{} command to identify level 2 heads.
%An example:
%\subsection{Level 2 Head}
%
% ---------------
% Level 3 head
%
% Use the \subsubsection{} command to identify level 3 heads
%An example:
%\subsubsection{Level 3 Head}
%
%---------------
% Level 4 head
%
% Use the \subsubsubsection{} command to identify level 3 heads
% An example:
%\subsubsubsection{Level 4 Head} An example.
%
%% ------------------------------------------------------------------------ %%
%
%  IN-TEXT LISTS
%
%% ------------------------------------------------------------------------ %%
%
% Do not use bulleted lists; enumerated lists are okay.
% \begin{enumerate}
% \item
% \item
% \item
% \end{enumerate}
%
%% ------------------------------------------------------------------------ %%
%
%  EQUATIONS
%
%% ------------------------------------------------------------------------ %%

% Single-line equations are centered.
% Equation arrays will appear left-aligned.

Math coded inside display math mode \[ ...\]
 will not be numbered, e.g.,:
 \[ x^2=y^2 + z^2\]

 Math coded inside \begin{equation} and \end{equation} will
 be automatically numbered, e.g.,:
 \begin{equation}
 x^2=y^2 + z^2
 \end{equation}

% To create multiline equations, use the
% \begin{eqnarray} and \end{eqnarray} environment
% as demonstrated below.
\begin{eqnarray}
  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
        && + (y - y_{0}) \sin \Theta  \nonumber \\
  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
        && + (y - y_{0}) \cos \Theta.
\end{eqnarray}

%If you don't want an equation number, use the star form:
%\begin{eqnarray*}...\end{eqnarray*}

% Break each line at a sign of operation
% (+, -, etc.) if possible, with the sign of operation
% on the new line.

% Indent second and subsequent lines to align with
% the first character following the equal sign on the
% first line.

% Use an \hspace{} command to insert horizontal space
% into your equation if necessary. Place an appropriate
% unit of measure between the curly braces, e.g.
% \hspace{1in}; you may have to experiment to achieve
% the correct amount of space.

%% ------------------------------------------------------------------------ %%
%
%  EQUATION NUMBERING: COUNTER
%
%% ------------------------------------------------------------------------ %%

% You may change equation numbering by resetting
% the equation counter or by explicitly numbering
% an equation.

% To explicitly number an equation, type \eqnum{}
% (with the desired number between the brackets)
% after the \begin{equation} or \begin{eqnarray}
% command.  The \eqnum{} command will affect only
% the equation it appears with; LaTeX will number
% any equations appearing later in the manuscript
% according to the equation counter.
%

% If you have a multiline equation that needs only
% one equation number, use a \nonumber command in
% front of the double backslashes (\\) as shown in
% the multiline equation above.

% If you are using line numbers, remember to surround
% equations with \begin{linenomath*}...\end{linenomath*}

%  To add line numbers to lines in equations:
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}

### 2019-06-11 doc/paper/body.tex
starting to work on james submission for pierre

@@
 stomatal conductance. We can use aforementioned semi-empirical results
-of \citet{Zhou_2015}, which were inspired optimal photosynthesis
+of \citet{Zhou_2015}, which were inspired by optimal photosynthesis
 theory, as a tool to approach this problem. \citet{Zhou_2015}, showed
@@
 terms \citep{Niu_2011, Franks_2017, Rogers_2017, Lawrence_2019}. Many
-models will assume a constant VPD slope term (e.g. g$_1$) term within
+models will assume a constant VPD slope term (e.g. g$_1$) within
 a given plant functional type, and research derived from these models

### 2019-06-11 doc/paper/agu-james-submission.tex
need to test new submission and make minor reformattting

@@
 \usepackage{lineno}
-\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
+% \usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
 \usepackage{soul}
 \linenumbers
+
 %%%%%%%
@@
 
-%% Enter journal name below.
-%% Choose from this list of Journals:
-%
-% JGR: Atmospheres
-% JGR: Biogeosciences
-% JGR: Earth Surface
-% JGR: Oceans
-% JGR: Planets
-% JGR: Solid Earth
-% JGR: Space Physics
-% Global Biogeochemical Cycles
-% Geophysical Research Letters
-% Paleoceanography and Paleoclimatology
-% Radio Science
-% Reviews of Geophysics
-% Tectonics
-% Space Weather
-% Water Resources Research
-% Geochemistry, Geophysics, Geosystems
-% Journal of Advances in Modeling Earth Systems (JAMES)
-% Earth's Future
-% Earth and Space Science
-% Geohealth
-%
-% ie, \journalname{Water Resources Research}
-
-\journalname{Enter journal name here}
+\journalname{Journal of Advances in Modeling Earth Systems (JAMES)}
 
@@
 
-%% ------------------------------------------------------------------------ %%
-%  Title
-%
-% (A title should be specific, informative, and brief. Use
-% abbreviations only if they are defined in the abstract. Titles that
-% start with general keywords then specific terms are optimized in
-% searches)
-%
-%% ------------------------------------------------------------------------ %%
+\title{When does vapor pressure deficit drive or reduce
+  evapotranspiration?}
 
-% Example: \title{This is a test title}
+\authors{A. Massmann\affil{1}, P. Gentine\affil{1}, C. Lin\affil{1,2}}
 
-\title{=enter title here=}
+\affiliation{1}{Department of Earth and Environmental Engineering,
+  Columbia University, New York, NY 10027}
+\affiliation{2}{State Key Laboratory of Hydroscience and Engineering, Department of Hydraulic
+  Engineering, Tsinghua University, Beijing, CN 100084}
 
-%% ------------------------------------------------------------------------ %%
-%
-%  AUTHORS AND AFFILIATIONS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Authors are individuals who have significantly contributed to the
-% research and preparation of the article. Group authors are allowed, if
-% each author in the group is separately identified in an appendix.)
-
-% List authors by first name or initial followed by last name and
-% separated by commas. Use \affil{} to number affiliations, and
-% \thanks{} for author notes.
-% Additional author notes should be indicated with \thanks{} (for
-% example, for current addresses).
-
-% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
-% Author\affil{3,4}\thanks{Also funded by Monsanto.}}
-
-\authors{=list all authors here=}
-
-% \affiliation{1}{First Affiliation}
-% \affiliation{2}{Second Affiliation}
-% \affiliation{3}{Third Affiliation}
-% \affiliation{4}{Fourth Affiliation}
-
-\affiliation{=number=}{=Affiliation Address=}
-%(repeat as many times as is necessary)
-
-%% Corresponding Author:
-% Corresponding author mailing address and e-mail address:
-
-% (include name and email addresses of the corresponding author.  More
-% than one corresponding author is allowed in this LaTeX file and for
-% publication; but only one corresponding author is allowed in our
-% editorial system.)
-
-% Example: \correspondingauthor{First and Last Name}{email@address.edu}
-
-\correspondingauthor{=name=}{=email address=}
-
-%% Keypoints, final entry on title page.
-
-%  List up to three key points (at least one is required)
-%  Key Points summarize the main points and conclusions of the article
-%  Each must be 100 characters or less with no special characters or punctuation and must be complete sentences
-
-% Example:
-% \begin{keypoints}
-% \item	List up to three key points (at least one is required)
-% \item	Key Points summarize the main points and conclusions of the article
-% \item	Each must be 100 characters or less with no special characters or punctuation and must be complete sentences
-% \end{keypoints}
+\correspondingauthor{Adam Massmann}{akm2203@columbia.edu}
 
@@
 
-%% ------------------------------------------------------------------------ %%
-%
-%  ABSTRACT and PLAIN LANGUAGE SUMMARY
-%
-% A good Abstract will begin with a short description of the problem
-% being addressed, briefly describe the new data or analyses, then
-% briefly states the main conclusion(s) and how they are supported and
-% uncertainties.
-
-% The Plain Language Summary should be written for a broad audience,
-% including journalists and the science-interested public, that will not have 
-% a background in your field.
-%
-% A Plain Language Summary is required in GRL, JGR: Planets, JGR: Biogeosciences,
-% JGR: Oceans, G-Cubed, Reviews of Geophysics, and JAMES.
-% see http://sharingscience.agu.org/creating-plain-language-summary/)
-%
-%% ------------------------------------------------------------------------ %%
-
-%% \begin{abstract} starts the second page
-
 \begin{abstract}
-[ enter your Abstract here ]
+\input{abstract.tex}
 \end{abstract}
@@
 \section*{Plain Language Summary}
-[ enter your Plain Language Summary here or delete this section]
 
-%% ------------------------------------------------------------------------ %%
-%
-%  TEXT
-%
-%% ------------------------------------------------------------------------ %%
+Plants can sense increasing dryness in the air and close up the pores
+on their leaves, preventing water loss. However, drier air also
+naturally demands more water from the land surface. Here we develop a
+simplified theory for when land surface water loss increases
+(atmospheric demand dominates) or decreases (plant response dominates)
+in response to increased dryness in the air. This theory provides
+intuition for how ecosystems regulate water in response to changes in
+atmospheric dryness. According to the theory ecosystems are capable of
+broad range of behavior in response to increased atmospheric dryness,
+from strongly reducing water loss to allowing large increases in water
+loss. Ecosystem behavior depends both on environmental conditions and
+plant type.
 
-%%% Suggested section heads:
-% \section{Introduction}
-%
-% The main text should start with an introduction. Except for short
-% manuscripts (such as comments and replies), the text should be divided
-% into sections, each with its own heading.
+\section{Introduction}
 
-% Headings should be sentence fragments and do not begin with a
-% lowercase letter or number. Examples of good headings are:
+\input{intro.tex}
 
-% \section{Materials and Methods}
-% Here is text on Materials and Methods.
-%
-% \subsection{A descriptive heading about methods}
-% More about Methods.
-%
-% \section{Data} (Or section title might be a descriptive heading about data)
-%
-% \section{Results} (Or section title might be a descriptive heading about the
-% results)
-%
-% \section{Conclusions}
+\input{body.tex}
 
-\section{= enter section title =}
+\section{Conclusions}
+
+\input{conclusions.tex}
+
 %Text here ===>>>
@@
 
-\acknowledgments
-Enter acknowledgments, including your data availability statement, here.
+\acknowledgments This work used eddy covariance data acquired and
+shared by the FLUXNET community, including these networks: AmeriFlux,
+AfriFlux, AsiaFlux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont,
+ChinaFlux, Fluxnet-Canada, GreenGrass, ICOS, KoFlux, LBA, NECC,
+OzFlux-TERN, TCOS-Siberia, and USCCC. The ERA-Interim reanalysis data
+are provided by ECMWF and processed by LSCE. The FLUXNET eddy
+covariance data processing and harmonization was carried out by the
+European Fluxes Database Cluster, AmeriFlux Management Project, and
+Fluxdata project of FLUXNET, with the support of CDIAC and ICOS
+Ecosystem Thematic Center, and the OzFlux, ChinaFlux and AsiaFlux
+offices. We would like to thank Dr. Trevor Keenan for providing tools
+for citation of the FLUXNET2015 dataset, available at
+\url{https://github.com/trevorkeenan/FLUXNET_citations}. This material
+is based upon work supported by the National Science Foundation
+Graduate Research Fellowship under Grant No. DGE 16-44869.
 
@@
 
-\bibliography{ enter your bibtex bibliography filename here }
+\bibliography{references}
 

### 2019-06-19 doc/paper/body.tex
james submission compiles, need to proof read and encorporate with makefile

@@
 \label{methods}
-The Penman-Monteith equation \citep [hereafter PM,][]{Penman_1948,
+The Penman-Monteith equation \cite<hereafter PM,>[]{Penman_1948,
   Monteith_1965} estimates ET as a function of observable atmospheric
@@
     $R_{net}$  & net radiation at land surface minus ground heat flux & W m$^{-2}$   & - \\
-    $g_a$  & aerodynamic conductance & m s$^{-1}$  & \citet{Shuttleworth_2012} \\
+    $g_a$  & aerodynamic conductance & m s$^{-1}$  & \citeA{Shuttleworth_2012} \\
     $\rho_a$  & air density & kg m$^{-3}$  & - \\
@@
     $g_{s}$  &  stomatal conductance & m s$^{-1}$
-                                   & \citet{Medlyn_2017} \\
-    $g_{1}$  & ecosystem-scale slope parameter & Pa$^{0.5}$ & \citet{Medlyn_2017} \\
+                                   & \citeA{Medlyn_2017} \\
+    $g_{1}$  & ecosystem-scale slope parameter & Pa$^{0.5}$ & \citeA{Medlyn_2017} \\
     $R$ & universal gas constant & J mol$^{-1}$ K$^{-1}$ & - \\
@@
                                             J$^{-1}$ ET  &
-                                                           \citet{Zhou_2015} \\
+                                                           \citeA{Zhou_2015} \\
     \hline
@@
 
-\citet{MEDLYN_2011} developed a model for stomatal conductance ($g_s$)
-by combining an optimal photosynthesis theory \citep{Cowan_1977} with
+\citeA{MEDLYN_2011} developed a model for stomatal conductance ($g_s$)
+by combining an optimal photosynthesis theory \cite{Cowan_1977} with
 an empirical approach, which describes the dependence of $g_s$ to
 VPD. They also extended this model to the ecosystem scale
-\citep{Medlyn_2017}:
+\cite{Medlyn_2017}:
 
@@
 ($\frac{\partial \; \text{transpiration}}{\partial \; CO_2 assimilation}$)
-\citep{Farquhar_1980, Katul_2009}. So,
+\cite{Farquhar_1980, Katul_2009}. So,
 $g_{1}$ is a leaf-scale term reflecting the trade-off of water for
@@
 stomatal conductance. We can use aforementioned semi-empirical results
-of \citet{Zhou_2015}, which were inspired by optimal photosynthesis
-theory, as a tool to approach this problem. \citet{Zhou_2015}, showed
+of \citeA{Zhou_2015}, which were inspired by optimal photosynthesis
+theory, as a tool to approach this problem. \citeA{Zhou_2015}, showed
 that underlying Water Use Efficiency (uWUE):
@@
   VPD over a diurnal cycle during the growing season
-  \citep{Zhou_2014}. The theoretical derivation of the square root VPD
+  \cite{Zhou_2014}. The theoretical derivation of the square root VPD
   dependence in uWUE leverages the same assumptions used in
@@
   with Eq. (\ref{et}) we have used recent results
-  \citep[][]{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017} to derive
+  \cite{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017} to derive
   ET explicitly as function of environmental variables and two
@@
 
-where $\Gamma$ is the CO$_2$ compensation point \citep[Eq. (5)
-in][]{Zhou_2014}. So uWUE is related to atmospheric CO$_2$
+where $\Gamma$ is the CO$_2$ compensation point \cite<Eq. (5)
+in>[]{Zhou_2014}. So uWUE is related to atmospheric CO$_2$
 concentration and compensation point, and is inversely proportional to
@@
   conditions and plant types the leaves can become decoupled from the
-  boundary layer \citep{De_2017, Medlyn_2017}. Therefore, our
+  boundary layer \cite{De_2017, Medlyn_2017}. Therefore, our
   derivation will be most applicable in times like the growing season
@@
   we would expect the surface to be generally well coupled. An
-  additional assumption in the formulation of uWUE \citep{Zhou_2014,
-    Zhou_2015} and \citet{Medlyn_2017}'s stomatal conductance model is
+  additional assumption in the formulation of uWUE \cite{Zhou_2014,
+    Zhou_2015} and \citeA{Medlyn_2017}'s stomatal conductance model is
   that direct soil evaporation contributions to ET remain small
@@
   perturbation. Both quantities have been shown to be relatively
-  constant with respect to changes in VPD \citep{Franks_2017,
+  constant with respect to changes in VPD \cite{Franks_2017,
     Zhou_2014}. These parameters will however vary with plant species
-  and characteristics \citep[e.g. wood density, ][]{Lin_2015}, as well
+  and characteristics \cite<e.g. wood density, >[]{Lin_2015}, as well
   as environmental conditions including soil water content and
-  temperature \citep{Lin_2015, Manzoni2013}. Exploring possible soil
+  temperature \cite{Lin_2015, Manzoni2013}. Exploring possible soil
   moisture (in)dependence of the plant parameters (g$_1$ and uWUE) is
@@
   scenarios, including observed compound events between high VPD and
-  low soil moisture \citep{Zhou_2019}. Supplementary material for this
+  low soil moisture \cite{Zhou_2019}. Supplementary material for this
   manuscript explores the soil moisture dependence of the plant
@@
 literature-based expectations for a range of growing season conditions
-and plant types \citep{Zhou_2015, Medlyn_2017}. Using this previous
+and plant types \cite{Zhou_2015, Medlyn_2017}. Using this previous
 literature we can connect the effect of varying plant parameters to
@@
 critical, for understanding complex responses when many quantities are
-varying simultaneously \citep{Zhou_2019}. If we cannot understand the
+varying simultaneously \cite{Zhou_2019}. If we cannot understand the
 contribution of one term to observed variability, independent of other
@@
 framed by our results on $\frac{\partial \; ET}{\partial \; VPD}$. In
-\citet{Zhou_2015} most plant types have uWUE similar to our median
+\citeA{Zhou_2015} most plant types have uWUE similar to our median
 value, under which the sign of the ET response depends on
@@
 we would then expect a decrease of ET in response to VPD. It is
-important to note that, while \citet{Zhou_2015} did not examine the
+important to note that, while \citeA{Zhou_2015} did not examine the
 role of soil moisture for within plant-type variability of uWUE, we
@@
 
-\citet{Medlyn_2017}'s results did not display the robust relationship
+\citeA{Medlyn_2017}'s results did not display the robust relationship
 between plant type and and plant parameter (g$_1$) that was
-demonstrated by \citet{Zhou_2015} for uWUE. Given that there is a
-robust relationship between g$_1$ and plant type in \citet{Lin_2015},
-and that \citet{Zhou_2015}'s ecosystem-scale results rely on the same
-underlying theory as \citet{Medlyn_2017}'s results, the ambiguous
-results in \citet{Medlyn_2017} could be due to noise in ecosystem
+demonstrated by \citeA{Zhou_2015} for uWUE. Given that there is a
+robust relationship between g$_1$ and plant type in \citeA{Lin_2015},
+and that \citeA{Zhou_2015}'s ecosystem-scale results rely on the same
+underlying theory as \citeA{Medlyn_2017}'s results, the ambiguous
+results in \citeA{Medlyn_2017} could be due to noise in ecosystem
 scale observations, the consequences of imposing a model structure in
@@
 increased variability at the ecosystem scale, as argued by
-\citet{Medlyn_2017}. However, if we assume that there is some real
-analogy between leaf scale results in \citet{Lin_2015} and ecosystem
-scale behavior, and that some of the ambiguity in \citet{Medlyn_2017}
+\citeA{Medlyn_2017}. However, if we assume that there is some real
+analogy between leaf scale results in \citeA{Lin_2015} and ecosystem
+scale behavior, and that some of the ambiguity in \citeA{Medlyn_2017}
 is due to model and/or observational error, then we can use the
-relationships between g$_1$ and plant characteristics defined by \citet{Lin_2015} to
+relationships between g$_1$ and plant characteristics defined by \citeA{Lin_2015} to
 frame the ET response to VPD. This is subject
 to the strong caveat that we still do not fully understand g$_1$
-behavior at the ecosystem scale. Extrapolating \citet{Lin_2015}'s
+behavior at the ecosystem scale. Extrapolating \citeA{Lin_2015}'s
 results (e.g. Fig. 2) to the ecosystem scale would reveal the
@@
 increases in VPD'' may be false depending on the water conservation
-strategy of a given ecosystem \citep{Lemordant_2018}.
+strategy of a given ecosystem \cite{Lemordant_2018}.
 
@@
 models make similar assumptions about the stationarity of VPD slope
-terms \citep{Niu_2011, Franks_2017, Rogers_2017, Lawrence_2019}. Many
+terms \cite{Niu_2011, Franks_2017, Rogers_2017, Lawrence_2019}. Many
 models will assume a constant VPD slope term (e.g. g$_1$) within
@@
 difficulty in the literature with quantifying an ecosystem scale slope
-parameter \citep{Medlyn_2017}. Given that something so fundamental as
+parameter \cite{Medlyn_2017}. Given that something so fundamental as
 the ET response to VPD varies strongly with g$_1$, we must invest more
@@
 carbon gain while minimizing water loss, which observations also
-generally support \citep{Lloyd_1991, MEDLYN_2011, Lin_2015, Zhou_2014,
+generally support \cite{Lloyd_1991, MEDLYN_2011, Lin_2015, Zhou_2014,
   Zhou_2015, Medlyn_2017}. However, some purely empirical results that
@@
 ecosystem-scale quantities based on stomata theory, may not always
-function optimally \citep{Zhou_2015, Lin_2018}. Specifically with
+function optimally \cite{Zhou_2015, Lin_2018}. Specifically with
 regards to uWUE, one would not expect that this ecosystem scale WUE
@@
 soil evaporation's contributions to ET should shift the exponent of
-the VPD dependence. \citet{Zhou_2015}'s results corroborate this: they
+the VPD dependence. \citeA{Zhou_2015}'s results corroborate this: they
 found a mean empirically fit exponential VPD dependence of 0.55,
 varying slightly from the theoretically optimal value of of 0.5 for
-AmeriFlux sites. \citet{Lin_2018}'s results also show variance in the
+AmeriFlux sites. \citeA{Lin_2018}'s results also show variance in the
 empirical exponent of the VPD dependence in the stomatal conductance
 model, but interpretation of this variance is more difficult as
-\citet{Lin_2018} do not handle the GPP dependence of stomatal
+\citeA{Lin_2018} do not handle the GPP dependence of stomatal
 conductance in a directly analogous manner to the optimal theory in
-\citet{MEDLYN_2011} and \citet{Medlyn_2017}. Regardless, given that
+\citeA{MEDLYN_2011} and \citeA{Medlyn_2017}. Regardless, given that
 these recent results on the relationship between VPD, GPP, and ET
-\citep{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017} form the
+\cite{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017} form the
 backbone of our analysis and are what allowed us to derive an explicit
@@
 used models utilize a VPD exponent other than the 1/2 suggested by
-optimal theory \citep[e.g. ][ which uses an exponent of
+optimal theory \cite<e.g. >[ which uses an exponent of
 1]{Leuning_1990}. It is useful to understand how the choice of model
@@
 models and empirical results have suggested exponents varying from 1/2
-\citep{Leuning_1990, Zhou_2015, Lin_2018}.
+\cite{Leuning_1990, Zhou_2015, Lin_2018}.
 
@@
 phloem and xylem. To accomplish this, stomata must remain slightly
-open \citep{De_2013, Nikinmaa_2013, Ryan_2014}. Furthermore, even if
+open \cite{De_2013, Nikinmaa_2013, Ryan_2014}. Furthermore, even if
 complete stomatal closure were possible, cuticular water loss and [at

### 2019-06-19 doc/paper/conclusions.tex
james submission compiles, need to proof read and encorporate with makefile

@@
 We derived a new form of Penman Monteith using the concept of
-semi-empirical optimal stomatal regulation \citep{Lin_2015,
-MEDLYN_2011} and near constant uWUE \citep{Zhou_2015} to remove the
+semi-empirical optimal stomatal regulation \cite{Lin_2015,
+MEDLYN_2011} and near constant uWUE \cite{Zhou_2015} to remove the
 implicit dependence of stomatal conductance on GPP and ET. With our
@@
 while decreasing temperature and increasing g$_a$ amplify the
-magnitude of the response. Previous literature \citep{Zhou_2015,
+magnitude of the response. Previous literature \cite{Zhou_2015,
 Lin_2015} suggests that crops, through association with higher g$_1$
@@
 ambiguity in applying leaf-scale estimations of g$_1$ to the ecosystem
-scale \citep{Medlyn_2017}.
+scale \cite{Medlyn_2017}.
 
@@
 which we currently believe is 1/2 for both uWUE and the stomatal
-conductance model \citep{MEDLYN_2011, Zhou_2014}. Indeed, this
+conductance model \cite{MEDLYN_2011, Zhou_2014}. Indeed, this
 sensitivity to the exponent of VPD dependence is an important result
@@
 exponent of VPD dependence in stomatal conductance
-\citep[e.g.,][]{Ball_1987, Leuning_1990, MEDLYN_2011}, and these
+\cite<e.g.,>[]{Ball_1987, Leuning_1990, MEDLYN_2011}, and these
 assumptions can fundamentally change the relationship between ET and
@@
 physiological responses to direct CO$_2$ effects
-\citep[e.g.,][]{Swann_2016, Lemordant_2018} receives more attention
+\cite<e.g.,>[]{Swann_2016, Lemordant_2018} receives more attention
 than physiological response to indirect effects like increased
-VPD \citep{Novick_2016}. Here, we provide a framework for understanding ET response to VPD
+VPD \cite{Novick_2016}. Here, we provide a framework for understanding ET response to VPD
 using a simplified model of two plant parameters. Feedbacks between the
@@
 climate. Any plant physiological heterogeneity or feedback that can be
-conceptualized with shifts in $g_1$ \citep[e.g.][]{Lin_2015,
-Medlyn_2017} and/or uWUE \citep[e.g.][]{Zhou_2014} are representable
+conceptualized with shifts in $g_1$ \cite<e.g.>[]{Lin_2015,
+Medlyn_2017} and/or uWUE \cite<e.g.>[]{Zhou_2014} are representable
 within our framework, which opens the door for a hierarchy of more

### 2019-06-19 doc/paper/intro.tex
james submission compiles, need to proof read and encorporate with makefile

@@
 depending on region, decreased relative humidity
-\citep{Byrne_2013}. Increases in VPD increase the atmospheric demand
-for evapotranspirated water \citep{Penman_1948, Monteith_1965}, but
+\cite{Byrne_2013}. Increases in VPD increase the atmospheric demand
+for evapotranspirated water \cite{Penman_1948, Monteith_1965}, but
 also reduce stomatal conductance through stomatal closure
-\citep{Rawson1977, Leuning_1990, Mott2007, Damour2010,
+\cite{Rawson1977, Leuning_1990, Mott2007, Damour2010,
 MEDLYN_2011}. Understanding the net evapotranspiration (ET) response
@@
 surface, and this drives an increase in ET
-\citep{Penman_1948}. However, plants' stomata have evolved to
+\cite{Penman_1948}. However, plants' stomata have evolved to
 optimally regulate the exchange of water and carbon, and tend to
 partially close in response to increased atmospheric dryness
-\citep{Farquhar_1978, Ball_1987, Leuning_1990, Katul_2009,
+\cite{Farquhar_1978, Ball_1987, Leuning_1990, Katul_2009,
 MEDLYN_2011}. This leads to a plant physiology perspective, in which
 an increase in VPD may actually correspond to a decrease in ET because
-of stomatal closure \citep[e.g.][]{Rigden_2017}.  In other words, the
+of stomatal closure \cite<e.g.>[]{Rigden_2017}.  In other words, the
 question ``When does VPD drive or reduce ET?'' can be related to
@@
 and atmospheric carbon can be accessed with a relatively smaller cost
-in water loss. If instead stomata were fully passive \citep [similar
-to soil pores, e.g. ][]{Or_2013}, increased atmospheric aridity would
-strongly reduce soil moisture \citep{Berg_2017}. This could further
+in water loss. If instead stomata were fully passive \cite<similar
+to soil pores, e.g. >[]{Or_2013}, increased atmospheric aridity would
+strongly reduce soil moisture \cite{Berg_2017}. This could further
 increase aridity as low soil moisture levels increases the Bowen
 ratio, leading to increased temperature and atmospheric drying
-\citep[][]{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006,
+\cite{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006,
 Salvucci_2013, Gentine_2016, Berg_2016, Zhou_2019}. Therefore, passive
@@
 were used to understand interactions between stomatal conductance,
-evapotranspiration and the environment \citep[e.g.,][]{Jarvis_1984,
+evapotranspiration and the environment \cite<e.g.,>[]{Jarvis_1984,
 Jarvis_1986, Mcnaughton_1991}. However these researchers did not
@@
 results drastically improving our understanding of VPD's impact on
-physiology, especially at the leaf level. \citet{MEDLYN_2011}
+physiology, especially at the leaf level. \citeA{MEDLYN_2011}
 developed a model for leaf-scale stomatal conductance ($g_s$),
 including VPD response, by combining an optimal photosynthesis theory
-\citep{Cowan_1977, Katul_2009, Katul_2010} with an empirical approach,
+\cite{Cowan_1977, Katul_2009, Katul_2010} with an empirical approach,
 and extended use of this model to the ecosystem scale in
-\citet{Medlyn_2017}. Additionally, \citet{Zhou_2014} demonstrated that
+\citeA{Medlyn_2017}. Additionally, \citeA{Zhou_2014} demonstrated that
 a quantity underlying water use efficiency $\left(uWUE = \frac{GPP\;
@@
 across space and time, within a plant functional type (PFT)
-\citep{Zhou_2015}.  While stomatal conductance parameterizations and
+\cite{Zhou_2015}.  While stomatal conductance parameterizations and
 uWUE greatly simplify complex plant physiological processes, they
  still capture ecosystem behavior for vegetated surfaces
-\citep{Medlyn_2017, Zhou_2014}, and are useful tools to transparently
+\cite{Medlyn_2017, Zhou_2014}, and are useful tools to transparently
 develop intuition for the behavior of complex land-atmosphere systems.
@@
 In this manuscript, we leverage uWUE and recent developments in
-stomatal conductance parameterizations \citep{MEDLYN_2011} with a
-Penman-Monteith framework \citep [hereafter PM,][]{Penman_1948,
+stomatal conductance parameterizations \cite{MEDLYN_2011} with a
+Penman-Monteith framework \cite<hereafter PM,>[]{Penman_1948,
 Monteith_1965} to derive the theoretical one-way response of ET to VPD
@@
 other variables as VPD is known to increase dramatically with future
-climate and limit ET more than soil moisture \citep{Novick_2016}, but
+climate and limit ET more than soil moisture \cite{Novick_2016}, but
 disentangling VPD effects from soil moisture effects has been
 difficult in previous research, given their co-variability
-\citep[][]{Lin_2018, Zhou_2019}.  We are
+\cite{Lin_2018, Zhou_2019}.  We are
 able to disentangle the effect of VPD on ET because we explicitly
@@
 conditions. This is relevant because previous research focusing on the
-leaf scale \citep{Rawson1977, Turner1984, Oren1999, Damour2010,
+leaf scale \cite{Rawson1977, Turner1984, Oren1999, Damour2010,
 Mott2013} does not consistently or analogously include the energetic
@@
 leaf-scale results agree that stomatal conductance decreases in
-response to VPD \citep{Oren1999, Damour2010}, which we expect to be
+response to VPD \cite{Oren1999, Damour2010}, which we expect to be
 true for an ecosystem as well. However leaf scale results also
 indicate that transpiration usually increases with increasing VPD in a
-concave downward shape \citep[e.g.,][]{Rawson1977, Turner1984,
+concave downward shape \cite<e.g.,>[]{Rawson1977, Turner1984,
 Mott2013}, which may not be true for ET at the ecosystem scale once

### 2019-06-19 doc/paper/agu-james-submission.tex
james submission compiles, need to proof read and encorporate with makefile
added key points

@@
 \usepackage{soul}
+
+%% added by adam
+\usepackage{amsmath}
+\usepackage{amssymb}
+\newcommand{\appropto}{\mathrel{\vcenter{
+      \offinterlineskip\halign{\hfil$##$\cr
+        \propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}
+
 \linenumbers
@@
 \begin{keypoints}
-\item enter point 1 here
-\item enter point 2 here
-\item enter point 3 here
+\item Using Penman-Monteith, a semi-empirical stomatal conductance
+  model, and underlying water use efficiency, we develop a simplified
+  analytical theory for when ecosystems will reduce or increase
+  evapotranspiration in response to vapor pressure deficit perturbations.
+\item The observed range of plant parameters result in a broad range
+  of ecosystem response to increases in vapor pressure deficit, from
+  strong reductions in evapotranspiration to strong increases in
+  evapotranspiration depending on ecosystem type and environment.
+\item The choice of stomatal conductance model fundamentally alters
+  the relationship between evapotranspiration and vapor pressure
+  deficit, highlighting the need for further research on stomatal
+  conductance and its representation in more sophisticated land
+  surface models.
 \end{keypoints}

### 2019-06-20 doc/paper/abstract.tex
read through once, just final read through and submit

@@
 Increasing vapor pressure deficit (VPD) increases atmospheric demand
-for water, and vapor pressure deficit is expected to rise with
-increasing greenhouse gases. While increased evapotranspiration (ET)
-in response to increased atmospheric demand seems intuitive, plants
-are capable of reducing ET in response to increased VPD by closing
-their stomata, in an effort to conserve water. Here we examine which
-effect dominates response to increasing VPD: atmospheric demand and
-increases in ET, or plant physiological response (stomata closure) and
-decreases in ET. We use Penman-Monteith, combined with semi-empirical
-optimal stomatal regulation theory and underlying water use
-efficiency, to develop a theoretical framework for understanding how
-ET responds to increases in VPD. The theoretical ET response to VPD
-over a range of reasonable environmental conditions and plant
-characteristics varies from a strong decrease in ET in response to
-increasing VPD (water conservative) to a strong increase in ET in
-response to VPD (water intensive), highlighting the diversity of plant
-water regulation strategies.
+for water. While increased evapotranspiration (ET) in response to
+increased atmospheric demand seems intuitive, plants are capable of
+reducing ET in response to increased VPD by closing their stomata, in
+an effort to conserve water. Here we examine which effect dominates
+response to increasing VPD: atmospheric demand and increases in ET, or
+plant response (stomata closure) and decreases in ET. We use
+Penman-Monteith, combined with semi-empirical optimal stomatal
+regulation theory and underlying water use efficiency, to develop a
+theoretical framework for understanding how ET responds to increases
+in VPD. The theoretical ET response to VPD over a range of reasonable
+environmental conditions and plant characteristics varies from a
+strong decrease in ET in response to increasing VPD (water
+conservative) to a strong increase in ET in response to VPD (water
+intensive), highlighting the diversity of plant water regulation
+strategies.
 

### 2019-06-20 doc/paper/agu-james-submission.tex
read through once, just final read through and submit

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
@@
   analytical theory for when ecosystems will reduce or increase
-  evapotranspiration in response to vapor pressure deficit perturbations.
-\item The observed range of plant parameters result in a broad range
-  of ecosystem response to increases in vapor pressure deficit, from
-  strong reductions in evapotranspiration to strong increases in
-  evapotranspiration depending on ecosystem type and environment.
+  evapotranspiration in response to increasing vapor pressure deficit.
+\item Variations in ecosystem characteristics and environmental
+  conditions result in a broad range of responses to increases in
+  vapor pressure deficit, from strong reductions to strong increases
+  in evapotranspiration.
 \item The choice of stomatal conductance model fundamentally alters

### 2019-06-20 doc/paper/body.tex
read through once, just final read through and submit

@@
   dependence in uWUE leverages the same assumptions used in
-  \cite{MEDLYN_2011} to derive the square-root VPD dependence of the
+  \citeA{MEDLYN_2011} to derive the square-root VPD dependence of the
   stomatal conductance model (Eq.  (\ref{medlyn})).  We can use uWUE
-  to remove the $GPP$ dependence from $g_s$ in a way that makes PM
+  to remove the $GPP$ dependence in $g_s$ in a way that makes PM
   analytically tractable:
@@
   factor and a photosynthesis model with a single parameter (uWUE). Both
-  the plant parameters reflect water conservation strategy.The
+  the plant parameters reflect water conservation strategy. The
   slope parameter is related to the willingness of stomata to trade
@@
 
-   Because Eq. (\ref{d_et}) is a partial derivative, it's utility is as a
-   conceptual model for the change in ET in response to VPD with all
-   other variables held fixed. This provides is a useful tool for
+  Because Eq. (\ref{d_et}) is a partial derivative, it's utility is as
+  a conceptual model for the change in ET in response to VPD with all
+  other variables held fixed. This provides a useful tool for
   identifying the effect of VPD on ET through atmospheric demand and
-   plant response, and allows a practitioner to disentangle
-   complicated feedbacks when many quantities co-vary. However users
-   of Eq. (\ref{d_et}) should take care that their interpretation
-   matches the assumptions inherent in a partial derivative. For
-   example, results will only be as valid as the assumption that g$_1$
-   and uWUE (and by extension $\lambda$) are fixed with respect to the
-   user's conceptual change in VPD. Care must also be taken with
-   possible indirect effects associated with a change in VPD: for
-   example, a change in ET induced by a change in VPD can cause a
-   change in surface temperature, which would drive a change in net
+  plant response, and allows a practitioner to disentangle complicated
+  feedbacks when many quantities co-vary. However users of
+  Eq. (\ref{d_et}) should take care that their interpretation matches
+  the assumptions inherent in a partial derivative. For example,
+  results will only be as valid as the assumption that g$_1$ and uWUE
+  (and by extension $\lambda$) are fixed with respect to the user's
+  conceptual change in VPD. Care must also be taken with possible
+  indirect effects associated with a change in VPD: for example, a
+  change in ET induced by a change in VPD can cause a change in
+  surface temperature, which would drive a change in net
   radiation. These types of indirect effects and feedbacks are not
@@
 generate the figures and tables, are publicly available at
-\url{https://github.com/massma/climate\_et}.
+\url{https://github.com/massma/climate_et}.
 %\hyperref[https://github.com/massma/climate\_et]{https://github.com/massma/climate_et}.
@@
   \centering
+  \footnotesize
   \begin{tabular}{l c c c c c}
@@
 increase and uWUE should decrease. Because high and low values of uWUE
-and g$_1$ were determined independently from each other in
-previous literature, plant types with anomalously high or low values
-for \textit{both} g$_1$ and uWUE should be relatively rare.
+and g$_1$ were determined independently from each other in previous
+literature, plant types with anomalously high or low values for
+\textit{both} g$_1$ and uWUE should be relatively rare. However, some
+plant types do exhibit co-occuring high values for g$_1$ and uWUE. For
+example, C3 crops have both the highest g$_1$ value in
+\citeA{Franks_2017}, and the highest uWUE value in \citeA{Zhou_2015}.
 
@@
     m$), whether the solution is concave up or concave down does not
-    depend on uWUE.
+    depend on $*WUE$.
   \item In general, increasing the exponent of VPD dependence for
-    either $uWUE*$ or $g_*$ increases the likelihood of a concave down
+    either $*WUE$ or $g_*$ increases the likelihood of a concave down
     result. Additionally, as the exponent of VPD dependence increases

### 2019-06-20 doc/paper/conclusions.tex
read through once, just final read through and submit

@@
 \cite<e.g.,>[]{Swann_2016, Lemordant_2018} receives more attention
-than physiological response to indirect effects like increased
-VPD \cite{Novick_2016}. Here, we provide a framework for understanding ET response to VPD
-using a simplified model of two plant parameters. Feedbacks between the
-land and the atmosphere may alter the net response to a long-timescale
-global VPD perturbation, but our focus on the one way plant response
-to a VPD perturbation in the atmospheric boundary layer is an
-important first step to disentangling such feedbacks, both in
-observations and model simulations of the present and future. By
-removing Penman Monteith's dependence on implicit relationships
-between GPP, VPD, and ET, we allow for explicit, accurate future
-analysis of plant-VPD feedbacks in the atmospheric boundary layer
-(Eq. (\ref{et})). Our approach can be extended to examine varying
-plant response to more nuanced consideration of plant type and
+than physiological response to indirect effects like increased VPD
+\cite{Novick_2016}. Here, we provide a framework for understanding ET
+response to VPD using a simplified model of two plant
+parameters. Feedbacks between the land and the atmosphere may alter
+the net response to a long-timescale global VPD perturbation, but our
+focus on the one way plant response to a VPD perturbation in the
+atmospheric boundary layer is an important first step to disentangling
+such feedbacks, both in observations and model simulations of the
+present and future. By removing Penman Monteith's dependence on
+implicit relationships between GPP, VPD, and ET, we allow for explicit
+future analysis of plant-VPD feedbacks in the atmospheric boundary
+layer (Eq. (\ref{et})). Our approach can be extended to examine
+varying plant response to more nuanced consideration of plant type and
 climate. Any plant physiological heterogeneity or feedback that can be

### 2019-09-03 doc/paper/abstract.tex
about to start revisions for JAMES
appears I forgot to make a final commit of what I originally submitted,
but here it is.

@@
 increased atmospheric demand seems intuitive, plants are capable of
-reducing ET in response to increased VPD by closing their stomata, in
-an effort to conserve water. Here we examine which effect dominates
-response to increasing VPD: atmospheric demand and increases in ET, or
-plant response (stomata closure) and decreases in ET. We use
-Penman-Monteith, combined with semi-empirical optimal stomatal
-regulation theory and underlying water use efficiency, to develop a
-theoretical framework for understanding how ET responds to increases
-in VPD. The theoretical ET response to VPD over a range of reasonable
-environmental conditions and plant characteristics varies from a
-strong decrease in ET in response to increasing VPD (water
-conservative) to a strong increase in ET in response to VPD (water
-intensive), highlighting the diversity of plant water regulation
-strategies.
+reducing ET in response to increased VPD by closing their stomata. We
+examine which effect dominates the response to increasing VPD:
+atmospheric demand and increases in ET, or plant response (stomata
+closure) and decreases in ET. We use Penman-Monteith, combined with
+semi-empirical optimal stomatal regulation theory and underlying water
+use efficiency, to develop a theoretical framework for assessing ET
+response to VPD. The theory suggests that depending on the environment
+and plant characteristics, ET response to increasing VPD can vary from
+strongly decreasing to increasing, highlighting the diversity of plant
+water regulation strategies.
 
@@
 understanding complex land atmosphere systems in terms of idealized
-scenarios in which evapotranspiration responds to VPD only. This
-response is otherwise challenging to assess in an environment where
-many processes co-evolve together.
+scenarios in which ET responds to VPD only. This response is otherwise
+challenging to assess in an environment where many processes co-evolve
+together.

### 2019-09-03 doc/paper/body.tex
about to start revisions for JAMES
appears I forgot to make a final commit of what I originally submitted,
but here it is.

@@
 responses with VPD (water intensive), and smaller g$_1$ values lead to
-a higher non-linear VPD dependence of the response. However care
-should be exercised when interpreting the range of ET responses,
-because g$_1$ and uWUE should generally be anti-correlated due to
-their dependencies on $\lambda$. As $\lambda$ increases, g$_1$ should
-increase and uWUE should decrease. Because high and low values of uWUE
-and g$_1$ were determined independently from each other in previous
-literature, plant types with anomalously high or low values for
-\textit{both} g$_1$ and uWUE should be relatively rare. However, some
-plant types do exhibit co-occuring high values for g$_1$ and uWUE. For
-example, C3 crops have both the highest g$_1$ value in
-\citeA{Franks_2017}, and the highest uWUE value in \citeA{Zhou_2015}.
+a larger VPD dependence of the response (i.e. ET response is a
+stronger function of VPD). However care should be exercised when
+interpreting the range of ET responses, because g$_1$ and uWUE should
+generally be anti-correlated due to their dependencies on
+$\lambda$. As $\lambda$ increases, g$_1$ should increase and uWUE
+should decrease. Because high and low values of uWUE and g$_1$ were
+determined independently from each other in previous literature, plant
+types with anomalously high or low values for \textit{both} g$_1$ and
+uWUE should be relatively rare. However, some plant types do exhibit
+co-occurring high values for g$_1$ and uWUE. For example, C3 crops
+have both the highest g$_1$ value in \citeA{Franks_2017}, and the
+highest uWUE value in \citeA{Zhou_2015}.
 
@@
 framed by our results on $\frac{\partial \; ET}{\partial \; VPD}$. In
-\citeA{Zhou_2015} most plant types have uWUE similar to our median
+\citeA{Zhou_2015} most plant types' uWUE is similar to our median
 value, under which the sign of the ET response depends on
@@
 
-\citeA{Medlyn_2017}'s results did not display the robust relationship
-between plant type and and plant parameter (g$_1$) that was
-demonstrated by \citeA{Zhou_2015} for uWUE. Given that there is a
-robust relationship between g$_1$ and plant type in \citeA{Lin_2015},
-and that \citeA{Zhou_2015}'s ecosystem-scale results rely on the same
-underlying theory as \citeA{Medlyn_2017}'s results, the ambiguous
-results in \citeA{Medlyn_2017} could be due to noise in ecosystem
-scale observations, the consequences of imposing a model structure in
-some of their g$_1$ estimations, or real ecosystem-scale within-plant
-type variability (e.g. multiple species impacting the relationship
-within the eddy-covariance footprint). These differences between
-leaf-scale behavior and ecosystem scale behavior highlight the
+\citeA{Medlyn_2017}'s ecosystem-scale results did not display the
+robust relationship between plant type and and plant parameter (g$_1$)
+that was demonstrated by \citeA{Zhou_2015} for uWUE. Given that there
+is a robust relationship between g$_1$ and plant type in
+\citeA{Lin_2015}'s leaf-scale results, and that \citeA{Zhou_2015}'s ecosystem-scale results
+rely on the same underlying theory as \citeA{Medlyn_2017}'s results,
+the ambiguous results in \citeA{Medlyn_2017} could be due to noise in
+ecosystem scale observations, the consequences of imposing a model
+structure in some of their g$_1$ estimations, or real ecosystem-scale
+within-plant type variability (e.g. multiple species impacting the
+relationship within the eddy-covariance footprint). These differences
+between leaf-scale behavior and ecosystem scale behavior highlight the
 importance of understanding how and why slope parameters show
@@
 is due to model and/or observational error, then we can use the
-relationships between g$_1$ and plant characteristics defined by \citeA{Lin_2015} to
-frame the ET response to VPD. This is subject
-to the strong caveat that we still do not fully understand g$_1$
-behavior at the ecosystem scale. Extrapolating \citeA{Lin_2015}'s
-results (e.g. Fig. 2) to the ecosystem scale would reveal the
-following relationships between g$_1$ and plant types:
+relationships between g$_1$ and plant characteristics defined by
+\citeA{Lin_2015} to frame the ET response to VPD. This is subject to
+the strong caveat that we still do not fully understand g$_1$ behavior
+at the ecosystem scale. Extrapolating \citeA{Lin_2015}'s results
+(e.g. Fig. 2) to the ecosystem scale would reveal the following
+relationships between g$_1$ and plant types:
 
@@
 imposed by our choice of stomatal conductance model. We now examine
-how the stomatal model choice, and specifically the exponent of the
-VPD dependence in the stomatal conductance model, can impact the
-general form of the ET-VPD relationship. There is a theoretical basis
-for the square root VPD dependence in both the stomatal conductance
-model and uWUE based on the assumption that stomata behave to maximize
-carbon gain while minimizing water loss, which observations also
-generally support \cite{Lloyd_1991, MEDLYN_2011, Lin_2015, Zhou_2014,
-  Zhou_2015, Medlyn_2017}. However, some purely empirical results that
-fit the exponent of the VPD dependence to data have shown that it may
-vary slightly from 1/2, suggesting that stomata, as well as
-ecosystem-scale quantities based on stomata theory, may not always
-function optimally \cite{Zhou_2015, Lin_2018}. Specifically with
-regards to uWUE, one would not expect that this ecosystem scale WUE
-quantity will respond to VPD exactly analogously to stomata. Direct
-soil evaporation's contributions to ET should shift the exponent of
-the VPD dependence. \citeA{Zhou_2015}'s results corroborate this: they
-found a mean empirically fit exponential VPD dependence of 0.55,
-varying slightly from the theoretically optimal value of of 0.5 for
-AmeriFlux sites. \citeA{Lin_2018}'s results also show variance in the
-empirical exponent of the VPD dependence in the stomatal conductance
-model, but interpretation of this variance is more difficult as
-\citeA{Lin_2018} do not handle the GPP dependence of stomatal
-conductance in a directly analogous manner to the optimal theory in
-\citeA{MEDLYN_2011} and \citeA{Medlyn_2017}. Regardless, given that
-these recent results on the relationship between VPD, GPP, and ET
-\cite{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017} form the
-backbone of our analysis and are what allowed us to derive an explicit
-ET expression for the first time (Eq. (\ref{et})), we will analyze if
-and how assumptions about the exponent of the VPD dependence impact
-the shape of the ET-VPD dependence. This analysis is also important to
-understand whether the choice of stomatal conductance model alters the
+how stomatal model choice, and specifically the exponent of the VPD
+dependence in the stomatal conductance model, can impact the general
+form of the ET-VPD relationship. There is a theoretical basis for the
+square root VPD dependence in both the stomatal conductance model and
+uWUE based on the assumption that stomata behave to maximize carbon
+gain while minimizing water loss, which observations also generally
+support \cite{Lloyd_1991, MEDLYN_2011, Lin_2015, Zhou_2014, Zhou_2015,
+  Medlyn_2017}. However, some purely empirical results that fit the
+exponent of the VPD dependence to data have shown that it may vary
+slightly from 1/2, suggesting that stomata, as well as ecosystem-scale
+quantities based on stomata theory, may not always function optimally
+\cite{Zhou_2015, Lin_2018}. Specifically with regards to uWUE, one
+would not expect that this ecosystem scale WUE quantity will respond
+to VPD exactly analogously to stomata. Direct soil evaporation's
+contributions to ET should shift the exponent of the VPD
+dependence. \citeA{Zhou_2015}'s results corroborate this: they found a
+mean empirically fit exponential VPD dependence of 0.55, varying
+slightly from the theoretically optimal value of of 0.5 for AmeriFlux
+sites. \citeA{Lin_2018}'s results also show variance in the empirical
+exponent of the VPD dependence in the stomatal conductance model, but
+interpretation of this variance is more difficult as \citeA{Lin_2018}
+do not handle the GPP dependence of stomatal conductance in a directly
+analogous manner to the optimal theory in \citeA{MEDLYN_2011} and
+\citeA{Medlyn_2017}. Regardless, given that these recent results on
+the relationship between VPD, GPP, and ET \cite{MEDLYN_2011,
+  Zhou_2014, Zhou_2015, Medlyn_2017} form the backbone of our analysis
+and are what allowed us to derive an explicit ET expression for the
+first time (Eq. (\ref{et})), we will analyze if and how assumptions
+about the exponent of the VPD dependence impact the shape of the
+ET-VPD dependence. This analysis is also important to understand
+whether the choice of stomatal conductance model alters the
 fundamental behavior of the ET-VPD relationships, as many commonly
@@
 
-While it is possible that in the future some other form of VPD
-dependence is derived, at present \cite{MEDLYN_2011} and
+While it is possible (and even likely) that in the future some other
+form of VPD dependence is derived, at present \cite{MEDLYN_2011} and
 \cite{Zhou_2014} established n=m=1/2 as the most likely candidate

### 2019-09-03 doc/paper/agu-james-original-submission.tex
seems easier to use latexdiff than agu's recommendation of trackchanges

--- 2019-06-11 doc/paper/agu-james-submission.tex
@@
 \begin{keypoints}
-\item enter point 1 here
-\item enter point 2 here
-\item enter point 3 here
+\item We derive a simplified analytical model for ecosystem-scale
+  evapotranspiration response to changes in vapor pressure deficit.
+\item Ecosystems exhibit a range of behavior, from reductions to
+  increases in evapotransipration, in response to increasing vapor
+  pressure deficit.
+\item The choice of stomatal conductance model fundamentally alters
+  the relationship between evapotranspiration and vapor pressure
+  deficit.
 \end{keypoints}

### 2019-09-03 doc/paper/james-supplement-original-submision.tex
seems easier to use latexdiff than agu's recommendation of trackchanges

--- 2019-03-01 doc/paper/hess-supplement.tex
@@
 The manuscript analyzes the partial derivative of ET with respect to
-VPD. This assumes that all other quantites remain fixed, including the
+VPD. This assumes that all other quantities remain fixed, including the
 plant parameters g$_1$ and uWUE (and by extension, $\lambda$). In
@@
 relationship with soil moisture so we do not include the figures in
-the body of the manuscript. But given that constant uWUE g$_1$
+the body of the manuscript. But given that constant uWUE and g$_1$
 assumptions can make our theory more useful to the reader we provide

### 2019-09-04 doc/paper/abstract.tex
started actual work on revisions, but note for latexdiff to work all edited content must be in same file

@@
 C3 plants more likely to exhibit a positive ET response than C4
-plants; and 3) due to plant type, with crops more likely to exhibit a
+plants; and 3) plant type, with crops more likely to exhibit a
 positive ET response, and shrubs and gymniosperm trees more likely to

### 2019-09-04 doc/paper/agu-james-original-submission.tex
started actual work on revisions, but note for latexdiff to work all edited content must be in same file

@@
 \begin{abstract}
-\input{abstract.tex}
+Increasing vapor pressure deficit (VPD) increases atmospheric demand
+for water. While increased evapotranspiration (ET) in response to
+increased atmospheric demand seems intuitive, plants are capable of
+reducing ET in response to increased VPD by closing their stomata. We
+examine which effect dominates the response to increasing VPD:
+atmospheric demand and increases in ET, or plant response (stomata
+closure) and decreases in ET. We use Penman-Monteith, combined with
+semi-empirical optimal stomatal regulation theory and underlying water
+use efficiency, to develop a theoretical framework for assessing ET
+response to VPD. The theory suggests that depending on the environment
+and plant characteristics, ET response to increasing VPD can vary from
+strongly decreasing to increasing, highlighting the diversity of plant
+water regulation strategies.
+
+The ET response varies due to: 1) climate, with tropical and temperate
+climates more likely to exhibit a positive ET response to increasing
+VPD than boreal and arctic climates; 2) photosynthesis strategy, with
+C3 plants more likely to exhibit a positive ET response than C4
+plants; and 3) due to plant type, with crops more likely to exhibit a
+positive ET response, and shrubs and gymniosperm trees more likely to
+exhibit a negative ET response. These results, derived from previous
+literature connecting plant parameters to plant and climate
+characteristics, highlight the utility of our simplified framework for
+understanding complex land atmosphere systems in terms of idealized
+scenarios in which ET responds to VPD only. This response is otherwise
+challenging to assess in an environment where many processes co-evolve
+together.
 \end{abstract}
@@
 
-\input{intro.tex}
+Vapor pressure deficit (VPD) is expected to rise over continents in
+the future due to the combination of increased temperature and,
+depending on region, decreased relative humidity
+\cite{Byrne_2013}. Increases in VPD increase the atmospheric demand
+for evapotranspirated water \cite{Penman_1948, Monteith_1965}, but
+also reduce stomatal conductance through stomatal closure
+\cite{Rawson1977, Leuning_1990, Mott2007, Damour2010,
+MEDLYN_2011}. Understanding the net evapotranspiration (ET) response
+to these two opposing effects of changes in VPD is crucial for
+assessing the impact of environmental VPD perturbations on the water
+cycle.
 
-\input{body.tex}
+The opposing effects of increased atmospheric demand and higher
+stomatal closure lead to two possible perspectives for how ET responds
+to shifts in VPD. The first, a hydrometeorological perspective, is
+that higher VPD increases atmospheric demand for water from the land
+surface, and this drives an increase in ET
+\cite{Penman_1948}. However, plants' stomata have evolved to
+optimally regulate the exchange of water and carbon, and tend to
+partially close in response to increased atmospheric dryness
+\cite{Farquhar_1978, Ball_1987, Leuning_1990, Katul_2009,
+MEDLYN_2011}. This leads to a plant physiology perspective, in which
+an increase in VPD may actually correspond to a decrease in ET because
+of stomatal closure \cite<e.g.>[]{Rigden_2017}.  In other words, the
+question ``When does VPD drive or reduce ET?'' can be related to
+whether plant regulation or atmospheric demand dominates the ET
+response.
+
+The ET response to changes in VPD alters water partitioning between
+the soil and atmosphere. If ecosystem plant response reduces ET with
+atmospheric drying then soil moisture will be better conserved. This
+represents a sensible evolutionary strategy to cope with aridity: save
+water for periods when atmospheric demand for water is relatively low,
+and atmospheric carbon can be accessed with a relatively smaller cost
+in water loss. If instead stomata were fully passive \cite<similar
+to soil pores, e.g. >[]{Or_2013}, increased atmospheric aridity would
+strongly reduce soil moisture \cite{Berg_2017}. This could further
+increase aridity as low soil moisture levels increases the Bowen
+ratio, leading to increased temperature and atmospheric drying
+\cite{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006,
+Salvucci_2013, Gentine_2016, Berg_2016, Zhou_2019}. Therefore, passive
+regulation and a lack of soil moisture conservation does not seem to
+be a sensible strategy for plants from an evolutionary
+standpoint. This simplified logic explains generally why plants
+evolved to respond to VPD, but also excludes many details and special
+cases (e.g. plant to plant interaction and highly specialized
+photosynthesis strategies like Crassulacean acid metabolism
+photosynthesis).
+
+We can use intuition about plant water conservation strategy to
+hypothesize about ET response to changes in VPD. Plants and ecosystems
+that evolved to conserve water, such as arid shrubs, should be more
+likely to reduce ET with increasing VPD, and plants that have evolved
+or have been engineered to prioritize carbon gain over water
+conservation, such as crops, will be more likely to increase ET with
+increasing VPD. Atmospheric conditions must matter as well. At the
+ecosystem scale, there are limits to plant water conservation
+strategies. As atmospheric demand for water (VPD) increases,
+ecosystems may begin to reach their water conservation limits and
+might not be able to entirely limit ET flux to the atmosphere. At this
+stage any further increase in VPD will most likely drive a (limited)
+increase in ET, because the increase in atmospheric demand for water
+overwhelms the limited plant response to conserve water.
+
+The objective of the present manuscript is to use reasonable
+approximations established in prior research as a tool to develop a
+framework for understanding plant responses to atmospheric drying and
+the the VPD dependence of ET while keeping other variables fixed. This
+framework will aid interpretation of observations, full complexity
+models, and facilitate the disentanglement of complex land-atmosphere
+feedbacks. In particular, our approach has applications for
+understanding climate change impacts, given expected increases in VPD
+with rising temperature.  In the past, similar simplified approaches
+were used to understand interactions between stomatal conductance,
+evapotranspiration and the environment \cite<e.g.,>[]{Jarvis_1984,
+Jarvis_1986, Mcnaughton_1991}. However these researchers did not
+explore explicitly the sensitivity of ET to VPD, including VPD's
+effect on stomatal conductance and plant function.
+
+Approaching the problem of ET response to VPD is aided by recent
+results drastically improving our understanding of VPD's impact on
+physiology, especially at the leaf level. \citeA{MEDLYN_2011}
+developed a model for leaf-scale stomatal conductance ($g_s$),
+including VPD response, by combining an optimal photosynthesis theory
+\cite{Cowan_1977, Katul_2009, Katul_2010} with an empirical approach,
+and extended use of this model to the ecosystem scale in
+\citeA{Medlyn_2017}. Additionally, \citeA{Zhou_2014} demonstrated that
+a quantity underlying water use efficiency $\left(uWUE = \frac{GPP\;
+\sqrt{VPD}}{ET}\right)$ properly captures a constant relationship
+between GPP, ET, and VPD over a diurnal cycle at the ecosystem
+scale. uWUE is also relatively well conserved in the growing season
+across space and time, within a plant functional type (PFT)
+\cite{Zhou_2015}.  While stomatal conductance parameterizations and
+uWUE greatly simplify complex plant physiological processes, they
+ still capture ecosystem behavior for vegetated surfaces
+\cite{Medlyn_2017, Zhou_2014}, and are useful tools to transparently
+develop intuition for the behavior of complex land-atmosphere systems.
+
+In this manuscript, we leverage uWUE and recent developments in
+stomatal conductance parameterizations \cite{MEDLYN_2011} with a
+Penman-Monteith framework \cite<hereafter PM,>[]{Penman_1948,
+Monteith_1965} to derive the theoretical one-way response of ET to VPD
+with other environmental variables properly controlled for, i.e. we
+develop a framework for evaluating the partial derivative of ET with
+respect to VPD. It is useful to disconnect the impact of VPD from
+other variables as VPD is known to increase dramatically with future
+climate and limit ET more than soil moisture \cite{Novick_2016}, but
+disentangling VPD effects from soil moisture effects has been
+difficult in previous research, given their co-variability
+\cite{Lin_2018, Zhou_2019}.  We are
+able to disentangle the effect of VPD on ET because we explicitly
+include VPD's full effect on stomatal conductance, including its
+impact on photosynthesis. The use of an energy balance framework (PM)
+allows us to include in our analysis the effects of the energy cost of
+evaporating water from a surface, which is an important factor in the
+natural environment, compared to prescribed in situ environmental
+conditions. This is relevant because previous research focusing on the
+leaf scale \cite{Rawson1977, Turner1984, Oren1999, Damour2010,
+Mott2013} does not consistently or analogously include the energetic
+constraints on evapotranspiration under which ecosystems operate. The
+leaf-scale results agree that stomatal conductance decreases in
+response to VPD \cite{Oren1999, Damour2010}, which we expect to be
+true for an ecosystem as well. However leaf scale results also
+indicate that transpiration usually increases with increasing VPD in a
+concave downward shape \cite<e.g.,>[]{Rawson1977, Turner1984,
+Mott2013}, which may not be true for ET at the ecosystem scale once
+energetic constraints on ET are included in the analysis.  Our
+approach allows us to estimate the expected ET response to VPD at the
+ecosystem scale, including the effects of surface energy balance
+constraints, and assess how ecosystem ET response to VPD deviates from
+previous leaf scale analyses.
+
+This manuscript presents the range of possible ecosystem-scale ET
+responses to VPD, given parameters previously established in peer
+reviewed literature. Additionally, we explore the sensitivity of the
+ET-VPD relationship to stomatal model and framework choice,
+highlighting the importance of: 1) future research on stomatal
+conductance and ecosystem scale modeling, and 2) thoughtful selection
+of photosynthesis and stomatal conductance models in more
+sophisticated land surface and earth system models.
+
+\section{Methods}
+\label{methods}
+The Penman-Monteith equation \cite<hereafter PM,>[]{Penman_1948,
+  Monteith_1965} estimates ET as a function of observable atmospheric
+variables and surface conductances:
+% \begin{linenomath*}
+  \begin{equation}
+    \label{orig_pen}
+    ET = \frac{\Delta R_{net} + g_a \rho_a c_p VPD}{\Delta + \gamma(1 + \frac{g_a}{g_s})},
+  \end{equation}
+% \end{linenomath*}
+  where $\Delta$ is the change in saturation vapor pressure with
+  temperature, given by Clausius-Clapeyron
+  ($\frac{d \; e_s}{d \; T}$), $R_{net}$ is the net radiation minus
+  ground heat flux, $g_a$ is aerodynamic conductance, $\rho_a$ is air
+  density, $c_p$ is specific heat of air at constant pressure,
+  $\gamma$ is the psychometric constant, and $g_s$ is the stomatal
+  conductance (Table \ref{definitions}). The issue with PM is that
+  $g_s$ is a function of carbon uptake, which is has a strong
+  functionally relation to ET through stomatal function. Therefore,
+  when PM is formulated in terms of $g_s$, it is an implicit function
+  of ET itself rather than an explicit function of ET. Here we will
+  derive a new form of PM in which ET is an explicit function of plant
+  parameters and environmental conditions, and use it to assess the
+  ecosystem scale response to VPD (by taking a partial derivative).
+
+\begin{table}
+  \caption{Definition of symbols and variables, with citation for how
+    values are calculated, if applicable.}
+  \label{definitions}
+  \centering \footnotesize
+  \begin{tabular}{l c c c}
+    \hline
+    Variable & Description & Units & Citation \\
+    \hline
+    $e_s$  & saturation vapor pressure & Pa  & - \\
+    $T$  & temperature  & K & - \\
+    $P$  & pressure & Pa  & - \\
+    $\Delta$  & $\frac{\partial e_s}{\partial T}$ & Pa K$^{-1}$ & - \\
+    $R_{net}$  & net radiation at land surface minus ground heat flux & W m$^{-2}$   & - \\
+    $g_a$  & aerodynamic conductance & m s$^{-1}$  & \citeA{Shuttleworth_2012} \\
+    $\rho_a$  & air density & kg m$^{-3}$  & - \\
+    $c_p$  & specific heat capacity of air at constant pressure & J K$^{-1}$ kg$^{-1}$ & - \\
+    $VPD$  & vapor pressure deficit & Pa  & - \\
+    $\gamma$  & psychometric constant & Pa K$^{-1}$   & - \\
+    $g_{s}$  &  stomatal conductance & m s$^{-1}$
+                                   & \citeA{Medlyn_2017} \\
+    $g_{1}$  & ecosystem-scale slope parameter & Pa$^{0.5}$ & \citeA{Medlyn_2017} \\
+    $R$ & universal gas constant & J mol$^{-1}$ K$^{-1}$ & - \\
+    $R_{air}$ & gas constant of air & J  K$^{-1}$ kg$^{-1}$ & - \\
+    $\sigma$ & uncertainty parameter & -& - \\
+    $c_a$ & CO$_2$ concentration & $\mu$ mol CO$_2$ mol$^{-1}$ air& - \\
+    $\lambda = \frac{\partial \; transpiration}{\partial\; CO_2\; assimilation}$ & marginal water cost of leaf carbon & mol H$_2$O mol$^{-1}$ CO$_2$ & - \\
+    $\Gamma$ & CO$_2$ compensation point & - & - \\
+    $\Gamma^*$ & CO$_2$ compensation point without dark respiration &
+                                                                      - & - \\
+    GPP & gross primary production & $\mu$-mol C s$^{-1}$ m$^{-2}$ & -
+    \\
+    ET & evapotranspiration & W m$^{-2}$ & - \\
+   uWUE & underlying water use efficiency & $\mu$-mol C Pa$^{0.5}$
+                                            J$^{-1}$ ET  &
+                                                           \citeA{Zhou_2015} \\
+    \hline
+  \end{tabular}
+\end{table}
+
+\citeA{MEDLYN_2011} developed a model for stomatal conductance ($g_s$)
+by combining an optimal photosynthesis theory \cite{Cowan_1977} with
+an empirical approach, which describes the dependence of $g_s$ to
+VPD. They also extended this model to the ecosystem scale
+\cite{Medlyn_2017}:
+
+% \begin{linenomath*}
+  \begin{equation}
+    g_s = \frac{R \,T}{P} \; 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{GPP}{c_a},
+    \label{medlyn}
+  \end{equation}
+% \end{linenomath*}
+
+  where $g_{1}$ is a VPD ``slope'' parameter, GPP is the ecosystem
+  scale gross primary production, and $c_a$ is the atmospheric CO$_2$
+  concentration at the canopy. \cite{MEDLYN_2011} relate the slope
+  parameter ($g_{1}$) to physical parameters as:
+% \begin{linenomath*}
+
+  \begin{equation}
+    g_{1}  \propto \sqrt{\frac{3 \, \Gamma^* \,
+        \lambda}{1.6}},\footnote{Note this expression has units of of
+      (mmol mol$^{-1}$)$^{1/2}$, but this can be converted to
+      Pa$^{1/2}$ using the ideal gas law.}
+    \label{slope}
+  \end{equation}
+% \end{linenomath*}
+
+where $\Gamma^*$ is the CO$_2$ compensation point for photosynthesis
+(without dark respiration), and $\lambda$ is the marginal water cost
+of leaf carbon
+($\frac{\partial \; \text{transpiration}}{\partial \; CO_2 assimilation}$)
+\cite{Farquhar_1980, Katul_2009}. So,
+$g_{1}$ is a leaf-scale term reflecting the trade-off of water for
+carbon uptake. The higher $g_{1}$, the more open the stomata and
+the more they release water in exchange for carbon.
+
+While Eq. (\ref{medlyn}) can be used in PM (Eq. (1)), it will make
+analytical work with the function intractable because GPP is
+functionally related to ET itself. Additionally, a perturbation to VPD
+should induce a physiological plant response that will alter GPP and
+cause an indirect change in stomatal conductance, in addition to the
+direct effect of VPD in Eq. (\ref{medlyn}). Therefore, in order to
+derive the full plant response of ET to VPD, we must account for the
+functional relationship between GPP, ET, and VPD, and its effect on
+stomatal conductance. We can use aforementioned semi-empirical results
+of \citeA{Zhou_2015}, which were inspired by optimal photosynthesis
+theory, as a tool to approach this problem. \citeA{Zhou_2015}, showed
+that underlying Water Use Efficiency (uWUE):
+
+% \begin{linenomath*}
+  \begin{equation}
+    uWUE = \frac{GPP \cdot \sqrt{VPD}}{ET}
+    \label{uwue}
+  \end{equation}
+% \end{linenomath*}
+
+  is relatively constant across time within a plant functional type,
+  and correctly captures a constant relationship between GPP, ET and
+  VPD over a diurnal cycle during the growing season
+  \cite{Zhou_2014}. The theoretical derivation of the square root VPD
+  dependence in uWUE leverages the same assumptions used in
+  \citeA{MEDLYN_2011} to derive the square-root VPD dependence of the
+  stomatal conductance model (Eq.  (\ref{medlyn})).  We can use uWUE
+  to remove the $GPP$ dependence in $g_s$ in a way that makes PM
+  analytically tractable:
+
+% \begin{linenomath*}
+  \begin{equation}
+    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{uWUE \; ET}{c_a \; \sqrt{VPD}}.
+    \label{new_g_s}
+  \end{equation}
+% \end{linenomath*}
+
+Plugging Eq. (\ref{new_g_s}) into Eq. (\ref{orig_pen}) and
+rearranging gives a new explicit expression for PM, in which
+dependence on $GPP$ is removed:
+
+% \begin{linenomath*}
+  \begin{equation}
+    ET = \frac{\Delta R_{net} + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R \; 1.6\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
+    \label{et}
+  \end{equation}
+% \end{linenomath*}
+
+  By accounting for photosynthesis changes in ecosystem conductance,
+  with Eq. (\ref{et}) we have used recent results
+  \cite{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017} to derive
+  ET explicitly as function of environmental variables and two
+  plant-specific constants, the slope parameter ($g_1$), and uWUE. For
+  the first time we have removed the implicit dependence of ET on
+  itself through the stomatal conductance term, and we have also
+  replaced the added complexity of a stomatal conductance reduction
+  factor and a photosynthesis model with a single parameter (uWUE). Both
+  the plant parameters reflect water conservation strategy. The
+  slope parameter is related to the willingness of stomata to trade
+  water for CO$_2$ and to keep stomata open (carbon cost in terms of
+  water)). uWUE is a semi-empirical ecosystem-scale constant related
+  to how WUE changes with VPD (specifically $VPD^{-1/2}$). It is also
+  roughly proportional to physical constants:
+
+\[uWUE \appropto \sqrt{\frac{c_a - \Gamma}{1.6 \lambda}},\]
+
+where $\Gamma$ is the CO$_2$ compensation point \cite<Eq. (5)
+in>[]{Zhou_2014}. So uWUE is related to atmospheric CO$_2$
+concentration and compensation point, and is inversely proportional to
+the marginal water cost of leaf carbon ($\lambda$). This relationship
+with $\lambda$ ($\propto \lambda^{-1/2}$) is important as it is the inverse of g$_1$'s
+relationship with $\lambda$ ($\propto \sqrt{\lambda}$). So, uWUE
+should increase as $\lambda$ decreases, and g$_1$ should decrease as
+$\lambda$ decreases.
+
+With Eq. (\ref{et}) we can take the partial derivative of ET with respect
+to VPD to understand whether VPD drives or reduces ecosystem-scale ET:
+
+% \begin{linenomath*}
+  \begin{equation}
+    \frac{\partial \;  ET}{\partial \; VPD} = \frac{2\; g_a \;
+      P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} -
+      \frac{\gamma c_a }{1.6 \; R\; \; \text{ uWUE }} \left(
+        \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right)
+    \right),
+    \label{d_et}
+  \end{equation}
+% \end{linenomath*}
+
+  providing analytical framework for ecosystem response to
+  atmospheric demand with environmental conditions held
+  fixed. There are a few subtleties to taking the derivative in
+  Eq. (\ref{d_et}): $\Delta$ ($\frac{d e_{s}}{d T}$) and $VPD$ are
+  functionally related, so while taking the derivative we evaluate
+  $\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET}
+  {\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD}
+  \Big|_{\text{RH fixed}} + \frac{\partial \; ET}{\partial \; RH}
+  \frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$
+      fixed}}$. $RH$ and $e_s$ are assumed to be approximately
+  independent, which is supported by data (not shown).
+
+  This derivation relied either implicitly or explicitly on several
+  assumptions. First, we assume that VPD at the leaf surface is the
+  same as VPD at measurement height; physically this implies that
+  leaves are perfectly coupled to the atmosphere. In reality, for some
+  conditions and plant types the leaves can become decoupled from the
+  boundary layer \cite{De_2017, Medlyn_2017}. Therefore, our
+  derivation will be most applicable in times like the growing season
+  (when we also expect uWUE to be most valid), when relatively high
+  insolation induces instability and convective boundary layers, and
+  we would expect the surface to be generally well coupled. An
+  additional assumption in the formulation of uWUE \cite{Zhou_2014,
+    Zhou_2015} and \citeA{Medlyn_2017}'s stomatal conductance model is
+  that direct soil evaporation contributions to ET remain small
+  relative to transpiration. Again, this should be more true during
+  the growing season. The ratio of evaporation to transpiration may
+  increase immediately after rainfall events due to high soil moisture
+  and ponding, but VPD is generally low anyways during these
+  times. However, some plant types allow for systematically larger
+  contributions of evaporation in ET, particularly those with sparse
+  canopies and smaller relative amounts of transpiration. We therefore
+  might expect that the theory will be most applicable to forest PFTs,
+  which will be most strongly coupled to the boundary layer due to
+  larger surface roughness, and will also generally have the highest
+  ratios of transpiration to evaporation. Finally, we assume that
+  $g_1$ and uWUE are constant with respect to the conceptual VPD
+  perturbation. Both quantities have been shown to be relatively
+  constant with respect to changes in VPD \cite{Franks_2017,
+    Zhou_2014}. These parameters will however vary with plant species
+  and characteristics \cite<e.g. wood density, >[]{Lin_2015}, as well
+  as environmental conditions including soil water content and
+  temperature \cite{Lin_2015, Manzoni2013}. Exploring possible soil
+  moisture (in)dependence of the plant parameters (g$_1$ and uWUE) is
+  particularly interesting because soil moisture only enters the
+  partial derivative directly through these plant parameters. If the
+  plant parameters are weak functions of soil moisture then the theory
+  can be directly applied to a broader range of conceptual VPD
+  scenarios, including observed compound events between high VPD and
+  low soil moisture \cite{Zhou_2019}. Supplementary material for this
+  manuscript explores the soil moisture dependence of the plant
+  parameters, but excessive noise and inconsistencies preclude
+  conclusions about the nature of any dependence. We provide it in
+  case it is of interest to the reader, and to motivate future
+  research.
+
+  Because Eq. (\ref{d_et}) is a partial derivative, it's utility is as
+  a conceptual model for the change in ET in response to VPD with all
+  other variables held fixed. This provides a useful tool for
+  identifying the effect of VPD on ET through atmospheric demand and
+  plant response, and allows a practitioner to disentangle complicated
+  feedbacks when many quantities co-vary. However users of
+  Eq. (\ref{d_et}) should take care that their interpretation matches
+  the assumptions inherent in a partial derivative. For example,
+  results will only be as valid as the assumption that g$_1$ and uWUE
+  (and by extension $\lambda$) are fixed with respect to the user's
+  conceptual change in VPD. Care must also be taken with possible
+  indirect effects associated with a change in VPD: for example, a
+  change in ET induced by a change in VPD can cause a change in
+  surface temperature, which would drive a change in net
+  radiation. These types of indirect effects and feedbacks are not
+  considered in Eq. (\ref{d_et}): temperature (a variable) is
+  mathematically fixed.
+
+\subsection{Framing Eq. \ref{d_et} with previous research }
+
+  To account for both the variability in plant parameters and the
+  environment, we will systematically analyze how the ET response to
+  VPD (Eq. (\ref{d_et})) varies. Eq. (\ref{d_et}) includes a
+  ``sign'' term that determines the sign of the response in addition
+  to magnitude:
+
+\begin{equation}
+  \label{sign}
+  \frac{c_p}{R_{air}} - \frac{\gamma c_a }{1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right),
+\end{equation}
+
+and a ``scaling'' term multiplying the ``sign'' term:
+
+\begin{equation}
+  \frac{g_a \; P}{T(\Delta + \gamma)}.
+\end{equation}
+
+In the ``sign term'' most of the quantities are relatively constant,
+except for VPD and the plant parameters g$_1$ and uWUE. In the scaling
+term, most of the terms are relatively constant with the exception of
+aerodynamic conductance ($g_a$) and temperature (especially its effect
+through $\Delta$). To determine the range of probable ET responses to
+VPD we will systematically vary these parameters according to Table
+\ref{param_varying}, while all other parameters are held fixed (Table
+\ref{param_fixed}). Physical variables ($g_a$, $T$) and plant
+physiological parameters (g$_1$, uWUE) are varied according to
+literature-based expectations for a range of growing season conditions
+and plant types \cite{Zhou_2015, Medlyn_2017}. Using this previous
+literature we can connect the effect of varying plant parameters to
+specific plant types and characteristics.
+
+All code and data used in this analysis, including those used to
+generate the figures and tables, are publicly available at
+\url{https://github.com/massma/climate_et}.
+%\hyperref[https://github.com/massma/climate\_et]{https://github.com/massma/climate_et}.
+
+\begin{table}
+  \caption{Variable quantities in the ET response to VPD. Each value
+    is varied to determine the effect of a range of expected plant and
+    environmental conditions on ET response to VPD. A citation is
+    provided in cases where the quantities are directly derived from
+    previous literature. Conceptually, min. values are extracted from
+    literature to correspond to approximately the 15th percentile of
+    observed conditions during the growing season, med. values
+    correspond to approximately the 50th percentile, and max. values
+    correspond to approximately the 85th percentile.}
+  \label{param_varying}
+  \centering
+  \footnotesize
+  \begin{tabular}{l c c c c c}
+    \hline
+    Symbol & units [units in citation] & min  & med & max
+    & citation  \\
+    \hline
+    \input{param_varying.tex}
+    \hline
+  \end{tabular}
+\end{table}
+
+\begin{table}
+  \caption{Quantities that are fixed in the ET response to VPD
+    (relative to those in Table \ref{param_varying}).}
+  \label{param_fixed}
+  \centering
+  \begin{tabular}{l c c}
+    \hline
+    Symbol & units & value \\
+    \hline
+    \input{param_fixed.tex}
+    \hline
+  \end{tabular}
+\end{table}
+
+\section{Results}
+\label{results}
+
+By varying four parameters (g$_a$, T, uWUE, g$_1$) at three different
+values (Table \ref{param_varying}) we generate nine different values
+for the scaling term, nine different curves (as a function of VPD) for
+the sign term, and 81 different curves for the ET response to VPD
+($\frac{\partial \; ET}{\partial \; VPD}$). To aid visualization we
+can examine a subset of nine of these 81 curves, defined by the
+minimum, median and maximum values for both the scaling and sign term
+(Figure \ref{full}). The range of ET responses to VPD vary from those
+where ET almost always decreases with increasing VPD (water
+conservative), to those where ET almost always increases with
+increasing VPD (water intensive). Additionally, for some parameters
+whether ET will increase or decrease with increasing VPD depends on
+atmospheric demand (Figure \ref{full}).
+
+\begin{figure}
+  \centering \includegraphics{./fully_idealized.pdf}
+  \caption{The functional form of $\frac{\partial \; ET}{\partial
+      \; VPD}$ for minimum, median and maximum values of both the sign
+    term and the scaling term.}
+  \label{full}
+\end{figure}
+
+\begin{figure}
+  \centering \includegraphics{./fully_idealized_sign.pdf}
+  \caption{The functional form of $\frac{\partial \; ET}{\partial
+      \; VPD}$ evaluated at the median value of the scaling term, for
+    varying values of g$_1$ and uWUE as given in Table \ref{param_varying}.}
+  \label{sign}
+\end{figure}
+
+Examining the sign term independent of the scaling term illuminates
+the role of the two plant parameters, g$_1$ and uWUE, in determining
+the degree to which the response is water conservative
+($\frac{\partial \; ET}{\partial \; VPD} < 0$) or water intensive
+($\frac{\partial \; ET}{\partial \; VPD} > 0$) (Figure \ref{sign})
+. Higher g$_1$ and uWUE shift the curve towards increasing ET
+responses with VPD (water intensive), and smaller g$_1$ values lead to
+a larger VPD dependence of the response (i.e. ET response is a
+stronger function of VPD). However care should be exercised when
+interpreting the range of ET responses, because g$_1$ and uWUE should
+generally be anti-correlated due to their dependencies on
+$\lambda$. As $\lambda$ increases, g$_1$ should increase and uWUE
+should decrease. Because high and low values of uWUE and g$_1$ were
+determined independently from each other in previous literature, plant
+types with anomalously high or low values for \textit{both} g$_1$ and
+uWUE should be relatively rare. However, some plant types do exhibit
+co-occurring high values for g$_1$ and uWUE. For example, C3 crops
+have both the highest g$_1$ value in \citeA{Franks_2017}, and the
+highest uWUE value in \citeA{Zhou_2015}.
+
+\begin{figure}
+  \centering \includegraphics{./fully_idealized_scaling.pdf}
+  \caption{The functional form of $\frac{\partial \; ET}{\partial
+      \; VPD}$ evaluated at the median value of the sign term, for
+    varying values of g$_a$ and T as given in Table \ref{param_varying}.}
+  \label{scaling}
+\end{figure}
+
+Examining the scaling term independent of the sign term shows how
+aerodynamic conductance (communication between the atmosphere and the
+surface) and temperature (controls the efficiency of energy conversion
+to latent heat through $\Delta$), amplify or suppress the plant
+response represented in the sign term (Figure \ref{scaling}). Both
+lower temperatures and higher aerodynamic conductance lead to
+amplified ET response to VPD, with the variability of aerodynamic
+conductance resulting in a slightly higher variability of ET response
+relative to temperature.
+
+\section{Discussion}
+\label{discussion}
+
+Interpreting both the derivation and results of this manuscript hinges
+on an understanding and appreciation for the usefulness of partial
+derivatives for understanding the behavior of complex systems in a
+simplified framework. Our results are a conceptual tool for answering
+the question, ``What is the ET response to VPD with all other
+quantities held fixed?'' This is useful, and we would argue is
+critical, for understanding complex responses when many quantities are
+varying simultaneously \cite{Zhou_2019}. If we cannot understand the
+contribution of one term to observed variability, independent of other
+confounding factors, there is little hope of disentangling and
+understanding the relative role of many co-varying quantities in
+determining observed variability.  Eq. (\ref{d_et}) explicitly
+provides an estimate of the ET response to VPD, given assumptions
+about ecosystem plant characteristics through the parameters g$_1$ and
+uWUE. Eq. (\ref{d_et}) is generic to external environmental factors
+and the timescale over which a VPD response acts, and has a
+straightforward conceptual interpretation. It provides the ET
+response, given other quantities and parameters held fixed, subject to
+the assumptions outlined in Sect. \ref{methods}. So, while the plant
+parameters g$_1$ and uWUE may vary with plant species and
+environmental conditions like soil moisture, we can still assess the
+ET response to VPD for a given ecosystem state, or for a given soil
+moisture condition.
+
+However, given that g$_1$ and uWUE can vary, and they both have a
+large impact on the ET response (Figure \ref{full}, \ref{sign}), it is
+useful to examine how these quantities vary in previous literature,
+framed by our results on $\frac{\partial \; ET}{\partial \; VPD}$. In
+\citeA{Zhou_2015} most plant types' uWUE is similar to our median
+value, under which the sign of the ET response depends on
+VPD. However, crops, which we expect to prioritize carbon uptake over
+water conservation, have a higher uWUE value closer to our maximum
+value. This higher uWUE results in a higher likelihood for an
+increasing ET response to VPD, which matches intuition given that we
+expect crops to keep stomata open for access to carbon, at the cost of
+increased water loss during high VPD. Shrubs, and to a lesser extent
+grass, have an uWUE closer to the minimum uWUE. For these plant types,
+we would then expect a decrease of ET in response to VPD. It is
+important to note that, while \citeA{Zhou_2015} did not examine the
+role of soil moisture for within plant-type variability of uWUE, we
+might expect some variation, especially in extreme cases when soil
+water becomes a limiting factor (see supplementary material).
+
+\citeA{Medlyn_2017}'s ecosystem-scale results did not display the
+robust relationship between plant type and and plant parameter (g$_1$)
+that was demonstrated by \citeA{Zhou_2015} for uWUE. Given that there
+is a robust relationship between g$_1$ and plant type in
+\citeA{Lin_2015}'s leaf-scale results, and that \citeA{Zhou_2015}'s ecosystem-scale results
+rely on the same underlying theory as \citeA{Medlyn_2017}'s results,
+the ambiguous results in \citeA{Medlyn_2017} could be due to noise in
+ecosystem scale observations, the consequences of imposing a model
+structure in some of their g$_1$ estimations, or real ecosystem-scale
+within-plant type variability (e.g. multiple species impacting the
+relationship within the eddy-covariance footprint). These differences
+between leaf-scale behavior and ecosystem scale behavior highlight the
+importance of understanding how and why slope parameters show
+increased variability at the ecosystem scale, as argued by
+\citeA{Medlyn_2017}. However, if we assume that there is some real
+analogy between leaf scale results in \citeA{Lin_2015} and ecosystem
+scale behavior, and that some of the ambiguity in \citeA{Medlyn_2017}
+is due to model and/or observational error, then we can use the
+relationships between g$_1$ and plant characteristics defined by
+\citeA{Lin_2015} to frame the ET response to VPD. This is subject to
+the strong caveat that we still do not fully understand g$_1$ behavior
+at the ecosystem scale. Extrapolating \citeA{Lin_2015}'s results
+(e.g. Fig. 2) to the ecosystem scale would reveal the following
+relationships between g$_1$ and plant types:
+
+\begin{itemize}
+  \item C3 plants would have a generally higher g$_1$ than C4 plants,
+  \item Crops would have a larger g$_1$ than shrubs, grass, and
+    angiosperm trees, which have a higher g$_1$ than gymnosperm
+    trees.
+  \item Tropical and temperate climates would generally be
+    characterized by plants with a higher g$_1$ than arctic and boreal
+    climates.
+\end{itemize}
+For all of the above relationships, a higher g$_1$ means a higher
+likelihood of a positive ET response to increasing VPD (water
+intensive), and a smaller VPD dependence of the ET response.
+
+Our theoretical results highlight the variability of ET
+response to VPD as a function of both climate  and plant
+characteristics, especially water usage strategy. Whether an
+ecosystem increases or decreases ET in response to VPD will vary
+depending on vegetation and climate. Generalizations about ET response
+to VPD therefore require thoughtful consideration of ecosystem
+characteristics; statements such as ``ET increases with warming due to
+increases in VPD'' may be false depending on the water conservation
+strategy of a given ecosystem \cite{Lemordant_2018}.
+
+Additionally, the sensitivity of ET response to plant parameters
+highlights the importance of understanding and developing stomatal
+conductance models and parameters for the ecosystem scale. In
+particular, why and how g$_1$ behavior at the leaf scale is not
+analogous to the ecosystem scale must be understood. While the
+derivation here was explicit about the assumption of a constant g$_1$
+term with respect to the VPD perturbation, many sophisticated land
+surface models and earth system models employing stomatal conductance
+models make similar assumptions about the stationarity of VPD slope
+terms \cite{Niu_2011, Franks_2017, Rogers_2017, Lawrence_2019}. Many
+models will assume a constant VPD slope term (e.g. g$_1$) within
+a given plant functional type, and research derived from these models
+often does not acknowledge limitations of this assumption, or the
+difficulty in the literature with quantifying an ecosystem scale slope
+parameter \cite{Medlyn_2017}. Given that something so fundamental as
+the ET response to VPD varies strongly with g$_1$, we must invest more
+in understanding g$_1$'s behavior at the ecosystem scale, and the
+efficacy of models using constant slope terms for ecosystem-scale
+fluxes. Fortunately, our framework, and the new approach of using
+robust ecosystem-scale ratios (uWUE) to remove the implicit dependence
+of stomatal conductance on ET, is flexible enough to be applied to any
+stomatal conductance model that contains a dependence on GPP.
+
+So far we have discussed the role of different parameters in
+modulating the ET response. However, the form of the ET response is
+imposed by our choice of stomatal conductance model. We now examine
+how stomatal model choice, and specifically the exponent of the VPD
+dependence in the stomatal conductance model, can impact the general
+form of the ET-VPD relationship. There is a theoretical basis for the
+square root VPD dependence in both the stomatal conductance model and
+uWUE based on the assumption that stomata behave to maximize carbon
+gain while minimizing water loss, which observations also generally
+support \cite{Lloyd_1991, MEDLYN_2011, Lin_2015, Zhou_2014, Zhou_2015,
+  Medlyn_2017}. However, some purely empirical results that fit the
+exponent of the VPD dependence to data have shown that it may vary
+slightly from 1/2, suggesting that stomata, as well as ecosystem-scale
+quantities based on stomata theory, may not always function optimally
+\cite{Zhou_2015, Lin_2018}. Specifically with regards to uWUE, one
+would not expect that this ecosystem scale WUE quantity will respond
+to VPD exactly analogously to stomata. Direct soil evaporation's
+contributions to ET should shift the exponent of the VPD
+dependence. \citeA{Zhou_2015}'s results corroborate this: they found a
+mean empirically fit exponential VPD dependence of 0.55, varying
+slightly from the theoretically optimal value of of 0.5 for AmeriFlux
+sites. \citeA{Lin_2018}'s results also show variance in the empirical
+exponent of the VPD dependence in the stomatal conductance model, but
+interpretation of this variance is more difficult as \citeA{Lin_2018}
+do not handle the GPP dependence of stomatal conductance in a directly
+analogous manner to the optimal theory in \citeA{MEDLYN_2011} and
+\citeA{Medlyn_2017}. Regardless, given that these recent results on
+the relationship between VPD, GPP, and ET \cite{MEDLYN_2011,
+  Zhou_2014, Zhou_2015, Medlyn_2017} form the backbone of our analysis
+and are what allowed us to derive an explicit ET expression for the
+first time (Eq. (\ref{et})), we will analyze if and how assumptions
+about the exponent of the VPD dependence impact the shape of the
+ET-VPD dependence. This analysis is also important to understand
+whether the choice of stomatal conductance model alters the
+fundamental behavior of the ET-VPD relationships, as many commonly
+used models utilize a VPD exponent other than the 1/2 suggested by
+optimal theory \cite<e.g. >[ which uses an exponent of
+1]{Leuning_1990}. It is useful to understand how the choice of model
+imposes the functional form of the ET-VPD relationship, given its
+fundamental importance for land-atmosphere interactions.
+
+\subsection{Functional form of ET dependence on VPD and its relation
+  to the VPD exponent}
+\label{functional_form}
+
+The results presented in Sect. \ref{results} indicates that for a
+given uWUE and $g_1$, the ET dependence on VPD should be concave
+upward. In other words, there should be some local minimum in ET at a
+critical VPD$_{crit}$, assuming the scaling and plant terms
+(e.g. aerodynamic conductance, $\Delta$, $g_1$ and uWUE) are held
+fixed. This result warrants further investigation, because to our
+knowledge no earlier work has derived the theoretical ecosystem-scale
+relationship between ET and VPD in an energy balance framework, while
+controlling for other environmental conditions. In particular there is
+an apparent lack of consensus over whether the shape of the ET-VPD
+curve should be concave upward (our result) or concave downward in the
+absence of dramatic water stress. Given that understanding the ET-VPD
+relationship of the one-way plant response is fundamental to
+hypothesizing about any feedbacks between the land surface and the
+atmosphere, we analyze why our derived ET-VPD relationship is concave
+upward, particularly with respect to the exponent of VPD dependence in
+uWUE and the Medlyn unified stomatal conductance model, as other
+models and empirical results have suggested exponents varying from 1/2
+\cite{Leuning_1990, Zhou_2015, Lin_2018}.
+
+By introducing $n$, the exponent of VPD in uWUE, and $m$, the exponent
+of VPD in the stomatal conductance model, we can free our theory from
+assumptions about VPD dependence:
+% \begin{linenomath*}
+  \begin{equation}
+    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g*}{VPD^m}\right) \frac{*WUE \; ET}{c_a \; VPD^n},
+    \label{m_n}
+  \end{equation}
+% \end{linenomath*}
+where:
+\[*WUE = \frac{GPP}{ET}VPD^n,\] and $g*$ is a generic slope parameter
+of units $VPD^m$. To determine how the exponent $n$ and $m$ alter the
+shape of the ET-VPD dependence we find the roots of the second
+derivative of ET, using Eq. (\ref{m_n}) for stomatal conductance
+($g_s$), with respect to VPD:
+
+\input{d2_solutions.tex}
+
+With this
+result we have defined the family of curves separating concave up from
+concave down ET solutions (Fig. \ref{concave}). These curves are
+only functions of the exponent of the VPD dependence and a quantity we
+call non-dimensional VPD ($VPD^m/g*$). Several important
+relations reveal themselves from Eq. (\ref{curves}):
+
+\begin{figure}
+  \centering
+  \centerline{\includegraphics[width=0.75\textwidth]{./concave.pdf}}
+  \caption{ Solutions corresponding to inflection points between
+    concave up and concave down ET-VPD curves (Eq. (\ref{curves}))
+    for three specific scenarios. Solutions are defined in terms of a
+    non-dimensional VPD ($VPD^m/g_*$), but to aide physical
+    interpretation the horizontal axis is additionally provided in terms of dimensionalized VPD assuming $m=1/2$ and
+    $g_*=110\; Pa^{1/2}$ (average of all PFT $g_1$). The vertical axis
+    has a different interpretation depending on the solution
+    curve. For the blue line ($m$ varying), it corresponds to $m$,
+    for the orange line ($n$ varying) it corresponds to $n$, and for
+    the green line it corresponds to the value of both $n$ and $m$
+    ($n=m$). Regions of the parameter space that correspond to
+    concave up and concave down results are labeled.}
+  \label{concave}
+\end{figure}
+
+\begin{itemize}
+  \item For optimal behavior (n, m = 1/2) the ET-VPD curve will be
+    concave up regardless of the magnitude of the plant constants
+    $g_1$ and uWUE. Therefore, the general concave up nature of our
+    results, given an assumption of optimal behavior, is insensitive
+    to plant type.
+  \item For all physically possible exponents of VPD dependence ($n,
+    m$), whether the solution is concave up or concave down does not
+    depend on $*WUE$.
+  \item In general, increasing the exponent of VPD dependence for
+    either $*WUE$ or $g_*$ increases the likelihood of a concave down
+    result. Additionally, as the exponent of VPD dependence increases
+    from the optimum value of 1/2, whether the curve is concave upward
+    or concave downward becomes a function of the plant specific slope
+    parameter $g_*$, through non-dimensional VPD
+    ($VPD^m/g_*$). Because the exponent of the VPD dependencies is
+    capable of altering the fundamental shape of ET-VPD dependence,
+    future research investment in understanding the exact VPD
+    dependence of stomatal conductance, and further reconciliation of
+    empirical and theoretical stomatal and ecosystem behavior should
+    be prioritized.
+\end{itemize}
+
+While it is possible (and even likely) that in the future some other
+form of VPD dependence is derived, at present \cite{MEDLYN_2011} and
+\cite{Zhou_2014} established n=m=1/2 as the most likely candidate
+given current theory and empirical data. Additionally, we argue that a
+concave up result matches physical intuition more than a concave down
+result. Plants must maintain nutrient and sugar transport through the
+phloem and xylem. To accomplish this, stomata must remain slightly
+open \cite{De_2013, Nikinmaa_2013, Ryan_2014}. Furthermore, even if
+complete stomatal closure were possible, cuticular water loss and [at
+the ecosystem-scale] direct soil evaporation are still sources of ET
+which increase with VPD, independent of stomatal closure. Therefore,
+in the limit as VPD becomes large and we assume plants are exercising
+all strategies to reduce ET, any further increase in VPD should result
+in an increase in ET through cuticular water loss and/or direct soil
+evaporation. This inevitable transition from conditions when stomata
+respond strongly to VPD to conditions when stomata response is
+asymptoting towards full closure would cause a concave up ET-VPD
+curve, which is matched by our theory. In short, plant response
+becomes more limited as VPD increases, while atmospheric demand
+monotonically increases with VPD, leading to the result that
+atmospheric demand dominates plant response when atmospheric demand is
+high.
+
+This analysis allows us to understand the theoretical shape of the ET
+response to VPD with environmental conditions held
+fixed. Accomplishing this with purely statistical methods applied to
+flux observations would be very difficult, given the relatively fast
+time scale of plant response and the non-stationarity of [solar
+forced] environmental conditions over the relatively coarse (half
+hourly) flux estimates (which is required to obtain robust
+eddy-covariance statistics). Our results on the shape of the ET-VPD
+curve with environmental conditions held fixed can be built upon with
+future work examining how changes in VPD and environmental conditions
+(e.g. soil water storage) feedback upon one another. In the soil water
+storage example, over very long time scales extremely high VPD
+perturbations coupled with no precipitation could result in decreases
+in soil water storage such that water becomes limiting. This could be
+represented by an extension of our framework in which $\lambda$ is
+allowed to decrease with decreasing soil water, increasing uWUE and
+decreasing g$_1$. Here, we focus our results by framing them with
+previous literature to build baseline intuition for ET-VPD dependence.
 
 \section{Conclusions}
+We derived a new form of Penman Monteith using the concept of
+semi-empirical optimal stomatal regulation \cite{Lin_2015,
+MEDLYN_2011} and near constant uWUE \cite{Zhou_2015} to remove the
+implicit dependence of stomatal conductance on GPP and ET. With our
+new form of Penman Monteith we developed a theory for when an
+ecosystem will tend to reduce or increase ET with increasing VPD,
+which we framed using previous literature exploring the relationship
+between plant parameters, plant types, and climate. The goal was to
+understand the range of possible ET responses to VPD and develop some
+intuition for how the ET response may vary with plant types and
+climate. This intuition can be used to disentangle land atmosphere
+feedbacks in more complicated scenarios, and will aid interpretation
+of observations and more sophisticated models.
 
-\input{conclusions.tex}
+ET response to VPD can vary from strongly water conservative (ET
+decreasing in response to increasing VPD), to strongly water intensive
+(ET increasing in response to VPD), which is indicative of the
+diversity of possible plant water conservation strategies. Higher uWUE
+and g$_1$ values increase the likelihood of a positive ET response,
+while decreasing temperature and increasing g$_a$ amplify the
+magnitude of the response. Previous literature \cite{Zhou_2015,
+Lin_2015} suggests that crops, through association with higher g$_1$
+and uWUE, are more likely to exhibit a positive ET response to
+VPD. Shrubs (lower uWUE), C4 plants (lower g$_1$), gymnosperm trees
+(lower g$_1$), and plants in arctic and boreal climates (lower g$_1$)
+are more likely to exhibit a negative ET response to VPD. However
+interpretation of g$_1$-induced variability is partially muddied by
+ambiguity in applying leaf-scale estimations of g$_1$ to the ecosystem
+scale \cite{Medlyn_2017}.
+
+Our paper builds intuition for how plants respond to VPD
+perturbations. We show that given optimal stomatal function and fixed
+environmental conditions, the ET-VPD dependence is theoretically
+concave upward, with ET increasing with increasing VPD as VPD
+increases past some critical value where $\frac{\partial \;
+ET}{\partial \; VPD} = 0$. However future research should focus on
+fully understanding the functional form of VPD dependence, as this
+concave up result is sensitive to the exponent of VPD dependence,
+which we currently believe is 1/2 for both uWUE and the stomatal
+conductance model \cite{MEDLYN_2011, Zhou_2014}. Indeed, this
+sensitivity to the exponent of VPD dependence is an important result
+itself: land surface models, including those used in earth system
+models for climate forecasts, employ different assumptions about the
+exponent of VPD dependence in stomatal conductance
+\cite<e.g.,>[]{Ball_1987, Leuning_1990, MEDLYN_2011}, and these
+assumptions can fundamentally change the relationship between ET and
+VPD from one that is concave upward (local minimum in ET) to one that
+is concave downward (local maximum in ET).
+
+Our results are also applicable to understanding the impact of
+expected increases in VPD induced by global change.  Plant
+physiological responses to direct CO$_2$ effects
+\cite<e.g.,>[]{Swann_2016, Lemordant_2018} receives more attention
+than physiological response to indirect effects like increased VPD
+\cite{Novick_2016}. Here, we provide a framework for understanding ET
+response to VPD using a simplified model of two plant
+parameters. Feedbacks between the land and the atmosphere may alter
+the net response to a long-timescale global VPD perturbation, but our
+focus on the one way plant response to a VPD perturbation in the
+atmospheric boundary layer is an important first step to disentangling
+such feedbacks, both in observations and model simulations of the
+present and future. By removing Penman Monteith's dependence on
+implicit relationships between GPP, VPD, and ET, we allow for explicit
+future analysis of plant-VPD feedbacks in the atmospheric boundary
+layer (Eq. (\ref{et})). Our approach can be extended to examine
+varying plant response to more nuanced consideration of plant type and
+climate. Any plant physiological heterogeneity or feedback that can be
+conceptualized with shifts in $g_1$ \cite<e.g.>[]{Lin_2015,
+Medlyn_2017} and/or uWUE \cite<e.g.>[]{Zhou_2014} are representable
+within our framework, which opens the door for a hierarchy of more
+sophisticated climate- and plant-specific analyses of ET sensitivity
+to environmental variables (including VPD). We argue that such
+simplified conceptual frameworks are critical tools for disentangling
+land-atmosphere feedbacks at various scales, from diurnal to seasonal
+and beyond, and to characterize ET response in a warmer,
+atmospherically drier, and enriched CO$_2$ world.
 

### 2019-09-04 doc/paper/agu-james-submission.tex
started actual work on revisions, but note for latexdiff to work all edited content must be in same file

--- 2019-06-11 doc/paper/agu-james-submission.tex
@@
 \begin{abstract}
-\input{abstract.tex}
+Increasing vapor pressure deficit (VPD) increases atmospheric demand
+for water. While increased evapotranspiration (ET) in response to
+increased atmospheric demand seems intuitive, plants are capable of
+reducing ET in response to increased VPD by closing their stomata. We
+examine which effect dominates the response to increasing VPD:
+atmospheric demand and increases in ET, or plant response (stomata
+closure) and decreases in ET. We use Penman-Monteith, combined with
+semi-empirical optimal stomatal regulation theory and underlying water
+use efficiency, to develop a theoretical framework for assessing ET
+response to VPD. The theory suggests that depending on the environment
+and plant characteristics, ET response to increasing VPD can vary from
+strongly decreasing to increasing, highlighting the diversity of plant
+water regulation strategies.
+
+The ET response varies due to: 1) climate, with tropical and temperate
+climates more likely to exhibit a positive ET response to increasing
+VPD than boreal and arctic climates; 2) photosynthesis strategy, with
+C3 plants more likely to exhibit a positive ET response than C4
+plants; and 3) plant type, with crops more likely to exhibit a
+positive ET response, and shrubs and gymniosperm trees more likely to
+exhibit a negative ET response. These results, derived from previous
+literature connecting plant parameters to plant and climate
+characteristics, highlight the utility of our simplified framework for
+understanding complex land atmosphere systems in terms of idealized
+scenarios in which ET responds to VPD only. This response is otherwise
+challenging to assess in an environment where many processes co-evolve
+together.
 \end{abstract}
@@
 
-\input{intro.tex}
+Vapor pressure deficit (VPD) is expected to rise over continents in
+the future due to the combination of increased temperature and,
+depending on region, decreased relative humidity
+\cite{Byrne_2013}. Increases in VPD increase the atmospheric demand
+for evapotranspirated water \cite{Penman_1948, Monteith_1965}, but
+also reduce stomatal conductance through stomatal closure
+\cite{Rawson1977, Leuning_1990, Mott2007, Damour2010,
+MEDLYN_2011}. Understanding the net evapotranspiration (ET) response
+to these two opposing effects of changes in VPD is crucial for
+assessing the impact of environmental VPD perturbations on the water
+cycle.
 
-\input{body.tex}
+The opposing effects of increased atmospheric demand and higher
+stomatal closure lead to two possible perspectives for how ET responds
+to shifts in VPD. The first, a hydrometeorological perspective, is
+that higher VPD increases atmospheric demand for water from the land
+surface, and this drives an increase in ET
+\cite{Penman_1948}. However, plants' stomata have evolved to
+optimally regulate the exchange of water and carbon, and tend to
+partially close in response to increased atmospheric dryness
+\cite{Farquhar_1978, Ball_1987, Leuning_1990, Katul_2009,
+MEDLYN_2011}. This leads to a plant physiology perspective, in which
+an increase in VPD may actually correspond to a decrease in ET because
+of stomatal closure \cite<e.g.>[]{Rigden_2017}.  In other words, the
+question ``When does VPD drive or reduce ET?'' can be related to
+whether plant regulation or atmospheric demand dominates the ET
+response.
+
+The ET response to changes in VPD alters water partitioning between
+the soil and atmosphere. If ecosystem plant response reduces ET with
+atmospheric drying then soil moisture will be better conserved. This
+represents a sensible evolutionary strategy to cope with aridity: save
+water for periods when atmospheric demand for water is relatively low,
+and atmospheric carbon can be accessed with a relatively smaller cost
+in water loss. If instead stomata were fully passive \cite<similar
+to soil pores, e.g. >[]{Or_2013}, increased atmospheric aridity would
+strongly reduce soil moisture \cite{Berg_2017}. This could further
+increase aridity as low soil moisture levels increases the Bowen
+ratio, leading to increased temperature and atmospheric drying
+\cite{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006,
+Salvucci_2013, Gentine_2016, Berg_2016, Zhou_2019}. Therefore, passive
+regulation and a lack of soil moisture conservation does not seem to
+be a sensible strategy for plants from an evolutionary
+standpoint. This simplified logic explains generally why plants
+evolved to respond to VPD, but also excludes many details and special
+cases \cite<e.g. plant to plant interaction and highly specialized
+photosynthesis strategies like Crassulacean acid metabolism
+photosynthesis>[]{Brooker_2006, Cushman_2001}.
+
+We can use intuition about plant water conservation strategy to
+hypothesize about ET response to changes in VPD. Plants and ecosystems
+that evolved to conserve water, such as arid shrubs, should be more
+likely to reduce ET with increasing VPD, and plants that have evolved
+or have been engineered to prioritize carbon gain over water
+conservation, such as crops, will be more likely to increase ET with
+increasing VPD. Atmospheric conditions must matter as well. At the
+ecosystem scale, there are limits to plant water conservation
+strategies. As atmospheric demand for water (VPD) increases,
+ecosystems may begin to reach their water conservation limits and
+might not be able to entirely limit ET flux to the atmosphere. At this
+stage any further increase in VPD will most likely drive a (limited)
+increase in ET, because the increase in atmospheric demand for water
+overwhelms the limited plant response to conserve water.
+
+The objective of the present manuscript is to use reasonable
+approximations established in prior research as a tool to develop a
+framework for understanding plant responses to atmospheric drying and
+the VPD dependence of ET while keeping other variables fixed. This
+framework will aid interpretation of observations, full complexity
+models, and facilitate the disentanglement of complex land-atmosphere
+feedbacks. In particular, our approach has applications for
+understanding climate change impacts, given expected increases in VPD
+with rising temperature.  In the past, similar simplified approaches
+were used to understand interactions between stomatal conductance,
+evapotranspiration and the environment \cite<e.g.,>[]{Jarvis_1984,
+Jarvis_1986, Mcnaughton_1991}. However these researchers did not
+explore explicitly the sensitivity of ET to VPD, including VPD's
+effect on stomatal conductance and plant function.
+
+Approaching the problem of ET response to VPD is aided by recent
+results drastically improving our understanding of VPD's impact on
+physiology, especially at the leaf level. \citeA{MEDLYN_2011}
+developed a model for leaf-scale stomatal conductance ($g_s$),
+including VPD response, by combining an optimal photosynthesis theory
+\cite{Cowan_1977, Katul_2009, Katul_2010} with an empirical approach,
+and extended use of this model to the ecosystem scale in
+\citeA{Medlyn_2017}. Additionally, \citeA{Zhou_2014} demonstrated that
+a quantity underlying water use efficiency $\left(uWUE = \frac{GPP\;
+\sqrt{VPD}}{ET}\right)$ properly captures a constant relationship
+between GPP, ET, and VPD over a diurnal cycle at the ecosystem
+scale. uWUE is also relatively well conserved in the growing season
+across space and time, within a plant functional type (PFT)
+\cite{Zhou_2015}.  While stomatal conductance parameterizations and
+uWUE greatly simplify complex plant physiological processes, they
+ still capture ecosystem behavior for vegetated surfaces
+\cite{Medlyn_2017, Zhou_2014}, and are useful tools to transparently
+develop intuition for the behavior of complex land-atmosphere systems.
+
+In this manuscript, we leverage uWUE and recent developments in
+stomatal conductance parameterizations \cite{MEDLYN_2011} with a
+Penman-Monteith framework \cite<hereafter PM,>[]{Penman_1948,
+Monteith_1965} to derive the theoretical one-way response of ET to VPD
+with other environmental variables properly controlled for, i.e. we
+develop a framework for evaluating the partial derivative of ET with
+respect to VPD. It is useful to disconnect the impact of VPD from
+other variables as VPD is known to increase dramatically with future
+climate and limit ET more than soil moisture \cite{Novick_2016}, but
+disentangling VPD effects from soil moisture effects has been
+difficult in previous research, given their co-variability
+\cite{Lin_2018, Zhou_2019}.  We are
+able to disentangle the effect of VPD on ET because we explicitly
+include VPD's full effect on stomatal conductance, including its
+impact on photosynthesis. The use of an energy balance framework (PM)
+allows us to include in our analysis the effects of the energy cost of
+evaporating water from a surface, which is an important factor in the
+natural environment, compared to prescribed in situ environmental
+conditions. This is relevant because previous research focusing on the
+leaf scale \cite{Rawson1977, Turner1984, Oren1999, Damour2010,
+Mott2013} does not consistently or analogously include the energetic
+constraints on evapotranspiration under which ecosystems operate. The
+leaf-scale results agree that stomatal conductance decreases in
+response to VPD \cite{Oren1999, Damour2010}, which we expect to be
+true for an ecosystem as well. However leaf scale results also
+indicate that transpiration usually increases with increasing VPD in a
+concave downward shape \cite<e.g.,>[]{Rawson1977, Turner1984,
+Mott2013}, which may not be true for ET at the ecosystem scale once
+energetic constraints on ET are included in the analysis.  Our
+approach allows us to estimate the expected ET response to VPD at the
+ecosystem scale, including the effects of surface energy balance
+constraints, and assess how ecosystem ET response to VPD deviates from
+previous leaf scale analyses.
+
+This manuscript presents the range of possible ecosystem-scale ET
+responses to VPD, given parameters previously established in peer
+reviewed literature. Additionally, we explore the sensitivity of the
+ET-VPD relationship to stomatal model and framework choice,
+highlighting the importance of: 1) future research on stomatal
+conductance and ecosystem scale modeling, and 2) thoughtful selection
+of photosynthesis and stomatal conductance models in more
+sophisticated land surface and earth system models.
+
+\section{Methods}
+\label{methods}
+The Penman-Monteith equation \cite<hereafter PM,>[]{Penman_1948,
+  Monteith_1965} estimates ET as a function of observable atmospheric
+variables and surface conductances:
+% \begin{linenomath*}
+  \begin{equation}
+    \label{orig_pen}
+    ET = \frac{\Delta R_{net} + g_a \rho_a c_p VPD}{\Delta + \gamma(1 + \frac{g_a}{g_s})},
+  \end{equation}
+% \end{linenomath*}
+  where $\Delta$ is the change in saturation vapor pressure with
+  temperature, given by Clausius-Clapeyron
+  ($\frac{d \; e_s}{d \; T}$), $R_{net}$ is the net radiation minus
+  ground heat flux, $g_a$ is aerodynamic conductance, $\rho_a$ is air
+  density, $c_p$ is specific heat of air at constant pressure,
+  $\gamma$ is the psychometric constant, and $g_s$ is the stomatal
+  conductance (Table \ref{definitions}). The issue with PM is that
+  $g_s$ is a function of carbon uptake, which is has a strong
+  functionally relation to ET through stomatal function. Therefore,
+  when PM is formulated in terms of $g_s$, it is an implicit function
+  of ET itself rather than an explicit function of ET. Here we will
+  derive a new form of PM in which ET is an explicit function of plant
+  parameters and environmental conditions, and use it to assess the
+  ecosystem scale response to VPD (by taking a partial derivative).
+
+\begin{table}
+  \caption{Definition of symbols and variables, with citation for how
+    values are calculated, if applicable.}
+  \label{definitions}
+  \centering \footnotesize
+  \begin{tabular}{l c c c}
+    \hline
+    Variable & Description & Units & Citation \\
+    \hline
+    $e_s$  & saturation vapor pressure & Pa  & - \\
+    $T$  & temperature  & K & - \\
+    $P$  & pressure & Pa  & - \\
+    $\Delta$  & $\frac{\partial e_s}{\partial T}$ & Pa K$^{-1}$ & - \\
+    $R_{net}$  & net radiation at land surface minus ground heat flux & W m$^{-2}$   & - \\
+    $g_a$  & aerodynamic conductance & m s$^{-1}$  & \citeA{Shuttleworth_2012} \\
+    $\rho_a$  & air density & kg m$^{-3}$  & - \\
+    $c_p$  & specific heat capacity of air at constant pressure & J K$^{-1}$ kg$^{-1}$ & - \\
+    $VPD$  & vapor pressure deficit & Pa  & - \\
+    $\gamma$  & psychometric constant & Pa K$^{-1}$   & - \\
+    $g_{s}$  &  stomatal conductance & m s$^{-1}$
+                                   & \citeA{Medlyn_2017} \\
+    $g_{1}$  & ecosystem-scale slope parameter & Pa$^{0.5}$ & \citeA{Medlyn_2017} \\
+    $R$ & universal gas constant & J mol$^{-1}$ K$^{-1}$ & - \\
+    $R_{air}$ & gas constant of air & J  K$^{-1}$ kg$^{-1}$ & - \\
+    $\sigma$ & uncertainty parameter & -& - \\
+    $c_a$ & CO$_2$ concentration & $\mu$ mol CO$_2$ mol$^{-1}$ air& - \\
+    $\lambda = \frac{\partial \; transpiration}{\partial\; CO_2\; assimilation}$ & marginal water cost of leaf carbon & mol H$_2$O mol$^{-1}$ CO$_2$ & - \\
+    $\Gamma$ & CO$_2$ compensation point & - & - \\
+    $\Gamma^*$ & CO$_2$ compensation point without dark respiration &
+                                                                      - & - \\
+    GPP & gross primary production & $\mu$-mol C s$^{-1}$ m$^{-2}$ & -
+    \\
+    ET & evapotranspiration & W m$^{-2}$ & - \\
+   uWUE & underlying water use efficiency & $\mu$-mol C Pa$^{0.5}$
+                                            J$^{-1}$ ET  &
+                                                           \citeA{Zhou_2015} \\
+    \hline
+  \end{tabular}
+\end{table}
+
+\citeA{MEDLYN_2011} developed a model for stomatal conductance ($g_s$)
+by combining an optimal photosynthesis theory \cite{Cowan_1977} with
+an empirical approach, which describes the dependence of $g_s$ to
+VPD. They also extended this model to the ecosystem scale
+\cite{Medlyn_2017}:
+
+% \begin{linenomath*}
+  \begin{equation}
+    g_s = \frac{R \,T}{P} \; 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{GPP}{c_a},
+    \label{medlyn}
+  \end{equation}
+% \end{linenomath*}
+
+  where $g_{1}$ is a VPD ``slope'' parameter, GPP is the ecosystem
+  scale gross primary production, and $c_a$ is the atmospheric CO$_2$
+  concentration at the canopy. \cite{MEDLYN_2011} relate the slope
+  parameter ($g_{1}$) to physical parameters as:
+% \begin{linenomath*}
+
+  \begin{equation}
+    g_{1}  \propto \sqrt{\frac{3 \, \Gamma^* \,
+        \lambda}{1.6}},\footnote{Note this expression has units of of
+      (mmol mol$^{-1}$)$^{1/2}$, but this can be converted to
+      Pa$^{1/2}$ using the ideal gas law.}
+    \label{slope}
+  \end{equation}
+% \end{linenomath*}
+
+where $\Gamma^*$ is the CO$_2$ compensation point for photosynthesis
+(without dark respiration), and $\lambda$ is the marginal water cost
+of leaf carbon
+($\frac{\partial \; \text{transpiration}}{\partial \; CO_2 assimilation}$)
+\cite{Farquhar_1980, Katul_2009}. So,
+$g_{1}$ is a leaf-scale term reflecting the trade-off of water for
+carbon uptake. The higher $g_{1}$, the more open the stomata and
+the more they release water in exchange for carbon.
+
+While Eq. (\ref{medlyn}) can be used in PM (Eq. (1)), it will make
+analytical work with the function intractable because GPP is
+functionally related to ET itself. Additionally, a perturbation to VPD
+should induce a physiological plant response that will alter GPP and
+cause an indirect change in stomatal conductance, in addition to the
+direct effect of VPD in Eq. (\ref{medlyn}). Therefore, in order to
+derive the full plant response of ET to VPD, we must account for the
+functional relationship between GPP, ET, and VPD, and its effect on
+stomatal conductance. We can use aforementioned semi-empirical results
+of \citeA{Zhou_2015}, which were inspired by optimal photosynthesis
+theory, as a tool to approach this problem. \citeA{Zhou_2015}, showed
+that underlying Water Use Efficiency (uWUE):
+
+% \begin{linenomath*}
+  \begin{equation}
+    uWUE = \frac{GPP \cdot \sqrt{VPD}}{ET}
+    \label{uwue}
+  \end{equation}
+% \end{linenomath*}
+
+  is relatively constant across time within a plant functional type,
+  and correctly captures a constant relationship between GPP, ET and
+  VPD over a diurnal cycle during the growing season
+  \cite{Zhou_2014}. The theoretical derivation of the square root VPD
+  dependence in uWUE leverages the same assumptions used in
+  \citeA{MEDLYN_2011} to derive the square-root VPD dependence of the
+  stomatal conductance model (Eq.  (\ref{medlyn})).  We can use uWUE
+  to remove the $GPP$ dependence in $g_s$ in a way that makes PM
+  analytically tractable:
+
+% \begin{linenomath*}
+  \begin{equation}
+    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{uWUE \; ET}{c_a \; \sqrt{VPD}}.
+    \label{new_g_s}
+  \end{equation}
+% \end{linenomath*}
+
+Plugging Eq. (\ref{new_g_s}) into Eq. (\ref{orig_pen}) and
+rearranging gives a new explicit expression for PM, in which
+dependence on $GPP$ is removed:
+
+% \begin{linenomath*}
+  \begin{equation}
+    ET = \frac{\Delta R_{net} + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R \; 1.6\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
+    \label{et}
+  \end{equation}
+% \end{linenomath*}
+
+  By accounting for photosynthesis changes in ecosystem conductance,
+  with Eq. (\ref{et}) we have used recent results
+  \cite{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017} to derive
+  ET explicitly as function of environmental variables and two
+  plant-specific constants, the slope parameter ($g_1$), and uWUE. For
+  the first time we have removed the implicit dependence of ET on
+  itself through the stomatal conductance term, and we have also
+  replaced the added complexity of a stomatal conductance reduction
+  factor and a photosynthesis model with a single parameter (uWUE). Both
+  the plant parameters reflect water conservation strategy. The
+  slope parameter is related to the willingness of stomata to trade
+  water for CO$_2$ and to keep stomata open (carbon cost in terms of
+  water)). uWUE is a semi-empirical ecosystem-scale constant related
+  to how WUE changes with VPD (specifically $VPD^{-1/2}$). It is also
+  roughly proportional to physical constants:
+
+\[uWUE \appropto \sqrt{\frac{c_a - \Gamma}{1.6 \lambda}},\]
+
+where $\Gamma$ is the CO$_2$ compensation point \cite<Eq. (5)
+in>[]{Zhou_2014}. So uWUE is related to atmospheric CO$_2$
+concentration and compensation point, and is inversely proportional to
+the marginal water cost of leaf carbon ($\lambda$). This relationship
+with $\lambda$ ($\propto \lambda^{-1/2}$) is important as it is the inverse of g$_1$'s
+relationship with $\lambda$ ($\propto \sqrt{\lambda}$). So, uWUE
+should increase as $\lambda$ decreases, and g$_1$ should decrease as
+$\lambda$ decreases.
+
+With Eq. (\ref{et}) we can take the partial derivative of ET with respect
+to VPD to understand whether VPD drives or reduces ecosystem-scale ET:
+
+% \begin{linenomath*}
+  \begin{equation}
+    \frac{\partial \;  ET}{\partial \; VPD} = \frac{2\; g_a \;
+      P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} -
+      \frac{\gamma c_a }{1.6 \; R\; \; \text{ uWUE }} \left(
+        \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right)
+    \right),
+    \label{d_et}
+  \end{equation}
+% \end{linenomath*}
+
+  providing analytical framework for ecosystem response to
+  atmospheric demand with environmental conditions held
+  fixed. There are a few subtleties to taking the derivative in
+  Eq. (\ref{d_et}): $\Delta$ ($\frac{d e_{s}}{d T}$) and $VPD$ are
+  functionally related, so while taking the derivative we evaluate
+  $\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET}
+  {\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD}
+  \Big|_{\text{RH fixed}} + \frac{\partial \; ET}{\partial \; RH}
+  \frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$
+      fixed}}$. $RH$ and $e_s$ are assumed to be approximately
+  independent, which is supported by data (not shown).
+
+  This derivation relied either implicitly or explicitly on several
+  assumptions. First, we assume that VPD at the leaf surface is the
+  same as VPD at measurement height; physically this implies that
+  leaves are perfectly coupled to the atmosphere. In reality, for some
+  conditions and plant types the leaves can become decoupled from the
+  boundary layer \cite{De_2017, Medlyn_2017}. Therefore, our
+  derivation will be most applicable in times like the growing season
+  (when we also expect uWUE to be most valid), when relatively high
+  insolation induces instability and convective boundary layers, and
+  we would expect the surface to be generally well coupled. An
+  additional assumption in the formulation of uWUE \cite{Zhou_2014,
+    Zhou_2015} and \citeA{Medlyn_2017}'s stomatal conductance model is
+  that direct soil evaporation contributions to ET remain small
+  relative to transpiration. Again, this should be more true during
+  the growing season. The ratio of evaporation to transpiration may
+  increase immediately after rainfall events due to high soil moisture
+  and ponding, but VPD is generally low anyways during these
+  times. However, some plant types allow for systematically larger
+  contributions of evaporation in ET, particularly those with sparse
+  canopies and smaller relative amounts of transpiration. We therefore
+  might expect that the theory will be most applicable to forest PFTs,
+  which will be most strongly coupled to the boundary layer due to
+  larger surface roughness, and will also generally have the highest
+  ratios of transpiration to evaporation. Finally, we assume that
+  $g_1$ and uWUE are constant with respect to the conceptual VPD
+  perturbation. Both quantities have been shown to be relatively
+  constant with respect to changes in VPD \cite{Franks_2017,
+    Zhou_2014}. These parameters will however vary with plant species
+  and characteristics \cite<e.g. wood density, >[]{Lin_2015}, as well
+  as environmental conditions including soil water content and
+  temperature \cite{Lin_2015, Manzoni2013}. Exploring possible soil
+  moisture (in)dependence of the plant parameters (g$_1$ and uWUE) is
+  particularly interesting because soil moisture only enters the
+  partial derivative directly through these plant parameters. If the
+  plant parameters are weak functions of soil moisture then the theory
+  can be directly applied to a broader range of conceptual VPD
+  scenarios, including observed compound events between high VPD and
+  low soil moisture \cite{Zhou_2019}. Supplementary material for this
+  manuscript explores the soil moisture dependence of the plant
+  parameters, but excessive noise and inconsistencies preclude
+  conclusions about the nature of any dependence. We provide it in
+  case it is of interest to the reader, and to motivate future
+  research.
+
+  Because Eq. (\ref{d_et}) is a partial derivative, it's utility is as
+  a conceptual model for the change in ET in response to VPD with all
+  other variables held fixed. This provides a useful tool for
+  identifying the effect of VPD on ET through atmospheric demand and
+  plant response, and allows a practitioner to disentangle complicated
+  feedbacks when many quantities co-vary. However users of
+  Eq. (\ref{d_et}) should take care that their interpretation matches
+  the assumptions inherent in a partial derivative. For example,
+  results will only be as valid as the assumption that g$_1$ and uWUE
+  (and by extension $\lambda$) are fixed with respect to the user's
+  conceptual change in VPD. Care must also be taken with possible
+  indirect effects associated with a change in VPD: for example, a
+  change in ET induced by a change in VPD can cause a change in
+  surface temperature, which would drive a change in net
+  radiation. These types of indirect effects and feedbacks are not
+  considered in Eq. (\ref{d_et}): temperature (a variable) is
+  mathematically fixed.
+
+\subsection{Framing Eq. \ref{d_et} with previous research }
+
+  To account for both the variability in plant parameters and the
+  environment, we will systematically analyze how the ET response to
+  VPD (Eq. (\ref{d_et})) varies. Eq. (\ref{d_et}) includes a
+  ``sign'' term that determines the sign of the response in addition
+  to magnitude:
+
+\begin{equation}
+  \label{sign}
+  \frac{c_p}{R_{air}} - \frac{\gamma c_a }{1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right),
+\end{equation}
+
+and a ``scaling'' term multiplying the ``sign'' term:
+
+\begin{equation}
+  \frac{g_a \; P}{T(\Delta + \gamma)}.
+\end{equation}
+
+In the ``sign term'' most of the quantities are relatively constant,
+except for VPD and the plant parameters g$_1$ and uWUE. In the scaling
+term, most of the terms are relatively constant with the exception of
+aerodynamic conductance ($g_a$) and temperature (especially its effect
+through $\Delta$). To determine the range of probable ET responses to
+VPD we will systematically vary these parameters according to Table
+\ref{param_varying}, while all other parameters are held fixed (Table
+\ref{param_fixed}). Physical variables ($g_a$, $T$) and plant
+physiological parameters (g$_1$, uWUE) are varied according to
+literature-based expectations for a range of growing season conditions
+and plant types \cite{Zhou_2015, Medlyn_2017}. Using this previous
+literature we can connect the effect of varying plant parameters to
+specific plant types and characteristics.
+
+All code and data used in this analysis, including those used to
+generate the figures and tables, are publicly available at
+\url{https://github.com/massma/climate_et}.
+%\hyperref[https://github.com/massma/climate\_et]{https://github.com/massma/climate_et}.
+
+\begin{table}
+  \caption{Variable quantities in the ET response to VPD. Each value
+    is varied to determine the effect of a range of expected plant and
+    environmental conditions on ET response to VPD. A citation is
+    provided in cases where the quantities are directly derived from
+    previous literature. Conceptually, min. values are extracted from
+    literature to correspond to approximately the 15th percentile of
+    observed conditions during the growing season, med. values
+    correspond to approximately the 50th percentile, and max. values
+    correspond to approximately the 85th percentile.}
+  \label{param_varying}
+  \centering
+  \footnotesize
+  \begin{tabular}{l c c c c c}
+    \hline
+    Symbol & units [units in citation] & min  & med & max
+    & citation  \\
+    \hline
+    \input{param_varying.tex}
+    \hline
+  \end{tabular}
+\end{table}
+
+\begin{table}
+  \caption{Quantities that are fixed in the ET response to VPD
+    (relative to those in Table \ref{param_varying}).}
+  \label{param_fixed}
+  \centering
+  \begin{tabular}{l c c}
+    \hline
+    Symbol & units & value \\
+    \hline
+    \input{param_fixed.tex}
+    \hline
+  \end{tabular}
+\end{table}
+
+\section{Results}
+\label{results}
+
+By varying four parameters (g$_a$, T, uWUE, g$_1$) at three different
+values (Table \ref{param_varying}) we generate nine different values
+for the scaling term, nine different curves (as a function of VPD) for
+the sign term, and 81 different curves for the ET response to VPD
+($\frac{\partial \; ET}{\partial \; VPD}$). To aid visualization we
+can examine a subset of nine of these 81 curves, defined by the
+minimum, median and maximum values for both the scaling and sign term
+(Figure \ref{full}). The range of ET responses to VPD vary from those
+where ET almost always decreases with increasing VPD (water
+conservative), to those where ET almost always increases with
+increasing VPD (water intensive). Additionally, for some parameters
+whether ET will increase or decrease with increasing VPD depends on
+atmospheric demand (Figure \ref{full}).
+
+\begin{figure}
+  \centering \includegraphics{./fully_idealized.pdf}
+  \caption{The functional form of $\frac{\partial \; ET}{\partial
+      \; VPD}$ for minimum, median and maximum values of both the sign
+    term and the scaling term.}
+  \label{full}
+\end{figure}
+
+\begin{figure}
+  \centering \includegraphics{./fully_idealized_sign.pdf}
+  \caption{The functional form of $\frac{\partial \; ET}{\partial
+      \; VPD}$ evaluated at the median value of the scaling term, for
+    varying values of g$_1$ and uWUE as given in Table \ref{param_varying}.}
+  \label{sign}
+\end{figure}
+
+Examining the sign term independent of the scaling term illuminates
+the role of the two plant parameters, g$_1$ and uWUE, in determining
+the degree to which the response is water conservative
+($\frac{\partial \; ET}{\partial \; VPD} < 0$) or water intensive
+($\frac{\partial \; ET}{\partial \; VPD} > 0$) (Figure \ref{sign})
+. Higher g$_1$ and uWUE shift the curve towards increasing ET
+responses with VPD (water intensive), and smaller g$_1$ values lead to
+a larger VPD dependence of the response (i.e. ET response is a
+stronger function of VPD). However care should be exercised when
+interpreting the range of ET responses, because g$_1$ and uWUE should
+generally be anti-correlated due to their dependencies on
+$\lambda$. As $\lambda$ increases, g$_1$ should increase and uWUE
+should decrease. Because high and low values of uWUE and g$_1$ were
+determined independently from each other in previous literature, plant
+types with anomalously high or low values for \textit{both} g$_1$ and
+uWUE should be relatively rare. However, some plant types do exhibit
+co-occurring high values for g$_1$ and uWUE. For example, C3 crops
+have both the highest g$_1$ value in \citeA{Franks_2017}, and the
+highest uWUE value in \citeA{Zhou_2015}.
+
+\begin{figure}
+  \centering \includegraphics{./fully_idealized_scaling.pdf}
+  \caption{The functional form of $\frac{\partial \; ET}{\partial
+      \; VPD}$ evaluated at the median value of the sign term, for
+    varying values of g$_a$ and T as given in Table \ref{param_varying}.}
+  \label{scaling}
+\end{figure}
+
+Examining the scaling term independent of the sign term shows how
+aerodynamic conductance (communication between the atmosphere and the
+surface) and temperature (controls the efficiency of energy conversion
+to latent heat through $\Delta$), amplify or suppress the plant
+response represented in the sign term (Figure \ref{scaling}). Both
+lower temperatures and higher aerodynamic conductance lead to
+amplified ET response to VPD, with the variability of aerodynamic
+conductance resulting in a slightly higher variability of ET response
+relative to temperature.
+
+\section{Discussion}
+\label{discussion}
+
+Interpreting both the derivation and results of this manuscript hinges
+on an understanding and appreciation for the usefulness of partial
+derivatives for understanding the behavior of complex systems in a
+simplified framework. Our results are a conceptual tool for answering
+the question, ``What is the ET response to VPD with all other
+quantities held fixed?'' This is useful, and we would argue is
+critical, for understanding complex responses when many quantities are
+varying simultaneously \cite{Zhou_2019}. If we cannot understand the
+contribution of one term to observed variability, independent of other
+confounding factors, there is little hope of disentangling and
+understanding the relative role of many co-varying quantities in
+determining observed variability.  Eq. (\ref{d_et}) explicitly
+provides an estimate of the ET response to VPD, given assumptions
+about ecosystem plant characteristics through the parameters g$_1$ and
+uWUE. Eq. (\ref{d_et}) is generic to external environmental factors
+and the timescale over which a VPD response acts, and has a
+straightforward conceptual interpretation. It provides the ET
+response, given other quantities and parameters held fixed, subject to
+the assumptions outlined in Sect. \ref{methods}. So, while the plant
+parameters g$_1$ and uWUE may vary with plant species and
+environmental conditions like soil moisture, we can still assess the
+ET response to VPD for a given ecosystem state, or for a given soil
+moisture condition.
+
+However, given that g$_1$ and uWUE can vary, and they both have a
+large impact on the ET response (Figure \ref{full}, \ref{sign}), it is
+useful to examine how these quantities vary in previous literature,
+framed by our results on $\frac{\partial \; ET}{\partial \; VPD}$. In
+\citeA{Zhou_2015} most plant types' uWUE is similar to our median
+value, under which the sign of the ET response depends on
+VPD. However, crops, which we expect to prioritize carbon uptake over
+water conservation, have a higher uWUE value closer to our maximum
+value. This higher uWUE results in a higher likelihood for an
+increasing ET response to VPD, which matches intuition given that we
+expect crops to keep stomata open for access to carbon, at the cost of
+increased water loss during high VPD. Shrubs, and to a lesser extent
+grass, have an uWUE closer to the minimum uWUE. For these plant types,
+we would then expect a decrease of ET in response to VPD. It is
+important to note that, while \citeA{Zhou_2015} did not examine the
+role of soil moisture for within plant-type variability of uWUE, we
+might expect some variation, especially in extreme cases when soil
+water becomes a limiting factor (see supplementary material).
+
+\citeA{Medlyn_2017}'s ecosystem-scale results did not display the
+robust relationship between plant type and and plant parameter (g$_1$)
+that was demonstrated by \citeA{Zhou_2015} for uWUE. Given that there
+is a robust relationship between g$_1$ and plant type in
+\citeA{Lin_2015}'s leaf-scale results, and that \citeA{Zhou_2015}'s ecosystem-scale results
+rely on the same underlying theory as \citeA{Medlyn_2017}'s results,
+the ambiguous results in \citeA{Medlyn_2017} could be due to noise in
+ecosystem scale observations, the consequences of imposing a model
+structure in some of their g$_1$ estimations, or real ecosystem-scale
+within-plant type variability (e.g. multiple species impacting the
+relationship within the eddy-covariance footprint). These differences
+between leaf-scale behavior and ecosystem scale behavior highlight the
+importance of understanding how and why slope parameters show
+increased variability at the ecosystem scale, as argued by
+\citeA{Medlyn_2017}. However, if we assume that there is some real
+analogy between leaf scale results in \citeA{Lin_2015} and ecosystem
+scale behavior, and that some of the ambiguity in \citeA{Medlyn_2017}
+is due to model and/or observational error, then we can use the
+relationships between g$_1$ and plant characteristics defined by
+\citeA{Lin_2015} to frame the ET response to VPD. This is subject to
+the strong caveat that we still do not fully understand g$_1$ behavior
+at the ecosystem scale. Extrapolating \citeA{Lin_2015}'s results
+(e.g. Fig. 2) to the ecosystem scale would reveal the following
+relationships between g$_1$ and plant types:
+
+\begin{itemize}
+  \item C3 plants would have a generally higher g$_1$ than C4 plants,
+  \item Crops would have a larger g$_1$ than shrubs, grass, and
+    angiosperm trees, which have a higher g$_1$ than gymnosperm
+    trees.
+  \item Tropical and temperate climates would generally be
+    characterized by plants with a higher g$_1$ than arctic and boreal
+    climates.
+\end{itemize}
+For all of the above relationships, a higher g$_1$ means a higher
+likelihood of a positive ET response to increasing VPD (water
+intensive), and a smaller VPD dependence of the ET response.
+
+Our theoretical results highlight the variability of ET
+response to VPD as a function of both climate  and plant
+characteristics, especially water usage strategy. Whether an
+ecosystem increases or decreases ET in response to VPD will vary
+depending on vegetation and climate. Generalizations about ET response
+to VPD therefore require thoughtful consideration of ecosystem
+characteristics; statements such as ``ET increases with warming due to
+increases in VPD'' may be false depending on the water conservation
+strategy of a given ecosystem \cite{Lemordant_2018}.
+
+Additionally, the sensitivity of ET response to plant parameters
+highlights the importance of understanding and developing stomatal
+conductance models and parameters for the ecosystem scale. In
+particular, why and how g$_1$ behavior at the leaf scale is not
+analogous to the ecosystem scale must be understood. While the
+derivation here was explicit about the assumption of a constant g$_1$
+term with respect to the VPD perturbation, many sophisticated land
+surface models and earth system models employing stomatal conductance
+models make similar assumptions about the stationarity of VPD slope
+terms \cite{Niu_2011, Franks_2017, Rogers_2017, Lawrence_2019}. Many
+models will assume a constant VPD slope term (e.g. g$_1$) within
+a given plant functional type, and research derived from these models
+often does not acknowledge limitations of this assumption, or the
+difficulty in the literature with quantifying an ecosystem scale slope
+parameter \cite{Medlyn_2017}. Given that something so fundamental as
+the ET response to VPD varies strongly with g$_1$, we must invest more
+in understanding g$_1$'s behavior at the ecosystem scale, and the
+efficacy of models using constant slope terms for ecosystem-scale
+fluxes. Fortunately, our framework, and the new approach of using
+robust ecosystem-scale ratios (uWUE) to remove the implicit dependence
+of stomatal conductance on ET, is flexible enough to be applied to any
+stomatal conductance model that contains a dependence on GPP.
+
+So far we have discussed the role of different parameters in
+modulating the ET response. However, the form of the ET response is
+imposed by our choice of stomatal conductance model. We now examine
+how stomatal model choice, and specifically the exponent of the VPD
+dependence in the stomatal conductance model, can impact the general
+form of the ET-VPD relationship. There is a theoretical basis for the
+square root VPD dependence in both the stomatal conductance model and
+uWUE based on the assumption that stomata behave to maximize carbon
+gain while minimizing water loss, which observations also generally
+support \cite{Lloyd_1991, MEDLYN_2011, Lin_2015, Zhou_2014, Zhou_2015,
+  Medlyn_2017}. However, some purely empirical results that fit the
+exponent of the VPD dependence to data have shown that it may vary
+slightly from 1/2, suggesting that stomata, as well as ecosystem-scale
+quantities based on stomata theory, may not always function optimally
+\cite{Zhou_2015, Lin_2018}. Specifically with regards to uWUE, one
+would not expect that this ecosystem scale WUE quantity will respond
+to VPD exactly analogously to stomata. Direct soil evaporation's
+contributions to ET should shift the exponent of the VPD
+dependence. \citeA{Zhou_2015}'s results corroborate this: they found a
+mean empirically fit exponential VPD dependence of 0.55, varying
+slightly from the theoretically optimal value of of 0.5 for AmeriFlux
+sites. \citeA{Lin_2018}'s results also show variance in the empirical
+exponent of the VPD dependence in the stomatal conductance model, but
+interpretation of this variance is more difficult as \citeA{Lin_2018}
+do not handle the GPP dependence of stomatal conductance in a directly
+analogous manner to the optimal theory in \citeA{MEDLYN_2011} and
+\citeA{Medlyn_2017}. Regardless, given that these recent results on
+the relationship between VPD, GPP, and ET \cite{MEDLYN_2011,
+  Zhou_2014, Zhou_2015, Medlyn_2017} form the backbone of our analysis
+and are what allowed us to derive an explicit ET expression for the
+first time (Eq. (\ref{et})), we will analyze if and how assumptions
+about the exponent of the VPD dependence impact the shape of the
+ET-VPD dependence. This analysis is also important to understand
+whether the choice of stomatal conductance model alters the
+fundamental behavior of the ET-VPD relationships, as many commonly
+used models utilize a VPD exponent other than the 1/2 suggested by
+optimal theory \cite<e.g. >[ which uses an exponent of
+1]{Leuning_1990}. It is useful to understand how the choice of model
+imposes the functional form of the ET-VPD relationship, given its
+fundamental importance for land-atmosphere interactions.
+
+\subsection{Functional form of ET dependence on VPD and its relation
+  to the VPD exponent}
+\label{functional_form}
+
+The results presented in Sect. \ref{results} indicates that for a
+given uWUE and $g_1$, the ET dependence on VPD should be concave
+upward. In other words, there should be some local minimum in ET at a
+critical VPD$_{crit}$, assuming the scaling and plant terms
+(e.g. aerodynamic conductance, $\Delta$, $g_1$ and uWUE) are held
+fixed. This result warrants further investigation, because to our
+knowledge no earlier work has derived the theoretical ecosystem-scale
+relationship between ET and VPD in an energy balance framework, while
+controlling for other environmental conditions. In particular there is
+an apparent lack of consensus over whether the shape of the ET-VPD
+curve should be concave upward (our result) or concave downward in the
+absence of dramatic water stress. Given that understanding the ET-VPD
+relationship of the one-way plant response is fundamental to
+hypothesizing about any feedbacks between the land surface and the
+atmosphere, we analyze why our derived ET-VPD relationship is concave
+upward, particularly with respect to the exponent of VPD dependence in
+uWUE and the Medlyn unified stomatal conductance model, as other
+models and empirical results have suggested exponents varying from 1/2
+\cite{Leuning_1990, Zhou_2015, Lin_2018}.
+
+By introducing $n$, the exponent of VPD in uWUE, and $m$, the exponent
+of VPD in the stomatal conductance model, we can free our theory from
+assumptions about VPD dependence:
+% \begin{linenomath*}
+  \begin{equation}
+    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g*}{VPD^m}\right) \frac{*WUE \; ET}{c_a \; VPD^n},
+    \label{m_n}
+  \end{equation}
+% \end{linenomath*}
+where:
+\[*WUE = \frac{GPP}{ET}VPD^n,\] and $g*$ is a generic slope parameter
+of units $VPD^m$. To determine how the exponent $n$ and $m$ alter the
+shape of the ET-VPD dependence we find the roots of the second
+derivative of ET, using Eq. (\ref{m_n}) for stomatal conductance
+($g_s$), with respect to VPD:
+
+\input{d2_solutions.tex}
+
+With this result we have defined the family of curves separating
+concave up from concave down ET solutions (Fig. \ref{concave}). These
+curves are only functions of the exponent of the VPD dependence and a
+quantity we call non-dimensional VPD ($VPD^m/g*$). Several important
+relations reveal themselves from Eq. (\ref{curves}):
+
+\begin{figure}
+  \centering
+  \centerline{\includegraphics[width=0.75\textwidth]{./concave.pdf}}
+  \caption{ Solutions corresponding to inflection points between
+    concave up and concave down ET-VPD curves (Eq. (\ref{curves}))
+    for three specific scenarios. Solutions are defined in terms of a
+    non-dimensional VPD ($VPD^m/g_*$), but to aide physical
+    interpretation the horizontal axis is additionally provided in terms of dimensionalized VPD assuming $m=1/2$ and
+    $g_*=110\; Pa^{1/2}$ (average of all PFT $g_1$). The vertical axis
+    has a different interpretation depending on the solution
+    curve. For the blue line ($m$ varying), it corresponds to $m$,
+    for the orange line ($n$ varying) it corresponds to $n$, and for
+    the green line it corresponds to the value of both $n$ and $m$
+    ($n=m$). Regions of the parameter space that correspond to
+    concave up and concave down results are labeled.}
+  \label{concave}
+\end{figure}
+
+\begin{itemize}
+  \item For optimal behavior (n, m = 1/2) the ET-VPD curve will be
+    concave up regardless of the magnitude of the plant constants
+    $g_1$ and uWUE. Therefore, the general concave up nature of our
+    results, given an assumption of optimal behavior, is insensitive
+    to plant type.
+  \item For all physically possible exponents of VPD dependence ($n,
+    m$), whether the solution is concave up or concave down does not
+    depend on $*WUE$.
+  \item In general, increasing the exponent of VPD dependence for
+    either $*WUE$ or $g_*$ increases the likelihood of a concave down
+    result. Additionally, as the exponent of VPD dependence increases
+    from the optimum value of 1/2, whether the curve is concave upward
+    or concave downward becomes a function of the plant specific slope
+    parameter $g_*$, through non-dimensional VPD
+    ($VPD^m/g_*$). Because the exponent of the VPD dependencies is
+    capable of altering the fundamental shape of ET-VPD dependence,
+    future research investment in understanding the exact VPD
+    dependence of stomatal conductance, and further reconciliation of
+    empirical and theoretical stomatal and ecosystem behavior should
+    be prioritized.
+\end{itemize}
+
+While it is possible (and even likely) that in the future some other
+form of VPD dependence is derived, at present \cite{MEDLYN_2011} and
+\cite{Zhou_2014} established n=m=1/2 as the most likely candidate
+given current theory and empirical data. Additionally, we argue that a
+concave up result matches physical intuition more than a concave down
+result. Plants must maintain nutrient and sugar transport through the
+phloem and xylem. To accomplish this, stomata must remain slightly
+open \cite{De_2013, Nikinmaa_2013, Ryan_2014}. Furthermore, even if
+complete stomatal closure were possible, cuticular water loss and [at
+the ecosystem-scale] direct soil evaporation are still sources of ET
+which increase with VPD, independent of stomatal closure. Therefore,
+in the limit as VPD becomes large and we assume plants are exercising
+all strategies to reduce ET, any further increase in VPD should result
+in an increase in ET through cuticular water loss and/or direct soil
+evaporation. This inevitable transition from conditions when stomata
+respond strongly to VPD to conditions when stomata response is
+asymptoting towards full closure would cause a concave up ET-VPD
+curve, which is matched by our theory. In short, plant response
+becomes more limited as VPD increases, while atmospheric demand
+monotonically increases with VPD, leading to the result that
+atmospheric demand dominates plant response when atmospheric demand is
+high.
+
+This analysis allows us to understand the theoretical shape of the ET
+response to VPD with environmental conditions held
+fixed. Accomplishing this with purely statistical methods applied to
+flux observations would be very difficult, given the relatively fast
+time scale of plant response and the non-stationarity of [solar
+forced] environmental conditions over the relatively coarse (half
+hourly) flux estimates (which is required to obtain robust
+eddy-covariance statistics). Our results on the shape of the ET-VPD
+curve with environmental conditions held fixed can be built upon with
+future work examining how changes in VPD and environmental conditions
+(e.g. soil water storage) feedback upon one another. In the soil water
+storage example, over very long time scales extremely high VPD
+perturbations coupled with no precipitation could result in decreases
+in soil water storage such that water becomes limiting. This could be
+represented by an extension of our framework in which $\lambda$ is
+allowed to decrease with decreasing soil water, increasing uWUE and
+decreasing g$_1$. Here, we focus our results by framing them with
+previous literature to build baseline intuition for ET-VPD dependence.
 
@@
 
-\input{conclusions.tex}
+We derived a new form of Penman Monteith using the concept of
+semi-empirical optimal stomatal regulation \cite{Lin_2015,
+MEDLYN_2011} and near constant uWUE \cite{Zhou_2015} to remove the
+implicit dependence of stomatal conductance on GPP and ET. With our
+new form of Penman Monteith we developed a theory for when an
+ecosystem will tend to reduce or increase ET with increasing VPD,
+which we framed using previous literature exploring the relationship
+between plant parameters, plant types, and climate. The goal was to
+understand the range of possible ET responses to VPD and develop some
+intuition for how the ET response may vary with plant types and
+climate. This intuition can be used to disentangle land atmosphere
+feedbacks in more complicated scenarios, and will aid interpretation
+of observations and more sophisticated models.
+
+ET response to VPD can vary from strongly water conservative (ET
+decreasing in response to increasing VPD), to strongly water intensive
+(ET increasing in response to VPD), which is indicative of the
+diversity of possible plant water conservation strategies. Higher uWUE
+and g$_1$ values increase the likelihood of a positive ET response,
+while decreasing temperature and increasing g$_a$ amplify the
+magnitude of the response. Previous literature \cite{Zhou_2015,
+Lin_2015} suggests that crops, through association with higher g$_1$
+and uWUE, are more likely to exhibit a positive ET response to
+VPD. Shrubs (lower uWUE), C4 plants (lower g$_1$), gymnosperm trees
+(lower g$_1$), and plants in arctic and boreal climates (lower g$_1$)
+are more likely to exhibit a negative ET response to VPD. However
+interpretation of g$_1$-induced variability is partially muddied by
+ambiguity in applying leaf-scale estimations of g$_1$ to the ecosystem
+scale \cite{Medlyn_2017}.
+
+Our paper builds intuition for how plants respond to VPD
+perturbations. We show that given optimal stomatal function and fixed
+environmental conditions, the ET-VPD dependence is theoretically
+concave upward, with ET increasing with increasing VPD as VPD
+increases past some critical value where $\frac{\partial \;
+ET}{\partial \; VPD} = 0$. However future research should focus on
+fully understanding the functional form of VPD dependence, as this
+concave up result is sensitive to the exponent of VPD dependence,
+which we currently believe is 1/2 for both uWUE and the stomatal
+conductance model \cite{MEDLYN_2011, Zhou_2014}. Indeed, this
+sensitivity to the exponent of VPD dependence is an important result
+itself: land surface models, including those used in earth system
+models for climate forecasts, employ different assumptions about the
+exponent of VPD dependence in stomatal conductance
+\cite<e.g.,>[]{Ball_1987, Leuning_1990, MEDLYN_2011}, and these
+assumptions can fundamentally change the relationship between ET and
+VPD from one that is concave upward (local minimum in ET) to one that
+is concave downward (local maximum in ET).
+
+Our results are also applicable to understanding the impact of
+expected increases in VPD induced by global change.  Plant
+physiological responses to direct CO$_2$ effects
+\cite<e.g.,>[]{Swann_2016, Lemordant_2018} receives more attention
+than physiological response to indirect effects like increased VPD
+\cite{Novick_2016}. Here, we provide a framework for understanding ET
+response to VPD using a simplified model of two plant
+parameters. Feedbacks between the land and the atmosphere may alter
+the net response to a long-timescale global VPD perturbation, but our
+focus on the one way plant response to a VPD perturbation in the
+atmospheric boundary layer is an important first step to disentangling
+such feedbacks, both in observations and model simulations of the
+present and future. By removing Penman Monteith's dependence on
+implicit relationships between GPP, VPD, and ET, we allow for explicit
+future analysis of plant-VPD feedbacks in the atmospheric boundary
+layer (Eq. (\ref{et})). Our approach can be extended to examine
+varying plant response to more nuanced consideration of plant type and
+climate. Any plant physiological heterogeneity or feedback that can be
+conceptualized with shifts in $g_1$ \cite<e.g.>[]{Lin_2015,
+Medlyn_2017} and/or uWUE \cite<e.g.>[]{Zhou_2014} are representable
+within our framework, which opens the door for a hierarchy of more
+sophisticated climate- and plant-specific analyses of ET sensitivity
+to environmental variables (including VPD). We argue that such
+simplified conceptual frameworks are critical tools for disentangling
+land-atmosphere feedbacks at various scales, from diurnal to seasonal
+and beyond, and to characterize ET response in a warmer,
+atmospherically drier, and enriched CO$_2$ world.
 

### 2019-09-04 doc/paper/intro.tex
started actual work on revisions, but note for latexdiff to work all edited content must be in same file

@@
 evolved to respond to VPD, but also excludes many details and special
-cases (e.g. plant to plant interaction and highly specialized
+cases \cite<e.g. plant to plant interaction and highly specialized
 photosynthesis strategies like Crassulacean acid metabolism
-photosynthesis).
+photosynthesis>[]{Brooker_2006, Cushman_2001}.
 
@@
 framework for understanding plant responses to atmospheric drying and
-the the VPD dependence of ET while keeping other variables fixed. This
+the VPD dependence of ET while keeping other variables fixed. This
 framework will aid interpretation of observations, full complexity

### 2019-09-04 doc/paper/agu-james-submission.tex
done with response to review 1

@@
     g_{1}  \propto \sqrt{\frac{3 \, \Gamma^* \,
-        \lambda}{1.6}},\footnote{Note this expression has units of of
-      (mmol mol$^{-1}$)$^{1/2}$, but this can be converted to
-      Pa$^{1/2}$ using the ideal gas law.}
+        \lambda}{1.6}},
     \label{slope}
@@
 
-where $\Gamma^*$ is the CO$_2$ compensation point for photosynthesis
-(without dark respiration), and $\lambda$ is the marginal water cost
-of leaf carbon
-($\frac{\partial \; \text{transpiration}}{\partial \; CO_2 assimilation}$)
-\cite{Farquhar_1980, Katul_2009}. So,
-$g_{1}$ is a leaf-scale term reflecting the trade-off of water for
-carbon uptake. The higher $g_{1}$, the more open the stomata and
-the more they release water in exchange for carbon.
+  where g$_1$ is presented units of $\sqrt{\text{mol fraction}}$ for
+  convenience (which can be converted to $\sqrt{Pa}$ using the ideal
+  gas law and ambient pressure), $\Gamma^*$ is the CO$_2$ compensation
+  point for photosynthesis (without dark respiration), and $\lambda$
+  is the marginal water cost of leaf carbon
+  ($\frac{\partial \; \text{transpiration}}{\partial \; CO_2
+    assimilation}$) \cite{Farquhar_1980, Katul_2009}. So, $g_{1}$ is a
+  leaf-scale term reflecting the trade-off of water for carbon
+  uptake. The higher $g_{1}$, the more open the stomata and the more
+  they release water in exchange for carbon.
 
@@
 
-  providing analytical framework for ecosystem response to
-  atmospheric demand with environmental conditions held
-  fixed. There are a few subtleties to taking the derivative in
-  Eq. (\ref{d_et}): $\Delta$ ($\frac{d e_{s}}{d T}$) and $VPD$ are
-  functionally related, so while taking the derivative we evaluate
+  providing analytical framework for ecosystem response to atmospheric
+  demand with environmental conditions held fixed. There are a few
+  subtleties to taking the derivative in Eq. (\ref{d_et}): $\Delta$
+  ($\frac{d e_{s}}{d T}$) and $VPD$ are functionally related, so while
+  taking the derivative we evaluate
   $\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET}
@@
       fixed}}$. $RH$ and $e_s$ are assumed to be approximately
-  independent, which is supported by data (not shown).
+  independent, which is supported by data (see supplemental material).
 
@@
   additional assumption in the formulation of uWUE \cite{Zhou_2014,
-    Zhou_2015} and \citeA{Medlyn_2017}'s stomatal conductance model is
-  that direct soil evaporation contributions to ET remain small
-  relative to transpiration. Again, this should be more true during
-  the growing season. The ratio of evaporation to transpiration may
-  increase immediately after rainfall events due to high soil moisture
-  and ponding, but VPD is generally low anyways during these
-  times. However, some plant types allow for systematically larger
-  contributions of evaporation in ET, particularly those with sparse
-  canopies and smaller relative amounts of transpiration. We therefore
-  might expect that the theory will be most applicable to forest PFTs,
-  which will be most strongly coupled to the boundary layer due to
-  larger surface roughness, and will also generally have the highest
-  ratios of transpiration to evaporation. Finally, we assume that
-  $g_1$ and uWUE are constant with respect to the conceptual VPD
-  perturbation. Both quantities have been shown to be relatively
-  constant with respect to changes in VPD \cite{Franks_2017,
-    Zhou_2014}. These parameters will however vary with plant species
-  and characteristics \cite<e.g. wood density, >[]{Lin_2015}, as well
-  as environmental conditions including soil water content and
-  temperature \cite{Lin_2015, Manzoni2013}. Exploring possible soil
-  moisture (in)dependence of the plant parameters (g$_1$ and uWUE) is
-  particularly interesting because soil moisture only enters the
-  partial derivative directly through these plant parameters. If the
-  plant parameters are weak functions of soil moisture then the theory
-  can be directly applied to a broader range of conceptual VPD
-  scenarios, including observed compound events between high VPD and
-  low soil moisture \cite{Zhou_2019}. Supplementary material for this
-  manuscript explores the soil moisture dependence of the plant
-  parameters, but excessive noise and inconsistencies preclude
-  conclusions about the nature of any dependence. We provide it in
-  case it is of interest to the reader, and to motivate future
-  research.
-
-  Because Eq. (\ref{d_et}) is a partial derivative, it's utility is as
+    Zhou_2015} and the stomatal conductance model of
+  \citeA{Medlyn_2017} is that direct soil evaporation contributions to
+  ET remain small relative to transpiration. Again, this should be
+  more true during the growing season. The ratio of evaporation to
+  transpiration may increase immediately after rainfall events due to
+  high soil moisture, ponding, and interception, but VPD is generally
+  low anyways during these times. However, some plant types allow for
+  systematically larger contributions of evaporation in ET,
+  particularly those with sparse canopies and smaller relative amounts
+  of transpiration. We therefore might expect that the theory will be
+  most applicable to forest PFTs, which will be most strongly coupled
+  to the boundary layer due to larger surface roughness, and will also
+  generally have the highest ratios of transpiration to
+  evaporation. Finally, we assume that $g_1$ and uWUE are constant
+  with respect to the conceptual VPD perturbation. Both quantities
+  have been shown to be relatively constant with respect to changes in
+  VPD \cite{Franks_2017, Zhou_2014}. These parameters will however
+  vary with plant species and characteristics \cite<e.g. wood density,
+  >[]{Lin_2015}, as well as environmental conditions including soil
+  water content and temperature \cite{Lin_2015,
+    Manzoni2013}. Exploring possible soil moisture (in)dependence of
+  the plant parameters (g$_1$ and uWUE) is particularly interesting
+  because soil moisture only enters the partial derivative directly
+  through these plant parameters. If the plant parameters are weak
+  functions of soil moisture then the theory can be directly applied
+  to a broader range of conceptual VPD scenarios, including observed
+  compound events between high VPD and low soil moisture
+  \cite{Zhou_2019}. Supplementary material for this manuscript
+  explores the soil moisture dependence of the plant parameters, but
+  excessive noise and inconsistencies preclude conclusions about the
+  nature of any dependence. We provide it in case it is of interest to
+  the reader, and to motivate future research.
+
+  Because Eq. (\ref{d_et}) is a partial derivative, its utility is as
   a conceptual model for the change in ET in response to VPD with all
@@
 
-\citeA{Medlyn_2017}'s ecosystem-scale results did not display the
+The ecosystem-scale results of \citeA{Medlyn_2017} did not display the
 robust relationship between plant type and and plant parameter (g$_1$)
 that was demonstrated by \citeA{Zhou_2015} for uWUE. Given that there
-is a robust relationship between g$_1$ and plant type in
-\citeA{Lin_2015}'s leaf-scale results, and that \citeA{Zhou_2015}'s ecosystem-scale results
-rely on the same underlying theory as \citeA{Medlyn_2017}'s results,
-the ambiguous results in \citeA{Medlyn_2017} could be due to noise in
-ecosystem scale observations, the consequences of imposing a model
-structure in some of their g$_1$ estimations, or real ecosystem-scale
-within-plant type variability (e.g. multiple species impacting the
-relationship within the eddy-covariance footprint). These differences
-between leaf-scale behavior and ecosystem scale behavior highlight the
-importance of understanding how and why slope parameters show
-increased variability at the ecosystem scale, as argued by
-\citeA{Medlyn_2017}. However, if we assume that there is some real
-analogy between leaf scale results in \citeA{Lin_2015} and ecosystem
-scale behavior, and that some of the ambiguity in \citeA{Medlyn_2017}
-is due to model and/or observational error, then we can use the
-relationships between g$_1$ and plant characteristics defined by
-\citeA{Lin_2015} to frame the ET response to VPD. This is subject to
-the strong caveat that we still do not fully understand g$_1$ behavior
-at the ecosystem scale. Extrapolating \citeA{Lin_2015}'s results
-(e.g. Fig. 2) to the ecosystem scale would reveal the following
-relationships between g$_1$ and plant types:
+is a robust relationship between g$_1$ and plant type in the
+leaf-scale results of \citeA{Lin_2015}, and the ecosystem-scale
+results of \citeA{Zhou_2015} rely on the same underlying theory as
+\citeA{Medlyn_2017}, the ambiguous results in \citeA{Medlyn_2017}
+could be due to noise in ecosystem scale observations, the
+consequences of imposing a model structure in some of their g$_1$
+estimations, or real ecosystem-scale within-plant type variability
+(e.g. multiple species impacting the relationship within the
+eddy-covariance footprint). These differences between leaf-scale
+behavior and ecosystem scale behavior highlight the importance of
+understanding how and why slope parameters show increased variability
+at the ecosystem scale, as argued by \citeA{Medlyn_2017}. However, if
+we assume that there is some real analogy between leaf scale results
+in \citeA{Lin_2015} and ecosystem scale behavior, and that some of the
+ambiguity in \citeA{Medlyn_2017} is due to model and/or observational
+error, then we can use the relationships between g$_1$ and plant
+characteristics defined by \citeA{Lin_2015} to frame the ET response
+to VPD. This is subject to the strong caveat that we still do not
+fully understand g$_1$ behavior at the ecosystem scale. Extrapolating
+the results of \citeA{Lin_2015} (e.g. Fig. 2) to the ecosystem scale
+would reveal the following relationships between g$_1$ and plant
+types:
 
@@
 contributions to ET should shift the exponent of the VPD
-dependence. \citeA{Zhou_2015}'s results corroborate this: they found a
-mean empirically fit exponential VPD dependence of 0.55, varying
-slightly from the theoretically optimal value of of 0.5 for AmeriFlux
-sites. \citeA{Lin_2018}'s results also show variance in the empirical
-exponent of the VPD dependence in the stomatal conductance model, but
-interpretation of this variance is more difficult as \citeA{Lin_2018}
-do not handle the GPP dependence of stomatal conductance in a directly
-analogous manner to the optimal theory in \citeA{MEDLYN_2011} and
-\citeA{Medlyn_2017}. Regardless, given that these recent results on
-the relationship between VPD, GPP, and ET \cite{MEDLYN_2011,
-  Zhou_2014, Zhou_2015, Medlyn_2017} form the backbone of our analysis
-and are what allowed us to derive an explicit ET expression for the
-first time (Eq. (\ref{et})), we will analyze if and how assumptions
-about the exponent of the VPD dependence impact the shape of the
-ET-VPD dependence. This analysis is also important to understand
-whether the choice of stomatal conductance model alters the
+dependence. The results from \citeA{Zhou_2015} corroborate this: they
+found a mean empirically fit exponential VPD dependence of 0.55,
+varying slightly from the theoretically optimal value of of 0.5 for
+AmeriFlux sites. Results in \citeA{Lin_2018} also show variance in the
+empirical exponent of the VPD dependence in the stomatal conductance
+model, but interpretation of this variance is more difficult as
+\citeA{Lin_2018} do not handle the GPP dependence of stomatal
+conductance in a directly analogous manner to the optimal theory in
+\citeA{MEDLYN_2011} and \citeA{Medlyn_2017}. Regardless, given that
+these recent results on the relationship between VPD, GPP, and ET
+\cite{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017} form the
+backbone of our analysis and are what allowed us to derive an explicit
+ET expression for the first time (Eq. (\ref{et})), we will analyze if
+and how assumptions about the exponent of the VPD dependence impact
+the shape of the ET-VPD dependence. This analysis is also important to
+understand whether the choice of stomatal conductance model alters the
 fundamental behavior of the ET-VPD relationships, as many commonly
@@
   \caption{ Solutions corresponding to inflection points between
-    concave up and concave down ET-VPD curves (Eq. (\ref{curves}))
-    for three specific scenarios. Solutions are defined in terms of a
+    concave up and concave down ET-VPD curves (Eq. (\ref{curves})) for
+    three specific scenarios. Solutions are defined in terms of a
     non-dimensional VPD ($VPD^m/g_*$), but to aide physical
-    interpretation the horizontal axis is additionally provided in terms of dimensionalized VPD assuming $m=1/2$ and
+    interpretation the horizontal axis is additionally provided in
+    terms of dimensionalized VPD assuming $m=1/2$ and
     $g_*=110\; Pa^{1/2}$ (average of all PFT $g_1$). The vertical axis
     has a different interpretation depending on the solution
-    curve. For the blue line ($m$ varying), it corresponds to $m$,
-    for the orange line ($n$ varying) it corresponds to $n$, and for
-    the green line it corresponds to the value of both $n$ and $m$
-    ($n=m$). Regions of the parameter space that correspond to
-    concave up and concave down results are labeled.}
+    curve. For the blue line ($m$ varying), it corresponds to $m$, for
+    the orange line ($n$ varying) it corresponds to $n$, and for the
+    green line it corresponds to the value of both $n$ and $m$
+    ($n=m$). Regions of the parameter space that correspond to concave
+    up and concave down results are labeled: for each curve, the
+    parameter space ``below'' the line corresponds to a concave up
+    relationship, which the parameter space ``above'' the line
+    corresponds to a concave down relationship.}
   \label{concave}
@@
 open \cite{De_2013, Nikinmaa_2013, Ryan_2014}. Furthermore, even if
-complete stomatal closure were possible, cuticular water loss and [at
-the ecosystem-scale] direct soil evaporation are still sources of ET
+complete stomatal closure were possible, cuticular water loss and (at
+the ecosystem-scale) direct soil evaporation are still sources of ET
 which increase with VPD, independent of stomatal closure. Therefore,
@@
 flux observations would be very difficult, given the relatively fast
-time scale of plant response and the non-stationarity of [solar
-forced] environmental conditions over the relatively coarse (half
+time scale of plant response and the non-stationarity of (solar
+forced) environmental conditions over the relatively coarse (half
 hourly) flux estimates (which is required to obtain robust
@@
 is based upon work supported by the National Science Foundation
-Graduate Research Fellowship under Grant No. DGE 16-44869.
+Graduate Research Fellowship under Grant No. DGE 16-44869. We would
+like to thank Emma Robinson and one anonymous reviewer for thoughtful
+comments that greatly improved the strength of the manuscript.
 

### 2019-09-04 doc/paper/james-supplement.tex
done with response to review 1

--- 2019-09-03 doc/paper/james-supplement-original-submision.tex
@@
 
+Additionally, we include a figure showing the joint distribution
+between saturation vapor pressure and relative humidity calculated
+from the FLUXNET-2015 data. Relative humidity and saturation vapor
+pressure are much more independent than saturation vapor pressure and
+VPD, and we use an assumption that relative humidity and saturation
+vapor pressure can be approximated as independent in order to evaluate
+$\frac{\partial ET}{\partial VPD}$. Please note that at a given site,
+the relationship may be more or less independent depending on the
+hydroclimate.
+
 \section{Description of data}
@@
 
+\begin{figure}
+  \centering \includegraphics{./supp-figs/0joint_rh_es.pdf}
+  \caption{The joint distribution of relative humidity and saturation
+    vapor pressure for the FLUXNET2015 dataset.}
+  \end{figure}
+
 \section{FLUXNET2015 sites}

### 2019-09-04 src/paper_figure_code/fully_idealized_figure.py

@@ -6,7 +6,7 @@ from shared_functions import *
 import matplotlib.lines as mlines
 
 # mean_df defined her at module-level
-linewidth=2.5
+linewidth=0.1 # for 0 line
 dashedlinewidth=3.0
 markersize=8
 
@@ -99,7 +99,7 @@ add_plots(df, ax)
 ax.set_xlabel('VPD (Pa)', fontsize=single_ax_fontsize)
 ax.set_ylabel(et_string, fontsize=single_ax_fontsize)
 xlims = ax.get_xlim()
-ax.plot(xlims, [0., 0.], 'k--', linewidth=linewidth)
+ax.plot(xlims, [0., 0.], 'k-', linewidth=linewidth)
 ax.set_xlim(xlims)
 color_handles = [mlines.Line2D([], [], color=c, label=r'%s: ($\frac{2 \; g_a \; P}{T(\Delta + \gamma)}$)' % p) for c,p in zip(colors, pdesc)]
 style_handles = [mlines.Line2D([], [], color='k', linestyle=l, label=(p + ": " + paren_string)) for l,p in zip(linestyles, pdesc)]
@@ -115,7 +115,7 @@ ax.set_xlabel('VPD (Pa)', fontsize=single_ax_fontsize)
 # ax.set_ylabel(paren_string, fontsize=single_ax_fontsize)
 ax.set_ylabel(et_string, fontsize=single_ax_fontsize)
 xlims = ax.get_xlim()
-ax.plot(xlims, [0., 0.], 'k--', linewidth=linewidth)
+ax.plot(xlims, [0., 0.], 'k-', linewidth=linewidth)
 ax.set_xlim(xlims)
 color_handles = [mlines.Line2D([], [], color=c, label=r'%s: g1' % p) for c,p in zip(colors, pdesc)]
 style_handles = [mlines.Line2D([], [], color='k', linestyle=l, label=(p + ": uWUE")) for l,p in zip(linestyles, pdesc)]
@@ -131,7 +131,7 @@ ax.set_xlabel('VPD (Pa)', fontsize=single_ax_fontsize)
 # ax.set_ylabel(paren_string, fontsize=single_ax_fontsize)
 ax.set_ylabel(et_string, fontsize=single_ax_fontsize)
 xlims = ax.get_xlim()
-ax.plot(xlims, [0., 0.], 'k--', linewidth=linewidth)
+ax.plot(xlims, [0., 0.], 'k-', linewidth=linewidth)
[…]

### 2019-09-05 doc/paper/agu-james-submission.tex
finished review 2 response

@@
 and atmospheric carbon can be accessed with a relatively smaller cost
-in water loss. If instead stomata were fully passive \cite<similar
-to soil pores, e.g. >[]{Or_2013}, increased atmospheric aridity would
+in water loss. If instead stomata were fully passive \cite<similar to
+soil pores, e.g. >[]{Or_2013}, increased atmospheric aridity would
 strongly reduce soil moisture \cite{Berg_2017}. This could further
@@
 \cite{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006,
-Salvucci_2013, Gentine_2016, Berg_2016, Zhou_2019}. Therefore, passive
-regulation and a lack of soil moisture conservation does not seem to
-be a sensible strategy for plants from an evolutionary
+  Salvucci_2013, Gentine_2016, Berg_2016, Zhou_2019}. Therefore,
+passive regulation and a lack of soil moisture conservation does not
+seem to be a sensible strategy for plants from an evolutionary
 standpoint. This simplified logic explains generally why plants
 evolved to respond to VPD, but also excludes many details and special
-cases \cite<e.g. plant to plant interaction and highly specialized
-photosynthesis strategies like Crassulacean acid metabolism
-photosynthesis>[]{Brooker_2006, Cushman_2001}.
+cases \cite<e.g. plant to plant interaction, thermal regulation with
+transpiration, and highly specialized photosynthesis strategies like
+Crassulacean acid metabolism photosynthesis>[]{Brooker_2006, Lin_2017,
+  Cushman_2001}.
 
@@
     environmental conditions on ET response to VPD. A citation is
-    provided in cases where the quantities are directly derived from
-    previous literature. Conceptually, min. values are extracted from
-    literature to correspond to approximately the 15th percentile of
-    observed conditions during the growing season, med. values
+    provided for values of g$_1$ and uWUE, which are directly derived
+    from previous literature. Conceptually, min. values are extracted
+    from literature to correspond to approximately the 15th percentile
+    of observed conditions during the growing season, med. values
     correspond to approximately the 50th percentile, and max. values
-    correspond to approximately the 85th percentile.}
+    correspond to approximately the 85th percentile. Values for $T$
+    and g$_a$ are calculated from FLUXNET-2015 data (see supplemental
+    material for description), rounding the 15th, 50th, and 85th
+    percentile to the nearest 5$^o$C and 0.005 m/s, respectively.}
   \label{param_varying}
@@
 value, under which the sign of the ET response depends on
-VPD. However, crops, which we expect to prioritize carbon uptake over
-water conservation, have a higher uWUE value closer to our maximum
-value. This higher uWUE results in a higher likelihood for an
+VPD. However, crops, shurbs, and grass are exceptions; they exhibit an
+uWUE different than the median value in Table
+\ref{param_varying}. Crops, which we expect to prioritize carbon
+uptake over water conservation, have a higher uWUE value closer to our
+maximum value. This higher uWUE results in a higher likelihood for an
 increasing ET response to VPD, which matches intuition given that we

### 2019-09-05 doc/paper/james-supplement.tex
finished review 2 response

@@
   humidity (\cite{Medlyn_2017}).
+\item For SWC measurements, we use the shallowest observed layer
+  available at each site.
 \end{itemize}
+
 Additionally, as in \citet{Lin_2018}, we restrict data to the daytime,
@@
 
-\input{supp-figs.tex}
-
 \begin{figure}
@@
   \end{figure}
+
+\input{supp-figs.tex}
 

### 2019-09-05 doc/paper/agu-james-submission.tex
submitting resubmission now

@@
 transpiration, and highly specialized photosynthesis strategies like
-Crassulacean acid metabolism photosynthesis>[]{Brooker_2006, Lin_2017,
+Crassulacean acid metabolism photosynthesis; see>[]{Brooker_2006, Lin_2017,
   Cushman_2001}.
@@
 
-  where g$_1$ is presented units of $\sqrt{\text{mol fraction}}$ for
+  where g$_1$ is presented in units of $\sqrt{\text{mol fraction}}$ for
   convenience (which can be converted to $\sqrt{Pa}$ using the ideal
@@
 value, under which the sign of the ET response depends on
-VPD. However, crops, shurbs, and grass are exceptions; they exhibit an
+VPD. However, crops, shrubs, and grass are exceptions; they exhibit an
 uWUE different than the median value in Table
@@
 leaf-scale results of \citeA{Lin_2015}, and the ecosystem-scale
-results of \citeA{Zhou_2015} rely on the same underlying theory as
+results of \citeA{Zhou_2015} rely on the same underlying theory as in
 \citeA{Medlyn_2017}, the ambiguous results in \citeA{Medlyn_2017}
@@
     up and concave down results are labeled: for each curve, the
-    parameter space ``below'' the line corresponds to a concave up
-    relationship, which the parameter space ``above'' the line
+    parameter space ``below'' the curve corresponds to a concave up
+    relationship, while the parameter space ``above'' the curve
     corresponds to a concave down relationship.}

### 2019-09-05 doc/paper/james-response-to-reviewers.tex
done with submission

\documentclass[12pt]{article}

\input{preamble}

\begin{document}

\large \textbf{Response to Reviewers: \textit{When does vapor pressure deficit drive or reduce
    evapotranspiration?}} \\
Adam Massmann, Pierre Gentine, and Changjie Lin \\
\today

\vspace{0.1in}

\normalsize
\section{Response to Reviewer 1}

Thank you very much for the thoughtful and thorough review. We
responded to all comments in the annotated PDF, but this is our first
time using that tool so if there are any issues please let us know. As
an aside, thanks for introducing us to the annotated PDF feature. We
were not aware of it, and it was pleasant to use.

\section{Response to Reviewer 2}

As with reviewer 1, thank you very much for the time spent on a
thoughtful review. Here are responses to your comments:

\begin{enumerate}

\item This paper actually discusses transpiration not
  evapotranspiration - while the role of soil evaporation is sometimes
  acknowledged (and then dismissed) - it would be much cleaner if
  references to evapotranspiration (ET) were replaced with
  transpiration (T). The role of soil evaporation and interception in
  a changing climate is a whole other (and difficult!) issue.

  \textit{Response}: Yes, we agree that the underlying theory was
  developed for transpiration, and that it would probably appear
  cleaner to just present the manuscript in terms of transpiration
  instead of evapotranspiration. However, we prefer to use ET, along
  with the language discussing the theory's likely limitation when
  evaporation is high relative to transpiration, for a few of
  reasons. The parameter ranges we use in Table 2 for uWUE and g1 and
  are derived from ecosystem ET measurements in Zhou et al. 2015 and
  Medlyn et al. 2017, rather than T measurements. Given the
  semi-empirical nature of the theory, and that the parameters were
  developed empirically from ET observations rather than T
  observations (and it would require the reader to dig into cited
  literature to figure this out), we believe it is more transparent to
  call the quantity ET rather than T, while including explicit
  language about the possible weaknesses of the theory when
  evaporation is large relative to transpiration. Additionally, Zhou
  et al. 2014 and Zhou et al. 2015 use "ET" rather than "T", so by
  using "ET" we are more consistent with that notation. However,
  Medlyn et al. 2017 use "E" for this quantity (their Eq. 5), but call
  it "canopy transpiration," which we actually did not prefer because
  we felt using "transpiration" obfuscated the fact that their
  estimate of $g$ was actually derived from ET measurements. So, while
  we agree that it could appear cleaner to just call the quantity T
  and remove our language discussing the theory's limitations and
  caveats with respect to evaporation, we also believe this could hide
  the fact the we use semi-empirical parameters derived for ET, and
  distance our work from the previous work we directly build upon (in
  particular Zhou et al. 2014 and 2015). While not relevant at all
  (but still helpful in the end), keeping ET rather than adopting T
  has a practical advantage that our title would not need to be
  changed, which would help match the final peer reviewed version of
  the manuscript to already publicly available pre-prints of the
  manuscript.

\item Lines 80-90. Plants also use evaporation to cool their leaves - is
  this significant? And does it suggest there should be some residual
  stomatal opening?

  \textit{Response}: Yes, some plants use transpiration to regulate
  temperature on their leaves in some conditions, and this could be a
  cause of stomatal opening/closing. We added language at lines 78-80
  acknowledging that the simplified logic in the opening excludes
  details, including how plants can use transpiration for thermal
  regulation: ``...excludes many details ... e.g., ... thermal
  regulation with transpiration ... Lin et al. 2017''

\item Line 210-211. It would be worth expanding on the RH - es
  relationship - what data?

\textit{Response}: Agreed, we added a figure to the supplemental material on
    the relationship between RH and es, and a reference to it in the text.

\item Table 4 legend. Needs a bit of clarification on the choice of ranges
  of T and ga.

  \textit{Response}: Agreed; added clarifying language to the caption
  for how T and g$_a$ ranges were estimated (FLUXNET data): ``Values
  for $T$ and g$_a$ are calculated from FLUXNET-2015 data (see
  supplemental material for description), rounding the 15th, 50th, and
  85th percentile to the nearest 5$^o$C and 0.005 m/s, respectively.''

\item Line 322-335. I found this paragraph a bit ambiguous. The second
  sentence suggests most ptfs are similar to the median type but the
  following sentences suggest this is not true. I think this is
  important - the reader needs to know whether the max and min values
  of parameters are common or outliers.

\textit{Response}: Thanks; we reworded this paragraph to make it more clear and
    explicit: "...However, crops, shrubs, and grass are exceptions; they
    exhibit an uWUE different than the median value in Table 2. Crops,
    which we expect to prioritize carbon uptake over water conservation,
    have a higher uWUE value closer to our maximum value. This higher uWUE
    results in a higher likelihood for an increasing ET response to VPD,
    which matches intuition given that we expect crops to keep stomata
    open for access to carbon, at the cost of increased water loss during
    high VPD. Shrubs, and to a lesser extent grass, have an uWUE closer to
    the minimum uWUE. For these plant types, we would then expect a
    decrease of ET in response to VPD."

\item Supplementary material. As I say above I commend the inclusion of
  this 'negative' result.  However I would like to know how soil
  water was derived - is this measured or modelled and is it at the
  surface or through the entire rooting zone?

\textit{Response}: Yes, more clarity about the source of the data is
    helpful. We added language to the supplement explicitly stating that
    SWC is taken from the FLUXNET-2015 data; specifically the shallowest
    observed layer at each site: "For SWC measurements, we use the
    shallowest observed layer available at each site."
\end{enumerate}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2019-09-05 doc/paper/agu-james-submission-single-file.tex
done with submission

--- 2019-09-05 doc/paper/agu-james-submission.tex
@@
     \hline
-    \input{param_varying.tex}
+g$_1$ & Pa$^{1/2}$ [kPa$^{1/2}$] & 63.25 [2.00] & 126.49 [4.00] & 189.74 [6.00] & Fig. 2, 7; \citeA{Medlyn_2017} \\
+uWUE & $\mu$-mol C Pa$^{0.5}$ J$^{-1}$ ET [g C hPa$^{1/2}$ kg$^{-1}$ H$_2$O] & 2.33 [6.99] & 3.17 [9.52] & 4.01 [12.05] & Table 4; \citeA{Zhou_2015} \\
+T & $^o$C & 10.00 & 20.00 & 30.00 & - \\
+g$_a$ & m/s & 0.015 & 0.035 & 0.055 & - \\
     \hline
@@
     \hline
-    \input{param_fixed.tex}
+P & Pa & 100000.00 \\
+$\gamma$ & Pa K$^{-1}$ & 64.50 \\
+R$_{air}$ & J  K$^{-1}$ kg$^{-1}$ & 288.00 \\
+c$_a$ & $\mu$ mol CO$_2$ mol$^{-1}$ air & 400.00 \\
     \hline
@@
 
-\input{d2_solutions.tex}
+\begin{equation}
+\frac{\partial^2 \; ET}{\partial \; VPD^2} = 0 \quad \forall \quad\frac{VPD^m}{g*} = \frac{m \left(m - 2 n - \sqrt{m^{2} - 4 m n + 2 m - 4 n^{2} + 4 n + 1} + 1\right)}{2 n \left(n - 1\right)} - 1.
+\label{curves}
+\end{equation}
 
@@
 
-\bibliography{references}
+% \bibliography{references}
+
+\begin{thebibliography}{}
+
+\bibitem [\protect \citeauthoryear {%
+Ball%
+, Woodrow%
+\BCBL {}\ \BBA {} Berry%
+}{%
+Ball%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {1987}}%
+}]{%
+Ball_1987}
+\APACinsertmetastar {%
+Ball_1987}%
+\begin{APACrefauthors}%
+Ball, J\BPBI T.%
+, Woodrow, I\BPBI E.%
+\BCBL {}\ \BBA {} Berry, J\BPBI A.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{1987}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {A Model Predicting Stomatal Conductance and its
+  Contribution to the Control of Photosynthesis under Different Environmental
+  Conditions} {A model predicting stomatal conductance and its contribution to
+  the control of photosynthesis under different environmental
+  conditions}.{\BBCQ}
+\newblock
+\BIn{} \APACrefbtitle {Progress in Photosynthesis Research} {Progress in
+  photosynthesis research}\ (\BPGS\ 221--224).
+\newblock
+\APACaddressPublisher{}{Springer Netherlands}.
+\newblock
+\begin{APACrefDOI} \doi{10.1007/978-94-017-0519-6\_48} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Berg%
+\ \protect \BOthers {.}}{%
+Berg%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2016}}%
+}]{%
+Berg_2016}
+\APACinsertmetastar {%
+Berg_2016}%
+\begin{APACrefauthors}%
+Berg, A.%
+, Findell, K.%
+, Lintner, B.%
+, Giannini, A.%
+, Seneviratne, S\BPBI I.%
+, van~den Hurk, B.%
+\BDBL {}Milly, P\BPBI C\BPBI D.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2016}{may}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Land{\textendash}atmosphere feedbacks amplify aridity
+  increase over land under global warming} {Land{\textendash}atmosphere
+  feedbacks amplify aridity increase over land under global warming}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Nature Climate Change}{6}{9}{869--874}.
+\newblock
+\begin{APACrefDOI} \doi{10.1038/nclimate3029} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Berg%
+, Sheffield%
+\BCBL {}\ \BBA {} Milly%
+}{%
+Berg%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2017}}%
+}]{%
+Berg_2017}
+\APACinsertmetastar {%
+Berg_2017}%
+\begin{APACrefauthors}%
+Berg, A.%
+, Sheffield, J.%
+\BCBL {}\ \BBA {} Milly, P\BPBI C\BPBI D.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2017}{jan}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Divergent surface and total soil moisture projections
+  under global warming} {Divergent surface and total soil moisture projections
+  under global warming}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Geophysical Research Letters}{44}{1}{236--244}.
+\newblock
+\begin{APACrefDOI} \doi{10.1002/2016gl071921} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Bouchet%
+}{%
+Bouchet%
+}{%
+{\protect \APACyear {1963}}%
+}]{%
+Bouchet_1963}
+\APACinsertmetastar {%
+Bouchet_1963}%
+\begin{APACrefauthors}%
+Bouchet, R\BPBI J.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{1963}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Evapotranspiration r{\'e}elle et potentielle,
+  signification climatique} {Evapotranspiration r{\'e}elle et potentielle,
+  signification climatique}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{IAHS Publ}{62}{}{134--142}.
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Brooker%
+}{%
+Brooker%
+}{%
+{\protect \APACyear {2006}}%
+}]{%
+Brooker_2006}
+\APACinsertmetastar {%
+Brooker_2006}%
+\begin{APACrefauthors}%
+Brooker, R\BPBI W.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2006}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Plant–plant interactions and environmental change}
+  {Plant–plant interactions and environmental change}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{New Phytologist}{171}{2}{271-284}.
+\newblock
+\begin{APACrefURL}
+  \url{https://nph.onlinelibrary.wiley.com/doi/abs/10.1111/j.1469-8137.2006.01752.x}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1111/j.1469-8137.2006.01752.x} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Brutsaert%
+}{%
+Brutsaert%
+}{%
+{\protect \APACyear {1999}}%
+}]{%
+Brutsaert_1999}
+\APACinsertmetastar {%
+Brutsaert_1999}%
+\begin{APACrefauthors}%
+Brutsaert, W.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{1999}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Aspects of bulk atmospheric boundary layer similarity
+  under free-convective conditions} {Aspects of bulk atmospheric boundary layer
+  similarity under free-convective conditions}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Reviews of {G}eophysics}{37}{4}{439--451}.
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Byrne%
+\ \BBA {} O'Gorman%
+}{%
+Byrne%
+\ \BBA {} O'Gorman%
+}{%
+{\protect \APACyear {2013}}%
+}]{%
+Byrne_2013}
+\APACinsertmetastar {%
+Byrne_2013}%
+\begin{APACrefauthors}%
+Byrne, M\BPBI P.%
+\BCBT {}\ \BBA {} O'Gorman, P\BPBI A.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2013}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Land-Ocean Warming Contrast over a Wide Range of
+  Climates: Convective Quasi-Equilibrium Theory and Idealized Simulations}
+  {Land-ocean warming contrast over a wide range of climates: Convective
+  quasi-equilibrium theory and idealized simulations}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Journal of Climate}{26}{12}{4000-4016}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1175/JCLI-D-12-00262.1}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1175/JCLI-D-12-00262.1} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Cowan%
+\ \BBA {} Farquhar%
+}{%
+Cowan%
+\ \BBA {} Farquhar%
+}{%
+{\protect \APACyear {1977}}%
+}]{%
+Cowan_1977}
+\APACinsertmetastar {%
+Cowan_1977}%
+\begin{APACrefauthors}%
+Cowan, I.%
+\BCBT {}\ \BBA {} Farquhar, G.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{1977}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Stomatal function in relation to leaf metabolism and
+  environment.} {Stomatal function in relation to leaf metabolism and
+  environment.}{\BBCQ}
+\newblock
+\BIn{} \APACrefbtitle {Symposia of the Society for Experimental Biology}
+  {Symposia of the society for experimental biology}\ (\BVOL~31, \BPG~471).
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Cushman%
+}{%
+Cushman%
+}{%
+{\protect \APACyear {2001}}%
+}]{%
+Cushman_2001}
+\APACinsertmetastar {%
+Cushman_2001}%
+\begin{APACrefauthors}%
+Cushman, J\BPBI C.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2001}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Crassulacean Acid Metabolism. A Plastic Photosynthetic
+  Adaptation to Arid Environments} {Crassulacean acid metabolism. a plastic
+  photosynthetic adaptation to arid environments}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Plant Physiology}{127}{4}{1439--1448}.
+\newblock
+\begin{APACrefURL} \url{http://www.plantphysiol.org/content/127/4/1439}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1104/pp.010818} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Damour%
+, Simonneau%
+, Cochard%
+\BCBL {}\ \BBA {} Urban%
+}{%
+Damour%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2010}}%
+}]{%
+Damour2010}
+\APACinsertmetastar {%
+Damour2010}%
+\begin{APACrefauthors}%
+Damour, G.%
+, Simonneau, T.%
+, Cochard, H.%
+\BCBL {}\ \BBA {} Urban, L.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2010}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {An overview of models of stomatal conductance at the
+  leaf level} {An overview of models of stomatal conductance at the leaf
+  level}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Plant, Cell \& Environment}{33}{9}{1419-1438}.
+\newblock
+\begin{APACrefURL}
+  \url{https://onlinelibrary.wiley.com/doi/abs/10.1111/j.1365-3040.2010.02181.x}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1111/j.1365-3040.2010.02181.x} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+De~Kauwe%
+, Medlyn%
+, Knauer%
+\BCBL {}\ \BBA {} Williams%
+}{%
+De~Kauwe%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2017}}%
+}]{%
+De_2017}
+\APACinsertmetastar {%
+De_2017}%
+\begin{APACrefauthors}%
+De~Kauwe, M\BPBI G.%
+, Medlyn, B\BPBI E.%
+, Knauer, J.%
+\BCBL {}\ \BBA {} Williams, C\BPBI A.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2017}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Ideas and perspectives: how coupled is the vegetation to
+  the boundary layer?} {Ideas and perspectives: how coupled is the vegetation
+  to the boundary layer?}{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Biogeosciences}{14}{19}{4435}.
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+De~Schepper%
+, De~Swaef%
+, Bauweraerts%
+\BCBL {}\ \BBA {} Steppe%
+}{%
+De~Schepper%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2013}}%
+}]{%
+De_2013}
+\APACinsertmetastar {%
+De_2013}%
+\begin{APACrefauthors}%
+De~Schepper, V.%
+, De~Swaef, T.%
+, Bauweraerts, I.%
+\BCBL {}\ \BBA {} Steppe, K.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2013}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Phloem transport: a review of mechanisms and controls}
+  {Phloem transport: a review of mechanisms and controls}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Journal of Experimental Botany}{64}{16}{4839--4850}.
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+G.~Farquhar%
+}{%
+G.~Farquhar%
+}{%
+{\protect \APACyear {1978}}%
+}]{%
+Farquhar_1978}
+\APACinsertmetastar {%
+Farquhar_1978}%
+\begin{APACrefauthors}%
+Farquhar, G.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{1978}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Feedforward responses of stomata to humidity}
+  {Feedforward responses of stomata to humidity}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Functional Plant Biology}{5}{6}{787--800}.
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+G\BPBI D.~Farquhar%
+, von Caemmerer%
+\BCBL {}\ \BBA {} Berry%
+}{%
+G\BPBI D.~Farquhar%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {1980}}%
+}]{%
+Farquhar_1980}
+\APACinsertmetastar {%
+Farquhar_1980}%
+\begin{APACrefauthors}%
+Farquhar, G\BPBI D.%
+, von Caemmerer, S.%
+\BCBL {}\ \BBA {} Berry, J\BPBI A.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{1980}{Jun}{01}.
+\newblock
+{\BBOQ}\APACrefatitle {A biochemical model of photosynthetic {CO}2 assimilation
+  in leaves of {C}3 species} {A biochemical model of photosynthetic {CO}2
+  assimilation in leaves of {C}3 species}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Planta}{149}{1}{78--90}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1007/BF00386231} \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1007/BF00386231} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Franks%
+, Berry%
+, Lombardozzi%
+\BCBL {}\ \BBA {} Bonan%
+}{%
+Franks%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2017}}%
+}]{%
+Franks_2017}
+\APACinsertmetastar {%
+Franks_2017}%
+\begin{APACrefauthors}%
+Franks, P\BPBI J.%
+, Berry, J\BPBI A.%
+, Lombardozzi, D\BPBI L.%
+\BCBL {}\ \BBA {} Bonan, G\BPBI B.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2017}{apr}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Stomatal Function across Temporal and Spatial Scales:
+  Deep-Time Trends, Land-Atmosphere Coupling and Global Models} {Stomatal
+  function across temporal and spatial scales: Deep-time trends,
+  land-atmosphere coupling and global models}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Plant Physiology}{174}{2}{583--602}.
+\newblock
+\begin{APACrefDOI} \doi{10.1104/pp.17.00287} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Gentine%
+, Chhang%
+, Rigden%
+\BCBL {}\ \BBA {} Salvucci%
+}{%
+Gentine%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2016}}%
+}]{%
+Gentine_2016}
+\APACinsertmetastar {%
+Gentine_2016}%
+\begin{APACrefauthors}%
+Gentine, P.%
+, Chhang, A.%
+, Rigden, A.%
+\BCBL {}\ \BBA {} Salvucci, G.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2016}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Evaporation estimates using weather station data and
+  boundary layer theory} {Evaporation estimates using weather station data and
+  boundary layer theory}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Geophysical Research Letters}{43}{22}{11,661--11,670}.
+\newblock
+\begin{APACrefURL} \url{http://dx.doi.org/10.1002/2016GL070819}
+  \end{APACrefURL}
+\newblock
+\APACrefnote{2016GL070819}
+\newblock
+\begin{APACrefDOI} \doi{10.1002/2016GL070819} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+P.~Jarvis%
+}{%
+P.~Jarvis%
+}{%
+{\protect \APACyear {1984}}%
+}]{%
+Jarvis_1984}
+\APACinsertmetastar {%
+Jarvis_1984}%
+\begin{APACrefauthors}%
+Jarvis, P.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{1984}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Coupling of transpiration to the atmosphere in
+  horticultural crops: the omega factor} {Coupling of transpiration to the
+  atmosphere in horticultural crops: the omega factor}.{\BBCQ}
+\newblock
+\BIn{} \APACrefbtitle {I International Symposium on Water Relations in Fruit
+  Crops 171} {I international symposium on water relations in fruit crops 171}\
+  (\BPGS\ 187--206).
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+P\BPBI G.~Jarvis%
+\ \BBA {} McNaughton%
+}{%
+P\BPBI G.~Jarvis%
+\ \BBA {} McNaughton%
+}{%
+{\protect \APACyear {1986}}%
+}]{%
+Jarvis_1986}
+\APACinsertmetastar {%
+Jarvis_1986}%
+\begin{APACrefauthors}%
+Jarvis, P\BPBI G.%
+\BCBT {}\ \BBA {} McNaughton, K.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{1986}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Stomatal control of transpiration: scaling up from leaf
+  to region} {Stomatal control of transpiration: scaling up from leaf to
+  region}.{\BBCQ}
+\newblock
+\BIn{} \APACrefbtitle {Advances in Ecological Research} {Advances in ecological
+  research}\ (\BVOL~15, \BPGS\ 1--49).
+\newblock
+\APACaddressPublisher{}{Elsevier}.
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+G.~Katul%
+, Manzoni%
+, Palmroth%
+\BCBL {}\ \BBA {} Oren%
+}{%
+G.~Katul%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2010}}%
+}]{%
+Katul_2010}
+\APACinsertmetastar {%
+Katul_2010}%
+\begin{APACrefauthors}%
+Katul, G.%
+, Manzoni, S.%
+, Palmroth, S.%
+\BCBL {}\ \BBA {} Oren, R.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2010}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {A stomatal optimization theory to describe the effects
+  of atmospheric {CO}2 on leaf photosynthesis and transpiration} {A stomatal
+  optimization theory to describe the effects of atmospheric {CO}2 on leaf
+  photosynthesis and transpiration}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Annals of Botany}{105}{3}{431-442}.
+\newblock
+\begin{APACrefURL} \url{+ http://dx.doi.org/10.1093/aob/mcp292}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1093/aob/mcp292} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+G\BPBI G.~Katul%
+, Palmroth%
+\BCBL {}\ \BBA {} Oren%
+}{%
+G\BPBI G.~Katul%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2009}}%
+}]{%
+Katul_2009}
+\APACinsertmetastar {%
+Katul_2009}%
+\begin{APACrefauthors}%
+Katul, G\BPBI G.%
+, Palmroth, S.%
+\BCBL {}\ \BBA {} Oren, R.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2009}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Leaf stomatal responses to vapour pressure deficit under
+  current and CO2-enriched atmosphere explained by the economics of gas
+  exchange} {Leaf stomatal responses to vapour pressure deficit under current
+  and co2-enriched atmosphere explained by the economics of gas
+  exchange}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Plant, Cell \& Environment}{32}{8}{968-979}.
+\newblock
+\begin{APACrefURL}
+  \url{https://onlinelibrary.wiley.com/doi/abs/10.1111/j.1365-3040.2009.01977.x}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1111/j.1365-3040.2009.01977.x} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Lawrence%
+\ \protect \BOthers {.}}{%
+Lawrence%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2019}}%
+}]{%
+Lawrence_2019}
+\APACinsertmetastar {%
+Lawrence_2019}%
+\begin{APACrefauthors}%
+Lawrence, D.%
+, Fisher, R.%
+, Koven, C.%
+, Oleson, K.%
+, Swenson, S.%
+, Bonan, G.%
+\BDBL {}Zeng, X.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2019}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {The Community Land Model version 5: Description of new
+  features, benchmarking, and impact of forcing uncertainty} {The community
+  land model version 5: Description of new features, benchmarking, and impact
+  of forcing uncertainty}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{submitted to Journal of Advances in Modeling Earth
+  Systems}{}{}{}.
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Lemordant%
+, Gentine%
+, Swann%
+, Cook%
+\BCBL {}\ \BBA {} Scheff%
+}{%
+Lemordant%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2018}}%
+}]{%
+Lemordant_2018}
+\APACinsertmetastar {%
+Lemordant_2018}%
+\begin{APACrefauthors}%
+Lemordant, L.%
+, Gentine, P.%
+, Swann, A\BPBI S.%
+, Cook, B\BPBI I.%
+\BCBL {}\ \BBA {} Scheff, J.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2018}{apr}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Critical impact of vegetation physiology on the
+  continental hydrologic cycle in response to increasing {CO}2} {Critical
+  impact of vegetation physiology on the continental hydrologic cycle in
+  response to increasing {CO}2}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Proceedings of the National Academy of
+  Sciences}{}{}{201720712}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1073\%2Fpnas.1720712115}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1073/pnas.1720712115} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Leuning%
+}{%
+Leuning%
+}{%
+{\protect \APACyear {1990}}%
+}]{%
+Leuning_1990}
+\APACinsertmetastar {%
+Leuning_1990}%
+\begin{APACrefauthors}%
+Leuning, R.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{1990}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Modelling Stomatal Behaviour and and Photosynthesis {of
+  Eucalyptus} grandis} {Modelling stomatal behaviour and and photosynthesis {of
+  Eucalyptus} grandis}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Australian Journal of Plant Physiology}{17}{2}{159}.
+\newblock
+\begin{APACrefDOI} \doi{10.1071/pp9900159} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+C.~Lin%
+\ \protect \BOthers {.}}{%
+C.~Lin%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2018}}%
+}]{%
+Lin_2018}
+\APACinsertmetastar {%
+Lin_2018}%
+\begin{APACrefauthors}%
+Lin, C.%
+, Gentine, P.%
+, Huang, Y.%
+, Guan, K.%
+, Kimm, H.%
+\BCBL {}\ \BBA {} Zhou, S.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2018}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Diel ecosystem conductance response to vapor pressure
+  deficit is suboptimal and independent of soil moisture} {Diel ecosystem
+  conductance response to vapor pressure deficit is suboptimal and independent
+  of soil moisture}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Agricultural and Forest Meteorology}{250}{}{24--34}.
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+H.~Lin%
+, Chen%
+, Zhang%
+, Fu%
+\BCBL {}\ \BBA {} Fan%
+}{%
+H.~Lin%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2017}}%
+}]{%
+Lin_2017}
+\APACinsertmetastar {%
+Lin_2017}%
+\begin{APACrefauthors}%
+Lin, H.%
+, Chen, Y.%
+, Zhang, H.%
+, Fu, P.%
+\BCBL {}\ \BBA {} Fan, Z.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2017}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Stronger cooling effects of transpiration and leaf
+  physical traits of plants from a hot dry habitat than from a hot wet habitat}
+  {Stronger cooling effects of transpiration and leaf physical traits of plants
+  from a hot dry habitat than from a hot wet habitat}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Functional Ecology}{31}{12}{2202-2211}.
+\newblock
+\begin{APACrefURL}
+  \url{https://besjournals.onlinelibrary.wiley.com/doi/abs/10.1111/1365-2435.12923}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1111/1365-2435.12923} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Y\BHBI S.~Lin%
+\ \protect \BOthers {.}}{%
+Y\BHBI S.~Lin%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2015}}%
+}]{%
+Lin_2015}
+\APACinsertmetastar {%
+Lin_2015}%
+\begin{APACrefauthors}%
+Lin, Y\BHBI S.%
+, Medlyn, B\BPBI E.%
+, Duursma, R\BPBI A.%
+, Prentice, I\BPBI C.%
+, Wang, H.%
+, Baig, S.%
+\BDBL {}Wingate, L.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2015}{mar}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Optimal stomatal behaviour around the world} {Optimal
+  stomatal behaviour around the world}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Nature Climate Change}{5}{5}{459--464}.
+\newblock
+\begin{APACrefDOI} \doi{10.1038/nclimate2550} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Lloyd%
+}{%
+Lloyd%
+}{%
+{\protect \APACyear {1991}}%
+}]{%
+Lloyd_1991}
+\APACinsertmetastar {%
+Lloyd_1991}%
+\begin{APACrefauthors}%
+Lloyd, J.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{1991}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Modelling stomatal responses to environment in
+  {M}acadamia integrifolia} {Modelling stomatal responses to environment in
+  {M}acadamia integrifolia}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Functional Plant Biology}{18}{6}{649--660}.
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Manzoni%
+, Vico%
+, Palmroth%
+, Porporato%
+\BCBL {}\ \BBA {} Katul%
+}{%
+Manzoni%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2013}}%
+}]{%
+Manzoni2013}
+\APACinsertmetastar {%
+Manzoni2013}%
+\begin{APACrefauthors}%
+Manzoni, S.%
+, Vico, G.%
+, Palmroth, S.%
+, Porporato, A.%
+\BCBL {}\ \BBA {} Katul, G.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2013}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Optimization of stomatal conductance for maximum carbon
+  gain under dynamic soil moisture} {Optimization of stomatal conductance for
+  maximum carbon gain under dynamic soil moisture}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Advances in Water Resources}{62}{}{90 - 105}.
+\newblock
+\begin{APACrefURL}
+  \url{http://www.sciencedirect.com/science/article/pii/S0309170813001814}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{https://doi.org/10.1016/j.advwatres.2013.09.020}
+  \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+McNaughton%
+\ \BBA {} Jarvis%
+}{%
+McNaughton%
+\ \BBA {} Jarvis%
+}{%
+{\protect \APACyear {1991}}%
+}]{%
+Mcnaughton_1991}
+\APACinsertmetastar {%
+Mcnaughton_1991}%
+\begin{APACrefauthors}%
+McNaughton, K.%
+\BCBT {}\ \BBA {} Jarvis, P.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{1991}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Effects of spatial scale on stomatal control of
+  transpiration} {Effects of spatial scale on stomatal control of
+  transpiration}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Agricultural and Forest
+  Meteorology}{54}{2-4}{279--302}.
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Medlyn%
+\ \protect \BOthers {.}}{%
+Medlyn%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2017}}%
+}]{%
+Medlyn_2017}
+\APACinsertmetastar {%
+Medlyn_2017}%
+\begin{APACrefauthors}%
+Medlyn, B\BPBI E.%
+, De~Kauwe, M\BPBI G.%
+, Lin, Y\BHBI S.%
+, Knauer, J.%
+, Duursma, R\BPBI A.%
+, Williams, C\BPBI A.%
+\BDBL {}others%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2017}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {How do leaf and ecosystem measures of water-use
+  efficiency compare?} {How do leaf and ecosystem measures of water-use
+  efficiency compare?}{\BBCQ}
+\newblock
+\APACjournalVolNumPages{New Phytologist}{216}{3}{758--770}.
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Medlyn%
+\ \protect \BOthers {.}}{%
+Medlyn%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2011}}%
+}]{%
+MEDLYN_2011}
+\APACinsertmetastar {%
+MEDLYN_2011}%
+\begin{APACrefauthors}%
+Medlyn, B\BPBI E.%
+, Duursma, R\BPBI A.%
+, Eamus, D.%
+, Ellsworth, D\BPBI S.%
+, Prentice, I\BPBI C.%
+, Barton, C\BPBI V\BPBI M.%
+\BDBL {}Wingate, L.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2011}{jan}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Reconciling the optimal and empirical approaches to
+  modelling stomatal conductance} {Reconciling the optimal and empirical
+  approaches to modelling stomatal conductance}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Global Change Biology}{17}{6}{2134--2144}.
+\newblock
+\begin{APACrefDOI} \doi{10.1111/j.1365-2486.2010.02375.x} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Monteith%
+\ \protect \BOthers {.}}{%
+Monteith%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {1965}}%
+}]{%
+Monteith_1965}
+\APACinsertmetastar {%
+Monteith_1965}%
+\begin{APACrefauthors}%
+Monteith, J\BPBI L.%
+\BCBT {}\ \BOthersPeriod {.}
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{1965}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Evaporation and environment} {Evaporation and
+  environment}.{\BBCQ}
+\newblock
+\BIn{} \APACrefbtitle {Symp. Soc. Exp. Biol} {Symp. soc. exp. biol}\ (\BVOL~19,
+  \BPG~4).
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Morton%
+}{%
+Morton%
+}{%
+{\protect \APACyear {1965}}%
+}]{%
+Morton_1965}
+\APACinsertmetastar {%
+Morton_1965}%
+\begin{APACrefauthors}%
+Morton, F\BPBI I.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{1965}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Potential evaporation and river basin evaporation}
+  {Potential evaporation and river basin evaporation}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Journal of the Hydraulics Division}{91}{6}{67--97}.
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Mott%
+}{%
+Mott%
+}{%
+{\protect \APACyear {2007}}%
+}]{%
+Mott2007}
+\APACinsertmetastar {%
+Mott2007}%
+\begin{APACrefauthors}%
+Mott, K\BPBI A.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2007}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Leaf hydraulic conductivity and stomatal responses to
+  humidity in amphistomatous leaves} {Leaf hydraulic conductivity and stomatal
+  responses to humidity in amphistomatous leaves}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Plant, Cell \& Environment}{30}{11}{1444-1449}.
+\newblock
+\begin{APACrefURL}
+  \url{https://onlinelibrary.wiley.com/doi/abs/10.1111/j.1365-3040.2007.01720.x}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1111/j.1365-3040.2007.01720.x} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Mott%
+\ \BBA {} Peak%
+}{%
+Mott%
+\ \BBA {} Peak%
+}{%
+{\protect \APACyear {2013}}%
+}]{%
+Mott2013}
+\APACinsertmetastar {%
+Mott2013}%
+\begin{APACrefauthors}%
+Mott, K\BPBI A.%
+\BCBT {}\ \BBA {} Peak, D.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2013}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Testing a vapour-phase model of stomatal responses to
+  humidity} {Testing a vapour-phase model of stomatal responses to
+  humidity}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Plant, Cell \& Environment}{36}{5}{936-944}.
+\newblock
+\begin{APACrefURL}
+  \url{https://onlinelibrary.wiley.com/doi/abs/10.1111/pce.12026}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1111/pce.12026} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Nikinmaa%
+\ \protect \BOthers {.}}{%
+Nikinmaa%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2013}}%
+}]{%
+Nikinmaa_2013}
+\APACinsertmetastar {%
+Nikinmaa_2013}%
+\begin{APACrefauthors}%
+Nikinmaa, E.%
+, H{\"o}ltt{\"a}, T.%
+, Hari, P.%
+, Kolari, P.%
+, M{\"a}kel{\"a}, A.%
+, Sevanto, S.%
+\BCBL {}\ \BBA {} Vesala, T.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2013}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Assimilate transport in phloem sets conditions for leaf
+  gas exchange} {Assimilate transport in phloem sets conditions for leaf gas
+  exchange}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Plant, Cell \& Environment}{36}{3}{655--669}.
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Niu%
+\ \protect \BOthers {.}}{%
+Niu%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2011}}%
+}]{%
+Niu_2011}
+\APACinsertmetastar {%
+Niu_2011}%
+\begin{APACrefauthors}%
+Niu, G\BHBI Y.%
+, Yang, Z\BHBI L.%
+, Mitchell, K\BPBI E.%
+, Chen, F.%
+, Ek, M\BPBI B.%
+, Barlage, M.%
+\BDBL {}Xia, Y.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2011}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {The community Noah land surface model with
+  multiparameterization options (Noah-MP): 1. Model description and evaluation
+  with local-scale measurements} {The community noah land surface model with
+  multiparameterization options (noah-mp): 1. model description and evaluation
+  with local-scale measurements}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Journal of Geophysical Research:
+  Atmospheres}{116}{D12}{}.
+\newblock
+\begin{APACrefURL}
+  \url{https://agupubs.onlinelibrary.wiley.com/doi/abs/10.1029/2010JD015139}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1029/2010JD015139} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Novick%
+\ \protect \BOthers {.}}{%
+Novick%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2016}}%
+}]{%
+Novick_2016}
+\APACinsertmetastar {%
+Novick_2016}%
+\begin{APACrefauthors}%
+Novick, K\BPBI A.%
+, Ficklin, D\BPBI L.%
+, Stoy, P\BPBI C.%
+, Williams, C\BPBI A.%
+, Bohrer, G.%
+, Oishi, A\BPBI C.%
+\BDBL {}others%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2016}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {The increasing importance of atmospheric demand for
+  ecosystem water and carbon fluxes} {The increasing importance of atmospheric
+  demand for ecosystem water and carbon fluxes}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Nature Climate Change}{6}{11}{1023}.
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Or%
+, Lehmann%
+, Shahraeeni%
+\BCBL {}\ \BBA {} Shokri%
+}{%
+Or%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2013}}%
+}]{%
+Or_2013}
+\APACinsertmetastar {%
+Or_2013}%
+\begin{APACrefauthors}%
+Or, D.%
+, Lehmann, P.%
+, Shahraeeni, E.%
+\BCBL {}\ \BBA {} Shokri, N.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2013}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Advances in soil evaporation physics- a review}
+  {Advances in soil evaporation physics- a review}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Vadose Zone Journal}{12}{4}{}.
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Oren%
+\ \protect \BOthers {.}}{%
+Oren%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {1999}}%
+}]{%
+Oren1999}
+\APACinsertmetastar {%
+Oren1999}%
+\begin{APACrefauthors}%
+Oren, R.%
+, Sperry, J\BPBI S.%
+, Katul, G\BPBI G.%
+, Pataki, D\BPBI E.%
+, Ewers, B\BPBI E.%
+, Phillips, N.%
+\BCBL {}\ \BBA {} Schäfer, K\BPBI V\BPBI R.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{1999}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Survey and synthesis of intra- and interspecific
+  variation in stomatal sensitivity to vapour pressure deficit} {Survey and
+  synthesis of intra- and interspecific variation in stomatal sensitivity to
+  vapour pressure deficit}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Plant, Cell \& Environment}{22}{12}{1515-1526}.
+\newblock
+\begin{APACrefURL}
+  \url{https://onlinelibrary.wiley.com/doi/abs/10.1046/j.1365-3040.1999.00513.x}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1046/j.1365-3040.1999.00513.x} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Ozdogan%
+, Salvucci%
+\BCBL {}\ \BBA {} Anderson%
+}{%
+Ozdogan%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2006}}%
+}]{%
+Ozdogan_2006}
+\APACinsertmetastar {%
+Ozdogan_2006}%
+\begin{APACrefauthors}%
+Ozdogan, M.%
+, Salvucci, G\BPBI D.%
+\BCBL {}\ \BBA {} Anderson, B\BPBI T.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2006}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Examination of the {B}ouchet-{M}orton Complementary
+  Relationship Using a Mesoscale Climate Model and Observations under a
+  Progressive Irrigation Scenario} {Examination of the {B}ouchet-{M}orton
+  complementary relationship using a mesoscale climate model and observations
+  under a progressive irrigation scenario}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Journal of Hydrometeorology}{7}{2}{235-251}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1175/JHM485.1} \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1175/JHM485.1} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Penman%
+}{%
+Penman%
+}{%
+{\protect \APACyear {1948}}%
+}]{%
+Penman_1948}
+\APACinsertmetastar {%
+Penman_1948}%
+\begin{APACrefauthors}%
+Penman, H\BPBI L.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{1948}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Natural evaporation from open water, bare soil and
+  grass} {Natural evaporation from open water, bare soil and grass}.{\BBCQ}
+\newblock
+\BIn{} \APACrefbtitle {Proc. R. Soc. Lond. A} {Proc. r. soc. lond. a}\
+  (\BVOL~193, \BPGS\ 120--145).
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Rawson%
+, Begg%
+\BCBL {}\ \BBA {} Woodward%
+}{%
+Rawson%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {1977}}%
+}]{%
+Rawson1977}
+\APACinsertmetastar {%
+Rawson1977}%
+\begin{APACrefauthors}%
+Rawson, H\BPBI M.%
+, Begg, J\BPBI E.%
+\BCBL {}\ \BBA {} Woodward, R\BPBI G.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{1977}{Jan}{01}.
+\newblock
+{\BBOQ}\APACrefatitle {The effect of atmospheric humidity on photosynthesis,
+  transpiration and water use efficiency of leaves of several plant species}
+  {The effect of atmospheric humidity on photosynthesis, transpiration and
+  water use efficiency of leaves of several plant species}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Planta}{134}{1}{5--10}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1007/BF00390086} \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1007/BF00390086} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Rigden%
+\ \BBA {} Salvucci%
+}{%
+Rigden%
+\ \BBA {} Salvucci%
+}{%
+{\protect \APACyear {2017}}%
+}]{%
+Rigden_2017}
+\APACinsertmetastar {%
+Rigden_2017}%
+\begin{APACrefauthors}%
+Rigden, A\BPBI J.%
+\BCBT {}\ \BBA {} Salvucci, G\BPBI D.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2017}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Stomatal response to humidity and {CO}2 implicated in
+  recent decline in US evaporation} {Stomatal response to humidity and {CO}2
+  implicated in recent decline in us evaporation}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Global Change Biology}{23}{3}{1140--1151}.
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Rogers%
+\ \protect \BOthers {.}}{%
+Rogers%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2017}}%
+}]{%
+Rogers_2017}
+\APACinsertmetastar {%
+Rogers_2017}%
+\begin{APACrefauthors}%
+Rogers, A.%
+, Medlyn, B\BPBI E.%
+, Dukes, J\BPBI S.%
+, Bonan, G.%
+, von Caemmerer, S.%
+, Dietze, M\BPBI C.%
+\BDBL {}Zaehle, S.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2017}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {A roadmap for improving the representation of
+  photosynthesis in Earth system models} {A roadmap for improving the
+  representation of photosynthesis in earth system models}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{New Phytologist}{213}{1}{22-42}.
+\newblock
+\begin{APACrefURL}
+  \url{https://nph.onlinelibrary.wiley.com/doi/abs/10.1111/nph.14283}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1111/nph.14283} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Ryan%
+\ \BBA {} Asao%
+}{%
+Ryan%
+\ \BBA {} Asao%
+}{%
+{\protect \APACyear {2014}}%
+}]{%
+Ryan_2014}
+\APACinsertmetastar {%
+Ryan_2014}%
+\begin{APACrefauthors}%
+Ryan, M\BPBI G.%
+\BCBT {}\ \BBA {} Asao, S.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2014}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Phloem transport in trees} {Phloem transport in
+  trees}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Tree Physiology}{34}{1}{1--4}.
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Salvucci%
+\ \BBA {} Gentine%
+}{%
+Salvucci%
+\ \BBA {} Gentine%
+}{%
+{\protect \APACyear {2013}}%
+}]{%
+Salvucci_2013}
+\APACinsertmetastar {%
+Salvucci_2013}%
+\begin{APACrefauthors}%
+Salvucci, G\BPBI D.%
+\BCBT {}\ \BBA {} Gentine, P.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2013}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Emergent relation between surface vapor conductance and
+  relative humidity profiles yields evaporation rates from weather data}
+  {Emergent relation between surface vapor conductance and relative humidity
+  profiles yields evaporation rates from weather data}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Proceedings of the National Academy of
+  Sciences}{110}{16}{6287--6291}.
+\newblock
+\begin{APACrefURL} \url{http://www.pnas.org/content/110/16/6287}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1073/pnas.1215844110} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Shuttleworth%
+}{%
+Shuttleworth%
+}{%
+{\protect \APACyear {2012}}%
+}]{%
+Shuttleworth_2012}
+\APACinsertmetastar {%
+Shuttleworth_2012}%
+\begin{APACrefauthors}%
+Shuttleworth, W\BPBI J.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYear{2012}.
+\newblock
+\APACrefbtitle {Terrestrial hydrometeorology} {Terrestrial hydrometeorology}.
+\newblock
+\APACaddressPublisher{}{John Wiley \& Sons}.
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Swann%
+, Hoffman%
+, Koven%
+\BCBL {}\ \BBA {} Randerson%
+}{%
+Swann%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2016}}%
+}]{%
+Swann_2016}
+\APACinsertmetastar {%
+Swann_2016}%
+\begin{APACrefauthors}%
+Swann, A\BPBI L\BPBI S.%
+, Hoffman, F\BPBI M.%
+, Koven, C\BPBI D.%
+\BCBL {}\ \BBA {} Randerson, J\BPBI T.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2016}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Plant responses to increasing {CO}2 reduce estimates of
+  climate impacts on drought severity} {Plant responses to increasing {CO}2
+  reduce estimates of climate impacts on drought severity}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Proceedings of the National Academy of
+  Sciences}{113}{36}{10019--10024}.
+\newblock
+\begin{APACrefURL} \url{http://www.pnas.org/content/113/36/10019}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1073/pnas.1604581113} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Turner%
+, Schulze%
+\BCBL {}\ \BBA {} Gollan%
+}{%
+Turner%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {1984}}%
+}]{%
+Turner1984}
+\APACinsertmetastar {%
+Turner1984}%
+\begin{APACrefauthors}%
+Turner, N\BPBI C.%
+, Schulze, E\BHBI D.%
+\BCBL {}\ \BBA {} Gollan, T.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{1984}{Aug}{01}.
+\newblock
+{\BBOQ}\APACrefatitle {The responses of stomata and leaf gas exchange to vapour
+  pressure deficits and soil water content} {The responses of stomata and leaf
+  gas exchange to vapour pressure deficits and soil water content}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Oecologia}{63}{3}{338--342}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1007/BF00390662} \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1007/BF00390662} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Zhou%
+, Yu%
+, Huang%
+\BCBL {}\ \BBA {} Wang%
+}{%
+Zhou%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2014}}%
+}]{%
+Zhou_2014}
+\APACinsertmetastar {%
+Zhou_2014}%
+\begin{APACrefauthors}%
+Zhou, S.%
+, Yu, B.%
+, Huang, Y.%
+\BCBL {}\ \BBA {} Wang, G.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2014}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {The effect of vapor pressure deficit on water use
+  efficiency at the subdaily time scale} {The effect of vapor pressure deficit
+  on water use efficiency at the subdaily time scale}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Geophysical Research Letters}{41}{14}{5005--5013}.
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Zhou%
+, Yu%
+, Huang%
+\BCBL {}\ \BBA {} Wang%
+}{%
+Zhou%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2015}}%
+}]{%
+Zhou_2015}
+\APACinsertmetastar {%
+Zhou_2015}%
+\begin{APACrefauthors}%
+Zhou, S.%
+, Yu, B.%
+, Huang, Y.%
+\BCBL {}\ \BBA {} Wang, G.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2015}{may}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Daily underlying water use efficiency for {AmeriFlux}
+  sites} {Daily underlying water use efficiency for {AmeriFlux} sites}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Journal of Geophysical Research:
+  Biogeosciences}{120}{5}{887--902}.
+\newblock
+\begin{APACrefDOI} \doi{10.1002/2015jg002947} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Zhou%
+, Zhang%
+, Park~Williams%
+\BCBL {}\ \BBA {} Gentine%
+}{%
+Zhou%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2019}}%
+}]{%
+Zhou_2019}
+\APACinsertmetastar {%
+Zhou_2019}%
+\begin{APACrefauthors}%
+Zhou, S.%
+, Zhang, Y.%
+, Park~Williams, A.%
+\BCBL {}\ \BBA {} Gentine, P.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2019}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Projected increases in intensity, frequency, and
+  terrestrial carbon costs of compound drought and aridity events} {Projected
+  increases in intensity, frequency, and terrestrial carbon costs of compound
+  drought and aridity events}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Science Advances}{5}{1}{}.
+\newblock
+\begin{APACrefURL} \url{http://advances.sciencemag.org/content/5/1/eaau5740}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1126/sciadv.aau5740} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\end{thebibliography}
 

### 2019-09-06 doc/paper/agu-james-submission-single-file.tex
need to use agu's template for supp material
also I added all of the fluxnet citations to the main mauscript,
per agu's guidelines

@@
         \propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}
+\usepackage{textcomp}
 
@@
 
+% include all the fluxnet citations I use in the supp material
+\nocite{AT-Neu, AU-ASM, AU-ASM, AU-Cpr, AU-DaP, AU-DaS, AU-Dry, AU-Gin,
+  AU-How, AU-Gin, AU-DaP, AU-Tum, AU-Whr, AU-Wom, BE-Lon, BE-Vie,
+  BR-Sa3, CA-Qfo, CA-SF1, CA-SF1, CH-Cha, CH-Dav, CH-Fru, DE-Geb,
+  DE-Gri, DE-Hai, DE-Gri, DE-Lkb, DE-Seh, DE-Tha, DK-Sor, FI-Hyy,
+  FI-Sod, FR-Gri, FR-LBr, IT-Col, IT-Cpz, IT-Lav, IT-MBo, IT-Noe,
+  IT-Ren, IT-Ro2, IT-SRo, IT-Tor, NL-Loo, RU-Fyo, US-AR1, US-AR1,
+  US-ARM, US-Blo, US-KS2, US-MMS, US-Me2, US-NR1, US-Ne1, US-Ne1,
+  US-Ne1, US-SRG, US-SRM, US-Syv, US-Ton, US-Var, US-WCr, US-Wkg,
+  ZA-Kru, ZM-Mon}
+
 %Text here ===>>>
@@
 \bibitem [\protect \citeauthoryear {%
+Anthoni%
+\ \protect \BOthers {.}}{%
+Anthoni%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2004}}%
+}]{%
+DE-Geb}
+\APACinsertmetastar {%
+DE-Geb}%
+\begin{APACrefauthors}%
+Anthoni, P\BPBI M.%
+, Knohl, A.%
+, Rebmann, C.%
+, Freibauer, A.%
+, Mund, M.%
+, Ziegler, W.%
+\BDBL {}Schulze, E\BHBI D.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2004}{dec}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Forest and agricultural land-use-dependent {CO}2
+  exchange in {T}huringia, {G}ermany} {Forest and agricultural
+  land-use-dependent {CO}2 exchange in {T}huringia, {G}ermany}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Global Change Biology}{10}{12}{2005--2019}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1111%2Fj.1365-2486.2004.00863.x}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1111/j.1365-2486.2004.00863.x} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Archibald%
+\ \protect \BOthers {.}}{%
+Archibald%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2009}}%
+}]{%
+ZA-Kru}
+\APACinsertmetastar {%
+ZA-Kru}%
+\begin{APACrefauthors}%
+Archibald, S\BPBI A.%
+, Kirton, A.%
+, van~der Merwe, M\BPBI R.%
+, Scholes, R\BPBI J.%
+, Williams, C\BPBI A.%
+\BCBL {}\ \BBA {} Hanan, N.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2009}{feb}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Drivers of inter-annual variability in Net Ecosystem
+  Exchange in a semi-arid savanna ecosystem, {S}outh {A}frica} {Drivers of
+  inter-annual variability in net ecosystem exchange in a semi-arid savanna
+  ecosystem, {S}outh {A}frica}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Biogeosciences}{6}{2}{251--266}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.5194%2Fbg-6-251-2009}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.5194/bg-6-251-2009} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Aubinet%
+\ \protect \BOthers {.}}{%
+Aubinet%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2001}}%
+}]{%
+BE-Vie}
+\APACinsertmetastar {%
+BE-Vie}%
+\begin{APACrefauthors}%
+Aubinet, M.%
+, Chermanne, B.%
+, Vandenhaute, M.%
+, Longdoz, B.%
+, Yernaux, M.%
+\BCBL {}\ \BBA {} Laitat, E.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2001}{jul}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Long term carbon dioxide exchange above a mixed forest
+  in the {B}elgian {A}rdennes} {Long term carbon dioxide exchange above a mixed
+  forest in the {B}elgian {A}rdennes}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Agricultural and Forest Meteorology}{108}{4}{293--315}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1016%2Fs0168-1923%2801%2900244-1}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1016/s0168-1923(01)00244-1} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Baldocchi%
+\ \protect \BOthers {.}}{%
+Baldocchi%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2010}}%
+}]{%
+US-Ton}
+\APACinsertmetastar {%
+US-Ton}%
+\begin{APACrefauthors}%
+Baldocchi, D.%
+, Chen, Q.%
+, Chen, X.%
+, Ma, S.%
+, Miller, G.%
+, Ryu, Y.%
+\BDBL {}Battles, J.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2010}{nov}{}.
+\newblock
+{\BBOQ}\APACrefatitle {The Dynamics of Energy, Water, and Carbon Fluxes in a
+  Blue Oak ({Q}uercus douglasii) Savanna in {C}alifornia} {The dynamics of
+  energy, water, and carbon fluxes in a blue oak ({Q}uercus douglasii) savanna
+  in {C}alifornia}.{\BBCQ}
+\newblock
+\BIn{} \APACrefbtitle {Ecosystem Function in {S}avannas} {Ecosystem function in
+  {S}avannas}\ (\BPGS\ 135--151).
+\newblock
+\APACaddressPublisher{}{{CRC} Press}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1201%2Fb10275-10} \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1201/b10275-10} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
 Ball%
@@
 \begin{APACrefDOI} \doi{10.1007/978-94-017-0519-6\_48} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Berbigier%
+, Bonnefond%
+\BCBL {}\ \BBA {} Mellmann%
+}{%
+Berbigier%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2001}}%
+}]{%
+FR-LBr}
+\APACinsertmetastar {%
+FR-LBr}%
+\begin{APACrefauthors}%
+Berbigier, P.%
+, Bonnefond, J\BHBI M.%
+\BCBL {}\ \BBA {} Mellmann, P.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2001}{jun}{}.
+\newblock
+{\BBOQ}\APACrefatitle {{CO}2 and water vapour fluxes for 2 years above
+  {E}uroflux forest site} {{CO}2 and water vapour fluxes for 2 years above
+  {E}uroflux forest site}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Agricultural and Forest Meteorology}{108}{3}{183--197}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1016%2Fs0168-1923%2801%2900240-4}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1016/s0168-1923(01)00240-4} \end{APACrefDOI}
 \PrintBackRefs{\CurrentBib}
@@
 \bibitem [\protect \citeauthoryear {%
+Bergeron%
+\ \protect \BOthers {.}}{%
+Bergeron%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2007}}%
+}]{%
+CA-Qfo}
+\APACinsertmetastar {%
+CA-Qfo}%
+\begin{APACrefauthors}%
+Bergeron, O.%
+, Margolis, H\BPBI A.%
+, Black, T\BPBI A.%
+, Coursolle, C.%
+, Dunn, A\BPBI L.%
+, Barr, A\BPBI G.%
+\BCBL {}\ \BBA {} Wofsy, S\BPBI C.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2007}{jan}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Comparison of carbon dioxide fluxes over three boreal
+  black spruce forests in {C}anada} {Comparison of carbon dioxide fluxes over
+  three boreal black spruce forests in {C}anada}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Global Change Biology}{13}{1}{89--107}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1111%2Fj.1365-2486.2006.01281.x}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1111/j.1365-2486.2006.01281.x} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Beringer%
+, Hutley%
+, Hacker%
+, Neininger%
+\BCBL {}\ \BBA {} U%
+}{%
+Beringer%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2011}}%
+}]{%
+AU-DaP}
+\APACinsertmetastar {%
+AU-DaP}%
+\begin{APACrefauthors}%
+Beringer, J.%
+, Hutley, L\BPBI B.%
+, Hacker, J\BPBI M.%
+, Neininger, B.%
+\BCBL {}\ \BBA {} U, K\BPBI T\BPBI P.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2011}{nov}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Patterns and processes of carbon, water and energy
+  cycles across northern {A}ustralian landscapes: From point to region}
+  {Patterns and processes of carbon, water and energy cycles across northern
+  {A}ustralian landscapes: From point to region}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Agricultural and Forest
+  Meteorology}{151}{11}{1409--1416}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1016%2Fj.agrformet.2011.05.003}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1016/j.agrformet.2011.05.003} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Beringer%
+\ \protect \BOthers {.}}{%
+Beringer%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2016}}%
+}]{%
+AU-Gin}
+\APACinsertmetastar {%
+AU-Gin}%
+\begin{APACrefauthors}%
+Beringer, J.%
+, Hutley, L\BPBI B.%
+, McHugh, I.%
+, Arndt, S\BPBI K.%
+, Campbell, D.%
+, Cleugh, H\BPBI A.%
+\BDBL {}Wardlaw, T.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2016}{oct}{}.
+\newblock
+{\BBOQ}\APACrefatitle {{An introduction to the Australian and New Zealand flux
+  tower network - OzFlux}} {{An introduction to the Australian and New Zealand
+  flux tower network - OzFlux}}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Biogeosciences}{13}{21}{5895--5916}.
+\newblock
+\begin{APACrefURL} \url{http://www.biogeosciences.net/13/5895/2016/
+  https://doi.org/10.5194 https://doi.org/10.5194{\%}2Fbg-13-5895-2016}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.5194/bg-13-5895-2016} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Beringer%
+, Hutley%
+, Tapper%
+\BCBL {}\ \BBA {} Cernusak%
+}{%
+Beringer%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2007}}%
+}]{%
+AU-How}
+\APACinsertmetastar {%
+AU-How}%
+\begin{APACrefauthors}%
+Beringer, J.%
+, Hutley, L\BPBI B.%
+, Tapper, N\BPBI J.%
+\BCBL {}\ \BBA {} Cernusak, L\BPBI A.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2007}{may}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Savanna fires and their impact on net ecosystem
+  productivity in {N}orth {A}ustralia} {Savanna fires and their impact on net
+  ecosystem productivity in {N}orth {A}ustralia}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Global Change Biology}{13}{5}{990--1004}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1111%2Fj.1365-2486.2007.01334.x}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1111/j.1365-2486.2007.01334.x} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
 Bouchet%
@@
 \bibitem [\protect \citeauthoryear {%
+Cernusak%
+, Hutley%
+, Beringer%
+, Holtum%
+\BCBL {}\ \BBA {} Turner%
+}{%
+Cernusak%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2011}}%
+}]{%
+AU-Dry}
+\APACinsertmetastar {%
+AU-Dry}%
+\begin{APACrefauthors}%
+Cernusak, L\BPBI A.%
+, Hutley, L\BPBI B.%
+, Beringer, J.%
+, Holtum, J\BPBI A.%
+\BCBL {}\ \BBA {} Turner, B\BPBI L.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2011}{nov}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Photosynthetic physiology of eucalypts along a
+  sub-continental rainfall gradient in northern {A}ustralia} {Photosynthetic
+  physiology of eucalypts along a sub-continental rainfall gradient in northern
+  {A}ustralia}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Agricultural and Forest
+  Meteorology}{151}{11}{1462--1470}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1016%2Fj.agrformet.2011.01.006}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1016/j.agrformet.2011.01.006} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Chiesi%
+\ \protect \BOthers {.}}{%
+Chiesi%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2005}}%
+}]{%
+IT-SRo}
+\APACinsertmetastar {%
+IT-SRo}%
+\begin{APACrefauthors}%
+Chiesi, M.%
+, Maselli, F.%
+, Bindi, M.%
+, Fibbi, L.%
+, Cherubini, P.%
+, Arlotta, E.%
+\BDBL {}Seufert, G.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2005}{dec}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Modelling carbon budget of {M}editerranean forests using
+  ground and remote sensing measurements} {Modelling carbon budget of
+  {M}editerranean forests using ground and remote sensing measurements}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Agricultural and Forest Meteorology}{135}{1-4}{22--34}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1016%2Fj.agrformet.2005.09.011}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1016/j.agrformet.2005.09.011} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Cleverly%
+\ \protect \BOthers {.}}{%
+Cleverly%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2013}}%
+}]{%
+AU-ASM}
+\APACinsertmetastar {%
+AU-ASM}%
+\begin{APACrefauthors}%
+Cleverly, J.%
+, Boulain, N.%
+, Villalobos-Vega, R.%
+, Grant, N.%
+, Faux, R.%
+, Wood, C.%
+\BDBL {}Eamus, D.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2013}{jul}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Dynamics of component carbon fluxes in a semi-arid
+  {A}cacia woodland, central {A}ustralia} {Dynamics of component carbon fluxes
+  in a semi-arid {A}cacia woodland, central {A}ustralia}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Journal of Geophysical Research:
+  Biogeosciences}{118}{3}{1168--1185}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1002%2Fjgrg.20101} \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1002/jgrg.20101} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Cook%
+\ \protect \BOthers {.}}{%
+Cook%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2004}}%
+}]{%
+US-WCr}
+\APACinsertmetastar {%
+US-WCr}%
+\begin{APACrefauthors}%
+Cook, B\BPBI D.%
+, Davis, K\BPBI J.%
+, Wang, W.%
+, Desai, A.%
+, Berger, B\BPBI W.%
+, Teclaw, R\BPBI M.%
+\BDBL {}Heilman, W.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2004}{nov}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Carbon exchange and venting anomalies in an upland
+  deciduous forest in northern {W}isconsin, {USA}} {Carbon exchange and venting
+  anomalies in an upland deciduous forest in northern {W}isconsin,
+  {USA}}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Agricultural and Forest
+  Meteorology}{126}{3-4}{271--295}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1016%2Fj.agrformet.2004.06.008}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1016/j.agrformet.2004.06.008} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
 Cowan%
@@
 \bibitem [\protect \citeauthoryear {%
+Desai%
+, Bolstad%
+, Cook%
+, Davis%
+\BCBL {}\ \BBA {} Carey%
+}{%
+Desai%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2005}}%
+}]{%
+US-Syv}
+\APACinsertmetastar {%
+US-Syv}%
+\begin{APACrefauthors}%
+Desai, A\BPBI R.%
+, Bolstad, P\BPBI V.%
+, Cook, B\BPBI D.%
+, Davis, K\BPBI J.%
+\BCBL {}\ \BBA {} Carey, E\BPBI V.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2005}{jan}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Comparing net ecosystem exchange of carbon dioxide
+  between an old-growth and mature forest in the upper {M}idwest, {USA}}
+  {Comparing net ecosystem exchange of carbon dioxide between an old-growth and
+  mature forest in the upper {M}idwest, {USA}}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Agricultural and Forest Meteorology}{128}{1-2}{33--55}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1016%2Fj.agrformet.2004.09.005}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1016/j.agrformet.2004.09.005} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
 De~Schepper%
@@
 \bibitem [\protect \citeauthoryear {%
+Dragoni%
+\ \protect \BOthers {.}}{%
+Dragoni%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2011}}%
+}]{%
+US-MMS}
+\APACinsertmetastar {%
+US-MMS}%
+\begin{APACrefauthors}%
+Dragoni, D.%
+, Schmid, H\BPBI P.%
+, Wayson, C\BPBI A.%
+, Potter, H.%
+, Grimmond, C\BPBI S\BPBI B.%
+\BCBL {}\ \BBA {} Randolph, J\BPBI C.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2011}{jan}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Evidence of increased net ecosystem productivity
+  associated with a longer vegetated season in a deciduous forest in
+  south-central {I}ndiana, {USA}} {Evidence of increased net ecosystem
+  productivity associated with a longer vegetated season in a deciduous forest
+  in south-central {I}ndiana, {USA}}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Global Change Biology}{17}{2}{886--897}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1111%2Fj.1365-2486.2010.02281.x}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1111/j.1365-2486.2010.02281.x} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
 G.~Farquhar%
@@
 \begin{APACrefDOI} \doi{10.1007/BF00386231} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Fischer%
+, Billesbach%
+, Berry%
+, Riley%
+\BCBL {}\ \BBA {} Torn%
+}{%
+Fischer%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2007}}%
+}]{%
+US-ARM}
+\APACinsertmetastar {%
+US-ARM}%
+\begin{APACrefauthors}%
+Fischer, M\BPBI L.%
+, Billesbach, D\BPBI P.%
+, Berry, J\BPBI A.%
+, Riley, W\BPBI J.%
+\BCBL {}\ \BBA {} Torn, M\BPBI S.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2007}{oct}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Spatiotemporal Variations in Growing Season Exchanges of
+  {CO}2, {H2O}, and Sensible Heat in Agricultural Fields of the Southern
+  {G}reat {P}lains} {Spatiotemporal variations in growing season exchanges of
+  {CO}2, {H2O}, and sensible heat in agricultural fields of the southern
+  {G}reat {P}lains}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Earth Interactions}{11}{17}{1--21}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1175%2Fei231.1} \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1175/ei231.1} \end{APACrefDOI}
 \PrintBackRefs{\CurrentBib}
@@
 \bibitem [\protect \citeauthoryear {%
+Galvagno%
+\ \protect \BOthers {.}}{%
+Galvagno%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2013}}%
+}]{%
+IT-Tor}
+\APACinsertmetastar {%
+IT-Tor}%
+\begin{APACrefauthors}%
+Galvagno, M.%
+, Wohlfahrt, G.%
+, Cremonese, E.%
+, Rossini, M.%
+, Colombo, R.%
+, Filippa, G.%
+\BDBL {}Migliavacca, M.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2013}{apr}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Phenology and carbon dioxide source/sink strength of a
+  subalpine grassland in response to an exceptionally short snow season}
+  {Phenology and carbon dioxide source/sink strength of a subalpine grassland
+  in response to an exceptionally short snow season}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Environmental Research Letters}{8}{2}{025008}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1088%2F1748-9326%2F8%2F2%2F025008}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1088/1748-9326/8/2/025008} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Garbulsky%
+, Pe{\~{n}}uelas%
+, Papale%
+\BCBL {}\ \BBA {} Filella%
+}{%
+Garbulsky%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2008}}%
+}]{%
+IT-Cpz}
+\APACinsertmetastar {%
+IT-Cpz}%
+\begin{APACrefauthors}%
+Garbulsky, M\BPBI F.%
+, Pe{\~{n}}uelas, J.%
+, Papale, D.%
+\BCBL {}\ \BBA {} Filella, I.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2008}{dec}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Remote estimation of carbon dioxide uptake by a
+  {M}editerranean forest} {Remote estimation of carbon dioxide uptake by a
+  {M}editerranean forest}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Global Change Biology}{14}{12}{2860--2867}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1111%2Fj.1365-2486.2008.01684.x}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1111/j.1365-2486.2008.01684.x} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
 Gentine%
@@
 \begin{APACrefDOI} \doi{10.1002/2016GL070819} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Goldstein%
+\ \protect \BOthers {.}}{%
+Goldstein%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2000}}%
+}]{%
+US-Blo}
+\APACinsertmetastar {%
+US-Blo}%
+\begin{APACrefauthors}%
+Goldstein, A.%
+, Hultman, N.%
+, Fracheboud, J.%
+, Bauer, M.%
+, Panek, J.%
+, Xu, M.%
+\BDBL {}Baugh, W.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2000}{mar}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Effects of climate variability on the carbon dioxide,
+  water, and sensible heat fluxes above a ponderosa pine plantation in the
+  {S}ierra {N}evada ({CA})} {Effects of climate variability on the carbon
+  dioxide, water, and sensible heat fluxes above a ponderosa pine plantation in
+  the {S}ierra {N}evada ({CA})}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Agricultural and Forest
+  Meteorology}{101}{2-3}{113--129}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1016%2Fs0168-1923%2899%2900168-9}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1016/s0168-1923(99)00168-9} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Gr{\"u}nwald%
+\ \BBA {} Bernhofer%
+}{%
+Gr{\"u}nwald%
+\ \BBA {} Bernhofer%
+}{%
+{\protect \APACyear {2007}}%
+}]{%
+DE-Tha}
+\APACinsertmetastar {%
+DE-Tha}%
+\begin{APACrefauthors}%
+Gr{\"u}nwald, T.%
+\BCBT {}\ \BBA {} Bernhofer, C.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2007}{jul}{}.
+\newblock
+{\BBOQ}\APACrefatitle {A decade of carbon, water and energy flux measurements
+  of an old spruce forest at the {A}nchor {S}tation {T}harandt} {A decade of
+  carbon, water and energy flux measurements of an old spruce forest at the
+  {A}nchor {S}tation {T}harandt}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Tellus B}{59}{3}{}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.3402%2Ftellusb.v59i3.17000}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.3402/tellusb.v59i3.17000} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Hinko-Najera%
+\ \protect \BOthers {.}}{%
+Hinko-Najera%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2017}}%
+}]{%
+AU-Wom}
+\APACinsertmetastar {%
+AU-Wom}%
+\begin{APACrefauthors}%
+Hinko-Najera, N.%
+, Isaac, P.%
+, Beringer, J.%
+, van Gorsel, E.%
+, Ewenz, C.%
+, McHugh, I.%
+\BDBL {}Arndt, S\BPBI K.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2017}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {{Net ecosystem carbon exchange of a dry temperate
+  eucalypt forest}} {{Net ecosystem carbon exchange of a dry temperate eucalypt
+  forest}}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Biogeosciences}{14}{16}{3781--3800}.
+\newblock
+\begin{APACrefURL} \url{https://www.biogeosciences.net/14/3781/2017/}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.5194/bg-14-3781-2017} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Hutley%
+, Beringer%
+, Isaac%
+, Hacker%
+\BCBL {}\ \BBA {} Cernusak%
+}{%
+Hutley%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2011}}%
+}]{%
+AU-DaS}
+\APACinsertmetastar {%
+AU-DaS}%
+\begin{APACrefauthors}%
+Hutley, L\BPBI B.%
+, Beringer, J.%
+, Isaac, P\BPBI R.%
+, Hacker, J\BPBI M.%
+\BCBL {}\ \BBA {} Cernusak, L\BPBI A.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2011}{nov}{}.
+\newblock
+{\BBOQ}\APACrefatitle {A sub-continental scale living laboratory: Spatial
+  patterns of savanna vegetation over a rainfall gradient in northern
+  {A}ustralia} {A sub-continental scale living laboratory: Spatial patterns of
+  savanna vegetation over a rainfall gradient in northern {A}ustralia}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Agricultural and Forest
+  Meteorology}{151}{11}{1417--1428}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1016%2Fj.agrformet.2011.03.002}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1016/j.agrformet.2011.03.002} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Imer%
+, Merbold%
+, Eugster%
+\BCBL {}\ \BBA {} Buchmann%
+}{%
+Imer%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2013}}%
+}]{%
+CH-Fru}
+\APACinsertmetastar {%
+CH-Fru}%
+\begin{APACrefauthors}%
+Imer, D.%
+, Merbold, L.%
+, Eugster, W.%
+\BCBL {}\ \BBA {} Buchmann, N.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2013}{sep}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Temporal and spatial variations of soil
+  {CO}\textsubscript{2}, {CH}\textsubscript{4} and {N}\textsubscript{2}{O}
+  fluxes at three differently managed grasslands} {Temporal and spatial
+  variations of soil {CO}\textsubscript{2}, {CH}\textsubscript{4} and
+  {N}\textsubscript{2}{O} fluxes at three differently managed
+  grasslands}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Biogeosciences}{10}{9}{5931--5945}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.5194%2Fbg-10-5931-2013}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.5194/bg-10-5931-2013} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Irvine%
+, Law%
+, Martin%
+\BCBL {}\ \BBA {} Vickers%
+}{%
+Irvine%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2008}}%
+}]{%
+US-Me2}
+\APACinsertmetastar {%
+US-Me2}%
+\begin{APACrefauthors}%
+Irvine, J.%
+, Law, B\BPBI E.%
+, Martin, J\BPBI G.%
+\BCBL {}\ \BBA {} Vickers, D.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2008}{dec}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Interannual variation in soil {CO}2 efflux and the
+  response of root respiration to climate and canopy gas exchange in mature
+  ponderosa pine} {Interannual variation in soil {CO}2 efflux and the response
+  of root respiration to climate and canopy gas exchange in mature ponderosa
+  pine}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Global Change Biology}{14}{12}{2848--2859}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1111%2Fj.1365-2486.2008.01682.x}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1111/j.1365-2486.2008.01682.x} \end{APACrefDOI}
 \PrintBackRefs{\CurrentBib}
@@
 \begin{APACrefDOI} \doi{10.1111/j.1365-3040.2009.01977.x} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Knohl%
+, Schulze%
+, Kolle%
+\BCBL {}\ \BBA {} Buchmann%
+}{%
+Knohl%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2003}}%
+}]{%
+DE-Hai}
+\APACinsertmetastar {%
+DE-Hai}%
+\begin{APACrefauthors}%
+Knohl, A.%
+, Schulze, E\BHBI D.%
+, Kolle, O.%
+\BCBL {}\ \BBA {} Buchmann, N.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2003}{sep}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Large carbon uptake by an unmanaged 250-year-old
+  deciduous forest in Central {G}ermany} {Large carbon uptake by an unmanaged
+  250-year-old deciduous forest in central {G}ermany}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Agricultural and Forest
+  Meteorology}{118}{3-4}{151--167}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1016%2Fs0168-1923%2803%2900115-1}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1016/s0168-1923(03)00115-1} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Kurbatova%
+, Li%
+, Varlagin%
+, Xiao%
+\BCBL {}\ \BBA {} Vygodskaya%
+}{%
+Kurbatova%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2008}}%
+}]{%
+RU-Fyo}
+\APACinsertmetastar {%
+RU-Fyo}%
+\begin{APACrefauthors}%
+Kurbatova, J.%
+, Li, C.%
+, Varlagin, A.%
+, Xiao, X.%
+\BCBL {}\ \BBA {} Vygodskaya, N.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2008}{jul}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Modeling carbon dynamics in two adjacent spruce forests
+  with different soil conditions in {R}ussia} {Modeling carbon dynamics in two
+  adjacent spruce forests with different soil conditions in {R}ussia}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Biogeosciences}{5}{4}{969--980}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.5194%2Fbg-5-969-2008}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.5194/bg-5-969-2008} \end{APACrefDOI}
 \PrintBackRefs{\CurrentBib}
@@
 \bibitem [\protect \citeauthoryear {%
+Leuning%
+, Cleugh%
+, Zegelin%
+\BCBL {}\ \BBA {} Hughes%
+}{%
+Leuning%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2005}}%
+}]{%
+AU-Tum}
+\APACinsertmetastar {%
+AU-Tum}%
+\begin{APACrefauthors}%
+Leuning, R.%
+, Cleugh, H\BPBI A.%
+, Zegelin, S\BPBI J.%
+\BCBL {}\ \BBA {} Hughes, D.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2005}{apr}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Carbon and water fluxes over a temperate Eucalyptus
+  forest and a tropical wet/dry savanna in {A}ustralia: measurements and
+  comparison with {MODIS} remote sensing estimates} {Carbon and water fluxes
+  over a temperate eucalyptus forest and a tropical wet/dry savanna in
+  {A}ustralia: measurements and comparison with {MODIS} remote sensing
+  estimates}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Agricultural and Forest
+  Meteorology}{129}{3-4}{151--173}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1016%2Fj.agrformet.2004.12.004}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1016/j.agrformet.2004.12.004} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
 C.~Lin%
@@
 \bibitem [\protect \citeauthoryear {%
+Lindauer%
+\ \protect \BOthers {.}}{%
+Lindauer%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2014}}%
+}]{%
+DE-Lkb}
+\APACinsertmetastar {%
+DE-Lkb}%
+\begin{APACrefauthors}%
+Lindauer, M.%
+, Schmid, H.%
+, Grote, R.%
+, Mauder, M.%
+, Steinbrecher, R.%
+\BCBL {}\ \BBA {} Wolpert, B.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2014}{oct}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Net ecosystem exchange over a non-cleared
+  wind-throw-disturbed upland spruce forest{\textemdash}Measurements and
+  simulations} {Net ecosystem exchange over a non-cleared wind-throw-disturbed
+  upland spruce forest{\textemdash}measurements and simulations}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Agricultural and Forest Meteorology}{197}{}{219--234}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1016%2Fj.agrformet.2014.07.005}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1016/j.agrformet.2014.07.005} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
 Lloyd%
@@
 \APACjournalVolNumPages{Functional Plant Biology}{18}{6}{649--660}.
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Loubet%
+\ \protect \BOthers {.}}{%
+Loubet%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2011}}%
+}]{%
+FR-Gri}
+\APACinsertmetastar {%
+FR-Gri}%
+\begin{APACrefauthors}%
+Loubet, B.%
+, Laville, P.%
+, Lehuger, S.%
+, Larmanou, E.%
+, Fl{\'{e}}chard, C.%
+, Mascher, N.%
+\BDBL {}Cellier, P.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2011}{mar}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Carbon, nitrogen and Greenhouse gases budgets over a
+  four years crop rotation in northern {F}rance} {Carbon, nitrogen and
+  greenhouse gases budgets over a four years crop rotation in northern
+  {F}rance}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Plant and Soil}{343}{1-2}{109--137}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1007%2Fs11104-011-0751-9}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1007/s11104-011-0751-9} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Ma%
+, Baldocchi%
+, Xu%
+\BCBL {}\ \BBA {} Hehn%
+}{%
+Ma%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2007}}%
+}]{%
+US-Var}
+\APACinsertmetastar {%
+US-Var}%
+\begin{APACrefauthors}%
+Ma, S.%
+, Baldocchi, D\BPBI D.%
+, Xu, L.%
+\BCBL {}\ \BBA {} Hehn, T.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2007}{dec}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Inter-annual variability in carbon dioxide exchange of
+  an oak/grass savanna and open grassland in {C}alifornia} {Inter-annual
+  variability in carbon dioxide exchange of an oak/grass savanna and open
+  grassland in {C}alifornia}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Agricultural and Forest
+  Meteorology}{147}{3-4}{157--171}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1016%2Fj.agrformet.2007.07.008}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1016/j.agrformet.2007.07.008} \end{APACrefDOI}
 \PrintBackRefs{\CurrentBib}
@@
   \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Marcolla%
+\ \protect \BOthers {.}}{%
+Marcolla%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2011}}%
+}]{%
+IT-MBo}
+\APACinsertmetastar {%
+IT-MBo}%
+\begin{APACrefauthors}%
+Marcolla, B.%
+, Cescatti, A.%
+, Manca, G.%
+, Zorer, R.%
+, Cavagna, M.%
+, Fiora, A.%
+\BDBL {}Zampedri, R.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2011}{sep}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Climatic controls and ecosystem responses drive the
+  inter-annual variability of the net ecosystem exchange of an alpine meadow}
+  {Climatic controls and ecosystem responses drive the inter-annual variability
+  of the net ecosystem exchange of an alpine meadow}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Agricultural and Forest
+  Meteorology}{151}{9}{1233--1243}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1016%2Fj.agrformet.2011.04.015}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1016/j.agrformet.2011.04.015} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Marcolla%
+, Pitacco%
+\BCBL {}\ \BBA {} Cescatti%
+}{%
+Marcolla%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2003}}%
+}]{%
+IT-Lav}
+\APACinsertmetastar {%
+IT-Lav}%
+\begin{APACrefauthors}%
+Marcolla, B.%
+, Pitacco, A.%
+\BCBL {}\ \BBA {} Cescatti, A.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2003}{jul}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Canopy Architecture and Turbulence Structure in a
+  Coniferous Forest} {Canopy architecture and turbulence structure in a
+  coniferous forest}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Boundary-Layer Meteorology}{108}{1}{39--59}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1023%2Fa%3A1023027709805}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1023/a:1023027709805} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+McHugh%
+\ \protect \BOthers {.}}{%
+McHugh%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2017}}%
+}]{%
+AU-Whr}
+\APACinsertmetastar {%
+AU-Whr}%
+\begin{APACrefauthors}%
+McHugh, I\BPBI D.%
+, Beringer, J.%
+, Cunningham, S\BPBI C.%
+, Baker, P\BPBI J.%
+, Cavagnaro, T\BPBI R.%
+, Nally, R\BPBI M.%
+\BCBL {}\ \BBA {} Thompson, R\BPBI M.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2017}{jun}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Interactions between nocturnal turbulent flux, storage
+  and advection at an {\textquotedblleft}ideal{\textquotedblright} eucalypt
+  woodland site} {Interactions between nocturnal turbulent flux, storage and
+  advection at an {\textquotedblleft}ideal{\textquotedblright} eucalypt
+  woodland site}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Biogeosciences}{14}{12}{3027--3050}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.5194%2Fbg-14-3027-2017}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.5194/bg-14-3027-2017} \end{APACrefDOI}
 \PrintBackRefs{\CurrentBib}
@@
 \bibitem [\protect \citeauthoryear {%
+Merbold%
+\ \protect \BOthers {.}}{%
+Merbold%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2009}}%
+}]{%
+ZM-Mon}
+\APACinsertmetastar {%
+ZM-Mon}%
+\begin{APACrefauthors}%
+Merbold, L.%
+, Ard{\"o}, J.%
+, Arneth, A.%
+, Scholes, R\BPBI J.%
+, Nouvellon, Y.%
+, de Grandcourt, A.%
+\BDBL {}Kutsch, W\BPBI L.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2009}{jun}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Precipitation as driver of carbon fluxes in 11 {A}frican
+  ecosystems} {Precipitation as driver of carbon fluxes in 11 {A}frican
+  ecosystems}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Biogeosciences}{6}{6}{1027--1041}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.5194%2Fbg-6-1027-2009}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.5194/bg-6-1027-2009} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Merbold%
+\ \protect \BOthers {.}}{%
+Merbold%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2014}}%
+}]{%
+CH-Cha}
+\APACinsertmetastar {%
+CH-Cha}%
+\begin{APACrefauthors}%
+Merbold, L.%
+, Eugster, W.%
+, Stieger, J.%
+, Zahniser, M.%
+, Nelson, D.%
+\BCBL {}\ \BBA {} Buchmann, N.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2014}{feb}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Greenhouse gas budget ({CO}2, {CH}4and {N2O}) of
+  intensively managed grassland following restoration} {Greenhouse gas budget
+  ({CO}2, {CH}4and {N2O}) of intensively managed grassland following
+  restoration}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Global Change Biology}{20}{6}{1913--1928}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1111%2Fgcb.12518} \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1111/gcb.12518} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Meyer%
+, Kondrlov{\`{a}}%
+\BCBL {}\ \BBA {} Koerber%
+}{%
+Meyer%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2015}}%
+}]{%
+AU-Cpr}
+\APACinsertmetastar {%
+AU-Cpr}%
+\begin{APACrefauthors}%
+Meyer, W\BPBI S.%
+, Kondrlov{\`{a}}, E.%
+\BCBL {}\ \BBA {} Koerber, G\BPBI R.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2015}{mar}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Evaporation of perennial semi-arid woodland in
+  southeastern {A}ustralia is adapted for irregular but common dry periods}
+  {Evaporation of perennial semi-arid woodland in southeastern {A}ustralia is
+  adapted for irregular but common dry periods}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Hydrological Processes}{29}{17}{3714--3726}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1002%2Fhyp.10467} \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1002/hyp.10467} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Mkhabela%
+\ \protect \BOthers {.}}{%
+Mkhabela%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2009}}%
+}]{%
+CA-SF1}
+\APACinsertmetastar {%
+CA-SF1}%
+\begin{APACrefauthors}%
+Mkhabela, M.%
+, Amiro, B.%
+, Barr, A.%
+, Black, T.%
+, Hawthorne, I.%
+, Kidston, J.%
+\BDBL {}Zha, T.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2009}{may}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Comparison of carbon dynamics and water use efficiency
+  following fire and harvesting in {C}anadian boreal forests} {Comparison of
+  carbon dynamics and water use efficiency following fire and harvesting in
+  {C}anadian boreal forests}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Agricultural and Forest Meteorology}{149}{5}{783--794}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1016%2Fj.agrformet.2008.10.025}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1016/j.agrformet.2008.10.025} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Monson%
+\ \protect \BOthers {.}}{%
+Monson%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2002}}%
+}]{%
+US-NR1}
+\APACinsertmetastar {%
+US-NR1}%
+\begin{APACrefauthors}%
+Monson, R\BPBI K.%
+, Turnipseed, A\BPBI A.%
+, Sparks, J\BPBI P.%
+, Harley, P\BPBI C.%
+, Scott-Denton, L\BPBI E.%
+, Sparks, K.%
+\BCBL {}\ \BBA {} Huxman, T\BPBI E.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2002}{may}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Carbon sequestration in a high-elevation, subalpine
+  forest} {Carbon sequestration in a high-elevation, subalpine forest}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Global Change Biology}{8}{5}{459--478}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1046%2Fj.1365-2486.2002.00480.x}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1046/j.1365-2486.2002.00480.x} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Montagnani%
+\ \protect \BOthers {.}}{%
+Montagnani%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2009}}%
+}]{%
+IT-Ren}
+\APACinsertmetastar {%
+IT-Ren}%
+\begin{APACrefauthors}%
+Montagnani, L.%
+, Manca, G.%
+, Canepa, E.%
+, Georgieva, E.%
+, Acosta, M.%
+, Feigenwinter, C.%
+\BDBL {}Ziegler, W.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2009}{apr}{}.
+\newblock
+{\BBOQ}\APACrefatitle {A new mass conservation approach to the study of {CO}2
+  advection in an alpine forest} {A new mass conservation approach to the study
+  of {CO}2 advection in an alpine forest}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Journal of Geophysical Research}{114}{D7}{}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1029%2F2008jd010650}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1029/2008jd010650} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
 Monteith%
@@
   \BPG~4).
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Moors%
+}{%
+Moors%
+}{%
+{\protect \APACyear {2012}}%
+}]{%
+NL-Loo}
+\APACinsertmetastar {%
+NL-Loo}%
+\begin{APACrefauthors}%
+Moors, E.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYear{2012}.
+\unskip\
+\newblock
+\APACrefbtitle {{Water Use of Forests in The Netherlands}} {{Water Use of
+  Forests in The Netherlands}}\ \APACtypeAddressSchool {\BUPhD}{}{}.
+\unskip\
+\newblock
+\APACaddressSchool {}{Vrije Universiteit Amsterdam}.
 \PrintBackRefs{\CurrentBib}
@@
 \begin{APACrefDOI} \doi{10.1111/pce.12026} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Moureaux%
+, Debacq%
+, Bodson%
+, Heinesch%
+\BCBL {}\ \BBA {} Aubinet%
+}{%
+Moureaux%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2006}}%
+}]{%
+BE-Lon}
+\APACinsertmetastar {%
+BE-Lon}%
+\begin{APACrefauthors}%
+Moureaux, C.%
+, Debacq, A.%
+, Bodson, B.%
+, Heinesch, B.%
+\BCBL {}\ \BBA {} Aubinet, M.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2006}{sep}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Annual net ecosystem carbon exchange by a sugar beet
+  crop} {Annual net ecosystem carbon exchange by a sugar beet crop}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Agricultural and Forest Meteorology}{139}{1-2}{25--39}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1016%2Fj.agrformet.2006.05.009}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1016/j.agrformet.2006.05.009} \end{APACrefDOI}
 \PrintBackRefs{\CurrentBib}
@@
 \bibitem [\protect \citeauthoryear {%
+Papale%
+\ \protect \BOthers {.}}{%
+Papale%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2014}}%
+}]{%
+IT-Noe}
+\APACinsertmetastar {%
+IT-Noe}%
+\begin{APACrefauthors}%
+Papale, D.%
+, Migliavacca, M.%
+, Cremonese, E.%
+, Cescatti, A.%
+, Alberti, G.%
+, Balzarolo, M.%
+\BDBL {}Valentini, R.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2014}{dec}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Carbon, Water and Energy Fluxes of Terrestrial
+  Ecosystems in {I}taly} {Carbon, water and energy fluxes of terrestrial
+  ecosystems in {I}taly}.{\BBCQ}
+\newblock
+\BIn{} \APACrefbtitle {The Greenhouse Gas Balance of Italy} {The greenhouse gas
+  balance of italy}\ (\BPGS\ 11--45).
+\newblock
+\APACaddressPublisher{}{Springer Berlin Heidelberg}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1007%2F978-3-642-32424-6\_2}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1007/978-3-642-32424-6\_2} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
 Penman%
@@
   (\BVOL~193, \BPGS\ 120--145).
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Pilegaard%
+, Ibrom%
+, Courtney%
+, Hummelsh{\o}j%
+\BCBL {}\ \BBA {} Jensen%
+}{%
+Pilegaard%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2011}}%
+}]{%
+DK-Sor}
+\APACinsertmetastar {%
+DK-Sor}%
+\begin{APACrefauthors}%
+Pilegaard, K.%
+, Ibrom, A.%
+, Courtney, M\BPBI S.%
+, Hummelsh{\o}j, P.%
+\BCBL {}\ \BBA {} Jensen, N\BPBI O.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2011}{jul}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Increasing net {CO}2 uptake by a {D}anish beech forest
+  during the period from 1996 to 2009} {Increasing net {CO}2 uptake by a
+  {D}anish beech forest during the period from 1996 to 2009}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Agricultural and Forest Meteorology}{151}{7}{934--946}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1016%2Fj.agrformet.2011.02.013}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1016/j.agrformet.2011.02.013} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Powell%
+\ \protect \BOthers {.}}{%
+Powell%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2006}}%
+}]{%
+US-KS2}
+\APACinsertmetastar {%
+US-KS2}%
+\begin{APACrefauthors}%
+Powell, T\BPBI L.%
+, Bracho, R.%
+, Li, J.%
+, Dore, S.%
+, Hinkle, C\BPBI R.%
+\BCBL {}\ \BBA {} Drake, B\BPBI G.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2006}{dec}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Environmental controls over net ecosystem carbon
+  exchange of scrub oak in central {F}lorida} {Environmental controls over net
+  ecosystem carbon exchange of scrub oak in central {F}lorida}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Agricultural and Forest Meteorology}{141}{1}{19--34}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1016%2Fj.agrformet.2006.09.002}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1016/j.agrformet.2006.09.002} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Prescher%
+, Gr{\"u}nwald%
+\BCBL {}\ \BBA {} Bernhofer%
+}{%
+Prescher%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2010}}%
+}]{%
+DE-Gri}
+\APACinsertmetastar {%
+DE-Gri}%
+\begin{APACrefauthors}%
+Prescher, A\BHBI K.%
+, Gr{\"u}nwald, T.%
+\BCBL {}\ \BBA {} Bernhofer, C.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2010}{jul}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Land use regulates carbon budgets in eastern {G}ermany:
+  From {NEE} to {NBP}} {Land use regulates carbon budgets in eastern {G}ermany:
+  From {NEE} to {NBP}}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Agricultural and Forest
+  Meteorology}{150}{7-8}{1016--1025}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1016%2Fj.agrformet.2010.03.008}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1016/j.agrformet.2010.03.008} \end{APACrefDOI}
 \PrintBackRefs{\CurrentBib}
@@
 \begin{APACrefDOI} \doi{10.1007/BF00390086} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Raz-Yaseef%
+\ \protect \BOthers {.}}{%
+Raz-Yaseef%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2015}}%
+}]{%
+US-AR1}
+\APACinsertmetastar {%
+US-AR1}%
+\begin{APACrefauthors}%
+Raz-Yaseef, N.%
+, Billesbach, D\BPBI P.%
+, Fischer, M\BPBI L.%
+, Biraud, S\BPBI C.%
+, Gunter, S\BPBI A.%
+, Bradford, J\BPBI A.%
+\BCBL {}\ \BBA {} Torn, M\BPBI S.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2015}{dec}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Vulnerability of crops and native grasses to summer
+  drying in the {U.S.} Southern {G}reat {P}lains} {Vulnerability of crops and
+  native grasses to summer drying in the {U.S.} southern {G}reat
+  {P}lains}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Agriculture, Ecosystems {\&}
+  Environment}{213}{}{209--218}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1016%2Fj.agee.2015.07.021}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1016/j.agee.2015.07.021} \end{APACrefDOI}
 \PrintBackRefs{\CurrentBib}
@@
 \bibitem [\protect \citeauthoryear {%
+Schmidt%
+, Reichenau%
+, Fiener%
+\BCBL {}\ \BBA {} Schneider%
+}{%
+Schmidt%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2012}}%
+}]{%
+DE-Seh}
+\APACinsertmetastar {%
+DE-Seh}%
+\begin{APACrefauthors}%
+Schmidt, M.%
+, Reichenau, T.%
+, Fiener, P.%
+\BCBL {}\ \BBA {} Schneider, K.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2012}{nov}{}.
+\newblock
+{\BBOQ}\APACrefatitle {The carbon budget of a winter wheat field: An eddy
+  covariance analysis of seasonal and inter-annual variability} {The carbon
+  budget of a winter wheat field: An eddy covariance analysis of seasonal and
+  inter-annual variability}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Agricultural and Forest Meteorology}{165}{}{114--126}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1016%2Fj.agrformet.2012.05.012}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1016/j.agrformet.2012.05.012} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Scott%
+, Biederman%
+, Hamerlynck%
+\BCBL {}\ \BBA {} Barron-Gafford%
+}{%
+Scott%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2015}}%
+}]{%
+US-SRG}
+\APACinsertmetastar {%
+US-SRG}%
+\begin{APACrefauthors}%
+Scott, R\BPBI L.%
+, Biederman, J\BPBI A.%
+, Hamerlynck, E\BPBI P.%
+\BCBL {}\ \BBA {} Barron-Gafford, G\BPBI A.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2015}{dec}{}.
+\newblock
+{\BBOQ}\APACrefatitle {The carbon balance pivot point of southwestern {U}.{S}.
+  semiarid ecosystems: {I}nsights from the 21st century drought} {The carbon
+  balance pivot point of southwestern {U}.{S}. semiarid ecosystems: {I}nsights
+  from the 21st century drought}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Journal of Geophysical Research:
+  Biogeosciences}{120}{12}{2612--2624}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1002%2F2015jg003181}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1002/2015jg003181} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Scott%
+, Hamerlynck%
+, Jenerette%
+, Moran%
+\BCBL {}\ \BBA {} Barron-Gafford%
+}{%
+Scott%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2010}}%
+}]{%
+US-Wkg}
+\APACinsertmetastar {%
+US-Wkg}%
+\begin{APACrefauthors}%
+Scott, R\BPBI L.%
+, Hamerlynck, E\BPBI P.%
+, Jenerette, G\BPBI D.%
+, Moran, M\BPBI S.%
+\BCBL {}\ \BBA {} Barron-Gafford, G\BPBI A.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2010}{sep}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Carbon dioxide exchange in a semidesert grassland
+  through drought-induced vegetation change} {Carbon dioxide exchange in a
+  semidesert grassland through drought-induced vegetation change}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Journal of Geophysical Research}{115}{G3}{}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1029%2F2010jg001348}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1029/2010jg001348} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Scott%
+, Jenerette%
+, Potts%
+\BCBL {}\ \BBA {} Huxman%
+}{%
+Scott%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2009}}%
+}]{%
+US-SRM}
+\APACinsertmetastar {%
+US-SRM}%
+\begin{APACrefauthors}%
+Scott, R\BPBI L.%
+, Jenerette, G\BPBI D.%
+, Potts, D\BPBI L.%
+\BCBL {}\ \BBA {} Huxman, T\BPBI E.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2009}{nov}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Effects of seasonal drought on net carbon dioxide
+  exchange from a woody-plant-encroached semiarid grassland} {Effects of
+  seasonal drought on net carbon dioxide exchange from a woody-plant-encroached
+  semiarid grassland}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Journal of Geophysical Research}{114}{G4}{}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1029%2F2008jg000900}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1029/2008jg000900} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
 Shuttleworth%
@@
 \APACaddressPublisher{}{John Wiley \& Sons}.
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Suni%
+\ \protect \BOthers {.}}{%
+Suni%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2003}}%
+}]{%
+FI-Hyy}
+\APACinsertmetastar {%
+FI-Hyy}%
+\begin{APACrefauthors}%
+Suni, T.%
+, Rinne, J.%
+, Reissel, A.%
+, Altimir, N.%
+, Keronen, P.%
+, Rannik, {\"{U}}.%
+\BDBL {}Vesala, T.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2003}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {{Long-term measurements of surface fluxes above a Scots
+  pine forest in Hyyti{\"{a}}l{\"{a}}, southern Finland}} {{Long-term
+  measurements of surface fluxes above a Scots pine forest in
+  Hyyti{\"{a}}l{\"{a}}, southern Finland}}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Boreal Environ. Res.}{4}{}{287--301}.
 \PrintBackRefs{\CurrentBib}
@@
 \bibitem [\protect \citeauthoryear {%
+Tedeschi%
+\ \protect \BOthers {.}}{%
+Tedeschi%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2006}}%
+}]{%
+IT-Ro2}
+\APACinsertmetastar {%
+IT-Ro2}%
+\begin{APACrefauthors}%
+Tedeschi, V.%
+, Rey, A.%
+, Manca, G.%
+, Valentini, R.%
+, Jarvis, P\BPBI G.%
+\BCBL {}\ \BBA {} Borghetti, M.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2006}{jan}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Soil respiration in a {M}editerranean oak forest at
+  different developmental stages after coppicing} {Soil respiration in a
+  {M}editerranean oak forest at different developmental stages after
+  coppicing}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Global Change Biology}{12}{1}{110--121}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1111%2Fj.1365-2486.2005.01081.x}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1111/j.1365-2486.2005.01081.x} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Thum%
+\ \protect \BOthers {.}}{%
+Thum%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2007}}%
+}]{%
+FI-Sod}
+\APACinsertmetastar {%
+FI-Sod}%
+\begin{APACrefauthors}%
+Thum, T.%
+, Aalto, T.%
+, Laurila, T.%
+, Aurela, M.%
+, Kolari, P.%
+\BCBL {}\ \BBA {} Hari, P.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2007}{nov}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Parametrization of two photosynthesis models at the
+  canopy scale in a northern boreal {S}cots pine forest} {Parametrization of
+  two photosynthesis models at the canopy scale in a northern boreal {S}cots
+  pine forest}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Tellus B}{59}{5}{}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.3402%2Ftellusb.v59i5.17066}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.3402/tellusb.v59i5.17066} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
 Turner%
@@
 \begin{APACrefDOI} \doi{10.1007/BF00390662} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Valentini%
+\ \protect \BOthers {.}}{%
+Valentini%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {1996}}%
+}]{%
+IT-Col}
+\APACinsertmetastar {%
+IT-Col}%
+\begin{APACrefauthors}%
+Valentini, R.%
+, Angelis, P.%
+, Matteucci, G.%
+, Monaco, R.%
+, Dore, S.%
+\BCBL {}\ \BBA {} Mucnozza, G\BPBI E\BPBI S.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{1996}{jun}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Seasonal net carbon dioxide exchange of a beech forest
+  with the atmosphere} {Seasonal net carbon dioxide exchange of a beech forest
+  with the atmosphere}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Global Change Biology}{2}{3}{199--207}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1111%2Fj.1365-2486.1996.tb00072.x}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1111/j.1365-2486.1996.tb00072.x} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Verma%
+\ \protect \BOthers {.}}{%
+Verma%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2005}}%
+}]{%
+US-Ne1}
+\APACinsertmetastar {%
+US-Ne1}%
+\begin{APACrefauthors}%
+Verma, S\BPBI B.%
+, Dobermann, A.%
+, Cassman, K\BPBI G.%
+, Walters, D\BPBI T.%
+, Knops, J\BPBI M.%
+, Arkebauer, T\BPBI J.%
+\BDBL {}Walter-Shea, E\BPBI A.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2005}{jul}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Annual carbon dioxide exchange in irrigated and rainfed
+  maize-based agroecosystems} {Annual carbon dioxide exchange in irrigated and
+  rainfed maize-based agroecosystems}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Agricultural and Forest Meteorology}{131}{1-2}{77--96}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1016%2Fj.agrformet.2005.05.003}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1016/j.agrformet.2005.05.003} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Wick%
+, Veldkamp%
+, de Mello%
+, Keller%
+\BCBL {}\ \BBA {} Crill%
+}{%
+Wick%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2005}}%
+}]{%
+BR-Sa3}
+\APACinsertmetastar {%
+BR-Sa3}%
+\begin{APACrefauthors}%
+Wick, B.%
+, Veldkamp, E.%
+, de Mello, W\BPBI Z.%
+, Keller, M.%
+\BCBL {}\ \BBA {} Crill, P.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2005}{aug}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Nitrous oxide fluxes and nitrogen cycling along a
+  pasture chronosequence in Central {A}mazonia, {B}razil} {Nitrous oxide fluxes
+  and nitrogen cycling along a pasture chronosequence in central {A}mazonia,
+  {B}razil}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Biogeosciences}{2}{2}{175--187}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.5194%2Fbg-2-175-2005}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.5194/bg-2-175-2005} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Wohlfahrt%
+\ \protect \BOthers {.}}{%
+Wohlfahrt%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2008}}%
+}]{%
+AT-Neu}
+\APACinsertmetastar {%
+AT-Neu}%
+\begin{APACrefauthors}%
+Wohlfahrt, G.%
+, Hammerle, A.%
+, Haslwanter, A.%
+, Bahn, M.%
+, Tappeiner, U.%
+\BCBL {}\ \BBA {} Cernusca, A.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2008}{apr}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Seasonal and inter-annual variability of the net
+  ecosystem {CO}2 exchange of a temperate mountain grassland: Effects of
+  weather and management} {Seasonal and inter-annual variability of the net
+  ecosystem {CO}2 exchange of a temperate mountain grassland: Effects of
+  weather and management}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Journal of Geophysical Research}{113}{D8}{}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1029%2F2007jd009286}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.1029/2007jd009286} \end{APACrefDOI}
 \PrintBackRefs{\CurrentBib}
@@
 \begin{APACrefDOI} \doi{10.1126/sciadv.aau5740} \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Zielis%
+\ \protect \BOthers {.}}{%
+Zielis%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2014}}%
+}]{%
+CH-Dav}
+\APACinsertmetastar {%
+CH-Dav}%
+\begin{APACrefauthors}%
+Zielis, S.%
+, Etzold, S.%
+, Zweifel, R.%
+, Eugster, W.%
+, Haeni, M.%
+\BCBL {}\ \BBA {} Buchmann, N.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2014}{mar}{}.
+\newblock
+{\BBOQ}\APACrefatitle {{NEP} of a {S}wiss subalpine forest is significantly
+  driven not only by current but also by previous year{\textquotesingle}s
+  weather} {{NEP} of a {S}wiss subalpine forest is significantly driven not
+  only by current but also by previous year{\textquotesingle}s weather}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Biogeosciences}{11}{6}{1627--1635}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.5194%2Fbg-11-1627-2014}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} \doi{10.5194/bg-11-1627-2014} \end{APACrefDOI}
 \PrintBackRefs{\CurrentBib}

### 2019-09-06 doc/paper/flux_sites.tex
need to use agu's template for supp material
also I added all of the fluxnet citations to the main mauscript,
per agu's guidelines

AT-Neu & GRA & 47.1167 & 11.3175 & Unk & 2002-2012 & \cite{AT-Neu} \\
AU-ASM & ENF & -22.2830 & 133.2490 & Unk & 2010-2013 & \cite{AU-ASM} \\
AU-Cpr & SAV & -34.0021 & 140.5891 & Unk & 2010-2014 & \cite{AU-Cpr} \\
AU-DaP & GRA & -14.0633 & 131.3181 & Aw  & 2007-2013 & \cite{AU-DaP} \\
AU-DaS & SAV & -14.1593 & 131.3881 & Aw  & 2008-2014 & \cite{AU-DaS} \\
AU-Dry & SAV & -15.2588 & 132.3706 & Unk & 2008-2014 & \cite{AU-Dry} \\
AU-Gin & WSA & -31.3764 & 115.7138 & Unk & 2011-2014 & \cite{AU-Gin} \\
AU-How & WSA & -12.4943 & 131.1523 & Aw  & 2001-2014 & \cite{AU-How} \\
AU-Rig & GRA & -36.6499 & 145.5759 & Unk & 2011-2014 & \cite{AU-Gin} \\
AU-Stp & GRA & -17.1507 & 133.3502 & Unk & 2008-2014 & \cite{AU-DaP} \\
AU-Tum & EBF & -35.6566 & 148.1517 & Cfb & 2001-2014 & \cite{AU-Tum} \\
AU-Whr & EBF & -36.6732 & 145.0294 & Unk & 2011-2014 & \cite{AU-Whr} \\
AU-Wom & EBF & -37.4222 & 144.0944 & Unk & 2010-2012 & \cite{AU-Wom} \\
BE-Lon & CRO & 50.5516 & 4.7461 & Cfb & 2004-2014 & \cite{BE-Lon} \\
BE-Vie & MF & 50.3051 & 5.9981 & Cfb & 1996-2014 & \cite{BE-Vie} \\
BR-Sa3 & EBF & -3.0180 & -54.9714 & Am & 2000-2004 & \cite{BR-Sa3} \\
CA-Qfo & ENF & 49.6925 & -74.3421 & Dfc & 2003-2010 & \cite{CA-Qfo} \\
CA-SF1 & ENF & 54.4850 & -105.8176 & Dfc & 2003-2006 & \cite{CA-SF1} \\
CA-SF2 & ENF & 54.2539 & -105.8775 & Dfc & 2001-2005 & \cite{CA-SF1} \\
CH-Cha & GRA & 47.2102 & 8.4104 & Unk & 2005-2014 & \cite{CH-Cha} \\
CH-Dav & ENF & 46.8153 & 9.8559 & Unk & 1997-2014 & \cite{CH-Dav} \\
CH-Fru & GRA & 47.1158 & 8.5378 & Unk & 2005-2014 & \cite{CH-Fru} \\
DE-Geb & CRO & 51.1001 & 10.9143 & Unk & 2001-2014 & \cite{DE-Geb} \\
DE-Gri & GRA & 50.9500 & 13.5126 & Cfb & 2004-2014 & \cite{DE-Gri} \\
DE-Hai & DBF & 51.0792 & 10.4530 & Unk & 2000-2012 & \cite{DE-Hai} \\
DE-Kli & CRO & 50.8931 & 13.5224 & Cfb & 2004-2014 & \cite{DE-Gri} \\
DE-Lkb & ENF & 49.0996 & 13.3047 & Unk & 2009-2013 & \cite{DE-Lkb} \\
DE-Obe & ENF & 50.7867 & 13.7213 & Cfb & 2008-2014 & {\textendash} \\
DE-Seh & CRO & 50.8706 & 6.4497 & Unk & 2007-2010 & \cite{DE-Seh} \\
DE-Tha & ENF & 50.9624 & 13.5652 & Cfb & 1996-2014 & \cite{DE-Tha} \\
DK-Sor & DBF & 55.4859 & 11.6446 & Unk & 1996-2014 & \cite{DK-Sor} \\
FI-Hyy & ENF & 61.8474 & 24.2948 & Unk & 1996-2014 & \cite{FI-Hyy} \\
FI-Sod & ENF & 67.3619 & 26.6378 & Unk & 2001-2014 & \cite{FI-Sod} \\
FR-Gri & CRO & 48.8442 & 1.9519 & Cfb & 2004-2013 & \cite{FR-Gri} \\
FR-LBr & ENF & 44.7171 & -0.7693 & Unk & 1996-2008 & \cite{FR-LBr} \\
IT-Col & DBF & 41.8494 & 13.5881 & Unk & 1996-2014 & \cite{IT-Col} \\
IT-Cpz & EBF & 41.7052 & 12.3761 & Unk & 1997-2009 & \cite{IT-Cpz} \\
IT-Lav & ENF & 45.9562 & 11.2813 & Unk & 2003-2014 & \cite{IT-Lav} \\
IT-MBo & GRA & 46.0147 & 11.0458 & Unk & 2003-2013 & \cite{IT-MBo} \\
IT-Noe & CSH & 40.6061 & 8.1515 & Unk & 2004-2014 & \cite{IT-Noe} \\
IT-Ren & ENF & 46.5869 & 11.4337 & Unk & 1998-2013 & \cite{IT-Ren} \\
IT-Ro2 & DBF & 42.3903 & 11.9209 & Unk & 2002-2012 & \cite{IT-Ro2} \\
IT-SRo & ENF & 43.7279 & 10.2844 & Unk & 1999-2012 & \cite{IT-SRo} \\
IT-Tor & GRA & 45.8444 & 7.5781 & Unk & 2008-2014 & \cite{IT-Tor} \\
NL-Loo & ENF & 52.1666 & 5.7436 & Unk & 1996-2013 & \cite{NL-Loo} \\
RU-Fyo & ENF & 56.4615 & 32.9221 & Unk & 1998-2014 & \cite{RU-Fyo} \\
US-AR1 & GRA & 36.4267 & -99.4200 & Dsa & 2009-2012 & \cite{US-AR1} \\
US-AR2 & GRA & 36.6358 & -99.5975 & Dsa & 2009-2012 & \cite{US-AR1} \\
US-ARM & CRO & 36.6058 & -97.4888 & Cfa & 2003-2012 & \cite{US-ARM} \\
US-Blo & ENF & 38.8953 & -120.6328 & Csa & 1997-2007 & \cite{US-Blo} \\
US-KS2 & CSH & 28.6086 & -80.6715 & Cwa & 2003-2006 & \cite{US-KS2} \\
US-MMS & DBF & 39.3232 & -86.4131 & Cfa & 1999-2014 & \cite{US-MMS} \\
US-Me2 & ENF & 44.4523 & -121.5574 & Csb & 2002-2014 & \cite{US-Me2} \\
US-NR1 & ENF & 40.0329 & -105.5464 & Dfc & 1998-2014 & \cite{US-NR1} \\
US-Ne1 & CRO & 41.1651 & -96.4766 & Dfa & 2001-2013 & \cite{US-Ne1} \\
US-Ne2 & CRO & 41.1649 & -96.4701 & Dfa & 2001-2013 & \cite{US-Ne1} \\
US-Ne3 & CRO & 41.1797 & -96.4397 & Dfa & 2001-2013 & \cite{US-Ne1} \\
US-SRG & GRA & 31.7894 & -110.8277 & Bsk & 2008-2014 & \cite{US-SRG} \\
US-SRM & WSA & 31.8214 & -110.8661 & Bsk & 2004-2014 & \cite{US-SRM} \\
US-Syv & MF & 46.2420 & -89.3477 & Dfb & 2001-2014 & \cite{US-Syv} \\
US-Ton & WSA & 38.4316 & -120.9660 & Csa & 2001-2014 & \cite{US-Ton} \\
US-Var & GRA & 38.4133 & -120.9507 & Csa & 2000-2014 & \cite{US-Var} \\
US-WCr & DBF & 45.8059 & -90.0799 & Dfb & 1999-2014 & \cite{US-WCr} \\
US-Wkg & GRA & 31.7365 & -109.9419 & Bsk & 2004-2014 & \cite{US-Wkg} \\
ZA-Kru & SAV & -25.0197 & 31.4969 & Unk & 2000-2010 & \cite{ZA-Kru} \\
ZM-Mon & DBF & -15.4378 & 23.2528 & Unk & 2000-2009 & \cite{ZM-Mon} \\

### 2019-09-06 doc/paper/si_template_2019.tex
need to use agu's template for supp material
also I added all of the fluxnet citations to the main mauscript,
per agu's guidelines

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AGUtmpl.tex: this template file is for articles formatted with LaTeX2e,
% Modified December 2018
%
% This template includes commands and instructions
% given in the order necessary to produce a final output that will
% satisfy AGU requirements.
%
% FOR FIGURES, DO NOT USE \psfrag
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% IMPORTANT NOTE:
%
% SUPPORTING INFORMATION DOCUMENTATION IS NOT COPYEDITED BEFORE PUBLICATION.
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Step 1: Set the \documentclass
%
%
% PLEASE USE THE DRAFT OPTION TO SUBMIT YOUR PAPERS.
% The draft option produces double spaced output.
%
% Choose the journal abbreviation for the journal you are
% submitting to:

% jgrga JOURNAL OF GEOPHYSICAL RESEARCH (use for all of them)
% gbc   GLOBAL BIOCHEMICAL CYCLES
% grl   GEOPHYSICAL RESEARCH LETTERS
% pal   PALEOCEANOGRAPHY
% ras   RADIO SCIENCE
% rog   REVIEWS OF GEOPHYSICS
% tec   TECTONICS
% wrr   WATER RESOURCES RESEARCH
% gc    GEOCHEMISTRY, GEOPHYSICS, GEOSYSTEMS
% sw    SPACE WEATHER
% ms    JAMES
% ef    EARTH'S FUTURE
%
%
%
% (If you are submitting to a journal other than jgrga,
% substitute the initials of the journal for "jgrga" below.)

\documentclass[draft,ms]{agutexSI2019}
\usepackage{longtable}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{xcolor}
\usepackage{hyperref}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `basename #1 .tif`.png}
\usepackage{color}
\usepackage{longtable}
\usepackage{textcomp}
\graphicspath{{../../../etc/plots/climate_et/test_plots/}}
\newcommand\newblock{\hskip .11em\@plus.33em\@minus.07em}

\RequirePackage{natbib}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  SUPPORTING INFORMATION TEMPLATE
%
%% ------------------------------------------------------------------------ %%
%
%
%Please use this template when formatting and submitting your Supporting Information.

%This template serves as both a “table of contents” for the supporting information for your article and as a summary of files.
%
%
%OVERVIEW
%
%Please note that all supporting information will be peer reviewed with your manuscript. It will not be copyedited if the paper is accepted.
%In general, the purpose of the supporting information is to enable authors to provide and archive auxiliary information such as data tables, method information, figures, video, or computer software, in digital formats so that other scientists can use it.
%The key criteria are that the data:
% 1. supplement the main scientific conclusions of the paper but are not essential to the conclusions (with the exception of
%    including %data so the experiment can be reproducible);
% 2. are likely to be usable or used by other scientists working in the field;
% 3. are described with sufficient precision that other scientists can understand them, and
% 4. are not exe files.
%
%USING THIS TEMPLATE
%
%***All references should be included in the reference list of the main paper so that they can be indexed, linked, and counted as citations.  The reference section does not count toward length limits.
%
%All Supporting text and figures should be included in this document. Insert supporting information content into each appropriate section of the template. To add additional captions, simply copy and paste each sample as needed.

%Tables may be included, but can also be uploaded separately, especially if they are larger than 1 page, or if necessary for retaining table formatting. Data sets, large tables, movie files, and audio files should be uploaded separately. Include their captions in this document and list the file name with the caption. You will be prompted to upload these files on the Upload Files tab during the submission process, using file type “Supporting Information (SI)”

%IMPORTANT NOTE ON FIGURES AND TABLES
% Placeholders for figures and tables appear after the \end{article} command, after references.
% DO NOT USE \psfrag or \subfigure commands.
%
 \usepackage{graphicx}
%
%  Uncomment the following command to allow illustrations to print
%   when using Draft:
 \setkeys{Gin}{draft=false}
%
% You may need to use one of these options for graphicx depending on the driver program you are using.
%
% [xdvi], [dvipdf], [dvipsone], [dviwindo], [emtex], [dviwin],
% [pctexps],  [pctexwin],  [pctexhp],  [pctex32], [truetex], [tcidvi],
% [oztex], [textures]
%
%
%% ------------------------------------------------------------------------ %%
%
%  ENTER PREAMBLE
%
%% ------------------------------------------------------------------------ %%

% Author names in capital letters:
%\authorrunninghead{BALES ET AL.}

% Shorter version of title entered in capital letters:
%\titlerunninghead{SHORT TITLE}

%Corresponding author mailing address and e-mail address:
%\authoraddr{Corresponding author: A. B. Smith,
%Department of Hydrology and Water Resources, University of
%Arizona, Harshbarger Building 11, Tucson, AZ 85721, USA.
%(a.b.smith@hwr.arizona.edu)}

\begin{document}

%% ------------------------------------------------------------------------ %%
%
%  TITLE
%
%% ------------------------------------------------------------------------ %%

%\includegraphics{agu_pubart-white_reduced.eps}

\title{Supporting Information for "When does vapor pressure deficit drive or reduce
  evapotranspiration?"}
%
% e.g., \title{Supporting Information for "Terrestrial ring current:
% Origin, formation, and decay $\alpha\beta\Gamma\Delta$"}
%
%DOI: 10.1002/%insert paper number here%

%% ------------------------------------------------------------------------ %%
%
%  AUTHORS AND AFFILIATIONS
%
%% ------------------------------------------------------------------------ %%

% List authors by first name or initial followed by last name and
% separated by commas. Use \affil{} to number affiliations, and
% \thanks{} for author notes.
% Additional author notes should be indicated with \thanks{} (for
% example, for current addresses).

% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
% Author\affil{3,4}\thanks{Also funded by Monsanto.}}

\authors{A. Massmann\affil{1}, P. Gentine\affil{1}, C. Lin\affil{1,2}}

\affiliation{1}{Department of Earth and Environmental Engineering,
  Columbia University, New York, NY 10027}
\affiliation{2}{State Key Laboratory of Hydroscience and Engineering, Department of Hydraulic
  Engineering, Tsinghua University, Beijing, CN 100084}

%% ------------------------------------------------------------------------ %%
%
%  BEGIN ARTICLE
%
%% ------------------------------------------------------------------------ %%

% The body of the article must start with a \begin{article} command
%
% \end{article} must follow the references section, before the figures
%  and tables.

\begin{article}

%% ------------------------------------------------------------------------ %%
%
%  TEXT
%
%% ------------------------------------------------------------------------ %%

\noindent\textbf{Contents of this file}
%%%Remove or add items as needed%%%
\begin{enumerate}
\item Figures S1 to S68
\item Table S1
\end{enumerate}

\noindent\textbf{Introduction}

The manuscript analyzes the partial derivative of ET with respect to
VPD. This assumes that all other quantities remain fixed, including the
plant parameters g$_1$ and uWUE (and by extension, $\lambda$). In
reality, these parameters may vary with environmental conditions, and
specifically soil moisture. However because soil moisture
\textit{only} enters the partial derivative directly through these
plant terms, if the plant parameters are weak functions of soil
moisture then the theory can be directly applied to a broader range of
conceptual VPD scenarios, including observed compound events between
high VPD and low soil moisture \cite{Zhou_2019}. To help the reader
assess the soil moisture dependence of uWUE (and partially by
extension $\lambda$ and g$_1$), we provide here figures showing the
distribution of uWUE with SWC for each of 66 FLUXNET sites from the
FLUXNET-2015 database. The functional relationship between uWUE and
SWC varies, with a mix of sites showing strong and weak
SWC-dependence. For all sites the ratio of signal to noise is very
low, an unfortunate consequence of taking a ratio of two highly
uncertain eddy-covariance derived fluxes. In general we find this
analysis inconclusive.

The paper does not rely on assumptions about uWUE's functional
relationship with soil moisture so we do not include the figures in
the body of the manuscript. But given that constant uWUE and g$_1$
assumptions can make our theory more useful to the reader we provide
the figures for their interpretation, and motivation for future
research.

Additionally, we include a figure showing the joint distribution
between saturation vapor pressure and relative humidity calculated
from the FLUXNET-2015 data. Relative humidity and saturation vapor
pressure are much more independent than saturation vapor pressure and
VPD, and we use an assumption that relative humidity and saturation
vapor pressure can be approximated as independent in order to evaluate
$\frac{\partial ET}{\partial VPD}$. Please note that at a given site,
the relationship may be more or less independent depending on the
hydroclimate.

\section{Description of data}
\subsection{Data}
We use both meteorological and eddy-covariance data from the
FLUXNET2015 database (data available at \sloppy
\url{https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/} \sloppy),
including all Tier 1 sites with at least four years of data. Sixty-six
sites met these requirements, and were grouped into nine plant
functional types (PFT) according to the International
Geosphere-Biosphere Programme vegetation classification scheme
\cite{Loveland_1999}: cropland (CRO), grass (GRA), deciduous
broadleaf forest (DBF), evergreen broadleaf forest (EBF), evergreen
needleleaf forest (ENF), mixed forest (MF), closed shrub (CSH),
savannah (SAV), and woody savannah (WSA).

\begin{figure}[h]
  \centering \includegraphics[trim={0 3cm 0 3cm}, clip]{./map.pdf}
  \caption{Plant functional type and location of FLUXNET2015 sites
    used in this analysis.}
  \label{map_fig}
\end{figure}
We filter and quality control the FLUXNET-2015 data using a similar
procedure as \cite{Zhou_2015}:

\begin{itemize}
\item Only measured or highest (``good'') quality gapfilled data,
  according to quality control flags, are used.
\item To isolate the growing season, we only use days in which the
  average Gross Primary Productivity (GPP) exceeds 10\% of the
  observed 95th percentile of GPP for a given site. GPP is calculated
  using the nighttime respiration partitioning method.
\item We remove days with rain and the day following to avoid issues
  with rain interception and sensor saturation at high relative
  humidity (\cite{Medlyn_2017}).
\item For SWC measurements, we use the shallowest observed layer
  available at each site.
\end{itemize}

Additionally, as in \citet{Lin_2018}, we restrict data to the daytime,
which is identified when downwelling shortwave radiation is greater
than 50 W m$^{-2}$ and sensible heat flux is greater than 5 W
m$^{-2}$. To reduce the chance of sensor saturation at high relative
humidity, we remove all time steps for which VPD is less than .01 kPa,
and to reduce errors at low windspeeds we remove all periods with wind
magnitudes less than 0.5 m s$^{-1}$. Timesteps with negative observed
GPP or ET are also removed, and we aggregate half hourly data to
hourly averages to reduce noise \cite{Lin_2018}.  After these quality
control procedures, 400,983 upscaled hourly observations remain.

\bibliography{references}

%\clearpage

%Delete all unused file types below. Copy/paste for multiples of each file type as needed.

%Repeat for any additional Supporting audio files

%%% End of body of article:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Optional Notation section goes here
%
% Notation -- End each entry with a period.
% \begin{notation}
% Term & definition.\\
% Second term & second definition.\\
% \end{notation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% ------------------------------------------------------------------------ %%
%%  REFERENCE LIST AND TEXT CITATIONS

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% \bibliography{<name of your .bib file>} do not specify file extension
%
% no need to specify bibliographystyle
%
% Note that ALL references in this supporting information file must also be referenced in the primary manuscript
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% if you get an error about newblock being undefined, uncomment this line:
%\newcommand{\newblock}{}

% \bibliography{ uncomment this line and enter the name of your bibtex file here }

%Reference citation instructions and examples:
%
% Please use ONLY \cite and \citeA for reference citations.
% \cite for parenthetical references
% ...as shown in recent studies (Simpson et al., 2019)
% \citeA for in-text citations
% ...Simpson et al (2019) have shown...
% DO NOT use other cite commands (e.g., \citet, \citep, \citeyear, \nocite, \citealp, etc.).
%
%
%...as shown by \citeA{jskilby}.
%...as shown by \citeA{lewin76}, \citeA{carson86}, \citeA{bartoldy02}, and \citeA{rinaldi03}.
%...has been shown \cite<e.g.,>{jskilbye}.
%...has been shown \cite{lewin76,carson86,bartoldy02,rinaldi03}.
%...has been shown \cite{lewin76,carson86,bartoldy02,rinaldi03}.
%
% apacite uses < > for prenotes, not [ ]
% DO NOT use other cite commands (e.g., \citet, \citep, \citeyear, \nocite, \citealp, etc.).
%

%% ------------------------------------------------------------------------ %%
%
%  END ARTICLE
%
%% ------------------------------------------------------------------------ %%
\end{article}
\clearpage

% Copy/paste for multiples of each file type as needed.

% enter figures and tables below here: %%%%%%%

\begin{figure}
  \centering \includegraphics{./supp-figs/0joint_rh_es.pdf}
  \caption{The joint distribution of relative humidity and saturation
    vapor pressure for the FLUXNET2015 dataset.}
  \end{figure}

\input{supp-figs.tex}

% tables

  % below taken adapted from Trevor Keenan's FLUXNET citations
  % at github.com/trevorkeenan/FLUXNET_citations
  Table S1: Metadata and citations for flux sites used in this analysis. All data are gathered from \url{www.fluxdata.org}, and citations are aggregated using tools available at \url{https://github.com/trevorkeenan/FLUXNET_citations}.
  \begin{longtable}{l l l l l l l}
    \hline
    \textbf{Site} &
    \textbf{PFT} &
    \textbf{Lat} &
    \textbf{Lon} &
    \textbf{Clim\textsuperscript{1}} &
    \textbf{Period} &
    \textbf{References} \\
    [0.5ex]
    \hline
    \input{flux_sites.tex}
    [1ex]
    \hline
  \end{longtable}
\textsuperscript{1} K{\"o}ppen Climate classification.

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

More Information and Advice:

%% ------------------------------------------------------------------------ %%
%
%  SECTION HEADS
%
%% ------------------------------------------------------------------------ %%

% Capitalize the first letter of each word (except for
% prepositions, conjunctions, and articles that are
% three or fewer letters).

% AGU follows standard outline style; therefore, there cannot be a section 1 without
% a section 2, or a section 2.3.1 without a section 2.3.2.
% Please make sure your section numbers are balanced.
% ---------------
% Level 1 head
%
% Use the \section{} command to identify level 1 heads;
% type the appropriate head wording between the curly
% brackets, as shown below.
%
%An example:
%\section{Level 1 Head: Introduction}
%
% ---------------
% Level 2 head
%
% Use the \subsection{} command to identify level 2 heads.
%An example:
%\subsection{Level 2 Head}
%
% ---------------
% Level 3 head
%
% Use the \subsubsection{} command to identify level 3 heads
%An example:
%\subsubsection{Level 3 Head}
%
%---------------
% Level 4 head
%
% Use the \subsubsubsection{} command to identify level 3 heads
% An example:
%\subsubsubsection{Level 4 Head} An example.
%
%% ------------------------------------------------------------------------ %%
%
%  IN-TEXT LISTS
%
%% ------------------------------------------------------------------------ %%
%
% Do not use bulleted lists; enumerated lists are okay.
% \begin{enumerate}
% \item
% \item
% \item
% \end{enumerate}
%
%% ------------------------------------------------------------------------ %%
%
%  EQUATIONS
%
%% ------------------------------------------------------------------------ %%

% Single-line equations are centered.
% Equation arrays will appear left-aligned.

Math coded inside display math mode \[ ...\]
 will not be numbered, e.g.,:
 \[ x^2=y^2 + z^2\]

 Math coded inside \begin{equation} and \end{equation} will
 be automatically numbered, e.g.,:
 \begin{equation}
 x^2=y^2 + z^2
 \end{equation}

% IF YOU HAVE MULTI-LINE EQUATIONS, PLEASE
% BREAK THE EQUATIONS INTO TWO OR MORE LINES
% OF SINGLE COLUMN WIDTH (20 pc, 8.3 cm)
% using double backslashes (\\).

% To create multiline equations, use the
% \begin{eqnarray} and \end{eqnarray} environment
% as demonstrated below.
\begin{eqnarray}
  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
        && + (y - y_{0}) \sin \Theta  \nonumber \\
  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
        && + (y - y_{0}) \cos \Theta.
\end{eqnarray}

%If you don't want an equation number, use the star form:
%\begin{eqnarray*}...\end{eqnarray*}

% Break each line at a sign of operation
% (+, -, etc.) if possible, with the sign of operation
% on the new line.

% Indent second and subsequent lines to align with
% the first character following the equal sign on the
% first line.

% Use an \hspace{} command to insert horizontal space
% into your equation if necessary. Place an appropriate
% unit of measure between the curly braces, e.g.
% \hspace{1in}; you may have to experiment to achieve
% the correct amount of space.

%% ------------------------------------------------------------------------ %%
%
%  EQUATION NUMBERING: COUNTER
%
%% ------------------------------------------------------------------------ %%

% You may change equation numbering by resetting
% the equation counter or by explicitly numbering
% an equation.

% To explicitly number an equation, type \eqnum{}
% (with the desired number between the brackets)
% after the \begin{equation} or \begin{eqnarray}
% command.  The \eqnum{} command will affect only
% the equation it appears with; LaTeX will number
% any equations appearing later in the manuscript
% according to the equation counter.
%

% If you have a multiline equation that needs only
% one equation number, use a \nonumber command in
% front of the double backslashes (\\) as shown in
% the multiline equation above.

%% ------------------------------------------------------------------------ %%
%
%  SIDEWAYS FIGURE AND TABLE EXAMPLES
%
%% ------------------------------------------------------------------------ %%
%
% For tables and figures, add \usepackage{rotating} to the paper and add the rotating.sty file to the folder.
% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
%
% \begin{sidewaysfigure}
% \includegraphics[width=20pc]{samplefigure.eps}
% \caption{caption here}
% \label{label_here}
% \end{sidewaysfigure}
%
%
%
% \begin{sidewaystable}
% \caption{}
% \begin{tabular}
% Table layout here.
% \end{tabular}
% \end{sidewaystable}
%
%

### 2019-09-18 doc/paper/ms.tex
spell check done and makefile updated
now proofread to send off to Pierre and Changjie
updated arxiv manuscript with published version
still need to test makefile on fresh pull (probably doesn't work
due to all the changes), and clean up repository (lots of old and
unused files laying around).

@@
 \maketitle
+
 \begin{abstract}
-  \input{abstract}
+Increasing vapor pressure deficit (VPD) increases atmospheric demand
+for water. While increased evapotranspiration (ET) in response to
+increased atmospheric demand seems intuitive, plants are capable of
+reducing ET in response to increased VPD by closing their stomata. We
+examine which effect dominates the response to increasing VPD:
+atmospheric demand and increases in ET, or plant response (stomata
+closure) and decreases in ET. We use Penman-Monteith, combined with
+semi-empirical optimal stomatal regulation theory and underlying water
+use efficiency, to develop a theoretical framework for assessing ET
+response to VPD. The theory suggests that depending on the environment
+and plant characteristics, ET response to increasing VPD can vary from
+strongly decreasing to increasing, highlighting the diversity of plant
+water regulation strategies.
+
+The ET response varies due to: 1) climate, with tropical and temperate
+climates more likely to exhibit a positive ET response to increasing
+VPD than boreal and arctic climates; 2) photosynthesis strategy, with
+C3 plants more likely to exhibit a positive ET response than C4
+plants; and 3) plant type, with crops more likely to exhibit a
+positive ET response, and shrubs and gymniosperm trees more likely to
+exhibit a negative ET response. These results, derived from previous
+literature connecting plant parameters to plant and climate
+characteristics, highlight the utility of our simplified framework for
+understanding complex land atmosphere systems in terms of idealized
+scenarios in which ET responds to VPD only. This response is otherwise
+challenging to assess in an environment where many processes co-evolve
+together.
 \end{abstract}
@@
 
-\input{intro.tex}
+Vapor pressure deficit (VPD) is expected to rise over continents in
+the future due to the combination of increased temperature and,
+depending on region, decreased relative humidity
+\cite{Byrne_2013}. Increases in VPD increase the atmospheric demand
+for evapotranspirated water \cite{Penman_1948, Monteith_1965}, but
+also reduce stomatal conductance through stomatal closure
+\cite{Rawson1977, Leuning_1990, Mott2007, Damour2010,
+MEDLYN_2011}. Understanding the net evapotranspiration (ET) response
+to these two opposing effects of changes in VPD is crucial for
+assessing the impact of environmental VPD perturbations on the water
+cycle.
 
-\input{body.tex}
+The opposing effects of increased atmospheric demand and higher
+stomatal closure lead to two possible perspectives for how ET responds
+to shifts in VPD. The first, a hydrometeorological perspective, is
+that higher VPD increases atmospheric demand for water from the land
+surface, and this drives an increase in ET
+\cite{Penman_1948}. However, plants' stomata have evolved to
+optimally regulate the exchange of water and carbon, and tend to
+partially close in response to increased atmospheric dryness
+\cite{Farquhar_1978, Ball_1987, Leuning_1990, Katul_2009,
+MEDLYN_2011}. This leads to a plant physiology perspective, in which
+an increase in VPD may actually correspond to a decrease in ET because
+of stomatal closure \cite<e.g.>[]{Rigden_2017}.  In other words, the
+question ``When does VPD drive or reduce ET?'' can be related to
+whether plant regulation or atmospheric demand dominates the ET
+response.
+
+The ET response to changes in VPD alters water partitioning between
+the soil and atmosphere. If ecosystem plant response reduces ET with
+atmospheric drying then soil moisture will be better conserved. This
+represents a sensible evolutionary strategy to cope with aridity: save
+water for periods when atmospheric demand for water is relatively low,
+and atmospheric carbon can be accessed with a relatively smaller cost
+in water loss. If instead stomata were fully passive \cite<similar to
+soil pores, e.g. >[]{Or_2013}, increased atmospheric aridity would
+strongly reduce soil moisture \cite{Berg_2017}. This could further
+increase aridity as low soil moisture levels increases the Bowen
+ratio, leading to increased temperature and atmospheric drying
+\cite{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006,
+  Salvucci_2013, Gentine_2016, Berg_2016, Zhou_2019}. Therefore,
+passive regulation and a lack of soil moisture conservation does not
+seem to be a sensible strategy for plants from an evolutionary
+standpoint. This simplified logic explains generally why plants
+evolved to respond to VPD, but also excludes many details and special
+cases \cite<e.g. plant to plant interaction, thermal regulation with
+transpiration, and highly specialized photosynthesis strategies like
+Crassulacean acid metabolism photosynthesis; see>[]{Brooker_2006, Lin_2017,
+  Cushman_2001}.
+
+We can use intuition about plant water conservation strategy to
+hypothesize about ET response to changes in VPD. Plants and ecosystems
+that evolved to conserve water, such as arid shrubs, should be more
+likely to reduce ET with increasing VPD, and plants that have evolved
+or have been engineered to prioritize carbon gain over water
+conservation, such as crops, will be more likely to increase ET with
+increasing VPD. Atmospheric conditions must matter as well. At the
+ecosystem scale, there are limits to plant water conservation
+strategies. As atmospheric demand for water (VPD) increases,
+ecosystems may begin to reach their water conservation limits and
+might not be able to entirely limit ET flux to the atmosphere. At this
+stage any further increase in VPD will most likely drive a (limited)
+increase in ET, because the increase in atmospheric demand for water
+overwhelms the limited plant response to conserve water.
+
+The objective of the present manuscript is to use reasonable
+approximations established in prior research as a tool to develop a
+framework for understanding plant responses to atmospheric drying and
+the VPD dependence of ET while keeping other variables fixed. This
+framework will aid interpretation of observations, full complexity
+models, and facilitate the disentanglement of complex land-atmosphere
+feedbacks. In particular, our approach has applications for
+understanding climate change impacts, given expected increases in VPD
+with rising temperature.  In the past, similar simplified approaches
+were used to understand interactions between stomatal conductance,
+evapotranspiration and the environment \cite<e.g.,>[]{Jarvis_1984,
+Jarvis_1986, Mcnaughton_1991}. However these researchers did not
+explore explicitly the sensitivity of ET to VPD, including VPD's
+effect on stomatal conductance and plant function.
+
+Approaching the problem of ET response to VPD is aided by recent
+results drastically improving our understanding of VPD's impact on
+physiology, especially at the leaf level. \citeA{MEDLYN_2011}
+developed a model for leaf-scale stomatal conductance ($g_s$),
+including VPD response, by combining an optimal photosynthesis theory
+\cite{Cowan_1977, Katul_2009, Katul_2010} with an empirical approach,
+and extended use of this model to the ecosystem scale in
+\citeA{Medlyn_2017}. Additionally, \citeA{Zhou_2014} demonstrated that
+a quantity underlying water use efficiency $\left(uWUE = \frac{GPP\;
+\sqrt{VPD}}{ET}\right)$ properly captures a constant relationship
+between GPP, ET, and VPD over a diurnal cycle at the ecosystem
+scale. uWUE is also relatively well conserved in the growing season
+across space and time, within a plant functional type (PFT)
+\cite{Zhou_2015}.  While stomatal conductance parameterizations and
+uWUE greatly simplify complex plant physiological processes, they
+ still capture ecosystem behavior for vegetated surfaces
+\cite{Medlyn_2017, Zhou_2014}, and are useful tools to transparently
+develop intuition for the behavior of complex land-atmosphere systems.
+
+In this manuscript, we leverage uWUE and recent developments in
+stomatal conductance parameterizations \cite{MEDLYN_2011} with a
+Penman-Monteith framework \cite<hereafter PM,>[]{Penman_1948,
+Monteith_1965} to derive the theoretical one-way response of ET to VPD
+with other environmental variables properly controlled for, i.e. we
+develop a framework for evaluating the partial derivative of ET with
+respect to VPD. It is useful to disconnect the impact of VPD from
+other variables as VPD is known to increase dramatically with future
+climate and limit ET more than soil moisture \cite{Novick_2016}, but
+disentangling VPD effects from soil moisture effects has been
+difficult in previous research, given their co-variability
+\cite{Lin_2018, Zhou_2019}.  We are
+able to disentangle the effect of VPD on ET because we explicitly
+include VPD's full effect on stomatal conductance, including its
+impact on photosynthesis. The use of an energy balance framework (PM)
+allows us to include in our analysis the effects of the energy cost of
+evaporating water from a surface, which is an important factor in the
+natural environment, compared to prescribed in situ environmental
+conditions. This is relevant because previous research focusing on the
+leaf scale \cite{Rawson1977, Turner1984, Oren1999, Damour2010,
+Mott2013} does not consistently or analogously include the energetic
+constraints on evapotranspiration under which ecosystems operate. The
+leaf-scale results agree that stomatal conductance decreases in
+response to VPD \cite{Oren1999, Damour2010}, which we expect to be
+true for an ecosystem as well. However leaf scale results also
+indicate that transpiration usually increases with increasing VPD in a
+concave downward shape \cite<e.g.,>[]{Rawson1977, Turner1984,
+Mott2013}, which may not be true for ET at the ecosystem scale once
+energetic constraints on ET are included in the analysis.  Our
+approach allows us to estimate the expected ET response to VPD at the
+ecosystem scale, including the effects of surface energy balance
+constraints, and assess how ecosystem ET response to VPD deviates from
+previous leaf scale analyses.
+
+This manuscript presents the range of possible ecosystem-scale ET
+responses to VPD, given parameters previously established in peer
+reviewed literature. Additionally, we explore the sensitivity of the
+ET-VPD relationship to stomatal model and framework choice,
+highlighting the importance of: 1) future research on stomatal
+conductance and ecosystem scale modeling, and 2) thoughtful selection
+of photosynthesis and stomatal conductance models in more
+sophisticated land surface and earth system models.
+
+\section{Methods}
+\label{methods}
+The Penman-Monteith equation \cite<hereafter PM,>[]{Penman_1948,
+  Monteith_1965} estimates ET as a function of observable atmospheric
+variables and surface conductances:
+% \begin{linenomath*}
+  \begin{equation}
+    \label{orig_pen}
+    ET = \frac{\Delta R_{net} + g_a \rho_a c_p VPD}{\Delta + \gamma(1 + \frac{g_a}{g_s})},
+  \end{equation}
+% \end{linenomath*}
+  where $\Delta$ is the change in saturation vapor pressure with
+  temperature, given by Clausius-Clapeyron
+  ($\frac{d \; e_s}{d \; T}$), $R_{net}$ is the net radiation minus
+  ground heat flux, $g_a$ is aerodynamic conductance, $\rho_a$ is air
+  density, $c_p$ is specific heat of air at constant pressure,
+  $\gamma$ is the psychometric constant, and $g_s$ is the stomatal
+  conductance (Table \ref{definitions}). The issue with PM is that
+  $g_s$ is a function of carbon uptake, which is has a strong
+  functionally relation to ET through stomatal function. Therefore,
+  when PM is formulated in terms of $g_s$, it is an implicit function
+  of ET itself rather than an explicit function of ET. Here we will
+  derive a new form of PM in which ET is an explicit function of plant
+  parameters and environmental conditions, and use it to assess the
+  ecosystem scale response to VPD (by taking a partial derivative).
+
+\begin{table}
+  \caption{Definition of symbols and variables, with citation for how
+    values are calculated, if applicable.}
+  \label{definitions}
+  \centering \footnotesize
+  \begin{adjustbox}{center=15cm}
+    \begin{tabular}{l c c c}
+    \hline
+    Variable & Description & Units & Citation \\
+    \hline
+    $e_s$  & saturation vapor pressure & Pa  & - \\
+    $T$  & temperature  & K & - \\
+    $P$  & pressure & Pa  & - \\
+    $\Delta$  & $\frac{\partial e_s}{\partial T}$ & Pa K$^{-1}$ & - \\
+    $R_{net}$  & net radiation at land surface minus ground heat flux & W m$^{-2}$   & - \\
+    $g_a$  & aerodynamic conductance & m s$^{-1}$  & \citeA{Shuttleworth_2012} \\
+    $\rho_a$  & air density & kg m$^{-3}$  & - \\
+    $c_p$  & specific heat capacity of air at constant pressure & J K$^{-1}$ kg$^{-1}$ & - \\
+    $VPD$  & vapor pressure deficit & Pa  & - \\
+    $\gamma$  & psychometric constant & Pa K$^{-1}$   & - \\
+    $g_{s}$  &  stomatal conductance & m s$^{-1}$
+                                   & \citeA{Medlyn_2017} \\
+    $g_{1}$  & ecosystem-scale slope parameter & Pa$^{0.5}$ & \citeA{Medlyn_2017} \\
+    $R$ & universal gas constant & J mol$^{-1}$ K$^{-1}$ & - \\
+    $R_{air}$ & gas constant of air & J  K$^{-1}$ kg$^{-1}$ & - \\
+    $\sigma$ & uncertainty parameter & -& - \\
+    $c_a$ & CO$_2$ concentration & $\mu$ mol CO$_2$ mol$^{-1}$ air& - \\
+    $\lambda = \frac{\partial \; transpiration}{\partial\; CO_2\; assimilation}$ & marginal water cost of leaf carbon & mol H$_2$O mol$^{-1}$ CO$_2$ & - \\
+    $\Gamma$ & CO$_2$ compensation point & - & - \\
+    $\Gamma^*$ & CO$_2$ compensation point without dark respiration &
+                                                                      - & - \\
+    GPP & gross primary production & $\mu$-mol C s$^{-1}$ m$^{-2}$ & -
+    \\
+    ET & evapotranspiration & W m$^{-2}$ & - \\
+   uWUE & underlying water use efficiency & $\mu$-mol C Pa$^{0.5}$
+                                            J$^{-1}$ ET  &
+                                                           \citeA{Zhou_2015} \\
+    \hline
+  \end{tabular}
+\end{adjustbox}
+\end{table}
+
+\citeA{MEDLYN_2011} developed a model for stomatal conductance ($g_s$)
+by combining an optimal photosynthesis theory \cite{Cowan_1977} with
+an empirical approach, which describes the dependence of $g_s$ to
+VPD. They also extended this model to the ecosystem scale
+\cite{Medlyn_2017}:
+
+% \begin{linenomath*}
+  \begin{equation}
+    g_s = \frac{R \,T}{P} \; 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{GPP}{c_a},
+    \label{medlyn}
+  \end{equation}
+% \end{linenomath*}
+
+  where $g_{1}$ is a VPD ``slope'' parameter, GPP is the ecosystem
+  scale gross primary production, and $c_a$ is the atmospheric CO$_2$
+  concentration at the canopy. \cite{MEDLYN_2011} relate the slope
+  parameter ($g_{1}$) to physical parameters as:
+% \begin{linenomath*}
+
+  \begin{equation}
+    g_{1}  \propto \sqrt{\frac{3 \, \Gamma^* \,
+        \lambda}{1.6}},
+    \label{slope}
+  \end{equation}
+% \end{linenomath*}
+
+  where g$_1$ is presented in units of $\sqrt{\text{mol fraction}}$ for
+  convenience (which can be converted to $\sqrt{Pa}$ using the ideal
+  gas law and ambient pressure), $\Gamma^*$ is the CO$_2$ compensation
+  point for photosynthesis (without dark respiration), and $\lambda$
+  is the marginal water cost of leaf carbon
+  ($\frac{\partial \; \text{transpiration}}{\partial \; CO_2
+    assimilation}$) \cite{Farquhar_1980, Katul_2009}. So, $g_{1}$ is a
+  leaf-scale term reflecting the trade-off of water for carbon
+  uptake. The higher $g_{1}$, the more open the stomata and the more
+  they release water in exchange for carbon.
+
+While Eq. (\ref{medlyn}) can be used in PM (Eq. (1)), it will make
+analytical work with the function intractable because GPP is
+functionally related to ET itself. Additionally, a perturbation to VPD
+should induce a physiological plant response that will alter GPP and
+cause an indirect change in stomatal conductance, in addition to the
+direct effect of VPD in Eq. (\ref{medlyn}). Therefore, in order to
+derive the full plant response of ET to VPD, we must account for the
+functional relationship between GPP, ET, and VPD, and its effect on
+stomatal conductance. We can use aforementioned semi-empirical results
+of \citeA{Zhou_2015}, which were inspired by optimal photosynthesis
+theory, as a tool to approach this problem. \citeA{Zhou_2015}, showed
+that underlying Water Use Efficiency (uWUE):
+
+% \begin{linenomath*}
+  \begin{equation}
+    uWUE = \frac{GPP \cdot \sqrt{VPD}}{ET}
+    \label{uwue}
+  \end{equation}
+% \end{linenomath*}
+
+  is relatively constant across time within a plant functional type,
+  and correctly captures a constant relationship between GPP, ET and
+  VPD over a diurnal cycle during the growing season
+  \cite{Zhou_2014}. The theoretical derivation of the square root VPD
+  dependence in uWUE leverages the same assumptions used in
+  \citeA{MEDLYN_2011} to derive the square-root VPD dependence of the
+  stomatal conductance model (Eq.  (\ref{medlyn})).  We can use uWUE
+  to remove the $GPP$ dependence in $g_s$ in a way that makes PM
+  analytically tractable:
+
+% \begin{linenomath*}
+  \begin{equation}
+    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{uWUE \; ET}{c_a \; \sqrt{VPD}}.
+    \label{new_g_s}
+  \end{equation}
+% \end{linenomath*}
+
+Plugging Eq. (\ref{new_g_s}) into Eq. (\ref{orig_pen}) and
+rearranging gives a new explicit expression for PM, in which
+dependence on $GPP$ is removed:
+
+% \begin{linenomath*}
+  \begin{equation}
+    ET = \frac{\Delta R_{net} + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R \; 1.6\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
+    \label{et}
+  \end{equation}
+% \end{linenomath*}
+
+  By accounting for photosynthesis changes in ecosystem conductance,
+  with Eq. (\ref{et}) we have used recent results
+  \cite{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017} to derive
+  ET explicitly as function of environmental variables and two
+  plant-specific constants, the slope parameter ($g_1$), and uWUE. For
+  the first time we have removed the implicit dependence of ET on
+  itself through the stomatal conductance term, and we have also
+  replaced the added complexity of a stomatal conductance reduction
+  factor and a photosynthesis model with a single parameter (uWUE). Both
+  the plant parameters reflect water conservation strategy. The
+  slope parameter is related to the willingness of stomata to trade
+  water for CO$_2$ and to keep stomata open (carbon cost in terms of
+  water)). uWUE is a semi-empirical ecosystem-scale constant related
+  to how WUE changes with VPD (specifically $VPD^{-1/2}$). It is also
+  roughly proportional to physical constants:
+
+\[uWUE \appropto \sqrt{\frac{c_a - \Gamma}{1.6 \lambda}},\]
+
+where $\Gamma$ is the CO$_2$ compensation point \cite<Eq. (5)
+in>[]{Zhou_2014}. So uWUE is related to atmospheric CO$_2$
+concentration and compensation point, and is inversely proportional to
+the marginal water cost of leaf carbon ($\lambda$). This relationship
+with $\lambda$ ($\propto \lambda^{-1/2}$) is important as it is the inverse of g$_1$'s
+relationship with $\lambda$ ($\propto \sqrt{\lambda}$). So, uWUE
+should increase as $\lambda$ decreases, and g$_1$ should decrease as
+$\lambda$ decreases.
+
+With Eq. (\ref{et}) we can take the partial derivative of ET with respect
+to VPD to understand whether VPD drives or reduces ecosystem-scale ET:
+
+% \begin{linenomath*}
+  \begin{equation}
+    \frac{\partial \;  ET}{\partial \; VPD} = \frac{2\; g_a \;
+      P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} -
+      \frac{\gamma c_a }{1.6 \; R\; \; \text{ uWUE }} \left(
+        \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right)
+    \right),
+    \label{d_et}
+  \end{equation}
+% \end{linenomath*}
+
+  providing analytical framework for ecosystem response to atmospheric
+  demand with environmental conditions held fixed. There are a few
+  subtleties to taking the derivative in Eq. (\ref{d_et}): $\Delta$
+  ($\frac{d e_{s}}{d T}$) and $VPD$ are functionally related, so while
+  taking the derivative we evaluate
+  $\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET}
+  {\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD}
+  \Big|_{\text{RH fixed}} + \frac{\partial \; ET}{\partial \; RH}
+  \frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$
+      fixed}}$. $RH$ and $e_s$ are assumed to be approximately
+  independent, which is supported by data (see supplementary material).
+
+  This derivation relied either implicitly or explicitly on several
+  assumptions. First, we assume that VPD at the leaf surface is the
+  same as VPD at measurement height; physically this implies that
+  leaves are perfectly coupled to the atmosphere. In reality, for some
+  conditions and plant types the leaves can become decoupled from the
+  boundary layer \cite{De_2017, Medlyn_2017}. Therefore, our
+  derivation will be most applicable in times like the growing season
+  (when we also expect uWUE to be most valid), when relatively high
+  insolation induces instability and convective boundary layers, and
+  we would expect the surface to be generally well coupled. An
+  additional assumption in the formulation of uWUE \cite{Zhou_2014,
+    Zhou_2015} and the stomatal conductance model of
+  \citeA{Medlyn_2017} is that direct soil evaporation contributions to
+  ET remain small relative to transpiration. Again, this should be
+  more true during the growing season. The ratio of evaporation to
+  transpiration may increase immediately after rainfall events due to
+  high soil moisture, ponding, and interception, but VPD is generally
+  low anyways during these times. However, some plant types allow for
+  systematically larger contributions of evaporation in ET,
+  particularly those with sparse canopies and smaller relative amounts
+  of transpiration. We therefore might expect that the theory will be
+  most applicable to forest PFTs, which will be most strongly coupled
+  to the boundary layer due to larger surface roughness, and will also
+  generally have the highest ratios of transpiration to
+  evaporation. Finally, we assume that $g_1$ and uWUE are constant
+  with respect to the conceptual VPD perturbation. Both quantities
+  have been shown to be relatively constant with respect to changes in
+  VPD \cite{Franks_2017, Zhou_2014}. These parameters will however
+  vary with plant species and characteristics \cite<e.g. wood density,
+  >[]{Lin_2015}, as well as environmental conditions including soil
+  water content and temperature \cite{Lin_2015,
+    Manzoni2013}. Exploring possible soil moisture (in)dependence of
+  the plant parameters (g$_1$ and uWUE) is particularly interesting
+  because soil moisture only enters the partial derivative directly
+  through these plant parameters. If the plant parameters are weak
+  functions of soil moisture then the theory can be directly applied
+  to a broader range of conceptual VPD scenarios, including observed
+  compound events between high VPD and low soil moisture
+  \cite{Zhou_2019}. Supplementary material for this manuscript
+  explores the soil moisture dependence of the plant parameters, but
+  excessive noise and inconsistencies preclude conclusions about the
+  nature of any dependence. We provide it in case it is of interest to
+  the reader, and to motivate future research.
+
+  Because Eq. (\ref{d_et}) is a partial derivative, its utility is as
+  a conceptual model for the change in ET in response to VPD with all
+  other variables held fixed. This provides a useful tool for
+  identifying the effect of VPD on ET through atmospheric demand and
+  plant response, and allows a practitioner to disentangle complicated
+  feedbacks when many quantities co-vary. However users of
+  Eq. (\ref{d_et}) should take care that their interpretation matches
+  the assumptions inherent in a partial derivative. For example,
+  results will only be as valid as the assumption that g$_1$ and uWUE
+  (and by extension $\lambda$) are fixed with respect to the user's
+  conceptual change in VPD. Care must also be taken with possible
+  indirect effects associated with a change in VPD: for example, a
+  change in ET induced by a change in VPD can cause a change in
+  surface temperature, which would drive a change in net
+  radiation. These types of indirect effects and feedbacks are not
+  considered in Eq. (\ref{d_et}): temperature (a variable) is
+  mathematically fixed.
+
+\subsection{Framing Eq. \ref{d_et} with previous research }
+
+  To account for both the variability in plant parameters and the
+  environment, we will systematically analyze how the ET response to
+  VPD (Eq. (\ref{d_et})) varies. Eq. (\ref{d_et}) includes a
+  ``sign'' term that determines the sign of the response in addition
+  to magnitude:
+
+\begin{equation}
+  \label{sign}
+  \frac{c_p}{R_{air}} - \frac{\gamma c_a }{1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right),
+\end{equation}
+
+and a ``scaling'' term multiplying the ``sign'' term:
+
+\begin{equation}
+  \frac{g_a \; P}{T(\Delta + \gamma)}.
+\end{equation}
+
+In the ``sign term'' most of the quantities are relatively constant,
+except for VPD and the plant parameters g$_1$ and uWUE. In the scaling
+term, most of the terms are relatively constant with the exception of
+aerodynamic conductance ($g_a$) and temperature (especially its effect
+through $\Delta$). To determine the range of probable ET responses to
+VPD we will systematically vary these parameters according to Table
+\ref{param_varying}, while all other parameters are held fixed (Table
+\ref{param_fixed}). Physical variables ($g_a$, $T$) and plant
+physiological parameters (g$_1$, uWUE) are varied according to
+literature-based expectations for a range of growing season conditions
+and plant types \cite{Zhou_2015, Medlyn_2017}. Using this previous
+literature we can connect the effect of varying plant parameters to
+specific plant types and characteristics.
+
+All code and data used in this analysis, including those used to
+generate the figures and tables, are publicly available at
+\url{https://github.com/massma/climate_et}.
+%\hyperref[https://github.com/massma/climate\_et]{https://github.com/massma/climate_et}.
+
+\begin{table}
+  \caption{Variable quantities in the ET response to VPD. Each value
+    is varied to determine the effect of a range of expected plant and
+    environmental conditions on ET response to VPD. A citation is
+    provided for values of g$_1$ and uWUE, which are directly derived
+    from previous literature. Conceptually, min. values are extracted
+    from literature to correspond to approximately the 15th percentile
+    of observed conditions during the growing season, med. values
+    correspond to approximately the 50th percentile, and max. values
+    correspond to approximately the 85th percentile. Values for $T$
+    and g$_a$ are calculated from FLUXNET-2015 data (see supplemental
+    material for description), rounding the 15th, 50th, and 85th
+    percentile to the nearest 5$^o$C and 0.005 m/s, respectively.}
+  \label{param_varying}
+  \centering
+  \footnotesize
+  \begin{adjustbox}{center=15cm}
+  \begin{tabular}{l c c c c c}
+    \hline
+    Symbol & units [units in citation] & min  & med & max
+    & citation  \\
+    \hline
+g$_1$ & Pa$^{1/2}$ [kPa$^{1/2}$] & 63.25 [2.00] & 126.49 [4.00] & 189.74 [6.00] & Fig. 2, 7; \citeA{Medlyn_2017} \\
+uWUE & $\mu$-mol C Pa$^{0.5}$ J$^{-1}$ ET [g C hPa$^{1/2}$ kg$^{-1}$ H$_2$O] & 2.33 [6.99] & 3.17 [9.52] & 4.01 [12.05] & Table 4; \citeA{Zhou_2015} \\
+T & $^o$C & 10.00 & 20.00 & 30.00 & - \\
+g$_a$ & m/s & 0.015 & 0.035 & 0.055 & - \\
+    \hline
+  \end{tabular}
+  \end{adjustbox}
+\end{table}
+
+\begin{table}
+  \caption{Quantities that are fixed in the ET response to VPD
+    (relative to those in Table \ref{param_varying}).}
+  \label{param_fixed}
+  \centering
+  \begin{tabular}{l c c}
+    \hline
+    Symbol & units & value \\
+    \hline
+P & Pa & 100000.00 \\
+$\gamma$ & Pa K$^{-1}$ & 64.50 \\
+R$_{air}$ & J  K$^{-1}$ kg$^{-1}$ & 288.00 \\
+c$_a$ & $\mu$ mol CO$_2$ mol$^{-1}$ air & 400.00 \\
+    \hline
+  \end{tabular}
+\end{table}
+
+\section{Results}
+\label{results}
+
+By varying four parameters (g$_a$, T, uWUE, g$_1$) at three different
+values (Table \ref{param_varying}) we generate nine different values
+for the scaling term, nine different curves (as a function of VPD) for
+the sign term, and 81 different curves for the ET response to VPD
+($\frac{\partial \; ET}{\partial \; VPD}$). To aid visualization we
+can examine a subset of nine of these 81 curves, defined by the
+minimum, median and maximum values for both the scaling and sign term
+(Figure \ref{full}). The range of ET responses to VPD vary from those
+where ET almost always decreases with increasing VPD (water
+conservative), to those where ET almost always increases with
+increasing VPD (water intensive). Additionally, for some parameters
+whether ET will increase or decrease with increasing VPD depends on
+atmospheric demand (Figure \ref{full}).
+
+\begin{figure}
+  \centering \includegraphics{./fully_idealized.pdf}
+  \caption{The functional form of $\frac{\partial \; ET}{\partial
+      \; VPD}$ for minimum, median and maximum values of both the sign
+    term and the scaling term.}
+  \label{full}
+\end{figure}
+
+\begin{figure}
+  \centering \includegraphics{./fully_idealized_sign.pdf}
+  \caption{The functional form of $\frac{\partial \; ET}{\partial
+      \; VPD}$ evaluated at the median value of the scaling term, for
+    varying values of g$_1$ and uWUE as given in Table \ref{param_varying}.}
+  \label{sign}
+\end{figure}
+
+Examining the sign term independent of the scaling term illuminates
+the role of the two plant parameters, g$_1$ and uWUE, in determining
+the degree to which the response is water conservative
+($\frac{\partial \; ET}{\partial \; VPD} < 0$) or water intensive
+($\frac{\partial \; ET}{\partial \; VPD} > 0$) (Figure \ref{sign})
+. Higher g$_1$ and uWUE shift the curve towards increasing ET
+responses with VPD (water intensive), and smaller g$_1$ values lead to
+a larger VPD dependence of the response (i.e. ET response is a
+stronger function of VPD). However care should be exercised when
+interpreting the range of ET responses, because g$_1$ and uWUE should
+generally be anti-correlated due to their dependencies on
+$\lambda$. As $\lambda$ increases, g$_1$ should increase and uWUE
+should decrease. Because high and low values of uWUE and g$_1$ were
+determined independently from each other in previous literature, plant
+types with anomalously high or low values for \textit{both} g$_1$ and
+uWUE should be relatively rare. However, some plant types do exhibit
+co-occurring high values for g$_1$ and uWUE. For example, C3 crops
+have both the highest g$_1$ value in \citeA{Franks_2017}, and the
+highest uWUE value in \citeA{Zhou_2015}.
+
+\begin{figure}
+  \centering \includegraphics{./fully_idealized_scaling.pdf}
+  \caption{The functional form of $\frac{\partial \; ET}{\partial
+      \; VPD}$ evaluated at the median value of the sign term, for
+    varying values of g$_a$ and T as given in Table \ref{param_varying}.}
+  \label{scaling}
+\end{figure}
+
+Examining the scaling term independent of the sign term shows how
+aerodynamic conductance (communication between the atmosphere and the
+surface) and temperature (controls the efficiency of energy conversion
+to latent heat through $\Delta$), amplify or suppress the plant
+response represented in the sign term (Figure \ref{scaling}). Both
+lower temperatures and higher aerodynamic conductance lead to
+amplified ET response to VPD, with the variability of aerodynamic
+conductance resulting in a slightly higher variability of ET response
+relative to temperature.
+
+\section{Discussion}
+\label{discussion}
+
+Interpreting both the derivation and results of this manuscript hinges
+on an understanding and appreciation for the usefulness of partial
+derivatives for understanding the behavior of complex systems in a
+simplified framework. Our results are a conceptual tool for answering
+the question, ``What is the ET response to VPD with all other
+quantities held fixed?'' This is useful, and we would argue is
+critical, for understanding complex responses when many quantities are
+varying simultaneously \cite{Zhou_2019}. If we cannot understand the
+contribution of one term to observed variability, independent of other
+confounding factors, there is little hope of disentangling and
+understanding the relative role of many co-varying quantities in
+determining observed variability.  Eq. (\ref{d_et}) explicitly
+provides an estimate of the ET response to VPD, given assumptions
+about ecosystem plant characteristics through the parameters g$_1$ and
+uWUE. Eq. (\ref{d_et}) is generic to external environmental factors
+and the timescale over which a VPD response acts, and has a
+straightforward conceptual interpretation. It provides the ET
+response, given other quantities and parameters held fixed, subject to
+the assumptions outlined in Sect. \ref{methods}. So, while the plant
+parameters g$_1$ and uWUE may vary with plant species and
+environmental conditions like soil moisture, we can still assess the
+ET response to VPD for a given ecosystem state, or for a given soil
+moisture condition.
+
+However, given that g$_1$ and uWUE can vary, and they both have a
+large impact on the ET response (Figure \ref{full}, \ref{sign}), it is
+useful to examine how these quantities vary in previous literature,
+framed by our results on $\frac{\partial \; ET}{\partial \; VPD}$. In
+\citeA{Zhou_2015} most plant types' uWUE is similar to our median
+value, under which the sign of the ET response depends on
+VPD. However, crops, shrubs, and grass are exceptions; they exhibit an
+uWUE different than the median value in Table
+\ref{param_varying}. Crops, which we expect to prioritize carbon
+uptake over water conservation, have a higher uWUE value closer to our
+maximum value. This higher uWUE results in a higher likelihood for an
+increasing ET response to VPD, which matches intuition given that we
+expect crops to keep stomata open for access to carbon, at the cost of
+increased water loss during high VPD. Shrubs, and to a lesser extent
+grass, have an uWUE closer to the minimum uWUE. For these plant types,
+we would then expect a decrease of ET in response to VPD. It is
+important to note that, while \citeA{Zhou_2015} did not examine the
+role of soil moisture for within plant-type variability of uWUE, we
+might expect some variation, especially in extreme cases when soil
+water becomes a limiting factor (see supplementary material).
+
+The ecosystem-scale results of \citeA{Medlyn_2017} did not display the
+robust relationship between plant type and and plant parameter (g$_1$)
+that was demonstrated by \citeA{Zhou_2015} for uWUE. Given that there
+is a robust relationship between g$_1$ and plant type in the
+leaf-scale results of \citeA{Lin_2015}, and the ecosystem-scale
+results of \citeA{Zhou_2015} rely on the same underlying theory as in
+\citeA{Medlyn_2017}, the ambiguous results in \citeA{Medlyn_2017}
+could be due to noise in ecosystem scale observations, the
+consequences of imposing a model structure in some of their g$_1$
+estimations, or real ecosystem-scale within-plant type variability
+(e.g. multiple species impacting the relationship within the
+eddy-covariance footprint). These differences between leaf-scale
+behavior and ecosystem scale behavior highlight the importance of
+understanding how and why slope parameters show increased variability
+at the ecosystem scale, as argued by \citeA{Medlyn_2017}. However, if
+we assume that there is some real analogy between leaf scale results
+in \citeA{Lin_2015} and ecosystem scale behavior, and that some of the
+ambiguity in \citeA{Medlyn_2017} is due to model and/or observational
+error, then we can use the relationships between g$_1$ and plant
+characteristics defined by \citeA{Lin_2015} to frame the ET response
+to VPD. This is subject to the strong caveat that we still do not
+fully understand g$_1$ behavior at the ecosystem scale. Extrapolating
+the results of \citeA{Lin_2015} (e.g. Fig. 2) to the ecosystem scale
+would reveal the following relationships between g$_1$ and plant
+types:
+
+\begin{itemize}
+  \item C3 plants would have a generally higher g$_1$ than C4 plants,
+  \item Crops would have a larger g$_1$ than shrubs, grass, and
+    angiosperm trees, which have a higher g$_1$ than gymnosperm
+    trees.
+  \item Tropical and temperate climates would generally be
+    characterized by plants with a higher g$_1$ than arctic and boreal
+    climates.
+\end{itemize}
+For all of the above relationships, a higher g$_1$ means a higher
+likelihood of a positive ET response to increasing VPD (water
+intensive), and a smaller VPD dependence of the ET response.
+
+Our theoretical results highlight the variability of ET
+response to VPD as a function of both climate  and plant
+characteristics, especially water usage strategy. Whether an
+ecosystem increases or decreases ET in response to VPD will vary
+depending on vegetation and climate. Generalizations about ET response
+to VPD therefore require thoughtful consideration of ecosystem
+characteristics; statements such as ``ET increases with warming due to
+increases in VPD'' may be false depending on the water conservation
+strategy of a given ecosystem \cite{Lemordant_2018}.
+
+Additionally, the sensitivity of ET response to plant parameters
+highlights the importance of understanding and developing stomatal
+conductance models and parameters for the ecosystem scale. In
+particular, why and how g$_1$ behavior at the leaf scale is not
+analogous to the ecosystem scale must be understood. While the
+derivation here was explicit about the assumption of a constant g$_1$
+term with respect to the VPD perturbation, many sophisticated land
+surface models and earth system models employing stomatal conductance
+models make similar assumptions about the stationarity of VPD slope
+terms \cite{Niu_2011, Franks_2017, Rogers_2017, Lawrence_2019}. Many
+models will assume a constant VPD slope term (e.g. g$_1$) within
+a given plant functional type, and research derived from these models
+often does not acknowledge limitations of this assumption, or the
+difficulty in the literature with quantifying an ecosystem scale slope
+parameter \cite{Medlyn_2017}. Given that something so fundamental as
+the ET response to VPD varies strongly with g$_1$, we must invest more
+in understanding g$_1$'s behavior at the ecosystem scale, and the
+efficacy of models using constant slope terms for ecosystem-scale
+fluxes. Fortunately, our framework, and the new approach of using
+robust ecosystem-scale ratios (uWUE) to remove the implicit dependence
+of stomatal conductance on ET, is flexible enough to be applied to any
+stomatal conductance model that contains a dependence on GPP.
+
+So far we have discussed the role of different parameters in
+modulating the ET response. However, the form of the ET response is
+imposed by our choice of stomatal conductance model. We now examine
+how stomatal model choice, and specifically the exponent of the VPD
+dependence in the stomatal conductance model, can impact the general
+form of the ET-VPD relationship. There is a theoretical basis for the
+square root VPD dependence in both the stomatal conductance model and
+uWUE based on the assumption that stomata behave to maximize carbon
+gain while minimizing water loss, which observations also generally
+support \cite{Lloyd_1991, MEDLYN_2011, Lin_2015, Zhou_2014, Zhou_2015,
+  Medlyn_2017}. However, some purely empirical results that fit the
+exponent of the VPD dependence to data have shown that it may vary
+slightly from 1/2, suggesting that stomata, as well as ecosystem-scale
+quantities based on stomata theory, may not always function optimally
+\cite{Zhou_2015, Lin_2018}. Specifically with regards to uWUE, one
+would not expect that this ecosystem scale WUE quantity will respond
+to VPD exactly analogously to stomata. Direct soil evaporation's
+contributions to ET should shift the exponent of the VPD
+dependence. The results from \citeA{Zhou_2015} corroborate this: they
+found a mean empirically fit exponential VPD dependence of 0.55,
+varying slightly from the theoretically optimal value of of 0.5 for
+AmeriFlux sites. Results in \citeA{Lin_2018} also show variance in the
+empirical exponent of the VPD dependence in the stomatal conductance
+model, but interpretation of this variance is more difficult as
+\citeA{Lin_2018} do not handle the GPP dependence of stomatal
+conductance in a directly analogous manner to the optimal theory in
+\citeA{MEDLYN_2011} and \citeA{Medlyn_2017}. Regardless, given that
+these recent results on the relationship between VPD, GPP, and ET
+\cite{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017} form the
+backbone of our analysis and are what allowed us to derive an explicit
+ET expression for the first time (Eq. (\ref{et})), we will analyze if
+and how assumptions about the exponent of the VPD dependence impact
+the shape of the ET-VPD dependence. This analysis is also important to
+understand whether the choice of stomatal conductance model alters the
+fundamental behavior of the ET-VPD relationships, as many commonly
+used models utilize a VPD exponent other than the 1/2 suggested by
+optimal theory \cite<e.g. >[ which uses an exponent of
+1]{Leuning_1990}. It is useful to understand how the choice of model
+imposes the functional form of the ET-VPD relationship, given its
+fundamental importance for land-atmosphere interactions.
+
+\subsection{Functional form of ET dependence on VPD and its relation
+  to the VPD exponent}
+\label{functional_form}
+
+The results presented in Sect. \ref{results} indicates that for a
+given uWUE and $g_1$, the ET dependence on VPD should be concave
+upward. In other words, there should be some local minimum in ET at a
+critical VPD$_{crit}$, assuming the scaling and plant terms
+(e.g. aerodynamic conductance, $\Delta$, $g_1$ and uWUE) are held
+fixed. This result warrants further investigation, because to our
+knowledge no earlier work has derived the theoretical ecosystem-scale
+relationship between ET and VPD in an energy balance framework, while
+controlling for other environmental conditions. In particular there is
+an apparent lack of consensus over whether the shape of the ET-VPD
+curve should be concave upward (our result) or concave downward in the
+absence of dramatic water stress. Given that understanding the ET-VPD
+relationship of the one-way plant response is fundamental to
+hypothesizing about any feedbacks between the land surface and the
+atmosphere, we analyze why our derived ET-VPD relationship is concave
+upward, particularly with respect to the exponent of VPD dependence in
+uWUE and the Medlyn unified stomatal conductance model, as other
+models and empirical results have suggested exponents varying from 1/2
+\cite{Leuning_1990, Zhou_2015, Lin_2018}.
+
+By introducing $n$, the exponent of VPD in uWUE, and $m$, the exponent
+of VPD in the stomatal conductance model, we can free our theory from
+assumptions about VPD dependence:
+% \begin{linenomath*}
+  \begin{equation}
+    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g*}{VPD^m}\right) \frac{*WUE \; ET}{c_a \; VPD^n},
+    \label{m_n}
+  \end{equation}
+% \end{linenomath*}
+where:
+\[*WUE = \frac{GPP}{ET}VPD^n,\] and $g*$ is a generic slope parameter
+of units $VPD^m$. To determine how the exponent $n$ and $m$ alter the
+shape of the ET-VPD dependence we find the roots of the second
+derivative of ET, using Eq. (\ref{m_n}) for stomatal conductance
+($g_s$), with respect to VPD:
+
+\begin{equation}
+\frac{\partial^2 \; ET}{\partial \; VPD^2} = 0 \quad \forall \quad\frac{VPD^m}{g*} = \frac{m \left(m - 2 n - \sqrt{m^{2} - 4 m n + 2 m - 4 n^{2} + 4 n + 1} + 1\right)}{2 n \left(n - 1\right)} - 1.
+\label{curves}
+\end{equation}
+
+With this result we have defined the family of curves separating
+concave up from concave down ET solutions (Fig. \ref{concave}). These
+curves are only functions of the exponent of the VPD dependence and a
+quantity we call non-dimensional VPD ($VPD^m/g*$). Several important
+relations reveal themselves from Eq. (\ref{curves}):
+
+\begin{figure}
+  \centering
+  \centerline{\includegraphics[width=0.75\textwidth]{./concave.pdf}}
+  \caption{ Solutions corresponding to inflection points between
+    concave up and concave down ET-VPD curves (Eq. (\ref{curves})) for
+    three specific scenarios. Solutions are defined in terms of a
+    non-dimensional VPD ($VPD^m/g_*$), but to aide physical
+    interpretation the horizontal axis is additionally provided in
+    terms of dimensionalized VPD assuming $m=1/2$ and
+    $g_*=110\; Pa^{1/2}$ (average of all PFT $g_1$). The vertical axis
+    has a different interpretation depending on the solution
+    curve. For the blue line ($m$ varying), it corresponds to $m$, for
+    the orange line ($n$ varying) it corresponds to $n$, and for the
+    green line it corresponds to the value of both $n$ and $m$
+    ($n=m$). Regions of the parameter space that correspond to concave
+    up and concave down results are labeled: for each curve, the
+    parameter space ``below'' the curve corresponds to a concave up
+    relationship, while the parameter space ``above'' the curve
+    corresponds to a concave down relationship.}
+  \label{concave}
+\end{figure}
+
+\begin{itemize}
+  \item For optimal behavior (n, m = 1/2) the ET-VPD curve will be
+    concave up regardless of the magnitude of the plant constants
+    $g_1$ and uWUE. Therefore, the general concave up nature of our
+    results, given an assumption of optimal behavior, is insensitive
+    to plant type.
+  \item For all physically possible exponents of VPD dependence ($n,
+    m$), whether the solution is concave up or concave down does not
+    depend on $*WUE$.
+  \item In general, increasing the exponent of VPD dependence for
+    either $*WUE$ or $g_*$ increases the likelihood of a concave down
+    result. Additionally, as the exponent of VPD dependence increases
+    from the optimum value of 1/2, whether the curve is concave upward
+    or concave downward becomes a function of the plant specific slope
+    parameter $g_*$, through non-dimensional VPD
+    ($VPD^m/g_*$). Because the exponent of the VPD dependencies is
+    capable of altering the fundamental shape of ET-VPD dependence,
+    future research investment in understanding the exact VPD
+    dependence of stomatal conductance, and further reconciliation of
+    empirical and theoretical stomatal and ecosystem behavior should
+    be prioritized.
+\end{itemize}
+
+While it is possible (and even likely) that in the future some other
+form of VPD dependence is derived, at present \cite{MEDLYN_2011} and
+\cite{Zhou_2014} established n=m=1/2 as the most likely candidate
+given current theory and empirical data. Additionally, we argue that a
+concave up result matches physical intuition more than a concave down
+result. Plants must maintain nutrient and sugar transport through the
+phloem and xylem. To accomplish this, stomata must remain slightly
+open \cite{De_2013, Nikinmaa_2013, Ryan_2014}. Furthermore, even if
+complete stomatal closure were possible, cuticular water loss and (at
+the ecosystem-scale) direct soil evaporation are still sources of ET
+which increase with VPD, independent of stomatal closure. Therefore,
+in the limit as VPD becomes large and we assume plants are exercising
+all strategies to reduce ET, any further increase in VPD should result
+in an increase in ET through cuticular water loss and/or direct soil
+evaporation. This inevitable transition from conditions when stomata
+respond strongly to VPD to conditions when stomata response is
+asymptoting towards full closure would cause a concave up ET-VPD
+curve, which is matched by our theory. In short, plant response
+becomes more limited as VPD increases, while atmospheric demand
+monotonically increases with VPD, leading to the result that
+atmospheric demand dominates plant response when atmospheric demand is
+high.
+
+This analysis allows us to understand the theoretical shape of the ET
+response to VPD with environmental conditions held
+fixed. Accomplishing this with purely statistical methods applied to
+flux observations would be very difficult, given the relatively fast
+time scale of plant response and the non-stationarity of (solar
+forced) environmental conditions over the relatively coarse (half
+hourly) flux estimates (which is required to obtain robust
+eddy-covariance statistics). Our results on the shape of the ET-VPD
+curve with environmental conditions held fixed can be built upon with
+future work examining how changes in VPD and environmental conditions
+(e.g. soil water storage) feedback upon one another. In the soil water
+storage example, over very long time scales extremely high VPD
+perturbations coupled with no precipitation could result in decreases
+in soil water storage such that water becomes limiting. This could be
+represented by an extension of our framework in which $\lambda$ is
+allowed to decrease with decreasing soil water, increasing uWUE and
+decreasing g$_1$. Here, we focus our results by framing them with
+previous literature to build baseline intuition for ET-VPD dependence.
 
@@
 
-\input{conclusions.tex}
+We derived a new form of Penman Monteith using the concept of
+semi-empirical optimal stomatal regulation \cite{Lin_2015,
+MEDLYN_2011} and near constant uWUE \cite{Zhou_2015} to remove the
+implicit dependence of stomatal conductance on GPP and ET. With our
+new form of Penman Monteith we developed a theory for when an
+ecosystem will tend to reduce or increase ET with increasing VPD,
+which we framed using previous literature exploring the relationship
+between plant parameters, plant types, and climate. The goal was to
+understand the range of possible ET responses to VPD and develop some
+intuition for how the ET response may vary with plant types and
+climate. This intuition can be used to disentangle land atmosphere
+feedbacks in more complicated scenarios, and will aid interpretation
+of observations and more sophisticated models.
+
+ET response to VPD can vary from strongly water conservative (ET
+decreasing in response to increasing VPD), to strongly water intensive
+(ET increasing in response to VPD), which is indicative of the
+diversity of possible plant water conservation strategies. Higher uWUE
+and g$_1$ values increase the likelihood of a positive ET response,
+while decreasing temperature and increasing g$_a$ amplify the
+magnitude of the response. Previous literature \cite{Zhou_2015,
+Lin_2015} suggests that crops, through association with higher g$_1$
+and uWUE, are more likely to exhibit a positive ET response to
+VPD. Shrubs (lower uWUE), C4 plants (lower g$_1$), gymnosperm trees
+(lower g$_1$), and plants in arctic and boreal climates (lower g$_1$)
+are more likely to exhibit a negative ET response to VPD. However
+interpretation of g$_1$-induced variability is partially muddied by
+ambiguity in applying leaf-scale estimations of g$_1$ to the ecosystem
+scale \cite{Medlyn_2017}.
+
+Our paper builds intuition for how plants respond to VPD
+perturbations. We show that given optimal stomatal function and fixed
+environmental conditions, the ET-VPD dependence is theoretically
+concave upward, with ET increasing with increasing VPD as VPD
+increases past some critical value where $\frac{\partial \;
+ET}{\partial \; VPD} = 0$. However future research should focus on
+fully understanding the functional form of VPD dependence, as this
+concave up result is sensitive to the exponent of VPD dependence,
+which we currently believe is 1/2 for both uWUE and the stomatal
+conductance model \cite{MEDLYN_2011, Zhou_2014}. Indeed, this
+sensitivity to the exponent of VPD dependence is an important result
+itself: land surface models, including those used in earth system
+models for climate forecasts, employ different assumptions about the
+exponent of VPD dependence in stomatal conductance
+\cite<e.g.,>[]{Ball_1987, Leuning_1990, MEDLYN_2011}, and these
+assumptions can fundamentally change the relationship between ET and
+VPD from one that is concave upward (local minimum in ET) to one that
+is concave downward (local maximum in ET).
+
+Our results are also applicable to understanding the impact of
+expected increases in VPD induced by global change.  Plant
+physiological responses to direct CO$_2$ effects
+\cite<e.g.,>[]{Swann_2016, Lemordant_2018} receives more attention
+than physiological response to indirect effects like increased VPD
+\cite{Novick_2016}. Here, we provide a framework for understanding ET
+response to VPD using a simplified model of two plant
+parameters. Feedbacks between the land and the atmosphere may alter
+the net response to a long-timescale global VPD perturbation, but our
+focus on the one way plant response to a VPD perturbation in the
+atmospheric boundary layer is an important first step to disentangling
+such feedbacks, both in observations and model simulations of the
+present and future. By removing Penman Monteith's dependence on
+implicit relationships between GPP, VPD, and ET, we allow for explicit
+future analysis of plant-VPD feedbacks in the atmospheric boundary
+layer (Eq. (\ref{et})). Our approach can be extended to examine
+varying plant response to more nuanced consideration of plant type and
+climate. Any plant physiological heterogeneity or feedback that can be
+conceptualized with shifts in $g_1$ \cite<e.g.>[]{Lin_2015,
+Medlyn_2017} and/or uWUE \cite<e.g.>[]{Zhou_2014} are representable
+within our framework, which opens the door for a hierarchy of more
+sophisticated climate- and plant-specific analyses of ET sensitivity
+to environmental variables (including VPD). We argue that such
+simplified conceptual frameworks are critical tools for disentangling
+land-atmosphere feedbacks at various scales, from diurnal to seasonal
+and beyond, and to characterize ET response in a warmer,
+atmospherically drier, and enriched CO$_2$ world.
+
+\nocite{AT-Neu, AU-ASM, AU-ASM, AU-Cpr, AU-DaP, AU-DaS, AU-Dry, AU-Gin,
+  AU-How, AU-Gin, AU-DaP, AU-Tum, AU-Whr, AU-Wom, BE-Lon, BE-Vie,
+  BR-Sa3, CA-Qfo, CA-SF1, CA-SF1, CH-Cha, CH-Dav, CH-Fru, DE-Geb,
+  DE-Gri, DE-Hai, DE-Gri, DE-Lkb, DE-Seh, DE-Tha, DK-Sor, FI-Hyy,
+  FI-Sod, FR-Gri, FR-LBr, IT-Col, IT-Cpz, IT-Lav, IT-MBo, IT-Noe,
+  IT-Ren, IT-Ro2, IT-SRo, IT-Tor, NL-Loo, RU-Fyo, US-AR1, US-AR1,
+  US-ARM, US-Blo, US-KS2, US-MMS, US-Me2, US-NR1, US-Ne1, US-Ne1,
+  US-Ne1, US-SRG, US-SRM, US-Syv, US-Ton, US-Var, US-WCr, US-Wkg,
+  ZA-Kru, ZM-Mon}
 
@@
 in this analysis, including those used to generate the figures and
-tables, are publicly available at \url{https://github.com/massma/climate\_et}
+tables, are publicly available at
+\url{https://github.com/massma/climate\_et}. We would
+like to thank Emma Robinson and one anonymous reviewer for thoughtful
+comments that greatly improved the strength of the manuscript.
 
-\begin{appendices}
-  \section{FLUXNET2015 sites}\label{flux_sites}
-  % below taken adapted from Trevor Keenan's FLUXNET citations
-  % at github.com/trevorkeenan/FLUXNET_citations
-  Table A1: Metadata and citations for flux sites used in this analysis.
-All data are gathered from \url{www.fluxdata.org}, and citations are aggregated using tools available at \url{https://github.com/trevorkeenan/FLUXNET_citations}.
-  \begin{longtable}{l l l l l l l}
-    \hline
-    \textbf{Site} &
-    \textbf{PFT} &
-    \textbf{Lat} &
-    \textbf{Lon} &
-    \textbf{Clim\textsuperscript{1}} &
-    \textbf{Period} &
-    \textbf{References} \\
-    [0.5ex]
-    \hline
-    \input{flux_sites.tex}
-    [1ex]
-    \hline
-  \end{longtable}
-  \textsuperscript{1}  K{\"o}ppen Climate classification.
-\end{appendices}
-
-\bibliography{references.bib}
+\bibliography{references}
 

### 2019-10-16 doc/paper/ms.tex
updated to fix typos in prrof

@@
   $g_s$ is a function of carbon uptake, which is has a strong
-  functionally relation to ET through stomatal function. Therefore,
+  functional relation to ET through stomatal function. Therefore,
   when PM is formulated in terms of $g_s$, it is an implicit function
@@
     $P$  & pressure & Pa  & - \\
-    $\Delta$  & $\frac{\partial e_s}{\partial T}$ & Pa K$^{-1}$ & - \\
+    $\Delta$  & $\frac{d \, e_s}{d \, T}$ & Pa K$^{-1}$ & - \\
     $R_{net}$  & net radiation at land surface minus ground heat flux & W m$^{-2}$   & - \\
@@
 g$_1$ & Pa$^{1/2}$ [kPa$^{1/2}$] & 63.25 [2.00] & 126.49 [4.00] & 189.74 [6.00] & Fig. 2, 7; \citeA{Medlyn_2017} \\
-uWUE & $\mu$-mol C Pa$^{0.5}$ J$^{-1}$ ET [g C hPa$^{1/2}$ kg$^{-1}$ H$_2$O] & 2.33 [6.99] & 3.17 [9.52] & 4.01 [12.05] & Table 4; \citeA{Zhou_2015} \\
+uWUE & $\mu$-mol C Pa$^{1/2}$ J$^{-1}$ ET [g C hPa$^{1/2}$ kg$^{-1}$ H$_2$O] & 2.33 [6.99] & 3.17 [9.52] & 4.01 [12.05] & Table 4; \citeA{Zhou_2015} \\
 T & $^o$C & 10.00 & 20.00 & 30.00 & - \\
@@
 
-\acknowledgments This work used eddy covariance data acquired and shared
-by the FLUXNET community, including these networks: AmeriFlux, AfriFlux,
-AsiaFlux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont,
+\acknowledgments This work used eddy covariance data acquired and
+shared by the FLUXNET community, including these networks: AmeriFlux,
+AfriFlux, AsiaFlux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont,
 ChinaFlux, Fluxnet-Canada, GreenGrass, ICOS, KoFlux, LBA, NECC,
 OzFlux-TERN, TCOS-Siberia, and USCCC. The ERA-Interim reanalysis data
-are provided by ECMWF and processed by LSCE. The FLUXNET eddy covariance
-data processing and harmonization was carried out by the European Fluxes
-Database Cluster, AmeriFlux Management Project, and Fluxdata project of
-FLUXNET, with the support of CDIAC and ICOS Ecosystem Thematic Center,
-and the OzFlux, ChinaFlux and AsiaFlux offices. We would like to thank
-Dr. Trevor Keenan for providing tools for citation of the FLUXNET2015
-dataset, available at
-\url{https://github.com/trevorkeenan/FLUXNET_citations}. This material is
-based upon work supported by the National Science Foundation Graduate
-Research Fellowship under Grant No. DGE 16-44869. All code and data used
-in this analysis, including those used to generate the figures and
-tables, are publicly available at
-\url{https://github.com/massma/climate\_et}. We would
-like to thank Emma Robinson and one anonymous reviewer for thoughtful
-comments that greatly improved the strength of the manuscript.
+are provided by ECMWF and processed by LSCE. The FLUXNET eddy
+covariance data processing and harmonization was carried out by the
+European Fluxes Database Cluster, AmeriFlux Management Project, and
+Fluxdata project of FLUXNET, with the support of CDIAC and ICOS
+Ecosystem Thematic Center, and the OzFlux, ChinaFlux and AsiaFlux
+offices. We would like to thank Dr. Trevor Keenan for providing tools
+for citation of the FLUXNET2015 dataset, available at
+\url{https://github.com/trevorkeenan/FLUXNET_citations}. This material
+is based upon work supported by the National Science Foundation
+Graduate Research Fellowship under Grant No. DGE 16-44869 and the
+National Aeronautics and Space Administration under grant
+80NSSC18K0998. All code and data used in this analysis, including
+those used to generate the figures and tables, are publicly available
+at \url{https://github.com/massma/climate\_et}. We would like to thank
+Emma Robinson and one anonymous reviewer for thoughtful comments that
+greatly improved the strength of the manuscript.
 
